### 2016-04-13 docs/papers/2015eurasip/2014eurasip (copy).tex
adding papers and their assets

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}
\usepackage[titletoc,title]{appendix}
%%% Load packages
\usepackage{amsthm,amsmath}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
%\RequirePackage{natbib}
%\RequirePackage{hyperref}
%\usepackage[utf8]{inputenc} %unicode support
\usepackage[mathletters]{ucs}
\usepackage[utf8x]{inputenc}
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

\usepackage{algorithm,algorithmic}% http://ctan.org/pkg/algorithms

\usepackage{natbib}

\usepackage{supertabular}
\usepackage{fancybox}
\usepackage{acronym}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{setspace}
\usepackage{placeins}
\usepackage{longtable}
\usepackage{mathtools}
\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
\usepackage{listings}
\usepackage{epstopdf}
\usepackage{subfigure}
\usepackage{courier}
\usepackage{amsfonts}
\usepackage{morefloats}
\usepackage{lipsum}
\newcommand*\justify{%
	\fontdimen2\font=0.4em% interword space
	\fontdimen3\font=0.2em% interword stretch
	\fontdimen4\font=0.1em% interword shrink
	\fontdimen7\font=0.1em% extra space
	%\hyphenchar\font=`\-% allowing hyphenation
}

% appendix
\usepackage{hyperref}
\usepackage{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\def\includegraphic{}
%\def%\includegraphics{}

% Algorithmic modifications
\makeatletter
\newcommand{\ALOOP}[1]{\ALC@it\algorithmicloop\ #1%
  \begin{ALC@loop}}
\newcommand{\ENDALOOP}{\end{ALC@loop}\ALC@it\algorithmicendloop}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\newcommand{\algorithmicbreak}{\textbf{break}}
\newcommand{\BREAK}{\STATE \algorithmicbreak}
\makeatother

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
	addressref={address1},                   % id's of addresses, e.g. {aff1,aff2}
	corref={address1},                       % id of corresponding address, if any	
	%noteref={n1},                        	% id's of article notes, if any
	email={tpbvieira@gmail.com} 				% email address
]{\inits{TPBV}\fnm{Thiago P. B.} \snm{Vieira}}
\author[
   	addressref={address1},
   	email={???????????????}
]{\inits{DFT}\fnm{Danilo F.} \snm{Tenório}}
\author[
   	addressref={address1},
   	email={???????????????}
]{\inits{JPCLC}\fnm{João Paulo C. L.} \snm{da Costa}}
\author[
   	addressref={address1,address2},
   	email={???????????????}
]{\inits{EPF}\fnm{Edison P.} \snm{de Freitas}}
%\author[
%   	addressref={address3},
%   	email={???????????????}
%]
%{\inits{CHC}\fnm{\\Carlos H.} \snm{Cardonha}}
\author[
   	addressref={address1},
   	email={???????????????}
]{\inits{RTSJ}\fnm{\\Rafael T.} \snm{de Sousa Júnior}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=address1]{                           								% unique id
	\orgname{Department of Electrical Engineering, University of Brasilia (UnB)},	% 	university, etc
%	\street{},                     												%
	\postcode{70910-900}                                							% post or zip code
	\city{Brasília-DF},                              							% city
	\cny{Brazil}                                    								% country  
}
\address[id=address2]{%
	\orgname{Graduate Program in Electrical Engineering, Federal University of Rio Grande do Sul (UFRGS)},
%	\street{},
	\postcode{90035-190}
	\city{Porto Alegre},
	\cny{Brazil}
}
%\address[id=address3]{                           								% unique id
%	\orgname{IBM Research},	% 	university, etc
%	\street{},                     												%
%	\postcode{}                                							% post or zip code
%	\city{São Paulo},                              							% city
%	\cny{Brazil}                                    								% country  
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
	%\note{Sample of title note}     	% note to the article
	%\note[id=n1]{Equal contributor} 		% note, connected to author
\end{artnotes}

\end{fmbox}								% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract

Novel schemes for attack detection are crucial to identify adaptive malicious traffic coming from sources that are quickly mobilized by attackers in high throughput communication networks. In this context, signal processing techniques have been applied to attack detection due to their capability to detect anomalies that are previously unknown, i.e. blind detection. This paper proposes a signal processing framework for the detection and identification of network attacks using concepts of model order selection (MOS), eigenvalues and similarity analysis. In order to validate our proposed framework, we consider a network traffic sample that contains malicious activity such as denial of service (DoS) and port probing. We propose to model the network traffic as a superposition of components, namely, user's operations (legitimate traffic), network service operation not related to the user (noise) and the malicious activity. The experiments performed in a real network show that the proposed blind detection approach achieves satisfactory levels of accuracy in terms of timely detection and identification of TCP/UDP ports under attack.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
	\kwd{Network Attack Detection}
	\kwd{Model Order Selection}
	\kwd{Eigen Analysis}
	\kwd{Similarity Analysis}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section{Introduction}
\label{sec:introduction}

Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, but may fail against innovative malicious techniques \cite{lakhina2005mining}. In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time. Moreover, it is also crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns. Recently, signal processing schemes have being applied to the detection of malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest}, showing advances in network traffic analysis.

Information security may consist of both technical and procedural aspects. The former includes equipment and security systems, while the latter corresponds to security rules and recommendations. Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks. Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.

In the context of anomaly-based schemes, this work proposes an approach for the detection of malicious traffic in computer networks. Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP). Our proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis. In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack. In addition, we also evaluate the accuracy of the proposed framework. Additionally, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.

The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of our proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.

This paper is organized as follows. In Section \ref{sec:relatedworks}, related works are discussed. Section \ref{sec:datamodel} presents the data model. Section \ref{sec:prop_getv} describes our proposed framework for blind and automatic detection of malicious traffic. Section \ref{sec:experimentalresults} discusses the experimental validation and presents the corresponding experimental results. Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work. The Appendix \ref{sec:mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.

\section{Related Works}
\label{sec:relatedworks}

Several methods have been proposed for the identification and characterization of malicious activity in computer networks. Classical methods typically employ data mining \cite{he2008applying,ghourabi2010data} and regular file analysis \cite{raynal2004honeypot} to detect patterns that indicate the presence of specific attacks in network traffic.

Data mining is often used to describe the process of extracting useful information from large databases. Multiple methods of data mining are used in \cite{he2008applying} to analyze data flow in a network with the aim of identifying characteristics of malicious traffic in large scale environments. Researchers have applied data mining techniques in log analysis \cite{ghourabi2010data} to improve intrusion detection performance. However, data mining techniques require prior collection of large data sets, which is a weakness of several schemes for online or low latency analysis \cite{Huang2009}.

Regular file analysis \cite{raynal2004honeypot} consists of using traffic analysis for detecting known patterns that indicate the presence of specific attacks, applying statistical analysis to the study of collected traffic. An essential feature of this method is that it depends on prior knowledge of the details of the attacks to be identified, and also depends on previous log collection for applying traffic analysis and reducing false positives.

Principal Component Analysis (PCA) is a statistical technique commonly used for dimensionality reduction. It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance. PCA has been used in attack detection \cite{almotairi2009technique}. However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques. Besides being prone to higher errors and false positives, such human intervention makes PCA not useful for real time applications. Therefore, in order to automate the analysis of eigenvalues profile, model order selection (MOS) schemes should be incorporated.

Signal processing techniques have been successfully applied to network anomaly detection \cite{Lu2009}. Lu and Ghorbani \cite{Lu2009} proposed a network anomaly detection model based on network flow, wavelet approximation, and system identification theory. However, their work did not address problems without significant outliers, such as port scan attacks. Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies. Zonglin \emph{et al} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work was not applied to port scan and denial of service (DoS) attack detection, simultaneously.

The data collected from honeypot systems, such as captured traffic and operating system logs, can be analyzed to obtain information about attack techniques, general trends of threats and exploits. Blind automatic detection of malicious traffic techniques have been developed for honeypots in \cite{david2011blind,da2012improved}. However, traffic on honeypot is simpler than real network traffic, because there are no running legitimate applications, due to the fact that honeypots emulate behavior of a host within a network to deceive and lure attackers \cite{zakaria2012review}. Since honeypots do not generate legitimate traffic, the amount of data captured in honeypots is significantly lower in comparison to a Network Intrusion Detection System (NIDS), which captures and analyzes the largest possible amount of network traffic \cite{david2011blind}. MOS for blind identification of malicious activities in honeypots was proposed by us in \cite{david2011blind}, which evaluated criteria for selecting the model order, through simulations and comparing the order of the resulting model with the true model order.

Lee \emph{et al} \cite{Lee2013} proposed OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling. In contrast to Lee \emph{et al}, our framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack. Additionally, Lee \emph{et al} only evaluated their proposed scheme for covariance analysis, while we adopted covariance and correlation analysis for DoS and probing attacks, respectively.

Our proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic. Our proposed solution is automatic and blind for detection of time frames under probing and DoS attacks through MOS and eigen analysis. Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.

\section{Data Model}
\label{sec:datamodel}

In this paper the scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}. The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively. We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz.

This section presents details of the simulated scenario along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic. The Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce DoS and probing attacks. The Subsection \ref{sec:ModelingData} presents how network traffic can be modelled as signal superposition, and the Subsection \ref{sec:AnalyzedScenario} details the network traffic of synflood, fraggle and port scan attacks.

\subsection{Analyzed Scenario and Data Collection}
\label{sec:DataCollection}

The environment of the analyzed scenario is composed of two computers and one router with access to Internet and to an internal network, where the simulation of legitimate traffic, noise, DoS and port scan attacks are performed. During the traffic generation, one computer assumes the role of attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:fig1}.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=5cm, width=7cm]{results/figures/fig09.png}
     \caption{\csentence{Analyzed Scenario.} Scenario to reproduce legitimate traffic, noise, DoS and port scan.}
     \label{fig:fig1}
\end{figure}

Note that the set of network traffic is modeled as legitimate, noise and malicious traffic, where the victim performs legitimate activities, that can be characterized by web access. In many organizations this type of traffic is predominant, since most of corporate services are web-based, such as: web pages, customized web-based systems and cloud services. It is possible to characterize the traffic of a DHCP service as an example of noise associated with the transport layer. For malicious traffic, three types of networks attacks are evaluated: synflood, fraggle and port scan. These attacks are reproduced using well-known security tools, such as Nmap to port scan, Metaspoit to synflood and Hping to lead the fraggle attack.

A network traffic log is commonly formed by timestamp, protocol, source IP address, source port, destination IP address, destination port and additional information, according to the type of transport protocol used. The following TCP traffic log is presented in order to exemplify the collected data:
\newline
\newline
\texttt{\justify21:00:34.099289 IP 192.168.1.102.34712 > 200.221.2.45.80: Flags [S], seq 2424058224, win 14600, options [mss 1460, sackOK,TS val 244136 ecr 0,nop,wscale 7], length 0}
\newline

and the following to exemplify UDP traffic log: 
\newline
\newline
\texttt{\justify21:24:42.484858 IP 192.168.1.102.68 > 192.168.1.1.67: BOOTP/DHCP, Request from 00:26:9e:b7:82:be, length 300}
\newline

In the proposed framework, the goal is to detect the anomalies only taking into account the traffic profile, i.e. specific information such as origin IP or day and time of the attack are not considered. Therefore, from the whole log information, we just consider the timestamp (for sequencing), port type and port number.

\subsection{Modeling Data}
\label{sec:ModelingData}

Modeling the dataset as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{S}), noise (\textbf{N}) and malicious traffic (\textbf{A}), according to the following expression:

\begin{equation}\label{eq:eq01}
\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{N}^{(q)} + \boldsymbol{A}^{(q)},
\end{equation}
where $q$ represents the $q$-th time frame, which is a time grouping of network traffic.

The matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns. Each row represents a communication port (TCP port or UDP port), and each column represents the time, whose unit can be minutes, for instance. Each element $x_{m,n}^{(q)}$ stands for the number of times that the port $m$ appears at the $n$-th minute, at the $q$-th time frame.

The legitimate traffic $\boldsymbol{U}^{(q)}$ is characterized by the traffic from user's operations. When a user accesses a web page, for example, there is the corresponding TCP/IP traffic to request the page, as well as there is the traffic required to domain name resolution. Figure \ref{fig:fig3} depicts an example of the legitimate traffic obtained during experiments, represented by traffic to the ports 67 and 68, which are related to the Bootstrap Protocol and to normal DHCP operations, respectively.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig03.png}
     \caption{\csentence{Legitimate Traffic.} The traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
     \label{fig:fig3}
\end{figure}

All traffic, that is not directly associated with users' operations, but it is not a malicious traffic, is considered as noise $\boldsymbol{N}^{(q)}$. The automatic acquisition service of logical IP network address (DHCP) is an example of noise. Independently of any user operation, the machine receives an IP address, since it is configured to perform a DHCP address request. Figure \ref{fig:fig4} depicts an example of noise in a network traffic.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig04.png}
     \caption{\csentence{Noise Traffic.} Network traffic of user independent operations for network management.}
     \label{fig:fig4}
\end{figure}

The traffic coming from a malicious activity, such as a synflood or fraggle attack, is represented by the matrix $\boldsymbol{A}^{(q)}$. For this work we only consider the traffic from port scanning and flood attacks, which aims to cause denial of service.

We define that if the obtained $\#\boldsymbol{A}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\boldsymbol{A}^{(q)} = 0$, then there is no malicious traffic. This paper shows how to detect the $\#\boldsymbol{A}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious network traffic.

\subsection{Synflood, Fraggle and Port scan}
\label{sec:SynfloodFraggleandPortscan}

The network attacks evaluated by this work are: synflood, fraggle and port scan. The first two attacks can be qualified as DoS attacks, while the last one can be qualified  as probing or port scanning attack.

The TCP protocol is a connection-oriented protocol, then a virtual connection must be established between two computers for an end-to-end TCP communication. This virtual connection requires a ``handshake'', that occurs in three steps, known as three-way handshake. If a computer needs to communicate with another computer, the requester sends a packet communication synchronization (SYN) to a specific destination port, which is in listening state. If the destination is active, running and accepting requests, it responds to the requester with a SYN/ACK confirmation message. After receiving this message, the requester sends an ACK message to the destination and then the connection is established.

With respect to the synflood attacks, the attacker sends a large quantity and concurrent successive SYN requests to a target, in order to consume resources and cause a DoS. Figure \ref{fig:fig5} depicts an example of a synflood attack carried out in a real computer network. In an interval of ten minutes, more than 210,000 packets are sent as a synflood attack. This network traffic behavior can be considered an abnormal behavior of network traffic, especially since it is concentrated in a short period of time and presents similar outstanding traffic during the time under attack.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig05.png}
     \caption{\csentence{Synflood Traffic.} A large quantity of SYN requests to a target, in order to cause a DoS.}
     \label{fig:fig5}
\end{figure}

With respect to the fraggle atack, large packets with ``UDP echo'' segments are sent to the broadcast address of a network. Every packet is modified to have the source address of the victim, in order to implement the source address spoofing technique. Therefore, each host receives a huge amount of requests ``UDP echo'' and all of them replies to the IP address of the victim, causing a packet flooding aiming a DoS. This attack can affect the entire network, since all hosts receive several requests ``UDP echo'' and respond with the ICMP protocol, therefore each host acts as an ``amplifier'' of the attack. This last part of the fraggle attack is not taken into account in this work, because the victim receives ICMP (network layer) packets originated from the hosts that are attacked with flooding packet ``UDP echo''. This occurs due to the UDP be not able to know if the segment sent has reached its destination.

Figure \ref{fig:fig6} depicts an example of the fraggle attack in a real computer network. More than 6,000,000 malicious packets can be counted in an interval of ten minutes, which can be considered an abnormal network traffic, especially due to the concentrated traffic in a short period of time and due to the similarity of the outstanding traffic.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig06.png}
     \caption{\csentence{Fraggle Traffic.} Large amount of “UDP echo” requests and replies, causing packet flooding.}
     \label{fig:fig6}
\end{figure}

Port scan is the attempt to establish a connection to TCP and UDP ports to identify what services are running or are in the listening state. There are several available port scanning techniques, including: TCP SYN scan, TCP ACK scan and UDP scan. This work evaluates the use of TCP SYN scan and UDP scan.

In TCP SYN scan, a SYN packet is sent to the destination and two types of responses may occur: SYN/ACK or RST/ACK. In the first case, the destination port is in the listening state, in the second case, the destination port is not listening. At the end of each port scanning, a RST/ACK packet is sent by the system that is performing the port  scan. Therefore, a full connection or a complete three-way handshake is never established, which makes the detection of the attack sender more difficult, and requires approaches able to identify probing attacks without connection stablishment. The UDP scan technique sends UDP packets to the destination port, and if it responds with a ``ICMP port unreachable'' message, then it indicates that the scanned port is closed. On the other hand, if a message is not received, then the port is considered as open.

\begin{figure}[h!]
     \centering 
     \includegraphics[height=5cm, width=9cm]{results/figures/fig07.png}
     \caption{\csentence{Port Scan Traffic.} Connection attempts in order to identify active ports.}
     \label{fig:fig7}
\end{figure}

Figure \ref{fig:fig7} depicts an example of the port scan attack in a real computer network. Note that the traffic is composed of two packets for each TCP port and one UDP packet to each port. The incoming and outgoing packets analysis, for each port, shows the high correlation and similarity of TCP and UDP traffic during the simulated port scan attack.

\section{Proposed Framework for Detection and Identification of Network Attacks}
\label{sec:prop_getv}

This section describes the proposed technique to detect synflood, fraggle and port scan. In Subsection \ref{sec:prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvelue for each $q$-th time frame, in order to apply a selected MOS scheme for attack detection, which is described in Subsection \ref{sec:prop_MOSSchemes}. In Subsection \ref{sec:prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:prop_EigenSimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.

\subsection{Largest Eigenvalue by Time Frames}
\label{sec:prop_LargestEigenvaluebyTimeFrames}

The proposed attack detection algorithm starts by the data pre-processing of a network traffic log containing IP, ports and timestamp of senders and receivers. During this step, the desired information is extracted in order to classify and count packets according to the origin and destination ports, and subsequently this information is grouped by minutes and by time frames.

With the data grouped into $Q$ time frames, the framework considers the time variations of the matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the attack.

According to DoS and port scan attacks' behavior, DoS attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively. These characteristics are substantiated by the results obtained through the covariation and correlation analysis described in Section \ref{sec:experimentalresults}, which shows that the main components of DoS attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents a highly correlated traffic.

Therefore, to detect DoS attacks, it is necessary to calculate the covariance matrix $\boldsymbol{R}_{yy}^{(q)}$ of the zero mean samples given by

\begin{equation}\label{eq:eq02}
\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
\end{equation}

The set of obtained vectors $\boldsymbol{y}_{m}^{(q)}$ composes the zero mean matrix $\boldsymbol{Y}^{(q)}$, then the covariance matrix $\boldsymbol{R}_{yy}^{(q)}$ can be calculated as follows

\begin{equation}\label{eq:eq03}
\boldsymbol{R}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
\end{equation}

For the port scan attack detection, it is necessary to calculate the correlation matrix $\boldsymbol{R}_{zz}^{(q)}$, instead of the covariance matrix $\boldsymbol{R}_{yy}^{(q)}$ used for DoS detection, since the main components are not dominated by the variables with large variance and since port scan attack presents a highly correlated network traffic. To obtain the correlation matrix $\boldsymbol{R}_{zz}^{(q)}$, it is required, for each variable, to make the standard deviation unitary as follows

\begin{equation}\label{eq:eq04}
\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
\end{equation}

The set of vectors $\boldsymbol{z}_{m}^{(q)}$ composes the matrix $\boldsymbol{Z}^{(q)}$, then the correlation matrix $\boldsymbol{R}_{zz}^{(q)}$ can be calculated via

\begin{equation}\label{eq:eq05}
\boldsymbol{R}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
\end{equation}

Once the $\boldsymbol{R}_{yy}^{(q)}$ and $\boldsymbol{R}_{zz}^{(q)}$ have been obtained for DoS and port scan attack detection, respectively, for the sake of simplicity, we refer to $\boldsymbol{R}_{yy}$ and $\boldsymbol{R}_{zz}$ as a matrix $\boldsymbol{C}$. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:eq06}), in order to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).

\begin{equation}\label{eq:eq06}
\boldsymbol{C}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
\end{equation}
\begin{equation}\label{eq:eq060}
\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}).
\end{equation}

The eigenvalues should be sorted in descending order, as defined by $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ... > \lambda_{m}^{(q)}$, to make possible the selection of the first eigenvalue in the obtained sequence, represented by $\lambda_{1}^{(q)}$, which is the largest eigenvalue of the $q$-th time frame evaluated for attack detection.

The matrix of eigenvalues of $\boldsymbol{C}^{(q)}$ can be represented in the following fashion

\begin{equation}\label{eq:eq07}
\boldsymbol{E} =
\begin{bmatrix}
  \lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
  \lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
  \lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
  \vdots & \vdots & \ddots & \vdots  \\
  \lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
\end{bmatrix}.
\end{equation}

The process of obtaining the $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, $q = 1, 2, 3, ..., Q$ and the matrix $\boldsymbol{C}^{(q)}$, finding the largest eigenvalue for each $q$-th time frame, should be repeated until $q = Q$, in order to obtain the largest eigenvalue of all time frames. Since $\lambda_1^{(q)} > \lambda_2^{(q)} > \lambda_3^{(q)} > \cdots > \lambda_{m-1}^{(q)} > \lambda_m^{(q)}$, then the first line of the matrix $\boldsymbol{E}$ contains the largest eigenvalues of each $q$-th time frame, which is the Greatest Eigenvalue Time
Vector (GETV) \cite{tenorio2013greatest}, denoted as

\begin{equation}\label{eq:eq08}
\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\}.
\end{equation}

\subsection{MOS Schemes}
\label{sec:prop_MOSSchemes}

Once obtained the largest eigenvalue of each $q$-th time frame, it is possible to apply a selected MOS scheme to estimate the model order $\hat{d}$, which is the estimated number of time frames under attack. Therefore, $\boldsymbol{e}_{\rm max}$ is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\boldsymbol{e}_{\rm max})$. Note that some MOS schemes may also require the number of minutes that compose a time frame, as $\hat{d} = \rm{MOS}(\boldsymbol{e}_{\rm max},\emph{Q})$. For more information about MOS, we refer to Appendix \ref{sec:mos}.

In our previous work \cite{tenorio2013greatest}, the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE schemes are evaluated for synflood and port scan attack detection, showing that EDC and EFT are effective for detecting this kind of attacks. The present work extends that evaluation to also analyse the effectiveness of the listed MOS schemes for fraggle attack detection, as shown in Section \ref{sec:experimentalresults}.

\subsection{Eigenvalue Analysis}
\label{sec:prop_EigenvalueAnalysis}

The selected MOS scheme can estimate the number of time frames under attack, denoted by $\hat{d}$, but it does not provide information of which $q$-th time frames are under attack. However, the time frames under attack can be estimated through the largest eigenvalue analysis, which indicated that the time frames with the $\hat{d}$ largest eigenvalues correspond to the $q$-th time frames under attack.

The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed as Algorithm \ref{alg:alg01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$. The Algorithm \ref{alg:alg01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$ (line 2) and its correspondent index, according to lines 4 and 5. Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$ (line 8) and a new iteration is performed until $\boldsymbol{e}_{\rm max} = \{\}$.

\begin{algorithm}[h!]
	\caption{Detection of Time Frames Under Attack}
  	\label{alg:alg01}
	\begin{algorithmic}[1]
		\show\LOOP
	    \ALOOP {$kk = 1$ until $kk == \hat{d}$} 
	    	    \STATE $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
    	    	    \ALOOP {$i = 1$ until $i == Q$} 
				\IF {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
				    \STATE $\boldsymbol{\hat{q}}_{\rm max}^{(kk)} = i$
				\ENDIF
        		\ENDALOOP
	    		\STATE $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(kk)}$
    		\ENDALOOP
	\end{algorithmic}
\end{algorithm}

After the estimation of the $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked. To deal with this problem, the adoption of a similarity analysis between legitimate traffic and the traffic of time frames estimated as under attack is evaluated, analysing the effectiveness of cosine similarity to highlight abnormalies inserted by network traffic attacks.

\subsection{Eigen Similarity Analysis}
\label{sec:prop_EigenSimilarityAnalysis}

Cosine similarity calculates the cosine of the angle between two vectors, which represents the similarity of values between the selected vectors. Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\boldsymbol{V}^{(q)}$ against the the most significant eigenvectors of time frame detected as under attack, to analyse similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.

\subsubsection{Time Similarity Analysis}
\label{sec:prop_TimeSimilarityAnalysis}

For eigen similarity analyis, we evaluate the cosine similarity in order to identify lacks of similarity between legitimate and malicious traffic, as follows

\begin{equation}\label{eq:eq11}
s_n = \abs{\frac{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}}}},
\end{equation}
where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, and $\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of traffic to be performed the DoS and port scan attack identification.

The most significant eigenvector $\boldsymbol{v}^{(q)}$, of a time frame $q$ without attack, can be derivated from (\ref{eq:eq06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the evaluated matrix. The same calculation shall be performed in order to obtain the target eigenvectors $\boldsymbol{v}_{(n)}$, calculated from a time frame without attack plus minutes of a time frame estimated as under attack, to evaluate the occurrence of network attacks.

The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic whithout attack, from a time frame $q$ composed of $Q$ minutes of legitimate network traffic. For the detailed attack identification, each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$, as represented by

\begin{equation}\label{eq:eq12}
\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
\end{equation}

The resultant $\boldsymbol{X}_{(n)}$ is necessary to obtain $\boldsymbol{v}_{(n)}$, through (\ref{eq:eq06}), for calculating the similarity degree $s_n$, ranging from 0 to 1, for each $n$-th minute. The $s_n$ denotes the absolute similarity degree of the $n$-th minute in comparison to a well-known traffic without attack, detected through MOS schemes and eigenvalue analysis.

The incremental approach for similarity analysis is based on the incremental appending of network traffic into $\boldsymbol{X}^{(q)}$, where the first evaluation is based on (\ref{eq:eq12}) and the subsequent evaluations is based on (\ref{eq:eq13}), incrementally appending each $n$-th minute until $n=N$.

\begin{equation}\label{eq:eq13}
\boldsymbol{X}_{n} = \{\boldsymbol{X}_{n} | \boldsymbol{x}^{(\hat{q})}_{(n)}\},
\end{equation}

The Figure \ref{fig:fig8} ilustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.

\begin{figure}[h!]
     \includegraphics[height=3.6cm, width=12.4cm]{results/figures/incremental.eps}
     \caption{\csentence{Scenario.} Traffic selection for incremental approach.}
     \label{fig:fig8}
\end{figure}

The eigen similarity analysis starts at $\boldsymbol{x}^{(\hat{3})}_{(41)}$ and is incrementally performed until $\boldsymbol{x}^{(\hat{3})}_{(60)}$, in order to calculate the $s_n$. We assume that $s_n < l$ means an attack identification, according the anomaly on similiarity of $s_n$ to a defined limiar $l$. Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}$ for eigen similarity analyis, the $s_n$ is calculated according to (\ref{eq:eq11}).

If $s_n = 1$, then the two eigenvectors are completely similar and no anomaly is detected. Smaller values of $s_n$ mean less similarity and can indicate an anomaly, according to a defined 
threshold $l$, assuming that if $s_n < l$, then a network attack is identified during the $n$-th minute. Therefore, the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack was identified, according to

\begin{equation}\label{eq:eq14}
  \boldsymbol{\hat{n}_{(n)}}=\left\{
  \begin{array}{@{}ll@{}}
    1, & \text{if}\ s_n < l \\
    0, & \text{otherwise}
  \end{array}\right.
\end{equation}
where $\boldsymbol{\hat{n}}_{(n)}$ denotes a vector of $n$-th minutes detected as under attack.

The eigen similarity analysis can also be applied in an individual fashion, where each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, as shown by Figure \ref{fig:fig9}.

\begin{figure}[h!]
     \includegraphics[height=3.6cm, width=12.4cm]{results/figures/individualized.eps}
     \caption{\csentence{Scenario.} Traffic selection for individual approach.}
     \label{fig:fig9}
\end{figure}

The incremental and the individual approaches can be combined to obtain the incremental individualized approach, where each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack. Subsequently, $\boldsymbol{X}_n$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:fig2}.

\begin{figure}[h!]
     \includegraphics[height=3.6cm, width=12.4cm]{results/figures/incremental_individualized.eps}
     \caption{\csentence{Scenario.} Traffic selection for incremental individualized approach.}
     \label{fig:fig2}
\end{figure}

This approach of incremental similarity analysis followed by individual analysis after an attack detection allows to identify the attack period, highlighting the first and last time under attack. This identification is possible due to the variation of the most significant eigenvectors, which becomes more significant when compared a traffic under attack against a traffic with no attack, according to results which are discussed in Section \ref{sec:experimentalresults}.

\subsubsection{Port Similarity Analysis}
\label{sec:prop_PortSimilarityAnalysis}

Given $\boldsymbol{\hat{n}}$, which is the set of $n$-th minutes under attack, it is still necessary to obtain more details about the identified network attack, such as the network ports that are attacked during each $n$-th minute identified as under attack. Hence, it is also applied the cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port during a $n$-th minute.

For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack shall be used as reference for similarity analysis against the $\boldsymbol{v}_{(n)}$ identified as under attack, individually evaluating the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.

Therefore, $\boldsymbol{v}^{(q)}$ should be calculated from the last $\boldsymbol{X}^{(q)}$ time frame without attack, and $\boldsymbol{v}_{(m,\hat{n})}$ should be calculated from the same traffic appened of all $n$-th minutes until the identified minute under attack, denoted as $\boldsymbol{X}_n$.

For similarity analyis, each $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, denoted as $x_{(m,n)}$, shall be individually replaced by the traffic of the evaluated $m$-th port of the $\hat{n}$-th minute under attack, denoted as $x^{(\hat{q})}_{(m,\hat{n})}$, in order to identify significant variation on similarity caused by the traffic of the $m$-th port.

This approach for detection of ports under attack via similarity analysis is given by

\begin{equation}\label{eq:eq15}
  \left\{
  \begin{array}{@{}ll@{}}
    x_{(m,n)} = x^{(\hat{q})}_{(m,\hat{n})} \\
    \\
    s_{m,\hat{n}} = \abs{\frac{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(m,\hat{n})}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(m,\hat{n})}}}},
  \end{array}\right.
\end{equation}
where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, which is used to calculate the $\boldsymbol{v}_{(m,\hat{n})}$ most significant eigenvectors that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.

Once $\boldsymbol{v}^{(q)}$ and $\boldsymbol{v}_{(m,\hat{n})}$ are obtained, then the $s_{m,\hat{n}}$ similarity degree can be calculated in order to identify if the traffic replacement highlights the adition of anomalous traffic by the evaluated $m$-th port during the $\hat{n}$-th minute previously identified as under attack.

This procedure should be repeated for each $m$-th target port of $\boldsymbol{\hat{n}}$, in order to individually identify the network ports under attack during each $\hat{q}$-th time frame.

%The algorithm described above can be summarized as following:
%\begin{algorithm}[h!]
%  \caption{GETV Attack Detection}
%  \label{alg:alg01}
%  \begin{algorithmic}[1]
%  \show\LOOP
%    \REQUIRE Network Traffic Log
%    \ENSURE GETV
%    \STATE $\textbf{X}$ = matrix of ports and its occurrences per time \emph{\COMMENT{Data Pre-Processing}}
%    \STATE $Q$ = number of time frames
%    \ALOOP {$q=1$ util $q=Q$} 
%        \STATE $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M\times{N}}$
%        \IF {$isDosAttack$}
%            \STATE $\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}$
%            \STATE $\boldsymbol{C}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}$
%        \ENDIF
%        \IF {$isPortscanAttack$}
%			\STATE $\boldsymbol{y}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}$
%			\STATE $\boldsymbol{R}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}$
%		\ENDIF
%		\STATE $\boldsymbol{e}^{(q)} = \boldsymbol{e}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{e}^{(q)^{\rm T}}$
%       \STATE $[\lambda_{m}^{(1)},\lambda_{m}^{(2)},\lambda_{m}^{(3)}, ..., \lambda_{m}^{(q)}] = \boldsymbol{e}^{(q)}$                        
%        \STATE $\lambda_{m}^{(1)} > \lambda_{m}^{(2)} > \lambda_{m}^{(3)} > ... > \lambda_{m}^{(q)}$
%    \ENDALOOP
%	\STATE $\boldsymbol{E} =
%		\begin{bmatrix}
%			\lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
%			\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
%			\lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
%			\vdots & \vdots & \ddots & \vdots  \\
%			\lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
%		\end{bmatrix}$
%	\STATE $\boldsymbol{E}_{1}^{q} = \lambda_1^{(1)}, \lambda_1^{(2)}, \lambda_1^{(3)}, \cdots, \lambda_1^{(q)}$
%	\STATE $\boldsymbol{e}_{\rm max} = \boldsymbol{E}_{1}^{q}$
%  \end{algorithmic}
%\end{algorithm}

\section{Experiments and Results}
\label{sec:experimentalresults}

This section presents the performed experiments and the acquired results. First, in Section \ref{sec:AnalyzedScenario}, the scenario adopted in the experiments is summarized. Then, the Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames. In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection. The Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, and the Section \ref{sec:EigenSimilarityAnalysis} shows the obtained results of similarity analysis for detailed DoS and port scan attack identification.

\subsection{Analyzed Scenario}
\label{sec:AnalyzedScenario}

The experiment time is 120 minutes, separated into six time frames, with each time frame corresponding to twenty minutes. Therefore, as the time of each sampling period is one minute, then $N = 20$. For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, as well as a covariance $\boldsymbol{R}_{yy}^{(q)} \in \mathbb{R}^{17 \times 17}$ (calculated via (\ref{eq:eq03})) and a correlation matrix $\boldsymbol{R}_{zz}^{(q)} \in \mathbb{R}^{17x17}$, assuming that $q = 1, 2, 3, 4, 5$ and $6$.

The simulation started at 21:00h, the first time frame was from 21:00h until 21:20h ($q = 1$), the second was from 21:20h until 21:40h ($q = 2$), the third was from 21:40h to 22:00h ($q = 3$), the fourth was from 22:00h until 22:20h ($q = 4$), the fifth was from 22:20h until 22:40h ($q = 5$), and finally, the sixth was from 22:40h until 23.00h ($q = 6$). During the simulation, the victim made legitimate access, and the attacker performed the following attacks: at 21:54h ($q = 3$) was performed a port scan, at the interval ranging from 22:10h to 22:20h ($q = 4$) a synflood attack was simulated, and at the interval from 22:30h to 22:40h ($q = 5$) a fraggle attack was performed.

\subsection{Largest Eigenvalues Analysis}
\label{sec:largesteigenvaluesanalysis}

For the evaluation of MOS Schemes acuracy for DoS and port scan detection, our framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance or correlation matrix calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}. Through eigenvalue analysis of traffic with DoS or port scan attacks, it is possible to visualize a significant difference between the largest eigenvalues and the remain eigenvalues, which can indicate a relationship between an outlier and time frames under attack.

Figure \ref{fig:fig10} depics the eigenvalues calculated from covariance matrix of the network traffic used to evaluate the synflood attack identification.

\begin{figure}[h!]
     \includegraphics[height=5cm, width=7cm]{results/figures/eigenvalues_synflood.eps} 
     \caption{\csentence{Synflood Eigenvalues.} Eigenvalues of the covariance matrix (synflood).}
     \label{fig:fig10}
\end{figure}

In this figure its possible to see that the largest eigenvalue, which is related to the simulated synflood attack ($q = 4$), stands out significantly from the others eigenvalues.

Figure \ref{fig:fig11} illustrates represents the eigenvalues calculated from covariance matrix of the matrix used for fraggle attack detection. In this figure it is possible to see that the largest eigenvalue, which is related to this attack ($q = 5$), stands out significantly from the others eigenvalues, in accordance with the result shown in Figure \ref{fig:fig10} for the synflood attack analysis.

\begin{figure}[h!]
     \includegraphics[height=5cm, width=7cm]{results/figures/eigenvalues_fraggle.eps}
     \caption{\csentence{Fraggle Eigenvalues.} Eigenvalues of the covariance matrix (fraggle).}
     \label{fig:fig11}
\end{figure}

Figure \ref{fig:fig12} depics the eigenvalues calculated from correlation matrix of the network traffic matrix evaluated for port scan detection.

\begin{figure}[h!]
     \includegraphics[height=5cm, width=7cm]{results/figures/eigenvalues_portscan.eps}
     \caption{\csentence{Port Scan Eigenvalues.} Eigenvalues of the correlation matrix (port scan).}
     \label{fig:fig12}
\end{figure}

As analyzed for the synflood and fraggle attacks, it is possible to observe that the largest eigenvalue, related to this attack ($q = 3$), stands out significantly from the others eigenvalues.

Table \ref{tab:tab3} presents the values of the largest eigenvalues of each time frame $q$-th for port scan, synflood and fraggle detection.

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Largest Eigenvalue related to attacks detection}
  \label{tab:tab3}
  \begin{tabular}{ c c c c c }
	\toprule
	\multirow{3}{*}{\textbf{Time Frame} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
			\hhline{~----}
		&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
		&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
	\midrule
	1 &1887545 &1887545 &1887545 &2,0734 \\
	2 &2341327 &2341327 &2341327 &2,1451 \\
	3 &3213867 &3213867 &3213867 &10,0718 \\
	4 &133238294 &133238294 &731229 &2,1620 \\
	5 &92384021611 &6367983 &92384021611 &2,4253 \\
	6 &708335 &708335 &708335 &1,7948 \\
    \bottomrule
  \end{tabular}
\end{table}

In Table \ref{tab:tab3}, note the significant variation of the eigenvalues associated with attacks, in comparison to the others. At $q = 4$, where the synflood attack occurred, the maximum eigenvalue obtained was approximately 21 times larger than the second one. At $q = 5$, where the fraggle attack occurred, the maximum eigenvalue obtained was about 29,000 times larger than the second one. At $q = 3$, where the port scan attack occurred, the maximum eigenvalue obtained was approximately 4 times larger than the second one. In the last case, for port scan attack detection, although the largest eigenvalue presented no too large variance to the second one, if compared to synflood or fraggle attacks, it clearly deviates from the remaining largest eigenvalues.

These results highlight that all $q$-th time frames, where a network attack was simulated, presented high significant variance between the largest eigenvalue and the remaining eigenvalues, obtained from covariance matrix, for DoS detection, or from correlation matrix, for port scan detection. Therefore, we propose to apply the vector of the largest eigenvalues to MOS schemes in order to evaluate their accuracy for identification of time frames under attack, motivated by the fact that it is relevant to apply MOS schemes to automate the attack detection process, taking into account the characteristics of the evaluated eigenvalues.

\subsection{MOS Schemes Evaluation}
\label{sec:MOSSchemesEvaluation}

In \cite{tenorio2013greatest}, we evaluated the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE MOS schemes \cite{da2009comparison,tenorio2013greatest} for synflood and port scan attack detection. In this work we extended that evaluation for fraggle attack detection, applying the same schemes to fraggle attack detection over the traffic presented in Section \ref{sec:datamodel}, as results shown on following Table \ref{tab:tab4}.

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{MOS schemes applied to port scan and DoS detection}
  \label{tab:tab4}
  \begin{tabular}{ c c c c c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{Real value (d)}}\\ 
			\hhline{~------~}
		&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
	\midrule
	Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
	Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
	\midrule
	Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
	Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
	\midrule
	Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
	Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
	\midrule
	Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
	Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
    \bottomrule
  \end{tabular}
\end{table}

It was expected 1 as the model order value $\hat{d}$ when there was only one attack, values greater than 1, returned by one MOS scheme, indicates that there is more than one attack. An example of this could be seen for attack detection via EFT for traffic containing synflood and fraggle attacks, showing $\hat{d} = 2$, which indicates the presence of two attacks, as expected by the $d$ real values of Table \ref{tab:tab4}.

In Table \ref{tab:tab4}, two MOS schemes outperforms from the others, EDC and EFT. Efficient Detection Criterion (EDC) and Exponential Fitting Test (EFT) are the most effective schemes, correctly estimating the number of attacks in comparison to the expected values for effetctive attack detection, as defined by the column of real values in Table \ref{tab:tab4}. The AIC and MDL schemes are satisfactory only for port scan detection, however SURE and RADOI schemes did not show effective results for port scan or DoS detection.

Although EDC and EFT presented the same accuracy on our evaluation, the EDC scheme requires less processing time than EFT, which is an important criteria to select EDC as the MOS scheme for DoS and port scan detection on our remain experiments.

According to Table \ref{tab:tab4}, EDC and EFT estimated correctly the number of attacks of a time frame vector, indicating that ocurred $\hat{d}$ network attacks, but not providing additional details, what highlights the necessity of complementary approaches in order to estimate the time and ports under attack. Hence, we propose apply eigen analysis to estimate the $q$-th time frames under attack and eigen similarity analysis to estimate the minutes and ports under attack.

\subsection{Eigenvalue Analysis}
\label{sec:EigenvalueAnalysis}

According to the results presented in Section \ref{sec:largesteigenvaluesanalysis}, the largest eigenvalue stands out significantly from the others eigenvalues of an evaluated $q$-th time frame. This behavior can also be observed in the largest eigenvalues analysis, according to results presented in Table \ref{tab:tab3}, where it is possible to observe that the $\hat{d}$ largest eigen values of the time frames under attacks stand out significantly from the others largest eigenvalues.

Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respectives $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:alg01}.

\subsection{Eigen Similarity Analysis}
\label{sec:EigenSimilarityAnalysis}

This paper proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$. Hence, the proposed framework was applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.

\subsubsection{Time Analysis}
\label{sec:TimeAnalysis}

Three approaches were evaluated for eigen similarity analysis: incremental, individual and incremental individualized approaches. For the incremental individualized approach, each minute was incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack. Subsequently, $\boldsymbol{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated. For the incremental approach, each $n$-th minute must be incrementally appended into $\boldsymbol{X}^{(q)}$, for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual time similarity analyis. For the individual approach, each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, without incremental append, but doing individual appended into $\boldsymbol{X}^{(q)}$ for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual similarity analyis.

The Table \ref{tab:tab5} presents the results of the evaluation of three approaches for similarity analysis of eigenvectors for port scan detection.

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Port Scan Detection}
  \label{tab:tab5}
  \begin{tabular}{ c c c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~~---~}
			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
	\midrule
	3 &1 &0.9946 &0.9946 &0.9946 &no \\
	3 &2 &0.9934 &0.9934 &0.9999 &no \\
	3 &3 &0.9912 &0.9912 &0.9999 &no \\
	3 &4 &0.9888 &0.9888 &0.9999 &no \\
	3 &5 &0.9856 &0.9856 &0.9998 &no \\
	3 &6 &0.9840 &0.9840 &0.9999 &no \\
	3 &7 &0.9824 &0.9824 &1.0000 &no \\
	3 &8 &0.9794 &0.9794 &0.9999 &no \\
	3 &9 &0.9673 &0.9673 &0.9926 &no \\
	3 &10 &0.9674 &0.9674 &0.9997 &no \\
	3 &11 &0.9733 &0.9733 &0.9993 &no \\
	3 &12 &0.9702 &0.9702 &0.9993 &no \\
	3 &13 &0.9677 &0.9677 &0.9999 &no \\
	3 &14 &0.9646 &0.9646 &0.9998 &no \\
	3 &15 &0.0216 &0.0216 &0.0276 &yes \\
	3 &16 &0.9621 &0.0209 &1.0000 &no \\
	3 &17 &0.9611 &0.0199 &0.9998 &no \\
	3 &18 &0.9612 &0.0191 &0.9999 &no \\
	3 &19 &0.9613 &0.0186 &0.9998 &no \\
	3 &20 &0.9638 &0.0190 &1.0000 &no \\
    \bottomrule
  \end{tabular}
\end{table}

The Table \ref{tab:tab5} shows the evaluation of the time frame $q=3$, when the port scan attack was simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis. According to the presented results, it is possible to observe the high similarity between network traffic without attack, which was larger than 0.9610 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with the simulated port scan attack ($n=15$), which was lower than 0.0276 for all evaluated approaches.

Comparing the approaches for similarity analysis, it is possible to observe that all approaches highlight the low similarity when evaluated the traffic under attack. However, the incremental approach figured out low similarity for times without attack, where $n=16, 17, 18, 19, 20$, what indicates that the incremental approach can produce false posive results. This behaviour occurs because the incremental approaches appends all selected traffic into the reference traffic for comparison against the original reference traffic, what makes more evident the first lack of similarity but reduces the changing detection capability after an attack detection.

The Table \ref{tab:tab6} presents the results of the evaluation of the similarity analysis of eigenvectors for synflood detection. It shows the evaluation of the time frame $q=4$, when the synflood attack is simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis. According to the  results, it is possible to observe the high similarity between network traffic without attack, which is larger than 0.9907 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with synflood attack (between $n=11$ and $n=20$), which is lower than 0.1244 for all evaluated approaches.

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Synflood Detection}
  \label{tab:tab6}
  \begin{tabular}{ c c c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~~---~}
			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
	\midrule
	4 &1 &1.0000 &1.0000 &1.0000 &no \\
	4 &2 &0.9999 &0.9999 &1.0000 &no \\
	4 &3 &0.9997 &0.9997 &0.9999 &no \\
	4 &4 &0.9998 &0.9998 &1.0000 &no \\
	4 &5 &0.9965 &0.9965 &0.9908 &no \\
	4 &6 &0.9975 &0.9975 &1.0000 &no \\
	4 &7 &0.9977 &0.9977 &1.0000 &no \\
	4 &8 &0.9980 &0.9980 &1.0000 &no \\
	4 &9 &0.9987 &0.9987 &0.9999 &no \\
	4 &10 &0.9991 &0.9991 &1.0000 &no \\
	4 &11 &0.0085 &0.0085 &0.0284 &yes \\
	4 &12 &0.0162 &0.0120 &0.0343 &yes \\
	4 &13 &0.0248 &0.0158 &0.0427 &yes \\
	4 &14 &0.1243 &0.0185 &0.1041 &yes \\
	4 &15 &0.0082 &0.0162 &0.0103 &yes \\
	4 &16 &0.0404 &0.0070 &0.0580 &yes \\
	4 &17 &0.0397 &0.0007 &0.0573 &yes \\
	4 &18 &0.0408 &0.0042 &0.0584 &yes \\
	4 &19 &0.0408 &0.0079 &0.0584 &yes \\
	4 &20 &0.0477 &0.0092 &0.0757 &yes \\
    \bottomrule
  \end{tabular}
\end{table}

The incremental approach produces better results if compared with other evaluated approaches, with lower values and maximum of 0.0185 for times under attack, but this approach presents change detection limitation after the first outlier of similarity, as shown in Table \ref{tab:tab5} for port scan detection.

Comparing the incremental individualized and the individual approaches for eigen similarity analysis, it is possible to observe that the incremental individualized approach obtain lowest values for almost all cases, except for the time $n=14$, where incremental individualized approach identified a larger similarity than the individual approach. The incremental individualized appends information about each evaluated traffic, therefore it incorporates traffic behaviours that can reduce the outlier capability detection, as ocurred for the time $n=14$.

The Table \ref{tab:tab7} presents the results of the eigen similarity analysis evaluation for fraggle detection.

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Fraggle Detection}
  \label{tab:tab7}
  \begin{tabular}{ c c c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~~---~}
			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
	\midrule
	5 &1 &1.0000 &1.0000 &1.0000 &no \\
	5 &2 &0.9999 &0.9999 &1.0000 &no \\
	5 &3 &1.0000 &1.0000 &1.0000 &no \\
	5 &4 &0.9999 &0.9999 &1.0000 &no \\
	5 &5 &0.9993 &0.9993 &0.9997 &no \\
	5 &6 &0.9993 &0.9993 &0.9997 &no \\
	5 &7 &0.9994 &0.9994 &1.0000 &no \\
	5 &8 &0.9995 &0.9995 &1.0000 &no \\
	5 &9 &0.9995 &0.9995 &1.0000 &no \\
	5 &10 &0.9995 &0.9995 &1.0000 &no \\
	5 &11 &0.0031 &0.0031 &0.0021 &yes \\
	5 &12 &0.0019 &0.0025 &0.0009 &yes \\
	5 &13 &0.0030 &0.0026 &0.0020 &yes \\
	5 &14 &0.0030 &0.0027 &0.0020 &yes \\
	5 &15 &0.0030 &0.0028 &0.0020 &yes \\
	5 &16 &0.0012 &0.0025 &0.0002 &yes \\
	5 &17 &0.0030 &0.0026 &0.0020 &yes \\
	5 &18 &0.0030 &0.0026 &0.0020 &yes \\
	5 &19 &0.0030 &0.0027 &0.0020 &yes \\
	5 &20 &0.0069 &0.0023 &0.0083 &yes \\
    \bottomrule
  \end{tabular}
\end{table}

For fraggle attack detection, the lack of similarity between legitimate and malicious traffic was more evident than for the evaluation of synflood and port scan detecion. This becaviour can be explained by the number of packets generated through the fraggle attack simulation, that was significative larger than the number of packets generated during the synflood simulation. Considering the three approaches, the largest value for times under attack was 0.0083, while the shortest value for times without attacks was 0.9993.

Therefore, considering the evaluation for port scan, synflood and fraggle detection, the incremental approach can produce false positive results, while the individual and incremental individualized approaches produce quite similar results, even though the individual approach be more simple and require less memory and processing time.

These results highlight the capabilitiy of change detection based on similarity between legitimate and malicious traffic from DoS or port scan attacks, endorsing the effectiveness and safety for adoption of threshold for attack detection through eigen similarity analysis.

\subsubsection{Port Analysis}
\label{sec:PortAnalysis}

Given $\hat{N}$, which is the set of estimated $n$-th minutes under attack, it is possible to apply cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port, during a $n$-th minute. Therefore, the incremental individualized and individual approaches of eigen similarity analysis were evaluated, for detection of ports under DoS and port scan attacks, according to results presented in following tables. For this evaluation, the $v$ last most significant eigenvectors without attack was used as reference for similarity analysis against each target port $m$-th.

The Table \ref{tab:tab8} presents the results of the evaluation of eigen similarity analysis for detection of ports under port scan attack, showing only the time frame $q=3$ and minute $n=15$, due to the simulated port scan attack occurred only at this time, although the remain time frame has been completely evaluated and presented high similarity to the reference of traffic without network attack.

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n=15)}
  \label{tab:tab8}
  \begin{tabular}{ c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~--~}
			&\textbf{Incremental Individualized} &\textbf{Individual}\\
	\midrule
	80 &0.9999 &0.9999 &no \\
	443 &0.9999 &0.9999 &no \\
	53 &0.9999 &0.9999 &no \\
	21 &0.9999 &0.9997 &yes \\
	22 &0.0298 &0.9997 &yes \\
	23 &0.0298 &0.9997 &yes \\
	25 &0.0298 &0.9997 &yes \\
	110 &0.0298 &0.9997 &yes \\
	143 &0.0298 &0.9997 &yes \\
	161 &0.0298 &0.9997 &yes \\
	69 &0.0298 &0.9997 &yes \\
	123 &0.0298 &0.9997 &yes \\
	445 &0.0298 &0.9997 &yes \\
	600 &0.9999 &0.9999 &no \\
	19 &0.9999 &0.9999 &no \\
	67 &0.9999 &0.9999 &no \\
	68 &0.9999 &0.9999 &no \\
    \bottomrule
  \end{tabular}
\end{table}

The incremental individualized approach presented more sensibility to anomaly detection than the individual approach, the former produced the identification of a low similarity of 0.0298 for almost all ports under attack, unless the port 21, although our simulation has attacked this port. The individual approach was not able to identify low similarity for ports under attack, resunting in values of 0.9997 for ports with anomalous traffic and 0.9999 for ports without network attack.

For the evaluation of the proposed approaches for identification of ports under synflood and fraggle attack, all minutes of each time frame, in which one attack location was estimated, were analyzed. Even though, due to space limitations, only the results of the firt minute where a low similarity was idenfified will be shown such as where $n=11$. Nevertheless, the results obtained for the evaluation of traffic without attack presented high similarity to the reference traffic, with similarities close to 0.9999, and the evaluation of the other minutes under attack presented results quite similar to the resuts shown in the Tables \ref{tab:tab9} and \ref{tab:tab10}.

The Table \ref{tab:tab9} presents the results of the evaluation of eigen similarity analysis for detection of ports under synflood attack, showing only the time frame $q=4$ and minute $n=11$.

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
  \label{tab:tab9}
  \begin{tabular}{ c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~--~}
			&\textbf{Incremental Individualized} &\textbf{Individual}\\
	\midrule
	80 &1.0000 &1.0000 &no \\
	443 &1.0000 &1.0000 &no \\
	53 &1.0000 &1.0000 &no \\
	21 &1.0000 &1.0000 &nos \\
	22 &1.0000 &1.0000 &no \\
	23 &1.0000 &1.0000 &no \\
	25 &1.0000 &1.0000 &no \\
	110 &1.0000 &1.0000 &no \\
	143 &1.0000 &1.0000 &no \\
	161 &1.0000 &1.0000 &no \\
	69 &1.0000 &1.0000 &no \\
	123 &1.0000 &1.0000 &no \\
	445 &1.0000 &1.0000 &no \\
	600 &0.0077 &0.0427 &yes \\
	19 &1.0000 &1.0000 &no \\
	67 &1.0000 &1.0000 &no \\
	68 &1.0000 &1.0000 &no \\
    \bottomrule
  \end{tabular}
\end{table}

According to results presented in Table \ref{tab:tab9}, both approches identifies low similarity for the traffic of port 600, which is the target port of our simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.

The Table \ref{tab:tab10} presents the results of the evaluation of eigen similarity analysis for detection of ports under fraggle attack, showing only the time frame $q=5$ and minute $n=11$.

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Detection of Ports Under Fraggle Attack (q=5 and t=11)}
  \label{tab:tab10}
  \begin{tabular}{ c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~--~}
			&\textbf{Incremental Individualized} &\textbf{Individual}\\
	\midrule
	80 &1.0000 &1.0000 &no \\
	443 &1.0000 &1.0000 &no \\
	53 &1.0000 &1.0000 &no \\
	21 &1.0000 &1.0000 &no \\
	22 &1.0000 &1.0000 &no \\
	23 &1.0000 &1.0000 &no \\
	25 &1.0000 &1.0000 &no \\
	110 &1.0000 &1.0000 &no \\
	143 &1.0000 &1.0000 &no \\
	161 &1.0000 &1.0000 &no \\
	69 &1.0000 &1.0000 &no \\
	123 &1.0000 &1.0000 &no \\
	445 &1.0000 &1.0000 &no \\
	600 &1.0000 &1.0000 &no \\
	19 &0.0031 &0.0004 &yes \\
	67 &1.0000 &1.0000 &no \\
	68 &1.0000 &1.0000 &no \\
    \bottomrule
  \end{tabular}
\end{table}

The results for the avaluation of ports under fraggle attack, shown in Table \ref{tab:tab10}, were similar to the results obtained for synflood analysis, with the identification of low similarity for traffic of the port under attack. Nevertheless, for fraggle analysis, the individual approach identified the lowest similarity, that is 0.0004 while the incremental individualized approach obtained a similarity of 0.0031.

The incremental individualized approach was able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches presented false positives or low sensibility to eigen similarity analyis for network attack detection. This approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.

According to the shown significant lack of similarity between legitimate and malicious traffic, it is possible to adopt safe thresholds for DoS and port scan detection through eigen similarity analysis.

\section{Conclusion and Future Works}
\label{sec:conclusionandfutureworks}

This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and DoS attacks, but still requiring more information for detailed attack detection. Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.

The proposed scheme is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and  blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack.

The incremental individualized approach of eigen similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to eigen similarity analyis for network attack detection. Therefore, the incremental individualized approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.

According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for DoS and port scan detection through eigen similarity analysis.

Future research is directed to the fully automated blind attack detection, extending the evaluation to performance and accuracy analyis of our approach to well-known datasets of network attack detection.

\appendix

\section{Model Order Selection (MOS)}
\label{sec:mos}
The model order selection is a key point in many digital signal processing applications, including radar, sonar, communications, channel modeling, medical imaging, among others. MOS allows analysis of reduced data set, through separating noise components of the main components, for example. Moreover, the model order is crucial for many parameter estimation techniques \cite{da2009comparison}, since the amount of parameters to be estimated depends on the model order.

The model selection procedure chooses the ``best'' model of a finite set of models, according to some criterias \cite{rajan1997model}. Therefore, given some data set, it is chosen a model which was evaluated as the best model to describe the specified data set.

The state of the art regarding estimation techniques of model order based on eigenvalues includes: Akaike's Information Theoretic Criterion - AIC \cite{akaike1974new,wax1985detection}; Minimum Description Length - MDL \cite{barron1998minimum,wax1985detection}; Efficient Detection Criterion - EDC \cite{zhao1986detection}; Stein's Unbiased Risk Estimator - SURE \cite{ulfarsson2008rank}; RADOI \cite{radoi2004new} and Exponential Fitting Test - EFT \cite{grouffaud1996some,quinlan2006model,david2011blind}.

In AIC, MDL and EDC techniques, the information criterion is a function of the geometric mean \emph{g(k)} and the arithmetic mean \emph{a(k)} relating to smaller \emph{k} eigenvalues, where \emph{k} is a candidate value for the model order \emph{d} \cite{da2009comparison}.

Basically, the difference between the AIC, MDL and EDC schemes is the penalty function \emph{p(k, N, α)}, so these techniques can be written in general as \cite{da2009comparison}:

\begin{equation}\label{eq:eq1}
\hat{d} = \argmin\limits_k \hspace{1 mm} J(k), 
\end{equation}

where

\begin{equation}\label{eq:eq2}
J(k) = -N(\alpha - k) \hspace{1 mm} {\rm log} \hspace{1 mm} (g(k)/a(k)) + p(k,N,\alpha),
\end{equation}

where $\hat{d}$ is an estimate \emph{d} of the model order, \emph{N} is the number of samples, \emph{α = M} and means the number of variables of the problem, and 0 ≤ \emph{k} ≤ min[\emph{M}, \emph{N}]. Penalty functions for AIC, MDL and EDC are given by the Table \ref{tab:tab2}.
%% ???

\begin{table}[h!]
  \centering
  \caption{Penalty functions for the schemes AIC, MDL and EDC}
  \label{tab:tab2}
  \begin{tabular}{*2c}
	\toprule
	\textbf{Scheme} &  \textbf{Penalty function} \\
	\textbf{} &  $p(k,N,\alpha)$ \\
	\midrule
    AIC	& $k(2\alpha - k)$ \\
    MDL	& $0.5k(2\alpha - k) \hspace{1 mm} {\rm log} (N)$ \\
    EDC	& $0.5k(2\alpha - k)\sqrt{N\hspace{1 mm}\rm ln(\rm ln N)}$ \\
    \bottomrule
  \end{tabular}
\end{table}

The Exponential Fitting Test (EFT) can effectively be used in cases where the number of samples \emph{N} is small. This technique is based on observations of data contaminated only with white noise, where the profile of eigenvalues can be approximated by an exponential decaying \cite{grouffaud1996some}.

Given $\lambda_i$ be the i-th eigenvalue, the exponential model can be expressed by:
\begin{equation}\label{eq:eq3}
E\{\lambda_i\} = E\{\lambda_1\} \cdot q(\alpha,\beta)^{i-1},
\end{equation}

where $E\{\cdot\} $ is the expectation operator, and it is considered that the eigenvalues are ordered in the that $\lambda_1$ represents the largest eigenvalue. The term \emph{q(α, β)} is defined as:

\begin{equation}\label{eq:eq4}
q(\alpha,\beta) = \exp\left\{-\sqrt{\frac{30}{\alpha^2 + 2} - \sqrt{\frac{900}{(\alpha^2 + 2)^2} - \frac{720\alpha}{\beta(\alpha^4 + \alpha^2 - 2)}}} \right\},
\end{equation}

where $0 < q(\alpha,\beta) < 1$. According to \cite{quinlan2006model}, if $M \leq N$, then $\beta = N$.

%Figure \ref{fig:fig1} shows a typical profile of eigenvalues​​. The last P – 1 eigenvalues are used to estimate the (M - P)-th eigenvalue, denoted by the yellow rectangle. The EFT method considers the discrepancy between the actual value and the estimated value obtained \cite{da2007enhanced}.

%\begin{figure}[h!]
 %    \centering 
     %\includegraphics[height=5cm, width=6cm]{results/figures/fig01.png}
  %   \caption{\csentence{EFT Example.} Example of application of EFT \cite{da2007enhanced}.}
   %  \label{fig:fig1}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

%\section*{Author's contributions}
%    Text for this section \ldots

\section*{Acknowledgements}
The authors thank the Brazilian Ministry of Planning, Budget and Management for the support during the development of this work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{references} 		% Style BST file
\bibliography{references}	      	% Bibliography file (usually '*.bib' )

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

%\section*{Figures}
%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig01.png}
%     \caption{\csentence{EFT Example.} Example of application of EFT \cite{da2007enhanced}.}
%     \label{fig:fig1}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig02.png}
%     \caption{\csentence{Traffic Matrix.} Obtaining the traffic matrix $\boldsymbol{X}^{(q)}$.}
%     \label{fig:fig2}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig03.png}
%     \caption{\csentence{Legitimate Traffic.} The legitimate traffic.}
%     \label{fig:fig3}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig04.png}
%     \caption{\csentence{Noise Traffic.} The noise traffic.}
%     \label{fig:fig4}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig05.png}
%     \caption{\csentence{Synflood Traffic.} The traffic characterized by synflood.}
%     \label{fig:fig5}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig06.png}
%     \caption{\csentence{Fraggle Traffic.} The traffic characterized by fraggle.}
%     \label{fig:fig6}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig07.png}
%     \caption{\csentence{Port scan Traffic.} The traffic characterized by port scan.}
%     \label{fig:fig7}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig08.png}
%     \caption{\csentence{GETV.} Proposed GETV attack detection approach.}
%     \label{fig:fig8}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig09.png}
%     \caption{\csentence{Scenario.} Analyzed Scenario.}
%     \label{fig:fig9}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig10.png}
%     \caption{\csentence{Synflood Eigenvalues.} Eigenvalues of the covariance matrix (synflood).}
%     \label{fig:fig10}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig11.png}
%     \caption{\csentence{Fraggle Eigenvalues.} Eigenvalues of the covariance matrix (fraggle).}
%     \label{fig:fig11}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig12.png}
%     \caption{\csentence{Port scan Eigenvalues.} Eigenvalues of the covariance matrix (port scan).}
%     \label{fig:fig12}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig13.png}
%     \caption{\csentence{Synflood Components.} The two first principal components (synflood).}
%     \label{fig:fig13}
%\end{figure}

%\begin{figure}[h!]
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig14.png}
%     \caption{\csentence{Fraggle Components.} The two first principal components (fraggle).}
%     \label{fig:fig14}
%\end{figure}

%\begin{figure}[h!]	
%     \centering 
%     %\includegraphics[height=10cm, width=10cm]{figures/fig15.png}
%     \caption{\csentence{Port scan Components.} The two first principal components (port scan).}
%     \label{fig:fig15}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
%\section*{Tables}
%\begin{table}[h!]
%  \centering
%  \footnotesize
%  \caption{Comparison between malicious traffic detection schemes of related works and of this paper}
%  \label{tab:tab1}
%  \begin{tabular}{ c c c c c c }
%	\toprule
%	\multirow{2}{*}{\textbf{Related Works}} &\multicolumn{5}{c }{\textbf{Techniques}}\\ 
%		&\textbf{Data Mining}	 &\textbf{Regular analysis of files}	 &\textbf{PCA}	 &\textbf{MOS}	 &\textbf{GETV}\\
%	\midrule
%	(He et al, 2008)		&x	&-	&-	&-	&- \\
%	(Raynal et al, 2004) 	&-	&x	&-	&-	&- \\
%	(Almotairi et al, 2009)	&-	&-	&x	&-	&- \\
%	(da Costa et al, 2012) 	&-	&-	&x	&x	&- \\
%	(Proposed) 				& -	& -	&x	&x	&x \\
%   \bottomrule
%  \end{tabular}
%\end{table}
%\begin{table}[h!]
%  \centering
%  \caption{Penalty functions for the schemes AIC, MDL and EDC}
%  \label{tab:tab2}
%  \begin{tabular}{*2c}
%	\toprule
%	\textbf{Scheme} &  \textbf{Penalty function} \\
%	\textbf{} &  $p(k,N,\alpha)$ \\
%	\midrule
%    AIC	& $k(2\alpha - k)$ \\
%    MDL	& $0.5k(2\alpha - k) \hspace{1 mm} log(N)$ \\
%    EDC	& $0.5k(2\alpha - k)\sqrt{N\hspace{1 mm}ln(lnN)}$ \\
%    \bottomrule
%  \end{tabular}
%\end{table}
%\begin{table}[h!]
%  \centering
%  \footnotesize
%  \caption{Greatest Eigenvalue related to attacks detection}
%  \label{tab:tab3}
%  \begin{tabular}{ c c c c c }
%	\toprule
%	\multirow{3}{*}{\textbf{Time Period} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
%			\hhline{~----}
%		&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
%		&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
%	\midrule
%	1 &1887545 &1887545 &1887545 &2,0734 \\
%	2 &2341327 &2341327 &2341327 &2,1451 \\
%	3 &3213867 &3213867 &3213867 &10,0718 \\
%	4 &133238294 &133238294 &731229 &2,1620 \\
%	5 &92384021611 &6367983 &92384021611 &2,4253 \\
%	6 &708335 &708335 &708335 &1,7948 \\
%    \bottomrule
%  \end{tabular}
%\end{table}
%\begin{table}[h!]
%  \centering
%  \footnotesize
%  \caption{MOS schemes applied to the GETV}
%  \label{tab:tab4}
%  \begin{tabular}{ c c c c c c c c }
%	\toprule
%	\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} %&{\textbf{Real value (d)}}\\ 
%			\hhline{~------~}
%		&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
%	\midrule
%	Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
%	Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
%	\midrule
%	Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
%	Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
%	\midrule
%	Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
%	Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
%	\midrule
%	Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
%	Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
%    \bottomrule
%  \end{tabular}
%\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section*{Additional Files}
%  \subsection*{Additional file 1 --- Sample additional file title}
%    Additional file descriptions text (including details of how to view the file, if it is in a non-standard format or the file extension). This might refer to a multi-page table or a figure.

%  \subsection*{Additional file 2 --- Sample additional file title}
%    Additional file descriptions text.

\end{backmatter}
\end{document}

### 2016-04-13 docs/papers/2015eurasip/code/eigen_analysis/html/eigencorrelation.tex
adding papers and their assets

% This LaTeX was auto-generated from MATLAB code.
% To make changes, update the MATLAB code and republish this document.

\documentclass{article}
\usepackage{graphicx}
\usepackage{color}

\sloppy
\definecolor{lightgray}{gray}{0.5}
\setlength{\parindent}{0pt}

\begin{document}

\section*{Input}

\begin{par}
Matrix
\end{par} \vspace{1em}

\subsection*{Contents}

\begin{itemize}
\setlength{\itemsep}{-1ex}
   \item Output
\end{itemize}

\subsection*{Output}

\begin{par}
(R) correlation matrix (E) eigenvalues diagonal matrix (V) eigenvectors (M) largest eigenvalue
\end{par} \vspace{1em}
\begin{verbatim}
function [R,E,V,M] = eigencorrelation(X)
    numLines = size(X,1);
    numColumns = size(X,2);
    Y = zeros(numLines,numColumns);

    for i = 1:numLines
        stdv = std(X(i,:),1);
        if stdv > 0
            Y(i,:) = (X(i,:) - mean(X(i,:))) / stdv;	% standardization, have zero mean and unit variance
        end
    end

    R = 1/numColumns*Y*Y';								% estimation of correlation matrix (R)
    [V,E] = eig(R);										% eigenvectors (V) and eigenvalues (E) of correlation matrix
    M(1) = max(diag(E));                                % largest eigenvalue (M)

    Ed  = diag(E);
    for i = 1:size(Ed,1);
        if M(1) == Ed(i)
            M(2) = i;                                   % get the position of largest eigenvalue
        end
    end
\end{verbatim}

        \color{lightgray} \begin{verbatim}Error using eigencorrelation (line 10)
Not enough input arguments.
\end{verbatim} \color{black}

\end{document}

### 2016-04-13 docs/papers/2016JNCA/2016JNCA.tex
adding papers and their assets

--- 2016-04-13 docs/papers/2015eurasip/2014eurasip (copy).tex
@@
-%% BioMed_Central_Tex_Template_v1.06
-%%                                      %
-%  bmc_article.tex            ver: 1.06 %
-%                                       %
-
-%%IMPORTANT: do not delete the first line of this template
-%%It must be present to enable the BMC Submission system to
-%%recognise this template!!
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                     %%
-%%  LaTeX template for BioMed Central  %%
-%%     journal article submissions     %%
-%%                                     %%
-%%          <8 June 2012>              %%
-%%                                     %%
-%%                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                                                 %%
-%% For instructions on how to fill out this Tex template           %%
-%% document please refer to Readme.html and the instructions for   %%
-%% authors page on the biomed central website                      %%
-%% http://www.biomedcentral.com/info/authors/                      %%
-%%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
-%% Submit your LaTeX manuscript as one .tex document.              %%
-%%                                                                 %%
-%% All additional figures and files should be attached             %%
-%% separately and not embedded in the \TeX\ document itself.       %%
-%%                                                                 %%
-%% BioMed Central currently use the MikTex distribution of         %%
-%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
-%% http://www.miktex.org                                           %%
-%%                                                                 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% additional documentclass options:
-%  [doublespacing]
-%  [linenumbers]   - put the line numbers on margins
-
-%%% loading packages, author definitions
-
-%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
-\documentclass{bmcart}
-\usepackage[titletoc,title]{appendix}
-%%% Load packages
-\usepackage{amsthm,amsmath}
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
-%\RequirePackage{natbib}
-%\RequirePackage{hyperref}
-%\usepackage[utf8]{inputenc} %unicode support
+\documentclass[review]{elsarticle}
+
+\usepackage{lineno,hyperref}
 \usepackage[mathletters]{ucs}
 \usepackage[utf8x]{inputenc}
-%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
-%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
-
 \usepackage{algorithm,algorithmic}% http://ctan.org/pkg/algorithms
-
 \usepackage{natbib}
-
 \usepackage{supertabular}
@@
 \usepackage{mathtools}
-\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
-\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
 \usepackage{listings}
@@
 \usepackage{lipsum}
+\usepackage{amsthm,amsmath}
+\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
+\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
 \newcommand*\justify{%
@@
 }
-
-% appendix
-\usepackage{hyperref}
-\usepackage{cleveref}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                             %%
-%%  If you wish to display your graphics for   %%
-%%  your own use using includegraphic or       %%
-%%  includegraphics, then comment out the      %%
-%%  following two lines of code.               %%
-%%  NB: These line *must* be included when     %%
-%%  submitting to BMC.                         %%
-%%  All figure files must be submitted as      %%
-%%  separate graphics through the BMC          %%
-%%  submission process, not included in the    %%
-%%  submitted article.                         %%
-%%                                             %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\def\includegraphic{}
-%\def%\includegraphics{}
 
@@
 
-%%% Put your definitions there:
-\startlocaldefs
-\endlocaldefs
-
-%%% Begin ...
+% appendix
+\usepackage[titletoc,title]{appendix}
+\usepackage{hyperref}
+\usepackage{cleveref}
+
+\modulolinenumbers[5]
+
+\journal{ Journal of Network and Computer Applications}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+%% Elsevier bibliography styles
+%%%%%%%%%%%%%%%%%%%%%%%
+%% To change the style, put a % in front of the second line of the current style and
+%% remove the % from the second line of the style you would like to use.
+%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Numbered
+%\bibliographystyle{model1-num-names}
+
+%% Numbered without titles
+%\bibliographystyle{model1a-num-names}
+
+%% Harvard
+%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+
+%% Vancouver numbered
+%\usepackage{numcompress}\bibliographystyle{model3-num-names}
+
+%% Vancouver name/year
+%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
+
+%% APA style
+%\bibliographystyle{model5-names}\biboptions{authoryear}
+
+%% AMA style
+%\usepackage{numcompress}\bibliographystyle{model6-num-names}
+
+%% `Elsevier LaTeX' style
+\bibliographystyle{elsarticle-num}
+%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{document}
 
-%%% Start of article front matter
 \begin{frontmatter}
 
-\begin{fmbox}
-\dochead{Research}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the title of your article here     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \title{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors here                   %%
-%%                                          %%
-%% Specify information, if available,       %%
-%% in the form:                             %%
-%%   <key>={<id1>,<id2>}                    %%
-%%   <key>=                                 %%
-%% Comment or delete the keys which are     %%
-%% not used. Repeat \author command as much %%
-%% as required.                             %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\author[
-	addressref={address1},                   % id's of addresses, e.g. {aff1,aff2}
-	corref={address1},                       % id of corresponding address, if any	
-	%noteref={n1},                        	% id's of article notes, if any
-	email={tpbvieira@gmail.com} 				% email address
-]{\inits{TPBV}\fnm{Thiago P. B.} \snm{Vieira}}
-\author[
-   	addressref={address1},
-   	email={???????????????}
-]{\inits{DFT}\fnm{Danilo F.} \snm{Tenório}}
-\author[
-   	addressref={address1},
-   	email={???????????????}
-]{\inits{JPCLC}\fnm{João Paulo C. L.} \snm{da Costa}}
-\author[
-   	addressref={address1,address2},
-   	email={???????????????}
-]{\inits{EPF}\fnm{Edison P.} \snm{de Freitas}}
-%\author[
-%   	addressref={address3},
-%   	email={???????????????}
-%]
-%{\inits{CHC}\fnm{\\Carlos H.} \snm{Cardonha}}
-\author[
-   	addressref={address1},
-   	email={???????????????}
-]{\inits{RTSJ}\fnm{\\Rafael T.} \snm{de Sousa Júnior}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors' addresses here        %%
-%%                                          %%
-%% Repeat \address commands as much as      %%
-%% required.                                %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\address[id=address1]{                           								% unique id
-	\orgname{Department of Electrical Engineering, University of Brasilia (UnB)},	% 	university, etc
-%	\street{},                     												%
-	\postcode{70910-900}                                							% post or zip code
-	\city{Brasília-DF},                              							% city
-	\cny{Brazil}                                    								% country  
-}
-\address[id=address2]{%
-	\orgname{Graduate Program in Electrical Engineering, Federal University of Rio Grande do Sul (UFRGS)},
-%	\street{},
-	\postcode{90035-190}
-	\city{Porto Alegre},
-	\cny{Brazil}
-}
-%\address[id=address3]{                           								% unique id
-%	\orgname{IBM Research},	% 	university, etc
-%	\street{},                     												%
-%	\postcode{}                                							% post or zip code
-%	\city{São Paulo},                              							% city
-%	\cny{Brazil}                                    								% country  
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter short notes here                   %%
-%%                                          %%
-%% Short notes will be after addresses      %%
-%% on first page.                           %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{artnotes}
-	%\note{Sample of title note}     	% note to the article
-	%\note[id=n1]{Equal contributor} 		% note, connected to author
-\end{artnotes}
-
-\end{fmbox}								% comment this for two column layout
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Abstract begins here                 %%
-%%                                          %%
-%% Please refer to the Instructions for     %%
-%% authors on http://www.biomedcentral.com  %%
-%% and include the section headings         %%
-%% accordingly for your article type.       %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{abstractbox}
-
-\begin{abstract} % abstract
-
+%% Group authors per affiliation:
+\author[unbaddress]{Thiago P. B. Vieira}
+\author[unbaddress]{Danilo F. Tenório}
+\author[unbaddress,Ilmenauaddress,Fraunhoferaddress]{João Paulo C. L. da Costa}
+\author[ufrgsaddress]{\\Edison P. de Freitas}
+\author[unbaddress]{Rafael T. de Sousa Júnior}
+
+\address[unbaddress]{Department of Electrical Engineering, University of Brasilia (UnB), 70910-900, Brasília-DF, Brazil}
+\address[Ilmenauaddress]{Institute for Information Technology, Ilmenau University of Technology, Ilmenau, Germany}
+\address[Fraunhoferaddress]{Fraunhofer Institute for Integrated Circuits IIS, Erlangen, Germany}
+\address[ufrgsaddress]{Graduate Program in Electrical Engineering, Federal University of Rio Grande do Sul (UFRGS), 90035-190, Porto Alegre, Brazil}
+
+\begin{abstract}
 Novel schemes for attack detection are crucial to identify adaptive malicious traffic coming from sources that are quickly mobilized by attackers in high throughput communication networks.
@@
 The experiments performed in a real network show that the proposed blind detection approach achieves satisfactory levels of accuracy in terms of timely detection and identification of TCP/UDP ports under attack.
-
 \end{abstract}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The keywords begin here                  %%
-%%                                          %%
-%% Put each keyword in separate \kwd{}.     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{keyword}
-	\kwd{Network Attack Detection}
-	\kwd{Model Order Selection}
-	\kwd{Eigen Analysis}
-	\kwd{Similarity Analysis}
+Network Attack Detection \sep Model Order Selection \sep Eigen Analysis \sep Similarity Analysis
 \end{keyword}
 
-% MSC classifications codes, if any
-%\begin{keyword}[class=AMS]
-%\kwd[Primary ]{}
-%\kwd{}
-%\kwd[; secondary ]{}
-%\end{keyword}
-
-\end{abstractbox}
-%
-%\end{fmbox}% uncomment this for twcolumn layout
-
 \end{frontmatter}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Main Body begins here                %%
-%%                                          %%
-%% Please refer to the instructions for     %%
-%% authors on:                              %%
-%% http://www.biomedcentral.com/info/authors%%
-%% and include the section headings         %%
-%% accordingly for your article type.       %%
-%%                                          %%
-%% See the Results and Discussion section   %%
-%% for details on how to create sub-sections%%
-%%                                          %%
-%% use \cite{...} to cite references        %%
-%%  \cite{koon} and                         %%
-%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
-%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
-% <put your article body there>
-
-%%%%%%%%%%%%%%%%
-%% Background %%
-%%
+\linenumbers
+
 \section{Introduction}
@@
 Section \ref{sec:prop_getv} describes our proposed framework for blind and automatic detection of malicious traffic.
-Section \ref{sec:experimentalresults} discusses the experimental validation and presents the corresponding experimental results.
+Section \ref{sec:experimentalresults} discusses the experimental validation and presents the experimental results.
 Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
-The Appendix \ref{sec:mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.
+The \ref{sec:mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.
 
@@
 Researchers have applied data mining techniques in log analysis \cite{ghourabi2010data} to improve intrusion detection performance.
-However, data mining techniques require prior collection of large data sets, which is a weakness of several schemes for online or low latency analysis \cite{Huang2009}.
-
-Regular file analysis \cite{raynal2004honeypot} consists of using traffic analysis for detecting known patterns that indicate the presence of specific attacks, applying statistical analysis to the study of collected traffic.
-An essential feature of this method is that it depends on prior knowledge of the details of the attacks to be identified, and also depends on previous log collection for applying traffic analysis and reducing false positives.
+However, data mining techniques used so far in network analysis require prior collection of large data sets, which is a limitation of several schemes for online analysis \cite{Huang2009}.
+
+Regular file analysis \cite{raynal2004honeypot} consists of traffic analysis for detecting known patterns that indicate the presence of attacks, applying statistical analysis to the study of collected traffic.
+An essential feature of this method is that it depends on prior knowledge of the details of the attacks to be identified, and also depends on previous log collection for traffic analysis and false positives reduction.
 
@@
 Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
-Zonglin \emph{et al} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work was not applied to port scan and denial of service (DoS) attack detection, simultaneously.
+Zonglin \emph{et al} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work was not applied to probing and denial of service (DoS) attack detection, simultaneously.
 
@@
 In contrast to Lee \emph{et al}, our framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
-Additionally, Lee \emph{et al} only evaluated their proposed scheme for covariance analysis, while we adopted covariance and correlation analysis for DoS and probing attacks, respectively.
+Additionally, Lee \emph{et al} only evaluated their proposed scheme for covariance analysis, while we adopted an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for DoS and probing attacks, respectively.
 
@@
 
-In this paper the scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
+In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
@@
 This section presents details of the simulated scenario along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic.
-The Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce DoS and probing attacks.
-The Subsection \ref{sec:ModelingData} presents how network traffic can be modelled as signal superposition, and the Subsection \ref{sec:AnalyzedScenario} details the network traffic of synflood, fraggle and port scan attacks.
+Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce DoS and probing attacks.
+Subsection \ref{sec:ModelingData} presents how network traffic can be modelled as signal superposition, and Subsection \ref{sec:AnalyzedScenario} details the traffic of synflood, fraggle and port scan attacks.
 
@@
      \includegraphics[height=5cm, width=7cm]{results/figures/fig09.png}
-     \caption{\csentence{Analyzed Scenario.} Scenario to reproduce legitimate traffic, noise, DoS and port scan.}
+     \caption{Scenario to reproduce legitimate traffic, noise, DoS and port scan.}
      \label{fig:fig1}
@@
 For malicious traffic, three types of networks attacks are evaluated: synflood, fraggle and port scan.
-These attacks are reproduced using well-known security tools, such as Nmap to port scan, Metaspoit to synflood and Hping to lead the fraggle attack.
+These attacks are reproduced using well-known security tools, such as Nmap\footnote{http://nmap.org} to port scan, Metasploit\footnote{http://www.metasploit.com} to synflood and Hping\footnote{http://hping.org} to lead the fraggle attack.
 
@@
 specific information such as origin IP or day and time of the attack are not considered.
-Therefore, from the whole log information, we just consider the timestamp (for sequencing), port type and port number.
+Therefore, from the entire log information, we just consider the timestamp (for sequencing), port type and port number.
 
@@
 
-Modeling the dataset as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{S}), noise (\textbf{N}) and malicious traffic (\textbf{A}), according to the following expression:
+By modeling the dataset as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{S}), noise (\textbf{N}) and malicious traffic (\textbf{A}), according to the following expression:
 
@@
 where $q$ represents the $q$-th time frame, which is a time grouping of network traffic.
-
 The matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns.
@@
 When a user accesses a web page, for example, there is the corresponding TCP/IP traffic to request the page, as well as there is the traffic required to domain name resolution.
-Figure \ref{fig:fig3} depicts an example of the legitimate traffic obtained during experiments, represented by traffic to the ports 67 and 68, which are related to the Bootstrap Protocol and to normal DHCP operations, respectively.
+Figure \ref{fig:fig3} depicts an example of the legitimate traffic obtained during experiments.
 
@@
      \includegraphics[height=4.5cm, width=9cm]{results/figures/fig03.png}
-     \caption{\csentence{Legitimate Traffic.} The traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
+     \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
      \label{fig:fig3}
@@
 
-All traffic, that is not directly associated with users' operations, but it is not a malicious traffic, is considered as noise $\boldsymbol{N}^{(q)}$.
+The traffic that is not associated with user's operations and with malicious traffic is modelled as noise $\boldsymbol{N}^{(q)}$.
 The automatic acquisition service of logical IP network address (DHCP) is an example of noise.
 Independently of any user operation, the machine receives an IP address, since it is configured to perform a DHCP address request.
-Figure \ref{fig:fig4} depicts an example of noise in a network traffic.
+Figure \ref{fig:fig4} depicts an example of noise in a network traffic, represented by traffic to ports 67 and 68.
 
@@
      \includegraphics[height=4.5cm, width=9cm]{results/figures/fig04.png}
-     \caption{\csentence{Noise Traffic.} Network traffic of user independent operations for network management.}
+     \caption{Network traffic of user independent operations for network management.}
      \label{fig:fig4}
@@
 The traffic coming from a malicious activity, such as a synflood or fraggle attack, is represented by the matrix $\boldsymbol{A}^{(q)}$.
-For this work we only consider the traffic from port scanning and flood attacks, which aims to cause denial of service.
+For this work we only consider the traffic from port scanning and flood attacks.
 
@@
 The first two attacks can be qualified as DoS attacks, while the last one can be qualified  as probing or port scanning attack.
-
-The TCP protocol is a connection-oriented protocol, then a virtual connection must be established between two computers for an end-to-end TCP communication.
-This virtual connection requires a ``handshake'', that occurs in three steps, known as three-way handshake.
-If a computer needs to communicate with another computer, the requester sends a packet communication synchronization (SYN) to a specific destination port, which is in listening state.
-If the destination is active, running and accepting requests, it responds to the requester with a SYN/ACK confirmation message.
-After receiving this message, the requester sends an ACK message to the destination and then the connection is established.
 
@@
      \includegraphics[height=4.5cm, width=9cm]{results/figures/fig05.png}
-     \caption{\csentence{Synflood Traffic.} A large quantity of SYN requests to a target, in order to cause a DoS.}
+     \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
      \label{fig:fig5}
@@
 
-With respect to the fraggle atack, large packets with ``UDP echo'' segments are sent to the broadcast address of a network.
+With respect to the fraggle atack, large packets with UDP echo segments are sent to the broadcast address of a network.
 Every packet is modified to have the source address of the victim, in order to implement the source address spoofing technique.
-Therefore, each host receives a huge amount of requests ``UDP echo'' and all of them replies to the IP address of the victim, causing a packet flooding aiming a DoS.
-This attack can affect the entire network, since all hosts receive several requests ``UDP echo'' and respond with the ICMP protocol, therefore each host acts as an ``amplifier'' of the attack.
-This last part of the fraggle attack is not taken into account in this work, because the victim receives ICMP (network layer) packets originated from the hosts that are attacked with flooding packet ``UDP echo''.
-This occurs due to the UDP be not able to know if the segment sent has reached its destination.
-
+Therefore, each host receives a huge amount of requests UDP echo and all of them replies to the IP address of the victim, causing a packet flooding aiming a DoS.
+This attack can affect the entire network, since all hosts receive several requests UDP echo and respond with the ICMP protocol, therefore each host acts as an amplifier of the attack.
+This last part of the fraggle attack is not taken into account in this work, because the victim receives ICMP (network layer) packets originated from the hosts that are attacked with flooding packet UDP echo.
 Figure \ref{fig:fig6} depicts an example of the fraggle attack in a real computer network.
+
+\begin{figure}[h!]
+     \centering 
+     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig06.png}
+     \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
+     \label{fig:fig6}
+\end{figure}
+
 More than 6,000,000 malicious packets can be counted in an interval of ten minutes, which can be considered an abnormal network traffic, especially due to the concentrated traffic in a short period of time and due to the similarity of the outstanding traffic.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig06.png}
-     \caption{\csentence{Fraggle Traffic.} Large amount of “UDP echo” requests and replies, causing packet flooding.}
-     \label{fig:fig6}
-\end{figure}
 
@@
 Therefore, a full connection or a complete three-way handshake is never established, which makes the detection of the attack sender more difficult, and requires approaches able to identify probing attacks without connection stablishment.
-The UDP scan technique sends UDP packets to the destination port, and if it responds with a ``ICMP port unreachable'' message, then it indicates that the scanned port is closed.
+The UDP scan technique sends UDP packets to the destination port, and if it responds with a \emph{ICMP port unreachable} message, then it indicates that the scanned port is closed.
 On the other hand, if a message is not received, then the port is considered as open.
@@
      \includegraphics[height=5cm, width=9cm]{results/figures/fig07.png}
-     \caption{\csentence{Port Scan Traffic.} Connection attempts in order to identify active ports.}
+     \caption{Connection attempts in order to identify active ports.}
      \label{fig:fig7}
@@
 This section describes the proposed technique to detect synflood, fraggle and port scan.
-In Subsection \ref{sec:prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvelue for each $q$-th time frame, in order to apply a selected MOS scheme for attack detection, which is described in Subsection \ref{sec:prop_MOSSchemes}.
+In Subsection \ref{sec:prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvelue for each $q$-th time frame.
+Next, in Subsection \ref{sec:prop_MOSSchemes}, we show how to apply the eiganvalues on the MOS scheme in order to detect the attack.
 In Subsection \ref{sec:prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:prop_EigenSimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.
@@
 According to DoS and port scan attacks' behavior, DoS attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively.
-These characteristics are substantiated by the results obtained through the covariation and correlation analysis described in Section \ref{sec:experimentalresults}, which shows that the main components of DoS attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents a highly correlated traffic.
-
-Therefore, to detect DoS attacks, it is necessary to calculate the covariance matrix $\boldsymbol{R}_{yy}^{(q)}$ of the zero mean samples given by
+These characteristics are substantiated by the results obtained through our analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:experimentalresults}, which shows that the main components of DoS attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
+
+Therefore, to detect DoS attacks, it is necessary to calculate the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
 
@@
 
-The set of obtained vectors $\boldsymbol{y}_{m}^{(q)}$ composes the zero mean matrix $\boldsymbol{Y}^{(q)}$, then the covariance matrix $\boldsymbol{R}_{yy}^{(q)}$ can be calculated as follows
+The set of obtained vectors $\boldsymbol{y}_{m}^{(q)}$ composes the zero mean matrix $\boldsymbol{Y}^{(q)}$, then the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
 
 \begin{equation}\label{eq:eq03}
-\boldsymbol{R}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
-\end{equation}
-
-For the port scan attack detection, it is necessary to calculate the correlation matrix $\boldsymbol{R}_{zz}^{(q)}$, instead of the covariance matrix $\boldsymbol{R}_{yy}^{(q)}$ used for DoS detection, since the main components are not dominated by the variables with large variance and since port scan attack presents a highly correlated network traffic.
-To obtain the correlation matrix $\boldsymbol{R}_{zz}^{(q)}$, it is required, for each variable, to make the standard deviation unitary as follows
+\boldsymbol{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
+\end{equation}
+
+For the detection of the port scan attack, the main components are not dominated by the variables with large variance.
+Moreover, the portscan traffic presents a highly correlated network traffic.
+In order to exploit such structure, we compute the sample covariance $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have zero mean and unitary standard deviation as follows
 
@@
 
-The set of vectors $\boldsymbol{z}_{m}^{(q)}$ composes the matrix $\boldsymbol{Z}^{(q)}$, then the correlation matrix $\boldsymbol{R}_{zz}^{(q)}$ can be calculated via
+The set of vectors $\boldsymbol{z}_{m}^{(q)}$ composes the matrix $\boldsymbol{Z}^{(q)}$, then the sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)}$ can be calculated via
 
 \begin{equation}\label{eq:eq05}
-\boldsymbol{R}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
-\end{equation}
-
-Once the $\boldsymbol{R}_{yy}^{(q)}$ and $\boldsymbol{R}_{zz}^{(q)}$ have been obtained for DoS and port scan attack detection, respectively, for the sake of simplicity, we refer to $\boldsymbol{R}_{yy}$ and $\boldsymbol{R}_{zz}$ as a matrix $\boldsymbol{C}$.
-Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:eq06}), in order to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).
+\boldsymbol{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
+\end{equation}
+
+Once the $\boldsymbol{\hat{R}}_{yy}^{(q)}$ and $\boldsymbol{\hat{R}}_{zz}^{(q)}$ have been obtained for DoS and port scan attack detection, respectively, and since the next steps are the same for both sample covariance matrices, we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{C}$.
+Therefore, the following step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:eq06}), in order to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).
 
@@
 \begin{equation}\label{eq:eq060}
-\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}).
-\end{equation}
-
-The eigenvalues should be sorted in descending order, as defined by $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ...
-> \lambda_{m}^{(q)}$, to make possible the selection of the first eigenvalue in the obtained sequence, represented by $\lambda_{1}^{(q)}$, which is the largest eigenvalue of the $q$-th time frame evaluated for attack detection.
-
-The matrix of eigenvalues of $\boldsymbol{C}^{(q)}$ can be represented in the following fashion
+\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
+\end{equation}
+
+where the operator diag$(\cdot)$ extracts the main diagonal of a matrix.
+
+The eigenvalues should be sorted in descending order, i.e.
+$\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ...
+> \lambda_{m}^{(q)}$.
+Therefore, the largest eigenvalue of the $q$-th time frame evaluated for the attack detect is given by $\lambda_{1}^{(q)}$.
+
+The concatenation of the eigenvalues vector $\boldsymbol{e}^{(q)}$ for $q = 1, \ldots, Q$ is represented by
 
@@
 
-The process of obtaining the $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, $q = 1, 2, 3, ..., Q$ and the matrix $\boldsymbol{C}^{(q)}$, finding the largest eigenvalue for each $q$-th time frame, should be repeated until $q = Q$, in order to obtain the largest eigenvalue of all time frames.
-Since $\lambda_1^{(q)} > \lambda_2^{(q)} > \lambda_3^{(q)} > \cdots > \lambda_{m-1}^{(q)} > \lambda_m^{(q)}$, then the first line of the matrix $\boldsymbol{E}$ contains the largest eigenvalues of each $q$-th time frame, which is the Greatest Eigenvalue Time
+Note that since $\lambda_1^{(q)} > \lambda_2^{(q)} > \lambda_3^{(q)} > \cdots > \lambda_{m-1}^{(q)} > \lambda_m^{(q)}$, then the first line of the matrix $\boldsymbol{E}$ contains the largest eigenvalues of each $q$-th time frame, which is the Greatest Eigenvalue Time
 Vector (GETV) \cite{tenorio2013greatest}, denoted as
@@
 \begin{equation}\label{eq:eq08}
-\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\}.
+\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
 \end{equation}
@@
 
-Once obtained the largest eigenvalue of each $q$-th time frame, it is possible to apply a selected MOS scheme to estimate the model order $\hat{d}$, which is the estimated number of time frames under attack.
-Therefore, $\boldsymbol{e}_{\rm max}$ is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\boldsymbol{e}_{\rm max})$.
+Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$.
+However, our goal here is to detect the variations of the eigenvalues for different values of $q$.
+Therefore, instead of using a certain $q$, our approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame, in order to identify variations and estimate the model order $\hat{d}$, which is the estimated number of time frames under attack.
+Therefore, $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
 Note that some MOS schemes may also require the number of minutes that compose a time frame, as $\hat{d} = \rm{MOS}(\boldsymbol{e}_{\rm max},\emph{Q})$.
-For more information about MOS, we refer to Appendix \ref{sec:mos}.
+For more information about MOS, we refer to \ref{sec:mos}.
 
@@
 
-The selected MOS scheme can estimate the number of time frames under attack, denoted by $\hat{d}$, but it does not provide information of which $q$-th time frames are under attack.
-However, the time frames under attack can be estimated through the largest eigenvalue analysis, which indicated that the time frames with the $\hat{d}$ largest eigenvalues correspond to the $q$-th time frames under attack.
-
-The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed as Algorithm \ref{alg:alg01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$.
-The Algorithm \ref{alg:alg01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$ (line 2) and its correspondent index, according to lines 4 and 5.
-Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$ (line 8) and a new iteration is performed until $\boldsymbol{e}_{\rm max} = \{\}$.
+After applying the MOS schemes to the vector $\sim\boldsymbol{e}_{\rm max}$, we obtain the an estimate of the $\#\boldsymbol{A}$.
+For instance, in the case of fraggle, synflood and portscan, if $\hat{d} = 1$, , then $\#\boldsymbol{A} = 1$, which means that during the during the $Q$ time frames one attack is present.
+However, if $\hat{d} = 0$, then $\#\boldsymbol{A} = 0$, and this means that none of these attacks are present.
+Note that $\hat{d}$ can be greater than 1, indicating the presence of more than one attack.
+
+In Subsection \ref{sec:prop_MOSSchemes}, we obtained only if $\hat{d} = 1$ or $\hat{d} = 0$.
+However, if $\hat{d} = 1$, the MOS schemes do not provide any information about the $q$-th time frame under attack.
+The identification of the $q$-th time frame under attack can be done through a eigenvalues analysis.
+
+The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed according to Algorithm \ref{alg:alg01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$.
+Algorithm \ref{alg:alg01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$, according to Line 2 of Algorithm \ref{alg:alg01}, and its correspondent index, according to Lines 4 and 5 of Algorithm \ref{alg:alg01}.
+Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$, according to Line 8 of Algorithm \ref{alg:alg01}, and a new iteration is performed until $\boldsymbol{e}_{\rm max} = \{\}$.
 
@@
 		\show\LOOP
-	    \ALOOP {$kk = 1$ until $kk == \hat{d}$} 
+	    \ALOOP {$f = 1$ until $f == \hat{d}$} 
 	    	    \STATE $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
@@
 				\IF {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
-				    \STATE $\boldsymbol{\hat{q}}_{\rm max}^{(kk)} = i$
+				    \STATE $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
 				\ENDIF
         		\ENDALOOP
-	    		\STATE $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(kk)}$
+	    		\STATE $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
     		\ENDALOOP
@@
 
+This subsection describes the proposed eigen similarity analysis for detailed attack identification, in complement to the attack estimation done through MOS schemes and eigenvalue analysis.
+In Subsection \ref{sec:prop_TimeSimilarityAnalysis} we present the eigen similarity analysis for identification of time under attack.
+Next, in Subsection \ref{sec:prop_PortSimilarityAnalysis}, we show how to apply the eigen similarity analysis in order to identify network ports under attack.
+
 \subsubsection{Time Similarity Analysis}
@@
 \begin{figure}[h!]
-     \includegraphics[height=3.6cm, width=12.4cm]{results/figures/incremental.eps}
-     \caption{\csentence{Scenario.} Traffic selection for incremental approach.}
+     \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental.eps}
+     \caption{Traffic selection for incremental approach.}
      \label{fig:fig8}
@@
 \begin{figure}[h!]
-     \includegraphics[height=3.6cm, width=12.4cm]{results/figures/individualized.eps}
-     \caption{\csentence{Scenario.} Traffic selection for individual approach.}
+     \includegraphics[height=4cm, width=12.4cm]{results/figures/individualized.eps}
+     \caption{Traffic selection for individual approach.}
      \label{fig:fig9}
@@
 \begin{figure}[h!]
-     \includegraphics[height=3.6cm, width=12.4cm]{results/figures/incremental_individualized.eps}
-     \caption{\csentence{Scenario.} Traffic selection for incremental individualized approach.}
+     \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental_individualized.eps}
+     \caption{Traffic selection for incremental individualized approach.}
      \label{fig:fig2}
@@
 
-%The algorithm described above can be summarized as following:
-%\begin{algorithm}[h!]
-%  \caption{GETV Attack Detection}
-%  \label{alg:alg01}
-%  \begin{algorithmic}[1]
-%  \show\LOOP
-%    \REQUIRE Network Traffic Log
-%    \ENSURE GETV
-%    \STATE $\textbf{X}$ = matrix of ports and its occurrences per time \emph{\COMMENT{Data Pre-Processing}}
-%    \STATE $Q$ = number of time frames
-%    \ALOOP {$q=1$ util $q=Q$} 
-%        \STATE $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M\times{N}}$
-%        \IF {$isDosAttack$}
-%            \STATE $\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}$
-%            \STATE $\boldsymbol{C}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}$
-%        \ENDIF
-%        \IF {$isPortscanAttack$}
-%			\STATE $\boldsymbol{y}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}$
-%			\STATE $\boldsymbol{R}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}$
-%		\ENDIF
-%		\STATE $\boldsymbol{e}^{(q)} = \boldsymbol{e}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{e}^{(q)^{\rm T}}$
-%       \STATE $[\lambda_{m}^{(1)},\lambda_{m}^{(2)},\lambda_{m}^{(3)}, ..., \lambda_{m}^{(q)}] = \boldsymbol{e}^{(q)}$                        
-%        \STATE $\lambda_{m}^{(1)} > \lambda_{m}^{(2)} > \lambda_{m}^{(3)} > ... > \lambda_{m}^{(q)}$
-%    \ENDALOOP
-%	\STATE $\boldsymbol{E} =
-%		\begin{bmatrix}
-%			\lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
-%			\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
-%			\lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
-%			\vdots & \vdots & \ddots & \vdots  \\
-%			\lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
-%		\end{bmatrix}$
-%	\STATE $\boldsymbol{E}_{1}^{q} = \lambda_1^{(1)}, \lambda_1^{(2)}, \lambda_1^{(3)}, \cdots, \lambda_1^{(q)}$
-%	\STATE $\boldsymbol{e}_{\rm max} = \boldsymbol{E}_{1}^{q}$
-%  \end{algorithmic}
-%\end{algorithm}
-
 \section{Experiments and Results}
@@
 First, in Section \ref{sec:AnalyzedScenario}, the scenario adopted in the experiments is summarized.
-Then, the Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames.
+Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames.
 In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection.
-The Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, and the Section \ref{sec:EigenSimilarityAnalysis} shows the obtained results of similarity analysis for detailed DoS and port scan attack identification.
+Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, and Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed DoS and port scan identification.
 
@@
 Therefore, as the time of each sampling period is one minute, then $N = 20$.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, as well as a covariance $\boldsymbol{R}_{yy}^{(q)} \in \mathbb{R}^{17 \times 17}$ (calculated via (\ref{eq:eq03})) and a correlation matrix $\boldsymbol{R}_{zz}^{(q)} \in \mathbb{R}^{17x17}$, assuming that $q = 1, 2, 3, 4, 5$ and $6$.
+For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, as well as a covariance $\boldsymbol{\hat{R}}_{yy}^{(q)} \in \mathbb{R}^{17 \times 17}$ (calculated via (\ref{eq:eq03})) and a sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)} \in \mathbb{R}^{17x17}$, assuming that $q = 1, 2, 3, 4, 5$ and $6$.
 
@@
 
-For the evaluation of MOS Schemes acuracy for DoS and port scan detection, our framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance or correlation matrix calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}.
+For the evaluation of MOS Schemes acuracy for DoS and port scan detection, our framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}.
 Through eigenvalue analysis of traffic with DoS or port scan attacks, it is possible to visualize a significant difference between the largest eigenvalues and the remain eigenvalues, which can indicate a relationship between an outlier and time frames under attack.
 
-Figure \ref{fig:fig10} depics the eigenvalues calculated from covariance matrix of the network traffic used to evaluate the synflood attack identification.
-
-\begin{figure}[h!]
-     \includegraphics[height=5cm, width=7cm]{results/figures/eigenvalues_synflood.eps} 
-     \caption{\csentence{Synflood Eigenvalues.} Eigenvalues of the covariance matrix (synflood).}
+\begin{figure}[h!]
+	\centering
+     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_synflood.eps} 
+     \caption{Eigenvalues of the sample covariance matrix (synflood).}
      \label{fig:fig10}
@@
 
-In this figure its possible to see that the largest eigenvalue, which is related to the simulated synflood attack ($q = 4$), stands out significantly from the others eigenvalues.
-
-Figure \ref{fig:fig11} illustrates represents the eigenvalues calculated from covariance matrix of the matrix used for fraggle attack detection.
-In this figure it is possible to see that the largest eigenvalue, which is related to this attack ($q = 5$), stands out significantly from the others eigenvalues, in accordance with the result shown in Figure \ref{fig:fig10} for the synflood attack analysis.
-
-\begin{figure}[h!]
-     \includegraphics[height=5cm, width=7cm]{results/figures/eigenvalues_fraggle.eps}
-     \caption{\csentence{Fraggle Eigenvalues.} Eigenvalues of the covariance matrix (fraggle).}
+Figure \ref{fig:fig10} depicts the eigenvalues calculated from sample covariance matrix of the network traffic used to evaluate the synflood attack identification.
+In Figure \ref{fig:fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 4$) stands out significantly from the other eigenvalues.
+
+Figure \ref{fig:fig11} illustrates the eigenvalues calculated from sample covariance matrix of the matrix used for fraggle attack detection.
+In Figure \ref{fig:fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 5$) stands out significantly from the other eigenvalues, in accordance with the result shown in Figure \ref{fig:fig10} for the synflood attack analysis.
+
+\begin{figure}[h!]
+	\centering
+     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_fraggle.eps}
+     \caption{Eigenvalues of the sample covariance matrix (fraggle).}
      \label{fig:fig11}
@@
 
-Figure \ref{fig:fig12} depics the eigenvalues calculated from correlation matrix of the network traffic matrix evaluated for port scan detection.
-
-\begin{figure}[h!]
-     \includegraphics[height=5cm, width=7cm]{results/figures/eigenvalues_portscan.eps}
-     \caption{\csentence{Port Scan Eigenvalues.} Eigenvalues of the correlation matrix (port scan).}
+Figure \ref{fig:fig12} depicts the eigenvalues calculated from covariance matrix of zero mean and unitary standard deviation variables, of the network traffic matrix evaluated for port scan detection.
+
+\begin{figure}[h!]
+	\centering
+     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_portscan.eps}
+     \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}
      \label{fig:fig12}
@@
 
-As analyzed for the synflood and fraggle attacks, it is possible to observe that the largest eigenvalue, related to this attack ($q = 3$), stands out significantly from the others eigenvalues.
+As analyzed for the synflood and fraggle attacks, note that the largest eigenvalue, related to this attack ($q = 3$), stands out significantly from the others eigenvalues.
 
@@
 In Table \ref{tab:tab3}, note the significant variation of the eigenvalues associated with attacks, in comparison to the others.
-At $q = 4$, where the synflood attack occurred, the maximum eigenvalue obtained was approximately 21 times larger than the second one.
-At $q = 5$, where the fraggle attack occurred, the maximum eigenvalue obtained was about 29,000 times larger than the second one.
-At $q = 3$, where the port scan attack occurred, the maximum eigenvalue obtained was approximately 4 times larger than the second one.
+At $q = 4$, where the synflood attack occurred, the maximum eigenvalue obtained is approximately 21 times larger than the second one.
+At $q = 5$, where the fraggle attack occurred, the maximum eigenvalue obtained is about 29,000 times larger than the second one.
+At $q = 3$, where the port scan attack occurred, the maximum eigenvalue obtained is approximately 4 times larger than the second one.
 In the last case, for port scan attack detection, although the largest eigenvalue presented no too large variance to the second one, if compared to synflood or fraggle attacks, it clearly deviates from the remaining largest eigenvalues.
 
-These results highlight that all $q$-th time frames, where a network attack was simulated, presented high significant variance between the largest eigenvalue and the remaining eigenvalues, obtained from covariance matrix, for DoS detection, or from correlation matrix, for port scan detection.
+These results highlight that all $q$-th time frames, where a network attack was simulated, present high significant variance between the largest eigenvalue and the remaining eigenvalues, obtained from sample covariance matrix, for DoS detection, or from covariance matrix of zero mean and unitary standard deviation variables, for port scan detection.
 Therefore, we propose to apply the vector of the largest eigenvalues to MOS schemes in order to evaluate their accuracy for identification of time frames under attack, motivated by the fact that it is relevant to apply MOS schemes to automate the attack detection process, taking into account the characteristics of the evaluated eigenvalues.
@@
 
-In \cite{tenorio2013greatest}, we evaluated the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE MOS schemes \cite{da2009comparison,tenorio2013greatest} for synflood and port scan attack detection.
-In this work we extended that evaluation for fraggle attack detection, applying the same schemes to fraggle attack detection over the traffic presented in Section \ref{sec:datamodel}, as results shown on following Table \ref{tab:tab4}.
+In \cite{tenorio2013greatest}, we evaluate the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE MOS schemes \cite{da2009comparison,tenorio2013greatest} for synflood and port scan attack detection.
+In this work we extend that evaluation for fraggle attack detection, applying the same schemes to fraggle attack detection over the traffic presented in Section \ref{sec:datamodel}, as results shown in Table \ref{tab:tab4}.
 
@@
   \centering
-  \footnotesize
+  \scriptsize
   \caption{MOS schemes applied to port scan and DoS detection}
@@
 	\toprule
-	\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{Real value (d)}}\\ 
+	\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
 			\hhline{~------~}
@@
 
-It was expected 1 as the model order value $\hat{d}$ when there was only one attack, values greater than 1, returned by one MOS scheme, indicates that there is more than one attack.
+Note that $\hat{d} = 1$, if there is attack, while $\hat{d} > 1$ indicates more than one attack.
 An example of this could be seen for attack detection via EFT for traffic containing synflood and fraggle attacks, showing $\hat{d} = 2$, which indicates the presence of two attacks, as expected by the $d$ real values of Table \ref{tab:tab4}.
@@
 
-%Figure \ref{fig:fig1} shows a typical profile of eigenvalues​​.
-The last P – 1 eigenvalues are used to estimate the (M - P)-th eigenvalue, denoted by the yellow rectangle.
-The EFT method considers the discrepancy between the actual value and the estimated value obtained \cite{da2007enhanced}.
-
-%\begin{figure}[h!]
- %    \centering 
-     %\includegraphics[height=5cm, width=6cm]{results/figures/fig01.png}
-  %   \caption{\csentence{EFT Example.} Example of application of EFT \cite{da2007enhanced}.}
-   %  \label{fig:fig1}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Backmatter begins here                   %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{backmatter}
-
-\section*{Competing interests}
-  The authors declare that they have no competing interests.
-
-%\section*{Author's contributions}
-%    Text for this section \ldots
-
-\section*{Acknowledgements}
-The authors thank the Brazilian Ministry of Planning, Budget and Management for the support during the development of this work.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                  The Bibliography                       %%
-%%                                                         %%
-%%  Bmc_mathpys.bst  will be used to                       %%
-%%  create a .BBL file for submission.                     %%
-%%  After submission of the .TEX file,                     %%
-%%  you will be prompted to submit your .BBL file.         %%
-%%                                                         %%
-%%                                                         %%
-%%  Note that the displayed Bibliography will not          %%
-%%  necessarily be rendered by Latex exactly as specified  %%
-%%  in the online Instructions for Authors.                %%
-%%                                                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% if your bibliography is in bibtex format, use those commands:
-\bibliographystyle{references} 		% Style BST file
-\bibliography{references}	      	% Bibliography file (usually '*.bib' )
-
-% or include bibliography directly:
-% \begin{thebibliography}
-% \bibitem{b1}
-% \end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Figures                       %%
-%%                               %%
-%% NB: this is for captions and  %%
-%% Titles. All graphics must be  %%
-%% submitted separately and NOT  %%
-%% included in the Tex document  %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Do not use \listoffigures as most will included as separate files
-
-%\section*{Figures}
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig01.png}
-%     \caption{\csentence{EFT Example.} Example of application of EFT \cite{da2007enhanced}.}
-%     \label{fig:fig1}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig02.png}
-%     \caption{\csentence{Traffic Matrix.} Obtaining the traffic matrix $\boldsymbol{X}^{(q)}$.}
-%     \label{fig:fig2}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig03.png}
-%     \caption{\csentence{Legitimate Traffic.} The legitimate traffic.}
-%     \label{fig:fig3}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig04.png}
-%     \caption{\csentence{Noise Traffic.} The noise traffic.}
-%     \label{fig:fig4}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig05.png}
-%     \caption{\csentence{Synflood Traffic.} The traffic characterized by synflood.}
-%     \label{fig:fig5}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig06.png}
-%     \caption{\csentence{Fraggle Traffic.} The traffic characterized by fraggle.}
-%     \label{fig:fig6}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig07.png}
-%     \caption{\csentence{Port scan Traffic.} The traffic characterized by port scan.}
-%     \label{fig:fig7}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig08.png}
-%     \caption{\csentence{GETV.} Proposed GETV attack detection approach.}
-%     \label{fig:fig8}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig09.png}
-%     \caption{\csentence{Scenario.} Analyzed Scenario.}
-%     \label{fig:fig9}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig10.png}
-%     \caption{\csentence{Synflood Eigenvalues.} Eigenvalues of the covariance matrix (synflood).}
-%     \label{fig:fig10}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig11.png}
-%     \caption{\csentence{Fraggle Eigenvalues.} Eigenvalues of the covariance matrix (fraggle).}
-%     \label{fig:fig11}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig12.png}
-%     \caption{\csentence{Port scan Eigenvalues.} Eigenvalues of the covariance matrix (port scan).}
-%     \label{fig:fig12}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig13.png}
-%     \caption{\csentence{Synflood Components.} The two first principal components (synflood).}
-%     \label{fig:fig13}
-%\end{figure}
-
-%\begin{figure}[h!]
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig14.png}
-%     \caption{\csentence{Fraggle Components.} The two first principal components (fraggle).}
-%     \label{fig:fig14}
-%\end{figure}
-
-%\begin{figure}[h!]	
-%     \centering 
-%     %\includegraphics[height=10cm, width=10cm]{figures/fig15.png}
-%     \caption{\csentence{Port scan Components.} The two first principal components (port scan).}
-%     \label{fig:fig15}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Tables                        %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Use of \listoftables is discouraged.
-%%
-%\section*{Tables}
-%\begin{table}[h!]
-%  \centering
-%  \footnotesize
-%  \caption{Comparison between malicious traffic detection schemes of related works and of this paper}
-%  \label{tab:tab1}
-%  \begin{tabular}{ c c c c c c }
-%	\toprule
-%	\multirow{2}{*}{\textbf{Related Works}} &\multicolumn{5}{c }{\textbf{Techniques}}\\ 
-%		&\textbf{Data Mining}	 &\textbf{Regular analysis of files}	 &\textbf{PCA}	 &\textbf{MOS}	 &\textbf{GETV}\\
-%	\midrule
-%	(He et al, 2008)		&x	&-	&-	&-	&- \\
-%	(Raynal et al, 2004) 	&-	&x	&-	&-	&- \\
-%	(Almotairi et al, 2009)	&-	&-	&x	&-	&- \\
-%	(da Costa et al, 2012) 	&-	&-	&x	&x	&- \\
-%	(Proposed) 				& -	& -	&x	&x	&x \\
-%   \bottomrule
-%  \end{tabular}
-%\end{table}
-%\begin{table}[h!]
-%  \centering
-%  \caption{Penalty functions for the schemes AIC, MDL and EDC}
-%  \label{tab:tab2}
-%  \begin{tabular}{*2c}
-%	\toprule
-%	\textbf{Scheme} &  \textbf{Penalty function} \\
-%	\textbf{} &  $p(k,N,\alpha)$ \\
-%	\midrule
-%    AIC	& $k(2\alpha - k)$ \\
-%    MDL	& $0.5k(2\alpha - k) \hspace{1 mm} log(N)$ \\
-%    EDC	& $0.5k(2\alpha - k)\sqrt{N\hspace{1 mm}ln(lnN)}$ \\
-%    \bottomrule
-%  \end{tabular}
-%\end{table}
-%\begin{table}[h!]
-%  \centering
-%  \footnotesize
-%  \caption{Greatest Eigenvalue related to attacks detection}
-%  \label{tab:tab3}
-%  \begin{tabular}{ c c c c c }
-%	\toprule
-%	\multirow{3}{*}{\textbf{Time Period} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
-%			\hhline{~----}
-%		&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
-%		&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
-%	\midrule
-%	1 &1887545 &1887545 &1887545 &2,0734 \\
-%	2 &2341327 &2341327 &2341327 &2,1451 \\
-%	3 &3213867 &3213867 &3213867 &10,0718 \\
-%	4 &133238294 &133238294 &731229 &2,1620 \\
-%	5 &92384021611 &6367983 &92384021611 &2,4253 \\
-%	6 &708335 &708335 &708335 &1,7948 \\
-%    \bottomrule
-%  \end{tabular}
-%\end{table}
-%\begin{table}[h!]
-%  \centering
-%  \footnotesize
-%  \caption{MOS schemes applied to the GETV}
-%  \label{tab:tab4}
-%  \begin{tabular}{ c c c c c c c c }
-%	\toprule
-%	\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} %&{\textbf{Real value (d)}}\\ 
-%			\hhline{~------~}
-%		&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
-%	\midrule
-%	Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-%	Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-%	\midrule
-%	Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-%	Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-%	\midrule
-%	Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
-%	Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
-%	\midrule
-%	Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
-%	Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-%    \bottomrule
-%  \end{tabular}
-%\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Additional Files              %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\section*{Additional Files}
-%  \subsection*{Additional file 1 --- Sample additional file title}
-%    Additional file descriptions text (including details of how to view the file, if it is in a non-standard format or the file extension). This might refer to a multi-page table or a figure.
-
-%  \subsection*{Additional file 2 --- Sample additional file title}
-%    Additional file descriptions text.
-
-\end{backmatter}
+\section*{References}
+
+\bibliography{references}
+
 \end{document}

### 2016-04-13 docs/papers/2015eurasip/code/eigen_analysis/correlacao.m

@@ -1,0 +1,27 @@
+%correlation matrix (R)
+%eigenvalues (E) of correlation matrix
+%eigenvectors (V) of correlation matrix 
+%maximum eigenvalue (M)
+%correlation coefficient or "Pearson's correlation coefficient" (Y)
+%map with original data (O)
+function [R,E,V,M,Y,O] = correlacao(X)
+numLines = size(X,1);
+numColumns = size(X,2);
+j = 0;
+
+for i = 1:numLines
+	if std(X(i,:),1) > 0
+		j = j + 1;
+        O(j) = i;
+		Y(j,:) = (X(i,:) - mean(X(i,:))) / std(X(i,:),1);	%correlation coefficient or "Pearson's correlation coefficient"
+    end
+end
+
+if j == 0
+	Y = [,];
+    O = [];
+end
+
+R = 1/numColumns*Y*Y';										%correlation matrix (R)
+[V,E] = eig(R);												%eigenvectors (V) e eigenvalues (E) of correlation matrix
+M = max(diag(E));											%maximum eigenvalue (M)

### 2016-04-13 docs/papers/2015eurasip/code/eigen_analysis/eigenAnalysisTest.m

@@ -1,0 +1,7 @@
+clc;
+clear all;
+close all;
+
+X = dlmread(['/media/thiago/shared/backup/doutorado/data/all/traffic/1.txt'], '\t');
+[Sc,Ec,Vrc,Mrc] = eigencorrelation(X);
+[Sc,Ec,Vrc,Mrc] = eigencovariance(X)

### 2016-04-13 docs/papers/2015eurasip/code/eigen_analysis/eigencorrelation.m

@@ -1,0 +1,30 @@
+%% Input
+% Matrix
+
+%% Output
+% (R) correlation matrix 
+% (E) eigenvalues diagonal matrix
+% (V) eigenvector matrix
+% (M) largest eigenvalue
+function [R,E,V,M] = eigencorrelation(X)
+    numLines = size(X,1);
+    numColumns = size(X,2);
+    Y = zeros(numLines,numColumns);
+    
+    for i = 1:numLines
+        stdv = std(X(i,:),1);
+        if stdv > 0          
+            Y(i,:) = (X(i,:) - mean(X(i,:))) / stdv;	% standardization, have zero mean and unit variance
+        end
+    end
+    
+    R = 1/numColumns*Y*Y';								% estimation of correlation matrix (R), with conjugate transpose
+    [V,E] = eig(R);										% eigenvectors (V) and eigenvalues (E) of correlation matrix
+    M(1) = max(diag(E));                                % value of the largest eigenvalue (M)
+    
+    Ed  = diag(E);    
+    for i = 1:size(Ed,1)
+        if M(1) == Ed(i)
+            M(2) = i;                                   % index of the largest eigenvalue
+        end        
+    end

### 2016-04-13 docs/papers/2015eurasip/code/eigen_analysis/eigencovariance.m

@@ -1,0 +1,27 @@
+%% Input
+% Matrix
+
+%% Output
+% (S) covariance matrix 
+% (E) eigenvalues diagonal matrix
+% (V) eigenvectors matrix
+% (M) largest eigenvalue
+function [S,E,V,M] = eigencovariance(X)
+    numLines = size(X,1);
+    numColumns = size(X,2);
+    Y = zeros(numLines,numColumns);
+    
+    for i = 1:numLines;
+        Y(i,:) = (X(i,:) - mean(X(i,:)));	% zero mean
+    end
+    
+    S = 1/numColumns*Y*Y';					% estimation of covariance matrices (S)    
+    [V,E] = eig(S);							% eigenvectors (V) and eigenvalues diagonal matrix (E)
+    M(1) = max(diag(E));                    % value of the largest eigenvalue (M)
+    
+    Ed  = diag(E);    
+    for i = 1:size(Ed,1);
+        if M(1) == Ed(i)
+            M(2) = i;                       % index of the largest eigenvalue
+        end        
+    end

### 2016-04-13 docs/papers/2015eurasip/code/eigendecomposition (copy)/eigencorrelation.m

@@ -1,0 +1,30 @@
+%% Input
+% Matrix
+
+%% Output
+% (R) correlation matrix 
+% (E) eigenvalues diagonal matrix
+% (V) eigenvectors 
+% (M) largest eigenvalue
+function [R,E,V,M] = eigencorrelation(X)
+    numLines = size(X,1);
+    numColumns = size(X,2);
+    Y = zeros(numLines,numColumns);
+    
+    for i = 1:numLines
+        stdv = std(X(i,:),1);
+        if stdv > 0          
+            Y(i,:) = (X(i,:) - mean(X(i,:))) / stdv;	% standardization or unit variation        
+        end
+    end
+    
+    R = 1/numColumns*Y*Y';								% estimation of correlation matrix (R)
+    [V,E] = eig(R);										% eigenvectors (V) and eigenvalues (E) of correlation matrix
+    M(1) = max(diag(E));                                % largest eigenvalue (M)
+    
+    Ed  = diag(E);    
+    for i = 1:size(Ed,1);
+        if M(1) == Ed(i)
+            M(2) = i;                                   % get the position of largest eigenvalue
+        end        
+    end

### 2016-04-13 docs/papers/2015eurasip/code/eigensimilarity.m

@@ -1,0 +1,150 @@
+clc;
+clear all;
+close all;
+
+addpath('eigen_analysis/');
+addpath('/home/thiago/Dropbox/dev/matlab/mos/');
+
+matrices = {'all';'signal';'noise';'portscan';'synflood';'fraggle'};
+numberOfMatrices = size(matrices,1); %6 matrices
+ports = [80,443,53,21,22,23,25,110,143,161,69,123,445,600,19,67,68];
+numberOfPorts = size(ports,2); %17 ports
+numPeriods = 6;
+periodsSize = 20;
+numMinutes = 120;
+threshold = 0.4;
+
+% incremental individualized
+for matrix = 4:numberOfMatrices    
+    switch matrix		
+        case 1 %all
+            q = 4;
+        case 2 %signal			
+            q = 4;
+		case 3 %noise            
+			q = 4;
+		case 4 %portscan
+            q = 3;
+		case 5 %synflood 
+            q = 4;
+   		case 6 %fraggle 
+			q = 5;
+    end    
+    X0 = dlmread(['/media/thiago/shared/backup/doutorado/data/all/traffic/' num2str(2) '.txt'], '\t');
+    if (q == 3)
+        [S0,E0,Vr0,Mr0] = eigencorrelation(X0);
+    else
+        [S0,E0,Vr0,Mr0] = eigencovariance(X0);
+    end
+    X = dlmread(['/media/thiago/shared/backup/doutorado/data/all/traffic/' num2str(q) '.txt'], '\t');
+    t_attacks = {};
+    for t = 1:periodsSize
+        Xc = cat(2,X0,X(:,1:t));
+        if (q == 3)
+            [Sc,Ec,Vrc,Mrc] = eigencorrelation(Xc);
+        else
+            [Sc,Ec,Vrc,Mrc] = eigencovariance(Xc);
+        end
+        cosTheta(t) = dot(Vr0(:,Mr0(2)),Vrc(:,Mrc(2)))/(norm(Vr0(:,Mr0(2)))*norm(Vrc(:,Mrc(2))));
+        cosTheta(t)  = abs(cosTheta(t));
+        warning('Matrix = %s, Period(q) = %s, Time(t) = %s, cosTheta= %s',matrices{matrix},num2str(q),num2str(t),cosTheta(t));
+        if(cosTheta(t) < threshold)
+            X0 = cat(2,X0,X(:,1:t-1));
+            ports = findport(X0,X(:,t),q == 3,threshold);
[…]

### 2016-04-13 docs/papers/2015eurasip/code/filters.sh

@@ -1,0 +1,67 @@
+echo ">Removing previous data"
+rm -rf data/all
+rm -rf data/signal
+rm -rf data/noise
+rm -rf data/synflood
+rm -rf data/fraggle
+rm -rf data/portscan
+
+echo ">Creating traffic directories"
+mkdir data/all
+mkdir data/signal
+mkdir data/noise
+mkdir data/synflood
+mkdir data/fraggle
+mkdir data/portscan
+mkdir data/all/traffic
+mkdir data/signal/traffic
+mkdir data/noise/traffic
+mkdir data/synflood/traffic
+mkdir data/fraggle/traffic
+mkdir data/portscan/traffic
+
+echo ">ALL = excludes ARP, ICMP and 224.0.0.251"
+cat data/traffic | grep -vP 'ARP|ICMP|224.0.0.251' > data/all/traffic/all.txt
+
+echo ">NOISE = excludes flags(service traffic) and known ports"
+cat data/all/traffic/all.txt | grep -vP 'Flags|\.53: |\.53 >|\.445: |\.445 >|\.69: |\.69 >|\.123: |\.123 >|\.19: |\.19 >' > data/noise/traffic/noise.txt
+
+echo ">SIGNAL = http, https and dns"
+cat data/all/traffic/all.txt | grep -P '\.80: |\.80 >|\.53: |\.53 >|\.443: |\.443 >' > data/signal/traffic/signal.txt
+
+echo ">SYNFLOOD = In/Out port 600 traffic"
+cat data/all/traffic/all.txt | grep -P '\.600: |\.600 >' > data/synflood/traffic/synflood.txt	
+
+echo ">FRAGGLE = In/Out port 19 traffic"
+cat data/all/traffic/all.txt | grep '\.19: ' > data/fraggle/traffic/fraggle.txt
+
+echo ">PORTSCAN = Ports without service running"
+cat data/all/traffic/all.txt | grep -vP '\.600: |\.600 >' | grep -P '\.21: |\.21 >|\.22: |\.22 >|\.23: |\.23 >|\.25: |\.25 >|\.110: |\.110 >|\.143: |\.143 >|\.161: |\.161 >|\.445: |\.69: |\.123: ' > data/portscan/traffic/portscan.txt
+
+for port in 80 443 53 21 22 23 25 110 143 161 69 123 445 600 19 67 68
+do
+	# selects traffic per port
[…]

### 2016-04-13 docs/papers/2015eurasip/code/findport.m

@@ -1,0 +1,42 @@
+function P = findport(X0,Xt,isCorr,threshold)
+    P = [];
+    numPorts = 0;
+    T0 = X0(:,size(X0,2));        % last column of X0
+    for p = 1:size(T0,1)
+        T = T0;
+        T(1:p) = Xt(1:p);           % concatenates until the first attack
+        if (isCorr)
+            [Sx,Ex,Vx,Mx] = eigencorrelation(X0);
+            [Sp,Ep,Vp,Mp] = eigencorrelation(cat(2,X0,T));            
+        else
+            [Sx,Ex,Vx,Mx] = eigencovariance(X0);
+            [Sp,Ep,Vp,Mp] = eigencovariance(cat(2,X0,T));
+        end
+        cosTheta = dot(Vx(:,Mx(2)),Vp(:,Mp(2)))/(norm(Vx(:,Mx(2)))*norm(Vp(:,Mp(2))));        
+        cosTheta  = abs(cosTheta)
+        if(cosTheta < threshold)
+            numPorts = numPorts + 1;
+            P(numPorts) = p;   
+            T = T0;
+            T(1:p-1) = Xt(1:p-1);      % copy ports without attack and use it as comparison
+            p = p + 1;
+            for pa = p:size(T0,1)
+                V0pa = T;
+                V0pa(pa) = Xt(pa);
+                if (isCorr)
+                    [Sx,Ex,Vx,Mx] = eigencorrelation(X0);
+                    [Sp,Ep,Vp,Mp] = eigencorrelation(cat(2,X0,V0pa));            
+                else
+                    [Sx,Ex,Vx,Mx] = eigencovariance(X0);
+                    [Sp,Ep,Vp,Mp] = eigencovariance(cat(2,X0,V0pa));
+                end
+                cosTheta = dot(Vx(:,Mx(2)),Vp(:,Mp(2)))/(norm(Vx(:,Mx(2)))*norm(Vp(:,Mp(2))));        
+                cosTheta  = abs(cosTheta)
+                if(cosTheta < threshold)
+                    numPorts = numPorts + 1;
+                    P(numPorts) = pa;   
+                end
+            end
+            break
+        end
+    end

### 2016-04-13 docs/papers/2015eurasip/code/generatefraggle.sh

@@ -1,0 +1,53 @@
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "01"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "02"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "03"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "04"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "05"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "06"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "07"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "07"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "09"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "10"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "11"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "12"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "13"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "14"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "15"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "16"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "17"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "17"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "19"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "20"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "21"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "22"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "23"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "24"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "25"
+cat fraggle/traffic/fraggle.txt >> traffic
+echo "26"
+

### 2016-04-13 docs/papers/2015eurasip/code/getv_getInfo.m

@@ -1,0 +1,202 @@
+%% Extracts GETV Information
+% It extracts traffic information from provided file and calculates GETV for kinds of traffic
+% Extracts GETV information by different granularities 
+function getv_info()
+    addpath('eigendecomposition/');
+    addpath('modelfit/allfitdist/');
+    addpath('modelfit/fitmethis/');
+
+    %matrix metadata
+    matrices = {'all';'signal';'noise';'portscan';'synflood';'fraggle'};
+    numberOfMatrices = size(matrices,1); %6 matrices
+    ports = [80,443,53,21,22,23,25,110,143,161,69,123,445,600,19,67,68];
+    numberOfPorts = size(ports,2); %17 ports
+    
+    %minutes
+    numPeriods = 6;
+    periodsSize = 20;
+    numMinutes = 120;
+    
+    %seconds
+    numSecPeriods = 120;
+    periodsSecSize = 60;
+    numSeconds = 7200;
+    
+    %time mapping
+    mapMatrixMinutes = containers.Map;
+    mapMatrixSeconds = containers.Map;
+
+    for matrix = 1:numberOfMatrices
+        portMinutes = zeros(numberOfPorts,numMinutes);
+        portSeconds = zeros(numberOfPorts,numMinutes * 60);
+        mapMinutes = containers.Map;
+        mapSeconds = containers.Map;
+        
+        %% Data extraction
+        % Creates matrix portMinutes, computing the number of packet per port x minute
+        [portMinutes, mapMinutes] = getv_getTimeFromFile(['data/' matrices{matrix} '/traffic/minutes_'], numberOfPorts, ports, 'minutes', portMinutes, mapMinutes);
+        %barplotfromlines(portMinutes, 'results/figures/', [matrices{matrix} 'PacketsPortMinutesBar'], 'png');
+        %histplotfromlines(portMinutes, 'results/figures/', [matrices{matrix} 'PacketsPortMinutesHist'], 'png');
+
+        %matrices{matrix}
+        %'Minutes'
+        %[D PD] = allfitdist(sum(portMinutes.'),'PDF'); %Compute and plot results 
+        %D(1)
+        
+        % Creates matrix X, computing the number of packet per port x second
+        [portSeconds, mapSeconds] = getv_getTimeFromFile(['data/' matrices{matrix} '/traffic/seconds_'], numberOfPorts, ports, 'seconds', portSeconds, mapSeconds);
[…]

### 2016-04-13 docs/papers/2015eurasip/code/getv_testInfo.m

@@ -1,0 +1,44 @@
+function getv_testInfo()
+    %matrix metadata
+    matrices = {'all';'signal';'noise';'portscan';'synflood';'fraggle'};
+    numberOfMatrices = size(matrices,1); %6 matrices
+    ports = [80,443,53,21,22,23,25,110,143,161,69,123,445,600,19,67,68];
+    numberOfPorts = size(ports,2); %17 ports    
+    %minutes
+    numPeriods = 6;
+    periodsSize = 20;
+    numMinutes = 120;    
+    %seconds
+    numSecPeriods = 120;
+    periodsSecSize = 60;
+    numSeconds = 7200;
+    
+    for matrix = 1:numberOfMatrices        
+        %% Testing extracted values
+        portMinutes = zeros(numberOfPorts,numMinutes);
+        portSeconds = zeros(numberOfPorts,numMinutes * 60);
+        mapMinutes = containers.Map;
+        mapSeconds = containers.Map;
+        [portMinutes, mapMinutes] = getv_getTimeFromFile(['data/' matrices{matrix} '/traffic/minutes_'], numberOfPorts, ports, 'minutes', portMinutes, mapMinutes);
+        [portSeconds, mapSeconds] = getv_getTimeFromFile(['data/' matrices{matrix} '/traffic/seconds_'], numberOfPorts, ports, 'seconds', portSeconds, mapSeconds);
+        if isequal(sum(sum(portMinutes)),sum(sum(portSeconds))) == 0
+            warning('Error! %s has different ocurrencies for CALCULATED minutes (%s) and seconds (%s)',matrices{matrix},num2str(sum(sum(portMinutes))),num2str(sum(sum(portSeconds))));
+        end
+
+        %% Testing saved port x times
+        portMinutes = 0;
+        portSeconds = 0;
+        for period = 1:numPeriods
+            X = dlmread(['data/' matrices{matrix} '/traffic/' num2str(period) '.txt'], '\t');
+            portMinutes = portMinutes + sum(sum(X));
+        end
+        
+        for period = 1:numSecPeriods
+            Xss = struct2cell(load(['data/' matrices{matrix} '/qs' num2str(period)]));
+            portSeconds = portSeconds + sum(sum(Xss{1}));
+        end
+        
+        if isequal(portMinutes,portSeconds) == 0
+            warning('Error! %s saved different values for SAVED minutes (%s) and seconds (%s)',matrices{matrix},num2str(portMinutes),num2str(portSeconds));
+        end
+    end

### 2016-04-13 docs/papers/2015eurasip/code/plot.py

@@ -1,0 +1,18 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+x = []
+y = np.cos(y1)
+
+
+1887545	2341327	3213867	133238294	6367983.40451472	708335
+21831	212783	44266	531843	4596480	66547
+5306	192	647	634322	32547	148
+0	0.442400000000000	1.39360000000000	491	3139	0
+0	0	0	0.322600000000000	0.0832000000000000	0
+
+plt.plot(x,'g--^', label='regional')
+plt.plot(y,'b-o', label='local')
+plt.legend( loc='upper left', numpoints = 1 )
+plt.show()
+plt.savefig('test.eps')

### 2016-04-13 docs/papers/2015eurasip/code/rgetv.m

@@ -1,0 +1,305 @@
+clc;
+clear all;
+close all;
+
+addpath('/home/thiago/dev/matlab/mos/');
+addpath('/home/thiago/dev/matlab/modelfit/allfitdist/');
+addpath('/home/thiago/dev/matlab/modelfit/fitmethis/');
+addpath('/home/thiago/dev/projects/2015eurasip/code/util/');
+addpath('/home/thiago/dev/projects/2015eurasip/code/eigendecomposition/');
+
+%[1]=80     
+%[2]=443    
+%[3]=53     
+%[4]=21     portscan
+%[5]=22     portscan
+%[6]=23     portscan
+%[7]=25     portscan
+%[8]=110    portscan
+%[9]=143    portscan
+%[10]=161   portscan
+%[11]=69    portscan
+%[12]=123   portscan
+%[13]=445	portscan
+%[14]=600	synflood
+%[15]=19	fraggle
+%[16]=67    
+%[17]=68    
+
+%portscan t=15
+%synfood t=11:20
+%fragle t=11:20
+
+%output_precision(4)
+%warning ('off', 'textscan: no data read');
+%warning ('off', 'division by zero');
+
+matrices = {'all';'signal';'noise';'portscan';'synflood';'fraggle'};
+numberOfMatrices = size(matrices,1); %6 matrices
+ports = [80,443,53,21,22,23,25,110,143,161,69,123,445,600,19,67,68];
+numberOfPorts = size(ports,2); %17 ports
+numPeriods = 6;
+periodsSize = 20;
+numMinutes = 120;
+
+threshold = 0.40;
+
+%% Extracts GETV Information
+% extracts traffic information from provided file and calculates GETV for kinds of traffic
+%getv_getInfo();
+
+
+%% eigen analysis
+for matrix = 1:numPeriods    
+    
+    %portscan 
+	if matrix == 4
+		Erqxx_plot = zeros(5,numPeriods);
+		periods = zeros(numPeriods,1);
+		for period = 1:numPeriods
+			periods(period) = period;            
+            Erqxx = struct2cell(load(['../data/' matrices{matrix} '/Erqxx']));
+			Erq = diag(Erqxx{1}{period});
+			for i = 1:size(Erq)
+				if i > 5
+					break;
+              end
+				Erqxx_plot(i,period) = Erq(i,1);
+            end
+        end
+		plot(Erqxx_plot,'linestyle','-','marker','x');
+        set(gca,'FontSize',16);
[…]

### 2016-04-13 docs/papers/2015eurasip/code/util/PC1_fraggle.m

@@ -1,0 +1,2 @@
+function [v] = PC1_fraggle(tcp80,tcp443,tcp53,udp19,udp67,udp68)
+v = 0.0044*tcp80 + 0.0002*tcp443 + 0.0001*tcp53 - 1.0000*udp19 + 0*udp67 + 0*udp68;

### 2016-04-13 docs/papers/2015eurasip/code/util/PC1_portscan.m

@@ -1,0 +1,2 @@
+function [v] = PC1_portscan(tcp80,tcp443,udp53,tcp21,tcp22,tcp23,tcp25,tcp110,tcp143,tcp161,udp69,udp123,udp445)
+v = 0.0345*tcp80 + 0.0488*tcp443 + 0.0667*udp53 - 0.3150*tcp21 - 0.3150*tcp22 - 0.3150*tcp23 - 0.3150*tcp25 - 0.3150*tcp110 - 0.3150*tcp143 - 0.3150*tcp161- 0.3150*udp69 - 0.3150*udp123 - 0.3150*udp445;

### 2016-04-13 docs/papers/2015eurasip/code/util/PC1_synflood.m

@@ -1,0 +1,2 @@
+function [v] = PC1_synflood(tcp80,tcp443,udp53,tcp600,udp67,udp68)
+v = 0.0242*tcp80 - 0.0290*tcp443 + 0.0006*udp53 + 0.9995*tcp600 + 0*udp67 + 0*udp68;

### 2016-04-13 docs/papers/2015eurasip/code/util/PC2_fraggle.m

@@ -1,0 +1,2 @@
+function [v] = PC2_fraggle(tcp80,tcp443,tcp53,udp19,udp67,udp68)
+v = 0.9999*tcp80 + 0.0086*tcp443 + 0.0124*tcp53 + 0.0044*udp19 - 0*udp67 - 0*udp68;

### 2016-04-13 docs/papers/2015eurasip/code/util/PC2_portscan.m

@@ -1,0 +1,3 @@
+function [v] = PC2_portscan(tcp80,tcp443,udp53,tcp21,tcp22,tcp23,tcp25,tcp110,tcp143,tcp161,udp69,udp123,udp445)
+v = 0.2770*tcp80 + 0.6243*tcp443 + 0.7251*udp53 + 0.0281*tcp21 + 0.0281*tcp22 + 0.0281*tcp23 + 0.0281*tcp25 + 0.0281*tcp110 + 0.0281*tcp143 + 0.0281*tcp161+ 0.0281*udp69 + 0.0281*udp123 + 0.0281*udp445;
+

### 2016-04-13 docs/papers/2015eurasip/code/util/PC2_synflood.m

@@ -1,0 +1,2 @@
+function [v] = PC2_synflood(tcp80,tcp443,udp53,tcp600,udp67,udp68)
+v = -0.0164*tcp80 + 0.9996*tcp443 + 0.0082*udp53 + 0.0213*tcp600 + 0*udp67 + 0*udp68;

### 2016-04-13 docs/papers/2015eurasip/code/util/vector_q.m

@@ -1,0 +1,11 @@
+j=0;
+for i=1:6 
+    j=j+20;
+    Z(i,:)= w(j-19:j);
+end
+A=Z(1,:);
+B=Z(2,:);
+C=Z(3,:);
+D=Z(4,:);
+E=Z(5,:);
+F=Z(6,:);

### 2016-05-15 docs/papers/2016JNCA/2016JNCA.tex
matlab code organization

@@
 \author[ufrgsaddress]{\\Edison P. de Freitas}
+\author[Ilmenauaddress,Fraunhoferaddress]{Giovanni Del Galdo}
 \author[unbaddress]{Rafael T. de Sousa Júnior}
@@
 
-This section describes the proposed technique to detect synflood, fraggle and port scan.
+This section describes the proposed technique to detect synflood, fraggle and port scan, according to Figure \ref{fig:fig80}, which represents the overview of the proposed framework for detection and identification of network attacks.
 In Subsection \ref{sec:prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvelue for each $q$-th time frame.
@@
 In Subsection \ref{sec:prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:prop_EigenSimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.
+
+\begin{figure}[h!]
+	\centering
+     \includegraphics[height=11cm, width=9cm]{mos_eigen_similarity.eps}
+     \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
+     \label{fig:fig80}
+\end{figure}
 
@@
 
+This paper models the network traffic as a signal processing formulation for applying to our framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
+
+The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and  blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
+The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack.
+
 This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and DoS attacks, but still requiring more information for detailed attack detection.
 Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.
-
-The proposed scheme is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and  blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack.
 

### 2016-05-15 docs/papers/2015eurasip/code/modelfit/fitmethis/fitmethis.m

@@ -1,257 +1,0 @@
-function F= fitmethis(data,varargin)
-% FITMETHIS finds best-fitting distribution
-%	F= fitmethis(X) fits all distributions available in MATLAB's function 
-%	MLE to data in vector X and returns them ordered by some criterion: 
-%	LogLikelihood or Akaike. Either 20 continuous or 5 discrete distributions 
-%	are used based on user input or the type of data supplied (see below)
-%
-%	The function returns a structure array with fields: 
-%	name: name of the distribution (see HELP MLE for a list)
-%	par:  vector of parameter estimates (1, 2 or 3 values)
-%	ci:   matrix of confidence limits, one column per parameter
-%	LL:   Log-Likelihood of the data
-%	aic:  Akaike Information Criterion
-%
-%	Optional arguments can be specified as name/value pairs. Argument 
-%	names are case insensitive and partial matches are allowed.
-%
-%	Name			Value
-%	'dtype'		character string that specifies if the data are continuous 
-% 				('cont') or discrete ('disc'). If missing, the function 
-% 				decides the data are discrete if all values of X are 
-% 				natural numbers.
-%
-%	'ntrials'	Specifies number of trials for the binomial distribution. 
-% 				NTRIALS must be either a scalar or a vector of the same 
-% 				size as X. If missing the binomial is not fitted.
-%
-%	'figure'		Either 'on' (default), or 'off'. If 'on' a plot of the 
-% 				data and the best fitting is produced (scaled to match the 
-%				data). Requires aditional function 'plotfitdist'.
-%
-%	'alpha'		A value between 0 and 1 specifying a confidence level
-%				for CI of 100*(1-alpha) (default is 0.05).
-%
-%	'criterion'	Criterion to use to order the fits. It can be: 'LL' for
-%				Log-Likelihood (default), or 'AIC' for Akaike.
[…]

### 2016-05-15 docs/papers/2015eurasip/code/modelfit/fitmethis/plotfitdist.m

@@ -1,87 +1,0 @@
-function plotfitdist(data,distname,par,dtype,varargin)
-%PLOTFITDIST plots data and fitted distribution
-%	PLOTFITDIST(DATA,NAME,PAR,DTYPE) plots the probability density 
-%	function (specified with name and parameters), and the histogram 
-%	of values in vector DATA scaled by total area using 'trapz'. 
-%
-%	NAME (char. string) is the name of the distribution
-%	(for examples of accepted names see 'HELP PDF'). PAR is a 
-%	vector parameters of the distribution, the size of PAR (1 to 3) 
-%	must match the requirements of function PDF for the specified 
-%	distribution. DTYPE is a character string, either 'cont' for continuous 
-%	distributions or 'disc' for discrete ones. Continuous 
-%	distributions are plotted as a line, while discrete ones are 
-%	plotted as an histogram superimposed on the data histogram.
-%
-%	If NAME is 'binomial' an additional parameter is required:
-%		PLOTFITDIST(...,'ntrials',NTRIALS) NTRIALS is a scalar
-%		specifying the number of trials. If NTRIALS is a vector (i.e.
-%		different number of trials for each value in DATA) the plot cannot be
-%		made due to mismatch in the number of simulated data and number of
-%		trials.
-%
-%	No input control is made, so the user is responsible for providing
-%	correct parameters.
-
-
-% Arguments
-for j= 1:2:length(varargin)
-	string= lower(varargin{j});
-	switch string(1:3)
-		case 'ntr'
-			ntrials= varargin{j+1};
-		case 'pre'
-			if ~isempty(varargin{j+1}) distname= varargin{j+1}; end
-		otherwise
-			error('Unknown argument name');
-	end
-end
-
-
-% Histogram for Discrete/Continuous
-if strcmp(dtype,'cont')
-	x = min(data):range(data)/100:max(data);
-	[bincount,binpos] = hist(data,min(100,numel(data)/5));
-else
-	x = unique(data);
-	[bincount,binpos] = hist(data,x);
-end
-
-
-% Calculate predicted
-switch numel(par)
-	case 1
-		if strcmp('binomial',distname)
-			y = pdf('bino',x,ntrials(1),par(1));
-		else
-			y = pdf(distname,x,par(1));
-		end
-	case 2
[…]

### 2016-05-15 fontes/mos-api/src/main/matlab/calc_coef_paper.m

@@ -1,0 +1,49 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Program to generate the coefficients for the EFT method %
+% Joao Paulo C. L. da Costa                               %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+function [local_coeff,prob_found] = calc_coef_paper(M,N,Pfa,coeff,q);
+    max_eig_numb = min([M N]);
+    % Calculating J_final vector since it depends only on M and N
+    for pp = 1:(max_eig_numb-1)
+        pp_aux = pp + 1;
+        a_mi_ff = 225/((pp_aux^2+2)^2);
+        a_mi_fs = 180*pp_aux / ( N * (pp_aux^2 - 1) * (pp_aux^2 +2) );
+        a_mi = (a_mi_ff - a_mi_fs)^0.5;
+        a_ex_fr = 15/(pp_aux^2 + 2);
+        a_diff = a_ex_fr - a_mi;
+        a_final(pp_aux) = (0.5*a_diff)^0.5;
+        r_final(pp_aux) = exp(-2*a_final(pp_aux));
+        J_final(pp_aux) = ( 1 - r_final(pp_aux) ) / ( 1 - (r_final(pp_aux)^pp_aux) );
+    end
+    barra = waitbar(0,'Please wait... Calculating thresholds for EFT');
+    for ii = 1:q
+        noise = randn(M,N);
+        noise_eig = (svd(noise).^2)/N;
+        noise_eig = sort(noise_eig,'descend');
+        for pp = 1:(max_eig_numb-1)
+            pp_aux = pp + 1;
+            sum_eig = sum(noise_eig((max_eig_numb-pp):max_eig_numb));
+            comp_q = noise_eig(max_eig_numb-pp)/sum_eig;
+            comp_q_tot = comp_q/J_final(pp_aux) - 1;
+            %histogram(ii,pp) = comp_q/J_final(pp_aux);
+            prob_q(pp,:) = coeff < comp_q_tot;
+            if (ii == 1)
+                prob_fim(pp,:) = 1*prob_q(pp,:);
+            else
+                prob_fim(pp,:) = ((ii-1)*prob_fim(pp,:)+1*prob_q(pp,:))/ii;
+            end
+        end
[…]

### 2016-05-15 fontes/mos-api/src/main/matlab/edc_short2.m

@@ -1,0 +1,35 @@
+% Function to find the number of users using Akaike equation:
+%
+% Find m, between 1 to M-1, which minimizes the following quantity:
+%
+% AIC(m) = -N(M-m) * log( g(m) / a(m) ) + m * (2M - m)
+%
+% g(m) is the geometric mean of the (M-m) smallest
+% eigenvalues of the covariance matrix of observation. - "geomean"
+% a(m) is the arithmetic mean. - function "mean"
+
+function [usernumber] = edc_short2(eigenvalues,N)
+    eigenvalues
+    N
+    [temp,max_size] = size(eigenvalues)
+	max_size = max(temp,max_size)
+	
+    % Putting in the ascendent order. It only works in the ascendent form
+	eigenvalues = eigenvalues.'
+	eigenvalues = sort(eigenvalues,'ascend')
+	c_N = sqrt(N*log(log(N)))
+	
+    for ii = 1:max_size
+		kk = max_size - ii + 1
+		trecho = eigenvalues(1:kk)
+		gii = geomean(trecho)
+		aii = mean(trecho)
+		% mm must vary from 0 to M - 1
+		% ii starts with 1 and ends in M
+		% wich means M-1 until 0
+		mm = max_size - kk
+		EDC_values(ii) = -2*N * ( max_size - mm ) * log(gii/aii) + mm * (2*max_size - mm) * c_N
+    end
+    
+	[EDC_min,position] = min( EDC_values (1:max_size) )
+	usernumber = position - 1

### 2016-05-15 fontes/mos-api/src/main/matlab/eft_coef.m

@@ -1,0 +1,43 @@
+#Cálculo dos coeficientes: EFT_CALC_COEF
+function [local_coeff,prob_found]=calc_coef_paper(M,N,Pfa,coeff,q);
+	max_eig_numb = min([M N]);
+	for pp = 1:(max_eig_numb-1)
+		pp_aux = pp+1;
+		a_mi_ff = 225/((pp_aux^2+2)^2);
+		a_mi_fs = 180*pp_aux/(N*(pp_aux^2-1)*(pp_aux^2+2));
+		a_mi = (a_mi_ff-a_mi_fs)^0.5;
+		a_ex_fr = 15/(pp_aux^2+2);
+		a_diff = a_ex_fr-a_mi;
+		a_final(pp_aux) = (0.5*a_diff)^0.5;
+		r_final(pp_aux) = exp(-2*a_final(pp_aux));
+		J_final(pp_aux)=(1-r_final(pp_aux))/(1-r_final(pp_aux)^pp_aux));
+	endfor
+	barra = waitbar(0,'Please wait... Calculating thresholds for EFT');
+	for ii = 1:q
+		noise = randn(M,N);
+		noise_eig = (svd(noise).^2)/N;
+		noise_eig = sort(noise_eig,'descend');
+		for pp = 1:(max_eig_numb-1)
+			pp_aux = pp+1;
+			sum_eig = sum(noise_eig((max_eig_numb-pp):max_eig_numb));
+			comp_q = noise_eig(max_eig_numb-pp)/sum_eig;
+			comp_q_tot = comp_q/J_final(pp_aux)-1;
+			prob_q(pp,:) = coeff<comp_q_tot;
+			if (ii == 1)
+				prob_fim(pp,:) = 1*prob_q(pp,:);
+			else
+				prob_fim(pp,:)=((ii-1)*prob_fim(pp,:)+prob_q(pp,:))/ii;
+			endif
+		endfor
+		waitbar(ii/q,barra);
+	endfor
+	close(barra);
+	for pp = 1:(max_eig_numb-1)
+		[temp,local_Pfa(pp)] = min(abs(prob_fim(pp,:)-Pfa));
+		local_coeff(pp) = coeff(local_Pfa(pp));
+		if (local_Pfa(pp) == 1)
+			prob_found(pp) = 0;
+			Else
+			prob_found(pp) = prob_fim(pp,local_Pfa(pp)-1);
+		endif
+	endfor

### 2016-05-15 fontes/mos-api/src/main/matlab/mdl_short2.m

@@ -1,0 +1,33 @@
+function [usernumber] = mdl_short2(eig_vec,N);
+%
+% Function to find the number of users
+% using Akaike equation:
+%
+% Find m, between 1 to M-1, which minimizes the
+% following quantity:
+%
+% AIC(m) = -N(M-m) * log( g(m) / a(m) ) + m * (2M - m)
+%
+% g(m) is the geometric mean of the (M-m) smallest
+% eigenvalues of the covariance matrix of observation. - "geomean"
+% a(m) is the arithmetic mean. - function "mean"
+[temp,M1] = size(eig_vec);
+M1 = max(temp,M1);
+M = M1;
+% Putting in the ascendent order
+% It only works in the ascendent form
+eig_vec = eig_vec.';
+eig_vec = sort(eig_vec,'ascend');
+for ii = 1:M1
+    kk = M1 - ii + 1;
+    trecho = eig_vec(1:kk);
+    gii = geomean(trecho);
+    aii = mean(trecho);
+    % mm must vary from 0 to M - 1
+    % ii starts with 1 and ends in M
+    % wich means M-1 until 0
+    mm = M1 - kk;
+    MDL_values(ii) = -N * (M - mm) * log(gii/aii) + 0.5* mm * (2*M - mm) * log(N);
+end
+[MDL_min,position] = min(MDL_values(1:M1));
+usernumber = position - 1;

### 2016-05-15 fontes/mos-api/src/main/matlab/noise_est_sure.m

@@ -1,0 +1,44 @@
+function [sigma] = noise_est_sure(eig_values,M,N);
+%
+% Taking the 25th percentile of l
+%
+num_eig = min(M,N);
+%
+% It is considered M < N
+%
+eig_desc = sort(eig_values,'descend');
+%
+gamma = N/M;
+%
+b_mp = (1+gamma^(-0.5))^2;
+%
+for ii = 1:num_eig
+    EDF_prob(ii) = (num_eig-ii+1)/num_eig;
+    factor_correc(ii) = mp_density(EDF_prob(ii),gamma);
+    eig_corr_eins(ii) = eig_desc(ii)/factor_correc(ii);
+end
+% Taking the 25th percentile
+percent_25 = round(0.25*num_eig);
+%
+eig_selection = eig_corr_eins((num_eig-percent_25):num_eig);
+%
+sigma_sqr_eins = median(eig_selection);
+%
+% Normalizing the eigenvalues:
+%
+norm_eig_eins = eig_desc/sigma_sqr_eins;
+%
+% Crude estimation for d
+mask_norm = norm_eig_eins > b_mp;
+% Give the position of the positive normalized eigenvalue closest to b_mp
+r_cru = sum(mask_norm);
+%
+for ii = (r_cru+1):num_eig
+    kk = ii - r_cru;
+    EDF_prob_zwei(kk) = (num_eig-ii+1)/(num_eig-r_cru);
+    factor_correc_zwei(kk) = mp_density(EDF_prob_zwei(kk),gamma);
+    eig_corr_zwei(kk) = eig_desc(ii)/factor_correc_zwei(kk);
+end
+%
+sigma = median(eig_corr_zwei);
+%

### 2016-05-15 fontes/mos-api/src/main/matlab/radoi_edc.m

@@ -1,0 +1,48 @@
+function d_est = ranoi_app(eig_values);
+% Discriminant approach
+eig_values = sort(eig_values,'descend');
+eig_num = length(eig_values);
+% Calculating mu_fct
+%
+%for ii = 1:M
+%    kk = M - ii + 1;
+%    trecho = eig_vec(1:kk);
+%    gii = geomean(trecho);
+%    aii = mean(trecho);
+%end
+%
+for kk = 1:(eig_num-1)
+    ii = eig_num - kk + 1;;
+    trecho = eig_values(1:ii);
+    gii = geomean(trecho);
+    aii = mean(trecho);
+    %mu_fct(kk) = (1/(eig_num-kk))*sum(eig_values(kk+1:eig_num));
+    mu_fct(kk) = geomean(trecho)/mean(trecho);
+end
+%
+% Calculating alpha_c
+for kk = 1:(eig_num-1)
+    trecho = eig_values(1:kk);  
+    alpha_var(kk) = (geomean(trecho) - mu_fct(kk))/mu_fct(kk);
+end
+% alpha_const is the position of the element
+[temp,alpha_const] = max(alpha_var);
+%
+% Calculating ksi
+for kk = 1:(eig_num-1)
+    trecho = eig_values(1:kk);  
+    ksi(kk) = 1-alpha_const*(geomean(trecho)-mu_fct(kk))/mu_fct(kk);
+end
+%
+for kk = 1:(eig_num-1)
+    trecho = eig_values(1:kk);  
+    discr_1(kk) = geomean(trecho)/mean(trecho);
+    %
+    discr_2(kk) = ksi(kk)/sum(ksi(1:(eig_num-1)));
+    %
+    discr_crit(kk) = discr_1(kk) - discr_2(kk);
+    %
+end
+%
+%
+[temp2,d_est]= min(discr_crit);

### 2016-05-15 fontes/mos-api/src/main/matlab/sure_method.m

@@ -1,0 +1,52 @@
+function d_est = sure_method(X_mat,M,N);
+% Model order selection
+sing_values = svd(X_mat);
+%
+eig_values = sing_values.^2;
+%
+[sigma_rmt] = noise_est_sure(eig_values,M,N);
+%
+max_d = min(M,N);
+%
+other_dim = max(M,N);
+%
+for ii = 1:(max_d-1)
+    %
+    sigma_r_sqr = (1/(max_d-ii))*sum(eig_values((ii+1):max_d));
+    %
+    C_part_one = 0;
+    C_part_three = 0;
+    for jj = 1:ii
+        %
+        for kk = (ii+1):max_d
+            C_part_one = (4*sigma_rmt/other_dim)* (eig_values(jj)-sigma_r_sqr)/(eig_values(jj)-eig_values(kk)) + C_part_one;
+        end
+        %
+        C_part_three = -(2*sigma_rmt/other_dim)*(max_d-1)*(1-sigma_r_sqr/eig_values(jj)) + C_part_three;
+        %
+    end
+    %
+    C_part_two = (2*sigma_rmt/other_dim)*ii*(ii-1);
+    %
+    C_total = C_part_one + C_part_two + C_part_three;
+    %
+    Risk_part_one = (max_d-ii)*sigma_r_sqr;
+    %
+    Risk_part_two = 0;
+    Risk_part_three = 2*sigma_rmt*ii;
+    Risk_part_four = 0;
+    Risk_part_five = 0;
+    %
+    for jj = 1:ii
+        %
+        Risk_part_two = (sigma_r_sqr^2)/eig_values(jj) + Risk_part_two;
+        %       
+        Risk_part_four = (-2*sigma_rmt*sigma_r_sqr)/eig_values(jj) + Risk_part_three;
+        %
+        Risk_part_five = (4*sigma_rmt*sigma_r_sqr/other_dim)/eig_values(jj) + Risk_part_four;
+        %
+    end
+    %
+    Risk_total(ii) = Risk_part_one + Risk_part_two + Risk_part_three + Risk_part_four + Risk_part_five + C_total;
+end
+[value,d_est] = min(Risk_total);

### 2016-05-26 docs/papers/2016JNCA/2016JNCA.tex
complexity analysis into paper

@@
 This paper proposes a signal processing framework for the detection and identification of network attacks using concepts of model order selection (MOS), eigenvalues and similarity analysis.
-In order to validate our proposed framework, we consider a network traffic sample that contains malicious activity such as denial of service (DoS) and port probing.
+In order to validate the proposed framework, we consider a network traffic sample that contains malicious activity such as denial of service (DoS) and port probing.
 We propose to model the network traffic as a superposition of components, namely, user's operations (legitimate traffic), network service operation not related to the user (noise) and the malicious activity.
@@
 
-Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, but may fail against innovative malicious techniques \cite{lakhina2005mining}.
+Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}.
 In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time.
-Moreover, it is also crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
+Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
 Recently, signal processing schemes have being applied to the detection of malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest}, showing advances in network traffic analysis.
@@
 Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
-Our proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
+To the best of our knowledge there is no similar model in the literature.
+The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
 In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack.
@@
 The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of our proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
+The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
 
@@
 Section \ref{sec:datamodel} presents the data model.
-Section \ref{sec:prop_getv} describes our proposed framework for blind and automatic detection of malicious traffic.
+Section \ref{sec:prop_getv} describes the proposed framework for blind and automatic detection of malicious traffic.
 Section \ref{sec:experimentalresults} discusses the experimental validation and presents the experimental results.
@@
 Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
-Zonglin \emph{et al} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work was not applied to probing and denial of service (DoS) attack detection, simultaneously.
+Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work was not applied to probing and denial of service (DoS) attack detection, simultaneously.
 
@@
 
-Lee \emph{et al} \cite{Lee2013} proposed OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
-In contrast to Lee \emph{et al}, our framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
-Additionally, Lee \emph{et al} only evaluated their proposed scheme for covariance analysis, while we adopted an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for DoS and probing attacks, respectively.
-
-Our proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic.
-Our proposed solution is automatic and blind for detection of time frames under probing and DoS attacks through MOS and eigen analysis.
+Lee \emph{et al.} \cite{Lee2013} proposed OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
+In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
+Additionally, Lee \emph{et al.} only evaluated their proposed scheme for covariance analysis, while we adopted an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for DoS and probing attacks, respectively.
+
+The proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic.
+The proposed solution is automatic and blind for detection of time frames under probing and DoS attacks through MOS and eigen analysis.
 Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.
@@
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
-We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz.
+We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz, and the operator $\sim$ that sorts the elements of a vector in ascending order.
 
@@
 Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce DoS and probing attacks.
-Subsection \ref{sec:ModelingData} presents how network traffic can be modelled as signal superposition, and Subsection \ref{sec:AnalyzedScenario} details the traffic of synflood, fraggle and port scan attacks.
+Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition, and Subsection \ref{sec:AnalyzedScenario} details the traffic of synflood, fraggle and port scan attacks.
 
@@
 
-The environment of the analyzed scenario is composed of two computers and one router with access to Internet and to an internal network, where the simulation of legitimate traffic, noise, DoS and port scan attacks are performed.
-During the traffic generation, one computer assumes the role of attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:fig1}.
+The environment of the analyzed scenario is composed of two computers and one router with access to the Internet and to an internal network, where the simulation of legitimate traffic, noise, DoS and port scan attacks are performed.
+During the traffic generation, one computer assumes the role of the attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:fig1}.
 
@@
 
-In the proposed framework, the goal is to detect the anomalies only taking into account the traffic profile, i.e.
-specific information such as origin IP or day and time of the attack are not considered.
+In the proposed framework, the goal is to detect the anomalies only taking into account the traffic profile, i.e., specific information such as origin IP or day and time of the attack are not considered.
 Therefore, from the entire log information, we just consider the timestamp (for sequencing), port type and port number.
@@
 The matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns.
-Each row represents a communication port (TCP port or UDP port), and each column represents the time, whose unit can be minutes, for instance.
+Each row represents a communication port (TCP port or UDP port), and each column represents time bins having a approtiate size, such as one minute.
 Each element $x_{m,n}^{(q)}$ stands for the number of times that the port $m$ appears at the $n$-th minute, at the $q$-th time frame.
@@
 
-The traffic that is not associated with user's operations and with malicious traffic is modelled as noise $\boldsymbol{N}^{(q)}$.
+The traffic that is not associated with user's operations and with malicious traffic is modeled as noise $\boldsymbol{N}^{(q)}$.
 The automatic acquisition service of logical IP network address (DHCP) is an example of noise.
@@
 
-With respect to the fraggle atack, large packets with UDP echo segments are sent to the broadcast address of a network.
+With respect to the fraggle attack, large packets with UDP echo segments are sent to the broadcast address of a network.
 Every packet is modified to have the source address of the victim, in order to implement the source address spoofing technique.
@@
 According to DoS and port scan attacks' behavior, DoS attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively.
-These characteristics are substantiated by the results obtained through our analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:experimentalresults}, which shows that the main components of DoS attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
+These characteristics are substantiated by the results obtained through the analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:experimentalresults}, which shows that the main components of DoS attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
 
@@
 
-The eigenvalues should be sorted in descending order, i.e.
-$\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ...
+The eigenvalues should be sorted in descending order, i.e., $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ...
 > \lambda_{m}^{(q)}$.
@@
 Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$.
-However, our goal here is to detect the variations of the eigenvalues for different values of $q$.
-Therefore, instead of using a certain $q$, our approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame, in order to identify variations and estimate the model order $\hat{d}$, which is the estimated number of time frames under attack.
+However, the goal here is to detect the variations of the eigenvalues for different values of $q$.
+Therefore, instead of using a certain $q$, the proposed approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame, in order to identify variations and estimate the model order $\hat{d}$, which is the estimated number of time frames under attack.
 Therefore, $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
@@
 
-After applying the MOS schemes to the vector $\sim\boldsymbol{e}_{\rm max}$, we obtain the an estimate of the $\#\boldsymbol{A}$.
+After applying the MOS schemes to the vector $\sim\boldsymbol{e}_{\rm max}$, we obtain the estimate of the $\#\boldsymbol{A}$.
 For instance, in the case of fraggle, synflood and portscan, if $\hat{d} = 1$, , then $\#\boldsymbol{A} = 1$, which means that during the during the $Q$ time frames one attack is present.
@@
 However, if $\hat{d} = 1$, the MOS schemes do not provide any information about the $q$-th time frame under attack.
-The identification of the $q$-th time frame under attack can be done through a eigenvalues analysis.
+The identification of the $q$-th time frame under attack can be carried out through a eigenvalues analysis.
 
@@
 
-This subsection describes the proposed eigen similarity analysis for detailed attack identification, in complement to the attack estimation done through MOS schemes and eigenvalue analysis.
+This subsection describes the proposed eigen similarity analysis for detailed attack identification, in complement to the attack estimation carried out through MOS schemes and eigenvalue analysis.
 In Subsection \ref{sec:prop_TimeSimilarityAnalysis} we present the eigen similarity analysis for identification of time under attack.
@@
 \begin{equation}\label{eq:eq11}
-s_n = \abs{\frac{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}}}},
+s_n = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}}},
 \end{equation}
@@
 
-The Figure \ref{fig:fig8} ilustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.
+Figure \ref{fig:fig8} ilustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.
 
@@
 
-The eigen similarity analysis starts at $\boldsymbol{x}^{(\hat{3})}_{(41)}$ and is incrementally performed until $\boldsymbol{x}^{(\hat{3})}_{(60)}$, in order to calculate the $s_n$.
+The eigen similarity analysis starts at $\boldsymbol{x}^{(3)}_{(41)}$ and is incrementally performed until $\boldsymbol{x}^{(\hat{3})}_{(60)}$, in order to calculate the $s_n$.
 We assume that $s_n < l$ means an attack identification, according the anomaly on similiarity of $s_n$ to a defined limiar $l$.
-Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}$ for eigen similarity analyis, the $s_n$ is calculated according to (\ref{eq:eq11}).
+Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}$ for eigen similarity analyis, the $s_n$ is calculated according to~(\ref{eq:eq11}).
 
@@
     \\
-    s_{m,\hat{n}} = \abs{\frac{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(m,\hat{n})}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(m,\hat{n})}}}},
+    s_{m,\hat{n}} = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(m,\hat{n})}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(m,\hat{n})}}},
   \end{array}\right.
@@
 In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection.
-Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, and Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed DoS and port scan identification.
+Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed DoS and port scan identification, and Section \ref{sec:Complexity} discuss the computational complexity of the proposed framework.
 
@@
 
-For the evaluation of MOS Schemes acuracy for DoS and port scan detection, our framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}.
+For the evaluation of MOS Schemes acuracy for DoS and port scan detection, the framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}.
 Through eigenvalue analysis of traffic with DoS or port scan attacks, it is possible to visualize a significant difference between the largest eigenvalues and the remain eigenvalues, which can indicate a relationship between an outlier and time frames under attack.
@@
 
-These results highlight that all $q$-th time frames, where a network attack was simulated, present high significant variance between the largest eigenvalue and the remaining eigenvalues, obtained from sample covariance matrix, for DoS detection, or from covariance matrix of zero mean and unitary standard deviation variables, for port scan detection.
+These results highlight that all $q$-th time frames where a network attack was simulated, present high significant variance between the largest eigenvalue and the remaining eigenvalues, obtained from sample covariance matrix, for DoS detection, or from covariance matrix of zero mean and unitary standard deviation variables, for port scan detection.
 Therefore, we propose to apply the vector of the largest eigenvalues to MOS schemes in order to evaluate their accuracy for identification of time frames under attack, motivated by the fact that it is relevant to apply MOS schemes to automate the attack detection process, taking into account the characteristics of the evaluated eigenvalues.
@@
 
-Although EDC and EFT presented the same accuracy on our evaluation, the EDC scheme requires less processing time than EFT, which is an important criteria to select EDC as the MOS scheme for DoS and port scan detection on our remain experiments.
+Although EDC and EFT presented the same accuracy on the evaluation, the EDC scheme requires less processing time than EFT, which is an important criteria to select EDC as the MOS scheme for DoS and port scan detection on the remain experiments.
 
@@
 
-The Table \ref{tab:tab5} presents the results of the evaluation of three approaches for similarity analysis of eigenvectors for port scan detection.
+Table \ref{tab:tab5} presents the results of the evaluation of three approaches for similarity analysis of eigenvectors for port scan detection.
 
@@
 
-The Table \ref{tab:tab5} shows the evaluation of the time frame $q=3$, when the port scan attack was simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
+Table \ref{tab:tab5} shows the evaluation of the time frame $q=3$, when the port scan attack was simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
 According to the presented results, it is possible to observe the high similarity between network traffic without attack, which was larger than 0.9610 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with the simulated port scan attack ($n=15$), which was lower than 0.0276 for all evaluated approaches.
@@
 However, the incremental approach figured out low similarity for times without attack, where $n=16, 17, 18, 19, 20$, what indicates that the incremental approach can produce false posive results.
-This behaviour occurs because the incremental approaches appends all selected traffic into the reference traffic for comparison against the original reference traffic, what makes more evident the first lack of similarity but reduces the changing detection capability after an attack detection.
-
-The Table \ref{tab:tab6} presents the results of the evaluation of the similarity analysis of eigenvectors for synflood detection.
+This behavior occurs because the incremental approaches appends all selected traffic into the reference traffic for comparison against the original reference traffic, what makes more evident the first lack of similarity but reduces the changing detection capability after an attack detection.
+
+Table \ref{tab:tab6} presents the results of the evaluation of the similarity analysis of eigenvectors for synflood detection.
 It shows the evaluation of the time frame $q=4$, when the synflood attack is simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
@@
 Comparing the incremental individualized and the individual approaches for eigen similarity analysis, it is possible to observe that the incremental individualized approach obtain lowest values for almost all cases, except for the time $n=14$, where incremental individualized approach identified a larger similarity than the individual approach.
-The incremental individualized appends information about each evaluated traffic, therefore it incorporates traffic behaviours that can reduce the outlier capability detection, as ocurred for the time $n=14$.
-
-The Table \ref{tab:tab7} presents the results of the eigen similarity analysis evaluation for fraggle detection.
+The incremental individualized appends information about each evaluated traffic, therefore it incorporates traffic behaviors that can reduce the outlier capability detection, as ocurred for the time $n=14$.
+
+Table \ref{tab:tab7} presents the results of the eigen similarity analysis evaluation for fraggle detection.
 
@@
 For fraggle attack detection, the lack of similarity between legitimate and malicious traffic was more evident than for the evaluation of synflood and port scan detecion.
-This becaviour can be explained by the number of packets generated through the fraggle attack simulation, that was significative larger than the number of packets generated during the synflood simulation.
+This behavior can be explained by the number of packets generated through the fraggle attack simulation, that was significative larger than the number of packets generated during the synflood simulation.
 Considering the three approaches, the largest value for times under attack was 0.0083, while the shortest value for times without attacks was 0.9993.
@@
 
-The Table \ref{tab:tab8} presents the results of the evaluation of eigen similarity analysis for detection of ports under port scan attack, showing only the time frame $q=3$ and minute $n=15$, due to the simulated port scan attack occurred only at this time, although the remain time frame has been completely evaluated and presented high similarity to the reference of traffic without network attack.
+Table \ref{tab:tab8} presents the results of the evaluation of eigen similarity analysis for detection of ports under port scan attack, showing only the time frame $q=3$ and minute $n=15$, due to the simulated port scan attack occurred only at this time, although the remain time frame has been completely evaluated and presented high similarity to the reference of traffic without network attack.
 
@@
 
-The incremental individualized approach presented more sensibility to anomaly detection than the individual approach, the former produced the identification of a low similarity of 0.0298 for almost all ports under attack, unless the port 21, although our simulation has attacked this port.
+The incremental individualized approach presented more sensibility to anomaly detection than the individual approach, the former produced the identification of a low similarity of 0.0298 for almost all ports under attack, unless the port 21, although the simulation has attacked this port.
 The individual approach was not able to identify low similarity for ports under attack, resunting in values of 0.9997 for ports with anomalous traffic and 0.9999 for ports without network attack.
@@
 
-The Table \ref{tab:tab9} presents the results of the evaluation of eigen similarity analysis for detection of ports under synflood attack, showing only the time frame $q=4$ and minute $n=11$.
+Table \ref{tab:tab9} presents the results of the evaluation of eigen similarity analysis for detection of ports under synflood attack, showing only the time frame $q=4$ and minute $n=11$.
 
@@
 
-According to results presented in Table \ref{tab:tab9}, both approches identifies low similarity for the traffic of port 600, which is the target port of our simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.
-
-The Table \ref{tab:tab10} presents the results of the evaluation of eigen similarity analysis for detection of ports under fraggle attack, showing only the time frame $q=5$ and minute $n=11$.
+According to results presented in Table \ref{tab:tab9}, both approches identifies low similarity for the traffic of port 600, which is the target port of the simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.
+
+Table \ref{tab:tab10} presents the results of the evaluation of eigen similarity analysis for detection of ports under fraggle attack, showing only the time frame $q=5$ and minute $n=11$.
 
@@
 
+\subsection{Complexity Analysis}
+\label{sec:Complexity}
+
+This subsection discuss the computational complexity of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
+
+The EVD, calculated according to (\ref{eq:eq06}), requires the previous calculation of covariance matrix, according to Equations \ref{eq:eq02}, \ref{eq:eq03}, \ref{eq:eq04} and \ref{eq:eq05}.
+The covariance matrix calculation is $O(M^2N)$ and the EVD is $O(N^3)$, where $M$ denotes the number of network ports and $N$ denotes the period time.
+Therefore, the computational complexity for all steps for EVD can be represented as $O(M^2N + N^3)$ and yields an $O(N^3)$ upper bound on the worst-case running time for EVD.
+
+EDC and EFT are the MOS schemes that presented accuracy on the evaluation for the network attack detection.
+The computational complexity evaluation for MOS focuses on EDC scheme, since EDC requires less processing time than EFT but presents the same accuracy for the evaluated scenario.
+EDC scheme is $O(Q log Q + Q + Q log Q)$ and its worst-case running time can be represented as $O(Q log Q)$, where $Q$ denotes the number of time frames.
+
+The largest eigenvalue analysis is $O(\hat{d}Q)$, where $\hat{d}$ denotes the number of time frame under attack, according to Algorithm \ref{alg:alg01}.
+Subsequently, the eigen similarity analysis relies on EVD and cosine similarity analysis, which is $O(N^2)$, for $\hat{d}$ time frames, therefore the eigen similarity analysis is which is $O(\hat{d}(M^2N + N^3 + N^2))$ and yields an $O(N^3)$ upper bound on the worst-case running time for eigen similarity analysis.
+
+Therefore, the proposed framework is $O(N^3 + Q log Q + \hat{d}Q + N^3)$ and its worst-case running time is $O(N^3)$.
+The computational complexity of EVD is predominant in the framework, but the approach splits the data into time frames with period time $N$, which makes possible to limitate the growth of $N$ even for evaluations of cases with total time larger than $N$, reducing the impact caused by the computational complexity of EVD.
+
 \section{Conclusion and Future Works}
@@
 
-This paper models the network traffic as a signal processing formulation for applying to our framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
+This paper models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
 
@@
 
-Future research is directed to the fully automated blind attack detection, extending the evaluation to performance and accuracy analyis of our approach to well-known datasets of network attack detection.
+Future research is directed to the fully automated blind attack detection, extending the evaluation to performance and accuracy analyis of the approach to well-known datasets of network attack detection, and analysing the performance according to parameter configuration in order to better understand the impact of $N, M$ and $Q$ and identify their best configuration for obtaining gains into processing time and acuracy.
 

### 2016-05-26 fontes/mos-eigen-similarity/src/main/matlab/mosEigenSimilarity.m

@@ -181,6 +181,7 @@ for matrix = 1:numberOfMatrices
     %% for each period under attack, gets the times t and ports with attack
     q_attacks = {};
     for pc = 1:numPCs
+        
         X0 = dlmread([dataPath matrices{matrix} '/traffic/' num2str(pcs_q(pc)-1) '.txt'], '\t');  % the reference traffic withtou attack
         if (pcs_q(pc) == 3)
             [S0,E0,Vr0,Mr0] = eigencorrelation(X0);
@@ -189,7 +190,9 @@ for matrix = 1:numberOfMatrices
         end
         X = dlmread([dataPath matrices{matrix} '/traffic/' num2str(pcs_q(pc)) '.txt'], '\t');     % traffi of the period q under attack
         t_attacks = {};
+        
         for t = 1:periodsSize
+            
             Xc = cat(2,X0,X(:,1:t));
             if (pcs_q(pc) == 3)
                 [Sc,Ec,Vrc,Mrc] = eigencorrelation(Xc);
@@ -197,7 +200,7 @@ for matrix = 1:numberOfMatrices
                 [Sc,Ec,Vrc,Mrc] = eigencovariance(Xc);
             end
             
-            cosTheta = dot(Vr0(:,Mr0(2)),Vrc(:,Mrc(2)))/(norm(Vr0(:,Mr0(2)))*norm(Vrc(:,Mrc(2))));            
+            cosTheta = dot(Vr0(:,Mr0(2)),Vrc(:,Mrc(2))) / (norm(Vr0(:,Mr0(2))) * norm(Vrc(:,Mrc(2))));            
             cosTheta  = abs(cosTheta);
             %warning('Matrix = %s, Period(q) = %s, Time(t) = %s, cosTheta= %s',matrices{matrix},num2str(pcs_q(pc)),num2str(t),cosTheta);
             if(cosTheta < threshold)

### 2016-06-11 docs/papers/2016JGC/2016JGC.tex
adding storgrid paper

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn

\smartqed  % flush right qed marks, e.g. at end of proof

\usepackage[utf8x]{inputenc}

% reference and bib issues
\usepackage[numbers]{natbib}
\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
\hypersetup{breaklinks=true}
\urlstyle{same}
\Urlmuskip=0mu plus 1mu\relax

\begin{document}

\title{Offline Mode for Corporate Mobile Client Security Architecture}

\author{Tatiana Galibus				\and
        Thiago P. B. Vieira			\and
        	Edison P. de Freitas			\and
        	Robson O. Albuquerque		\and
        	João Paulo C. L da Costa		\and
        	Rafael T. de Souza Júnior	\and
        	V. Krasnoproshin1			\and
        	Anton Zaleski				\and
        	H.E.R.M. Vissia				\and
        	Giovanni del Galdo
}

\institute{Tatiana Galibus \at
              Belarusian State University, Minsk, 4, Nezavisimosti, 220030, Belarus
           \and
           Tatiana Galibus \and Thiago P. B. Vieira \and Robson O. Albuquerque \and João Paulo C. L da Costa \and Rafael T. de Souza Júnior  \at
              University of Brasilia, UnB, FT, ENE, CP: 4386, 70910-900, Brasília, Brazil
           \and
           Edison P. de Freitas \at
				Federal University of Rio Grande do Sul, UFRGS, INF, CP: 15064, 91501-970, Porto Alegre, Brazil
           \and
           João Paulo C. L da Costa \and Anton Zaleski \and H.E.R.M. Vissia \and 	Giovanni del Galdo \at
				Institute for Information Technology, Ilmenau University of Technology, Ilmenau, Germany
           \and
           João Paulo C. L da Costa \and Giovanni del Galdo \at
				Fraunhofer Institute for Integrated Circuits IIS, Erlangen, Germany \\
           \and
           ???? \at
				Byelex Multimedia Products BV Argon 1, 4751 XC Oud Gastel, The Netherlands
}

\date{Received: date / Accepted: date}

\maketitle

\begin{abstract}
Preventing data leakage on the mobile device is a crucial security problem. Therefore, additional control and protection measures should be taken for the confidential data on the mobile devices that leave the boundaries of the organization. In this scenario, such devices can be used on the offline mode for the convenience and traffic reduce or even for a particular characteristic of the app or of the business. This paper presents a novel approach to the security of the corporate mobile devices, in particular the offline mode. The important part of the presented approach is the conceptualization and the definition of the core methodology to solve the problem of offline mobile security, i.e. the protection of the confidential data in use when the mobile client is not connected to the corporate cloud. The protection of the sensitive data is provided by the combination of the cryptographic methods, such as AES file encryption, ABE authorization based both on user and share attributes, user key secret sharing (SS) based protection between the device and the user as well as MOS-based analytics methods to prevent the malicious user behavior. The proposed security architecture supports the basic mobile device protection principles: minimized traffic load and reduced communication with the cloud; usage of the light-weighted operations and an optimized combination of the security methods.
\end{abstract}

\keywords{ mobile security \and offline mode ABE \and MOS \and user behavior analysis \and secret sharing \and protected cloud }

\section{Introduction}
\label{sec_intro}
Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation. The basic features of the cloud environment is providing the elastic, on-demand and secure services for the end-users. While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}. Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized leading to obscure and incomplete security paradigms and solutions.
A common practice to provide a cloud solution with stable security is to use a specific type of cloud service, such as Cloud Access Security Broker (CASB) or Cloud Access Control (CAC). These services are specifically designed to bring security at a single access point and provide the coordination of the most important security measures. It is estimated by Gartner \cite{skyhigh2015} that such systems will be used by 85\%  of companies by 2020. Obviously, the reason for this is that the organization of security measures at a single control point allows to control and to monitor the level of cloud protection much more effectively. The basic features of the CASB are: discovery of cloud services, encryption along with tokenization for better search properties, access control, Data Loss Prevention (DLP) services, authentication, and auditing and alerting services \cite{lawson2015}. The protection of the confidential data, according to the standards of CASB deployment should be provided elsewhere, i.e. in transit and at the end-user, i.e. end-client, side \cite{campbell2015}.
Additional security issues and requirements have to be considered when mobile devices are actively used in corporate cloud environment \cite{yovel2014}. Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm. The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS). While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, such as CASB and CAC, the corporate mobile app is usually light-weighted and generally less protected. The protection scheme used on a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}. Therefore, encrypting files and generating keys on a mobile device is not considered a good solution. On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}. The common practice is the shadow user activities monitoring \cite{yovel2014}. However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode.
Suppose that a SME uses CASB in order to protect data at rest, i.e. while stored on the server-side, in transit, communicating with server, and, in use, while the client is connected to the network. In this case, when the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile app has to secure itself with its own limited resources. Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection. For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
Observing the above described landscape, this paper outlines the concept of the offline mobile client security. This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing \cite{galibus2007} and ABE encryption \cite{goyal2006attribute}. Moreover, our proposed approach proposal includes the usage of the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to highlight possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack. The key expiration period is safely incorporated into the proposed system solution in order to enhance security. Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors. The main target of our proposed solution is to provide a maximum defense at the minimal resource cost.
The paper is structured as follows. Section 2 analyzes the most common security problems in the mobile cloud environment and the proposed solutions for them related to the offline protection in the BYOD world. Section 3 outlines the basic processes and the module security infrastructure of the proposed solution as well as the outline of online mode security. Section 4 presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section 5 provides the schemes of the main security blocks including encryption and key usage hierarchies as well as explanations of the core security methods: SSS, ABE and MOS. Section 6 discusses the security proofs including the outline of adversary model, common threat scenarios and the proposed analytics to discover abnormal usage. Still in Section 6, complexity analysis and practical implementation results are included. Section 7 concludes the paper.

\section{Related works}
\label{sec_related_works}
The increasing usage of BYOD demands more sophisticated data protection services compared to ordinary computing environment. A common practice is to provide additional contextual methods apart from authentication, DLP services, and encryption, which can be at rest, in transit and in use \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. The contextual methods increase the security of the app at a maximum level with minimum resource requirements. The most commonly used are:
Using geolocation of the device to trace its usage;
Setting up the expiration period of an app;
Secured transfer politics between apps;
Restricting access to the corporate app;
Setting up the expiration period of app pass pin;
Setting up the counter of failed tries;
Restricted or prohibited offline access;
Logging and auditing.
Preventing data leakage on the mobile device is a crucial security problem. Therefore, administrators should take additional control and protection measures for the confidential data on the mobile devices that leave the boundaries of the organization. Generally, the most sensitive and confidential data should not be permitted to be transferred to the mobile device. However, what if the SME need to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reduce or even for a particular characteristic of the app or the business itself?
From the theoretical point of view of this problem, there are several surveys, whose common point is the mobile cloud computing as a rapidly developing paradigm that poses many security and complexity problems \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. An analysis of the new models of mobile cloud computing and new ways of using data storage services is presented in \cite{khan2015cloud, khan2013towards}. Commonly, all the models and protection schemes concentrate on the encryption properties and either perform the computations on their own \cite{zhao2010trusted, yang2011provable} or use the cloud provider to off-load the cryptographic operations \cite{itani2010energy, ren2011lightweight}. Obviously, it is natural the mobile device cannot handle all operations securely without the assistance of a cloud provider, due to resources constraint and battery limitation.
The necessity to use schemes that function without putting load on a provider arises when it is desired to make the device less dependent on the cloud provider, i.e. corporate app continues to provide the secure access to the sensitive data without connection to the network. As discussed in \cite{khan2015cloud}, all the currently known schemes of encryption, performing the computation, either use a cloud provider \cite{zhao2010trusted}, a third party trusted agent \cite{yang2011provable} or a combination of both \cite{itani2010energy}. In some cases, they concentrate on computational complexity without taking care of user privacy and security \cite{ren2011lightweight}.
Therefore, according to \cite{khan2015cloud, tenorio2013greatest, khan2013towards}, the state-of-the-art mobile cloud security models do not consider the problem of the offline security mode. In many cases the industrial providers of the mobile security API avoid the problem by completely forbidding the offline access to the protected app, i.e. SAP Mocona, which operates as a secure app wrapping layer \cite{Lelyveld2013}. Due to various constraints such as traffic load, travelling and ease of access, the SME business procedure may require such access. To the best of our knowledge, the offline mode security problem has not yet been deployed, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}. Therefore, the main concern of this proposal is the protection of the device and app in offline mode when the functions of data protection cannot be offloaded to a cloud or a trusted party.

\section{The client security architecture}
\label{sec_the_client_security_architecture}
The approach proposed in this work describes and implements the complete lifecycle of the mobile app security infrastructure. The security processes depend on the key expiry period, and are used to access the protected storage. Once the user keys expire, the user is requested to enter his valid credentials, i.e. PIN and password. The client app then sends the credentials to the server for verification. Once the new set of access keys is received, the user can access the protected files in the offline mode, without the access to the server. This means that no further communication with the server is needed until the key expires. The core set of functions and protocols can be divided into three sets of operations as shown in Figure 3.1.

Figure 3.1. Block diagram of the core set of functions and protocols of the mobile app security infrastructure
These operations function based on the mobile client architecture. This architecture consists of the modules of cryptographic functions, threat intelligence infrastructure, communication with server and storage. 
The cryptographic functions include the decryption (for AES and ABE) key restoration for SSS and EKE token generation procedure. While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode and the EKE generation and executions is used in the online mode. These operations are described in Sections 5.1-5.4 and are part of the Figure 3.1 related to online and offline mode operations.

The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and more elaborate MOS-inspired statistics analyzer. Functions such as alerting and deleting the expired key belong to this block as well. These functions are described in Section 5.5.

The communication with server includes the separate sender and receiver to check the user credentials and receiver new keys which acts based on the EKE protocol. The remaining data like access list and files can be sent via unprotected channels. Section 3.1 briefly discusses the communication with the server, but it is also an intention to make it a subject of our next work.

The storage is divided in two parts – temporary and permanent. While the files and file keys are stored permanently in order to reduce unnecessary traffic and the resource usage on client, the ABE keys and the key storage protection data are temporary. The unprotected files stored in the mobile client app memory are not the subject of the security architecture.

Figure 3.2 illustrates the client security architecture:

Figure 3.2: Mobile security infrastructure

\subsection{Online mode}
\label{sec_online_mode}
This section briefly outlines the main processes and functions in the online mode. The connection to the network allows verifying the user credentials with the help of the protected cloud server. The communication with the cloud is secured by means of the encrypted key exchange protocol \cite{hao2010j, hao2008password} and serves to receive the new set of keys, files and send the collected log information. The client app operations in the online mode are as represented in Figure 3.3.
Figure 3.3: Online mode operations
The communication with server is necessary once the key is expired. This allows optimizing the load on the network once the traffic is reduced to the discrete sessions. In such environment it is important to keep track of the time synchronization both on server and on client in parallel in order to avoid malicious user behavior messing up with time attacks.
As depicted in Figure 3.4, the operation of the online mode starts from the client app sending a token to server by means of EKE: DEV\_PASS+TIME+PASS+PIN in order to validate his identity and the identity of the user to the server. This value serves as a proof that the device and the user are the ones with which server communicated previously. After performing the verification the server sends the new calculated DEV\_PASS and TIME to the mobile app by means of established EKE session. Afterwards the server sends the new encrypted key set and encrypted files to the client app in the clear.
Figure 3.4: Online authorization workflow diagram.
This model proposes to use light-weighted EKE for the fast communication with server communication like J-PAKE \cite{hao2010j, hao2008password} or SIS-based EKE \cite{bogos2013primeless}. SIS is a public key encryption that avoids generating the big primes, thus it can be used for the secure key exchange \cite{bogos2013primeless,bellovin1992encrypted}. EKE is performed as soon as the key is expired and the client asks the user to synchronize with the server. Note that the EKE is not required to send the encrypted files or statistics. The current implementation of mobile app uses J-PAKE protocol. We refer to \cite{hao2010j, hao2008password, toorani2014security} for a complete description and analysis of the J-PAKE protocol.

\section{Offline mode: our proposed model and solution}
\label{sec_offline_mode}
This section proposes an open model of mechanisms for the mobile device protection in which the security is supported both in online and offline modes. Currently and to the best of our knowledge, the systems of mobile device protection follow a model where the protected mobile client can operate only when it is connected to the cloud, which is not always convenient for the end-user. The basic principles of the mobile device protection herein proposed are: 
Optimized communication with the cloud when the device does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
Optimized combination of the security mechanisms so that the mobile client does not need to perform complex computation like encryption and key generation due to its resource constraint;
Behavioral analysis of user's operations on mobile client, which can indicate anomalous or automated activities performed by attackers.
The most important security issues in the proposed model arise when the device goes to the offline mode and the user is still allowed to get the access to the protected SME documents. In this case, the server can neither monitor the user activity nor provide the protection methods. The security should be performed at the mobile client. Additionally, the maximum protection should be provided at the minimum resource cost. 
In the online mode the mobile device uses the secure communication with the server in order to verify the validity of user’s credentials. On the contrary, the offline protection model should be approached independently. Thus, the proposal is that the authentication/authorization mechanisms in the offline mode should utilize the derived proof of the user identity. The requirements for the proof are as follows:
The proof is derived from the previous session, in order to verify that the user is still authorized;
The proof should not give access to user password, i.e. it should not be stored on the device.
The proof should be temporary and have an expiration period;
It should not be directly used in communication with the server, in order to prevent the malicious user from mimicking;
It should be resilient to offline dictionary attacks;
It has to stay effective both in the scenarios of the malicious outsider and leakage of information when the formerly authorized user leaves the group.
Such proof cannot be stored on the client device, as it is not possible to guarantee its protection in the offline mode. Therefore, the most effective way to secure the proof on the offline mode without performing complex computation is to share this proof between the client and the user in a protected manner. In this case, there is no need to store the function of the user password. The additional argument against the traditional password verification is the necessity to check the PIN, which is very small, so the construction of valid one-way function resilient to the offline dictionary attacks is a difficult task.
The proposed offline mode architecture is represented in Figure 4.1.

Figure 4.1: Offline security architecture

The infrastructure in Figure 4.1 includes:
The protected storage: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files. The outline of the hybrid encryption scheme is presented in the Section 5.
Threat Intelligence Manager (TIM): most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode. Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios. Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior according to the scenarios in Section 6. Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones.
Key Management center: it includes the functions for maintaining the key expiry period and deleting the expired keys.

\subsection{Offline mode workflow}
\label{sec_offline_mode_workflow}
The user performs the following operations in the offline mode:

1. Enters the PASS and the 4-digit PIN;
2. Views the list of authorized files kept on the mobile storage;
3. Open the protected encrypted files;
4. Modify the protected files and save them in an unprotected storage.

The client app performs all the cryptographic calculations in the shadow. These calculations include the key storage, key restoring, decryption and showing the decrypted files in the client area. Note that the client app does not check the password or PIN validity as it does not hold the verification proof for the above. Also, the app keeps track of the user activities and the key expiry. The complete list of the app activities is presented in the Figure 4.2.

Figure 4.2: Offline mode operations chart

To finalize the description of the offline mode, Figure 4.3 shows the complete workflow of the proposed mobile application in the offline-mode.

Figure 4.3: Offline mode authorization workflow

The core cryptographic module of the implemented solution is based on the combination of AES-ABE-SSS methods, described in the Section 5. The key feature of the offline security is that the client app does not actually store any part of the user password to be verified. The client app combines its own key with the user share (PIN and password-derived) in order to restore the initial KEY\_SET\_KEY. If the user provides the wrong share the client will not be able to recognize it, but will decrypt the files incorrectly. Additionally, the password entering is tracked and too many tries in a short time are considered a threat.

\subsection{Offline Behavioral Analysis}
\label{sec_offline_behavioral_analysis}
The Log Analysis Center (LAC) is responsible for the behavioral analysis done by the mobile client, in order to identify malicious behaviors through MOS schemes applied to selected features, that shall be incrementally logged by TIM. Therefore, the selected features are modeled as matrices, composed of the number of occurrences of the selected feature by its location and by the occurrence time. Subsequently, the matrices are submitted to the MOS algorithm for threat intelligence, which is detailed in Subsection 5.4. 
The Figure 4.4 depicts the TIM and LAC for offline behavioral analysis.

Figure 4.4: TIM and LAC workflow for offline behavioral analysis

\section{The algorithms, key usage and data protection methods}
\label{sec_algorithms}
In the proposed approach, the kernel encryption scheme in the mobile device is a combination of several methods of security. The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption. The set of expiring ABE keys corresponding to the set of files accessible by user in encrypted with a single expiring AES key (KEY\_SET\_KEY). This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing. The encryption workflow is outlined in the Figure 5.1.

Figure 5.1: Encryption workflow

The key hierarchy can be explained by Figure 5.2 where the key represented in Figure 5.1 serves for decrypting the key for each level represented in Figure 5.2.

Figure 5.2: Encryption key hierarchy
This proposed encryption scheme is called as a hybrid encryption since it uses a combination of different encryption methods in order to support all requirements to the encryption scheme used in the proposed infrastructure. Then it has:
AES provides a fast encryption of the data files;
ABE permits to support the authorization policy on the encryption level;
3)  SSS allows protecting the key storage by avoiding the necessity to store the key proof.

In Subsections 5.1.-5.3, a detailed description of the building blocks of the hybrid encryption is provided.

\subsection{AES file encryption }
\label{sec_aes}
The protected files are encrypted on the server side with the secure 128/256 bit AES encryption, which is currently an industrial standard. Nevertheless, domain administrator can choose to encrypt the file storage with Blowfish or Serpent which also provide a high level of safety \cite{nist2000}. The symmetric encryption on a server side is performed in two steps. First there is a single AES key for preserving the server storage. Second, when the encrypted file is sent from the server to the mobile app or desktop device another unique AES key is generated. This key serves for the encryption on the client side. In other words, along with the encrypted file the user gets his own unique AES key for decrypting the files received. This key is protected by attribute-based public key encryption (ABE). The scheme of the server-side AES encryption is represented in Figure 5.3.

Figure 5.3: Server-side encryption
Usually, The AES encryption key is used as a session key. In our setting, it is not desirable to re-encrypt the files stored in the device unless there is a certain specific condition (for example, the user leaves the domain or a specific group in the domain (the group of users in the domain contains the users with the equal access rights, i.e. the same set of accessible file shares) and the file should not be accessible for this user anymore). This is why AES key in the presented notation is permanent and defined as FILE\_KEY. The randomly generated FILE\_KEY is unique for each file or set of files stored on a client.

\subsection{ABE encryption to protect the FILE\_KEY }
\label{sec_abe}
With each user session, the permanent FILE\_KEY (unique AES key) is re-encrypted. The set of FILE\_KEYs is protected with the corresponding ABE keys. The unique ABE model that we propose supports the attribute policy based on user groups and on file shares, i.e. the key attributes correspond both to the groups and the file shares.
The model supports the simple selective ABE scheme \cite{galibus2015cloud,galibus2014}. The selective scheme for attribute-based encryption is as follows: if at least one attribute in the set{ti}U is equal to the attribute in the set{ti}M, the corresponding user U can decrypt the text M. In other words, as soon as the user and share have one attribute in common – the user can get access to the share. The components of the ABE encryption are:

Master-key (MK) which is kept safely on server and accessible only for the domain administrator and is given by
MK = (t1, t2,...,tn, y),
where the values ti are randomly selected from the huge group Zp. They are the private keys corresponding to the group attributes. Note, that this is different from the usual PK encryption: the private keys are controlled by the admin and not by the users.

Public key (PK) depends on the master key values and is kept in the clear allowing users to access the information:

PK = (gt1, gt2,...,gtn, e(g,g)y),
Here e(g,g) is the bilinear pairing function corresponding to an elliptic curve.

Secret user KEY\_SET depends on his attribute set. Here each Di (GROUP\_KEY) serves for decryption of the data of a single group of users, for example, related to some project:
{ti}UD={Di=gywti}

Encrypted text M, in our context, M = FILE\_KEY, or the permanent AES symmetric key, which allows to avoid the file re-encryption. Encryption procedure is multiplication. The set of the public keys Ei (PUBLIC\_SHARE\_KEY) corresponding to the set of groups able to access the text is kept along with the encrypted text E:

E=Me(g,g)ys,{ Ei =gtis/w }, i{ti}M

Decryption is division:
M=E/Ys

In order to perform this operation the user needs the pair of private key Di and public key Ei corresponding to the attribute ti :
Ys=e(g,g)ys=e(Ei,Di)=e(gywti,gtis/w)=e(g,g)ys

The result of decryption is the FILE\_KEY - the symmetric AES key that permits to decrypt the contents of protected file.

\subsection{Secret sharing scheme to protect the key storage }
\label{sec_secret}

The attribute-based private keys Di should be protected while being stored in the device memory. Therefore, server encrypts the set of Di with a single AES key before sending it to the user device. This AES key (master key) is denoted by the value KEY\_SET\_KEY in our notation. KEY\_SET\_KEY is a secret value and it is split by the secure method of polynomial modular secret sharing \cite{galibus2015mobile, galibus2008some} into the set of 4 shares:

KEY\_SET\_KEY = PASS+PIN+TIME+DEV\_PASS.

Since the underlying sharing scheme is perfect \cite{galibus2008some}, the adversary cannot get any information of the KEY\_SET\_KEY unless he possessing all 4 key parts. Here the values PASS and PIN are predefined similar to the construction in \cite{galibus2015mobile}.

The proposed authentication system is based on the shared storing of the user key. Also, the device acts as a dealer in the SSS. Using the SSS ensures that the key can only be accessed by an authenticated user. The participants of the (2, 2)-threshold SSS are the user and device. The user share s1(x) is computed based on the PIN and the PASS entered by the user. Additionally, the current time value TIME is used in the calculation of the share. Let s(x)=d and s1(x)=f(PIN+PASS+TIME), 
Where f is a one-way function that transforms the data into the string of the desired length:

According to the CRT:

Thus calculated DEV\_PASS is written to the permanent device memory. The user share is not saved. Otherwise, it would allow an attacker to locally validate the restored private key.

\section{MOS for Threat Intelligence}
\label{sec_mos}
In the context of anomaly-based schemes for attack detection, the proposed behavioral analysis approach applies signal processing techniques, such as Principal Component Analysis and Model Order Selection schemes \cite{tenorio2013greatest}, for automatic identification of attacks or malicious behaviors.

Model Order Selection is an effective signal processing technique for several applications, allowing separating the only noise components from the principal components applying a rank reduction of the data. Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. MOS for threat intelligence requires that the target features, which can be user operations, are modeled as a matrix composed by the number of occurrences by location and time. For example, one element can represent the number of file readings on folder m during the minute n.

The MOS schemes can rely on correlation or covariance analysis, where the correlation analysis identifies anomalies on sparse or subtle number of file operations, while the MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period.

Classical approaches to model order selection require the computation of the sample covariance matrix Vxx and of its eigenvalues, obtained from the measurement matrix X of the zero mean samples given by

The set of obtained vectors zm composes the zero mean matrix Z, then the covariance matrix Vxx can be calculated as follows

Vxx =1NZZT

For MOS based on correlation analysis, instead of calculate the zero mean samples, to obtain the correlation matrix Rxx , it is required, for each variable, to make the standard deviation unitary as follows

The set of vectors sm composes the matrix S, then the correlation matrix Rxx can be calculated via 
. Once the Vxx or Rxx have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to Vxx and Rxx as a matrix C. Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to , in order to obtain the vector of eigenvalues e, as following:

The eigenvalues should be sorted in descending order, as defined by , to make possible the selection of the first eigenvalue in the obtained sequence, represented by, which is the largest eigenvalue of the data evaluated for attack detection.

The process of obtaining the and the matrix C(t), finding the largest eigenvalue for each t-th time frame, should be repeated until t = T, in order to obtain the largest eigenvalue of all time frames, as presented by

E =

Since   , then the first line of the matrix E contains the largest eigenvalues of each t-th time frame, which is the expected input for MOS schemes and can be expressed as 
emax=E{:,1}

Once obtained the largest eigenvalues of each q-th time frame, it is possible to apply a selected MOS scheme to estimate the model order $\hat{d}$, which is the estimated number of time frames with malicious behavior. Therefore, emax is used as input parameter for MOS schemes, according to the equation.

Note that some MOS schemes may also require the amount of time that compose a time frame, such as . For more information about MOS schemes, interested readers are referred to \cite{da2009comparison}.

\section{Results and analysis}
\label{sec_mos}

... add some text

\subsection{Security analysis}
\label{sec_sec_analysis}
The security analysis of the proposed model was performed both from the point of implemented cryptographic mechanisms and from the user behavioral analysis. Two common attack scenarios were analyzed. First, the malicious outsider trying to infect or steal the important files. Second, the malicious expired user trying to steal the important files.

\subsubsection{Adversary model}
\label{sec_adversary_model}
The proposal is to use a restricted adversary model. In other words, the actions of the adversary are predicted within the outlines model and scenarios. For example, it is not possible to guarantee the security of the decrypted document once the user copies its content to another file. In the analyzed scenarios, it is assumed that the adversary can:

Steal the user keys one-by-one while the app performs decryption, if he is outsider;
Use expired key to decrypt the files, if he is insider;
Does not have possibility to perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the client app;
Steal the encrypted files but not the decrypted ones.

\subsubsection{Security analysis}
\label{sec_sec_analysis}
Offline mode
In the offline mode the client does not store the user password (see section 3 for details), i.e. no information about the password leaks, and therefore there is no possibility for the malicious user to check if the password he is trying to enter is correct or not. The only possible scenario for the information leakage in the case of the malicious outsider is if:
The hacker steals all the parts of KEY\_SET\_KEY;
The hacker steals the KEY\_SET;
The hacker try the brute force offline dictionary attack on all the previous values, they have to belong to ONE TIME SESSION (the values of ”a” and ”b” belong to one period of time i.e. TIME, DEV\_PASS, KEY\_SET);
Steal the permanent FILE\_KEY (this is protected by the KEY\_SET);
Steal the file and try to decrypt it with offline dictionary attacks.
        Still, the hacker has to get 4 values: TIME, DEV\_PASS, KEY\_SET, FILE\_KEY - from one session. At the same time he should try the offline dictionary attack on PIN+PASS. Moreover, the 4 values provide access only to 1 single file. So practically, it is very difficult to perform such attack due to the key expiration period. In accordance with the conditions 1) and 3) of our adversary model, the above actions are not possible for such adversary in the conditions presented.

The temporary nature of all parameters obliges the user to connect to server when necessary and prevents the malicious actions from the user side. The only possible scenario when malicious or expired user wishes to prolong his old credentials is:
He has to steal the DEV\_PASS and TIME synchronized with his credentials;
He has to be able to combine;
He has to steal the KEY\_SET synchronized with his credentials;
He has to steal the protected FILE\_KEY (also synchronized) and encrypted file;
He has to do everything without the client (because the client checks the TIME and renews the PUBLIC\_SHARE\_KEY).
Basically, for a malicious user there is practically no way to use the client with the old keys. The fact that the client does not contain any data to be checked or validated does not prevent the user from seeing the contents of decrypted files, but with the wrong password the decrypted files will be different from the original ones.

The client still has to count the tries (to prevent the hacker to perform brute force attack) within one session. In accordance with the conditions 2) and 4) of the adversary model, the above actions are not possible for such adversary.
Online mode
In the online mode the protection of the mobile system is backed up by the communication with server. The key expiry period and the usage of J-PAKE \cite{hao2010j,hao2008password} guarantees that the sensitive data sent over the network cannot leak. The detailed security analysis of J-PAKE is presented in \cite{toorani2014security}. The data sent in the clear i.e. encrypted files and the public keys of ABE does not need to be protected additionally as the AES and ABE guarantee its security.

\subsubsection{Common threat scenarios}
\label{sec_common}
This section provides the detailed description of the common scenarios in which the log and behavioral analysis is provided. The behavioral analysis can help to keep the user or administrator informed of the threat and take actions, as well as it can be useful in order to implement threat preventions or reactive actions to avoid threat propagation.

\subsubsection{An attacker uses a valid password to perform operations on a bulk of files}
\label{sec_attacker_valid}
The session time defines the period when operations can be performed until the next session renewing. During this period, it is still necessary to identify attacks and malicious behavior on file operations, in order to avoid fast attacks to perform unauthorized access to information or data modification. Some attacks present behavioral patterns based on abrupt number of operations, such as the ransomware attack, which is a growing attack \cite{McAfee2015} that blocks the access to valuable resources and requires a payment in order to unblock the content. The access to the resources can be blocked by the attacker through some techniques, when the content is encrypted by the attacker, the ransomware attack can be called cryptoransomware \cite{kaspersky2014}.
MOS schemes based on covariance analysis are effective to reveal abrupt changing of behaviors over time \cite{tenorio2013greatest}, making possible to identify intense malicious behaviors on offline mode of mobile clients, such in case of ransomware attack or bulk access to sensitive data.
The large number of operations over time is a well-known pattern of some attacks, due to the efforts on security measures to make the attacks infeasible over time. In this context, the operations can also be evaluated in contrast to the estimated required time for operations done by legitimate behaviors, such as the evaluation of the mean time between operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities.
Sparse or subtle file operations, with low number of operations distributed over different files or directories, during short period of time can indicate anomalies in contrast to the required time for legitimate directory navigation. MOS and correlation analysis can be suitable if applied to evaluate the time and location of operations, in order to identify unreachable navigation, if compared to legitimate navigation
The MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period. The eigenvalue analysis based on the correlation matrix highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
This threat scenario, where an attacker uses a valid password and session to perform operations on a bulk of files, can have its steps described as:
The hacker has access to the mobile client and is able to perform operations;
The session time is valid;
The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
The client incrementally append each operation attempt time into the logging;
The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time;

\subsubsection{Usage of expired password to perform unauthorized operations }
\label{sec_attacker_expired}
In the offline mode, the session time is used to restrict the operations during a specified period, although it is possible to manipulate the current time in mobile clients, to emulate a period in which the session was valid. The log analysis by MOS can deal with this kind of threat, through the incremental logging of the time when each operation was performed, followed by the behavioral evaluation of operations over time. 
The incremental logging assumes that new logged operations shall have equal or bigger time than the last logged operation, the violation of this rule means that the system is out of sync and indicates a malicious behavior. Additionally, a large amount or sparse operation performed at the same time, or during a short period, can indicate the use of backtrack techniques to maintain a valid session during necessary time to perform an attack. Massive, subtle or sparse malicious operation performed during a valid session time can be identified by MOS schemes based on covariance and correlation analysis.
Applying MOS to the analysis of the time between user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session. The MOS based on correlation analysis identifies anomalies on sparse or subtle number of file operations, since the eigenvalue analysis based on the correlation matrix highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations. Consequently, the result of the eigenvalue analysis is applied to MOS schemes, in order to identify the occurrence of malicious behaviors during the valid session.
The MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period. The eigenvalue analysis based on correlation matrix highlights massive or concentrated operations over time or location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
This threat scenario, where the attacker uses expired password to perform unauthorized operations, can have its steps described as:
The hacker steals the operating system;
The hacker modifies the time of the operating system to a period when the session was valid;
The hacker has access to the mobile client and is able to perform operations;
The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
The client incrementally append each operation attempt time into the logging;
The client verifies if one logged time is older than the last operation time. If it is true, the MOS module classifies the evaluated operation as malicious;
The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations;
The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time;

\subsubsection{Data Modeling for Behavioral Analysis }
\label{sec_data}
 MOS schemes are used in order to identify anomalous behavior that can indicate an attack and be used to prevent or avoid attack propagation. Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to MOS schemes, according to described in Section 5.4.
The selected features shall be modeled as matrices which represents a signal superposition containing noise, legitimate and malicious behavior \cite{tenorio2013greatest}, grouped into time frames , where M defines the decomposition of a selected feature, N defines the time decomposition and represents the number of occurrences of the feature m during the time n.
In offline mode, the user is still allowed to get access to operations that do not require communication with the server side. These operations and their selected features shall be incrementally logged by the mobile client, in order to be evaluated to identify malicious behaviors. This work proposes to evaluate the following features:
File Access (Time and File System Location), i.e. data access to selected files in offline mode, accessing the data stored on the mobile client. The file access feature can be decomposed into:
file decryption, 
decrypted file reading 
decrypted file execution, 
Therefore, it is necessary to generate three matrices for malicious behaviors analysis, such as: 
massive file access, which can reveal data leakage and be identified by MOS schemes based on covariance analysis; 
low file access into several folders, characterized by sparse operations that can reveal unreachable navigation performed by automated file accesses in order to avoid the massive file access characterization. 
Malicious sparse file accesses can be identified by MOS schemes base correlation analysis.
File Update (Time and File System Location), i.e. writing operations into selected files in offline mode, writing the data stored on the mobile client. The update feature can be decomposed into:
 file encryption 
decrypted file writing, 
Therefore, it is necessary to generate two matrices for malicious behaviors analysis, such as: 
massive file update, which can reveal ransomware or similar attacks and be identified by MOS schemes based on covariance analysis; 
low number of file update into several folders, characterized by sparse operations that can reveal unreachable navigation performed by automated file accesses in order to avoid the massive file access characterization. Malicious sparse file accesses can be identified by MOS schemes base correlation analysis.
File Download (Start Time, End Time and File System Location), i.e download requests in online mode, evaluated by the mobile client. The file download feature shall be modeled as the matrix of number downloads by file location over time, in order to perform malicious behaviors analysis, such as:
massive data leakage or similar attacks, which can be identified by MOS schemes based on covariance analysis; 
low number of file download from several folders, characterized by sparse operations, which can reveal unreachable navigation performed by automated file download in order to avoid the massive file download characterization. Malicious sparse file download can be identified by MOS schemes base correlation analysis.
File Upload (Start Time, End Time and File System Location), i.e upload requests in online mode, evaluated by the mobile client. The file upload feature can reveal attempts of ransomware or similar attacks and be identified by MOS schemes based on covariance analysis. Therefore, it is necessary model the matrix of number uploads by file location over time, in order to perform malicious behaviors analysis, such as:
massive file upload, similar to ransomware attack, which can be identified by MOS schemes based on covariance analysis; 
low number of file upload to several folders, characterized by sparse operations, which can reveal unreachable navigation performed by automated file upload in order to avoid the massive file upload characterization. Malicious sparse file upload can be identified by MOS schemes base correlation analysis.

These features provide the necessary information to apply MOS schemes for evaluating malicious behaviors according to the following scenarios and steps:
The hacker accesses the mobile client during a valid session and read/write few files at several folders:
The hacker steals the device;
The hacker changes the device to offline mode;
The hacker accesses the mobile client;
The current session is valid;
The hacker read/writes few files at several folders;
MOS based on correlation analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
MOS indicates the number of time frames with anomalous behavior;
The hacker accesses the mobile client during a valid session and read/write several files at few folders:
The hacker steals the device;
The hacker changes the device to offline mode;
The hacker accesses the mobile client;
The current session is valid;
The hacker read/writes several files at few folders
MOS based on covariance analysis is applied to the matrix of number of file operations, grouped by the file system location over time
MOS indicates the number of time frames with anomalous behavior;
The hacker emulates a valid session and read/write few files at several folders: 
The hacker steals the device;
The hacker changes the device to offline mode;
The current session is expired;
The hacker changes the device time to a valid period for the current session;
The hacker accesses the mobile client;
The current session is valid;
The hacker perform sparse operations during the valid period;
MOS based on correlation analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
MOS indicates the number of time frames with anomalous behavior;
The hacker emulates a valid session and read/write several files at few folders: 
The hacker steals the device;
The hacker changes the device to offline mode;
The current session is expired;
The hacker changes the device time to a valid period for the current session;
The hacker accesses the mobile client;
The current session is valid;
The hacker perform bulk operations during the valid session;
MOS based on covariance analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
MOS indicates the number of time frames with anomalous behavior;

\subsection{Complexity analysis }
\label{sec_complex}

The complexity analysis depends on the operation that the user and the mobile app perform in order to keep the device protected. In the offline mode the client app performs the following actions:

Combine the PASS+PIN+TIME+DEV\_PASS=KEY\_SET\_KEY   ->  SSS secret restoring;
Decrypt the KEY\_SET with the KEY\_SET\_KEY -> symmetric 128/256 AES decryption;
Select the SHARE\_KEY from the KEY\_SET -> no calculation;
Decrypt the FILE\_KEY with the SHARE\_KEY  -> ABE decryption;
Decrypt the file with the FILE\_KEY ->symmetric 128/256 AES d -> ecryption;
Modify the TIME periodically -> timer;
Count the tries within the TIME ->count;
Modify or delete PUBLIC\_SHARE\_KEY-> no calculation;
Malicious behavior detection -> eigenvalues calculation.
It is easy to check that in the offline mode the client app does not perform complex calculations and does not use the resources extensively due to the fact that the initial key is shared and the client app only performs decryption, which is not a time-consuming operation. Thus, the proposal supports the concept of the light-weighted client, i.e. the most consuming operations are ABE and AES decryption.

Similarly, in the online mode the client app does not perform resource-consuming operations apart from J-PAKE construction which is used to renew the user/app master key (DEV\_PASS):

Generates keys for J-PAKE -> this can be a resource-consuming operation. In the future we suggest to replace J-PAKE with SIS-based PKE \cite{bogos2013primeless,bellovin1992encrypted};
Sends and receives data in the clear;
Performs operations 1)-7);
Performs the operation 8) and sends the log data for server side analysis.

The proposed concept of mobile client security has been implemented in the Storgrid protected cloud environment \cite{storgrid2016}. Therefore, the approach is correlated with the practical usability requirements:  the corporate user continues to use the mobile storage app in offline and does not need to reload the files every time the key is renewed. This methodology can be used in other mobile apps. The common advantage is that the mobile client performs the operations both in the offline and online mode and uses the key expiry and ABE to protect the privacy of the corporate data. The table below presents the results of mobile app security workflow testing. Table 1 demonstrates that the device decrypts the data with the acceptable speed. One can compare the speed of ABE/AES decryption (keys and actual data) and the speed of the whole decryption process.

Table 1. Speed of decryption on a client device.

Table 2 demonstrates the results of testing the speed of the mobile device for receiving and displaying the list of files. The app was tested for displaying up to 1000 files. Around 100 requests were generated from 10 threads.

Table 2. Speed of receiving the list of files

The device receives and displays the list of files depending on the user attributes with acceptable average speed and the proposed protection scheme is usable. Table3 demonstrates the results of performance testing of the keys generation procedure on server:

Table 3. Speed of generating the user keys.

It is easy to observe that the performance does not decrease drastically with the increase of particular user shares. Each key sample generated contains the domain member code, key data and the sharing key value for each file sharing code. We may conclude that the proposed client is not overloaded with calculations due to the carefully selected mathematical operations. It can be successfully used and provides acceptable level of security.
The log analysis of the Log Analysis Center has been implemented and evaluated for offline anomaly detection in mobile clients, making it possible to apply anomaly detection techniques in a lightweight fashion, considering low processing requirements for deal with the resource constraints of mobile clients. The evaluation considered the required processing time for anomaly detection from log analysis, measuring the data modelling time, eigenvalue decomposition time and the required time for the EDC MOS scheme execution. The experiments were performed in two mobile devices, with variations of log size and window size. 
Table 4 presents the data modelling time and the processing time of eigenvalues decomposition and MOS calculations applied to anomaly detection from user operation logs of Storgrid mobile client. The results show that the lower window size leads to the larger eigenvalue decomposition time, but the largest eigenvalue decomposition time, which was 421 milliseconds, highlights an acceptable speed even for the worst evaluated scenario.

Table 4. Data Modelling and Eigenvalue (Correlation and Covariance) Decomposition Time

Table 5 presents the processing time of EDC MOS calculations applied to anomaly detection from user operation logs of Storgrid mobile client. It is possible to observe that the processing time increases with the window size decreasing, similar to the results for eigenvalue decomposition time.

Table 5. EDC MOS scheme processing time for anomaly detection

The longest processing time measured is lower than 200 milliseconds, even considering window size of 10 minutes or processing 6 MB of user operation log, what represents a acceptable processing time for anomaly detection in mobile devices.

\section{Conclusion and future work }
\label{sec_conclusion}
An important security issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile devices. Once a mobile device releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device. Aware of this problem and its importance, this work presented a proposal to address the offline mobile security problem combining a different cryptographic methods. Moreover, this proposed approach also prevents malicious user behavior by applying a MOS-based analytic method. As prove of concept, a fully working mobile application was developed to test the proposed security solution and acquired results provide evidence that besides achieving the desired security features, the solution also has positive results in terms of performance. This fact is due to the usage of lightweight operations and the optimized combination of the selected security methods. 
Future works in the area can further explore enhancements in the analytics methods as well as to extended the approach to be used by mobile devices with even more severe resources constraints.

Our approach is a practical application to be used in the corporate mobile environment. It is implemented as a fully working mobile app. WE suggest that it can be used for any type of enterprise. Also, part of concept is sed in the big data apps. We plan to enhance the analytics and implement the effective protocols.

\begin{acknowledgements}
The authors thank the Brazilian research and innovation agencies FAPDF (Research Support Foundation of the Federal District), FINEP (Agreement RENASIC / PROTO 01.12.0555.00), CAPES and CNPq under the FORTE Project - CAPES Forensic Sciences Announcement 25/2014, under the productivity grant number 303905 / 2014-0, under the PVE project number 88881.030392 / 2013-01 for their financial support on this research, and under the project number 207644/2015-2 for the PDE scholarship in the context of the program CSF- Aerospace Technology.\end{acknowledgements}

% BibTeX users please use one of
\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}     % mathematics and physical sciences
%\bibliographystyle{spphys}      % APS-like style for physics
\bibliography{references}   		% name your BibTeX data base

\end{document}

### 2016-06-11 docs/papers/2016JNCA/2016JNCA.tex
adding storgrid paper

@@
 
-This subsection discuss the computational complexity of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
+This subsection discusses the computational complexity of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
 

### 2016-06-11 docs/papers/2016JGC/2016JGC.tex
figure redesign

@@
 
+% text
 \usepackage[utf8x]{inputenc}
+
+% figure
+\usepackage{graphicx}
+\usepackage{epstopdf}
 
@@
 
-\keywords{ mobile security \and offline mode ABE \and MOS \and user behavior analysis \and secret sharing \and protected cloud }
+\keywords{ mobile security \and offline mode ABE \and Model Order Selection (MOS) \and user behavior analysis \and secret sharing \and protected cloud }
 
@@
 This means that no further communication with the server is needed until the key expires.
-The core set of functions and protocols can be divided into three sets of operations as shown in Figure 3.1.
+The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:1}.
 
-Figure 3.1. Block diagram of the core set of functions and protocols of the mobile app security infrastructure
-These operations function based on the mobile client architecture.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=4cm]{figures/coresetoffunctionsandprotocols.eps}
+	\caption{Block diagram of the core set of functions and protocols of the mobile app security infrastructure}
+	\label{fig:1}
+\end{figure}
+
+These operations function based on the mobile client architecture, which is depicted in Figure \ref{fig:2}.
 This architecture consists of the modules of cryptographic functions, threat intelligence infrastructure, communication with server and storage.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8.7cm]{figures/mobileclientarchitecture.eps}
+	\caption{The Mobile Client Architecture}
+	\label{fig:2}
+\end{figure}
+
 The cryptographic functions include the decryption (for AES and ABE) key restoration for SSS and EKE token generation procedure.
 While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode and the EKE generation and executions is used in the online mode.
-These operations are described in Sections 5.1-5.4 and are part of the Figure 3.1 related to online and offline mode operations.
+These operations are described in Subsections \ref{sec_aes}-\ref{sec_secret} and are part of the Figure \ref{fig:1}, related to online and offline mode operations.
 
@@
 Functions such as alerting and deleting the expired key belong to this block as well.
-These functions are described in Section 5.5.
+These functions are described in Subsection \ref{sec_mos}.
 
@@
 The remaining data like access list and files can be sent via unprotected channels.
-Section 3.1 briefly discusses the communication with the server, but it is also an intention to make it a subject of our next work.
+Subsection \ref{sec_online_mode} briefly discusses the communication with the server, but it is also an intention to make it a subject of our next work.
 
@@
 
-Figure 3.2 illustrates the client security architecture:
-
-Figure 3.2: Mobile security infrastructure
-
 \subsection{Online mode}
 \label{sec_online_mode}
-This section briefly outlines the main processes and functions in the online mode.
+This subsection briefly outlines the main processes and functions in the online mode.
 The connection to the network allows verifying the user credentials with the help of the protected cloud server.
@@
 The client app operations in the online mode are as represented in Figure 3.3.
-Figure 3.3: Online mode operations
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8.8cm]{figures/onlinemodeoperationss.eps}
+	\caption{Online Mode Operations}
+	\label{fig:3}
+\end{figure}
+
 The communication with server is necessary once the key is expired.
@@
 In such environment it is important to keep track of the time synchronization both on server and on client in parallel in order to avoid malicious user behavior messing up with time attacks.
-As depicted in Figure 3.4, the operation of the online mode starts from the client app sending a token to server by means of EKE: DEV\_PASS+TIME+PASS+PIN in order to validate his identity and the identity of the user to the server.
+
+As depicted in Figure \ref{fig:4}, the operation of the online mode starts from the client app sending a token to server by means of EKE: DEV\_PASS + TIME + PASS + PIN in order to validate his identity and the identity of the user to the server.
 This value serves as a proof that the device and the user are the ones with which server communicated previously.
@@
 Afterwards the server sends the new encrypted key set and encrypted files to the client app in the clear.
-Figure 3.4: Online authorization workflow diagram.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8.7cm]{figures/OnlineAuthorizationWorkflow.eps}
+	\caption{Online Authorization Workflow}
+	\label{fig:4}
+\end{figure}
+
 This model proposes to use light-weighted EKE for the fast communication with server communication like J-PAKE \cite{hao2010j, hao2008password} or SIS-based EKE \cite{bogos2013primeless}.
@@
 The basic principles of the mobile device protection herein proposed are:
-Optimized communication with the cloud when the device does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
-Optimized combination of the security mechanisms so that the mobile client does not need to perform complex computation like encryption and key generation due to its resource constraint;
-Behavioral analysis of user's operations on mobile client, which can indicate anomalous or automated activities performed by attackers.
+
+\begin{enumerate}
+	\item Optimized communication with the cloud when the device does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
+	\item Optimized combination of the security mechanisms so that the mobile client does not need to perform complex computation like encryption and key generation due to its resource constraint;
+	\item Behavioral analysis of user's operations on mobile client, which can indicate anomalous or automated activities performed by attackers.
+\end{enumerate}
+
 The most important security issues in the proposed model arise when the device goes to the offline mode and the user is still allowed to get the access to the protected SME documents.
@@
 Additionally, the maximum protection should be provided at the minimum resource cost.
+
 In the online mode the mobile device uses the secure communication with the server in order to verify the validity of user’s credentials.
@@
 The requirements for the proof are as follows:
-The proof is derived from the previous session, in order to verify that the user is still authorized;
-The proof should not give access to user password, i.e. it should not be stored on the device.
-The proof should be temporary and have an expiration period;
-It should not be directly used in communication with the server, in order to prevent the malicious user from mimicking;
-It should be resilient to offline dictionary attacks;
-It has to stay effective both in the scenarios of the malicious outsider and leakage of information when the formerly authorized user leaves the group.
+
+\begin{enumerate}
+	\item The proof is derived from the previous session, in order to verify that the user is still authorized;
+	\item The proof should not give access to user password, i.e. it should not be stored on the device;
+	\item The proof should be temporary and have an expiration period;
+	\item It should not be directly used in communication with the server, in order to prevent the malicious user from mimicking;
+	\item It should be resilient to offline dictionary attacks;
+	\item It has to stay effective both in the scenarios of the malicious outsider and leakage of information when the formerly authorized user leaves the group.
+\end{enumerate}
+
 Such proof cannot be stored on the client device, as it is not possible to guarantee its protection in the offline mode.
@@
 The additional argument against the traditional password verification is the necessity to check the PIN, which is very small, so the construction of valid one-way function resilient to the offline dictionary attacks is a difficult task.
-The proposed offline mode architecture is represented in Figure 4.1.
 
-Figure 4.1: Offline security architecture
+The proposed offline mode architecture is represented in Figure \ref{fig:5}.
 
-The infrastructure in Figure 4.1 includes:
-The protected storage: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files.
-The outline of the hybrid encryption scheme is presented in the Section 5.
-Threat Intelligence Manager (TIM): most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode.
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8.7cm]{figures/offlinesecurityarchitecture.eps}
+	\caption{Online Authorization Workflow}
+	\label{fig:5}
+\end{figure}
+
+The infrastructure in Figure \ref{fig:5} includes:
+
+\begin{enumerate}
+	\item \textbf{The protected storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files.
+The outline of the hybrid encryption scheme is presented in the Section \ref{sec_algorithms}.
+	\item \textbf{Threat Intelligence Manager (TIM)}: most attacks incur into significant variation on the legitimate behavior of information systems, or they adopt well-known patterns that can be easily detected by monitoring the system in the case of the offline mode.
 Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios.
-Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior according to the scenarios in Section 6.
+Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior according to the scenarios in Section \ref{sec_results}.
 Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones.
-Key Management center: it includes the functions for maintaining the key expiry period and deleting the expired keys.
+	\item \textbf{Key Management center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
+\end{enumerate}
 
@@
 \label{sec_offline_mode_workflow}
+
 The user performs the following operations in the offline mode:
 
-1. Enters the PASS and the 4-digit PIN;
-2. Views the list of authorized files kept on the mobile storage;
-3. Open the protected encrypted files;
-4. Modify the protected files and save them in an unprotected storage.
+\begin{enumerate}
+	\item Enters the PASS and the 4-digit PIN;
+	\item Views the list of authorized files kept on the mobile storage;
+	\item Open the protected encrypted files;
+	\item Modify the protected files and save them in an unprotected storage.
+\end{enumerate}
 
@@
 Also, the app keeps track of the user activities and the key expiry.
-The complete list of the app activities is presented in the Figure 4.2.
+The complete list of the app activities is presented in the Figure \ref{fig:6}.
 
-Figure 4.2: Offline mode operations chart
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8.8cm]{figures/offlinemodeoperationss.eps}
+	\caption{Offline Mode Operations}
+	\label{fig:6}
+\end{figure}
 
-To finalize the description of the offline mode, Figure 4.3 shows the complete workflow of the proposed mobile application in the offline-mode.
+To finalize the description of the offline mode, Figure \ref{fig:7} shows the complete workflow of the proposed mobile application in the offline-mode.
 
-Figure 4.3: Offline mode authorization workflow
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8.7cm]{figures/OfflineAuthorizationWorkflow.eps}
+	\caption{Offline mode authorization workflow}
+	\label{fig:7}
+\end{figure}
 
-The core cryptographic module of the implemented solution is based on the combination of AES-ABE-SSS methods, described in the Section 5.
+The core cryptographic module of the implemented solution is based on the combination of AES-ABE-SSS methods, described in the Section \ref{sec_algorithms}.
 The key feature of the offline security is that the client app does not actually store any part of the user password to be verified.
@@
 Therefore, the selected features are modeled as matrices, composed of the number of occurrences of the selected feature by its location and by the occurrence time.
-Subsequently, the matrices are submitted to the MOS algorithm for threat intelligence, which is detailed in Subsection 5.4.
-The Figure 4.4 depicts the TIM and LAC for offline behavioral analysis.
+Subsequently, the matrices are submitted to the MOS algorithm for threat intelligence, which is detailed in Subsection \ref{sec_mos}.
+The Figure \ref{fig:8} depicts the TIM and LAC for offline behavioral analysis.
 
-Figure 4.4: TIM and LAC workflow for offline behavioral analysis
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8.7cm]{figures/mosworkflow.eps}
+	\caption{TIM and LAC workflow for offline behavioral analysis}
+	\label{fig:8}
+\end{figure}
 
@@
 
-In Subsections 5.1.-5.3, a detailed description of the building blocks of the hybrid encryption is provided.
+In Subsections \ref{sec_aes}-\ref{sec_secret}, a detailed description of the building blocks of the hybrid encryption is provided.
 
@@
 
-\section{MOS for Threat Intelligence}
+\subsection{MOS for Threat Intelligence}
 \label{sec_mos}
@@
 \section{Results and analysis}
-\label{sec_mos}
+\label{sec_results}
 

### 2016-06-12 docs/papers/2016JGC/2016JGC.tex
new improvements

@@
 	\centering
-	\includegraphics[width=4cm]{figures/coresetoffunctionsandprotocols.eps}
-	\caption{Block diagram of the core set of functions and protocols of the mobile app security infrastructure}
+	\includegraphics[width=6cm]{figures/coresetoffunctionsandprotocols.eps}
+	\caption{The core set of functions and protocols of the mobile app security infrastructure}
 	\label{fig:1}
@@
 	\centering
-	\includegraphics[width=8.7cm]{figures/mobileclientarchitecture.eps}
+	\includegraphics[width=8.8cm]{figures/mobileclientarchitecture.eps}
 	\caption{The Mobile Client Architecture}
@@
 \begin{figure}[h!]
-	\centering
-	\includegraphics[width=8.8cm]{figures/onlinemodeoperationss.eps}
+	\includegraphics[width=8.6cm]{figures/onlinemodeoperations.eps}
 	\caption{Online Mode Operations}
@@
 	\centering
-	\includegraphics[width=8.7cm]{figures/OnlineAuthorizationWorkflow.eps}
+	\includegraphics[width=8.6cm]{figures/onlineauthorizationworkflow.eps}
 	\caption{Online Authorization Workflow}
@@
 	\centering
-	\includegraphics[width=8.7cm]{figures/offlinesecurityarchitecture.eps}
+	\includegraphics[width=8.6cm]{figures/offlinesecurityarchitecture.eps}
 	\caption{Online Authorization Workflow}
@@
 	\centering
-	\includegraphics[width=8.8cm]{figures/offlinemodeoperationss.eps}
+	\includegraphics[width=8.6cm]{figures/offlinemodeoperations.eps}
 	\caption{Offline Mode Operations}
@@
 	\centering
-	\includegraphics[width=8.7cm]{figures/OfflineAuthorizationWorkflow.eps}
+	\includegraphics[width=8.6cm]{figures/offlineauthorizationworkflow.eps}
 	\caption{Offline mode authorization workflow}
@@
 \label{sec_offline_behavioral_analysis}
-The Log Analysis Center (LAC) is responsible for the behavioral analysis done by the mobile client, in order to identify malicious behaviors through MOS schemes applied to selected features, that shall be incrementally logged by TIM.
-Therefore, the selected features are modeled as matrices, composed of the number of occurrences of the selected feature by its location and by the occurrence time.
-Subsequently, the matrices are submitted to the MOS algorithm for threat intelligence, which is detailed in Subsection \ref{sec_mos}.
-The Figure \ref{fig:8} depicts the TIM and LAC for offline behavioral analysis.
+
+In the proposed client security architecture, the Threat Intelligence Manager (TIM) is resposibible for receiving logged user operations, feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode.
+The Figure \ref{fig:8} depicts the Threat Intelligence Manager for offline behavioral analysis.
 
@@
 	\centering
-	\includegraphics[width=8.7cm]{figures/mosworkflow.eps}
-	\caption{TIM and LAC workflow for offline behavioral analysis}
+	\includegraphics[width=8.6cm]{figures/mosworkflow.eps}
+	\caption{The Threat Intelligence Manager Workflow}
 	\label{fig:8}
 \end{figure}
+
+The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats.
+Therefore, the user operations is monitored by the client app, which sends the information of the captured operations to the User Activity Logging (UAL),
+
+The UAL is responsible for the incremental logging, feature extraction and data modeling for malicious behavior detection, through the Log Analysis Center (LAC).
+The logged information shall be decomposed into selected features and modeled as matrices, composed of the number of occurrences of the selected features by its location and by the occurrence time.
+The resultant data is submited to the LAC, for anomaly detection.
+
+The LAC performs the behavioral analysis through eigenvalue analysis and MOS schemes, which identify anomalies on sparse, subtle or abrupt number of user operations.
+The malicious behavior detection is detail described in \ref{sec_mos}.
 
@@
 This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing.
-The encryption workflow is outlined in the Figure 5.1.
+The encryption workflow is outlined in the Figure \ref{fig:9}.
 
-Figure 5.1: Encryption workflow
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8.6cm]{figures/encryptionworkflow.eps}
+	\caption{Encryption workflow}
+	\label{fig:9}
+\end{figure}
 
-The key hierarchy can be explained by Figure 5.2 where the key represented in Figure 5.1 serves for decrypting the key for each level represented in Figure 5.2.
+The key hierarchy can be explained by Figure \ref{fig:10} where the key represented in Figure \ref{fig:9} serves for decrypting the key for each level represented in Figure \ref{fig:10}.
 
-Figure 5.2: Encryption key hierarchy
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8cm]{figures/encryptionkeyhierarchy.eps}
+	\caption{Encryption key hierarchy}
+	\label{fig:10}
+\end{figure}
+
 This proposed encryption scheme is called as a hybrid encryption since it uses a combination of different encryption methods in order to support all requirements to the encryption scheme used in the proposed infrastructure.
 Then it has:
-AES provides a fast encryption of the data files;
-ABE permits to support the authorization policy on the encryption level;
-3)  SSS allows protecting the key storage by avoiding the necessity to store the key proof.
+\begin{enumerate}
+	\item AES provides a fast encryption of the data files;
+	\item ABE permits to support the authorization policy on the encryption level;
+	\item SSS allows protecting the key storage by avoiding the necessity to store the key proof.
+\end{enumerate}
 
-In Subsections \ref{sec_aes}-\ref{sec_secret}, a detailed description of the building blocks of the hybrid encryption is provided.
+In Subsections \ref{sec_aes}-\ref{sec_secret}, a detailed description of the building blocks of the hybrid encryption is provided, and the proposed malicious behavior detection scheme is detail described in \ref{sec_mos}.
 
@@
 This key is protected by attribute-based public key encryption (ABE).
-The scheme of the server-side AES encryption is represented in Figure 5.3.
+The scheme of the server-side AES encryption is represented in Figure \ref{fig:11}.
 
-Figure 5.3: Server-side encryption
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8cm]{figures/server-sideencryption.eps}
+	\caption{Server-side encryption}
+	\label{fig:11}
+\end{figure}
+
 Usually, The AES encryption key is used as a session key.

### 2016-06-14 docs/papers/2016JGC/2016JGC.tex
one more step

@@
 Then it has:
+
 \begin{enumerate}
@@
 	\centering
-	\includegraphics[width=8cm]{figures/server-sideencryption.eps}
+	\includegraphics[width=8.6cm]{figures/server-sideencryption.eps}
 	\caption{Server-side encryption}
@@
 
-Usually, The AES encryption key is used as a session key.
+Usually, the AES encryption key is used as a session key.
 In our setting, it is not desirable to re-encrypt the files stored in the device unless there is a certain specific condition (for example, the user leaves the domain or a specific group in the domain (the group of users in the domain contains the users with the equal access rights, i.e.
@@
 \label{sec_abe}
-With each user session, the permanent FILE\_KEY (unique AES key) is re-encrypted.
+With each user session, the permanent FILE\_KEY (uni\-que AES key) is re-encrypted.
 The set of FILE\_KEYs is protected with the corresponding ABE keys.
@@
 the key attributes correspond both to the groups and the file shares.
+
 The model supports the simple selective ABE scheme \cite{galibus2015cloud,galibus2014}.
-The selective scheme for attribute-based encryption is as follows: if at least one attribute in the set{ti}U is equal to the attribute in the set{ti}M, the corresponding user U can decrypt the text M.
+The selective scheme for attribute-based encryption is as follows: if at least one attribute in the set $\{t_i\}_M$ is equal to the attribute in the set{ti}M, the corresponding user U can decrypt the text M.
 In other words, as soon as the user and share have one attribute in common – the user can get access to the share.
@@
 
-Master-key (MK) which is kept safely on server and accessible only for the domain administrator and is given by
-MK = (t1, t2,...,tn, y),
-where the values ti are randomly selected from the huge group Zp.
+\begin{enumerate}
+	\item \textbf{Master-key (MK)} which is kept safely on server and accessible only for the domain administrator and is given by $MK = (t_1, t_2,...,t_n, y)$, where the values $t_i$ are randomly selected from the huge group $Zp$.
 They are the private keys corresponding to the group attributes.
 Note, that this is different from the usual PK encryption: the private keys are controlled by the admin and not by the users.
+	\item \textbf{Public key (PK)} depends on the master key values and is kept in the clear allowing users to access the information: $PK = (g^{t_1}, g^{t_2}",...,g^{t_n}, e(g,g)^y)$, Here $e(g,g)$ is the bilinear pairing function corresponding to an elliptic curve.
+	\item Secret user \textbf{KEY\_SET} depends on his attribute set.
+Here each $D_i$ (\textbf{GROUP\_KEY}) serves for decryption of the data of a single group of users, for example, related to some project:
 
-Public key (PK) depends on the master key values and is kept in the clear allowing users to access the information:
+	$\{t_i\}_U \rightarrow D = \{D_i = g^{\frac{yw}{t_i}}\}$.
 
-PK = (gt1, gt2,...,gtn, e(g,g)y),
-Here e(g,g) is the bilinear pairing function corresponding to an elliptic curve.
+	\item Encrypted text $M$, in our context, $M = FILE\_KEY$, or the permanent AES symmetric key, which allows to avoid the file re-encryption.
+Encryption procedure is multiplication.
+The set of the public keys $E_i$ (PUBLIC\_SHARE\_KEY) corresponding to the set of groups able to access the text is kept along with the encrypted text $E$:
 
-Secret user KEY\_SET depends on his attribute set. Here each Di (GROUP\_KEY) serves for decryption of the data of a single group of users, for example, related to some project:
-{ti}UD={Di=gywti}
+	$E=Me(g,g)^{ys}, \{E_i = g^{\frac{t_is}{w}}\}, \forall i \in \{t_i\}_M$.
 
-Encrypted text M, in our context, M = FILE\_KEY, or the permanent AES symmetric key, which allows to avoid the file re-encryption.
-Encryption procedure is multiplication.
-The set of the public keys Ei (PUBLIC\_SHARE\_KEY) corresponding to the set of groups able to access the text is kept along with the encrypted text E:
+	\item Decryption is division: $M = \frac{E}{Y^s}$. 
+\end{enumerate}
 
-E=Me(g,g)ys,{ Ei =gtis/w }, i{ti}M
+In order to perform this operation the user needs the pair of private key $D_i$ and public key $E_i$ corresponding to the attribute $t_i$:
 
-Decryption is division:
-M=E/Ys
-
-In order to perform this operation the user needs the pair of private key Di and public key Ei corresponding to the attribute ti :
-Ys=e(g,g)ys=e(Ei,Di)=e(gywti,gtis/w)=e(g,g)ys
+$Y^s = e(g,g)^{ys} = e(E_i,D_i) = e(g^{\frac{yw}{t_i}},g^{\frac{t_is}{w}})=e(g,g)^{ys}$.
 
@@
 
-The attribute-based private keys Di should be protected while being stored in the device memory.
-Therefore, server encrypts the set of Di with a single AES key before sending it to the user device.
+The attribute-based private keys $D_i$ should be protected while being stored in the device memory.
+Therefore, server encrypts the set of $Di$ with a single AES key before sending it to the user device.
 This AES key (master key) is denoted by the value KEY\_SET\_KEY in our notation.
-KEY\_SET\_KEY is a secret value and it is split by the secure method of polynomial modular secret sharing \cite{galibus2015mobile, galibus2008some} into the set of 4 shares:
-
-KEY\_SET\_KEY = PASS+PIN+TIME+DEV\_PASS.
+KEY\_SET\_KEY is a secret value and it is split by the secure method of polynomial modular secret sharing \cite{galibus2015mobile, galibus2008some} into the set of 4 shares: KEY\_SET\_KEY = PASS + PIN + TIME + DEV\_PASS.
 
@@
 The participants of the (2, 2)-threshold SSS are the user and device.
-The user share s1(x) is computed based on the PIN and the PASS entered by the user.
+The user share $s_1(x)$ is computed based on the PIN and the PASS entered by the user.
 Additionally, the current time value TIME is used in the calculation of the share.
-Let s(x)=d and s1(x)=f(PIN+PASS+TIME),
-Where f is a one-way function that transforms the data into the string of the desired length:
+Let $s(x) = d$ and $s_1(x) = f(PIN + PASS + TIME)$, where $f$ is a one-way function that transforms the data into the string of the desired length:
+
+\begin{equation}\label{eq:eq1}
+  \left\{
+  \begin{array}{@{}ll@{}}
+    S = s \textrm{ mod } p_0
+    \\
+    S = s_1 \textrm{ mod } p_1
+  \end{array}\right.
+\end{equation}
 
 According to the CRT:
+\begin{equation}\label{eq:eq2}
+S \equiv {s_{1}p_{0}{p_0}^{-1}}_{p1} + {s_{1}p_{1}{p_1}^{-1}}_{p0} + \textrm{ mod } p_0p_1
+\end{equation}
 
@@
 MOS for threat intelligence requires that the target features, which can be user operations, are modeled as a matrix composed by the number of occurrences by location and time.
-For example, one element can represent the number of file readings on folder m during the minute n.
+For example, one element can represent the number of file readings on folder $m$ during the minute $n$.
 

### 2016-06-14 docs/papers/2016JNCA/2016JNCA.tex
one more step

@@
 \end{equation}
+
 where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, which is used to calculate the $\boldsymbol{v}_{(m,\hat{n})}$ most significant eigenvectors that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.

### 2016-06-16 docs/papers/2016JGC/2016JGC.tex
more...

@@
 \usepackage[utf8x]{inputenc}
+\usepackage{enumitem}
+\newtheorem{thm}{Scenario}
+
+% math
+\usepackage{amsfonts}
+\usepackage{amsmath}
 
@@
 
-% reference and bib issues
+% reference and bib
 \usepackage[numbers]{natbib}
@@
 Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized leading to obscure and incomplete security paradigms and solutions.
+
 A common practice to provide a cloud solution with stable security is to use a specific type of cloud service, such as Cloud Access Security Broker (CASB) or Cloud Access Control (CAC).
@@
 end-client, side \cite{campbell2015}.
+
 Additional security issues and requirements have to be considered when mobile devices are actively used in corporate cloud environment \cite{yovel2014}.
@@
 However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode.
+
 Suppose that a SME uses CASB in order to protect data at rest, i.e.
@@
 The main target of our proposed solution is to provide a maximum defense at the minimal resource cost.
+
 The paper is structured as follows.
@@
 \begin{enumerate}
-	\item \textbf{Master-key (MK)} which is kept safely on server and accessible only for the domain administrator and is given by $MK = (t_1, t_2,...,t_n, y)$, where the values $t_i$ are randomly selected from the huge group $Zp$.
+	\item \textbf{Master-key (MK)} which is kept safely on server and accessible only for the domain administrator and is given by $MK = (t_1, t_2, \cdots ,t_n, y)$, where the values $t_i$ are randomly selected from the huge group $Zp$.
 They are the private keys corresponding to the group attributes.
 Note, that this is different from the usual PK encryption: the private keys are controlled by the admin and not by the users.
-	\item \textbf{Public key (PK)} depends on the master key values and is kept in the clear allowing users to access the information: $PK = (g^{t_1}, g^{t_2}",...,g^{t_n}, e(g,g)^y)$, Here $e(g,g)$ is the bilinear pairing function corresponding to an elliptic curve.
+	\item \textbf{Public key (PK)} depends on the master key values and is kept in the clear allowing users to access the information: $PK = (g^{t_1}, g^{t_2}", \cdots ,g^{t_n}, e(g,g)^y)$, Here $e(g,g)$ is the bilinear pairing function corresponding to an elliptic curve.
 	\item Secret user \textbf{KEY\_SET} depends on his attribute set.
@@
 
-Since the underlying sharing scheme is perfect \cite{galibus2008some}, the adversary cannot get any information of the KEY\_SET\_KEY unless he possessing all 4 key parts.
+Since the underlying sharing scheme is perfect \cite{galibus2008some}, hence the adversary cannot get any information of the KEY\_SET\_KEY unless he possessing all 4 key parts.
 Here the values PASS and PIN are predefined similar to the construction in \cite{galibus2015mobile}.
@@
 
-\begin{equation}\label{eq:eq1}
+\begin{equation}\label{eq:eq01}
   \left\{
@@
 According to the CRT:
-\begin{equation}\label{eq:eq2}
+\begin{equation}\label{eq:eq02}
 S \equiv {s_{1}p_{0}{p_0}^{-1}}_{p1} + {s_{1}p_{1}{p_1}^{-1}}_{p0} + \textrm{ mod } p_0p_1
@@
 Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session.
-MOS for threat intelligence requires that the target features, which can be user operations, are modeled as a matrix composed by the number of occurrences by location and time.
-For example, one element can represent the number of file readings on folder $m$ during the minute $n$.
+MOS for threat intelligence requires that the target features, which can be user operations, are modeled as a matrix composed by the number of occurrences by location and time, and grouped into $Q$ time frames.
+Therefore, the framework considers the time variations of the matrix $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the occurrence of malicious behaviors.
+For example, one element of $\mathbf{X}^{(q)}$ can represent the number of file readings on folder $m$ during the minute $n$, from file operations logged by the client app.
 
-The MOS schemes can rely on correlation or covariance analysis, where the correlation analysis identifies anomalies on sparse or subtle number of file operations, while the MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period.
+The MOS schemes can rely on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, where the former is useful to identify abnormalities caused by large amounts of operations during a period, while the latter is applied to identify anomalies on sparse or subtle number of file operations.
 
-Classical approaches to model order selection require the computation of the sample covariance matrix Vxx and of its eigenvalues, obtained from the measurement matrix X of the zero mean samples given by
+Classical approaches to model order selection require the computation of the sample covariance matrix $\mathbf{\hat{R}}_{yy}^{(q)}$ and of its eigenvalues, obtained from the measurement matrix $\mathbf{X}$ of the zero mean samples given by
 
-The set of obtained vectors zm composes the zero mean matrix Z, then the covariance matrix Vxx can be calculated as follows
+\begin{equation}\label{eq:eq03}
+\mathbf{y}_{m}^{(q)} = \mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}.
+\end{equation}
 
-Vxx =1NZZT
+The set of obtained vectors $\mathbf{y}_{m}^{(q)}$ composes the zero mean matrix $\mathbf{Y}^{(q)}$, then the sample covariance matrix $\mathbf{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
 
-For MOS based on correlation analysis, instead of calculate the zero mean samples, to obtain the correlation matrix Rxx , it is required, for each variable, to make the standard deviation unitary as follows
+\begin{equation}\label{eq:eq04}
+\mathbf{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\mathbf{Y}^{(q)}\mathbf{Y}^{(q)^{\rm T}}.
+\end{equation}
 
-The set of vectors sm composes the matrix S, then the correlation matrix Rxx can be calculated via 
-.
-Once the Vxx or Rxx have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to Vxx and Rxx as a matrix C.
-Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to , in order to obtain the vector of eigenvalues e, as following:
+For MOS based on sample covariance of zero mean and unitary standard deviation, it is required, for each variable, to make the standard deviation unitary as follows
 
-The eigenvalues should be sorted in descending order, as defined by , to make possible the selection of the first eigenvalue in the obtained sequence, represented by, which is the largest eigenvalue of the data evaluated for attack detection.
+\begin{equation}\label{eq:eq05}
+\mathbf{z}_{m}^{(q)} = \frac{\mathbf{x}_{m}^{(q)} - \bar{\mathbf{x}}_{m}^{(q)}}{\mathbf{\sigma}_{m}^{(q)}}.
+\end{equation}
 
-The process of obtaining the and the matrix C(t), finding the largest eigenvalue for each t-th time frame, should be repeated until t = T, in order to obtain the largest eigenvalue of all time frames, as presented by
+The set of vectors $\mathbf{z}_{m}^{(q)}$ composes the matrix $\mathbf{Z}^{(q)}$, then the sample covariance matrix $\mathbf{\hat{R}}_{zz}^{(q)}$ can be calculated via
 
-E =
+\begin{equation}\label{eq:eq06}
+\mathbf{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\mathbf{Z}^{(q)}\mathbf{Z}^{(q)^{\rm T}}.
+\end{equation}
 
-Since   , then the first line of the matrix E contains the largest eigenvalues of each t-th time frame, which is the expected input for MOS schemes and can be expressed as 
-emax=E{:,1}
+Once the $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ as a matrix $\mathbf{C}$.
+Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\mathbf{C}^{(q)} = \mathbf{V}^{(q)}\mathbf{\Lambda}^{(q)}\mathbf{V}^{(q)^{\rm T}},$, in order to obtain the vector of eigenvalues $e$, as following:
 
-Once obtained the largest eigenvalues of each q-th time frame, it is possible to apply a selected MOS scheme to estimate the model order $\hat{d}$, which is the estimated number of time frames with malicious behavior.
-Therefore, emax is used as input parameter for MOS schemes, according to the equation.
+\begin{equation}\label{eq:eq07}
+\mathbf{e}^{(q)} = \rm diag(\mathbf{\Lambda}^{(q)}),
+\end{equation}
 
-Note that some MOS schemes may also require the amount of time that compose a time frame, such as . For more information about MOS schemes, interested readers are referred to \cite{da2009comparison}.
+The eigenvalues should be sorted in descending order, as defined by $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > \cdots > \lambda_{m}^{(q)}$, to make possible the selection of the first eigenvalue in the obtained sequence, represented by $\lambda_{1}^{(q)}$, which is the largest eigenvalue of the data evaluated for attack detection.
+
+The process of obtaining the $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$ and the matrix $\mathbf{C}^{(q)}$, finding the largest eigenvalue for each $q$-th time frame, should be repeated until $q = Q$, in order to obtain the largest eigenvalue of all time frames, as presented by
+
+\begin{equation}\label{eq:eq08}
+\mathbf{E} =
+\begin{bmatrix}
+  \lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
+  \lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
+  \lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
+  \vdots & \vdots & \ddots & \vdots  \\
+  \lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
+\end{bmatrix}.
+\end{equation}
+
+Since $\lambda_1^{(q)} > \lambda_2^{(q)} > \lambda_3^{(q)} > \cdots > \lambda_{m-1}^{(q)} > \lambda_m^{(q)}$, then the first line of the matrix $\mathbf{E}$ contains the largest eigenvalues of each $q$-th time frame, which is the expected input for MOS schemes and can be expressed as
+
+\begin{equation}\label{eq:eq09}
+\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} \cdots \lambda_1^{(Q)}]
+\end{equation}
+
+Once obtained the largest eigenvalues of each $q$-th time frame, it is possible to apply a selected MOS scheme to estimate the model order $\hat{d}$, which is the estimated number of time frames with malicious behavior.
+Therefore, $\boldsymbol{e}_{\rm max}$ is used as input parameter for MOS schemes, according to the equation
+
+\begin{equation}\label{eq:eq09}
+\hat{d} = \textrm{MOS}(\boldsymbol{e}_{\rm max})
+\end{equation}
+
+Note that some MOS schemes may also require the amount of time that compose a time frame, such as $\hat{d} = \textrm{MOS}(\boldsymbol{e}_{\rm max}, M)$.
+For more information about MOS schemes, interested readers are referred to \cite{da2009comparison}.
 
@@
 
-... add some text
+This section provides the detailed analysis of the results the proposed approach regarding security and complexity analysis.
 
@@
 
-Steal the user keys one-by-one while the app performs decryption, if he is outsider;
-Use expired key to decrypt the files, if he is insider;
-Does not have possibility to perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the client app;
-Steal the encrypted files but not the decrypted ones.
+\begin{enumerate}
+	\item Steal the user keys one-by-one while the app performs decryption, if he is outsider;
+	\item Use expired key to decrypt the files, if he is insider;
+	\item Does not have possibility to perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the client app;
+	\item Steal the encrypted files but not the decrypted ones.
+\end{enumerate}
 
@@
 \label{sec_sec_analysis}
-Offline mode
-In the offline mode the client does not store the user password (see section 3 for details), i.e.
+
+\textbf{Offline Mode}\\
+
+In the offline mode the client does not store the user password (see Section \ref{sec_offline_mode_workflow} for details), i.e.
 no information about the password leaks, and therefore there is no possibility for the malicious user to check if the password he is trying to enter is correct or not.
 The only possible scenario for the information leakage in the case of the malicious outsider is if:
-The hacker steals all the parts of KEY\_SET\_KEY;
-The hacker steals the KEY\_SET;
-The hacker try the brute force offline dictionary attack on all the previous values, they have to belong to ONE TIME SESSION (the values of ”a” and ”b” belong to one period of time i.e.
+
+\begin{enumerate}[label=(\alph*)]
+	\item The hacker steals all the parts of KEY\_SET\_KEY;
+	\item The hacker steals the KEY\_SET;
+	\item The hacker try the brute force offline dictionary attack on all the previous values, they have to belong to ONE TIME SESSION (the values of ”a” and ”b” belong to one period of time i.e.
 TIME, DEV\_PASS, KEY\_SET);
-Steal the permanent FILE\_KEY (this is protected by the KEY\_SET);
-Steal the file and try to decrypt it with offline dictionary attacks.
-        Still, the hacker has to get 4 values: TIME, DEV\_PASS, KEY\_SET, FILE\_KEY - from one session.
+	\item Steal the permanent FILE\_KEY (this is protected by the KEY\_SET);
+	\item Steal the file and try to decrypt it with offline dictionary attacks.
+\end{enumerate}
+
+Still, the hacker has to get 4 values from one session: TIME, DEV\_PASS, KEY\_SET, FILE\_KEY.
 At the same time he should try the offline dictionary attack on PIN+PASS.
@@
 So practically, it is very difficult to perform such attack due to the key expiration period.
-In accordance with the conditions 1) and 3) of our adversary model, the above actions are not possible for such adversary in the conditions presented.
+In accordance with the conditions 1 and 3 of our adversary model, the above actions are not possible for such adversary in the conditions presented.
 
@@
 The only possible scenario when malicious or expired user wishes to prolong his old credentials is:
-He has to steal the DEV\_PASS and TIME synchronized with his credentials;
-He has to be able to combine;
-He has to steal the KEY\_SET synchronized with his credentials;
-He has to steal the protected FILE\_KEY (also synchronized) and encrypted file;
-He has to do everything without the client (because the client checks the TIME and renews the PUBLIC\_SHARE\_KEY).
+
+\begin{enumerate}[label=(\alph*)]
+	\item Steal the file and try to decrypt it with offline dictionary attacks.
+	\item He has to be able to combine;
+	\item He has to steal the KEY\_SET synchronized with his credentials;
+	\item He has to steal the protected FILE\_KEY (also synchronized) and encrypted file;
+	\item He has to do everything without the client (because the client checks the TIME and renews the PUBLIC\_SHARE\_KEY).
+\end{enumerate}
+
 Basically, for a malicious user there is practically no way to use the client with the old keys.
@@
 The client still has to count the tries (to prevent the hacker to perform brute force attack) within one session.
-In accordance with the conditions 2) and 4) of the adversary model, the above actions are not possible for such adversary.
-Online mode
+In accordance with the conditions 2 and 4 of the adversary model, the above actions are not possible for such adversary.
+\\
+\\
+\textbf{Online mode}\\
+
 In the online mode the protection of the mobile system is backed up by the communication with server.
@@
 
-\subsubsection{An attacker uses a valid password to perform operations on a bulk of files}
-\label{sec_attacker_valid}
+\begin{thm}
+An attacker uses a valid password to perform operations on a bulk of files.
+\end{thm}
+
 The session time defines the period when operations can be performed until the next session renewing.
@@
 The access to the resources can be blocked by the attacker through some techniques, when the content is encrypted by the attacker, the ransomware attack can be called cryptoransomware \cite{kaspersky2014}.
+
 MOS schemes based on covariance analysis are effective to reveal abrupt changing of behaviors over time \cite{tenorio2013greatest}, making possible to identify intense malicious behaviors on offline mode of mobile clients, such in case of ransomware attack or bulk access to sensitive data.
+
 The large number of operations over time is a well-known pattern of some attacks, due to the efforts on security measures to make the attacks infeasible over time.
 In this context, the operations can also be evaluated in contrast to the estimated required time for operations done by legitimate behaviors, such as the evaluation of the mean time between operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities.
+
 Sparse or subtle file operations, with low number of operations distributed over different files or directories, during short period of time can indicate anomalies in contrast to the required time for legitimate directory navigation.
 MOS and correlation analysis can be suitable if applied to evaluate the time and location of operations, in order to identify unreachable navigation, if compared to legitimate navigation
+
 The MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period.
 The eigenvalue analysis based on the correlation matrix highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
+
 This threat scenario, where an attacker uses a valid password and session to perform operations on a bulk of files, can have its steps described as:
-The hacker has access to the mobile client and is able to perform operations;
-The session time is valid;
-The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
-The client incrementally append each operation attempt time into the logging;
-The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
-The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time;
 
-\subsubsection{Usage of expired password to perform unauthorized operations }
-\label{sec_attacker_expired}
+\begin{enumerate}[label=(\alph*)]
+	\item The hacker has access to the mobile client and is able to perform operations;
+	\item The session time is valid;
+	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
+	\item The client incrementally append each operation attempt time into the logging;
+	\item The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
+	\item The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time.
+\end{enumerate}
+
+\begin{thm}
+Usage of expired password to perform unauthorized operations.
+\end{thm}
+
 In the offline mode, the session time is used to restrict the operations during a specified period, although it is possible to manipulate the current time in mobile clients, to emulate a period in which the session was valid.
 The log analysis by MOS can deal with this kind of threat, through the incremental logging of the time when each operation was performed, followed by the behavioral evaluation of operations over time.
+
 The incremental logging assumes that new logged operations shall have equal or bigger time than the last logged operation, the violation of this rule means that the system is out of sync and indicates a malicious behavior.
@@
 Massive, subtle or sparse malicious operation performed during a valid session time can be identified by MOS schemes based on covariance and correlation analysis.
+
 Applying MOS to the analysis of the time between user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session.
@@
 Consequently, the result of the eigenvalue analysis is applied to MOS schemes, in order to identify the occurrence of malicious behaviors during the valid session.
+
 The MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period.
 The eigenvalue analysis based on correlation matrix highlights massive or concentrated operations over time or location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
+
 This threat scenario, where the attacker uses expired password to perform unauthorized operations, can have its steps described as:
-The hacker steals the operating system;
-The hacker modifies the time of the operating system to a period when the session was valid;
-The hacker has access to the mobile client and is able to perform operations;
-The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
-The client incrementally append each operation attempt time into the logging;
-The client verifies if one logged time is older than the last operation time. If it is true, the MOS module classifies the evaluated operation as malicious;
-The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations;
-The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time;
+
+\begin{enumerate}[label=(\alph*)]
+	\item The hacker steals the operating system;
+	\item The hacker modifies the time of the operating system to a period when the session was valid;
+	\item The hacker has access to the mobile client and is able to perform operations;
+	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
+	\item The client incrementally append each operation attempt time into the logging;
+	\item The client verifies if one logged time is older than the last operation time. If it is true, the MOS module classifies the evaluated operation as malicious;
+	\item The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations;
+	\item The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time;
+\end{enumerate}
 
@@
 \label{sec_data}
+
 MOS schemes are used in order to identify anomalous behavior that can indicate an attack and be used to prevent or avoid attack propagation.
-Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to MOS schemes, according to described in Section 5.4.
-The selected features shall be modeled as matrices which represents a signal superposition containing noise, legitimate and malicious behavior \cite{tenorio2013greatest}, grouped into time frames , where M defines the decomposition of a selected feature, N defines the time decomposition and represents the number of occurrences of the feature m during the time n.
+Therefore, it is necessary to analyze the data that can be collected from user operations on mobile client, to identify features that can be modeled and submitted to MOS schemes, according to described in Section \ref{sec_mos}.
+
+The selected features shall be modeled as matrices which represents a signal superposition containing noise, legitimate and malicious behavior \cite{tenorio2013greatest}, grouped into time frames $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, 2, 3, \ldots, Q$ , where $M$ defines the decomposition of a selected feature, $N$ defines the time decomposition and represents the number of occurrences of the feature $m$ during the time $n$.
+
 In offline mode, the user is still allowed to get access to operations that do not require communication with the server side.
@@
 This work proposes to evaluate the following features:
+
 File Access (Time and File System Location), i.e. data access to selected files in offline mode, accessing the data stored on the mobile client. The file access feature can be decomposed into:
+
 file decryption, 
@@
 decrypted file execution,
+
 Therefore, it is necessary to generate three matrices for malicious behaviors analysis, such as: 

### 2016-06-17 docs/papers/2016JGC/2016JGC.tex
initing tables

@@
 These operations and their selected features shall be incrementally logged by the mobile client, in order to be evaluated to identify malicious behaviors.
-This work proposes to evaluate the following features:
+This work proposes to evaluate the following features.
 
-File Access (Time and File System Location), i.e. data access to selected files in offline mode, accessing the data stored on the mobile client. The file access feature can be decomposed into:
+\paragraph{\textbf{File Access (Time and File System Location)},}i.e.
+data access to selected files in offline mode, accessing the data stored on the mobile client.
+The file access feature can be decomposed into more detailed features, which are:
 
-file decryption, 
-decrypted file reading 
-decrypted file execution,
+\begin{enumerate}
+	\item file decryption; 
+	\item decrypted file reading;
+	\item decrypted file execution. 
+\end{enumerate}
 
 Therefore, it is necessary to generate three matrices for malicious behaviors analysis, such as:
-massive file access, which can reveal data leakage and be identified by MOS schemes based on covariance analysis; 
-low file access into several folders, characterized by sparse operations that can reveal unreachable navigation performed by automated file accesses in order to avoid the massive file access characterization.
+
+\begin{enumerate}[label=(\alph*)]
+	\item massive file access, which can reveal data leakage and be identified by MOS schemes based on covariance analysis; 
+	\item low file access into several folders, characterized by sparse operations that can reveal unreachable navigation performed by automated file accesses in order to avoid the massive file access characterization;
+	\item Malicious sparse file accesses can be identified by MOS schemes base correlation analysis. 
+\end{enumerate}
+
+\paragraph{\textbf{File Update (Time and File System Location)},}i.e.
+writing operations into selected files in offline mode, writing the data stored on the mobile client.
+The update feature can be decomposed into:
+
+\begin{enumerate}
+	\item file encryption;
+	\item decrypted file writing.
+\end{enumerate}
+
+Therefore, it is necessary to generate two matrices for malicious behaviors analysis, such as:
+
+\begin{enumerate}[label=(\alph*)]
+	\item massive file update, which can reveal ransomware or similar attacks and be identified by MOS schemes based on covariance analysis; 
+	\item low number of file update into several folders, characterized by sparse operations that can reveal unreachable navigation performed by automated file accesses in order to avoid the massive file access characterization.
 Malicious sparse file accesses can be identified by MOS schemes base correlation analysis.
-File Update (Time and File System Location), i.e. writing operations into selected files in offline mode, writing the data stored on the mobile client. The update feature can be decomposed into:
- file encryption 
-decrypted file writing, 
-Therefore, it is necessary to generate two matrices for malicious behaviors analysis, such as: 
-massive file update, which can reveal ransomware or similar attacks and be identified by MOS schemes based on covariance analysis; 
-low number of file update into several folders, characterized by sparse operations that can reveal unreachable navigation performed by automated file accesses in order to avoid the massive file access characterization.
-Malicious sparse file accesses can be identified by MOS schemes base correlation analysis.
-File Download (Start Time, End Time and File System Location), i.e download requests in online mode, evaluated by the mobile client.
+\end{enumerate}
+
+\paragraph{\textbf{File Download (Start Time, End Time and File System Location)},}i.e.
+download requests in online mode, evaluated by the mobile client.
 The file download feature shall be modeled as the matrix of number downloads by file location over time, in order to perform malicious behaviors analysis, such as:
-massive data leakage or similar attacks, which can be identified by MOS schemes based on covariance analysis; 
-low number of file download from several folders, characterized by sparse operations, which can reveal unreachable navigation performed by automated file download in order to avoid the massive file download characterization.
+
+\begin{enumerate}
+	\item massive data leakage or similar attacks, which can be identified by MOS schemes based on covariance analysis;
+	\item low number of file download from several folders, characterized by sparse operations, which can reveal unreachable navigation performed by automated file download in order to avoid the massive file download characterization.
 Malicious sparse file download can be identified by MOS schemes base correlation analysis.
-File Upload (Start Time, End Time and File System Location), i.e upload requests in online mode, evaluated by the mobile client.
+\end{enumerate}
+
+\paragraph{\textbf{File Upload (Start Time, End Time and File System Location)},}i.e.
+upload requests in online mode, evaluated by the mobile client.
 The file upload feature can reveal attempts of ransomware or similar attacks and be identified by MOS schemes based on covariance analysis.
 Therefore, it is necessary model the matrix of number uploads by file location over time, in order to perform malicious behaviors analysis, such as:
-massive file upload, similar to ransomware attack, which can be identified by MOS schemes based on covariance analysis; 
-low number of file upload to several folders, characterized by sparse operations, which can reveal unreachable navigation performed by automated file upload in order to avoid the massive file upload characterization.
+
+\begin{enumerate}
+	\item massive file upload, similar to ransomware attack, which can be identified by MOS schemes based on covariance analysis; 
+	\item low number of file upload to several folders, characterized by sparse operations, which can reveal unreachable navigation performed by automated file upload in order to avoid the massive file upload characterization.
 Malicious sparse file upload can be identified by MOS schemes base correlation analysis.
+\end{enumerate}
 
-These features provide the necessary information to apply MOS schemes for evaluating malicious behaviors according to the following scenarios and steps:
-The hacker accesses the mobile client during a valid session and read/write few files at several folders:
-The hacker steals the device;
-The hacker changes the device to offline mode;
-The hacker accesses the mobile client;
-The current session is valid;
-The hacker read/writes few files at several folders;
-MOS based on correlation analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
-MOS indicates the number of time frames with anomalous behavior;
-The hacker accesses the mobile client during a valid session and read/write several files at few folders:
-The hacker steals the device;
-The hacker changes the device to offline mode;
-The hacker accesses the mobile client;
-The current session is valid;
-The hacker read/writes several files at few folders
-MOS based on covariance analysis is applied to the matrix of number of file operations, grouped by the file system location over time
-MOS indicates the number of time frames with anomalous behavior;
-The hacker emulates a valid session and read/write few files at several folders: 
-The hacker steals the device;
-The hacker changes the device to offline mode;
-The current session is expired;
-The hacker changes the device time to a valid period for the current session;
-The hacker accesses the mobile client;
-The current session is valid;
-The hacker perform sparse operations during the valid period;
-MOS based on correlation analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
-MOS indicates the number of time frames with anomalous behavior;
-The hacker emulates a valid session and read/write several files at few folders: 
-The hacker steals the device;
-The hacker changes the device to offline mode;
-The current session is expired;
-The hacker changes the device time to a valid period for the current session;
-The hacker accesses the mobile client;
-The current session is valid;
-The hacker perform bulk operations during the valid session;
-MOS based on covariance analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
-MOS indicates the number of time frames with anomalous behavior;
+%These features provide the necessary information to apply MOS schemes for evaluating malicious behaviors according to the following scenarios and steps:
+%
+%\begin{enumerate}
+%	\item The hacker accesses the mobile client during a valid session and read/write few files at several folders:
+%	\begin{enumerate}
+%		\item The hacker steals the device;
+%		\item The hacker changes the device to offline mode;
+%		\item The hacker accesses the mobile client;
+%		\item The current session is valid;
+%		\item The hacker read/writes few files at several folders;
+%		\item MOS based on correlation analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
+%		\item MOS indicates the number of time frames with anomalous behavior;
+%	\end{enumerate}
+%	\item The hacker accesses the mobile client during a valid session and read/write several files at few folders:
+%	\begin{enumerate}
+%		\item The hacker steals the device;
+%		\item The hacker changes the device to offline mode;
+%		\item The hacker accesses the mobile client;
+%		\item The current session is valid;
+%		\item The hacker read/writes several files at few folders
+%		\item MOS based on covariance analysis is applied to the matrix of number of file operations, grouped by the file system location over time
+%		\item MOS indicates the number of time frames with anomalous behavior;
+%	\end{enumerate}
+%	\item The hacker emulates a valid session and read/write few files at several folders: 
+%	\begin{enumerate}
+%		\item The hacker steals the device;
+%		\item The hacker changes the device to offline mode;
+%		\item The current session is expired;
+%		\item The hacker changes the device time to a valid period for the current session;
+%		\item The hacker accesses the mobile client;
+%		\item The current session is valid;
+%		\item The hacker perform sparse operations during the valid period;
+%		\item MOS based on correlation analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
+%		\item MOS indicates the number of time frames with anomalous behavior;
+%	\end{enumerate}
+%	\item The hacker emulates a valid session and read/write several files at few folders: 
+%	\begin{enumerate}
+%		\item The hacker steals the device;
+%		\item The hacker changes the device to offline mode;
+%		\item The current session is expired;
+%		\item The hacker changes the device time to a valid period for the current session;
+%		\item The hacker accesses the mobile client;
+%		\item The current session is valid;
+%		\item The hacker perform bulk operations during the valid session;
+%		\item MOS based on covariance analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
+%		\item MOS indicates the number of time frames with anomalous behavior;
+%	\end{enumerate}
+%\end{enumerate}
 
@@
 
-Combine the PASS+PIN+TIME+DEV\_PASS=KEY\_SET\_KEY   ->  SSS secret restoring;
-Decrypt the KEY\_SET with the KEY\_SET\_KEY -> symmetric 128/256 AES decryption;
-Select the SHARE\_KEY from the KEY\_SET -> no calculation;
-Decrypt the FILE\_KEY with the SHARE\_KEY  -> ABE decryption;
-Decrypt the file with the FILE\_KEY ->symmetric 128/256 AES d -> ecryption;
-Modify the TIME periodically -> timer;
-Count the tries within the TIME ->count;
-Modify or delete PUBLIC\_SHARE\_KEY-> no calculation;
-Malicious behavior detection -> eigenvalues calculation.
+\begin{enumerate}
+	\item Combine the PASS + PIN + TIME + DEV\_PASS = KEY\_SET\_KEY $\rightarrow$ SSS secret restoring;
+	\item Decrypt the KEY\_SET with the KEY\_SET\_KEY $\rightarrow$ symmetric 128/256 AES decryption;
+	\item Select the SHARE\_KEY from the KEY\_SET $\rightarrow$ no calculation;
+	\item Decrypt the FILE\_KEY with the SHARE\_KEY $\rightarrow$ ABE decryption;
+	\item Decrypt the file with the FILE\_KEY ->symmetric 128/256 AES $\rightarrow$ decryption;
+	\item Modify the TIME periodically $\rightarrow$ timer;
+	\item Count the tries within the TIME $\rightarrow$ count;
+	\item Modify or delete PUBLIC\_SHARE\_KEY $\rightarrow$ no calculation;
+	\item Malicious behavior detection $\rightarrow$ eigenvalues calculation.
+\end{enumerate}
+
 It is easy to check that in the offline mode the client app does not perform complex calculations and does not use the resources extensively due to the fact that the initial key is shared and the client app only performs decryption, which is not a time-consuming operation.
@@
 
-Generates keys for J-PAKE -> this can be a resource-consuming operation. In the future we suggest to replace J-PAKE with SIS-based PKE \cite{bogos2013primeless,bellovin1992encrypted};
-Sends and receives data in the clear;
-Performs operations 1)-7);
-Performs the operation 8) and sends the log data for server side analysis.
+\begin{enumerate}
+	\item Generates keys for J-PAKE -> this can be a resource-consuming operation. In the future we suggest to replace J-PAKE with SIS-based PKE \cite{bogos2013primeless,bellovin1992encrypted};
+	\item Sends and receives data in the clear;
+	\item Performs operations 1)-7);
+	\item Performs the operation 8) and sends the log data for server side analysis.
+\end{enumerate}
 
@@
 The common advantage is that the mobile client performs the operations both in the offline and online mode and uses the key expiry and ABE to protect the privacy of the corporate data.
-The table below presents the results of mobile app security workflow testing.
-Table 1 demonstrates that the device decrypts the data with the acceptable speed.
+The Table \ref{tab:01} presents the results of mobile app security workflow testing.
+Table \ref{tab:01} demonstrates that the device decrypts the data with the acceptable speed.
 One can compare the speed of ABE/AES decryption (keys and actual data) and the speed of the whole decryption process.
 
-Table 1. Speed of decryption on a client device.
+\begin{table}
+	\caption{Speed of decryption on a client device}
+	\label{tab:01}
+	\centering
+	\begin{tabular}{l|l|l|l|l|l}
+		Device Model	& File set size	& Mean time of getting decrypted stream (ABE)	& Mean time of decrypting content (AES)	& Mean time of copying file without decrypt	& Mean time of decryption (the whole file)  \\
+		\noalign{\smallskip}\hline\noalign{\smallskip}
+		Samsung SM-T320 Galaxy Tab Pro 8.4	& 10 mb	& 1041 ms	& 4230 ms	& 298 ms		& 5271 ms	\\
+		Samsung SM-T320 Galaxy Tab Pro 8.4	& 1 mb	& 811 ms		& 319 ms		& 23 ms		& 1130 ms	\\
+		Samsung SM-T320 Galaxy Tab Pro 8.4	& 1 kb	& 1078 ms	& 1 ms		& 0 ms		& 1080 ms	\\
+		Samsung GT-I9190 Galaxy S4 Mini		& 1 b	& 1167ms		& 2 ms		& 0 ms		& 1169 ms	\\
+		Samsung GT-I9190 Galaxy S4 Mini		& 1 kb	& 1141ms		& 2 ms		& 0 ms		& 1144 ms	\\
+		Samsung GT-I9190 Galaxy S4 Mini		& 1 mb	& 1138 ms	& 368 ms		& 35 ms		& 1506 ms	\\
+		Samsung GT-I9190 Galaxy S4 Mini		& 10 mb	& 1131 ms	& 3618 ms	& 324 ms		& 4750 ms	\\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
 
-Table 2 demonstrates the results of testing the speed of the mobile device for receiving and displaying the list of files.
+Table \ref{tab:02} demonstrates the results of testing the speed of the mobile device for receiving and displaying the list of files.
 The app was tested for displaying up to 1000 files.
@@
 
-Table 2. Speed of receiving the list of files
+\begin{table}
+	\caption{Speed of receiving the list of files}
+	\label{tab:02}
+	\centering
+	\begin{tabular}{l|l|l|l|l|l|l|l|l|l}
+		Files quantity 	& \# Samples 	& Average (ms) 	& Min (ms) 	& Max (ms) 	& Std. Dev. 	& Error (\%)		& Throughput 	& KB/sec 	& Avg. Bytes \\
+		\noalign{\smallskip}\hline\noalign{\smallskip}
+		1		& 1000	& 131	& 29		& 321	& 31.56		& 0.00\%		& 74		& 161.32		& 2233	\\
+		10		& 1000	& 154	& 37		& 269	& 32.24		& 0.00\%		& 63.4	& 659.79		& 10656	\\
+		100		& 1000	& 396	& 68	 	& 1055	& 87.56		& 0.00\%		& 25		& 2323.2		& 95308	\\
+		1000		& 1000	& 2879	& 475	& 4634	& 496.02		& 0.00\%		& 3.4	& 3178		& 946311	\\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
 
 The device receives and displays the list of files depending on the user attributes with acceptable average speed and the proposed protection scheme is usable.
-Table3 demonstrates the results of performance testing of the keys generation procedure on server:
+Table \ref{tab:03} demonstrates the results of performance testing of the keys generation procedure on server:
 
-Table 3. Speed of generating the user keys.
+\begin{table}
+	\caption{Speed of generating the user keys}
+	\label{tab:03}
+	\centering
+	\begin{tabular}{l|l|l|l|l|l|l|l|l}
+		Label	& \# Samples	& Average (ms)	& Min (ms)	& Max (ms)	& Std. Dev.	& Throughput		& KB/sec		& Avg. Bytes \\
+		\noalign{\smallskip}\hline\noalign{\smallskip}
+		Login 5 shares available		& 1000	& 982	& 225	& 1990	& 271.88		& 20.1	& 146.36		& 7459	\\
+		Login 10 shares available	& 1000	& 1140	& 340	& 4176	& 459.3		& 17.3	& 223.88		& 13218	\\
+		Login 25 shares available	& 1000	& 1267	& 887	& 2303	& 192.75		& 15.5	& 294.55		& 19466 \\
+		\noalign{\smallskip}\hline
+	\end{tabular}
+\end{table}
 
@@
 It can be successfully used and provides acceptable level of security.
+
 The log analysis of the Log Analysis Center has been implemented and evaluated for offline anomaly detection in mobile clients, making it possible to apply anomaly detection techniques in a lightweight fashion, considering low processing requirements for deal with the resource constraints of mobile clients.
@@
 The experiments were performed in two mobile devices, with variations of log size and window size.
+
 Table 4 presents the data modelling time and the processing time of eigenvalues decomposition and MOS calculations applied to anomaly detection from user operation logs of Storgrid mobile client.

### 2016-06-18 docs/papers/2016JGC/2016JGC.tex
table fix

@@
 \newtheorem{thm}{Scenario}
+
+% table
+\usepackage{color, colortbl}
+\definecolor{Gray}{gray}{0.9}
 
@@
 
+\begin{table*}[!t]
+	\caption{Speed of decryption on a client device}
+	\label{tab:01}
+	\centering
+	\begin{tabular}{|l|l|l|l|l|l|}
+		\hline \rowcolor{Gray} Device Model	& File set size	& \begin{tabular}[x]{@{}l@{}}Mean time of\\getting decrypted\\stream (ABE)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Mean time of\\decrypting \\content (AES)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Mean time of\\copying file\\without decrypt\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Mean time of\\decryption\\(the whole file)\end{tabular}	\\ \hline
+
+		\begin{tabular}[x]{@{}l@{}}Samsung SM-T320 Galaxy\\Tab Pro 8.4\end{tabular}	& 10 mb	& 1041 ms	& 4230 ms	& 298 ms		& 5271 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}Samsung SM-T320 Galaxy\\Tab Pro 8.4\end{tabular}	& 1 mb	& 811 ms		& 319 ms		& 23 ms		& 1130 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}Samsung SM-T320 Galaxy\\Tab Pro 8.4\end{tabular}	& 1 kb	& 1078 ms	& 1 ms		& 0 ms		& 1080 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}Samsung GT-I9190 Galaxy\\S4 Mini\end{tabular}		& 1 b	& 1167ms		& 2 ms		& 0 ms		& 1169 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}Samsung GT-I9190 Galaxy\\S4 Mini\end{tabular}		& 1 kb	& 1141ms		& 2 ms		& 0 ms		& 1144 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}Samsung GT-I9190 Galaxy\\S4 Mini\end{tabular}		& 1 mb	& 1138 ms	& 368 ms		& 35 ms		& 1506 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}Samsung GT-I9190 Galaxy\\S4 Mini\end{tabular}		& 10 mb	& 1131 ms	& 3618 ms	& 324 ms		& 4750 ms	\\ \hline
+	\end{tabular}
+\end{table*}
+
 \subsection{Complexity analysis }
@@
 
-\begin{table}
-	\caption{Speed of decryption on a client device}
-	\label{tab:01}
+\begin{table*}[!t]
+	\caption{Speed of receiving the list of files}
+	\label{tab:02}
 	\centering
-	\begin{tabular}{l|l|l|l|l|l}
-		Device Model	& File set size	& Mean time of getting decrypted stream (ABE)	& Mean time of decrypting content (AES)	& Mean time of copying file without decrypt	& Mean time of decryption (the whole file)  \\
-		\noalign{\smallskip}\hline\noalign{\smallskip}
-		Samsung SM-T320 Galaxy Tab Pro 8.4	& 10 mb	& 1041 ms	& 4230 ms	& 298 ms		& 5271 ms	\\
-		Samsung SM-T320 Galaxy Tab Pro 8.4	& 1 mb	& 811 ms		& 319 ms		& 23 ms		& 1130 ms	\\
-		Samsung SM-T320 Galaxy Tab Pro 8.4	& 1 kb	& 1078 ms	& 1 ms		& 0 ms		& 1080 ms	\\
-		Samsung GT-I9190 Galaxy S4 Mini		& 1 b	& 1167ms		& 2 ms		& 0 ms		& 1169 ms	\\
-		Samsung GT-I9190 Galaxy S4 Mini		& 1 kb	& 1141ms		& 2 ms		& 0 ms		& 1144 ms	\\
-		Samsung GT-I9190 Galaxy S4 Mini		& 1 mb	& 1138 ms	& 368 ms		& 35 ms		& 1506 ms	\\
-		Samsung GT-I9190 Galaxy S4 Mini		& 10 mb	& 1131 ms	& 3618 ms	& 324 ms		& 4750 ms	\\
-		\noalign{\smallskip}\hline
+	\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Files\\quantity\end{tabular}	 	& Samples 	& \begin{tabular}[x]{@{}l@{}}Average\\(ms)\end{tabular}	 	& Min (ms) 	& Max (ms) 	& Std.
+Dev.
+	& Error (\%)		& Throughput 	& KB/sec 	& Avg.
+Bytes \\ \hline
+		1		& 1000	& 131	& 29		& 321	& 31.56		& 0.00\%		& 74		& 161.32		& 2233	\\ \hline		
+		10		& 1000	& 154	& 37		& 269	& 32.24		& 0.00\%		& 63.4	& 659.79		& 10656	\\ \hline
+		100		& 1000	& 396	& 68	 	& 1055	& 87.56		& 0.00\%		& 25		& 2323.2		& 95308	\\ \hline
+		1000		& 1000	& 2879	& 475	& 4634	& 496.02		& 0.00\%		& 3.4	& 3178		& 946311	\\ \hline
 	\end{tabular}
-\end{table}
+\end{table*}
+
+\begin{table*}
+	\caption{Speed of generating the user keys}
+	\label{tab:03}
+	\centering
+	\begin{tabular}{|l|l|l|l|l|l|l|l|l|}
+		\hline \rowcolor{Gray} Label	& Samples	& \begin{tabular}[x]{@{}l@{}}Average\\(ms)\end{tabular}	& Min (ms)	& Max (ms)	& Std. Dev.	& Throughput		& KB/sec		& Avg. Bytes \\ \hline
+		\begin{tabular}[x]{@{}l@{}}Login 5\\shares\\available\end{tabular}		& 1000	& 982	& 225	& 1990	& 271.88		& 20.1	& 146.36		& 7459	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}Login 10\\shares\\available\end{tabular}	& 1000	& 1140	& 340	& 4176	& 459.3		& 17.3	& 223.88		& 13218	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}Login 25\\shares\\available\end{tabular}	& 1000	& 1267	& 887	& 2303	& 192.75		& 15.5	& 294.55		& 19466 \\ \hline
+	\end{tabular}
+\end{table*}
 
@@
 
-\begin{table}
-	\caption{Speed of receiving the list of files}
-	\label{tab:02}
-	\centering
-	\begin{tabular}{l|l|l|l|l|l|l|l|l|l}
-		Files quantity 	& \# Samples 	& Average (ms) 	& Min (ms) 	& Max (ms) 	& Std. Dev. 	& Error (\%)		& Throughput 	& KB/sec 	& Avg. Bytes \\
-		\noalign{\smallskip}\hline\noalign{\smallskip}
-		1		& 1000	& 131	& 29		& 321	& 31.56		& 0.00\%		& 74		& 161.32		& 2233	\\
-		10		& 1000	& 154	& 37		& 269	& 32.24		& 0.00\%		& 63.4	& 659.79		& 10656	\\
-		100		& 1000	& 396	& 68	 	& 1055	& 87.56		& 0.00\%		& 25		& 2323.2		& 95308	\\
-		1000		& 1000	& 2879	& 475	& 4634	& 496.02		& 0.00\%		& 3.4	& 3178		& 946311	\\
-		\noalign{\smallskip}\hline
-	\end{tabular}
-\end{table}
-
 The device receives and displays the list of files depending on the user attributes with acceptable average speed and the proposed protection scheme is usable.
 Table \ref{tab:03} demonstrates the results of performance testing of the keys generation procedure on server:
-
-\begin{table}
-	\caption{Speed of generating the user keys}
-	\label{tab:03}
-	\centering
-	\begin{tabular}{l|l|l|l|l|l|l|l|l}
-		Label	& \# Samples	& Average (ms)	& Min (ms)	& Max (ms)	& Std. Dev.	& Throughput		& KB/sec		& Avg. Bytes \\
-		\noalign{\smallskip}\hline\noalign{\smallskip}
-		Login 5 shares available		& 1000	& 982	& 225	& 1990	& 271.88		& 20.1	& 146.36		& 7459	\\
-		Login 10 shares available	& 1000	& 1140	& 340	& 4176	& 459.3		& 17.3	& 223.88		& 13218	\\
-		Login 25 shares available	& 1000	& 1267	& 887	& 2303	& 192.75		& 15.5	& 294.55		& 19466 \\
-		\noalign{\smallskip}\hline
-	\end{tabular}
-\end{table}
 

### 2016-08-26 docs/papers/2016JGC/2016JGC.tex
reorganizacao do projeto

@@
 	\centering
-	\includegraphics[width=6cm]{figures/coresetoffunctionsandprotocols.eps}
+	\includegraphics[width=7cm]{figures/coresetoffunctionsandprotocols.eps}
 	\caption{The core set of functions and protocols of the mobile app security infrastructure}
@@
 
-In the proposed client security architecture, the Threat Intelligence Manager (TIM) is resposibible for receiving logged user operations, feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode.
-The Figure \ref{fig:8} depicts the Threat Intelligence Manager for offline behavioral analysis.
+In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode.
+Figure \ref{fig:8} depicts the Threat Intelligence Manager for offline behavioral analysis.
 
@@
 
+As depicted in Figure \ref{fig:8}, users request operations are logged so that the main features can be monitored in the client app.
 The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats.
-Therefore, the user operations is monitored by the client app, which sends the information of the captured operations to the User Activity Logging (UAL),
+Therefore, the user operations are monitored by the client app, which sends the information of the captured operations to the User Activity Logging (UAL).
 
@@
 	\centering
-	\includegraphics[width=8cm]{figures/encryptionkeyhierarchy.eps}
+	\includegraphics[width=8.6cm]{figures/encryptionkeyhierarchy.eps}
 	\caption{Encryption key hierarchy}
@@
 Nevertheless, domain administrator can choose to encrypt the file storage with Blowfish or Serpent which also provide a high level of safety \cite{nist2000}.
+
 The symmetric encryption on a server side is performed in two steps.
@@
 This key is protected by attribute-based public key encryption (ABE).
+
 The scheme of the server-side AES encryption is represented in Figure \ref{fig:11}.
@@
 	\item Secret user \textbf{KEY\_SET} depends on his attribute set.
-Here each $D_i$ (\textbf{GROUP\_KEY}) serves for decryption of the data of a single group of users, for example, related to some project:
-
-	$\{t_i\}_U \rightarrow D = \{D_i = g^{\frac{yw}{t_i}}\}$.
-
+Here each $D_i$ (\textbf{GROUP\_KEY}) serves for decryption of the data of a single group of users, for example, related to some project: 	$\{t_i\}_U \rightarrow D = \{D_i = g^{\frac{yw}{t_i}}\}$.
 	\item Encrypted text $M$, in our context, $M = FILE\_KEY$, or the permanent AES symmetric key, which allows to avoid the file re-encryption.
 Encryption procedure is multiplication.
-The set of the public keys $E_i$ (PUBLIC\_SHARE\_KEY) corresponding to the set of groups able to access the text is kept along with the encrypted text $E$:
-
-	$E=Me(g,g)^{ys}, \{E_i = g^{\frac{t_is}{w}}\}, \forall i \in \{t_i\}_M$.
-
+The set of the public keys $E_i$ (PUBLIC\_SHARE\_KEY) corresponding to the set of groups able to access the text is kept along with the encrypted text $E$: $E=Me(g,g)^{ys}, \{E_i = g^{\frac{t_is}{w}}\}, \forall i \in \{t_i\}_M$.
 	\item Decryption is division: $M = \frac{E}{Y^s}$. 
@@
 
-$Y^s = e(g,g)^{ys} = e(E_i,D_i) = e(g^{\frac{yw}{t_i}},g^{\frac{t_is}{w}})=e(g,g)^{ys}$.
+\begin{equation}\label{eq:eq00}
+Y^s = e(g,g)^{ys} = e(E_i,D_i) = e(g^{\frac{yw}{t_i}},g^{\frac{t_is}{w}})=e(g,g)^{ys}.
+\end{equation}
 
@@
 Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session.
+
 MOS for threat intelligence requires that the target features, which can be user operations, are modeled as a matrix composed by the number of occurrences by location and time, and grouped into $Q$ time frames.
@@
 
-The MOS schemes can rely on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, where the former is useful to identify abnormalities caused by large amounts of operations during a period, while the latter is applied to identify anomalies on sparse or subtle number of file operations.
+The MOS schemes can rely on sample covariance of zero mean variables (called as zero mean covariance for the sake of simplicity) and sample covariance of zero mean and unitary standard deviation (called as zero mean and standardized covariance for the sake of simplicity) variables, where the former is useful to identify abnormalities caused by large amounts of operations during a period, while the latter is applied to identify anomalies on sparse or subtle number of file operations.
 
@@
 
-The set of obtained vectors $\mathbf{y}_{m}^{(q)}$ composes the zero mean matrix $\mathbf{Y}^{(q)}$, then the sample covariance matrix $\mathbf{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
+The set of obtained vectors $\mathbf{y}_{m}^{(q)}$ composes the zero mean matrix $\mathbf{Y}^{(q)}$, then the zero mean covariance matrix $\mathbf{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
 
@@
 
-For MOS based on sample covariance of zero mean and unitary standard deviation, it is required, for each variable, to make the standard deviation unitary as follows
+For MOS based on sample covariance of zero mean and unitary standard deviation, in order to identify anomalies with sparse or subtle behavior, it is required, for each variable, to make the standard deviation unitary as follows
 
@@
 
-The set of vectors $\mathbf{z}_{m}^{(q)}$ composes the matrix $\mathbf{Z}^{(q)}$, then the sample covariance matrix $\mathbf{\hat{R}}_{zz}^{(q)}$ can be calculated via
+The set of vectors $\mathbf{z}_{m}^{(q)}$ composes the matrix $\mathbf{Z}^{(q)}$, then the zero mean and standardized covariance matrix $\mathbf{\hat{R}}_{zz}^{(q)}$ can be calculated via
 
@@
 
-MOS schemes based on covariance analysis are effective to reveal abrupt changing of behaviors over time \cite{tenorio2013greatest}, making possible to identify intense malicious behaviors on offline mode of mobile clients, such in case of ransomware attack or bulk access to sensitive data.
+MOS schemes based on zero mean covariance analysis are effective to reveal abrupt changing of behaviors over time \cite{tenorio2013greatest}, making possible to identify intense malicious behaviors on offline mode of mobile clients, such in case of ransomware attack or bulk access to sensitive data.
 
@@
 Sparse or subtle file operations, with low number of operations distributed over different files or directories, during short period of time can indicate anomalies in contrast to the required time for legitimate directory navigation.
-MOS and correlation analysis can be suitable if applied to evaluate the time and location of operations, in order to identify unreachable navigation, if compared to legitimate navigation
-
-The MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period.
-The eigenvalue analysis based on the correlation matrix highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
+MOS and zero mean and standardized covariance analysis can be suitable if applied to evaluate the time and location of operations, in order to identify unreachable navigation, if compared to legitimate navigation
+
+The MOS based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period.
+The eigenvalue analysis based on the zero mean and standardized covariance highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
 
@@
 	\item The client incrementally append each operation attempt time into the logging;
-	\item The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
-	\item The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time.
+	\item The MOS module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
+	\item The MOS module evaluates the logging of legitimate operations, applying zero mean covariance analysis to identify abnormalities caused by massive operations during the session time.
 \end{enumerate}
@@
 Additionally, a large amount or sparse operation performed at the same time, or during a short period, can indicate the use of backtrack techniques to maintain a valid session during necessary time to perform an attack.
-Massive, subtle or sparse malicious operation performed during a valid session time can be identified by MOS schemes based on covariance and correlation analysis.
+Massive, subtle or sparse malicious operation performed during a valid session time can be identified by MOS schemes based on covariance analysis.
 
 Applying MOS to the analysis of the time between user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session.
-The MOS based on correlation analysis identifies anomalies on sparse or subtle number of file operations, since the eigenvalue analysis based on the correlation matrix highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations.
+The MOS based on zero mean and standardized covariance analysis identifies anomalies on sparse or subtle variation in the number of file operations, since the this eigenvalue analysis highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations.
 Consequently, the result of the eigenvalue analysis is applied to MOS schemes, in order to identify the occurrence of malicious behaviors during the valid session.
 
-The MOS based on covariance analysis indicates abnormalities caused by large amounts of operations during a period.
-The eigenvalue analysis based on correlation matrix highlights massive or concentrated operations over time or location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
+The MOS based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period.
+The eigenvalue analysis based on the zero mean covariance matrix highlights massive or concentrated operations over time or location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
 
@@
 	\item The client verifies if one logged time is older than the last operation time. If it is true, the MOS module classifies the evaluated operation as malicious;
-	\item The MOS module evaluates the logging of legitimate operations, applying correlation analysis to identify anomalies on sparse or subtle number of file operations;
-	\item The MOS module evaluates the logging of legitimate operations, applying covariance analysis to identify abnormalities caused by massive operations during the session time;
+	\item The MOS module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations;
+	\item The MOS module evaluates the logging of legitimate operations, applying zero mean covariance analysis to identify abnormalities caused by massive operations during the session time;
 \end{enumerate}
@@
 \begin{enumerate}[label=(\alph*)]
-	\item massive file access, which can reveal data leakage and be identified by MOS schemes based on covariance analysis; 
+	\item massive file access, which can reveal data leakage and be identified by MOS schemes based on zero mean covariance analysis; 
 	\item low file access into several folders, characterized by sparse operations that can reveal unreachable navigation performed by automated file accesses in order to avoid the massive file access characterization;
-	\item Malicious sparse file accesses can be identified by MOS schemes base correlation analysis. 
+	\item Malicious sparse file accesses can be identified by MOS schemes based on zero mean and standardized covariance analysis. 
 \end{enumerate}
@@
 \begin{enumerate}[label=(\alph*)]
-	\item massive file update, which can reveal ransomware or similar attacks and be identified by MOS schemes based on covariance analysis; 
+	\item massive file update, which can reveal ransomware or similar attacks and be identified by MOS schemes based on zero mean covariance analysis; 
 	\item low number of file update into several folders, characterized by sparse operations that can reveal unreachable navigation performed by automated file accesses in order to avoid the massive file access characterization.
-Malicious sparse file accesses can be identified by MOS schemes base correlation analysis.
+Malicious sparse file accesses can be identified by MOS schemes based on zero mean and standardized covariance analysis.
 \end{enumerate}
@@
 \begin{enumerate}
-	\item massive data leakage or similar attacks, which can be identified by MOS schemes based on covariance analysis;
+	\item massive data leakage or similar attacks, which can be identified by MOS schemes based on zero mean covariance analysis;
 	\item low number of file download from several folders, characterized by sparse operations, which can reveal unreachable navigation performed by automated file download in order to avoid the massive file download characterization.
-Malicious sparse file download can be identified by MOS schemes base correlation analysis.
+Malicious sparse file download can be identified by MOS schemes based on zero mean and standardized covariance analysis.
 \end{enumerate}
@@
 \begin{enumerate}
-	\item massive file upload, similar to ransomware attack, which can be identified by MOS schemes based on covariance analysis; 
+	\item massive file upload, similar to ransomware attack, which can be identified by MOS schemes based on zero mean covariance analysis; 
 	\item low number of file upload to several folders, characterized by sparse operations, which can reveal unreachable navigation performed by automated file upload in order to avoid the massive file upload characterization.
-Malicious sparse file upload can be identified by MOS schemes base correlation analysis.
+Malicious sparse file upload can be identified by MOS schemes based on zero mean and standardized covariance analysis.
 \end{enumerate}
@@
 Each key sample generated contains the domain member code, key data and the sharing key value for each file sharing code.
-We may conclude that the proposed client is not overloaded with calculations due to the carefully selected mathematical operations.
+It is possible to conclude that the proposed client is not overloaded with calculations due to the carefully selected mathematical operations.
 It can be successfully used and provides acceptable level of security.
 
-The log analysis of the Log Analysis Center has been implemented and evaluated for offline anomaly detection in mobile clients, making it possible to apply anomaly detection techniques in a lightweight fashion, considering low processing requirements for deal with the resource constraints of mobile clients.
-The evaluation considered the required processing time for anomaly detection from log analysis, measuring the data modelling time, eigenvalue decomposition time and the required time for the EDC MOS scheme execution.
-The experiments were performed in two mobile devices, with variations of log size and window size.
-
-Table 4 presents the data modelling time and the processing time of eigenvalues decomposition and MOS calculations applied to anomaly detection from user operation logs of Storgrid mobile client.
-The results show that the lower window size leads to the larger eigenvalue decomposition time, but the largest eigenvalue decomposition time, which was 421 milliseconds, highlights an acceptable speed even for the worst evaluated scenario.
-
-Table 4. Data Modelling and Eigenvalue (Correlation and Covariance) Decomposition Time
-
-Table 5 presents the processing time of EDC MOS calculations applied to anomaly detection from user operation logs of Storgrid mobile client.
+The log analysis of the Log Analysis Center (LAC) has been implemented and evaluated for offline anomaly detection in mobile clients, making it possible to apply anomaly detection techniques in a lightweight fashion, considering low processing requirements for deal with the resource constraints of mobile clients.
+The evaluation considered the required processing time for anomaly detection from log analysis, measuring the data modeling time through the UAL, the eigenvalue decomposition time and the required time for the EDC MOS scheme execution, which is the scheme that requires less processing capacity and provides more anomaly identification accuracy \cite{da2009comparison,tenorio2013greatest}.
+The experiments were performed in two mobile devices, Galaxy GT-I9300 and Galaxy Tab SM-T800, with variations of log size and window size.
+The Galaxy GT-I9300 has Quad-core 1.4 GHz Cortex-A9 processor and 1 GB RAM, while the Galaxy Tab SM-T800 has its processing capacity composed by Quad-core 1.9 GHz Cortex-A15 and quad-core 1.3 GHz Cortex-A7, and 3 GB RAM.
+
+Table \ref{tab:04} presents the data modeling time and the processing time of eigenvalues decomposition calculations to be applied to anomaly detection from user operation logs of Storgrid mobile client.
+The information presented by column are the device model, the log size in megabytes, the window size in minutes, the data modeling time in milliseconds, the average of eigenvalue decomposition time in milliseconds, the standard deviation of eigenvalue decomposition time in milliseconds, the minimum of eigenvalue decomposition time in milliseconds and the maximum of eigenvalue decomposition time in milliseconds.
+
+The results show that the lower window size leads to the larger eigenvalue decomposition time, but the largest eigenvalue decomposition time, which was the maximum of 421 milliseconds with average of 347.42 milliseconds.
+This result highlights an acceptable speed even for the worst evaluated scenario, which is the Galaxy GT-I9300 processing 6MB with window size of 10 minutes.
+
+\begin{table*}
+	\caption{Data Modeling and Eigenvalue Decomposition Time }
+	\label{tab:04}
+	\centering
+	\begin{tabular}{|l|l|l|l|l|l|l|l|}
+		\hline \rowcolor{Gray} Device	& \begin{tabular}[x]{@{}l@{}}Log Size\\(MB)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Window\\(min)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Modeling\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Avg.
+Eig.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Std.
+Eig.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Eig.
+Min.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Eig.
+Max.\\(ms)\end{tabular}	\\ \hline
+		Galaxy GT-I9300	& 6	& 60	& 107	& 209.52	& 18.58	& 183	& 276	\\ \hline
+		Galaxy GT-I9300	& 6	& 40	& 115	& 227.26	& 18.13	& 191	& 289	\\ \hline
+		Galaxy GT-I9300	& 6	& 20	& 89	& 268.14	& 21.94	& 229	& 315	\\ \hline
+		Galaxy GT-I9300	& 6	& 10	& 90	& 347.42	& 24.11	& 304	& 421	\\ \hline
+		Galaxy GT-I9300	& 4.1	& 60	& 20	& 60.90	& 15.19	& 37	& 106	\\ \hline
+		Galaxy GT-I9300	& 4.1	& 40	& 20	& 68.72	& 15.71	& 43	& 114	\\ \hline
+		Galaxy GT-I9300	& 4.1	& 20	& 34	& 89.04	& 16.78	& 54	& 133	\\ \hline
+		Galaxy GT-I9300	& 4.1	& 10	& 21	& 117.24	& 14.36	& 96	& 171	\\ \hline
+		Galaxy GT-I9300	& 1.4	& 60	& 10	& 159.82	& 15.82	& 125	& 197	\\ \hline
+		Galaxy GT-I9300	& 1.4	& 40	& 10	& 168.06	& 15.90	& 139	& 220	\\ \hline
+		Galaxy GT-I9300	& 1.4	& 20	& 11	& 204.4	& 20.46	& 176	& 269	\\ \hline
+		Galaxy GT-I9300	& 1.4	& 10	& 13	& 259.00	& 21.34	& 220	& 315	\\ \hline
+		Galaxy Tab SM-T800	& 6	& 60	& 7	& 59.30	& 6.55	& 54	& 74	\\ \hline
+		Galaxy Tab SM-T800	& 6	& 40	& 8	& 62.56	& 7.05	& 56	& 80	\\ \hline
+		Galaxy Tab SM-T800	& 6	& 20	& 10	& 73.28	& 8.59	& 65	& 95	\\ \hline
+		Galaxy Tab SM-T800	& 6	& 10	& 8	& 93.48	& 9.13	& 83	& 130	\\ \hline
+		Galaxy Tab SM-T800	& 4.1	& 60	& 11	& 18.64	& 4.51	& 16	& 38	\\ \hline
+		Galaxy Tab SM-T800	& 4.1	& 40	& 11	& 19.64	& 5.12	& 17	& 38	\\ \hline
+		Galaxy Tab SM-T800	& 4.1	& 20	& 12	& 25.12	& 5.55	& 21	& 46	\\ \hline
+		Galaxy Tab SM-T800	& 4.1	& 10	& 12	& 32.32	& 7.29	& 27	& 55	\\ \hline
+		Galaxy Tab SM-T800	& 1.4	& 60	& 4	& 49.08	& 6.01	& 42	& 62	\\ \hline
+		Galaxy Tab SM-T800	& 1.4	& 40	& 5	& 51.42	& 7.36	& 44	& 74	\\ \hline
+		Galaxy Tab SM-T800	& 1.4	& 20	& 5	& 51.12	& 7.80	& 54	& 91	\\ \hline
+		Galaxy Tab SM-T800	& 1.4	& 10	& 7	& 75.24	& 7.71	& 65	& 90	\\ \hline
+	\end{tabular}
+\end{table*}
+
+Table \ref{tab:05} presents the processing time of EDC MOS calculations applied to anomaly detection from user operation logs of Storgrid mobile client.
+Table \ref{tab:05} respectively presents the device model, the log size in megabytes, the window size in minutes, the average of EDC calculation time in milliseconds, the standard deviation of EDC calculation time in milliseconds, the minimum of EDC calculation time in milliseconds and the maximum of EDC calculation time in milliseconds.
+
+\begin{table*}
+	\caption{EDC MOS scheme processing time for anomaly detection}
+	\label{tab:05}
+	\centering
+	\begin{tabular}{|l|l|l|l|l|l|l|l|}
+		\hline \rowcolor{Gray} Device	& \begin{tabular}[x]{@{}l@{}}Log Size\\(MB)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Window\\(min)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Avg.
+EDC.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Std.
+EDC.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Min.
+EDC.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Max.
+EDC.\\(ms)\end{tabular}	\\ \hline
+		Galaxy GT-I9300	& 6	& 60	& 5.27	& 4.04	& 3	& 20	\\ \hline
+		Galaxy GT-I9300	& 6	& 40	& 10.78	& 6.37	& 6	& 34	\\ \hline
+		Galaxy GT-I9300	& 6	& 20	& 32.62	& 12.44	& 21	& 88	\\ \hline
+		Galaxy GT-I9300	& 6	& 10	& 115.08	& 17.45	& 88	& 158	\\ \hline
+		Galaxy GT-I9300	& 4.1	& 60	& 5.68	& 4.18	& 3	& 23	\\ \hline
+		Galaxy GT-I9300	& 4.1	& 40	& 10.76	& 5.31	& 7	& 27	\\ \hline
+		Galaxy GT-I9300	& 4.1	& 20	& 37.58	& 10.30	& 23	& 61	\\ \hline
+		Galaxy GT-I9300	& 4.1	& 10	& 125.98	& 18.56	& 101	& 191	\\ \hline
+		Galaxy GT-I9300	& 1.4	& 60	& 4.92	& 3.49	& 3	& 17	\\ \hline
+		Galaxy GT-I9300	& 1.4	& 40	& 9.00	& 4.23	& 6	& 25	\\ \hline
+		Galaxy GT-I9300	& 1.4	& 20	& 30.14	& 9.21	& 19	& 62	\\ \hline
+		Galaxy GT-I9300	& 1.4	& 10	& 100.62	& 15.83	& 69	& 163	\\ \hline
+		Galaxy Tab SM-T800	& 6	& 60	& 1.84	& 0.65	& 1	& 3	\\ \hline
+		Galaxy Tab SM-T800	& 6	& 40	& 3.26	& 1.24	& 2	& 7	\\ \hline
+		Galaxy Tab SM-T800	& 6	& 20	& 10.90	& 2.40	& 9	& 21	\\ \hline
+		Galaxy Tab SM-T800	& 6	& 10	& 41.86	& 7.33	& 34	& 60	\\ \hline
+		Galaxy Tab SM-T800	& 4.1	& 60	& 1.85	& 0.60	& 1	& 3	\\ \hline
+		Galaxy Tab SM-T800	& 4.1	& 40	& 3.62	& 1.10	& 2	& 8	\\ \hline
+		Galaxy Tab SM-T800	& 4.1	& 20	& 12.04	& 2.79	& 9	& 22	\\ \hline
+		Galaxy Tab SM-T800	& 4.1	& 10	& 40.16	& 6.48	& 35	& 60	\\ \hline
+		Galaxy Tab SM-T800	& 1.4	& 60	& 1.98	& 0.89	& 1	& 6	\\ \hline
+		Galaxy Tab SM-T800	& 1.4	& 40	& 3.30	& 1.16	& 2	& 7	\\ \hline
+		Galaxy Tab SM-T800	& 1.4	& 20	& 10.48	& 2.90	& 8	& 21	\\ \hline
+		Galaxy Tab SM-T800	& 1.4	& 10	& 34.52	& 4.08	& 30	& 45	\\ \hline
+	\end{tabular}
+\end{table*}
+
 It is possible to observe that the processing time increases with the window size decreasing, similar to the results for eigenvalue decomposition time.
-
-Table 5. EDC MOS scheme processing time for anomaly detection
-
-The longest processing time measured is lower than 200 milliseconds, even considering window size of 10 minutes or processing 6 MB of user operation log, what represents a acceptable processing time for anomaly detection in mobile devices.
+The longest processing time measured is lower than 200 milliseconds, even considering window size of 10 minutes or processing 6 MB of user operation log.
+This result represents an acceptable processing time for anomaly detection in mobile devices.
 
@@
 Moreover, this proposed approach also prevents malicious user behavior by applying a MOS-based analytic method.
+
 As prove of concept, a fully working mobile application was developed to test the proposed security solution and acquired results provide evidence that besides achieving the desired security features, the solution also has positive results in terms of performance.
 This fact is due to the usage of lightweight operations and the optimized combination of the selected security methods.
+The proposed approach is a practical application to be used in the corporate mobile environment.
+It is implemented as a fully working mobile app and can be used for any type of enterprise.
+Also, part of concept is seed for new security solutions for big data apps.
+
 Future works in the area can further explore enhancements in the analytics methods as well as to extended the approach to be used by mobile devices with even more severe resources constraints.
-
-Our approach is a practical application to be used in the corporate mobile environment.
-It is implemented as a fully working mobile app.
-WE suggest that it can be used for any type of enterprise.
-Also, part of concept is sed in the big data apps.
-We plan to enhance the analytics and implement the effective protocols.
 

### 2016-08-26 lab/src/main/matlab/dofba_tensor.m

@@ -1,0 +1,41 @@
+function Z = dofba_tensor(X,dim)
+
+% DOFBA_TENSOR   Apply forward-backward averaging to a measurement tensor
+%
+% Syntax:
+%   Z = DOFBA_TENSOR(X,dim)
+% 
+% Input:
+%   X - measurement tensor of size M_1 x M_2 x ... x M_R x N
+%   dim - Number of dimensions in X. Optional, defaults to NDIMS(X)-1. For
+%   an R-D problem, dim must be equal to R+1.
+%
+% Output:
+%   Z - forward-backward averaged measurement tensor of 
+%       size M_1 x M_2 x ... x M_R x 2N
+%
+%
+% Note:
+%    If N = 1, then dim needs to be specified otherwise DOFBA_TENSOR will
+%    produce wrong results. Since singleton dimensions are ignored the
+%    number of dimensions will otherwise be wrong.
+
+S = size(X);
+if nargin > 1
+    if length(S) < dim
+        S = [S, ones(1,dim - length(S))];
+    end
+end    
+R = length(S)-1;
+M = S(1:R);
+N = S(R+1);
+
+
+thesubs_r = cell(1,R+1);
+for r = 1:R
+    thesubs_r{r} = M(r):-1:1;
+end
+thesubs_r{R+1} = N:-1:1;
+Z = cat(R+1,X,subsref(conj(X),struct('type','()','subs',{thesubs_r})));
+    
+% alternative: X = invunfold(fba_2d(unfold(X,3).'),3,M(1)) <- slower!

### 2016-08-26 lab/src/main/matlab/tensorlab3/demo_basic.m

@@ -1,0 +1,164 @@
+%% Basic usage of Tensorlab demo
+%    Tensorlab is a Matlab package for complex optimization and tensor
+%    computations. This demo will discuss the basics of Tensorlab. It
+%    consists of three consecutive parts. A first section will explain how
+%    a tensor can be defined and visualized. The second section handles the
+%    elementary use of the ``cpd`` command, while the third section
+%    discusses more advanced items such as customized initialization
+%    methods and algorithms.
+%
+%    A more detailed discussion can be found <a href="matlab:
+%    web('http://www.tensorlab.net/demos/basic.html', '-browser')">online</a>.
+%
+%    See also cpd, cpderr, cpd_nls, mlsvd, lmlra
+
+% Authors: Otto Debals         (Otto.Debals@esat.kuleuven.be)
+%          Lieven De Lathauwer (Lieven.DeLathauwer@kuleuven-kulak.be)
+%
+% Version History:
+% - 2016/03/26   NV      Added visualize
+% - 2015/07/12   OD      Initial version
+
+clear variables; close all; clc
+rng('default'); rng(50)
+
+fprintf(['Tensorlab is a Matlab package for complex optimization and tensor\n' ...
+    'computations. This demo will discuss the basics of Tensorlab. It\n' ...
+    'consists of three consecutive parts. A first section ''Tensor\n' ...
+    'construction and visualization'' will explain how a tensor can be\n' ...
+    'defined and visualized. The second section ''Canonical polyadic\n' ...
+    'decomposition for beginners'' handles the elementary use of the\n' ...
+    'cpd command, while the third section ''CPD for pros'' discusses more\n' ...
+    'advanced items such as customized initialization methods and algorithms.\n']);
+fprintf('An accompanying text can be found %s.\n\n',gendemolink('basic','online'));
+fprintf(['Go to <a href="matlab: web(''http://www.tensorlab.net'', ''-browser'')">www.tensorlab.net</a> for Tensorlab\''s documentation and further demos.\n' ...
+    '--------------------------------------------------------------------------\n\n']);
+
[…]

### 2016-08-26 lab/src/main/matlab/tensorlab3/gendemolink.m

@@ -1,0 +1,12 @@
+function url = gendemolink(name, text)
+%GENDEMOLINK Generate a weblink for a demo
+    
+% Author(s): Nico Vervliet       (Nico.Vervliet@esat.kuleuven.be)
+%
+% Version History:
+% - 2015/07/10   NV      Initial version
+    
+    baseurl = 'http://www.tensorlab.net';
+    url = sprintf('<a href="matlab: web(''%s/demos/%s.html'', ''-browser'')">%s</a>', ...
+                  baseurl, name, text);
+end

### 2016-08-26 ransomdroid/app/src/main/java/br/unb/lasp/ransomdroid/AsymmetricAlgorithmRSA.java

@@ -1,0 +1,74 @@
+package br.unb.lasp.ransomdroid;
+
+import android.app.Activity;
+import android.os.Bundle;
+import android.util.Base64;
+import android.util.Log;
+import android.widget.TextView;
+
+import java.security.Key;
+import java.security.KeyPair;
+import java.security.KeyPairGenerator;
+
+import javax.crypto.Cipher;
+
+import br.unb.ransomdroid.R;
+
+public class AsymmetricAlgorithmRSA extends Activity {
+    static final String TAG = "AsymmetricAlgorithmRSA";
+
+    TextView originalTextView;
+    TextView encodedTextView;
+    TextView decodedTextView;
+
+    @Override
+    public void onCreate(Bundle savedInstanceState) {
+        super.onCreate(savedInstanceState);
+
+        setContentView(R.layout.activity_cryptest);
+
+        // Original text
+        String originalText = "This is just a simple test!";
+        originalTextView = (TextView) findViewById(R.id.original);
+        originalTextView.setText("\n[ORIGINAL]:\n" + originalText + "\n");
+
+        // Generate key pair for 1024-bit RSA encryption and decryption
+        Key publicKey = null;
+        Key privateKey = null;
+        try {
+            KeyPairGenerator keygen = KeyPairGenerator.getInstance("RSA");
+            keygen.initialize(1024);
+            KeyPair keyPair = keygen.genKeyPair();
+            publicKey = keyPair.getPublic();
+            privateKey = keyPair.getPrivate();
+        } catch (Exception e) {
+            Log.e(TAG, "RSA key pair error");
+        }
+
+        // Encode the original data with RSA private key
+        byte[] encodedBytes = null;
+        try {
+            Cipher rsaCipher = Cipher.getInstance("RSA");
+            rsaCipher.init(Cipher.ENCRYPT_MODE, privateKey);
+            encodedBytes = rsaCipher.doFinal(originalText.getBytes());
+        } catch (Exception e) {
+            Log.e(TAG, "RSA encryption error");
+        }
+        encodedTextView = (TextView) findViewById(R.id.encoded);
+        encodedTextView.setText("[ENCODED]:\n" + Base64.encodeToString(encodedBytes, Base64.DEFAULT) + "\n");
+
+        // Decode the encoded data with RSA public key
+        byte[] decodedBytes = null;
+        try {
[…]

### 2016-08-28 docs/papers/2016JGC/2016JGC.tex
primeiros passos para remocao da parte online o paper

@@
 \begin{abstract}
-Preventing data leakage on the mobile device is a crucial security problem.
+Preventing data leakage on the mobile app is a crucial security problem.
 Therefore, additional control and protection measures should be taken for the confidential data on the mobile devices that leave the boundaries of the organization.
-In this scenario, such devices can be used on the offline mode for the convenience and traffic reduce or even for a particular characteristic of the app or of the business.
-This paper presents a novel approach to the security of the corporate mobile devices, in particular the offline mode.
+In this scenario, such apps can be used on the offline mode for the convenience and traffic reducing or even for a particular characteristic of the app or of the business.
+This paper presents a novel approach to the security of the corporate mobile apps, in particular the offline mode.
 The important part of the presented approach is the conceptualization and the definition of the core methodology to solve the problem of offline mobile security, i.e.
 the protection of the confidential data in use when the mobile client is not connected to the corporate cloud.
-The protection of the sensitive data is provided by the combination of the cryptographic methods, such as AES file encryption, ABE authorization based both on user and share attributes, user key secret sharing (SS) based protection between the device and the user as well as MOS-based analytics methods to prevent the malicious user behavior.
-The proposed security architecture supports the basic mobile device protection principles: minimized traffic load and reduced communication with the cloud; usage of the light-weighted operations and an optimized combination of the security methods.
+The protection of the sensitive data is provided by the combination of the cryptographic methods and analytics methods to detect malicious user behavior.
+The proposed security architecture supports the basic mobile app protection principles: minimized traffic load and reduced communication with the cloud; usage of the light-weighted operations and an optimized combination of the security methods.
 \end{abstract}
 
-\keywords{ mobile security \and offline mode ABE \and Model Order Selection (MOS) \and user behavior analysis \and secret sharing \and protected cloud }
+\keywords{ offline mobile cloud security \and offline ABE \and Model Order Selection (MOS) \and user behavior analysis \and secret sharing}
 
@@
 
-Additional security issues and requirements have to be considered when mobile devices are actively used in corporate cloud environment \cite{yovel2014}.
+Additional security issues and requirements have to be considered when mobile apps are actively used in corporate cloud environment \cite{yovel2014}.
 Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm.
@@
 For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
-Observing the above described landscape, this paper outlines the concept of the offline mobile client security.
-This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing \cite{galibus2007} and ABE encryption \cite{goyal2006attribute}.
-Moreover, our proposed approach proposal includes the usage of the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to highlight possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack.
+Observing the above described landscape, this paper outlines the concept of the offline mobile cloud security.
+This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing (SSS) \cite{galibus2007} and ABE encryption \cite{goyal2006attribute}.
+Moreover, our proposed approach includes the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack.
 The key expiration period is safely incorporated into the proposed system solution in order to enhance security.
@@
 The paper is structured as follows.
-Section 2 analyzes the most common security problems in the mobile cloud environment and the proposed solutions for them related to the offline protection in the BYOD world.
-Section 3 outlines the basic processes and the module security infrastructure of the proposed solution as well as the outline of online mode security.
-Section 4 presents the detailed scheme of the proposed solution to the problem of offline mobile client security.
-Section 5 provides the schemes of the main security blocks including encryption and key usage hierarchies as well as explanations of the core security methods: SSS, ABE and MOS.
+Section \ref{sec_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions related to the offline protection in the BYOD world.
+Section \ref{sec_the_client_security_architecture} outlines the basic processes and the module security infrastructure of the proposed solution.
+Section \ref{sec_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security.
+Section \ref{sec_algorithms} provides the schemes of the main security blocks including encryption and key usage hierarchies as well as explanations of the core security methods: SSS, ABE and MOS.
 Section 6 discusses the security proofs including the outline of adversary model, common threat scenarios and the proposed analytics to discover abnormal usage.
@@
 The most commonly used are:
-Using geolocation of the device to trace its usage;
-Setting up the expiration period of an app;
-Secured transfer politics between apps;
-Restricting access to the corporate app;
-Setting up the expiration period of app pass pin;
-Setting up the counter of failed tries;
-Restricted or prohibited offline access;
-Logging and auditing.
+
+\begin{enumerate}
+	\item Using geolocation of the device to trace its usage;
+	\item Setting up the expiration period of an app;
+	\item Setting up the expiration period of app pass pin;
+	\item Setting up the counter of failed tries;
+	\item Secured transfer politics between apps;
+	\item Restricting access to the corporate app;
+	\item Restricted or prohibited offline access;
+	\item Logging and auditing.
+\end{enumerate}
+
 Preventing data leakage on the mobile device is a crucial security problem.
-Therefore, administrators should take additional control and protection measures for the confidential data on the mobile devices that leave the boundaries of the organization.
+Therefore, it is necessary to take additional control and protection measures for the confidential data on the mobile devices that leave the boundaries of the organization.
 Generally, the most sensitive and confidential data should not be permitted to be transferred to the mobile device.
-However, what if the SME need to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reduce or even for a particular characteristic of the app or the business itself?
+However, what if the SME need to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reducing, or even for a particular characteristic of the app or the business itself?
+
 From the theoretical point of view of this problem, there are several surveys, whose common point is the mobile cloud computing as a rapidly developing paradigm that poses many security and complexity problems \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}.
@@
 Commonly, all the models and protection schemes concentrate on the encryption properties and either perform the computations on their own \cite{zhao2010trusted, yang2011provable} or use the cloud provider to off-load the cryptographic operations \cite{itani2010energy, ren2011lightweight}.
-Obviously, it is natural the mobile device cannot handle all operations securely without the assistance of a cloud provider, due to resources constraint and battery limitation.
-The necessity to use schemes that function without putting load on a provider arises when it is desired to make the device less dependent on the cloud provider, i.e.
+Obviously, it is natural the mobile app cannot handle all operations securely without the assistance of a cloud provider, due to resources constraint and battery limitation.
+
+The necessity to use schemes that function without putting load on a provider arises when it is desired to make the app less dependent on the cloud provider, i.e.
 corporate app continues to provide the secure access to the sensitive data without connection to the network.
@@
 Therefore, according to \cite{khan2015cloud, tenorio2013greatest, khan2013towards}, the state-of-the-art mobile cloud security models do not consider the problem of the offline security mode.
+
 In many cases the industrial providers of the mobile security API avoid the problem by completely forbidding the offline access to the protected app, i.e.
@@
 To the best of our knowledge, the offline mode security problem has not yet been deployed, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}.
-Therefore, the main concern of this proposal is the protection of the device and app in offline mode when the functions of data protection cannot be offloaded to a cloud or a trusted party.
-
-\section{The client security architecture}
+Therefore, the main concern of this proposal is the protection of the app and its data in offline mode, when the functions of data protection cannot be offloaded to a cloud or a trusted party.
+
+\section{The mobile security architecture}
 \label{sec_the_client_security_architecture}
 The approach proposed in this work describes and implements the complete lifecycle of the mobile app security infrastructure.
-The security processes depend on the key expiry period, and are used to access the protected storage.
+The mobile security processes depend on the key expiry period, and are used to access the protected storage.
 Once the user keys expire, the user is requested to enter his valid credentials, i.e.
@@
 	\centering
-	\includegraphics[width=7cm]{figures/coresetoffunctionsandprotocols.eps}
-	\caption{The core set of functions and protocols of the mobile app security infrastructure}
+	\includegraphics[width=5cm]{figures/coresetoffunctionsandprotocols.eps}
+	\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
 	\label{fig:1}
@@
 The cryptographic functions include the decryption (for AES and ABE) key restoration for SSS and EKE token generation procedure.
-While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode and the EKE generation and executions is used in the online mode.
-These operations are described in Subsections \ref{sec_aes}-\ref{sec_secret} and are part of the Figure \ref{fig:1}, related to online and offline mode operations.
-
-The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and more elaborate MOS-inspired statistics analyzer.
+While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode, while the EKE generation and executions is used in the online mode.
+These operations are described in Subsections \ref{sec_aes}-\ref{sec_secret} and are part of the Figure \ref{fig:1}.
+
+The threat intelligence infrastructure takes into account simple actors such as the time counter for the key expiry period, the counter of unsuccessful tries in order to protect from brute force attacks, and MOS-inspired statistics analyzer.
 Functions such as alerting and deleting the expired key belong to this block as well.
@@
 
-The communication with server includes the separate sender and receiver to check the user credentials and receiver new keys which acts based on the EKE protocol.
+The communication with server includes the separate sender and receiver to check the user credentials and receive new keys, which acts based on the EKE protocol.
 The remaining data like access list and files can be sent via unprotected channels.
-Subsection \ref{sec_online_mode} briefly discusses the communication with the server, but it is also an intention to make it a subject of our next work.
-
-The storage is divided in two parts – temporary and permanent.
-While the files and file keys are stored permanently in order to reduce unnecessary traffic and the resource usage on client, the ABE keys and the key storage protection data are temporary.
-The unprotected files stored in the mobile client app memory are not the subject of the security architecture.
-
-\subsection{Online mode}
-\label{sec_online_mode}
-This subsection briefly outlines the main processes and functions in the online mode.
-The connection to the network allows verifying the user credentials with the help of the protected cloud server.
-The communication with the cloud is secured by means of the encrypted key exchange protocol \cite{hao2010j, hao2008password} and serves to receive the new set of keys, files and send the collected log information.
-The client app operations in the online mode are as represented in Figure 3.3.
-
-\begin{figure}[h!]
-	\includegraphics[width=8.6cm]{figures/onlinemodeoperations.eps}
-	\caption{Online Mode Operations}
-	\label{fig:3}
-\end{figure}
-
-The communication with server is necessary once the key is expired.
-This allows optimizing the load on the network once the traffic is reduced to the discrete sessions.
-In such environment it is important to keep track of the time synchronization both on server and on client in parallel in order to avoid malicious user behavior messing up with time attacks.
-
-As depicted in Figure \ref{fig:4}, the operation of the online mode starts from the client app sending a token to server by means of EKE: DEV\_PASS + TIME + PASS + PIN in order to validate his identity and the identity of the user to the server.
-This value serves as a proof that the device and the user are the ones with which server communicated previously.
-After performing the verification the server sends the new calculated DEV\_PASS and TIME to the mobile app by means of established EKE session.
-Afterwards the server sends the new encrypted key set and encrypted files to the client app in the clear.
-
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=8.6cm]{figures/onlineauthorizationworkflow.eps}
-	\caption{Online Authorization Workflow}
-	\label{fig:4}
-\end{figure}
-
-This model proposes to use light-weighted EKE for the fast communication with server communication like J-PAKE \cite{hao2010j, hao2008password} or SIS-based EKE \cite{bogos2013primeless}.
-SIS is a public key encryption that avoids generating the big primes, thus it can be used for the secure key exchange \cite{bogos2013primeless,bellovin1992encrypted}.
-EKE is performed as soon as the key is expired and the client asks the user to synchronize with the server.
-Note that the EKE is not required to send the encrypted files or statistics.
-The current implementation of mobile app uses J-PAKE protocol.
-We refer to \cite{hao2010j, hao2008password, toorani2014security} for a complete description and analysis of the J-PAKE protocol.
-
-\section{Offline mode: our proposed model and solution}
+
+\section{The proposed solution for offline mobile security}
 \label{sec_offline_mode}
-This section proposes an open model of mechanisms for the mobile device protection in which the security is supported both in online and offline modes.
-Currently and to the best of our knowledge, the systems of mobile device protection follow a model where the protected mobile client can operate only when it is connected to the cloud, which is not always convenient for the end-user.
-The basic principles of the mobile device protection herein proposed are:
-
-\begin{enumerate}
-	\item Optimized communication with the cloud when the device does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
+This section proposes an approach for the mobile app protection in which the security is supported both in online and offline modes.
+Currently and to the best of our knowledge, the systems of mobile app protection follow a model where the protected mobile app can operate only when it is connected to the cloud, which is not always convenient for the end-user.
+The basic principles of the mobile app protection herein proposed are:
+
+\begin{enumerate}
+	\item Optimized communication with the cloud when the app does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
 	\item Optimized combination of the security mechanisms so that the mobile client does not need to perform complex computation like encryption and key generation due to its resource constraint;
@@
 
-The most important security issues in the proposed model arise when the device goes to the offline mode and the user is still allowed to get the access to the protected SME documents.
+The most important security issues in the proposed model arise when the app goes to the offline mode and the user is still allowed to get the access to the protected SME documents.
 In this case, the server can neither monitor the user activity nor provide the protection methods.
@@
 
-In the online mode the mobile device uses the secure communication with the server in order to verify the validity of user’s credentials.
+In the online mode the mobile app uses the secure communication with the server in order to verify the validity of user’s credentials.
 On the contrary, the offline protection model should be approached independently.
@@
 	\item The proof is derived from the previous session, in order to verify that the user is still authorized;
-	\item The proof should not give access to user password, i.e. it should not be stored on the device;
+	\item The proof should not give access to user password, i.e. it should not be stored on the app;
 	\item The proof should be temporary and have an expiration period;
@@
 
-Such proof cannot be stored on the client device, as it is not possible to guarantee its protection in the offline mode.
+Such proof cannot be stored on the mobile app, as it is not possible to guarantee its protection in the offline mode.
 Therefore, the most effective way to secure the proof on the offline mode without performing complex computation is to share this proof between the client and the user in a protected manner.
 In this case, there is no need to store the function of the user password.
-The additional argument against the traditional password verification is the necessity to check the PIN, which is very small, so the construction of valid one-way function resilient to the offline dictionary attacks is a difficult task.
-
-The proposed offline mode architecture is represented in Figure \ref{fig:5}.
+The additional argument against the traditional password verification is the necessity to check the PIN, which is very small, so the construction of valid one-way function resilient to the offline dictionary attacks is difficult.
+
+The proposed architecture for offline mobile security is represented in Figure \ref{fig:5}.
 
@@
 	\includegraphics[width=8.6cm]{figures/offlinesecurityarchitecture.eps}
-	\caption{Online Authorization Workflow}
+	\caption{Proposed Architecture for Offline Mobile Security }
 	\label{fig:5}
@@
 \begin{enumerate}
-	\item \textbf{The protected storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files.
+	\item \textbf{Protected Storage}: the storage is protected with the shared user key and contains the ABE keys giving access to the file keys which allow decrypting the stored files.
 The outline of the hybrid encryption scheme is presented in the Section \ref{sec_algorithms}.
@@
 Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones.
-	\item \textbf{Key Management center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
+	\item \textbf{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
 \end{enumerate}
@@
 \label{sec_algorithms}
-In the proposed approach, the kernel encryption scheme in the mobile device is a combination of several methods of security.
+In the proposed approach, the kernel encryption scheme in the mobile app is a combination of several methods of security.
 The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption.
@@
 The proposed authentication system is based on the shared storing of the user key.
-Also, the device acts as a dealer in the SSS.
+Also, the mobile app acts as a dealer in the SSS.
 Using the SSS ensures that the key can only be accessed by an authenticated user.
@@
 
-The complexity analysis depends on the operation that the user and the mobile app perform in order to keep the device protected. In the offline mode the client app performs the following actions:
+The complexity analysis depends on the operation that the user and the mobile app perform in order to keep the app protected. In the offline mode the client app performs the following actions:
 
@@
 The Table \ref{tab:01} presents the results of mobile app security workflow testing.
-Table \ref{tab:01} demonstrates that the device decrypts the data with the acceptable speed.
+Table \ref{tab:01} demonstrates that the mobile app decrypts the data with the acceptable speed.
 One can compare the speed of ABE/AES decryption (keys and actual data) and the speed of the whole decryption process.
@@
 
-Table \ref{tab:02} demonstrates the results of testing the speed of the mobile device for receiving and displaying the list of files.
+Table \ref{tab:02} demonstrates the results of testing the speed of the mobile app for receiving and displaying the list of files.
 The app was tested for displaying up to 1000 files.
@@
 
-The device receives and displays the list of files depending on the user attributes with acceptable average speed and the proposed protection scheme is usable.
+The mobile app receives and displays the list of files depending on the user attributes with acceptable average speed and the proposed protection scheme is usable.
 Table \ref{tab:03} demonstrates the results of performance testing of the keys generation procedure on server:
@@
 The longest processing time measured is lower than 200 milliseconds, even considering window size of 10 minutes or processing 6 MB of user operation log.
-This result represents an acceptable processing time for anomaly detection in mobile devices.
+This result represents an acceptable processing time for anomaly detection in mobile apps.
 
@@
 \label{sec_conclusion}
-An important security issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile devices.
-Once a mobile device releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device.
+An important security issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile apps.
+Once a mobile app releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device.
 Aware of this problem and its importance, this work presented a proposal to address the offline mobile security problem combining a different cryptographic methods.

### 2016-09-01 docs/papers/2016JGC/2016JGC.tex
adicao de prototipo de tela splash

@@
 \begin{abstract}
-Preventing data leakage on the mobile app is a crucial security problem.
-Therefore, additional control and protection measures should be taken for the confidential data on the mobile devices that leave the boundaries of the organization.
-In this scenario, such apps can be used on the offline mode for the convenience and traffic reducing or even for a particular characteristic of the app or of the business.
-This paper presents a novel approach to the security of the corporate mobile apps, in particular the offline mode.
+Preventing data leakage on the mobile client is a crucial security problem.
+Therefore, additional control and protection should be taken for the confidential data on the mobile clients that leave the boundaries of the organization.
+In this scenario, such apps can be used on the offline mode for the convenience and traffic reducing or even for a particular characteristic of the mobile client or of the business.
+This paper presents a novel approach to the security of the corporate mobile clients, in particular the offline mode.
 The important part of the presented approach is the conceptualization and the definition of the core methodology to solve the problem of offline mobile security, i.e.
 the protection of the confidential data in use when the mobile client is not connected to the corporate cloud.
-The protection of the sensitive data is provided by the combination of the cryptographic methods and analytics methods to detect malicious user behavior.
-The proposed security architecture supports the basic mobile app protection principles: minimized traffic load and reduced communication with the cloud; usage of the light-weighted operations and an optimized combination of the security methods.
+The protection of the sensitive data is provided by the combination of the cryptographic and analytics methods to detect malicious user behavior.
+The proposed security architecture supports the basic mobile client protection principles: minimized traffic load and reduced communication with the cloud; usage of the light-weighted operations and an optimized combination of the security methods.
 \end{abstract}
@@
 
-Additional security issues and requirements have to be considered when mobile apps are actively used in corporate cloud environment \cite{yovel2014}.
+Additional security issues and requirements have to be considered when mobile clients are actively used in corporate cloud environment \cite{yovel2014}.
 Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm.
 The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS).
-While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, such as CASB and CAC, the corporate mobile app is usually light-weighted and generally less protected.
+While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, such as CASB and CAC, the corporate mobile client is usually light-weighted and generally less protected.
 The protection scheme used on a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}.
@@
 Suppose that a SME uses CASB in order to protect data at rest, i.e.
-while stored on the server-side, in transit, communicating with server, and, in use, while the client is connected to the network.
-In this case, when the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile app has to secure itself with its own limited resources.
+while stored on the server-side, in transit, communicating with server, and, in use, while the mobile client is connected to the network.
+In this case, when the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile client has to secure itself with its own limited resources.
 Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection.
@@
 Observing the above described landscape, this paper outlines the concept of the offline mobile cloud security.
-This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing (SSS) \cite{galibus2007} and ABE encryption \cite{goyal2006attribute}.
+
+This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing (SSS) \cite{galibus2007} and ABE encryption \cite{goyal2006attribute} for securing mobile clients.
 Moreover, our proposed approach includes the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack.
@@
 Section \ref{sec_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions related to the offline protection in the BYOD world.
-Section \ref{sec_the_client_security_architecture} outlines the basic processes and the module security infrastructure of the proposed solution.
+Section \ref{sec_the_client_security_architecture} outlines the mobile security architecture of the proposed solution.
 Section \ref{sec_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security.
-Section \ref{sec_algorithms} provides the schemes of the main security blocks including encryption and key usage hierarchies as well as explanations of the core security methods: SSS, ABE and MOS.
-Section 6 discusses the security proofs including the outline of adversary model, common threat scenarios and the proposed analytics to discover abnormal usage.
-Still in Section 6, complexity analysis and practical implementation results are included.
-Section 7 concludes the paper.
+Section \ref{sec_algorithms} explains the algorithms for the core security methods: AES, ABE, SSS and MOS.
+Section \ref{sec_results} discusses the security proofs including the outline of adversary model, the common threat scenarios, the data modeling, the complexity analysis and dicussion of the practical implementation results.
+Section \ref{sec_conclusion} concludes the paper.
 
@@
 A common practice is to provide additional contextual methods apart from authentication, DLP services, and encryption, which can be at rest, in transit and in use \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}.
-The contextual methods increase the security of the app at a maximum level with minimum resource requirements.
+The contextual methods increase the security of the mobile client at a maximum level with minimum resource requirements.
 The most commonly used are:
@@
 	\item Setting up the expiration period of an app;
-	\item Setting up the expiration period of app pass pin;
+	\item Setting up the expiration period of client pass pin;
 	\item Setting up the counter of failed tries;
@@
 Generally, the most sensitive and confidential data should not be permitted to be transferred to the mobile device.
-However, what if the SME need to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reducing, or even for a particular characteristic of the app or the business itself?
-
-From the theoretical point of view of this problem, there are several surveys, whose common point is the mobile cloud computing as a rapidly developing paradigm that poses many security and complexity problems \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}.
+However, what if the SME need to allow their employees to work on such devices and even use them on the offline mode for the convenience and traffic reducing, or even for a particular characteristic of the mobile client or the business itself?
+
+From the theoretical point of view of this problem, there are several surveys, whose common point is the mobile cloud as a rapidly developing paradigm that poses many security and complexity problems \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}.
 An analysis of the new models of mobile cloud computing and new ways of using data storage services is presented in \cite{khan2015cloud, khan2013towards}.
 Commonly, all the models and protection schemes concentrate on the encryption properties and either perform the computations on their own \cite{zhao2010trusted, yang2011provable} or use the cloud provider to off-load the cryptographic operations \cite{itani2010energy, ren2011lightweight}.
-Obviously, it is natural the mobile app cannot handle all operations securely without the assistance of a cloud provider, due to resources constraint and battery limitation.
-
-The necessity to use schemes that function without putting load on a provider arises when it is desired to make the app less dependent on the cloud provider, i.e.
-corporate app continues to provide the secure access to the sensitive data without connection to the network.
+Obviously, it is natural the mobile client cannot handle all operations securely without the assistance of a cloud provider, due to resources constraint and battery limitation.
+
+The necessity to use schemes that function without putting load on a provider arises when it is desired to make the mobile client less dependent on the cloud provider, i.e.
+corporate client continues to provide the secure access to the sensitive data without connection to the network.
 As discussed in \cite{khan2015cloud}, all the currently known schemes of encryption, performing the computation, either use a cloud provider \cite{zhao2010trusted}, a third party trusted agent \cite{yang2011provable} or a combination of both \cite{itani2010energy}.
@@
 In many cases the industrial providers of the mobile security API avoid the problem by completely forbidding the offline access to the protected app, i.e.
-SAP Mocona, which operates as a secure app wrapping layer \cite{Lelyveld2013}.
+SAP Mocona, which operates as a secure client wrapping layer \cite{Lelyveld2013}.
 Due to various constraints such as traffic load, travelling and ease of access, the SME business procedure may require such access.
 To the best of our knowledge, the offline mode security problem has not yet been deployed, neither in academia nor in the industry \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}.
-Therefore, the main concern of this proposal is the protection of the app and its data in offline mode, when the functions of data protection cannot be offloaded to a cloud or a trusted party.
+Therefore, the main concern of this proposal is the protection of the mobile client and its data in offline mode, when the functions of data protection cannot be offloaded to a cloud or a trusted party.
 
@@
 \label{sec_the_client_security_architecture}
-The approach proposed in this work describes and implements the complete lifecycle of the mobile app security infrastructure.
+The approach proposed in this work describes and implements the complete lifecycle of the mobile client security infrastructure.
 The mobile security processes depend on the key expiry period, and are used to access the protected storage.
@@
 PIN and password.
-The client app then sends the credentials to the server for verification.
+The mobile client then sends the credentials to the server for verification.
 Once the new set of access keys is received, the user can access the protected files in the offline mode, without the access to the server.
@@
 \label{sec_offline_mode}
-This section proposes an approach for the mobile app protection in which the security is supported both in online and offline modes.
-Currently and to the best of our knowledge, the systems of mobile app protection follow a model where the protected mobile app can operate only when it is connected to the cloud, which is not always convenient for the end-user.
-The basic principles of the mobile app protection herein proposed are:
-
-\begin{enumerate}
-	\item Optimized communication with the cloud when the app does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
+This section proposes an approach for the mobile client protection in which the security is supported both in online and offline modes.
+Currently and to the best of our knowledge, the systems of mobile client protection follow a model where the protected client can operate only when it is connected to the cloud, which is not always convenient for the end-user.
+The basic principles of the mobile client protection herein proposed are:
+
+\begin{enumerate}
+	\item Optimized communication with the cloud when the mobile client does not need to be constantly connected to the server due to the resource constraint and necessity to secure this communication;
 	\item Optimized combination of the security mechanisms so that the mobile client does not need to perform complex computation like encryption and key generation due to its resource constraint;
@@
 
-The most important security issues in the proposed model arise when the app goes to the offline mode and the user is still allowed to get the access to the protected SME documents.
+The most important security issues in the proposed model arise when the client goes to the offline mode and the user is still allowed to get the access to the protected SME documents.
 In this case, the server can neither monitor the user activity nor provide the protection methods.
@@
 
-In the online mode the mobile app uses the secure communication with the server in order to verify the validity of user’s credentials.
+In the online mode the mobile client uses the secure communication with the server in order to verify the validity of user’s credentials.
 On the contrary, the offline protection model should be approached independently.
@@
 
-Such proof cannot be stored on the mobile app, as it is not possible to guarantee its protection in the offline mode.
-Therefore, the most effective way to secure the proof on the offline mode without performing complex computation is to share this proof between the client and the user in a protected manner.
+Such proof cannot be stored on the client, as it is not possible to guarantee its protection in the offline mode.
+Therefore, the most effective way to secure the proof on the offline mode without performing complex computation is to share this proof between the mobile client and the user in a protected manner.
 In this case, there is no need to store the function of the user password.
@@
 Signal processing techniques have been successfully applied to anomaly detection \cite{lu2009network, huang2009signal} and have become a solution to a problem of improving detection accuracy, adaptability and computational cost for application on resource-constrained scenarios.
-Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior according to the scenarios in Section \ref{sec_results}.
+Therefore, signal processing can be applied in offline mobile client security, for evaluating anomalies on user's behavior, according to the scenarios in Section \ref{sec_results}.
 Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones.
@@
 
-The client app performs all the cryptographic calculations in the shadow.
-These calculations include the key storage, key restoring, decryption and showing the decrypted files in the client area.
-Note that the client app does not check the password or PIN validity as it does not hold the verification proof for the above.
-Also, the app keeps track of the user activities and the key expiry.
-The complete list of the app activities is presented in the Figure \ref{fig:6}.
+The mobile client performs all the cryptographic calculations in the shadow.
+These calculations include the key storage, key restoring, decryption and showing the decrypted files in the mobile client area.
+Note that the mobile client does not check the password or PIN validity as it does not hold the verification proof for the above.
+Also, the mobile client keeps track of the user activities and the key expiry.
+The complete list of the mobile client activities is presented in the Figure \ref{fig:6}.
 
@@
 The core cryptographic module of the implemented solution is based on the combination of AES-ABE-SSS methods, described in the Section \ref{sec_algorithms}.
-The key feature of the offline security is that the client app does not actually store any part of the user password to be verified.
-The client app combines its own key with the user share (PIN and password-derived) in order to restore the initial KEY\_SET\_KEY.
-If the user provides the wrong share the client will not be able to recognize it, but will decrypt the files incorrectly.
+The key feature of the offline security is that the mobile client does not actually store any part of the user password to be verified.
+The mobile client combines its own key with the user share (PIN and password-derived) in order to restore the initial KEY\_SET\_KEY.
+If the user provides the wrong share the mobile client will not be able to recognize it, but will decrypt the files incorrectly.
 Additionally, the password entering is tracked and too many tries in a short time are considered a threat.
@@
 
-In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode.
-Figure \ref{fig:8} depicts the Threat Intelligence Manager for offline behavioral analysis.
+In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode.
+Figure \ref{fig:8} depicts the TIM for offline behavioral analysis.
 
@@
 
-As depicted in Figure \ref{fig:8}, users request operations are logged so that the main features can be monitored in the client app.
+As depicted in Figure \ref{fig:8}, users request operations are logged so that the main features can be monitored in the mobile client.
 The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats.
-Therefore, the user operations are monitored by the client app, which sends the information of the captured operations to the User Activity Logging (UAL).
+Therefore, the user operations are monitored by the client app, which sends the information to the User Activity Logging (UAL).
 
 The UAL is responsible for the incremental logging, feature extraction and data modeling for malicious behavior detection, through the Log Analysis Center (LAC).
-The logged information shall be decomposed into selected features and modeled as matrices, composed of the number of occurrences of the selected features by its location and by the occurrence time.
+The logged information shall be decomposed into selected features and modeled as matrices, composed of the number of occurrences of the selected features by its location and by time.
 The resultant data is submited to the LAC, for anomaly detection.
@@
 \label{sec_algorithms}
-In the proposed approach, the kernel encryption scheme in the mobile app is a combination of several methods of security.
+In the proposed approach, the kernel encryption scheme in the mobile client is a combination of several methods of security.
 The files are encrypted with 128/256-bit AES, while the permanent file keys are encrypted with the attribute-based encryption.
@@
 
-The key hierarchy can be explained by Figure \ref{fig:10} where the key represented in Figure \ref{fig:9} serves for decrypting the key for each level represented in Figure \ref{fig:10}.
+The key hierarchy is explained by Figure \ref{fig:10} where the key represented in Figure \ref{fig:9} serves for decrypting the key for each level represented in Figure \ref{fig:10}.
 
@@
 The protected files are encrypted on the server side with the secure 128/256 bit AES encryption, which is currently an industrial standard.
-Nevertheless, domain administrator can choose to encrypt the file storage with Blowfish or Serpent which also provide a high level of safety \cite{nist2000}.
+Nevertheless, it is possible to encrypt the file storage with Blowfish or Serpent which also provide a high level of safety \cite{nist2000}.
 
@@
 First there is a single AES key for preserving the server storage.
-Second, when the encrypted file is sent from the server to the mobile app or desktop device another unique AES key is generated.
-This key serves for the encryption on the client side.
+Second, when the encrypted file is sent from the server to the mobile client another unique AES key is generated, which serves for the encryption on the client side.
 In other words, along with the encrypted file the user gets his own unique AES key for decrypting the files received.
@@
 The model supports the simple selective ABE scheme \cite{galibus2015cloud,galibus2014}.
-The selective scheme for attribute-based encryption is as follows: if at least one attribute in the set $\{t_i\}_M$ is equal to the attribute in the set{ti}M, the corresponding user U can decrypt the text M.
+The selective scheme for attribute-based encryption is as follows: if at least one attribute in the set $\{t_i\}_U$ is equal to the attribute in the set $\{t_i\}_M$, the corresponding user $U$ can decrypt the text $M$.
 In other words, as soon as the user and share have one attribute in common – the user can get access to the share.
@@
 
-Since the underlying sharing scheme is perfect \cite{galibus2008some}, hence the adversary cannot get any information of the KEY\_SET\_KEY unless he possessing all 4 key parts.
-Here the values PASS and PIN are predefined similar to the construction in \cite{galibus2015mobile}.
+The adversary cannot get any information of the KEY\_SET\_KEY unless he possessing all 4 key parts. Here the values PASS and PIN are predefined similar to the construction in \cite{galibus2015mobile}.
 
 The proposed authentication system is based on the shared storing of the user key.
-Also, the mobile app acts as a dealer in the SSS.
+Also, the mobile client acts as a dealer in the SSS.
 Using the SSS ensures that the key can only be accessed by an authenticated user.
@@
 Model Order Selection is an effective signal processing technique for several applications, allowing separating the only noise components from the principal components applying a rank reduction of the data.
+
 Applying MOS to the analysis of user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session.
-
-MOS for threat intelligence requires that the target features, which can be user operations, are modeled as a matrix composed by the number of occurrences by location and time, and grouped into $Q$ time frames.
+MOS for threat intelligence requires that the target features, which can be user operations, should be modeled as a matrix composed by the number of occurrences by location and time, and grouped into $Q$ time frames.
 Therefore, the framework considers the time variations of the matrix $\mathbf{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the occurrence of malicious behaviors.
-For example, one element of $\mathbf{X}^{(q)}$ can represent the number of file readings on folder $m$ during the minute $n$, from file operations logged by the client app.
+For example, one element of $\mathbf{X}^{(q)}$ can represent the number of file readings on folder $m$ during the minute $n$, from file operations logged by the mobile client.
 
@@
 Once the $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ have been obtained for MOS in order to anomaly detection, for the sake of simplicity, we refer to $\mathbf{\hat{R}}_{yy}$ or $\mathbf{\hat{R}}_{zz}$ as a matrix $\mathbf{C}$.
-Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\mathbf{C}^{(q)} = \mathbf{V}^{(q)}\mathbf{\Lambda}^{(q)}\mathbf{V}^{(q)^{\rm T}},$, in order to obtain the vector of eigenvalues $e$, as following:
+Therefore, the next step of the algorithm is the eigenvalue decomposition (EVD), calculated according to $\mathbf{C}^{(q)} = \mathbf{V}^{(q)}\mathbf{\Lambda}^{(q)}\mathbf{V}^{(q)^{\rm T}}$, in order to obtain the vector of eigenvalues $e$, as following:
 
@@
 \begin{enumerate}
-	\item Steal the user keys one-by-one while the app performs decryption, if he is outsider;
-	\item Use expired key to decrypt the files, if he is insider;
-	\item Does not have possibility to perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the client app;
+	\item Steal the user keys one-by-one while the mobile client performs decryption, if the attacker is outsider;
+	\item Use expired key to decrypt the files, if the attacker is insider;
+	\item Does not have possibility to perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the mobile client;
 	\item Steal the encrypted files but not the decrypted ones.
@@
 
-\textbf{Offline Mode}\\
-
-In the offline mode the client does not store the user password (see Section \ref{sec_offline_mode_workflow} for details), i.e.
+In the offline mode the mobile client does not store the user password (see Section \ref{sec_offline_mode_workflow} for details), i.e.
 no information about the password leaks, and therefore there is no possibility for the malicious user to check if the password he is trying to enter is correct or not.
-The only possible scenario for the information leakage in the case of the malicious outsider is if:
+The possible scenario for the information leakage in the case of the malicious outsider is if:
 
@@
 The temporary nature of all parameters obliges the user to connect to server when necessary and prevents the malicious actions from the user side.
-The only possible scenario when malicious or expired user wishes to prolong his old credentials is:
+The possible scenario when malicious or expired user wishes to prolong his old credentials is:
 
@@
 	\item He has to steal the protected FILE\_KEY (also synchronized) and encrypted file;
-	\item He has to do everything without the client (because the client checks the TIME and renews the PUBLIC\_SHARE\_KEY).
-\end{enumerate}
-
-Basically, for a malicious user there is practically no way to use the client with the old keys.
+	\item He has to do everything without the mobile client (because the client checks the TIME and renews the PUBLIC\_SHARE\_KEY).
+\end{enumerate}
+
+Basically, for a malicious user there is practically no way to use the mobile client with the old keys.
 The fact that the client does not contain any data to be checked or validated does not prevent the user from seeing the contents of decrypted files, but with the wrong password the decrypted files will be different from the original ones.
 
-The client still has to count the tries (to prevent the hacker to perform brute force attack) within one session.
+The mobile client still has to count the tries (to prevent the hacker to perform brute force attack) within one session.
 In accordance with the conditions 2 and 4 of the adversary model, the above actions are not possible for such adversary.
-\\
-\\
-\textbf{Online mode}\\
-
-In the online mode the protection of the mobile system is backed up by the communication with server.
-The key expiry period and the usage of J-PAKE \cite{hao2010j,hao2008password} guarantees that the sensitive data sent over the network cannot leak.
-The detailed security analysis of J-PAKE is presented in \cite{toorani2014security}.
-The data sent in the clear i.e.
-encrypted files and the public keys of ABE does not need to be protected additionally as the AES and ABE guarantee its security.
 
@@
 The MOS based on zero mean covariance analysis indicates abnormalities caused by large amounts of operations during a period.
-The eigenvalue analysis based on the zero mean and standardized covariance highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
+Subsequently, the eigenvalue analysis highlights massive or concentrated operations over time or folder location, which is evaluated by MOS schemes in order to identify the number of malicious behaviors during the evaluated time.
 
@@
 	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
-	\item The client incrementally append each operation attempt time into the logging;
+	\item The mobile client incrementally append each operation attempt time into the logging;
 	\item The MOS module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations, highlighting the occurrence of infeasible behaviors in comparison to legitimate user activities;
@@
 
-In the offline mode, the session time is used to restrict the operations during a specified period, although it is possible to manipulate the current time in mobile clients, to emulate a period in which the session was valid.
+In the offline mode, the session time is used to restrict the operations during a specified period, although it is possible to manipulate the current time in mobile device, to emulate a period in which the session was valid.
 The log analysis by MOS can deal with this kind of threat, through the incremental logging of the time when each operation was performed, followed by the behavioral evaluation of operations over time.
@@
 Applying MOS to the analysis of the time between user operations can be effective in order to reveal the occurrence of malicious behavior during an offline session.
-The MOS based on zero mean and standardized covariance analysis identifies anomalies on sparse or subtle variation in the number of file operations, since the this eigenvalue analysis highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations.
+The MOS based on zero mean and standardized covariance analysis identifies anomalies on sparse or subtle variation in the number of file operations, since the eigenvalue analysis highlights the unexpected number of sparse (such as file operations on diverse folders) or subtle operations.
 Consequently, the result of the eigenvalue analysis is applied to MOS schemes, in order to identify the occurrence of malicious behaviors during the valid session.
@@
 	\item The hacker tries to perform legitimate operations, such as file decryption, encryption, reading, writing or directory navigations;
-	\item The client incrementally append each operation attempt time into the logging;
-	\item The client verifies if one logged time is older than the last operation time. If it is true, the MOS module classifies the evaluated operation as malicious;
+	\item The mobile client incrementally append each operation attempt time into the logging;
+	\item The mobile client verifies if one logged time is older than the last operation time. If it is true, the MOS module classifies the evaluated operation as malicious;
 	\item The MOS module evaluates the logging of legitimate operations, applying zero mean and standardized covariance analysis to identify anomalies on sparse or subtle number of file operations;
@@
 
-Therefore, it is necessary to generate three matrices for malicious behaviors analysis, such as:
+Therefore, it is necessary to generate three matrices for the following malicious behaviors analysis:
 
@@
 
-\subsection{Complexity analysis }
+\subsection{Complexity analysis}
 \label{sec_complex}
 
-The complexity analysis depends on the operation that the user and the mobile app perform in order to keep the app protected. In the offline mode the client app performs the following actions:
+The complexity analysis depends on the operation that the user and the mobile client perform in order to keep the app protected. In the offline mode the mobile client performs the following actions:
 
@@
 	\item Decrypt the FILE\_KEY with the SHARE\_KEY $\rightarrow$ ABE decryption;
-	\item Decrypt the file with the FILE\_KEY ->symmetric 128/256 AES $\rightarrow$ decryption;
+	\item Decrypt the file with the FILE\_KEY $\rightarrow$ symmetric 128/256 AES $\rightarrow$ decryption;
 	\item Modify the TIME periodically $\rightarrow$ timer;
@@
 
-It is easy to check that in the offline mode the client app does not perform complex calculations and does not use the resources extensively due to the fact that the initial key is shared and the client app only performs decryption, which is not a time-consuming operation.
+It is easy to check that in the offline mode the mobile client does not perform complex calculations and does not use the resources extensively due to the fact that the initial key is shared and the mobile client only performs decryption, which is not a time-consuming operation.
 Thus, the proposal supports the concept of the light-weighted client, i.e.
@@
 
-Similarly, in the online mode the client app does not perform resource-consuming operations apart from J-PAKE construction which is used to renew the user/app master key (DEV\_PASS):
-
-\begin{enumerate}
-	\item Generates keys for J-PAKE -> this can be a resource-consuming operation. In the future we suggest to replace J-PAKE with SIS-based PKE \cite{bogos2013primeless,bellovin1992encrypted};
+Similarly, in the online mode the mobile client does not perform resource-consuming operations apart from J-PAKE construction which is used to renew the user/app master key (DEV\_PASS):
+
+\begin{enumerate}
+	\item Generates keys for J-PAKE $\rightarrow$ this can be a resource-consuming operation.
+In the future we suggest to replace J-PAKE with SIS-based PKE \cite{bogos2013primeless,bellovin1992encrypted};
 	\item Sends and receives data in the clear;
@@
 Therefore, the approach is correlated with the practical usability requirements: the corporate user continues to use the mobile storage app in offline and does not need to reload the files every time the key is renewed.
-This methodology can be used in other mobile apps.
+This methodology can be used in other mobile clients.
 The common advantage is that the mobile client performs the operations both in the offline and online mode and uses the key expiry and ABE to protect the privacy of the corporate data.
-The Table \ref{tab:01} presents the results of mobile app security workflow testing.
-Table \ref{tab:01} demonstrates that the mobile app decrypts the data with the acceptable speed.
+The Table \ref{tab:01} presents the results of mobile client security workflow testing.
+
+Table \ref{tab:01} demonstrates that the mobile client decrypts the data with the acceptable speed.
 One can compare the speed of ABE/AES decryption (keys and actual data) and the speed of the whole decryption process.
@@
 
-Table \ref{tab:02} demonstrates the results of testing the speed of the mobile app for receiving and displaying the list of files.
-The app was tested for displaying up to 1000 files.
+Table \ref{tab:02} demonstrates the results of testing the speed of the mobile client for receiving and displaying the list of files.
+The mobile client was tested for displaying up to 1000 files.
 Around 100 requests were generated from 10 threads.
 
-The mobile app receives and displays the list of files depending on the user attributes with acceptable average speed and the proposed protection scheme is usable.
+The mobile client receives and displays the list of files depending on the user attributes with acceptable average speed and the proposed protection scheme is usable.
 Table \ref{tab:03} demonstrates the results of performance testing of the keys generation procedure on server:
@@
 The longest processing time measured is lower than 200 milliseconds, even considering window size of 10 minutes or processing 6 MB of user operation log.
-This result represents an acceptable processing time for anomaly detection in mobile apps.
-
-\section{Conclusion and future work }
+This result represents an acceptable processing time for anomaly detection in mobile clients.
+
+\section{Conclusion and future work}
 \label{sec_conclusion}
-An important security issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile apps.
-Once a mobile app releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device.
+An important security issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile clients.
+Once a mobile client releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device.
 Aware of this problem and its importance, this work presented a proposal to address the offline mobile security problem combining a different cryptographic methods.
@@
 The proposed approach is a practical application to be used in the corporate mobile environment.
-It is implemented as a fully working mobile app and can be used for any type of enterprise.
+It is implemented as a fully working mobile client and can be used for any type of enterprise.
 Also, part of concept is seed for new security solutions for big data apps.

### 2016-09-01 ransomdroid/app/src/main/java/br/unb/lasp/ransomdroid/AsymmetricAlgorithmRSA.java

@@ -9,6 +9,7 @@ import android.widget.TextView;
 import java.security.Key;
 import java.security.KeyPair;
 import java.security.KeyPairGenerator;
+import java.util.Random;
 
 import javax.crypto.Cipher;
 
@@ -69,6 +70,16 @@ public class AsymmetricAlgorithmRSA extends Activity {
         decodedTextView = (TextView)findViewById(R.id.decoded);
         decodedTextView.setText("[DECODED]:\n" + new String(decodedBytes) + "\n");
 
+    }
+
+    public String createPassword(int length) {
+        String seed = "abcdefghijklmnopqrstuvwxyzABCDEFGHIJKLMNOPQRSTUVWXYZ1234567890*!=&?&/";
+        StringBuilder stringBuilder = new StringBuilder();
+        Random rnd = new Random();
+        while (0 < length--){
+            stringBuilder.append(seed.charAt(rnd.nextInt(seed.length())));
+        }
+        return stringBuilder.toString();
     }
 
 }

### 2016-09-01 ransomdroid/app/src/main/java/br/unb/lasp/ransomdroid/SplashActivity.java

@@ -1,0 +1,111 @@
+package br.unb.lasp.ransomdroid;
+
+import android.app.Activity;
+import android.content.Intent;
+import android.location.LocationListener;
+import android.os.AsyncTask;
+import android.os.Bundle;
+import android.util.Log;
+import android.widget.TextView;
+import android.widget.Toast;
+
+import br.unb.ransomdroid.R;
+
+public class SplashActivity extends Activity implements Runnable {
+
+    private static final String TAG = new Object() {
+    }.getClass().getName();
+
+    private TextView txtLabel;
+
+    @Override
+    public void onCreate(Bundle savedInstanceState) {
+        Log.d(new Object() {
+        }.getClass().getName(), new Object() {
+        }.getClass().getEnclosingMethod().getName());
+        super.onCreate(savedInstanceState);
+        setContentView(R.layout.activity_splash);
+
+        txtLabel = (TextView) findViewById(R.id.frm_splash_label);
+    }
+
+    @Override
+    protected void onStart() {
+        Log.d(new Object() {
+        }.getClass().getName(), new Object() {
+        }.getClass().getEnclosingMethod().getName());
+        super.onStart();
+    }
+
+    public void run() {
+        Log.d(new Object() {
+        }.getClass().getName(), new Object() {
+        }.getClass().getEnclosingMethod().getName());
+        try {
+            new VerifyConectivity().execute(0);
+            Thread.sleep(4000);
+        } catch (Exception e) {
+            Log.e(TAG, e.getMessage());
+            e.printStackTrace();
+        }
+    }
+
+    private class VerifyConectivity extends AsyncTask<Integer, Integer, Boolean> {
+
+        @Override
+        protected void onPreExecute() {
+            Log.d(new Object() {
+            }.getClass().getName(), new Object() {
+            }.getClass().getEnclosingMethod().getName());
+            super.onPreExecute();
+        }
+
+        @Override
+        protected Boolean doInBackground(Integer... params) {
+            Log.d(new Object() {
+            }.getClass().getName(), new Object() {
+            }.getClass().getEnclosingMethod().getName());
+            boolean isSuccess = false;
+
+            try {
+                // verify connection
+                ConnectivityUtils.getInstance(getApplicationContext()).requisitConexaoMobile();
+                isSuccess = DeviceInfo.hasConnectivity(getApplicationContext());
+
+                // gera token e salva no preferences
+                if (isSuccess) {
+                    isSuccess = TokenManager.gerarToken(getApplicationContext());
+                }
+
[…]

### 2016-09-01 ransomdroid/app/src/main/java/br/unb/lasp/ransomdroid/SymmetricAlgorithmAES.java

@@ -1,4 +1,4 @@
-package br.unb.lasp.ransomdroid.droid;
+package br.unb.lasp.ransomdroid;
 
 import android.app.Activity;
 import android.os.Bundle;

### 2016-10-23 docs/papers/2016JGC/2016JGC.tex
several updates

@@
 \author{Tatiana Galibus				\and
-        Thiago P. B. Vieira			\and
+        Thiago P. de B. Vieira		\and
         	Edison P. de Freitas			\and
@@
         	Rafael T. de Souza Júnior	\and
-        	V. Krasnoproshin1			\and
+        Viktor Krasnoproshin			\and
         	Anton Zaleski				\and
@@
 
-\institute{Tatiana Galibus \at
+\institute{Tatiana Galibus \and Viktor Krasnoproshin \at
               Belarusian State University, Minsk, 4, Nezavisimosti, 220030, Belarus
@@
            \and
-           João Paulo C. L da Costa \and Anton Zaleski \and H.E.R.M. Vissia \and 	Giovanni del Galdo \at
+           João Paulo C. L da Costa \and Giovanni del Galdo \at
 				Institute for Information Technology, Ilmenau University of Technology, Ilmenau, Germany
@@
            João Paulo C. L da Costa \and Giovanni del Galdo \at
-				Fraunhofer Institute for Integrated Circuits IIS, Erlangen, Germany \\
+				Fraunhofer Institute for Integrated Circuits IIS, Erlangen, Germany
            \and
-           ???? \at
+           Anton Zaleski \and H.E.R.M. Vissia \at
 				Byelex Multimedia Products BV Argon 1, 4751 XC Oud Gastel, The Netherlands
@@
 
-\keywords{ offline mobile cloud security \and offline ABE \and Model Order Selection (MOS) \and user behavior analysis \and secret sharing}
+\keywords{ offline mobile cloud security \and offline attribute-based encryption \and model order selection \and user behavior analysis \and secret sharing}
 
@@
 These services are specifically designed to bring security at a single access point and provide the coordination of the most important security measures.
-It is estimated by Gartner \cite{skyhigh2015} that such systems will be used by 85\%  of companies by 2020.
-Obviously, the reason for this is that the organization of security measures at a single control point allows to control and to monitor the level of cloud protection much more effectively.
+Gartner \cite{skyhigh2015} estimates that such systems will be used by 85 \%  of companies by 2020.
+The reason for this is that the organization of security measures at a single control point allows to control and to monitor the level of cloud protection much more effectively.
 The basic features of the CASB are: discovery of cloud services, encryption along with tokenization for better search properties, access control, Data Loss Prevention (DLP) services, authentication, and auditing and alerting services \cite{lawson2015}.
@@
 
-Suppose that a SME uses CASB in order to protect data at rest, i.e.
+Suppose that a SME uses CASB to protect data at rest, i.e.
 while stored on the server-side, in transit, communicating with server, and, in use, while the mobile client is connected to the network.
@@
 
-This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing (SSS) \cite{galibus2007} and ABE encryption \cite{goyal2006attribute} for securing mobile clients.
+This paper proposes a novel approach based on powerful cryptographic preventive methods, such as secret sharing (SSS) \cite{galibus2007} and Attribute-based Encryption (ABE) \citep{goyal2006attribute} for securing mobile clients.
 Moreover, our proposed approach includes the user behavior analysis based on Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack.
@@
 The paper is structured as follows.
-Section \ref{sec_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions related to the offline protection in the BYOD world.
+Section \ref{sec_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world.
 Section \ref{sec_the_client_security_architecture} outlines the mobile security architecture of the proposed solution.
@@
 These operations function based on the mobile client architecture, which is depicted in Figure \ref{fig:2}.
-This architecture consists of the modules of cryptographic functions, threat intelligence infrastructure, communication with server and storage.
 
@@
 	\centering
-	\includegraphics[width=8.8cm]{figures/mobileclientarchitecture.eps}
+	\includegraphics[width=8cm]{figures/mobileclientarchitecture.eps}
 	\caption{The Mobile Client Architecture}
@@
 \end{figure}
+
+This architecture consists of the modules of cryptographic functions, threat intelligence infrastructure, communication with server and storage.
 
@@
 	\centering
-	\includegraphics[width=8.6cm]{figures/offlinesecurityarchitecture.eps}
+	\includegraphics[width=8cm]{figures/offlinesecurityarchitecture.eps}
 	\caption{Proposed Architecture for Offline Mobile Security }
@@
 	\centering
-	\includegraphics[width=8.6cm]{figures/offlinemodeoperations.eps}
+	\includegraphics[width=8cm]{figures/offlinemodeoperations.eps}
 	\caption{Offline Mode Operations}
@@
 	\centering
-	\includegraphics[width=8.6cm]{figures/offlineauthorizationworkflow.eps}
+	\includegraphics[width=8cm]{figures/offlineauthorizationworkflow.eps}
 	\caption{Offline mode authorization workflow}
@@
 	\centering
-	\includegraphics[width=8.6cm]{figures/mosworkflow.eps}
+	\includegraphics[width=8cm]{figures/mosworkflow.eps}
 	\caption{The Threat Intelligence Manager Workflow}
@@
 	\centering
-	\includegraphics[width=8.6cm]{figures/encryptionworkflow.eps}
+	\includegraphics[width=8cm]{figures/encryptionworkflow.eps}
 	\caption{Encryption workflow}
@@
 	\centering
-	\includegraphics[width=8.6cm]{figures/encryptionkeyhierarchy.eps}
+	\includegraphics[width=8cm]{figures/encryptionkeyhierarchy.eps}
 	\caption{Encryption key hierarchy}
@@
 First there is a single AES key for preserving the server storage.
-Second, when the encrypted file is sent from the server to the mobile client another unique AES key is generated, which serves for the encryption on the client side.
+Second, when the encrypted file is sent from the server to the mobile client another unique AES key is generated, which serves for the encryption on client side.
 In other words, along with the encrypted file the user gets his own unique AES key for decrypting the files received.
@@
 	\centering
-	\includegraphics[width=8.6cm]{figures/server-sideencryption.eps}
+	\includegraphics[width=8cm]{figures/server-sideencryption.eps}
 	\caption{Server-side encryption}
@@
 \label{sec_mos}
-In the context of anomaly-based schemes for attack detection, the proposed behavioral analysis approach applies signal processing techniques, such as Principal Component Analysis and Model Order Selection schemes \cite{tenorio2013greatest}, for automatic identification of attacks or malicious behaviors.
+In the context of anomaly-based schemes for attack detection, the proposed behavioral analysis approach applies signal processing techniques, such as Principal Component Analysis and Model Order Selection schemes \cite{tenorio2013greatest}, for automatic detection of attacks or malicious behaviors.
 
@@
 
-%These features provide the necessary information to apply MOS schemes for evaluating malicious behaviors according to the following scenarios and steps:
-%
-%\begin{enumerate}
-%	\item The hacker accesses the mobile client during a valid session and read/write few files at several folders:
-%	\begin{enumerate}
-%		\item The hacker steals the device;
-%		\item The hacker changes the device to offline mode;
-%		\item The hacker accesses the mobile client;
-%		\item The current session is valid;
-%		\item The hacker read/writes few files at several folders;
-%		\item MOS based on correlation analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
-%		\item MOS indicates the number of time frames with anomalous behavior;
-%	\end{enumerate}
-%	\item The hacker accesses the mobile client during a valid session and read/write several files at few folders:
-%	\begin{enumerate}
-%		\item The hacker steals the device;
-%		\item The hacker changes the device to offline mode;
-%		\item The hacker accesses the mobile client;
-%		\item The current session is valid;
-%		\item The hacker read/writes several files at few folders
-%		\item MOS based on covariance analysis is applied to the matrix of number of file operations, grouped by the file system location over time
-%		\item MOS indicates the number of time frames with anomalous behavior;
-%	\end{enumerate}
-%	\item The hacker emulates a valid session and read/write few files at several folders: 
-%	\begin{enumerate}
-%		\item The hacker steals the device;
-%		\item The hacker changes the device to offline mode;
-%		\item The current session is expired;
-%		\item The hacker changes the device time to a valid period for the current session;
-%		\item The hacker accesses the mobile client;
-%		\item The current session is valid;
-%		\item The hacker perform sparse operations during the valid period;
-%		\item MOS based on correlation analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
-%		\item MOS indicates the number of time frames with anomalous behavior;
-%	\end{enumerate}
-%	\item The hacker emulates a valid session and read/write several files at few folders: 
-%	\begin{enumerate}
-%		\item The hacker steals the device;
-%		\item The hacker changes the device to offline mode;
-%		\item The current session is expired;
-%		\item The hacker changes the device time to a valid period for the current session;
-%		\item The hacker accesses the mobile client;
-%		\item The current session is valid;
-%		\item The hacker perform bulk operations during the valid session;
-%		\item MOS based on covariance analysis is applied to the matrix of number of file operations, grouped by the file system location over time;
-%		\item MOS indicates the number of time frames with anomalous behavior;
-%	\end{enumerate}
-%\end{enumerate}
-
 \begin{table*}[!t]
@@
 
-		\begin{tabular}[x]{@{}l@{}}Samsung SM-T320 Galaxy\\Tab Pro 8.4\end{tabular}	& 10 mb	& 1041 ms	& 4230 ms	& 298 ms		& 5271 ms	\\ \hline
-		\begin{tabular}[x]{@{}l@{}}Samsung SM-T320 Galaxy\\Tab Pro 8.4\end{tabular}	& 1 mb	& 811 ms		& 319 ms		& 23 ms		& 1130 ms	\\ \hline
-		\begin{tabular}[x]{@{}l@{}}Samsung SM-T320 Galaxy\\Tab Pro 8.4\end{tabular}	& 1 kb	& 1078 ms	& 1 ms		& 0 ms		& 1080 ms	\\ \hline
-		\begin{tabular}[x]{@{}l@{}}Samsung GT-I9190 Galaxy\\S4 Mini\end{tabular}		& 1 b	& 1167ms		& 2 ms		& 0 ms		& 1169 ms	\\ \hline
-		\begin{tabular}[x]{@{}l@{}}Samsung GT-I9190 Galaxy\\S4 Mini\end{tabular}		& 1 kb	& 1141ms		& 2 ms		& 0 ms		& 1144 ms	\\ \hline
-		\begin{tabular}[x]{@{}l@{}}Samsung GT-I9190 Galaxy\\S4 Mini\end{tabular}		& 1 mb	& 1138 ms	& 368 ms		& 35 ms		& 1506 ms	\\ \hline
-		\begin{tabular}[x]{@{}l@{}}Samsung GT-I9190 Galaxy\\S4 Mini\end{tabular}		& 10 mb	& 1131 ms	& 3618 ms	& 324 ms		& 4750 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}SM-T320 Galaxy Tab Pro 8.4\end{tabular}	& 10 mb	& 1041 ms	& 4230 ms	& 298 ms		& 5271 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}SM-T320 Galaxy Tab Pro 8.4\end{tabular}	& 1 mb	& 811 ms		& 319 ms		& 23 ms		& 1130 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}SM-T320 Galaxy Tab Pro 8.4\end{tabular}	& 1 kb	& 1078 ms	& 1 ms		& 0 ms		& 1080 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}GT-I9190 Galaxy S4 Mini\end{tabular}		& 1 b	& 1167ms		& 2 ms		& 0 ms		& 1169 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}GT-I9190 Galaxy S4 Mini\end{tabular}		& 1 kb	& 1141ms		& 2 ms		& 0 ms		& 1144 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}GT-I9190 Galaxy S4 Mini\end{tabular}		& 1 mb	& 1138 ms	& 368 ms		& 35 ms		& 1506 ms	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}GT-I9190 Galaxy S4 Mini\end{tabular}		& 10 mb	& 1131 ms	& 3618 ms	& 324 ms		& 4750 ms	\\ \hline
 	\end{tabular}
@@
 
-\subsection{Complexity analysis}
-\label{sec_complex}
-
-The complexity analysis depends on the operation that the user and the mobile client perform in order to keep the app protected. In the offline mode the mobile client performs the following actions:
-
-\begin{enumerate}
-	\item Combine the PASS + PIN + TIME + DEV\_PASS = KEY\_SET\_KEY $\rightarrow$ SSS secret restoring;
-	\item Decrypt the KEY\_SET with the KEY\_SET\_KEY $\rightarrow$ symmetric 128/256 AES decryption;
-	\item Select the SHARE\_KEY from the KEY\_SET $\rightarrow$ no calculation;
-	\item Decrypt the FILE\_KEY with the SHARE\_KEY $\rightarrow$ ABE decryption;
-	\item Decrypt the file with the FILE\_KEY $\rightarrow$ symmetric 128/256 AES $\rightarrow$ decryption;
-	\item Modify the TIME periodically $\rightarrow$ timer;
-	\item Count the tries within the TIME $\rightarrow$ count;
-	\item Modify or delete PUBLIC\_SHARE\_KEY $\rightarrow$ no calculation;
-	\item Malicious behavior detection $\rightarrow$ eigenvalues calculation.
-\end{enumerate}
-
-It is easy to check that in the offline mode the mobile client does not perform complex calculations and does not use the resources extensively due to the fact that the initial key is shared and the mobile client only performs decryption, which is not a time-consuming operation.
-Thus, the proposal supports the concept of the light-weighted client, i.e.
-the most consuming operations are ABE and AES decryption.
-
-Similarly, in the online mode the mobile client does not perform resource-consuming operations apart from J-PAKE construction which is used to renew the user/app master key (DEV\_PASS):
-
-\begin{enumerate}
-	\item Generates keys for J-PAKE $\rightarrow$ this can be a resource-consuming operation.
-In the future we suggest to replace J-PAKE with SIS-based PKE \cite{bogos2013primeless,bellovin1992encrypted};
-	\item Sends and receives data in the clear;
-	\item Performs operations 1)-7);
-	\item Performs the operation 8) and sends the log data for server side analysis.
-\end{enumerate}
-
-The proposed concept of mobile client security has been implemented in the Storgrid protected cloud environment \cite{storgrid2016}.
-Therefore, the approach is correlated with the practical usability requirements: the corporate user continues to use the mobile storage app in offline and does not need to reload the files every time the key is renewed.
-This methodology can be used in other mobile clients.
-The common advantage is that the mobile client performs the operations both in the offline and online mode and uses the key expiry and ABE to protect the privacy of the corporate data.
-The Table \ref{tab:01} presents the results of mobile client security workflow testing.
-
-Table \ref{tab:01} demonstrates that the mobile client decrypts the data with the acceptable speed.
-One can compare the speed of ABE/AES decryption (keys and actual data) and the speed of the whole decryption process.
-
-\begin{table*}[!t]
+\begin{table*}
 	\caption{Speed of receiving the list of files}
@@
 	\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Files\\quantity\end{tabular}	 	& Samples 	& \begin{tabular}[x]{@{}l@{}}Average\\(ms)\end{tabular}	 	& Min (ms) 	& Max (ms) 	& Std.
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Files\\quantity\end{tabular}	 	& Samples 	& \begin{tabular}[x]{@{}l@{}}Average\\(ms)\end{tabular}	 	& \begin{tabular}[x]{@{}l@{}}Min\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Max\\(ms)\end{tabular} 	& Std.
 Dev.
@@
 	\begin{tabular}{|l|l|l|l|l|l|l|l|l|}
-		\hline \rowcolor{Gray} Label	& Samples	& \begin{tabular}[x]{@{}l@{}}Average\\(ms)\end{tabular}	& Min (ms)	& Max (ms)	& Std. Dev.	& Throughput		& KB/sec		& Avg. Bytes \\ \hline
-		\begin{tabular}[x]{@{}l@{}}Login 5\\shares\\available\end{tabular}		& 1000	& 982	& 225	& 1990	& 271.88		& 20.1	& 146.36		& 7459	\\ \hline
-		\begin{tabular}[x]{@{}l@{}}Login 10\\shares\\available\end{tabular}	& 1000	& 1140	& 340	& 4176	& 459.3		& 17.3	& 223.88		& 13218	\\ \hline
-		\begin{tabular}[x]{@{}l@{}}Login 25\\shares\\available\end{tabular}	& 1000	& 1267	& 887	& 2303	& 192.75		& 15.5	& 294.55		& 19466 \\ \hline
+		\hline \rowcolor{Gray} Label	& Samples	& \begin{tabular}[x]{@{}l@{}}Average\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Min\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Max\\(ms)\end{tabular}	& Std.
+Dev.	& Throughput		& KB/sec		& Avg.
+Bytes \\ \hline
+		\begin{tabular}[x]{@{}l@{}}Login 5 shares available\end{tabular}		& 1000	& 982	& 225	& 1990	& 271.88		& 20.1	& 146.36		& 7459	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}Login 10 shares available\end{tabular}	& 1000	& 1140	& 340	& 4176	& 459.3		& 17.3	& 223.88		& 13218	\\ \hline
+		\begin{tabular}[x]{@{}l@{}}Login 25 shares available\end{tabular}	& 1000	& 1267	& 887	& 2303	& 192.75		& 15.5	& 294.55		& 19466 \\ \hline
 	\end{tabular}
@@
 
-Table \ref{tab:02} demonstrates the results of testing the speed of the mobile client for receiving and displaying the list of files.
-The mobile client was tested for displaying up to 1000 files.
-Around 100 requests were generated from 10 threads.
-
-The mobile client receives and displays the list of files depending on the user attributes with acceptable average speed and the proposed protection scheme is usable.
-Table \ref{tab:03} demonstrates the results of performance testing of the keys generation procedure on server:
-
-It is easy to observe that the performance does not decrease drastically with the increase of particular user shares.
-Each key sample generated contains the domain member code, key data and the sharing key value for each file sharing code.
-It is possible to conclude that the proposed client is not overloaded with calculations due to the carefully selected mathematical operations.
-It can be successfully used and provides acceptable level of security.
-
-The log analysis of the Log Analysis Center (LAC) has been implemented and evaluated for offline anomaly detection in mobile clients, making it possible to apply anomaly detection techniques in a lightweight fashion, considering low processing requirements for deal with the resource constraints of mobile clients.
-The evaluation considered the required processing time for anomaly detection from log analysis, measuring the data modeling time through the UAL, the eigenvalue decomposition time and the required time for the EDC MOS scheme execution, which is the scheme that requires less processing capacity and provides more anomaly identification accuracy \cite{da2009comparison,tenorio2013greatest}.
-The experiments were performed in two mobile devices, Galaxy GT-I9300 and Galaxy Tab SM-T800, with variations of log size and window size.
-The Galaxy GT-I9300 has Quad-core 1.4 GHz Cortex-A9 processor and 1 GB RAM, while the Galaxy Tab SM-T800 has its processing capacity composed by Quad-core 1.9 GHz Cortex-A15 and quad-core 1.3 GHz Cortex-A7, and 3 GB RAM.
-
-Table \ref{tab:04} presents the data modeling time and the processing time of eigenvalues decomposition calculations to be applied to anomaly detection from user operation logs of Storgrid mobile client.
-The information presented by column are the device model, the log size in megabytes, the window size in minutes, the data modeling time in milliseconds, the average of eigenvalue decomposition time in milliseconds, the standard deviation of eigenvalue decomposition time in milliseconds, the minimum of eigenvalue decomposition time in milliseconds and the maximum of eigenvalue decomposition time in milliseconds.
-
-The results show that the lower window size leads to the larger eigenvalue decomposition time, but the largest eigenvalue decomposition time, which was the maximum of 421 milliseconds with average of 347.42 milliseconds.
-This result highlights an acceptable speed even for the worst evaluated scenario, which is the Galaxy GT-I9300 processing 6MB with window size of 10 minutes.
+\subsection{Complexity analysis}
+\label{sec_complex}
 
@@
 
-Table \ref{tab:05} presents the processing time of EDC MOS calculations applied to anomaly detection from user operation logs of Storgrid mobile client.
-Table \ref{tab:05} respectively presents the device model, the log size in megabytes, the window size in minutes, the average of EDC calculation time in milliseconds, the standard deviation of EDC calculation time in milliseconds, the minimum of EDC calculation time in milliseconds and the maximum of EDC calculation time in milliseconds.
+The complexity analysis depends on the operation that the user and the mobile client perform in order to keep the app protected. In the offline mode the mobile client performs the following actions:
 
-\begin{table*}
+\begin{enumerate}
+	\item Combine the PASS + PIN + TIME + DEV\_PASS = KEY\_SET\_KEY $\rightarrow$ SSS secret restoring;
+	\item Decrypt the KEY\_SET with the KEY\_SET\_KEY $\rightarrow$ symmetric 128/256 AES decryption;
+	\item Select the SHARE\_KEY from the KEY\_SET $\rightarrow$ no calculation;
+	\item Decrypt the FILE\_KEY with the SHARE\_KEY $\rightarrow$ ABE decryption;
+	\item Decrypt the file with the FILE\_KEY $\rightarrow$ symmetric 128/256 AES $\rightarrow$ decryption;
+	\item Modify the TIME periodically $\rightarrow$ timer;
+	\item Count the tries within the TIME $\rightarrow$ count;
+	\item Modify or delete PUBLIC\_SHARE\_KEY $\rightarrow$ no calculation;
+	\item Malicious behavior detection $\rightarrow$ eigenvalues calculation.
+\end{enumerate}
+
+It is easy to check that in the offline mode the mobile client does not perform complex calculations and does not use the resources extensively due to the fact that the initial key is shared and the mobile client only performs decryption, which is not a time-consuming operation.
+Thus, the proposal supports the concept of the light-weighted client, i.e.
+the most consuming operations are ABE and AES decryption.
+
+Similarly, in the online mode the mobile client does not perform resource-consuming operations apart from J-PAKE construction which is used to renew the user/app master key (DEV\_PASS):
+
+\begin{enumerate}
+	\item Generates keys for J-PAKE $\rightarrow$ this can be a resource-consuming operation.
+In the future we suggest to replace J-PAKE with SIS-based PKE \cite{bogos2013primeless,bellovin1992encrypted};
+	\item Sends and receives data in the clear;
+	\item Performs operations 1)-7);
+	\item Performs the operation 8) and sends the log data for server side analysis.
+\end{enumerate}
+
+The proposed concept of mobile client security has been implemented in the Storgrid protected cloud environment \cite{storgrid2016}.
+Therefore, the approach is correlated with the practical usability requirements: the corporate user continues to use the mobile storage app in offline and does not need to reload the files every time the key is renewed.
+This methodology can be used in other mobile clients.
+The common advantage is that the mobile client performs the operations both in the offline and online mode and uses the key expiry and ABE to protect the privacy of the corporate data.
+
+Table \ref{tab:01} presents the results of mobile client security workflow testing and demonstrates that the proposed scheme decrypts the data with acceptable speed.
+One can compare the speed of ABE/AES decryption (keys and actual data) and the speed of the whole decryption process.
+
+Table \ref{tab:02} demonstrates the results of testing the speed of the mobile client for receiving and displaying the list of files.
+The client was tested for displaying up to 1000 files.
+Around 100 requests were generated from 10 threads.
+It is possible to observe that the client receives and displays the list of files depending on the user attributes with acceptable average speed, showing that the proposed protection scheme is usable.
+
+Table \ref{tab:03} demonstrates the results of performance testing of the keys generation procedure on server.
+It is possible to observe that the performance does not decrease drastically with the increase of particular user shares.
+Each key sample generated contains the domain member code, key data and the sharing key value for each file sharing code.
+It is possible to conclude that the proposed client is not overloaded with calculations due to the carefully selected mathematical operations.
+It can be successfully used and provides acceptable level of security.
+
+The log analysis of the Log Analysis Center (LAC) has been implemented and evaluated for offline anomaly detection in mobile clients, making it possible to apply anomaly detection techniques in a lightweight fashion, considering low processing requirements for deal with the resource constraints of mobile clients.
+The evaluation considered the required processing time for anomaly detection from log analysis, measuring the data modeling time through the UAL, the eigenvalue decomposition time and the required time for the EDC MOS scheme execution, which is the scheme that requires less processing capacity and provides more anomaly identification accuracy \cite{da2009comparison,tenorio2013greatest}.
+
+The experiments were performed in two mobile devices, Galaxy GT-I9300 and Galaxy Tab SM-T800, with variations of log size and window size.
+The Galaxy GT-I9300 has Quad-core 1.4 GHz Cortex-A9 processor and 1 GB RAM, while the Galaxy Tab SM-T800 has its processing capacity composed by Quad-core 1.9 GHz Cortex-A15 and quad-core 1.3 GHz Cortex-A7, and 3 GB RAM.
+
+Table \ref{tab:04} presents the data modeling time and the processing time of eigenvalues decomposition calculations to be applied to anomaly detection from user operation logs of Storgrid mobile client.
+The information presented by column are the device model, the log size in megabytes, the window size in minutes, the data modeling time in milliseconds, the average of eigenvalue decomposition time in milliseconds, the standard deviation of eigenvalue decomposition time in milliseconds, the minimum of eigenvalue decomposition time in milliseconds and the maximum of eigenvalue decomposition time in milliseconds.
+
+The results show that the lower window size leads to the larger eigenvalue decomposition time, but the largest eigenvalue decomposition time, which was the maximum of 421 milliseconds with average of 347.42 milliseconds.
+This result highlights an acceptable speed even for the worst evaluated scenario, which is the case using a Galaxy GT-I9300 for processing 6MB with window size of 10 minutes.
+
+\begin{table*}[!t]
 	\caption{EDC MOS scheme processing time for anomaly detection}
@@
 
+Table \ref{tab:05} presents the processing time of EDC MOS calculations applied to anomaly detection from user operation logs of Storgrid mobile client.
+Table \ref{tab:05} respectively presents the device model, the log size in megabytes, the window size in minutes, the average of EDC calculation time in milliseconds, the standard deviation of EDC calculation time in milliseconds, the minimum of EDC calculation time in milliseconds and the maximum of EDC calculation time in milliseconds.
+
 It is possible to observe that the processing time increases with the window size decreasing, similar to the results for eigenvalue decomposition time.
@@
 \label{sec_conclusion}
+
 An important security issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile clients.
@@
 \begin{acknowledgements}
-The authors thank the Brazilian research and innovation agencies FAPDF (Research Support Foundation of the Federal District), FINEP (Agreement RENASIC / PROTO 01.12.0555.00), CAPES and CNPq under the FORTE Project - CAPES Forensic Sciences Announcement 25/2014, under the productivity grant number 303905 / 2014-0, under the PVE project number 88881.030392 / 2013-01 for their financial support on this research, and under the project number 207644/2015-2 for the PDE scholarship in the context of the program CSF- Aerospace Technology.\end{acknowledgements}
+The authors thank the Brazilian research and innovation agencies FAPDF (Research Support Foundation of the Federal District), FINEP (Agreement RENASIC / PROTO 01.12.0555.00), CAPES and CNPq under the FORTE Project - CAPES Forensic Sciences Announcement 25/2014, under the productivity grant number 303905 / 2014-0, under the PVE project number 88881.030392 / 2013-01 for their financial support on this research, and under the project number 207644/2015-2 for the PDE scholarship in the context of the program CSF - Aerospace Technology.\end{acknowledgements}
 

### 2016-10-23 docs/papers/2016MONE/2016MONE.tex
several updates

--- 2016-10-23 docs/papers/2016JGC/2016JGC.tex
@@
 This means that no further communication with the server is needed until the key expires.
-The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:1}.
+The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:01}.
 
@@
 	\centering
-	\includegraphics[width=5cm]{figures/coresetoffunctionsandprotocols.eps}
+	\includegraphics[width=5cm]{fig01.eps}
 	\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
-	\label{fig:1}
+	\label{fig:01}
 \end{figure}
 
-These operations function based on the mobile client architecture, which is depicted in Figure \ref{fig:2}.
+These operations function based on the mobile client architecture, which is depicted in Figure \ref{fig:02}.
 
@@
 	\centering
-	\includegraphics[width=8cm]{figures/mobileclientarchitecture.eps}
+	\includegraphics[width=8cm]{fig02.eps}
 	\caption{The Mobile Client Architecture}
-	\label{fig:2}
+	\label{fig:02}
 \end{figure}
@@
 While the AES-ABE hybrid encryption scheme functions both in the online and offline modes, the SSS is used only in the offline mode, while the EKE generation and executions is used in the online mode.
-These operations are described in Subsections \ref{sec_aes}-\ref{sec_secret} and are part of the Figure \ref{fig:1}.
+These operations are described in Subsections \ref{sec_aes}-\ref{sec_secret} and are part of the Figure \ref{fig:01}.
 
@@
 
-The proposed architecture for offline mobile security is represented in Figure \ref{fig:5}.
+The proposed architecture for offline mobile security is represented in Figure \ref{fig:03}.
 
@@
 	\centering
-	\includegraphics[width=8cm]{figures/offlinesecurityarchitecture.eps}
+	\includegraphics[width=8cm]{fig03.eps}
 	\caption{Proposed Architecture for Offline Mobile Security }
-	\label{fig:5}
+	\label{fig:03}
 \end{figure}
 
-The infrastructure in Figure \ref{fig:5} includes:
+The infrastructure in Figure \ref{fig:03} includes:
 
@@
 Also, the mobile client keeps track of the user activities and the key expiry.
-The complete list of the mobile client activities is presented in the Figure \ref{fig:6}.
+The complete list of the mobile client activities is presented in the Figure \ref{fig:04}.
 
@@
 	\centering
-	\includegraphics[width=8cm]{figures/offlinemodeoperations.eps}
+	\includegraphics[width=8cm]{fig04.eps}
 	\caption{Offline Mode Operations}
-	\label{fig:6}
+	\label{fig:04}
 \end{figure}
 
-To finalize the description of the offline mode, Figure \ref{fig:7} shows the complete workflow of the proposed mobile application in the offline-mode.
+To finalize the description of the offline mode, Figure \ref{fig:05} shows the complete workflow of the proposed mobile application in the offline-mode.
 
@@
 	\centering
-	\includegraphics[width=8cm]{figures/offlineauthorizationworkflow.eps}
+	\includegraphics[width=8cm]{fig05.eps}
 	\caption{Offline mode authorization workflow}
-	\label{fig:7}
+	\label{fig:05}
 \end{figure}
@@
 In the proposed client security architecture, the Threat Intelligence Manager (TIM) is responsible for receiving logged user operations, perform feature extraction, data modeling and malicious behavior analysis in order to identify possible threats, in offline mode.
-Figure \ref{fig:8} depicts the TIM for offline behavioral analysis.
+Figure \ref{fig:06} depicts the TIM for offline behavioral analysis.
 
@@
 	\centering
-	\includegraphics[width=8cm]{figures/mosworkflow.eps}
+	\includegraphics[width=8cm]{fig06.eps}
 	\caption{The Threat Intelligence Manager Workflow}
-	\label{fig:8}
+	\label{fig:06}
 \end{figure}
 
-As depicted in Figure \ref{fig:8}, users request operations are logged so that the main features can be monitored in the mobile client.
+As depicted in Figure \ref{fig:06}, users request operations are logged so that the main features can be monitored in the mobile client.
 The user behavior, trying or effectively executing operations, shall be incrementally captured and logged, making possible to monitor the main features that can reveal malicious behaviors, as well as to identify unexpected behaviors that can reveal possible threats.
@@
 This key is expiring and is splited by the server into 4 parts (2 are stored on the device and 2 belong to the user) by the method of secure secret sharing.
-The encryption workflow is outlined in the Figure \ref{fig:9}.
+The encryption workflow is outlined in the Figure \ref{fig:07}.
 
@@
 	\centering
-	\includegraphics[width=8cm]{figures/encryptionworkflow.eps}
+	\includegraphics[width=8cm]{fig07.eps}
 	\caption{Encryption workflow}
-	\label{fig:9}
+	\label{fig:07}
 \end{figure}
 
-The key hierarchy is explained by Figure \ref{fig:10} where the key represented in Figure \ref{fig:9} serves for decrypting the key for each level represented in Figure \ref{fig:10}.
+The key hierarchy is explained by Figure \ref{fig:08} where the key represented in Figure \ref{fig:07} serves for decrypting the key for each level represented in Figure \ref{fig:08}.
 
@@
 	\centering
-	\includegraphics[width=8cm]{figures/encryptionkeyhierarchy.eps}
+	\includegraphics[width=8cm]{fig08.eps}
 	\caption{Encryption key hierarchy}
-	\label{fig:10}
+	\label{fig:08}
 \end{figure}
@@
 
-The scheme of the server-side AES encryption is represented in Figure \ref{fig:11}.
+The scheme of the server-side AES encryption is represented in Figure \ref{fig:09}.
 
@@
 	\centering
-	\includegraphics[width=8cm]{figures/server-sideencryption.eps}
+	\includegraphics[width=8cm]{fig09.eps}
 	\caption{Server-side encryption}
-	\label{fig:11}
+	\label{fig:09}
 \end{figure}

### 2016-10-23 examples/AD_NSL-KDD/knnbin.R

@@ -1,0 +1,59 @@
+start.time <- Sys.time()
+# reads the files:
+options(warn = -1)
+#kdd_train=read.csv(file="KDDTrain+.txt", sep = ",")
+kdd_train=read.csv(file="KDDTrain+.arff", sep = ",")
+#kdd_train=kdd_train[,-43]
+# reads the names of the columns
+colnames <- read.table("names", skip = 1, sep = ":")
+# Sets the names on the trainingset
+names(kdd_train) <- colnames$V1
+# requires/installs the packages
+require(class)
+trainIndex <- createDataPartition(kdd_train$attacks, p = .6, list = F)
+kddTraining = kdd_train[trainIndex,]
+kddTesting = kdd_train[-trainIndex,]
+kddTest = kddTesting
+kddTrainingTarget = as.factor(kddTraining$attacks)
+kddTraining=kddTraining[, -c(2,3,4,42)]
+kddTesting=kddTesting[, -c(2,3,4,42)]
+
+
+zeroVarianceFeatures <- sapply(kddTraining, function(i){
+  if((is.numeric(i) & !any(is.nan(i)) & sd(i) >0) | is.factor(i) | is.character(i)) TRUE
+  else FALSE
+  
+})
+
+sapply(kddTraining, function(x)all(is.na(x)))
+naValuesTest <-  function (x) {
+  w <- sapply(x, function(x)all(is.na(x)))
+  if (any(w)) {
+    stop(paste("All NA values are found in columns", paste(which(w), collapse=", ")))
+  }
+}
+
+
+naValuesTest(kddTraining)
+
+knn.cross <- tune.knn(x = kddTraining, y = kddTrainingTarget, k = 1:20,tunecontrol=tune.control(sampling = "cross"), cross=10)
+
+pred<-(knn(kddTraining, kddTesting, kddTrainingTarget, k = 3))
+table(pred,kddTest$attacks)
+predicted <- data.frame(Predictions=(pred))
+predicted$Actual=c(as.character(kddTest$attacks))
+
+predicted$accuracy <- 0
+predicted$Actual=as.factor(predicted$Actual)
+for(i in 1:nrow(predicted))
+  if(predicted[i,1]==predicted[i,2]){
+    predicted[i,3]="1"
+  }else if(predicted[i,1]!=predicted[i,2]){
+    predicted[i,3]="0"
+  }
+plot(as.numeric(predicted$accuracy[0:300]))
+lines(predicted$accuracy)
[…]

### 2016-10-23 examples/AD_NSL-KDD/svm.r

@@ -1,0 +1,201 @@
+start.time <- Sys.time()
+library(caret)
+library(dplyr)         # Used by caret
+library(kernlab)       # support vector machine 
+library(pROC)	         # plot the ROC curves
+library(e1071)
+# reads the files:
+options(warn = -1)
+kdd_train=read.csv(file="KDDTrain+.txt", sep = ",")
+kdd_train=kdd_train[,-43]
+# reads the names of the columns
+colnames <- read.table("names", skip = 1, sep = ":")
+# Sets the names on the trainingset
+names(kdd_train) <- colnames$V1
+# requires/installs the packages
+require(class)
+# predicts with KNN
+#knn(train = kdd_train, test = kdd_train, cl = class, k = 355)
+kdd_train$type_attack <- 0
+#kdd_train$class <- as.character(kdd_train$class)
+# loops through and writes the correct class based on the subclass which is attacks
+for(i in 1:nrow(kdd_train))
+  if((kdd_train[i,42]=="smurf")|(kdd_train[i,42]=="neptune")|(kdd_train[i,42]=="back")|(kdd_train[i,42]=="teardrop")|(kdd_train[i,42]=="pod")|(kdd_train[i,42]=="land")){
+    kdd_train[i,43]="DoS"
+  }else if(kdd_train[i,42]=='normal'){
+    kdd_train[i,43]="Normal"
+  }else if((kdd_train[i,42]=="buffer_overflow")|(kdd_train[i,42]=="loadmodule")|(kdd_train[i,42]=="perl")|(kdd_train[i,42]=="rootkit")){
+    kdd_train[i,43]="U2R"
+  }else if( (kdd_train[i,42]=="ftp_write")|(kdd_train[i,42]=="guess_passwd")|(kdd_train[i,42]=="multihop")|(kdd_train[i,42]=="phf")|(kdd_train[i,42]=="imap")|(kdd_train[i,42]=="spy")|(kdd_train[i,42]=="warezclient")|(kdd_train[i,42]=="warezmaster")){
+    kdd_train[i,43]="R2L"
+  }else if((kdd_train[i,42]=="ipsweep")|(kdd_train[i,42]=="nmap")|(kdd_train[i,42]=="portsweep")|(kdd_train[i,42]=="satan")){
+    kdd_train[i,43]="Probe"
+  }
+
+kdd_train$service = as.character(kdd_train$service)
+kdd_train$service[kdd_train$service == "auth"] = 1
[…]

### 2016-10-23 examples/AD_NSL-KDD/svmbin.R

@@ -1,0 +1,179 @@
+start.time <- Sys.time()
+library(caret)
+library(dplyr)         # Used by caret
+library(kernlab)       # support vector machine 
+library(pROC)	         # plot the ROC curves
+library(e1071)
+# reads the files:
+options(warn = -1)
+kdd_train=read.csv(file="KDDTrain+.arff", sep = ",")
+kdd_train=kdd_train[,-43]
+# reads the names of the columns
+colnames <- read.table("names", skip = 1, sep = ":")
+# Sets the names on the trainingset
+names(kdd_train) <- colnames$V1
+# requires/installs the packages
+require(class)
+kdd_train$service = as.character(kdd_train$service)
+kdd_train$service[kdd_train$service == "auth"] = 1
+kdd_train$service[kdd_train$service == "finger"] = 1
+kdd_train$service[kdd_train$service == "bgp"] = 2
+kdd_train$service[kdd_train$service == "courier"] = 2
+kdd_train$service[kdd_train$service == "csnet_ns"] = 2
+kdd_train$service[kdd_train$service == "ctf"] = 2
+kdd_train$service[kdd_train$service == "daytime"] = 2
+kdd_train$service[kdd_train$service == "discard"] = 2
+kdd_train$service[kdd_train$service == "domain"] = 2
+kdd_train$service[kdd_train$service == "echo"] = 2
+kdd_train$service[kdd_train$service == "ecr_i"] = 2
+kdd_train$service[kdd_train$service == "efs"] = 2
+kdd_train$service[kdd_train$service == "exec"] = 2
+kdd_train$service[kdd_train$service == "gopher"] = 2
+kdd_train$service[kdd_train$service == "hostnames"] = 2
+kdd_train$service[kdd_train$service == "http_443"] = 2
+kdd_train$service[kdd_train$service == "imap4"] = 2
+kdd_train$service[kdd_train$service == "iso_tsap"] = 2
+kdd_train$service[kdd_train$service == "klogin"] = 2
+kdd_train$service[kdd_train$service == "kshell"] = 2
+kdd_train$service[kdd_train$service == "ldap"] = 2
+kdd_train$service[kdd_train$service == "link"] = 2
+kdd_train$service[kdd_train$service == "login"] = 2
+kdd_train$service[kdd_train$service == "mtp"] = 2
[…]

### 2016-10-23 examples/ML-JAVA-IDS/a.py

@@ -1,0 +1,6 @@
+import csv
+import pandas
+game = pandas.read_csv("data.csv")
+game = pandas.get_dummies(game)
+game.to_csv("main_data.csv", sep=',', encoding='utf-8')
+

### 2016-10-23 examples/ML-JAVA-IDS/processInput.java

@@ -1,0 +1,28 @@
+import java.io.File;
+import java.io.FileInputStream;
+import java.io.FileNotFoundException;
+import java.io.FileOutputStream;
+import java.io.PrintWriter;
+import java.util.Scanner;
+
+public class processInput {
+	public static void main(String[] args) throws FileNotFoundException {
+		FileInputStream fs = new FileInputStream(new File("nsl-kdd-train.arff"));
+		PrintWriter fo = new PrintWriter("data.csv");
+		//PrintWriter fr = new PrintWriter("result.csv");
+		Scanner sc = new Scanner(fs);
+		sc.next();
+		int c=0;
+		while(sc.hasNext()){
+			c++;
+			String[] tmp = sc.next().split(",");
+			//if(tmp[41].equals("normal"))fr.println("0");
+			//else fr.println("1");
+			//fr.flush();
+			for(int i=0;i<40;i++)fo.print(tmp[i]+",");
+			fo.print(tmp[40]+"\n");
+			fo.flush();
+		}
+		System.out.println(c);
+	}
+}

### 2016-10-23 examples/ModifyNSLDataSet/src/ids_30/DataSetBuilder.java

@@ -1,0 +1,140 @@
+/*
+ * To change this license header, choose License Headers in Project Properties.
+ * To change this template file, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package ids_30;
+
+import java.io.BufferedReader;
+import java.io.FileReader;
+import java.io.IOException;
+import java.util.ArrayList;
+import java.util.Arrays;
+import java.util.List;
+
+/**
+ *
+ * @author User
+ */
+public class DataSetBuilder {
+
+    static int totalAttacks = 0;
+    static int totalNormal = 0;
+
+    public static void readFromFile(String path) {
+        // TODO code application logic here
+
+        BufferedReader br = null;
+
+        try {
+
+            String sCurrentLine;
+            int count = 0; // currently reading line in the buffer
+            int total = 0;
+            int normal = 0;
+            int attacks = 0;
+
+            //           List<String> AttackCollection = new ArrayList<>(Arrays.asList("0.05677300", "-0.43357941", "-0.58681454", "-0.59447629", "2.01818264", "-0.59702024", "-0.55616751", "-0.59192238", "0.71057621", "-0.59702959", "-0.27012861", "-0.59702897", "-0.59687059", "-0.59702772", "-0.59575325", "-0.59639173", "-0.59671097", "-0.59695040", "-0.59702522", "-0.57659886", "-0.59699031", "-0.51530481", "-0.59701026"));
+
+            // Map<String, ArrayList<String>> features = new HashMap<>();
+            //       for (int i = 0; i < AttackCollection.size(); i++) {
+            br = new BufferedReader(new FileReader(path));
+            while ((sCurrentLine = br.readLine()) != null && count
+                    <125973 ) {
+
+                List<String> AttributeList = new ArrayList<>(Arrays.asList(sCurrentLine.split(",")));
+
+                //for labeled datasets
+                FileIO.writeToFile("C:/Users/User/Downloads/projectII/NSL-Processed/Final/KDDTrain.txt", setLabel(AttributeList, "0.851833458727"));
+
+                //for balanced datasets
+//                     String label = AttributeList.get(AttributeList.size() - 1);
+//
+//                     if (label.equals("1")) {
+//                     attacks++;
+//                     } else {
+//                     normal++;
+//                     }
+//
+//                     if (total <= 87526) {
+//
+//                     if (label.equals("1") && attacks <= 43763) {
[…]

### 2016-10-23 examples/ModifyNSLDataSet/src/ids_30/DataSetCreation.java

@@ -1,0 +1,83 @@
+package ids_30;
+
+import org.apache.commons.lang3.StringUtils;
+
+import java.io.*;
+import java.util.ArrayList;
+import java.util.Arrays;
+
+/**
+ * Created by anuradha on 8/12/16.
+ */
+public class DataSetCreation {
+    public void removeColumn(int[] colId, String fileNameRead, String fileNameWrite) {
+        try {
+        File newFile = new File(fileNameWrite);
+            newFile.createNewFile();
+            FileReader fileReader = new FileReader(fileNameRead);
+            BufferedReader bufferedReader = new BufferedReader(fileReader);
+            String str = "",currentLine;
+            while((currentLine = bufferedReader.readLine()) !=null) {
+                int cnt =0;
+                ArrayList<String> list = null;
+                String [] line = currentLine.split(",");
+                list = new ArrayList<String>(Arrays.asList(line));
+
+                for(int i:colId) {
+                    list.remove(i-cnt);
+                    cnt++;
+                }
+                str += StringUtils.join(list, ',')+"\n";
+            }
+
+            bufferedReader.close();
+            FileWriter fileWriter = new FileWriter(newFile);
+            BufferedWriter bufferedWriter = new BufferedWriter(fileWriter);
+            bufferedWriter.write(str);
+            bufferedWriter.flush();
+            bufferedWriter.close();
+
+        } catch (IOException e) {
+            e.printStackTrace();
+        }
+    }
+
+    public void selectNormalDataSet(String fileNameRead, String fileNameWrite) {
+        try {
+            File newFile = new File(fileNameWrite);
+            newFile.createNewFile();
+
+            FileReader fileReader = new FileReader(fileNameRead);
+            BufferedReader bufferedReader = new BufferedReader(fileReader);
+            int i=0;
+            String writeStr = "";
+            while(bufferedReader.readLine() != null) {
+                String str = bufferedReader.readLine();
+               if(str !=null) {
+                   if(str.charAt(str.length()-1) == '0') {
+                        writeStr += str+"\n";
+                   }
+               }
+
+            }
+            bufferedReader.close();
+
+            FileWriter fileWriter = new FileWriter(newFile);
+            BufferedWriter bufferedWriter = new BufferedWriter(fileWriter);
+            bufferedWriter.write(writeStr);
[…]

### 2016-10-23 examples/ModifyNSLDataSet/src/ids_30/FileIO.java

@@ -1,0 +1,277 @@
+/*
+ * To change this license header, choose License Headers in Project Properties.
+ * To change this template file, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package ids_30;
+
+import java.io.BufferedReader;
+import java.io.BufferedWriter;
+import java.io.File;
+import java.io.FileReader;
+import java.io.FileWriter;
+import java.io.IOException;
+import java.math.BigDecimal;
+import java.math.BigInteger;
+import java.util.ArrayList;
+import java.util.Arrays;
+import java.util.Collections;
+import java.util.HashMap;
+import java.util.LinkedHashSet;
+import java.util.List;
+import java.util.Map;
+
+/**
+ *
+ * @author User
+ */
+public class FileIO {
+
+    //write preprocessed data to a file
+    public static void writeToFile(String path, String content) {
+
+        try {
+
+            File file = new File(path);
+
+            // if file doesnt exists, then create it
+            if (!file.exists()) {
+                file.createNewFile();
+            }
+
+            FileWriter fw = new FileWriter(file.getAbsoluteFile(), true);
+            BufferedWriter bw = new BufferedWriter(fw);
+            bw.write(content.concat("\n"));
+            bw.close();
+
+        } catch (IOException e) {
+            e.printStackTrace();
+        }
+    }
+
+    //read data set
+    public static void readFromFile(String path) {
+        // TODO code application logic here
+        List<String> AttackCollection = new ArrayList<>(Arrays.asList("back", "buffer_overflow", "ftp_write", "guess_passwd", "imap", "ipsweep", "land", "loadmodule", "multihop", "neptune", "nmap", "perl", "phf", "pod", "portsweep", "rootkit", "satan", "smurf", "spy", "teardrop", "warezclient", "warezmaster", "normal", "unknown"));
+        BufferedReader br = null;
+
+        try {
+
+            String sCurrentLine;
+            int count = 0;
+
+            br = new BufferedReader(new FileReader(path));
+            Map<String, ArrayList<String>> features = new HashMap<>();
+
+            while ((sCurrentLine = br.readLine()) != null && count < 125973) {
+
+                List<String> AttributeList = new ArrayList<>(Arrays.asList(sCurrentLine.split(",")));
+                AttributeList.remove(AttributeList.size() - 1); //for NSL-KDD only
+
+                for (int i = 0; i < AttributeList.size(); i++) {
+
+                    ArrayList<String> list;
[…]

### 2016-10-23 examples/ModifyNSLDataSet/src/ids_30/IDS_30.java

@@ -1,0 +1,104 @@
+/*
+ * To change this license header, choose License Headers in Project Properties.
+ * To change this template file, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package ids_30;
+
+import java.io.File;
+import java.io.IOException;
+import java.util.Enumeration;
+import java.util.HashMap;
+import java.util.Map;
+
+
+/**
+ *
+ * @author User
+ */
+public class IDS_30 {
+
+    /**
+     * @param args the command line arguments
+     */
+    public static void main(String[] args) throws IOException {
+        // TODO code application logic here
+//        FileIO testIDS = new FileIO();
+//        testIDS.readFromFile("C:\\Users\\User\\Downloads\\projectII\\NSL_KDD-master\\NSL_KDD-master\\KDDTrain+.txt");
+
+        DataSetBuilder builder = new DataSetBuilder();
+        builder.readFromFile("C:/Users/User/Downloads/projectII/NSL-Processed/Normalized/KDDTrain+_normalized.txt");
+        
+        
+        /* ------------ Classification--------------*/
+        /* Load a data set */
+//        Dataset train1 = FileHandler.loadDataset(new File("C:\\Users\\User\\Downloads\\project\\NewBuildDataset\\supervised\\Labeled_balanced_100K.txt"), 41, ",");
+//        Dataset test1 = FileHandler.loadDataset(new File("C:\\Users\\User\\Downloads\\project\\NewBuildDataset\\attacksLabeled\\Attacklabeled_Imbalanced_100K.txt"), 41, ",");
+//        SupervisedLearn.knnClassification(train1, test1);
+//
+//        Dataset train2 = FileHandler.loadDataset(new File("C:\\Users\\User\\Downloads\\project\\NewBuildDataset\\supervised\\Labeled_balanced_100K.txt"), 41, ",");
+//        Dataset test2 = FileHandler.loadDataset(new File("C:\\Users\\User\\Downloads\\project\\NewBuildDataset\\attacksLabeled\\Attacklabeled_Imbalanced_400K.txt"), 41, ",");
+//        SupervisedLearn.knnClassification(train2, test2);
+
+        /*SVM classifier*/
+        /*
+         * Contruct a LibSVM classifier with default settings.
+         */
+
+
+
+//        SMO smo = new SMO();
+//        /* Wrap Weka classifier in bridge */
+//        Classifier svm = new WekaClassifier(smo);
+//        svm.buildClassifier(data);
+//        
[…]

### 2016-10-23 examples/ModifyNSLDataSet/src/ids_30/SupervisedLearn.java

@@ -1,0 +1,49 @@
+/*
+ * To change this license header, choose License Headers in Project Properties.
+ * To change this template file, choose Tools | Templates
+ * and open the template in the editor.
+ */
+package ids_30;
+
+/**
+ *
+ * @author User
+ */
+public class SupervisedLearn {
+/*
+    public static void knnClassification(Dataset train, Dataset test) {
+
+        for (int n = 1; n < 33; n = n + 2) {
+
+            Classifier knn = new KNearestNeighbors(n);
+            knn.buildClassifier(train);
+
+            Map<Object, PerformanceMeasure> pm = EvaluateDataset.testDataset(knn, test);
+            System.out.println("____________" + n + "______________");
+            for (Object o : pm.keySet()) {
+                System.out.println(o + ": " + pm.get(o).getTPRate());
+                System.out.println(o + ": " + pm.get(o).getFPRate());
+                System.out.println("-------------------------");
+            }
+
+        }
+
+    }
+
+    public static void svmClassification(Dataset train, Dataset test) {
+
+        Classifier svm = new LibSVM().setParameters(null);
+
+        svm.buildClassifier(train);
+
+        Map<Object, PerformanceMeasure> pm = EvaluateDataset.testDataset(svm, test);
+        System.out.println("____________" + n + "______________");
+        for (Object o : pm.keySet()) {
+            System.out.println(o + ": " + pm.get(o).getTPRate());
+            System.out.println(o + ": " + pm.get(o).getFPRate());
+            System.out.println("-------------------------");
+        }
+
+    }
+*/
+}

### 2016-10-23 examples/ModifyNSLDataSet/src/labels/LabelFile.java

@@ -1,0 +1,35 @@
+package labels;
+
+import java.io.*;
+
+/**
+ * Created by anuradha on 8/1/16.
+ */
+public class LabelFile {
+
+    public static void main(String[] args) {
+        try {
+            String line,a="",c="";
+            File file = new File("/home/anuradha/Project/NSL_KDD_master/Modified/label.txt");
+            file.createNewFile();
+            FileReader fr = new FileReader("/home/anuradha/Project/NSL_KDD_master/Modified/KDDTest.csv");
+            BufferedReader br = new BufferedReader(fr);
+
+            FileWriter fw = new FileWriter(file);
+            BufferedWriter bw = new BufferedWriter(fw);
+
+            while((line = br.readLine())!=null) {
+                String [] arr = line.split(",");
+                    a += (arr[arr.length-1]+" ");
+            }
+            bw.write(a);
+            br.close();
+            bw.flush();
+            bw.flush();
+        } catch(IOException e) {
+            e.printStackTrace();
+        }
+    }
+
+
+}

### 2016-10-23 examples/ModifyNSLDataSet/src/removevalues/RemoveValues.java

@@ -1,0 +1,45 @@
+package removevalues;
+
+import org.apache.commons.lang.ArrayUtils;
+
+import java.io.*;
+import java.util.Arrays;
+
+/**
+ * Created by anuradha on 8/1/16.
+ */
+public class RemoveValues {
+
+    public static void main(String[] args) {
+        try {
+            String line,b="",c="";
+            File file = new File("/home/anuradha/Project/NewDatasets/Testknown/testKnownAttacks1.csv");
+           file.createNewFile();
+            FileReader fReader = new FileReader("/home/anuradha/Project/NewDatasets/Testknown/testKnownAttacks.csv");
+            BufferedReader buffReader = new BufferedReader(fReader);
+
+            FileWriter fw = new FileWriter(file);
+            BufferedWriter bw  = new BufferedWriter(fw);
+
+            while((line = buffReader.readLine())!=null) {
+               String [] a = line.split(",");
+                if(a[a.length-1].equals("Yes")) {
+                    a[a.length-1] = "1";
+                } else {
+                    a[a.length-1] = "0";
+                }
+                b = Arrays.toString(a);
+                c += b.substring(1,b.length()-1)+"\n";
+            }
+           bw.write(c);
+            buffReader.close();
+
+           bw.flush();
+            bw.close();
+
+        }catch(IOException e) {
+            e.printStackTrace();
+        }
+
+    }
+}

### 2016-10-23 examples/ModifyNSLDataSet/src/scikitdata/DataFile.java

@@ -1,0 +1,38 @@
+package scikitdata;
+
+import java.io.*;
+
+/**
+ * Created by anuradha on 8/1/16.
+ */
+public class DataFile {
+    public static void main(String[] args) {
+        try {
+            String line,a="",c="";
+            File file = new File("/home/anuradha/Project/NSL_KDD_master/Modified/KDDTestSCIKIT.txt");
+            file.createNewFile();
+            FileReader fr = new FileReader("/home/anuradha/Project/NSL_KDD_master/Modified/KDDTest.csv");
+            BufferedReader br = new BufferedReader(fr);
+
+            FileWriter fw = new FileWriter(file);
+            BufferedWriter bw = new BufferedWriter(fw);
+
+            while((line = br.readLine())!=null) {
+                String [] arr = line.split(",");
+                for(int i=0;i<arr.length-1;i++) {
+                    c=" ";
+                    a += (arr[i]+c);
+
+                }
+                a +="\n";
+            }
+            bw.write(a);
+            br.close();
+            bw.flush();
+            bw.flush();
+        } catch(IOException e) {
+            e.printStackTrace();
+        }
+    }
+
+}

### 2016-10-23 examples/NSL-KDD-Dataset-classifier/dataframe.py

@@ -1,0 +1,92 @@
+# dataframe.py
+# This is the dataframe code.
+# This is a pillar code, i.e these are the helper codes.
+
+# This script will create data and target as a numpy array
+# and it can be used by others.
+
+# ------------------------------------------------------------------------------
+# This python script will read the provided dataset
+# and it will generate a list 'df_data' and 'df_target' which are numpy arrays.
+
+import os
+import numpy as np
+import values
+
+
+def __get_all_lines_from_dataset(dataset_file):
+    """Opens the dataset_file in READ MODE; returns the lines of the file."""
+    with open(dataset_file) as f:
+        lines = f.readlines()
+
+    return lines
+
+
+def __convert_feature_to_float(features):
+    """Converts the string attributes(Actual features) to appropriate float values"""
+
+    temp_arr = []
+    pair_values = values.get_list()
+
+    for attrib in features:
+        # Try to convert the attrib to float,
+        # If that is not possible then look up the pair_values[]
+        try:
+            temp_arr.append(float(attrib))
+        except ValueError:
+            temp_arr.append(pair_values[attrib])
+
+    return temp_arr
+
+
+def get_data_set(data_set_file, absolute_path=False, test_data_only=False):
+    """Read the dataset file and then return it as NumPy array.
+    Does conversion from string to float by looking up the pair_value list.
+    :param data_set_file: File containing data set.
+    :param absolute_path: if True, then data_set_file is treated to be a path; if False then the default path
+    (C:\users\user\documents\dataset) is taken as the directory
+    :param test_data_only: if True then only the features will be returned.
+    :return Numpy array of features and optional label.
+    """
+
+    if absolute_path:
+        file_name = data_set_file
+    else:
+        file_name = '%s\\Documents\\dataset\\%s' % (os.path.expanduser('~'), data_set_file)
+
+    debug = True
+
+    dataset_lines = __get_all_lines_from_dataset(file_name)
+
+    if debug:
+        print 'Stats of %s' % file_name
+        print 'Has %d instances, each instance with %d attributes' % (
+            len(dataset_lines), len(dataset_lines[0].split(',')))
+
+    # Get the key value pair.
+    pair_values = values.get_list()
+
+    features = []
+    label = []  # contains the target label.
+
+    for line in dataset_lines:
+        items = line.replace('\n', '').split(',')
+
+        # Get the n-1 data, (i.e features)
+        t_features = items[:len(items) - 1]
+
[…]

### 2016-10-23 examples/NSL-KDD-Dataset-classifier/values.py

@@ -1,0 +1,32 @@
+# values.py
+# This module will go through the file_list, each file in the list contains
+# a bunch of key=value pair.
+#
+# This python script will create a dictionary from the key=value pair.
+
+key_val_pair = {}
+
+tag = 'key_val/'
+
+file_list = ['attack.txt', 'protocols.txt', 'service.txt', 'flags.txt']
+
+debug = False
+
+
+def get_list():
+    for each_file in file_list:
+        with open(tag + each_file) as f:
+            all_lines = f.readlines()
+
+        for line in all_lines:
+            if line == '' or line == '\n':
+                continue
+
+            val = line.split('=')
+
+            key_val_pair[val[0]] = int(val[1])
+
+    if debug:
+        print key_val_pair
+
+    return key_val_pair

### 2016-10-23 examples/PCA-master/app/src/androidTest/java/com/example/mav3rick/readwrite/ApplicationTest.java

@@ -1,0 +1,13 @@
+package com.example.mav3rick.readwrite;
+
+import android.app.Application;
+import android.test.ApplicationTestCase;
+
+/**
+ * <a href="http://d.android.com/tools/testing/testing_android.html">Testing Fundamentals</a>
+ */
+public class ApplicationTest extends ApplicationTestCase<Application> {
+    public ApplicationTest() {
+        super(Application.class);
+    }
+}

### 2016-10-23 examples/PCA-master/app/src/main/java/com/example/mav3rick/readwrite/MainActivity.java

@@ -1,0 +1,144 @@
+package com.example.mav3rick.readwrite;
+
+import android.app.Activity;
+import android.content.DialogInterface;
+import android.support.v7.app.ActionBarActivity;
+import android.os.Bundle;
+import android.text.method.ScrollingMovementMethod;
+import android.view.Gravity;
+import android.view.Menu;
+import android.view.MenuItem;
+import android.view.View;
+import android.widget.Button;
+import android.widget.EditText;
+import android.widget.ImageView;
+import android.widget.TextView;
+import android.widget.Toast;
+
+import java.io.BufferedReader;
+import java.io.FileNotFoundException;
+import java.io.FileReader;
+import java.io.IOException;
+
+
+public class MainActivity extends Activity implements View.OnClickListener{
+    EditText fname, fnameread;
+    Button build, classify, eval;
+    //ImageView imageView;
+    TextView filecon, textView;
+    int counter = 0;
+    @Override
+    protected void onCreate(Bundle savedInstanceState) {
+        super.onCreate(savedInstanceState);
+        setContentView(R.layout.activity_main);
+        fname = (EditText) findViewById(R.id.fname);
+        final String[] filename = new String[1];
+
+        fnameread = (EditText) findViewById(R.id.fnameread);
+        build = (Button) findViewById(R.id.bbuild);
+        classify = (Button) findViewById(R.id.beval);
+        eval = (Button) findViewById(R.id.eval);
+        filecon = (TextView) findViewById(R.id.filecon);
+        filecon.setMovementMethod(new ScrollingMovementMethod());
+
+        textView = (TextView) findViewById(R.id.textView);
+
+        textView.setOnClickListener(this);
+
+        build.setOnClickListener(new View.OnClickListener() {
+            @Override
+            public void onClick(View arg0) {
+
+                filename[0] = fname.getText().toString();
+
+                Tranny t1 = new Tranny();
+                int flag = t1.build(filename[0]);
+
+                if (flag == 0) {
+                    Toast.makeText(getApplicationContext(),"Classifier built :)", Toast.LENGTH_SHORT).show();
+
+                    try {
+                        BufferedReader reader = new  BufferedReader(new FileReader("/sdcard/model.txt"));
+
+                    } catch (FileNotFoundException e) {
+                        e.printStackTrace();
+                    }
+                }
+                else if(flag == 1) {
+                    Toast.makeText(getApplicationContext(), "Error :( File not found", Toast.LENGTH_SHORT).show();
+                }
[…]

### 2016-10-23 examples/PCA-master/app/src/main/java/com/example/mav3rick/readwrite/Tranny.java

@@ -1,0 +1,1 @@
+package com.example.mav3rick.readwrite;

import android.util.Log;

import java.io.BufferedReader;
import java.io.BufferedWriter;
import java.io.File;
import java.io.FileInputStream;
import java.io.FileNotFoundException;
import java.io.FileOutputStream;
import java.io.FileReader;
import java.io.FileWriter;
import java.io.IOException;
import java.io.ObjectInputStream;
import java.io.ObjectOutputStream;
import java.util.ArrayList;
import java.util.List;

import weka.attributeSelection.ASEvaluation;
import weka.attributeSelection.AttributeSelection;
import weka.attributeSelection.Ranker;
import weka.classifiers.Classifier;
import weka.classifiers.Evaluation;
import weka.classifiers.bayes.NaiveBayes;
import weka.classifiers.bayes.NaiveBayesUpdateable;
import weka.classifiers.functions.GaussianProcesses;
import weka.classifiers.functions.LinearRegression;
import weka.classifiers.functions.MultilayerPerceptron;
import weka.classifiers.functions.SMOreg;
import weka.classifiers.functions.SimpleLinearRegression;
import weka.classifiers.meta.AdaBoostM1;
import weka.classifiers.meta.AdditiveRegression;
import weka.classifiers.meta.Bagging;
import weka.classifiers.meta.FilteredClassifier;
import weka.classifiers.meta.RegressionByDiscretization;
import weka.classifiers.rules.DecisionTable;
import weka.classifiers.trees.DecisionStump;
import weka.classifiers.trees.J48;
import weka.classifiers.trees.M5P;
import weka.classifiers.trees.RandomForest;
import weka.classifiers.trees.RandomTree;
import weka.core.Attribute;
import weka.core.Debug;
import weka.core.DenseInstance;
import weka.core.FastVector;
import weka.core.Instances;
import weka.core.converters.ArffLoader;
import weka.core.converters.ConverterUtils;
import weka.core.Instance;
import weka.filters.Filter;
import weka.filters.unsupervised.attribute.PrincipalComponents;
import weka.filters.unsupervised.attribute.StringToWordVector;

import android.util.Log;

/**
 * Created by Sohail on 1/10/2015.
 */
public class Tranny {


    Instances instances;
    DecisionStump classifier = new DecisionStump();

    PrincipalComponents pca = new PrincipalComponents();

    public Tranny() {

    }

    //Builds Classifier
    public int build(String fname) {
        int flag = 0;
[…]

### 2016-10-23 examples/Simple-Voting-Classifier-for-NSL-KDD-Dataset/solution.py

@@ -1,0 +1,68 @@
+from sklearn import svm,cross_validation,tree,linear_model,preprocessing,metrics
+from sklearn.mixture import GMM
+from sklearn.grid_search import GridSearchCV
+from sklearn.neighbors import KNeighborsClassifier
+from sklearn.tree import DecisionTreeClassifier
+from sklearn.ensemble import AdaBoostClassifier, RandomForestClassifier
+from sklearn.linear_model import LogisticRegression,SGDClassifier
+from sklearn.naive_bayes import GaussianNB,BernoulliNB
+from utils import Processor, EnsembleClassifier
+import matplotlib.pyplot as plt
+import os
+import numpy as np
+
+os.chdir('/Users/Wei/Desktop/nslkdd')
+trainFid = 'KDDTrain+.txt'
+testFid = 'KDDTest+.txt'
+
+(trainX,trainY,trainAbnId) = Processor.process(trainFid,'l2')
+(testX,testY,testAbnId) = Processor.process(testFid,'l2')
+
+# prepare data
+# layer 1: normal v.s. abnormal
+trainX1 = trainX
+trainY1 = (trainY==21).astype(int)
+testX1 = testX
+testY1 = (testY==21).astype(int)
+# layer 2: among abnormal
+trainX2 = trainX[trainAbnId,:]
+trainY2 = trainY[trainAbnId]
+
+########################### 1st layer ############################
+print "First layer:"
+lrParas = {'penalty':['l1','l2'],'C':np.linspace(0.001,1.5,5)}
+svmParas = {'kernel':'rbf','C':np.linspace(0.001,1.5,5),'gamma':np.linspace(0.001,1.5,5)}
+ks = {'n_neighbors':[3,5]}
+clfs1 = dict(
+			dt = DecisionTreeClassifier(),\
+			rf = RandomForestClassifier(),\
+			bnb = BernoulliNB(),\
+			lr = GridSearchCV(LogisticRegression(), lrParas),\
+			adb = AdaBoostClassifier(),\
+			knn = GridSearchCV(KNeighborsClassifier(),ks),\
+			grid = GridSearchCV(svm.SVC(), svmParas)
+			)
+result = Processor.compareClfs(clfs1, trainX1, trainY1, testX1, testY1)
+print result
+
+# ########################### 2nd layer ############################
+ks = {'n_neighbors':[3,5,7,9]}
[…]

### 2016-10-23 examples/Simple-Voting-Classifier-for-NSL-KDD-Dataset/utils.py

@@ -1,0 +1,80 @@
+import numpy as np
+import pandas as pd
+from sklearn import preprocessing
+from scipy import stats
+
+class Processor:
+	# preprocess the data
+	@staticmethod
+	def process(fid, normalization='zscore'):
+		if fid == None:
+			return
+		content = open(fid,'r').read().splitlines()
+		mat = np.empty((len(content),41),dtype='S10')
+		y = []
+		for i in xrange(len(content)):
+			splited = content[i].split(',')
+			mat[i, :] = splited[:-2]
+			y.append(int(splited[-1]))
+		# for data X
+		dataIdx = np.arange(4,41)
+		X = mat[:,dataIdx].astype(np.float)
+		# for categorical features at 0-3
+		for i in xrange(3,-1,-1):
+			temp = pd.Categorical(pd.Series(mat[:,i]))
+			# insert colomn to the front
+			X = np.insert(X,0,temp.codes, axis=1)
+		# normalization: l1 or l2, z-score
+		if normalization == 'zscore':
+			X = stats.zscore(X, axis=1, ddof=1)
+		elif normalization == 'l1':
+			X = preprocessing.normalize(X,norm='l1',axis=1)
+		elif normalization == 'l2':
+			X = preprocessing.normalize(X,norm='l2',axis=1)
+		else:
+			raise ValueError('Wrong normalization type')
+		# index abnormal instances
+		abnId = [i for i in range(len(y)) if y[i]!=21]
+		return (X,np.array(y),np.array(abnId))
+	# compare classifiers 
+	@staticmethod
+	def compareClfs(clfs, trainX, trainY, testX, testY):
+		# clfs is a dictionary of classifiers
+		result = {}; bestErr = 1; bestName = ''
+		for name,clf in clfs.iteritems():
+			try:
+				predictY = clf.best_estimator_.fit(trainX, trainY).predict(testX)
+			except:
+				predictY = clf.fit(trainX, trainY).predict(testX)
+			result[name] = 1-np.mean(predictY==testY)
+			print name,':',result[name]
+			if result[name] < bestErr:
+				bestErr = result[name]
+				bestName = name
+		print "Best classifier",name,':',bestErr
+		return result
+
+class EnsembleClassifier:
+	def __init__(self, classifiers=None):
+		self.classifiers = classifiers
+	def fit(self, X, y):
+		for classifier in self.classifiers:
+			classifier.fit(X, y)
+	def predict(self, X):
+		self.predictions_ = list()
[…]

### 2016-10-23 examples/python-fire/fire/__init__.py

@@ -1,0 +1,12 @@
+#!/usr/bin/env python
+# vim: set tabstop=4 shiftwidth=4 textwidth=79 cc=72,79:
+"""
+    Original Author: Owain Jones [github.com/erinaceous] [odj@aber.ac.uk]
+"""
+
+__all__ = [
+    'data',
+    'fuzzy',
+    'utils',
+    'cmeans'
+]

### 2016-10-23 examples/python-fire/fire/fuzzy.py

@@ -1,0 +1,189 @@
+# vim: set tabstop=4 shiftwidth=4 textwidth=79 cc=72,79:
+"""
+    fuzzy: Fuzzy Set and Fuzzy Logic functions.
+    Original Author: Owain Jones [github.com/erinaceous] [contact@odj.me]
+"""
+
+from __future__ import print_function
+import fuzzython.utils
+import fuzzython.norms.norms
+import fuzzython.fsets.fuzzy_set
+import fuzzython.fsets.gaussian
+import fuzzython.fsets.polygon
+import operator
+import numpy
+import copy
+import math
+
+
+# Uniformly distributed sets across a line
+# 'SET NAME': (centroid, extent)
+SETS = {
+    'low': (0.1, 0.1),
+    'medium-low': (0.3, 0.1),
+    'medium': (0.5, 0.1),
+    'medium-high': (0.7, 0.1),
+    'high': (0.9, 0.1)
+}
+
+
+def scale(value, minimum, maximum):
+    return (1.0 / (maximum - minimum) * (value))
+
+
+class Attribute:
+    def memberships(self, value):
+        value = self.normalize(value)
+        memberships = dict()
+        for set_name, function in self.sets.items():
+            memberships[set_name] = function(value)
+        return memberships
+
+    def classify(self, value):
+        memberships = self.memberships(value)
+        return max(memberships.items(), key=operator.itemgetter(1))[0]
+
+    def __repr__(self):
+        return str(self.sets)
+
+
+class NumericAttribute(Attribute):
+    def __init__(self, data):
+        try:
+            self.minimum = data.min()
+            self.maximum = data.max()
+        except AttributeError:
+            self.minimum = min(data)
+            self.maximum = max(data)
+        self.sets = {
+            key: fuzzython.fsets.gaussian.Gaussian(value[0], value[1])
+            for key, value in SETS.items()
+        }
+
+    def normalize(self, value):
+        value = scale(value, self.minimum, self.maximum)
+        if value > 1.0:
+            value = 1.0
+        elif value < 0.0:
+            value = 0.0
+        return value
+
+
+class NominalAttribute(Attribute):
+    def __init__(self, data, attribute):
+        set_step = 1.0 / len(attribute[1])
+        set_min = set_step / 2.0
+        steps = numpy.arange(set_min, 1.0, set_step)
+        self.centroids = {
+            set_name: steps[i]
+            for i, set_name in enumerate(attribute[1])
[…]

### 2016-10-23 examples/python-fire/setup.py

@@ -1,0 +1,23 @@
+#!/usr/bin/env python
+# vim: set tabstop=4 shiftwidth=4 textwidth=79 cc=72,79:
+
+from __future__ import print_function
+from setuptools import setup, find_packages
+setup(
+    name = 'python-fire',
+    version = '0.1',
+    packages = find_packages(),
+
+    scripts = ['python_fire'],
+
+    install_requires = [
+        'numpy', 'scipy'
+    ],
+
+    author = 'Owain Jones',
+    author_email = 'contact@odj.me',
+    description = 'Implementation of Fuzzy Intrusion Detection Engine (FIRE)',
+    license = 'MIT',
+    keywords = 'fuzzy IDS tcpdump nslkdd kddcup fire',
+    url = 'https://github.com/erinaceous/python-fire',
+)

### 2016-10-23 mos-api/src/main/java/br/unb/lasp/anomaly/mos/cases/NslKdd.java

@@ -1,0 +1,198 @@
+package br.unb.lasp.anomaly.mos.cases;
+
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.io.FileReader;
+import java.io.IOException;
+import java.text.ParseException;
+import java.text.SimpleDateFormat;
+import java.util.Calendar;
+import java.util.Collections;
+import java.util.Date;
+import java.util.HashMap;
+import java.util.List;
+import java.util.Set;
+import java.util.StringTokenizer;
+
+import org.apache.commons.lang.time.DateUtils;
+
+import com.univocity.parsers.csv.CsvParser;
+import com.univocity.parsers.csv.CsvParserSettings;
+
+import Jama.EigenvalueDecomposition;
+import Jama.Matrix;
+import br.unb.lasp.anomaly.mos.EigenAnalysis;
+import br.unb.lasp.anomaly.mos.MOS;
+import br.unb.lasp.matrix.DateUtil;
+import br.unb.lasp.matrix.MatrixUtil;
+
+public class NslKdd {
+
+	public static void main(String[] args) { 
+		long initTime, startTime, endTime;
+		short windowSize = 60;// minutes
+		
+		String filePath = "/home/thiago/tmp/OriginalNSLKDDZip";
+		SimpleDateFormat sdf = new SimpleDateFormat(DateUtil.usDateTimeMS);
+
+		// Parsing
+		initTime = startTime = System.currentTimeMillis();
+		HashMap<String,Integer> matrixRowIndeces = new HashMap<String,Integer>();
+		HashMap<Long,HashMap<String,Integer>> values = StoregridCase.parseCSV(filePath, sdf, matrixRowIndeces);
+		System.out.println(values.size());
+		endTime = System.currentTimeMillis();
+		System.out.println("ParsingTime: " + (endTime - startTime));
+
+		// Data Modeling
+		startTime = System.currentTimeMillis();
+		HashMap<Long,Integer> matrixColumnIndeces = new HashMap<Long,Integer>();		
+		Matrix[] matrices = StoregridCase.modelIntoMatrices(windowSize, matrixRowIndeces, values, matrixColumnIndeces);
+		endTime = System.currentTimeMillis();
+		System.out.println("DataModelingTime: " + (endTime - startTime));
+
+		// EigenAnalysis
+		startTime = System.currentTimeMillis();
+		double[][] largestEigValCov = getLargestEigValCov(matrices);
+		endTime = System.currentTimeMillis();
[…]

### 2016-10-23 mos-api/src/main/java/br/unb/lasp/anomaly/mos/cases/StoregridCase.java

@@ -1,0 +1,199 @@
+package br.unb.lasp.anomaly.mos.cases;
+
+import java.io.File;
+import java.io.FileNotFoundException;
+import java.io.FileReader;
+import java.io.IOException;
+import java.text.ParseException;
+import java.text.SimpleDateFormat;
+import java.util.Calendar;
+import java.util.Collections;
+import java.util.Date;
+import java.util.HashMap;
+import java.util.List;
+import java.util.Set;
+import java.util.StringTokenizer;
+
+import org.apache.commons.lang.time.DateUtils;
+
+import com.univocity.parsers.csv.CsvParser;
+import com.univocity.parsers.csv.CsvParserSettings;
+
+import Jama.EigenvalueDecomposition;
+import Jama.Matrix;
+import br.unb.lasp.anomaly.mos.EigenAnalysis;
+import br.unb.lasp.anomaly.mos.MOS;
+import br.unb.lasp.matrix.DateUtil;
+import br.unb.lasp.matrix.MatrixUtil;
+
+public class StoregridCase {
+
+	public static void main(String[] args) { 
+		long initTime, startTime, endTime;
+		short windowSize = 60;// minutes
+		
+		//		String folderPath = "/home/thiago/Dropbox/doutorado/tanya/storgrid.log/";
+		String filePath = "/home/thiago/Dropbox/doutorado/tanya/storgrid.log/4.1.log";
+		SimpleDateFormat sdf = new SimpleDateFormat(DateUtil.usDateTimeMS);
+
+		// Parsing
+		initTime = startTime = System.currentTimeMillis();
+		HashMap<String,Integer> matrixRowIndeces = new HashMap<String,Integer>();
+		HashMap<Long,HashMap<String,Integer>> values = StoregridCase.parseCSV(filePath, sdf, matrixRowIndeces);
+		System.out.println(values.size());
+		endTime = System.currentTimeMillis();
+		System.out.println("ParsingTime: " + (endTime - startTime));
+
+		// Data Modeling
+		startTime = System.currentTimeMillis();
+		HashMap<Long,Integer> matrixColumnIndeces = new HashMap<Long,Integer>();		
+		Matrix[] matrices = StoregridCase.modelIntoMatrices(windowSize, matrixRowIndeces, values, matrixColumnIndeces);
+		endTime = System.currentTimeMillis();
+		System.out.println("DataModelingTime: " + (endTime - startTime));
+
+		// EigenAnalysis
+		startTime = System.currentTimeMillis();
[…]

### 2016-10-23 mos-api/src/main/java/br/unb/lasp/parser/JxtaMessageViewDriver.java

@@ -1,0 +1,265 @@
+package br.unb.lasp.parser;
+
+import java.io.ByteArrayOutputStream;
+import java.io.DataInputStream;
+import java.io.IOException;
+import java.nio.BufferUnderflowException;
+import java.nio.ByteBuffer;
+import java.util.Arrays;
+import java.util.HashMap;
+import java.util.Set;
+import java.util.SortedMap;
+
+import org.jnetpcap.Pcap;
+import org.jnetpcap.packet.JPacket;
+import org.jnetpcap.packet.JPacketHandler;
+import org.jnetpcap.protocol.network.Ip4;
+import org.jnetpcap.protocol.tcpip.Tcp;
+
+public class JxtaMessageViewDriver {
+
+	public static void main(String[] args) {
+		
+	    try{
+            System.loadLibrary("jnetpcap");
+        }catch(Exception ex){
+            ex.printStackTrace();
+        }
+		
+		long t0,t1;
+		final String FILENAME = "/media/thiago/ubuntu/datasets/darpa/1998/01_sample_data/sample_data01.tcpdump";
+		final StringBuilder errbuf = new StringBuilder();
+		
+		t0 = System.currentTimeMillis();
+		final Pcap pcap = Pcap.openOffline(FILENAME, errbuf);
+		pcap.close();
+		t1 = System.currentTimeMillis();
+		
+	}
+
+//	public static HashMap<Integer,Jxta> generateFullSocketFlows(final StringBuilder errbuf, final Pcap pcap) {
+//
+//		final HashMap<Integer,Jxta> frags = new HashMap<Integer,Jxta>();
+//
+//		pcap.loop(Pcap.LOOP_INFINITE, new JPacketHandler<StringBuilder>() {					
+//
+//			Tcp tcp = new Tcp();
+//			Jxta jxta = new Jxta();
+//
+//			public void nextPacket(JPacket packet, StringBuilder errbuf) {				
+//
+//				if(packet.hasHeader(Tcp.ID)){
+//					packet.getHeader(tcp);
+//					long seqNumber = tcp.seq();
+//
+//					// Looking for tcp fragmentation 
+//					if(frags.size() > 0 && tcp.getPayloadLength() > 0){
+//						System.out.println("\n### Frame: " + packet.getFrameNumber());
+//
+//						Ip4 ip = new Ip4();
+//						packet.getHeader(ip);
+//
+//						// id = IP and port of source and destiny
[…]

### 2016-10-23 ransomdroid/app/src/main/java/br/unb/lasp/Global.java

@@ -11,7 +11,7 @@ public class Global {
     public static String pass = null;
     public static String uid = null;
     public static SecretKeySpec key = null;
-//    public static final String serverURL = "http://172.31.130.84:8080/mobisaude-services/mobile";
-    public static final String serverURL = "http://192.168.1.102:8080/mobisaude-services/mobile";
+    public static final String serverURL = "http://172.31.130.123:8080/mobisaude-services/mobile";
+//    public static final String serverURL = "http://192.168.1.102:8080/mobisaude-services/mobile";
 
 }

### 2016-11-02 docs/papers/2016JNCA/2016JNCA.tex
several updates
baseline para as modificacoes do paper JNCA
updates of introduction, related work and dataset description

@@
 This paper proposes a signal processing framework for the detection and identification of network attacks using concepts of model order selection (MOS), eigenvalues and similarity analysis.
-In order to validate the proposed framework, we consider a network traffic sample that contains malicious activity such as denial of service (DoS) and port probing.
+In order to validate the proposed framework, we consider network traffic datasets that contains malicious activity such as flood and port probing attacks.
 We propose to model the network traffic as a superposition of components, namely, user's operations (legitimate traffic), network service operation not related to the user (noise) and the malicious activity.
-The experiments performed in a real network show that the proposed blind detection approach achieves satisfactory levels of accuracy in terms of timely detection and identification of TCP/UDP ports under attack.
+The experiments performed in a real network and in a well known public dataset show that the proposed blind detection approach achieves satisfactory levels of accuracy in terms of timely detection and identification of TCP/UDP ports under attack.
 \end{abstract}
@@
 Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.
-
-In the context of anomaly-based schemes, this work proposes an approach for the detection of malicious traffic in computer networks.
+Additionally, anomaly detection techniques can be categorized in classification, statistical, Information theory and clustering based \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
+
+In the context of anomaly-based schemes, this work proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
 Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
@@
 In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack.
-In addition, we also evaluate the accuracy of the proposed framework.
-Additionally, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.
+In addition, we also evaluate the accuracy and performance of the proposed framework applied to a simulated scenario and to the DARPA 1998 Dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset.
+Furthermore, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.
 
@@
 Several methods have been proposed for the identification and characterization of malicious activity in computer networks.
-Classical methods typically employ data mining \cite{he2008applying,ghourabi2010data} and regular file analysis \cite{raynal2004honeypot} to detect patterns that indicate the presence of specific attacks in network traffic.
+Classical methods typically employ data mining \cite{he2008applying,ghourabi2010data,osanaiye2016distributed} and regular file analysis \cite{raynal2004honeypot} to detect patterns that indicate the presence of specific attacks in network traffic.
 
 Data mining is often used to describe the process of extracting useful information from large databases.
-Multiple methods of data mining are used in \cite{he2008applying} to analyze data flow in a network with the aim of identifying characteristics of malicious traffic in large scale environments.
+Multiple methods of data mining are used in \cite{he2008applying,osanaiye2016distributed} to analyze data flow in a network with the aim of identifying characteristics of malicious traffic in large scale environments.
 Researchers have applied data mining techniques in log analysis \cite{ghourabi2010data} to improve intrusion detection performance.
@@
 However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques.
+
+Callegari \emph{et al} \cite{Zonglin2009} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
+However, Callegari \emph{et al} focused in flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
+
+Lee \emph{et al.} \cite{Lee2013} proposed OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
+In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
+Additionally, Lee \emph{et al.} only evaluated their proposed scheme for covariance analysis, while we adopted an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
+
 Besides being prone to higher errors and false positives, such human intervention makes PCA not useful for real time applications.
@@
 Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
-Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work was not applied to probing and denial of service (DoS) attack detection, simultaneously.
+Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work was not applied to probe and flood attack detection, simultaneously.
 
@@
 
-Lee \emph{et al.} \cite{Lee2013} proposed OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
-In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
-Additionally, Lee \emph{et al.} only evaluated their proposed scheme for covariance analysis, while we adopted an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for DoS and probing attacks, respectively.
-
 The proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic.
-The proposed solution is automatic and blind for detection of time frames under probing and DoS attacks through MOS and eigen analysis.
+The proposed solution is automatic and blind for detection of time frames under probe and flood attacks through MOS and eigen analysis.
 Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.
+
+Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks.
+Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
+NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications.
+Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations.
+Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD datasets adopt a traffic aggregation by connections, we consider the use of a simulated scenario on a real network and the DARPA 1998 dataset, which is the source for the creation of the KDD 99 and NSL-KDD datasets.
+It is important to note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD datasets.
 
@@
 
-This section presents details of the simulated scenario along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic.
-Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce DoS and probing attacks.
-Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition, and Subsection \ref{sec:AnalyzedScenario} details the traffic of synflood, fraggle and port scan attacks.
+This section presents details of the simulated scenario and the selected cases of the DARPA dataset, along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic.
+Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
+Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition.
+Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discuss the use of the DARPA dataset for evaluation of the proposed approach.
 
@@
 
-The environment of the analyzed scenario is composed of two computers and one router with access to the Internet and to an internal network, where the simulation of legitimate traffic, noise, DoS and port scan attacks are performed.
+The environment of the analyzed scenario is composed of two computers and one router with access to the Internet and to an internal network, where the simulation of legitimate traffic, noise, flood and port scan attacks are performed.
 During the traffic generation, one computer assumes the role of the attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:fig1}.
@@
      \includegraphics[height=5cm, width=7cm]{results/figures/fig09.png}
-     \caption{Scenario to reproduce legitimate traffic, noise, DoS and port scan.}
+     \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
      \label{fig:fig1}
@@
 
-By modeling the dataset as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{S}), noise (\textbf{N}) and malicious traffic (\textbf{A}), according to the following expression:
+By modeling the dataset as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{U}), noise (\textbf{N}) and malicious traffic (\textbf{A}), according to the following expression:
 
@@
 The network attacks evaluated by this work are: synflood, fraggle and port scan.
-The first two attacks can be qualified as DoS attacks, while the last one can be qualified  as probing or port scanning attack.
+The first two attacks can be qualified as flood or denial of service (DoS) attacks, while the last one can be qualified as probe or port scanning attack.
+DoS attempts to block access to system or network resources, it is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks.
+Probe attacks scan the networks to to collect information about host, such as IP addresses, ports and services.
 
@@
 At the end of each port scanning, a RST/ACK packet is sent by the system that is performing the port scan.
-Therefore, a full connection or a complete three-way handshake is never established, which makes the detection of the attack sender more difficult, and requires approaches able to identify probing attacks without connection stablishment.
+Therefore, a full connection or a complete three-way handshake is never established, which makes the detection of the attack sender more difficult, and requires approaches able to identify probe attacks without connection stablishment.
 The UDP scan technique sends UDP packets to the destination port, and if it responds with a \emph{ICMP port unreachable} message, then it indicates that the scanned port is closed.
@@
 The incoming and outgoing packets analysis, for each port, shows the high correlation and similarity of TCP and UDP traffic during the simulated port scan attack.
+
+\subsection{The DARPA Dataset}
+\label{sec:DarpaDataset}
+
+The DARPA 1998 dataset\footnote{https://www.ll.mit.edu/ideval/data/} includes 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks.
+The attacks in this dataset
+can be grouped into: denial-of-service (DoS); remote to local (R2L), which is characterized by unauthorized access from a
+remote machine; user to root (U2R), which is characterized by unauthorized access to local
+super-user privileges; and probe attack.
+Since the proposed approach focus on flood and probe attack, we analyze the attacks that present behaviors similar to flood or probe attack.
+We observe that the most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack.
+One example is the occurrence of a neptune attack which sends 20 only SYN packets, what is a behavior that differs of the expected flood attack behavior.
+Therefore, we select the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey,osanaiye2016distributed}, and the cases that scan ports sending just a few packets.
+From the simulated probe attacks, we select the cases that rely on TCP or UDP connections.
+
+The data modeling follows the method described by the Subsection \ref{sec:ModelingData}, with time frames of 20 minutes, packet counting aggregation by minute and considering the traffic to the following ports: 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389, 443.
 
@@
 	\centering
-     \includegraphics[height=11cm, width=9cm]{mos_eigen_similarity.eps}
+     \includegraphics[height=11cm, width=9cm]{results/figures/mos_eigen_similarity.eps}
      \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
@@
 
-According to DoS and port scan attacks' behavior, DoS attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively.
-These characteristics are substantiated by the results obtained through the analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:experimentalresults}, which shows that the main components of DoS attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
-
-Therefore, to detect DoS attacks, it is necessary to calculate the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
+According to flood and port scan attacks' behavior, flood attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively.
+These characteristics are substantiated by the results obtained through the analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
+
+Therefore, to detect flood attacks, it is necessary to calculate the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
 
@@
 
-Once the $\boldsymbol{\hat{R}}_{yy}^{(q)}$ and $\boldsymbol{\hat{R}}_{zz}^{(q)}$ have been obtained for DoS and port scan attack detection, respectively, and since the next steps are the same for both sample covariance matrices, we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{C}$.
+Once the $\boldsymbol{\hat{R}}_{yy}^{(q)}$ and $\boldsymbol{\hat{R}}_{zz}^{(q)}$ have been obtained for flood and port scan attack detection, respectively, and since the next steps are the same for both sample covariance matrices, we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{C}$.
 Therefore, the following step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:eq06}), in order to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).
@@
 \end{equation}
+
 \begin{equation}\label{eq:eq060}
@@
 \end{equation}
-where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, and $\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of traffic to be performed the DoS and port scan attack identification.
+where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, and $\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of traffic to be performed the flood and port scan attack identification.
 
@@
 In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection.
-Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed DoS and port scan identification, and Section \ref{sec:Complexity} discuss the computational complexity of the proposed framework.
+Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification, and Section \ref{sec:Complexity} discuss the computational complexity of the proposed framework.
 
@@
 
-For the evaluation of MOS Schemes acuracy for DoS and port scan detection, the framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}.
-Through eigenvalue analysis of traffic with DoS or port scan attacks, it is possible to visualize a significant difference between the largest eigenvalues and the remain eigenvalues, which can indicate a relationship between an outlier and time frames under attack.
+For the evaluation of MOS Schemes accuracy for flood and port scan detection, the framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}.
+Through eigenvalue analysis of traffic with flood or port scan attacks, it is possible to visualize a significant difference between the largest eigenvalues and the remain eigenvalues, which can indicate a relationship between an outlier and time frames under attack.
 
@@
 
-These results highlight that all $q$-th time frames where a network attack was simulated, present high significant variance between the largest eigenvalue and the remaining eigenvalues, obtained from sample covariance matrix, for DoS detection, or from covariance matrix of zero mean and unitary standard deviation variables, for port scan detection.
+These results highlight that all $q$-th time frames where a network attack was simulated, present high significant variance between the largest eigenvalue and the remaining eigenvalues, obtained from sample covariance matrix, for flood detection, or from covariance matrix of zero mean and unitary standard deviation variables, for port scan detection.
 Therefore, we propose to apply the vector of the largest eigenvalues to MOS schemes in order to evaluate their accuracy for identification of time frames under attack, motivated by the fact that it is relevant to apply MOS schemes to automate the attack detection process, taking into account the characteristics of the evaluated eigenvalues.
@@
   \scriptsize
-  \caption{MOS schemes applied to port scan and DoS detection}
+  \caption{MOS schemes applied to port scan and flood detection}
   \label{tab:tab4}
@@
 Efficient Detection Criterion (EDC) and Exponential Fitting Test (EFT) are the most effective schemes, correctly estimating the number of attacks in comparison to the expected values for effetctive attack detection, as defined by the column of real values in Table \ref{tab:tab4}.
-The AIC and MDL schemes are satisfactory only for port scan detection, however SURE and RADOI schemes did not show effective results for port scan or DoS detection.
-
-Although EDC and EFT presented the same accuracy on the evaluation, the EDC scheme requires less processing time than EFT, which is an important criteria to select EDC as the MOS scheme for DoS and port scan detection on the remain experiments.
+The AIC and MDL schemes are satisfactory only for port scan detection, however SURE and RADOI schemes did not show effective results for port scan or flood detection.
+
+Although EDC and EFT presented the same accuracy on the evaluation, the EDC scheme requires less processing time than EFT, which is an important criteria to select EDC as the MOS scheme for flood and port scan detection on the remain experiments.
 
@@
 
-These results highlight the capabilitiy of change detection based on similarity between legitimate and malicious traffic from DoS or port scan attacks, endorsing the effectiveness and safety for adoption of threshold for attack detection through eigen similarity analysis.
+These results highlight the capability of change detection based on similarity between legitimate and malicious traffic from flood or port scan attacks, endorsing the effectiveness and safety for adoption of threshold for attack detection through eigen similarity analysis.
 
@@
 Given $\hat{N}$, which is the set of estimated $n$-th minutes under attack, it is possible to apply cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port, during a $n$-th minute.
-Therefore, the incremental individualized and individual approaches of eigen similarity analysis were evaluated, for detection of ports under DoS and port scan attacks, according to results presented in following tables.
+Therefore, the incremental individualized and individual approaches of eigen similarity analysis were evaluated, for detection of ports under flood and port scan attacks, according to results presented in following tables.
 For this evaluation, the $v$ last most significant eigenvectors without attack was used as reference for similarity analysis against each target port $m$-th.
@@
 
-According to the shown significant lack of similarity between legitimate and malicious traffic, it is possible to adopt safe thresholds for DoS and port scan detection through eigen similarity analysis.
+According to the shown significant lack of similarity between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
 
@@
 
-This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and DoS attacks, but still requiring more information for detailed attack detection.
+This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
 Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.
@@
 
-According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for DoS and port scan detection through eigen similarity analysis.
+According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
 

### 2016-11-02 docs/papers/2016JNCA/2016JNCA.tex
version with results discussion

@@
 In Section \ref{sec:relatedworks}, related works are discussed.
-Section \ref{sec:datamodel} presents the data model.
-Section \ref{sec:prop_getv} describes the proposed framework for blind and automatic detection of malicious traffic.
-Section \ref{sec:experimentalresults} discusses the experimental validation and presents the experimental results.
+Section \ref{sec:datamodel} presents the data model and the evaluated datasets.
+Section \ref{sec:prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks.
+Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:Complexity} discuss the computational complexity of the proposed framework and evaluates the required processing time for simulated scenarios.
 Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
@@
 Lu and Ghorbani \cite{Lu2009} proposed a network anomaly detection model based on network flow, wavelet approximation, and system identification theory.
-However, their work did not address problems without significant outliers, such as port scan attacks.
+However, their work requires a training method to produce a prediction model for normal daily traffic and presents limitations on identification of behaviors without significant outliers, such as port scan attacks.
 Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
@@
 This section presents the performed experiments and the acquired results.
-First, in Section \ref{sec:AnalyzedScenario}, the scenario adopted in the experiments is summarized.
-Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames.
-In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection.
-Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification, and Section \ref{sec:Complexity} discuss the computational complexity of the proposed framework.
-
-\subsection{Analyzed Scenario}
+First, in Section \ref{sec:AnalyzedScenario}, the simulated scenario adopted in the experiments is summarized.
+Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the simulated scenario.
+In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated dataset.
+Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the simulated scenario.
+Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 dataset.
+
+\subsection{Simulated Scenario}
 \label{sec:AnalyzedScenario}
@@
 
-\subsection{Complexity Analysis}
+\subsection{DARPA Scenario}
+\label{sec:DarpaEvaluation}
+
+This subsection presents a summarized view of results obtained from the application of the proposed framework, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 dataset.
+Since the proposed framework is detailed explained in previous sections, here we focus on the parameter selection, dataset evaluation and results for flood and probe attack identification.
+
+The DARPA dataset includes 7 weeks of sniffed traffic saved into raw packet data.
+The traffic and the labeled attacks are grouped by week and day, with information of the number and types of attacks per day, but also providing the start time for each labeled attack.
+For this evaluation, we performed an evaluation per day, considering the network traffic of 24 hours split into $Q$ time frames of 60 minutes ($N = 60$) and aggregated by minute and by port number.
+For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.
+
+Since the proposed framework focus on flood and probe attack detection, we evaluated only the attacks with behavior similar to flood or probe attack.
+Initially we selected all DoS and probe attacks, but we observed that the most cases of DoS focus on exploit system vulnerabilities instead of flooding attack, and most of probe attacks focus on ICMP intead of port scanning.
+Therefore, for evaluation of the proposed approach for flood and probe attack detection, it is necessary to select cases that implements flood or port scan behaviors.
+We select the following week-day-attack cases:
+
+\begin{enumerate}
+	\item week3-thursday-neptune;
+	\item week4-friday-portsweep;
+	\item week5-thursday-neptune;
+	\item week5-thursday-portsweep;
+	\item week5-friday-portsweep;
+	\item week6-wednesday-neptune;
+	\item week6-thursday-neptune;
+	\item week7-wednesday-portsweep.
+\end{enumerate}
+
+The analysis based on sample covariation of zero mean variables was evaluated for flooding behavior of netpune attacks, obtaining rates of 100\% for true positive (TP) detection and 60\% for false positive (FP) detection from 30 time frames.
+The results also show 50\% of misclassification rate \cite{bhuyan2014network}, which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
+The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time  from one source to one target, with no variation on IP source or target port.
+This observation corroborates with previous evaluations of the DARPA dataset that highlight issues regarding traffic redundancy.
+
+The analysis based on covariance of zero mean and unitary standard deviation variables was evaluated for port scan attacks, including probe attacks and DoS attacks that send few packets for several ports in order to exploit some vulnerability.
+The results show rates of 76.9\% for TP detection and 23\% for FP detection from 94 time frames.
+The observed misclassification rate for this scenario was 32.7\%.
+We observed that all FN cases are probe attacks with a time delay between scanning one port and start scanning the next port, what can be called as sparse probe attacks.
+Cases with a delay of one minute or more were not detected by the proposed approach.
+
+The performance of detection rate of flooding attacks is compared with the method proposed by Callegari \emph{et al} \cite{Zonglin2009}.
+This work is a statistical method, based on PCA, without training or learning methods, even though it relies on visual analysis for principal components selection.
+The best detection rate of \cite{Zonglin2009} was 82\% for detection of synthetically added flood attacks, while our proposal obtains 100\% of detection rate for detection of flood attacks of DARPA dataset.
+
+Due to the lack of statistical techniques without training or learning methods for detection of probe attacks, we compare our approach with Lu and Ghorbani's \cite{Lu2009} proposal, which is a network anomaly detection model based on signal processing techniques that uses DARPA dataset for evaluation.
+The results of \cite{Lu2009} show the best detection rate of 94.67\% in terms of general attack instance detection, but shows a case case with 50\% of attack instance detection for the portsweep attack.
+The proposed approach presents 76.9\% of detection rate measured specifically for probe attacks, without the requirement of learning or training methods, in contrast to Lu and Ghorbani's \cite{Lu2009} work.
+
+\section{Complexity Analysis}
 \label{sec:Complexity}
 
-This subsection discusses the computational complexity of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
+This section discusses the computational complexity of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
 
@@
 According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
+
+falar sobre as melhorias para reduzir falso positivo e também para a detecção de probe atacks disperos
 

### 2016-11-04 docs/papers/2016JNCA/2016JNCA.diff1.tex
first round of corrections

--- 2016-11-02 docs/papers/2016JNCA/2016JNCA.tex
@@
 
+% table
+\usepackage{color, colortbl}
+\definecolor{Gray}{gray}{0.9}
+
 % Algorithmic modifications
@@
 
-\section{Complexity Analysis}
+\section{Complexity Analysis and Performance Evaluation}
 \label{sec:Complexity}
 
-This section discusses the computational complexity of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
+This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
 
@@
 
+For better understanding the scalability and impact of configurations of $N$, $M$ and $Q$, we evaluated the processing time requited for different scenarios of parameter configurations and dataset, measuring the processing time of:
+
+\begin{enumerate}
+	\item Eigen analysis based on sample covariance of zero mean;
+	\item Eigen analysis based on sample covariance of zero mean and unitary standard deviation;
+	\item EDC MOS scheme;
+\end{enumerate}
+
+The performance evaluation focus on the main steps, which were previously discussed regarding the complexity analysis.
+The data modeling is also a time consuming step, however its processing can be optimized trough distributed processing techniques, such as MapReduce, achieving high throughput for packet counting or even for deep packet inspection \cite{Vieira2013}.
+
+The experiments were performed in a desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM, considering: variations on the network traffic time; the frame size denoted as $N$; the number of network ports denoted as $M$; the mean processing time for eigen analysis based on sample covariance of zero mean, denoted as 1-time; the mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation, denoted as 2-time; and the mean processing time for EDC MOS scheme, denoted as 3-time.
+The mean time calculations was obtained from 200 measurement repetitions, in order to obtain reliable values.
+
+Table \ref{tab:11} presents the measured results.
+The experiment considered traffic time of 16, 20 and 22 hours, according to the selected traffic time per day available by the DARPA dataset.
+It is possible to observe the processing time increases according to the increment in traffic time, around 2 or 3 times for 1-time and 2-time, but the worst measured processing time is 4.7250 milliseconds.
+
+\begin{table*}[!t]
+	\caption{Processing time of the main steps for anomaly detection}
+	\label{tab:11}
+	\centering
+	\begin{tabular}{|r|r|r|r|r|r|r|}
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Traffic Time\\(hour)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Frame Size\\(min)\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Num.
+Ports\end{tabular}	& \begin{tabular}[x]{@{}l@{}}1-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}2-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}3-time\\(ms)\end{tabular}	\\ \hline
+16	& 10	& 17	& 0.7900	& 0.8100	& 0.0650\\ \hline
+16	& 20	& 17	& 0.5250	& 0.5950	& 0.0100\\ \hline
+16	& 60	& 17	& 0.9700	& 1.1400	& 0.0250\\ \hline
+16	& 120	& 17	& 0.6050	& 0.6100	& 0.0050\\ \hline
+16	& 60	& 34	& 1.2750	& 1.2200	& 0.0050\\ \hline
+16	& 120	& 34	& 1.1200	& 1.1700	& 0.0050\\ \hline
+20	& 10	& 17	& 2.7950	& 2.8950	& 1.1000\\ \hline
+20	& 20	& 17	& 2.0700	& 2.0200	& 0.3500\\ \hline
+20	& 60	& 17	& 1.0250	& 1.0450	& 0.0650\\ \hline
+20	& 120	& 17	& 1.0000	& 1.0700	& 0.0350\\ \hline
+20	& 60	& 34	& 2.9650	& 3.2100	& 0.0400\\ \hline
+20	& 120	& 34	& 2.9950	& 3.1150	& 0.0200\\ \hline
+22	& 10	& 17	& 4.7250	& 4.0850	& 1.4600\\ \hline
+22	& 20	& 17	& 2.3200	& 2.6800	& 0.2450\\ \hline
+22	& 60	& 17	& 1.0700	& 1.1200	& 0.0300\\ \hline
+22	& 120	& 17	& 0.9900	& 1.0500	& 0.0250\\ \hline
+22	& 60	& 34	& 3.0850	& 3.1250	& 0.0650\\ \hline
+22	& 120	& 34	& 2.8100	& 2.9600	& 0.0250\\ \hline
+	\end{tabular}
+\end{table*}
+
+It is possible to observe that the processing time increases with the frame size $N$ decreasing, therefore it is possible to evaluate the frame size that produces better identification rates and acceptable processing time.
+The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations, since the significant increase of processing time observed between scenarios considering 17 or 34 ports, with growth between 7\% and 199\%.
+
 \section{Conclusion and Future Works}
@@
 The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack.
+The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
 
@@
 
-falar sobre as melhorias para reduzir falso positivo e também para a detecção de probe atacks disperos
-
-Future research is directed to the fully automated blind attack detection, extending the evaluation to performance and accuracy analyis of the approach to well-known datasets of network attack detection, and analysing the performance according to parameter configuration in order to better understand the impact of $N, M$ and $Q$ and identify their best configuration for obtaining gains into processing time and acuracy.
+Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks.
+Additionally, future research can evaluate the application of the proposed approach to different domains, considering cases that are aware to behavioral analysis.
 

### 2016-12-22 docs/papers/2016MONE/submission/respose01/mone-response_to_reviewers.txt
preparacao para respostas aos revisores

Response to the Handling Editor and Reviewers

Submission No.: MONE-D-16-00333
Title: Offline Mode for Corporate Mobile Client Security Architecture
SM 188 – Sustainable Computing Techniques for Mobile Networks
Mobile Networks and Applications

Dear Editor-in-Chief and reviewers,
The authors would like to thank the Associate Editor-in-Chief and the Reviewers for volunteering their time in reviewing our manuscript and providing us with valuable comments which allowed us to significantly improve the paper. We have carefully revised the manuscript based on the comments and suggestions. All changes in the paper have been marked with a color. The following is our point-by-point responses to the raised comments.

Kind regards,

The authors.

1. REVIEWER #1
1.1. COMMENT: The paper presents a reasonable approach to offline mode for reading secure documents on mobile clients without consuming much of the device resources under a restricted adversary model.
1.1.1. REPLY: Agradecer
1.2. COMMENT: The adversary model seems not complete. (1) The adversary model mentioned in Subsection 6.1 is not clear because affirmative statements are mixed with negative statements. (2) May the adversary disable the TIM?
1.2.1. REPLY: Aguardando resposta de Tanya
1.3. COMMENT: In Subsection 5.2, the description of ABE components is unclear: (1) typo of redundant double prime('') in PK, (2) no description of basic properties of bilinear paring function, (3) no description of the meanings of the notations w and s. (4) bad-form notation "for all i in {t_i}_M" (because "i" cannot belong to the set).
1.3.1. REPLY: Aguardando resposta de Tanya
1.4. COMMENT: In Subsection 5.3, the description of SSS (modular secret sharing) is unclear: (1) Di is a typo, (2) the notation based on Chinese Remainder Theory (CRT) is not well presented, (3) the meanings of d, S, p_0, and p_1 are not clear, (4) What are the reasons for S being equivalent to s and s_1 in modulo p_0 and p_1, respectively?
1.4.1. REPLY: Aguardando resposta de Tanya
1.5. COMMENT: In Subsection 6.2, it may be better to change "operations 1)-7)" to "actions 1)-7) (as mentioned in the offline mode)".
1.5.1. REPLY: Aguardando resposta de Tanya
1.6. COMMENT: There is little description about how to retrieve the logs from the Galaxy mobile devices and the correctness of anomaly detection in the experiments.
1.6.1. REPLY: Interagir com anton. Correctness vai ser tratado em trabalhos futuros

2. REVIEWER #3
2.1. COMMENT: In the related work, these following references need to be considered to enhance the completeness of the work. Heydari, Mohammad, et al. "An efficient password-based authenticated key exchange protocol with provable security for mobile client-client networks." Wireless Personal Communications 88.2 (2016). Mayrhofer, René. "An architecture for secure mobile devices." Security and Communication Networks 8.10 (2015): 1958-1970. Shila, Devu Manikantan, et al. "AMCloud: Towards a Secure Autonomic Mobile Ad Hoc Cloud Computing System." IEEE Wireless Communications(2016). Chang, Hyunseok, et al. "Design and Architecture of a Software Defined Proximity Cloud." Advances in Mobile Cloud Computing Systems (2015): 123. Xia, Yubin, et al. "TinMan: eliminating confidential mobile data exposure with security oriented offloading." Proceedings of the Tenth European Conference on Computer Systems. ACM, 2015. Kulkarni, Pallavi, and Rajashri Khanai. "Addressing mobile Cloud Computing security issues: a survey." Communications and Signal Processing (ICCSP), 2015 International Conference on. IEEE, 2015.
2.1.1. REPLY: Fazer
2.2. COMMENT: More discussions on the figure 1 and 2 should be added. The current version is not easy to understand.
2.2.1. REPLY: Tanya
2.3. COMMENT: The figure 8 should be explained more in detail.
2.3.1. REPLY: Tanya
2.4. COMMENT: The formula 5 should be discussed more, the related sentence should be introduced with more theoretical support. The set of obtained vectors y(q)m composes the zero mean matrix Y(q), then the zero mean covariance ma-trix ^R(q)=yy can be calculated as follows.
2.4.1. REPLY: Fazer

*** Definir data de retorno que seja suficiente para podermos fazer por conta própria se Tanya não responder, considerando o deadline de 11/01

### 2016-12-22 docs/papers/2016JNCA/2016JNCA.tex
preparacao para respostas aos revisores

@@
 This paper proposes a signal processing framework for the detection and identification of network attacks using concepts of model order selection (MOS), eigenvalues and similarity analysis.
-In order to validate the proposed framework, we consider network traffic datasets that contains malicious activity such as flood and port probing attacks.
+In order to validate the proposed framework, we consider network traffic datasets that contain malicious activity such as flood and port probing attacks.
 We propose to model the network traffic as a superposition of components, namely, user's operations (legitimate traffic), network service operation not related to the user (noise) and the malicious activity.
-The experiments performed in a real network and in a well known public dataset show that the proposed blind detection approach achieves satisfactory levels of accuracy in terms of timely detection and identification of TCP/UDP ports under attack.
+The experiments performed in a real network and also using the DARPA 1998 public dataset show that the proposed blind detection approach achieves satisfactory levels of accuracy in terms of timely detection and identification of TCP/UDP ports under attack.
 \end{abstract}
@@
 Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.
-Additionally, anomaly detection techniques can be categorized in classification, statistical, Information theory and clustering based \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
+Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
 
@@
 In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack.
-In addition, we also evaluate the accuracy and performance of the proposed framework applied to a simulated scenario and to the DARPA 1998 Dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset.
+In addition, we also evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset.
 Furthermore, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.
@@
 Section \ref{sec:prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks.
-Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:Complexity} discuss the computational complexity of the proposed framework and evaluates the required processing time for simulated scenarios.
+Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:Complexity} discusses the computational complexity of the proposed framework and evaluates the required processing time for tested scenarios.
 Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
@@
 
-Callegari \emph{et al} \cite{Zonglin2009} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
-However, Callegari \emph{et al} focused in flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
-
-Lee \emph{et al.} \cite{Lee2013} proposed OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
+Callegari \emph{et al} \cite{Zonglin2009} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
+However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
+
+Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
 In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
-Additionally, Lee \emph{et al.} only evaluated their proposed scheme for covariance analysis, while we adopted an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
+Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
 
@@
 Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
-Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work was not applied to probe and flood attack detection, simultaneously.
+Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work is not applied to probe and flood attack detection, simultaneously.
 
@@
 Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations.
-Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD datasets adopt a traffic aggregation by connections, we consider the use of a simulated scenario on a real network and the DARPA 1998 dataset, which is the source for the creation of the KDD 99 and NSL-KDD datasets.
-It is important to note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD datasets.
+Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD datasets adopt a traffic aggregation by connections, we consider the use of a experimental scenario on a real network and the DARPA 1998 dataset, which is the source for the creation of the KDD 99 and NSL-KDD datasets.
+Note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD datasets.
 
@@
 
-This section presents details of the simulated scenario and the selected cases of the DARPA dataset, along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic.
+This section presents details of the experimental scenario and the selected cases of the DARPA dataset, along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic.
 Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
 Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition.
-Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discuss the use of the DARPA dataset for evaluation of the proposed approach.
+Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
 
@@
      \centering 
-     \includegraphics[height=5cm, width=7cm]{results/figures/fig09.png}
+     \includegraphics[height=6cm, width=9cm]{results/figures/fig09.png}
      \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
@@
 
-A network traffic log is commonly formed by timestamp, protocol, source IP address, source port, destination IP address, destination port and additional information, according to the type of transport protocol used.
+A network traffic log is commonly formed by timestamp, protocol, source IP address, source port, destination IP address, destination port and additional information, according to the type of the used transport protocol.
 The following TCP traffic log is presented in order to exemplify the collected data:
@@
 The first two attacks can be qualified as flood or denial of service (DoS) attacks, while the last one can be qualified as probe or port scanning attack.
-DoS attempts to block access to system or network resources, it is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks.
-Probe attacks scan the networks to to collect information about host, such as IP addresses, ports and services.
+DoS attempts to block access to system or network resources and is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks.
+Probe attacks scan the networks to collect information about host, such as IP addresses, ports and services.
 
@@
 super-user privileges; and probe attack.
-Since the proposed approach focus on flood and probe attack, we analyze the attacks that present behaviors similar to flood or probe attack.
+Since the proposed approach focus on flood and probe attack, the analysis concentrates on the attacks that present behaviors similar to flood or probe attack.
 We observe that the most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack.
 One example is the occurrence of a neptune attack which sends 20 only SYN packets, what is a behavior that differs of the expected flood attack behavior.
-Therefore, we select the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey,osanaiye2016distributed}, and the cases that scan ports sending just a few packets.
+Therefore, there were selected the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey,osanaiye2016distributed}, and the cases that scan ports sending just a few packets.
 From the simulated probe attacks, we select the cases that rely on TCP or UDP connections.
@@
 threshold $l$, assuming that if $s_n < l$, then a network attack is identified during the $n$-th minute.
-Therefore, the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack was identified, according to
+Therefore, the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack is identified, according to
 
@@
 This section presents the performed experiments and the acquired results.
-First, in Section \ref{sec:AnalyzedScenario}, the simulated scenario adopted in the experiments is summarized.
-Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the simulated scenario.
+First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
+Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario.
 In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated dataset.
-Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the simulated scenario.
+Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario.
 Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 dataset.
 
-\subsection{Simulated Scenario}
+\subsection{Experimental Scenario}
 \label{sec:AnalyzedScenario}
@@
 This paper proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$.
-Hence, the proposed framework was applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.
+Hence, the proposed framework is applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.
 
@@
 Three approaches were evaluated for eigen similarity analysis: incremental, individual and incremental individualized approaches.
-For the incremental individualized approach, each minute was incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
+For the incremental individualized approach, each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
 Subsequently, $\boldsymbol{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated.
@@
 This subsection presents a summarized view of results obtained from the application of the proposed framework, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 dataset.
-Since the proposed framework is detailed explained in previous sections, here we focus on the parameter selection, dataset evaluation and results for flood and probe attack identification.
+Since the proposed framework is detailed in Section \ref{sec:prop_getv} and in Subsections \ref{sec:AnalyzedScenario}, \ref{sec:largesteigenvaluesanalysis}, \ref{sec:MOSSchemesEvaluation}, \ref{sec:EigenvalueAnalysis} and \ref{sec:EigenSimilarityAnalysis}, here the focus is on the parameter selection, dataset evaluation and results for flood and probe attack identification.
 
@@
 The traffic and the labeled attacks are grouped by week and day, with information of the number and types of attacks per day, but also providing the start time for each labeled attack.
-For this evaluation, we performed an evaluation per day, considering the network traffic of 24 hours split into $Q$ time frames of 60 minutes ($N = 60$) and aggregated by minute and by port number.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.
-
-Since the proposed framework focus on flood and probe attack detection, we evaluated only the attacks with behavior similar to flood or probe attack.
-Initially we selected all DoS and probe attacks, but we observed that the most cases of DoS focus on exploit system vulnerabilities instead of flooding attack, and most of probe attacks focus on ICMP intead of port scanning.
+For this evaluation, an evaluation per day was performed, considering the network traffic of 24 hours split into $Q$ time frames of 60 minutes ($N = 60$) and aggregate by minute and by port number.
+For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ is obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.
+
+Since the proposed framework focus on flood and probe attack detection, only the attacks with behavior similar to flood or probe attack were evaluated.
+Initially all DoS and probe attacks were selected, but it was observed that the most cases of DoS focus on exploit system vulnerabilities instead of flooding attack, and most of probe attacks focus on ICMP intead of port scanning.
 Therefore, for evaluation of the proposed approach for flood and probe attack detection, it is necessary to select cases that implements flood or port scan behaviors.
-We select the following week-day-attack cases:
+The following week-day-attack cases  were selected:
 
@@
 
-The analysis based on sample covariation of zero mean variables was evaluated for flooding behavior of netpune attacks, obtaining rates of 100\% for true positive (TP) detection and 60\% for false positive (FP) detection from 30 time frames.
-The results also show 50\% of misclassification rate \cite{bhuyan2014network}, which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
+\begin{table*}[!t]
+	\caption{Results of the attack detection evaluation}
+	\label{tab:tab12}
+	\centering
+	\begin{tabular}{|c|c|c|c|c|}
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Solution\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Attack Type\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Metric\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Result\end{tabular} \\ \hline
+	Proposed Work	&Flooding	&True Positive	&100.00 \%\\ \hline
+	Proposed Work	&Flooding	&False Positive	&60.00 \%\\ \hline
+	Proposed Work	&Flooding	&Misclassification	&50.00 \%\\ \hline
+	Proposed Work	&Probe	&True Positive	&76.92 \%\\ \hline
+	Proposed Work	&Probe	&False Positive	&18.52 \%\\ \hline
+	Proposed Work	&Probe	&Misclassification	&32.73 \%\\ \hline
+	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&True Positive	&82.00 \%\\ \hline
+	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&False Positive	&-\\ \hline
+	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&Misclassification	&-\\ \hline
+	Lu and Ghorbani \cite{Lu2009}	&Overall	&True Positive	&94.67 \%\\ \hline
+	Lu and Ghorbani \cite{Lu2009}	&Overall	&False Positive	&-\\ \hline
+	Lu and Ghorbani \cite{Lu2009}	&Overall	&Misclassification	&-\\ \hline
+	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&True Positive	&50.00 \%\\ \hline
+	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&False Positive	&-\\ \hline
+	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&Misclassification	&-\\ \hline
+	\end{tabular}
+\end{table*}
+
+Table \ref{tab:tab12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}.
+The analysis based on sample covariation of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames.
+The results also show 50.00 \% of misclassification rate , which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
 The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time  from one source to one target, with no variation on IP source or target port.
@@
 The analysis based on covariance of zero mean and unitary standard deviation variables was evaluated for port scan attacks, including probe attacks and DoS attacks that send few packets for several ports in order to exploit some vulnerability.
-The results show rates of 76.9\% for TP detection and 23\% for FP detection from 94 time frames.
-The observed misclassification rate for this scenario was 32.7\%.
-We observed that all FN cases are probe attacks with a time delay between scanning one port and start scanning the next port, what can be called as sparse probe attacks.
+The results show rates of 76.92 \% for TP detection and 18.52 \% for FP detection from 94 time frames.
+The observed misclassification rate for this scenario is 32.73 \%.
+It was observed that all FN cases are probe attacks with a time delay between scanning one port and start scanning the next port, what can be called as sparse probe attacks.
 Cases with a delay of one minute or more were not detected by the proposed approach.
@@
 This work is a statistical method, based on PCA, without training or learning methods, even though it relies on visual analysis for principal components selection.
-The best detection rate of \cite{Zonglin2009} was 82\% for detection of synthetically added flood attacks, while our proposal obtains 100\% of detection rate for detection of flood attacks of DARPA dataset.
-
-Due to the lack of statistical techniques without training or learning methods for detection of probe attacks, we compare our approach with Lu and Ghorbani's \cite{Lu2009} proposal, which is a network anomaly detection model based on signal processing techniques that uses DARPA dataset for evaluation.
-The results of \cite{Lu2009} show the best detection rate of 94.67\% in terms of general attack instance detection, but shows a case case with 50\% of attack instance detection for the portsweep attack.
-The proposed approach presents 76.9\% of detection rate measured specifically for probe attacks, without the requirement of learning or training methods, in contrast to Lu and Ghorbani's \cite{Lu2009} work.
-
-\section{Complexity Analysis}
+The best detection rate of \cite{Zonglin2009} was 82.00 \% for detection of synthetically added flood attacks, while this current proposal obtains 100.00 \% of detection rate for detection of flood attacks of DARPA dataset.
+
+Due to the lack of statistical techniques without training or learning methods for detection of probe attacks, this proposed approach is compared to the Lu and Ghorbani's \cite{Lu2009} proposal, which is a network anomaly detection model based on signal processing techniques that uses DARPA dataset for evaluation.
+The results of \cite{Lu2009} show the best detection rate of 94.67 \% in terms of general attack instance detection, but shows a case case with 50.00 \% of attack instance detection for the portsweep attack.
+The proposed approach presents 76.92 \% of detection rate measured specifically for probe attacks, without the requirement of learning or training methods, in contrast to Lu and Ghorbani's \cite{Lu2009} work.
+
+\section{Performance Evaluation}
 \label{sec:Complexity}
 
-This section discusses the computational complexity of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
-
-The EVD, calculated according to (\ref{eq:eq06}), requires the previous calculation of covariance matrix, according to Equations \ref{eq:eq02}, \ref{eq:eq03}, \ref{eq:eq04} and \ref{eq:eq05}.
+This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
+
+\subsection{Complexity Analysis}
+\label{sec:ComplexityAnalysis}
+The EVD, calculated according to (\ref{eq:eq06}), requires the previous calculation of covariance matrix, according to Equations (\ref{eq:eq02}), (\ref{eq:eq03}), (\ref{eq:eq04}) and (\ref{eq:eq05}).
 The covariance matrix calculation is $O(M^2N)$ and the EVD is $O(N^3)$, where $M$ denotes the number of network ports and $N$ denotes the period time.
@@
 The computational complexity evaluation for MOS focuses on EDC scheme, since EDC requires less processing time than EFT but presents the same accuracy for the evaluated scenario.
-EDC scheme is $O(Q log Q + Q + Q log Q)$ and its worst-case running time can be represented as $O(Q log Q)$, where $Q$ denotes the number of time frames.
+EDC scheme is $O(Q \log Q + Q + Q \log Q)$ and its worst-case running time can be represented as $O(Q \log Q)$, where $Q$ denotes the number of time frames.
 
@@
 
-Therefore, the proposed framework is $O(N^3 + Q log Q + \hat{d}Q + N^3)$ and its worst-case running time is $O(N^3)$.
+Therefore, the proposed framework is $O(N^3 + Q \log Q + \hat{d}Q + N^3)$ and its worst-case running time is $O(N^3)$.
 The computational complexity of EVD is predominant in the framework, but the approach splits the data into time frames with period time $N$, which makes possible to limitate the growth of $N$ even for evaluations of cases with total time larger than $N$, reducing the impact caused by the computational complexity of EVD.
+
+\subsection{Processing Time Analysis}
+\label{sec:ProcessingAnalysis}
+For better understanding the scalability and impact of configurations of $N$, $M$ and $Q$, the processing time required for different scenarios of parameter configurations and dataset were evaluated, measuring the processing time of:
+
+\begin{enumerate}
+	\item Eigen analysis based on sample covariance of zero mean;
+	\item Eigen analysis based on sample covariance of zero mean and unitary standard deviation;
+	\item EDC MOS scheme;
+\end{enumerate}
+
+The performance evaluation focus on the main steps, which are discussed in subsection \ref{sec:ComplexityAnalysis} regarding the complexity analysis.
+The data modeling is also a time consuming step, however its processing can be optimized through distributed processing techniques, such as MapReduce, achieving high throughput for packet counting or even for deep packet inspection \cite{Vieira2013}.
+
+The experiments were performed in a desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM, considering: variations on the network traffic time; the frame size denoted as $N$; the number of network ports denoted as $M$; the mean processing time for eigen analysis based on sample covariance of zero mean, denoted as 1-time; the mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation, denoted as 2-time; and the mean processing time for EDC MOS scheme, denoted as 3-time.
+The mean time calculations was obtained from 200 measurement repetitions, in order to obtain reliable values.
+
+\begin{table*}[!t]
+	\caption{Processing time of the main steps for anomaly detection}
+	\label{tab:11}
+	\centering
+	\begin{tabular}{|r|r|r|r|r|r|r|}
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Traffic Time\\(hour)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Frame Size\\(min)\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Num.
+Ports\end{tabular}	& \begin{tabular}[x]{@{}l@{}}1-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}2-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}3-time\\(ms)\end{tabular}	\\ \hline
+16	& 10	& 17	& 0.7900	& 0.8100	& 0.0650\\ \hline
+16	& 20	& 17	& 0.5250	& 0.5950	& 0.0100\\ \hline
+16	& 60	& 17	& 0.9700	& 1.1400	& 0.0250\\ \hline
+16	& 120	& 17	& 0.6050	& 0.6100	& 0.0050\\ \hline
+16	& 60	& 34	& 1.2750	& 1.2200	& 0.0050\\ \hline
+16	& 120	& 34	& 1.1200	& 1.1700	& 0.0050\\ \hline
+20	& 10	& 17	& 2.7950	& 2.8950	& 1.1000\\ \hline
+20	& 20	& 17	& 2.0700	& 2.0200	& 0.3500\\ \hline
+20	& 60	& 17	& 1.0250	& 1.0450	& 0.0650\\ \hline
+20	& 120	& 17	& 1.0000	& 1.0700	& 0.0350\\ \hline
+20	& 60	& 34	& 2.9650	& 3.2100	& 0.0400\\ \hline
+20	& 120	& 34	& 2.9950	& 3.1150	& 0.0200\\ \hline
+22	& 10	& 17	& 4.7250	& 4.0850	& 1.4600\\ \hline
+22	& 20	& 17	& 2.3200	& 2.6800	& 0.2450\\ \hline
+22	& 60	& 17	& 1.0700	& 1.1200	& 0.0300\\ \hline
+22	& 120	& 17	& 0.9900	& 1.0500	& 0.0250\\ \hline
+22	& 60	& 34	& 3.0850	& 3.1250	& 0.0650\\ \hline
+22	& 120	& 34	& 2.8100	& 2.9600	& 0.0250\\ \hline
+	\end{tabular}
+\end{table*}
+
+Table \ref{tab:11} presents the measured results.
+The experiment considered traffic time of 16, 20 and 22 hours, according to the selected traffic time per day available by the DARPA dataset.
+Note that the processing time increases according to the increment in traffic time, around 2 or 3 times for 1-time and 2-time, but the worst measured processing time is 4.7250 milliseconds.
+
+According to Table \ref{tab:11}, the processing time increases with the frame size $N$ decreasing, therefore it is possible to evaluate the frame size that produces better identification rates and acceptable processing time.
+The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations, since the significant increase of processing time observed between scenarios considering 17 or 34 ports, with growth between 7 \% and 199 \%.
 
@@
 
-falar sobre as melhorias para reduzir falso positivo e também para a detecção de probe atacks disperos
-
-Future research is directed to the fully automated blind attack detection, extending the evaluation to performance and accuracy analyis of the approach to well-known datasets of network attack detection, and analysing the performance according to parameter configuration in order to better understand the impact of $N, M$ and $Q$ and identify their best configuration for obtaining gains into processing time and acuracy.
+Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks, considering the evaluation of novel datasets.
+Additionally, future research can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.
 

### 2017-01-03 docs/papers/2016JNCA/2016JNCA.tex
inclusao de trabalhos futuros para atender as consideracoes da revisao R1

@@
 In our previous work \cite{tenorio2013greatest}, the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE schemes are evaluated for synflood and port scan attack detection, showing that EDC and EFT are effective for detecting this kind of attacks.
-The present work extends that evaluation to also analyse the effectiveness of the listed MOS schemes for fraggle attack detection, as shown in Section \ref{sec:experimentalresults}.
+The present work extends that evaluation to also analyze the effectiveness of the listed MOS schemes for fraggle attack detection, as shown in Section \ref{sec:experimentalresults}.
 
@@
 Cosine similarity calculates the cosine of the angle between two vectors, which represents the similarity of values between the selected vectors.
-Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\boldsymbol{V}^{(q)}$ against the the most significant eigenvectors of time frame detected as under attack, to analyse similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.
+Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\boldsymbol{V}^{(q)}$ against the the most significant eigenvectors of time frame detected as under attack, to analyze similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.
 
@@
 
-Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks, considering the evaluation of novel datasets.
+Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets.
+Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high thoughput network traffic.
 Additionally, future research can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.

### 2017-01-09 docs/papers/2016MONE/2016MONE.tex
versao com as correcoes de tanya

@@
 
-These operations function based on the mobile client architecture, which is depicted in Figure \ref{fig:02}.
+Figure \ref{fig:01} describes the mobile client protection both in online and offline mode.
+In online mode, the client has the possibility to connect to the server and the security of the client is enhanced by the server-backed up mechanisms.
+On the other hand, in offline mode the client’s security is supported by the standalone mechanisms.
+Additionally, the mobile client protection is enhanced by the threat intelligence unit providing the constant monitoring and analysis.
+
+Figure \ref{fig:02} depicts the client-side protection mechanisms.
+The client should support 4 subsystems: i) encryption subsystem that contains the procedures of encryption and decryption; ii) storage subsystem that provides the downloaded shares and key storage protection; iii) threat intelligence unit that provides the constant monitoring, and iv) the communication subsystem.
+In short, all security procedures are connected to 4 groups of operations: file request and receiving; encryption and decryption; file and key storing; monitoring and analysis.
 
@@
 	\includegraphics[width=8cm]{fig03.eps}
-	\caption{Proposed Architecture for Offline Mobile Security }
+	\caption{Proposed Architecture for Offline Mobile Security}
 	\label{fig:03}
@@
 
-The infrastructure in Figure \ref{fig:03} includes:
+The Figure \ref{fig:03} describes the proposed architecture for offline mobile security, showing the modules responsible for securing the mobile client, which includes:
 
@@
 Moreover, Model Order Selection (MOS), which is an effective signal processing technique to separate noise components from the principal components, can be applied into anomaly and attack detection \cite{tenorio2013greatest}, to identify and separate malicious behaviors from the legitimate ones.
+The TIM is an internal module of the mobile client that implements offline anomaly detection through signal processing techniques.
 	\item \textbf{Key Management Center}: it includes the functions for maintaining the key expiry period and deleting the expired keys. 
@@
 
-The UAL is responsible for the incremental logging, feature extraction and data modeling for malicious behavior detection, through the Log Analysis Center (LAC).
+The UAL is responsible for the incremental logging of activities of the mobile client, feature extraction and data modeling for malicious behavior detection, through the Log Analysis Center (LAC).
+As an internal module of the mobile client, the UAL implements monitors of selected events of the application, such as a login attempt or a file decryption, and logs the desired information for further analysis.
 The logged information shall be decomposed into selected features and modeled as matrices, composed of the number of occurrences of the selected features by its location and by time.
-The resultant data is submited to the LAC, for anomaly detection.
+The resultant data is submitted to the LAC, for anomaly detection.
 
@@
 The key hierarchy is explained by Figure \ref{fig:08} where the key represented in Figure \ref{fig:07} serves for decrypting the key for each level represented in Figure \ref{fig:08}.
+As it is possible to be observed in Figure \ref{fig:08}, the master key on the upper level is computed by means of SSS (Subsection \ref{sec_secret}).
+It serves for decrypting the ABE key storage (section \ref{sec_abe}).
+The ABE key allows the user finally decrypt the permanent file share keys.
+Further, the file share can be decrypted by means of AES.
 
@@
 Note, that this is different from the usual PK encryption: the private keys are controlled by the admin and not by the users.
-	\item \textbf{Public key (PK)} depends on the master key values and is kept in the clear allowing users to access the information: $PK = (g^{t_1}, g^{t_2}", \cdots ,g^{t_n}, e(g,g)^y)$, Here $e(g,g)$ is the bilinear pairing function corresponding to an elliptic curve.
+	\item \textbf{Public key (PK)} depends on the master key values and is kept in clear allowing users to access the information: $PK = (g^{t_1}, g^{t_2}, \cdots ,g^{t_n}, e(g,g)^y)$.
+Here $e(g,g)$ is the bilinear pairing function corresponding to an elliptic curve.
+In order to explain the properties of bilinear paring function, let $\mathbb{G}_1$ and $\mathbb{G}_2$ be two multiplicative cyclic groups of prime order $p$.
+Let $g$ be a generator of $\mathbb{G}_1$ and $e$ be a bilinear map, $e$: $\mathbb{G}_1 X \mathbb{G}_1 \rightarrow \mathbb{G}_2$.
+The bilinear map $e$ has two properties, the non-degeneracy, which defines that $e(g,g) \neq 1$, and the bilinearity, which defines that for all $u,v \in \mathbb{G}_1$ and $a,b \in \mathbb{Z}_p$, we have $e(u^a, v^b) = e(u,v)^{ab}$ \citep{goyal2006attribute}.
+Here $w \in \mathbb{Z}$ is the unique security parameter allowing to control the key generation for each user group.
 	\item Secret user \textbf{KEY\_SET} depends on his attribute set.
 Here each $D_i$ (\textbf{GROUP\_KEY}) serves for decryption of the data of a single group of users, for example, related to some project: 	$\{t_i\}_U \rightarrow D = \{D_i = g^{\frac{yw}{t_i}}\}$.
-	\item Encrypted text $M$, in our context, $M = FILE\_KEY$, or the permanent AES symmetric key, which allows to avoid the file re-encryption.
-Encryption procedure is multiplication.
-The set of the public keys $E_i$ (PUBLIC\_SHARE\_KEY) corresponding to the set of groups able to access the text is kept along with the encrypted text $E$: $E=Me(g,g)^{ys}, \{E_i = g^{\frac{t_is}{w}}\}, \forall i \in \{t_i\}_M$.
+	\item Encrypted text $M$, in this context, $M = FILE\_KEY$, or the permanent AES symmetric key, which allows to avoid the file re-encryption.
+The encryption procedure is performed by a multiplication.
+The set of the public keys $E_i$ (PUBLIC\_SHARE\_KEY) corresponding to the set of groups able to access the text is kept along with the encrypted text $E$: $E=Me(g,g)^{ys}, \{E_i = g^{\frac{t_is}{w}}\}, \forall i, t_i \in \{t_i\}_U$.
+Here $s \in \mathbb{Z}$ is the unique security parameter allowing to control the key generation for each file share.
 	\item Decryption is division: $M = \frac{E}{Y^s}$. 
@@
 The attribute-based private keys $D_i$ should be protected while being stored in the device memory.
-Therefore, server encrypts the set of $Di$ with a single AES key before sending it to the user device.
+Therefore, server encrypts the set of $D_i$ with a single AES key before sending it to the user device.
 This AES key (master key) is denoted by the value KEY\_SET\_KEY in our notation.
@@
 
-The adversary cannot get any information of the KEY\_SET\_KEY unless he possessing all 4 key parts. Here the values PASS and PIN are predefined similar to the construction in \cite{galibus2015mobile}.
+The adversary cannot get any information of the KEY\_SET\_KEY unless he possesses all 4 key parts. Here the values PASS and PIN are predefined similar to the construction in \cite{galibus2015mobile}.
+
+The PASS and PIN values are predefined, similar to the key storing construction based on the polynomial secret sharing scheme (SSS) in \cite{galibus2015cloud}.
+
+In this SSS, the following notation is used: $S(x) \in F_2[x]$ denotes the secret polynomial with coefficients from the finite field $F_2$ to be computed by pooling the shares together.
+$s_1(x)$, $s_2(x) \in F_2[x]$ denote the shares of the SSS, that are used to compute the secret $s(x)$.
 
@@
 Additionally, the current time value TIME is used in the calculation of the share.
-Let $s(x) = d$ and $s_1(x) = f(PIN + PASS + TIME)$, where $f$ is a one-way function that transforms the data into the string of the desired length:
+Let $s(x) = d$ and $s_1(x) = f(PIN + PASS + TIME)$, where $f$ is a one-way (hash) function that transforms the data into the string of the desired length:
 
@@
 \end{equation}
+
+Here, $p_0$ and $p_1 \in F_2[x]$ denote the public moduli of the polynomial SSS [10, 11]  and $s \in F_2[x]$ denotes the intermediate secret value.
+The formulae for these moduli, secret and participants’ shares directly follow from the definition of polynomial SSS construction.
 
@@
 
-The set of obtained vectors $\mathbf{y}_{m}^{(q)}$ composes the zero mean matrix $\mathbf{Y}^{(q)}$, then the zero mean covariance matrix $\mathbf{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
+The set of obtained vectors $\mathbf{y}_{m}^{(q)}$ composes the zero mean matrix $\mathbf{Y}^{(q)}$, then the zero mean covariance matrix $\mathbf{\hat{R}}_{yy}^{(q)}$ can be estimated as follows
 
 \begin{equation}\label{eq:eq04}
-\mathbf{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\mathbf{Y}^{(q)}\mathbf{Y}^{(q)^{\rm T}}.
+\mathbf{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\mathbf{Y}^{(q)}\mathbf{Y}^{(q)^{\rm T}},
 \end{equation}
+where $\mathbf{\hat{R}}_{yy}^{(q)}$ means the estimation of the sample covariance matrix from the measured zero mean matrix $\mathbf{Y}^{(q)}$ over $N$ minutes of the time frame $q$.
 
@@
 For example, it is not possible to guarantee the security of the decrypted document once the user copies its content to another file.
+
 In the analyzed scenarios, it is assumed that the adversary can:
@@
 	\item Use expired key to decrypt the files, if the attacker is insider;
-	\item Does not have possibility to perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the mobile client;
-	\item Steal the encrypted files but not the decrypted ones.
+	\item Steal the encrypted files.
+\end{enumerate}
+
+It is also assumed that the adversary can not:
+
+\begin{enumerate}
+	\item Perform offline dictionary attack and cryptanalysis, i.e. does not have the possibility to verify user credentials apart from entering them in the mobile client;
+	\item Steal the decrypted files.
 \end{enumerate}
@@
 In offline mode, the user is still allowed to get access to operations that do not require communication with the server side.
-These operations and their selected features shall be incrementally logged by the mobile client, in order to be evaluated to identify malicious behaviors.
-This work proposes to evaluate the following features.
+These operations and their selected features shall be incrementally logged by the UAL of the mobile client, in order to be analyzed by the TIM to identify malicious behaviors.
+
+This work proposes to evaluate the following features, which represents events of the user operating the mobile client.
 
@@
 Thus, the proposal supports the concept of the light-weighted client, i.e.
-the most consuming operations are ABE and AES decryption.
+the most consuming actions are ABE and AES decryption.
 
-Similarly, in the online mode the mobile client does not perform resource-consuming operations apart from J-PAKE construction which is used to renew the user/app master key (DEV\_PASS):
+Similarly, in the online mode the mobile client does not perform resource-consuming actions apart from J-PAKE construction which is used to renew the user/app master key (DEV\_PASS):
 
 \begin{enumerate}
-	\item Generates keys for J-PAKE $\rightarrow$ this can be a resource-consuming operation.
-In the future we suggest to replace J-PAKE with SIS-based PKE \cite{bogos2013primeless,bellovin1992encrypted};
+	\item Generates keys for J-PAKE $\rightarrow$ this can be a resource-consuming action. In the future we suggest to replace J-PAKE with SIS-based PKE \cite{bogos2013primeless,bellovin1992encrypted};
 	\item Sends and receives data in the clear;
-	\item Performs operations 1)-7);
-	\item Performs the operation 8) and sends the log data for server side analysis.
+	\item Performs actions 1)-7);
+	\item Performs the action 8) and sends the log data for server side analysis.
 \end{enumerate}

### 2017-01-09 docs/papers/2016MONE/2016MONE.tex
correcao do nome do prof. rafael

@@
         	João Paulo C. L da Costa		\and
-        	Rafael T. de Souza Júnior	\and
+        	Rafael T. de Sousa Júnior	\and
         Viktor Krasnoproshin			\and
@@
            \and
-           Tatiana Galibus \and Thiago P. B. Vieira \and Robson O. Albuquerque \and João Paulo C. L da Costa \and Rafael T. de Souza Júnior  \at
+           Tatiana Galibus \and Thiago P. B. Vieira \and Robson O. Albuquerque \and João Paulo C. L da Costa \and Rafael T. de Sousa Júnior  \at
               University of Brasilia, UnB, FT, ENE, CP: 4386, 70910-900, Brasília, Brazil

### 2017-01-10 docs/papers/2016MONE/2016MONE.tex
correcoes do prof rafael e adicao de referencias recomendadas

@@
         	Edison P. de Freitas			\and
-        	Robson O. Albuquerque		\and
+        	Robson de O. Albuquerque		\and
         	João Paulo C. L da Costa		\and
@@
            \and
-           Tatiana Galibus \and Thiago P. B. Vieira \and Robson O. Albuquerque \and João Paulo C. L da Costa \and Rafael T. de Sousa Júnior  \at
+           Tatiana Galibus \and Thiago P. de B. Vieira \and Robson de O. Albuquerque \and João Paulo C. L da Costa \and Rafael T. de Sousa Júnior  \at
               University of Brasilia, UnB, FT, ENE, CP: 4386, 70910-900, Brasília, Brazil
@@
            João Paulo C. L da Costa \and Giovanni del Galdo \at
-				Institute for Information Technology, Ilmenau University of Technology, Ilmenau, Germany
-           \and
-           João Paulo C. L da Costa \and Giovanni del Galdo \at
-				Fraunhofer Institute for Integrated Circuits IIS, Erlangen, Germany
+				Institute for Information Technology, Ilmenau University of Technology, Ilmenau, Germany, and Fraunhofer Institute for Integrated Circuits IIS, Erlangen, Germany
            \and
@@
 Therefore, additional control and protection should be taken for the confidential data on the mobile clients that leave the boundaries of the organization.
-In this scenario, such apps can be used on the offline mode for the convenience and traffic reducing or even for a particular characteristic of the mobile client or of the business.
-This paper presents a novel approach to the security of the corporate mobile clients, in particular the offline mode.
-The important part of the presented approach is the conceptualization and the definition of the core methodology to solve the problem of offline mobile security, i.e.
+This paper presents a novel approach to the security of the corporate mobile clients, in particular when they operate in the offline mode.
+The presented approach includes the essential conceptualization and the definition of the core methodology to solve the problem of offline mobile security, i.e.
 the protection of the confidential data in use when the mobile client is not connected to the corporate cloud.
-The protection of the sensitive data is provided by the combination of the cryptographic and analytics methods to detect malicious user behavior.
-The proposed security architecture supports the basic mobile client protection principles: minimized traffic load and reduced communication with the cloud; usage of the light-weighted operations and an optimized combination of the security methods.
+The protection of the sensitive data is provided by the combination of cryptographic means and analytics methods to detect malicious user behavior.
+The proposed security architecture supports the basic mobile client protection principles: minimized traffic load and reduced communication with the cloud; usage of light-weighted operations and an optimized combination of the security methods.
 \end{abstract}
 
-\keywords{ offline mobile cloud security \and offline attribute-based encryption \and model order selection \and user behavior analysis \and secret sharing}
+\keywords{offline mobile cloud security \and offline attribute-based encryption \and model order selection \and user behavior analysis \and secret sharing}
 
@@
 While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, such as CASB and CAC, the corporate mobile client is usually light-weighted and generally less protected.
+
 The protection scheme used on a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}.
@@
 
-From the theoretical point of view of this problem, there are several surveys, whose common point is the mobile cloud as a rapidly developing paradigm that poses many security and complexity problems \cite{yovel2014, khan2015cloud, khan2014survey, khan2013towards}.
-An analysis of the new models of mobile cloud computing and new ways of using data storage services is presented in \cite{khan2015cloud, khan2013towards}.
-Commonly, all the models and protection schemes concentrate on the encryption properties and either perform the computations on their own \cite{zhao2010trusted, yang2011provable} or use the cloud provider to off-load the cryptographic operations \cite{itani2010energy, ren2011lightweight}.
+From the theoretical point of view of this problem, there are several surveys, whose common point is the mobile cloud as a rapidly developing paradigm that poses many security and complexity problems \cite{khan2013towards, yovel2014, khan2014survey, khan2015cloud, mayrhofer2015architecture, chang2015design, xia2015tinman, kulkarni2015addressing, shila2016amcloud, heydari2016efficient}.
+Kulkarni and Khanai \cite{kulkarni2015addressing} discuss the most important threats related to Mobile Cloud Computing and argue that there is a need for a lightweight secure framework that provides security with minimum communication and processing overhead on mobile devices.
+
+An analysis of the new models of mobile cloud computing and new ways of using data storage services is presented in \cite{khan2015cloud, khan2013towards, mayrhofer2015architecture, chang2015design}.
+Commonly, the models and protection schemes concentrate on the encryption properties and either perform the computations on their own \cite{zhao2010trusted, yang2011provable} or use the cloud provider to off-load the cryptographic operations \cite{itani2010energy, ren2011lightweight}.
 Obviously, it is natural the mobile client cannot handle all operations securely without the assistance of a cloud provider, due to resources constraint and battery limitation.
@@
 corporate client continues to provide the secure access to the sensitive data without connection to the network.
-As discussed in \cite{khan2015cloud}, all the currently known schemes of encryption, performing the computation, either use a cloud provider \cite{zhao2010trusted}, a third party trusted agent \cite{yang2011provable} or a combination of both \cite{itani2010energy}.
+As discussed in \cite{khan2015cloud}, all the currently known schemes of encryption, performing the computation, either use a cloud provider \cite{zhao2010trusted}, a third party trusted agent \cite{yang2011provable, heydari2016efficient}, a combination of both \cite{itani2010energy} or ad hoc approaches \cite{shila2016amcloud}.
 In some cases, they concentrate on computational complexity without taking care of user privacy and security \cite{ren2011lightweight}.
-Therefore, according to \cite{khan2015cloud, tenorio2013greatest, khan2013towards}, the state-of-the-art mobile cloud security models do not consider the problem of the offline security mode.
+Therefore, according to \cite{khan2013towards, khan2015cloud}, the state-of-the-art mobile cloud security models do not consider the problem of the offline security mode.
 
@@
 This means that no further communication with the server is needed until the key expires.
+
 The core set of functions and protocols can be divided into three sets of operations as shown in Figure \ref{fig:01}.
@@
 \begin{acknowledgements}
-The authors thank the Brazilian research and innovation agencies FAPDF (Research Support Foundation of the Federal District), FINEP (Agreement RENASIC / PROTO 01.12.0555.00), CAPES and CNPq under the FORTE Project - CAPES Forensic Sciences Announcement 25/2014, under the productivity grant number 303905 / 2014-0, under the PVE project number 88881.030392 / 2013-01 for their financial support on this research, and under the project number 207644/2015-2 for the PDE scholarship in the context of the program CSF - Aerospace Technology.\end{acknowledgements}
+The authors thank the Brazilian research and innovation agencies CAPES (Grant FORTE \\23038.007604/2014-69), FAPDF (Research Support Foundation of the Federal District), FINEP (Grant RENASIC PROTO 01.12.0555.00), CNPq (Productivity grant 303905/2014-0 and PVE Grant 88881.030392/2013-01), as well as the European CSF - Aerospace Technology Program (PDE scholarship within project 207644/2015-2), for their support to this research.\end{acknowledgements}
 

### 2017-02-21 docs/papers/2017EUSIPCO/2017eusipco.tex
data modeling

\documentclass[conference]{IEEEtran}

\ifCLASSINFOpdf
\else
\fi

\hyphenation{op-tical net-works semi-conduc-tor}

\usepackage{graphicx}
\usepackage{array}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{epstopdf}
\usepackage[mathletters]{ucs}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{morefloats}
\usepackage{lipsum}
\usepackage{amsthm,amsmath}

\usepackage{algorithm,algorithmic}% http://ctan.org/pkg/algorithms

\begin{document}

\title{A Distributed Tensor-based Approach for Multidimensional Dictionary Learning}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Thiago P de B Vieira}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Andr ́e L. F. de Almeida}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{Florian Roemer\\ and João Paulo Lustosa}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678--2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

\maketitle

\begin{abstract}
The abstract goes here.
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}
Dictionary Learning is a signal processing technique useful for sparse representation of signals throught the estimation of basis vectors and learning sparse representations of training data. The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and in solving problems such as denoising, compression, classification, data decomposition, feature extraction and image processing.

In some applications the data and its dictionary are multidimensional, e.g., when estimating jointly bahavior of users in social networks or when evaluating correlation between audio and video. Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in data analytics problems. Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}. However, tensor-based algorithms face challenges to achieve reasonable performance to handle large-scale tensor factorizations, demanding efforts in order to explore distributed techniques for tensor-based big data analytics.

	Distributed tensor-based...

	Distributed tensor based dictionary learning...

\begin{figure}[!htb]
     \centering 
	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_1000_100.eps}
     \caption{Matrix of 1.000 elements, 2000 sample training, 100 iterations}
     \label{fig:facets_cross}
\end{figure}

\begin{figure}[!htb]
     \centering 
	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_4000_100.eps}
     \caption{Matrix of 4.000 elements, 2000 sample training, 100 iterations}
     \label{fig:facets_cross}
\end{figure}

\begin{figure}[!htb]
     \centering 
	 \includegraphics[width=0.4\textwidth]{figures/5_20_500_16000_100.eps}
     \caption{Matrix of 16.000 elements, 500 sample training, 100 iterations}
     \label{fig:facets_cross}
\end{figure}

\begin{figure}[!htb]
     \centering 
	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_24000_100.eps}
     \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}
     \label{fig:facets_cross}
\end{figure}

\section{Data Model}

\begin{equation}\label{eq:eq01}
\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W},
\end{equation}

where $\boldsymbol{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\boldsymbol{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary, $\boldsymbol{S} \in \mathbb{C}^{N \times T}$ represents the sparse coefficient matrix and $\boldsymbol{W} \in \mathbb{C}^{M \times T}$ is the additive noise.

\begin{equation}\label{eq:eq02}
\boldsymbol{X} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}) \cdot \boldsymbol{S} + \boldsymbol{W},
\end{equation}

where $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$.

\begin{equation}\label{eq:eq03}
\boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} +  \boldsymbol{\mathcal{W}},
\end{equation}

where $\boldsymbol{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\boldsymbol{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\boldsymbol{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\boldsymbol{X}$, $\boldsymbol{S}$, and $\boldsymbol{W}$ such that $\boldsymbol{X} = [\boldsymbol{\mathcal{X}}]_{(3)}^T$, $\boldsymbol{S} = [\boldsymbol{\mathcal{S}}]_{(3)}^T$, and $\boldsymbol{W} = [\boldsymbol{\mathcal{W}}]_{(3)}^T$, respectively.

We adopt the signal-to-noise ratio (snr) of 20 and the sparseness factor of 5 in order to generate the separable dictionary $\boldsymbol{A}$ according to $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$ of a gaussian distributed data $\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}$.

\begin{equation}\label{eq:eq04}
\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{N}^{(q)} + \boldsymbol{A}^{(q)},
\end{equation}

\section{Proposed Approach}

\section{Experiments}

\section{Conclusion}
The conclusion goes here.

\section*{Acknowledgment}
{\small This research was supported by ...}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2017-02-21 distributed-tensor-dictionary-learning/dtmod.m

@@ -1,4 +1,3 @@
-function A_hat = dtmod(noIt, X, A_hat1, A_hat2, solver, varargin)
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     %%% distributed tensor-based Method of Optimized Directions (MOD)   %%%
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@ -34,11 +33,31 @@ function A_hat = dtmod(noIt, X, A_hat1, A_hat2, solver, varargin)
     %    solver     = selected solver
     %    varargin	= additional parameters
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    clc;
+    clf;
+    clear all;
+
+    s = 5;
+    snr = 20;
+    L = 500;
+    nofTrials = 1;
+    noIt = 5;
+    I = 8;
+    J = 6;
+    K = 2;
     
-    %% data initialization
-    [M1,N1] = size(A_hat1);
-    [M2,N2] = size(A_hat2);
+    M1 = 4;
+    M2 = 2;
+    N1 = 4;
+    N2 = 3;
+    
+    % data generation
+    [A, A1, A2] = makedict2(M1, M2, N1, N2, 'G');
+    X = datamake(A, L, s, snr, 'G');
+    A_hat = dictnormalize( X(:, floor(0.85 * L - K) + (1:K)) );         % Normalize and arrange the vectors for an initial estimated dictionary        
+    [A_hat1, A_hat2] = krondecomp(A_hat, M1, M2, N1, N2);               % Make the data separable decomposing the approximation of A_hat and generating new A_hat*
     T = size(X,2);
+    
     Xt = permute(reshape(X.',[T,M2,M1]),[3,2,1]);    
     Xt1 = reshape(permute(Xt,[1,3,2]),M1,[]);
     Xt2 = reshape(permute(Xt,[2,1,3]),M2,[]);
@@ -46,7 +65,9 @@ function A_hat = dtmod(noIt, X, A_hat1, A_hat2, solver, varargin)
     
     %% minimize
     for it = 1:noIt
[…]

### 2017-02-21 distributed-tensor-dictionary-learning/execDL.m

@@ -135,7 +135,7 @@ function results = execDL(L, N, K, M1, M2, N1, N2, snr, methodChar, s, noIt, nof
         elseif (strcmpi(methodChar,'B'))                                    % MiniBatch
             mb = [1,25; 1,50; 1,125; 1,300];                                % building block in minibatch
             v2p = sum( mb(:,1).*mb(:,2) );                                  % vectors to process (500)
-            mb = repmat([ceil((L*noIt)/(v2p)),1],4,1).*mb;                  % we want: v2p >= (L*noIt)
+            mb = repmat([ceil((L * noIt)/(v2p)),1],4,1).*mb;                  % we want: v2p >= (L*noIt)
             MBopt = struct('K',K, ...
                        'samet','mexomp', ...
                        'saopt',struct('tnz',s, 'verbose',0), ...

### 2017-02-21 distributed-tensor-dictionary-learning/project.m

@@ -20,8 +20,8 @@ scriptName = 'dictionary_learning';
 %% parameters
 s = 5;           % sparseness
 snr = 20;        % snr for added noise
-L = 2000;        % number of training vectors to use
-nofTrials = 3;  % enought trials to obtain reliable results
+L = 500;         % number of training vectors to use
+nofTrials = 3;   % enought trials to obtain reliable results
 noIt = 100;      % number of iterations in each trial
 N = 80;
 K = 200;

### 2017-07-12 docs/papers/2016JNCA/submition/response02/jnca-response_to_reviewers.txt
returning dictionary and sparse coding. implementation of method evaluation saving into files

Response to the Handling Editor and Reviewers

Submission No.: JNCA-D-16-00617R1
Title: Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks
Journal of Network and Computer Applications (JNCA)
Research Paper

Dear Mohammed Atiquzzaman, Ph.D., Editor-in-Chief, and reviewers,
The authors would like to thank the Associate Editor-in-Chief and the Reviewers for volunteering their time in reviewing our manuscript and providing us with valuable comments which allowed us to significantly improve the paper. We have carefully revised the manuscript based on the reviewer’s comments and suggestions. All changes in the paper have been marked with a color. The following is our point-by-point responses to the raised comments.

Kind regards,

The authors.

1. EDITOR
1.1. COMMENT: The referees have suggested that certain aspects of the paper should be re-examined, before it can be considered further for publication.
1.1.1. REPLY: We thank the Editor for the time, work, and positive evaluation of our paper.

2. REVIEWER #3
2.1. COMMENT: The reviewer appreciates the efforts that the authors have made to revise this paper. The revised version has improved a lot compared to the previous version. Most of the questions raised from the reviewers have been clarified. The reviewer has three additional comments regarding the revised version.
2.1.1. REPLY: Obrigado pela revisão positiva de maneira geral. Esperamos poder responder os comentários adicionais da maneira esperada.
2.2. COMMENT: The mentioned dataset used for additional evaluation is almost twenty years ago, and it might be able to reflect the current state of arts of modern attacks.
2.2.1. REPLY: Entendemos esta limitação e até inclimos comentários sobre isto durante a revisão da literatura. Mesmo este dataset sendo velho e podendo nao refletir ataques modernos, este dataset ainda vem sendo utilizado por lacunas em datasets novos e públicos, conforme surveys recentes.
2.3 COMMENT: The authors have provided complexity and running time evaluation in the revision. However, given that the EVD complexity in O(M^2*N+N^3), it looks like the proposed approach might be DoS'ed if an attacker attempts to flood using randomized port numbers, e.g., (M=) 64K network ports. This is because performing bandwidth consumption flooding attacks does not need to work with a dedicated (or a monitored) port number.
2.3.1. REPLY: A proposta é a aplicação da técnica em portas selecionadas de acordo com seu interese e seu uso em uma eventual rede interna. Nossos cenários consideraram ataques provinientes de de uma rede interna e também de uma rede externa para uma rede interna. "performing bandwidth consumption flooding attacks" não são algo da nossa análise, mas nossa técnica foi capaz de detectar flood atack em casos que utilizam diveras portas , utilizando análise de correlação, como nos casos avaliados para DARPA 98 nas seções X e Y
2.4 COMMENT: The authors have revised the paper to focus exactly on probe and scanning attacks. This may greatly limited the contribution of the proposed work, as we already have a number of commercial solutions to deal with the problem. 
2.4.1. REPLY: Nosso trabalho foca em flooding, probe and scannnig, buscando desenvolver uma nova tecnica sem dependência de conhecimento prévio, análise de comportamentos conhecidos, dicionários, e etc. Nosso objetivo é utilizar uma téction unsupervisioned e blind

3. REVIEWER #4
3.1. COMMENT: The authors have satisfactorily addressed the queries raised by the reviewers. However, only DARPA 1998 dataset has been considered to conduct the experiments. The authors can also conduct the experiments on other publicly available datasets such as CAIDA (https://www.caida.org/data/) and UNB-ISCX (http://www.unb.ca/research/iscx/dataset/iscx-IDS-dataset.html). Such datasets contain more recent attack vectors and types. Conducting experiments on such datasets and discussion of analysis on such datasets can provide in depth understanding of the proposed work.
3.1.1. REPLY: Obrigado pela construtuva comentário... Estes datasets foram considerados e citados em X e Y, entretanto os ataques monitorados por estes casos não consideram probe attack e flooding em portas tcp/udp, como também são baseados em fluxo enquanto nossa técnica é baseada em pacotes. Por fim, mas nao menos importante, estes datasets não dispõem de uma classificação prévia (label) dos ataques. Como UNB-ISCX

4 REVIEWER #5
4.1. COMMENT: Since only one user (computer) was used, the amount of traffic noise (DHCP) would be very limited.
4.1.1. REPLY: Como forma de adicionar cenários mais abrangentes e realistas e bem conhecidos, adicionamos a análise da nossa tecnica aplicada ao DARPA 98
4.2. COMMENT: There is no description of the systems used, because attacks such as the fraggle attack are easily eliminated by both systems and routers. Plus, the network is very limited with the number of devices used and type of traffic observed.
4.3.1. REPLY: Para o fraggle utilizamos a ferramenta X, uma descrição sobre isto foi adicionada no em Z
4.4. COMMENT: The authors stated """The UDP scan technique sends UDP packets to the destination port, and if it responds with a ICMP port 230 unreachable message, then it indicates that the scanned port is closed. On the other hand, if a message is not received, then the port is considered as open."""* lots of systems today disable this by default, that is why it is unclear which systems were used for testing.
4.4.1. REPLY: Esta descrição se refere a uma forma de portscan avaliada em nosso cenário sintético, mas os cenários de portscan avaliados no dataset darpa incluem casos em que não há resposta para o envio de pacotes
4.5. COMMENT: What if the port scan was not sent in a sequence fashion and sent with time frames higher than the samples selected for evaluation and checking? The information considered in the proposed framework will not help.
4.5.1. REPLY: Casos como este foram avaliados e não foram identificados, conforme descritos em XX
4.6. COMMENT: Why didn't the authors use flow analysis or even compare to such related work. Flow analysis has been used a lot for such detection, especially when you're looking at time and ports.
4.6.1. REPLY: Flow analysis implica em custo adicional para a identificação e montagem de fluxos para uma eventual análise. Entretanto, iremos considerar uma análise baseada em fluxo como trabalhos futuros.
4.7. COMMENT: How can such module detect new novel attacks, when the only experiments tested were SYN(flood), fraggle (rare today), and two port scanning techniques? ex: traffic like DHCP are considered noise, so how would this framework detect an ARP Scan? O framework proposto foca em flood e probe attacks na camada de transporte, entretanto a nossa técnica pode highlight ataques inesperados que tenham comportamentos semelhantes ao que selecionados, sendo então possível selecionar executar análises mais detalhadas nestes casos específicos. Dado que ARP scan foca em uma camada abaixo da camada de transporte, este não seria detectado por nossa proposta
4.7.1. REPLY: 
4.8. COMMENT: The Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack shows that port 80, 443, and 53 were not an attack, even though all of these traffic are currently used for exfiltration or covert communication. So, how will the proposed framework detect such attacks? Detecting such attacks could be considered novel if they could achieve such detection.
4.8.1. REPLY: No cenário considerado o tráfego nas portas 80, 443, and 53 foi modelado como legítimo, representando tráfego de usuários aos serviços HTTP, HTTPS e DNS para demonstrar a capacidade de detecção de ataques mixed into legitimate traffic. Nosso framework não é capaz de detectar exfiltration, covert communication ou exploiting, como demonstrado ao analisarmos nossa solução para o dataset DARPA. Podemos considerar iste ponto como um foco para nosso trabalho de doutorado, com uma análise de correlação para identificar casos de exfiltration or covert communication

### 2017-07-12 docs/papers/2017EUSIPCO/2017eusipco.tex
returning dictionary and sparse coding. implementation of method evaluation saving into files

@@
 
+%% packages
+% graph
 \usepackage{graphicx}
+\usepackage{epstopdf}
 \usepackage{array}
@@
 \usepackage{tabularx}
-\usepackage{epstopdf}
-\usepackage[mathletters]{ucs}
+% math
+\usepackage{amsfonts}
 \usepackage{mathtools}
-\usepackage{amsfonts}
-\usepackage{morefloats}
-\usepackage{lipsum}
-\usepackage{amsthm,amsmath}
-
+% algorithm
 \usepackage{algorithm,algorithmic}% http://ctan.org/pkg/algorithms
@@
 
-\title{A Distributed Tensor-based Approach for Multidimensional Dictionary Learning}
+\title{A Distributed and Tensor-based Approach for Multidimensional Dictionary Learning}
 
@@
 \author{\IEEEauthorblockN{Thiago P de B Vieira}
-\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
-Georgia Institute of Technology\\
-Atlanta, Georgia 30332--0250\\
-Email: http://www.michaelshell.org/contact.html}
+\IEEEauthorblockA{
+	Department of Electrical Engineering\\
+	University of Brasilia\\
+	Bras\'ilia-DF, Brazil\\
+	Emai	l: tpbvieira@gmail.com
+}
+
 \and
-\IEEEauthorblockN{Andr ́e L. F. de Almeida}
-\IEEEauthorblockA{Twentieth Century Fox\\
-Springfield, USA\\
-Email: homer@thesimpsons.com}
+
+\IEEEauthorblockN{Andr\'e L. F. de Almeida}
+\IEEEauthorblockA{
+	Department of Teleinformatics Engineering\\
+	Federal University of Cear\'a\\
+	Fortaleza, Brazil\\
+	Emai	l: andre.almeida@ieee.org
+}
+
 \and
-\IEEEauthorblockN{Florian Roemer\\ and João Paulo Lustosa}
-\IEEEauthorblockA{Starfleet Academy\\
-San Francisco, California 96678--2391\\
-Telephone: (800) 555--1212\\
-Fax: (888) 555--1212}}
+
+\IEEEauthorblockN{Florian Roemer\\ and Jo\~ao Paulo C. L. da Costa}
+\IEEEauthorblockA{
+	Institute for Information Technology\\
+	Ilmenau University of Technology\\
+	Ilmenau, Germany
+}
+
+}
 
@@
 \section{Introduction}
-Dictionary Learning is a signal processing technique useful for sparse representation of signals throught the estimation of basis vectors and learning sparse representations of training data.
-The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and in solving problems such as denoising, compression, classification, data decomposition, feature extraction and image processing.
 
-In some applications the data and its dictionary are multidimensional, e.g., when estimating jointly bahavior of users in social networks or when evaluating correlation between audio and video.
-Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in data analytics problems.
+Dictionary Learning is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
+The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, classification, data decomposition, feature extraction and image processing \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.
+
+In some applications the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks.
+Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in multidimensional data analytics problems \cite{kolda2009tensor}.
 Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}.
-However, tensor-based algorithms face challenges to achieve reasonable performance to handle large-scale tensor factorizations, demanding efforts in order to explore distributed techniques for tensor-based big data analytics.
 
-	Distributed tensor-based...
+Multidimensional parameter estimation and learning multidimensional separable dictionaries are growing research problems.
+Roemer \emph{et al.} \cite{roemer2014tensor} show that the multidimensional dictionary estimation problem can be efficiently formulated in terms of tensors, and that their results outperform existing schemes by exploiting the multilinear structure of the problem.
 
-	Distributed tensor based dictionary learning...
+Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results.
+However, the performance of tensor-based algorithms outperform existing schemes when dealing with growing multidimensional datasets, as can be seen in Figures \ref{fig:fig1} and \ref{fig:fig2}, which show that the detection rating of traditional algorithms decreases over the dataset increasing.
 
@@
      \caption{Matrix of 1.000 elements, 2000 sample training, 100 iterations}
-     \label{fig:facets_cross}
+     \label{fig:fig1}
 \end{figure}
+
+\begin{figure}[!htb]
+     \centering 
+	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_24000_100.eps}
+     \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}
+     \label{fig:fig2}
+\end{figure}
+
+Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based and traditional dictionary learning algorithms for dictionary reconstruction of a multidimensional and separable data.
+Figure \ref{fig:fig1} presents the results for dictionary reconstruction from a matrix of 1000 elements, while Figures \ref{fig:fig2} presents the results from a matrix with 24000 entries.
+It is possible to observe better detection rating for tensor-based algorithms.
+However, tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.
+
+We propose a distributed and tensor-based approach for multidimensional dictionary learning in order to obtain better processing time for larger datasets.
+We focus on a distributed implementation of T-MOD \cite{roemer2014tensor} algorithm, based on a modification of Almeida and Kibangou's \cite{de2014distributed} approach to use Tucker-2 decomposition.
+We perform experiments to test different methods for learning dictionaries, evaluating the processing time and the reconstruction of a dictionary of multidimensional and separable data.
+
+\section{Data Model}
+
+Consider a generic sparse recovery problem of the following form
+
+\begin{equation}\label{eq:eq01}
+\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W},
+\end{equation}
+
+where $\boldsymbol{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\boldsymbol{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary, $\boldsymbol{S} \in \mathbb{•}hbb{C}^{N \times T}$ represents the sparse coefficient matrix, $\boldsymbol{W} \in \mathbb{C}^{M \times T}$ is the additive noise, and $M < N < T$.
+
+Consider a sparse recovery problem for a separable 2-D manifold, i.e., a manifold that can be written as the product of two 1-D manifolds.
+If we choose a separable 2-D sampling grid, we can write the 2D sparse recovery problem as
+
+\begin{equation}\label{eq:eq02}
+\boldsymbol{X} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}) \cdot \boldsymbol{S} + \boldsymbol{W},
+\end{equation}
+
+where $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$, $\boldsymbol{A}^{(1)} \in \mathbb{C}^{M_1 \times N_1}$, $\boldsymbol{A}^{(2)} \in \mathbb{C}^{M_2 \times N_2}$, $M = M_1 \times M_2$, and $N = N_1 \times N_2$.
+
+The Kronecker model (\ref{eq:eq02}) can be rewritten in an equivalent tensor form.
+Applying the algebraic rules for unfoldings of n-mode products \cite{roemer2014tensor} we can rewrite (\ref{eq:eq02}) into a "Tucker-2" decomposition, as
+
+\begin{equation}\label{eq:eq03}
+\boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} +  \boldsymbol{\mathcal{W}},
+\end{equation}
+
+where $\boldsymbol{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\boldsymbol{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\boldsymbol{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\boldsymbol{X}$, $\boldsymbol{S}$, and $\boldsymbol{W}$ such that $\boldsymbol{X} = [\boldsymbol{\mathcal{X}}]_{(3)}^T$, $\boldsymbol{S} = [\boldsymbol{\mathcal{S}}]_{(3)}^T$, and $\boldsymbol{W} = [\boldsymbol{\mathcal{W}}]_{(3)}^T$, respectively.
+
+We consider a randomly generated data in order to evaluate the proposed approach, where initially $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\boldsymbol{A}$ is obtained.
+In order to be able to evaluate the dictionary reconstruction, the sparse data $\boldsymbol{\mathcal{X}}$ is generated from the dictionary $\boldsymbol{A}$, which is used for further comparisons.
+We adopt 20 as the signal-to-noise ratio (snr) and 5 as the sparseness factor of the data $\boldsymbol{\mathcal{X}}$.
+
+The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\boldsymbol{A}$ from the sparse data $\boldsymbol{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.
+
+\section{Proposed Approach}
+
+...
+
+\section{Experiments}
+
+...
 
@@
 
-\begin{figure}[!htb]
-     \centering 
-	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_24000_100.eps}
-     \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}
-     \label{fig:facets_cross}
-\end{figure}
-
-\section{Data Model}
-
-\begin{equation}\label{eq:eq01}
-\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W},
-\end{equation}
-
-where $\boldsymbol{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\boldsymbol{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary, $\boldsymbol{S} \in \mathbb{C}^{N \times T}$ represents the sparse coefficient matrix and $\boldsymbol{W} \in \mathbb{C}^{M \times T}$ is the additive noise.
-
-\begin{equation}\label{eq:eq02}
-\boldsymbol{X} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}) \cdot \boldsymbol{S} + \boldsymbol{W},
-\end{equation}
-
-where $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$.
-
-\begin{equation}\label{eq:eq03}
-\boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} +  \boldsymbol{\mathcal{W}},
-\end{equation}
-
-where $\boldsymbol{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\boldsymbol{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\boldsymbol{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\boldsymbol{X}$, $\boldsymbol{S}$, and $\boldsymbol{W}$ such that $\boldsymbol{X} = [\boldsymbol{\mathcal{X}}]_{(3)}^T$, $\boldsymbol{S} = [\boldsymbol{\mathcal{S}}]_{(3)}^T$, and $\boldsymbol{W} = [\boldsymbol{\mathcal{W}}]_{(3)}^T$, respectively.
-
-We adopt the signal-to-noise ratio (snr) of 20 and the sparseness factor of 5 in order to generate the separable dictionary $\boldsymbol{A}$ according to $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$ of a gaussian distributed data $\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}$.
-
-\begin{equation}\label{eq:eq04}
-\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{N}^{(q)} + \boldsymbol{A}^{(q)},
-\end{equation}
-
-\section{Proposed Approach}
-
-\section{Experiments}
-
 \section{Conclusion}

### 2017-07-12 distributed-tensor-dictionary-learning/execDL.m

@@ -11,7 +11,6 @@ function results = execDL(L, N, K, M1, M2, N1, N2, snr, methodChar, s, noIt, nof
     % This code is a version of Karl Skretting work.
     %-------------------------------------------------------------------------
     % parameters:
-    %   res         = a struct which is also stored in 'ex210xsyynn.mat'
     %   L           = number of training vectors to use
     %   N           = number of lines of the file where the set of training vectors, X, is stored during design
     %   K           = number of columns of the file where the set of training vectors, X, is stored during design
@@ -29,6 +28,8 @@ function results = execDL(L, N, K, M1, M2, N1, N2, snr, methodChar, s, noIt, nof
     %   s           = sparseness, number of non-zero coefficients, default 5 
     %   noIt        = number of iterations to do for each trial, default 200
     %   nofTrials   = number of trials to do, default 1
+	%
+    %   res         = a struct which is also stored in 'ex210xsyynn.mat'
     %-------------------------------------------------------------------------    
     % Exemples:
     % res = ex210(L, snr, methodChar, s, noIt, nofTrials, makeFig);

### 2017-07-12 distributed-tensor-dictionary-learning/khosvd.m

@@ -1,4 +1,4 @@
-function A_hat = khosvd2(noIt, X, A_hat, M1, M2, solver, varargin)
+function [A_hat,S_hat] = khosvd(noIt, X, A_hat, M1, M2, solver, varargin)
     % What I suggest is to start from the K-SVD algorithm you have and only 
     % replace the matrix rank-one approximation (SVD-based) by the tensor-based 
     % rank-one approximation (HOSVD-based). If the data is multidimensional, 

### 2017-07-12 distributed-tensor-dictionary-learning/ksvd.m

@@ -1,4 +1,4 @@
-function A_hat = ksvd(noIt, X, A_hat, solver, varargin)
+function [A_hat, S_hat] = ksvd(noIt, X, A_hat, solver, varargin)
     % ...
     %
     % This code is a version of Karl Skretting work.
@@ -25,5 +25,4 @@ function A_hat = ksvd(noIt, X, A_hat, solver, varargin)
                 R(:,I) = Ri - A_hat(:,k) * S_hat(k,I);
             end
         end
-    end    
-    return
+    end

### 2017-07-12 distributed-tensor-dictionary-learning/makedict2.m

@@ -26,8 +26,8 @@ function [A, A1, A2] = makedict2(M1, M2, N1, N2, met)
 
     % make the generating frame/dictionary
     if met(1)=='U';
-        A1 = 2 * rand(M1,N1) -1;
-        A2 = 2 * rand(M2,N2) -1;
+        A1 = 2 * rand(M1,N1) - 1;
+        A2 = 2 * rand(M2,N2) - 1;
     elseif met(1)=='u';
         A1 = rand(M1,N1);
         A2 = rand(M2,N2);

### 2017-07-12 distributed-tensor-dictionary-learning/my_matlab_path.m

@@ -84,7 +84,7 @@ else     % her for PC, Dropbox may be mounted different places
         p = [p0,lower(name)];   
     elseif (strcmpi(name,'java') || strcmpi(name,'javaclasses'))
         p = [p0,'javaclasses'];   
-    elseif strcmpi(name,'maxflow') 
+    elseif strcmpi(name,'maxflow')
         p = [p0,'maxflow-v3.01'];   
     elseif strcmpi(name,'spams') 
         p = [p0,'spams-matlab'];           

### 2017-07-12 distributed-tensor-dictionary-learning/project.m

@@ -34,15 +34,15 @@ N2 = 10;
 fileNameInfo = sprintf('%1i_%li_%li_%li_%li',s,snr,L,N*K,noIt);
 fileNameSufix = sprintf('%s.mat',fileNameInfo);
 dataFiles = [
-             ['L', fileNameSufix] % 'L', 'Q', 'C', 'H' or 'E' = RLS-DLA (java),
-             ['T', fileNameSufix] % 'T' = K-HOSVD
-             ['O', fileNameSufix] % 'O' = T-MOD,             
-             ['K', fileNameSufix] % 'K' = K-SVD,   
-             ['D', fileNameSufix] % 'D' = MOD,
-             %['A', fileNameSufix] % 'A' = AK-SVD,
-             %['M', fileNameSufix] % 'M' = ILS-DLA MOD,             
-             %['I', fileNameSufix] % 'I' = ILS-DLA MOD (java),             
-             %['B', fileNameSufix] % 'B' = RLS-DLA miniBatch             
+             ['L', fileNameSufix] % 'L', 'Q', 'C', 'H' or 'E' =             RLS-DLA (java),
+             ['T', fileNameSufix] % 'T' =                                   K-HOSVD
+             ['O', fileNameSufix] % 'O' =                                   T-MOD,             
+             ['K', fileNameSufix] % 'K' =                                   K-SVD,   
+             ['D', fileNameSufix] % 'D' =                                   MOD,
+             %['A', fileNameSufix] % 'A' =                                  AK-SVD,
+             %['M', fileNameSufix] % 'M' =                                  ILS-DLA MOD,             
+             %['I', fileNameSufix] % 'I' =                                  ILS-DLA MOD (java),             
+             %['B', fileNameSufix] % 'B' =                                  RLS-DLA miniBatch             
              ];
 numMethods = size(dataFiles,1);
 methodNames = cell(numMethods,1);
@@ -63,7 +63,7 @@ for i=1:numMethods;
     if exist(dataFiles(i,:),'file')
         fileName = dir(dataFiles(i,:));
         disp([scriptName,': results of ',dataFiles(i,:),' at ',fileName.date,'.']);
-        load(dataFiles(i,:));                                               % try load privious results
+        load(dataFiles(i,:));                                               % try load previous results
         trialsDone = results.nofTrials;
         if results.noIt ~= noIt
             disp(['WARN : Number of iterations in file is ',int2str(results.noIt),...

### 2017-07-12 distributed-tensor-dictionary-learning/projectImageDenoising.m

@@ -1,0 +1,307 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Project to evaluate Dictionary Learning methods for image       %%%
+%%% denoising                                                       %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% Author: Thiago Vieira, UnB, 2017                                %%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% General goal: ...
+%
+% For details, please see:
+% TENSOR-BASED ALGORITHMS FOR LEARNING MULTIDIMENSIONAL SEPARABLE 
+% DICTIONARIES
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+clc; clf;
+
+javaclasspath('-dynamic')                                                   % java configuration
+filePath = '/media/thiago/ubuntu/datasets/imageDenoising/dl/';              % mudar para o local onde deseja guardar os arquivos
+L = 94500;                                                                  % observações utilizadas no aprendizado
+K = 100; N1 = 20; N2 = 5;                                                   % atoms do dicionário
+N = 49; M1 = 7; M2 = 7;                                                     % features/variáveis/componentes
+noIt = 1;                                                                   % iterações para aprendizado
+solver = 'javaORMP';                                                        % mais rápido que OMP
+sparsity = 2;                                                               % grau de espasidade ( quantidade desejada de valores não-zedo por coluna )
[…]

### 2017-07-12 distributed-tensor-dictionary-learning/tmod.m

@@ -1,4 +1,4 @@
-function A_hat = tmod(noIt, X, A_hat1, A_hat2, solver, varargin)
+function [A_hat,S_hat] = tmod(noIt, X, A_hat1, A_hat2, solver, varargin)
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     %%% tensor-based Method of Optimized Directions (MOD)                   %%%
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@ -37,4 +37,3 @@ function A_hat = tmod(noIt, X, A_hat1, A_hat2, solver, varargin)
         A_hat2 = dictnormalize(A_hat2);
     end
     A_hat = kron(A_hat1,A_hat2);
-    return;

### 2017-09-14 docs/thesis/thesis.tex
thesis baseline

\documentclass[a4paper,12pt,openright,titlepage,oneside]{book}

%\usepackage[english,brazil]{babel} 
%Define regra de gramática para separar síbalas {babel}
%e altera os títulos (como chapters, sections, references) para português
\usepackage[brazil]{babel}

% Define uso de caracteres acentuados no PDF gerado
% e permite copiar corretamente o texto do PDF.
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% Adição do pacote do template da UnB
\usepackage{template-FT-UnB/ft2unb}

\DeclareGraphicsExtensions{.jpg,.pdf,.mps,.png,.gif, .eps}
\graphicspath{imagens} %define diretório de imagens

%Arquivo com lista com hifenização correta de algumas palavras.
%Defina novas palavras no arquivo a medida que verificar que a hifenização automática
%etá errada para tais palavras.
\input{hifenizacao}

% ALTERE OS VALORES DENTRO DAS CHAVES DOS COMANDOS NESTA SEÇÃO PARA INCLUIR OS SEUS DADOS E DADOS DA SUA
% DISSERTAÇÃO DE MESTRADO OU TESE DE DOUTORADO
% -----------------------------------------------------------------------------------------------------

%\onehalfspacing
\title{DISCRIMINATIVE SENSING BASED ON SIGNAL PROCESSING}
\author{THIAGO PEREIRA DE BRITO VIEIRA}
\date{2017-06-30} %data da defesa

\grau{Ph.D} %Mestre ou Doutor
\area{Eletrical Engennering} %Nome do curso
\siglaarea{ENE} %Sigla do departamento
\tipodemonografia{Thesis} %Dissertação ou Tese
\programa{Mestrado} %Mestrado ou Doutorado
\autorendereco{THIAGO PEREIRA DE BRITO VIEIRA.} %Endereço do autor da dissertação/tese
\totalpgs{147} %total de páginas atualmente na sua dissertação
\dia{16} %dia da defesa
\mes{Junho} %mês da defesa
\ano{2011} %ano da defesa
\numpublicacao{xxx/AAAA} %número da publicação (após a defesa, tal número deve ser obtido na secretaria)

%PPGENE.DM  = Programa de Pós Graduação em ENgenharia Elétrica.Dissertação de Mestrado
%PPGENE.TD  = Programa de Pós Graduação em ENgenharia Elétrica.Tese de Doutorado
\siglapublicacao{PPGENE.DM}

\titulolinhai{DISCRIMINATIVE SENSING}
\titulolinhaii{BASED ON SIGNAL PROCESSING}
\titulolinhaiv{}

\autori{SEU NOME COMPLETO AQUI}
%Caso seu nome não caiba em uma única linha, divida ele nos comandos abaixo
%\autorii{} 
%\autoriii{}

\membrodabancai{Prof. Dr. NOME DO SEU ORIENTADOR, ENE/UnB}
\membrodabancaifuncao{Orientador}
\membrodabancaii{Prof. Fulano de Tal 2, ENE/UnB}
\membrodabancaiifuncao{Examinador interno}
\membrodabancaiii{Prof. Fulano de Tal 3, ENE/UnB}
\membrodabancaiiifuncao{Examinador interno}
\membrodabancaiv{Prof. Fulano de Tal 4, EESC/USP}
\membrodabancaivfuncao{Examinador externo}
\membrodabancav{}
\membrodabancavfuncao{}
% -----------------------------------------------------------------------------------------------------

%line-numbers, inputencoding=utf8/latin1
%Define o estilo para listagens de código fonte
\lstset{
  numbers=left, %numeração de linhas à esquerda
  stepnumber=1,
  firstnumber=1,
  numberstyle=\tiny,
  extendedchars=true,
  frame=none,
  basicstyle=\footnotesize,
  stringstyle=\ttfamily,
  showstringspaces=false,
  %language=Java, %deve ser definida na inclusão de cada trecho de código, pois podem existir linguagens diferentes em exemplos diferentes
  breaklines=true,
  breakautoindent=true,
  %estilos de comentário de uma e várias linhas
  morecomment=[l]{--}, morecomment=[s]{/*}{*/}, morecomment=[s]{<!--}{-->}, morecomment=[s]{--[[}{--]]}
}

% Adição de metadados no PDF (propriedades do documento PDF)
\makeatletter
	 \hypersetup{
		 pdftoolbar=true,        % show Acrobat’s toolbar?
		 pdfmenubar=true,        % show Acrobat’s menu?
		 pdffitwindow=false,     % window fit to page when opened
		 pdfstartview={FitH},    % fits the width of the page to the window	 
		 pdftitle={\@title},
		 pdfauthor={\@author},
		 pdfsubject={\tipodemonografianome \ de\ \programastr \ em\ \areastr},   % subject of the document
		 pdfcreationdate={\pdfdate}
	 }
\makeatother

\makeindex
\makenomenclature %Necessário para gerar lista de siglas

\begin{document}

	\pdfbookmark[0]{Agradecimentos}{agradecimentos}
	\include{agradecimentos}
	\include{resumo}

	\pdfbookmark[0]{Sumário}{sumario}
	\sumario

	\pdfbookmark[0]{Lista de Figuras}{listafiguras}
	\listadefiguras

	\pdfbookmark[0]{Lista de Tabelas}{listatabelas}
	\listadetabelas

	\pdfbookmark[0]{Lista de Códigos Fonte}{listacodigosfonte}
	\listadecodigosfonte

	\renewcommand{\nomname}{LISTA DE TERMOS E SIGLAS} %Define um caption à lista de siglas
	%Inclui a lista de siglas 
	\pdfbookmark[0]{Lista de Termos e Siglas}{nomenclatura}
	\printnomenclature[2.5cm]

	\mainmatter %Inicia a numeracao normal cardinal
	\setcounter{page}{1} \pagenumbering{arabic} \pagestyle{plain}

	\include{introducao}
	\include{capitulo1}
	\include{capitulo2} 
	\include{conclusao}

	%\bibliographystyle{abnt-num} % estilo bibliográfico ABNT numérico
	%\bibliographystyle{abnt-alf} % estilo bibliográfico ABNT alfabético
	\bibliographystyle{sbc}  % estilo bibliográfico da Sociedade Brasileira de Computação (SBC)

	%\renewcommand{\bibname}{REFERÊNCIAS BIBLIOGRÁFICAS} %Define o Caption da seção de bibliografia
	%\addcontentsline{toc}{chapter}{REFERÊNCIAS BIBLIOGRÁFICAS}

	%não pode ter espaço entre os nomes dos arquivos bib
	\bibliography{referencias}

	\include{apendice}
\end{document}

### 2017-09-19 docs/thesis/abstract.tex
compiling version

Distributed systems has been adopted for building modern Internet services and cloud computing infrastructures, in order to obtain services with high performance, scalability, and reliability. Cloud computing SLAs require low time to identify, diagnose and solve problems in a cloud computing production infrastructure, in order to avoid negative impacts into the quality of service provided for its clients. Thus, the detection of error causes, diagnose and reproduction of errors are challenges that motivate efforts to the development of less intrusive mechanisms for monitoring and debugging distributed applications at runtime.

Network traffic analysis is one option to the distributed systems measurement, although there are limitations on capacity to process large amounts of network traffic in short time, and on scalability to process network traffic where there is variation of resource demand.

The goal of this dissertation is to analyse the processing capacity problem for measuring distributed systems through network traffic analysis, in order to evaluate the performance of distributed systems at a data center, using commodity hardware and cloud computing services, in a minimally intrusive way.

We propose a new approach based on MapReduce, for deep inspection of distributed application traffic, in order to evaluate the performance of distributed systems at runtime, using commodity hardware. In this dissertation we evaluated the effectiveness of MapReduce for a deep packet inspection algorithm, its processing capacity, completion time speedup, processing capacity scalability, and the behavior followed by MapReduce phases, when applied to deep packet inspection for extracting indicators of distributed applications.

\begin{keywords}
Distributed Application Measurement, Profiling, MapReduce, Network Traffic Analysis, Packet Level Analysis, Deep Packet Inspection
\end{keywords}

### 2017-09-19 docs/thesis/chapters/1_introduction/introduction.tex
compiling version

% Introdução: cenario, motivacao, problema, solucoes atuais, escopo do artigo, lista de contribuicoes, organizacao do resto do artigo. 
% Chapter Structure
% 	Motivation
% 	Problems Statement
%	Contributions
% 	Dissertation Organization
\chapter{Introduction}
\label{ch:introduction}

\begin{quotation}[]{Chico Xavier}
Though nobody can go back and make a new beginning, anyone can start over and make a new ending.
\end{quotation}

\section{Motivation}
\label{sc:motivation}
...

\section{Problem Statement}
\label{sc:problems}
...

\section{Contributions}
\label{sc:contributions}
...

\section{Thesis Organization}
\label{sc:organization}
...

### 2017-09-19 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
compiling version

\chapter{PCA and SVM for Critical Factors Analysis}
\label{ch:2_pca_svm_cfa}

\begin{quotation}[]{Paulo Freire}
No one knows it all. No one is ignorant of everything. We all know something. We are all ignorant of something.
\end{quotation}

There is a significant continuous public spending on information technology (IT) by the Brazilian Federal Public Administration. In order to monitor and to diagnose the concerned public institutions IT governance (ITG), the Federal Court of Accounts, in Portuguese Tribunal de Contas da União (TCU), surveys data from these institutions practices as well as best practices for IT in government and business. These surveys started in 2007 with 39 questions and nowadays they contain more than 100 questions impacting on the ITG of Brazilian public institutions. Moreover, these questions encourage public institutions to adopt best practices not only in IT but also in all other management areas. In this paper, we propose the identification and the verification of the critical success factors (CSFs) of the TCU survey variables by means of statistical analysis. The CSFs are defined as the factors that most contributed for the high performance of the institutions and they are measured by using the ITG index created by TCU. Therefore, in order to validate our results, the statistically identified CSFs are compared with those mentioned by the public administration IT executives in interviews. Besides the statistical analysis, we successfully apply the Support Vector Classification (SVC) algorithm to classify public institutions in terms of their ITG index. By comparing the SVC based classification with the CSFs obtained from the interviews, we show that there is a very high level of similarity. Hence, the CSFs identified with SCV provide a very high impact, regarding the public administration ITG.

2.3 Análise de Componentes Principais

Segundo Sabin, Ferrão e Furtado (2004) e Silva et al. (2012), a Análise de Componentes Principais (ACP), do inglês Principal Component Analysis – PCA, é utilizada para compressão de dados para identificação das relações entre características dos dados. Tal compressão é obtida por meio da substituição das variáveis originais por um novo conjunto de variáveis, conhecidas como Componentes Principais (CPs). As CPs são obtidas pela combinação linear das variáveis que apresentam a maior variabilidade na matriz de covariância. Assim, objetivo principal da ACP é identificar as CPs responsáveis pelas maiores variações entre os resultados. Em termos práticos eliminam-se algumas variáveis com pouca informação e determinam-se as variáveis de maior influência na formação de cada CP (Vicini, 2005).
Segundo Vicini (2005), o conjunto ortogonal de eixos não correlacionados representam as variáveis que mantém ao máximo a variabilidade do conjunto, em outras palavras, a menor perda de informação. 
Conforme Vicini (2005:29), para o processo de determinação das CPs 
“é necessário calcular matriz de variância-covariância (∑), ou na matriz de correlação (), encontrar os autovalores e autovetores e, por fim, escrever as combinações lineares, que serão as novas variáveis, denominadas de componentes principais, sendo que cada componente principal é a combinação linear de todas as variáveis originais [...] em ordem de estimação e em termos da variância total, contida nos dados iniciais”. 
Ressalta-se que não são todos as CPs a serem analisadas, mas apenas as com maior variância, alguns autores (Vicini, 2005) determinam os CPs por meio da porcentagem que representam, geralmente, mais de 70\% da informação. 
Conforme Silva et al. (2012) para a geração dos CPs, considera-se uma matriz X = (X1, X2, ..., Xp) de dimensões nxp (a Seção 4 mostra as dimensões utilizadas nesta pesquisa) originada a partir de um vetor aleatório. A matriz de variância-covariância  das amostras da matriz X é calculada por meio da seguinte expressão , onde é o operador de transposição de uma matriz. Nota-se que a matriz de variância-covariância é uma matriz quadrada e não negativa. Portanto, pode-se aplica a decomposição em autovalores, do Eigenvalue Decomposition (EVD), da matriz  resultando em um produto de três matrizes.
,
onde  é a matriz de autovetores e  é a matriz cuja diagonal principal contém os autovalores. Para cada autovalor λi da matriz  existe um autovetor unitário e ortognal vi.
vi = 
A i-nésima coluna da matriz  é o autovetor normalizado  correspondente ao autovalor . Desta forma, define-se i-ésima componente principal (Yi) como sendo
Yi= λ i vi
Salienta-se que a ACP é baseada na EVD já que parte dos autovalores e dos autovetores representam as componentes principais (CP). 
A Figura 1 mostra geometricamente a ACP de duas variáveis X1 e X2. A nuvem de pontos representa o diagrama de espalhamento destas duas variáveis. Conforme explicado, o primeiro passo é o cálculo da matriz de variância-covariância e em seguida se aplica a EVD. Verifica-se que a CP1, denominado CP principal (com a maior variância), é ortogonal à segunda CP (CP2 com a menor variância). Note que as duas variáveis são perpendiculares porque as variáveis obtidas pela ACP são independentes. Logo, após a aplicação da ACP, ao invés de representar os dados por meio de duas variáveis correlacionadas X1 e X2, eles podem ser representados por duas variáveis independentes Y1 e Y2. Dada a elevada correlação entre X1 e X2, verifica-se que as duas variáveis X1 e X2 podem ser aproximadas por uma única variável Y1 na Figura 1.

Figura 1 – Representação e Rotação do Componente Principal
Fonte: Desenvolvido pelos autores

A Figura 1 reforça a ideia da ACP cujo objetivo inicial é a identificação de planos e linhas que representem um conjunto de pontos em um espaço com um número menor de variáveis tendo em vista as possíveis correlações que podem existir entre as variáveis originais.
Assim, para o método da estatística multivariada ACP utilizar todas variáveis, separa-se a informação útil da informação redundante (Finkler, 2003). Ressalta-se que a ACP é um dos primeiros passos para outras análises multivariadas, sendo um dos métodos mais comuns empregados na análise de informações (Sabin, Ferrão e Furtado, 2004; Silva et al.,2012).

2.4 Máquinas de Vetores de Suporte (MVS)

Máquina de Vetores de Suporte (MVS), do inglês Support Vector Machine (SVM), é um método supervisionado de aprendizagem de máquina, não probabilístico, baseado na teoria de aprendizagem estatística, usado para classificação, regressão e detecção de padrões. MVS pode ser aplicado por meio de dois passos: o primeiro passo é o treinamento de um modelo a partir de um determinado conjunto de dados; o segundo consiste em estimar a classificação de dados a partir da aplicação do modelo treinado. 
Quando aplicado para classificação, MVS busca a identificação de um hiperplano que separa os dados em classes distintas, por meio de uma margem máxima entre duas classes de dados. Desta forma, um conjunto de dados é linearmente separável se for possível dividir seus dados em duas classes, por meio um hiperplano, conforme na figura a seguir.

Figura 2 – Separação Linear por Margem Máxima

Dado um conjunto de dados previamente classificados, (xi, yi), xi ∈ Rn, yi ∈ {1, −1}, i = 1, . . . , l, um classificador linear pode ser definido por: wT ⋅ x + b = 0, desta forma o hiperplano ótimo é definido pelos valores ótimos do vetor de pesos wi e do bias bi, de forma que wT ⋅ x + b = 1 e wT ⋅ x + b = -1 representam respectivamente os vetores de suporte positivos e negativos, que são os pontos próximos à margem do hiperplano ideal. A margem máxima, entre os vetores de suporte, é definida por:

As funções de kernel têm o objetivo de projetar os vetores de variáveis de um conjunto de dados em um espaço de maior dimensão, para a classificação de classes originalmente apresentadas em espaços não separáveis linearmente. Com o aumento da dimensão, aumenta a probabilidade desses dados poderem ser linearmente separáveis. 
Uma função kernel K(xi, xj) = φ(xi) T φ(xj) pode ser usada para treinar a MVS. Uma MVS linear tem φ(x) = x, assim uma função kernel linear pode ser representada por K(xi, xj) = xTi xj.
Foi adotada a estratégia "um contra um" (Knerr, Personnaz e Dreyfus, 1990), para a utilização de MVS para a classificação de muitas classes. Esta estratégia consiste em construir uma máquina vetor de suporte para cada par de classes. Para um problema com c classes, c(c-1)/2 MVSs são treinados para classificar as classes entre as c classes possíveis.

2 Eliminação Recursiva de Variáveis (ERV)

Dado um algoritmo de classificação, que possa estimar pesos para as variáveis de um conjunto de dados, o objetivo da Eliminação Recursiva de Variáveis (ERV), do inglês Recursive Feature Elimination (RFE) (Guyon, 2002), é selecionar variáveis por meio da redução recursiva da quantidade de variáveis, eliminando recursivamente as variáveis de menor peso para classificação por meio do algoritmo adotado. 
Primeiramente, um modelo é treinado, utilizando o conjunto de dados inicial e o algoritmo selecionado, durante o treinamento são atribuídos pesos para cada variável, representando a importância de cada variável para a classificação. Em seguida, as variáveis com menor peso são eliminadas do conjunto de dados. Este processo, de treinamento, ordenamento de variáveis e eliminação de variáveis menos importantes, é repetido recursivamente até a obtenção do número desejado de variáveis, ou até a satisfação de alguma condição, como um limiar de taxa de erro de um algoritmo de classificação.
As variáveis com maiores pesos apresentam maior influência na classificação (Guyon, 2002). Desta forma, se um algoritmo de classificação apresenta boa acurácia, as variáveis com maiores pesos representam as variáveis que apresentam maior influência para a classificação.
MVS-ERV (Guyon, 2002) é uma aplicação da ERV utilizando os pesos obtidos por meio do treinamento utilizando MVS como algoritmo de classificação, para identificar as variáveis mais importantes para predições de classificação e eliminar recursivamente as variáveis que menos influenciam na classificação.

4.3 Análise das componentes principais

Para a análise eficiente da matriz original  é relevante separar as informações úteis das redundantes. Segundo Johnson e Wichern (1992), há vários instrumentos para essa finalidade, mas a ACP é a que melhor desempenha este papel. 
Uma vez calculados os autovalores da matriz original, temos um resultado que mostra que aproximadamente 70\% da variabilidade dos dados é explicado por 51 componentes principais. O critério para a escolha desses fatores foi o de identificar os autovalores que possuem variância acumulada em torno de 70\% (Mardia, 1979). Tal valor também será utilizado nesta pesquisa (Quadro 7).

 Quadro 7 – Autovalores
Ordem dos autovalores
Autovalores
\% da variância explicada
Autovalores acumulados
\% da variância explicada acumulada
1. 
24,88966
12,38292
24,8897
12,3829
2. 
10,33305
5,14082
35,2227
17,5237
3. 
7,49873
3,73071
42,7214
21,2545
4. 
5,27875
2,62625
48,0002
23,8807
5. 
4,78008
2,37815
52,7803
26,2588
6. 
4,49112
2,23439
57,2714
28,4932
7. 
3,73598
1,85870
61,0074
30,3519
8. 
3,59164
1,78689
64,5990
32,1388
9. 
3,19618
1,59014
67,7952
33,7290
10. 
3,11010
1,54731
70,9053
35,2763
11. 
2,66175
1,32426
73,5671
36,6005
12. 
2,54482
1,26608
76,1119
37,8666
13. 
2,51698
1,25223
78,6289
39,1188
14. 
2,40926
1,19864
81,0381
40,3175
15. 
2,35614
1,17221
83,3943
41,4897
16. 
2,30653
1,14753
85,7008
42,6372
17. 
2,25284
1,12081
87,9536
43,7580
18. 
2,11772
1,05359
90,0714
44,8116
19. 
2,06312
1,02643
92,1345
45,8380
20. 
2,01891
1,00443
94,1534
46,8425
21. 
1,99258
0,99133
96,1460
47,8338
22. 
1,97266
0,98142
98,1186
48,8152
23. 
1,95377
0,97203
100,0724
49,7873
24. 
1,86320
0,92696
101,9356
50,7142
25. 
1,83124
0,91107
103,7668
51,6253
26. 
1,80163
0,89633
105,5685
52,5216
27. 
1,76688
0,87905
107,3354
53,4007
28. 
1,70876
0,85013
109,0441
54,2508
29. 
1,69776
0,84465
110,7419
55,0955
30. 
1,66091
0,82633
112,4028
55,9218
31. 
1,62693
0,80942
114,0297
56,7312
32. 
1,60039
0,79621
115,6301
57,5274
33. 
1,54662
0,76946
117,1767
58,2969
34. 
1,51306
0,75277
118,6898
59,0496
35. 
1,51005
0,75127
120,1998
59,8009
36. 
1,47437
0,73352
121,6742
60,5344
37. 
1,43597
0,71441
123,1102
61,2488
38. 
1,42247
0,70770
124,5326
61,9565
39. 
1,40464
0,69882
125,9373
62,6554
40. 
1,35176
0,67252
127,2890
63,3279
41. 
1,33162
0,66250
128,6207
63,9904
42. 
1,32118
0,65731
129,9418
64,6477
43. 
1,29010
0,64184
131,2320
65,2895
44. 
1,26013
0,62693
132,4921
65,9165
45. 
1,25589
0,62482
133,7480
66,5413
46. 
1,22165
0,60779
134,9696
67,1491
47. 
1,20095
0,59749
136,1706
67,7466
48. 
1,18070
0,58741
137,3513
68,3340
49. 
1,17744
0,58579
138,5287
68,9198
50. 
1,16657
0,58038
139,6953
69,5001
51. 
1,14340
0,56886
140,8387
70,0690
Fonte: Dados da Pesquisa
Depois da extração dos autovalores e percentual da variância explicada, é decidida a quantidade de fatores a serem retirados para análise. Para isso, o Gráfico 1 em que o total de autovalores está no eixo das ordenadas e os autovalores  no eixo das abscissas, auxilia na identificação. Esse gráfico consiste no ranking dos autovalores (eixo x), relacionado com o valor de cada autovalor (eixo y). Verifica-se que uma queda menos acentuada ocorreu entre o quarto e o quinto autovalor e analisando-se os autovalores superiores a 2, observa-se que pode-se considerar até o vigésimo valor já que a partir daí os valores dos autovalores sucessivos são praticamente constantes.

Gráfico 1 – Ranking dos autovalores
Fonte: Dados da pesquisa

Visando encontrar os planos fatoriais realizou-se uma rotação dos eixos, onde as cargas fatoriais mais elevadas são as responsáveis pelas denominações das componentes e são estatisticamente significativas. As rotações de eixos melhor expressam a dispersão de dados. No modelo fatorial final, as variáveis das medidas estão maximizadas e as relações entre dimensões suavizadas (Vicini, 2005).
Para esta análise buscam-se valores que possuem significância maior que 0,7, mostrando que a correlação entre as variáveis está de moderada a forte (Vicini, 2005). Essa identificação não seria possível sem a rotação dos eixos, possibilitando assim a melhor visualização das variáveis mais significativas em cada componente. Tal rotação mantem os eixos perpendiculares entre si, ou seja, ortogonais e a variabilidade do sistema não é alterada, apenas as coordenadas dos eixos são rotacionadas e a inércia do sistema fica inalterada. 
A partir dos valores obtidos pela rotação das componentes principais, podem-se obter valores com significância maior que 0,7. Logo, foi possível a identificação das variáveis significantes de cada componente principal.
Para visualização desses fatores foi utilizado o gráfico de dispersão (Gráfico 2). O Gráfico 2 mostra a caixa de seleção de variáveis e comandos para ACP em que se utilizam os fatores 1 e 2, eixo x e eixo y respectivamente. O objetivo deste gráfico é fazer os planos principais com a nuvem de pontos dos indivíduos, no caso as 349 instituições, destacando o posicionamento das respostas das instituições classificadas pelo iGovTI. Tal gráfico se baseia na rotação dos componentes principais. Para a elaboração do Gráfico 2 apenas os  e  foram utilizados.

Gráfico 2 – Relação entre os  e  
Fonte: Dados da pesquisa

No Gráfico 2, para analisar apenas a , projetam-se os pontos sobre o eixo da e se tem três grupos (V, II e I). O mesmo processo se aplica a analise da . Assim, observam-se três grupos (V, IV e III). Para cada grupo se seleciona apenas uma variável representativa, o restante é desconsiderado, uma vez que cada grupo significa correlação.
Para o grupo I, tomam-se as variáveis Q16d, Q16e e Q16f que são descorrelacionadas, já no grupo II a questão Q16g é questão selecionada. Ambos os grupos não apresentam FCS. Essas questões denotadas Q16, referem-se a Subdimensão 1.6 (A Alta Administração Utilizou Informações Fornecidas pela Auditoria Interna). Tal subdimensão verifica a participação da auditoria interna das instituições para o preenchimento do questionário. 
Já no grupo III, que não tem FCS, a variável Q53_Relev é a selecionada; precedidas do grupo IV que também não apresenta FCS. Ressalta-se que ambas as questões apuram a relevância de determinado item. Pela análise ACP, o grupo V mostra as questões que não são representadas pelas  e , é justamente este grupo que contem os FCS.
Depois, buscou-se por meio da análise das ,  e a identificação de FCS (Gráfico 3). O Gráfico 3 mostra a análise dessas componentes principais que revela a existência de três grupos.

Gráfico 3 – Relação entre os ,  e 
Fonte: Dados da pesquisa

A análise do Gráfico 3 mostra revela a existência de três grupos (VI, VII e VIII). O grupo I contém a variável Q51l que corresponde a FCS e o grupo II também mostra FCS, a variável Q23b.
Na identificação das variáveis significativas das primeiras 20 componentes principais, com os autovalores superiores a 1, as questões que mais contribuem são Q16b (para responder às questões do grupo 2. Estratégias e planos), Q16c (para responder às questões do grupo 3. Informação e conhecimento), Q16d (para responder às questões do grupo 4. Pessoas), Q16e (para responder às questões do grupo 5. Processos) e Q16f (para responder às questões do grupo 6. Resultados da gestão), todas da dimensão “Governança corporativa e de TI, a alta administração utilizou informações fornecidas pela auditoria interna (ou instância equivalente)”. Nota-se que as variáveis Q16, citadas anteriormente, fazem parte do grupo I do Gráfico 2. Ao total foram identificadas 29 variáveis dos 20 primeiras componentes principais (autovalores maiores que 1) das variáveis com significância maior que 0,7. O Quadro 8 mostra quais são essas variáveis significativas.

Quadro 8 – Variáveis com alta significância CP1 a CP20
CP
Dimensão
Var.1
Sig.
Var. 2
Sig.
FCS
CP 01
D1
Q16b
0,8108
Q16c
0,7496
não
CP 01
D1
Q16d
0,7923
Q16e
0,8469
não
CP 01
D1
Q16f
0,7998

não
CP 02
D5
Q51_Relev
0,8941
Q53_Relev

0,8176
não
CP 03
D2
Q23a
-0,8833
Q23b
0,9009
não/sim
CP 04
D5
Q51l
0,7925
 Q51l1
0,8254
sim/não
CP 05
D1
Q13a
0,8842

não
CP 06
D5
Q57h
0,8285

sim
CP 07
D1
Q11f2
0,8362

não
CP 08
D5
Q53e
0,8684

sim
CP 09
D7
Q71
0,9111

não
CP 10
D4
Q45b
-0,8960

não
CP 11
D1
Q14c
-0,8605

não
CP 12
D5
Q510d2
0,8322

não
CP 13
D6
Q64h
0,9240

não
CP 14
D8
Q83a
0,8822

não
CP 15
D1
Q15d
0,8738
Q15g 
-0,7003 
não
CP 16
D3
Q31b
0,9098
Q31e 
-0,9070 
não
CP 17
D2
Q23i
0,9286

não
CP 18
D1
Q11c1
0,9204

não
CP 19
D5
Q510e1
0,9362

não
CP 20
D4
Q45j
0,9536

não
Total
29
4
Fonte: Dados da pesquisa

No Quadro 8, verificou-se as variáveis que são FCS, de acordo com resultado da pesquisa apresentada na Subseção 4.1. Das 30 questões apenas 5 são FCS, de acordo com classificação obtida por meio de pesquisa qualitativa, quais sejam: Q12c (designou representantes de todas as áreas relevantes para o negócio institucional para compor o Comitê de TI); Q23b (a instituição aprovou e  publicou PDTI interna e externamente); Q51l (gestão de configuração de ativos); Q53e (formalizou a política corporativa de segurança da informação); Q57h (os pagamentos são feitos em função da mensuração objetiva dos resultados entregues e aceitos.
Ainda no Quadro 8, chama-se atenção as ,  e. Nelas constam variáveis com valor positivo ou negativo. Tal fato significa que quanto maior for a resposta de uma questão menor será a da outra. Frisa-se que isto só vale para questões que estejam em uma mesma variável ACP. Caso as questões tenham sinais diferentes (i.e. os elementos dos autovetores), mas em ACP diferentes o sinal positivo e negativo não importa. 
A análise das categorias (Quadro 5) relacionadas às variáveis revela: Q53e referente à Categoria Processos de Gestão de Serviços de TI em Desenho do Serviço que mostrou porcentual igual a 53,85\%; Q57h se liga à Categoria Gestão de Contratos que teve porcentual de 38,46\%; Q23b se vincula à Categoria PDTI que teve 38,46\%. A variável Q51l se relaciona à Categoria Processos de Gestão de Serviços de TI em Transição de Serviços com 19,23\%.
Devido ao baixo percentual de 16,66\% de identificação de FCS do Quadro 8, realizou-se a análise das variáveis que completam as 51 componentes principais, referentes aos 51 autovalores do Quadro 7. Nesta identificação, foram identificadas 33 variáveis, sendo que 8 são FCS, 24,24\% (Quadro 9). Somando os totais obtidos pelo Quadro 8 e pelo se tem, 20,63\% de variáveis que são FCS.

Quadro 9 – Variáveis com alta significância CP 21 a CP 51
CP
Dimensão
Var.1
Sig.
Var. 2
Sig.
FCS
CP 21
D1
Q12e 
0,7210
 Q12e1 
0,8314
sim/não
CP 22
D5
 Q55a
0,8256

não
CP 23
D8
 Q83b2
0,8816

não
CP 24
D7
Q72a3 
0,9220

não
CP 25
D4
 Q45g
-0,9334

não
CP 26
D5
Q510b1
0,8942

não
CP 27
D1
Q12e3 
0,9294

não
CP 28
D1
Q12b 
0,8415
 Q12c
0,8934
sim/sim
CP 29
D1
 Q12e2
0,9621

não
CP 30
D1
 Q11_Relev
0,8400

não
CP 31
D2
 Q24
0,9425

sim
CP 32
D5
 Q510b2
0,9024

não
CP 33
D5
 Q53a1
0,8840

não
CP 34
D1
 Q13e
0,8986

não
CP 35
D4
 Q45d
0,9487

não
CP 36
D1
Q16_relev 
0,8364

não
CP 37
D5
 Q51h
0,8175

sim
CP 38
D5
 Q53d
0,8864

não
CP 39
D1
 Q11c2
0,9258

não
CP 40
D1
Q11a 
0,8674

não
CP 41
D1
 Q15f
0,9201

não
CP 42
D4
 Q45f
0,9450

não
CP 43
D5
 Q53f
0,7871

não
CP 44
D4
Q45h 
-0,9612

não
CP 45
D7
 Q72a4
0,9401

não
CP 46
D3
 Q31d
0,8714

sim
CP 47
D4
 Q45i
0,9462

não
CP 48
D5
 Q51e
0,8432

sim
CP 49
D4
 Q45e
0,9406

não
CP 50
D1
 Q12a1
0,8635

sim
CP 51
D8
 Q83b5
0,9237

não
Total
33
8
Fonte: Dados da pesquisa

Já o Gráfico 4 mostra a caixa de seleção de variáveis e comandos para ACP em que se utilizam os fatores 1 e 2, eixo x e eixo y respectivamente. O objetivo deste gráfico é fazer os planos principais com a nuvem de pontos dos indivíduos, no caso as 349 instituições.

Gráfico 4 – Dados brutos da pesquisa
Fonte: Dados da pesquisa

O Gráfico 4 representa a relação entre os 2 principais componentes, que representam as questões ou variáveis com maiores autovalores, assim cada ponto representa a relação entre uma instituição e seus 2 principais componentes obtidos por meio de ACP. Cada ponto deste gráfico representa a classificação da instituição em relação a sua classificação no iGovTI de 2012. A partir do resultado apresentado no gráfico 4 é possível identificar um padrão formado pela localização dos componentes principais das organizações com maior índice, onde as organizações com maior iGovTI tiveram valores relativamente semelhantes para seus 2 componentes principais, entretanto ainda não é possível obter uma separação clara entre os resultados. Fazendo-se um corte no eixo das abcissas, a partir do ponto (-8,8, 0), destacam-se 23 instituições à esquerda desse ponto (Quadro 5).

Instituição
Posição
iGovTI

33
5
0,7954

35
3
0,7984

36
11
0,7371

39
2
0,8226

44
19
0,7244

57
1
0,88

63
18
0,7247

77
4
0,7963

80
12
0,736

92
10
0,7615

122
8
0,7723

142
21
0,7131

158
6
0,7944

186
7
0,7913

188
17
0,7269

197
40
0,6403

201
13
0,7356

205
9
0,7687

311
22
0,7107

346
75
0,5644

347
16
0,7322

349
39
0,6427

351
28
0,6817
Gráfico 5 – Destaque de corte nas instituições 
Fonte: Dados da pesquisa

Dessas instituições, 22, são consideradas pelo TCU como aprimoradas, ou seja, 95,65\%. Apenas a instituição 346 é avaliada como intermediária.
Após a análise ACP, verifica-se que a mesma não é útil para a análise da Hipótese 2. Para responder a essa hipótese, é necessária a execução dos algoritmos de classificação da Subseção 4.4 e da Subseção 4.5.

4.4 Máquinas de Vetores de Suporte

Para prever a classificação de uma instituição de acordo com suas respostas para as questões do questionário iGovTI, foram avaliados algoritmos de classificação que pudessem apresentar acurácia para a classificação de organizações de acordo com o iGovTI. Uma vez encontrado um algoritmo capaz desta classificação, é possível utilizar o algoritmo selecionado em conjunto com técnicas de feature selection para identificar as questões mais relevantes para a classificação da instituição de acordo com o iGovTI.
Assim, para a definição do algoritmo de classificação foram realizados testes em 21 algoritmos, a seguir é apresentada uma listagem dos algoritmos avaliados e a taxa de acerto de classificação obtida: 
1. KNN: 0.714286
2. ElasticNet: 0.155336
3. ElasticNetCV: 0.831531
4. LassoCV: 0.827440
5. LassoLarsIC: 0.713763
6. LinearRegression: 0.360878
7. LogisticRegression: 0.771429
8. OrthogonalMatchingPursuit: 0.769630
9. PassiveAggressiveClassifier: 0.800000
10. PassiveAggressiveRegressor: 0.852184
11. Perceptron: 0.800000
12. Ridge: 0.529922
13. RidgeClassifier: 0.657143
14. RidgeClassifierCV: 0.742857
15. RidgeCV: 0.750149
16. SGDClassifier: 0.828571
17. MultinomialNB: 0.742857
18. lda.LDA: 0.628571
19. SVM.SVR: 0.826885
20. SVM.SVC: 0.914286
21. SVM.LinearSVC: 0.714286
22. O algoritmo que obteve maior sucesso para classificação foi o SVC, que é uma implementação de Máquina de Vetores de Suporte aplicado para a classificação. SVC apresentou uma taxa de acerto de 91,4\%. Para a avaliação dos algoritmos, utilizamos uma metodologia que divide os dados dos questionários pelo iGovTI entre questões que serão utilizadas para aprendizado do algoritmo e questões que serão utilizadas para comparativo de predições. Para treinar o algoritmo utilizou-se 90\% dos dados e os 10\% restantes foram utilizados para avaliar a eficiência de classificação dos algoritmos, comparando a taxa de acerto entre as predições feitas e os valores reais de classificações de organizações pelo iGovTI. 
23. Uma vez identificado um algoritmo capaz de efetuar a classificação desejada, o próximo passo é identificar as variáveis mais relevantes para esta classificação. Para isso, será utilizado o algoritmo ERV. 
24. 
4.5 Eliminação Recursiva de Variáveis
25. 
26. A ERV pode usar vários algoritmos de classificação como critério de seleção das variáveis mais importantes, escolhemos o algoritmo SVC por ele ter apresentado maior acurácia entre os algoritmos de classificação avaliados. Ainda é necessário definir o quantitativo de variáveis mais importantes a serem selecionadas. A partir da pesquisa de identificação por meio de entrevistas na APF, identificou-se 54 variáveis consideradas FCS, este critério foi utilizado para determinar o quantitativo de variáveis mais importantes para a classificação.
27. Assim, aplicaram-se o algoritmo ERV utilizando o SVC como critério para a seleção das variáveis mais importantes para a classificação, selecionando as 54 variáveis que correspondem aos FCS levantados nas entrevistas com os executivos de TI. Os resultados mostraram que 69,9\% das variáveis foram classificadas da mesma forma que os FCS identificados anteriormente (Quadro 10) por meio de pesquisa qualitativa.

### 2017-09-19 docs/thesis/chapters/3_eigen_similarity/capitulo3.tex
compiling version

--- 2017-01-03 docs/papers/2016JNCA/2016JNCA.tex
@@
-\documentclass[review]{elsarticle}
-
-\usepackage{lineno,hyperref}
-\usepackage[mathletters]{ucs}
-\usepackage[utf8x]{inputenc}
-\usepackage{algorithm,algorithmic}% http://ctan.org/pkg/algorithms
-\usepackage{natbib}
-\usepackage{supertabular}
-\usepackage{fancybox}
-\usepackage{acronym}
-\usepackage{array}
-\usepackage{booktabs}
-\usepackage{graphicx}
-\usepackage{rotating}
-\usepackage{tabularx}
-\usepackage{colortbl}
-\usepackage{multirow}
-\usepackage{hhline}
-\usepackage{setspace}
-\usepackage{placeins}
-\usepackage{longtable}
-\usepackage{mathtools}
-\usepackage{listings}
-\usepackage{epstopdf}
-\usepackage{subfigure}
-\usepackage{courier}
-\usepackage{amsfonts}
-\usepackage{morefloats}
-\usepackage{lipsum}
-\usepackage{amsthm,amsmath}
-\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
-\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\newcommand*\justify{%
-	\fontdimen2\font=0.4em% interword space
-	\fontdimen3\font=0.2em% interword stretch
-	\fontdimen4\font=0.1em% interword shrink
-	\fontdimen7\font=0.1em% extra space
-	%\hyphenchar\font=`\-% allowing hyphenation
-}
-
-% table
-\usepackage{color, colortbl}
-\definecolor{Gray}{gray}{0.9}
-
-% Algorithmic modifications
-\makeatletter
-\newcommand{\ALOOP}[1]{\ALC@it\algorithmicloop\ #1%
-  \begin{ALC@loop}}
-\newcommand{\ENDALOOP}{\end{ALC@loop}\ALC@it\algorithmicendloop}
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\newcommand{\algorithmicbreak}{\textbf{break}}
-\newcommand{\BREAK}{\STATE \algorithmicbreak}
-\makeatother
-
-% appendix
-\usepackage[titletoc,title]{appendix}
-\usepackage{hyperref}
-\usepackage{cleveref}
-
-\modulolinenumbers[5]
-
-\journal{ Journal of Network and Computer Applications}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
-%% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
-
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
-%% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
-%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-\begin{frontmatter}
-
-\title{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
-
-%% Group authors per affiliation:
-\author[unbaddress]{Thiago P. B. Vieira}
-\author[unbaddress]{Danilo F. Tenório}
-\author[unbaddress,Ilmenauaddress,Fraunhoferaddress]{João Paulo C. L. da Costa}
-\author[ufrgsaddress]{\\Edison P. de Freitas}
-\author[Ilmenauaddress,Fraunhoferaddress]{Giovanni Del Galdo}
-\author[unbaddress]{Rafael T. de Sousa Júnior}
-
-\address[unbaddress]{Department of Electrical Engineering, University of Brasilia (UnB), 70910-900, Brasília-DF, Brazil}
-\address[Ilmenauaddress]{Institute for Information Technology, Ilmenau University of Technology, Ilmenau, Germany}
-\address[Fraunhoferaddress]{Fraunhofer Institute for Integrated Circuits IIS, Erlangen, Germany}
-\address[ufrgsaddress]{Graduate Program in Electrical Engineering, Federal University of Rio Grande do Sul (UFRGS), 90035-190, Porto Alegre, Brazil}
-
-\begin{abstract}
-Novel schemes for attack detection are crucial to identify adaptive malicious traffic coming from sources that are quickly mobilized by attackers in high throughput communication networks.
-In this context, signal processing techniques have been applied to attack detection due to their capability to detect anomalies that are previously unknown, i.e.
-blind detection.
-This paper proposes a signal processing framework for the detection and identification of network attacks using concepts of model order selection (MOS), eigenvalues and similarity analysis.
-In order to validate the proposed framework, we consider network traffic datasets that contain malicious activity such as flood and port probing attacks.
-We propose to model the network traffic as a superposition of components, namely, user's operations (legitimate traffic), network service operation not related to the user (noise) and the malicious activity.
-The experiments performed in a real network and also using the DARPA 1998 public dataset show that the proposed blind detection approach achieves satisfactory levels of accuracy in terms of timely detection and identification of TCP/UDP ports under attack.
-\end{abstract}
-
-\begin{keyword}
-Network Attack Detection \sep Model Order Selection \sep Eigen Analysis \sep Similarity Analysis
-\end{keyword}
-
-\end{frontmatter}
-
-\linenumbers
+\chapter{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks} 
+\label{chapter3}
 
@@
 where $0 < q(\alpha,\beta) < 1$. According to \cite{quinlan2006model}, if $M \leq N$, then $\beta = N$.
-
-\section*{References}
-
-\bibliography{references}
-
-\end{document}

### 2017-09-19 docs/thesis/chapters/3_eigen_similarity/evaluation.tex
compiling version

\chapter{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
\label{ch:evaluation}

\begin{quotation}[]{Lao Tzu}
All difficult things have their origin in that which is easy, and great things in that which is small. 
\end{quotation}

The use of MapReduce for distributed data processing has been growing and achieving benefits with its application for different workloads. MapReduce can be used for distributed traffic analysis, although network traffic traces present characteristics which are not similar to the data type commonly processed through MapReduce, which in general is divisible and text-like data, while network traces are binary and may present restrictions about splitting, when processed through distributed approaches.

Due to the lack of evaluation of MapReduce for traffic analysis and the peculiarity of this kind of data, this chapter deeply evaluates the performance of MapReduce in packet level analysis and DPI of distributed application traffic, evaluating its scalability, speed-up and the behaviour followed by MapReduce phases. The experiments provide evidences for the predominant phases in this kind of MapReduce job, and show the impact of input size, block size and number of nodes, on completion time and scalability.

This chapter is organized as follows. We first describe the motivation for a MapReduce performance evaluation for network traffic analysis in Section \ref{sc:eval_motivation}. Then we present the evaluation plan and methodology adopted in Section \ref{sc:eval_evaluation}, and the results are presented in Section \ref{sc:eval_results}. Section \ref{sc:eval_discussion} discusses the results and Section \ref{sc:eval_summary} summarizes the chapter.

\section{Motivation}
\label{sc:eval_motivation}

It is possible to measure, evaluate and diagnose distributed applications through the evaluation of information from communication protocols, flows, throughput, and load distribution \citep{Mi2012, Nagaraj2012, Sambasivan2011, Aguilera2003, Yu2011}. This information can be collected through network traffic analysis, but to retrieve application information from network traces, it is necessary to recognize the application protocol and deeply inspect the traffic to retrieve details about its behaviour, session and states.

MapReduce can be used for offline evaluation of distributed applications, analysing application traffic inside a data center, through packet level analysis \citep{Lee2011}, evaluating each packet individually, and through DPI \citep{Vieira2012_icpads,Vieira2012_icsea}, adopting a different approach for data splitting, where a whole block is processed without division into individual packets, due to the necessity to reassemble two or more packages to retrieve information of the application layer, in order to evaluate application messages and protocols.

The kind of workload submitted for processing by MapReduce impacts on the behaviour and performance of MapReduce \citep{Tan2012, Groot2012}, requiring specific configuration to obtain an optimal performance. Information about the occupation of MapReduce phases, the processing characteristics (if the job is I/O or CPU bound), and the mean duration of Map and Reduce tasks, can be used to optimize parameter configurations, and to improve resource allocation and task scheduling.

The main evaluations of MapReduce are in text processing \citep{Zaharia2008, Yanpei2011, Jiang2010, Guanying2009}, where the input data are split into blocks and into records, to be processed by parallel and independent Map functions. For distributed processing of network traffic traces, which are usually binary, the data splitting into packets is a concern and, in some cases, may require data without splitting, especially when packet reassembly is required to extract application information from the application layer.

Although work has been done to understand, analyse and improve workload management decisions in MapReduce \citep{Lu2012, Groot2012}, there is no evaluation to characterize the MapReduce behaviour or to identify its optimal configuration to achieve the best performance for packet level analysis and DPI.

Due to the lack of evaluation of MapReduce for traffic analysis and the peculiarity of this kind of data, it is necessary to understand the behaviour of MapReduce to process network traces and to understand what optimizations can be done to better explore the potential provided by MapReduce for packet level analysis and DPI.

This chapter evaluates MapReduce performance for network packet level analysis and DPI using Hadoop, characterizing the behaviour followed by MapReduce phases, scalability and speed-up, over variation of input, block and cluster sizes. The main contributions of this chapter are:

\begin{enumerate}
	\item Characterization of MapReduce phases behaviour for packet level analysis and DPI;
	\item Description of scalability behaviour and its relation with important MapReduce factors;
	\item Identification of the performance provided by the block sizes adopted for different cluster size;
	\item Description of speed-up obtained for DPI.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% na secao de metodologia, coloque uma figura para ilustrar todo o processo de coleta e analise. além disso, coloque numa tabela (ou mais de uma) a descricao das metricas de interesse, alem dos fatores e niveis selecionados em outra tabela. explique o motivo desses fatores influenciarem as metricas de interesse.
% Falar sobre o planejamento da avaliacao, destacando como sera feita an avaliacao e por que sera feito desta maneira. eh importante destacar que sera utilizada uma versao de p3 para comparar com a solucao desenvolvida.
% Descrever que avaliacoes de speedup e comportamento das fases foram aplicados para ambos os tipos analisados, packet analysis e dpi. Um comparativo especifico foi feito para identificar o desempenho de non-spliting contra p3
\section{Evaluation}
\label{sc:eval_evaluation}

The goal of this evaluation is to characterize the behaviour of MapReduce phases, its scalability characteristics over node addition and the speed-up achieved with MapReduce for packet level analysis and DPI. Thus, we performed a performance measurement and evaluation of MapReduce jobs that execute packet level analysis and DPI algorithms.

To evaluate MapReduce for DPI, the Algorithm \ref{alg:JXTAperfmapper} implemented by \textit{JXTAPerfMapper} and \textit{JXTAPerfReducer} was used, and applied to new factors and levels. To evaluate MapReduce for packet level analysis, a port counter algorithm developed by \cite{Lee2011} was used, which divides a block into packets and processes each packet individually to count the number of occurrence of TCP and UDP port numbers. The same algorithm was evaluated using the splitting approach that processes a whole block per Map function, without the division of a block into records or packets. A comparison was done between these two approaches for packet level analysis.

\subsection{Evaluation Methodology}
\label{ssc:eval_evaluation_methodology}

For this evaluation, we adopted a methodology based on the systematic approach to performance evaluation defined by \cite{Jain1991}, which consists of the definition of the goal, metrics, factors and levels for a performance study.

The goal of this evaluation is to characterize the behaviour of MapReduce phases, its scalability through node addition and the speed-up achieved with MapReduce for packet level analysis and DPI, to understand the impact of each factor on MapReduce performance for this kind of input data, in order to be able to configure the MapReduce and obtain an optimal performance over the evaluated factors.

\begin{table}[htb]
	\centering
	\caption{Metrics for evaluating MapReduce for DPI and packet level analysis}
	\label{tab:Metrics2}
    \begin{tabular}{|l|p{11cm}|} \hline
        \textbf{Metrics}	&\textbf{Description}\\ 																		\hline
        Completion Time   	&Completion time of MapReduce jobs\\ 															\hline
        Phases Time			&Time consumed by each MapReduce phase in total completion time of MapReduce jobs\\ 			\hline
        Phases Occupation	&Relative time consumed by each MapReduce phase in total completion time of MapReduce jobs\\ 	\hline
        Scalability			&Processing capacity increasing obtained with node addition in a MapReduce cluster\\			\hline
        Speed-up 			&Improvement in completion time against the same algorithm implemented without distributed processing\\ 									\hline
    \end{tabular}
\end{table}

Table \ref{tab:Metrics2} describes the metrics evaluated, that is the completion time of MapReduce jobs, the relative and absolute time of each MapReduce phases from the total job time, the processing capacity scalability, and the speed-up against non-distributed processing.

The experiments adopt the factors and level described in Table \ref{tab:FactorsLevels2}. The selected factors were chosen due to its importance for MapReduce performance evaluations and its adoption by relevant previous researches \citep{Jiang2010, Yanpei2011, Shafer2010, Guanying2009}.

\begin{table}[htb]
	\centering
	\caption{Factors and Levels}
	\label{tab:FactorsLevels2}	
	\begin{tabular}{|l|l|} 								\hline
	    \textbf{Factors}		&\textbf{Levels}		\\ \hline
   		Number of Worker Nodes	&2 up to 29				\\ \hline
	    Block Size				&32MB, 64MB and 128MB	\\ \hline
   		Input Size				&90Gb and 30Gb			\\ \hline
	\end{tabular}
\end{table}

Hadoop logs are a valuable information source about its environment and execution jobs; important MapReduce indicators and information about jobs, tasks, attempts, failures and topology are logged by Hadoop during its execution. The collected data to perform this performance evaluation was extracted from Hadoop logs.

To extract information from Hadoop logs and to evaluate the selected metricis, we developed the \textit{Hadoop-Analyzer} \citep{Hadoopanalyzer}, an open source and publicly available tool to extract and evaluate MapReduce indicators, such as job completion time and MapReduce phases distribution, from logs generated by Hadoop from its job executions. With \textit{Hadoop-Analyzer} is possible to generate graphs with the extracted indicators and thereby evaluate the desired metrics.

\textit{Hadoop-Analyzer} relies on \cite{Rumen} to extract raw data from Hadoop logs and generate structured information, which is processed and shown in graphs generated through \textit{R} \citep{Eddelbuettel2012} and \textit{Gnuplot} \citep{Janert2010}, such as the results presented in Section \ref{sc:eval_results}

\subsection{Experiment Setup}
\label{ssc:eval_experiment_setup}

Network traffic traces of distributed applications were captured to be used as input for MapReduce jobs of our experiments; these traces were divided into files with size defined by the block size adopted in each experiment, and then the files were stored into HDFS, following the process described in the previous chapter. The packets were captured using Tcpdump and were split into files with sizes of 32MB, 64MB and 128MB.

For packet level analysis and DPI evaluation, two sizes of datasets were captured from network traffic transferred between some nodes of distributed systems. One dataset was 30Gb of network traffic, with data divided in 30 files of 128MB, 60 files of 64MB and 120 files of 32MB. The other dataset was 90Gb of network traffic, split in 90 files of 128MB, 180 files of 64MB and 360 files of 32MB.

For the experiments of DPI through MapReduce, we used network traffic captured from the same JXTA-based application described in Section \ref{ssc:prof_setup}, but with different sizes of traces and files. To evaluate MapReduce for packet level analysis, we processed network traffic captured from data transferred between 5 clients and one server of a data storage service provided through the Internet, known as Dropbox\footnote{http://www.dropbox.com/}.

To evaluate MapReduce for packet level analysis and DPI, one driver was developed for each case of network traffic analysis, with one version using MapReduce and another without it.

\textit{CountUpDriver} implements packet level analysis for a port counter of network traces, which records how many times a port appears in TCP or UDP packets; its implementation is based on processing a whole block as input for Map functions, without splitting and with block size defined by the block size of HDFS. Furthermore a port counter implemented with \textit{P3} was evaluated; this implementation is a version of the tool presented by \cite{Lee2011}, which adopts an approach that divides a block into packets and processes each packet individually, without dependent information between packets.

\textit{JxtaSocketPerfDriver} implements DPI to extract, from a JXTA \citep{Duigou2003} network traffic, the round-trip time of JXTA messages, the number of connection requisitions per time and the number of JXTA Socket messages from JXTA clients and a JXTA Socket server. \textit{JxtaSocketPerfDriver} uses whole files as input for each Map function, with size defined by the HDFS block size, in order to reassemble JXTA messages with its content divided into many TCP packets.

One TCP packet can transport one or more JXTA messages per time, which makes it necessary to evaluate the full content of TCP segments to recognize all possible JXTA messages, instead of to evaluate only a message header or signature. The round-trip time of JXTA messages is calculated from the time between a client peer sending a JXTA message and receiving the JXTA message arrival confirmation. To evaluate the round-trip time it is necessary to keep the information of requests and which responses correspond to each request, and thus, it is necessary to analyse several packages to retrieve and evaluate information of the application behaviour and its states.

To analyse the speed-up provided by MapReduce against a single machine execution, two drivers were developed that use the same dataset and implement the same algorithms implemented by \textit{CountUpDriver} and \textit{JxtaSocketPerfDriver}, but without distributed processing. These drivers are respectively \textit{CountUpMono} and \textit{JxtaSocketPerfMono}.

The source code of all implemented drivers and other implementations to support the use of MapReduce for network traffic analysis, are open source and publicly available at \cite{Hadoopdpi}.

The experiments were performed on a 30-node Hadoop-1.0.3 cluster composed of nodes with four 3.2GHz cores, 8 GB RAM and 260GB of available hard disk space, running Linux kernel 3.2.0-29. Hadoop was used as our MapReduce implementation, and configured to permit a maximum of 4 Map and 1 Reduce task per node; also, we defined the value \textit{-Xmx1500m} as child option of JVM and 400 as $io.Sort.mb$ value.

For drivers \textit{CountUpDriver} and \textit{JxtaSocketPerfDriver}, the number of Reducers was defined as a function of the number of slots of Reducers per node, defined by $numReducers = (0.95)(numNodes)(maxReducersPerNode)$ \citep{Kavulya2010}. The driver implemented with \textit{P3} \citep{Lee2011} adopts a fixed number of Reducers, defined as 10 by the available version of \textit{P3}. Each experiment was executed 20 times to obtain reliable values \citep{Yanpei2011}, within confidence interval of 95\% and a maximum error ratio of 5\%.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sc:eval_results}

Two dataset sizes of network traffic were used during the experiments, with 30Gb and 90Gb. Each dataset was processed by MapReduce jobs that implement packet level analysis and DPI, in Hadoop clusters with variation in number of worker nodes between 2 and 29, and block size of 32MB, 64MB and 128MB.

Each dataset was processed by algorithms implemented through MapReduce and without distributed processing, to evaluate the speed-up achieved. The Table \ref{tab:monoExecution} shows the execution times obtained by the non-distributed processing, implemented and executed through \textit{JxtaSocketPerfMono} and \textit{CountUpMono}, using a single machine, with the resource configuration described in Subsection \ref{ssc:eval_experiment_setup}.

\begin{table}[!htbp]
  \centering
  \caption{Non-Distributed Execution Time in seconds}
  \label{tab:monoExecution}
  \begin{tabular}{*5c}
	\toprule
	\textbf{Block} &  \multicolumn{2}{c}{\textbf{JxtaSocketPerfMono}} & \multicolumn{2}{c}{\textbf{CountUpMono}} \\
	\midrule
    {}		& \textbf{90Gb}	& \textbf{30Gb}	& \textbf{90Gb}	& \textbf{30Gb}\\
   	\midrule
    32MB	& 1.745,35		& 584,92		& 872,40		& 86,71\\
    64MB	& 1.755,40		& 587,02		& 571,33 		& 91,76\\
    128MB	& 1.765,50		& 606,50		& 745,25		& 94,82\\
    \bottomrule
  \end{tabular}
\end{table}

Figure \ref{fig:JxtaMaxSpeedupBarLines} shows the completion time and speed-up of the DPI Algorithm \ref{alg:JXTAperfmapper} to extract indicators from a JXTA-based distributed application. The \textit{Completion Time} represents the job time of \textit{JxtaSocketPerfDriver} and the \textit{Speed-up} represents gains in execution time of \textit{JxtaSocketPerfDriver} against \textit{JxtaSocketPerfMono} to process 90Gb of network traffic.

\begin{figure}[htb]
	\centering
	\includegraphics[width=0.6\textwidth]{images/JxtaSocketPerfDriverMaxJobSpeedupBarLines.eps}
	\caption{DPI Completion Time and Speed-up of MapReduce for 90Gb of a JXTA-application network traffic}
	\label{fig:JxtaMaxSpeedupBarLines}
\end{figure}

According to Figure \ref{fig:JxtaMaxSpeedupBarLines}, \textit{JxtaSocketPerfDriver} performs better than \textit{JxtaSocketPerfMono} over all factors variation. Initially we observed the best speed-up of 3.70 times with 2 nodes and block of 128MB, lastly we observed a maximum speed-up of 16.19 times with 29 nodes and block of 64MB. The speed-up achieved with block size of 32MB was the worst case initially, but its speed-up increased with node addition and became better than blocks with 128MB and near of the speed-up achieved with block of 64MB, for a cluster with 29 nodes.

The completion time scalability behaviour of 32MB block size showed reduction in completion time for all node additions, although cases with block size of 64MB and 128MB present no significant reduction in completion time in clusters with more than 25 nodes. According to Figure \ref{fig:JxtaMaxSpeedupBarLines}, the completion time does not reduce linearly with node addition, and the improvement in completion time was less significant when the dataset was processed by more than 14 nodes, specially for cases that adopted blocks of 64MB and 128MB.

Figure \ref{fig:JxtaMaxProcessing} shows the processing capacity of MapReduce applied to DPI of 90Gb of a JXTA-based application traffic, over variation of cluster size and block size. The processing capacity was evaluated by the throughput of network traffic processed, and by the relative throughput, defined by the processing capacity achieved per number of allocated nodes.

\begin{figure}[htb]
	\centering
	\includegraphics[width=0.6\textwidth]{images/JxtaSocketPerfDriverMaxJobProcessingLines.eps}
	\caption{DPI Processing Capacity for 90Gb}
	\label{fig:JxtaMaxProcessing}
\end{figure}

The processing capacity achieved for DPI of 90Gb using block size of 64MB was 159.89 Mbps with 2 worker nodes, increasing up to 869.43 Mbps with 29 worker nodes. For the same case, the relative processing capacity achieved was 79.94 Mbps/node with 2 nodes and 29.98 Mbps/node with 29 nodes, showing a decrease of relative processing capacity with the growth of the MapReduce cluster size.

Although the processing capacity increased, the relative processing capacity, defined by the processing capacity per allocated node, decreased with all node addition. This behaviour indicates that MapReduce presents a reduction of efficiency with the increase of cluster size \citep{Gunther2006}, which highlights the importance of evaluation about the cost of node allocation and its benefits for completion time and processing capacity.

Figures \ref{fig:JxtaMaxSpeedupBarLines} and \ref{fig:JxtaMaxProcessing} also show the difference of performance achieved with different blocks sizes, and its relation to the cluster size. It was observed that blocks of 128MB achieved a higher throughput in cluster sizes up to 14 nodes, and that blocks of 64MB performed better in clusters bigger than 14 worker nodes.

Figures \ref{fig:JxtaMaxPhases} and \ref{fig:JxtaMaxPhasesPercent} show the behaviour of MapReduce phases to DPI of 90Gb.

\begin{figure}[!htb]
	\centering
	\subfigure[Phases Time for DPI]{
		\includegraphics[width=0.6\textwidth]{images/JxtaSocketPerfDriverMaxPhasesEvaluation.eps}
		\label{fig:JxtaMaxPhases}
	}
	\subfigure[Phases Distribution for DPI]{
		\includegraphics[width=0.6\textwidth]{images/JxtaSocketPerfDriverMaxPhasesPercentEvaluation.eps}
		\label{fig:JxtaMaxPhasesPercent}
	}
	\caption[MapReduce Phases Behaviour for DPI of 90Gb]{MapReduce Phases Behaviour for DPI of 90Gb}
	\label{fig:JxtaMaxPhasesBehaviour}
\end{figure}

MapReduce execution can be divided into Map, Shuffle, Sort and Reduce phases, although Shuffle tasks can be executed before the conclusion of all Map tasks, thereby Map and Shuffle tasks can overlap. According to the Hadoop default configuration, the overlapping between Map and Shuffle tasks starts after 5\% of Map tasks are concluded; and then Shuffle tasks are started and run until the Map phase ends.

In Figures \ref{fig:JxtaMaxPhases} and \ref{fig:JxtaMaxPhasesPercent} we showed the overlapping between Map and Shuffle tasks as a specific MapReduce phase, represented as "Map and Shuffle" phase. The time consumed by Setup and Cleanup tasks was considered too, for a better visualization of the execution time division in Hadoop jobs.

Figure \ref{fig:JxtaMaxPhases} shows the cumulative time of each MapReduce phase in total job time. For DPI, Map time, which is the Map and the "Map and Shuffle" phases, consumes the major part of a job execution time and it is the phase that exhibits more variation with the number of nodes variation, but no significant time reduction is achieved with more than 21 nodes and block size of 64MB or 128MB.

The Shuffle time, which happens after all Map tasks are completed, presented low variation with node addition. Sort and Reduce phases required relatively low execution times and did not appear in some bars of the graph. Setup and Cleanup tasks consumed an almost constant time, independently of cluster size or block size variation.

Figure \ref{fig:JxtaMaxPhasesPercent} shows the percentage of each MapReduce phase in total job completion time. We also considered an additional phase, called \textit{others}, which represents the time consumed by cluster management tasks, like scheduling and tasks assignment. The behaviour followed by phases occupation is similar over all block sizes evaluated, with the exception of the case where Map time does not decrease with node addition, in clusters using block size of 128MB and with more than 21 nodes.

With cluster size variation, a relative reduction in Map time was observed and a relative increase in the time of Shuffle, Setup and Cleanup phases. During the evaluation of Figure \ref{fig:JxtaMaxPhases}, it was observed that Setup and Cleanup phases consume an almost absolute constant time, independently of cluster size and block size, and thereby with node addition and completion time decreasing, the time consumed by Setup and Cleanup tasks became more significant in relation to the total execution time, due to the taotal job completion time decreasing and the time of Setup and Cleanup tasks remaining almost the same; therefore, the Setup and Cleanup percentage time increased and became more significant over the total job completion time reduction, by node addition into the MapReduce cluster.

According to Figures \ref{fig:JxtaMaxPhases} and \ref{fig:JxtaMaxPhasesPercent}, Map phase is predominant in MapReduce jobs for DPI, and the reduction of the total job completion time over node addition is related to the decreasing of Map phase time. Thus, improvements in Map phase execution for DPI workloads can produce more significant gains, in order to reduce the total job completion time for DPI.

Figure \ref{fig:CountUpDriverMaxJobTimeComp} shows the comparison between completion time of \textit{CountUpDriver} and \textit{P3} to packet level analysis of 90Gb of network traffic, over variation of cluster size and block size.

\begin{figure}[htb]
	\centering
	\includegraphics[width=0.7\textwidth]{images/CountUpDriverMaxJobTimeBarComp.eps}
	\caption{Completion time comparison of MapReduce for packet level analysis, evaluating the approach with and without splitting into packets}
	\label{fig:CountUpDriverMaxJobTimeComp}
\end{figure}

\textit{P3} achieves better completion time than \textit{CountUpDriver} over all factors, showing that a divisible files approach performs better for packet level analysis and that block size is a significant factor for both approaches, due to significant impact on completion time caused by adoption of blocks with different sizes.

With variation in the number of nodes, it was observed that using a block size of 128MB was achieved better completion time up to 10 nodes, but that no more improvement in completion time was achieved with node addition on clusters with more than 10 nodes. Blocks of 32MB and 64MB only present significant completion time difference in a cluster up to 14 nodes; for a cluster bigger than 14 nodes a similar completion time was achieved for both block sizes adopted, but these are still better completion times than the completion time achieved with blocks of 128MB.

Figures \ref{fig:P3MaxSpeedupBarLines} and \ref{fig:CountUpDriverMaxSpeedupBarLines} show respectively the completion time and speed-up of \textit{P3} and \textit{CountUpDriver} against \textit{CountUpMono}, for packet level analysis, with variation on the number of nodes and block size. For both cases, the use of a block size of 128MB provides the best completion time in smaller clusters, up to 10 nodes, but it provides a worse completion time in a cluster with more than 21 nodes. For both evaluations, the speed-up adopting block of 128MB scales up to 10 nodes, but for a bigger cluster a speed-up gain was not achieved with node addition.

\begin{figure}[!htb]
	\centering
	\subfigure[P3 evaluation]{
		\includegraphics[width=0.6\textwidth]{images/P3MaxJobSpeedupBarLines.eps}
		\label{fig:P3MaxSpeedupBarLines}
	}
	\subfigure[CountUpDriver evaluation]{
		\includegraphics[width=0.6\textwidth]{images/CountUpDriverMaxJobSpeedupBarLines.eps}
		\label{fig:CountUpDriverMaxSpeedupBarLines}
	}
	\caption[CountUp completion time and speed-up of 90Gb]{CountUp completion time and speed-up of 90Gb}
	\label{fig:CountUpMaxSpeedup}
\end{figure}

Using blocks of 32MB was achieved improvement in completion time for all node addition, which causes improvement of speed-up for all cluster size, although the use of this block size did not present better completion time than others block sizes in any case.

The adoption of 32MB blocks provided better speed-up than other block sizes in a cluster with more than 14 nodes, due to the time consumed by \textit{CountUpMono} for processing of 90Gb divided into 32MB files which was bigger than the time consumed by cases with other block sizes, as shown in Table \ref{tab:monoExecution}.

Figures \ref{fig:P3MaxThroughput} and \ref{fig:CountUpDriverThroughput} show the processing capacity of \textit{P3} and \textit{CountUpDriver} to perform packet level analysis of 90Gb of network traffic, over variation of cluster size and block size.

\begin{figure}[!htb]
	\centering
	\subfigure[P3 processing capacity]{
		\includegraphics[width=0.6\textwidth]{images/P3MaxJobProcessingLines.eps}
		\label{fig:P3MaxThroughput}
	}
	\subfigure[CountUpDriver processing capacity]{
		\includegraphics[width=0.6\textwidth]{images/CountUpDriverMaxJobProcessingLines.eps}
		\label{fig:CountUpDriverThroughput}
	}
	\caption[CountUp processing capacity for 90Gb]{CountUp processing capacity for 90Gb}
	\label{fig:CountUpThroughput}
\end{figure}

Using block size of 64MB, \textit{P3} achieved throughput of 413.16 Mbps with 2 nodes and the maximum of 1606.13 Mbps with 28 nodes, while its relative throughput for the same configuration was 206.58 Mbps and 55.38 Mbps. The processing capacity for packet level analysis, evaluated for \textit{P3} and \textit{CountUpDriver}, follows the same behaviour showed in Figure \ref{fig:JxtaMaxProcessing}. Additionally, it is possible to observe a convergent decreasing of relative processing capacity for all block sizes evaluated, starting in a cluster size of 14 nodes, where the relative throughput achieved by all block sizes is quite similar.

Figure \ref{fig:CountUpDriverThroughput} shows a relative processing capacity increasing with the addition of 2 nodes into a cluster with 4 nodes. For packet level analysis of 90Gb, MapReduce achieved a better processing capacity efficiency per node using 6 nodes, which provides 24 Mappers and 5 Reducers per Map and Reduce waves. With the adopted variation in number of Reducers, according to cluster size, using 5 Reducers was achieved better processing efficiency and significant reduction on Reduce time, as shown in Figure \ref{fig:CountUpDriverMaxPhases}.

Figures \ref{fig:P3MaxPhases} and \ref{fig:CountUpDriverMaxPhases} show the cumulative time per phase during a job execution.

\begin{figure}[!htb]
	\centering
	\subfigure[MapReduce Phases Times of P3]{
		\includegraphics[width=0.6\textwidth]{images/P3MaxPhasesEvaluation.eps}
		\label{fig:P3MaxPhases}
	}
	\subfigure[MapReduce Phases Times for CountUpDriver]{
		\includegraphics[width=0.6\textwidth]{images/CountUpDriverMaxPhasesEvaluation.eps}
		\label{fig:CountUpDriverMaxPhases}
	}
	\caption[MapReduce Phases time of CountUp for 90Gb]{MapReduce Phases time of CountUp for 90Gb}
	\label{fig:CountUpThroughput}
\end{figure}

The behaviour of MapReduce phases for packet level analysis  is similar to the behaviour observed for DPI; with Map time being predominant, Map and Shuffle time do not decreasing with node addition in a cluster bigger than a specific size, and Sort and Reduce phases consuming low execution time. The exception is that Shuffle phase consumes more time in packet level analysis jobs than in DPI, specially in smaller clusters.

For packet level analysis, the amount of intermediate data generated by Map functions is bigger than the amount generated through the use of MapReduce for DPI; packet level analysis generates an intermediate data for each packet evaluated, but for DPI it is necessary to evaluate more than one packet to generate intermediate data. Shuffle phase is responsible for sorting and transferring the Map outputs to the Reducers as inputs; then the amount of intermediate data generated by Map tasks and the network transfer cost, will impact on the Shuffle phase time.

Figures \ref{fig:P3MaxPhasesPercent} and \ref{fig:CountUpDriverMaxPhasesPercent} show the percentage of each phase on job completion time of \textit{P3} and \textit{CountUpDriver}, respectively.

\begin{figure}[!htb]
	\centering
	\subfigure[Phases Distribution for P3]{
		\includegraphics[width=0.6\textwidth]{images/P3MaxPhasesPercentEvaluation.eps}
		\label{fig:P3MaxPhasesPercent}
	}
	\subfigure[Phases Distribution for CountUpDriver]{
		\includegraphics[width=0.6\textwidth]{images/CountUpDriverMaxPhasesPercentEvaluation.eps}
		\label{fig:CountUpDriverMaxPhasesPercent}
	}
	\caption[MapReduce Phases Distribution for CountUp of 90Gb]{MapReduce Phases Distribution for CountUp of 90Gb}
	\label{fig:CountUpThroughput}
\end{figure}

As the behaviour observed in Figure \ref{fig:JxtaMaxPhasesPercent} and followed by these cases, Map and Shuffle phases consume more relative time than all others phases, over all factors. But, for packet level analysis, Map phase occupation decreases significantly, with node addition, only when block sizes are 32MB or 64MB, and following the completion time behaviour observed in Figures \ref{fig:P3MaxSpeedupBarLines} and \ref{fig:CountUpDriverMaxSpeedupBarLines}.

For the dataset of 30Gb of network traffic the same experiments were conducted, and the results about MapReduce phases evaluation presented a behaviour quite similar to the results already presented by 90Gb experiments, for DPI and packet level analysis.

Relevant differences were identified for speed-up, completion time and scalability evaluation, as shown by Figures \ref{fig:JxtaMinSpeedupBarLines} and \ref{fig:JxtaMinThroughput}, which exhibit the completion time and processing capacity scalability of MapReduce for DPI of 30Gb of network traffic, with variation in cluster size and block size.

\begin{figure}[!htb]
	\centering
	\subfigure[DPI Completion Time and Speed-up of MapReduce for 30Gb of a JXTA-application network traffic]{
		\includegraphics[width=0.6\textwidth]{images/JxtaSocketPerfDriverMinJobSpeedupBarLines.eps}
		\label{fig:JxtaMinSpeedupBarLines}
	}
	\subfigure[DPI Processing Capacity of 30Gb]{
		\includegraphics[width=0.6\textwidth]{images/JxtaSocketPerfDriverMinJobProcessingLines.eps}
		\label{fig:JxtaMinThroughput}
	}
	\caption[MapReduce Phases Distribution for CountUp of 90Gb]{DPI Completion Time and Processing capacity for 90Gb}
	\label{fig:JxtaMinSpeedupThroughput}
\end{figure}

The completion time of DPI of 30Gb scales significantly up to 10 nodes; then the experiment presents no more gains with node addition using block size of 128MB, and presents a low increase of completion time in cases using blocks of 32MB and 64MB. This behaviour is the same observed for job completion time for 90Gb, as shown in Figures \ref{fig:P3MaxSpeedupBarLines} and \ref{fig:CountUpDriverMaxSpeedupBarLines}, but with significant scaling just up to 10 nodes for the dataset of 30Gb, while it was achieved scaling up to 25 nodes for 90Gb.

Figure \ref{fig:JxtaMinSpeedupBarLines} shows that a completion time of 199.12 seconds was obtained with 2 nodes using block of 128MB, scaling up to 87.33 seconds with 10 nodes and the same block size, while it was achieved respectively 474.44 and 147.12 seconds by the same configuration for DPI of 90Gb, as shown in Figure \ref{fig:JxtaMaxSpeedupBarLines}.

Although the case for processing of 90Gb (Figure \ref{fig:JxtaMaxSpeedupBarLines}) had processed a dataset 3 times bigger than the case of 30Gb (Figure \ref{fig:JxtaMinSpeedupBarLines}), the completion time achieved in all cases for processing of 90Gb was smaller than 3 times the completion time for processing of 30Gb. For the cases with 2 and 10 nodes using block of 128MB, it was consumed respectively 2.38 and 1.68 times more time to process a dataset 90Gb, which is a dataset 3 times bigger than the dataset of 30Gb.

Figure \ref{fig:JxtaMinThroughput} shows the processing capacity for DPI of 30Gb. The maximum speed-up achieved for DPI of 30Gb was 7.90 times, using block of 32MB and 29 worker nodes, while it was achieved the maximum speed-up of 16.19 times for DPI of 90Gb with 28 nodes.

From these results, it is possible to conclude that the MapReduce efficiency fits better for bigger data and in some cases can be more efficient to accumulate input data to process a bigger amount of data. Therefore it is important to analyse the dataset size to be processed, and to quantify the ideal number of allocated nodes for each job, in order to avoid wasting resources.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% explicar por que MapReduce alcanca um speedup de 3.7 com dois nos. Isto ocorre por que o MapReduce ja eh programado para explorar concorrencia diponivel nos nos, alocando uma task por core, diferente da implementacao monolitica avaliada
%destacar os beneficios da avaliação das aplicacoes, verificação da distruicao de carga entre servidores, avaliar as metricas, taxa de transferência, tempo de resposta, identificar padrões nos tempos que ocorrem as requisições e respostas
\section{Discussion}
\label{sc:eval_discussion}

In this section, we discuss the measured results and evaluate their meaning, restrictions and opportunities. We also discuss possible threats to the validity of our experiment.

\subsection{Results Discussion}
\label{ssc:eval_results_discussion}

According to the processing capacity presented in our experimental results, of the evaluation of MapReduce for packet level analysis and DPI, it is possible to see that the MapReduce adoption for packet level analysis and DPI provided high processing capacity and speed-up of completion time, when compared with a solution without distributed processing, making it possible to evaluate a large amount of network traffic and extract information from distributed applications of an evaluated data center.

The block size adopted and the number of nodes allocated to data processing are important factors for obtaining an efficient job completion time and processing capacity scalability. Some benchmarks show that MapReduce performance can be improved by an optimal block size choice \citep{Jiang2010}, showing better performance with the adoption of bigger block sizes. We evaluated the impact of the block size for packet level analysis and DPI workloads; blocks with 128MB provided a better completion time for smaller clusters, but blocks with 64MB performed better in bigger clusters. Thus, in order to obtain an optimal completion time adopting bigger block size it is necessary also to evaluate the node allocation for the MapReduce job, due to the variation in block size and cluster size can cause significant impact into completion time.

The different processing capacity achieved for processing datasets of 30Gb and 90Gb highlights the efficiency of MapReduce for dealing with bigger data processing, and that can be more efficient to accumulate input data, to process a larger amount of data. Therefore, it is important to analyse the dataset size to be processed, and to quantify the ideal number of allocated nodes for each job, in order to avoid wasting resources.

The evaluation of the dataset size and the optimal number of nodes is important to understand how to schedule MapReduce jobs and resource allocation through specific Hadoop schedulers, such as \textit{Capacity Scheduler} and \textit{Fair Scheduler} \citep{Zaharia2010}, in order to avoid resource wasting with allocation of nodes that will not produce significant gains \citep{Verma20122}. Thus, the variation of processing capacity achieved in out experiments highlights the importance of evaluation of the cost of node allocation and its benefits, and the need to evaluate the ideal size of pools in the Hadoop cluster, to obtain efficiency between the cluster size allocated to process an input size, and the resource sharing of a Hadoop cluster.

The MapReduce processing capacity does not scale proportionally to node addition; in some cases there is not significant processing capacity which increases with node addition, as shown in Figure \ref{fig:JxtaMaxSpeedupBarLines}, where jobs using block size of 64MB and 128MB in clusters with more than 14 nodes for DPI of 90Gb, present no significant completion time gain with node addition.

The number of execution waves is a factor that must be evaluated \citep{Kavulya2010} when MapReduce scalability is analysed, due to the execution time decrease wis related to the number of execution waves necessary to process all input data. The number of execution waves is defined by the available slots for execution of Map and Reduce tasks; for example, if a MapReduce job is divided into 10 tasks in a cluster with 5 available slots, then it will be necessary to have 2 execution waves for all tasks be executed.

Figure \ref{fig:JxtaMinSpeedupBarLines} shows a case of DPI of 30Gb, using block size of 128MB, in which there was not a reduction of completion time with cluster size bigger than 10 nodes, because there was not a reduction in the number of execution waves. But in our experiments, cases with a reduction of execution waves also presented no significant reduction of completion time, as cases using block size of 128MB in clusters with 21 nodes or more, for DPI and packet level analysis, showed in Figure \ref{fig:JxtaMaxSpeedupBarLines}. Thus, node addition or tasks distribution must be evaluated for resources usage optimization and to avoid additional or unnecessary costs with machines and power consumption.

The comparison of completion time between \textit{CountUpDriver} and \textit{P3} shows that \textit{P3}, which splits the data into packets, performs better than \textit{CountUpDriver}, which processes a whole block without splitting. Processing a whole block as input,  the local node parallelism is limited to the number of slots per node, while in the divisible approach each split can be processed by an independent thread, increasing the possible parallelism. Because some cases require data without splitting, such as DPI and video processing cases \citep{Pereira2010}, improvements for this issue must be evaluated, considering better schedulers, data location and task assignment.

The behavioural evaluation of MapReduce phases showed that the Map phase is predominant in total execution time for packet level analysis and DPI, with Shuffle being the second most expressive phase. Shuffle can overlap Map phase, and this condition must be considered in MapReduce evaluations, specially in our case, due to the overlap of Map and Shuffle which represents more than 50\% of total time execution.

The long time of "Map and Shuffle" phase represents a long time of Shuffle tasks being executed in parallel with Map tasks, and a long time of slots allocated for Shuffle tasks that only will be concluded after all Map tasks are finished, although these Shuffle tasks can be longer than the time required to read and process the generated intermediate data. If there are slots allocated for Shuffle tasks that are only waiting for Map phase conclusion, these slots could be used for other task executions, which could accelerate the job completion time.

With the increase of cluster size and reduction of job completion time, it was observed that the Map phase showed a proportional decrease, while Shuffle phase increased with the growth of the number of nodes. With more nodes, the intermediate data generated by Map tasks is placed in more nodes, which are responsible for shuffling the data and sending them to specific Reducers, increasing the amount of remote I/O from Mappers to Reducers and the number of data sources for each Reducer. Shuffle phase may represent a bottleneck \citep{Shubin2009} for scalability and could be optimized, due to I/O restrictions \citep{Lee2012, Akram2012} and data locality issues for Reduce phase \citep{Hammoud2011}.

Information extracted from the analysed results about the performance obtained with specific cluster, block and input sizes, is important for configuring MapReduce resource allocation and specialized schedulers, such as the Fair Scheduler \citep{Zaharia2008}, which defines pool sizes and resource share for MapReduce jobs. Thus, with the information of the performance achieved with specific resources, it is possible to configure MapReduce parameters in order to obtain efficiency between the resource allocation, and the expected completion time or resource sharing \citep{Zaharia2008, Zaharia2010}.

\subsection{Possible threats to validity}
\label{ssc:eval_threats}

In this chapter we evaluated for packet level analysis, a port counter implemented with \textit{P3}. It was used a version of this implementation published in \cite{Lee2011} website\footnote{https://sites.google.com/a/networks.cnu.ac.kr/yhlee/p3}, obtained on 2012 February, when a complete binary version was available, which was used in our experiments, but this binary version is currently not available.

Part of the \textit{P3} source code was published later, but not all necessary code to compile all binary libraries necessary to evaluate the \textit{P3} implementation of a port counter. Thereby it is important to highlight that the obtained results, through our evaluation, was for the \textit{P3} version obtained on 2012 February from \cite{Lee2011} website.

It is also important to highlight that the DPI can present restrictions to evaluate encrypted messages, and that the obtained results are specifics for the input datasets, factors, levels and experiment setup used in our evaluation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Chapter Summary}
\label{sc:eval_summary}

In this chapter, we evaluated the performance of MapReduce for packet level analysis and DPI of applications traffic. We evaluated how data input, block and cluster sizes, impacts on MapReduce phases, job completion time, processing capacity scalability and on the speed-up achieved in comparison with the same algorithm executed by a non-distributed implementation.

The results show that MapReduce presents high processing capacity for dealing with massive application traffic analysis. The behaviour of MapReduce phases over variation of block size and cluster size was evaluated; we verified that packet level analysis and DPI are Map-intensive jobs, and that Map phase consumes more than 70\% of execution time, with Shuffle phase being the second predominant phase.

We showed that input size, block size and cluster size are important factors to be considered to achieve better job completion time and to explore MapReduce scalability and efficient resource allocation, due to the variation in completion time provided by the block size adopted and, in some cases, due to the processing capacity which does not increase with node addition into the cluster.

We also showed that using a whole block as input for Map functions, achieved a poorer performance than using divisible data, thereby more evaluation is necessary to understand how it can be handled and improved.

### 2017-09-19 docs/thesis/chapters/6_conclusion_future_work/conclusion.tex
compiling version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% na conclusao, evite apresentar os resultados mais fáceis de concluir. bons revisores olham o casamento do que vc escreve no abstract/intro com a conclusao. mantenha coerente. 
% em geral, resultados que sao aparentemente obvios, enfraquecem as outras contribuicoes. p.ex, "Os resultados mostraram que o MapReduce apresenta uma boa solução para utilizar computadores comuns para obter alta capacidade de processamento e lidar com a análise massiva de tráfego de rede." pode ser reformulado ou retirado da conclusao.
\chapter{Conclusion and Future Work}
\label{ch:conclusionfuturework}

\begin{quotation}[]{Lao Tzu}
The softest things in the world overcome the hardest things in the world.
\end{quotation}

Distributed systems has been adopted for building modern Internet services and cloud computing infrastructure. The detection of error causes, diagnose and reproduction of errors of distributed systems are challenges that motivate efforts to develop less intrusive mechanisms for monitoring and debugging distributed applications at runtime.

Network traffic analysis is one option for distributed systems measurement, although there are limitations on capacity to process large amounts of network traffic in short time, and on scalability to process network traffic where there is variation of resource demand.

In this dissertation we proposed an approach to perform deep inspection in distributed applications network traffic, in order to evaluate distributed systems at a data center through network traffic analysis, using commodity hardware and cloud computing services, in a minimally intrusive way. Thus we developed an approach based on MapReduce, to evaluate the behavior of a JXTA-based distributed system through DPI.

We evaluated the effectiveness of MapReduce to implement a DPI algorithm and its completion time scalability to measure a JXTA-based application, using virtual machines of a cloud computing provider. Also, was deeply evaluated the performance of MapReduce for packet-level analysis and DPI, characterizing the behavior followed by MapReduce phases, its processing capacity scalability and speed-up, over variations of input size, block size and cluster size.

\section{Conclusion}
\label{sc:conc_conclusion}

With our proposed approach, it is possible to measure the network traffic behavior of distributed applications with intensive network traffic generation, through the offline evaluation of information from the production environment of a distributed system, making it possible to use the information from the evaluated indicators, to diagnose problems and analyse performance of distributed systems.

We showed that MapReduce programming model can express algorithms for DPI, as the Algorithm \ref{alg:JXTAperfmapper}, implemented to extract application indicators from the network traffic of a JXTA-based distributed application. We analysed the completion time scalability achieved for different number of nodes in a Hadoop cluster composed of virtual machines, with different size of network traffic used as input. We showed the processing capacity and the completion time scalability achieved, and also was showed the influence of the number of nodes and the data input size in the processing capacity for DPI using virtual machines of Amazon EC2, for a selected scenario.

We evaluated the performance of MapReduce for packet level analysis and DPI of applications traffic, using commodity hardware, and showed how data input size, block size and cluster size cause relevant impacts into MapReduce phases, job completion time, processing capacity scalability and in the speedup achieved in comparison against the same execution by a non distributed implementation.

The results showed that although MapReduce presents a good processing capacity using cloud services or commodity computers for dealing with massive application traffic analysis, but it is necessary to evaluate the behaviour of MapReduce to process specifics data type, in order to understand its relation with the available resources and the configuration of MapReduce parameters, and to obtain an optimal performance for specific environments.

We showed that MapReduce processing capacity scalability is not proportional to number of allocated nodes, and the relative processing capacity decreases with node addition. We showed that input size, block size and cluster size are important factors to be considered to achieve better job completion time and to explore MapReduce scalability, due to the observed variation in completion time provided by different block size adopted. Also, in some cases, the processing capacity does not scale with node addition into the cluster, what highlights the importance of allocating resources according with the workload and input data, in order to avoid wasting resources.

We verified that packet level analysis and DPI are Map-intensive jobs, due to Map phase consumes more than 70\% of the total job completion time, and shuffle phase is the second predominant phase. We also showed that using whole block as input for Map functions, it was achieved a poor completion time than the approach which splits the block into records.

\section{Contributions}
\label{sc:conc_contributions}

We attempt to analyse the processing capacity problem of measurement of distributed systems through network traffic analysis, the results of the work presented in this dissertation provide the contributions below:

\begin{enumerate}
	\item We proposed \textbf{an approach to implements DPI algorithms through MapReduce}, using whole blocks as input for Map functions. It was \textbf{shown the effectiveness of MapReduce for a DPI algorithm} to extract indicators from a distributed application traffic, also it was \textbf{shown the completion time scalability of MapReduce for DPI}, using virtual machines of a cloud provider;
	\item We developed \textit{JNetPCAP-JXTA} \citep{jnetpcapjxta}, an open source \textbf{parser to extract JXTA messages from network traffic traces};
	\item We developed \textit{Hadoop-Analyzer} \citep{Hadoopanalyzer}, an open source \textbf{tool to extract indicators from Hadoop logs and generate graphs} of specified metrics.
	\item We \textbf{characterized the behavior followed by MapReduce phases for packet level analysis and DPI}, showing that this kind of job is intense in Map phase and highlighting points that can be improved;
	\item We \textbf{described the processing capacity scalability of MapReduce for packet level analysis and DPI}, evaluating the \textbf{impact caused by variations in input size, cluster size and block size};
	\item We \textbf{showed the speed-up obtained with MapReduce for DPI}, with variations in input size, cluster size and block size;
	\item We \textbf{published two papers} reporting our results, as follows:
	\begin{enumerate}
		\item Vieira, T., Soares, P., Machado, M., Assad, R., and Garcia, V. \textit{Evaluating Performance of Distributed Systems with MapReduce and Network Traffic Analysis}. In ICSEA 2012, The Seventh International Conference on Software Engineering Advances. Xpert Publishing Services.
		\item Vieira, T., Soares, P., Machado, M., Assad, R., and Garcia, V. \textit{Measuring Distributed Applications Through MapReduce and Traffic Analysis}. In Parallel and Distributed Systems (ICPADS), 2012 IEEE 18th International Conference on, pages 704 - 705.	
	\end{enumerate}
\end{enumerate}

\subsection{Lessons Learned}
\label{sc:chaptersummary}

The contributions cited are of scientific and academic scope, with implementations and evaluations little explored in the literature. However, with the development of this work, some important lessons were learned.

During this research, different approaches for evaluating distributed systems of cloud computing providers were studied. In this period, we could see the importance of the performance evaluation in a cloud computing environment, and the recent efforts to diagnose and evaluate system at production environment of a data center. Also, the growth of the Internet and resource utilization make necessary solutions to be able to evaluate large amounts of data in short time, with low performance degradation of the evaluated system.

MapReduce has grown as a general purpose solution for big data processing, but it is not a solution for all kind of problems, and its performance is dependent of several parameters. Some researches has been done in order to improve MapReduce performance, through analytical modelling, simulation and measurement, but the most relevant contributions in this direction was guided by realistic workload evaluations, from large MapReduce clusters.

We learned that although the facilities provided by the MapReduce for distributed processing, its performance is influenced by the environment, network topology, workload, data type and by several specific parameter configurations. Therefore, an evaluation of the MapReduce behavior using data of a realistic environment will provide more accurate and wide results, while in controlled experiments the results are more restricted and limited to the evaluated metrics and factors.

\section{Future Work}
\label{sc:conc_futurework}

Because of time constraints imposed on the master degree, this dissertation addresses some problems, but some problems are still open and others are emerging from current results. Thus, the following issues should be investigated as future work:

\begin{itemize}
	\item \textbf{Evaluating of all components of the proposed approach}. This dissertation evaluated the \textit{JNetPCAP-JXTA}, the \textit{AppAnalyzer} and its implementation to evaluate a JXTA-based distributed application, it is necessary to evaluate the \textit{SnifferServer}, \textit{Manager} and the whole system working together, analysing their impact into the measured distributed system and the scalability achieved;	
	\item Development of a \textbf{technique for the efficient evaluation of distributed systems through information extracted from network traffic}. This dissertation addressed the problem of processing capacity for measuring distributed systems through network traffic analysis, but it is necessary an efficient approach to diagnose problems of distributed systems, using information of flows, connections, throughput and response time obtained from network traffic analysis;
	\item Development of a \textbf{analytic model and simulations}, using information of MapReduce behavior for network traffic analysis, measured by this dissertation, to reproduce its characteristics and enable the evaluation and prediction of some cases of MapReduce for network traffic analysis;
\end{itemize}

### 2017-09-19 docs/thesis/resumo.tex
compiling version

Sistemas distribuídos têm sido utilizados na construção de modernos serviços da Internet e infraestrutura de computação em núvem, com o intuito de obter serviços com alto desempenho, escalabilidade e confiabilidade. Os acordos de níves de serviço adotados pela computação na núvem requerem um reduzido tempo para identificar, diagnosticar e solucionar problemas em sua infraestrutura, de modo a evitar que problemas gerem impactos negativos na qualidade dos serviços prestados aos seus clientes. Então, a detecção de causas de erros, diagnóstico e reprodução de erros provenientes de sistemas distribuídos são desafios que motivam esforços para o desenvolvimento de mecanismos menos intrusivos e mais eficientes, para o monitoramento e depuração de aplicações distribuídas em tempo de execução.

A análise de tráfego de rede é uma opção para a medição de sistemas distribuídos, embora haja limitações na capacidade de processar grande quantidade de tráfego de rede em curto tempo, e na escalabilidade para processar tráfego de rede sob variação de demanda de recursos.

O objetivo desta dissertação é analisar o problema da capacidade de processamento para mensurar sistemas distribuídos através da análise de  tráfego de rede, com o intuito de avaliar o desempenho de sistemas distribuídos de um data center, usando hardware não especializado e serviços de computação em núvem, de uma forma minimamente intrusiva.

Nós propusemos uma nova abordagem baseada em MapReduce para profundamente inspecionar tráfego de rede de aplicações distribuídas, com o objetivo de avaliar o desempenho de sistemas distribuídos em tempo de execução, usando hardware não especializado. Nesta dissertação nós avaliamos a eficácia do MapReduce para um algoritimo de avaliação profunda de pacotes, sua capacidade de processamento, o ganho no tempo de conclusão de tarefas, a escalabilidade na capacidade de processamento, e o comportamento seguido pelas fases do MapReduce, quando aplicado à inspeção profunda de pacotes, para extrair indicadores de aplicações distribuídas.

\begin{keywords}
Medição de Aplicações Distribuídas, Depuração, MapReduce, Análise de Tráfego de Rede, Análise em Nível de Pacotes, Análise Profunda de Pacotes
\end{keywords}

### 2017-09-19 docs/thesis/thesis.tex
compiling version

@@
-\documentclass[a4paper,12pt,openright,titlepage,oneside]{book}
+\documentclass[en, oneside, onehalfspacing]{risethesis}
 
-%\usepackage[english,brazil]{babel} 
-%Define regra de gramática para separar síbalas {babel}
-%e altera os títulos (como chapters, sections, references) para português
-\usepackage[brazil]{babel}
-
-% Define uso de caracteres acentuados no PDF gerado
-% e permite copiar corretamente o texto do PDF.
-\usepackage[T1]{fontenc}
-\usepackage[utf8]{inputenc}
-
-% Adição do pacote do template da UnB
-\usepackage{template-FT-UnB/ft2unb}
-
-\DeclareGraphicsExtensions{.jpg,.pdf,.mps,.png,.gif, .eps}
-\graphicspath{imagens} %define diretório de imagens
-
-%Arquivo com lista com hifenização correta de algumas palavras.
-%Defina novas palavras no arquivo a medida que verificar que a hifenização automática
-%etá errada para tais palavras.
-\input{hifenizacao}
-
-% ALTERE OS VALORES DENTRO DAS CHAVES DOS COMANDOS NESTA SEÇÃO PARA INCLUIR OS SEUS DADOS E DADOS DA SUA
-% DISSERTAÇÃO DE MESTRADO OU TESE DE DOUTORADO
-% -----------------------------------------------------------------------------------------------------
-
-%\onehalfspacing
-\title{DISCRIMINATIVE SENSING BASED ON SIGNAL PROCESSING}
-\author{THIAGO PEREIRA DE BRITO VIEIRA}
-\date{2017-06-30} %data da defesa
-
-\grau{Ph.D} %Mestre ou Doutor
-\area{Eletrical Engennering} %Nome do curso
-\siglaarea{ENE} %Sigla do departamento
-\tipodemonografia{Thesis} %Dissertação ou Tese
-\programa{Mestrado} %Mestrado ou Doutorado
-\autorendereco{THIAGO PEREIRA DE BRITO VIEIRA.} %Endereço do autor da dissertação/tese
-\totalpgs{147} %total de páginas atualmente na sua dissertação
-\dia{16} %dia da defesa
-\mes{Junho} %mês da defesa
-\ano{2011} %ano da defesa
-\numpublicacao{xxx/AAAA} %número da publicação (após a defesa, tal número deve ser obtido na secretaria)
-
-%PPGENE.DM  = Programa de Pós Graduação em ENgenharia Elétrica.Dissertação de Mestrado
-%PPGENE.TD  = Programa de Pós Graduação em ENgenharia Elétrica.Tese de Doutorado
-\siglapublicacao{PPGENE.DM}
-
-\titulolinhai{DISCRIMINATIVE SENSING}
-\titulolinhaii{BASED ON SIGNAL PROCESSING}
-\titulolinhaiv{}
-
-\autori{SEU NOME COMPLETO AQUI}
-%Caso seu nome não caiba em uma única linha, divida ele nos comandos abaixo
-%\autorii{} 
-%\autoriii{}
-
-\membrodabancai{Prof. Dr. NOME DO SEU ORIENTADOR, ENE/UnB}
-\membrodabancaifuncao{Orientador}
-\membrodabancaii{Prof. Fulano de Tal 2, ENE/UnB}
-\membrodabancaiifuncao{Examinador interno}
-\membrodabancaiii{Prof. Fulano de Tal 3, ENE/UnB}
-\membrodabancaiiifuncao{Examinador interno}
-\membrodabancaiv{Prof. Fulano de Tal 4, EESC/USP}
-\membrodabancaivfuncao{Examinador externo}
-\membrodabancav{}
-\membrodabancavfuncao{}
-% -----------------------------------------------------------------------------------------------------
-
-%line-numbers, inputencoding=utf8/latin1
-%Define o estilo para listagens de código fonte
-\lstset{
-  numbers=left, %numeração de linhas à esquerda
+\usepackage{natbib}
+\usepackage{babel}
+\usepackage{supertabular}
+\usepackage{fancybox}
+\usepackage{acronym}
+\usepackage{array}
+\usepackage{booktabs}
+\usepackage{graphicx}
+\usepackage{rotating}
+\usepackage{tabularx}
+\usepackage{colortbl}
+\usepackage{multirow}
+\usepackage{setspace}
+\usepackage{placeins}
+\usepackage{longtable}
+\usepackage{mathtools}
+\usepackage{listings}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage{epstopdf}
+\usepackage{subfigure}
+\lstset{numbers=left,
 	stepnumber=1,
@@
 	extendedchars=true,
-  frame=none,
+	breaklines=true,
+	frame=tb,
 	basicstyle=\footnotesize,
 	stringstyle=\ttfamily,
-  showstringspaces=false,
-  %language=Java, %deve ser definida na inclusão de cada trecho de código, pois podem existir linguagens diferentes em exemplos diferentes
-  breaklines=true,
-  breakautoindent=true,
-  %estilos de comentário de uma e várias linhas
-  morecomment=[l]{--}, morecomment=[s]{/*}{*/}, morecomment=[s]{<!--}{-->}, morecomment=[s]{--[[}{--]]}
+	showstringspaces=false
 }
+\renewcommand{\lstlistingname}{Code}
+\renewcommand{\lstlistlistingname}{Lista de Listagens}
+\definecolor{lightgray}{gray}{0.8}
+\usepackage[linkcolor=blue, citecolor=blue, urlcolor=blue, colorlinks, pdfpagelabels, pdftitle={ycc-msc}, pdfauthor={Thiago Pereira de Brito Vieira}]{hyperref}
 
-% Adição de metadados no PDF (propriedades do documento PDF)
-\makeatletter
-	 \hypersetup{
-		 pdftoolbar=true,        % show Acrobat’s toolbar?
-		 pdfmenubar=true,        % show Acrobat’s menu?
-		 pdffitwindow=false,     % window fit to page when opened
-		 pdfstartview={FitH},    % fits the width of the page to the window	 
-		 pdftitle={\@title},
-		 pdfauthor={\@author},
-		 pdfsubject={\tipodemonografianome \ de\ \programastr \ em\ \areastr},   % subject of the document
-		 pdfcreationdate={\pdfdate}
-	 }
-\makeatother
+\address{BRASÍLIA}
 
-\makeindex
-\makenomenclature %Necessário para gerar lista de siglas
+\universitypt{Universidade de Brasília}
+\universityen{Universidade de Brasília}
+
+\departmentpt{Departamento de Engenharia Elétrica - ENE/FT}
+\departmenten{Departamento de Engenharia Elétrica - ENE/FT}
+
+\programpt{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}
+\programen{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}
+
+\majorfieldpt{Engenharia Elétrica}
+\majorfielden{Electrical Engineering}
+
+\title{Discriminative Sensing Based on Signal Processing}
+\date{June/2018}
+
+\author{Thiago Pereira de Brito Vieira}
+\adviser{Adviser}
+\coadviser{coadviser}
 
@@
 
-	\pdfbookmark[0]{Agradecimentos}{agradecimentos}
-	\include{agradecimentos}
-	\include{resumo}
+\frontmatter
+\frontpage
+\presentationpage
 
-	\pdfbookmark[0]{Sumário}{sumario}
-	\sumario
+\begin{dedicatory} Ficha Catalográfica\end{dedicatory}
 
-	\pdfbookmark[0]{Lista de Figuras}{listafiguras}
-	\listadefiguras
+\begin{dedicatory} Signatures\end{dedicatory}
 
-	\pdfbookmark[0]{Lista de Tabelas}{listatabelas}
-	\listadetabelas
+\begin{dedicatory} Dedicatory.\end{dedicatory}
 
-	\pdfbookmark[0]{Lista de Códigos Fonte}{listacodigosfonte}
-	\listadecodigosfonte
+\agradecimentos
+First and foremost, I would like to thank God for giving me the life, strength, knowledge and opportunity to undertake this research study and to ..
 
-	\renewcommand{\nomname}{LISTA DE TERMOS E SIGLAS} %Define um caption à lista de siglas
-	%Inclui a lista de siglas 
-	\pdfbookmark[0]{Lista de Termos e Siglas}{nomenclatura}
-	\printnomenclature[2.5cm]
+(...)
 
-	\mainmatter %Inicia a numeracao normal cardinal
-	\setcounter{page}{1} \pagenumbering{arabic} \pagestyle{plain}
+Thank You!!!
 
-	\include{introducao}
-	\include{capitulo1}
-	\include{capitulo2} 
-	\include{conclusao}
+\begin{epigraph}[]{Confucius}
+Wherever you go, go with all your heart.
+\end{epigraph}
 
-	%\bibliographystyle{abnt-num} % estilo bibliográfico ABNT numérico
-	%\bibliographystyle{abnt-alf} % estilo bibliográfico ABNT alfabético
-	\bibliographystyle{sbc}  % estilo bibliográfico da Sociedade Brasileira de Computação (SBC)
+\resumo
+\input{resumo}
 
-	%\renewcommand{\bibname}{REFERÊNCIAS BIBLIOGRÁFICAS} %Define o Caption da seção de bibliografia
-	%\addcontentsline{toc}{chapter}{REFERÊNCIAS BIBLIOGRÁFICAS}
+\abstract
+\input{abstract}
 
-	%não pode ter espaço entre os nomes dos arquivos bib
-	\bibliography{referencias}
+\tableofcontents
 
-	\include{apendice}
+\makeatletter
+\renewcommand{\@thesubfigure}{\thesubfigure:\hskip\subfiglabelskip}
+\makeatother
+\setcounter{lofdepth}{2}
+
+\listoffigures
+\listoftables
+
+\input{acronyms}
+\mainmatter
+
+\include{chapters/1_introduction/introduction}
+%\include{chapters/2_pca_svm_cfa/pca_svm_cfa}
+%\include{chapters/3_eigen_similarity/eigen_similarity}
+%\include{chapters/4_mobile/mobile}
+\include{chapters/6_conclusion_future_work/conclusion}
+
+\bibliographystyle{natbib}
+\addcontentsline{toc}{chapter}{Bibliography}
+\bibliography{references}
+
+\clearpage
+\addappheadtotoc
+
 \end{document}

### 2017-09-20 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
compiling version with all chapters

@@
 ,
-onde  é a matriz de autovetores e  é a matriz cuja diagonal principal contém os autovalores. Para cada autovalor λi da matriz  existe um autovetor unitário e ortognal vi.
+% onde  é a matriz de autovetores e  é a matriz cuja diagonal principal contém os autovalores. Para cada autovalor λi da matriz  existe um autovetor unitário e ortognal vi.
 vi = 
 A i-nésima coluna da matriz  é o autovetor normalizado  correspondente ao autovalor . Desta forma, define-se i-ésima componente principal (Yi) como sendo
-Yi= λ i vi
+% Yi= λ i vi
 Salienta-se que a ACP é baseada na EVD já que parte dos autovalores e dos autovetores representam as componentes principais (CP). 
@@
 Com o aumento da dimensão, aumenta a probabilidade desses dados poderem ser linearmente separáveis.
-Uma função kernel K(xi, xj) = φ(xi) T φ(xj) pode ser usada para treinar a MVS. Uma MVS linear tem φ(x) = x, assim uma função kernel linear pode ser representada por K(xi, xj) = xTi xj.
+% Uma função kernel K(xi, xj) = φ(xi) T φ(xj) pode ser usada para treinar a MVS. Uma MVS linear tem φ(x) = x, assim uma função kernel linear pode ser representada por K(xi, xj) = xTi xj.
 Foi adotada a estratégia "um contra um" (Knerr, Personnaz e Dreyfus, 1990), para a utilização de MVS para a classificação de muitas classes.
@@
 
- Quadro 7 – Autovalores
-Ordem dos autovalores
-Autovalores
-\% da variância explicada
-Autovalores acumulados
-\% da variância explicada acumulada
-1. 
-24,88966
-12,38292
-24,8897
-12,3829
-2. 
-10,33305
-5,14082
-35,2227
-17,5237
-3. 
-7,49873
-3,73071
-42,7214
-21,2545
-4. 
-5,27875
-2,62625
-48,0002
-23,8807
-5. 
-4,78008
-2,37815
-52,7803
-26,2588
-6. 
-4,49112
-2,23439
-57,2714
-28,4932
-7. 
-3,73598
-1,85870
-61,0074
-30,3519
-8. 
-3,59164
-1,78689
-64,5990
-32,1388
-9. 
-3,19618
-1,59014
-67,7952
-33,7290
-10. 
-3,11010
-1,54731
-70,9053
-35,2763
-11. 
-2,66175
-1,32426
-73,5671
-36,6005
-12. 
-2,54482
-1,26608
-76,1119
-37,8666
-13. 
-2,51698
-1,25223
-78,6289
-39,1188
-14. 
-2,40926
-1,19864
-81,0381
-40,3175
-15. 
-2,35614
-1,17221
-83,3943
-41,4897
-16. 
-2,30653
-1,14753
-85,7008
-42,6372
-17. 
-2,25284
-1,12081
-87,9536
-43,7580
-18. 
-2,11772
-1,05359
-90,0714
-44,8116
-19. 
-2,06312
-1,02643
-92,1345
-45,8380
-20. 
-2,01891
-1,00443
-94,1534
-46,8425
-21. 
-1,99258
-0,99133
-96,1460
-47,8338
-22. 
-1,97266
-0,98142
-98,1186
-48,8152
-23. 
-1,95377
-0,97203
-100,0724
-49,7873
-24. 
-1,86320
-0,92696
-101,9356
-50,7142
-25. 
-1,83124
-0,91107
-103,7668
-51,6253
-26. 
-1,80163
-0,89633
-105,5685
-52,5216
-27. 
-1,76688
-0,87905
-107,3354
-53,4007
-28. 
-1,70876
-0,85013
-109,0441
-54,2508
-29. 
-1,69776
-0,84465
-110,7419
-55,0955
-30. 
-1,66091
-0,82633
-112,4028
-55,9218
-31. 
-1,62693
-0,80942
-114,0297
-56,7312
-32. 
-1,60039
-0,79621
-115,6301
-57,5274
-33. 
-1,54662
-0,76946
-117,1767
-58,2969
-34. 
-1,51306
-0,75277
-118,6898
-59,0496
-35. 
-1,51005
-0,75127
-120,1998
-59,8009
-36. 
-1,47437
-0,73352
-121,6742
-60,5344
-37. 
-1,43597
-0,71441
-123,1102
-61,2488
-38. 
-1,42247
-0,70770
-124,5326
-61,9565
-39. 
-1,40464
-0,69882
-125,9373
-62,6554
-40. 
-1,35176
-0,67252
-127,2890
-63,3279
-41. 
-1,33162
-0,66250
-128,6207
-63,9904
-42. 
-1,32118
-0,65731
-129,9418
-64,6477
-43. 
-1,29010
-0,64184
-131,2320
-65,2895
-44. 
-1,26013
-0,62693
-132,4921
-65,9165
-45. 
-1,25589
-0,62482
-133,7480
-66,5413
-46. 
-1,22165
-0,60779
-134,9696
-67,1491
-47. 
-1,20095
-0,59749
-136,1706
-67,7466
-48. 
-1,18070
-0,58741
-137,3513
-68,3340
-49. 
-1,17744
-0,58579
-138,5287
-68,9198
-50. 
-1,16657
-0,58038
-139,6953
-69,5001
-51. 
-1,14340
-0,56886
-140,8387
-70,0690
+ Quadro 7 – Autovalores  Ordem dos autovalores  Autovalores da variância explicada Autovalores acumulados da variância explicada acumulada
 Fonte: Dados da Pesquisa
+
 Depois da extração dos autovalores e percentual da variância explicada, é decidida a quantidade de fatores a serem retirados para análise.
@@
 Tal subdimensão verifica a participação da auditoria interna das instituições para o preenchimento do questionário.
-Já no grupo III, que não tem FCS, a variável Q53_Relev é a selecionada; precedidas do grupo IV que também não apresenta FCS.
+% Já no grupo III, que não tem FCS, a variável Q53_Relev é a selecionada; precedidas do grupo IV que também não apresenta FCS.
 Ressalta-se que ambas as questões apuram a relevância de determinado item.
@@
 Quadro 8 – Variáveis com alta significância CP1 a CP20
-CP
-Dimensão
-Var.1
-Sig.
-Var. 2
-Sig.
-FCS
-CP 01
-D1
-Q16b
-0,8108
-Q16c
-0,7496
-não
-CP 01
-D1
-Q16d
-0,7923
-Q16e
-0,8469
-não
-CP 01
-D1
-Q16f
-0,7998
-
-não
-CP 02
-D5
-Q51_Relev
-0,8941
-Q53_Relev
-
-0,8176
-não
-CP 03
-D2
-Q23a
--0,8833
-Q23b
-0,9009
-não/sim
-CP 04
-D5
-Q51l
-0,7925
- Q51l1
-0,8254
-sim/não
-CP 05
-D1
-Q13a
-0,8842
-
-não
-CP 06
-D5
-Q57h
-0,8285
-
-sim
-CP 07
-D1
-Q11f2
-0,8362
-
-não
-CP 08
-D5
-Q53e
-0,8684
-
-sim
-CP 09
-D7
-Q71
-0,9111
-
-não
-CP 10
-D4
-Q45b
--0,8960
-
-não
-CP 11
-D1
-Q14c
--0,8605
-
-não
-CP 12
-D5
-Q510d2
-0,8322
-
-não
-CP 13
-D6
-Q64h
-0,9240
-
-não
-CP 14
-D8
-Q83a
-0,8822
-
-não
-CP 15
-D1
-Q15d
-0,8738
-Q15g 
--0,7003 
-não
-CP 16
-D3
-Q31b
-0,9098
-Q31e 
--0,9070 
-não
-CP 17
-D2
-Q23i
-0,9286
-
-não
-CP 18
-D1
-Q11c1
-0,9204
-
-não
-CP 19
-D5
-Q510e1
-0,9362
-
-não
-CP 20
-D4
-Q45j
-0,9536
-
-não
-Total
-29
-4
 Fonte: Dados da pesquisa
@@
 Quadro 9 – Variáveis com alta significância CP 21 a CP 51
-CP
-Dimensão
-Var.1
-Sig.
-Var. 2
-Sig.
-FCS
-CP 21
-D1
-Q12e 
-0,7210
- Q12e1 
-0,8314
-sim/não
-CP 22
-D5
- Q55a
-0,8256
-
-não
-CP 23
-D8
- Q83b2
-0,8816
-
-não
-CP 24
-D7
-Q72a3 
-0,9220
-
-não
-CP 25
-D4
- Q45g
--0,9334
-
-não
-CP 26
-D5
-Q510b1
-0,8942
-
-não
-CP 27
-D1
-Q12e3 
-0,9294
-
-não
-CP 28
-D1
-Q12b 
-0,8415
- Q12c
-0,8934
-sim/sim
-CP 29
-D1
- Q12e2
-0,9621
-
-não
-CP 30
-D1
- Q11_Relev
-0,8400
-
-não
-CP 31
-D2
- Q24
-0,9425
-
-sim
-CP 32
-D5
- Q510b2
-0,9024
-
-não
-CP 33
-D5
- Q53a1
-0,8840
-
-não
-CP 34
-D1
- Q13e
-0,8986
-
-não
-CP 35
-D4
- Q45d
-0,9487
-
-não
-CP 36
-D1
-Q16_relev 
-0,8364
-
-não
-CP 37
-D5
- Q51h
-0,8175
-
-sim
-CP 38
-D5
- Q53d
-0,8864
-
-não
-CP 39
-D1
- Q11c2
-0,9258
-
-não
-CP 40
-D1
-Q11a 
-0,8674
-
-não
-CP 41
-D1
- Q15f
-0,9201
-
-não
-CP 42
-D4
- Q45f
-0,9450
-
-não
-CP 43
-D5
- Q53f
-0,7871
-
-não
-CP 44
-D4
-Q45h 
--0,9612
-
-não
-CP 45
-D7
- Q72a4
-0,9401
-
-não
-CP 46
-D3
- Q31d
-0,8714
-
-sim
-CP 47
-D4
- Q45i
-0,9462
-
-não
-CP 48
-D5
- Q51e
-0,8432
-
-sim
-CP 49
-D4
- Q45e
-0,9406
-
-não
-CP 50
-D1
- Q12a1
-0,8635
-
-sim
-CP 51
-D8
- Q83b5
-0,9237
-
-não
-Total
-33
-8
 Fonte: Dados da pesquisa
@@
 
-Instituição
-Posição
-iGovTI
-
-33
-5
-0,7954
-
-35
-3
-0,7984
-
-36
-11
-0,7371
-
-39
-2
-0,8226
-
-44
-19
-0,7244
-
-57
-1
-0,88
-
-63
-18
-0,7247
-
-77
-4
-0,7963
-
-80
-12
-0,736
-
-92
-10
-0,7615
-
-122
-8
-0,7723
-
-142
-21
-0,7131
-
-158
-6
-0,7944
-
-186
-7
-0,7913
-
-188
-17
-0,7269
-
-197
-40
-0,6403
-
-201
-13
-0,7356
-
-205
-9
-0,7687
-
-311
-22
-0,7107
-
-346
-75
-0,5644
-
-347
-16
-0,7322
-
-349
-39
-0,6427
-
-351
-28
-0,6817
 Gráfico 5 – Destaque de corte nas instituições 

### 2017-09-20 docs/thesis/chapters/3_eigen_similarity/eigen_similarity.tex
compiling version with all chapters

--- 2017-09-19 docs/thesis/chapters/3_eigen_similarity/capitulo3.tex
@@
 \chapter{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
-\label{chapter3}
-
-\section{Introduction}
-\label{sec:introduction}
+\label{ch:evaluation}
+
+\begin{quotation}[]{Lao Tzu}
+All difficult things have their origin in that which is easy, and great things in that which is small. 
+\end{quotation}
 
@@
 
-In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
+% In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
@@
 \begin{algorithm}[h!]
-	\caption{Detection of Time Frames Under Attack}
-  	\label{alg:alg01}
-	\begin{algorithmic}[1]
-		\show\LOOP
-	    \ALOOP {$f = 1$ until $f == \hat{d}$} 
-	    	    \STATE $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
-    	    	    \ALOOP {$i = 1$ until $i == Q$} 
-				\IF {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
-				    \STATE $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
-				\ENDIF
-        		\ENDALOOP
-	    		\STATE $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
-    		\ENDALOOP
-	\end{algorithmic}
+	% \caption{Detection of Time Frames Under Attack}
+ %  	\label{alg:alg01}
+	% \begin{algorithmic}[1]
+	% 	\show\LOOP
+	%     \ALOOP {$f = 1$ until $f == \hat{d}$} 
+	%     	    \STATE $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
+ %    	    	    \ALOOP {$i = 1$ until $i == Q$} 
+	% 			\IF {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
+	% 			    \STATE $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
+	% 			\ENDIF
+ %        		\ENDALOOP
+	%     		\STATE $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
+ %   		\ENDALOOP
+	% \end{algorithmic}
 \end{algorithm}
@@
 Additionally, future research can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.
-
-\appendix
-
-\section{Model Order Selection (MOS)}
-\label{sec:mos}
-The model order selection is a key point in many digital signal processing applications, including radar, sonar, communications, channel modeling, medical imaging, among others.
-MOS allows analysis of reduced data set, through separating noise components of the main components, for example.
-Moreover, the model order is crucial for many parameter estimation techniques \cite{da2009comparison}, since the amount of parameters to be estimated depends on the model order.
-
-The model selection procedure chooses the ``best'' model of a finite set of models, according to some criterias \cite{rajan1997model}.
-Therefore, given some data set, it is chosen a model which was evaluated as the best model to describe the specified data set.
-
-The state of the art regarding estimation techniques of model order based on eigenvalues includes: Akaike's Information Theoretic Criterion - AIC \cite{akaike1974new,wax1985detection}; Minimum Description Length - MDL \cite{barron1998minimum,wax1985detection}; Efficient Detection Criterion - EDC \cite{zhao1986detection}; Stein's Unbiased Risk Estimator - SURE \cite{ulfarsson2008rank}; RADOI \cite{radoi2004new} and Exponential Fitting Test - EFT \cite{grouffaud1996some,quinlan2006model,david2011blind}.
-
-In AIC, MDL and EDC techniques, the information criterion is a function of the geometric mean \emph{g(k)} and the arithmetic mean \emph{a(k)} relating to smaller \emph{k} eigenvalues, where \emph{k} is a candidate value for the model order \emph{d} \cite{da2009comparison}.
-
-Basically, the difference between the AIC, MDL and EDC schemes is the penalty function \emph{p(k, N, α)}, so these techniques can be written in general as \cite{da2009comparison}:
-
-\begin{equation}\label{eq:eq1}
-\hat{d} = \argmin\limits_k \hspace{1 mm} J(k), 
-\end{equation}
-
-where
-
-\begin{equation}\label{eq:eq2}
-J(k) = -N(\alpha - k) \hspace{1 mm} {\rm log} \hspace{1 mm} (g(k)/a(k)) + p(k,N,\alpha),
-\end{equation}
-
-where $\hat{d}$ is an estimate \emph{d} of the model order, \emph{N} is the number of samples, \emph{α = M} and means the number of variables of the problem, and 0 ≤ \emph{k} ≤ min[\emph{M}, \emph{N}].
-Penalty functions for AIC, MDL and EDC are given by the Table \ref{tab:tab2}.
-%% ???
-
-\begin{table}[h!]
-  \centering
-  \caption{Penalty functions for the schemes AIC, MDL and EDC}
-  \label{tab:tab2}
-  \begin{tabular}{*2c}
-	\toprule
-	\textbf{Scheme} &  \textbf{Penalty function} \\
-	\textbf{} &  $p(k,N,\alpha)$ \\
-	\midrule
-    AIC	& $k(2\alpha - k)$ \\
-    MDL	& $0.5k(2\alpha - k) \hspace{1 mm} {\rm log} (N)$ \\
-    EDC	& $0.5k(2\alpha - k)\sqrt{N\hspace{1 mm}\rm ln(\rm ln N)}$ \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The Exponential Fitting Test (EFT) can effectively be used in cases where the number of samples \emph{N} is small.
-This technique is based on observations of data contaminated only with white noise, where the profile of eigenvalues can be approximated by an exponential decaying \cite{grouffaud1996some}.
-
-Given $\lambda_i$ be the i-th eigenvalue, the exponential model can be expressed by:
-\begin{equation}\label{eq:eq3}
-E\{\lambda_i\} = E\{\lambda_1\} \cdot q(\alpha,\beta)^{i-1},
-\end{equation}
-
-where $E\{\cdot\} $ is the expectation operator, and it is considered that the eigenvalues are ordered in the that $\lambda_1$ represents the largest eigenvalue. The term \emph{q(α, β)} is defined as:
-
-\begin{equation}\label{eq:eq4}
-q(\alpha,\beta) = \exp\left\{-\sqrt{\frac{30}{\alpha^2 + 2} - \sqrt{\frac{900}{(\alpha^2 + 2)^2} - \frac{720\alpha}{\beta(\alpha^4 + \alpha^2 - 2)}}} \right\},
-\end{equation}
-
-where $0 < q(\alpha,\beta) < 1$. According to \cite{quinlan2006model}, if $M \leq N$, then $\beta = N$.

### 2017-09-20 docs/thesis/chapters/5_tensor_dl/tensor_dl.tex
compiling version with all chapters

\chapter{Tensor-Based Discriminative Dictionary Learning}
\label{ch:tensor_dl}

Tensor-Based Discriminative Dictionary Learning for Fraud Detection

### 2017-09-20 docs/thesis/thesis.tex
compiling version with all chapters

@@
+% class
 \documentclass[en, oneside, onehalfspacing]{risethesis}
 
-\usepackage{natbib}
+% packages and configurations
+\usepackage{enumitem}
+\usepackage{lineno, hyperref}
+\usepackage[numbers]{natbib}
 \usepackage{babel}
@@
 \usepackage{tabularx}
-\usepackage{colortbl}
+\usepackage{color, colortbl}
 \usepackage{multirow}
+\usepackage{hhline}
 \usepackage{setspace}
@@
 \usepackage{longtable}
+\usepackage{amsthm,amsmath}
 \usepackage{mathtools}
+\usepackage{algorithm}% http://ctan.org/pkg/algorithms
+\usepackage{algpseudocode}
 \usepackage{listings}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
 \usepackage{epstopdf}
 \usepackage{subfigure}
+\usepackage{courier}
+\usepackage{amsfonts}
+\usepackage{morefloats}
+\usepackage{lipsum}
+\usepackage{cleveref}
+\usepackage{hyperref}
+
+\newtheorem{thm}{Scenario}
+\hypersetup{breaklinks=true}
+\urlstyle{same}
+\Urlmuskip=0mu plus 1mu\relax
 \lstset{numbers=left, 
@@
 }
+
+\definecolor{lightgray}{gray}{0.8}
+\definecolor{Gray}{gray}{0.9}
 \renewcommand{\lstlistingname}{Code}
 \renewcommand{\lstlistlistingname}{Lista de Listagens}
-\definecolor{lightgray}{gray}{0.8}
-\usepackage[linkcolor=blue, citecolor=blue, urlcolor=blue, colorlinks, pdfpagelabels, pdftitle={ycc-msc}, pdfauthor={Thiago Pereira de Brito Vieira}]{hyperref}
 
+\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
+\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
+\DeclareMathOperator*{\argmin}{arg\,min}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\newcommand*\justify{%
+	\fontdimen2\font=0.4em% interword space
+	\fontdimen3\font=0.2em% interword stretch
+	\fontdimen4\font=0.1em% interword shrink
+	\fontdimen7\font=0.1em% extra space
+	%\hyphenchar\font=`\-% allowing hyphenation
+}
+
+% Algorithmic modifications
+\makeatletter
+\newcommand{\ALOOP}[1]{\ALC@it\algorithmicloop\ #1%
+  \begin{ALC@loop}}
+\newcommand{\ENDALOOP}{\end{ALC@loop}\ALC@it\algorithmicendloop}
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
+\newcommand{\algorithmicbreak}{\textbf{break}}
+\newcommand{\BREAK}{\STATE \algorithmicbreak}
+\makeatother
+
+% 
 \address{BRASÍLIA}
@@
 \author{Thiago Pereira de Brito Vieira}
-\adviser{Adviser}
-\coadviser{coadviser}
+\adviser{João Paulo Carvalho Lustosa da Costa}
+\coadviser{?}
 
@@
 
-\begin{dedicatory} Ficha Catalográfica\end{dedicatory}
+\begin{dedicatory}Ficha Catalográfica\end{dedicatory}
 
-\begin{dedicatory} Signatures\end{dedicatory}
+\begin{dedicatory}Signatures\end{dedicatory}
 
-\begin{dedicatory} Dedicatory.\end{dedicatory}
+\begin{dedicatory}Dedicatory.\end{dedicatory}
 
@@
 
-Thank You!!!
+Thank You!!
 
@@
 \resumo
-\input{resumo}
+\input{chapters/resumo}
 
 \abstract
-\input{abstract}
+\input{chapters/abstract}
 
@@
 
-\input{acronyms}
+\input{chapters/acronyms}
 \mainmatter
@@
 \include{chapters/1_introduction/introduction}
-%\include{chapters/2_pca_svm_cfa/pca_svm_cfa}
-%\include{chapters/3_eigen_similarity/eigen_similarity}
-%\include{chapters/4_mobile/mobile}
+\include{chapters/2_pca_svm_cfa/pca_svm_cfa}
+\include{chapters/3_eigen_similarity/eigen_similarity}
+\include{chapters/4_mobile/mobile}
+\include{chapters/5_tensor_dl/tensor_dl}
 \include{chapters/6_conclusion_future_work/conclusion}

### 2017-09-30 docs/papers/2016MONE-Final/2016MONE.tex
adding all

--- 2017-01-10 docs/papers/2016MONE/2016MONE.tex
@@
 \begin{acknowledgements}
-The authors thank the Brazilian research and innovation agencies CAPES (Grant FORTE \\23038.007604/2014-69), FAPDF (Research Support Foundation of the Federal District), FINEP (Grant RENASIC PROTO 01.12.0555.00), CNPq (Productivity grant 303905/2014-0 and PVE Grant 88881.030392/2013-01), as well as the European CSF - Aerospace Technology Program (PDE scholarship within project 207644/2015-2), for their support to this research.\end{acknowledgements}
+This research work has the support of the Brazilian research and innovation Agencies CAPES - Coordination for the Improvement of Higher Education Personnel (Grant 23038.007604/2014-69 FORTE - Tempestive Forensics Project), FINEP - Funding  Authority for Studies and Projects (Grant 01.12.0555.00 RENASIC/PROTO - Secure Protocols Laboratory of the National Information Security and Cryptography Network) and FAPDF - Research Support Foundation of the Federal District (Grants 0193.001366/\\2016 UIoT - Universal Internet of Things and 0193.001365/\\2016 - Secure Software Defined Data Center - SSDDC), as well as the Ministry of Planning, Development and Management (Grants 005/2016 DIPLA - Planning and Management Directorate, and 11/2016 SEST - Secretariat of State-owned Federal Companies) and the DPGU - Brazilian Union Public Defender (Grant 066/2016).
+\end{acknowledgements}
 

### 2017-09-30 docs/papers/2017SBSEG/2017sbseg.tex
adding all

\documentclass[12pt]{article}

\usepackage{sbc-template}
\usepackage{graphicx,url}
\usepackage[brazil]{babel}   
%\usepackage[latin1]{inputenc}  
\usepackage[utf8]{inputenc}	% UTF-8 encoding is recommended by ShareLaTex

\sloppy

\title{Detecção de ransomware em dispositivos móveis através de análise comportamental}

\author{}

\address{}

\begin{document}

\maketitle

\begin{abstract}
A quantidade e os danos causados por ransomwares tem crescido significativamente nos últimos anos, evidenciando as necessidades de melhorias nas técnicas de mitigação deste tipo de ataque. A maioria das técnicas para mitigar malwares são baseadas em análise estática ou dependem de bases de dados de comportamentos para treinar um classificador de ameaças, o que limita a capacidade de detecção versões novas ou modificadas de um ransomware. Nós propomos uma abordagem de detecção de ransomwares em dispositivos móveis baseada em análise comportamental através de técnicas de processamento de sinais e avaliamos resultados dos experimentos de sua implementação.
\end{abstract}

\begin{resumo} 
A quantidade e os danos causados por ransomwares tem crescido significativamente nos últimos anos, evidenciando as necessidades de melhorias nas técnicas de mitigação deste tipo de ataque. A maioria das técnicas para mitigar malwares são baseadas em análise estática ou dependem de bases de dados de comportamentos para treinar um classificador de ameaças, o que limita a capacidade de detecção versões novas ou modificadas de um ransomware. Nós propomos uma abordagem de detecção de ransomwares em dispositivos móveis baseada em análise comportamental através de técnicas de processamento de sinais e avaliamos resultados dos experimentos de sua implementação.
\end{resumo}

\section{General Information}

All full papers and posters (short papers) submitted to some SBC conference,
including any supporting documents, should be written in English or in
Portuguese. The format paper should be A4 with single column, 3.5 cm for upper
margin, 2.5 cm for bottom margin and 3.0 cm for lateral margins, without
headers or footers. The main font must be Times, 12 point nominal size, with 6
points of space before each paragraph. Page numbers must be suppressed.

Full papers must respect the page limits defined by the conference.
Conferences that publish just abstracts ask for \textbf{one}-page texts.

\section{First Page} \label{sec:firstpage}

The first page must display the paper title, the name and address of the
authors, the abstract in English and ``resumo'' in Portuguese (``resumos'' are
required only for papers written in Portuguese). The title must be centered
over the whole page, in 16 point boldface font and with 12 points of space
before itself. Author names must be centered in 12 point font, bold, all of
them disposed in the same line, separated by commas and with 12 points of
space after the title. Addresses must be centered in 12 point font, also with
12 points of space after the authors' names. E-mail addresses should be
written using font Courier New, 10 point nominal size, with 6 points of space
before and 6 points of space after.

The abstract and ``resumo'' (if is the case) must be in 12 point Times font,
indented 0.8cm on both sides. The word \textbf{Abstract} and \textbf{Resumo},
should be written in boldface and must precede the text.

\section{CD-ROMs and Printed Proceedings}

In some conferences, the papers are published on CD-ROM while only the
abstract is published in the printed Proceedings. In this case, authors are
invited to prepare two final versions of the paper. One, complete, to be
published on the CD and the other, containing only the first page, with
abstract and ``resumo'' (for papers in Portuguese).

\section{Sections and Paragraphs}

Section titles must be in boldface, 13pt, flush left. There should be an extra
12 pt of space before each title. Section numbering is optional. The first
paragraph of each section should not be indented, while the first lines of
subsequent paragraphs should be indented by 1.27 cm.

\subsection{Subsections}

The subsection titles must be in boldface, 12pt, flush left.

\section{Figures and Captions}\label{sec:figs}

Figure and table captions should be centered if less than one line
(Figure~\ref{fig:exampleFig1}), otherwise justified and indented by 0.8cm on
both margins, as shown in Figure~\ref{fig:exampleFig2}. The caption font must
be Helvetica, 10 point, boldface, with 6 points of space before and after each
caption.

\begin{figure}[ht]
\centering
\includegraphics[width=.5\textwidth]{fig1.jpg}
\caption{A typical figure}
\label{fig:exampleFig1}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=.3\textwidth]{fig2.jpg}
\caption{This figure is an example of a figure caption taking more than one
  line and justified considering margins mentioned in Section~\ref{sec:figs}.}
\label{fig:exampleFig2}
\end{figure}

In tables, try to avoid the use of colored or shaded backgrounds, and avoid
thick, doubled, or unnecessary framing lines. When reporting empirical data,
do not use more decimal digits than warranted by their precision and
reproducibility. Table caption must be placed before the table (see Table 1)
and the font used must also be Helvetica, 10 point, boldface, with 6 points of
space before and after each caption.

\begin{table}[ht]
\centering
\caption{Variables to be considered on the evaluation of interaction
  techniques}
\label{tab:exTable1}
\smallskip
\begin{tabular}{|l|c|c|}
\hline
& Value 1 & Value 2\\[0.5ex]
\hline
&&\\[-2ex]
Case 1 & 1.0 $\pm$ 0.1 & 1.75$\times$10$^{-5}$ $\pm$ 5$\times$10$^{-7}$\\[0.5ex]
\hline
&&\\[-2ex]
Case 2 & 0.003(1) & 100.0\\[0.5ex]
\hline
\end{tabular}
\end{table}

\section{Images}

All images and illustrations should be in black-and-white, or gray tones,
excepting for the papers that will be electronically available (on CD-ROMs,
internet, etc.). The image resolution on paper should be about 600 dpi for
black-and-white images, and 150-300 dpi for grayscale images.  Do not include
images with excessive resolution, as they may take hours to print, without any
visible difference in the result.

\section{References}

Bibliographic references must be unambiguous and uniform.  We recommend giving
the author names references in brackets, e.g. \cite{knuth:84},
\cite{boulic:91}, and \cite{smith:99}.

The references must be listed using 12 point font size, with 6 points of space
before each reference. The first line of each reference should not be
indented, while the subsequent should be indented by 0.5 cm.

\bibliographystyle{sbc}
\bibliography{sbc-template}

\end{document}

### 2017-09-30 fraudDetection/fraudDetectionDLCV.py

@@ -1,0 +1,94 @@
+import numpy as np
+import pandas as pd  																									# data processing, CSV file I/O (e.g. pd.read_csv)
+# import matlab.engine
+from time import time
+from sklearn.decomposition import MiniBatchDictionaryLearning
+from skimage.measure import (compare_mse, compare_nrmse)
+
+
+# settings
+filePath = "/media/thiago/ubuntu/datasets/fraudDetection/"
+# matlab = matlab.engine.start_matlab()
+
+
+# Function to plot image difference
+def print_comparison(target, reference):
+	difference = target - reference
+	mse = compare_mse(reference, target)
+	nrmse = compare_nrmse(reference, target)
+	psnr = 0
+	text = 'norm: %(norm).4f\tMSE: %(MSE).4fs\tNRMSE: %(NRMSE).4fs\tPSNR: %(PSNR).4fs' % {'norm': np.sqrt(np.sum(difference ** 2)), 'MSE': mse, 'NRMSE': nrmse, 'PSNR': psnr}
+	print >> results, text
+
+
+results = open('results/dictionaryLearningCV_fraudDetection.txt', 'w')
+print >> results, '\n## Loading saved data...'
+# data = pd.read_csv('/media/thiago/ubuntu/datasets/fraudDetection/train_under_data.csv', index_col=0)
+data = pd.read_csv('/media/thiago/ubuntu/datasets/fraudDetection/train_data.csv', index_col=0)
+print >> results, 'Data Shape: ' + str(data.shape)
+
+# matlab.cd(r'/home/thiago/dev/projects/anomaly-detection/distributed-tensor-dictionary-learning', nargout=0)
+# matlab.project(nargout=0)
+
+L = data.shape[0]
+N = data.shape[1]
+K_range = [10, 50, 100, 200, 500]
+noIt_range = [10, 50, 100]
+tnz_range = [2, 3, 5, 7]
+
+# ToDo: Evaluate differents sparse coding solvers, such as OMP, Lasso, JavaOMP and others
+for K in K_range:
+	for tnz in tnz_range:
+		for noIt in noIt_range:
+
+			fileNameSufix = 'L={:d}_K={:d}_noIt={:d}_solver=javaORMP_tnz={:d}.csv'.format(L, N, tnz, K, noIt)
+
+			transform_algorithms = [
+				('MiniBatchDL OMP', 'omp', {'transform_n_nonzero_coefs': tnz})
[…]

### 2017-09-30 imageDenoising/Decomposable_Nonlocal_Tensor_Dictionary_Learning_for_Multispectral_Image_Denoising_-_2014/tdl_demo/Comparison.m

@@ -1,0 +1,1 @@
+%==========================================================================
% This script compares multi-spectral imagery (MSI) noise removal methods
% listed as follows:
%   1. band-wise K-SVD
%   2. band-wise BM3D
%   3. Integral K-SVD
%   4. 3D NLM
%   5. BM4D
%   6. LRTA
%   7. PARAFAC
%   8. tensor dictionary learning
%
% Five quality assessment (QA) indices -- PSNR, SSIM, FSIM, ERGAS and SAM
% -- are calculated for each methods after denoising.
%
% See also TensorDL, ksvddenoise, bm3d, NLM3D, bm4d, LRTA, PARAFAC and MSIQA
%
% by Yi Peng
%==========================================================================

clear; clc;

addpath(genpath('..\'));

%% Set enable bits
EN_BWKSVD = 1;
EN_BWBM3D = 1;
EN_KSVD = 1;
EN_NLM3D = 1;
EN_BM4D = 1;
EN_LRTA = 1;
EN_PARAFAC = 1;

%% Set noise level
kappa = 4;       % smaller kappa <--> heavier Poisson noise
sigma_ratio = 0.1;  % higher sigma_ratio <--> heavier Gaussian noise

%% Load the MSI
% NOTE: make sure the MSI is of size height x width x nbands and in range [0, 1]
% In this example, the MSI is stored in a tensor named 'msi'
filename = 'pompoms.mat';
load(filename);
[height, width, nbands] = size(msi);
msi_sz = [height, width, nbands];

psnr = zeros(9, 1);
ssim = zeros(9, 1);
fsim = zeros(9, 1);
ergas = zeros(9, 1);
sam = zeros(9, 1);
time = zeros(9, 1);

%% Add Poisson and Gaussian noise
peak = 2^kappa;                                                             % expected peak value
sigma = peak * sigma_ratio;                                                 % sigma of Gaussian distribution
noisy_msi = poissrnd(msi * peak);                                           % add Poisson noise
[…]

### 2017-09-30 imageDenoising/Decomposable_Nonlocal_Tensor_Dictionary_Learning_for_Multispectral_Image_Denoising_-_2014/tdl_demo/compete_methods/ksvdbox/ksvddenoise.m

@@ -1,0 +1,342 @@
+function [y,D,usage,nz] = ksvddenoise(params,msgdelta)
+%KSVDDENOISE K-SVD denoising.
+%  [Y,D] = KSVDDENOISE(PARAMS) denoises the specified (possibly
+%  multi-dimensional) signal using K-SVD denoising. Y is the denoised
+%  signal and D is the trained dictionary produced by K-SVD.
+%
+%  [Y,D] = KSVDDENOISE(PARAMS,MSGDELTA) specifies the frequency of message
+%  printing during the process. MSGDELTA should be a positive number
+%  representing the interval in seconds between messages. A zero or
+%  negative value cancels all messages. Default is MSGDELTA=5.
+%
+%  [Y,D,NZ] = KSVDDENOISE(...) also returns the average number of non-zero
+%  coefficients in the representations of the denoised blocks.
+%
+%
+%  Required fields in PARAMS:
+%  --------------------------
+%
+%    'x' - Noisy signal.
+%      The signal to denoise (can be multi-dimensional). Should be of type
+%      double, and (for PSNR computations) with values within [0,1] (to
+%      specify a different range, see parameter 'maxval' below).
+%
+%    'blocksize' - Size of block.
+%      Indicates the size of the blocks to operate on. Should be either an
+%      array of the form [N1 N2 ... Np], where p is the number of
+%      dimensions of x, or simply a scalar N, representing the square block
+%      [N N ... N]. See parameter 'stepsize' below to specify the amount of
+%      overlap between the blocks.
+%
+%    'dictsize' - Size of dictionary to train.
+%      Specifies the number of dictionary atoms to train by K-SVD.
+%
+%    'psnr' / 'sigma' - Noise power.
+%      Specifies the noise power in dB (psnr) or the noise standard
+%      deviation (sigma), used to determine the target error for
+%      sparse-coding each block. If both fields are present, sigma is used
+%      unless the field 'noisemode' is specified (below). When specifying
+%      the noise power in psnr, make sure to set the 'maxval' parameter
+%      as well (below) if the signal values are not within [0,1].
+%
+%    'trainnum' - Number of training blocks.
+%      Specifies the number of training blocks to extract from the noisy
+%      signal for K-SVD training.
+%
+%
+%  Optional fields in PARAMS:
+%  --------------------------
+%
[…]

### 2017-09-30 imageDenoising/Decomposable_Nonlocal_Tensor_Dictionary_Learning_for_Multispectral_Image_Denoising_-_2014/tdl_demo/compete_methods/ksvdbox/ompdenoise2.m

@@ -1,0 +1,227 @@
+function [y,nz] = ompdenoise2(params,msgdelta)
+%OMPDENOISE2 OMP denoising of 2-D signals.
+%  OMPDENOISE2 denoises a 2-dimensional signal using OMP denoising. The
+%  function syntax is identical to OMPDENOISE, but it runs significantly
+%  faster on 2-D signals. OMPDENOISE2 requires somewhat more memory than
+%  OMPDENOISE (approximately the size of the input signal), so if memory is
+%  limited, OMPDENOISE can be used instead.
+%
+%  See also OMPDENOISE.
+
+
+%  Ron Rubinstein
+%  Computer Science Department
+%  Technion, Haifa 32000 Israel
+%  ronrubin@cs
+%
+%  August 2009
+
+
+% parse input arguments %
+
+x = params.x;
+D = params.dict;
+blocksize = params.blocksize;
+
+
+% blocksize %
+if (numel(blocksize)==1)
+  blocksize = ones(1,2)*blocksize;
+end
+
+
+% maxval %
+if (isfield(params,'maxval'))
+  maxval = params.maxval;
+else
+  maxval = 1;
+end
+
+
+% gain %
+if (isfield(params,'gain'))
+  gain = params.gain;
+else
+  gain = 1.15;
+end
+
+
+% maxatoms %
+if (isfield(params,'maxatoms'))
+  maxatoms = params.maxatoms;
+else
+  maxatoms = floor(prod(blocksize)/2);
+end
+
+
+% stepsize %
+if (isfield(params,'stepsize'))
+  stepsize = params.stepsize;
+  if (numel(stepsize)==1)
+    stepsize = ones(1,2)*stepsize;
+  end
+else
+  stepsize = ones(1,2);
+end
+if (any(stepsize<1))
+  error('Invalid step size.');
+end
+
+
+% noise mode %
+if (isfield(params,'noisemode'))
+  switch lower(params.noisemode)
+    case 'psnr'
+      sigma = maxval / 10^(params.psnr/20);
+    case 'sigma'
+      sigma = params.sigma;
+    otherwise
+      error('Invalid noise mode specified');
+  end
+elseif (isfield(params,'sigma'))
+  sigma = params.sigma;
+elseif (isfield(params,'psnr'))
+  sigma = maxval / 10^(params.psnr/20);
+else
[…]

### 2017-09-30 imageDenoising/print_gray_image_denoising_DL_CV.py

@@ -7,7 +7,7 @@ from sklearn.decomposition import MiniBatchDictionaryLearning
 from sklearn.feature_extraction.image import extract_patches_2d, reconstruct_from_patches_2d
 
 
-# Function to plot image difference
+## Function to plot image difference
 def print_comparison(image, reference, method_name):
 	difference = image - reference
 	mse = compare_mse(reference, image)
@@ -18,19 +18,19 @@ def print_comparison(image, reference, method_name):
 	print(text)
 
 
-# settings
+## settings
 filePath = "/media/thiago/ubuntu/datasets/imageDenoising/gray_192_512/"
 results = open('results/print_gray_192_512_image_denoising_DL_CV.txt', 'w')
 # matlab = matlab.engine.start_matlab()
 
 
-# Generating matlab data
+## Generating matlab data
 # print('\n## Generating matlab data...')
 # matlab.cd(r'/home/thiago/dev/projects/anomaly-detection/distributed-tensor-dictionary-learning', nargout=0)
 # matlab.projectImageDenoising(nargout=0)
 
 
-# print >> results, '\n## Loading saved data...'
+## Loading saved data
 print('\n## Loading saved data...')
 face = np.loadtxt(filePath + 'face.csv', delimiter=';')
 distorted = np.loadtxt(filePath + 'distortedFace.csv', delimiter=';')
@@ -39,6 +39,7 @@ refPatches = np.loadtxt(filePath + 'refPatches.csv', delimiter=';')
 # print('Data Shape: ' + str(refPatches.shape))
 
 
+## Parameters
 patch_size = (7, 7)
 L = refPatches.shape[0]
 N = refPatches.shape[1]
@@ -47,10 +48,11 @@ noIt_range = [10, 50, 100]
 tnz_range = [2, 5]
 
 
+## Compare distorted image to original one
 print_comparison(distorted, face, 'Distorted image')
 
 
-# Extract noisy patches and reconstruct them using the dictionary
+## Extract noisy patches
 print('\nExtracting noisy patches... ')
 height, width = face.shape
 noisyPatches = extract_patches_2d(distorted[:, width // 2:], patch_size)
@@ -60,6 +62,7 @@ noisyPatches -= noiseMean
 # print('noisyPatches: ' + str(noisyPatches.shape))
 
 
+## For each DL method: learn the dictionary, obtain the sparse coding and reconstruc the denoised image
[…]

### 2017-10-15 docs/papers/2018ICASSP/2018icassp.tex
baseline for icassp 2018 paper

--- 2017-07-12 docs/papers/2017EUSIPCO/2017eusipco.tex
@@
-\documentclass[conference]{IEEEtran}
-
-\ifCLASSINFOpdf
-\else
-\fi
-
-\hyphenation{op-tical net-works semi-conduc-tor}
+% Template for ICASSP-2018 paper; to be used with:
+%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
+%          IEEEbib.bst - IEEE bibliography style file.
+% --------------------------------------------------------------------------
+\documentclass{article}
 
 %% packages
+\usepackage{spconf}
 % graph
@@
 \usepackage{amsfonts}
+\usepackage{amsmath}
 \usepackage{mathtools}
@@
 
+% Example definitions.
+% --------------------
+\def\x{{\mathbf x}}
+\def\L{{\cal L}}
+
+% Title.
+% ------
+\title{TENDOR-BASED DICTIONARY LEARNING FOR FRAUD DETECTION FROM UNBALANCED DATA}
+%
+% Single address.
+% ---------------
+\name{Thiago P. de B. Vieira and Jo\~ao Paulo C. L. da Costa}
+\address{University of Bras\'ilia}
+
+%
+% For example:
+% ------------
+%\address{School\\
+%	Department\\
+%	Address}
+%
+% Two addresses (uncomment and modify for two-address case).
+% ----------------------------------------------------------
+%\twoauthors
+%  {A. Author-one, B. Author-two\sthanks{Thanks to XYZ agency for funding.}}
+%	{School A-B\\
+%	Department A-B\\
+%	Address A-B}
+%  {C. Author-three, D. Author-four\sthanks{The fourth author performed the work
+%	while at ...}}
+%	{School C-D\\
+%	Department C-D\\
+%	Address C-D}
+%
 \begin{document}
-
-\title{A Distributed and Tensor-based Approach for Multidimensional Dictionary Learning}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-\author{\IEEEauthorblockN{Thiago P de B Vieira}
-\IEEEauthorblockA{
-	Department of Electrical Engineering\\
-	University of Brasilia\\
-	Bras\'ilia-DF, Brazil\\
-	Emai	l: tpbvieira@gmail.com
-}
-
-\and
-
-\IEEEauthorblockN{Andr\'e L. F. de Almeida}
-\IEEEauthorblockA{
-	Department of Teleinformatics Engineering\\
-	Federal University of Cear\'a\\
-	Fortaleza, Brazil\\
-	Emai	l: andre.almeida@ieee.org
-}
-
-\and
-
-\IEEEauthorblockN{Florian Roemer\\ and Jo\~ao Paulo C. L. da Costa}
-\IEEEauthorblockA{
-	Institute for Information Technology\\
-	Ilmenau University of Technology\\
-	Ilmenau, Germany
-}
-
-}
-
+%\ninept
+%
 \maketitle
@@
 \begin{abstract}
-The abstract goes here.
+The abstract goes here...
 \end{abstract}
 
-\IEEEpeerreviewmaketitle
+\begin{keywords}
+Tensor-Based Dictionary Learning, Fraud Detection, Classification, Unbalanced Data
+\end{keywords}
 
 \section{Introduction}
-
+\label{sec:introduction}
 Dictionary Learning is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
-The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, classification, data decomposition, feature extraction and image processing \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.
+The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.
 
@@
 
-\section*{Acknowledgment}
-{\small This research was supported by ...}
-
-\bibliographystyle{IEEEtran}
+% References should be produced using the bibtex program from suitable BiBTeX files (here: strings, refs, manuals).
+The IEEEbib.bst bibliography style file from IEEE produces unsorted bibliography list.
+% -------------------------------------------------------------------------
+\vfill\pagebreak
+\bibliographystyle{IEEEbib}
 \bibliography{references}

### 2017-10-16 docs/papers/2018ICASSP/2018icassp.tex
minor improvement

@@
 The abstract goes here...
+Compressive Sensing (CS) allows to acquire signals at sampling rates significantly lower than the Nyquist rate, provided that the signals possess a sparse representation in an appropriate basis.
+However, in some applications of CS, the dictionary providing the sparse description is partially or entirely unknown.
+It has been shown that dictionary learning algorithms are able to estimate the basis vectors from a set of training samples.
+In some applications the dictionary is multidimensional, e.g., when estimating jointly azimuth and elevation in a 2-D direction of arrival (DOA) estimation context.
+In this paper we show that existing dictionary learning algorithms can be extended to exploit this structure, thereby providing a more accurate estimate of the dictionary.
+As examples we choose two prominent dictionary learning algorithms, the method of optimal directions (MOD) and the K-SVD algorithm.
+We propose tensor-based multidimensional extensions for both algorithms and show their improved performances numerically.
 \end{abstract}
@@
 
+% There are two major approaches for dictionary learning.
+First is the analytic approach, in which DCT bases, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries.
+Second is the learning-based approaches, such as the unsupervised learning for dictionary construction [9] and the online dictionary learning [11], [10], which use machine learning methods to construct the dictionary.
+In the analytic approach, some pre-defined functions are used to construct the dictionary.
+Curvelets [24], which tracks the shape of the discontinuity set, offers efficient and near-optimal representation of smooth objects.
+Shearlets [25], which is obtained from dilations, action of translations and shear transformations, has nice geometric properties and mathematical properties for image representation.
+Bandelets [26], which specifies the geometry as a vector field, is de-signed to improve the image compression and noise reduction performance.
+In the learning-based approach, machine learning methods are used to construct the dictionary from the training data.
+The least square error is used by the method of the optimal directions (MOD) [27] to update the dictionary iteratively:
+
 In some applications the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks.
@@
 Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results.
-However, the performance of tensor-based algorithms outperform existing schemes when dealing with growing multidimensional datasets, as can be seen in Figures \ref{fig:fig1} and \ref{fig:fig2}, which show that the detection rating of traditional algorithms decreases over the dataset increasing.
+However, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets, as can be seen in Figures \ref{fig:fig1} and \ref{fig:fig2}, which show that the recovering rating of traditional algorithms decreases over the dataset increasing.
 
@@
 Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based and traditional dictionary learning algorithms for dictionary reconstruction of a multidimensional and separable data.
-Figure \ref{fig:fig1} presents the results for dictionary reconstruction from a matrix of 1000 elements, while Figures \ref{fig:fig2} presents the results from a matrix with 24000 entries.
-It is possible to observe better detection rating for tensor-based algorithms.
-However, tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.
+Figure \ref{fig:fig1} presents the results for dictionary reconstruction from a matrix of 1000 elements by 50 atoms, while Figures \ref{fig:fig2} presents the results from a matrix with 24000 entries by 300 atoms.
+It is possible to observe better reconstruction rating for tensor-based algorithms.
 
+Imbalanced data...
+Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate.
+This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security.
+However, the high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes.
+The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms
+
+A key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset.
+Therefore, we introduce a tensor-based dictionary learning method for fraud detecton from unbalanced data.
+Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification.
+In this paper, we propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds.
+We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
+
+%However, tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.
 We propose a distributed and tensor-based approach for multidimensional dictionary learning in order to obtain better processing time for larger datasets.
@@
 
-\begin{figure}[!htb]
-     \centering 
-	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_4000_100.eps}
-     \caption{Matrix of 4.000 elements, 2000 sample training, 100 iterations}
-     \label{fig:facets_cross}
-\end{figure}
-
-\begin{figure}[!htb]
-     \centering 
-	 \includegraphics[width=0.4\textwidth]{figures/5_20_500_16000_100.eps}
-     \caption{Matrix of 16.000 elements, 500 sample training, 100 iterations}
-     \label{fig:facets_cross}
-\end{figure}
-
 \section{Conclusion}

### 2017-10-17 docs/thesis/chapters/3_eigen_similarity/eigen_similarity.tex
compiling version

@@
      \centering 
-     \includegraphics[height=6cm, width=9cm]{results/figures/fig09.png}
+     \includegraphics[height=6cm, width=9cm]{figures/fig09.png}
      \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
@@
      \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig03.png}
+     \includegraphics[height=4.5cm, width=9cm]{figures/fig03.png}
      \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
@@
      \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig04.png}
+     \includegraphics[height=4.5cm, width=9cm]{figures/fig04.png}
      \caption{Network traffic of user independent operations for network management.}
@@
      \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig05.png}
+     \includegraphics[height=4.5cm, width=9cm]{figures/fig05.png}
      \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
@@
      \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig06.png}
+     \includegraphics[height=4.5cm, width=9cm]{figures/fig06.png}
      \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
@@
      \centering 
-     \includegraphics[height=5cm, width=9cm]{results/figures/fig07.png}
+     \includegraphics[height=5cm, width=9cm]{figures/fig07.png}
      \caption{Connection attempts in order to identify active ports.}
@@
 	\centering
-     \includegraphics[height=11cm, width=9cm]{results/figures/mos_eigen_similarity.eps}
+     \includegraphics[height=11cm, width=9cm]{figures/mos_eigen_similarity.eps}
      \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
@@
 \begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental.eps}
+     \includegraphics[height=4cm, width=12.4cm]{figures/incremental.eps}
      \caption{Traffic selection for incremental approach.}
@@
 \begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/individualized.eps}
+     \includegraphics[height=4cm, width=12.4cm]{figures/individualized.eps}
      \caption{Traffic selection for individual approach.}
@@
 \begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental_individualized.eps}
+     \includegraphics[height=4cm, width=12.4cm]{figures/incremental_individualized.eps}
      \caption{Traffic selection for incremental individualized approach.}
@@
 	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_synflood.eps} 
+     \includegraphics[height=6cm, width=8cm]{figures/eigenvalues_synflood.eps} 
      \caption{Eigenvalues of the sample covariance matrix (synflood).}
@@
 	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_fraggle.eps}
+     \includegraphics[height=6cm, width=8cm]{figures/eigenvalues_fraggle.eps}
      \caption{Eigenvalues of the sample covariance matrix (fraggle).}
@@
 	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_portscan.eps}
+     \includegraphics[height=6cm, width=8cm]{figures/eigenvalues_portscan.eps}
      \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}

### 2017-10-17 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
minor improvements

@@
-\chapter{PCA and SVM for Critical Factors Analysis}
+\chapter{Critical Success Factor Analysis Based on Feaure Selection}
 \label{ch:2_pca_svm_cfa}
@@
 
-There is a significant continuous public spending on information technology (IT) by the Brazilian Federal Public Administration.
-In order to monitor and to diagnose the concerned public institutions IT governance (ITG), the Federal Court of Accounts, in Portuguese Tribunal de Contas da União (TCU), surveys data from these institutions practices as well as best practices for IT in government and business.
-These surveys started in 2007 with 39 questions and nowadays they contain more than 100 questions impacting on the ITG of Brazilian public institutions.
-Moreover, these questions encourage public institutions to adopt best practices not only in IT but also in all other management areas.
-In this paper, we propose the identification and the verification of the critical success factors (CSFs) of the TCU survey variables by means of statistical analysis.
-The CSFs are defined as the factors that most contributed for the high performance of the institutions and they are measured by using the ITG index created by TCU.
-Therefore, in order to validate our results, the statistically identified CSFs are compared with those mentioned by the public administration IT executives in interviews.
-Besides the statistical analysis, we successfully apply the Support Vector Classification (SVC) algorithm to classify public institutions in terms of their ITG index.
-By comparing the SVC based classification with the CSFs obtained from the interviews, we show that there is a very high level of similarity.
-Hence, the CSFs identified with SCV provide a very high impact, regarding the public administration ITG.
+Critical success factor (CSF) is a management term for an element that is necessary for an organization or project to achieve its mission.
+CSFs represent those managerial or enterprise areas that must be given special and continual attention to bring about high performance.
+CSF analysis is one challenger strategic management tool, but if properly applied, CSF analysis does provide a robust and very practical assessment for strategic planners.
 
-2.3 Análise de Componentes Principais
+The Tribunal de Contas da União (Federal Court of Accounts, often referred to as TCU) is the Brazilian federal accountability office that surveys data from public organizations in order to classify the maturity of IT governance.
+
+Feature selection...
+
+We propose a critical factors analysis based on Principal Compoment Analytis (PCA) and recurive feature elimination (RFE) combined with Support Vector Machine (SVM), applied to the surveys data from public organizations, in order to identify the CSF for IT governance measured by TCU.
+
+The resusts show that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn the classification adopted by TCU for IT governance evaluation.
+Therefore, SVM is used to highlight the more significant features for RFE, which corroborates with previous qualitative analysis for critical success factor identification.
+
+\section{Related Works}
+\label{sec:relatedworks}
+
+...
+
+\section{Data Model}
+\label{sec:datamodel}
+
+...
+
+\section{Proposed Framework for Detection and Identification of Network Attacks}
+\label{sec:prop_getv}
+
+\subsection{Análise de Componentes Principais}
 
@@
 
-2.4 Máquinas de Vetores de Suporte (MVS)
+\subsection{Máquinas de Vetores de Suporte}
 
@@
 
-2 Eliminação Recursiva de Variáveis (ERV)
+\subsection{Eliminação Recursiva de Variáveis (ERV)}
 
@@
 
-4.3 Análise das componentes principais
+\section{Experiments and Results}
+\label{sec:experimentalresults}
+
+\subsection{Análise das componentes principais}
 
@@
 
-4.4 Máquinas de Vetores de Suporte
+\subsection{Máquinas de Vetores de Suporte}
 

### 2017-10-18 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
improvements on csf, pca and rfe

@@
 \chapter{Critical Success Factor Analysis Based on Feaure Selection}
-\label{ch:2_pca_svm_cfa}
+\label{ch:2_csf_fs}
 
@@
 Critical success factor (CSF) is a management term for an element that is necessary for an organization or project to achieve its mission.
-CSFs represent those managerial or enterprise areas that must be given special and continual attention to bring about high performance.
-CSF analysis is one challenger strategic management tool, but if properly applied, CSF analysis does provide a robust and very practical assessment for strategic planners.
+CSFs represent the principal assets or areas that must be given investments to achieve better results.
+CSF analysis is one challenger strategic management tool, wich can provide a robust and very practical assessment for strategic planners.
 
-The Tribunal de Contas da União (Federal Court of Accounts, often referred to as TCU) is the Brazilian federal accountability office that surveys data from public organizations in order to classify the maturity of IT governance.
+The identification of the most significative information for one problem is referred to as feature selection by the signal processing and data mining areas, as well as it can be formulated as a principal component problem, which is a widely adopted signal processing technique.
+Feature selection aims to select a subset of relevant information from a larger dataset, in order to improve: data visualization and data understanding, storage requirements, dimensionality, processing time, discriminative sensing, and to overcome overfitting problems to improve prediction and classification performance \cite{chandrashekar2014survey}.
 
-Feature selection...
+Recursive feature elimination (RFE) is a feature selection method for small sample classification problems.
+RFE seeks to improve generalization performance by recursivelly removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features \cite{chen2007enhanced}.
 
-We propose a critical factors analysis based on Principal Compoment Analytis (PCA) and recurive feature elimination (RFE) combined with Support Vector Machine (SVM), applied to the surveys data from public organizations, in order to identify the CSF for IT governance measured by TCU.
-
-The resusts show that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn the classification adopted by TCU for IT governance evaluation.
-Therefore, SVM is used to highlight the more significant features for RFE, which corroborates with previous qualitative analysis for critical success factor identification.
+We propose a critical factors analysis based on Principal Compoment Analytis (PCA) and recurive feature elimination (RFE) combined with Support Vector Machine (SVM), applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
+Resusts show how PCA can make the data discriminative and that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of brazilian public sector.
+Finally, SVM is used to highlight the more significant features identified by RFE, wich are limilar to CSFs previsously identified by a qualitative analysis of the same datased.
 
@@
 
-\section{Proposed Framework for Detection and Identification of Network Attacks}
+\section{Discriminative}
 \label{sec:prop_getv}
 
-\subsection{Análise de Componentes Principais}
+In this section we evaluate the Principal Component Analysis (Principal Component Analysis) for data visualization and identification of critical factors of the evaluated dataset.
+Futhermore, we propose a
 
-Segundo Sabin, Ferrão e Furtado (2004) e Silva et al.
-(2012), a Análise de Componentes Principais (ACP), do inglês Principal Component Analysis – PCA, é utilizada para compressão de dados para identificação das relações entre características dos dados.
+\subsection{Principal Component Analysis}
+
+Principal Component Analysis – PCA, é utilizada para compressão de dados para identificação das relações entre características dos dados.
 Tal compressão é obtida por meio da substituição das variáveis originais por um novo conjunto de variáveis, conhecidas como Componentes Principais (CPs).
@@
 
-\subsection{Máquinas de Vetores de Suporte}
+\subsection{Suport Vector Machines}
 
@@
 
-\subsection{Eliminação Recursiva de Variáveis (ERV)}
+\subsection{Recursive Feature Elimination}
 
@@
 Os resultados mostraram que 69,9\% das variáveis foram classificadas da mesma forma que os FCS identificados anteriormente (Quadro 10) por meio de pesquisa qualitativa.
+
+\section{Conclusion}
+
+support vector machines (SVMs) which have been shown to generalize well even for small sample classification[8].

### 2017-10-20 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
some fixes and text for tensor-dl

@@
 
-Critical success factor (CSF) is a management term for an element that is necessary for an organization or project to achieve its mission.
+Critical Success Factor (CSF) is a management term for an element that is necessary for an organization or project to achieve its mission.
 CSFs represent the principal assets or areas that must be given investments to achieve better results.
@@
 
-Recursive feature elimination (RFE) is a feature selection method for small sample classification problems.
+Recursive Feature Elimination (RFE) is a feature selection method for small sample classification problems.
 RFE seeks to improve generalization performance by recursivelly removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features \cite{chen2007enhanced}.
 
-We propose a critical factors analysis based on Principal Compoment Analytis (PCA) and recurive feature elimination (RFE) combined with Support Vector Machine (SVM), applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
+We propose a critical factors analysis based on Principal Compoment Analytis (PCA) and RFE combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
 Resusts show how PCA can make the data discriminative and that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of brazilian public sector.
@@
 
-...
+This section presents details of the experimental scenario and describes the dataset model used to evaluate our proposal.
+Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
+Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition.
+Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
 
-\section{Discriminative}
+\section{Discriminative...}
 \label{sec:prop_getv}
 
-In this section we evaluate the Principal Component Analysis (Principal Component Analysis) for data visualization and identification of critical factors of the evaluated dataset.
-Futhermore, we propose a
+In this section we evaluate the PCA for data visualization and identification of critical factors of the evaluated dataset. Futhermore, we propose a
 
@@
 
-Principal Component Analysis – PCA, é utilizada para compressão de dados para identificação das relações entre características dos dados.
-Tal compressão é obtida por meio da substituição das variáveis originais por um novo conjunto de variáveis, conhecidas como Componentes Principais (CPs).
-As CPs são obtidas pela combinação linear das variáveis que apresentam a maior variabilidade na matriz de covariância.
-Assim, objetivo principal da ACP é identificar as CPs responsáveis pelas maiores variações entre os resultados.
-Em termos práticos eliminam-se algumas variáveis com pouca informação e determinam-se as variáveis de maior influência na formação de cada CP (Vicini, 2005).
-Segundo Vicini (2005), o conjunto ortogonal de eixos não correlacionados representam as variáveis que mantém ao máximo a variabilidade do conjunto, em outras palavras, a menor perda de informação. 
-Conforme Vicini (2005:29), para o processo de determinação das CPs 
-“é necessário calcular matriz de variância-covariância (∑), ou na matriz de correlação (), encontrar os autovalores e autovetores e, por fim, escrever as combinações lineares, que serão as novas variáveis, denominadas de componentes principais, sendo que cada componente principal é a combinação linear de todas as variáveis originais [...] em ordem de estimação e em termos da variância total, contida nos dados iniciais”.
-Ressalta-se que não são todos as CPs a serem analisadas, mas apenas as com maior variância, alguns autores (Vicini, 2005) determinam os CPs por meio da porcentagem que representam, geralmente, mais de 70\% da informação.
-Conforme Silva et al.
-(2012) para a geração dos CPs, considera-se uma matriz X = (X1, X2, ..., Xp) de dimensões nxp (a Seção 4 mostra as dimensões utilizadas nesta pesquisa) originada a partir de um vetor aleatório.
-A matriz de variância-covariância  das amostras da matriz X é calculada por meio da seguinte expressão , onde é o operador de transposição de uma matriz.
-Nota-se que a matriz de variância-covariância é uma matriz quadrada e não negativa.
-Portanto, pode-se aplica a decomposição em autovalores, do Eigenvalue Decomposition (EVD), da matriz  resultando em um produto de três matrizes.
-,
-% onde  é a matriz de autovetores e  é a matriz cuja diagonal principal contém os autovalores. Para cada autovalor λi da matriz  existe um autovetor unitário e ortognal vi.
-vi = 
-A i-nésima coluna da matriz  é o autovetor normalizado  correspondente ao autovalor . Desta forma, define-se i-ésima componente principal (Yi) como sendo
-% Yi= λ i vi
-Salienta-se que a ACP é baseada na EVD já que parte dos autovalores e dos autovetores representam as componentes principais (CP). 
+PCA is a statistical technique commonly used for dimensionality reduction.
+It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
+PCA is a common technique for finding patterns in data of high dimension and for highlighting its discrinative characteristics and strucutres, throught an orthogonal basis transformation into new basis, by diagonalizing the centered covariance matrix of a data set {Xk E RNlk = 1, ...
+,f}, defined by C = ((Xi - (Xk))(Xi - (Xk))T).
+The coordinates in the Eigenvector basis are called principal components.
+The size of an Eigenvalue >.
+corresponding to an Eigenvector v of C equals the amount of variance in the direction of v.
+PCA is usually used for signal denoising, blind source separation, data compression, data visualization, feature extraction and dimensionality reduction, where a reduced number of features is extrated retaining as much information as possible \cite{jolliffe1986principal}.
+
+describe PCA matematically here? acho que talvez seja melhor fazer uma descricao mais alto nível e voltar depois se sobrar tempo
+
+descrever cmo os dados são aplicados ao pca e apresentar os resultados obtidos
+
 A Figura 1 mostra geometricamente a ACP de duas variáveis X1 e X2.
@@
 
-Máquina de Vetores de Suporte (MVS), do inglês Support Vector Machine (SVM), é um método supervisionado de aprendizagem de máquina, não probabilístico, baseado na teoria de aprendizagem estatística, usado para classificação, regressão e detecção de padrões.
+SVM é um método supervisionado de aprendizagem de máquina, não probabilístico, baseado na teoria de aprendizagem estatística, usado para classificação, regressão e detecção de padrões.
 MVS pode ser aplicado por meio de dois passos: o primeiro passo é o treinamento de um modelo a partir de um determinado conjunto de dados; o segundo consiste em estimar a classificação de dados a partir da aplicação do modelo treinado.

### 2017-10-20 docs/thesis/chapters/5_tensor_dl/tensor_dl.tex
some fixes and text for tensor-dl

@@
-\chapter{Tensor-Based Discriminative Dictionary Learning}
+\chapter{Tensor-based Discriminative Sensing for Fraud Detection}
 \label{ch:tensor_dl}
 
-Tensor-Based Discriminative Dictionary Learning for Fraud Detection
+\section{Introduction}
+\label{sec:introduction}
+
+Compressive sensing is emergent for signal processing and hás been applied to Fields such as signal restoration, image processing and classification (sparce representation classification src )
+
+Iterative dictionary learning hás been used as a blind technique for feature extraction, improving classification without require feature selection or principal conponent analysis
+
+Big data analytics require techniques to Deal with multidimensional data, making Sense of strucutre and relationship of many dimensios.
+Pca is commonly adopted for dimensionality reduction and feature extraction.
+Tensor hás been adopted for multidimensional data analysis and can be explored for dictionary learning
+
+Adaptive techniques for high dimensional and high thoughput
+
+we propose to apply the tensor deconposition for dictionary learning in order to evidentiate the discriminative sensing of a fraud detection dataset
+
+Dictionary Learning is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
+The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.
+
+% There are two major approaches for dictionary learning.
+First is the analytic approach, in which DCT bases, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries.
+Second is the learning-based approaches, such as the unsupervised learning for dictionary construction [9] and the online dictionary learning [11], [10], which use machine learning methods to construct the dictionary.
+In the analytic approach, some pre-defined functions are used to construct the dictionary.
+Curvelets [24], which tracks the shape of the discontinuity set, offers efficient and near-optimal representation of smooth objects.
+Shearlets [25], which is obtained from dilations, action of translations and shear transformations, has nice geometric properties and mathematical properties for image representation.
+Bandelets [26], which specifies the geometry as a vector field, is de-signed to improve the image compression and noise reduction performance.
+In the learning-based approach, machine learning methods are used to construct the dictionary from the training data.
+The least square error is used by the method of the optimal directions (MOD) [27] to update the dictionary iteratively:
+% Dictionaries are either available analytically, or can be learned from a suitable training set.
+While analytic dictionaries permit to capture the global structure of a signal and allow a fast implementation, learned dictionaries often perform better in applications as they are more adapted to the considered class of signals.
+
+In some applications the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks.
+Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in multidimensional data analytics problems \cite{kolda2009tensor}.
+Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}.
+
+% In imagery, the numerical burden for (i) learning a dictionary and for (ii) employing the dictionary for reconstruction tasks only allows to deal with relatively small image patches that only capture local image information.
+Separable dictionaries aims at overcoming these drawbacks by allowing a separable structure on the dictionary throughout the learning process.
+On the one hand, this permits larger patch-sizes for the learning phase, on the other hand, the dictionary is applied efficiently in reconstruction tasks.
+The crucial idea is to allow the dictionary to have a separable structure, where separable means that the dictionary D is given by the Kronecker product of two smaller dictionaries A ∈ R h×a and B ∈ R w×b
+% Without loss of generality we choose square image patches with w = h = 8, which is in accordance to the patch-sizes mostly used in the literature
+
+Multidimensional parameter estimation and learning multidimensional separable dictionaries are growing research problems.
+Roemer \emph{et al.} \cite{roemer2014tensor} show that the multidimensional dictionary estimation problem can be efficiently formulated in terms of tensors, and that their results outperform existing schemes by exploiting the multilinear structure of the problem.
+
+Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results.
+However, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets, as can be seen in Figures \ref{fig:fig1} and \ref{fig:fig2}, which show that the recovering rating of traditional algorithms decreases over the dataset increasing.
+
+\begin{figure}[!htb]
+     \centering 
+	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_1000_100.eps}
+     \caption{Matrix of 1.000 elements, 2000 sample training, 100 iterations}
+     \label{fig:fig1}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering 
+	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_24000_100.eps}
+     \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}
+     \label{fig:fig2}
+\end{figure}
+
+Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based and traditional dictionary learning algorithms for dictionary reconstruction of a multidimensional and separable data.
+Figure \ref{fig:fig1} presents the results for dictionary reconstruction from a matrix of 1000 elements by 50 atoms, while Figures \ref{fig:fig2} presents the results from a matrix with 24000 entries by 300 atoms.
+It is possible to observe better reconstruction rating for tensor-based algorithms.
+
+Imbalanced data...
+Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate.
+This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security.
+However, the high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes.
+The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms
+
+A key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset.
+Therefore, we introduce a tensor-based dictionary learning method for fraud detecton from unbalanced data.
+Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification.
+In this paper, we propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds.
+We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
+
+%However, tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.
+We propose a distributed and tensor-based approach for multidimensional dictionary learning in order to obtain better processing time for larger datasets.
+We focus on a distributed implementation of T-MOD \cite{roemer2014tensor} algorithm, based on a modification of Almeida and Kibangou's \cite{de2014distributed} approach to use Tucker-2 decomposition.
+We perform experiments to test different methods for learning dictionaries, evaluating the processing time and the reconstruction of a dictionary of multidimensional and separable data.
+
+\section{Related Works}
+
+Face recognition...
+
+Fisher...
+
+Florian...
+
+Tem que ver outro...
+
+\section{Data Model}
+
+Consider a generic sparse recovery problem of the following form
+
+\begin{equation}\label{eq:eq01}
+\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W},
+\end{equation}
+
+where $\boldsymbol{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\boldsymbol{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary, $\boldsymbol{S} \in \mathbb{•}hbb{C}^{N \times T}$ represents the sparse coefficient matrix, $\boldsymbol{W} \in \mathbb{C}^{M \times T}$ is the additive noise, and $M < N < T$.
+
+Consider a sparse recovery problem for a separable 2-D manifold, i.e., a manifold that can be written as the product of two 1-D manifolds.
+If we choose a separable 2-D sampling grid, we can write the 2D sparse recovery problem as
+
+Imbalance data where zero is predominante
+
+Data divided into learning and test, with isfraud labels with 1 or 0
+
+\begin{equation}\label{eq:eq02}
+\boldsymbol{X} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}) \cdot \boldsymbol{S} + \boldsymbol{W},
+\end{equation}
+
+where $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$, $\boldsymbol{A}^{(1)} \in \mathbb{C}^{M_1 \times N_1}$, $\boldsymbol{A}^{(2)} \in \mathbb{C}^{M_2 \times N_2}$, $M = M_1 \times M_2$, and $N = N_1 \times N_2$.
+
+The Kronecker model (\ref{eq:eq02}) can be rewritten in an equivalent tensor form.
+Applying the algebraic rules for unfoldings of n-mode products \cite{roemer2014tensor} we can rewrite (\ref{eq:eq02}) into a "Tucker-2" decomposition, as
+
+\begin{equation}\label{eq:eq03}
+\boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} +  \boldsymbol{\mathcal{W}},
+\end{equation}
+
+where $\boldsymbol{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\boldsymbol{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\boldsymbol{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\boldsymbol{X}$, $\boldsymbol{S}$, and $\boldsymbol{W}$ such that $\boldsymbol{X} = [\boldsymbol{\mathcal{X}}]_{(3)}^T$, $\boldsymbol{S} = [\boldsymbol{\mathcal{S}}]_{(3)}^T$, and $\boldsymbol{W} = [\boldsymbol{\mathcal{W}}]_{(3)}^T$, respectively.
+
+We consider a randomly generated data in order to evaluate the proposed approach, where initially $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\boldsymbol{A}$ is obtained.
+In order to be able to evaluate the dictionary reconstruction, the sparse data $\boldsymbol{\mathcal{X}}$ is generated from the dictionary $\boldsymbol{A}$, which is used for further comparisons.
+We adopt 20 as the signal-to-noise ratio (snr) and 5 as the sparseness factor of the data $\boldsymbol{\mathcal{X}}$.
+
+The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\boldsymbol{A}$ from the sparse data $\boldsymbol{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.
+
+\section{Proposed Approach}
+
+...
+
+\section{Experiments}
+
+rocauc is usually used for classification evaluation, but It os not the best option for imbalance data, and Also because It does not consider the unclassified cases, such as false negatives
+
+Due to the inherent complex characteristics of imbalanced data sets, learning from such data requires new understandings, principles, algorithms, and tools to transform vast amounts of raw data efficiently into information and knowledge representation.
+
+We adopt precision recall curve and compare with logistic regression based on original and pca data
+
+Classification based on reconstruction or other value, such as similarity analysis
+
+Evaluate what metric is more discriminative, mse, tsnr or other
+
+\section{Conclusion}
+
+The conclusion goes here.
+
+The importance of the right metric for classification of imbalanced data
+
+Tensor based analysis explores the extructure and relationship of multidimensional data and can improve the discriminative power of imbalanced data for classification
+
+Can be used for feature extraction and used for well known algorithm,  and can be used for discriminative classification based on heuristics
+
+Our proposal is better for imbalanced data than others techniques

### 2017-10-21 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
related works for pca

@@
 
-The identification of the most significative information for one problem is referred to as feature selection by the signal processing and data mining areas, as well as it can be formulated as a principal component problem, which is a widely adopted signal processing technique.
+The identification of the most significative information for one problem is referred to as feature selection by the signal processing and data mining areas, as well as it can be formulated as a principal component problem, which is a widely adopted signal processing technique for data visualization and feature extraction.
 Feature selection aims to select a subset of relevant information from a larger dataset, in order to improve: data visualization and data understanding, storage requirements, dimensionality, processing time, discriminative sensing, and to overcome overfitting problems to improve prediction and classification performance \cite{chandrashekar2014survey}.
@@
 
-We propose a critical factors analysis based on Principal Compoment Analytis (PCA) and RFE combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
-Resusts show how PCA can make the data discriminative and that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of brazilian public sector.
+We propose a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminative analysis and based on RFE combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
+Results show how PCA can make the data discriminative and that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of brazilian public sector.
 Finally, SVM is used to highlight the more significant features identified by RFE, wich are limilar to CSFs previsously identified by a qualitative analysis of the same datased.
@@
 
-...
+Fink and Sukenik \cite{fink2011effect} explore the relationships among IT infrastructure capability and IT business value using PCA applied to all indicators of their study, resulting into 11 factors, with the first factor accounting for only 27.9\% of the variance.
+This technique was used because the PCA extracts orthogonal factors that overcome the problem of multicollinearity.
+
+Ramos \emph{et al} \cite{ramos2016information} propose an overview regarding the evolution of scientific research on IT Governance critical success factors within the domain of public administration.
+By means of bibliometric analysis it was investigated seminal works regarding this theme, considering the characteristic key words found during our analysis.
+The results present 64 critical success factors with high impact on IT governance.
+
+Guyon \emph{et al} \cite{guyon2002gene} propose a method of gene selection utilizing Support Vector Machine methods based on Recursive Feature Elimination (RFE) and demonstrate that the selected genes yield better classification performance and are biologically relevant to cancer.
+The proposed method eliminates gene redundancy automatically and yields better and more compact gene subsets.
 
@@
 \label{sec:datamodel}
+
+In order to audit IT governance of public institutions, the Federal Court of Accounts (TCU, in Portuguese) surveys data of IT practices from these institutions.
+Although these surveys started in 2007 with 39 questions, nowadays they contain more than 100 questions that allow TCU to rank these institutions by their IT governance index.
 
@@
 Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
+
+To the best of our knowledge, we are the first to propose a critical factors analysis based on PCA for visual discriminative analysis and based on RFE combined with SVM for CSF identification.
 
@@
 PCA is usually used for signal denoising, blind source separation, data compression, data visualization, feature extraction and dimensionality reduction, where a reduced number of features is extrated retaining as much information as possible \cite{jolliffe1986principal}.
+
+PCA combines similar (correlated) attributes and creates new ones. Superior to original attributes.
+
+Feature selection doesn't combine attributes. Just evaluates their quality, predictive power and select the best set.
+
+Benefit in PCA is that combination of N attributes is better than any individual attribute. Disadvantage is in harder explanation what exactly that PCA component means.
 
@@
 MVS-ERV (Guyon, 2002) é uma aplicação da ERV utilizando os pesos obtidos por meio do treinamento utilizando MVS como algoritmo de classificação, para identificar as variáveis mais importantes para predições de classificação e eliminar recursivamente as variáveis que menos influenciam na classificação.
+
+Recursive feature elimination is based on the idea to repeatedly construct a model (for example an SVM or a regression model) and choose either the best or worst performing feature (for example based on coefficients), setting the feature aside and then repeating the process with the rest of the features.
+This process is applied until all features in the dataset are exhausted.
+Features are then ranked according to when they were eliminated.
+As such, it is a greedy optimization for finding the best performing subset of features.
+
+The stability of RFE depends heavily on the type of model that is used for feature ranking at each iteration.
+Just as non-regularized regression can be unstable, so can RFE when utilizing it, while using ridge regression can provide more stable results.
+
+%Given an external estimator that assigns weights to features e.g., the coefficients of a linear model, recursive feature elimination RFE is to select features by recursively considering smaller and smaller sets of features.
+First, the estimator is trained on the initial set of features and the importance of each feature is obtained either through a coef attribute or through a feature_importances attribute.
+Then, the least important features are pruned from current set of features.That procedure is recursively repeated on the pruned set until the desired number of features to select is eventually reached.
 

### 2017-10-21 docs/thesis/chapters/5_tensor_dl/tensor_dl.tex
related works for pca

@@
 
+Compressed sensing (also known as compressive sensing, compressive sampling, or sparse sampling) is a signal processing technique for efficiently acquiring and reconstructing a signal, by finding solutions to underdetermined linear systems.
+This is based on the principle that, through optimization, the sparsity of a signal can be exploited to recover it from far fewer samples than required by the Shannon-Nyquist sampling theorem.
+There are two conditions under which recovery is possible.[1] The first one is sparsity which requires the signal to be sparse in some domain.
+The second one is incoherence which is applied through the isometric property which is sufficient for sparse signals.[2][3]
+
+A common goal of the engineering field of signal processing is to reconstruct a signal from a series of sampling measurements.
+In general, this task is impossible because there is no way to reconstruct a signal during the times that the signal is not measured.
+
+Around 2004, Emmanuel Candès, Terence Tao, and David Donoho proved that given knowledge about a signal's sparsity, the signal may be reconstructed with even fewer samples than the sampling theorem requires.[4][5] This idea is the basis of compressed sensing.
+	https://www.wired.com/2008/03/engineers-test-highly-accurate-face-recognition/
+
+Sparse dictionary learning is a representation learning method which aims at finding a sparse representation of the input data (also known as sparse coding) in the form of a linear combination of basic elements as well as those basic elements themselves.
+These elements are called atoms and they compose a dictionary.
+Atoms in the dictionary are not required to be orthogonal, and they may be an over-complete spanning set.
+This problem setup also allows the dimensionality of the signals being represented to be higher than the one of the signals being observed.
+The above two properties lead to having seemingly redundant atoms that allow multiple representations of the same signal but also provide an improvement in sparsity and flexibility of the representation.
+
+sparse coding is a sparse representation of one data in the form of a linear combination of basic elements. These elements are called atoms and they compose a dictionary.
+
+One of the key principles of dictionary learning is that the dictionary has to be inferred from the input data.
+The emergence of sparse dictionary learning methods was stimulated by the fact that in signal processing one typically wants to represent the input data using as few components as possible.
+Before this approach the general practice was to use predefined dictionaries (such as fourier or wavelet transforms).
+However, in certain cases a dictionary that is trained to fit the input data can significantly improve the sparsity, which has applications in data decomposition, compression and analysis and has been used in the fields of image denoising and classification, video and audio processing.
+Sparsity and overcomplete dictionaries have immense applications in image compression, image fusion and inpainting.
+
 Face recognition...
@@
 
+Receiver Operating Characteristic Curve (ROC curve) http://www.chioka.in/differences-between-roc-auc-and-pr-auc/
+
+A ROC curve is plotting True Positive Rate (TPR) against False Positive Rate (FPR).
+
+TPR is defined as:
+
+tpr-formula
+
+FPR is defined as:
+
+fpr-formula
+
+where TP = true positive, TN = true negative, FP = false positive, FN = false negative.
+
+A typical ROC curve looks like this, which shows two ROC curves for Algorithm 1 and Algorithm 2.
+
+sample-ROC-curve
+
+The goal is to have a model be at the upper left corner, which is basically getting no false positives – a perfect classifier.
+
+The receiver operating characteristic area under curve (ROC AUC) is just the area under the ROC curve. The higher it is, the better the model is.
+
+Precision Recall Curve (PR Curve)
+A PR curve is plotting Precision against Recall.
+
+Precision is defined as:
+
+precision-formula
+
+Recall is defined as:
+
+recall-formulaA typical PR curve looks like this, which shows two PR curves for Algorithm 1 and Algorithm 2.
+
+sample-PR-curve
+
+The goal is to have a model be at the upper right corner, which is basically getting only the true positives with no false positives and no false negatives – a perfect classifier.
+
+The precision recall area under curve (PR AUC) is just the area under the PR curve. The higher it is, the better the model is.
+
+Differences between the ROC AUC and PR AUC
+Randy has given a great explanation here, plus a little of my understanding.
+
+Since PR does not account for true negatives (as TN is not a component of either Precision or Recall), or there are many more negatives than positives (a characteristic of class imbalance problem), use PR.
+If not, use ROC.
+
 Due to the inherent complex characteristics of imbalanced data sets, learning from such data requires new understandings, principles, algorithms, and tools to transform vast amounts of raw data efficiently into information and knowledge representation.

### 2017-10-21 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
minor

@@
 
+To the best of our knowledge, we are the first to propose a critical factors analysis based on PCA for visual discriminative analysis and based on RFE combined with SVM for CSF identification from IT governance data.
+
 \section{Data Model}
@@
 
-In order to audit IT governance of public institutions, the Federal Court of Accounts (TCU, in Portuguese) surveys data of IT practices from these institutions.
-Although these surveys started in 2007 with 39 questions, nowadays they contain more than 100 questions that allow TCU to rank these institutions by their IT governance index.
+The Federal Court of Accounts (TCU, in Portuguese) surveys data of IT practices of brazilian public organizations in order to audit IT governance.
+This resulting dataset is called iGovTI.
+We propose to use the iGovTI of 2012 for CSF analysis.
+The iGovTI is composed by 201 questions used for organization ranking according to their responses and their IT governance.
 
@@
 Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
-
-To the best of our knowledge, we are the first to propose a critical factors analysis based on PCA for visual discriminative analysis and based on RFE combined with SVM for CSF identification.
 
@@
 Assim, para a definição do algoritmo de classificação foram realizados testes em 21 algoritmos, a seguir é apresentada uma listagem dos algoritmos avaliados e a taxa de acerto de classificação obtida:
-1. KNN: 0.714286
-2. ElasticNet: 0.155336
-3. ElasticNetCV: 0.831531
-4. LassoCV: 0.827440
-5. LassoLarsIC: 0.713763
-6. LinearRegression: 0.360878
-7. LogisticRegression: 0.771429
-8. OrthogonalMatchingPursuit: 0.769630
-9. PassiveAggressiveClassifier: 0.800000
-10. PassiveAggressiveRegressor: 0.852184
-11. Perceptron: 0.800000
-12. Ridge: 0.529922
-13. RidgeClassifier: 0.657143
-14. RidgeClassifierCV: 0.742857
-15. RidgeCV: 0.750149
-16. SGDClassifier: 0.828571
-17. MultinomialNB: 0.742857
-18. lda.LDA: 0.628571
-19. SVM.SVR: 0.826885
-20. SVM.SVC: 0.914286
-21. SVM.LinearSVC: 0.714286
-22.
+
+k-nearest neighbors \cite{fukunaga1975branch}	0.714286
+Elastic Net \cite{zou2005regularization}	0.831531
+Lasso \cite{tibshirani1996regression}	0.827440
+LinearRegression \cite{draper2014applied}	0.360878
+LogisticRegression \cite{hosmer2013applied}	0.771429
+LDA \cite{martinez2001pca}	0.628571
+SVM \cite{hearst1998support}	0.914286
+SVM Regression \cite{smola2004tutorial}	0.826885
+Linear SVM \cite{fan2008liblinear}	0.714286
+
 O algoritmo que obteve maior sucesso para classificação foi o SVC, que é uma implementação de Máquina de Vetores de Suporte aplicado para a classificação.
@@
 Para treinar o algoritmo utilizou-se 90\% dos dados e os 10\% restantes foram utilizados para avaliar a eficiência de classificação dos algoritmos, comparando a taxa de acerto entre as predições feitas e os valores reais de classificações de organizações pelo iGovTI.
-23.
+
 Uma vez identificado um algoritmo capaz de efetuar a classificação desejada, o próximo passo é identificar as variáveis mais relevantes para esta classificação.
 Para isso, será utilizado o algoritmo ERV.
-24. 
-4.5 Eliminação Recursiva de Variáveis
-25. 
-26.
+
 A ERV pode usar vários algoritmos de classificação como critério de seleção das variáveis mais importantes, escolhemos o algoritmo SVC por ele ter apresentado maior acurácia entre os algoritmos de classificação avaliados.
@@
 A partir da pesquisa de identificação por meio de entrevistas na APF, identificou-se 54 variáveis consideradas FCS, este critério foi utilizado para determinar o quantitativo de variáveis mais importantes para a classificação.
-27.
+
 Assim, aplicaram-se o algoritmo ERV utilizando o SVC como critério para a seleção das variáveis mais importantes para a classificação, selecionando as 54 variáveis que correspondem aos FCS levantados nas entrevistas com os executivos de TI.
@@
 
-support vector machines (SVMs) which have been shown to generalize well even for small sample classification[8].
+pca make possible visual discriminative analysis of reduced dataset but it is not easy make correlation betwee extracted features and its original, therefore it is necessary adictional techniques to identify CSF.
+
+SVM have been shown to generalize well even for small sample classification[8]. SVM presents high performance to reproduce the iGovTI classification and can be used for RFE in order to identify CSF
+
+The selected features revealed by RFE are quite similar to the qualitative analysis of Ramos \emph{et al} \cite{ramos2016information}.

### 2017-10-21 docs/thesis/chapters/6_conclusion_future_work/conclusion.tex
compiling version with all chapters
minor

@@
 	\item We \textbf{showed the speed-up obtained with MapReduce for DPI}, with variations in input size, cluster size and block size;
-	\item We \textbf{published two papers} reporting our results, as follows:
+	\item We \textbf{published five papers} reporting our results, as follows:
 	\begin{enumerate}
-		\item Vieira, T., Soares, P., Machado, M., Assad, R., and Garcia, V.
-\textit{Evaluating Performance of Distributed Systems with MapReduce and Network Traffic Analysis}.
-In ICSEA 2012, The Seventh International Conference on Software Engineering Advances.
-Xpert Publishing Services.
-		\item Vieira, T., Soares, P., Machado, M., Assad, R., and Garcia, V.
-\textit{Measuring Distributed Applications Through MapReduce and Traffic Analysis}.
-In Parallel and Distributed Systems (ICPADS), 2012 IEEE 18th International Conference on, pages 704 - 705.
+		\item T.
+P.
+B.
+Vieira, D.
+F.
+Ten\'orio, J.
+P.
+C.
+L.
+da Costa, E.
+P.
+de Freitas, G.
+Del Galdo, and R.
+T.
+de Sousa J\'unior, \textit{Model Order Selection and Eigen Similarity based Framework for Detection and Identication of Network Attacks}.
+Journal of Networking and Computer Applications (JNCA), Vol 90, Jul 2017, Pages 26–41.
+		\item T.
+Galibus, T.
+P.
+B.
+Vieira, E.
+P.
+de Freitas, R.
+Albuquerque, J.
+P.
+C.
+L.
+da Costa, R.
+T.
+de Sousa Jr, V.
+Krasnoproshin, A.
+Zaleski, H.
+Vissia, and G.
+del Galdo, \textit{Offline Mode for Corporate Mobile Client Security Architecture}.
+Mobile Networks and Applications, Springer, Mar 2017, Pages 1-17.
+		\item K.
+H.
+C.
+Ramos, R.
+T.
+Sousa Jr, T.
+P.
+B.
+Vieira, and J.
+P.
+C.
+L.
+da Costa, \textit{DISCOVERING CRITICAL SUCCESS FACTORS FOR INFORMATION TECHNOLOGIES GOVERNANCE THROUGH BIBLIOMETRIC ANALYSIS OF RESEARCH PUBLICATIONS IN THIS DOMAIN}.
+Information (Yamaguchi), 2016
+		\item K.
+H.
+C.
+Ramos, R.
+T.
+de Sousa Jr., T.
+P.
+B.
+Vieira, and J.
+P.
+C.
+L.
+da Costa, \textit{Identifying relevant IT governance critical success factors by means of bibliometric analysis}.
+7th International Conference on Information, November 2015, Taipei, Taiwan.
+		\item K.
+H.
+C.
+Ramos, T.
+P.
+B.
+Vieira, J.
+P.
+C.
+L.
+da Costa, and R.
+T.
+Sousa Jr., \textit{ANÁLISE MULTIDIMENSIONAL DE FATORES CRÍTICOS DE SUCESSO EM GOVERNANÇA DE TI NA ADMINISTRAÇÃO PÚBLICA FEDERAL À LUZ DOS DADOS DE CONTROLE EXTERNO}.
+in Proc.
+12th International Conference on Management of Technology and Information Systems (CONTECSI), 2015, São Paulo, Brazil.
 	\end{enumerate}
 \end{enumerate}
-
-\subsection{Lessons Learned}
-\label{sc:chaptersummary}
-
-The contributions cited are of scientific and academic scope, with implementations and evaluations little explored in the literature.
-However, with the development of this work, some important lessons were learned.
-
-During this research, different approaches for evaluating distributed systems of cloud computing providers were studied.
-In this period, we could see the importance of the performance evaluation in a cloud computing environment, and the recent efforts to diagnose and evaluate system at production environment of a data center.
-Also, the growth of the Internet and resource utilization make necessary solutions to be able to evaluate large amounts of data in short time, with low performance degradation of the evaluated system.
-
-MapReduce has grown as a general purpose solution for big data processing, but it is not a solution for all kind of problems, and its performance is dependent of several parameters.
-Some researches has been done in order to improve MapReduce performance, through analytical modelling, simulation and measurement, but the most relevant contributions in this direction was guided by realistic workload evaluations, from large MapReduce clusters.
-
-We learned that although the facilities provided by the MapReduce for distributed processing, its performance is influenced by the environment, network topology, workload, data type and by several specific parameter configurations.
-Therefore, an evaluation of the MapReduce behavior using data of a realistic environment will provide more accurate and wide results, while in controlled experiments the results are more restricted and limited to the evaluated metrics and factors.
 
@@
 
-Because of time constraints imposed on the master degree, this dissertation addresses some problems, but some problems are still open and others are emerging from current results.
+Because of time constraints imposed on the Ph.D.
+degree, this thesis addresses some problems, but some problems are still open and others are emerging from current results.
 Thus, the following issues should be investigated as future work:
@@
 \begin{itemize}
-	\item \textbf{Evaluating of all components of the proposed approach}.
-This dissertation evaluated the \textit{JNetPCAP-JXTA}, the \textit{AppAnalyzer} and its implementation to evaluate a JXTA-based distributed application, it is necessary to evaluate the \textit{SnifferServer}, \textit{Manager} and the whole system working together, analysing their impact into the measured distributed system and the scalability achieved;
-	\item Development of a \textbf{technique for the efficient evaluation of distributed systems through information extracted from network traffic}.
-This dissertation addressed the problem of processing capacity for measuring distributed systems through network traffic analysis, but it is necessary an efficient approach to diagnose problems of distributed systems, using information of flows, connections, throughput and response time obtained from network traffic analysis;
-	\item Development of a \textbf{analytic model and simulations}, using information of MapReduce behavior for network traffic analysis, measured by this dissertation, to reproduce its characteristics and enable the evaluation and prediction of some cases of MapReduce for network traffic analysis;
+	\item ...;
+	\item ....
 \end{itemize}

### 2017-10-22 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
one more step

@@
 
-We propose a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminative analysis and based on RFE combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
+We propose a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminant analysis and based on RFE combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
 Results show how PCA can make the data discriminative and that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of brazilian public sector.
 Finally, SVM is used to highlight the more significant features identified by RFE, wich are limilar to CSFs previsously identified by a qualitative analysis of the same datased.
+
+This chapter is organized as follows.
+In Section \ref{sec:relatedworks}, related works are discussed.
+Section \ref{sec:datamodel} presents the data model and the evaluated datasets.
+Section \ref{sec:csf_fs} describes the proposed approach for critical success factors analysis.
+Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:conclusion} draws the conclusions.
 
@@
 
-To the best of our knowledge, we are the first to propose a critical factors analysis based on PCA for visual discriminative analysis and based on RFE combined with SVM for CSF identification from IT governance data.
+To the best of our knowledge, we are the first to propose a critical factors analysis based on PCA for visual discriminant analysis and based on RFE combined with SVM for CSF identification from IT governance data.
 
@@
 
-The Federal Court of Accounts (TCU, in Portuguese) surveys data of IT practices of brazilian public organizations in order to audit IT governance.
-This resulting dataset is called iGovTI.
-We propose to use the iGovTI of 2012 for CSF analysis.
-The iGovTI is composed by 201 questions used for organization ranking according to their responses and their IT governance.
+The brazilian Federal Court of Accounts (TCU, in Portuguese) surveys data regarding IT practices of brazilian public organizations in order to audit IT governance.
+The dataset with a consolidate view about the answers for this surgey the IT governance index is called iGovTI.
+The iGovTI is composed by 201 multiple choice questions, used for ranking according to their IT governance, submmited to 349 organizations.
+The TCU computes the IT governance index and classifies the IT governace of each organization.
+Additionally, Ramos \emph{et al} \cite{ramos2016information} classifies each question regarding its relevance for IT governance through a qualitative analysis, and identify the CSFs for selected IT managers regarding IT governance.
 
-This section presents details of the experimental scenario and describes the dataset model used to evaluate our proposal.
-Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
-Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition.
-Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
+\section{An approach for Critical Success Factors Analysis}
+\label{sec:csf_fs}
 
-\section{Discriminative...}
-\label{sec:prop_getv}
+In this section we propose an approach for Critical Success Factors Analysis based on visual discriminant analysis and based on feature selection, in orde to identify the CSFs for IT governance according to iGovTI.
+Initially we conduct an analysis based on PCA to evaluate the relevante of each feature according to their variance, and use the 2 most relevant features for a visualization of the iGovTI ranking.
+Furthermore, we propose a critical success factors analysis based on SVM for classification and based on RFE for identification of the most relevant factors.
 
-In this section we evaluate the PCA for data visualization and identification of critical factors of the evaluated dataset. Futhermore, we propose a
-
-\subsection{Principal Component Analysis}
+\subsection{Visual Discriminant Analysis based on PCA}
 
@@
 It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
-PCA is a common technique for finding patterns in data of high dimension and for highlighting its discrinative characteristics and strucutres, throught an orthogonal basis transformation into new basis, by diagonalizing the centered covariance matrix of a data set {Xk E RNlk = 1, ...
+
+PCA is also used for finding patterns in data of high dimension and for highlighting its discrinative characteristics and strucutres, throught an orthogonal basis transformation into new basis, by diagonalizing the centered covariance matrix of a data set {Xk E RNlk = 1, ...
 ,f}, defined by C = ((Xi - (Xk))(Xi - (Xk))T).
@@
 
-\subsection{Suport Vector Machines}
+\subsection{iGovTI classification}
 
@@
 
-\subsection{Recursive Feature Elimination}
+\subsection{CSF analysis based on SVM-RFE}
 
@@
 
- Quadro 7 – Autovalores  Ordem dos autovalores  Autovalores da variância explicada Autovalores acumulados da variância explicada acumulada
-Fonte: Dados da Pesquisa
+\begin{figure}[h!]
+     \centering 
+     \includegraphics[height=6cm, width=9cm]{figures/pca-svm-rfe_raw_variance_ecdf.eps}
+     \caption{Empirical CDF of variance.}
+     \label{fig:fig1}
+\end{figure}
 
@@
 Verifica-se que uma queda menos acentuada ocorreu entre o quarto e o quinto autovalor e analisando-se os autovalores superiores a 2, observa-se que pode-se considerar até o vigésimo valor já que a partir daí os valores dos autovalores sucessivos são praticamente constantes.
-
-Gráfico 1 – Ranking dos autovalores
-Fonte: Dados da pesquisa
 
@@
 
-Gráfico 2 – Relação entre os  e  
-Fonte: Dados da pesquisa
+\begin{figure}[h!]
+     \centering 
+     \includegraphics[height=6cm, width=9cm]{figures/pca-svm-rfe_raw_igovti_ranking_pc2.eps}
+     \caption{iGovTI ranking from 2 principal components.}
+     \label{fig:fig1}
+\end{figure}
 
@@
 
-Gráfico 3 – Relação entre os ,  e 
-Fonte: Dados da pesquisa
-
 A análise do Gráfico 3 mostra revela a existência de três grupos (VI, VII e VIII). O grupo I contém a variável Q51l que corresponde a FCS e o grupo II também mostra FCS, a variável Q23b.
@@
 O Quadro 8 mostra quais são essas variáveis significativas.
-
-Quadro 8 – Variáveis com alta significância CP1 a CP20
-Fonte: Dados da pesquisa
 
@@
 
-Quadro 9 – Variáveis com alta significância CP 21 a CP 51
-Fonte: Dados da pesquisa
-
 Já o Gráfico 4 mostra a caixa de seleção de variáveis e comandos para ACP em que se utilizam os fatores 1 e 2, eixo x e eixo y respectivamente.
 O objetivo deste gráfico é fazer os planos principais com a nuvem de pontos dos indivíduos, no caso as 349 instituições.
-
-Gráfico 4 – Dados brutos da pesquisa
-Fonte: Dados da pesquisa
 
@@
 Fazendo-se um corte no eixo das abcissas, a partir do ponto (-8,8, 0), destacam-se 23 instituições à esquerda desse ponto (Quadro 5).
-
-Gráfico 5 – Destaque de corte nas instituições 
-Fonte: Dados da pesquisa
 
@@
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2017-10-22 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
section 2.3

@@
 This chapter is organized as follows.
-In Section \ref{sec:relatedworks}, related works are discussed.
-Section \ref{sec:datamodel} presents the data model and the evaluated datasets.
-Section \ref{sec:csf_fs} describes the proposed approach for critical success factors analysis.
-Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:conclusion} draws the conclusions.
+In Section \ref{sec:ch2_relatedworks}, related works are discussed.
+Section \ref{sec:ch2_datamodel} presents the data model and the evaluated datasets.
+Section \ref{sec:ch2_csf_fs} describes the proposed approach for critical success factors analysis.
+Section \ref{sec:ch2_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:ch2_conclusion} draws the conclusions.
 
 \section{Related Works}
-\label{sec:relatedworks}
+\label{sec:ch2_relatedworks}
 
@@
 \section{Data Model}
-\label{sec:datamodel}
+\label{sec:ch2_datamodel}
 
@@
 \section{An approach for Critical Success Factors Analysis}
-\label{sec:csf_fs}
+\label{sec:ch2_csf_fs}
 
-In this section we propose an approach for Critical Success Factors Analysis based on visual discriminant analysis and based on feature selection, in orde to identify the CSFs for IT governance according to iGovTI.
-Initially we conduct an analysis based on PCA to evaluate the relevante of each feature according to their variance, and use the 2 most relevant features for a visualization of the iGovTI ranking.
+In this section we propose an approach for CSF Analysis based on visual discriminant analysis and based on feature selection, in order to identify the CSFs for IT governance according to iGovTI.
+Initially we conduct an analysis based on PCA to evaluate the relevance of each question according to their variance, and use the 2 most relevant features for a visualization of the iGovTI ranking.
 Furthermore, we propose a critical success factors analysis based on SVM for classification and based on RFE for identification of the most relevant factors.
@@
 \subsection{Visual Discriminant Analysis based on PCA}
+\label{sec:ch2_pca}
 
-PCA is a statistical technique commonly used for dimensionality reduction.
+PCA is a statistical technique commonly used for signal denoising, blind source separation, data compression, data visualization, feature extraction and dimensionality reduction, where a reduced number of features is extrated retaining as much information as possible \cite{jolliffe1986principal}.
 It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
 
-PCA is also used for finding patterns in data of high dimension and for highlighting its discrinative characteristics and strucutres, throught an orthogonal basis transformation into new basis, by diagonalizing the centered covariance matrix of a data set {Xk E RNlk = 1, ...
-,f}, defined by C = ((Xi - (Xk))(Xi - (Xk))T).
-The coordinates in the Eigenvector basis are called principal components.
-The size of an Eigenvalue >.
-corresponding to an Eigenvector v of C equals the amount of variance in the direction of v.
-PCA is usually used for signal denoising, blind source separation, data compression, data visualization, feature extraction and dimensionality reduction, where a reduced number of features is extrated retaining as much information as possible \cite{jolliffe1986principal}.
+PCA is an orthogonal basis transformation into new basis, by diagonalizing the centered covariance matrix of a data set \{$\mathbf{x}_j \in \mathbb{R}^m, j = 1, ...
+,n$\}, defined by $\mathbf{C} = \frac{\mathbf{X}^\intercal\mathbf{X}}{n}$, where $\mathbf{X} = (\mathbf{x}_1, ...
+, \mathbf{x}_n)^\intercal$ and the samples are assumed to have zero mean.
+The eigenvectors $\mathbf{v}_i$ of $\mathbf{C}$ are called principal components (PC), and the sample variance along $\mathbf{v}_i$ of $\mathbf{C}$ is given by the corresponding eigenvalue $\lambda_i$.
+Projecting onto the eigenvectors with the largest eigenvalues assumes that minimal information is lost, considering that in many applications these directions contain the most interesting information, such as in in data compression and de-noising.
 
-PCA combines similar (correlated) attributes and creates new ones. Superior to original attributes.
+Initially, we compute the covariance matrix $\mathbf{C}$ of a zero mean data and visualize the data relationship for the sambple covariance.
+Sample covariance is calculated by computing deviations of each measurement from the average of all measurements for that variable.
+Then the deviations for the two measurements are multiplied together separately for each subject.
+Finally these values are averaged.
+After that, the eigenvectors $\mathbf{v}_i$ and eigenvalues $\lambda_i$ of $\mathbf{C}$ are computed through Singular Value Decomposition (SVD), then it is possible to evaluate the variance distribution of the extracted components through an empirical cumulative distribution function (ECDF).
+Evaluating the variance distribution we expect to visualize if some features concentrates the variance and indicates advantages for dimensionality reduction.
 
-Feature selection doesn't combine attributes. Just evaluates their quality, predictive power and select the best set.
+Finally, we propose to select the two features with the largest variance and evaluate the relationship between the two principal components and the iGovTI classification, in order to visualize if there is a segmentation according to the iGovTI ranking.
 
-Benefit in PCA is that combination of N attributes is better than any individual attribute. Disadvantage is in harder explanation what exactly that PCA component means.
+Considering that PCA combines attributes and creates new ones, with measurements from all of the original variables, it is hard to identify what original variables are most relevant.
+Therefore, it is still necessary an additional method to reveal what are the CSF for iGovTI.
+For this problem, we propose a CSF analysis based on RFE.
 
-describe PCA matematically here? acho que talvez seja melhor fazer uma descricao mais alto nível e voltar depois se sobrar tempo
+\subsection{CSF analysis based on RFE}
+\label{sec:ch2_csfa}
 
-descrever cmo os dados são aplicados ao pca e apresentar os resultados obtidos
+Feature selection aims the identification of the most significative information for one problem or algorithm, such as classification or prediction problems.
+RFE is a feature selection method for small sample classification problems by recursivelly removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features through a selected classifier.
 
-A Figura 1 mostra geometricamente a ACP de duas variáveis X1 e X2.
-A nuvem de pontos representa o diagrama de espalhamento destas duas variáveis.
-Conforme explicado, o primeiro passo é o cálculo da matriz de variância-covariância e em seguida se aplica a EVD.
-Verifica-se que a CP1, denominado CP principal (com a maior variância), é ortogonal à segunda CP (CP2 com a menor variância).
-Note que as duas variáveis são perpendiculares porque as variáveis obtidas pela ACP são independentes.
-Logo, após a aplicação da ACP, ao invés de representar os dados por meio de duas variáveis correlacionadas X1 e X2, eles podem ser representados por duas variáveis independentes Y1 e Y2.
-Dada a elevada correlação entre X1 e X2, verifica-se que as duas variáveis X1 e X2 podem ser aproximadas por uma única variável Y1 na Figura 1.
+Initially, it is necessary to identify one classifier that can identify the iGovTI classificatin of one organization according to a trainning dataset.
+Therefore, we propose an algorithm evaluatio in order to identify wich one presents the best accuracy for iGovTI classification.
+The selected algorithm should be used by RFE to identify the CSF, that should be compared to results of the CSF qualitative analysis conducted by Ramos \emph{et al} \cite{ramos2016information}.
 
-Figura 1 – Representação e Rotação do Componente Principal
-Fonte: Desenvolvido pelos autores
+Feature selection doesn't combine attributes, as PCA, but just evaluates their informative quality, predictive power and select the best set.
+Given an external estimator or classifier, that assigns weights to features, RFE is able to select features by recursively considering smaller and smaller sets of features.
+First, the estimator is trained on the initial set of features and the importance of each feature is obtained either through a coefficient attribute or through a feature importances attribute.
+Then, the least important features are pruned from current set of features.
+That procedure is recursively repeated on the pruned set until the desired number of features to select is eventually reached.
+RFE can also rank all features areaccording to when they were eliminated.
 
-A Figura 1 reforça a ideia da ACP cujo objetivo inicial é a identificação de planos e linhas que representem um conjunto de pontos em um espaço com um número menor de variáveis tendo em vista as possíveis correlações que podem existir entre as variáveis originais.
-Assim, para o método da estatística multivariada ACP utilizar todas variáveis, separa-se a informação útil da informação redundante (Finkler, 2003).
-Ressalta-se que a ACP é um dos primeiros passos para outras análises multivariadas, sendo um dos métodos mais comuns empregados na análise de informações (Sabin, Ferrão e Furtado, 2004; Silva et al.,2012).
-
-\subsection{iGovTI classification}
-
-SVM é um método supervisionado de aprendizagem de máquina, não probabilístico, baseado na teoria de aprendizagem estatística, usado para classificação, regressão e detecção de padrões.
-MVS pode ser aplicado por meio de dois passos: o primeiro passo é o treinamento de um modelo a partir de um determinado conjunto de dados; o segundo consiste em estimar a classificação de dados a partir da aplicação do modelo treinado.
-Quando aplicado para classificação, MVS busca a identificação de um hiperplano que separa os dados em classes distintas, por meio de uma margem máxima entre duas classes de dados.
-Desta forma, um conjunto de dados é linearmente separável se for possível dividir seus dados em duas classes, por meio um hiperplano, conforme na figura a seguir.
-
-Figura 2 – Separação Linear por Margem Máxima
-
-Dado um conjunto de dados previamente classificados, (xi, yi), xi ∈ Rn, yi ∈ {1, −1}, i = 1, .
-.
-.
-, l, um classificador linear pode ser definido por: wT ⋅ x + b = 0, desta forma o hiperplano ótimo é definido pelos valores ótimos do vetor de pesos wi e do bias bi, de forma que wT ⋅ x + b = 1 e wT ⋅ x + b = -1 representam respectivamente os vetores de suporte positivos e negativos, que são os pontos próximos à margem do hiperplano ideal.
-A margem máxima, entre os vetores de suporte, é definida por:
-
-As funções de kernel têm o objetivo de projetar os vetores de variáveis de um conjunto de dados em um espaço de maior dimensão, para a classificação de classes originalmente apresentadas em espaços não separáveis linearmente.
-Com o aumento da dimensão, aumenta a probabilidade desses dados poderem ser linearmente separáveis.
-% Uma função kernel K(xi, xj) = φ(xi) T φ(xj) pode ser usada para treinar a MVS. Uma MVS linear tem φ(x) = x, assim uma função kernel linear pode ser representada por K(xi, xj) = xTi xj.
-Foi adotada a estratégia "um contra um" (Knerr, Personnaz e Dreyfus, 1990), para a utilização de MVS para a classificação de muitas classes.
-Esta estratégia consiste em construir uma máquina vetor de suporte para cada par de classes.
-Para um problema com c classes, c(c-1)/2 MVSs são treinados para classificar as classes entre as c classes possíveis.
-
-\subsection{CSF analysis based on SVM-RFE}
-
-Dado um algoritmo de classificação, que possa estimar pesos para as variáveis de um conjunto de dados, o objetivo da Eliminação Recursiva de Variáveis (ERV), do inglês Recursive Feature Elimination (RFE) (Guyon, 2002), é selecionar variáveis por meio da redução recursiva da quantidade de variáveis, eliminando recursivamente as variáveis de menor peso para classificação por meio do algoritmo adotado.
-Primeiramente, um modelo é treinado, utilizando o conjunto de dados inicial e o algoritmo selecionado, durante o treinamento são atribuídos pesos para cada variável, representando a importância de cada variável para a classificação.
-Em seguida, as variáveis com menor peso são eliminadas do conjunto de dados.
-Este processo, de treinamento, ordenamento de variáveis e eliminação de variáveis menos importantes, é repetido recursivamente até a obtenção do número desejado de variáveis, ou até a satisfação de alguma condição, como um limiar de taxa de erro de um algoritmo de classificação.
-As variáveis com maiores pesos apresentam maior influência na classificação (Guyon, 2002).
-Desta forma, se um algoritmo de classificação apresenta boa acurácia, as variáveis com maiores pesos representam as variáveis que apresentam maior influência para a classificação.
-MVS-ERV (Guyon, 2002) é uma aplicação da ERV utilizando os pesos obtidos por meio do treinamento utilizando MVS como algoritmo de classificação, para identificar as variáveis mais importantes para predições de classificação e eliminar recursivamente as variáveis que menos influenciam na classificação.
-
-Recursive feature elimination is based on the idea to repeatedly construct a model (for example an SVM or a regression model) and choose either the best or worst performing feature (for example based on coefficients), setting the feature aside and then repeating the process with the rest of the features.
-This process is applied until all features in the dataset are exhausted.
-Features are then ranked according to when they were eliminated.
-As such, it is a greedy optimization for finding the best performing subset of features.
-
-The stability of RFE depends heavily on the type of model that is used for feature ranking at each iteration.
-Just as non-regularized regression can be unstable, so can RFE when utilizing it, while using ridge regression can provide more stable results.
-
-%Given an external estimator that assigns weights to features e.g., the coefficients of a linear model, recursive feature elimination RFE is to select features by recursively considering smaller and smaller sets of features.
-First, the estimator is trained on the initial set of features and the importance of each feature is obtained either through a coef attribute or through a feature_importances attribute.
-Then, the least important features are pruned from current set of features.That procedure is recursively repeated on the pruned set until the desired number of features to select is eventually reached.
+The variables with the least effect on training errors or largest weights indicates more relevance for a classifier.
+Therefore, we propose to assume that the selected most relevant features are the CSF for iGovTI and also propose to validate this assumption against the results of the CSF qualitative analysis conducted by Ramos \emph{et al} \cite{ramos2016information}.
 
 \section{Experiments and Results}
-\label{sec:experimentalresults}
+\label{sec:ch2_experimentalresults}
 
-\subsection{Análise das componentes principais}
+In this section we present the experiements and results for the visual discriminant analysis based on PCA and for CSF analysis based on RFE, and discuss the results.
 
-Para a análise eficiente da matriz original  é relevante separar as informações úteis das redundantes.
-Segundo Johnson e Wichern (1992), há vários instrumentos para essa finalidade, mas a ACP é a que melhor desempenha este papel.
-Uma vez calculados os autovalores da matriz original, temos um resultado que mostra que aproximadamente 70\% da variabilidade dos dados é explicado por 51 componentes principais.
-O critério para a escolha desses fatores foi o de identificar os autovalores que possuem variância acumulada em torno de 70\% (Mardia, 1979).
-Tal valor também será utilizado nesta pesquisa (Quadro 7).
+For the visual discriminant analysis we initially we compute the zero mean and then compute the correlation matrix $\mathbf{C}$ is computed from $\mathbf{X}$, which is shown by Figure \ref{fig:ch2_fig1}.
 
@@
      \centering 
-     \includegraphics[height=6cm, width=9cm]{figures/pca-svm-rfe_raw_variance_ecdf.eps}
-     \caption{Empirical CDF of variance.}
-     \label{fig:fig1}
+     \includegraphics[height=10cm, width=10cm]{figures/ch2/raw_igovti_covariance}
+     \caption{Covariance matrix of iGovTI questions.}
+     \label{fig:ch2_fig1}
 \end{figure}
+
+The correlation matrix of 201 questions reveals
 
@@
 Verifica-se que uma queda menos acentuada ocorreu entre o quarto e o quinto autovalor e analisando-se os autovalores superiores a 2, observa-se que pode-se considerar até o vigésimo valor já que a partir daí os valores dos autovalores sucessivos são praticamente constantes.
+
+\begin{figure}[h!]
+     \centering 
+     \includegraphics[height=8cm, width=11cm]{figures/ch2/raw_variance_ecdf.eps}
+     \caption{Empirical CDF of variance.}
+     \label{fig:ch2_fig2}
+\end{figure}
 
@@
      \centering 
-     \includegraphics[height=6cm, width=9cm]{figures/pca-svm-rfe_raw_igovti_ranking_pc2.eps}
+     \includegraphics[height=8cm, width=11cm]{figures/ch2/raw_igovti_ranking_pc2.eps}
      \caption{iGovTI ranking from 2 principal components.}
-     \label{fig:fig1}
+     \label{fig:ch2_fig3}
 \end{figure}
@@
 
-\subsection{Máquinas de Vetores de Suporte}
-
 Para prever a classificação de uma instituição de acordo com suas respostas para as questões do questionário iGovTI, foram avaliados algoritmos de classificação que pudessem apresentar acurácia para a classificação de organizações de acordo com o iGovTI.
@@
 
-k-nearest neighbors \cite{fukunaga1975branch}	0.714286
-Elastic Net \cite{zou2005regularization}	0.831531
-Lasso \cite{tibshirani1996regression}	0.827440
-LinearRegression \cite{draper2014applied}	0.360878
-LogisticRegression \cite{hosmer2013applied}	0.771429
-LDA \cite{martinez2001pca}	0.628571
-SVM \cite{hearst1998support}	0.914286
-SVM Regression \cite{smola2004tutorial}	0.826885
-Linear SVM \cite{fan2008liblinear}	0.714286
+\begin{table*}[!t]
+	\caption{Evaluation of classification accuracy for iGovTI}
+  	\label{tab:ch2_tab3}
+	\centering
+	\begin{tabular}{|l|r|}
+		\hline \rowcolor{Gray} Algorithm	& Mean Accuracy\\\hline
+		Linear Regression \cite{draper2014applied}	&0.3608\\ \hline
+		LDA \cite{martinez2001pca}	&0.6285\\ \hline
+		K-Nearest Neighbors \cite{fukunaga1975branch}	&0.7142\\ \hline
+		Linear SVM \cite{fan2008liblinear}	&0.7142\\ \hline
+		Logistic Regression \cite{hosmer2013applied}	&0.7714\\ \hline
+		SVM Regression \cite{smola2004tutorial}	&0.8268\\ \hline
+		Lasso \cite{tibshirani1996regression}	&0.8274\\ \hline
+		Elastic Net \cite{zou2005regularization}	&0.8315\\ \hline
+		SVM \cite{hearst1998support}	&0.9142\\ \hline
+	\end{tabular}
+\end{table*}
+
+MVS-ERV (Guyon, 2002) é uma aplicação da ERV utilizando os pesos obtidos por meio do treinamento utilizando MVS como algoritmo de classificação, para identificar as variáveis mais importantes para predições de classificação e eliminar recursivamente as variáveis que menos influenciam na classificação..
+Falar que SVM-RFE é um algoritmo bem estabelecido e que concidentemente foi o melhor avaliado para este caso.
+SVM já é proposto para small datasets, o que foi validado com o nosso caso.
 
@@
 \section{Conclusion}
-\label{sec:conclusion}
+\label{sec:ch2_conclusion}
 
@@
 The selected features revealed by RFE are quite similar to the qualitative analysis of Ramos \emph{et al} \cite{ramos2016information}.
+
+Question with low variance can indicate that it is irrelevant for evaluation, since its value is almost the same all the time

### 2017-10-23 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
section 2.3

@@
      \centering 
-     \includegraphics[height=10cm, width=10cm]{figures/ch2/raw_igovti_covariance}
+     \includegraphics[height=10cm, width=10cm]{figures/ch2/raw_igovti_covariance.eps}
      \caption{Covariance matrix of iGovTI questions.}
@@
 
-The correlation matrix of 201 questions reveals
+Figure \ref{fig:ch2_fig1} presents the matrix presents of 201 questions of iGovTI, where is possible to observe low covariance for the majority questions and high variance for just a few variables.
+The low covariance reveals
 

### 2017-10-23 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
section 2.4...

@@
 Figure \ref{fig:ch2_fig1} presents the matrix presents of 201 questions of iGovTI, where is possible to observe low covariance for the majority questions and high variance for just a few variables.
-The low covariance reveals
 
-Depois da extração dos autovalores e percentual da variância explicada, é decidida a quantidade de fatores a serem retirados para análise.
-Para isso, o Gráfico 1 em que o total de autovalores está no eixo das ordenadas e os autovalores  no eixo das abscissas, auxilia na identificação.
-Esse gráfico consiste no ranking dos autovalores (eixo x), relacionado com o valor de cada autovalor (eixo y).
-Verifica-se que uma queda menos acentuada ocorreu entre o quarto e o quinto autovalor e analisando-se os autovalores superiores a 2, observa-se que pode-se considerar até o vigésimo valor já que a partir daí os valores dos autovalores sucessivos são praticamente constantes.
+In the next step the eigenvectors $\mathbf{v}_i$ and eigenvalues $\lambda_i$ of $\mathbf{C}$ are computed through SVD.
+Considering that the first principal components have the largest variance indicated by their eigenvalues, it is necessary to evaluate the value distribution of the measured eigenvalues.
+Therefore, we present the Figure \ref{fig:ch2_fig2}, wich shows the variance ECDF of the iGovTI questions and shows that 97\% of the questions have variance lesser than 2, while just 3\% of the questions have variance between 2 and 11.
+This result indicates that just a few principal components concentrates the most significative information and motivates the evaluation regarding the relationship between the principal components and the iGovTI classification.
 
@@
 
-Visando encontrar os planos fatoriais realizou-se uma rotação dos eixos, onde as cargas fatoriais mais elevadas são as responsáveis pelas denominações das componentes e são estatisticamente significativas.
-As rotações de eixos melhor expressam a dispersão de dados.
-No modelo fatorial final, as variáveis das medidas estão maximizadas e as relações entre dimensões suavizadas (Vicini, 2005).
-Para esta análise buscam-se valores que possuem significância maior que 0,7, mostrando que a correlação entre as variáveis está de moderada a forte (Vicini, 2005).
-Essa identificação não seria possível sem a rotação dos eixos, possibilitando assim a melhor visualização das variáveis mais significativas em cada componente.
-Tal rotação mantem os eixos perpendiculares entre si, ou seja, ortogonais e a variabilidade do sistema não é alterada, apenas as coordenadas dos eixos são rotacionadas e a inércia do sistema fica inalterada.
-A partir dos valores obtidos pela rotação das componentes principais, podem-se obter valores com significância maior que 0,7.
-Logo, foi possível a identificação das variáveis significantes de cada componente principal.
-Para visualização desses fatores foi utilizado o gráfico de dispersão (Gráfico 2).
-O Gráfico 2 mostra a caixa de seleção de variáveis e comandos para ACP em que se utilizam os fatores 1 e 2, eixo x e eixo y respectivamente.
-O objetivo deste gráfico é fazer os planos principais com a nuvem de pontos dos indivíduos, no caso as 349 instituições, destacando o posicionamento das respostas das instituições classificadas pelo iGovTI.
-Tal gráfico se baseia na rotação dos componentes principais.
-Para a elaboração do Gráfico 2 apenas os  e  foram utilizados.
+Finally, we select the two largest eigenvalues and their correspondent componets in order to evaluate the relationship between the iGovTI classification and the two most informative variables.
+The Figure \ref{fig:ch2_fig3} shows the scatter diagram that plots the organizations with higher iGovTI with larger circunferences and colors near of red, while organizations with lower iGovTI have colors near of blue and shorter circunferences.
 
@@
 
-No Gráfico 2, para analisar apenas a , projetam-se os pontos sobre o eixo da e se tem três grupos (V, II e I).
-O mesmo processo se aplica a analise da .
-Assim, observam-se três grupos (V, IV e III).
-Para cada grupo se seleciona apenas uma variável representativa, o restante é desconsiderado, uma vez que cada grupo significa correlação.
-Para o grupo I, tomam-se as variáveis Q16d, Q16e e Q16f que são descorrelacionadas, já no grupo II a questão Q16g é questão selecionada.
-Ambos os grupos não apresentam FCS.
-Essas questões denotadas Q16, referem-se a Subdimensão 1.6 (A Alta Administração Utilizou Informações Fornecidas pela Auditoria Interna).
-Tal subdimensão verifica a participação da auditoria interna das instituições para o preenchimento do questionário.
-% Já no grupo III, que não tem FCS, a variável Q53_Relev é a selecionada; precedidas do grupo IV que também não apresenta FCS.
-Ressalta-se que ambas as questões apuram a relevância de determinado item.
-Pela análise ACP, o grupo V mostra as questões que não são representadas pelas  e , é justamente este grupo que contem os FCS.
-Depois, buscou-se por meio da análise das ,  e a identificação de FCS (Gráfico 3). O Gráfico 3 mostra a análise dessas componentes principais que revela a existência de três grupos.
+O objetivo deste gráfico é fazer os planos principais com a nuvem de pontos dos indivíduos, no caso as 349 instituições, destacando o posicionamento das respostas das instituições classificadas pelo iGovTI.
 
-A análise do Gráfico 3 mostra revela a existência de três grupos (VI, VII e VIII). O grupo I contém a variável Q51l que corresponde a FCS e o grupo II também mostra FCS, a variável Q23b.
-Na identificação das variáveis significativas das primeiras 20 componentes principais, com os autovalores superiores a 1, as questões que mais contribuem são Q16b (para responder às questões do grupo 2.
-Estratégias e planos), Q16c (para responder às questões do grupo 3.
-Informação e conhecimento), Q16d (para responder às questões do grupo 4.
-Pessoas), Q16e (para responder às questões do grupo 5.
-Processos) e Q16f (para responder às questões do grupo 6.
-Resultados da gestão), todas da dimensão “Governança corporativa e de TI, a alta administração utilizou informações fornecidas pela auditoria interna (ou instância equivalente)”.
-Nota-se que as variáveis Q16, citadas anteriormente, fazem parte do grupo I do Gráfico 2.
-Ao total foram identificadas 29 variáveis dos 20 primeiras componentes principais (autovalores maiores que 1) das variáveis com significância maior que 0,7.
-O Quadro 8 mostra quais são essas variáveis significativas.
-
-No Quadro 8, verificou-se as variáveis que são FCS, de acordo com resultado da pesquisa apresentada na Subseção 4.1.
-Das 30 questões apenas 5 são FCS, de acordo com classificação obtida por meio de pesquisa qualitativa, quais sejam: Q12c (designou representantes de todas as áreas relevantes para o negócio institucional para compor o Comitê de TI); Q23b (a instituição aprovou e  publicou PDTI interna e externamente); Q51l (gestão de configuração de ativos); Q53e (formalizou a política corporativa de segurança da informação); Q57h (os pagamentos são feitos em função da mensuração objetiva dos resultados entregues e aceitos.
-Ainda no Quadro 8, chama-se atenção as ,  e.
-Nelas constam variáveis com valor positivo ou negativo.
-Tal fato significa que quanto maior for a resposta de uma questão menor será a da outra.
-Frisa-se que isto só vale para questões que estejam em uma mesma variável ACP.
-Caso as questões tenham sinais diferentes (i.e.
-os elementos dos autovetores), mas em ACP diferentes o sinal positivo e negativo não importa.
-A análise das categorias (Quadro 5) relacionadas às variáveis revela: Q53e referente à Categoria Processos de Gestão de Serviços de TI em Desenho do Serviço que mostrou porcentual igual a 53,85\%; Q57h se liga à Categoria Gestão de Contratos que teve porcentual de 38,46\%; Q23b se vincula à Categoria PDTI que teve 38,46\%.
-A variável Q51l se relaciona à Categoria Processos de Gestão de Serviços de TI em Transição de Serviços com 19,23\%.
-Devido ao baixo percentual de 16,66\% de identificação de FCS do Quadro 8, realizou-se a análise das variáveis que completam as 51 componentes principais, referentes aos 51 autovalores do Quadro 7.
-Nesta identificação, foram identificadas 33 variáveis, sendo que 8 são FCS, 24,24\% (Quadro 9).
-Somando os totais obtidos pelo Quadro 8 e pelo se tem, 20,63\% de variáveis que são FCS.
-
-Já o Gráfico 4 mostra a caixa de seleção de variáveis e comandos para ACP em que se utilizam os fatores 1 e 2, eixo x e eixo y respectivamente.
-O objetivo deste gráfico é fazer os planos principais com a nuvem de pontos dos indivíduos, no caso as 349 instituições.
-
-O Gráfico 4 representa a relação entre os 2 principais componentes, que representam as questões ou variáveis com maiores autovalores, assim cada ponto representa a relação entre uma instituição e seus 2 principais componentes obtidos por meio de ACP.
-Cada ponto deste gráfico representa a classificação da instituição em relação a sua classificação no iGovTI de 2012.
-A partir do resultado apresentado no gráfico 4 é possível identificar um padrão formado pela localização dos componentes principais das organizações com maior índice, onde as organizações com maior iGovTI tiveram valores relativamente semelhantes para seus 2 componentes principais, entretanto ainda não é possível obter uma separação clara entre os resultados.
-Fazendo-se um corte no eixo das abcissas, a partir do ponto (-8,8, 0), destacam-se 23 instituições à esquerda desse ponto (Quadro 5).
-
-Dessas instituições, 22, são consideradas pelo TCU como aprimoradas, ou seja, 95,65\%. Apenas a instituição 346 é avaliada como intermediária.
-Após a análise ACP, verifica-se que a mesma não é útil para a análise da Hipótese 2.
-Para responder a essa hipótese, é necessária a execução dos algoritmos de classificação da Subseção 4.4 e da Subseção 4.5.
+...
 
@@
 
-MVS-ERV (Guyon, 2002) é uma aplicação da ERV utilizando os pesos obtidos por meio do treinamento utilizando MVS como algoritmo de classificação, para identificar as variáveis mais importantes para predições de classificação e eliminar recursivamente as variáveis que menos influenciam na classificação..
-Falar que SVM-RFE é um algoritmo bem estabelecido e que concidentemente foi o melhor avaliado para este caso.
-SVM já é proposto para small datasets, o que foi validado com o nosso caso.
-
 O algoritmo que obteve maior sucesso para classificação foi o SVC, que é uma implementação de Máquina de Vetores de Suporte aplicado para a classificação.
@@
 Para isso, será utilizado o algoritmo ERV.
+MVS-ERV (Guyon, 2002) é uma aplicação da ERV utilizando os pesos obtidos por meio do treinamento utilizando MVS como algoritmo de classificação, para identificar as variáveis mais importantes para predições de classificação e eliminar recursivamente as variáveis que menos influenciam na classificação.
+Falar que SVM-RFE é um algoritmo bem estabelecido e que concidentemente foi o melhor avaliado para este caso.
+SVM já é proposto para small datasets, o que foi validado com o nosso caso.
 

### 2017-10-24 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
section 2.4

@@
 
-For the visual discriminant analysis we initially we compute the zero mean and then compute the correlation matrix $\mathbf{C}$ is computed from $\mathbf{X}$, which is shown by Figure \ref{fig:ch2_fig1}.
+For the visual discriminant analysis we initially we compute the zero mean and the the covariance matrix $\mathbf{C}$ is computed from $\mathbf{X}$.The resuls are shown by Figure \ref{fig:ch2_fig1}.
 
@@
 
-Figure \ref{fig:ch2_fig1} presents the matrix presents of 201 questions of iGovTI, where is possible to observe low covariance for the majority questions and high variance for just a few variables.
+Figure \ref{fig:ch2_fig1} presents the covariance matrix of 201 questions of iGovTI, where is possible to observe low covariance for the majority questions and high variance for just a few questions.
 
 In the next step the eigenvectors $\mathbf{v}_i$ and eigenvalues $\lambda_i$ of $\mathbf{C}$ are computed through SVD.
-Considering that the first principal components have the largest variance indicated by their eigenvalues, it is necessary to evaluate the value distribution of the measured eigenvalues.
+Considering that the first principal components have the largest variance indicated by their eigenvalues, it is necessary to evaluate the variance distribution of the measured eigenvalues.
 Therefore, we present the Figure \ref{fig:ch2_fig2}, wich shows the variance ECDF of the iGovTI questions and shows that 97\% of the questions have variance lesser than 2, while just 3\% of the questions have variance between 2 and 11.
@@
 
-Finally, we select the two largest eigenvalues and their correspondent componets in order to evaluate the relationship between the iGovTI classification and the two most informative variables.
+Finally, we select the two largest eigenvalues and their correspondent componets in order to evaluate the relationship between the iGovTI classification for 349 organizations and the two most informative variables.
 The Figure \ref{fig:ch2_fig3} shows the scatter diagram that plots the organizations with higher iGovTI with larger circunferences and colors near of red, while organizations with lower iGovTI have colors near of blue and shorter circunferences.
+It is possible to observe that organizations with higher iGovTI are concentrated in the top left quadrant, with a visual separation of organizations with lower iGovTI, but without a clear distance separating the classes.
 
@@
 
-O objetivo deste gráfico é fazer os planos principais com a nuvem de pontos dos indivíduos, no caso as 349 instituições, destacando o posicionamento das respostas das instituições classificadas pelo iGovTI.
+The visual discriminant analysis indicates that just a few principal components concentrates the most significative information and that the two principal components show a visual classification of organizations according to iGovTI classification, however it is not clear what are the CSF.
 
-...
+Hence, it is necessary additional techniques for CSF identification and we propose a critical success factors analysis based on RFE for identification of the most relevant factors for iGovTI.
 
-Para prever a classificação de uma instituição de acordo com suas respostas para as questões do questionário iGovTI, foram avaliados algoritmos de classificação que pudessem apresentar acurácia para a classificação de organizações de acordo com o iGovTI.
-Uma vez encontrado um algoritmo capaz desta classificação, é possível utilizar o algoritmo selecionado em conjunto com técnicas de feature selection para identificar as questões mais relevantes para a classificação da instituição de acordo com o iGovTI.
-Assim, para a definição do algoritmo de classificação foram realizados testes em 21 algoritmos, a seguir é apresentada uma listagem dos algoritmos avaliados e a taxa de acerto de classificação obtida:
+RFE requires a classifier or predictor to evaluate its generalization performance by recursivelly removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features.
+We perform an algorithm evaluation in order to identify the classifier with the highest accuracy for iGovTI classification.
+The evaluated dataset is composed of 201 questions (features) and 349 organizations (observations), where each organization is classified according to iGovTI classes, which are initial, intermediate and enhanced.
+We divide the dataset into train and test, with the training data equivalent to 90\% of the whole dataset, while the test has 10 \% of the whole dataset.
+
+The Table \ref{tab:ch2_tab1} present the selected classification algoritms and the measured accuracy for iGovTI classification.
+According to results, SVM \cite{hearst1998support} is the algorithm with highest classification accuracy, with 91.42\%, while the second is Elastic Net \cite{zou2005regularization}, with 83.15\% of accuracy.
+The results for iGovTI classification corroborates previous evaluations that highlight the advantages of SVM for classifications from small datasets \cite{guyon2002gene}.
+Additionally, our results indicate that SVM is the best classifier for RFE, which corroborates the previous research of Guyon \emph{et al} \cite{guyon2002gene}, where is proposed SVM-RFE, that is a method of gene selection utilizing SVM methods based on RFE.
 
@@
 	\caption{Evaluation of classification accuracy for iGovTI}
-  	\label{tab:ch2_tab3}
+  	\label{tab:ch2_tab1}
 	\centering
@@
 
-O algoritmo que obteve maior sucesso para classificação foi o SVC, que é uma implementação de Máquina de Vetores de Suporte aplicado para a classificação.
-SVC apresentou uma taxa de acerto de 91,4\%.
-Para a avaliação dos algoritmos, utilizamos uma metodologia que divide os dados dos questionários pelo iGovTI entre questões que serão utilizadas para aprendizado do algoritmo e questões que serão utilizadas para comparativo de predições.
-Para treinar o algoritmo utilizou-se 90\% dos dados e os 10\% restantes foram utilizados para avaliar a eficiência de classificação dos algoritmos, comparando a taxa de acerto entre as predições feitas e os valores reais de classificações de organizações pelo iGovTI.
+The next step is to use the selected algorithm as validator for recursive feature elimination in order to identify the CSF for iGovTI.
+Additionally, the RFE algorithm requires the number of desired most significatives features.
+We adopt 54 for this variable, considering that is the number of CSF identified by Ramos \emph{et al} \cite{ramos2016information} in a qualitative analysis, that is the proposed target to validate the results of the CSF identification for iGovTI.
 
-Uma vez identificado um algoritmo capaz de efetuar a classificação desejada, o próximo passo é identificar as variáveis mais relevantes para esta classificação.
-Para isso, será utilizado o algoritmo ERV.
-MVS-ERV (Guyon, 2002) é uma aplicação da ERV utilizando os pesos obtidos por meio do treinamento utilizando MVS como algoritmo de classificação, para identificar as variáveis mais importantes para predições de classificação e eliminar recursivamente as variáveis que menos influenciam na classificação.
-Falar que SVM-RFE é um algoritmo bem estabelecido e que concidentemente foi o melhor avaliado para este caso.
-SVM já é proposto para small datasets, o que foi validado com o nosso caso.
-
-A ERV pode usar vários algoritmos de classificação como critério de seleção das variáveis mais importantes, escolhemos o algoritmo SVC por ele ter apresentado maior acurácia entre os algoritmos de classificação avaliados.
-Ainda é necessário definir o quantitativo de variáveis mais importantes a serem selecionadas.
-A partir da pesquisa de identificação por meio de entrevistas na APF, identificou-se 54 variáveis consideradas FCS, este critério foi utilizado para determinar o quantitativo de variáveis mais importantes para a classificação.
-
-Assim, aplicaram-se o algoritmo ERV utilizando o SVC como critério para a seleção das variáveis mais importantes para a classificação, selecionando as 54 variáveis que correspondem aos FCS levantados nas entrevistas com os executivos de TI.
-Os resultados mostraram que 69,9\% das variáveis foram classificadas da mesma forma que os FCS identificados anteriormente (Quadro 10) por meio de pesquisa qualitativa.
+The RFE algorith presents an ranking according to the significance of each feature, with an accuracy of 69,9\% for CSF classification when compared to the CSF identified by Ramos \emph{et al} \cite{ramos2016information} in a qualitative analysis.
 

### 2017-10-24 docs/thesis/chapters/appendix/a_mos/a_mos.tex
appendix configuration

--- 2017-09-19 docs/thesis/chapters/3_eigen_similarity/capitulo3.tex
@@
-\chapter{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks} 
-\label{chapter3}
+\chapter{Model Order Selection (MOS)}
+\label{sec:a_mos}
 
-\section{Introduction}
-\label{sec:introduction}
-
-Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}.
-In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time.
-Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
-Recently, signal processing schemes have being applied to the detection of malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest}, showing advances in network traffic analysis.
-
-Information security may consist of both technical and procedural aspects.
-The former includes equipment and security systems, while the latter corresponds to security rules and recommendations.
-Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks.
-Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.
-Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
-
-In the context of anomaly-based schemes, this work proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
-Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
-To the best of our knowledge there is no similar model in the literature.
-The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
-In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack.
-In addition, we also evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset.
-Furthermore, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.
-
-The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
-
-This paper is organized as follows.
-In Section \ref{sec:relatedworks}, related works are discussed.
-Section \ref{sec:datamodel} presents the data model and the evaluated datasets.
-Section \ref{sec:prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks.
-Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:Complexity} discusses the computational complexity of the proposed framework and evaluates the required processing time for tested scenarios.
-Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
-The \ref{sec:mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.
-
-\section{Related Works}
-\label{sec:relatedworks}
-
-Several methods have been proposed for the identification and characterization of malicious activity in computer networks.
-Classical methods typically employ data mining \cite{he2008applying,ghourabi2010data,osanaiye2016distributed} and regular file analysis \cite{raynal2004honeypot} to detect patterns that indicate the presence of specific attacks in network traffic.
-
-Data mining is often used to describe the process of extracting useful information from large databases.
-Multiple methods of data mining are used in \cite{he2008applying,osanaiye2016distributed} to analyze data flow in a network with the aim of identifying characteristics of malicious traffic in large scale environments.
-Researchers have applied data mining techniques in log analysis \cite{ghourabi2010data} to improve intrusion detection performance.
-However, data mining techniques used so far in network analysis require prior collection of large data sets, which is a limitation of several schemes for online analysis \cite{Huang2009}.
-
-Regular file analysis \cite{raynal2004honeypot} consists of traffic analysis for detecting known patterns that indicate the presence of attacks, applying statistical analysis to the study of collected traffic.
-An essential feature of this method is that it depends on prior knowledge of the details of the attacks to be identified, and also depends on previous log collection for traffic analysis and false positives reduction.
-
-Principal Component Analysis (PCA) is a statistical technique commonly used for dimensionality reduction.
-It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
-PCA has been used in attack detection \cite{almotairi2009technique}.
-However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques.
-
-Callegari \emph{et al} \cite{Zonglin2009} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
-However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
-
-Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
-In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
-Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
-
-Besides being prone to higher errors and false positives, such human intervention makes PCA not useful for real time applications.
-Therefore, in order to automate the analysis of eigenvalues profile, model order selection (MOS) schemes should be incorporated.
-
-Signal processing techniques have been successfully applied to network anomaly detection \cite{Lu2009}.
-Lu and Ghorbani \cite{Lu2009} proposed a network anomaly detection model based on network flow, wavelet approximation, and system identification theory.
-However, their work requires a training method to produce a prediction model for normal daily traffic and presents limitations on identification of behaviors without significant outliers, such as port scan attacks.
-Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
-Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work is not applied to probe and flood attack detection, simultaneously.
-
-The data collected from honeypot systems, such as captured traffic and operating system logs, can be analyzed to obtain information about attack techniques, general trends of threats and exploits.
-Blind automatic detection of malicious traffic techniques have been developed for honeypots in \cite{david2011blind,da2012improved}.
-However, traffic on honeypot is simpler than real network traffic, because there are no running legitimate applications, due to the fact that honeypots emulate behavior of a host within a network to deceive and lure attackers \cite{zakaria2012review}.
-Since honeypots do not generate legitimate traffic, the amount of data captured in honeypots is significantly lower in comparison to a Network Intrusion Detection System (NIDS), which captures and analyzes the largest possible amount of network traffic \cite{david2011blind}.
-MOS for blind identification of malicious activities in honeypots was proposed by us in \cite{david2011blind}, which evaluated criteria for selecting the model order, through simulations and comparing the order of the resulting model with the true model order.
-
-The proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic.
-The proposed solution is automatic and blind for detection of time frames under probe and flood attacks through MOS and eigen analysis.
-Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.
-
-Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks.
-Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
-NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications.
-Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations.
-Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD datasets adopt a traffic aggregation by connections, we consider the use of a experimental scenario on a real network and the DARPA 1998 dataset, which is the source for the creation of the KDD 99 and NSL-KDD datasets.
-Note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD datasets.
-
-\section{Data Model}
-\label{sec:datamodel}
-
-In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
-The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
-We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz, and the operator $\sim$ that sorts the elements of a vector in ascending order.
-
-This section presents details of the experimental scenario and the selected cases of the DARPA dataset, along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic.
-Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
-Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition.
-Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
-
-\subsection{Analyzed Scenario and Data Collection}
-\label{sec:DataCollection}
-
-The environment of the analyzed scenario is composed of two computers and one router with access to the Internet and to an internal network, where the simulation of legitimate traffic, noise, flood and port scan attacks are performed.
-During the traffic generation, one computer assumes the role of the attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:fig1}.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=6cm, width=9cm]{results/figures/fig09.png}
-     \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
-     \label{fig:fig1}
-\end{figure}
-
-Note that the set of network traffic is modeled as legitimate, noise and malicious traffic, where the victim performs legitimate activities, that can be characterized by web access.
-In many organizations this type of traffic is predominant, since most of corporate services are web-based, such as: web pages, customized web-based systems and cloud services.
-It is possible to characterize the traffic of a DHCP service as an example of noise associated with the transport layer.
-For malicious traffic, three types of networks attacks are evaluated: synflood, fraggle and port scan.
-These attacks are reproduced using well-known security tools, such as Nmap\footnote{http://nmap.org} to port scan, Metasploit\footnote{http://www.metasploit.com} to synflood and Hping\footnote{http://hping.org} to lead the fraggle attack.
-
-A network traffic log is commonly formed by timestamp, protocol, source IP address, source port, destination IP address, destination port and additional information, according to the type of the used transport protocol.
-The following TCP traffic log is presented in order to exemplify the collected data:
-\newline
-\newline
-\texttt{\justify21:00:34.099289 IP 192.168.1.102.34712 > 200.221.2.45.80: Flags [S], seq 2424058224, win 14600, options [mss 1460, sackOK,TS val 244136 ecr 0,nop,wscale 7], length 0}
-\newline
-
-and the following to exemplify UDP traffic log: 
-\newline
-\newline
-\texttt{\justify21:24:42.484858 IP 192.168.1.102.68 > 192.168.1.1.67: BOOTP/DHCP, Request from 00:26:9e:b7:82:be, length 300}
-\newline
-
-In the proposed framework, the goal is to detect the anomalies only taking into account the traffic profile, i.e., specific information such as origin IP or day and time of the attack are not considered.
-Therefore, from the entire log information, we just consider the timestamp (for sequencing), port type and port number.
-
-\subsection{Modeling Data}
-\label{sec:ModelingData}
-
-By modeling the dataset as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{U}), noise (\textbf{N}) and malicious traffic (\textbf{A}), according to the following expression:
-
-\begin{equation}\label{eq:eq01}
-\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{N}^{(q)} + \boldsymbol{A}^{(q)},
-\end{equation}
-where $q$ represents the $q$-th time frame, which is a time grouping of network traffic.
-The matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns.
-Each row represents a communication port (TCP port or UDP port), and each column represents time bins having a approtiate size, such as one minute.
-Each element $x_{m,n}^{(q)}$ stands for the number of times that the port $m$ appears at the $n$-th minute, at the $q$-th time frame.
-
-The legitimate traffic $\boldsymbol{U}^{(q)}$ is characterized by the traffic from user's operations.
-When a user accesses a web page, for example, there is the corresponding TCP/IP traffic to request the page, as well as there is the traffic required to domain name resolution.
-Figure \ref{fig:fig3} depicts an example of the legitimate traffic obtained during experiments.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig03.png}
-     \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
-     \label{fig:fig3}
-\end{figure}
-
-The traffic that is not associated with user's operations and with malicious traffic is modeled as noise $\boldsymbol{N}^{(q)}$.
-The automatic acquisition service of logical IP network address (DHCP) is an example of noise.
-Independently of any user operation, the machine receives an IP address, since it is configured to perform a DHCP address request.
-Figure \ref{fig:fig4} depicts an example of noise in a network traffic, represented by traffic to ports 67 and 68.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig04.png}
-     \caption{Network traffic of user independent operations for network management.}
-     \label{fig:fig4}
-\end{figure}
-
-The traffic coming from a malicious activity, such as a synflood or fraggle attack, is represented by the matrix $\boldsymbol{A}^{(q)}$.
-For this work we only consider the traffic from port scanning and flood attacks.
-
-We define that if the obtained $\#\boldsymbol{A}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\boldsymbol{A}^{(q)} = 0$, then there is no malicious traffic.
-This paper shows how to detect the $\#\boldsymbol{A}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious network traffic.
-
-\subsection{Synflood, Fraggle and Port scan}
-\label{sec:SynfloodFraggleandPortscan}
-
-The network attacks evaluated by this work are: synflood, fraggle and port scan.
-The first two attacks can be qualified as flood or denial of service (DoS) attacks, while the last one can be qualified as probe or port scanning attack.
-DoS attempts to block access to system or network resources and is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks.
-Probe attacks scan the networks to collect information about host, such as IP addresses, ports and services.
-
-With respect to the synflood attacks, the attacker sends a large quantity and concurrent successive SYN requests to a target, in order to consume resources and cause a DoS.
-Figure \ref{fig:fig5} depicts an example of a synflood attack carried out in a real computer network.
-In an interval of ten minutes, more than 210,000 packets are sent as a synflood attack.
-This network traffic behavior can be considered an abnormal behavior of network traffic, especially since it is concentrated in a short period of time and presents similar outstanding traffic during the time under attack.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig05.png}
-     \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
-     \label{fig:fig5}
-\end{figure}
-
-With respect to the fraggle attack, large packets with UDP echo segments are sent to the broadcast address of a network.
-Every packet is modified to have the source address of the victim, in order to implement the source address spoofing technique.
-Therefore, each host receives a huge amount of requests UDP echo and all of them replies to the IP address of the victim, causing a packet flooding aiming a DoS.
-This attack can affect the entire network, since all hosts receive several requests UDP echo and respond with the ICMP protocol, therefore each host acts as an amplifier of the attack.
-This last part of the fraggle attack is not taken into account in this work, because the victim receives ICMP (network layer) packets originated from the hosts that are attacked with flooding packet UDP echo.
-Figure \ref{fig:fig6} depicts an example of the fraggle attack in a real computer network.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig06.png}
-     \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
-     \label{fig:fig6}
-\end{figure}
-
-More than 6,000,000 malicious packets can be counted in an interval of ten minutes, which can be considered an abnormal network traffic, especially due to the concentrated traffic in a short period of time and due to the similarity of the outstanding traffic.
-
-Port scan is the attempt to establish a connection to TCP and UDP ports to identify what services are running or are in the listening state.
-There are several available port scanning techniques, including: TCP SYN scan, TCP ACK scan and UDP scan.
-This work evaluates the use of TCP SYN scan and UDP scan.
-
-In TCP SYN scan, a SYN packet is sent to the destination and two types of responses may occur: SYN/ACK or RST/ACK.
-In the first case, the destination port is in the listening state, in the second case, the destination port is not listening.
-At the end of each port scanning, a RST/ACK packet is sent by the system that is performing the port scan.
-Therefore, a full connection or a complete three-way handshake is never established, which makes the detection of the attack sender more difficult, and requires approaches able to identify probe attacks without connection stablishment.
-The UDP scan technique sends UDP packets to the destination port, and if it responds with a \emph{ICMP port unreachable} message, then it indicates that the scanned port is closed.
-On the other hand, if a message is not received, then the port is considered as open.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=5cm, width=9cm]{results/figures/fig07.png}
-     \caption{Connection attempts in order to identify active ports.}
-     \label{fig:fig7}
-\end{figure}
-
-Figure \ref{fig:fig7} depicts an example of the port scan attack in a real computer network.
-Note that the traffic is composed of two packets for each TCP port and one UDP packet to each port.
-The incoming and outgoing packets analysis, for each port, shows the high correlation and similarity of TCP and UDP traffic during the simulated port scan attack.
-
-\subsection{The DARPA Dataset}
-\label{sec:DarpaDataset}
-
-The DARPA 1998 dataset\footnote{https://www.ll.mit.edu/ideval/data/} includes 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks.
-The attacks in this dataset
-can be grouped into: denial-of-service (DoS); remote to local (R2L), which is characterized by unauthorized access from a
-remote machine; user to root (U2R), which is characterized by unauthorized access to local
-super-user privileges; and probe attack.
-Since the proposed approach focus on flood and probe attack, the analysis concentrates on the attacks that present behaviors similar to flood or probe attack.
-We observe that the most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack.
-One example is the occurrence of a neptune attack which sends 20 only SYN packets, what is a behavior that differs of the expected flood attack behavior.
-Therefore, there were selected the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey,osanaiye2016distributed}, and the cases that scan ports sending just a few packets.
-From the simulated probe attacks, we select the cases that rely on TCP or UDP connections.
-
-The data modeling follows the method described by the Subsection \ref{sec:ModelingData}, with time frames of 20 minutes, packet counting aggregation by minute and considering the traffic to the following ports: 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389, 443.
-
-\section{Proposed Framework for Detection and Identification of Network Attacks}
-\label{sec:prop_getv}
-
-This section describes the proposed technique to detect synflood, fraggle and port scan, according to Figure \ref{fig:fig80}, which represents the overview of the proposed framework for detection and identification of network attacks.
-In Subsection \ref{sec:prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvelue for each $q$-th time frame.
-Next, in Subsection \ref{sec:prop_MOSSchemes}, we show how to apply the eiganvalues on the MOS scheme in order to detect the attack.
-In Subsection \ref{sec:prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:prop_EigenSimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=11cm, width=9cm]{results/figures/mos_eigen_similarity.eps}
-     \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
-     \label{fig:fig80}
-\end{figure}
-
-\subsection{Largest Eigenvalue by Time Frames}
-\label{sec:prop_LargestEigenvaluebyTimeFrames}
-
-The proposed attack detection algorithm starts by the data pre-processing of a network traffic log containing IP, ports and timestamp of senders and receivers.
-During this step, the desired information is extracted in order to classify and count packets according to the origin and destination ports, and subsequently this information is grouped by minutes and by time frames.
-
-With the data grouped into $Q$ time frames, the framework considers the time variations of the matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the attack.
-
-According to flood and port scan attacks' behavior, flood attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively.
-These characteristics are substantiated by the results obtained through the analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
-
-Therefore, to detect flood attacks, it is necessary to calculate the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
-
-\begin{equation}\label{eq:eq02}
-\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
-\end{equation}
-
-The set of obtained vectors $\boldsymbol{y}_{m}^{(q)}$ composes the zero mean matrix $\boldsymbol{Y}^{(q)}$, then the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
-
-\begin{equation}\label{eq:eq03}
-\boldsymbol{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
-\end{equation}
-
-For the detection of the port scan attack, the main components are not dominated by the variables with large variance.
-Moreover, the portscan traffic presents a highly correlated network traffic.
-In order to exploit such structure, we compute the sample covariance $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have zero mean and unitary standard deviation as follows
-
-\begin{equation}\label{eq:eq04}
-\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
-\end{equation}
-
-The set of vectors $\boldsymbol{z}_{m}^{(q)}$ composes the matrix $\boldsymbol{Z}^{(q)}$, then the sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)}$ can be calculated via
-
-\begin{equation}\label{eq:eq05}
-\boldsymbol{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
-\end{equation}
-
-Once the $\boldsymbol{\hat{R}}_{yy}^{(q)}$ and $\boldsymbol{\hat{R}}_{zz}^{(q)}$ have been obtained for flood and port scan attack detection, respectively, and since the next steps are the same for both sample covariance matrices, we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{C}$.
-Therefore, the following step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:eq06}), in order to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).
-
-\begin{equation}\label{eq:eq06}
-\boldsymbol{C}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
-\end{equation}
-
-\begin{equation}\label{eq:eq060}
-\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
-\end{equation}
-
-where the operator diag$(\cdot)$ extracts the main diagonal of a matrix.
-
-The eigenvalues should be sorted in descending order, i.e., $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ...
-> \lambda_{m}^{(q)}$.
-Therefore, the largest eigenvalue of the $q$-th time frame evaluated for the attack detect is given by $\lambda_{1}^{(q)}$.
-
-The concatenation of the eigenvalues vector $\boldsymbol{e}^{(q)}$ for $q = 1, \ldots, Q$ is represented by
-
-\begin{equation}\label{eq:eq07}
-\boldsymbol{E} =
-\begin{bmatrix}
-  \lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
-  \lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
-  \lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
-  \vdots & \vdots & \ddots & \vdots  \\
-  \lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
-\end{bmatrix}.
-\end{equation}
-
-Note that since $\lambda_1^{(q)} > \lambda_2^{(q)} > \lambda_3^{(q)} > \cdots > \lambda_{m-1}^{(q)} > \lambda_m^{(q)}$, then the first line of the matrix $\boldsymbol{E}$ contains the largest eigenvalues of each $q$-th time frame, which is the Greatest Eigenvalue Time
-Vector (GETV) \cite{tenorio2013greatest}, denoted as
-
-\begin{equation}\label{eq:eq08}
-\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
-\end{equation}
-
-\subsection{MOS Schemes}
-\label{sec:prop_MOSSchemes}
-
-Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$.
-However, the goal here is to detect the variations of the eigenvalues for different values of $q$.
-Therefore, instead of using a certain $q$, the proposed approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame, in order to identify variations and estimate the model order $\hat{d}$, which is the estimated number of time frames under attack.
-Therefore, $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
-Note that some MOS schemes may also require the number of minutes that compose a time frame, as $\hat{d} = \rm{MOS}(\boldsymbol{e}_{\rm max},\emph{Q})$.
-For more information about MOS, we refer to \ref{sec:mos}.
-
-In our previous work \cite{tenorio2013greatest}, the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE schemes are evaluated for synflood and port scan attack detection, showing that EDC and EFT are effective for detecting this kind of attacks.
-The present work extends that evaluation to also analyze the effectiveness of the listed MOS schemes for fraggle attack detection, as shown in Section \ref{sec:experimentalresults}.
-
-\subsection{Eigenvalue Analysis}
-\label{sec:prop_EigenvalueAnalysis}
-
-After applying the MOS schemes to the vector $\sim\boldsymbol{e}_{\rm max}$, we obtain the estimate of the $\#\boldsymbol{A}$.
-For instance, in the case of fraggle, synflood and portscan, if $\hat{d} = 1$, , then $\#\boldsymbol{A} = 1$, which means that during the during the $Q$ time frames one attack is present.
-However, if $\hat{d} = 0$, then $\#\boldsymbol{A} = 0$, and this means that none of these attacks are present.
-Note that $\hat{d}$ can be greater than 1, indicating the presence of more than one attack.
-
-In Subsection \ref{sec:prop_MOSSchemes}, we obtained only if $\hat{d} = 1$ or $\hat{d} = 0$.
-However, if $\hat{d} = 1$, the MOS schemes do not provide any information about the $q$-th time frame under attack.
-The identification of the $q$-th time frame under attack can be carried out through a eigenvalues analysis.
-
-The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed according to Algorithm \ref{alg:alg01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$.
-Algorithm \ref{alg:alg01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$, according to Line 2 of Algorithm \ref{alg:alg01}, and its correspondent index, according to Lines 4 and 5 of Algorithm \ref{alg:alg01}.
-Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$, according to Line 8 of Algorithm \ref{alg:alg01}, and a new iteration is performed until $\boldsymbol{e}_{\rm max} = \{\}$.
-
-\begin{algorithm}[h!]
-	\caption{Detection of Time Frames Under Attack}
-  	\label{alg:alg01}
-	\begin{algorithmic}[1]
-		\show\LOOP
-	    \ALOOP {$f = 1$ until $f == \hat{d}$} 
-	    	    \STATE $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
-    	    	    \ALOOP {$i = 1$ until $i == Q$} 
-				\IF {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
-				    \STATE $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
-				\ENDIF
-        		\ENDALOOP
-	    		\STATE $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
-    		\ENDALOOP
-	\end{algorithmic}
-\end{algorithm}
-
-After the estimation of the $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked.
-To deal with this problem, the adoption of a similarity analysis between legitimate traffic and the traffic of time frames estimated as under attack is evaluated, analysing the effectiveness of cosine similarity to highlight abnormalies inserted by network traffic attacks.
-
-\subsection{Eigen Similarity Analysis}
-\label{sec:prop_EigenSimilarityAnalysis}
-
-Cosine similarity calculates the cosine of the angle between two vectors, which represents the similarity of values between the selected vectors.
-Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\boldsymbol{V}^{(q)}$ against the the most significant eigenvectors of time frame detected as under attack, to analyze similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.
-
-This subsection describes the proposed eigen similarity analysis for detailed attack identification, in complement to the attack estimation carried out through MOS schemes and eigenvalue analysis.
-In Subsection \ref{sec:prop_TimeSimilarityAnalysis} we present the eigen similarity analysis for identification of time under attack.
-Next, in Subsection \ref{sec:prop_PortSimilarityAnalysis}, we show how to apply the eigen similarity analysis in order to identify network ports under attack.
-
-\subsubsection{Time Similarity Analysis}
-\label{sec:prop_TimeSimilarityAnalysis}
-
-For eigen similarity analyis, we evaluate the cosine similarity in order to identify lacks of similarity between legitimate and malicious traffic, as follows
-
-\begin{equation}\label{eq:eq11}
-s_n = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}}},
-\end{equation}
-where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, and $\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of traffic to be performed the flood and port scan attack identification.
-
-The most significant eigenvector $\boldsymbol{v}^{(q)}$, of a time frame $q$ without attack, can be derivated from (\ref{eq:eq06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the evaluated matrix.
-The same calculation shall be performed in order to obtain the target eigenvectors $\boldsymbol{v}_{(n)}$, calculated from a time frame without attack plus minutes of a time frame estimated as under attack, to evaluate the occurrence of network attacks.
-
-The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic whithout attack, from a time frame $q$ composed of $Q$ minutes of legitimate network traffic.
-For the detailed attack identification, each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$, as represented by
-
-\begin{equation}\label{eq:eq12}
-\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
-\end{equation}
-
-The resultant $\boldsymbol{X}_{(n)}$ is necessary to obtain $\boldsymbol{v}_{(n)}$, through (\ref{eq:eq06}), for calculating the similarity degree $s_n$, ranging from 0 to 1, for each $n$-th minute.
-The $s_n$ denotes the absolute similarity degree of the $n$-th minute in comparison to a well-known traffic without attack, detected through MOS schemes and eigenvalue analysis.
-
-The incremental approach for similarity analysis is based on the incremental appending of network traffic into $\boldsymbol{X}^{(q)}$, where the first evaluation is based on (\ref{eq:eq12}) and the subsequent evaluations is based on (\ref{eq:eq13}), incrementally appending each $n$-th minute until $n=N$.
-
-\begin{equation}\label{eq:eq13}
-\boldsymbol{X}_{n} = \{\boldsymbol{X}_{n} | \boldsymbol{x}^{(\hat{q})}_{(n)}\},
-\end{equation}
-
-Figure \ref{fig:fig8} ilustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.
-
-\begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental.eps}
-     \caption{Traffic selection for incremental approach.}
-     \label{fig:fig8}
-\end{figure}
-
-The eigen similarity analysis starts at $\boldsymbol{x}^{(3)}_{(41)}$ and is incrementally performed until $\boldsymbol{x}^{(\hat{3})}_{(60)}$, in order to calculate the $s_n$.
-We assume that $s_n < l$ means an attack identification, according the anomaly on similiarity of $s_n$ to a defined limiar $l$.
-Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}$ for eigen similarity analyis, the $s_n$ is calculated according to~(\ref{eq:eq11}).
-
-If $s_n = 1$, then the two eigenvectors are completely similar and no anomaly is detected. Smaller values of $s_n$ mean less similarity and can indicate an anomaly, according to a defined 
-threshold $l$, assuming that if $s_n < l$, then a network attack is identified during the $n$-th minute.
-Therefore, the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack is identified, according to
-
-\begin{equation}\label{eq:eq14}
-  \boldsymbol{\hat{n}_{(n)}}=\left\{
-  \begin{array}{@{}ll@{}}
-    1, & \text{if}\ s_n < l \\
-    0, & \text{otherwise}
-  \end{array}\right.
-\end{equation}
-where $\boldsymbol{\hat{n}}_{(n)}$ denotes a vector of $n$-th minutes detected as under attack.
-
-The eigen similarity analysis can also be applied in an individual fashion, where each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, as shown by Figure \ref{fig:fig9}.
-
-\begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/individualized.eps}
-     \caption{Traffic selection for individual approach.}
-     \label{fig:fig9}
-\end{figure}
-
-The incremental and the individual approaches can be combined to obtain the incremental individualized approach, where each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
-Subsequently, $\boldsymbol{X}_n$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:fig2}.
-
-\begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental_individualized.eps}
-     \caption{Traffic selection for incremental individualized approach.}
-     \label{fig:fig2}
-\end{figure}
-
-This approach of incremental similarity analysis followed by individual analysis after an attack detection allows to identify the attack period, highlighting the first and last time under attack.
-This identification is possible due to the variation of the most significant eigenvectors, which becomes more significant when compared a traffic under attack against a traffic with no attack, according to results which are discussed in Section \ref{sec:experimentalresults}.
-
-\subsubsection{Port Similarity Analysis}
-\label{sec:prop_PortSimilarityAnalysis}
-
-Given $\boldsymbol{\hat{n}}$, which is the set of $n$-th minutes under attack, it is still necessary to obtain more details about the identified network attack, such as the network ports that are attacked during each $n$-th minute identified as under attack.
-Hence, it is also applied the cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port during a $n$-th minute.
-
-For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack shall be used as reference for similarity analysis against the $\boldsymbol{v}_{(n)}$ identified as under attack, individually evaluating the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.
-
-Therefore, $\boldsymbol{v}^{(q)}$ should be calculated from the last $\boldsymbol{X}^{(q)}$ time frame without attack, and $\boldsymbol{v}_{(m,\hat{n})}$ should be calculated from the same traffic appened of all $n$-th minutes until the identified minute under attack, denoted as $\boldsymbol{X}_n$.
-
-For similarity analyis, each $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, denoted as $x_{(m,n)}$, shall be individually replaced by the traffic of the evaluated $m$-th port of the $\hat{n}$-th minute under attack, denoted as $x^{(\hat{q})}_{(m,\hat{n})}$, in order to identify significant variation on similarity caused by the traffic of the $m$-th port.
-
-This approach for detection of ports under attack via similarity analysis is given by
-
-\begin{equation}\label{eq:eq15}
-  \left\{
-  \begin{array}{@{}ll@{}}
-    x_{(m,n)} = x^{(\hat{q})}_{(m,\hat{n})} \\
-    \\
-    s_{m,\hat{n}} = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(m,\hat{n})}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(m,\hat{n})}}},
-  \end{array}\right.
-\end{equation}
-
-where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, which is used to calculate the $\boldsymbol{v}_{(m,\hat{n})}$ most significant eigenvectors that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.
-
-Once $\boldsymbol{v}^{(q)}$ and $\boldsymbol{v}_{(m,\hat{n})}$ are obtained, then the $s_{m,\hat{n}}$ similarity degree can be calculated in order to identify if the traffic replacement highlights the adition of anomalous traffic by the evaluated $m$-th port during the $\hat{n}$-th minute previously identified as under attack.
-
-This procedure should be repeated for each $m$-th target port of $\boldsymbol{\hat{n}}$, in order to individually identify the network ports under attack during each $\hat{q}$-th time frame.
-
-\section{Experiments and Results}
-\label{sec:experimentalresults}
-
-This section presents the performed experiments and the acquired results.
-First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
-Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario.
-In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated dataset.
-Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario.
-Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 dataset.
-
-\subsection{Experimental Scenario}
-\label{sec:AnalyzedScenario}
-
-The experiment time is 120 minutes, separated into six time frames, with each time frame corresponding to twenty minutes.
-Therefore, as the time of each sampling period is one minute, then $N = 20$.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, as well as a covariance $\boldsymbol{\hat{R}}_{yy}^{(q)} \in \mathbb{R}^{17 \times 17}$ (calculated via (\ref{eq:eq03})) and a sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)} \in \mathbb{R}^{17x17}$, assuming that $q = 1, 2, 3, 4, 5$ and $6$.
-
-The simulation started at 21:00h, the first time frame was from 21:00h until 21:20h ($q = 1$), the second was from 21:20h until 21:40h ($q = 2$), the third was from 21:40h to 22:00h ($q = 3$), the fourth was from 22:00h until 22:20h ($q = 4$), the fifth was from 22:20h until 22:40h ($q = 5$), and finally, the sixth was from 22:40h until 23.00h ($q = 6$).
-During the simulation, the victim made legitimate access, and the attacker performed the following attacks: at 21:54h ($q = 3$) was performed a port scan, at the interval ranging from 22:10h to 22:20h ($q = 4$) a synflood attack was simulated, and at the interval from 22:30h to 22:40h ($q = 5$) a fraggle attack was performed.
-
-\subsection{Largest Eigenvalues Analysis}
-\label{sec:largesteigenvaluesanalysis}
-
-For the evaluation of MOS Schemes accuracy for flood and port scan detection, the framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}.
-Through eigenvalue analysis of traffic with flood or port scan attacks, it is possible to visualize a significant difference between the largest eigenvalues and the remain eigenvalues, which can indicate a relationship between an outlier and time frames under attack.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_synflood.eps} 
-     \caption{Eigenvalues of the sample covariance matrix (synflood).}
-     \label{fig:fig10}
-\end{figure}
-
-Figure \ref{fig:fig10} depicts the eigenvalues calculated from sample covariance matrix of the network traffic used to evaluate the synflood attack identification.
-In Figure \ref{fig:fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 4$) stands out significantly from the other eigenvalues.
-
-Figure \ref{fig:fig11} illustrates the eigenvalues calculated from sample covariance matrix of the matrix used for fraggle attack detection.
-In Figure \ref{fig:fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 5$) stands out significantly from the other eigenvalues, in accordance with the result shown in Figure \ref{fig:fig10} for the synflood attack analysis.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_fraggle.eps}
-     \caption{Eigenvalues of the sample covariance matrix (fraggle).}
-     \label{fig:fig11}
-\end{figure}
-
-Figure \ref{fig:fig12} depicts the eigenvalues calculated from covariance matrix of zero mean and unitary standard deviation variables, of the network traffic matrix evaluated for port scan detection.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_portscan.eps}
-     \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}
-     \label{fig:fig12}
-\end{figure}
-
-As analyzed for the synflood and fraggle attacks, note that the largest eigenvalue, related to this attack ($q = 3$), stands out significantly from the others eigenvalues.
-
-Table \ref{tab:tab3} presents the values of the largest eigenvalues of each time frame $q$-th for port scan, synflood and fraggle detection.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Largest Eigenvalue related to attacks detection}
-  \label{tab:tab3}
-  \begin{tabular}{ c c c c c }
-	\toprule
-	\multirow{3}{*}{\textbf{Time Frame} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
-			\hhline{~----}
-		&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
-		&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
-	\midrule
-	1 &1887545 &1887545 &1887545 &2,0734 \\
-	2 &2341327 &2341327 &2341327 &2,1451 \\
-	3 &3213867 &3213867 &3213867 &10,0718 \\
-	4 &133238294 &133238294 &731229 &2,1620 \\
-	5 &92384021611 &6367983 &92384021611 &2,4253 \\
-	6 &708335 &708335 &708335 &1,7948 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-In Table \ref{tab:tab3}, note the significant variation of the eigenvalues associated with attacks, in comparison to the others.
-At $q = 4$, where the synflood attack occurred, the maximum eigenvalue obtained is approximately 21 times larger than the second one.
-At $q = 5$, where the fraggle attack occurred, the maximum eigenvalue obtained is about 29,000 times larger than the second one.
-At $q = 3$, where the port scan attack occurred, the maximum eigenvalue obtained is approximately 4 times larger than the second one.
-In the last case, for port scan attack detection, although the largest eigenvalue presented no too large variance to the second one, if compared to synflood or fraggle attacks, it clearly deviates from the remaining largest eigenvalues.
-
-These results highlight that all $q$-th time frames where a network attack was simulated, present high significant variance between the largest eigenvalue and the remaining eigenvalues, obtained from sample covariance matrix, for flood detection, or from covariance matrix of zero mean and unitary standard deviation variables, for port scan detection.
-Therefore, we propose to apply the vector of the largest eigenvalues to MOS schemes in order to evaluate their accuracy for identification of time frames under attack, motivated by the fact that it is relevant to apply MOS schemes to automate the attack detection process, taking into account the characteristics of the evaluated eigenvalues.
-
-\subsection{MOS Schemes Evaluation}
-\label{sec:MOSSchemesEvaluation}
-
-In \cite{tenorio2013greatest}, we evaluate the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE MOS schemes \cite{da2009comparison,tenorio2013greatest} for synflood and port scan attack detection.
-In this work we extend that evaluation for fraggle attack detection, applying the same schemes to fraggle attack detection over the traffic presented in Section \ref{sec:datamodel}, as results shown in Table \ref{tab:tab4}.
-
-\begin{table}[h!]
-  \centering
-  \scriptsize
-  \caption{MOS schemes applied to port scan and flood detection}
-  \label{tab:tab4}
-  \begin{tabular}{ c c c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
-			\hhline{~------~}
-		&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
-	\midrule
-	Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-	Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-	\midrule
-	Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-	Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-	\midrule
-	Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
-	Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
-	\midrule
-	Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
-	Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-Note that $\hat{d} = 1$, if there is attack, while $\hat{d} > 1$ indicates more than one attack.
-An example of this could be seen for attack detection via EFT for traffic containing synflood and fraggle attacks, showing $\hat{d} = 2$, which indicates the presence of two attacks, as expected by the $d$ real values of Table \ref{tab:tab4}.
-
-In Table \ref{tab:tab4}, two MOS schemes outperforms from the others, EDC and EFT.
-Efficient Detection Criterion (EDC) and Exponential Fitting Test (EFT) are the most effective schemes, correctly estimating the number of attacks in comparison to the expected values for effetctive attack detection, as defined by the column of real values in Table \ref{tab:tab4}.
-The AIC and MDL schemes are satisfactory only for port scan detection, however SURE and RADOI schemes did not show effective results for port scan or flood detection.
-
-Although EDC and EFT presented the same accuracy on the evaluation, the EDC scheme requires less processing time than EFT, which is an important criteria to select EDC as the MOS scheme for flood and port scan detection on the remain experiments.
-
-According to Table \ref{tab:tab4}, EDC and EFT estimated correctly the number of attacks of a time frame vector, indicating that ocurred $\hat{d}$ network attacks, but not providing additional details, what highlights the necessity of complementary approaches in order to estimate the time and ports under attack.
-Hence, we propose apply eigen analysis to estimate the $q$-th time frames under attack and eigen similarity analysis to estimate the minutes and ports under attack.
-
-\subsection{Eigenvalue Analysis}
-\label{sec:EigenvalueAnalysis}
-
-According to the results presented in Section \ref{sec:largesteigenvaluesanalysis}, the largest eigenvalue stands out significantly from the others eigenvalues of an evaluated $q$-th time frame.
-This behavior can also be observed in the largest eigenvalues analysis, according to results presented in Table \ref{tab:tab3}, where it is possible to observe that the $\hat{d}$ largest eigen values of the time frames under attacks stand out significantly from the others largest eigenvalues.
-
-Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respectives $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:alg01}.
-
-\subsection{Eigen Similarity Analysis}
-\label{sec:EigenSimilarityAnalysis}
-
-This paper proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$.
-Hence, the proposed framework is applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.
-
-\subsubsection{Time Analysis}
-\label{sec:TimeAnalysis}
-
-Three approaches were evaluated for eigen similarity analysis: incremental, individual and incremental individualized approaches.
-For the incremental individualized approach, each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
-Subsequently, $\boldsymbol{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated.
-For the incremental approach, each $n$-th minute must be incrementally appended into $\boldsymbol{X}^{(q)}$, for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual time similarity analyis.
-For the individual approach, each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, without incremental append, but doing individual appended into $\boldsymbol{X}^{(q)}$ for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual similarity analyis.
-
-Table \ref{tab:tab5} presents the results of the evaluation of three approaches for similarity analysis of eigenvectors for port scan detection.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Port Scan Detection}
-  \label{tab:tab5}
-  \begin{tabular}{ c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~~---~}
-			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
-	\midrule
-	3 &1 &0.9946 &0.9946 &0.9946 &no \\
-	3 &2 &0.9934 &0.9934 &0.9999 &no \\
-	3 &3 &0.9912 &0.9912 &0.9999 &no \\
-	3 &4 &0.9888 &0.9888 &0.9999 &no \\
-	3 &5 &0.9856 &0.9856 &0.9998 &no \\
-	3 &6 &0.9840 &0.9840 &0.9999 &no \\
-	3 &7 &0.9824 &0.9824 &1.0000 &no \\
-	3 &8 &0.9794 &0.9794 &0.9999 &no \\
-	3 &9 &0.9673 &0.9673 &0.9926 &no \\
-	3 &10 &0.9674 &0.9674 &0.9997 &no \\
-	3 &11 &0.9733 &0.9733 &0.9993 &no \\
-	3 &12 &0.9702 &0.9702 &0.9993 &no \\
-	3 &13 &0.9677 &0.9677 &0.9999 &no \\
-	3 &14 &0.9646 &0.9646 &0.9998 &no \\
-	3 &15 &0.0216 &0.0216 &0.0276 &yes \\
-	3 &16 &0.9621 &0.0209 &1.0000 &no \\
-	3 &17 &0.9611 &0.0199 &0.9998 &no \\
-	3 &18 &0.9612 &0.0191 &0.9999 &no \\
-	3 &19 &0.9613 &0.0186 &0.9998 &no \\
-	3 &20 &0.9638 &0.0190 &1.0000 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-Table \ref{tab:tab5} shows the evaluation of the time frame $q=3$, when the port scan attack was simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
-According to the presented results, it is possible to observe the high similarity between network traffic without attack, which was larger than 0.9610 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with the simulated port scan attack ($n=15$), which was lower than 0.0276 for all evaluated approaches.
-
-Comparing the approaches for similarity analysis, it is possible to observe that all approaches highlight the low similarity when evaluated the traffic under attack.
-However, the incremental approach figured out low similarity for times without attack, where $n=16, 17, 18, 19, 20$, what indicates that the incremental approach can produce false posive results.
-This behavior occurs because the incremental approaches appends all selected traffic into the reference traffic for comparison against the original reference traffic, what makes more evident the first lack of similarity but reduces the changing detection capability after an attack detection.
-
-Table \ref{tab:tab6} presents the results of the evaluation of the similarity analysis of eigenvectors for synflood detection.
-It shows the evaluation of the time frame $q=4$, when the synflood attack is simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
-According to the results, it is possible to observe the high similarity between network traffic without attack, which is larger than 0.9907 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with synflood attack (between $n=11$ and $n=20$), which is lower than 0.1244 for all evaluated approaches.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Synflood Detection}
-  \label{tab:tab6}
-  \begin{tabular}{ c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~~---~}
-			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
-	\midrule
-	4 &1 &1.0000 &1.0000 &1.0000 &no \\
-	4 &2 &0.9999 &0.9999 &1.0000 &no \\
-	4 &3 &0.9997 &0.9997 &0.9999 &no \\
-	4 &4 &0.9998 &0.9998 &1.0000 &no \\
-	4 &5 &0.9965 &0.9965 &0.9908 &no \\
-	4 &6 &0.9975 &0.9975 &1.0000 &no \\
-	4 &7 &0.9977 &0.9977 &1.0000 &no \\
-	4 &8 &0.9980 &0.9980 &1.0000 &no \\
-	4 &9 &0.9987 &0.9987 &0.9999 &no \\
-	4 &10 &0.9991 &0.9991 &1.0000 &no \\
-	4 &11 &0.0085 &0.0085 &0.0284 &yes \\
-	4 &12 &0.0162 &0.0120 &0.0343 &yes \\
-	4 &13 &0.0248 &0.0158 &0.0427 &yes \\
-	4 &14 &0.1243 &0.0185 &0.1041 &yes \\
-	4 &15 &0.0082 &0.0162 &0.0103 &yes \\
-	4 &16 &0.0404 &0.0070 &0.0580 &yes \\
-	4 &17 &0.0397 &0.0007 &0.0573 &yes \\
-	4 &18 &0.0408 &0.0042 &0.0584 &yes \\
-	4 &19 &0.0408 &0.0079 &0.0584 &yes \\
-	4 &20 &0.0477 &0.0092 &0.0757 &yes \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The incremental approach produces better results if compared with other evaluated approaches, with lower values and maximum of 0.0185 for times under attack, but this approach presents change detection limitation after the first outlier of similarity, as shown in Table \ref{tab:tab5} for port scan detection.
-
-Comparing the incremental individualized and the individual approaches for eigen similarity analysis, it is possible to observe that the incremental individualized approach obtain lowest values for almost all cases, except for the time $n=14$, where incremental individualized approach identified a larger similarity than the individual approach.
-The incremental individualized appends information about each evaluated traffic, therefore it incorporates traffic behaviors that can reduce the outlier capability detection, as ocurred for the time $n=14$.
-
-Table \ref{tab:tab7} presents the results of the eigen similarity analysis evaluation for fraggle detection.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Fraggle Detection}
-  \label{tab:tab7}
-  \begin{tabular}{ c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~~---~}
-			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
-	\midrule
-	5 &1 &1.0000 &1.0000 &1.0000 &no \\
-	5 &2 &0.9999 &0.9999 &1.0000 &no \\
-	5 &3 &1.0000 &1.0000 &1.0000 &no \\
-	5 &4 &0.9999 &0.9999 &1.0000 &no \\
-	5 &5 &0.9993 &0.9993 &0.9997 &no \\
-	5 &6 &0.9993 &0.9993 &0.9997 &no \\
-	5 &7 &0.9994 &0.9994 &1.0000 &no \\
-	5 &8 &0.9995 &0.9995 &1.0000 &no \\
-	5 &9 &0.9995 &0.9995 &1.0000 &no \\
-	5 &10 &0.9995 &0.9995 &1.0000 &no \\
-	5 &11 &0.0031 &0.0031 &0.0021 &yes \\
-	5 &12 &0.0019 &0.0025 &0.0009 &yes \\
-	5 &13 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &14 &0.0030 &0.0027 &0.0020 &yes \\
-	5 &15 &0.0030 &0.0028 &0.0020 &yes \\
-	5 &16 &0.0012 &0.0025 &0.0002 &yes \\
-	5 &17 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &18 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &19 &0.0030 &0.0027 &0.0020 &yes \\
-	5 &20 &0.0069 &0.0023 &0.0083 &yes \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-For fraggle attack detection, the lack of similarity between legitimate and malicious traffic was more evident than for the evaluation of synflood and port scan detecion.
-This behavior can be explained by the number of packets generated through the fraggle attack simulation, that was significative larger than the number of packets generated during the synflood simulation.
-Considering the three approaches, the largest value for times under attack was 0.0083, while the shortest value for times without attacks was 0.9993.
-
-Therefore, considering the evaluation for port scan, synflood and fraggle detection, the incremental approach can produce false positive results, while the individual and incremental individualized approaches produce quite similar results, even though the individual approach be more simple and require less memory and processing time.
-
-These results highlight the capability of change detection based on similarity between legitimate and malicious traffic from flood or port scan attacks, endorsing the effectiveness and safety for adoption of threshold for attack detection through eigen similarity analysis.
-
-\subsubsection{Port Analysis}
-\label{sec:PortAnalysis}
-
-Given $\hat{N}$, which is the set of estimated $n$-th minutes under attack, it is possible to apply cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port, during a $n$-th minute.
-Therefore, the incremental individualized and individual approaches of eigen similarity analysis were evaluated, for detection of ports under flood and port scan attacks, according to results presented in following tables.
-For this evaluation, the $v$ last most significant eigenvectors without attack was used as reference for similarity analysis against each target port $m$-th.
-
-Table \ref{tab:tab8} presents the results of the evaluation of eigen similarity analysis for detection of ports under port scan attack, showing only the time frame $q=3$ and minute $n=15$, due to the simulated port scan attack occurred only at this time, although the remain time frame has been completely evaluated and presented high similarity to the reference of traffic without network attack.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n=15)}
-  \label{tab:tab8}
-  \begin{tabular}{ c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~--~}
-			&\textbf{Incremental Individualized} &\textbf{Individual}\\
-	\midrule
-	80 &0.9999 &0.9999 &no \\
-	443 &0.9999 &0.9999 &no \\
-	53 &0.9999 &0.9999 &no \\
-	21 &0.9999 &0.9997 &yes \\
-	22 &0.0298 &0.9997 &yes \\
-	23 &0.0298 &0.9997 &yes \\
-	25 &0.0298 &0.9997 &yes \\
-	110 &0.0298 &0.9997 &yes \\
-	143 &0.0298 &0.9997 &yes \\
-	161 &0.0298 &0.9997 &yes \\
-	69 &0.0298 &0.9997 &yes \\
-	123 &0.0298 &0.9997 &yes \\
-	445 &0.0298 &0.9997 &yes \\
-	600 &0.9999 &0.9999 &no \\
-	19 &0.9999 &0.9999 &no \\
-	67 &0.9999 &0.9999 &no \\
-	68 &0.9999 &0.9999 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The incremental individualized approach presented more sensibility to anomaly detection than the individual approach, the former produced the identification of a low similarity of 0.0298 for almost all ports under attack, unless the port 21, although the simulation has attacked this port.
-The individual approach was not able to identify low similarity for ports under attack, resunting in values of 0.9997 for ports with anomalous traffic and 0.9999 for ports without network attack.
-
-For the evaluation of the proposed approaches for identification of ports under synflood and fraggle attack, all minutes of each time frame, in which one attack location was estimated, were analyzed.
-Even though, due to space limitations, only the results of the firt minute where a low similarity was idenfified will be shown such as where $n=11$.
-Nevertheless, the results obtained for the evaluation of traffic without attack presented high similarity to the reference traffic, with similarities close to 0.9999, and the evaluation of the other minutes under attack presented results quite similar to the resuts shown in the Tables \ref{tab:tab9} and \ref{tab:tab10}.
-
-Table \ref{tab:tab9} presents the results of the evaluation of eigen similarity analysis for detection of ports under synflood attack, showing only the time frame $q=4$ and minute $n=11$.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
-  \label{tab:tab9}
-  \begin{tabular}{ c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~--~}
-			&\textbf{Incremental Individualized} &\textbf{Individual}\\
-	\midrule
-	80 &1.0000 &1.0000 &no \\
-	443 &1.0000 &1.0000 &no \\
-	53 &1.0000 &1.0000 &no \\
-	21 &1.0000 &1.0000 &nos \\
-	22 &1.0000 &1.0000 &no \\
-	23 &1.0000 &1.0000 &no \\
-	25 &1.0000 &1.0000 &no \\
-	110 &1.0000 &1.0000 &no \\
-	143 &1.0000 &1.0000 &no \\
-	161 &1.0000 &1.0000 &no \\
-	69 &1.0000 &1.0000 &no \\
-	123 &1.0000 &1.0000 &no \\
-	445 &1.0000 &1.0000 &no \\
-	600 &0.0077 &0.0427 &yes \\
-	19 &1.0000 &1.0000 &no \\
-	67 &1.0000 &1.0000 &no \\
-	68 &1.0000 &1.0000 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-According to results presented in Table \ref{tab:tab9}, both approches identifies low similarity for the traffic of port 600, which is the target port of the simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.
-
-Table \ref{tab:tab10} presents the results of the evaluation of eigen similarity analysis for detection of ports under fraggle attack, showing only the time frame $q=5$ and minute $n=11$.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Fraggle Attack (q=5 and t=11)}
-  \label{tab:tab10}
-  \begin{tabular}{ c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~--~}
-			&\textbf{Incremental Individualized} &\textbf{Individual}\\
-	\midrule
-	80 &1.0000 &1.0000 &no \\
-	443 &1.0000 &1.0000 &no \\
-	53 &1.0000 &1.0000 &no \\
-	21 &1.0000 &1.0000 &no \\
-	22 &1.0000 &1.0000 &no \\
-	23 &1.0000 &1.0000 &no \\
-	25 &1.0000 &1.0000 &no \\
-	110 &1.0000 &1.0000 &no \\
-	143 &1.0000 &1.0000 &no \\
-	161 &1.0000 &1.0000 &no \\
-	69 &1.0000 &1.0000 &no \\
-	123 &1.0000 &1.0000 &no \\
-	445 &1.0000 &1.0000 &no \\
-	600 &1.0000 &1.0000 &no \\
-	19 &0.0031 &0.0004 &yes \\
-	67 &1.0000 &1.0000 &no \\
-	68 &1.0000 &1.0000 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The results for the avaluation of ports under fraggle attack, shown in Table \ref{tab:tab10}, were similar to the results obtained for synflood analysis, with the identification of low similarity for traffic of the port under attack.
-Nevertheless, for fraggle analysis, the individual approach identified the lowest similarity, that is 0.0004 while the incremental individualized approach obtained a similarity of 0.0031.
-
-The incremental individualized approach was able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches presented false positives or low sensibility to eigen similarity analyis for network attack detection.
-This approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.
-
-According to the shown significant lack of similarity between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
-
-\subsection{DARPA Scenario}
-\label{sec:DarpaEvaluation}
-
-This subsection presents a summarized view of results obtained from the application of the proposed framework, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 dataset.
-Since the proposed framework is detailed in Section \ref{sec:prop_getv} and in Subsections \ref{sec:AnalyzedScenario}, \ref{sec:largesteigenvaluesanalysis}, \ref{sec:MOSSchemesEvaluation}, \ref{sec:EigenvalueAnalysis} and \ref{sec:EigenSimilarityAnalysis}, here the focus is on the parameter selection, dataset evaluation and results for flood and probe attack identification.
-
-The DARPA dataset includes 7 weeks of sniffed traffic saved into raw packet data.
-The traffic and the labeled attacks are grouped by week and day, with information of the number and types of attacks per day, but also providing the start time for each labeled attack.
-For this evaluation, an evaluation per day was performed, considering the network traffic of 24 hours split into $Q$ time frames of 60 minutes ($N = 60$) and aggregate by minute and by port number.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ is obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.
-
-Since the proposed framework focus on flood and probe attack detection, only the attacks with behavior similar to flood or probe attack were evaluated.
-Initially all DoS and probe attacks were selected, but it was observed that the most cases of DoS focus on exploit system vulnerabilities instead of flooding attack, and most of probe attacks focus on ICMP intead of port scanning.
-Therefore, for evaluation of the proposed approach for flood and probe attack detection, it is necessary to select cases that implements flood or port scan behaviors.
-The following week-day-attack cases  were selected:
-
-\begin{enumerate}
-	\item week3-thursday-neptune;
-	\item week4-friday-portsweep;
-	\item week5-thursday-neptune;
-	\item week5-thursday-portsweep;
-	\item week5-friday-portsweep;
-	\item week6-wednesday-neptune;
-	\item week6-thursday-neptune;
-	\item week7-wednesday-portsweep.
-\end{enumerate}
-
-\begin{table*}[!t]
-	\caption{Results of the attack detection evaluation}
-	\label{tab:tab12}
-	\centering
-	\begin{tabular}{|c|c|c|c|c|}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Solution\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Attack Type\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Metric\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Result\end{tabular} \\ \hline
-	Proposed Work	&Flooding	&True Positive	&100.00 \%\\ \hline
-	Proposed Work	&Flooding	&False Positive	&60.00 \%\\ \hline
-	Proposed Work	&Flooding	&Misclassification	&50.00 \%\\ \hline
-	Proposed Work	&Probe	&True Positive	&76.92 \%\\ \hline
-	Proposed Work	&Probe	&False Positive	&18.52 \%\\ \hline
-	Proposed Work	&Probe	&Misclassification	&32.73 \%\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&True Positive	&82.00 \%\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&False Positive	&-\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&Misclassification	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Overall	&True Positive	&94.67 \%\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Overall	&False Positive	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Overall	&Misclassification	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&True Positive	&50.00 \%\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&False Positive	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&Misclassification	&-\\ \hline
-	\end{tabular}
-\end{table*}
-
-Table \ref{tab:tab12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}.
-The analysis based on sample covariation of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames.
-The results also show 50.00 \% of misclassification rate , which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
-The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time  from one source to one target, with no variation on IP source or target port.
-This observation corroborates with previous evaluations of the DARPA dataset that highlight issues regarding traffic redundancy.
-
-The analysis based on covariance of zero mean and unitary standard deviation variables was evaluated for port scan attacks, including probe attacks and DoS attacks that send few packets for several ports in order to exploit some vulnerability.
-The results show rates of 76.92 \% for TP detection and 18.52 \% for FP detection from 94 time frames.
-The observed misclassification rate for this scenario is 32.73 \%.
-It was observed that all FN cases are probe attacks with a time delay between scanning one port and start scanning the next port, what can be called as sparse probe attacks.
-Cases with a delay of one minute or more were not detected by the proposed approach.
-
-The performance of detection rate of flooding attacks is compared with the method proposed by Callegari \emph{et al} \cite{Zonglin2009}.
-This work is a statistical method, based on PCA, without training or learning methods, even though it relies on visual analysis for principal components selection.
-The best detection rate of \cite{Zonglin2009} was 82.00 \% for detection of synthetically added flood attacks, while this current proposal obtains 100.00 \% of detection rate for detection of flood attacks of DARPA dataset.
-
-Due to the lack of statistical techniques without training or learning methods for detection of probe attacks, this proposed approach is compared to the Lu and Ghorbani's \cite{Lu2009} proposal, which is a network anomaly detection model based on signal processing techniques that uses DARPA dataset for evaluation.
-The results of \cite{Lu2009} show the best detection rate of 94.67 \% in terms of general attack instance detection, but shows a case case with 50.00 \% of attack instance detection for the portsweep attack.
-The proposed approach presents 76.92 \% of detection rate measured specifically for probe attacks, without the requirement of learning or training methods, in contrast to Lu and Ghorbani's \cite{Lu2009} work.
-
-\section{Performance Evaluation}
-\label{sec:Complexity}
-
-This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
-
-\subsection{Complexity Analysis}
-\label{sec:ComplexityAnalysis}
-The EVD, calculated according to (\ref{eq:eq06}), requires the previous calculation of covariance matrix, according to Equations (\ref{eq:eq02}), (\ref{eq:eq03}), (\ref{eq:eq04}) and (\ref{eq:eq05}).
-The covariance matrix calculation is $O(M^2N)$ and the EVD is $O(N^3)$, where $M$ denotes the number of network ports and $N$ denotes the period time.
-Therefore, the computational complexity for all steps for EVD can be represented as $O(M^2N + N^3)$ and yields an $O(N^3)$ upper bound on the worst-case running time for EVD.
-
-EDC and EFT are the MOS schemes that presented accuracy on the evaluation for the network attack detection.
-The computational complexity evaluation for MOS focuses on EDC scheme, since EDC requires less processing time than EFT but presents the same accuracy for the evaluated scenario.
-EDC scheme is $O(Q \log Q + Q + Q \log Q)$ and its worst-case running time can be represented as $O(Q \log Q)$, where $Q$ denotes the number of time frames.
-
-The largest eigenvalue analysis is $O(\hat{d}Q)$, where $\hat{d}$ denotes the number of time frame under attack, according to Algorithm \ref{alg:alg01}.
-Subsequently, the eigen similarity analysis relies on EVD and cosine similarity analysis, which is $O(N^2)$, for $\hat{d}$ time frames, therefore the eigen similarity analysis is which is $O(\hat{d}(M^2N + N^3 + N^2))$ and yields an $O(N^3)$ upper bound on the worst-case running time for eigen similarity analysis.
-
-Therefore, the proposed framework is $O(N^3 + Q \log Q + \hat{d}Q + N^3)$ and its worst-case running time is $O(N^3)$.
-The computational complexity of EVD is predominant in the framework, but the approach splits the data into time frames with period time $N$, which makes possible to limitate the growth of $N$ even for evaluations of cases with total time larger than $N$, reducing the impact caused by the computational complexity of EVD.
-
-\subsection{Processing Time Analysis}
-\label{sec:ProcessingAnalysis}
-For better understanding the scalability and impact of configurations of $N$, $M$ and $Q$, the processing time required for different scenarios of parameter configurations and dataset were evaluated, measuring the processing time of:
-
-\begin{enumerate}
-	\item Eigen analysis based on sample covariance of zero mean;
-	\item Eigen analysis based on sample covariance of zero mean and unitary standard deviation;
-	\item EDC MOS scheme;
-\end{enumerate}
-
-The performance evaluation focus on the main steps, which are discussed in subsection \ref{sec:ComplexityAnalysis} regarding the complexity analysis.
-The data modeling is also a time consuming step, however its processing can be optimized through distributed processing techniques, such as MapReduce, achieving high throughput for packet counting or even for deep packet inspection \cite{Vieira2013}.
-
-The experiments were performed in a desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM, considering: variations on the network traffic time; the frame size denoted as $N$; the number of network ports denoted as $M$; the mean processing time for eigen analysis based on sample covariance of zero mean, denoted as 1-time; the mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation, denoted as 2-time; and the mean processing time for EDC MOS scheme, denoted as 3-time.
-The mean time calculations was obtained from 200 measurement repetitions, in order to obtain reliable values.
-
-\begin{table*}[!t]
-	\caption{Processing time of the main steps for anomaly detection}
-	\label{tab:11}
-	\centering
-	\begin{tabular}{|r|r|r|r|r|r|r|}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Traffic Time\\(hour)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Frame Size\\(min)\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Num.
-Ports\end{tabular}	& \begin{tabular}[x]{@{}l@{}}1-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}2-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}3-time\\(ms)\end{tabular}	\\ \hline
-16	& 10	& 17	& 0.7900	& 0.8100	& 0.0650\\ \hline
-16	& 20	& 17	& 0.5250	& 0.5950	& 0.0100\\ \hline
-16	& 60	& 17	& 0.9700	& 1.1400	& 0.0250\\ \hline
-16	& 120	& 17	& 0.6050	& 0.6100	& 0.0050\\ \hline
-16	& 60	& 34	& 1.2750	& 1.2200	& 0.0050\\ \hline
-16	& 120	& 34	& 1.1200	& 1.1700	& 0.0050\\ \hline
-20	& 10	& 17	& 2.7950	& 2.8950	& 1.1000\\ \hline
-20	& 20	& 17	& 2.0700	& 2.0200	& 0.3500\\ \hline
-20	& 60	& 17	& 1.0250	& 1.0450	& 0.0650\\ \hline
-20	& 120	& 17	& 1.0000	& 1.0700	& 0.0350\\ \hline
-20	& 60	& 34	& 2.9650	& 3.2100	& 0.0400\\ \hline
-20	& 120	& 34	& 2.9950	& 3.1150	& 0.0200\\ \hline
-22	& 10	& 17	& 4.7250	& 4.0850	& 1.4600\\ \hline
-22	& 20	& 17	& 2.3200	& 2.6800	& 0.2450\\ \hline
-22	& 60	& 17	& 1.0700	& 1.1200	& 0.0300\\ \hline
-22	& 120	& 17	& 0.9900	& 1.0500	& 0.0250\\ \hline
-22	& 60	& 34	& 3.0850	& 3.1250	& 0.0650\\ \hline
-22	& 120	& 34	& 2.8100	& 2.9600	& 0.0250\\ \hline
-	\end{tabular}
-\end{table*}
-
-Table \ref{tab:11} presents the measured results.
-The experiment considered traffic time of 16, 20 and 22 hours, according to the selected traffic time per day available by the DARPA dataset.
-Note that the processing time increases according to the increment in traffic time, around 2 or 3 times for 1-time and 2-time, but the worst measured processing time is 4.7250 milliseconds.
-
-According to Table \ref{tab:11}, the processing time increases with the frame size $N$ decreasing, therefore it is possible to evaluate the frame size that produces better identification rates and acceptable processing time.
-The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations, since the significant increase of processing time observed between scenarios considering 17 or 34 ports, with growth between 7 \% and 199 \%.
-
-\section{Conclusion and Future Works}
-\label{sec:conclusionandfutureworks}
-
-This paper models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
-
-The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
-
-This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
-Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.
-
-The incremental individualized approach of eigen similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to eigen similarity analyis for network attack detection.
-Therefore, the incremental individualized approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.
-
-According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
-
-Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets.
-Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high thoughput network traffic.
-Additionally, future research can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.
-
-\appendix
-
-\section{Model Order Selection (MOS)}
-\label{sec:mos}
 The model order selection is a key point in many digital signal processing applications, including radar, sonar, communications, channel modeling, medical imaging, among others.
@@
 
-In AIC, MDL and EDC techniques, the information criterion is a function of the geometric mean \emph{g(k)} and the arithmetic mean \emph{a(k)} relating to smaller \emph{k} eigenvalues, where \emph{k} is a candidate value for the model order \emph{d} \cite{da2009comparison}.
+In AIC, MDL and EDC techniques, the information criterion is a function of the geometric mean $g(k)$ and the arithmetic mean $a(k)$ relating to smaller $k$ eigenvalues, where $k$ is a candidate value for the model order $d$ \cite{da2009comparison}.
 
-Basically, the difference between the AIC, MDL and EDC schemes is the penalty function \emph{p(k, N, α)}, so these techniques can be written in general as \cite{da2009comparison}:
+Basically, the difference between the AIC, MDL and EDC schemes is the penalty function $p(k, N, \alpha)$, so these techniques can be written in general as \cite{da2009comparison}:
 
@@
 
-where $\hat{d}$ is an estimate \emph{d} of the model order, \emph{N} is the number of samples, \emph{α = M} and means the number of variables of the problem, and 0 ≤ \emph{k} ≤ min[\emph{M}, \emph{N}].
+where $\hat{d}$ is an estimate $d$ of the model order, $N$ is the number of samples, $\alpha = M$ and means the number of variables of the problem, and $0 \leqslant k \leqslant min[M, N]$.
 Penalty functions for AIC, MDL and EDC are given by the Table \ref{tab:tab2}.
@@
 
-The Exponential Fitting Test (EFT) can effectively be used in cases where the number of samples \emph{N} is small.
+The Exponential Fitting Test (EFT) can effectively be used in cases where the number of samples $N$ is small.
 This technique is based on observations of data contaminated only with white noise, where the profile of eigenvalues can be approximated by an exponential decaying \cite{grouffaud1996some}.
@@
 
-where $E\{\cdot\} $ is the expectation operator, and it is considered that the eigenvalues are ordered in the that $\lambda_1$ represents the largest eigenvalue. The term \emph{q(α, β)} is defined as:
+where $E\{\cdot\} $ is the expectation operator, and it is considered that the eigenvalues are ordered in the that $\lambda_1$ represents the largest eigenvalue.
+The term $q(\alpha, \beta)$ is defined as:
 

### 2017-10-24 docs/thesis/chapters/appendix/b_pca_svm_cfa/pca_svm_cfa.tex
appendix configuration

--- 2017-10-24 docs/thesis/chapters/2_pca_svm_cfa/pca_svm_cfa.tex
@@
 \chapter{Critical Success Factor Analysis Based on Feaure Selection}
-\label{ch:2_csf_fs}
-
-\begin{quotation}[]{Paulo Freire}
-No one knows it all. No one is ignorant of everything. We all know something. We are all ignorant of something.
-\end{quotation}
+\label{ch:b_csf_fs}
 
@@
 This chapter is organized as follows.
-In Section \ref{sec:ch2_relatedworks}, related works are discussed.
-Section \ref{sec:ch2_datamodel} presents the data model and the evaluated datasets.
-Section \ref{sec:ch2_csf_fs} describes the proposed approach for critical success factors analysis.
-Section \ref{sec:ch2_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:ch2_conclusion} draws the conclusions.
+In Section \ref{sec:b_relatedworks}, related works are discussed.
+Section \ref{sec:b_datamodel} presents the data model and the evaluated datasets.
+Section \ref{sec:b_csf_fs} describes the proposed approach for critical success factors analysis.
+Section \ref{sec:b_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:b_conclusion} draws the conclusions.
 
 \section{Related Works}
-\label{sec:ch2_relatedworks}
+\label{sec:b_relatedworks}
 
@@
 \section{Data Model}
-\label{sec:ch2_datamodel}
+\label{sec:b_datamodel}
 
@@
 \section{An approach for Critical Success Factors Analysis}
-\label{sec:ch2_csf_fs}
+\label{sec:b_csf_fs}
 
@@
 \subsection{Visual Discriminant Analysis based on PCA}
-\label{sec:ch2_pca}
+\label{sec:b_pca}
 
@@
 \subsection{CSF analysis based on RFE}
-\label{sec:ch2_csfa}
+\label{sec:b_csfa}
 
@@
 \section{Experiments and Results}
-\label{sec:ch2_experimentalresults}
+\label{sec:b_experimentalresults}
 
@@
 \section{Conclusion}
-\label{sec:ch2_conclusion}
+\label{sec:b_conclusion}
 

### 2017-10-25 docs/thesis/chapters/appendix/b_pca_svm_cfa/pca_svm_cfa.tex
pca_rfe conclusion

@@
 Results show how PCA can make the data discriminative and that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of brazilian public sector.
-Finally, SVM is used to highlight the more significant features identified by RFE, wich are limilar to CSFs previsously identified by a qualitative analysis of the same datased.
+Finally, SVM is used to highlight the more significant features identified by RFE, wich are similar to CSFs previsously identified by a qualitative analysis of the same datased.
 
@@
 
-pca make possible visual discriminative analysis of reduced dataset but it is not easy make correlation betwee extracted features and its original, therefore it is necessary adictional techniques to identify CSF.
+The proposed approach is evaluated and the experimental results show that the visual discriminant analysis, throught PCA, highlights important characteristcs of the data, and that a selected classifier and RFE can be applied for iGovTI classification and for CSF identification.
 
-SVM have been shown to generalize well even for small sample classification[8]. SVM presents high performance to reproduce the iGovTI classification and can be used for RFE in order to identify CSF
+Results show how PCA can make the data discriminative, however it is hard to identify what original variables are most relevant.
+Therefore, it is still necessary an additional method to reveal what are the CSF for iGovTI.
+For this problem, we propose a CSF analysis based on RFE.
+Since RFE depends on a classifier to select the most relevant features, we perform an algorithm evaluation to identify the classifier with the highest performance, and results show that SVM \cite{hearst1998support} presents the best accuracy, with 91.42\%.
 
-The selected features revealed by RFE are quite similar to the qualitative analysis of Ramos \emph{et al} \cite{ramos2016information}.
-
-Question with low variance can indicate that it is irrelevant for evaluation, since its value is almost the same all the time
+Finally, SVM is used to select the more significant features identified by RFE, wich are 69,9\% similar to CSFs previsously identified by a qualitative analysis of the same datased.

### 2017-10-26 docs/thesis/chapters/1_introduction/introduction.tex
compiling version with all chapters
introducion...

@@
-% Introdução: cenario, motivacao, problema, solucoes atuais, escopo do artigo, lista de contribuicoes, organizacao do resto do artigo. 
+% Introdução: cenario, motivacao, problema, solucoes atuais, escopo do trabalho, lista de contribuicoes, organizacao do resto do artigo. 
 % Chapter Structure
@@
 \label{sc:motivation}
-...
+
+Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}.
+In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time.
+Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
+Recently, signal processing schemes have being applied to the detection of malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest}, showing advances in network traffic analysis.
+
+Information security may consist of both technical and procedural aspects.
+The former includes equipment and security systems, while the latter corresponds to security rules and recommendations.
+Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks.
+Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.
+Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
+
+Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation.
+The basic features of the cloud environment is providing the elastic, on-demand and secure services for the end-users.
+While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}.
+Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized leading to obscure and incomplete security paradigms and solutions.
+
+Additional security issues and requirements have to be considered when mobile clients are actively used in corporate cloud environment \cite{yovel2014}.
+Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm.
+The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS).
+While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, the corporate mobile client is usually light-weighted and generally less protected.
+
+The protection scheme used on a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}.
+Therefore, encrypting files and generating keys on a mobile device is not considered a good solution.
+On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}.
+The common practice is the shadow user activities monitoring \cite{yovel2014}.
+However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode.
+When the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile client has to secure itself with its own limited resources.
+Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection.
+For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
+
+The field of Mobile Money Transfer (MMT) is a growing market segment and has been motivating investiments into security defense and fraud detection.
+Obtaining access to data sets of mobile transactions for research is a very hard task due to the intrinsic private nature of such transactions.
+Scientists and researchers must today spend time and effort in obtaining permits and access to relevant data sets before they can research on such data set.
+However, PaySim is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.\footnote{https://www.kaggle.com/ntnu-testimon/paysim1}
+
+Despite the actual high availability of information, the relevant information of some observations is generally of under much reduced dimensionality compared to available data sets.
+The extraction of relevant information by identifying the generating causes within classes of signals is useful for classification problems and for security analysis.
+Big data analytics require techniques to deal with multidimensional data, making sense of strucutre and relationship of many dimensions.
+Dictionary learning is a signal processing technique based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world.
+Dictionary Learning is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
+The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.
+
+Tensor decomposition has been adopted for multidimensional data analysis and can be explored for dictionary learning...principalmente por tensor decompositio extrair mais informações e relacionamentos que técnicas tradicionais, podendo ser melhores que técnicas tradicionais de dictionary learning
+
+Iterative dictionary learning has been used as a blind technique for feature extraction, improving classification without require feature selection or principal conponent analysis
+
+We propose methods for information security analysis
+
+Observing the above described landscape, this paper outlines the concept of the offline mobile cloud security.
+
+Compressive sensing is emergent for signal processing and has been applied to fields such as signal restoration, image processing and classification (sparce representation classification src )
+
+Adaptive techniques for high dimensional and high thoughput
+
+% There are two major approaches for dictionary learning.
+First is the analytic approach, in which DCT bases, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries.
+Second is the learning-based approaches, such as the unsupervised learning for dictionary construction [9] and the online dictionary learning [11], [10], which use machine learning methods to construct the dictionary.
+In the analytic approach, some pre-defined functions are used to construct the dictionary.
+Curvelets [24], which tracks the shape of the discontinuity set, offers efficient and near-optimal representation of smooth objects.
+Shearlets [25], which is obtained from dilations, action of translations and shear transformations, has nice geometric properties and mathematical properties for image representation.
+Bandelets [26], which specifies the geometry as a vector field, is de-signed to improve the image compression and noise reduction performance.
+In the learning-based approach, machine learning methods are used to construct the dictionary from the training data.
+The least square error is used by the method of the optimal directions (MOD) [27] to update the dictionary iteratively:
+% Dictionaries are either available analytically, or can be learned from a suitable training set.
+While analytic dictionaries permit to capture the global structure of a signal and allow a fast implementation, learned dictionaries often perform better in applications as they are more adapted to the considered class of signals.
+
+In some applications the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks.
+Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in multidimensional data analytics problems \cite{kolda2009tensor}.
+Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}.
+
+% In imagery, the numerical burden for (i) learning a dictionary and for (ii) employing the dictionary for reconstruction tasks only allows to deal with relatively small image patches that only capture local image information.
+Separable dictionaries aims at overcoming these drawbacks by allowing a separable structure on the dictionary throughout the learning process.
+On the one hand, this permits larger patch-sizes for the learning phase, on the other hand, the dictionary is applied efficiently in reconstruction tasks.
+The crucial idea is to allow the dictionary to have a separable structure, where separable means that the dictionary D is given by the Kronecker product of two smaller dictionaries A ∈ R h×a and B ∈ R w×b
+% Without loss of generality we choose square image patches with w = h = 8, which is in accordance to the patch-sizes mostly used in the literature
+
+Multidimensional parameter estimation and learning multidimensional separable dictionaries are growing research problems.
+Roemer \emph{et al.} \cite{roemer2014tensor} show that the multidimensional dictionary estimation problem can be efficiently formulated in terms of tensors, and that their results outperform existing schemes by exploiting the multilinear structure of the problem.
+
+Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results.
+However, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets, as can be seen in Figures \ref{fig:fig1} and \ref{fig:fig2}, which show that the recovering rating of traditional algorithms decreases over the dataset increasing.
+
+\begin{figure}[!htb]
+     \centering 
+	 \includegraphics[height=8cm, width=11cm]{figures/5_20_2000_1000_100.eps}
+     \caption{Matrix of 1.000 elements, 2000 sample training, 100 iterations}
+     \label{fig:fig1}
+\end{figure}
+
+\begin{figure}[!htb]
+     \centering 
+	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_24000_100.eps}
+     \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}
+     \label{fig:fig2}
+\end{figure}
+
+Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based and traditional dictionary learning algorithms for dictionary reconstruction of a multidimensional and separable data.
+Figure \ref{fig:fig1} presents the results for dictionary reconstruction from a matrix of 1000 elements by 50 atoms, while Figures \ref{fig:fig2} presents the results from a matrix with 24000 entries by 300 atoms.
+It is possible to observe better reconstruction rating for tensor-based algorithms.
+
+Imbalanced data...
+Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate.
+This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security.
+However, the high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes.
+The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms
+
+A key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset.
+Therefore, we introduce a tensor-based dictionary learning method for fraud detecton from unbalanced data.
+Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification.
+In this paper, we propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds.
+We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
+
+%However, tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.
+We propose a distributed and tensor-based approach for multidimensional dictionary learning in order to obtain better processing time for larger datasets.
+We focus on a distributed implementation of T-MOD \cite{roemer2014tensor} algorithm, based on a modification of Almeida and Kibangou's \cite{de2014distributed} approach to use Tucker-2 decomposition.
+We perform experiments to test different methods for learning dictionaries, evaluating the processing time and the reconstruction of a dictionary of multidimensional and separable data.
 
@@
 \label{sc:problems}
-...
+
+In the context of anomaly-based schemes, this work proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
+Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
+The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
+In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack.
+In addition, we also evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset.
+Furthermore, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.
+
+This thesis also proposes a novel approach based on user behavior analysis through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack.
+Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.
+The main target of this proposed solution is to provide a maximum defense at the minimal resource cost.
+
+Finally, we propose a fraud detection system for mobile payment transactions...
+we propose to apply the tensor deconposition for dictionary learning in order to evidentiate the discriminative sensing of a fraud detection dataset.
 
@@
 \label{sc:contributions}
-...
+
+The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
+The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
 
@@
 \label{sc:organization}
-...
+
+This thesis is organized as follows.
+In Chapter \ref{ch:mos_eig_sim}, we propose a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
+Chapter \ref{ch:mobile} presents an evaluation of an approach and architecture based on user behavior analysis through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats in a mobile application.
+Chapter \ref{ch:tensor_dl} proposes a tensor-based dictionary learning approach for fraud detection from mobile payment transactions.
+Chapter \ref{ch:conclusionfuturework} draws the conclusions and the suggestions for future work.
+Futhermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminant analysis and based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.

### 2017-10-27 docs/thesis/chapters/1_introduction/introduction.tex
conclusion and introduction

@@
 \chapter{Introduction}
-\label{ch:introduction}
+\label{ch:1_introduction}
 
-\begin{quotation}[]{Chico Xavier}
-Though nobody can go back and make a new beginning, anyone can start over and make a new ending.
+\begin{quotation}[]{Lao Tzu}
+The softest things in the world overcome the hardest things in the world.
 \end{quotation}
@@
 Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection.
-For example, offline mode does not allow performing the extensive computation and encryption on the mobile device.
 
@@
 Big data analytics require techniques to deal with multidimensional data, making sense of strucutre and relationship of many dimensions.
+
 Dictionary learning is a signal processing technique based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world.
@@
 
-Observing the above described landscape, this paper outlines the concept of the offline mobile cloud security.
+Observing the above described landscape, this thesis outlines the concept of the offline mobile cloud security.
 
@@
 
-\begin{figure}[!htb]
-     \centering 
-	 \includegraphics[height=8cm, width=11cm]{figures/5_20_2000_1000_100.eps}
-     \caption{Matrix of 1.000 elements, 2000 sample training, 100 iterations}
-     \label{fig:fig1}
-\end{figure}
-
-\begin{figure}[!htb]
-     \centering 
-	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_24000_100.eps}
-     \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}
-     \label{fig:fig2}
-\end{figure}
-
 Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based and traditional dictionary learning algorithms for dictionary reconstruction of a multidimensional and separable data.
@@
 
-In the context of anomaly-based schemes, this work proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
-Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
+In the context of anomaly-based schemes, this thesis proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
+Inspired by \cite{david2011blind,da2012improved}, this thesis models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
 The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
@@
 This thesis is organized as follows.
-In Chapter \ref{ch:mos_eig_sim}, we propose a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
-Chapter \ref{ch:mobile} presents an evaluation of an approach and architecture based on user behavior analysis through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats in a mobile application.
-Chapter \ref{ch:tensor_dl} proposes a tensor-based dictionary learning approach for fraud detection from mobile payment transactions.
-Chapter \ref{ch:conclusionfuturework} draws the conclusions and the suggestions for future work.
+In Chapter \ref{ch:2_mos_eig_sim}, we propose a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
+Chapter \ref{ch:3_mobile} presents an evaluation of an approach and architecture based on user behavior analysis through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats in a mobile application.
+Chapter \ref{ch:4_tensor_dl} proposes a tensor-based dictionary learning approach for fraud detection from mobile payment transactions.
+Chapter \ref{ch:5_conclusionfuturework} draws the conclusions and the suggestions for future work.
 Futhermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminant analysis and based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.

### 2017-10-27 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
introducion...
conclusion and introduction

--- 2017-10-17 docs/thesis/chapters/3_eigen_similarity/eigen_similarity.tex
@@
 \chapter{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
-\label{ch:evaluation}
+\label{ch:2_mos_eig_sim}
 
@@
 
-In the context of anomaly-based schemes, this work proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
-Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
+In the context of anomaly-based schemes, this thesis proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
+Inspired by \cite{david2011blind,da2012improved}, this thesis models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
 To the best of our knowledge there is no similar model in the literature.
@@
 
-This paper is organized as follows.
-In Section \ref{sec:relatedworks}, related works are discussed.
-Section \ref{sec:datamodel} presents the data model and the evaluated datasets.
-Section \ref{sec:prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks.
-Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:Complexity} discusses the computational complexity of the proposed framework and evaluates the required processing time for tested scenarios.
-Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
-The \ref{sec:mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.
+This chapter is organized as follows.
+In Section \ref{sec:2_relatedworks}, related works are discussed.
+Section \ref{sec:2_datamodel} presents the data model and the evaluated datasets.
+Section \ref{sec:2_prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks.
+Section \ref{sec:2_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:2_Complexity} discusses the computational complexity of the proposed framework and evaluates the required processing time for tested scenarios.
+Section \ref{sec:2_conclusionandfutureworks} draws the conclusions and the suggestions for future work.
+The \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.
 
 \section{Related Works}
-\label{sec:relatedworks}
+\label{sec:2_relatedworks}
 
@@
 \section{Data Model}
-\label{sec:datamodel}
+\label{sec:2_datamodel}
 
@@
 This section presents details of the experimental scenario and the selected cases of the DARPA dataset, along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic.
-Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
-Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition.
-Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
+Subsection \ref{sec:2_DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
+Subsection \ref{sec:2_ModelingData} presents how network traffic can be modeled as signal superposition.
+Subsection \ref{sec:2_SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:2_DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
 
 \subsection{Analyzed Scenario and Data Collection}
-\label{sec:DataCollection}
+\label{sec:2_DataCollection}
 
 The environment of the analyzed scenario is composed of two computers and one router with access to the Internet and to an internal network, where the simulation of legitimate traffic, noise, flood and port scan attacks are performed.
-During the traffic generation, one computer assumes the role of the attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:fig1}.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=6cm, width=9cm]{figures/fig09.png}
+During the traffic generation, one computer assumes the role of the attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:2_fig1}.
+
+\begin{figure}[h!]
+     \centering 
+     \includegraphics[width=11cm]{figures/fig09.png}
      \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
-     \label{fig:fig1}
+     \label{fig:2_fig1}
 \end{figure}
@@
 \subsection{Modeling Data}
-\label{sec:ModelingData}
+\label{sec:2_ModelingData}
 
@@
 When a user accesses a web page, for example, there is the corresponding TCP/IP traffic to request the page, as well as there is the traffic required to domain name resolution.
-Figure \ref{fig:fig3} depicts an example of the legitimate traffic obtained during experiments.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{figures/fig03.png}
+Figure \ref{fig:2_fig3} depicts an example of the legitimate traffic obtained during experiments.
+
+\begin{figure}[h!]
+     \centering 
+     \includegraphics[width=11cm]{figures/fig03.png}
      \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
-     \label{fig:fig3}
+     \label{fig:2_fig3}
 \end{figure}
@@
 Independently of any user operation, the machine receives an IP address, since it is configured to perform a DHCP address request.
-Figure \ref{fig:fig4} depicts an example of noise in a network traffic, represented by traffic to ports 67 and 68.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{figures/fig04.png}
+Figure \ref{fig:2_fig4} depicts an example of noise in a network traffic, represented by traffic to ports 67 and 68.
+
+\begin{figure}[h!]
+     \centering 
+     \includegraphics[width=11cm]{figures/fig04.png}
      \caption{Network traffic of user independent operations for network management.}
-     \label{fig:fig4}
+     \label{fig:2_fig4}
 \end{figure}
@@
 \subsection{Synflood, Fraggle and Port scan}
-\label{sec:SynfloodFraggleandPortscan}
+\label{sec:2_SynfloodFraggleandPortscan}
 
@@
 With respect to the synflood attacks, the attacker sends a large quantity and concurrent successive SYN requests to a target, in order to consume resources and cause a DoS.
-Figure \ref{fig:fig5} depicts an example of a synflood attack carried out in a real computer network.
+Figure \ref{fig:2_fig5} depicts an example of a synflood attack carried out in a real computer network.
 In an interval of ten minutes, more than 210,000 packets are sent as a synflood attack.
@@
      \centering 
-     \includegraphics[height=4.5cm, width=9cm]{figures/fig05.png}
+     \includegraphics[width=11cm]{figures/fig05.png}
      \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
-     \label{fig:fig5}
+     \label{fig:2_fig5}
 \end{figure}
@@
 This last part of the fraggle attack is not taken into account in this work, because the victim receives ICMP (network layer) packets originated from the hosts that are attacked with flooding packet UDP echo.
-Figure \ref{fig:fig6} depicts an example of the fraggle attack in a real computer network.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{figures/fig06.png}
+Figure \ref{fig:2_fig6} depicts an example of the fraggle attack in a real computer network.
+
+\begin{figure}[h!]
+     \centering 
+     \includegraphics[width=11cm]{figures/fig06.png}
      \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
-     \label{fig:fig6}
+     \label{fig:2_fig6}
 \end{figure}
@@
      \centering 
-     \includegraphics[height=5cm, width=9cm]{figures/fig07.png}
+     \includegraphics[width=11cm]{figures/fig07.png}
      \caption{Connection attempts in order to identify active ports.}
-     \label{fig:fig7}
-\end{figure}
-
-Figure \ref{fig:fig7} depicts an example of the port scan attack in a real computer network.
+     \label{fig:2_fig7}
+\end{figure}
+
+Figure \ref{fig:2_fig7} depicts an example of the port scan attack in a real computer network.
 Note that the traffic is composed of two packets for each TCP port and one UDP packet to each port.
@@
 \subsection{The DARPA Dataset}
-\label{sec:DarpaDataset}
+\label{sec:2_DarpaDataset}
 
@@
 
-The data modeling follows the method described by the Subsection \ref{sec:ModelingData}, with time frames of 20 minutes, packet counting aggregation by minute and considering the traffic to the following ports: 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389, 443.
+The data modeling follows the method described by the Subsection \ref{sec:2_ModelingData}, with time frames of 20 minutes, packet counting aggregation by minute and considering the traffic to the following ports: 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389, 443.
 
 \section{Proposed Framework for Detection and Identification of Network Attacks}
-\label{sec:prop_getv}
-
-This section describes the proposed technique to detect synflood, fraggle and port scan, according to Figure \ref{fig:fig80}, which represents the overview of the proposed framework for detection and identification of network attacks.
-In Subsection \ref{sec:prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvelue for each $q$-th time frame.
-Next, in Subsection \ref{sec:prop_MOSSchemes}, we show how to apply the eiganvalues on the MOS scheme in order to detect the attack.
-In Subsection \ref{sec:prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:prop_EigenSimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=11cm, width=9cm]{figures/mos_eigen_similarity.eps}
+\label{sec:2_prop_getv}
+
+This section describes the proposed technique to detect synflood, fraggle and port scan, according to Figure \ref{fig:2_fig80}, which represents the overview of the proposed framework for detection and identification of network attacks.
+In Subsection \ref{sec:2_prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvelue for each $q$-th time frame.
+Next, in Subsection \ref{sec:2_prop_MOSSchemes}, we show how to apply the eiganvalues on the MOS scheme in order to detect the attack.
+In Subsection \ref{sec:2_prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:2_prop_EigenSimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.
+
+\begin{figure}[h!]
+	\centering
+     \includegraphics[width=11cm]{figures/mos_eigen_similarity.eps}
      \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
-     \label{fig:fig80}
+     \label{fig:2_fig80}
 \end{figure}
@@
 \subsection{Largest Eigenvalue by Time Frames}
-\label{sec:prop_LargestEigenvaluebyTimeFrames}
+\label{sec:2_prop_LargestEigenvaluebyTimeFrames}
 
@@
 According to flood and port scan attacks' behavior, flood attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively.
-These characteristics are substantiated by the results obtained through the analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
+These characteristics are substantiated by the results obtained through the analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:2_experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
 
@@
 \subsection{MOS Schemes}
-\label{sec:prop_MOSSchemes}
+\label{sec:2_prop_MOSSchemes}
 
@@
 Note that some MOS schemes may also require the number of minutes that compose a time frame, as $\hat{d} = \rm{MOS}(\boldsymbol{e}_{\rm max},\emph{Q})$.
-For more information about MOS, we refer to \ref{sec:mos}.
+For more information about MOS, we refer to \ref{apx:a_mos}.
 
 In our previous work \cite{tenorio2013greatest}, the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE schemes are evaluated for synflood and port scan attack detection, showing that EDC and EFT are effective for detecting this kind of attacks.
-The present work extends that evaluation to also analyze the effectiveness of the listed MOS schemes for fraggle attack detection, as shown in Section \ref{sec:experimentalresults}.
+The present work extends that evaluation to also analyze the effectiveness of the listed MOS schemes for fraggle attack detection, as shown in Section \ref{sec:2_experimentalresults}.
 
 \subsection{Eigenvalue Analysis}
-\label{sec:prop_EigenvalueAnalysis}
+\label{sec:2_prop_EigenvalueAnalysis}
 
@@
 
-In Subsection \ref{sec:prop_MOSSchemes}, we obtained only if $\hat{d} = 1$ or $\hat{d} = 0$.
+In Subsection \ref{sec:2_prop_MOSSchemes}, we obtained only if $\hat{d} = 1$ or $\hat{d} = 0$.
 However, if $\hat{d} = 1$, the MOS schemes do not provide any information about the $q$-th time frame under attack.
@@
 
-The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed according to Algorithm \ref{alg:alg01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$.
-Algorithm \ref{alg:alg01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$, according to Line 2 of Algorithm \ref{alg:alg01}, and its correspondent index, according to Lines 4 and 5 of Algorithm \ref{alg:alg01}.
-Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$, according to Line 8 of Algorithm \ref{alg:alg01}, and a new iteration is performed until $\boldsymbol{e}_{\rm max} = \{\}$.
+The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed according to Algorithm \ref{alg:2_alg01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$.
+Algorithm \ref{alg:2_alg01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$, according to Line 2 of Algorithm \ref{alg:2_alg01}, and its correspondent index, according to Lines 4 and 5 of Algorithm \ref{alg:2_alg01}.
+Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$, according to Line 8 of Algorithm \ref{alg:2_alg01}, and a new iteration is performed until $\boldsymbol{e}_{\rm max} = \{\}$.
 
 \begin{algorithm}[h!]
-	% \caption{Detection of Time Frames Under Attack}
- %  	\label{alg:alg01}
+	\caption{Detection of Time Frames Under Attack}
+  	\label{alg:2_alg01}
 	% \begin{algorithmic}[1]
@@
 \subsection{Eigen Similarity Analysis}
-\label{sec:prop_EigenSimilarityAnalysis}
+\label{sec:2_prop_EigenSimilarityAnalysis}
 
@@
 This subsection describes the proposed eigen similarity analysis for detailed attack identification, in complement to the attack estimation carried out through MOS schemes and eigenvalue analysis.
-In Subsection \ref{sec:prop_TimeSimilarityAnalysis} we present the eigen similarity analysis for identification of time under attack.
-Next, in Subsection \ref{sec:prop_PortSimilarityAnalysis}, we show how to apply the eigen similarity analysis in order to identify network ports under attack.
+In Subsection \ref{sec:2_prop_TimeSimilarityAnalysis} we present the eigen similarity analysis for identification of time under attack.
+Next, in Subsection \ref{sec:2_prop_PortSimilarityAnalysis}, we show how to apply the eigen similarity analysis in order to identify network ports under attack.
 
 \subsubsection{Time Similarity Analysis}
-\label{sec:prop_TimeSimilarityAnalysis}
+\label{sec:2_prop_TimeSimilarityAnalysis}
 
@@
 
-Figure \ref{fig:fig8} ilustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.
-
-\begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{figures/incremental.eps}
+Figure \ref{fig:2_fig8} ilustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.
+
+\begin{figure}[h!]
+     \includegraphics[width=15cm]{figures/incremental.eps}
      \caption{Traffic selection for incremental approach.}
-     \label{fig:fig8}
+     \label{fig:2_fig8}
 \end{figure}
@@
 
-The eigen similarity analysis can also be applied in an individual fashion, where each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, as shown by Figure \ref{fig:fig9}.
-
-\begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{figures/individualized.eps}
+The eigen similarity analysis can also be applied in an individual fashion, where each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, as shown by Figure \ref{fig:2_fig9}.
+
+\begin{figure}[h!]
+     \includegraphics[width=15cm]{figures/individualized.eps}
      \caption{Traffic selection for individual approach.}
-     \label{fig:fig9}
+     \label{fig:2_fig9}
 \end{figure}
@@
 The incremental and the individual approaches can be combined to obtain the incremental individualized approach, where each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
-Subsequently, $\boldsymbol{X}_n$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:fig2}.
-
-\begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{figures/incremental_individualized.eps}
+Subsequently, $\boldsymbol{X}_n$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:2_fig2}.
+
+\begin{figure}[h!]
+     \includegraphics[width=15cm]{figures/incremental_individualized.eps}
      \caption{Traffic selection for incremental individualized approach.}
-     \label{fig:fig2}
+     \label{fig:2_fig2}
 \end{figure}
@@
 This approach of incremental similarity analysis followed by individual analysis after an attack detection allows to identify the attack period, highlighting the first and last time under attack.
-This identification is possible due to the variation of the most significant eigenvectors, which becomes more significant when compared a traffic under attack against a traffic with no attack, according to results which are discussed in Section \ref{sec:experimentalresults}.
+This identification is possible due to the variation of the most significant eigenvectors, which becomes more significant when compared a traffic under attack against a traffic with no attack, according to results which are discussed in Section \ref{sec:2_experimentalresults}.
 
 \subsubsection{Port Similarity Analysis}
-\label{sec:prop_PortSimilarityAnalysis}
+\label{sec:2_prop_PortSimilarityAnalysis}
 
@@
 \section{Experiments and Results}
-\label{sec:experimentalresults}
+\label{sec:2_experimentalresults}
 
 This section presents the performed experiments and the acquired results.
-First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
-Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario.
-In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated dataset.
-Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario.
-Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 dataset.
+First, in Section \ref{sec:2_AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
+Then, Section \ref{sec:2_largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario.
+In Section \ref{sec:2_MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated dataset.
+Section \ref{sec:2_EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:2_EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario.
+Section \ref{sec:2_DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 dataset.
 
 \subsection{Experimental Scenario}
-\label{sec:AnalyzedScenario}
+\label{sec:2_AnalyzedScenario}
 
@@
 \subsection{Largest Eigenvalues Analysis}
-\label{sec:largesteigenvaluesanalysis}
-
-For the evaluation of MOS Schemes accuracy for flood and port scan detection, the framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}.
+\label{sec:2_largesteigenvaluesanalysis}
+
+For the evaluation of MOS Schemes accuracy for flood and port scan detection, the framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:2_prop_getv}.
 Through eigenvalue analysis of traffic with flood or port scan attacks, it is possible to visualize a significant difference between the largest eigenvalues and the remain eigenvalues, which can indicate a relationship between an outlier and time frames under attack.
@@
 	\centering
-     \includegraphics[height=6cm, width=8cm]{figures/eigenvalues_synflood.eps} 
+     \includegraphics[width=11cm]{figures/eigenvalues_synflood.eps} 
      \caption{Eigenvalues of the sample covariance matrix (synflood).}
-     \label{fig:fig10}
-\end{figure}
-
-Figure \ref{fig:fig10} depicts the eigenvalues calculated from sample covariance matrix of the network traffic used to evaluate the synflood attack identification.
-In Figure \ref{fig:fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 4$) stands out significantly from the other eigenvalues.
-
-Figure \ref{fig:fig11} illustrates the eigenvalues calculated from sample covariance matrix of the matrix used for fraggle attack detection.
-In Figure \ref{fig:fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 5$) stands out significantly from the other eigenvalues, in accordance with the result shown in Figure \ref{fig:fig10} for the synflood attack analysis.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=6cm, width=8cm]{figures/eigenvalues_fraggle.eps}
+     \label{fig:2_fig10}
+\end{figure}
+
+Figure \ref{fig:2_fig10} depicts the eigenvalues calculated from sample covariance matrix of the network traffic used to evaluate the synflood attack identification.
+In Figure \ref{fig:2_fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 4$) stands out significantly from the other eigenvalues.
+
+Figure \ref{fig:2_fig11} illustrates the eigenvalues calculated from sample covariance matrix of the matrix used for fraggle attack detection.
+In Figure \ref{fig:2_fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 5$) stands out significantly from the other eigenvalues, in accordance with the result shown in Figure \ref{fig:2_fig10} for the synflood attack analysis.
+
+\begin{figure}[h!]
+	\centering
+     \includegraphics[width=11cm]{figures/eigenvalues_fraggle.eps}
      \caption{Eigenvalues of the sample covariance matrix (fraggle).}
-     \label{fig:fig11}
-\end{figure}
-
-Figure \ref{fig:fig12} depicts the eigenvalues calculated from covariance matrix of zero mean and unitary standard deviation variables, of the network traffic matrix evaluated for port scan detection.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=6cm, width=8cm]{figures/eigenvalues_portscan.eps}
+     \label{fig:2_fig11}
+\end{figure}
+
+Figure \ref{fig:2_fig12} depicts the eigenvalues calculated from covariance matrix of zero mean and unitary standard deviation variables, of the network traffic matrix evaluated for port scan detection.
+
+\begin{figure}[h!]
+	\centering
+     \includegraphics[width=11cm]{figures/eigenvalues_portscan.eps}
      \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}
-     \label{fig:fig12}
+     \label{fig:2_fig12}
 \end{figure}
@@
   \centering
-  \footnotesize
+  % \footnotesize
   \caption{Largest Eigenvalue related to attacks detection}
@@
 \subsection{MOS Schemes Evaluation}
-\label{sec:MOSSchemesEvaluation}
+\label{sec:2_MOSSchemesEvaluation}
 
 In \cite{tenorio2013greatest}, we evaluate the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE MOS schemes \cite{da2009comparison,tenorio2013greatest} for synflood and port scan attack detection.
-In this work we extend that evaluation for fraggle attack detection, applying the same schemes to fraggle attack detection over the traffic presented in Section \ref{sec:datamodel}, as results shown in Table \ref{tab:tab4}.
+In this work we extend that evaluation for fraggle attack detection, applying the same schemes to fraggle attack detection over the traffic presented in Section \ref{sec:2_datamodel}, as results shown in Table \ref{tab:tab4}.
 
@@
   \centering
-  \scriptsize
+  % \scriptsize
   \caption{MOS schemes applied to port scan and flood detection}
@@
 \subsection{Eigenvalue Analysis}
-\label{sec:EigenvalueAnalysis}
-
-According to the results presented in Section \ref{sec:largesteigenvaluesanalysis}, the largest eigenvalue stands out significantly from the others eigenvalues of an evaluated $q$-th time frame.
+\label{sec:2_EigenvalueAnalysis}
+
+According to the results presented in Section \ref{sec:2_largesteigenvaluesanalysis}, the largest eigenvalue stands out significantly from the others eigenvalues of an evaluated $q$-th time frame.
 This behavior can also be observed in the largest eigenvalues analysis, according to results presented in Table \ref{tab:tab3}, where it is possible to observe that the $\hat{d}$ largest eigen values of the time frames under attacks stand out significantly from the others largest eigenvalues.
 
-Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respectives $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:alg01}.
+Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respectives $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:2_alg01}.
 
 \subsection{Eigen Similarity Analysis}
-\label{sec:EigenSimilarityAnalysis}
+\label{sec:2_EigenSimilarityAnalysis}
 
@@
 \subsubsection{Time Analysis}
-\label{sec:TimeAnalysis}
+\label{sec:2_TimeAnalysis}
 
@@
 \subsubsection{Port Analysis}
-\label{sec:PortAnalysis}
+\label{sec:2_PortAnalysis}
 
@@
   \centering
-  \footnotesize
+  % \footnotesize
   \caption{Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n=15)}
@@
   \centering
-  \footnotesize
+  % \footnotesize
   \caption{Eigen Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
@@
   \centering
-  \footnotesize
+  % \footnotesize
   \caption{Eigen Similarity Analysis for Detection of Ports Under Fraggle Attack (q=5 and t=11)}
@@
 \subsection{DARPA Scenario}
-\label{sec:DarpaEvaluation}
+\label{sec:2_DarpaEvaluation}
 
 This subsection presents a summarized view of results obtained from the application of the proposed framework, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 dataset.
-Since the proposed framework is detailed in Section \ref{sec:prop_getv} and in Subsections \ref{sec:AnalyzedScenario}, \ref{sec:largesteigenvaluesanalysis}, \ref{sec:MOSSchemesEvaluation}, \ref{sec:EigenvalueAnalysis} and \ref{sec:EigenSimilarityAnalysis}, here the focus is on the parameter selection, dataset evaluation and results for flood and probe attack identification.
+Since the proposed framework is detailed in Section \ref{sec:2_prop_getv} and in Subsections \ref{sec:2_AnalyzedScenario}, \ref{sec:2_largesteigenvaluesanalysis}, \ref{sec:2_MOSSchemesEvaluation}, \ref{sec:2_EigenvalueAnalysis} and \ref{sec:2_EigenSimilarityAnalysis}, here the focus is on the parameter selection, dataset evaluation and results for flood and probe attack identification.
 
@@
 \section{Performance Evaluation}
-\label{sec:Complexity}
-
-This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
+\label{sec:2_Complexity}
+
+This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:2_fig80} and equations presented in Section \ref{sec:2_prop_getv}.
 
 \subsection{Complexity Analysis}
-\label{sec:ComplexityAnalysis}
+\label{sec:2_ComplexityAnalysis}
 The EVD, calculated according to (\ref{eq:eq06}), requires the previous calculation of covariance matrix, according to Equations (\ref{eq:eq02}), (\ref{eq:eq03}), (\ref{eq:eq04}) and (\ref{eq:eq05}).
@@
 
-The largest eigenvalue analysis is $O(\hat{d}Q)$, where $\hat{d}$ denotes the number of time frame under attack, according to Algorithm \ref{alg:alg01}.
+The largest eigenvalue analysis is $O(\hat{d}Q)$, where $\hat{d}$ denotes the number of time frame under attack, according to Algorithm \ref{alg:2_alg01}.
 Subsequently, the eigen similarity analysis relies on EVD and cosine similarity analysis, which is $O(N^2)$, for $\hat{d}$ time frames, therefore the eigen similarity analysis is which is $O(\hat{d}(M^2N + N^3 + N^2))$ and yields an $O(N^3)$ upper bound on the worst-case running time for eigen similarity analysis.
@@
 \subsection{Processing Time Analysis}
-\label{sec:ProcessingAnalysis}
+\label{sec:2_ProcessingAnalysis}
 For better understanding the scalability and impact of configurations of $N$, $M$ and $Q$, the processing time required for different scenarios of parameter configurations and dataset were evaluated, measuring the processing time of:
@@
 
-The performance evaluation focus on the main steps, which are discussed in subsection \ref{sec:ComplexityAnalysis} regarding the complexity analysis.
+The performance evaluation focus on the main steps, which are discussed in subsection \ref{sec:2_ComplexityAnalysis} regarding the complexity analysis.
 The data modeling is also a time consuming step, however its processing can be optimized through distributed processing techniques, such as MapReduce, achieving high throughput for packet counting or even for deep packet inspection \cite{Vieira2013}.
@@
 \section{Conclusion and Future Works}
-\label{sec:conclusionandfutureworks}
+\label{sec:2_conclusionandfutureworks}
 

### 2017-10-27 docs/thesis/chapters/5_conclusion_future_work/conclusion.tex
conclusion and introduction

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% na conclusao, evite apresentar os resultados mais fáceis de concluir. bons revisores olham o casamento do que vc escreve no abstract/intro com a conclusao. mantenha coerente. 
% em geral, resultados que sao aparentemente obvios, enfraquecem as outras contribuicoes. p.ex, "Os resultados mostraram que o MapReduce apresenta uma boa solução para utilizar computadores comuns para obter alta capacidade de processamento e lidar com a análise massiva de tráfego de rede." pode ser reformulado ou retirado da conclusao.
\chapter{Conclusion and Future Work}
\label{ch:5_conclusionfuturework}

\begin{quotation}[]{Chico Xavier}
Though nobody can go back and make a new beginning, anyone can start over and make a new ending.
\end{quotation}

% falar sobre uma contextualização de forma bem resumida'

In the context of anomaly-based schemes, this thesis proposed a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks. The proposed technique is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis, where MOS and eigenvalue analysis are applied to detect time frames under attack. In addition, we evaluated the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset.

This thesis also proposes a novel architecture and approach based on user behavior analysis through Model Order Selection (MOS), in order to detect possible threats, and to reduce the risks and the harm of the most common threats. The behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.

Finally, we proposed a fraud detection approach composed by tensor-based dictionary learning for discriminant analysis of a mobile payment dataset.

\section{Conclusion}
\label{sc:conc_conclusion}

This thesis models the network traffic as a signal processing formulation for applying the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.

The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of this thesis were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.

This thesis evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection. Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.

The incremental individualized approach of eigen similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to eigen similarity analyis for network attack detection. Therefore, the incremental individualized approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.

According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.

Considering the offline mobile security context, an important issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile clients. Once a mobile client releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device. Aware of this problem and its importance, this thesis presented a proposal to address the offline mobile security problem combining a different cryptographic methods. Moreover, this proposed approach also prevents malicious user behavior by applying a MOS-based analytic method.

As prove of concept, a fully working mobile application was developed to test the proposed security solution and acquired results provide evidence that besides achieving the desired security features, the solution also has positive results in terms of performance. This fact is due to the usage of lightweight operations and the optimized combination of the selected security methods. The proposed approach is a practical application to be used in the corporate mobile environment. It is implemented as a fully working mobile client and can be used for any type of enterprise. Also, part of concept is seed for new security solutions for big data apps.

\section{Contributions}
\label{sc:conc_contributions}

We attempt to analyse the processing capacity problem of measurement of distributed systems through network traffic analysis, the results of the work presented in this thesis provide the contributions below:

\begin{enumerate}
	\item We proposed an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset;
	\item We discussed the computational complexity of the proposed framework and evaluated the required processing time for tested scenarios;
	\item We proposed an architechture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
	\item We discussed the processing time of the proposed framework for mobile devices;
	\item We proposed an tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
	\item We \textbf{published five papers} reporting our results, as follows:
	\begin{enumerate}
		\item T. P. B. Vieira, D. F. Ten\'orio, J. P. C. L. da Costa, E. P. de Freitas, G. Del Galdo, and R. T. de Sousa J\'unior, \textit{Model Order Selection and Eigen Similarity based Framework for Detection and Identication of Network Attacks}. Journal of Networking and Computer Applications (JNCA), Vol 90, Jul 2017, Pages 26–41.
		\item T. Galibus, T. P. B. Vieira, E. P. de Freitas, R. Albuquerque, J. P. C. L. da Costa, R. T. de Sousa Jr, V. Krasnoproshin, A. Zaleski, H. Vissia, and G. del Galdo, \textit{Offline Mode for Corporate Mobile Client Security Architecture}. Mobile Networks and Applications, Springer, Mar 2017, Pages 1-17.
		\item K. H. C. Ramos, R. T. Sousa Jr, T. P. B. Vieira, and J. P. C. L. da Costa, \textit{DISCOVERING CRITICAL SUCCESS FACTORS FOR INFORMATION TECHNOLOGIES GOVERNANCE THROUGH BIBLIOMETRIC ANALYSIS OF RESEARCH PUBLICATIONS IN THIS DOMAIN}. Information (Yamaguchi), 2016
		\item K. H. C. Ramos, R. T. de Sousa Jr., T. P. B. Vieira, and J. P. C. L. da Costa, \textit{Identifying relevant IT governance critical success factors by means of bibliometric analysis}. 7th International Conference on Information, November 2015, Taipei, Taiwan.
		\item K. H. C. Ramos, T. P. B. Vieira, J. P. C. L. da Costa, and R. T. Sousa Jr., \textit{ANÁLISE MULTIDIMENSIONAL DE FATORES CRÍTICOS DE SUCESSO EM GOVERNANÇA DE TI NA ADMINISTRAÇÃO PÚBLICA FEDERAL À LUZ DOS DADOS DE CONTROLE EXTERNO}. in Proc. 12th International Conference on Management of Technology and Information Systems (CONTECSI), 2015, São Paulo, Brazil.
	\end{enumerate}
\end{enumerate}

\section{Future Work}
\label{sc:conc_futurework}

Because of time constraints imposed on the Ph.D. degree, this thesis addresses some problems, but some problems are still open and others are emerging from current results. Thus, the following issues should be investigated as future work:

\begin{itemize}
	\item Improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets;
	\item Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high thoughput network traffic;
	\item Evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis;
	\item Explore enhancements in the analytics methods as well as to extended the approach to be used by mobile devices with even more severe resources constraints;
\end{itemize}

### 2017-11-08 docs/thesis/chapters/1_introduction/introduction.tex
justificatives and problem statement

@@
 Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
-Recently, signal processing schemes have being applied to the detection of malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest}, showing advances in network traffic analysis.
+Recently, signal processing schemes have being applied to detect malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest, vieira2017model}, showing advances in network traffic analysis.
 
@@
 Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation.
-The basic features of the cloud environment is providing the elastic, on-demand and secure services for the end-users.
+The basic features of the cloud environment is providing elastic, on-demand and secure services for the end-users.
 While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}.
@@
 Additional security issues and requirements have to be considered when mobile clients are actively used in corporate cloud environment \cite{yovel2014}.
-Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm.
+Today organizations and enterprises are adopting the Bring-Your-Own-Device (BYOD) paradigm.
 The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS).
@@
 Obtaining access to data sets of mobile transactions for research is a very hard task due to the intrinsic private nature of such transactions.
-Scientists and researchers must today spend time and effort in obtaining permits and access to relevant data sets before they can research on such data set.
+Scientists and researchers must today spend time and effort in obtaining access to relevant data sets before they can research on such data set.
 However, PaySim is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.\footnote{https://www.kaggle.com/ntnu-testimon/paysim1}
 
+Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate.
+This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security.
 Despite the actual high availability of information, the relevant information of some observations is generally of under much reduced dimensionality compared to available data sets.
 The extraction of relevant information by identifying the generating causes within classes of signals is useful for classification problems and for security analysis.
-Big data analytics require techniques to deal with multidimensional data, making sense of strucutre and relationship of many dimensions.
 
-Dictionary learning is a signal processing technique based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world.
-Dictionary Learning is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
+The high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes.
+The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms.
+Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection.
+Considering that imbalanced data can be composed by rare or unknown events, adaptive techniques can be useful in order to improve models and be able to react to changes or unknown patterns and maintain acceptable performance for classification algorithms.
+Adaptive techniques are also important for online data analysis or streaming processing, where an incremental processing approach can be applied to deal with high thoughput data analysis.
+
+Compressed Sensing (CS) aims to obtain the most relevant information of a dataset, what makes it useful for compression, signal reconstruction, image processing and classification, such as in sparce representation classification (SRC) problems.
+Dictionary learning is related to CS but seeks to obtain compact and meaningful signal representations, known as dictionary, from learning algorithms or models.
+DL is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
+DL is based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world.
 The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.
 
-Tensor decomposition has been adopted for multidimensional data analysis and can be explored for dictionary learning...principalmente por tensor decompositio extrair mais informações e relacionamentos que técnicas tradicionais, podendo ser melhores que técnicas tradicionais de dictionary learning
-
+Dictionaries are either available analytically, or can be learned from a suitable training set.
+DL can be formulated as an adptive algorithm that learns from traning data that is updated to represent changes or to adpat to new patterns.
 Iterative dictionary learning has been used as a blind technique for feature extraction, improving classification without require feature selection or principal conponent analysis
 
-We propose methods for information security analysis
-
-Observing the above described landscape, this thesis outlines the concept of the offline mobile cloud security.
-
-Compressive sensing is emergent for signal processing and has been applied to fields such as signal restoration, image processing and classification (sparce representation classification src )
-
-Adaptive techniques for high dimensional and high thoughput
-
-% There are two major approaches for dictionary learning.
-First is the analytic approach, in which DCT bases, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries.
-Second is the learning-based approaches, such as the unsupervised learning for dictionary construction [9] and the online dictionary learning [11], [10], which use machine learning methods to construct the dictionary.
-In the analytic approach, some pre-defined functions are used to construct the dictionary.
-Curvelets [24], which tracks the shape of the discontinuity set, offers efficient and near-optimal representation of smooth objects.
-Shearlets [25], which is obtained from dilations, action of translations and shear transformations, has nice geometric properties and mathematical properties for image representation.
-Bandelets [26], which specifies the geometry as a vector field, is de-signed to improve the image compression and noise reduction performance.
-In the learning-based approach, machine learning methods are used to construct the dictionary from the training data.
-The least square error is used by the method of the optimal directions (MOD) [27] to update the dictionary iteratively:
-% Dictionaries are either available analytically, or can be learned from a suitable training set.
-While analytic dictionaries permit to capture the global structure of a signal and allow a fast implementation, learned dictionaries often perform better in applications as they are more adapted to the considered class of signals.
-
-In some applications the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks.
+In some applications, the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks.
 Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in multidimensional data analytics problems \cite{kolda2009tensor}.
 Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}.
-
-% In imagery, the numerical burden for (i) learning a dictionary and for (ii) employing the dictionary for reconstruction tasks only allows to deal with relatively small image patches that only capture local image information.
-Separable dictionaries aims at overcoming these drawbacks by allowing a separable structure on the dictionary throughout the learning process.
-On the one hand, this permits larger patch-sizes for the learning phase, on the other hand, the dictionary is applied efficiently in reconstruction tasks.
-The crucial idea is to allow the dictionary to have a separable structure, where separable means that the dictionary D is given by the Kronecker product of two smaller dictionaries A ∈ R h×a and B ∈ R w×b
-% Without loss of generality we choose square image patches with w = h = 8, which is in accordance to the patch-sizes mostly used in the literature
-
-Multidimensional parameter estimation and learning multidimensional separable dictionaries are growing research problems.
-Roemer \emph{et al.} \cite{roemer2014tensor} show that the multidimensional dictionary estimation problem can be efficiently formulated in terms of tensors, and that their results outperform existing schemes by exploiting the multilinear structure of the problem.
-
 Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results.
-However, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets, as can be seen in Figures \ref{fig:fig1} and \ref{fig:fig2}, which show that the recovering rating of traditional algorithms decreases over the dataset increasing.
-
-Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based and traditional dictionary learning algorithms for dictionary reconstruction of a multidimensional and separable data.
-Figure \ref{fig:fig1} presents the results for dictionary reconstruction from a matrix of 1000 elements by 50 atoms, while Figures \ref{fig:fig2} presents the results from a matrix with 24000 entries by 300 atoms.
-It is possible to observe better reconstruction rating for tensor-based algorithms.
-
-Imbalanced data...
-Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate.
-This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security.
-However, the high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes.
-The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms
-
-A key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset.
-Therefore, we introduce a tensor-based dictionary learning method for fraud detecton from unbalanced data.
-Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification.
-In this paper, we propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds.
-We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
-
-%However, tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.
-We propose a distributed and tensor-based approach for multidimensional dictionary learning in order to obtain better processing time for larger datasets.
-We focus on a distributed implementation of T-MOD \cite{roemer2014tensor} algorithm, based on a modification of Almeida and Kibangou's \cite{de2014distributed} approach to use Tucker-2 decomposition.
-We perform experiments to test different methods for learning dictionaries, evaluating the processing time and the reconstruction of a dictionary of multidimensional and separable data.
+However, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets.
 
@@
 
-In the context of anomaly-based schemes, this thesis proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
+Considering the above described landscape, this thesis outlines the development and evaluation of approaches based on signal processing for information security analysis, through methods to make the data discriminative and able to identify structures, hidden patterns and the most relevant information for network attack detection, mobile malicious behavior analysis and fraud detection in mobile money transactions.
+
+In the context of anomaly-based schemes, this thesis proposes a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks.
 Inspired by \cite{david2011blind,da2012improved}, this thesis models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
@@
 
-This thesis also proposes a novel approach based on user behavior analysis through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common threats, which are the expired user misusing password and the intruder attack.
+This thesis also proposes an architecture and approach based on user behavior analysis of mobile apps through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common security threats to corporate mobile apps, which are the expired user misusing password and the intruder attack.
 Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.
@@
 
-Finally, we propose a fraud detection system for mobile payment transactions...
-we propose to apply the tensor deconposition for dictionary learning in order to evidentiate the discriminative sensing of a fraud detection dataset.
+Considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, and also considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, we also propose a tensor-based dictionary learning method for fraud detecton from unbalanced data, in order to apply the tensor deconposition for dictionary learning methods in order to evidentiate the discriminative sensing of a fraud detection dataset.
+Specifically, we propose to apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification of mobile money transactions.
+We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
 
@@
 The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
+
+architecture and performance evaluation for mobile behavioral analysis
+
+an adaptive fraud detecion technique for imbalanced data
+
+applied to security analysis, such as network attack detection, malicious behavios in mobile apps and fraud detection in mobile payment transacions
 

### 2017-11-08 docs/thesis/chapters/4_tensor_dl/tensor_dl.tex
introducion...
conclusion and introduction
justificatives and problem statement

--- 2017-10-21 docs/thesis/chapters/5_tensor_dl/tensor_dl.tex
@@
 \chapter{Tensor-based Discriminative Sensing for Fraud Detection}
-\label{ch:tensor_dl}
+\label{ch:4_tensor_dl}
 
-\section{Introduction}
-\label{sec:introduction}
+\begin{quotation}[]{Paulo Freire}
+No one knows it all. No one is ignorant of everything. We all know something. We are all ignorant of something.
+\end{quotation}
 
@@
 
-% There are two major approaches for dictionary learning.
+There are two major approaches for dictionary learning.
 First is the analytic approach, in which DCT bases, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries.
@@
 The least square error is used by the method of the optimal directions (MOD) [27] to update the dictionary iteratively:
-% Dictionaries are either available analytically, or can be learned from a suitable training set.
+
+Dictionaries are either available analytically, or can be learned from a suitable training set.
 While analytic dictionaries permit to capture the global structure of a signal and allow a fast implementation, learned dictionaries often perform better in applications as they are more adapted to the considered class of signals.
@@
 
-% In imagery, the numerical burden for (i) learning a dictionary and for (ii) employing the dictionary for reconstruction tasks only allows to deal with relatively small image patches that only capture local image information.
+In imagery, the numerical burden for (i) learning a dictionary and for (ii) employing the dictionary for reconstruction tasks only allows to deal with relatively small image patches that only capture local image information.
 Separable dictionaries aims at overcoming these drawbacks by allowing a separable structure on the dictionary throughout the learning process.
@@
 The crucial idea is to allow the dictionary to have a separable structure, where separable means that the dictionary D is given by the Kronecker product of two smaller dictionaries A ∈ R h×a and B ∈ R w×b
+
 % Without loss of generality we choose square image patches with w = h = 8, which is in accordance to the patch-sizes mostly used in the literature
@@
      \centering 
-	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_1000_100.eps}
+	 \includegraphics[width=12cm]{figures/5_20_2000_1000_100.eps}
      \caption{Matrix of 1.000 elements, 2000 sample training, 100 iterations}
@@
      \centering 
-	 \includegraphics[width=0.4\textwidth]{figures/5_20_2000_24000_100.eps}
+	 \includegraphics[width=12cm]{figures/5_20_2000_24000_100.eps}
      \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}

### 2017-11-08 docs/thesis/chapters/5_conclusion_future_work/conclusion.tex
justificatives and problem statement

@@
 	\item We proposed an tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
-	\item We \textbf{published five papers} reporting our results, as follows:
+	\item We published the following papers reporting our results:
 	\begin{enumerate}
@@
 de Sousa J\'unior, \textit{Model Order Selection and Eigen Similarity based Framework for Detection and Identication of Network Attacks}.
-Journal of Networking and Computer Applications (JNCA), Vol 90, Jul 2017, Pages 26–41.
+Journal of Networking and Computer Applications (JNCA), Vol 90, Jul 2017, Pages 26–41 \cite{vieira2017model}.
 		\item T.
@@
 del Galdo, \textit{Offline Mode for Corporate Mobile Client Security Architecture}.
-Mobile Networks and Applications, Springer, Mar 2017, Pages 1-17.
+Mobile Networks and Applications, Springer, Mar 2017, Pages 1-17 \cite{galibus2017offline}.
 		\item K.
@@
 L.
-da Costa, \textit{DISCOVERING CRITICAL SUCCESS FACTORS FOR INFORMATION TECHNOLOGIES GOVERNANCE THROUGH BIBLIOMETRIC ANALYSIS OF RESEARCH PUBLICATIONS IN THIS DOMAIN}.
-Information (Yamaguchi), 2016
-		\item K.
-H.
-C.
-Ramos, R.
-T.
-de Sousa Jr., T.
-P.
-B.
-Vieira, and J.
-P.
-C.
-L.
-da Costa, \textit{Identifying relevant IT governance critical success factors by means of bibliometric analysis}.
-7th International Conference on Information, November 2015, Taipei, Taiwan.
+da Costa, \textit{Discovering Critical Success Factors for Information Technologies Governance through Bibliometric Analysis of Research Publications in This Domain}.
+Information (Yamaguchi), 2016 \cite{ramos2016information}.
 		\item K.
@@
 in Proc.
-12th International Conference on Management of Technology and Information Systems (CONTECSI), 2015, São Paulo, Brazil.
+12th International Conference on Management of Technology and Information Systems (CONTECSI), 2015, São Paulo, Brazil \cite{}.
 	\end{enumerate}

### 2017-11-09 docs/thesis/chapters/1_introduction/introduction.tex
chapter 2 ok

@@
 While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}.
-Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized leading to obscure and incomplete security paradigms and solutions.
+Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized, leading to obscure and incomplete security paradigms and solutions \cite{galibus2017offline}.
 
@@
 Today organizations and enterprises are adopting the Bring-Your-Own-Device (BYOD) paradigm.
-The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS).
+The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms, being the bottleneck of the corporate information security system (ISS).
 While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, the corporate mobile client is usually light-weighted and generally less protected.
@@
 Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection.
+
 Considering that imbalanced data can be composed by rare or unknown events, adaptive techniques can be useful in order to improve models and be able to react to changes or unknown patterns and maintain acceptable performance for classification algorithms.
@@
 Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results.
-However, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets.
+Additionally, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets.
 
@@
 
-This thesis also proposes an architecture and approach based on user behavior analysis of mobile apps through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common security threats to corporate mobile apps, which are the expired user misusing password and the intruder attack.
+This thesis also proposes an architecture and approach based on user behavior analysis of mobile apps through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common security threats to offline corporate mobile apps, which are the expired user misusing password and the intruder attack.
 Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.
@@
 
-Considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, and also considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, we also propose a tensor-based dictionary learning method for fraud detecton from unbalanced data, in order to apply the tensor deconposition for dictionary learning methods in order to evidentiate the discriminative sensing of a fraud detection dataset.
+Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, we also propose a tensor-based dictionary learning method for fraud detecton from imbalanced data, in order to apply the tensor deconposition for dictionary learning methods to evidentiate the discriminative sensing of a fraud detection dataset.
 Specifically, we propose to apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification of mobile money transactions.
@@
 
-The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
+We analyse problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques.
+The results of the work presented in this thesis provide the following contributions:
 
-architecture and performance evaluation for mobile behavioral analysis
-
-an adaptive fraud detecion technique for imbalanced data
-
-applied to security analysis, such as network attack detection, malicious behavios in mobile apps and fraud detection in mobile payment transacions
+\begin{enumerate}
+	\item We propose an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset;
+	\item We discuss the computational complexity of the proposed framework and evaluate the required processing time for tested scenarios;
+	\item We propose an architechture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
+	\item We discuss the processing time of the proposed framework for mobile devices;
+	\item We propose a tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
+\end{enumerate}
 
@@
 Chapter \ref{ch:5_conclusionfuturework} draws the conclusions and the suggestions for future work.
-Futhermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminant analysis and based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.
+Futhermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminant analysis, and presents an approach based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.

### 2017-11-09 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
chapter 2 ok

@@
 
-% In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
+In this work, scalars are denoted by italic letters ($a, b, A, B, \alpha, \beta$), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
@@
 We define that if the obtained $\#\boldsymbol{A}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\boldsymbol{A}^{(q)} = 0$, then there is no malicious traffic.
-This paper shows how to detect the $\#\boldsymbol{A}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious network traffic.
+This work shows how to detect the $\#\boldsymbol{A}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious network traffic.
 
@@
   	\label{alg:2_alg01}
-	% \begin{algorithmic}[1]
-	% 	\show\LOOP
-	%     \ALOOP {$f = 1$ until $f == \hat{d}$} 
-	%     	    \STATE $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
- %    	    	    \ALOOP {$i = 1$ until $i == Q$} 
-	% 			\IF {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
-	% 			    \STATE $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
-	% 			\ENDIF
- %        		\ENDALOOP
-	%     		\STATE $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
- %   		\ENDALOOP
-	% \end{algorithmic}
+	\begin{algorithmic}[1]
+		\For{$f = 1$ until $f == \hat{d}$} 
+    	    \State $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
+	    	\For{$i = 1$ until $i == Q$} 
+				\If {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
+			    	\State $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
+				\EndIf
+    		\EndFor
+    		\State $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
+    	\EndFor
+	\end{algorithmic}
 \end{algorithm}
 
-After the estimation of the $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked.
+After the estimation of the $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected Attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked.
 To deal with this problem, the adoption of a similarity analysis between legitimate traffic and the traffic of time frames estimated as under attack is evaluated, analysing the effectiveness of cosine similarity to highlight abnormalies inserted by network traffic attacks.
@@
 
-This paper proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$.
+This work proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$.
 Hence, the proposed framework is applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.
@@
 
-This paper models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
+This work models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
 
@@
 
-This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
+This work evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
 Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.

### 2017-11-09 docs/thesis/thesis.tex
some fixes and text for tensor-dl
appendix configuration
introducion...
justificatives and problem statement
chapter 2 ok

@@
 \usepackage{algorithm}% http://ctan.org/pkg/algorithms
+\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
 \usepackage{algpseudocode}
@@
 
-% Algorithmic modifications
-\makeatletter
-\newcommand{\ALOOP}[1]{\ALC@it\algorithmicloop\ #1%
-  \begin{ALC@loop}}
-\newcommand{\ENDALOOP}{\end{ALC@loop}\ALC@it\algorithmicendloop}
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\newcommand{\algorithmicbreak}{\textbf{break}}
-\newcommand{\BREAK}{\STATE \algorithmicbreak}
-\makeatother
-
-% 
 \address{BRASÍLIA}
@@
 
-\title{Discriminative Sensing Based on Signal Processing}
-\date{June/2018}
+\title{Discriminative Sensing Based on Signal Processing for Information Security Analysis}
+\date{November/2017}
 
@@
 \adviser{João Paulo Carvalho Lustosa da Costa}
-\coadviser{?}
+% \coadviser{?}
 
@@
 \include{chapters/1_introduction/introduction}
-\include{chapters/2_pca_svm_cfa/pca_svm_cfa}
-\include{chapters/3_eigen_similarity/eigen_similarity}
-\include{chapters/4_mobile/mobile}
-\include{chapters/5_tensor_dl/tensor_dl}
-\include{chapters/6_conclusion_future_work/conclusion}
+\include{chapters/2_eigen_similarity/eigen_similarity}
+\include{chapters/3_mobile/mobile}
+\include{chapters/4_tensor_dl/tensor_dl}
+\include{chapters/5_conclusion_future_work/conclusion}
+
+\appendix
+\include{chapters/appendix/a_mos/a_mos}
+\include{chapters/appendix/b_pca_svm_cfa/pca_svm_cfa}
 

### 2017-11-11 docs/thesis/chapters/4_tensor_dl/tensor_dl.tex
initing tensor-dl

@@
 
-Compressive sensing is emergent for signal processing and hás been applied to Fields such as signal restoration, image processing and classification (sparce representation classification src )
+% Compressed Sensing | Dictionary Learning | Multidimensional data | Tensor DL and Separable dictionaries
 
-Iterative dictionary learning hás been used as a blind technique for feature extraction, improving classification without require feature selection or principal conponent analysis
+Compressed Sensing (CS) aims to obtain the most relevant information of a dataset, what makes it useful for compression, signal reconstruction, image processing and classification, such as in sparce representation classification (SRC) problems.
 
-Big data analytics require techniques to Deal with multidimensional data, making Sense of strucutre and relationship of many dimensios.
-Pca is commonly adopted for dimensionality reduction and feature extraction.
-Tensor hás been adopted for multidimensional data analysis and can be explored for dictionary learning
-
-Adaptive techniques for high dimensional and high thoughput
-
-we propose to apply the tensor deconposition for dictionary learning in order to evidentiate the discriminative sensing of a fraud detection dataset
-
-Dictionary Learning is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
+Dictionary learning is related to CS but seeks to obtain compact and meaningful signal representations, known as dictionary, from learning algorithms or models.
+DL is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
+DL is based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world.
 The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.
@@
 There are two major approaches for dictionary learning.
-First is the analytic approach, in which DCT bases, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries.
+First is the analytic approach, in which Discrete Cosine Transform basis, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries.
 Second is the learning-based approaches, such as the unsupervised learning for dictionary construction [9] and the online dictionary learning [11], [10], which use machine learning methods to construct the dictionary.
+
 In the analytic approach, some pre-defined functions are used to construct the dictionary.
@@
 Bandelets [26], which specifies the geometry as a vector field, is de-signed to improve the image compression and noise reduction performance.
+
 In the learning-based approach, machine learning methods are used to construct the dictionary from the training data.
-The least square error is used by the method of the optimal directions (MOD) [27] to update the dictionary iteratively:
-
-Dictionaries are either available analytically, or can be learned from a suitable training set.
+The least square error is used by the method of the optimal directions (MOD) [27] to update the dictionary iteratively.
 While analytic dictionaries permit to capture the global structure of a signal and allow a fast implementation, learned dictionaries often perform better in applications as they are more adapted to the considered class of signals.
 
-In some applications the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks.
+In some applications, the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks.
 Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in multidimensional data analytics problems \cite{kolda2009tensor}.
@@
 On the one hand, this permits larger patch-sizes for the learning phase, on the other hand, the dictionary is applied efficiently in reconstruction tasks.
-The crucial idea is to allow the dictionary to have a separable structure, where separable means that the dictionary D is given by the Kronecker product of two smaller dictionaries A ∈ R h×a and B ∈ R w×b
-
-% Without loss of generality we choose square image patches with w = h = 8, which is in accordance to the patch-sizes mostly used in the literature
 
 Multidimensional parameter estimation and learning multidimensional separable dictionaries are growing research problems.
+The crucial idea about separable dictionaries is to allow the dictionary to have a separable structure, where separable means that the dictionary $\textbf{D}$ is given by the Kronecker product of two smaller dictionaries $\textbf{A} \in \mathbb{R}^{h \times a}$ and $\textbf{B} \in \mathbb{R}^{w \times b}$, for example.
 Roemer \emph{et al.} \cite{roemer2014tensor} show that the multidimensional dictionary estimation problem can be efficiently formulated in terms of tensors, and that their results outperform existing schemes by exploiting the multilinear structure of the problem.
 
+Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, this work propose a tensor-based dictionary learning method for fraud detecton from imbalanced data, in order to apply the tensor deconposition for dictionary learning methods to evidentiate the discriminative sensing of a fraud detection dataset.
+
+This chapter is organized as follows.
+Section \ref{sec:4_motivation} presents the motivation for the proposed work.
+In Section \ref{sec:4_relatedworks}, related works are discussed.
+Section \ref{sec:4_datamodel} presents the data model and the evaluated datasets.
+Section \ref{sec:4_proposal} describes the proposed approach for for fraud detecton from mobile money transacion.
+Section \ref{sec:2_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.
+
+\section{Motivation}
+\label{sec:4_motivation}
+
 Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results.
-However, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets, as can be seen in Figures \ref{fig:fig1} and \ref{fig:fig2}, which show that the recovering rating of traditional algorithms decreases over the dataset increasing.
+However, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets.
+
+Considering that DL aims to obtain meaningful dictionaries, it is possible to evaluate DL algorithms by their performance to generate dictionaries that better represent one target data.
+Therefore, we can compare dictionary learning algorithms regarding recovering of a known dictionary.
+
+A true dictionary of size 20x50 is made in each trial.
+A set of trainning data is made by using the true dictionary, each training vector is a sparse combination of some, here 5, vectors from the true dictionary and random noise.
+The number of training vectors can be chosen and the level of noise can be given.
+Then a dictionary is trained from the training set using one of the three methods, ILS-DLA (MOD), K-SVD, and RLS-DLA.
+The trained dictionary is then compared to the true dictionary, this is done by pairing the dictionaries atoms in the best possible way and using the angle between a true atom and a trained atom as a distance measure.
+The number of identified atoms is then found by comparing these angles to a limit, βlim.
+
+Dictionary recovery problem
+
+Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based (K-HOSVD \cite{roemer2014tensor} and T-MOD \cite{roemer2014tensor}) and traditional dictionary learning algorithms (RLS-DLA \cite{skretting2010recursive}, K-SVD \cite{aharon2006rm} and MOD \cite{engan1999method}) for dictionary reconstruction of a multidimensional and separable data.
 
@@
      \centering 
-	 \includegraphics[width=12cm]{figures/5_20_2000_1000_100.eps}
+	 \includegraphics[width=10cm]{figures/5_20_2000_1000_100.eps}
      \caption{Matrix of 1.000 elements, 2000 sample training, 100 iterations}
@@
 
+Figure \ref{fig:fig1} presents the results for dictionary reconstruction from a matrix of 1000 elements by 50 atoms.
+We see that the RLS-DLA method here performs better than ILS-DLA (MOD) and K-SVD which both perform quite similar.
+The most remarkable difference is that RLS-DLA is able to identify all atoms for a quite small angle βlim.
+ILS-DLA (MOD) and K-SVD always have some atoms that are far away from the true atom.
+Having βlim=8.11 degrees corresponds to the case where the inner product of the two compared atoms is equal to 0.99, each atom has 2-norm equal to 1.
+
+Figure \ref{fig:fig2} presents the results from a matrix with 24000 entries by 300 atoms. It is possible to observe better reconstruction rating for tensor-based algorithms
+
 \begin{figure}[!htb]
      \centering 
-	 \includegraphics[width=12cm]{figures/5_20_2000_24000_100.eps}
+	 \includegraphics[width=10cm]{figures/5_20_2000_24000_100.eps}
      \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}
@@
 
-Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based and traditional dictionary learning algorithms for dictionary reconstruction of a multidimensional and separable data.
-Figure \ref{fig:fig1} presents the results for dictionary reconstruction from a matrix of 1000 elements by 50 atoms, while Figures \ref{fig:fig2} presents the results from a matrix with 24000 entries by 300 atoms.
-It is possible to observe better reconstruction rating for tensor-based algorithms.
+Imbalanced data...
 
-Imbalanced data...
 Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate.
@@
 
+OMP \cite{davis1997adaptive}
+
+Specifically, we propose to apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification of mobile money transactions.
+We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
+
 \section{Related Works}
+\label{sec:4_relatedworks}
 
@@
 
+RLS-DLA \cite{skretting2010recursive}, K-SVD \cite{aharon2006rm} and MOD \cite{engan1999method}
+
+T-MOD and K-HOSVD \cite{roemer2014tensor}
+
 Face recognition...
@@
 \section{Data Model}
+\label{sec:4_datamodel}
 
@@
 \section{Proposed Approach}
+\label{sec:4_proposal}
 
@@
 \section{Experiments}
+\label{sec:4_experiments}
 
@@
 \section{Conclusion}
+\label{sec:4_conclusion}
 

### 2017-11-11 docs/thesis/chapters/4_tensor_dl/tensor_dl.tex
tensor dl motivation

@@
 
+% \section{Introduction}
 % Compressed Sensing | Dictionary Learning | Multidimensional data | Tensor DL and Separable dictionaries
@@
 
-There are two major approaches for dictionary learning.
+There are two major approaches for DL.
 First is the analytic approach, in which Discrete Cosine Transform basis, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries.
@@
 Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in multidimensional data analytics problems \cite{kolda2009tensor}.
-Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}.
+Tensor-based algorithms for DL can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}.
 
@@
 Multidimensional parameter estimation and learning multidimensional separable dictionaries are growing research problems.
-The crucial idea about separable dictionaries is to allow the dictionary to have a separable structure, where separable means that the dictionary $\textbf{D}$ is given by the Kronecker product of two smaller dictionaries $\textbf{A} \in \mathbb{R}^{h \times a}$ and $\textbf{B} \in \mathbb{R}^{w \times b}$, for example.
+The crucial idea about separable dictionaries is to allow the dictionary to have a separable structure, where separable means that the dictionary $\textbf{D}$ is given by the Kronecker product of two smaller dictionaries $\boldsymbol{A}^{(1)} \in \mathbb{R}^{h \times a}$ and $\boldsymbol{A}^{(2)} \in \mathbb{R}^{w \times b}$, for example.
 Roemer \emph{et al.} \cite{roemer2014tensor} show that the multidimensional dictionary estimation problem can be efficiently formulated in terms of tensors, and that their results outperform existing schemes by exploiting the multilinear structure of the problem.
 
-Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, this work propose a tensor-based dictionary learning method for fraud detecton from imbalanced data, in order to apply the tensor deconposition for dictionary learning methods to evidentiate the discriminative sensing of a fraud detection dataset.
+Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, this work propose a tensor-based DL for fraud detecton from imbalanced data, in order to apply the tensor deconposition for DL methods to evidentiate the discriminative sensing of a fraud detection dataset.
 
@@
 
-Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results.
+Existing DL schemes can be applied to multidimensional analysis and obtain valuable results.
 However, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets.
@@
 Considering that DL aims to obtain meaningful dictionaries, it is possible to evaluate DL algorithms by their performance to generate dictionaries that better represent one target data.
-Therefore, we can compare dictionary learning algorithms regarding recovering of a known dictionary.
+Therefore, we can compare DL algorithms regarding recovering of a known dictionary.
+We conduct a evaluation of tensor-based DL algorithms for recovering a known dictionary in order to analyze the gains of tensor-based approaches in comparison to well known algorithms.
 
-A true dictionary of size 20x50 is made in each trial.
-A set of trainning data is made by using the true dictionary, each training vector is a sparse combination of some, here 5, vectors from the true dictionary and random noise.
-The number of training vectors can be chosen and the level of noise can be given.
-Then a dictionary is trained from the training set using one of the three methods, ILS-DLA (MOD), K-SVD, and RLS-DLA.
-The trained dictionary is then compared to the true dictionary, this is done by pairing the dictionaries atoms in the best possible way and using the angle between a true atom and a trained atom as a distance measure.
-The number of identified atoms is then found by comparing these angles to a limit, βlim.
+\subsection{Data Model and Experiments}
+\label{sec:4_motivation_datamodel}
 
-Dictionary recovery problem
+Consider a generic sparse recovery problem of the following form
 
-Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based (K-HOSVD \cite{roemer2014tensor} and T-MOD \cite{roemer2014tensor}) and traditional dictionary learning algorithms (RLS-DLA \cite{skretting2010recursive}, K-SVD \cite{aharon2006rm} and MOD \cite{engan1999method}) for dictionary reconstruction of a multidimensional and separable data.
+\begin{equation}\label{eq:eq01}
+\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W},
+\end{equation}
+
+where $\boldsymbol{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\boldsymbol{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary where $N \ll T$, $\boldsymbol{S} \in \mathbb{C}^{N \times T}$ represents the sparse coefficient matrix, $\boldsymbol{W} \in \mathbb{C}^{M \times T}$ is the additive noise, and $M < N < T$.
+
+Consider a sparse recovery problem for a separable 2-D dictionary that we can write as
+
+\begin{equation}\label{eq:eq02}
+\boldsymbol{X} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}) \cdot \boldsymbol{S} + \boldsymbol{W},
+\end{equation}
+
+where $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$, $\boldsymbol{A}^{(1)} \in \mathbb{C}^{M_1 \times N_1}$, $\boldsymbol{A}^{(2)} \in \mathbb{C}^{M_2 \times N_2}$, $M = M_1 \times M_2$, and $N = N_1 \times N_2$.
+
+In order to evaluate tensor-based DL algorithms for recovering a known dictionary, we generate two random dictionaries $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ from an i.i.d.
+zero mean Gaussian random process, and calculate the dictionary $\boldsymbol{A}$ through the Kronecker product, according to $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$.
+Therefore, $\boldsymbol{A}$ is the known dictionary that shall be used for recovery evaluation.
+
+Subsequently, we generate a synthetic data set $\boldsymbol{X}$ using the given dictionary $\boldsymbol{A}$, according to $\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W}$, where $\boldsymbol{S}$ is a random sparse coefficient matrix for what we assume that each column has $K = 5$ non-zero entries, and $\boldsymbol{W}$ is the additive white gaussian noise.
+Thus, we estimate the initial $\boldsymbol{\hat{A}}$ dictionary from a normalized subset of $\boldsymbol{X}$ and make the dictionary separable decomposing the approximation of $\boldsymbol{\hat{A}}$ and generating $\boldsymbol{\hat{A}}^{(1)}$ and $\boldsymbol{\hat{A}}^{(2)}$ according to \cite{van1993approximation}.
+Afterward, we update $\boldsymbol{\hat{A}}$ through $\boldsymbol{\hat{A}} = (\boldsymbol{\hat{A}}^{(1)} \otimes \boldsymbol{\hat{A}}^{(2)})$ and use it for DL experiments using MOD \cite{engan1999method}, K-SVD \cite{aharon2006rm}, RLS-DLA \cite{skretting2010recursive}, T-MOD \cite{roemer2014tensor} and K-HOSVD \cite{roemer2014tensor}.
+Due to its low computational complexity, we employ the Orthogonal Matching Pursuit (OMP) algorithm\cite{davis1997adaptive} for the sparse recovery stage in all the dictionary learning algorithms, assuming that K is known.
+
+The final estimated dictionaries $\boldsymbol{\hat{A}}$ recovered by each DL algorithm are compared to the known dictionary $\boldsymbol{A}$, through a measure of the distance between two dictionaries.
+The recovered dictionary is then compared to the known dictionary, by using the angle between a known atom and a recovered atom as a distance measure.
+The number of identified atoms is computed by comparing these angles to a limit of $\beta_{lim} = 8.11$ degrees, that corresponds to the case where the inner product of the two compared atoms is equal to 0.99, and where each atom has 2-norm equal to 1.
+
+For this experiments we adopt an additive noise with $snr = 20$ and 100 iterations for dictionary learning. All experiments run 100 trials in order to obtain confident results.
+
+This evaluation extends Roemer \emph{et al.} \cite{roemer2014tensor} including RLS-DLA for comparison, and for adding the evaluation of the degree distribution for atom identification and the evaluation of the atom identification according to the dictionary size variation.
+Additionally, we conduct a experimental evaluation, while Roemer \emph{et al.} \cite{roemer2014tensor} presents numerical results obtained via Monte Carlo simulations.
+
+\subsection{Results}
+\label{sec:4_motivation_results}
+
+Figures \ref{fig:fig1} and \ref{fig:fig2} show the evaluation of tensor-based (K-HOSVD \cite{roemer2014tensor} and T-MOD \cite{roemer2014tensor}) and traditional DL algorithms (RLS-DLA, K-SVD and MOD) for dictionary reconstruction of a multidimensional and separable data.
+These figures show the distribution of the cumulative atom identification over the degree rates.
+
+Figure \ref{fig:fig1} presents the results for recovering a dictionary of $M=20$ and $N=50$.
+We see that the RLS-DLA method performs better than other methods, although T-MOD presents the best results for identification of up to 15 atoms of 50.
+It is possible to observe that T-MOD and K-HOSVD are better than MOD and K-SVD, and that the tensor-based methods outperforms MOD and K-SVD significantly for this experiment.
+Finally, Figure \ref{fig:fig1} shows that RLS-DLA recovered almost 40 atoms in $\beta_{lim} = 8.11$ degrees, while other methods present results lower results.
 
@@
 	 \includegraphics[width=10cm]{figures/5_20_2000_1000_100.eps}
-     \caption{Matrix of 1.000 elements, 2000 sample training, 100 iterations}
+     \caption{Cumulative atom identificatin per degree rates. $T=2000$, $M=20$, $N=50$}
      \label{fig:fig1}
@@
 
-Figure \ref{fig:fig1} presents the results for dictionary reconstruction from a matrix of 1000 elements by 50 atoms.
-We see that the RLS-DLA method here performs better than ILS-DLA (MOD) and K-SVD which both perform quite similar.
-The most remarkable difference is that RLS-DLA is able to identify all atoms for a quite small angle βlim.
-ILS-DLA (MOD) and K-SVD always have some atoms that are far away from the true atom.
-Having βlim=8.11 degrees corresponds to the case where the inner product of the two compared atoms is equal to 0.99, each atom has 2-norm equal to 1.
-
-Figure \ref{fig:fig2} presents the results from a matrix with 24000 entries by 300 atoms. It is possible to observe better reconstruction rating for tensor-based algorithms
+Figure \ref{fig:fig2} presents the results for recovering a dictionary of $M=80$ and $N=30$.
+It is possible to observe that the tensor-base methods outperform MOD, RLS-DLA and K-SVD in number of identified atoms and the required degrees for identification.
+Note that MOD performs quite silimar to RLS-DLA, while K-SVD is the worst evaluated case for this experiment.
+Considering the tensor-based methods, T-MOD outperforms K-HOSVD with higher atom identification over lower degrees.
+The experimental results verify that the tensor-based dictionary learning algorithms are able to identify more dictionary atoms when dealing with larger dictionaries and that they converge faster.
 
@@
 	 \includegraphics[width=10cm]{figures/5_20_2000_24000_100.eps}
-     \caption{Matrix of 24.000 elements, 2000 sample training, 100 iterations}
+     \caption{Cumulative atom identificatin per degree rates. $T=2000$, $M=80$, $N=30$}
      \label{fig:fig2}
@@
 
-Imbalanced data...
+The results reveals that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases.
+Since dictionary learning and reconstruction error are usually used for sparce representation classification (SRC) problems, it is possible to supose that tensor-based algorithms can improve results of classification problems, as well as can extract hidden patterns and structure in multidimensional data analytics problems.
 
@@
 
-A key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset.
-Therefore, we introduce a tensor-based dictionary learning method for fraud detecton from unbalanced data.
+A key challenge to use SC and DL for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset.
+Therefore, we introduce a tensor-based DL method for fraud detecton from unbalanced data.
 Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification.
 In this paper, we propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds.
-We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
-
-%However, tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.
-We propose a distributed and tensor-based approach for multidimensional dictionary learning in order to obtain better processing time for larger datasets.
-We focus on a distributed implementation of T-MOD \cite{roemer2014tensor} algorithm, based on a modification of Almeida and Kibangou's \cite{de2014distributed} approach to use Tucker-2 decomposition.
-We perform experiments to test different methods for learning dictionaries, evaluating the processing time and the reconstruction of a dictionary of multidimensional and separable data.
-
-OMP \cite{davis1997adaptive}
-
-Specifically, we propose to apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification of mobile money transactions.
-We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
+We propose to use tensor-based DL for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
 
@@
 
-Consider a generic sparse recovery problem of the following form
-
-\begin{equation}\label{eq:eq01}
-\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W},
-\end{equation}
-
-where $\boldsymbol{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\boldsymbol{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary, $\boldsymbol{S} \in \mathbb{•}hbb{C}^{N \times T}$ represents the sparse coefficient matrix, $\boldsymbol{W} \in \mathbb{C}^{M \times T}$ is the additive noise, and $M < N < T$.
-
-Consider a sparse recovery problem for a separable 2-D manifold, i.e., a manifold that can be written as the product of two 1-D manifolds.
-If we choose a separable 2-D sampling grid, we can write the 2D sparse recovery problem as
-
 Imbalance data where zero is predominante
@@
 Data divided into learning and test, with isfraud labels with 1 or 0
-
-\begin{equation}\label{eq:eq02}
-\boldsymbol{X} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}) \cdot \boldsymbol{S} + \boldsymbol{W},
-\end{equation}
-
-where $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$, $\boldsymbol{A}^{(1)} \in \mathbb{C}^{M_1 \times N_1}$, $\boldsymbol{A}^{(2)} \in \mathbb{C}^{M_2 \times N_2}$, $M = M_1 \times M_2$, and $N = N_1 \times N_2$.
 
@@
 Our proposal is better for imbalanced data than others techniques
+
+%However, tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.
+We propose a distributed and tensor-based approach for multidimensional dictionary learning in order to obtain better processing time for larger datasets.
+We focus on a distributed implementation of T-MOD \cite{roemer2014tensor} algorithm, based on a modification of Almeida and Kibangou's \cite{de2014distributed} approach to use Tucker-2 decomposition.
+We perform experiments to test different methods for learning dictionaries, evaluating the processing time and the reconstruction of a dictionary of multidimensional and separable data.

### 2017-11-12 docs/thesis/chapters/1_introduction/introduction.tex
tensor-related works

@@
 Scientists and researchers must today spend time and effort in obtaining access to relevant data sets before they can research on such data set.
-However, PaySim is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.\footnote{https://www.kaggle.com/ntnu-testimon/paysim1}
+However, PaySim \cite{lopez2016paysim} is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.
 

### 2017-11-12 docs/thesis/chapters/4_tensor_dl/tensor_dl.tex
tensor-related works

@@
 DL is based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world.
-The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.
+The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification, when the learning includes a class separability criteria in the objective function \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.
 
@@
 
-\begin{equation}\label{eq:eq01}
+\begin{equation}\label{eq:4_eq01}
 	\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W},
@@
 
-\begin{equation}\label{eq:eq02}
+\begin{equation}\label{eq:4_eq02}
 	\boldsymbol{X} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}) \cdot \boldsymbol{S} + \boldsymbol{W},
@@
 In order to evaluate tensor-based DL algorithms for recovering a known dictionary, we generate two random dictionaries $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ from an i.i.d.
-zero mean Gaussian random process, and calculate the dictionary $\boldsymbol{A}$ through the Kronecker product, according to $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$.
+zero mean Gaussian random process, and calculate the dictionary $\boldsymbol{A}$ through the Kronecker product, according to
+
+\begin{equation}\label{eq:4_eq03}
+	\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}).
+\end{equation}
+
 Therefore, $\boldsymbol{A}$ is the known dictionary that shall be used for recovery evaluation.
 
-Subsequently, we generate a synthetic data set $\boldsymbol{X}$ using the given dictionary $\boldsymbol{A}$, according to $\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W}$, where $\boldsymbol{S}$ is a random sparse coefficient matrix for what we assume that each column has $K = 5$ non-zero entries, and $\boldsymbol{W}$ is the additive white gaussian noise.
+Subsequently, we generate a synthetic data set $\boldsymbol{X}$ using the given dictionary $\boldsymbol{A}$, according to Equation \ref{eq:4_eq01}, where $\boldsymbol{S}$ is a random sparse coefficient matrix for what we assume that each column has $K = 5$ non-zero entries, and $\boldsymbol{W}$ is the additive white gaussian noise.
 Thus, we estimate the initial $\boldsymbol{\hat{A}}$ dictionary from a normalized subset of $\boldsymbol{X}$ and make the dictionary separable decomposing the approximation of $\boldsymbol{\hat{A}}$ and generating $\boldsymbol{\hat{A}}^{(1)}$ and $\boldsymbol{\hat{A}}^{(2)}$ according to \cite{van1993approximation}.
-Afterward, we update $\boldsymbol{\hat{A}}$ through $\boldsymbol{\hat{A}} = (\boldsymbol{\hat{A}}^{(1)} \otimes \boldsymbol{\hat{A}}^{(2)})$ and use it for DL experiments using MOD \cite{engan1999method}, K-SVD \cite{aharon2006rm}, RLS-DLA \cite{skretting2010recursive}, T-MOD \cite{roemer2014tensor} and K-HOSVD \cite{roemer2014tensor}.
+Afterward, we update $\boldsymbol{\hat{A}}$ through
+
+\begin{equation}\label{eq:4_eq04}
+	\boldsymbol{\hat{A}} = (\boldsymbol{\hat{A}}^{(1)} \otimes \boldsymbol{\hat{A}}^{(2)}),
+\end{equation}
+
+and use it for DL experiments using MOD \cite{engan1999method}, K-SVD \cite{aharon2006rm}, RLS-DLA \cite{skretting2010recursive}, T-MOD \cite{roemer2014tensor} and K-HOSVD \cite{roemer2014tensor}.
 Due to its low computational complexity, we employ the Orthogonal Matching Pursuit (OMP) algorithm\cite{davis1997adaptive} for the sparse recovery stage in all the dictionary learning algorithms, assuming that K is known.
+However, other solver for sparse recovery can be used, such as Basis Pursuit \cite{chen2001atomic}, the LASSO algorithm \cite{tibshirani1996regression} or greedy \cite{davis1997adaptive}.
 
@@
 Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification.
-In this paper, we propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds.
+
+In this chapter, we propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds.
 We propose to use tensor-based DL for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
@@
 
+% Compressed Sensing | Dictionary Learning | Classification | Tensor-Based DL
+
 Compressed sensing (also known as compressive sensing, compressive sampling, or sparse sampling) is a signal processing technique for efficiently acquiring and reconstructing a signal, by finding solutions to underdetermined linear systems.
 This is based on the principle that, through optimization, the sparsity of a signal can be exploited to recover it from far fewer samples than required by the Shannon-Nyquist sampling theorem.
-There are two conditions under which recovery is possible.[1] The first one is sparsity which requires the signal to be sparse in some domain.
-The second one is incoherence which is applied through the isometric property which is sufficient for sparse signals.[2][3]
 
-A common goal of the engineering field of signal processing is to reconstruct a signal from a series of sampling measurements.
-In general, this task is impossible because there is no way to reconstruct a signal during the times that the signal is not measured.
-
+%There are two conditions under which recovery is possible.[1] The first one is sparsity which requires the signal to be sparse in some domain.
+The second one is incoherence which is applied through the isometric property which is sufficient for sparse signals.[2][3].
 Around 2004, Emmanuel Candès, Terence Tao, and David Donoho proved that given knowledge about a signal's sparsity, the signal may be reconstructed with even fewer samples than the sampling theorem requires.[4][5] This idea is the basis of compressed sensing.
@@
 
-Sparse dictionary learning is a representation learning method which aims at finding a sparse representation of the input data (also known as sparse coding) in the form of a linear combination of basic elements as well as those basic elements themselves.
+Sparse coding is a sparse representation of one data in the form of a linear combination of basic elements.
+These elements are called atoms and they compose a dictionary.
+Dictionary learning is a representation learning method which aims at finding a sparse representation of the input data (also known as sparse coding) in the form of a linear combination of basic elements as well as those basic elements themselves.
 These elements are called atoms and they compose a dictionary.
@@
 The above two properties lead to having seemingly redundant atoms that allow multiple representations of the same signal but also provide an improvement in sparsity and flexibility of the representation.
-
-sparse coding is a sparse representation of one data in the form of a linear combination of basic elements. These elements are called atoms and they compose a dictionary.
 
@@
 
-RLS-DLA \cite{skretting2010recursive}, K-SVD \cite{aharon2006rm} and MOD \cite{engan1999method}
+MOD \cite{engan1999method} is a DL algorithm that is commonly used for extensions of Least-Squares (LS) type schemes, also known as iterative least squares dictionary learning algorithm (ILS-DLA) \cite{engan2007family}.
+MOD defines that if we know $S$ in equation \ref{eq:4_eq01}, $A$ can be estimated solving a linear LS problem.
+Additionally, knowing $A$ we obtain $S$ by solving a sparse recovery problem.
+MOD iterates between the two steps, starting with an initial guess of $A$, denoted as $\hat{A}$, which can be estimated, generated by previous dictionary learning or generated analitically.
+Despite - simplicity, the MOD algorithm has been criticized for its slow convergence, which is due to the fact that $A$ and $S$ are updated separately, which may lead to a slow convergence \cite{aharon2006rm}.
 
-T-MOD and K-HOSVD \cite{roemer2014tensor}
+K-SVD \cite{aharon2006rm} is a generalization of the K-means algorithm for dictionary learning.
+After the sparse approximation step, the dictionary update is performed by sequentially updating each column of $A$ using SVD to minimize the approximation error.
+The update step is hence a generalized K-means algorithm since each patch can be represented by multiple atoms and with different weights.
+In practice, dictionaries learned with K-SVD
+have shown excellent performance in image denoising.
 
-Face recognition...
+Most DL algorithms update the dictionary after a batch of training vectors has been processed, usually using the whole set of training vectors as one batch.
+The training set is used iteratively to gradually improve the dictionary.
+The approach in RLS-DLA \cite{skretting2010recursive} is a continuous update of the dictionary as each training vector is being processed.
+The algorithm is based on recursive least squares (RLS) algorithm for adaptive filtering, what makes the algorithm less dependent on the initial dictionary and it improves both convergence properties of RLS-DLA as well as the representation ability of the resulting dictionary.
 
-Fisher...
+Dimensionality reduction can first be achieved by selecting a subset of functions from a large, fixed dictionary that is used for the analysis of particular signals.
+These functions then determine a subspace of reduced dimension, where classification can be performed by computing the nearest neighbor points among the projected data.
+A simple method to build such a subspace consists in modifying the sparse approximation methods such that the objective function is augmented with a discrimination term that represents the separability properties of the projection subspace.
+This kind of approach typically tries to maximize the variance between the active atoms from $A$ that represent signals in different classes.
+The reduced subspace used for classification is finally formed by the subset of atoms in $A$ whose corresponding coefficients in $S$ are nonzero.
+The subset selection problem can be interpreted as the inference step in the dictionary learning methods when the objective function is modified to include a discriminative term \cite{tosic2011dictionary}.
 
-Florian...
+An important advantage of redundant dictionaries for classification is that signal analysis can be performed with functions that are likely to match the data characteristics in different classes of signals.
+Similarly to data approximation problems, data analysis applications can further benefit from dictionary learning methods.
+The previous section describes subspace selection methods from predefined dictionaries.
+DL algorithms can improve the classification performance, as it leads to a better adaptation of the dictionary by enforcing sparsity in the representation of data in the different classes.
+The atoms in a dictionary $A$ that is computed with dictionary learning methods generally capture the most important constitutive components of the signals.
+They naturally permit to classify the data into the corresponding linear subspace.
+However, there is no guarantee that the subspace built on a learned dictionary $A$ is truly optimal for classification, as it targets efficient representation but not necessarily class separability \cite{tosic2011dictionary}.
 
-Tem que ver outro...
+Dictionary learning methods should rather be modified so that they become simultaneously reconstructive and discriminative.
+The addition of a discriminative term into the dictionary learning algorithms requires supervision, where labels of training data are used to ensure that the data representation is sufficiently different in each class.
+It can be achieved by modifying the sparse coding step in the learning algorithms, so that it optimizes an objective function that favors the sparsest representation of a given signal and simultaneously the representation that is also the most different from the one of signals in other data classes.
+The supervised dictionary learning problem can be cast as a mixed formulation that minimizes the average value of the sparse approximation errors over different classes and also enforces discrimination between classes \cite{tosic2011dictionary}.
+
+The use of one learned dictionary for all the data classes leads to a straightforward classification stage where the dictionary vectors and the coefficients in the signal representation are used directly to make classification decisions.
+Alternatively, one may want to improve the discrimination by building a distinctive projection subspace for each data class.
+Classification is then performed by selecting the subspace that is the nearest to the test signal, or equivalently the subspace that leads to the best representation of the test signal.
+A simple way to build adaptive dictionaries for each class is to use the signals in the training set for the class dictionary.
+Sparsity constraints are then rather applied within the classification process, where the sparsest representation of the test signal determines its class label \cite{tosic2011dictionary}.
+
+Wright \emph{et al.} \cite{wright2009robust} have proposed a face recognition method that uses training face images as dictionaries and an $l_1$ sparse optimization method in the classification stage.
+The authors show that the recognition task can be successfully accomplished even using random features at first.
+Furthermore, the algorithm is robust to a certain amount of noise due to the sparsity constraints \cite{tosic2011dictionary}.
+
+It is often preferable, however, to construct adapted dictionaries that can lead to an efficient classification process based on simple subspace projections.
+The construction of class dictionaries can be performed with learning methods where the sparse coding step in the iterative learning algorithms is modified, so that sparse coding is computed independently within each class.
+Such a sparse coding stage can be implemented by class-supervised versions of simultaneous pursuit algorithms \cite{tosic2011dictionary}.
+
+Tensor data are typically highly structured, a perfect match for compressive sampling, so that the CS framework relaxes data acquisition requirements, enables compact storage, and facilitates data completion (i.e., inpainting of missing samples due to a faulty sensor or unreliable measurement) \cite{cichocki2015tensor}.
+Roemer \emph{et al.} \cite{roemer2014tensor} have proposed a tensor-based multidimensional extensions for MOD and K-SVD algorithms, which are respectivelly T-MOD and K-HOSVD, and show their improved performances numerically, for the dictionary reconstruction problem.
+
+Considering that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases, accordint to subsection \ref{sec:4_motivation_results}, we believe that tensor-based DL algorithms can improve the discriminative sensing for classification classification problems.
+Therefore, we propose use tensor-based DL for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics
 
@@
 \label{sec:4_datamodel}
+
+\subsection{Dataset}
+\label{sec:4_dataset}
+
+We adopt a synthetic financial datasets for fraud detection\footnote{https://www.kaggle.com/ntnu-testimon/paysim1} in order to validate our proposal, this dataset is synthetically generated by  PaySim \cite{lopez2016paysim} , which is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.
+PaySim uses aggregated data from the private dataset to generate a synthetic dataset that resembles the normal operation of transactions and injects malicious behaviour to later evaluate the performance of fraud detection methods.
+PaySim simulates mobile money transactions based on a sample of real transactions extracted from one month of financial logs from a mobile money service implemented in an African country.
+The original logs were provided by a multinational company
+
+The selected dataset is composed by 9 features (variables) and 6.362.620 tramsactions (observations), it is also provided one lable that classifies each transaction as fraudulent or not.
+As the majority dataset of anomaly or fraud detection, the data is imbalanced, where the number of fraudlent transactions is ZZ against of NN normal transactions.
 
@@
 
-The Kronecker model (\ref{eq:eq02}) can be rewritten in an equivalent tensor form.
-Applying the algebraic rules for unfoldings of n-mode products \cite{roemer2014tensor} we can rewrite (\ref{eq:eq02}) into a "Tucker-2" decomposition, as
+\subsection{Tensor-based Model}
+\label{sec:4_tensor_model}
 
-\begin{equation}\label{eq:eq03}
+The Kronecker model (\ref{eq:4_eq02}) can be rewritten in an equivalent tensor form.
+Applying the algebraic rules for unfoldings of n-mode products \cite{roemer2014tensor} we can rewrite (\ref{eq:4_eq02}) into a "Tucker-2" decomposition, as
+
+\begin{equation}\label{eq:4_eq04}
 \boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} + \boldsymbol{\mathcal{W}},
@@
 
-%However, tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.
-We propose a distributed and tensor-based approach for multidimensional dictionary learning in order to obtain better processing time for larger datasets.
-We focus on a distributed implementation of T-MOD \cite{roemer2014tensor} algorithm, based on a modification of Almeida and Kibangou's \cite{de2014distributed} approach to use Tucker-2 decomposition.
-We perform experiments to test different methods for learning dictionaries, evaluating the processing time and the reconstruction of a dictionary of multidimensional and separable data.
+Tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.
+We propose as future work a distributed and tensor-based approach for multidimensional dictionary learning in order to obtain better processing time for larger datasets, based on a modification of Almeida and Kibangou's \cite{de2014distributed} approach to use Tucker-2 decomposition.

### 2017-11-14 docs/thesis/chapters/4_tensor_dl/tensor_dl.tex
improving the approach and experiment description

@@
 
-Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, this work propose a tensor-based DL for fraud detecton from imbalanced data, in order to apply the tensor deconposition for DL methods to evidentiate the discriminative sensing of a fraud detection dataset.
+Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, this work propose a tensor-based DL for fraud detection from imbalanced data, in order to apply the tensor deconposition for DL methods to evidentiate the discriminative sensing of a fraud detection dataset.
 
 This chapter is organized as follows.
-Section \ref{sec:4_motivation} presents the motivation for the proposed work.
+Section \ref{sec:4_motivation} presents one experiment and results that motivate the proposed work.
 In Section \ref{sec:4_relatedworks}, related works are discussed.
@@
 MOD \cite{engan1999method} is a DL algorithm that is commonly used for extensions of Least-Squares (LS) type schemes, also known as iterative least squares dictionary learning algorithm (ILS-DLA) \cite{engan2007family}.
-MOD defines that if we know $S$ in equation \ref{eq:4_eq01}, $A$ can be estimated solving a linear LS problem.
-Additionally, knowing $A$ we obtain $S$ by solving a sparse recovery problem.
-MOD iterates between the two steps, starting with an initial guess of $A$, denoted as $\hat{A}$, which can be estimated, generated by previous dictionary learning or generated analitically.
-Despite - simplicity, the MOD algorithm has been criticized for its slow convergence, which is due to the fact that $A$ and $S$ are updated separately, which may lead to a slow convergence \cite{aharon2006rm}.
+MOD defines that if we know $\boldsymbol{S}$ in equation \ref{eq:4_eq01}, $\boldsymbol{A}$ can be estimated solving a linear LS problem.
+Additionally, knowing $\boldsymbol{A}$ we obtain $\boldsymbol{S}$ by solving a sparse recovery problem.
+MOD iterates between the two steps, starting with an initial guess of $\boldsymbol{A}$, denoted as $\hat{\boldsymbol{A}}$, which can be estimated, generated by previous dictionary learning or generated analitically.
+Despite - simplicity, the MOD algorithm has been criticized for its slow convergence, which is due to the fact that $\boldsymbol{A}$ and $\boldsymbol{S}$ are updated separately, which may lead to a slow convergence \cite{aharon2006rm}.
 
 K-SVD \cite{aharon2006rm} is a generalization of the K-means algorithm for dictionary learning.
-After the sparse approximation step, the dictionary update is performed by sequentially updating each column of $A$ using SVD to minimize the approximation error.
+After the sparse approximation step, the dictionary update is performed by sequentially updating each column of $\boldsymbol{A}$ using SVD to minimize the approximation error.
 The update step is hence a generalized K-means algorithm since each patch can be represented by multiple atoms and with different weights.
@@
 A simple method to build such a subspace consists in modifying the sparse approximation methods such that the objective function is augmented with a discrimination term that represents the separability properties of the projection subspace.
-This kind of approach typically tries to maximize the variance between the active atoms from $A$ that represent signals in different classes.
-The reduced subspace used for classification is finally formed by the subset of atoms in $A$ whose corresponding coefficients in $S$ are nonzero.
+This kind of approach typically tries to maximize the variance between the active atoms from $\boldsymbol{A}$ that represent signals in different classes.
+The reduced subspace used for classification is finally formed by the subset of atoms in $\boldsymbol{A}$ whose corresponding coefficients in $\boldsymbol{S}$ are nonzero.
 The subset selection problem can be interpreted as the inference step in the dictionary learning methods when the objective function is modified to include a discriminative term \cite{tosic2011dictionary}.
@@
 DL algorithms can improve the classification performance, as it leads to a better adaptation of the dictionary by enforcing sparsity in the representation of data in the different classes.
-The atoms in a dictionary $A$ that is computed with dictionary learning methods generally capture the most important constitutive components of the signals.
+The atoms in a dictionary $\boldsymbol{A}$ that is computed with dictionary learning methods generally capture the most important constitutive components of the signals.
 They naturally permit to classify the data into the corresponding linear subspace.
-However, there is no guarantee that the subspace built on a learned dictionary $A$ is truly optimal for classification, as it targets efficient representation but not necessarily class separability \cite{tosic2011dictionary}.
+However, there is no guarantee that the subspace built on a learned dictionary $\boldsymbol{A}$ is truly optimal for classification, as it targets efficient representation but not necessarily class separability \cite{tosic2011dictionary}.
 
@@
 
-\subsection{Dataset}
-\label{sec:4_dataset}
+% \subsection{Dataset}
+% \label{sec:4_dataset}
 
-We adopt a synthetic financial datasets for fraud detection\footnote{https://www.kaggle.com/ntnu-testimon/paysim1} in order to validate our proposal, this dataset is synthetically generated by  PaySim \cite{lopez2016paysim} , which is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.
+We adopt a synthetic financial datasets for fraud detection\footnote{https://www.kaggle.com/ntnu-testimon/paysim1} in order to validate our proposal.
+This dataset is synthetically generated by  PaySim \cite{lopez2016paysim}, which is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.
 PaySim uses aggregated data from the private dataset to generate a synthetic dataset that resembles the normal operation of transactions and injects malicious behaviour to later evaluate the performance of fraud detection methods.
@@
 
-The selected dataset is composed by 9 features (variables) and 6.362.620 tramsactions (observations), it is also provided one lable that classifies each transaction as fraudulent or not.
-As the majority dataset of anomaly or fraud detection, the data is imbalanced, where the number of fraudlent transactions is ZZ against of NN normal transactions.
+The selected dataset is composed by 9 features (variables) and 2.770.408 transactions (observations).
+It is also provided one lable that classifies each transaction as fraudulent or not.
+As the majority dataset of anomaly or fraud detection, the data is imbalanced, where the number of fraudulent transactions is 8.213 against 2.762.196 of normal transactions.
+Therefore, the fraud class represents 0,30 \% of the entire datased.
 
-Imbalance data where zero is predominante
+For evaluation of the proposed classification approach we adopt a data splitting into training and test data, where the training data has 70 \% of the whole dataset.
+We denote the traning data as $X_r$ and the test data as $X_e$.
+The dataset can also be divided by normal and fraudulent transactions, where the normal train data is denoted as $X_r^o$, the fraud train data is denoted as $X_r^f$.
+The same notation of fraud, normal, train and test is applied to qualify $A$ and $S$.
 
-Data divided into learning and test, with isfraud labels with 1 or 0
+% \subsection{Tensor-based Model}
+% \label{sec:4_tensor_model}
 
-\subsection{Tensor-based Model}
-\label{sec:4_tensor_model}
-
-The Kronecker model (\ref{eq:4_eq02}) can be rewritten in an equivalent tensor form.
+% The Kronecker model (\ref{eq:4_eq02}) can be rewritten in an equivalent tensor form.
 Applying the algebraic rules for unfoldings of n-mode products \cite{roemer2014tensor} we can rewrite (\ref{eq:4_eq02}) into a "Tucker-2" decomposition, as
 
-\begin{equation}\label{eq:4_eq04}
-\boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} + \boldsymbol{\mathcal{W}},
-\end{equation}
+% \begin{equation}\label{eq:4_eq04}
+% \boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} + \boldsymbol{\mathcal{W}},
+% \end{equation}
 
-where $\boldsymbol{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\boldsymbol{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\boldsymbol{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\boldsymbol{X}$, $\boldsymbol{S}$, and $\boldsymbol{W}$ such that $\boldsymbol{X} = [\boldsymbol{\mathcal{X}}]_{(3)}^T$, $\boldsymbol{S} = [\boldsymbol{\mathcal{S}}]_{(3)}^T$, and $\boldsymbol{W} = [\boldsymbol{\mathcal{W}}]_{(3)}^T$, respectively.
+% where $\boldsymbol{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\boldsymbol{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\boldsymbol{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\boldsymbol{X}$, $\boldsymbol{S}$, and $\boldsymbol{W}$ such that $\boldsymbol{X} = [\boldsymbol{\mathcal{X}}]_{(3)}^T$, $\boldsymbol{S} = [\boldsymbol{\mathcal{S}}]_{(3)}^T$, and $\boldsymbol{W} = [\boldsymbol{\mathcal{W}}]_{(3)}^T$, respectively.
 
-We consider a randomly generated data in order to evaluate the proposed approach, where initially $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\boldsymbol{A}$ is obtained.
+% We consider a randomly generated data in order to evaluate the proposed approach, where initially $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\boldsymbol{A}$ is obtained.
 In order to be able to evaluate the dictionary reconstruction, the sparse data $\boldsymbol{\mathcal{X}}$ is generated from the dictionary $\boldsymbol{A}$, which is used for further comparisons.
@@
 
-The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\boldsymbol{A}$ from the sparse data $\boldsymbol{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.
+% The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\boldsymbol{A}$ from the sparse data $\boldsymbol{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.
 
-\section{Proposed Approach}
+\section{Tensor-based Sparce Representation Classification}
 \label{sec:4_proposal}
 
-...
+We propose a Tensor-based Sparce Representation Classification approach to use tensor-based DL for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
+
+Our proposal requires supervision, where labels of training data are used to classify each transaction as fraud or normal.
+The first step is to split the training dataset into classes, as fraud or normal, and create respectively the data matrices $\boldsymbol{X}_r^f$ and $\boldsymbol{X}_r^o$.
+For each data matrix we compute the data normalization and subsequently the initial dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$.
+Afterward, we apply the dictionaries and the data matrices to the selected DL algorithms, in order to learn the best dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$ that respectivelly represents fraud and normal classes as basis vectors.
+
+With the trainned dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$ and the test data $\boldsymbol{X}_e$ it is possible to process each transacion of $\boldsymbol{X}_e$, denoted as $\boldsymbol{x}_e^t$ where $t$ represents the transaction index, in order to compute its sparse coding from fraud and normal dictionaries, generating respectively $\boldsymbol{S}_e^f$ and $\boldsymbol{S}_e^o$.
+Subsequently, we compute the signal rescontruction as $\boldsymbol{x}_e^f = \boldsymbol{A}_r^f \cdot \boldsymbol{S}_e^f$ and $\boldsymbol{x}_e^o = \boldsymbol{A}_r^o \cdot \boldsymbol{S}_e^o$.
+
+Finally, we compare $\boldsymbol{x}_e^f$ and $\boldsymbol{x}_e^o$ to $\boldsymbol{x}_e^t$ in order to evaluate the reconstruction error to classify the transaction as fraud or normal according to the minimum reconstruction error from $\boldsymbol{x}_e^f$ and $\boldsymbol{x}_e^o$, respectivelly.
+We adopt the cosine similarity to compare the reconstructed signal to the original one.
 
@@
 
-rocauc is usually used for classification evaluation, but It os not the best option for imbalance data, and Also because It does not consider the unclassified cases, such as false negatives
+For data analysis of imbalanced classes it is usually applied a data undersample method, where an under sized sample is selected from the original datased, with similar amount of occurrences for each class of the data.
+This method is efficient for algorithm trainning, but it faces generalization problems when tries to classify new data.
+Therefore, we propose to evaluate the our proposal for under sampled data and for the complete dataset.
 
-Receiver Operating Characteristic Curve (ROC curve) http://www.chioka.in/differences-between-roc-auc-and-pr-auc/
+The Receiver Operating Characteristics (ROC) curve makes use of the proportion of two metrics: true positives rate (TPR) and false positives rate (FPR), which are defined as:
 
-A ROC curve is plotting True Positive Rate (TPR) against False Positive Rate (FPR).
+\begin{equation}\label{eq:4_eq04}
+	TPR = \frac{TP}{Pc},
+\end{equation}
 
-TPR is defined as:
+\begin{equation}\label{eq:4_eq05}
+	FPR = \frac{FP}{Nc},
+\end{equation}
 
-tpr-formula
+where $TP$ denotes true positive, $Pc$ is the total positive values, $FP$ denotes the false positive and $Nc$ the total negative values.
 
-FPR is defined as:
+ROC curves are commonly used to present results for binary decision problems in machine learning.
+However, when dealing with highly skewed datasets, Precision-Recall (PR) curves give a more informative picture of an algorithm’s performance \cite{davis2006relationship, he2009learning}.
+The PR curve is defined by plotting Precision Rate (PR) over the Recall Rate (RR), which are defined as:
 
-fpr-formula
+\begin{equation}\label{eq:4_eq06}
+	PR = \frac{TP}{TP + FP},
+\end{equation}
 
-where TP = true positive, TN = true negative, FP = false positive, FN = false negative.
+\begin{equation}\label{eq:4_eq07}
+	RR = \frac{TP}{TP + FN}.
+\end{equation}
 
-A typical ROC curve looks like this, which shows two ROC curves for Algorithm 1 and Algorithm 2.
+PR curves exhibit a strong correspondence to ROC curves: A curve dominates in ROC space if and only if it dominates in PR space.
+However, an algorithm that optimizes the AUC in the ROC space is not guaranteed to optimize the AUC in PR space.
+Considering that our data is very imbalanced, with many more negatives than positives values, we adopt the averate precision recall for the classification evaluation
 
-sample-ROC-curve
+Logistic regression is the most common method used to model binary response data \cite{hilbe2011logistic}, it is useful for situations in where it is necessary to predict the presence or absence of a characteristic, with binary classification, such as in anomaly or fraud detection \cite{bhattacharyya2011data}.
+Therefore, we compare the results of out proposed approach with the logistic regression algorithm applied for fraud detection from the adopted datase.
 
-The goal is to have a model be at the upper left corner, which is basically getting no false positives – a perfect classifier.
+\section{Results}
+\label{sec:4_results}
 
-The receiver operating characteristic area under curve (ROC AUC) is just the area under the ROC curve. The higher it is, the better the model is.
+% \section{Discussion}
+% \label{sec:4_Discussion}
 
-Precision Recall Curve (PR Curve)
-A PR curve is plotting Precision against Recall.
-
-Precision is defined as:
-
-precision-formula
-
-Recall is defined as:
-
-recall-formulaA typical PR curve looks like this, which shows two PR curves for Algorithm 1 and Algorithm 2.
-
-sample-PR-curve
-
-The goal is to have a model be at the upper right corner, which is basically getting only the true positives with no false positives and no false negatives – a perfect classifier.
-
-The precision recall area under curve (PR AUC) is just the area under the PR curve. The higher it is, the better the model is.
-
-Differences between the ROC AUC and PR AUC
-Randy has given a great explanation here, plus a little of my understanding.
-
-Since PR does not account for true negatives (as TN is not a component of either Precision or Recall), or there are many more negatives than positives (a characteristic of class imbalance problem), use PR.
-If not, use ROC.
-
-Due to the inherent complex characteristics of imbalanced data sets, learning from such data requires new understandings, principles, algorithms, and tools to transform vast amounts of raw data efficiently into information and knowledge representation.
-
-We adopt precision recall curve and compare with logistic regression based on original and pca data
-
-Classification based on reconstruction or other value, such as similarity analysis
-
-Evaluate what metric is more discriminative, mse, tsnr or other
+%problem of imbalanced data
 
@@
 
+incorporate discriminative funcions
+
+evaluate other metrics beyond reconstruction, using techniques such as reconstruction with dictionary concatenation
+
 Tensor-based algorithms face challenges to achieve reasonable processing time to handle large-scale tensor factorizations \cite{de2014distributed}, demanding efforts in order to explore distributed processing techniques for tensor-based analytics of large datasets.

### 2017-11-14 distributed-tensor-dictionary-learning/projectDictRecover.m

@@ -22,7 +22,7 @@ s = 5;                                                                      % sp
 snr = 20;                                                                   % snr for added noise
 L = 2000;                                                                   % number of training vectors to use
 noIt = 100;                                                                 % number of iterations in each trial
-nofTrials = 1;                                                              % enought trials to obtain reliable results
+nofTrials = 100;                                                              % enought trials to obtain reliable results
 N = 20; M1 = 5; M2 = 4;                                                     % features/variables/components
 K = 50; N1 = 10; N2 = 5;                                                    % dictionary's atoms (K << L)
 % N = 80; M1 = 10; M2 = 8;                                                  % features/variables/components
@@ -33,9 +33,9 @@ fileNameInfo = sprintf('s=%1i_snr=%li_L=%li_noIt=%li_trials=%li_N=%li_K=%li',s,s
 fileNameSufix = sprintf('%s.mat',fileNameInfo);
 dataFiles = [
              ['L', fileNameSufix] % 'L', 'Q', 'C', 'H' or 'E' =             RLS-DLA (java),
-             ['T', fileNameSufix] % 'T' =                                   K-HOSVD
-             ['O', fileNameSufix] % 'O' =                                   T-MOD,             
-             ['K', fileNameSufix] % 'K' =                                   K-SVD,   
+             ['T', fileNameSufix] % 'T' =                                   K-HOSVD,
+             ['O', fileNameSufix] % 'O' =                                   T-MOD,
+             ['K', fileNameSufix] % 'K' =                                   K-SVD,
              ['D', fileNameSufix] % 'D' =                                   MOD,
              %['A', fileNameSufix] % 'A' =                                  AK-SVD,
              %['M', fileNameSufix] % 'M' =                                  ILS-DLA MOD,             

### 2017-11-14 fraudDetection/featureSelection/plot_pca_iris.py

@@ -1,6 +1,6 @@
 # coding: utf-8
 ########################################################################################################################
-# PCA example with Iris Data-set
+## PCA example with Iris Data-set
 ########################################################################################################################
 import warnings
 import numpy as np
@@ -12,17 +12,17 @@ warnings.filterwarnings("ignore", category=DeprecationWarning)
 
 np.random.seed(5)
 
-# original data
+## original data
 iris = datasets.load_iris()
 X = iris.data
 y = iris.target
 
-# transformed data
+## transformed data
 pca = decomposition.PCA(n_components=3)
 pca.fit(X)
 X = pca.transform(X)
 
-# plot
+## plot
 fig = plt.figure(1, figsize=(4, 3))
 plt.clf()
 ax = Axes3D(fig, rect=[0, 0, .95, 1], elev=48, azim=134)
@@ -33,9 +33,9 @@ for name, label in [('Setosa', 0), ('Versicolour', 1), ('Virginica', 2)]:							
 				horizontalalignment='center',
 				bbox=dict(alpha=.5, edgecolor='w', facecolor='w'))
 
-# Reorder the labels to have colors matching the y (i.e iris.target)
+## Reorder the labels to have colors matching the y (i.e iris.target)
 y = np.choose(y, [1, 2, 0]).astype(np.float)
 
-# plot
+## plot
 ax.scatter(X[:, 0], X[:, 1], X[:, 2], c=y, cmap=plt.cm.spectral)
 plt.show()

### 2017-11-15 docs/thesis/chapters/1_introduction/introduction.tex
tensor-dl and spell check

@@
 \begin{enumerate}
-	\item We propose an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset;
+	\item We propose an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset;
 	\item We discuss the computational complexity of the proposed framework and evaluate the required processing time for tested scenarios;

### 2017-11-15 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
tensor-dl and spell check

@@
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
-We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz, and the operator $\sim$ that sorts the elements of a vector in ascending order.
+We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matrix, and the operator $\sim$ that sorts the elements of a vector in ascending order.
 
@@
 The matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns.
-Each row represents a communication port (TCP port or UDP port), and each column represents time bins having a approtiate size, such as one minute.
+Each row represents a communication port (TCP port or UDP port), and each column represents time bins having a appropriate size, such as one minute.
 Each element $x_{m,n}^{(q)}$ stands for the number of times that the port $m$ appears at the $n$-th minute, at the $q$-th time frame.
@@
 At the end of each port scanning, a RST/ACK packet is sent by the system that is performing the port scan.
-Therefore, a full connection or a complete three-way handshake is never established, which makes the detection of the attack sender more difficult, and requires approaches able to identify probe attacks without connection stablishment.
+Therefore, a full connection or a complete three-way handshake is never established, which makes the detection of the attack sender more difficult, and requires approaches able to identify probe attacks without connection establishment.
 The UDP scan technique sends UDP packets to the destination port, and if it responds with a \emph{ICMP port unreachable} message, then it indicates that the scanned port is closed.
@@
 This section describes the proposed technique to detect synflood, fraggle and port scan, according to Figure \ref{fig:2_fig80}, which represents the overview of the proposed framework for detection and identification of network attacks.
-In Subsection \ref{sec:2_prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvelue for each $q$-th time frame.
-Next, in Subsection \ref{sec:2_prop_MOSSchemes}, we show how to apply the eiganvalues on the MOS scheme in order to detect the attack.
+In Subsection \ref{sec:2_prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvalue for each $q$-th time frame.
+Next, in Subsection \ref{sec:2_prop_MOSSchemes}, we show how to apply the eigenvalues on the MOS scheme in order to detect the attack.
 In Subsection \ref{sec:2_prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:2_prop_EigenSimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.
@@
 After the estimation of the $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected Attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked.
-To deal with this problem, the adoption of a similarity analysis between legitimate traffic and the traffic of time frames estimated as under attack is evaluated, analysing the effectiveness of cosine similarity to highlight abnormalies inserted by network traffic attacks.
+To deal with this problem, the adoption of a similarity analysis between legitimate traffic and the traffic of time frames estimated as under attack is evaluated, analyzing the effectiveness of cosine similarity to highlight abnormalities inserted by network traffic attacks.
 
@@
 
-For eigen similarity analyis, we evaluate the cosine similarity in order to identify lacks of similarity between legitimate and malicious traffic, as follows
+For eigen similarity analysis, we evaluate the cosine similarity in order to identify lacks of similarity between legitimate and malicious traffic, as follows
 
@@
 
-The most significant eigenvector $\boldsymbol{v}^{(q)}$, of a time frame $q$ without attack, can be derivated from (\ref{eq:eq06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the evaluated matrix.
+The most significant eigenvector $\boldsymbol{v}^{(q)}$, of a time frame $q$ without attack, can be derivate from (\ref{eq:eq06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the evaluated matrix.
 The same calculation shall be performed in order to obtain the target eigenvectors $\boldsymbol{v}_{(n)}$, calculated from a time frame without attack plus minutes of a time frame estimated as under attack, to evaluate the occurrence of network attacks.
 
-The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic whithout attack, from a time frame $q$ composed of $Q$ minutes of legitimate network traffic.
+The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic without attack, from a time frame $q$ composed of $Q$ minutes of legitimate network traffic.
 For the detailed attack identification, each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$, as represented by
@@
 
-Figure \ref{fig:2_fig8} ilustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.
+Figure \ref{fig:2_fig8} illustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.
 
@@
 The eigen similarity analysis starts at $\boldsymbol{x}^{(3)}_{(41)}$ and is incrementally performed until $\boldsymbol{x}^{(\hat{3})}_{(60)}$, in order to calculate the $s_n$.
-We assume that $s_n < l$ means an attack identification, according the anomaly on similiarity of $s_n$ to a defined limiar $l$.
-Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}$ for eigen similarity analyis, the $s_n$ is calculated according to~(\ref{eq:eq11}).
+We assume that $s_n < l$ means an attack identification, according the anomaly on similarity of $s_n$ to a defined limiar $l$.
+Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}$ for eigen similarity analysis, the $s_n$ is calculated according to~(\ref{eq:eq11}).
 
@@
 
-For similarity analyis, each $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, denoted as $x_{(m,n)}$, shall be individually replaced by the traffic of the evaluated $m$-th port of the $\hat{n}$-th minute under attack, denoted as $x^{(\hat{q})}_{(m,\hat{n})}$, in order to identify significant variation on similarity caused by the traffic of the $m$-th port.
+For similarity analysis, each $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, denoted as $x_{(m,n)}$, shall be individually replaced by the traffic of the evaluated $m$-th port of the $\hat{n}$-th minute under attack, denoted as $x^{(\hat{q})}_{(m,\hat{n})}$, in order to identify significant variation on similarity caused by the traffic of the $m$-th port.
 
@@
 In Table \ref{tab:tab4}, two MOS schemes outperforms from the others, EDC and EFT.
-Efficient Detection Criterion (EDC) and Exponential Fitting Test (EFT) are the most effective schemes, correctly estimating the number of attacks in comparison to the expected values for effetctive attack detection, as defined by the column of real values in Table \ref{tab:tab4}.
+Efficient Detection Criterion (EDC) and Exponential Fitting Test (EFT) are the most effective schemes, correctly estimating the number of attacks in comparison to the expected values for effective attack detection, as defined by the column of real values in Table \ref{tab:tab4}.
 The AIC and MDL schemes are satisfactory only for port scan detection, however SURE and RADOI schemes did not show effective results for port scan or flood detection.
@@
 
-According to Table \ref{tab:tab4}, EDC and EFT estimated correctly the number of attacks of a time frame vector, indicating that ocurred $\hat{d}$ network attacks, but not providing additional details, what highlights the necessity of complementary approaches in order to estimate the time and ports under attack.
+According to Table \ref{tab:tab4}, EDC and EFT estimated correctly the number of attacks of a time frame vector, indicating that occurred $\hat{d}$ network attacks, but not providing additional details, what highlights the necessity of complementary approaches in order to estimate the time and ports under attack.
 Hence, we propose apply eigen analysis to estimate the $q$-th time frames under attack and eigen similarity analysis to estimate the minutes and ports under attack.
@@
 
-Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respectives $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:2_alg01}.
+Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respective $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:2_alg01}.
 
@@
 Comparing the approaches for similarity analysis, it is possible to observe that all approaches highlight the low similarity when evaluated the traffic under attack.
-However, the incremental approach figured out low similarity for times without attack, where $n=16, 17, 18, 19, 20$, what indicates that the incremental approach can produce false posive results.
+However, the incremental approach figured out low similarity for times without attack, where $n=16, 17, 18, 19, 20$, what indicates that the incremental approach can produce false positive results.
 This behavior occurs because the incremental approaches appends all selected traffic into the reference traffic for comparison against the original reference traffic, what makes more evident the first lack of similarity but reduces the changing detection capability after an attack detection.
@@
 Comparing the incremental individualized and the individual approaches for eigen similarity analysis, it is possible to observe that the incremental individualized approach obtain lowest values for almost all cases, except for the time $n=14$, where incremental individualized approach identified a larger similarity than the individual approach.
-The incremental individualized appends information about each evaluated traffic, therefore it incorporates traffic behaviors that can reduce the outlier capability detection, as ocurred for the time $n=14$.
+The incremental individualized appends information about each evaluated traffic, therefore it incorporates traffic behaviors that can reduce the outlier capability detection, as occurred for the time $n=14$.
 
@@
 
-For fraggle attack detection, the lack of similarity between legitimate and malicious traffic was more evident than for the evaluation of synflood and port scan detecion.
-This behavior can be explained by the number of packets generated through the fraggle attack simulation, that was significative larger than the number of packets generated during the synflood simulation.
+For fraggle attack detection, the lack of similarity between legitimate and malicious traffic was more evident than for the evaluation of synflood and port scan detection.
+This behavior can be explained by the number of packets generated through the fraggle attack simulation, that was significantly larger than the number of packets generated during the synflood simulation.
 Considering the three approaches, the largest value for times under attack was 0.0083, while the shortest value for times without attacks was 0.9993.
@@
 The incremental individualized approach presented more sensibility to anomaly detection than the individual approach, the former produced the identification of a low similarity of 0.0298 for almost all ports under attack, unless the port 21, although the simulation has attacked this port.
-The individual approach was not able to identify low similarity for ports under attack, resunting in values of 0.9997 for ports with anomalous traffic and 0.9999 for ports without network attack.
+The individual approach was not able to identify low similarity for ports under attack, resulting in values of 0.9997 for ports with anomalous traffic and 0.9999 for ports without network attack.
 
 For the evaluation of the proposed approaches for identification of ports under synflood and fraggle attack, all minutes of each time frame, in which one attack location was estimated, were analyzed.
-Even though, due to space limitations, only the results of the firt minute where a low similarity was idenfified will be shown such as where $n=11$.
-Nevertheless, the results obtained for the evaluation of traffic without attack presented high similarity to the reference traffic, with similarities close to 0.9999, and the evaluation of the other minutes under attack presented results quite similar to the resuts shown in the Tables \ref{tab:tab9} and \ref{tab:tab10}.
+Even though, due to space limitations, only the results of the first minute where a low similarity was identified will be shown such as where $n=11$.
+Nevertheless, the results obtained for the evaluation of traffic without attack presented high similarity to the reference traffic, with similarities close to 0.9999, and the evaluation of the other minutes under attack presented results quite similar to the results shown in the Tables \ref{tab:tab9} and \ref{tab:tab10}.
 
@@
 
-According to results presented in Table \ref{tab:tab9}, both approches identifies low similarity for the traffic of port 600, which is the target port of the simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.
+According to results presented in Table \ref{tab:tab9}, both approaches identifies low similarity for the traffic of port 600, which is the target port of the simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.
 
@@
 
-The results for the avaluation of ports under fraggle attack, shown in Table \ref{tab:tab10}, were similar to the results obtained for synflood analysis, with the identification of low similarity for traffic of the port under attack.
+The results for the evaluation of ports under fraggle attack, shown in Table \ref{tab:tab10}, were similar to the results obtained for synflood analysis, with the identification of low similarity for traffic of the port under attack.
 Nevertheless, for fraggle analysis, the individual approach identified the lowest similarity, that is 0.0004 while the incremental individualized approach obtained a similarity of 0.0031.
@@
 Since the proposed framework focus on flood and probe attack detection, only the attacks with behavior similar to flood or probe attack were evaluated.
-Initially all DoS and probe attacks were selected, but it was observed that the most cases of DoS focus on exploit system vulnerabilities instead of flooding attack, and most of probe attacks focus on ICMP intead of port scanning.
+Initially all DoS and probe attacks were selected, but it was observed that the most cases of DoS focus on exploit system vulnerabilities instead of flooding attack, and most of probe attacks focus on ICMP instead of port scanning.
 Therefore, for evaluation of the proposed approach for flood and probe attack detection, it is necessary to select cases that implements flood or port scan behaviors.
@@
 Therefore, the proposed framework is $O(N^3 + Q \log Q + \hat{d}Q + N^3)$ and its worst-case running time is $O(N^3)$.
-The computational complexity of EVD is predominant in the framework, but the approach splits the data into time frames with period time $N$, which makes possible to limitate the growth of $N$ even for evaluations of cases with total time larger than $N$, reducing the impact caused by the computational complexity of EVD.
+The computational complexity of EVD is predominant in the framework, but the approach splits the data into time frames with period time $N$, which makes possible to limit the growth of $N$ even for evaluations of cases with total time larger than $N$, reducing the impact caused by the computational complexity of EVD.
 
@@
 Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets.
-Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high thoughput network traffic.
+Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high throughput network traffic.
 Additionally, future research can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.

### 2017-11-15 docs/thesis/chapters/4_tensor_dl/tensor_dl.tex
tensor-dl and spell check

@@
 
-Compressed Sensing (CS) aims to obtain the most relevant information of a dataset, what makes it useful for compression, signal reconstruction, image processing and classification, such as in sparce representation classification (SRC) problems.
+Compressed Sensing (CS) aims to obtain the most relevant information of a dataset, what makes it useful for compression, signal reconstruction, image processing and classification, such as in sparse representation classification (SRC) problems.
 
@@
 Shearlets [25], which is obtained from dilations, action of translations and shear transformations, has nice geometric properties and mathematical properties for image representation.
-Bandelets [26], which specifies the geometry as a vector field, is de-signed to improve the image compression and noise reduction performance.
+Bandelets [26], which specifies the geometry as a vector field, is designed to improve the image compression and noise reduction performance.
 
@@
 
-Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, this work propose a tensor-based DL for fraud detection from imbalanced data, in order to apply the tensor deconposition for DL methods to evidentiate the discriminative sensing of a fraud detection dataset.
+Considering that big data problems require techniques to deal with multidimensional data in order to make sense of structure and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highlight discriminative structure and relationships of one dataset, this work propose a tensor-based DL for fraud detection from imbalanced data, in order to apply the tensor decomposition for DL methods to highlight the discriminative sensing of a fraud detection dataset.
 
@@
 Section \ref{sec:4_datamodel} presents the data model and the evaluated datasets.
-Section \ref{sec:4_proposal} describes the proposed approach for for fraud detecton from mobile money transacion.
+Section \ref{sec:4_proposal} describes the proposed approach for fraud detection from mobile money transactions.
 Section \ref{sec:2_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.
@@
 	\includegraphics[width=10cm]{figures/5_20_2000_1000_100.eps}
-	\caption{Cumulative atom identificatin per degree rates. $T=2000$, $M=20$, $N=50$}
+	\caption{Cumulative atom identification per degree rates. $T=2000$, $M=20$, $N=50$}
 	\label{fig:fig1}
@@
 It is possible to observe that the tensor-base methods outperform MOD, RLS-DLA and K-SVD in number of identified atoms and the required degrees for identification.
-Note that MOD performs quite silimar to RLS-DLA, while K-SVD is the worst evaluated case for this experiment.
+Note that MOD performs quite similar to RLS-DLA, while K-SVD is the worst evaluated case for this experiment.
 Considering the tensor-based methods, T-MOD outperforms K-HOSVD with higher atom identification over lower degrees.
@@
 	\includegraphics[width=10cm]{figures/5_20_2000_24000_100.eps}
-	\caption{Cumulative atom identificatin per degree rates. $T=2000$, $M=80$, $N=30$}
+	\caption{Cumulative atom identification per degree rates. $T=2000$, $M=80$, $N=300$}
 	\label{fig:fig2}
@@
 The results reveals that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases.
-Since dictionary learning and reconstruction error are usually used for sparce representation classification (SRC) problems, it is possible to supose that tensor-based algorithms can improve results of classification problems, as well as can extract hidden patterns and structure in multidimensional data analytics problems.
+Since dictionary learning and reconstruction error are usually used for sparse representation classification (SRC) problems, it is possible to suppose that tensor-based algorithms can improve results of classification problems, as well as can extract hidden patterns and structure in multidimensional data analytics problems.
 
@@
 
-A key challenge to use SC and DL for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset.
-Therefore, we introduce a tensor-based DL method for fraud detecton from unbalanced data.
+A key challenge to use SC and DL for classification is how to find proper dictionaries and coefficients that highlight discriminative structure and relationships of one dataset.
+Therefore, we introduce a tensor-based DL method for fraud detection from unbalanced data.
 Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification.
@@
 In this chapter, we propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds.
-We propose to use tensor-based DL for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
+We propose to use tensor-based DL for learning fraudulent and legitimate data separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics.
 
@@
 
-Compressed sensing (also known as compressive sensing, compressive sampling, or sparse sampling) is a signal processing technique for efficiently acquiring and reconstructing a signal, by finding solutions to underdetermined linear systems.
+Compressed sensing (also known as compressive sensing, compressive sampling, or sparse sampling) is a signal processing technique for efficiently acquiring and reconstructing a signal, by finding solutions to undetermined linear systems.
 This is based on the principle that, through optimization, the sparsity of a signal can be exploited to recover it from far fewer samples than required by the Shannon-Nyquist sampling theorem.
@@
 The emergence of sparse dictionary learning methods was stimulated by the fact that in signal processing one typically wants to represent the input data using as few components as possible.
-Before this approach the general practice was to use predefined dictionaries (such as fourier or wavelet transforms).
+Before this approach the general practice was to use predefined dictionaries (such as Fourier or wavelet transforms).
 However, in certain cases a dictionary that is trained to fit the input data can significantly improve the sparsity, which has applications in data decomposition, compression and analysis and has been used in the fields of image denoising and classification, video and audio processing.
@@
 Additionally, knowing $\boldsymbol{A}$ we obtain $\boldsymbol{S}$ by solving a sparse recovery problem.
-MOD iterates between the two steps, starting with an initial guess of $\boldsymbol{A}$, denoted as $\hat{\boldsymbol{A}}$, which can be estimated, generated by previous dictionary learning or generated analitically.
+MOD iterates between the two steps, starting with an initial guess of $\boldsymbol{A}$, denoted as $\hat{\boldsymbol{A}}$, which can be estimated, generated by previous dictionary learning or generated analytically.
 Despite - simplicity, the MOD algorithm has been criticized for its slow convergence, which is due to the fact that $\boldsymbol{A}$ and $\boldsymbol{S}$ are updated separately, which may lead to a slow convergence \cite{aharon2006rm}.
@@
 Tensor data are typically highly structured, a perfect match for compressive sampling, so that the CS framework relaxes data acquisition requirements, enables compact storage, and facilitates data completion (i.e., inpainting of missing samples due to a faulty sensor or unreliable measurement) \cite{cichocki2015tensor}.
-Roemer \emph{et al.} \cite{roemer2014tensor} have proposed a tensor-based multidimensional extensions for MOD and K-SVD algorithms, which are respectivelly T-MOD and K-HOSVD, and show their improved performances numerically, for the dictionary reconstruction problem.
+Roemer \emph{et al.} \cite{roemer2014tensor} have proposed a tensor-based multidimensional extensions for MOD and K-SVD algorithms, which are respectively T-MOD and K-HOSVD, and show their improved performances numerically, for the dictionary reconstruction problem.
 
-Considering that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases, accordint to subsection \ref{sec:4_motivation_results}, we believe that tensor-based DL algorithms can improve the discriminative sensing for classification classification problems.
-Therefore, we propose use tensor-based DL for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics
+Considering that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases, according to subsection \ref{sec:4_motivation_results}, we believe that tensor-based DL algorithms can improve the discriminative sensing for classification classification problems.
+Therefore, we propose use tensor-based DL for learning fraudulent and legitimate data separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics
 
@@
 This dataset is synthetically generated by  PaySim \cite{lopez2016paysim}, which is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.
-PaySim uses aggregated data from the private dataset to generate a synthetic dataset that resembles the normal operation of transactions and injects malicious behaviour to later evaluate the performance of fraud detection methods.
+PaySim uses aggregated data from the private dataset to generate a synthetic dataset that resembles the normal operation of transactions and injects malicious behavior to later evaluate the performance of fraud detection methods.
 PaySim simulates mobile money transactions based on a sample of real transactions extracted from one month of financial logs from a mobile money service implemented in an African country.
@@
 The selected dataset is composed by 9 features (variables) and 2.770.408 transactions (observations).
-It is also provided one lable that classifies each transaction as fraudulent or not.
-As the majority dataset of anomaly or fraud detection, the data is imbalanced, where the number of fraudulent transactions is 8.213 against 2.762.196 of normal transactions.
-Therefore, the fraud class represents 0,30 \% of the entire datased.
+It is also provided one label that classifies each transaction as fraudulent or not.
+As the majority dataset of anomaly or fraud detection, the dataset of mobile money transactions is imbalanced, where the number of fraudulent transactions is 8.213 against 2.762.196 of normal transactions.
+Therefore, the fraud class represents 0,30 \% of the total transactions.
 
 For evaluation of the proposed classification approach we adopt a data splitting into training and test data, where the training data has 70 \% of the whole dataset.
-We denote the traning data as $X_r$ and the test data as $X_e$.
-The dataset can also be divided by normal and fraudulent transactions, where the normal train data is denoted as $X_r^o$, the fraud train data is denoted as $X_r^f$.
+We denote the training data as $X_r$ and the test data as $X_e$.
+The dataset can also be divided by normal and fraudulent transactions, where the normal train data is denoted as $X_r^o$ and the fraud train data is denoted as $X_r^f$.
 The same notation of fraud, normal, train and test is applied to qualify $A$ and $S$.
@@
 
-\section{Tensor-based Sparce Representation Classification}
+\section{Tensor-based Sparse Representation Classification}
 \label{sec:4_proposal}
 
-We propose a Tensor-based Sparce Representation Classification approach to use tensor-based DL for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
+We propose a Tensor-based Sparse Representation Classification approach to use tensor-based DL for learning fraudulent and legitimate dictionaries separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics.
 
 Our proposal requires supervision, where labels of training data are used to classify each transaction as fraud or normal.
-The first step is to split the training dataset into classes, as fraud or normal, and create respectively the data matrices $\boldsymbol{X}_r^f$ and $\boldsymbol{X}_r^o$.
-For each data matrix we compute the data normalization and subsequently the initial dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$.
-Afterward, we apply the dictionaries and the data matrices to the selected DL algorithms, in order to learn the best dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$ that respectivelly represents fraud and normal classes as basis vectors.
+The first step is to split the training dataset into fraud and normal classes, and create the data matrices $\boldsymbol{X}_r^f$ and $\boldsymbol{X}_r^o$, respectively.
+For each data matrix, we randomly select $N$ columns into a matrix and compute its data normalization in order to estimate the initial dictionaries $\boldsymbol{\hat{A}}_r^f$ and $\boldsymbol{\hat{A}}_r^o$.
+Afterward, we apply the dictionaries and the data matrices to the selected DL algorithms for learning the best dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$ that respectively represents fraud and normal classes as basis vectors.
 
-With the trainned dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$ and the test data $\boldsymbol{X}_e$ it is possible to process each transacion of $\boldsymbol{X}_e$, denoted as $\boldsymbol{x}_e^t$ where $t$ represents the transaction index, in order to compute its sparse coding from fraud and normal dictionaries, generating respectively $\boldsymbol{S}_e^f$ and $\boldsymbol{S}_e^o$.
-Subsequently, we compute the signal rescontruction as $\boldsymbol{x}_e^f = \boldsymbol{A}_r^f \cdot \boldsymbol{S}_e^f$ and $\boldsymbol{x}_e^o = \boldsymbol{A}_r^o \cdot \boldsymbol{S}_e^o$.
+With the trained dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$ and the test data $\boldsymbol{X}_e$ it is possible to process each transaction of $\boldsymbol{X}_e$, denoted as $\boldsymbol{x}_t$, where $t$ represents the transaction index, in order to compute its sparse coding from fraud and normal dictionaries, generating respectively $\boldsymbol{S}_e^f$ and $\boldsymbol{S}_e^o$.
+Subsequently, we compute the signal reconstruction for each test transaction $t$, as $\boldsymbol{x}_e^f = \boldsymbol{A}_r^f \cdot \boldsymbol{S}_e^f$ and $\boldsymbol{x}_e^o = \boldsymbol{A}_r^o \cdot \boldsymbol{S}_e^o$.
 
-Finally, we compare $\boldsymbol{x}_e^f$ and $\boldsymbol{x}_e^o$ to $\boldsymbol{x}_e^t$ in order to evaluate the reconstruction error to classify the transaction as fraud or normal according to the minimum reconstruction error from $\boldsymbol{x}_e^f$ and $\boldsymbol{x}_e^o$, respectivelly.
+Finally, we compare $\boldsymbol{x}_t^f$ and $\boldsymbol{x}_t^o$ to $\boldsymbol{x}_t$ in order to evaluate the reconstruction error and to classify the transaction as fraud or normal according to the minimum reconstruction error from $\boldsymbol{x}_t^f$ and $\boldsymbol{x}_t^o$, respectively.
 We adopt the cosine similarity to compare the reconstructed signal to the original one.
@@
 
-For data analysis of imbalanced classes it is usually applied a data undersample method, where an under sized sample is selected from the original datased, with similar amount of occurrences for each class of the data.
-This method is efficient for algorithm trainning, but it faces generalization problems when tries to classify new data.
-Therefore, we propose to evaluate the our proposal for under sampled data and for the complete dataset.
+For data analysis of imbalanced classes it is usually applied a data undersample method, where an under sized sample is selected from the original dataset, with similar amount of occurrences for each class of the dataset.
+This method is efficient for classification algorithm training, but it faces generalization problems when tries to classify new data.
 
-The Receiver Operating Characteristics (ROC) curve makes use of the proportion of two metrics: true positives rate (TPR) and false positives rate (FPR), which are defined as:
+Considering that fraud or other kind of malicious events are rare when compared to legitimate events, we propose to evaluate our proposal for undersampled data and for the complete dataset, evaluating the results for training using undersample data ans testing with undersample test, as well as using the undersample data for training and the complete testing data to evaluate the classification accuracy.
+
+The approach described in Section \ref{sec:4_proposal} can adopt any DL algorithm, however we propose to evaluate the K-HOSVD in order to analyze how a tensor-based DL algorithm can highlight the discriminative structure and relationships of one dataset.
+We propose to compare the proposed approach against the Support Vector Machines (SVM) algorithm and against the Logistic Regression (LR) algorithm, using features extracted trough Principal Component Analysis (PCA).
+
+SVM is one classifier algorithm that has been successfully applied to areas ranging from handwriting recognition to text classification.
+SVM has also received considerable attention for classifying high-dimensional non-Gaussian data, with applications for anomaly detection.
+SVMs yield good generalization performance on such problems by directly estimating a decision boundary with maximal separability \cite{akbani2004applying, banerjee2006support}.
+LR is the most common method used to model binary response data \cite{hilbe2011logistic}, it is useful for situations in where it is necessary to predict the presence or absence of a characteristic, with binary classification, such as in anomaly or fraud detection \cite{bhattacharyya2011data}.
+
+We adopt the area under the Receiver Operating Characteristic (ROC) curve and the area under the Precision-Recall (PR) curve in order to evaluate our approach for fraud detection from a mobile money transaction dataset, and to compare the results against the logistic regression algorithm applied to the same problem.
+The ROC curve makes use of the proportion of two metrics: true positives rate (TPR) and false positives rate (FPR), which are defined as:
 
@@
 ROC curves are commonly used to present results for binary decision problems in machine learning.
-However, when dealing with highly skewed datasets, Precision-Recall (PR) curves give a more informative picture of an algorithm’s performance \cite{davis2006relationship, he2009learning}.
-The PR curve is defined by plotting Precision Rate (PR) over the Recall Rate (RR), which are defined as:
+However, when dealing with highly skewed datasets, PR curves give a more informative picture of an algorithm’s performance \cite{davis2006relationship, he2009learning}.
+The PR curve is defined by plotting Precision Rate (Pr) over the Recall Rate (Rr), which are defined as:
 
 \begin{equation}\label{eq:4_eq06}
-	PR = \frac{TP}{TP + FP},
+	Pr = \frac{TP}{TP + FP},
 \end{equation}
@@
 \begin{equation}\label{eq:4_eq07}
-	RR = \frac{TP}{TP + FN}.
+	Rr = \frac{TP}{TP + FN}.
 \end{equation}
 
-PR curves exhibit a strong correspondence to ROC curves: A curve dominates in ROC space if and only if it dominates in PR space.
-However, an algorithm that optimizes the AUC in the ROC space is not guaranteed to optimize the AUC in PR space.
-Considering that our data is very imbalanced, with many more negatives than positives values, we adopt the averate precision recall for the classification evaluation
+PR curves exhibit a strong correspondence to ROC curves, where a curve dominates in ROC space if and only if it dominates in PR space.
+However, an algorithm that optimizes the Area Under the Curve (AUC) in the ROC space is not guaranteed to optimize the AUC in PR space.
+Considering that our data is very imbalanced, with many more negatives than positives (frauds) values, we adopt the area under the PR curve for classification evaluation, but we also compute the area under the ROC curve as a secondary classification metric.
 
-Logistic regression is the most common method used to model binary response data \cite{hilbe2011logistic}, it is useful for situations in where it is necessary to predict the presence or absence of a characteristic, with binary classification, such as in anomaly or fraud detection \cite{bhattacharyya2011data}.
-Therefore, we compare the results of out proposed approach with the logistic regression algorithm applied for fraud detection from the adopted datase.
+Additionally, we perform a cross-validation analysis in order to optimize the parameters $K$ and $N$, evaluating the values of 2, 3 and 5 for $K$ and 10, 20, 40 and 100 for $N$.
 
@@
 
+Table \ref{tab:4_tab1} presents the results for fraud detection from mobile money transactions through SVM, LR with PCA and the proposed K-HOSVD-based classification.
+
+\begin{table}[h!]
+  \centering
+  % \footnotesize
+  \caption{Results for fraud detection from mobile money transactions} 
+  \label{tab:4_tab1}
+  \begin{tabular}{ c c c c c }
+	\toprule
+	\multirow{2}{*}{\textbf{Classifier}}   &\multicolumn{2}{c}{\textbf{undersample/undersample}} &\multicolumn{2}{c}{\textbf{undersample/complete}}\\ 
+			\hhline{~----}
+			&\textbf{ROC-AUC} &\textbf{PR-AUC} &\textbf{ROC-AUC} &\textbf{PR-AUC}\\
+	\midrule
+	SVM &0.5039 &0.7493 &0.8496 &0.8497 \\
+	LR with PCA &0.9767 &0.9793 &0.5633 &0.0051 \\
+	K-HOSVD ($K=3$, $N=40$) &0.8622 &0.8917 &0.8652 &0.4578 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+According to the results, the SVM algorithm presents the best PR-AUC for the scenario that evaluates the generalization performance for fraud detection of imbalanced data.
+However, the PR-AUC of the SVM algorithm was the worst case for training and testing with undersample data.
+The LR algorithm applied to PCA data presents the best results for training and testing with undersample data, although its results for the evaluation of the complete testing data presents a very low PR-AUC.
+
+The proposed Tensor-based Sparse Representation Classification, using K-HOSVD presents higher PR-AUC than SVM for training and testing with undersample data, but its results are worst than LR with PCA for the same scenario.
+For training the undersample data and test with complete data, the proposed approach presents better PR-AUC than LR with PCA and the best ROC-AUC.
+Therefore, it is possible to observe that our approach presents average results if compared to two well known and highly adopted classifier algorithms.
+
 % \section{Discussion}
-% \label{sec:4_Discussion}
+% \undersample{sec:4_Discussion}
 
@@
 
-The conclusion goes here.
+The results shows how imbalanced data and its instance for fraud detection in mobile money transactions are challenging for classification algorithms, regarding the classification performance for undersampled data and for generalization from undersample data training to complete data testing.
 
-The importance of the right metric for classification of imbalanced data
+It is also possible to observe that the proposed tensor-based sparse representation classification, using K-HOSVD, presents average performance and highlights some enhancement possibilities, such as:
 
-Tensor based analysis explores the extructure and relationship of multidimensional data and can improve the discriminative power of imbalanced data for classification
-
-Can be used for feature extraction and used for well known algorithm, and can be used for discriminative classification based on heuristics
-
-Our proposal is better for imbalanced data than others techniques
-
-incorporate discriminative funcions
-
-evaluate other metrics beyond reconstruction, using techniques such as reconstruction with dictionary concatenation
+\begin{enumerate}
+	\item Extend the proposed approach to adopt one dictionary with basis vectors that represent legitimate and fraudulent data, as a concatenation or as a kronecker product, with the classification according to a sparse code comparison;
+	\item Extend the proposed approach to adopt discriminative discriminative functions for T-MOD and K-HOSVD;
+	\item Extend the evaluation for other DL algorithms.
+\end{enumerate}
 

### 2017-11-15 docs/thesis/chapters/5_conclusion_future_work/conclusion.tex
chapter 2 ok
tensor-dl and spell check

@@
 The proposed technique is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis, where MOS and eigenvalue analysis are applied to detect time frames under attack.
-In addition, we evaluated the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset.
+In addition, we evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset.
 
@@
 
-Finally, we proposed a fraud detection approach composed by tensor-based dictionary learning for discriminant analysis of a mobile payment dataset.
+Finally, we proposed a fraud detection approach composed by tensor-based dictionary learning for discriminant analysis of a mobile payment dataset, where the problem of imbalanced data is discussed and it is presented one analysis regarding the classification accuracy of the proposed approach against well known classifiers.
 
@@
 
-The incremental individualized approach of eigen similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to eigen similarity analyis for network attack detection.
+The incremental individualized approach of eigen similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to eigen similarity analysis for network attack detection.
 Therefore, the incremental individualized approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.
@@
 It is implemented as a fully working mobile client and can be used for any type of enterprise.
-Also, part of concept is seed for new security solutions for big data apps.
+Also, part of concept is seed for new security solutions for big data applications.
+
+This thesis shows that a tensor-based approach can be used for fraud detection from mobile money transactions, with average results when compared to well known classifiers in one scenario with highly skewed datasets.
 
@@
 
-We attempt to analyse the processing capacity problem of measurement of distributed systems through network traffic analysis, the results of the work presented in this thesis provide the contributions below:
+We analyze problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques.
+The results of the work presented in this thesis provide the following contributions:
 
 \begin{enumerate}
-	\item We proposed an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset;
+	\item We proposed an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset;
 	\item We discussed the computational complexity of the proposed framework and evaluated the required processing time for tested scenarios;
-	\item We proposed an architechture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
+	\item We proposed an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
 	\item We discussed the processing time of the proposed framework for mobile devices;
-	\item We proposed an tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
+	\item We proposed a tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
 	\item We published the following papers reporting our results:
@@
 	\item Improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets;
-	\item Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high thoughput network traffic;
+	\item Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high throughput network traffic;
 	\item Evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis;

### 2017-11-15 docs/thesis/chapters/appendix/b_pca_svm_cfa/pca_svm_cfa.tex
introducion...
tensor-dl and spell check

@@
-\chapter{Critical Success Factor Analysis Based on Feaure Selection}
-\label{ch:b_csf_fs}
+\chapter{Critical Success Factor Analysis Based on Feature Selection}
+\label{apx:b_csf_fs}
 
@@
 CSFs represent the principal assets or areas that must be given investments to achieve better results.
-CSF analysis is one challenger strategic management tool, wich can provide a robust and very practical assessment for strategic planners.
+CSF analysis is one challenger strategic management tool, which can provide a robust and very practical assessment for strategic planners.
 
-The identification of the most significative information for one problem is referred to as feature selection by the signal processing and data mining areas, as well as it can be formulated as a principal component problem, which is a widely adopted signal processing technique for data visualization and feature extraction.
+The identification of the most significant information for one problem is referred to as feature selection by the signal processing and data mining areas, as well as it can be formulated as a principal component problem, which is a widely adopted signal processing technique for data visualization and feature extraction.
 Feature selection aims to select a subset of relevant information from a larger dataset, in order to improve: data visualization and data understanding, storage requirements, dimensionality, processing time, discriminative sensing, and to overcome overfitting problems to improve prediction and classification performance \cite{chandrashekar2014survey}.
@@
 Recursive Feature Elimination (RFE) is a feature selection method for small sample classification problems.
-RFE seeks to improve generalization performance by recursivelly removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features \cite{chen2007enhanced}.
+RFE seeks to improve generalization performance by recursively removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features \cite{chen2007enhanced}.
 
-We propose a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminant analysis and based on RFE combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
-Results show how PCA can make the data discriminative and that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of brazilian public sector.
-Finally, SVM is used to highlight the more significant features identified by RFE, wich are similar to CSFs previsously identified by a qualitative analysis of the same datased.
+We propose a critical factors analysis based on Principal Component Analysis (PCA) for visual discriminant analysis and based on RFE combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of Brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
+Results show how PCA can make the data discriminative and that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of Brazilian public sector.
+Finally, SVM is used to highlight the more significant features identified by RFE, which are similar to CSFs previously identified by a qualitative analysis of the same dataset.
 
@@
 
-The brazilian Federal Court of Accounts (TCU, in Portuguese) surveys data regarding IT practices of brazilian public organizations in order to audit IT governance.
-The dataset with a consolidate view about the answers for this surgey the IT governance index is called iGovTI.
-The iGovTI is composed by 201 multiple choice questions, used for ranking according to their IT governance, submmited to 349 organizations.
-The TCU computes the IT governance index and classifies the IT governace of each organization.
+The Brazilian Federal Court of Accounts (TCU, in Portuguese) surveys data regarding IT practices of Brazilian public organizations in order to audit IT governance.
+The dataset with a consolidate view about the answers for this survey the IT governance index is called iGovTI.
+The iGovTI is composed by 201 multiple choice questions, used for ranking according to their IT governance, submitted to 349 organizations.
+The TCU computes the IT governance index and classifies the IT governance of each organization.
 Additionally, Ramos \emph{et al} \cite{ramos2016information} classifies each question regarding its relevance for IT governance through a qualitative analysis, and identify the CSFs for selected IT managers regarding IT governance.
@@
 
-PCA is a statistical technique commonly used for signal denoising, blind source separation, data compression, data visualization, feature extraction and dimensionality reduction, where a reduced number of features is extrated retaining as much information as possible \cite{jolliffe1986principal}.
+PCA is a statistical technique commonly used for signal denoising, blind source separation, data compression, data visualization, feature extraction and dimensionality reduction, where a reduced number of features is extracted retaining as much information as possible \cite{jolliffe1986principal}.
 It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
@@
 The eigenvectors $\mathbf{v}_i$ of $\mathbf{C}$ are called principal components (PC), and the sample variance along $\mathbf{v}_i$ of $\mathbf{C}$ is given by the corresponding eigenvalue $\lambda_i$.
-Projecting onto the eigenvectors with the largest eigenvalues assumes that minimal information is lost, considering that in many applications these directions contain the most interesting information, such as in in data compression and de-noising.
+Projecting onto the eigenvectors with the largest eigenvalues assumes that minimal information is lost, considering that in many applications these directions contain the most interesting information, such as in in data compression and denoising.
 
-Initially, we compute the covariance matrix $\mathbf{C}$ of a zero mean data and visualize the data relationship for the sambple covariance.
+Initially, we compute the covariance matrix $\mathbf{C}$ of a zero mean data and visualize the data relationship for the sample covariance.
 Sample covariance is calculated by computing deviations of each measurement from the average of all measurements for that variable.
@@
 
-Feature selection aims the identification of the most significative information for one problem or algorithm, such as classification or prediction problems.
-RFE is a feature selection method for small sample classification problems by recursivelly removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features through a selected classifier.
+Feature selection aims the identification of the most significant information for one problem or algorithm, such as classification or prediction problems.
+RFE is a feature selection method for small sample classification problems by recursively removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features through a selected classifier.
 
-Initially, it is necessary to identify one classifier that can identify the iGovTI classificatin of one organization according to a trainning dataset.
-Therefore, we propose an algorithm evaluatio in order to identify wich one presents the best accuracy for iGovTI classification.
+Initially, it is necessary to identify one classifier that can identify the iGovTI classification of one organization according to a training dataset.
+Therefore, we propose an algorithm evaluation in order to identify which one presents the best accuracy for iGovTI classification.
 The selected algorithm should be used by RFE to identify the CSF, that should be compared to results of the CSF qualitative analysis conducted by Ramos \emph{et al} \cite{ramos2016information}.
@@
 That procedure is recursively repeated on the pruned set until the desired number of features to select is eventually reached.
-RFE can also rank all features areaccording to when they were eliminated.
+RFE can also rank all features according to when they were eliminated.
 
@@
 
-In this section we present the experiements and results for the visual discriminant analysis based on PCA and for CSF analysis based on RFE, and discuss the results.
+In this section we present the experiments and results for the visual discriminant analysis based on PCA and for CSF analysis based on RFE, and discuss the results.
 
-For the visual discriminant analysis we initially we compute the zero mean and the the covariance matrix $\mathbf{C}$ is computed from $\mathbf{X}$.The resuls are shown by Figure \ref{fig:ch2_fig1}.
+For the visual discriminant analysis we initially we compute the zero mean and the the covariance matrix $\mathbf{C}$ is computed from $\mathbf{X}$.The results are shown by Figure \ref{fig:ch2_fig1}.
 
@@
 Therefore, we present the Figure \ref{fig:ch2_fig2}, wich shows the variance ECDF of the iGovTI questions and shows that 97\% of the questions have variance lesser than 2, while just 3\% of the questions have variance between 2 and 11.
-This result indicates that just a few principal components concentrates the most significative information and motivates the evaluation regarding the relationship between the principal components and the iGovTI classification.
+This result indicates that just a few principal components concentrates the most significant information and motivates the evaluation regarding the relationship between the principal components and the iGovTI classification.
 
@@
 
-Finally, we select the two largest eigenvalues and their correspondent componets in order to evaluate the relationship between the iGovTI classification for 349 organizations and the two most informative variables.
-The Figure \ref{fig:ch2_fig3} shows the scatter diagram that plots the organizations with higher iGovTI with larger circunferences and colors near of red, while organizations with lower iGovTI have colors near of blue and shorter circunferences.
+Finally, we select the two largest eigenvalues and their correspondent components in order to evaluate the relationship between the iGovTI classification for 349 organizations and the two most informative variables.
+The Figure \ref{fig:ch2_fig3} shows the scatter diagram that plots the organizations with higher iGovTI with larger circumferences and colors near of red, while organizations with lower iGovTI have colors near of blue and shorter circumferences.
 It is possible to observe that organizations with higher iGovTI are concentrated in the top left quadrant, with a visual separation of organizations with lower iGovTI, but without a clear distance separating the classes.
@@
 
-The visual discriminant analysis indicates that just a few principal components concentrates the most significative information and that the two principal components show a visual classification of organizations according to iGovTI classification, however it is not clear what are the CSF.
+The visual discriminant analysis indicates that just a few principal components concentrates the most significant information and that the two principal components show a visual classification of organizations according to iGovTI classification, however it is not clear what are the CSF.
 
@@
 
-RFE requires a classifier or predictor to evaluate its generalization performance by recursivelly removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features.
+RFE requires a classifier or predictor to evaluate its generalization performance by recursively removing the least significant features whose deletion will have the least effect on training errors, according to the higher variance measured from the features.
 We perform an algorithm evaluation in order to identify the classifier with the highest accuracy for iGovTI classification.
@@
 
-The Table \ref{tab:ch2_tab1} present the selected classification algoritms and the measured accuracy for iGovTI classification.
+The Table \ref{tab:ch2_tab1} present the selected classification algorithms and the measured accuracy for iGovTI classification.
 According to results, SVM \cite{hearst1998support} is the algorithm with highest classification accuracy, with 91.42\%, while the second is Elastic Net \cite{zou2005regularization}, with 83.15\% of accuracy.
@@
 The next step is to use the selected algorithm as validator for recursive feature elimination in order to identify the CSF for iGovTI.
-Additionally, the RFE algorithm requires the number of desired most significatives features.
+Additionally, the RFE algorithm requires the number of desired most significants features.
 We adopt 54 for this variable, considering that is the number of CSF identified by Ramos \emph{et al} \cite{ramos2016information} in a qualitative analysis, that is the proposed target to validate the results of the CSF identification for iGovTI.
 
-The RFE algorith presents an ranking according to the significance of each feature, with an accuracy of 69,9\% for CSF classification when compared to the CSF identified by Ramos \emph{et al} \cite{ramos2016information} in a qualitative analysis.
+The RFE algorithm presents an ranking according to the significance of each feature, with an accuracy of 69,9\% for CSF classification when compared to the CSF identified by Ramos \emph{et al} \cite{ramos2016information} in a qualitative analysis.
 
@@
 
-The proposed approach is evaluated and the experimental results show that the visual discriminant analysis, throught PCA, highlights important characteristcs of the data, and that a selected classifier and RFE can be applied for iGovTI classification and for CSF identification.
+The proposed approach is evaluated and the experimental results show that the visual discriminant analysis, through PCA, highlights important characteristics of the data, and that a selected classifier and RFE can be applied for iGovTI classification and for CSF identification.
 
@@
 
-Finally, SVM is used to select the more significant features identified by RFE, wich are 69,9\% similar to CSFs previsously identified by a qualitative analysis of the same datased.
+Finally, SVM is used to select the more significant features identified by RFE, which are 69,9\% similar to CSFs previously identified by a qualitative analysis of the same dataset.

### 2017-11-15 distributed-tensor-dictionary-learning/execDLMethods.m

@@ -69,7 +69,8 @@ function results = execDLMethods(L, N, K, M1, M2, N1, N2, snr, methodChar, s, no
     javaclasspath('-dynamic')
 
     %% prapare output
-    fileName = [methodChar, sprintf('%1i_%li_%li_%li_%li.mat',s ,snr ,L ,N*K, noIt)];
+    fileName = [methodChar, sprintf('_s=%1i_snr=%li_L=%li_noIt=%li_N=%li_K=%li.mat', s, snr, L, noIt, N, K)];
+    fileName = strcat(destPath,fileName);
     results = struct('beta', zeros(K, nofTrials), ...
                  'times', zeros(nofTrials,1), ...
                  'detection', zeros(nofTrials,1), ...
@@ -87,23 +88,25 @@ function results = execDLMethods(L, N, K, M1, M2, N1, N2, snr, methodChar, s, no
     %% try load previous files and verify if the execution can be avoided
     if exist(fileName,'file')
         oldFile = dir(fileName);
-        disp([scriptName,': add to results stored in ', fileName,', (created ', oldFile.date,').']);
+        disp([scriptName,': loading ', fileName,', (created ', oldFile.date,').']);
         load(fileName);
         trialsDone = results.nofTrials;
+        disp([scriptName,': ', int2str(trialsDone) ,' trials of ', int2str(nofTrials)]);
         if (nofTrials > 0)
             results.beta = [results.beta, zeros(K, nofTrials)];
             results.times = vertcat(results.times, zeros(nofTrials,1));
             results.detection = vertcat(results.detection, zeros(nofTrials,1));
-            results.nofTrials = trialsDone + nofTrials;
+            results.nofTrials = nofTrials;
         end
     else
+        disp([scriptName,': Created ', fileName]);
         trialsDone = 0;
     end
 
     %% For each trial: generate data, estimate dictionary and 
-    disp([scriptName,': noIt=',int2str(noIt),', nofTrials=',int2str(nofTrials)]);    
     timestart = now();
-    for trial = 1:nofTrials
+    nextTrial = trialsDone + 1;
+    for trial = nextTrial:nofTrials
         % logging        
         disp([scriptName,': ',method,' L=',int2str(L), ...
             ' snr=',num2str(snr), ', s=',int2str(s), ...
@@ -152,7 +155,7 @@ function results = execDLMethods(L, N, K, M1, M2, N1, N2, snr, methodChar, s, no
[…]

### 2017-11-15 distributed-tensor-dictionary-learning/projectFraudDetectionDL.m

@@ -30,14 +30,14 @@ lineformat = [repmat('%*s',1,HC) repmat('%f',1,NF)];
 train_data = cell2mat(textscan(fid, lineformat, 'HeaderLines', HL, 'Delimiter', ','));
 fclose (fid);
 train_data = train_data.';
-L = size(train_data, 2);                                                    % number of observations
-% Ks = [[10, 5, 2]; [50, 10, 5]; [100, 20, 5]; [200, 20, 10]; [500, 50, 10]]; % dictionary's atoms (K << L)
-Ks = [[10, 5, 2]; [50, 10, 5]; [100, 20, 5]];                               % dictionary's atoms (K << L)
-N = 7; N1 = 7; N2 = 1;                                                     % features/variables/components
-sparsities = [2, 5];                                                          % sparsity degrees
-noIts = [10, 50, 100];                                                      % number of iterations
-solver = 'javaORMP';                                                        % solver for sparse approximation
-methodChar = 'H';                                                           % method for RLS-DLA
+L = size(train_data, 2);                                                                    % number of observations
+% Ks = [[10, 5, 2]; [50, 10, 5]; [100, 20, 5]; [200, 20, 10]; [500, 50, 10]];               % dictionary's atoms (K << L)
+Ks = [[10, 5, 2]; [50, 10, 5]; [100, 20, 5]];                                               % dictionary's atoms (K << L)
+N = 7; N1 = 7; N2 = 1;                                                                      % features/variables/components
+sparsities = [2, 5];                                                                        % sparsity degrees
+noIts = [10, 50, 100];                                                                      % number of iterations
+solver = 'javaORMP';                                                                        % solver for sparse approximation
+methodChar = 'H';                                                                           % method for RLS-DLA
 
 %% Dictionary Learning Cross-Validation for fraud detection data
 fprintf('## Dictionary Learning from fraud detection data');

### 2017-11-27 docs/thesis/slides/apresentacao_unb.tex
slides

% qual o tempo da apresentação? apresento os apêndices? apresento os trabalhos relacionados? apresento as referências (90)?  
%	Name			:: 	sthlm Beamer Theme  HEAVILY based on the hsrmbeamer theme (Benjamin Weiss)
%	Author			:: 	Mark Hendry Olson (mark@hendryolson.com)
%	Created			::	2013-07-31
%	Updated	    	::	[[April]] 04, 2017 at 16:26:39
%	Version			:: 	2.0.2
%	Email			:: 	hendryolson@gmail.com
%	Website			:: 	http://markolson.se
%	Twitter			:: 	markolsonse
%	Instagram		:: 	markolsonse
%
%	License			:: 	This file may be distributed and/or modified under the
%					GNU Public License.
%
%	Description		::	This presentation is a demonstration of the sthlm beamer
%					theme, which is HEAVILY based on the HSRM beamer theme created by Benjamin Weiss
%					(benjamin.weiss@student.hs-rm.de), which can be found on GitHub
%					<https://github.com/hsrmbeamertheme/hsrmbeamertheme>.  It also borrows heavily
%					from the work of Matthias Vogelgesang, (https://bloerg.net) and his Metropolis Mtheme,
%					<https://github.com/matze/mtheme>.
%
%	Theme			::	newPxFont
%	Options			::	progressbar
%					::	sectionpages
%					::	numfooter
%					::	fullfooter
%					::	dovaligncolumns
%					::	protectframetitle
%					::	greybg
%					::	cblock
%					::	minimal

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%        LOADING DOCUMENT
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\documentclass[newPxFont, numfooter, sectionpages]{beamer}
\usepackage[utf8]{inputenc}
\usetheme{sthlm}
\usepackage{pgfplots}
\pgfplotsset{compat=1.9}
\usepackage{cancel}
\usepackage{amsthm,amsmath}
\usepackage{mathtools}

\usepackage{booktabs}
\usepackage{supertabular}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{color, colortbl}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\definecolor{lightgray}{gray}{0.8}
\definecolor{Gray}{gray}{0.9}

\title{Discriminative Sensing Based on Signal Processing for Information Security Analysis}
\subtitle{\small{Ph.D. Thesis Qualification}}
\date{\today}
\author{\texttt{Thiago P. de B. Vieira \newline \textbf{Advisor:} João Paulo C. L. da Costa} }
\institute{Universidade de Brasília 
	\par \scriptsize{Departamento de Engenharia Elétrica - ENE/FT}
	\par \scriptsize{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}
}

\hypersetup{
	pdfauthor = {Thiago P. de B. Vieira},
	pdfsubject = {Discriminative Sensing Based on Signal Processing for Information Security Analysis},
	pdfkeywords = {Signal Processing, Information Security Analysis},
	pdfmoddate= {/home/thiago/dev/projects/discriminative-sensing/docs/thesis/slides},
	pdfcreator = {Sublime and Latex}
}

\begin{document}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%	TITLE PAGE
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\maketitle
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Theme Package Requirements
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begingroup
\setbeamercolor{normal text}{fg=\cnDarkGrey,bg=white}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%	TABLE OF CONTENTS: OVERVIEW
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section*{Outline}
\begin{frame}{Outline}
	\tableofcontents[hideallsubsections]% For longer presentations use hideallsubsections option
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%	TABLE OF CONTENTS: OVERVIEW
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Introduction}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Motivation
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Motivation}
	It is necessary investments to detect \textbf{novel attacks and their variations}
	\begin{itemize}
		\item Signal processing schemes have being applied to detect malicious traffic;		
	\end{itemize}
	Secure mobile cloud architecture is challenging, specially in \textbf{offline mode}
	\begin{itemize}
		\item Signal processing can be applied to \textbf{behavioral analysis of mobile apps in offline mode};		
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Motivation
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Motivation}
	\textbf{Fraud detection} of \textbf{Mobile Money Transactions} has been motivating investments
	\begin{itemize}
		\item The \textbf{extraction of relevant information} from \textbf{raw or big} data is useful for classification problems and for security analysis.;
		\item Imbalanced data is challenging for classification related to anomaly detection, novelty detection, fraud detection and attack detection;
		\item \textbf{Tensor-based} algorithms for \textbf{dictionary learning} can improve \textbf{discriminative sensing} for classification in information security analysis;
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Problem Statement
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Problem Statement}
	This thesis outlines the development and evaluation of approaches based on \textbf{signal processing for information security analysis}
	\begin{itemize}
		\item To make the data discriminative;
		\item For network attack detection, mobile malicious behavior analysis and fraud detection in mobile money transactions.
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Problem Statement
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Problem Statement}
	We propose an approach based on \textbf{signal processing techniques for detection of malicious traffic in computer networks}
	\begin{itemize}
		\item Signal processing formulation as legitimate traffic, malicious traffic and noise;
		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
		\item Eigen similarity analysis for extracting detailed information about accurate time and network ports under attack;
		\item Accuracy and performance evaluation through experimental scenario and the DARPA 1998 dataset.
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Problem Statement
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Problem Statement}
	This thesis also proposes an \textbf{architecture and approach based on user behavior analysis of mobile apps}
	\begin{itemize}
		\item An architecture for mobile security analysis;
		\item Offline user behavior analysis based on MOS;
		\item Scenario analysis for feature selection;
		\item Processing time evaluation in mobile devices.
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Problem Statement
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Problem Statement}
	We also propose a \textbf{tensor-based dictionary learning method for fraud detection from imbalanced data}, in order to apply the tensor decomposition for dictionary learning methods to evidentiate the \textbf{discriminative sensing of a fraud detection dataset}
	\begin{itemize}
		\item A sparse representation based classification (SRC) method through learning a tensor-based dictionary;
		\item Apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate;
		\item Evaluate the reconstruction error for fraud classification of mobile money transactions.
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Contributions
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Contributions}
	\begin{enumerate}
		\item An approach based on \textbf{eigen similarity analysis} for extracting detailed information about accurate time and network ports under network attack;
		\item Computational \textbf{complexity analysis} of the proposed framework;
		\item An architecture and techniques for \textbf{offline behavioral analysis} of a corporate mobile client security architecture;
		\item A \textbf{tensor-based dictionary learning} approach for \textbf{fraud detection} in mobile payment transactions;
	\end{enumerate}
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%	SECTION: Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Introduction
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Introduction}
	We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise})
	\begin{itemize}
		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
		\item Eigen similarity analysis for extracting detailed time and network ports under attack. 
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Introduction
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Introduction}
	We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise})
	\begin{itemize}
		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
		\item Eigen similarity analysis for extracting detailed time and network ports under attack. 
	\end{itemize}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Related works
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Related works}

	Classical methods typically employ data mining [39, 34, 70] and regular file analysis [76] to detect patterns that indicate the presence of specific attacks in network traffic;

	Callegari \emph{et al} [104] propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data. However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Related works
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Related works}

	Lee \emph{et al.} [61] propose OverSampling PCA (osPCA), determines the anomaly of the target according to the variation of eigenvector by similarity analysis and over sampling. In contrast, we apply MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack, for flood and probe attacks;

	Lu and Ghorbani [63] proposed a network anomaly detection model based on network flow, wavelet approximation, and system identification theory. However, their work requires a training method and presents limitations on identification of behaviors without significant outliers;

	Zonglin \emph{et al} [104] proposed a method to detect traffic anomaly with correlation analysis, but the work is not applied to probe and flood attack detection.

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Data Model
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Data Model}

	Focus on \textbf{probe} and \textbf{flooding} attacks;

	Two scenarios: A \textbf{synthetic dataset} as a signal superposition of legitimate traffic, noise and malicious traffic, and selected cases of \textbf{DARPA dataset} that reproduce probe and flooding attacks.

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Data Model
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Data Model}

	Network traffic:

	\begin{equation}\label{eq:eq01}
		\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{N}^{(q)} + \boldsymbol{A}^{(q)},
	\end{equation}

	$\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ has \emph{M} rows and \emph{N} columns.

	Each row is a TCP or UDP port, and each column represents time bins (minute)

	Each element $x_{m,n}^{(q)}$ stands for the number of times that the port $m$ appears at the $n$-th minute, at the $q$-th time frame.

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Synthetic Dataset
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Synthetic Dataset}

	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=9cm]{../figures/fig09.png}
	     \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
	     \label{fig:2_fig1}
	\end{figure}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Synthetic Dataset
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Synthetic Dataset}

	\begin{figure}[h!]
	     \centering 
	     \includegraphics[width=9cm]{../figures/fig03.png}
	     \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
	     \label{fig:2_fig3}
	\end{figure}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Synthetic Dataset
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Synthetic Dataset}

	\begin{figure}[h!]
	     \centering 
	     \includegraphics[width=9cm]{../figures/fig04.png}
	     \caption{Network traffic of user independent operations for network management.}
	     \label{fig:2_fig4}
	\end{figure}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Synthetic Dataset
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Synthetic Dataset}

	\begin{figure}[h!]
	     \centering 
	     \includegraphics[width=9cm]{../figures/fig05.png}
	     \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
	     \label{fig:2_fig5}
	\end{figure}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Synthetic Dataset
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Synthetic Dataset}

	\begin{figure}[h!]
	     \centering 
	     \includegraphics[width=9cm]{../figures/fig06.png}
	     \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
	     \label{fig:2_fig6}
	\end{figure}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Synthetic Dataset
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}[c]{Synthetic Dataset}

	\begin{figure}[h!]
	     \centering 
	     \includegraphics[width=9cm]{../figures/fig07.png}
	     \caption{Connection attempts in order to identify active ports.}
	     \label{fig:2_fig7}
	\end{figure}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: DARPA Dataset
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{DARPA Dataset}

	7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks.

	The attacks in this dataset can be grouped into: denial-of-service (DoS); remote to local (R2L); user to root (U2R); and probe attack.

	The most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack.

	We select the cases that reproduce probe and flooding attacks.

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Proposed Framework
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Proposed Framework}

	\begin{figure}[h!]
		\centering
	     \includegraphics[width=7cm]{../figures/mos_eigen_similarity.eps}
	     \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
	     \label{fig:2_fig80}
	\end{figure}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Proposed Framework
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Proposed Framework}

	For flooding detection, calculate the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
	\begin{equation}\label{eq:eq02}
		\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
	\end{equation}

	the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
	\begin{equation}\label{eq:eq03}
		\boldsymbol{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
	\end{equation}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Proposed Framework
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Proposed Framework}

	For probing detection, compute the sample covariance $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have zero mean and unitary standard deviation as follows
	\begin{equation}\label{eq:eq04}
		\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
	\end{equation}

	The sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)}$ can be calculated via 
	\begin{equation}\label{eq:eq05}
		\boldsymbol{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
	\end{equation}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Proposed Framework
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Proposed Framework}

	we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{C}$

	Compute eigenvalue decomposition (EVD) to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).
	\begin{equation}\label{eq:eq06}
	\boldsymbol{C}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
	\end{equation}
	\begin{equation}\label{eq:eq060}
	\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
	\end{equation}

	The eigenvalues should be sorted in descending order, i.e., $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ... > \lambda_{m}^{(q)}$.

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Proposed Framework
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Proposed Framework}

	\begin{equation}\label{eq:eq07}
		\boldsymbol{E} =
		\begin{bmatrix}
			\lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
			\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
			\lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
			\vdots & \vdots & \ddots & \vdots  \\
			\lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
		\end{bmatrix}.
	\end{equation}

	\begin{equation}\label{eq:eq08}
		\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
	\end{equation}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: MOS Scheme
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{MOS Scheme}

	Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$.

	The proposed approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame to identify variations and estimate the model order $\hat{d}$ (estimated number of time frames under attack)

	$\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to

	$\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Eigen Similarity Analysis
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Eigen Similarity Analysis}

	\begin{figure}[h!]
	     \centering 
	     \includegraphics[width=11cm]{alg.png}
	     \label{fig:2_fig9}
	\end{figure}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Eigenvalue Analysis
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Eigen Similarity Analysis}

	For eigen similarity analysis, we evaluate the cosine similarity to identify legitimate and malicious traffic,
	\begin{equation}
		\label{eq:eq11}
		s_n = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}}},
	\end{equation}

	$s_n$ denotes the absolute similarity degree of the $n$-th minute

	$\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack

	$\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of traffic to be performed the flood and port scan attack identification.

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Eigenvalue Analysis
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Eigen Similarity Analysis}

	The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic without attack. Each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$, as represented by
	\begin{equation}\label{eq:eq12}
		\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
	\end{equation}

	The resultant $\boldsymbol{X}_{(n)}$ is necessary to obtain $\boldsymbol{v}_{(n)}$, through (\ref{eq:eq06}).

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Eigenvalue Analysis
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Eigen Similarity Analysis}

	We evaluate three approaches for eigen similarity analysis: 
	\begin{itemize}
		\item \textbf{incremental:} concatenates eigenvectors incrementally;
		\item \textbf{individual:} concatenates each eigenvector individually and discards;
		\item \textbf{incremental individualized:} incremental until detect attach and individual after.
	\end{itemize}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Eigenvalue Analysis
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Eigen Similarity Analysis}

	\begin{figure}[h!]
		\centering
	    \includegraphics[width=11.5cm]{../figures/incremental.eps}
	    \caption{Traffic selection for incremental approach.}
	    \label{fig:2_fig8}
	\end{figure}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Eigenvalue Analysis
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Eigen Similarity Analysis}

	\begin{figure}[h!]
	     \includegraphics[width=11.5cm]{../figures/individualized.eps}
	     \caption{Traffic selection for individual approach.}
	     \label{fig:2_fig9}
	\end{figure}
\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Eigenvalue Analysis
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Eigen Similarity Analysis}

	\begin{figure}[h!]
	     \includegraphics[width=11.5cm]{../figures/incremental_individualized.eps}
	     \caption{Traffic selection for incremental individualized approach.}
	     \label{fig:2_fig2}
	\end{figure}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Eigenvalue Analysis
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Eigen Similarity Analysis}

	For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack is compared against the $\boldsymbol{v}_{(n)}$ identified as under attack, individually evaluating the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.

	\begin{equation}\label{eq:eq15}
		\left\{
		\begin{array}{@{}ll@{}}
			x_{(m,n)} = x^{(\hat{q})}_{(m,\hat{n})} \\
			\\
			s_{m,\hat{n}} = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(m,\hat{n})}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(m,\hat{n})}}},
		\end{array}\right.
	\end{equation}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Results
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Results}

	\begin{table}[h!]
	  \centering
	  \scriptsize
	  \caption{Largest Eigenvalue related to attacks detection}
	  \label{tab:tab3}
	  \begin{tabular}{ c c c c c }
		\toprule
		\multirow{3}{*}{\textbf{Time Frame} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
				\hhline{~----}
			&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
			&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
		\midrule
		1 &1887545 &1887545 &1887545 &2,0734 \\
		2 &2341327 &2341327 &2341327 &2,1451 \\
		3 &3213867 &3213867 &3213867 &10,0718 \\
		4 &133238294 &133238294 &731229 &2,1620 \\
		5 &92384021611 &6367983 &92384021611 &2,4253 \\
		6 &708335 &708335 &708335 &1,7948 \\
	    \bottomrule
	  \end{tabular}
	\end{table}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Results
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Results}

	\begin{table}[h!]
	  \centering
	  % \scriptsize
	  \caption{MOS schemes applied to port scan and flood detection}
	  \label{tab:tab4}
	  \begin{tabular}{ c c c c c c c c }
		\toprule
		\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
				\hhline{~------~}
			&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
		\midrule
		Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
		Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
		\midrule
		Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
		Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
		\midrule
		Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
		Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
		\midrule
		Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
		Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
	    \bottomrule
	  \end{tabular}
	\end{table}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Results
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Results}

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Port Scan Detection}
  \label{tab:tab5}
  \begin{tabular}{ c c c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~~---~}
			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
	\midrule
	3 &1 &0.9946 &0.9946 &0.9946 &no \\
	3 &2 &0.9934 &0.9934 &0.9999 &no \\
	3 &3 &0.9912 &0.9912 &0.9999 &no \\
	3 &4 &0.9888 &0.9888 &0.9999 &no \\
	3 &5 &0.9856 &0.9856 &0.9998 &no \\
	3 &6 &0.9840 &0.9840 &0.9999 &no \\
	3 &7 &0.9824 &0.9824 &1.0000 &no \\
	3 &8 &0.9794 &0.9794 &0.9999 &no \\
	3 &9 &0.9673 &0.9673 &0.9926 &no \\
	3 &10 &0.9674 &0.9674 &0.9997 &no \\
	3 &11 &0.9733 &0.9733 &0.9993 &no \\
	3 &12 &0.9702 &0.9702 &0.9993 &no \\
	3 &13 &0.9677 &0.9677 &0.9999 &no \\
	3 &14 &0.9646 &0.9646 &0.9998 &no \\
	3 &15 &0.0216 &0.0216 &0.0276 &yes \\
	3 &16 &0.9621 &0.0209 &1.0000 &no \\
	3 &17 &0.9611 &0.0199 &0.9998 &no \\
	3 &18 &0.9612 &0.0191 &0.9999 &no \\
	3 &19 &0.9613 &0.0186 &0.9998 &no \\
	3 &20 &0.9638 &0.0190 &1.0000 &no \\
    \bottomrule
  \end{tabular}
\end{table}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Results
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Results}

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Synflood Detection}
  \label{tab:tab6}
  \begin{tabular}{ c c c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~~---~}
			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
	\midrule
	4 &1 &1.0000 &1.0000 &1.0000 &no \\
	4 &2 &0.9999 &0.9999 &1.0000 &no \\
	4 &3 &0.9997 &0.9997 &0.9999 &no \\
	4 &4 &0.9998 &0.9998 &1.0000 &no \\
	4 &5 &0.9965 &0.9965 &0.9908 &no \\
	4 &6 &0.9975 &0.9975 &1.0000 &no \\
	4 &7 &0.9977 &0.9977 &1.0000 &no \\
	4 &8 &0.9980 &0.9980 &1.0000 &no \\
	4 &9 &0.9987 &0.9987 &0.9999 &no \\
	4 &10 &0.9991 &0.9991 &1.0000 &no \\
	4 &11 &0.0085 &0.0085 &0.0284 &yes \\
	4 &12 &0.0162 &0.0120 &0.0343 &yes \\
	4 &13 &0.0248 &0.0158 &0.0427 &yes \\
	4 &14 &0.1243 &0.0185 &0.1041 &yes \\
	4 &15 &0.0082 &0.0162 &0.0103 &yes \\
	4 &16 &0.0404 &0.0070 &0.0580 &yes \\
	4 &17 &0.0397 &0.0007 &0.0573 &yes \\
	4 &18 &0.0408 &0.0042 &0.0584 &yes \\
	4 &19 &0.0408 &0.0079 &0.0584 &yes \\
	4 &20 &0.0477 &0.0092 &0.0757 &yes \\
    \bottomrule
  \end{tabular}
\end{table}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Results
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Results}

\begin{table}[h!]
  \centering
  \footnotesize
  \caption{Eigen Similarity Analysis for Fraggle Detection}
  \label{tab:tab7}
  \begin{tabular}{ c c c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~~---~}
			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
	\midrule
	5 &1 &1.0000 &1.0000 &1.0000 &no \\
	5 &2 &0.9999 &0.9999 &1.0000 &no \\
	5 &3 &1.0000 &1.0000 &1.0000 &no \\
	5 &4 &0.9999 &0.9999 &1.0000 &no \\
	5 &5 &0.9993 &0.9993 &0.9997 &no \\
	5 &6 &0.9993 &0.9993 &0.9997 &no \\
	5 &7 &0.9994 &0.9994 &1.0000 &no \\
	5 &8 &0.9995 &0.9995 &1.0000 &no \\
	5 &9 &0.9995 &0.9995 &1.0000 &no \\
	5 &10 &0.9995 &0.9995 &1.0000 &no \\
	5 &11 &0.0031 &0.0031 &0.0021 &yes \\
	5 &12 &0.0019 &0.0025 &0.0009 &yes \\
	5 &13 &0.0030 &0.0026 &0.0020 &yes \\
	5 &14 &0.0030 &0.0027 &0.0020 &yes \\
	5 &15 &0.0030 &0.0028 &0.0020 &yes \\
	5 &16 &0.0012 &0.0025 &0.0002 &yes \\
	5 &17 &0.0030 &0.0026 &0.0020 &yes \\
	5 &18 &0.0030 &0.0026 &0.0020 &yes \\
	5 &19 &0.0030 &0.0027 &0.0020 &yes \\
	5 &20 &0.0069 &0.0023 &0.0083 &yes \\
    \bottomrule
  \end{tabular}
\end{table}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Results
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Results}

\begin{table}[h!]
  \centering
  % \footnotesize
  \caption{Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n=15)}
  \label{tab:tab8}
  \begin{tabular}{ c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~--~}
			&\textbf{Incremental Individualized} &\textbf{Individual}\\
	\midrule
	80 &0.9999 &0.9999 &no \\
	443 &0.9999 &0.9999 &no \\
	53 &0.9999 &0.9999 &no \\
	21 &0.9999 &0.9997 &yes \\
	22 &0.0298 &0.9997 &yes \\
	23 &0.0298 &0.9997 &yes \\
	25 &0.0298 &0.9997 &yes \\
	110 &0.0298 &0.9997 &yes \\
	143 &0.0298 &0.9997 &yes \\
	161 &0.0298 &0.9997 &yes \\
	69 &0.0298 &0.9997 &yes \\
	123 &0.0298 &0.9997 &yes \\
	445 &0.0298 &0.9997 &yes \\
	600 &0.9999 &0.9999 &no \\
	19 &0.9999 &0.9999 &no \\
	67 &0.9999 &0.9999 &no \\
	68 &0.9999 &0.9999 &no \\
    \bottomrule
  \end{tabular}
\end{table}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Results
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Results}

\begin{table}[h!]
  \centering
  % \footnotesize
  \caption{Eigen Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
  \label{tab:tab9}
  \begin{tabular}{ c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~--~}
			&\textbf{Incremental Individualized} &\textbf{Individual}\\
	\midrule
	80 &1.0000 &1.0000 &no \\
	443 &1.0000 &1.0000 &no \\
	53 &1.0000 &1.0000 &no \\
	21 &1.0000 &1.0000 &nos \\
	22 &1.0000 &1.0000 &no \\
	23 &1.0000 &1.0000 &no \\
	25 &1.0000 &1.0000 &no \\
	110 &1.0000 &1.0000 &no \\
	143 &1.0000 &1.0000 &no \\
	161 &1.0000 &1.0000 &no \\
	69 &1.0000 &1.0000 &no \\
	123 &1.0000 &1.0000 &no \\
	445 &1.0000 &1.0000 &no \\
	600 &0.0077 &0.0427 &yes \\
	19 &1.0000 &1.0000 &no \\
	67 &1.0000 &1.0000 &no \\
	68 &1.0000 &1.0000 &no \\
    \bottomrule
  \end{tabular}
\end{table}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Results
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Results}

\begin{table}[h!]
  \centering
  % \footnotesize
  \caption{Eigen Similarity Analysis for Detection of Ports Under Fraggle Attack (q=5 and t=11)}
  \label{tab:tab10}
  \begin{tabular}{ c c c c }
	\toprule
	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
			\hhline{~--~}
			&\textbf{Incremental Individualized} &\textbf{Individual}\\
	\midrule
	80 &1.0000 &1.0000 &no \\
	443 &1.0000 &1.0000 &no \\
	53 &1.0000 &1.0000 &no \\
	21 &1.0000 &1.0000 &no \\
	22 &1.0000 &1.0000 &no \\
	23 &1.0000 &1.0000 &no \\
	25 &1.0000 &1.0000 &no \\
	110 &1.0000 &1.0000 &no \\
	143 &1.0000 &1.0000 &no \\
	161 &1.0000 &1.0000 &no \\
	69 &1.0000 &1.0000 &no \\
	123 &1.0000 &1.0000 &no \\
	445 &1.0000 &1.0000 &no \\
	600 &1.0000 &1.0000 &no \\
	19 &0.0031 &0.0004 &yes \\
	67 &1.0000 &1.0000 &no \\
	68 &1.0000 &1.0000 &no \\
    \bottomrule
  \end{tabular}
\end{table}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Results
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Results}

\begin{table}[!t]
	\caption{Results of the attack detection evaluation}
	\label{tab:tab12}
	\centering
	\begin{tabular}{|c|c|c|c|c|}
		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Solution\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Attack Type\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Metric\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Result\end{tabular} \\ \hline
		Proposed Work	&Flooding	&True Positive	&100.00 \%\\ \hline
		Proposed Work	&Flooding	&False Positive	&60.00 \%\\ \hline
		Proposed Work	&Flooding	&Misclassification	&50.00 \%\\ \hline
		Proposed Work	&Probe	&True Positive	&76.92 \%\\ \hline
		Proposed Work	&Probe	&False Positive	&18.52 \%\\ \hline
		Proposed Work	&Probe	&Misclassification	&32.73 \%\\ \hline
		Callegari \emph{et al}	&Flooding	&True Positive	&82.00 \%\\ \hline
		Callegari \emph{et al}	&Flooding	&False Positive	&-\\ \hline
		Callegari \emph{et al}	&Flooding	&Misclassification	&-\\ \hline
		Lu and Ghorbani	&Overall	&True Positive	&94.67 \%\\ \hline
		Lu and Ghorbani	&Overall	&False Positive	&-\\ \hline
		Lu and Ghorbani	&Overall	&Misclassification	&-\\ \hline
		Lu and Ghorbani	&Portsweep	&True Positive	&50.00 \%\\ \hline
		Lu and Ghorbani	&Portsweep	&False Positive	&-\\ \hline
		Lu and Ghorbani	&Portsweep	&Misclassification	&-\\ \hline
	\end{tabular}
\end{table}

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Performance Evaluation
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Performance Evaluation}

	complexity

\end{frame}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Performance Evaluation
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begin{frame}{Performance Evaluation}

	processing

\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	SECTION: Offline Mode for Corporate Mobile Client Security Architecture
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Offline Mode for Corporate Mobile Client Security Architecture}
\label{Blocks}
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Blocks
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}{Blocks}

\begin{block}{Block Title Here}
		Great for definitions
\end{block}

\begin{alertblock}{Alert Title Here}
		Great for definitions
\end{alertblock}

\begin{exampleblock}{Example Title Here}
		Great for examples
\end{exampleblock}
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Blocks
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}{Blocks}

\begin{block}{Block Title Here}
	\begin{itemize}
		\item point 1
		\item point 2
	\end{itemize}
\end{block}

\begingroup
\setbeamercolor{block title}{fg=white, bg=\cnBlue}
\setbeamercolor{block body}{bg=\cnLightBlue}
\begin{block}{Blue Colored Blocks}
	Produced by using the \texttt{cblock} theme option
\end{block}
\endgroup
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Additional Blocks
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}{Additional Blocks}
\begin{alertblock}{Alert Block}
	Highlight important information.
\end{alertblock}

\begingroup
\setbeamercolor{block title}{fg=white, bg=\cnRed}
\setbeamercolor{block body}{bg=\cnLightRed}
\begin{block}{Red Colored Blocks}
	Produced by using the \texttt{cblock} theme option
\end{block}
\endgroup
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Additional Blocks
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}{Additional Blocks}

\begin{exampleblock}{Example Block}
	Examples can be good.
\end{exampleblock}

\begingroup
\setbeamercolor{block title}{fg=white, bg=\cnGreen}
\setbeamercolor{block body}{bg=\cnLightGreen}
\begin{block}{Green Colored Blocks}
	Produced by using the \texttt{cblock} theme option
\end{block}
\endgroup

\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Blocks
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}{Custom Blocks}
\begingroup
\setbeamercolor{block title}{fg=white, bg=\cnPurple}
\setbeamercolor{block body}{bg=\cnLightPurple}
\begin{block}{Purple customization}
	Using the theme colors to generate colored blocks.
\end{block}
\endgroup

\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%	SECTION: Tensor-based Discriminative Sensing for Fraud Detection
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Tensor-based Discriminative Sensing for Fraud Detection}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: 
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}[c]{No Special Fonts Required}

This theme was originally made to work with \texttt{pdflatex} and the default latex fonts.\\
\vspace{1em}
\cRed{sthlm} does comes with a \texttt{pdflatex} font option, \cRed{newPxFont}, which loads the following fonts:

\begin{itemize}
	\item \emph{newpxtext} for text
	\item \emph{cantarell} for sans-serif
	\item \emph{inconsolata} for sans-serif monospaced
	\item \emph{newpxmath} for math
\end{itemize}

Please refer to the \texttt{beamerfontthememsthlm.sty} for the package requirements.

\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%	SECTION: Conclusion and Future Work
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\section{Conclusion and Future Work}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Tables
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}{Tables}
\begin{table}[]
	\caption{Selection of window function and their properties}
	\begin{tabular}[]{lrrr}
		\toprule
		\textbf{Window}			& \multicolumn{1}{c}{\textbf{First side lobe}}
		                    & \multicolumn{1}{c}{\textbf{3\,dB bandwidth}}
		                    & \multicolumn{1}{c}{\textbf{Roll-off}} \\
		\midrule
		Rectangular				& 13.2\,dB	& 0.886\,Hz/bin	& 6\,dB/oct		\\[0.25em]
		Triangular				& 26.4\,dB	& 1.276\,Hz/bin	& 12\,dB/oct	\\[0.25em]
		Hann					& 31.0\,dB	& 1.442\,Hz/bin	& 18\,dB/oct	\\[0.25em]
		Hamming					& 41.0\,dB	& 1.300\,Hz/bin	& 6\,dB/oct		\\
		\bottomrule
	\end{tabular}
	\label{tab:WindowFunctions}
\end{table}
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME:
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}[c]{Mathematics Step by Step}
Show $[x^n]'=nx^{n-1}$ by using first principles.
\begin{align*}
\visible<+->{f'(x) & = \displaystyle \lim_{\Delta x \to 0} \frac{f(x+\Delta x)-f(x)}{\Delta x} \\}
\visible<+->{&= \displaystyle \lim_{\Delta x \to 0} \frac{(\cRed{x+\Delta x})^n - (\cRed{x})^n}{\Delta x}\\}
\visible<+->{&= \displaystyle \lim_{\Delta x \to 0} \frac{\binom{n}{0}x^n\Delta x^0 +\binom{n}{1}x^{n-1}\Delta x^1 + \cdots + \binom{n}{n}x^0\Delta x^n - \cRed{x^n} }{\Delta x}\\}
\visible<+->{&= \displaystyle \lim_{\Delta x \to 0} \frac{\cRed{1}x^n(\cRed{1}) +\cRed{n}x^{n-1}\Delta x^1 + \cdots + \cRed{1}(\cRed{1})\Delta x^n - \cRed{x^n} }{\Delta x}\\}
\visible<+->{&= \displaystyle \lim_{\Delta x \to 0} \frac{\cRed{x^n} +nx^{n-1}\Delta x + \cdots + \Delta x^n - \cRed{x^n} }{\Delta x}\\}
\visible<+->{&= \displaystyle \lim_{\Delta x \to 0} \frac{\cancel{\cRed{\Delta x}} (nx^{n-1} + \cdots + \Delta x^{n-1} ) }{\cancel{\Delta x}}\\}
\visible<+->{&=nx^{n-1}}
\end{align*}

\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Formulas
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}{Functions}
\begin{block}{Gaussian Probability Density Function}
\[
f \left(x \mid \mu, \sigma^2 \right) = \dfrac{1}{\sqrt{2 \sigma^2 \pi}} e^{- \dfrac{(x-\mu)^2}{2\sigma^2}}
\]
\end{block}
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: 
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}[c]{More Mathematics}

\begin{columns}[c]
\begin{column}{0.5\textwidth}
\begin{tikzpicture}[scale=0.7]
	\begin{axis}[
            domain=0:2, range=0:6,ymax=6,ymin=0,
            axis lines =left,
            every axis y label/.style={at=(current axis.above origin),anchor=south},
            every axis x label/.style={at=(current axis.right of origin),anchor=west},
            xtick={1,1.666}, ytick={1,3},
            xticklabels={$\cPurple{x}$,$\cPurple{x}+\cRed{\Delta x}$}, yticklabels={$f(\cPurple{x})$,$f(\cPurple{x}+\cRed{\Delta x})$},
            axis on top,
          ]
          \addplot [very thick,\cnRed, domain=(0:2)] {-1.994+2.994*x};S
	  	  \addplot [\cnDarkGrey,dashed] plot coordinates {(1,0) (1,1)};
          \addplot [\cnDarkGrey,dashed] plot coordinates {(0,1) (1,1)};
          \addplot [\cnDarkGrey,dashed] plot coordinates {(0,3) (1.666,3)};
          \addplot [\cnDarkGrey,dashed] plot coordinates {(1.666,0) (1.666,3)};
          \addplot [very thick,\cnBlue, smooth,domain=(0:1.833)] {-1/(x-2)};
          \addplot[color=\cnDarkGrey,fill=\cnDarkGrey,only marks,mark=*] coordinates{(1.666,3)};  %% closed hole
          \addplot[color=\cnDarkGrey,fill=\cnDarkGrey,only marks,mark=*] coordinates{(1,1)};  %% closed hole
        \end{axis}
\end{tikzpicture}
\end{column}

\begin{column}{0.5\textwidth}
\begin{align*}
m &= \frac{\Delta y}{\Delta x}	\\
\\
m & =  \frac{\visible<2->{f(\cPurple{x}+\cRed{\Delta x})}\visible<3->{-}\visible<4->{f(\cPurple{x})}}{\visible<5->{\cPurple{x}+\cRed{\Delta x}}\visible<6->{-}\visible<7->{\cPurple{x}}}\\
m  &=  \underbrace{\frac{\visible<8->{f(\cPurple{x}+\cRed{\Delta x})-f(\cPurple{x})}}{\visible<9->{\cRed{\Delta x}}}}_{\visible<10->{\text{\cRed{difference quotient}}}}
\end{align*}
\end{column}
\end{columns}

\visible<10->{The slope of the secant line }
\begin{itemize}
	\item<10-> can be found using the \cRed{difference quotient}
	\item<11-> represents a \cBlue{function}'s \cRed{average} slope on the interval $[\cPurple{x}, \cRed{x+\Delta x}]$
\end{itemize}

\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME:
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}[c, fragile]{Fragile Frames Not a Problem}
\begin{tikzpicture}
	\begin{axis}[
            domain=0:6, range=0:7,
            ymin=-.2,ymax=7,
            width=\textwidth,
            height=7cm, %% Hard coded height! Moreover this effects the aspect ratio of the zoom--sort of BAD
            axis lines=none,
          ]
          \only<3->{\addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(.8,1.6) (2.834,5)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(2.834,5) (4.166,5)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=white] plot coordinates {(1.2,1.6) (4.166,5)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=white] plot coordinates {(.8,1.6) (1.2,1.6)} \closedcycle; %% zoom fill
		  }
          \only<4->{\addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(3.3,3.6) (5.334,5)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(5.334,5) (6.666,5)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=white] plot coordinates {(3.7,3.6) (6.666,5)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=white] plot coordinates {(3.3,3.6) (3.7,3.6)} \closedcycle; %% zoom fill
  }
          \only<4->{\addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(3.7,2.4) (6.666,1)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(3.3,2.4) (3.7,2.4)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=white] plot coordinates {(3.3,2.4) (5.334,1)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=white] plot coordinates {(5.334,1) (6.666,1)} \closedcycle; %% zoom fill
 }

          \only<3->{\addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(.8,.4) (2.834,1)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(2.834,1) (4.166,1)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=white] plot coordinates {(1.2,.4) (4.166,1)} \closedcycle; %% zoom fill
          \addplot [draw=none, fill=white] plot coordinates {(.8,.4) (1.2,.4)} \closedcycle; %% zoom fill
}
          \addplot[very thick,\cnRed, smooth,domain=(0:1.833)] {-1/(x-2)};
          \only<3->{\addplot[very thick,\cnRed, smooth,domain=(2.834:4.166)] {3.333/(2.050-.3*x)-0.333}; %% 2.5 to 4.333
          %\addplot[very thick,\cnRed, smooth,domain=(5.334:6.666)] {11.11/(1.540-.09*x)-8.109}; %% 5 to 6.833
          \only<4->\addplot[very thick,\cnRed, smooth,domain=(5.334:6.666)] {x-3}; %% 5 to 6.833
 }
         \only<2->{ \addplot[color=\cnRed,fill=\cnRed,only marks,mark=*] coordinates{(1,1)};}  %% point to be zoomed
          \only<3->{\addplot[color=\cnRed,fill=\cnRed,only marks,mark=*] coordinates{(3.5,3)}; } %% zoomed pt 1
          \only<4->{\addplot[color=\cnRed,fill=\cnRed,only marks,mark=*] coordinates{(6,3)}; } %% zoomed pt 2

          \addplot [->,\cnDarkGrey] plot coordinates {(0,0) (0,6)}; %% axis
          \addplot [->,\cnDarkGrey] plot coordinates {(0,0) (2,0)}; %% axis

          \only<2->{\addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(.8,.4) (.8,1.6)}; %% box around pt
          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(1.2,.4) (1.2,1.6)}; %% box around pt
          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(.8,1.6) (1.2,1.6)}; %% box around pt
          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(.8,.4) (1.2,.4)}; %% box around pt
 }
          \only<3->{\addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(2.834,1) (2.834,5)}; %% zoomed box 1
          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(4.166,1) (4.166,5)}; %% zoomed box 1
          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(2.834,1) (4.166,1)}; %% zoomed box 1
          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(2.834,5) (4.166,5)}; %% zoomed box 1
}
          \only<4->{\addplot [\cnDarkGrey] plot coordinates {(3.3,2.4) (3.3,3.6)}; %% box around zoomed pt
          \addplot [\cnDarkGrey] plot coordinates {(3.7,2.4) (3.7,3.6)}; %% box around zoomed pt
          \addplot [\cnDarkGrey] plot coordinates {(3.3,3.6) (3.7,3.6)}; %% box around zoomed pt
          \addplot [\cnDarkGrey] plot coordinates {(3.3,2.4) (3.7,2.4)}; %% box around zoomed pt
}
          \only<4->{\addplot [\cnDarkGrey] plot coordinates {(5.334,1) (5.334,5)}; %% zoomed box 2
          \addplot [\cnDarkGrey] plot coordinates {(6.666,1) (6.666,5)}; %% zoomed box 2
          \addplot [\cnDarkGrey] plot coordinates {(5.334,1) (6.666,1)}; %% zoomed box 2
          \addplot [\cnDarkGrey] plot coordinates {(5.334,5) (6.666,5)}; %% zoomed box 2
}
          \node at (axis cs:2.4,0) [anchor=east] {$x$};
          \node at (axis cs:0,6.9) [anchor=north] {$y$};
        \end{axis}
\end{tikzpicture}
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: PGFPlots
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}{PGFPlots Example}
	\begin{figure}[h]
		\centering
		\begin{tikzpicture}
		\begin{axis}[
		    ybar,
		    enlarge x limits=0.15,
		    legend style={at={(-.5,0.5)},
		      anchor=north,legend columns=1},
		    ylabel={\% students},
		    symbolic x coords={A, B, C, D, E, F, DNF},
		    xtick=data,
		     bar width=2mm,
		     width=0.7\textwidth
		    ]
		    \legend{2012, 2013};
		    % Spring 2012 results
			\addplot[fill=\cnGrey]  coordinates {(A,0) (B,0) (C,3.85) (D,23.07) (E,43.31) (F,30.77) (DNF,0.00)};
			% Spring 2013 results
			\addplot[fill=\cnBlue]   coordinates {(A,0) (B,3.70) (C,22.22) (D,22.22) (E,40.74) (F,3.70) (DNF,7.41)};
		\end{axis}
		\end{tikzpicture}
		\caption{Consistent improvement over the last year}
		\end{figure}
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Multiple Columns
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}{Multiple Columns}
\begin{columns}
\begin{column}{.48\linewidth}
		Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
		tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam.
\end{column}
\begin{column}{.48\linewidth}
		\begin{itemize}
        	\item Point 1
			\begin{itemize}
				\item Sub point a
				\item Sub point b
			\end{itemize}
        	\item Point 2
		\end{itemize}
	\end{column}
	\end{columns}
\end{frame}

\begin{frame}{References}
	\begin{thebibliography}{10}

	\beamertemplatebookbibitems
	\bibitem{Oppenheim2009}
	Alan V. Oppenheim
	\newblock Discrete - Time Signal Processing
	\newblock Prentice Hall Press, 2009

	\beamertemplatearticlebibitems
	\bibitem{EBU2011}
	European Broadcasting Union
	\newblock Specification of the Broadcast Wave Format (BWF)
	\newblock 2011

  \end{thebibliography}
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%	SECTION: Conclusion
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\begin{frame}{About}

This sthlm beamer theme is free software: you can redistribute it and/or modify
it under the terms of the GNU General Public License as published by
the Free Software Foundation, either version 3 of the License, or
(at your option) any later version.\\
\vspace{1cm}
If you have any questions or comments
\begin{itemize}
	\item Website: \cBlue{markolson.se}
	\item Twitter: \cBlue{@markolsonse}
	\item Instagram: \cBlue{@markolson.se}
\end{itemize}
\end{frame}

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME:
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begingroup
\setbeamercolor{background canvas}{bg=\cnDarkGrey}
\begin{frame}[plain]

\centering{\cGrey{\Huge{THE \newline END}}}

\end{frame}
\endgroup

\end{document}

### 2017-11-27 docs/thesis/slides/apresentacao_unb.tex
slides

@@
 	  \centering
-	  % \scriptsize
+	  \tiny
 	  \caption{MOS schemes applied to port scan and flood detection}
@@
 
+	MOS schemes for probing and flooding attack detection:
+	\begin{itemize}
+		\item The largest eigenvalues visually indicate the time frame under probing and flooding attack detection, but MOS schemes can estimate it mathematically;
+		\item EDC and EFT estimate correctly the number of attacks $d$;
+		\item EDC is the MOS scheme that requires less processing capacity.
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Results}
+
 	\begin{table}[h!]
 	  \centering
-  \footnotesize
+	  \tiny
 	  \caption{Eigen Similarity Analysis for Port Scan Detection}
@@
 	  \centering
-  \footnotesize
+	  \tiny
 	  \caption{Eigen Similarity Analysis for Synflood Detection}
@@
 	  \centering
-  \footnotesize
+	  \tiny
 	  \caption{Eigen Similarity Analysis for Fraggle Detection}
@@
 	  \centering
-  % \footnotesize
+	  \tiny
 	  \caption{Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n=15)}
@@
 	  \centering
-  % \footnotesize
+	  \tiny
 	  \caption{Eigen Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
@@
 	  \centering
-  % \footnotesize
+	  \tiny
 	  \caption{Eigen Similarity Analysis for Detection of Ports Under Fraggle Attack (q=5 and t=11)}
@@
 
+	For time attack detection:
+	\begin{itemize}
+		\item High similarity between network traffic without attack (0.9610) and low similarity with attack (0.0276);
+		\item The incremental approach produces false positive results;
+		\item Capability of change detection based on similarity between legitimate and malicious traffic (flood or probe);
+	\end{itemize}
+
+	For port attack detection:
+	\begin{itemize}
+		\item The incremental individualized has more sensibility to anomaly detection;
+		\item The individual approach was not able to identify low similarity for ports under attack;
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Results}
+
+	Conclusion:
+	\begin{itemize}
+		\item \textbf{Incremental individualized can detect} low similarity for all evaluated network attacks, while the \textbf{other approaches presented false positives or low sensibility};
+		\item This approach is able to \textbf{gradually and incrementally adapt to network traffic changing}.
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results - DARPA Dataset
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Results - DARPA Dataset}
+
 	\begin{table}[!t]
@@
 		\centering
+		\scriptsize
 		\begin{tabular}{|c|c|c|c|c|}
@@
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results - DARPA Dataset
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Results - DARPA Dataset}
+
+	Misclassification is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$;
+
+	High FP and misclassification due to legitimate traffic be massive (such as an attack) sometimes;
+
+	Callegari \emph{et al} is a statistical method, based on PCA, without training or learning methods
+	\begin{itemize}
+		\item 82 \% while we obtain 100 \%;
+		\item FP and misclassification was not evaluated.
+	\end{itemize}
+
+	Ghorbani's is based on signal processing techniques and uses DARPA
+	\begin{itemize}
+		\item 94.67 \% and 50.00 \%, while we obtain 100 \% and 76.92 \%;
+		\item FP and misclassification was not evaluated.
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
 %	FRAME: Performance Evaluation
@@
 
-	complexity
+	Complexity Analysis:
+	\begin{itemize}
+		\item the proposed framework is $O(N^3 + Q \log Q + \hat{d}Q + N^3)$ and its worst-case running time is $O(N^3)$;
+		\item The computational complexity of EVD is predominant in the framework;
+		\item However, the approach splits the data into time frames with period time $N$, which makes possible to limit the growth of $N$ even for evaluations of cases with total time larger than $N$;
+		\item It reduces the impact caused by the computational complexity of EVD.
+	\end{itemize}
 
@@
 
-	processing
+	Processing Time Analysis:
+	\begin{itemize}
+		\item desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM;
+		\item variations on the network traffic time; 
+		\item the frame size denoted as $N$; 
+		\item the number of network ports denoted as $M$; 
+		\item the mean processing time for eigen analysis based on sample covariance of zero mean (1-time); 
+		\item the mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation (2-time); 
+		\item the mean processing time for EDC MOS scheme, (3-time).
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Performance Evaluation
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Performance Evaluation}
+
+	\begin{table}[!t]
+		\caption{Processing time of the main steps for anomaly detection}
+		\label{tab:11}
+		\centering
+		\tiny
+		\begin{tabular}{|r|r|r|r|r|r|r|}
+			\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Traffic Time\\(hour)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Frame Size\\(min)\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Num.
+Ports\end{tabular}	& \begin{tabular}[x]{@{}l@{}}1-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}2-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}3-time\\(ms)\end{tabular}	\\ \hline
+			16	& 10	& 17	& 0.7900	& 0.8100	& 0.0650\\ \hline
+			16	& 20	& 17	& 0.5250	& 0.5950	& 0.0100\\ \hline
+			16	& 60	& 17	& 0.9700	& 1.1400	& 0.0250\\ \hline
+			16	& 120	& 17	& 0.6050	& 0.6100	& 0.0050\\ \hline
+			16	& 60	& 34	& 1.2750	& 1.2200	& 0.0050\\ \hline
+			16	& 120	& 34	& 1.1200	& 1.1700	& 0.0050\\ \hline
+			20	& 10	& 17	& 2.7950	& 2.8950	& 1.1000\\ \hline
+			20	& 20	& 17	& 2.0700	& 2.0200	& 0.3500\\ \hline
+			20	& 60	& 17	& 1.0250	& 1.0450	& 0.0650\\ \hline
+			20	& 120	& 17	& 1.0000	& 1.0700	& 0.0350\\ \hline
+			20	& 60	& 34	& 2.9650	& 3.2100	& 0.0400\\ \hline
+			20	& 120	& 34	& 2.9950	& 3.1150	& 0.0200\\ \hline
+			22	& 10	& 17	& 4.7250	& 4.0850	& 1.4600\\ \hline
+			22	& 20	& 17	& 2.3200	& 2.6800	& 0.2450\\ \hline
+			22	& 60	& 17	& 1.0700	& 1.1200	& 0.0300\\ \hline
+			22	& 120	& 17	& 0.9900	& 1.0500	& 0.0250\\ \hline
+			22	& 60	& 34	& 3.0850	& 3.1250	& 0.0650\\ \hline
+			22	& 120	& 34	& 2.8100	& 2.9600	& 0.0250\\ \hline
+		\end{tabular}
+	\end{table}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Performance Evaluation
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Performance Evaluation}
+
+	Processing Time Analysis:
+	\begin{itemize}
+		\item The processing time increases according to the increment in traffic time, around 2 or 3 times for 1-time and 2-time, but the worst measured processing time is 4.7250 milliseconds;
+		\item the processing time increases with the frame size $N$ decreasing;
+		\item The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations.
+	\end{itemize}
 

### 2017-11-27 docs/thesis/slides/apresentacao_unb.tex
eigen similarity

--- 2017-11-27 docs/thesis/slides/apresentacao_unb.tex
@@
 \begin{frame}[c]{Problem Statement}
-	This thesis outlines the development and evaluation of approaches based on \textbf{signal processing for information security analysis}
-	\begin{itemize}
-		\item To make the data discriminative;
-		\item For network attack detection, mobile malicious behavior analysis and fraud detection in mobile money transactions.
+	\textbf{This thesis} outlines the development and evaluation of approaches based on \textbf{signal processing for information security analysis}...
+	\begin{itemize}
+		\item To make the data \textbf{discriminative} for network attack detection, mobile malicious behavior analysis and fraud detection in mobile money transactions.
 	\end{itemize}
@@
 		\item Signal processing formulation as legitimate traffic, malicious traffic and noise;
-		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
-		\item Eigen similarity analysis for extracting detailed information about accurate time and network ports under attack;
-		\item Accuracy and performance evaluation through experimental scenario and the DARPA 1998 dataset.
+		\item MOS and eigenvalue analysis are applied to network attack detection;
+		\item Eigen similarity analysis identify accurate time and network ports under attack;
+		\item Evaluation through experimental scenario and the DARPA 1998 dataset.
 	\end{itemize}
@@
 \begin{frame}[c]{Problem Statement}
-	This thesis also proposes an \textbf{architecture and approach based on user behavior analysis of mobile apps}
+	This thesis also proposes an \textbf{approach malicious behavior detection in mobile apps}
 	\begin{itemize}
@@
 \begin{frame}[c]{Problem Statement}
-	We also propose a \textbf{tensor-based dictionary learning method for fraud detection from imbalanced data}, in order to apply the tensor decomposition for dictionary learning methods to evidentiate the \textbf{discriminative sensing of a fraud detection dataset}
-	\begin{itemize}
-		\item A sparse representation based classification (SRC) method through learning a tensor-based dictionary;
+	We also propose a \textbf{tensor-based dictionary learning method for fraud detection from imbalanced data} 
+	\begin{itemize}
+		\item To evidentiate the \textbf{discriminative sensing of a fraud detection dataset};
+		\item Sparse representation based classification (SRC) through learning a tensor-based dictionary;
 		\item Apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate;
-		\item Evaluate the reconstruction error for fraud classification of mobile money transactions.
+		\item Classify fraud classification of mobile money transactions according to the reconstruction error.
 	\end{itemize}
@@
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Contributions}
-	\begin{enumerate}
-		\item An approach based on \textbf{eigen similarity analysis} for extracting detailed information about accurate time and network ports under network attack;
-		\item Computational \textbf{complexity analysis} of the proposed framework;
-		\item An architecture and techniques for \textbf{offline behavioral analysis} of a corporate mobile client security architecture;
-		\item A \textbf{tensor-based dictionary learning} approach for \textbf{fraud detection} in mobile payment transactions;
-	\end{enumerate}
-\end{frame}
+% \begin{frame}{Contributions}
+% 	\begin{enumerate}
+% 		\item An approach based on \textbf{eigen similarity analysis} for extracting detailed information about accurate time and network ports under network attack;
+% 		\item Computational \textbf{complexity analysis} of the proposed framework;
+% 		\item An architecture and techniques for \textbf{offline behavioral analysis} of a corporate mobile client security architecture;
+% 		\item A \textbf{tensor-based dictionary learning} approach for \textbf{fraud detection} in mobile payment transactions;
+% 	\end{enumerate}
+% \end{frame}
 
@@
 \begin{frame}[c]{Introduction}
-	We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise})
-	\begin{itemize}
+	\begin{itemize}
+		\item We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
 		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
 		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
-		\item Eigen similarity analysis for extracting detailed time and network ports under attack. 
-	\end{itemize}
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Introduction
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Introduction}
-	We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise})
-	\begin{itemize}
-		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
-		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
-		\item Eigen similarity analysis for extracting detailed time and network ports under attack. 
+		\item Eigen similarity analysis for extracting detailed time and network ports under attack;
+		\item We conduct a performance evaluation for complexity analysis and processing time.
 	\end{itemize}
@@
 
-	Classical methods typically employ data mining [39, 34, 70] and regular file analysis [76] to detect patterns that indicate the presence of specific attacks in network traffic;
-
-	Callegari \emph{et al} [104] propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
-However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
+	\begin{itemize}
+		\item Classical methods typically employ data mining [39, 34, 70] and regular file analysis [76] to detect patterns that indicate the presence of specific attacks in network traffic;
+		\item Callegari \emph{et al} [104] propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
+However, Callegari \emph{et al} \textbf{focus on flood attack detection}, not addressing probe attack detection, and their approach relies on \textbf{visual analysis}.
+	\end{itemize}
 
@@
 
-	Lee \emph{et al.} [61] propose OverSampling PCA (osPCA), determines the anomaly of the target according to the variation of eigenvector by similarity analysis and over sampling.
-In contrast, we apply MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack, for flood and probe attacks;
-
-	Lu and Ghorbani [63] proposed a network anomaly detection model based on network flow, wavelet approximation, and system identification theory.
-However, their work requires a training method and presents limitations on identification of behaviors without significant outliers;
-
-	Zonglin \emph{et al} [104] proposed a method to detect traffic anomaly with correlation analysis, but the work is not applied to probe and flood attack detection.
-
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Data Model
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Data Model}
-
-	Focus on \textbf{probe} and \textbf{flooding} attacks;
-
-	Two scenarios: A \textbf{synthetic dataset} as a signal superposition of legitimate traffic, noise and malicious traffic, and selected cases of \textbf{DARPA dataset} that reproduce probe and flooding attacks.
+	\begin{itemize}
+		\item Lee \emph{et al.} [61] propose OverSampling PCA, determines the \textbf{anomaly} according to the variation of \textbf{eigenvector by similarity analysis and over sampling}.
+We apply \textbf{MOS for attack detection} and \textbf{similarity analysis for identification} of time and ports under attack;
+		\item Lu and Ghorbani [63] proposed a \textbf{network anomaly} detection model based on \textbf{network flow, wavelet approximation, and system identification theory}.
+However, their work requires a \textbf{training} and presents \textbf{limitations} on identification of behaviors without \textbf{significant outliers}.
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Datasets
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Datasets}
+
+	We focus on \textbf{probe} and \textbf{flooding} attacks;
+
+	Two scenarios: 
+	\begin{itemize}
+		\item A \textbf{synthetic dataset} as a signal superposition of legitimate traffic, noise and malicious traffic;
+		\item Selected cases of \textbf{DARPA dataset} that reproduce probe and flooding attacks.
+	\end{itemize}
 
@@
 
-	7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks.
-
-	The attacks in this dataset can be grouped into: denial-of-service (DoS); remote to local (R2L); user to root (U2R); and probe attack.
-
-	The most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack.
-
-	We select the cases that reproduce probe and flooding attacks.
+	\begin{itemize}
+		\item 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks;
+		\item The most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack;
+		\item We select the cases that reproduce probe and flooding attacks.
+	\end{itemize}
 
@@
 
-	For flooding detection, calculate the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
+	For \textbf{flooding} detection, calculate the \textbf{sample covariance matrix} $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
 	\begin{equation}\label{eq:eq02}
@@
 
-	For probing detection, compute the sample covariance $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have zero mean and unitary standard deviation as follows
+	For \textbf{probing} detection, compute the \textbf{sample covariance} $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have \textbf{zero mean and unitary standard deviation} as follows
 	\begin{equation}\label{eq:eq04}
@@
 
-	Compute eigenvalue decomposition (EVD) to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).
+	Compute eigenvalue decomposition (EVD) to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix (time frame $q$), according to (\ref{eq:eq060}).
 	\begin{equation}\label{eq:eq06}
@@
 		\begin{bmatrix}
-			\lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
+			\color{red}\lambda_1^{(1)} & \color{red}\lambda_1^{(2)} & \color{red}\lambda_1^{(3)} & \cdots & \color{red}\lambda_1^{(Q)} \\
 			\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
@@
 
-	Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$.
-
-	The proposed approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame to identify variations and estimate the model order $\hat{d}$ (estimated number of time frames under attack)
-
-	$\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to
-
-	$\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
-
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Eigen Similarity Analysis
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Eigen Similarity Analysis}
+	\begin{itemize}
+		\item Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$;
+		\item $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
+		\item $\hat{d}$ is the estimated number of time frames under attack.
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Eigenvalue Analysis
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Eigenvalue Analysis}
+
+	Through eigenvalue analysis, we identify the time frames $q$ under attack
 
@@
 \begin{frame}{Eigen Similarity Analysis}
+
+	The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic without attack.
+
+	Each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$
+	\begin{equation}\label{eq:eq12}
+		\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
+	\end{equation}
+
+	The resultant $\boldsymbol{X}_{(n)}$ is necessary to obtain $\boldsymbol{v}_{(n)}$, through (\ref{eq:eq06}).
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Eigenvalue Analysis
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Eigen Similarity Analysis}
+
+	$s_n$ denotes the absolute similarity degree of the $n$-th minute
+
+	$\boldsymbol{v}^{(q)}$ is the most significant eigenvectors
+
+	$\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of one network traffic
 
@@
 
-	$s_n$ denotes the absolute similarity degree of the $n$-th minute
-
-	$\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack
-
-	$\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of traffic to be performed the flood and port scan attack identification.
-
 \end{frame}
@@
 
-	The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic without attack.
-Each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$, as represented by
-	\begin{equation}\label{eq:eq12}
-		\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
-	\end{equation}
-
-	The resultant $\boldsymbol{X}_{(n)}$ is necessary to obtain $\boldsymbol{v}_{(n)}$, through (\ref{eq:eq06}).
-
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Eigenvalue Analysis
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Eigen Similarity Analysis}
-
-	We evaluate three approaches for eigen similarity analysis: 
+	We evaluate \textbf{three approaches} for eigen similarity analysis: 
 	\begin{itemize}
@@
 
-	For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack is compared against the $\boldsymbol{v}_{(n)}$ identified as under attack, individually evaluating the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.
+	For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack is compared against the $\boldsymbol{v}_{(n)}$ identified as under attack,
+
+	Evaluate the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.
 
@@
 		2 &2341327 &2341327 &2341327 &2,1451 \\
-		3 &3213867 &3213867 &3213867 &10,0718 \\
-		4 &133238294 &133238294 &731229 &2,1620 \\
-		5 &92384021611 &6367983 &92384021611 &2,4253 \\
+		3 &3213867 &3213867 &3213867 &\color{red}10,0718 \\
+		4 &\color{red}133238294 &\color{red}133238294 &731229 &2,1620 \\
+		5 &\color{red}92384021611 &6367983 &\color{red}92384021611 &2,4253 \\
 		6 &708335 &708335 &708335 &1,7948 \\
@@
 		\midrule
-		Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-		Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
+		Detection of synflood \\(presence of attack) &2 &1 &\textbf{\color{red}1} &5 &\textbf{\color{red}1} &4 &\textbf{\color{red}1} \\
+		Detection of synflood \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
 		\midrule
-		Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-		Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
+		Detection of fraggle \\(presence of attack) &1 &1 &\textbf{\color{red}1} &5 &\textbf{\color{red}1} &4 &\textbf{\color{red}1} \\
+		Detection of fraggle \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
 		\midrule
-		Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
-		Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
+		Detection of port scan \\(presence of attack) &1 &1 &\textbf{\color{red}1} &1 &\textbf{\color{red}1} &9 &\textbf{\color{red}1} \\
+		Detection of port scan \\(absence of attack) &0 &0 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} \\
 		\midrule
-		Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
-		Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
+		Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{\color{red}2} &5 &\textbf{\color{red}2} &5 &\textbf{\color{red}2} \\
+		Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
 	    \bottomrule
@@
 
+	MOS schemes for probing and flooding attack detection:
+	\begin{itemize}
+		\item The \textbf{largest eigenvalues visually} indicate the time frame under probing and flooding attack detection;
+		\item MOS schemes can estimate it mathematically;
+		\begin{itemize}
+			\item \textbf{EDC and EFT} estimate \textbf{correctly} the number of attacks $d$;
+		\end{itemize}
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Results}
+
 	\begin{table}[h!]
 	  \centering
-  \footnotesize
+	  \tiny
 	  \caption{Eigen Similarity Analysis for Port Scan Detection}
@@
 		3 &14 &0.9646 &0.9646 &0.9998 &no \\
-	3 &15 &0.0216 &0.0216 &0.0276 &yes \\
-	3 &16 &0.9621 &0.0209 &1.0000 &no \\
-	3 &17 &0.9611 &0.0199 &0.9998 &no \\
-	3 &18 &0.9612 &0.0191 &0.9999 &no \\
-	3 &19 &0.9613 &0.0186 &0.9998 &no \\
-	3 &20 &0.9638 &0.0190 &1.0000 &no \\
+		3 &15 &\color{red}0.0216 &\color{red}0.0216 &\color{red}0.0276 &\color{red}yes \\
+		3 &16 &0.9621 &\color{red}0.0209 &1.0000 &no \\
+		3 &17 &0.9611 &\color{red}0.0199 &0.9998 &no \\
+		3 &18 &0.9612 &\color{red}0.0191 &0.9999 &no \\
+		3 &19 &0.9613 &\color{red}0.0186 &0.9998 &no \\
+		3 &20 &0.9638 &\color{red}0.0190 &1.0000 &no \\
 	    \bottomrule
@@
 		4 &10 &0.9991 &0.9991 &1.0000 &no \\
-	4 &11 &0.0085 &0.0085 &0.0284 &yes \\
-	4 &12 &0.0162 &0.0120 &0.0343 &yes \\
-	4 &13 &0.0248 &0.0158 &0.0427 &yes \\
-	4 &14 &0.1243 &0.0185 &0.1041 &yes \\
-	4 &15 &0.0082 &0.0162 &0.0103 &yes \\
-	4 &16 &0.0404 &0.0070 &0.0580 &yes \\
-	4 &17 &0.0397 &0.0007 &0.0573 &yes \\
-	4 &18 &0.0408 &0.0042 &0.0584 &yes \\
-	4 &19 &0.0408 &0.0079 &0.0584 &yes \\
-	4 &20 &0.0477 &0.0092 &0.0757 &yes \\
+		4 &11 &\color{red}0.0085 &\color{red}0.0085 &\color{red}0.0284 &\color{red}yes \\
+		4 &12 &\color{red}0.0162 &\color{red}0.0120 &\color{red}0.0343 &\color{red}yes \\
+		4 &13 &\color{red}0.0248 &\color{red}0.0158 &\color{red}0.0427 &\color{red}yes \\
+		4 &14 &\color{red}0.1243 &\color{red}0.0185 &\color{red}0.1041 &\color{red}yes \\
+		4 &15 &\color{red}0.0082 &\color{red}0.0162 &\color{red}0.0103 &\color{red}yes \\
+		4 &16 &\color{red}0.0404 &\color{red}0.0070 &\color{red}0.0580 &\color{red}yes \\
+		4 &17 &\color{red}0.0397 &\color{red}0.0007 &\color{red}0.0573 &\color{red}yes \\
+		4 &18 &\color{red}0.0408 &\color{red}0.0042 &\color{red}0.0584 &\color{red}yes \\
+		4 &19 &\color{red}0.0408 &\color{red}0.0079 &\color{red}0.0584 &\color{red}yes \\
+		4 &20 &\color{red}0.0477 &\color{red}0.0092 &\color{red}0.0757 &\color{red}yes \\
 	    \bottomrule
@@
 		5 &10 &0.9995 &0.9995 &1.0000 &no \\
-	5 &11 &0.0031 &0.0031 &0.0021 &yes \\
-	5 &12 &0.0019 &0.0025 &0.0009 &yes \\
-	5 &13 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &14 &0.0030 &0.0027 &0.0020 &yes \\
-	5 &15 &0.0030 &0.0028 &0.0020 &yes \\
-	5 &16 &0.0012 &0.0025 &0.0002 &yes \\
-	5 &17 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &18 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &19 &0.0030 &0.0027 &0.0020 &yes \\
-	5 &20 &0.0069 &0.0023 &0.0083 &yes \\
+		5 &11 &\color{red}0.0031 &\color{red}0.0031 &\color{red}0.0021 &\color{red}yes \\
+		5 &12 &\color{red}0.0019 &\color{red}0.0025 &\color{red}0.0009 &\color{red}yes \\
+		5 &13 &\color{red}0.0030 &\color{red}0.0026 &\color{red}0.0020 &\color{red}yes \\
+		5 &14 &\color{red}0.0030 &\color{red}0.0027 &\color{red}0.0020 &\color{red}yes \\
+		5 &15 &\color{red}0.0030 &\color{red}0.0028 &\color{red}0.0020 &\color{red}yes \\
+		5 &16 &\color{red}0.0012 &\color{red}0.0025 &\color{red}0.0002 &\color{red}yes \\
+		5 &17 &\color{red}0.0030 &\color{red}0.0026 &\color{red}0.0020 &\color{red}yes \\
+		5 &18 &\color{red}0.0030 &\color{red}0.0026 &\color{red}0.0020 &\color{red}yes \\
+		5 &19 &\color{red}0.0030 &\color{red}0.0027 &\color{red}0.0020 &\color{red}yes \\
+		5 &20 &\color{red}0.0069 &\color{red}0.0023 &\color{red}0.0083 &\color{red}yes \\
 	    \bottomrule
@@
 		53 &0.9999 &0.9999 &no \\
-	21 &0.9999 &0.9997 &yes \\
-	22 &0.0298 &0.9997 &yes \\
-	23 &0.0298 &0.9997 &yes \\
-	25 &0.0298 &0.9997 &yes \\
-	110 &0.0298 &0.9997 &yes \\
-	143 &0.0298 &0.9997 &yes \\
-	161 &0.0298 &0.9997 &yes \\
-	69 &0.0298 &0.9997 &yes \\
-	123 &0.0298 &0.9997 &yes \\
-	445 &0.0298 &0.9997 &yes \\
+		21 &0.9999 &0.9997 &\color{red}yes \\
+		22 &\color{red}0.0298 &0.9997 &\color{red}yes \\
+		23 &\color{red}0.0298 &0.9997 &\color{red}yes \\
+		25 &\color{red}0.0298 &0.9997 &\color{red}yes \\
+		110 &\color{red}0.0298 &0.9997 &\color{red}yes \\
+		143 &\color{red}0.0298 &0.9997 &\color{red}yes \\
+		161 &\color{red}0.0298 &0.9997 &\color{red}yes \\
+		69 &\color{red}0.0298 &0.9997 &\color{red}yes \\
+		123 &\color{red}0.0298 &0.9997 &\color{red}yes \\
+		445 &\color{red}0.0298 &0.9997 &\color{red}yes \\
 		600 &0.9999 &0.9999 &no \\
@@
 		445 &1.0000 &1.0000 &no \\
-	600 &0.0077 &0.0427 &yes \\
+		600 &\color{red}0.0077 &\color{red}0.0427 &\color{red}yes \\
 		19 &1.0000 &1.0000 &no \\
@@
 		600 &1.0000 &1.0000 &no \\
-	19 &0.0031 &0.0004 &yes \\
+		19 &\color{red}0.0031 &\color{red}0.0004 &\color{red}yes \\
 		67 &1.0000 &1.0000 &no \\
@@
 
+	For \textbf{time} attack detection:
+	\begin{itemize}
+		\item \textbf{High similarity} between network traffic \textbf{without attack} (0.9610) and \textbf{low similarity} with \textbf{attack} (0.0276);
+		\item The \textbf{incremental} approach produces \textbf{false positive} results;
+		\item Capability of \textbf{change detection} based on similarity between legitimate and malicious traffic (flood or probe);
+	\end{itemize}
+
+	For \textbf{port} attack detection:
+	\begin{itemize}
+		\item The \textbf{incremental individualized} has more \textbf{sensibility} to anomaly detection;
+		\item The \textbf{individual} approach \textbf{was not able} to identify low similarity for ports under attack;
+	\end{itemize}
+
+\end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Results
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}{Results}
+
+% 	Conclusion:
+% 	\begin{itemize}
+% 		\item \textbf{Incremental individualized can detect} low similarity for all evaluated network attacks, while the \textbf{other approaches presented false positives or low sensibility};
+% 		\item This approach is able to \textbf{gradually and incrementally adapt to network traffic changing}.
+% 	\end{itemize}
+
+% \end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results - DARPA Dataset
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Results - DARPA Dataset}
+
 	\begin{table}[!t]
@@
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results - DARPA Dataset
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Results - DARPA Dataset}
+
+	% Misclassification is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$;
+	\begin{itemize}
+		\item High FP and misclassification due to legitimate traffic be massive (such as an attack) sometimes;
+		\item Callegari \emph{et al} is a statistical method, based on PCA, without training or learning methods
+		\begin{itemize}
+			\item 82 \% while we obtain 100 \%;
+			\item FP and misclassification was not evaluated.
+		\end{itemize}
+		\item Ghorbani's is based on signal processing techniques and uses DARPA
+		\begin{itemize}
+			\item 94.67 \% and 50.00 \%, while we obtain 100 \% and 76.92 \%;
+			\item FP and misclassification was not evaluated.
+		\end{itemize}
+	\end{itemize}
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
 %	FRAME: Performance Evaluation
@@
 
-	complexity
+	Complexity Analysis:
+	\begin{itemize}
+		\item the proposed framework is $O(N^3 + Q \log Q + \hat{d}Q + N^3)$ and its worst-case running time is $O(N^3)$;
+		\item The computational complexity of \textbf{EVD is predominant};
+		\item \textbf{However}, the approach \textbf{splits the data into time frames} with period time $N$, which makes possible to \textbf{limit the growth of $N$} even when \textbf{total time is larger} than $N$;
+	\end{itemize}
 
@@
 
-	processing
+	Processing Time Analysis:
+	\begin{itemize}
+		\item desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM;
+		\item Network traffic time; 
+		\item Frame size denoted as $N$; 
+		\item Number of network ports denoted as $M$; 
+		\item Mean processing time for eigen analysis based on sample covariance of zero mean (1-time); 
+		\item Mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation (2-time); 
+		\item Mean processing time for EDC MOS scheme, (3-time).
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Performance Evaluation
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Performance Evaluation}
+
+	\begin{table}[!t]
+		\caption{Processing time of the main steps for anomaly detection}
+		\label{tab:11}
+		\centering
+		\tiny
+		\begin{tabular}{|r|r|r|r|r|r|r|}
+			\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Traffic Time\\(hour)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Frame Size\\(min)\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Num.
+Ports\end{tabular}	& \begin{tabular}[x]{@{}l@{}}1-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}2-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}3-time\\(ms)\end{tabular}	\\ \hline
+			16	& 10	& 17	& 0.7900	& 0.8100	& 0.0650\\ \hline
+			16	& 20	& 17	& 0.5250	& 0.5950	& 0.0100\\ \hline
+			16	& 60	& 17	& 0.9700	& 1.1400	& 0.0250\\ \hline
+			16	& 120	& 17	& 0.6050	& 0.6100	& 0.0050\\ \hline
+			16	& 60	& 34	& 1.2750	& 1.2200	& 0.0050\\ \hline
+			16	& 120	& 34	& 1.1200	& 1.1700	& 0.0050\\ \hline
+			20	& 10	& 17	& 2.7950	& 2.8950	& 1.1000\\ \hline
+			20	& 20	& 17	& 2.0700	& 2.0200	& 0.3500\\ \hline
+			20	& 60	& 17	& 1.0250	& 1.0450	& 0.0650\\ \hline
+			20	& 120	& 17	& 1.0000	& 1.0700	& 0.0350\\ \hline
+			20	& 60	& 34	& 2.9650	& 3.2100	& 0.0400\\ \hline
+			20	& 120	& 34	& 2.9950	& 3.1150	& 0.0200\\ \hline
+			22	& 10	& 17	& 4.7250	& 4.0850	& 1.4600\\ \hline
+			22	& 20	& 17	& 2.3200	& 2.6800	& 0.2450\\ \hline
+			22	& 60	& 17	& 1.0700	& 1.1200	& 0.0300\\ \hline
+			22	& 120	& 17	& 0.9900	& 1.0500	& 0.0250\\ \hline
+			22	& 60	& 34	& 3.0850	& 3.1250	& 0.0650\\ \hline
+			22	& 120	& 34	& 2.8100	& 2.9600	& 0.0250\\ \hline
+		\end{tabular}
+	\end{table}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Performance Evaluation
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}{Performance Evaluation}
+
+	Processing Time Analysis:
+	\begin{itemize}
+		\item The processing time increases according to traffic time, but the \textbf{worst} processing time is \textbf{4.7250 milliseconds};
+		\item The processing time \textbf{increases} with the frame size $N$ \textbf{decreasing};
+		\item The number of ports $M$ increases the processing time.
+	\end{itemize}
 
@@
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Blocks
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}{Blocks}
-
-\begin{block}{Block Title Here}
-		Great for definitions
-\end{block}
-
-\begin{alertblock}{Alert Title Here}
-		Great for definitions
-\end{alertblock}
-
-\begin{exampleblock}{Example Title Here}
-		Great for examples
-\end{exampleblock}
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Blocks
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}{Blocks}
-
-\begin{block}{Block Title Here}
-	\begin{itemize}
-		\item point 1
-		\item point 2
-	\end{itemize}
-\end{block}
-
-\begingroup
-\setbeamercolor{block title}{fg=white, bg=\cnBlue}
-\setbeamercolor{block body}{bg=\cnLightBlue}
-\begin{block}{Blue Colored Blocks}
-	Produced by using the \texttt{cblock} theme option
-\end{block}
-\endgroup
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Additional Blocks
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}{Additional Blocks}
-\begin{alertblock}{Alert Block}
-	Highlight important information.
-\end{alertblock}
-
-\begingroup
-\setbeamercolor{block title}{fg=white, bg=\cnRed}
-\setbeamercolor{block body}{bg=\cnLightRed}
-\begin{block}{Red Colored Blocks}
-	Produced by using the \texttt{cblock} theme option
-\end{block}
-\endgroup
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Additional Blocks
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}{Additional Blocks}
-
-\begin{exampleblock}{Example Block}
-	Examples can be good.
-\end{exampleblock}
-
-\begingroup
-\setbeamercolor{block title}{fg=white, bg=\cnGreen}
-\setbeamercolor{block body}{bg=\cnLightGreen}
-\begin{block}{Green Colored Blocks}
-	Produced by using the \texttt{cblock} theme option
-\end{block}
-\endgroup
-
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Blocks
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}{Custom Blocks}
-\begingroup
-\setbeamercolor{block title}{fg=white, bg=\cnPurple}
-\setbeamercolor{block body}{bg=\cnLightPurple}
-\begin{block}{Purple customization}
-	Using the theme colors to generate colored blocks.
-\end{block}
-\endgroup
-
+%	FRAME: Introduction
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Introduction}
+	\begin{itemize}
+		\item We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
+		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
+		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
+		\item Eigen similarity analysis for extracting detailed time and network ports under attack;
+		\item We conduct a performance evaluation for complexity analysis and processing time.
+	\end{itemize}
 \end{frame}
@@
 \section{Tensor-based Discriminative Sensing for Fraud Detection}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: 
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}[c]{No Special Fonts Required}
-
-This theme was originally made to work with \texttt{pdflatex} and the default latex fonts.\\
-\vspace{1em}
-\cRed{sthlm} does comes with a \texttt{pdflatex} font option, \cRed{newPxFont}, which loads the following fonts:
-
-\begin{itemize}
-	\item \emph{newpxtext} for text
-	\item \emph{cantarell} for sans-serif
-	\item \emph{inconsolata} for sans-serif monospaced
-	\item \emph{newpxmath} for math
-\end{itemize}
-
-Please refer to the \texttt{beamerfontthememsthlm.sty} for the package requirements.
-
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Introduction
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Introduction}
+	\begin{itemize}
+		\item We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
+		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
+		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
+		\item Eigen similarity analysis for extracting detailed time and network ports under attack;
+		\item We conduct a performance evaluation for complexity analysis and processing time.
+	\end{itemize}
 \end{frame}
@@
 \section{Conclusion and Future Work}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Tables
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}{Tables}
-\begin{table}[]
-	\caption{Selection of window function and their properties}
-	\begin{tabular}[]{lrrr}
-		\toprule
-		\textbf{Window}			& \multicolumn{1}{c}{\textbf{First side lobe}}
-		                    & \multicolumn{1}{c}{\textbf{3\,dB bandwidth}}
-		                    & \multicolumn{1}{c}{\textbf{Roll-off}} \\
-		\midrule
-		Rectangular				& 13.2\,dB	& 0.886\,Hz/bin	& 6\,dB/oct		\\[0.25em]
-		Triangular				& 26.4\,dB	& 1.276\,Hz/bin	& 12\,dB/oct	\\[0.25em]
-		Hann					& 31.0\,dB	& 1.442\,Hz/bin	& 18\,dB/oct	\\[0.25em]
-		Hamming					& 41.0\,dB	& 1.300\,Hz/bin	& 6\,dB/oct		\\
-		\bottomrule
-	\end{tabular}
-	\label{tab:WindowFunctions}
-\end{table}
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME:
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}[c]{Mathematics Step by Step}
-Show $[x^n]'=nx^{n-1}$ by using first principles.
-\begin{align*}
-\visible<+->{f'(x) & = \displaystyle \lim_{\Delta x \to 0} \frac{f(x+\Delta x)-f(x)}{\Delta x} \\}
-\visible<+->{&= \displaystyle \lim_{\Delta x \to 0} \frac{(\cRed{x+\Delta x})^n - (\cRed{x})^n}{\Delta x}\\}
-\visible<+->{&= \displaystyle \lim_{\Delta x \to 0} \frac{\binom{n}{0}x^n\Delta x^0 +\binom{n}{1}x^{n-1}\Delta x^1 + \cdots + \binom{n}{n}x^0\Delta x^n - \cRed{x^n} }{\Delta x}\\}
-\visible<+->{&= \displaystyle \lim_{\Delta x \to 0} \frac{\cRed{1}x^n(\cRed{1}) +\cRed{n}x^{n-1}\Delta x^1 + \cdots + \cRed{1}(\cRed{1})\Delta x^n - \cRed{x^n} }{\Delta x}\\}
-\visible<+->{&= \displaystyle \lim_{\Delta x \to 0} \frac{\cRed{x^n} +nx^{n-1}\Delta x + \cdots + \Delta x^n - \cRed{x^n} }{\Delta x}\\}
-\visible<+->{&= \displaystyle \lim_{\Delta x \to 0} \frac{\cancel{\cRed{\Delta x}} (nx^{n-1} + \cdots + \Delta x^{n-1} ) }{\cancel{\Delta x}}\\}
-\visible<+->{&=nx^{n-1}}
-\end{align*}
-
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Formulas
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}{Functions}
-\begin{block}{Gaussian Probability Density Function}
-\[
-f \left(x \mid \mu, \sigma^2 \right) = \dfrac{1}{\sqrt{2 \sigma^2 \pi}} e^{- \dfrac{(x-\mu)^2}{2\sigma^2}}
-\]
-\end{block}
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: 
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}[c]{More Mathematics}
-
-\begin{columns}[c]
-\begin{column}{0.5\textwidth}
-\begin{tikzpicture}[scale=0.7]
-	\begin{axis}[
-            domain=0:2, range=0:6,ymax=6,ymin=0,
-            axis lines =left,
-            every axis y label/.style={at=(current axis.above origin),anchor=south},
-            every axis x label/.style={at=(current axis.right of origin),anchor=west},
-            xtick={1,1.666}, ytick={1,3},
-            xticklabels={$\cPurple{x}$,$\cPurple{x}+\cRed{\Delta x}$}, yticklabels={$f(\cPurple{x})$,$f(\cPurple{x}+\cRed{\Delta x})$},
-            axis on top,
-          ]
-          \addplot [very thick,\cnRed, domain=(0:2)] {-1.994+2.994*x};S
-	  	  \addplot [\cnDarkGrey,dashed] plot coordinates {(1,0) (1,1)};
-          \addplot [\cnDarkGrey,dashed] plot coordinates {(0,1) (1,1)};
-          \addplot [\cnDarkGrey,dashed] plot coordinates {(0,3) (1.666,3)};
-          \addplot [\cnDarkGrey,dashed] plot coordinates {(1.666,0) (1.666,3)};
-          \addplot [very thick,\cnBlue, smooth,domain=(0:1.833)] {-1/(x-2)};
-          \addplot[color=\cnDarkGrey,fill=\cnDarkGrey,only marks,mark=*] coordinates{(1.666,3)};  %% closed hole
-          \addplot[color=\cnDarkGrey,fill=\cnDarkGrey,only marks,mark=*] coordinates{(1,1)};  %% closed hole
-        \end{axis}
-\end{tikzpicture}
-\end{column}
-
-\begin{column}{0.5\textwidth}
-\begin{align*}
-m &= \frac{\Delta y}{\Delta x}	\\
-\\
-m & =  \frac{\visible<2->{f(\cPurple{x}+\cRed{\Delta x})}\visible<3->{-}\visible<4->{f(\cPurple{x})}}{\visible<5->{\cPurple{x}+\cRed{\Delta x}}\visible<6->{-}\visible<7->{\cPurple{x}}}\\
-m  &=  \underbrace{\frac{\visible<8->{f(\cPurple{x}+\cRed{\Delta x})-f(\cPurple{x})}}{\visible<9->{\cRed{\Delta x}}}}_{\visible<10->{\text{\cRed{difference quotient}}}}
-\end{align*}
-\end{column}
-\end{columns}
-
-\visible<10->{The slope of the secant line }
-\begin{itemize}
-	\item<10-> can be found using the \cRed{difference quotient}
-	\item<11-> represents a \cBlue{function}'s \cRed{average} slope on the interval $[\cPurple{x}, \cRed{x+\Delta x}]$
-\end{itemize}
-
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME:
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}[c, fragile]{Fragile Frames Not a Problem}
-\begin{tikzpicture}
-	\begin{axis}[
-            domain=0:6, range=0:7,
-            ymin=-.2,ymax=7,
-            width=\textwidth,
-            height=7cm, %% Hard coded height! Moreover this effects the aspect ratio of the zoom--sort of BAD
-            axis lines=none,
-          ]
-          \only<3->{\addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(.8,1.6) (2.834,5)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(2.834,5) (4.166,5)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=white] plot coordinates {(1.2,1.6) (4.166,5)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=white] plot coordinates {(.8,1.6) (1.2,1.6)} \closedcycle; %% zoom fill
-		  }
-          \only<4->{\addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(3.3,3.6) (5.334,5)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(5.334,5) (6.666,5)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=white] plot coordinates {(3.7,3.6) (6.666,5)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=white] plot coordinates {(3.3,3.6) (3.7,3.6)} \closedcycle; %% zoom fill
-  }
-          \only<4->{\addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(3.7,2.4) (6.666,1)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(3.3,2.4) (3.7,2.4)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=white] plot coordinates {(3.3,2.4) (5.334,1)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=white] plot coordinates {(5.334,1) (6.666,1)} \closedcycle; %% zoom fill
- }
-
-          \only<3->{\addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(.8,.4) (2.834,1)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=\cnDarkGrey!10!\cnGrey] plot coordinates {(2.834,1) (4.166,1)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=white] plot coordinates {(1.2,.4) (4.166,1)} \closedcycle; %% zoom fill
-          \addplot [draw=none, fill=white] plot coordinates {(.8,.4) (1.2,.4)} \closedcycle; %% zoom fill
-}
-          \addplot[very thick,\cnRed, smooth,domain=(0:1.833)] {-1/(x-2)};
-          \only<3->{\addplot[very thick,\cnRed, smooth,domain=(2.834:4.166)] {3.333/(2.050-.3*x)-0.333}; %% 2.5 to 4.333
-          %\addplot[very thick,\cnRed, smooth,domain=(5.334:6.666)] {11.11/(1.540-.09*x)-8.109}; %% 5 to 6.833
-          \only<4->\addplot[very thick,\cnRed, smooth,domain=(5.334:6.666)] {x-3}; %% 5 to 6.833
- }
-         \only<2->{ \addplot[color=\cnRed,fill=\cnRed,only marks,mark=*] coordinates{(1,1)};}  %% point to be zoomed
-          \only<3->{\addplot[color=\cnRed,fill=\cnRed,only marks,mark=*] coordinates{(3.5,3)}; } %% zoomed pt 1
-          \only<4->{\addplot[color=\cnRed,fill=\cnRed,only marks,mark=*] coordinates{(6,3)}; } %% zoomed pt 2
-
-          \addplot [->,\cnDarkGrey] plot coordinates {(0,0) (0,6)}; %% axis
-          \addplot [->,\cnDarkGrey] plot coordinates {(0,0) (2,0)}; %% axis
-
-          \only<2->{\addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(.8,.4) (.8,1.6)}; %% box around pt
-          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(1.2,.4) (1.2,1.6)}; %% box around pt
-          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(.8,1.6) (1.2,1.6)}; %% box around pt
-          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(.8,.4) (1.2,.4)}; %% box around pt
- }
-          \only<3->{\addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(2.834,1) (2.834,5)}; %% zoomed box 1
-          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(4.166,1) (4.166,5)}; %% zoomed box 1
-          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(2.834,1) (4.166,1)}; %% zoomed box 1
-          \addplot [\cnDarkGrey!50!\cnGrey] plot coordinates {(2.834,5) (4.166,5)}; %% zoomed box 1
-}
-          \only<4->{\addplot [\cnDarkGrey] plot coordinates {(3.3,2.4) (3.3,3.6)}; %% box around zoomed pt
-          \addplot [\cnDarkGrey] plot coordinates {(3.7,2.4) (3.7,3.6)}; %% box around zoomed pt
-          \addplot [\cnDarkGrey] plot coordinates {(3.3,3.6) (3.7,3.6)}; %% box around zoomed pt
-          \addplot [\cnDarkGrey] plot coordinates {(3.3,2.4) (3.7,2.4)}; %% box around zoomed pt
-}
-          \only<4->{\addplot [\cnDarkGrey] plot coordinates {(5.334,1) (5.334,5)}; %% zoomed box 2
-          \addplot [\cnDarkGrey] plot coordinates {(6.666,1) (6.666,5)}; %% zoomed box 2
-          \addplot [\cnDarkGrey] plot coordinates {(5.334,1) (6.666,1)}; %% zoomed box 2
-          \addplot [\cnDarkGrey] plot coordinates {(5.334,5) (6.666,5)}; %% zoomed box 2
-}
-          \node at (axis cs:2.4,0) [anchor=east] {$x$};
-          \node at (axis cs:0,6.9) [anchor=north] {$y$};
-        \end{axis}
-\end{tikzpicture}
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: PGFPlots
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}{PGFPlots Example}
-	\begin{figure}[h]
-		\centering
-		\begin{tikzpicture}
-		\begin{axis}[
-		    ybar,
-		    enlarge x limits=0.15,
-		    legend style={at={(-.5,0.5)},
-		      anchor=north,legend columns=1},
-		    ylabel={\% students},
-		    symbolic x coords={A, B, C, D, E, F, DNF},
-		    xtick=data,
-		     bar width=2mm,
-		     width=0.7\textwidth
-		    ]
-		    \legend{2012, 2013};
-		    % Spring 2012 results
-			\addplot[fill=\cnGrey]  coordinates {(A,0) (B,0) (C,3.85) (D,23.07) (E,43.31) (F,30.77) (DNF,0.00)};
-			% Spring 2013 results
-			\addplot[fill=\cnBlue]   coordinates {(A,0) (B,3.70) (C,22.22) (D,22.22) (E,40.74) (F,3.70) (DNF,7.41)};
-		\end{axis}
-		\end{tikzpicture}
-		\caption{Consistent improvement over the last year}
-		\end{figure}
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Multiple Columns
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}{Multiple Columns}
-\begin{columns}
-\begin{column}{.48\linewidth}
-		Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod
-		tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam.
-\end{column}
-\begin{column}{.48\linewidth}
-		\begin{itemize}
-        	\item Point 1
-			\begin{itemize}
-				\item Sub point a
-				\item Sub point b
-			\end{itemize}
-        	\item Point 2
-		\end{itemize}
-	\end{column}
-	\end{columns}
-\end{frame}
-
-\begin{frame}{References}
-	\begin{thebibliography}{10}
-
-	\beamertemplatebookbibitems
-	\bibitem{Oppenheim2009}
-	Alan V. Oppenheim
-	\newblock Discrete - Time Signal Processing
-	\newblock Prentice Hall Press, 2009
-
-	\beamertemplatearticlebibitems
-	\bibitem{EBU2011}
-	European Broadcasting Union
-	\newblock Specification of the Broadcast Wave Format (BWF)
-	\newblock 2011
-
-  \end{thebibliography}
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%
-%	SECTION: Conclusion
-%
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-\begin{frame}{About}
-
-This sthlm beamer theme is free software: you can redistribute it and/or modify
-it under the terms of the GNU General Public License as published by
-the Free Software Foundation, either version 3 of the License, or
-(at your option) any later version.\\
-\vspace{1cm}
-If you have any questions or comments
-\begin{itemize}
-	\item Website: \cBlue{markolson.se}
-	\item Twitter: \cBlue{@markolsonse}
-	\item Instagram: \cBlue{@markolson.se}
-\end{itemize}
-\end{frame}
-
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME:
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begingroup
-\setbeamercolor{background canvas}{bg=\cnDarkGrey}
-\begin{frame}[plain]
-
-\centering{\cGrey{\Huge{THE \newline END}}}
-
-\end{frame}
-\endgroup
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Introduction
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Introduction}
+	\begin{itemize}
+		\item We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
+		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
+		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
+		\item Eigen similarity analysis for extracting detailed time and network ports under attack;
+		\item We conduct a performance evaluation for complexity analysis and processing time.
+	\end{itemize}
+\end{frame}
 

### 2017-11-27 docs/thesis/slides/apresentacao_unb.tex
mobile

@@
 	     \includegraphics[width=11cm]{alg.png}
-	     \label{fig:2_fig9}
+	     \label{fig:2_fig91}
 	\end{figure}
@@
 	\begin{itemize}
+		\item We present the \textbf{mobile security architecture for offline mobile client security};
+		\item We discuss the common \textbf{threat} scenarios for \textbf{file cloud apps};
+		\item We evaluate the \textbf{performance} for \textbf{mobile} offline processing.
+	\end{itemize}
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Related works
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Related works}
+
+	\begin{itemize}
+		\item According to [53, 55], the \textbf{state-of-the-art mobile cloud security models do not consider} the problem of the offline security mode;
+		\item Providers of the mobile security \textbf{avoid} the problem by completely \textbf{forbidding} the offline access.
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: The mobile security architecture
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{The mobile security architecture}
+
+	\begin{figure}[h!]
+		\centering
+		\includegraphics[width=5cm]{../figures/fig01.eps}
+		\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
+		\label{fig:3_01}
+	\end{figure}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: The Client-side Architecture
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{The Client-side Architecture}
+
+	\begin{figure}[h!]
+		\centering
+		\includegraphics[width=9cm]{../figures/fig02.eps}
+		\caption{The Client-side Architecture}
+		\label{fig:3_02}
+	\end{figure}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Proposal for offline mobile security
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Proposal for offline mobile security}
+
+	\begin{figure}[h!]
+		\centering
+		\includegraphics[width=11cm]{../figures/fig03.eps}
+		\caption{Proposed Architecture for Offline Mobile Security}
+		\label{fig:3_03}
+	\end{figure}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Offline Behavioral Analysis
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Offline Behavioral Analysis}
+
+	\begin{figure}[h!]
+		\centering
+		\includegraphics[width=7cm]{../figures/fig06.eps}
+		\caption{The Threat Intelligence Manager Workflow}
+		\label{fig:3_06}
+	\end{figure}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: MOS for Threat Intelligence
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{MOS for Threat Intelligence}
+
+	\begin{itemize}
+		\item Previously presented;
+		\item Applied to user behavior analysis;
+		\begin{itemize}
+			\item File operations modeled as signals into matrices;
+			\item Selected features can highlight malicious behaviors;
+			\item Implemented and evaluated into mobile devices.
+		\end{itemize}
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Common threat scenarios
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Common threat scenarios}
+
+	\begin{enumerate}
+		\item An attacker uses a \textbf{valid password} to perform operations on a \textbf{bulk} of files.;
+		\item Usage of \textbf{expired password} to perform unauthorized \textbf{operations};
+	\end{enumerate}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Data Modeling for Behavioral Analysis
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Data Modeling for Behavioral Analysis}
+
+	Selected features:
+	\begin{itemize}
+		\item File Access (Time and File System Location);
+		\item File Update (Time and File System Location);
+		\item File Download (Start Time, End Time and File System Location);
+		\item File Upload (Start Time, End Time and File System Location).
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Data Modeling and Eigenvalue Decomposition Time}
+
+	\begin{table}
+		\tiny
+		\label{tab:3_04}
+		\centering
+		\begin{tabular}{|l|l|l|l|l|l|l|l|}
+			\hline \rowcolor{Gray} Device	& \begin{tabular}[x]{@{}l@{}}Log Size\\(MB)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Window\\(min)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Modeling\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Avg.
+Eig.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Std.
+Eig.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Eig.
+Min.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Eig.
+Max.\\(ms)\end{tabular}	\\ \hline
+			Galaxy GT-I9300	& 6	& 60	& 107	& 209.52	& 18.58	& 183	& 276	\\ \hline
+			Galaxy GT-I9300	& 6	& 40	& 115	& 227.26	& 18.13	& 191	& 289	\\ \hline
+			Galaxy GT-I9300	& 6	& 20	& 89	& 268.14	& 21.94	& 229	& 315	\\ \hline
+			Galaxy GT-I9300	& 6	& 10	& 90	& 347.42	& 24.11	& 304	& 421	\\ \hline
+			Galaxy GT-I9300	& 4.1	& 60	& 20	& 60.90	& 15.19	& 37	& 106	\\ \hline
+			Galaxy GT-I9300	& 4.1	& 40	& 20	& 68.72	& 15.71	& 43	& 114	\\ \hline
+			Galaxy GT-I9300	& 4.1	& 20	& 34	& 89.04	& 16.78	& 54	& 133	\\ \hline
+			Galaxy GT-I9300	& 4.1	& 10	& 21	& 117.24	& 14.36	& 96	& 171	\\ \hline
+			Galaxy GT-I9300	& 1.4	& 60	& 10	& 159.82	& 15.82	& 125	& 197	\\ \hline
+			Galaxy GT-I9300	& 1.4	& 40	& 10	& 168.06	& 15.90	& 139	& 220	\\ \hline
+			Galaxy GT-I9300	& 1.4	& 20	& 11	& 204.4	& 20.46	& 176	& 269	\\ \hline
+			Galaxy GT-I9300	& 1.4	& 10	& 13	& 259.00	& 21.34	& 220	& 315	\\ \hline
+			Galaxy Tab SM-T800	& 6	& 60	& 7	& 59.30	& 6.55	& 54	& 74	\\ \hline
+			Galaxy Tab SM-T800	& 6	& 40	& 8	& 62.56	& 7.05	& 56	& 80	\\ \hline
+			Galaxy Tab SM-T800	& 6	& 20	& 10	& 73.28	& 8.59	& 65	& 95	\\ \hline
+			Galaxy Tab SM-T800	& 6	& 10	& 8	& 93.48	& 9.13	& 83	& 130	\\ \hline
+			Galaxy Tab SM-T800	& 4.1	& 60	& 11	& 18.64	& 4.51	& 16	& 38	\\ \hline
+			Galaxy Tab SM-T800	& 4.1	& 40	& 11	& 19.64	& 5.12	& 17	& 38	\\ \hline
+			Galaxy Tab SM-T800	& 4.1	& 20	& 12	& 25.12	& 5.55	& 21	& 46	\\ \hline
+			Galaxy Tab SM-T800	& 4.1	& 10	& 12	& 32.32	& 7.29	& 27	& 55	\\ \hline
+			Galaxy Tab SM-T800	& 1.4	& 60	& 4	& 49.08	& 6.01	& 42	& 62	\\ \hline
+			Galaxy Tab SM-T800	& 1.4	& 40	& 5	& 51.42	& 7.36	& 44	& 74	\\ \hline
+			Galaxy Tab SM-T800	& 1.4	& 20	& 5	& 51.12	& 7.80	& 54	& 91	\\ \hline
+			Galaxy Tab SM-T800	& 1.4	& 10	& 7	& 75.24	& 7.71	& 65	& 90	\\ \hline
+		\end{tabular}
+	\end{table}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{EDC MOS scheme processing time for anomaly detection}
+
+	\begin{table}[!t]
+		\tiny
+		\label{tab:3_05}
+		\centering
+		\begin{tabular}{|l|l|l|l|l|l|l|l|}
+			\hline \rowcolor{Gray} Device	& \begin{tabular}[x]{@{}l@{}}Log Size\\(MB)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Window\\(min)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Avg.
+EDC.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Std.
+EDC.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Min.
+EDC.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Max.
+EDC.\\(ms)\end{tabular}	\\ \hline
+			Galaxy GT-I9300	& 6	& 60	& 5.27	& 4.04	& 3	& 20	\\ \hline
+			Galaxy GT-I9300	& 6	& 40	& 10.78	& 6.37	& 6	& 34	\\ \hline
+			Galaxy GT-I9300	& 6	& 20	& 32.62	& 12.44	& 21	& 88	\\ \hline
+			Galaxy GT-I9300	& 6	& 10	& 115.08	& 17.45	& 88	& 158	\\ \hline
+			Galaxy GT-I9300	& 4.1	& 60	& 5.68	& 4.18	& 3	& 23	\\ \hline
+			Galaxy GT-I9300	& 4.1	& 40	& 10.76	& 5.31	& 7	& 27	\\ \hline
+			Galaxy GT-I9300	& 4.1	& 20	& 37.58	& 10.30	& 23	& 61	\\ \hline
+			Galaxy GT-I9300	& 4.1	& 10	& 125.98	& 18.56	& 101	& 191	\\ \hline
+			Galaxy GT-I9300	& 1.4	& 60	& 4.92	& 3.49	& 3	& 17	\\ \hline
+			Galaxy GT-I9300	& 1.4	& 40	& 9.00	& 4.23	& 6	& 25	\\ \hline
+			Galaxy GT-I9300	& 1.4	& 20	& 30.14	& 9.21	& 19	& 62	\\ \hline
+			Galaxy GT-I9300	& 1.4	& 10	& 100.62	& 15.83	& 69	& 163	\\ \hline
+			Galaxy Tab SM-T800	& 6	& 60	& 1.84	& 0.65	& 1	& 3	\\ \hline
+			Galaxy Tab SM-T800	& 6	& 40	& 3.26	& 1.24	& 2	& 7	\\ \hline
+			Galaxy Tab SM-T800	& 6	& 20	& 10.90	& 2.40	& 9	& 21	\\ \hline
+			Galaxy Tab SM-T800	& 6	& 10	& 41.86	& 7.33	& 34	& 60	\\ \hline
+			Galaxy Tab SM-T800	& 4.1	& 60	& 1.85	& 0.60	& 1	& 3	\\ \hline
+			Galaxy Tab SM-T800	& 4.1	& 40	& 3.62	& 1.10	& 2	& 8	\\ \hline
+			Galaxy Tab SM-T800	& 4.1	& 20	& 12.04	& 2.79	& 9	& 22	\\ \hline
+			Galaxy Tab SM-T800	& 4.1	& 10	& 40.16	& 6.48	& 35	& 60	\\ \hline
+			Galaxy Tab SM-T800	& 1.4	& 60	& 1.98	& 0.89	& 1	& 6	\\ \hline
+			Galaxy Tab SM-T800	& 1.4	& 40	& 3.30	& 1.16	& 2	& 7	\\ \hline
+			Galaxy Tab SM-T800	& 1.4	& 20	& 10.48	& 2.90	& 8	& 21	\\ \hline
+			Galaxy Tab SM-T800	& 1.4	& 10	& 34.52	& 4.08	& 30	& 45	\\ \hline
+		\end{tabular}
+	\end{table}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Results}
+
+	\begin{itemize}
+		\item The lower window size leads to the larger eigenvalue decomposition time;
+		\item The largest eigenvalue decomposition time, which was the maximum of 421 milliseconds with average of 347.42 milliseconds;
+		\item The MOS processing time increases with the window size decreasing;
+		\item The longest processing time measured is lower than 200 milliseconds;
+		\item This results represent an acceptable processing time for anomaly detection in mobile clients
+	\end{itemize}
+
+\end{frame}
+
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%
+%	SECTION: Tensor-based Discriminative Sensing for Fraud Detection
+%
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\section{Tensor-based Discriminative Sensing for Fraud Detection}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Introduction
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Introduction}
+	\begin{itemize}
 		\item We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
@@
 %
-%	SECTION: Tensor-based Discriminative Sensing for Fraud Detection
-%
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\section{Tensor-based Discriminative Sensing for Fraud Detection}
+%	SECTION: Conclusion and Future Work
+%
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\section{Conclusion and Future Work}
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
@@
 
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%
-%	SECTION: Conclusion and Future Work
-%
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\section{Conclusion and Future Work}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Introduction
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Introduction}
-	\begin{itemize}
-		\item We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
-		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
-		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
-		\item Eigen similarity analysis for extracting detailed time and network ports under attack;
-		\item We conduct a performance evaluation for complexity analysis and processing time.
-	\end{itemize}
-\end{frame}
-
 \end{document}

### 2017-11-28 docs/thesis/slides/apresentacao_unb.tex
first full version

@@
 		\item Number of network ports denoted as $M$; 
-		\item Mean processing time for eigen analysis based on sample covariance of zero mean (1-time); 
-		\item Mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation (2-time); 
-		\item Mean processing time for EDC MOS scheme, (3-time).
+		\item Mean processing time for eigen analysis based on sample covariance of zero mean (\textbf{1-time}); 
+		\item Mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation (\textbf{2-time}); 
+		\item Mean processing time for EDC MOS scheme (\textbf{3-time}).
 	\end{itemize}
@@
 	\begin{itemize}
-		\item We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
-		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
-		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
-		\item Eigen similarity analysis for extracting detailed time and network ports under attack;
-		\item We conduct a performance evaluation for complexity analysis and processing time.
-	\end{itemize}
+		\item A \textbf{key challenge} to use sparse coding and dictionary learning for \textbf{classification} is how to find \textbf{proper dictionaries and coefficients};
+		\item The data and its dictionary can be \textbf{multidimensional};
+		\item \textbf{Tensor decompositions} can be useful to extract \textbf{hidden patterns} and structure in \textbf{multidimensional} data analytics problems;
+		\item The \textbf{fundamental issue} with the \textbf{imbalanced learning} problem is the ability of imbalanced data to \textbf{compromise the performance of algorithms};
+	\end{itemize}
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Introduction
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Introduction}
+	\begin{itemize}
+		\item We propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds; 
+		\item Tensor-based DL for learning fraudulent and legitimate data separately
+		\item Apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate.
+	\end{itemize}
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Motivation: Tensor-based Dictionary Recovery
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Motivation: Tensor-based Dictionary Recovery}
+	\begin{itemize}
+		\item Tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets.
+	\end{itemize}
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Motivation: Tensor-based Dictionary Recovery
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Motivation: Tensor-based Dictionary Recovery}
+
+	Consider a generic sparse recovery problem
+	\begin{equation}\label{eq:4_eq01}
+		\textbf{X} = \textbf{A} \cdot \textbf{S} + \textbf{W},
+	\end{equation}
+	where $\textbf{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\textbf{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary where $N \ll T$, $\textbf{S} \in \mathbb{C}^{N \times T}$ represents the sparse coefficient matrix, $\textbf{W} \in \mathbb{C}^{M \times T}$ is the additive noise, and $M < N < T$.
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Motivation: Tensor-based Dictionary Recovery
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Motivation: Tensor-based Dictionary Recovery}
+
+	Consider a sparse recovery problem for a separable 2-D dictionary that we can write as
+	\begin{equation}\label{eq:4_eq02}
+		\textbf{X} = (\textbf{A}^{(1)} \otimes \textbf{A}^{(2)}) \cdot \textbf{S} + \textbf{W},
+	\end{equation}
+	where $\textbf{A} = (\textbf{A}^{(1)} \otimes \textbf{A}^{(2)})$, $\textbf{A}^{(1)} \in \mathbb{C}^{M_1 \times N_1}$, $\textbf{A}^{(2)} \in \mathbb{C}^{M_2 \times N_2}$, $M = M_1 \times M_2$, and $N = N_1 \times N_2$.
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Motivation: Tensor-based Dictionary Recovery
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Motivation: Tensor-based Dictionary Recovery}
+
+	we generate two random dictionaries $\textbf{A}^{(1)}$ and $\textbf{A}^{(2)}$ from an i.i.d. zero mean Gaussian random process, and calculate the known dictionary $\textbf{A}$
+	\begin{equation}\label{eq:4_eq03}
+		\textbf{A} = (\textbf{A}^{(1)} \otimes \textbf{A}^{(2)}).
+	\end{equation}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Motivation: Tensor-based Dictionary Recovery
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Motivation: Tensor-based Dictionary Recovery}
+
+	Generate a synthetic data set $\textbf{X}$ from $\textbf{A}$, estimate the initial $\hat{\textbf{A}}$ and make the dictionary separable decomposing the approximation of $\hat{\textbf{A}}$ and generating $\hat{\textbf{A}}^{(1)}$ and $\hat{\textbf{A}}^{(2)}$
+	\begin{equation}\label{eq:4_eq04}
+		\hat{\textbf{A}} = (\hat{\textbf{A}}^{(1)} \otimes \hat{\textbf{A}}^{(2)}),
+	\end{equation}
+	and use it for DL recovery experiments using MOD, K-SVD, RLS-DLA, T-MOD and K-HOSVD.
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Motivation: Tensor-based Dictionary Recovery
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Motivation: Tensor-based Dictionary Recovery}
+	\begin{figure}[!htb]
+		\centering 
+		\includegraphics[width=10cm]{../figures/ch4/s=5_snr=20_L=2000_noIt=100_N=20_K=50.eps}
+		\caption{Cumulative atom identification per degree rates. $T=2000$, $M=20$, $N=50$}
+		\label{fig:fig1}
+	\end{figure}
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Motivation: Tensor-based Dictionary Recovery
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Motivation: Tensor-based Dictionary Recovery}
+	\begin{figure}[!htb]
+		\centering 
+		\includegraphics[width=10cm]{../figures/ch4/5_20_2000_24000_100.eps}
+		\caption{Cumulative atom identification per degree rates. $T=2000$, $M=80$, $N=300$}
+		\label{fig:fig2}
+	\end{figure}
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Related Works
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Related Works}
+
+	\begin{itemize}
+		\item Wright \emph{et al} [93] have proposed a face \textbf{recognition method} that uses training face images as dictionaries and an $l_1$ sparse optimization method in the classification stage.
+The authors show that the recognition task can be \textbf{successfully} accomplished even using \textbf{random features};
+		\item Roemer \emph{et al} [78] have proposed a \textbf{tensor-based multidimensional extensions for MOD and K-SVD algorithms} and show their improved performances numerically, for the dictionary reconstruction problem;
+		\item To the best of our knowledge \textbf{there is no similar model} applied to \textbf{fraud detection}.
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Data Model
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Data Model}
+
+	\begin{itemize}
+		\item We adopt a synthetic financial datasets for fraud detection generated by PaySim;
+		\begin{itemize}
+			\item 9 features (variables);
+			\item 2.770.408 transactions(observations);
+			\item labels to classify fraudulent or legitimate transactions;
+			\item 2.762.195 of normal transactions.
+			\item 8.213 fraudulent transactions(0,30 \%);
+		\end{itemize}
+	\end{itemize}
+
+	Training data as $\textbf{X}_r$,  test data as $\textbf{X}_e$.
+
+	Normal train data is $\textbf{X}_r^o$ and fraud train data is $\textbf{X}_r^f$.
+
+	The same notation is applied to qualify $\textbf{A}$ and $\textbf{S}$.
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Tensor-based Sparse Representation Classification
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Tensor-based Sparse Representation Classification}
+
+	\begin{itemize}
+		\item Split the training data into fraud and normal ($\textbf{X}_r^f$ and $\textbf{X}_r^o$);
+		\item Estimate the initial dictionaries $\hat{\textbf{A}}_r^f$ and $\hat{\textbf{A}}_r^o$;
+		\item Apply the dictionaries and the data matrices to the selected DL algorithms for learning the best dictionaries $\hat{\textbf{A}}_r^f$ and $\hat{\textbf{A}}_r^o$;
+		\item Process each transaction of $\textbf{X}_e$, denoted as $\textbf{x}_t$, to generate  $\hat{\textbf{s}}_t^f$ and $\hat{\textbf{s}}_t^o$;
+		\item Compute signal reconstruction for each test transaction $t$
+			\begin{equation}\label{eq:4_eq041}
+				\hat{\textbf{x}}_t^f = \hat{\textbf{A}}_r^f \cdot \hat{\textbf{s}}_t^f
+			\end{equation}
+			\begin{equation}\label{eq:4_eq042}
+				\hat{\textbf{x}}_e^o = \hat{\textbf{A}}_r^o \cdot \hat{\textbf{s}}_e^o
+			\end{equation}
+		\item Compare $\hat{\textbf{x}}_t^f$ and $\hat{\textbf{x}}_t^o$ to $\textbf{x}_t$ to classify the transaction as fraud or normal according to the minimum reconstruction error.
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Experiments
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Experiments}
+
+	\begin{itemize}
+		\item We propose to evaluate the \textbf{K-HOSVD} to analyze how a tensor-based DL algorithm can highlight the \textbf{discriminative structure for fraud detection};
+		\item We compare the proposed approach against the \textbf{Support Vector Machines (SVM) Logistic Regression (LR) using PCA data};
+		\item We adopt the area under the Receiver Operating Characteristic (\textbf{ROC}) curve and the area under the Precision-Recall (\textbf{PR}) curve.
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Results}
+
+	\begin{table}[h!]
+	  \centering
+	  \scriptsize
+	  \caption{Results for fraud detection from mobile money transactions} 
+	  \label{tab:4_tab1}
+	  \begin{tabular}{ c c c c c }
+		\toprule
+		\multirow{2}{*}{\textbf{Classifier}}   &\multicolumn{2}{c}{\textbf{undersample/undersample}} &\multicolumn{2}{c}{\textbf{undersample/complete}}\\ 
+				\hhline{~----}
+				&\textbf{ROC-AUC} &\textbf{PR-AUC} &\textbf{ROC-AUC} &\textbf{PR-AUC}\\
+		\midrule
+		SVM &0.5039 &0.7493 &0.8496 &\color{red}0.8497 \\
+		LR with PCA &\color{red}0.9767 &\color{red}0.9793 &0.5633 &0.0051 \\
+		K-HOSVD ($K=3$, $N=40$) &0.8622 &0.8917 &\color{red}0.8652 &0.4578 \\
+	    \bottomrule
+	  \end{tabular}
+	\end{table}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Results}
+
+	\begin{itemize}
+		\item SVM presents the best PR-AUC for generalization, but its PR-AUC was the worst for training and testing with undersampled data;
+		\item LR with PCA data presents the best results for training and testing with undersampled data, but the complete testing data presents a very low PR-AUC;
+		\item K-HOSVD presents higher PR-AUC than SVM for undersampled data, but its results are worst than LR with PCA for the same scenario;
+		\item K-HOSVD presents better PR-AUC than LR with PCA and the best ROC-AUC for training the undersampled data and test with complete data.
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Results
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Results}
+
+	Our approach presents average results if compared to two well known and highly adopted classifier algorithms for fraud detection from mobile money transactions.
+
 \end{frame}
@@
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Introduction
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Introduction}
-	\begin{itemize}
-		\item We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
-		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
-		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
-		\item Eigen similarity analysis for extracting detailed time and network ports under attack;
-		\item We conduct a performance evaluation for complexity analysis and processing time.
-	\end{itemize}
+%	FRAME: Conclusion
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Conclusion}
+
+	\begin{itemize}
+		\item Synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis;
+		\item This thesis presented a proposal to address the offline mobile security problem through malicious behavior analysis. The solution has positive results in terms of performance;
+		\item A tensor-based approach can be used for fraud detection from mobile money transactions, with average results
+	\end{itemize}
+
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Contributions
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}[c]{Contributions}
+% 	\begin{enumerate}
+% 		\item We proposed an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset;
+% 		\item We discussed the computational complexity of the proposed framework and evaluated the required processing time for tested scenarios;
+% 		\item We proposed an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
+% 		\item We discussed the processing time of the proposed framework for mobile devices.
+% 	\end{enumerate}
+% \end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Contributions
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Contributions}
+	\begin{enumerate}
+		% \item We proposed a tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
+		\item We published the following papers reporting our results:
+		\begin{enumerate}
+			\item T.
+P.
+B.
+Vieira, D.
+F.
+Ten\'orio, J.
+P.
+C.
+L.
+da Costa, E.
+P.
+de Freitas, G.
+Del Galdo, and R.
+T.
+de Sousa J\'unior, \textit{Model Order Selection and Eigen Similarity based Framework for Detection and Identication of Network Attacks}.
+Journal of Networking and Computer Applications (JNCA), Vol 90, Jul 2017, Pages 26–41.
+			\item T.
+Galibus, T.
+P.
+B.
+Vieira, E.
+P.
+de Freitas, R.
+Albuquerque, J.
+P.
+C.
+L.
+da Costa, R.
+T.
+de Sousa Jr, V.
+Krasnoproshin, A.
+Zaleski, H.
+Vissia, and G.
+del Galdo, \textit{Offline Mode for Corporate Mobile Client Security Architecture}.
+Mobile Networks and Applications, Springer, Mar 2017, Pages 1-17.
+		\end{enumerate}
+	\end{enumerate}
+\end{frame}
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Future Work
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[c]{Future Work}
+	\begin{enumerate}
+		\item To adopt one dictionary with basis vectors that represent legitimate and fraudulent data, as a concatenation or as a kronecker product;
+		\item To adopt discriminative discriminative functions for T-MOD and K-HOSVD, such as the Fisher discrimination criterion;
+		\item To extend the evaluation for other DL algorithms.
+		% \item Improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets;
+		% \item Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high throughput network traffic;
+		% \item Evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis;
+		% \item Explore enhancements in the analytics methods as well as to extended the approach to be used by mobile devices with even more severe resources constraints;
+	\end{enumerate}
 \end{frame}

### 2017-11-29 docs/thesis/slides/apresentacao_unb.tex
reduced full version

@@
 
+\usepackage[backend=bibtex, style=numeric, defernumbers=true]{biblatex}
+\addbibresource{../references.bib}
+
 \DeclarePairedDelimiter\abs{\lvert}{\rvert}%
 \DeclarePairedDelimiter\norm{\lVert}{\rVert}%
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
 
@@
 \subtitle{\small{Ph.D. Thesis Qualification}}
-\date{\today}
 \author{\texttt{Thiago P. de B. Vieira \newline \textbf{Advisor:} João Paulo C. L. da Costa} }
-\institute{Universidade de Brasília 
-	\par \scriptsize{Departamento de Engenharia Elétrica - ENE/FT}
-	\par \scriptsize{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}
-}
-
-\hypersetup{
-	pdfauthor = {Thiago P. de B. Vieira},
-	pdfsubject = {Discriminative Sensing Based on Signal Processing for Information Security Analysis},
-	pdfkeywords = {Signal Processing, Information Security Analysis},
-	pdfmoddate= {/home/thiago/dev/projects/discriminative-sensing/docs/thesis/slides},
-	pdfcreator = {Sublime and Latex}
-}
+\institute{
+		\scriptsize{Universidade de Brasília}
+	\\ 	\scriptsize{Departamento de Engenharia Elétrica - ENE/FT}
+	\\ \scriptsize{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}}
 
@@
 \begin{frame}[c]{Motivation}
-	It is necessary investments to detect \textbf{novel attacks and their variations}
-	\begin{itemize}
-		\item Signal processing schemes have being applied to detect malicious traffic;		
-	\end{itemize}
-	Secure mobile cloud architecture is challenging, specially in \textbf{offline mode}
-	\begin{itemize}
-		\item Signal processing can be applied to \textbf{behavioral analysis of mobile apps in offline mode};		
-	\end{itemize}
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Motivation
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Motivation}
-	\textbf{Fraud detection} of \textbf{Mobile Money Transactions} has been motivating investments
-	\begin{itemize}
-		\item The \textbf{extraction of relevant information} from \textbf{raw or big} data is useful for classification problems and for security analysis.;
-		\item Imbalanced data is challenging for classification related to anomaly detection, novelty detection, fraud detection and attack detection;
-		\item \textbf{Tensor-based} algorithms for \textbf{dictionary learning} can improve \textbf{discriminative sensing} for classification in information security analysis;
+	\begin{itemize}
+		\item It is necessary investments to detect \textbf{novel attacks and their variations};
+		\item Secure mobile cloud architecture is challenging, specially in \textbf{offline mode}
+		\item \textbf{Fraud detection} of \textbf{Mobile Money Transactions} has been motivating investments
+		\begin{itemize}
+			\item The \textbf{extraction of relevant information} from \textbf{raw or big data} is useful for classification problems and for security analysis.;
+			\item \textbf{Imbalanced data is challenging} for classification related to anomaly detection, novelty detection, fraud detection and attack detection.
+		\end{itemize}
 	\end{itemize}
@@
 \begin{frame}[c]{Problem Statement}
-	\textbf{This thesis} outlines the development and evaluation of approaches based on \textbf{signal processing for information security analysis}...
-	\begin{itemize}
-		\item To make the data \textbf{discriminative} for network attack detection, mobile malicious behavior analysis and fraud detection in mobile money transactions.
-	\end{itemize}
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Problem Statement
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Problem Statement}
-	We propose an approach based on \textbf{signal processing techniques for detection of malicious traffic in computer networks}
-	\begin{itemize}
-		\item Signal processing formulation as legitimate traffic, malicious traffic and noise;
-		\item MOS and eigenvalue analysis are applied to network attack detection;
-		\item Eigen similarity analysis identify accurate time and network ports under attack;
-		\item Evaluation through experimental scenario and the DARPA 1998 dataset.
-	\end{itemize}
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Problem Statement
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Problem Statement}
-	This thesis also proposes an \textbf{approach malicious behavior detection in mobile apps}
-	\begin{itemize}
-		\item An architecture for mobile security analysis;
-		\item Offline user behavior analysis based on MOS;
-		\item Scenario analysis for feature selection;
-		\item Processing time evaluation in mobile devices.
-	\end{itemize}
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Problem Statement
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Problem Statement}
-	We also propose a \textbf{tensor-based dictionary learning method for fraud detection from imbalanced data} 
-	\begin{itemize}
-		\item To evidentiate the \textbf{discriminative sensing of a fraud detection dataset};
-		\item Sparse representation based classification (SRC) through learning a tensor-based dictionary;
-		\item Apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate;
-		\item Classify fraud classification of mobile money transactions according to the reconstruction error.
-	\end{itemize}
-\end{frame}
+	\textbf{This thesis} outlines the \textbf{development} and evaluation of approaches based on \textbf{signal processing for information security analysis}...
+	\begin{itemize}
+		\item To make the data \textbf{discriminative for network attack detection, mobile malicious behavior analysis and fraud detection.}
+	\end{itemize}
+\end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Problem Statement
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}[c]{Problem Statement}
+% 	We propose an approach for \textbf{identification of malicious traffic in computer networks}
+% 	\begin{itemize}
+% 		\item We use\textbf{largest eigenvalues and MOS} to \textbf{network attack detection};
+% 		\item We propose \textbf{Eigen similarity analysis} identify \textbf{accurate time and ports} under attack;
+% 	\end{itemize}
+% \end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Problem Statement
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}[c]{Problem Statement}
+% 	This thesis also proposes an \textbf{approach for malicious behavior detection in mobile apps}
+% 	\begin{itemize}
+% 		\item An \textbf{architecture} for mobile security analysis;
+% 		\item \textbf{Offline user behavior analysis} based on MOS;
+% 		\item Scenario analysis for \textbf{feature selection};
+% 		\item \textbf{Processing time} evaluation in \textbf{mobile} devices.
+% 	\end{itemize}
+% \end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Problem Statement
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}[c]{Problem Statement}
+% 	We finally propose a \textbf{tensor-based dictionary learning method for fraud detection} 
+% 	\begin{itemize}
+% 		\item Sparse representation based \textbf{classification} (SRC) through learning a \textbf{tensor-based} dictionary;
+% 		\item Apply the \textbf{learned dictionaries to reconstruct} a test signal;
+% 		\item \textbf{Fraud classification} of mobile money transactions according to the \textbf{reconstruction error}.
+% 	\end{itemize}
+% \end{frame}
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
@@
 	\begin{itemize}
-		\item We model the network traffic as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
-		\item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
+		\item The network traffic is modeled as a signal processing formulation (\textbf{legitimate traffic, malicious traffic and noise});
+		% \item The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis;
 		\item MOS and eigenvalue analysis are applied to detect time frames under attack;
-		\item Eigen similarity analysis for extracting detailed time and network ports under attack;
-		\item We conduct a performance evaluation for complexity analysis and processing time.
+		\item We \textbf{propose Eigen similarity analysis} for extracting detailed time and network ports under attack;
+		\item We \textbf{conduct a performance evaluation} for complexity analysis and processing time.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Classical methods typically employ data mining [39, 34, 70] and regular file analysis [76] to detect patterns that indicate the presence of specific attacks in network traffic;
-		\item Callegari \emph{et al} [104] propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
-However, Callegari \emph{et al} \textbf{focus on flood attack detection}, not addressing probe attack detection, and their approach relies on \textbf{visual analysis}.
+		% \item \textbf{Classical methods typically employ data mining} \cite{he2008applying,ghourabi2010data,osanaiye2016distributed} and regular file analysis \cite{raynal2004honeypot} to detect patterns that indicate the presence of specific attacks in network traffic;
+		\item Callegari \emph{et al} \cite{Zonglin2009} propose a \textbf{PCA-based} method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
+However, Callegari \emph{et al} \textbf{focus on flood attack detection, not addressing probe} attack detection, and their approach relies on \textbf{visual analysis}.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Lee \emph{et al.} [61] propose OverSampling PCA, determines the \textbf{anomaly} according to the variation of \textbf{eigenvector by similarity analysis and over sampling}.
+		\item Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA, determines the \textbf{anomaly} according to the variation of \textbf{eigenvector by similarity analysis and over sampling}.
 We apply \textbf{MOS for attack detection} and \textbf{similarity analysis for identification} of time and ports under attack;
-		\item Lu and Ghorbani [63] proposed a \textbf{network anomaly} detection model based on \textbf{network flow, wavelet approximation, and system identification theory}.
+		\item Lu and Ghorbani \cite{Lu2009} proposed a \textbf{network anomaly} detection model based on \textbf{network flow, wavelet approximation, and system identification theory}.
 However, their work requires a \textbf{training} and presents \textbf{limitations} on identification of behaviors without \textbf{significant outliers}.
@@
 
-	We focus on \textbf{probe} and \textbf{flooding} attacks;
-
-	Two scenarios: 
+	We focus on \textbf{probe} and \textbf{flooding} attacks for two scenarios: 
 	\begin{itemize}
@@
 
-	Network traffic:
-
+	Network traffic as a signal superposition matrix:
 	\begin{equation}\label{eq:eq01}
@@
 
+	The data is divided into $Q$ time frames.
+
 	$\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ has \emph{M} rows and \emph{N} columns.
 
-	Each row is a TCP or UDP port, and each column represents time bins (minute)
+	Each \textbf{row} is a \textbf{TCP or UDP port}, and each \textbf{column} represents \textbf{time} bins (minute).
 
@@
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Synthetic Dataset}
-
-	\begin{figure}[h!]
-	     \centering 
-	     \includegraphics[width=9cm]{../figures/fig04.png}
-	     \caption{Network traffic of user independent operations for network management.}
-	     \label{fig:2_fig4}
-	\end{figure}
-
-\end{frame}
+% \begin{frame}[c]{Synthetic Dataset}
+
+% 	\begin{figure}[h!]
+% 	     \centering 
+% 	     \includegraphics[width=9cm]{../figures/fig04.png}
+% 	     \caption{Network traffic of user independent operations for network management.}
+% 	     \label{fig:2_fig4}
+% 	\end{figure}
+
+% \end{frame}
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
@@
 \end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Synthetic Dataset
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Synthetic Dataset}
-
-	\begin{figure}[h!]
-	     \centering 
-	     \includegraphics[width=9cm]{../figures/fig06.png}
-	     \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
-	     \label{fig:2_fig6}
-	\end{figure}
-
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Synthetic Dataset
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Synthetic Dataset}
-
-	\begin{figure}[h!]
-	     \centering 
-	     \includegraphics[width=9cm]{../figures/fig07.png}
-	     \caption{Connection attempts in order to identify active ports.}
-	     \label{fig:2_fig7}
-	\end{figure}
-
-\end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Synthetic Dataset
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}[c]{Synthetic Dataset}
+
+% 	\begin{figure}[h!]
+% 	     \centering 
+% 	     \includegraphics[width=9cm]{../figures/fig06.png}
+% 	     \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
+% 	     \label{fig:2_fig6}
+% 	\end{figure}
+
+% \end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Synthetic Dataset
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}[c]{Synthetic Dataset}
+
+% 	\begin{figure}[h!]
+% 	     \centering 
+% 	     \includegraphics[width=9cm]{../figures/fig07.png}
+% 	     \caption{Connection attempts in order to identify active ports.}
+% 	     \label{fig:2_fig7}
+% 	\end{figure}
+
+% \end{frame}
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
@@
 	\begin{itemize}
-		\item 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks;
-		\item The most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack;
-		\item We select the cases that reproduce probe and flooding attacks.
+		\item \textbf{7 weeks} of sniffed traffic saved into raw TCPDUMP packet data, from \textbf{inside and outside} origins, with \textbf{labeled} attacks;
+		\item The most cases of \textbf{DoS focus on exploit} system vulnerabilities instead of on \textbf{flooding} attack;
+		\item We select the cases that reproduce \textbf{probe and flooding} attacks.
 	\end{itemize}
@@
 \end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Proposed Framework
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Proposed Framework}
-
-	For \textbf{flooding} detection, calculate the \textbf{sample covariance matrix} $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
-	\begin{equation}\label{eq:eq02}
-		\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
-	\end{equation}
-
-	the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
-	\begin{equation}\label{eq:eq03}
-		\boldsymbol{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
-	\end{equation}
-
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Proposed Framework
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Proposed Framework}
-
-	For \textbf{probing} detection, compute the \textbf{sample covariance} $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have \textbf{zero mean and unitary standard deviation} as follows
-	\begin{equation}\label{eq:eq04}
-		\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
-	\end{equation}
-
-	The sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)}$ can be calculated via 
-	\begin{equation}\label{eq:eq05}
-		\boldsymbol{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
-	\end{equation}
-
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Proposed Framework
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Proposed Framework}
-
-	we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{C}$
-
-	Compute eigenvalue decomposition (EVD) to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix (time frame $q$), according to (\ref{eq:eq060}).
-	\begin{equation}\label{eq:eq06}
-	\boldsymbol{C}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
-	\end{equation}
-	\begin{equation}\label{eq:eq060}
-	\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
-	\end{equation}
-
-	The eigenvalues should be sorted in descending order, i.e., $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ... > \lambda_{m}^{(q)}$.
-
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Proposed Framework
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Proposed Framework}
-
-	\begin{equation}\label{eq:eq07}
-		\boldsymbol{E} =
-		\begin{bmatrix}
-			\color{red}\lambda_1^{(1)} & \color{red}\lambda_1^{(2)} & \color{red}\lambda_1^{(3)} & \cdots & \color{red}\lambda_1^{(Q)} \\
-			\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
-			\lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
-			\vdots & \vdots & \ddots & \vdots  \\
-			\lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
-		\end{bmatrix}.
-	\end{equation}
-
-	\begin{equation}\label{eq:eq08}
-		\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
-	\end{equation}
-
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: MOS Scheme
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{MOS Scheme}
-
-	\begin{itemize}
-		\item Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$;
-		\item $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
-		\item $\hat{d}$ is the estimated number of time frames under attack.
-	\end{itemize}
-
-\end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Proposed Framework
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}{Proposed Framework}
+
+% 	For \textbf{flooding} detection, calculate the \textbf{sample covariance matrix} $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
+% 	\begin{equation}\label{eq:eq02}
+% 		\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
+% 	\end{equation}
+
+% 	the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
+% 	\begin{equation}\label{eq:eq03}
+% 		\boldsymbol{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
+% 	\end{equation}
+
+% \end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Proposed Framework
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}{Proposed Framework}
+
+% 	For \textbf{probing} detection, compute the \textbf{sample covariance} $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have \textbf{zero mean and unitary standard deviation} as follows
+% 	\begin{equation}\label{eq:eq04}
+% 		\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
+% 	\end{equation}
+
+% 	The sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)}$ can be calculated via 
+% 	\begin{equation}\label{eq:eq05}
+% 		\boldsymbol{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
+% 	\end{equation}
+
+% \end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Proposed Framework
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}{Proposed Framework}
+
+% 	we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{C}$
+
+% 	Compute eigenvalue decomposition (EVD) to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix (time frame $q$), according to (\ref{eq:eq060}).
+% 	\begin{equation}\label{eq:eq06}
+% 	\boldsymbol{C}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
+% 	\end{equation}
+% 	\begin{equation}\label{eq:eq060}
+% 	\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
+% 	\end{equation}
+
+% 	The eigenvalues should be sorted in descending order, i.e., $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ... > \lambda_{m}^{(q)}$.
+
+% \end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Proposed Framework
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}{Proposed Framework}
+
+% 	\begin{equation}\label{eq:eq07}
+% 		\boldsymbol{E} =
+% 		\begin{bmatrix}
+% 			\color{red}\lambda_1^{(1)} & \color{red}\lambda_1^{(2)} & \color{red}\lambda_1^{(3)} & \cdots & \color{red}\lambda_1^{(Q)} \\
+% 			\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
+% 			\lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
+% 			\vdots & \vdots & \ddots & \vdots  \\
+% 			\lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
+% 		\end{bmatrix}.
+% 	\end{equation}
+
+% 	\begin{equation}\label{eq:eq08}
+% 		\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
+% 	\end{equation}
+
+% \end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: MOS Scheme
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}{MOS Scheme}
+
+% 	\begin{itemize}
+% 		\item Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$;
+% 		\item $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
+% 		\item $\hat{d}$ is the estimated number of time frames under attack.
+% 	\end{itemize}
+
+% \end{frame}
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
@@
 
-	Through eigenvalue analysis, we identify the time frames $q$ under attack
+	\textbf{MOS} reveal the \textbf{number of attacks} and through \textbf{eigenvalue analysis} we identify the \textbf{time frames} $q$ under attack
 
@@
 
-	The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic without attack.
+	\textbf{Eigen Similarity} Analysis attempts to identify \textbf{time and ports}.
+
+	The \textbf{reference eigenvectors} $\boldsymbol{v}^{(q)}$ is calculated from the \textbf{traffic without attack}.
 
@@
 
-	The resultant $\boldsymbol{X}_{(n)}$ is necessary to obtain $\boldsymbol{v}_{(n)}$, through (\ref{eq:eq06}).
+	The resultant $\boldsymbol{X}_{(n)}$ is necessary to obtain $\boldsymbol{v}_{(n)}$
 
@@
 
-	For eigen similarity analysis, we evaluate the cosine similarity to identify legitimate and malicious traffic,
+	we evaluate the cosine similarity to identify legitimate and malicious traffic,
 	\begin{equation}
@@
 
-	We evaluate \textbf{three approaches} for eigen similarity analysis: 
+	We propose and evaluate \textbf{three approaches} for eigen similarity analysis: 
 	\begin{itemize}
@@
 
-	For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack is compared against the $\boldsymbol{v}_{(n)}$ identified as under attack,
+	\textbf{For detection of ports under attack}, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack is compared against the $\boldsymbol{v}_{(n)}$ identified as under attack,
 
@@
 \end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Results
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Results}
-
-	\begin{table}[h!]
-	  \centering
-	  \scriptsize
-	  \caption{Largest Eigenvalue related to attacks detection}
-	  \label{tab:tab3}
-	  \begin{tabular}{ c c c c c }
-		\toprule
-		\multirow{3}{*}{\textbf{Time Frame} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
-				\hhline{~----}
-			&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
-			&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
-		\midrule
-		1 &1887545 &1887545 &1887545 &2,0734 \\
-		2 &2341327 &2341327 &2341327 &2,1451 \\
-		3 &3213867 &3213867 &3213867 &\color{red}10,0718 \\
-		4 &\color{red}133238294 &\color{red}133238294 &731229 &2,1620 \\
-		5 &\color{red}92384021611 &6367983 &\color{red}92384021611 &2,4253 \\
-		6 &708335 &708335 &708335 &1,7948 \\
-	    \bottomrule
-	  \end{tabular}
-	\end{table}
-
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Results
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Results}
-
-	\begin{table}[h!]
-	  \centering
-	  \tiny
-	  \caption{MOS schemes applied to port scan and flood detection}
-	  \label{tab:tab4}
-	  \begin{tabular}{ c c c c c c c c }
-		\toprule
-		\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
-				\hhline{~------~}
-			&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
-		\midrule
-		Detection of synflood \\(presence of attack) &2 &1 &\textbf{\color{red}1} &5 &\textbf{\color{red}1} &4 &\textbf{\color{red}1} \\
-		Detection of synflood \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
-		\midrule
-		Detection of fraggle \\(presence of attack) &1 &1 &\textbf{\color{red}1} &5 &\textbf{\color{red}1} &4 &\textbf{\color{red}1} \\
-		Detection of fraggle \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
-		\midrule
-		Detection of port scan \\(presence of attack) &1 &1 &\textbf{\color{red}1} &1 &\textbf{\color{red}1} &9 &\textbf{\color{red}1} \\
-		Detection of port scan \\(absence of attack) &0 &0 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} \\
-		\midrule
-		Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{\color{red}2} &5 &\textbf{\color{red}2} &5 &\textbf{\color{red}2} \\
-		Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
-	    \bottomrule
-	  \end{tabular}
-	\end{table}
-
-\end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Results
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Results}
-
-	MOS schemes for probing and flooding attack detection:
-	\begin{itemize}
-		\item The \textbf{largest eigenvalues visually} indicate the time frame under probing and flooding attack detection;
-		\item MOS schemes can estimate it mathematically;
-		\begin{itemize}
-			\item \textbf{EDC and EFT} estimate \textbf{correctly} the number of attacks $d$;
-		\end{itemize}
-	\end{itemize}
-
-\end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Results
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}{Results}
+
+% 	\begin{table}[h!]
+% 	  \centering
+% 	  \scriptsize
+% 	  \caption{Largest Eigenvalue related to attacks detection}
+% 	  \label{tab:tab3}
+% 	  \begin{tabular}{ c c c c c }
+% 		\toprule
+% 		\multirow{3}{*}{\textbf{Time Frame} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
+% 				\hhline{~----}
+% 			&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
+% 			&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
+% 		\midrule
+% 		1 &1887545 &1887545 &1887545 &2,0734 \\
+% 		2 &2341327 &2341327 &2341327 &2,1451 \\
+% 		3 &3213867 &3213867 &3213867 &\color{red}10,0718 \\
+% 		4 &\color{red}133238294 &\color{red}133238294 &731229 &2,1620 \\
+% 		5 &\color{red}92384021611 &6367983 &\color{red}92384021611 &2,4253 \\
+% 		6 &708335 &708335 &708335 &1,7948 \\
+% 	    \bottomrule
+% 	  \end{tabular}
+% 	\end{table}
+
+% \end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Results
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}{Results}
+
+% 	\begin{table}[h!]
+% 	  \centering
+% 	  \tiny
+% 	  \caption{MOS schemes applied to port scan and flood detection}
+% 	  \label{tab:tab4}
+% 	  \begin{tabular}{ c c c c c c c c }
+% 		\toprule
+% 		\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
+% 				\hhline{~------~}
+% 			&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
+% 		\midrule
+% 		Detection of synflood \\(presence of attack) &2 &1 &\textbf{\color{red}1} &5 &\textbf{\color{red}1} &4 &\textbf{\color{red}1} \\
+% 		Detection of synflood \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
+% 		\midrule
+% 		Detection of fraggle \\(presence of attack) &1 &1 &\textbf{\color{red}1} &5 &\textbf{\color{red}1} &4 &\textbf{\color{red}1} \\
+% 		Detection of fraggle \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
+% 		\midrule
+% 		Detection of port scan \\(presence of attack) &1 &1 &\textbf{\color{red}1} &1 &\textbf{\color{red}1} &9 &\textbf{\color{red}1} \\
+% 		Detection of port scan \\(absence of attack) &0 &0 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} \\
+% 		\midrule
+% 		Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{\color{red}2} &5 &\textbf{\color{red}2} &5 &\textbf{\color{red}2} \\
+% 		Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
+% 	    \bottomrule
+% 	  \end{tabular}
+% 	\end{table}
+
+% \end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Results
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}{Results}
+
+% 	MOS schemes for probing and flooding attack detection:
+% 	\begin{itemize}
+% 		\item The \textbf{largest eigenvalues visually} indicate the time frame under probing and flooding attack detection;
+% 		\item MOS schemes can estimate it mathematically;
+% 		\begin{itemize}
+% 			\item \textbf{EDC and EFT} estimate \textbf{correctly} the number of attacks $d$;
+% 		\end{itemize}
+% 	\end{itemize}
+
+% 	It is still necessary \textbf{to identify} the \textbf{times and ports} under attack.
+
+% \end{frame}
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
@@
 	\begin{itemize}
-		\item High FP and misclassification due to legitimate traffic be massive (such as an attack) sometimes;
-		\item Callegari \emph{et al} is a statistical method, based on PCA, without training or learning methods
+		\item \textbf{High FP and misclassification} due to \textbf{massive legitimate traffic} (such as an attack) sometimes;
+		\item Callegari \emph{et al} is based on PCA, without training
 		\begin{itemize}
 			\item 82 \% while we obtain 100 \%;
-			\item FP and misclassification was not evaluated.
-		\end{itemize}
-		\item Ghorbani's is based on signal processing techniques and uses DARPA
+			\item \textbf{FP and misclassification was not evaluated}.
+		\end{itemize}
+		\item Ghorbani's is based on signal processing and uses DARPA
 		\begin{itemize}
 			\item 94.67 \% and 50.00 \%, while we obtain 100 \% and 76.92 \%;
-			\item FP and misclassification was not evaluated.
+			\item \textbf{FP and misclassification was not evaluated}.
 		\end{itemize}
@@
 	\begin{itemize}
-		\item the proposed framework is $O(N^3 + Q \log Q + \hat{d}Q + N^3)$ and its worst-case running time is $O(N^3)$;
+		\item The worst-case running time is $O(N^3)$;
 		\item The computational complexity of \textbf{EVD is predominant};
@@
 \end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: Performance Evaluation
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}{Performance Evaluation}
-
-	Processing Time Analysis:
-	\begin{itemize}
-		\item desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM;
-		\item Network traffic time; 
-		\item Frame size denoted as $N$; 
-		\item Number of network ports denoted as $M$; 
-		\item Mean processing time for eigen analysis based on sample covariance of zero mean (\textbf{1-time}); 
-		\item Mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation (\textbf{2-time}); 
-		\item Mean processing time for EDC MOS scheme (\textbf{3-time}).
-	\end{itemize}
-
-\end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: Performance Evaluation
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}{Performance Evaluation}
+
+% 	Processing Time Analysis:
+% 	\begin{itemize}
+% 		% \item desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM;
+% 		\item Network traffic time; 
+% 		\item Frame size denoted as $N$; 
+% 		\item Number of network ports denoted as $M$; 
+% 		\item Mean processing time for eigen analysis based on sample covariance of zero mean (\textbf{1-time}); 
+% 		\item Mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation (\textbf{2-time}); 
+% 		\item Mean processing time for \textbf{EDC MOS}	 scheme (\textbf{3-time}).
+% 	\end{itemize}
+
+% \end{frame}
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
@@
 	\begin{itemize}
-		\item The processing time increases according to traffic time, but the \textbf{worst} processing time is \textbf{4.7250 milliseconds};
-		\item The processing time \textbf{increases} with the frame size $N$ \textbf{decreasing};
+		\item The processing time \textbf{increases according to traffic time}, but the \textbf{worst} processing time is \textbf{4.7250 milliseconds};
+		% \item The processing time \textbf{increases} with the frame size $N$ \textbf{decreasing};
 		\item The number of ports $M$ increases the processing time.
@@
 \begin{frame}[c]{Introduction}
+	In this work:
 	\begin{itemize}
@@
 	\begin{itemize}
-		\item According to [53, 55], the \textbf{state-of-the-art mobile cloud security models do not consider} the problem of the offline security mode;
+		\item According to \cite{khan2013towards, khan2015cloud}, the \textbf{state-of-the-art mobile cloud security models do not consider} the problem of the \textbf{offline security mode};
 		\item Providers of the mobile security \textbf{avoid} the problem by completely \textbf{forbidding} the offline access.
@@
 \end{frame}
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-%	FRAME: The Client-side Architecture
-%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{The Client-side Architecture}
-
-	\begin{figure}[h!]
-		\centering
-		\includegraphics[width=9cm]{../figures/fig02.eps}
-		\caption{The Client-side Architecture}
-		\label{fig:3_02}
-	\end{figure}
-
-\end{frame}
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% %	FRAME: The Client-side Architecture
+% %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+% \begin{frame}[c]{The Client-side Architecture}
+
+% 	\begin{figure}[h!]
+% 		\centering
+% 		\includegraphics[width=9cm]{../figures/fig02.eps}
+% 		\caption{The Client-side Architecture}
+% 		\label{fig:3_02}
+% 	\end{figure}
+
+% \end{frame}
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
@@
 	\begin{itemize}
-		\item Previously presented;
+		\item Largest eigenvalue analysis and MOS (Previously presented);
 		\item Applied to user behavior analysis;
 		\begin{itemize}
-			\item File operations modeled as signals into matrices;
-			\item Selected features can highlight malicious behaviors;
-			\item Implemented and evaluated into mobile devices.
+			\item We model file operations as signals (matrices);
+			\item We Select features that can highlight malicious behaviors;
+			\item We implement and evaluate into mobile devices.
 		\end{itemize}
@@
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{Data Modeling and Eigenvalue Decomposition Time}
+\begin{frame}[c]{Result: Data Modeling and Eigenvalue Analysis}
 
@@
 %-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-\begin{frame}[c]{EDC MOS scheme processing time for anomaly detection}
+\begin{frame}[c]{Result: EDC MOS}
 
@@
 	\begin{itemize}
-		\item The lower window size leads to the larger eigenvalue decomposition time;
-		\item The largest eigenvalue decomposition time, which was the maximum of 421 milliseconds with average of 347.42 milliseconds;
-		\item The MOS processing time increases with the window size decreasing;
-		\item The longest processing time measured is lower than 200 milliseconds;
-		\item This results represent an acceptable processing time for anomaly detection in mobile clients
+		\item The \textbf{lower window} size leads to the \textbf{larger eigenvalue analysis time};
+		\item The largest eigenvalue analysis time: 421 milliseconds with average of 347.42 milliseconds;
+		\item The \textbf{MOS} processing time \textbf{increases} with the \textbf{window size decreasing};
+		\item The longest MOS processing time: lower than 200 milliseconds;
+		\item This results represent an \textbf{acceptable processing time} for \textbf{anomaly detection in mobile clients}.
 	\end{itemize}
@@
 		\item \textbf{Tensor decompositions} can be useful to extract \textbf{hidden patterns} and structure in \textbf{multidimensional} data analytics problems;
-		\item The \textbf{fundamental issue} with the \textbf{imbalanced learning} problem is the ability of imbalanced data to \textbf{compromise the performance of algorithms};
+		\item The \textbf{fundamental issue} with the \textbf{imbalanced learning} problem is the ability of imbalanced data to \textbf{compromise the performance of algorithms} \cite{he2009learning};
 	\end{itemize}
@@
 	\begin{itemize}
-		\item We propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds; 
-		\item Tensor-based DL for learning fraudulent and legitimate data separately
-		\item Apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate.
+		\item We propose a tensor-based sparse representation and dictionary learning technique to analyze mobile money transactions in order to identify frauds.
+	% 	\item Tensor-based DL for learning fraudulent and legitimate data separately
+	% 	\item Apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate.
 	\end{itemize}
@@
 
-	Consider a generic sparse recovery problem
+	We consider a \textbf{generic sparse recovery problem}
 	\begin{equation}\label{eq:4_eq01}
@@
 
-	Consider a sparse recovery problem for a separable 2-D dictionary that we can write as
+	Based on Roemer \emph{et al} \cite{roemer2014tensor}, we consider a sparse recovery problem for a \textbf{separable 2-D dictionary} that we can write as
 	\begin{equation}\label{eq:4_eq02}
@@
 
-	we generate two random dictionaries $\textbf{A}^{(1)}$ and $\textbf{A}^{(2)}$ from an i.i.d. zero mean Gaussian random process, and calculate the known dictionary $\textbf{A}$
+	We \textbf{generate two random dictionaries} $\textbf{A}^{(1)}$ and $\textbf{A}^{(2)}$ from an i.i.d. zero mean Gaussian random process, and calculate the known dictionary $\textbf{A}$
 	\begin{equation}\label{eq:4_eq03}
@@
 
-	Generate a synthetic data set $\textbf{X}$ from $\textbf{A}$, estimate the initial $\hat{\textbf{A}}$ and make the dictionary separable decomposing the approximation of $\hat{\textbf{A}}$ and generating $\hat{\textbf{A}}^{(1)}$ and $\hat{\textbf{A}}^{(2)}$
+	Generate a synthetic data set $\textbf{X}$ from $\textbf{A}$
+
+	Estimate the initial $\hat{\textbf{A}}$ and make the dictionary separable decomposing the approximation of $\hat{\textbf{A}}$ and generating $\hat{\textbf{A}}^{(1)}$ and $\hat{\textbf{A}}^{(2)}$
 	\begin{equation}\label{eq:4_eq04}
@@
 	\begin{itemize}
-		\item Wright \emph{et al} [93] have proposed a face \textbf{recognition method} that uses training face images as dictionaries and an $l_1$ sparse optimization method in the classification stage.
+		\item Wright \emph{et al} \cite{wright2009robust} have proposed a face \textbf{recognition method} that uses training face images as dictionaries and an $l_1$ sparse optimization method in the classification stage.
 The authors show that the recognition task can be \textbf{successfully} accomplished even using \textbf{random features};
-		\item Roemer \emph{et al} [78] have proposed a \textbf{tensor-based multidimensional extensions for MOD and K-SVD algorithms} and show their improved performances numerically, for the dictionary reconstruction problem;
+		\item Roemer \emph{et al} \cite{roemer2014tensor} have proposed a \textbf{tensor-based multidimensional extensions for MOD and K-SVD algorithms} and show their improved performances numerically, for the dictionary reconstruction problem;
 		\item To the best of our knowledge \textbf{there is no similar model} applied to \textbf{fraud detection}.
@@
 	\begin{itemize}
-		\item Split the training data into fraud and normal ($\textbf{X}_r^f$ and $\textbf{X}_r^o$);
-		\item Estimate the initial dictionaries $\hat{\textbf{A}}_r^f$ and $\hat{\textbf{A}}_r^o$;
-		\item Apply the dictionaries and the data matrices to the selected DL algorithms for learning the best dictionaries $\hat{\textbf{A}}_r^f$ and $\hat{\textbf{A}}_r^o$;
+		\item \textbf{Split} the training data into \textbf{fraud and normal} ($\textbf{X}_r^f$ and $\textbf{X}_r^o$);
+		\item \textbf{Estimate} the initial dictionaries $\hat{\textbf{A}}_r^f$ and $\hat{\textbf{A}}_r^o$;
+		\item Apply the \textbf{dictionaries} and the \textbf{data matrices} to the selected \textbf{DL algorithm} for learning the \textbf{best dictionaries} $\hat{\textbf{A}}_r^f$ and $\hat{\textbf{A}}_r^o$;
 		\item Process each transaction of $\textbf{X}_e$, denoted as $\textbf{x}_t$, to generate  $\hat{\textbf{s}}_t^f$ and $\hat{\textbf{s}}_t^o$;
@@
 	\begin{itemize}
-		\item SVM presents the best PR-AUC for generalization, but its PR-AUC was the worst for training and testing with undersampled data;
-		\item LR with PCA data presents the best results for training and testing with undersampled data, but the complete testing data presents a very low PR-AUC;
-		\item K-HOSVD presents higher PR-AUC than SVM for undersampled data, but its results are worst than LR with PCA for the same scenario;
-		\item K-HOSVD presents better PR-AUC than LR with PCA and the best ROC-AUC for training the undersampled data and test with complete data.
+		\item \textbf{SVM} presents the \textbf{best PR-AUC for generalization}, but its \textbf{PR-AUC} was the \textbf{worst} for training and testing with \textbf{undersampled} data;
+		\item \textbf{LR with PCA} data presents the \textbf{best} results for training and testing with \textbf{undersampled} data, but the \textbf{complete} testing data presents a \textbf{very low PR-AUC};
+		\item \textbf{K-HOSVD} presents \textbf{higher PR-AUC than SVM} for \textbf{undersampled} data, but its results are \textbf{worst than LR with PCA} for the same scenario;
+		\item \textbf{K-HOSVD} presents \textbf{better PR-AUC than LR with PCA} and the \textbf{best ROC-AUC} for training the undersampled data and test with \textbf{complete} data.
 	\end{itemize}
@@
 
-	Our approach presents average results if compared to two well known and highly adopted classifier algorithms for fraud detection from mobile money transactions.
+	Our \textbf{approach} presents \textbf{average results} if compared to two well known and highly adopted \textbf{classifier algorithms for fraud detection} from mobile money transactions.
 
@@
 	\begin{itemize}
-		\item Synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis;
-		\item This thesis presented a proposal to address the offline mobile security problem through malicious behavior analysis. The solution has positive results in terms of performance;
-		\item A tensor-based approach can be used for fraud detection from mobile money transactions, with average results
+		\item \textbf{Synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion}, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis;
+		\item This thesis presented a proposal to address the \textbf{offline mobile security} problem through malicious behavior analysis.
+The solution has \textbf{positive results in terms of performance};
+		\item A \textbf{tensor-based DL} approach can be used for \textbf{fraud detection} from mobile money transactions, with \textbf{average results}.
 	\end{itemize}
@@
 	\begin{enumerate}
-		\item To adopt one dictionary with basis vectors that represent legitimate and fraudulent data, as a concatenation or as a kronecker product;
-		\item To adopt discriminative discriminative functions for T-MOD and K-HOSVD, such as the Fisher discrimination criterion;
-		\item To extend the evaluation for other DL algorithms.
+		\item To adopt one dictionary with basis vectors that represent legitimate and fraudulent data, as a \textbf{concatenation or as a kronecker product};
+		\item To adopt \textbf{discriminative functions} for T-MOD and K-HOSVD, such as the Fisher discrimination criterion;
+		\item To extend the evaluation for other \textbf{DL algorithms}.
 		% \item Improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets;
@@
 
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+%	FRAME: Bibliography
+%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
+\begin{frame}[plain, allowframebreaks]{Bibliography}
+	\setbeamertemplate{bibliography item}[text]
+	\printbibliography
+\end{frame}
+
 \end{document}

### 2018-05-15 docs/thesis/chapters/1_introduction/introduction.tex
new project for adaptive multivariage gaussian applied to network attack detection

@@
 
-The field of Mobile Money Transfer (MMT) is a growing market segment and has been motivating investiments into security defense and fraud detection.
+The field of Mobile Money Transfer (MMT) is a growing market segment and has been motivating investments into security defense and fraud detection.
 Obtaining access to data sets of mobile transactions for research is a very hard task due to the intrinsic private nature of such transactions.
@@
 Considering that imbalanced data can be composed by rare or unknown events, adaptive techniques can be useful in order to improve models and be able to react to changes or unknown patterns and maintain acceptable performance for classification algorithms.
-Adaptive techniques are also important for online data analysis or streaming processing, where an incremental processing approach can be applied to deal with high thoughput data analysis.
+Adaptive techniques are also important for online data analysis or streaming processing, where an incremental processing approach can be applied to deal with high throughput data analysis.
 
-Compressed Sensing (CS) aims to obtain the most relevant information of a dataset, what makes it useful for compression, signal reconstruction, image processing and classification, such as in sparce representation classification (SRC) problems.
+Compressed Sensing (CS) aims to obtain the most relevant information of a dataset, what makes it useful for compression, signal reconstruction, image processing and classification, such as in sparse representation classification (SRC) problems.
 Dictionary learning is related to CS but seeks to obtain compact and meaningful signal representations, known as dictionary, from learning algorithms or models.
@@
 Dictionaries are either available analytically, or can be learned from a suitable training set.
-DL can be formulated as an adptive algorithm that learns from traning data that is updated to represent changes or to adpat to new patterns.
-Iterative dictionary learning has been used as a blind technique for feature extraction, improving classification without require feature selection or principal conponent analysis
+DL can be formulated as an adaptive algorithm that learns from training data that is updated to represent changes or to adapt to new patterns.
+Iterative dictionary learning has been used as a blind technique for feature extraction, improving classification without require feature selection or principal component analysis
 
@@
 
-Considering that big data problems require techniques to deal with multidimensional data in order to make sense of strucutre and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highligh discriminative structure and relationships of one dataset, we also propose a tensor-based dictionary learning method for fraud detecton from imbalanced data, in order to apply the tensor deconposition for dictionary learning methods to evidentiate the discriminative sensing of a fraud detection dataset.
+Considering that big data problems require techniques to deal with multidimensional data in order to make sense of structure and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highlight discriminative structure and relationships of one dataset, we also propose a tensor-based dictionary learning method for fraud detection from imbalanced data, in order to apply the tensor decomposition for dictionary learning methods to evidentiate the discriminative sensing of a fraud detection dataset.
 Specifically, we propose to apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification of mobile money transactions.
-We propose to use tensor-based dictionary learning for learning fraudulent and legitimante data separetelly, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or ligitimate, according to the minimum reconstruction error measured by some metrics.
+We propose to use tensor-based dictionary learning for learning fraudulent and legitimate data separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics.
 
@@
 
-We analyse problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques.
+We analyze problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques.
 The results of the work presented in this thesis provide the following contributions:
@@
 	\item We discuss the computational complexity of the proposed framework and evaluate the required processing time for tested scenarios;
-	\item We propose an architechture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
+	\item We propose an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
 	\item We discuss the processing time of the proposed framework for mobile devices;
@@
 Chapter \ref{ch:5_conclusionfuturework} draws the conclusions and the suggestions for future work.
-Futhermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Compoment Analytis (PCA) for visual discriminant analysis, and presents an approach based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.
+Furthermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Component Analysis (PCA) for visual discriminant analysis, and presents an approach based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of Brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.

### 2018-05-15 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
new project for adaptive multivariage gaussian applied to network attack detection

@@
 The DARPA 1998 dataset\footnote{https://www.ll.mit.edu/ideval/data/} includes 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks.
-The attacks in this dataset
-can be grouped into: denial-of-service (DoS); remote to local (R2L), which is characterized by unauthorized access from a
-remote machine; user to root (U2R), which is characterized by unauthorized access to local
-super-user privileges; and probe attack.
+The attacks in this dataset can be grouped into: denial-of-service (DoS); remote to local (R2L), which is characterized by unauthorized access from a remote machine; user to root (U2R), which is characterized by unauthorized access to local super-user privileges; and probe attack.
 Since the proposed approach focus on flood and probe attack, the analysis concentrates on the attacks that present behaviors similar to flood or probe attack.
@@
 According to flood and port scan attacks' behavior, flood attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively.
-These characteristics are substantiated by the results obtained through the analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:2_experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
+These characteristics are substantiated by the results obtained through the analysis based on sample covariance of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:2_experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
 
@@
 
-Therefore, $\boldsymbol{v}^{(q)}$ should be calculated from the last $\boldsymbol{X}^{(q)}$ time frame without attack, and $\boldsymbol{v}_{(m,\hat{n})}$ should be calculated from the same traffic appened of all $n$-th minutes until the identified minute under attack, denoted as $\boldsymbol{X}_n$.
+Therefore, $\boldsymbol{v}^{(q)}$ should be calculated from the last $\boldsymbol{X}^{(q)}$ time frame without attack, and $\boldsymbol{v}_{(m,\hat{n})}$ should be calculated from the same traffic appended of all $n$-th minutes until the identified minute under attack, denoted as $\boldsymbol{X}_n$.
 
@@
 Table \ref{tab:tab12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}.
-The analysis based on sample covariation of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames.
-The results also show 50.00 \% of misclassification rate , which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
+The analysis based on sample covariance of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames.
+The results also show 50.00 \% of misclassification rate, which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
 The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time from one source to one target, with no variation on IP source or target port.

### 2018-05-15 docs/thesis/chapters/5_conclusion_future_work/conclusion.tex
new project for adaptive multivariage gaussian applied to network attack detection

@@
 The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of this thesis were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
+The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
 
@@
 
-This thesis shows that a tensor-based approach can be used for fraud detection from mobile money transactions, with average results when compared to well known classifiers in one scenario with highly skewed datasets.
+Finally, this thesis shows that a tensor-based approach can be used for fraud detection from mobile money transactions, with average results when compared to well known classifiers in one scenario with highly skewed datasets.
 
@@
 T.
-Sousa Jr., \textit{ANÁLISE MULTIDIMENSIONAL DE FATORES CRÍTICOS DE SUCESSO EM GOVERNANÇA DE TI NA ADMINISTRAÇÃO PÚBLICA FEDERAL À LUZ DOS DADOS DE CONTROLE EXTERNO}.
+Sousa Jr., \textit{Análise Multidimensional de Fatores Críticos de Sucesso em Governança de TI na Administração Pública Federal à Luz dos Dados de Controle Externo}.
 in Proc.
-12th International Conference on Management of Technology and Information Systems (CONTECSI), 2015, São Paulo, Brazil \cite{}.
+12th International Conference on Management of Technology and Information Systems (CONTECSI), 2015, São Paulo, Brazil \cite{ramos2015}.
 	\end{enumerate}

### 2018-05-15 faceRecognition/FDDL/demo.m

@@ -1,0 +1,69 @@
+close all;
+clear all;
+clc;
+
+addpath([cd '/utilies']);
+load(['AR_EigenFace']);
+workPath = '/home/thiago/dev/projects/discriminative-sensing/faceRecognition/FDDL/';
+logFileName = 'FDDL.log';
+
+% logging
+logFile = fopen(fullfile(workPath, logFileName), 'wt');
+if logFile == -1
+  error('Cannot open log file.');
+end
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+% FDDL parameter
+%%%%%%%%%%%%%%%%%%%%%%%%
+fprintf(logFile, '[demo] FDDL execution\n');
+opts.nClass = 100;
+opts.wayInit = 'PCA';
+opts.lambda1 = 0.005;
+opts.lambda2 = 0.05;
+opts.nIter = 15;
+opts.show = true;
+% Data = 300 features, 700 occurences, 100 classes, 7 occurences of each
+% class, dict_class_size = 300x7, dict_total_size = 300x700,
+% coef_class_size = 700x7, coef_total_size = 700x700
+[Dict, Drls, CoefM, CMlabel] = FDDL(tr_dat, trls, opts, logFile);
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+% Sparse Classification
+%%%%%%%%%%%%%%%%%%%%%%%%
+fprintf(logFile, '[demo] Sparse Classification\n');
+lambda = 0.005;
+nClass = opts.nClass;
+weight = 0.5;
+td1_ipts.D = Dict;
+td1_ipts.tau1 = lambda;
+if size(td1_ipts.D,1)>=size(td1_ipts.D,2)
+   td1_par.eigenv = eigs(td1_ipts.D'*td1_ipts.D,1);
+else
+   td1_par.eigenv = eigs(td1_ipts.D*td1_ipts.D',1);  
+end
+
+ID = [];
+for indTest = 1:size(tt_dat,2)
+    fprintf(logFile, '[demo] Classification: %i of %i\n', indTest, size(tt_dat,2));
+    td1_ipts.y = tt_dat(:,indTest);
+    [opts] = IPM_SC(td1_ipts,td1_par);
+    s = opts.x;
+    
+    for indClass = 1:nClass
+        temp_s = zeros(size(s));
+        temp_s(indClass==Drls) = s(indClass==Drls);
[…]

### 2018-05-15 faceRecognition/FDDL/utilies/FDDL.m

@@ -1,0 +1,184 @@
+function [Dict,Drls,CoefM,CMlabel] = FDDL(TrainData, TrainLabel, opts, logFile)
+% ========================================================================
+% Fisher Discriminative Dictionary Learning (FDDL), Version 1.0
+% Copyright(c) 2011  Meng YANG, Lei Zhang, Xiangchu Feng and David Zhang
+% All Rights Reserved.
+%
+% ----------------------------------------------------------------------
+% Permission to use, copy, or modify this software and its documentation
+% for educational and research purposes only and without fee is here
+% granted, provided that this copyright notice and the original authors'
+% names appear on all copies and supporting documentation. This program
+% shall not be used, rewritten, or adapted as the basis of a commercial
+% software or hardware product without first obtaining permission of the
+% authors. The authors make no representations about the suitability of
+% this software for any purpose. It is provided "as is" without express
+% or implied warranty.
+%----------------------------------------------------------------------
+%
+% This is an implementation of the algorithm for learning the
+% Fisher Discriminative Dictionary from a labeled training data
+%
+% Please refer to the following paper
+%
+% Meng Yang, Lei Zhang, Xiangchu Feng, and David Zhang,"Fisher Discrimination 
+% Dictionary Learning for Sparse Representation", In IEEE Int. Conf. on
+% Computer Vision, 2011.
+% 
+%----------------------------------------------------------------------
+%
+%Input : (1) TrainDat: the training data matrix. 
+%                      Each column is a training sample
+%        (2) TrainDabel: the training data labels
+%        (3) opts      : the struture of parameters
+%               .nClass   the number of classes
+%               .wayInit  the way to initialize the dictionary
[…]

### 2018-05-15 faceRecognition/FDDL/utilies/FDDL_INID.m

@@ -1,0 +1,29 @@
+function D = FDDL_INID(data, nCol, wayInit)
+% ========================================================================
+% Dictionary Initialization of FDDL, Version 1.0
+% Copyright(c) 2011  Meng YANG, Lei Zhang, Xiangchu Feng and David Zhang
+% All Rights Reserved.
+%
+% -----------------------------------------------------------------------    
+%   
+% Input :   (1) data :  the data matrix 
+%           (2) nCol :  the number of dictioanry's columns
+%           (3) wayInit:  the method to initialize dictionary
+% 
+% Outputs : (1) D  :    the initialized dictionary
+%
+%------------------------------------------------------------------------
+m = size(data,1);
+
+switch lower(wayInit)
+    case {'pca'}
+        [D, disc_value, Mean_Image] = Eigenface_f(data, nCol-1);
+        D = [D Mean_Image./norm(Mean_Image)];
+    case {'random'}
+        phi = randn(m, nCol);
+        phinorm = sqrt(sum(phi.*phi, 2));
+        D = phi ./ repmat(phinorm, 1, nCol);
+    otherwise 
+        error{'Unkonw method.'}
+end
+return;

### 2018-05-15 faceRecognition/FDDL/utilies/FDDL_SpaCoef.m

@@ -1,0 +1,238 @@
+function [opts] = FDDL_SpaCoef(ipts, par)
+% ========================================================================
+% Coefficient updating of FDDL, Version 1.0
+% Copyright(c) 2011  Meng YANG, Lei Zhang, Xiangchu Feng and David Zhang
+% All Rights Reserved.
+%
+% -----------------------------------------------------------------------
+% Permission to use, copy, or modify this software and its documentation
+% for educational and research purposes only and without fee is here
+% granted, provided that this copyright notice and the original authors'
+% names appear on all copies and supporting documentation. This program
+% shall not be used, rewritten, or adapted as the basis of a commercial
+% software or hardware product without first obtaining permission of the
+% authors. The authors make no representations about the suitability of
+% this software for any purpose. It is provided "as is" without express
+% or implied warranty.
+%----------------------------------------------------------------------
+%
+% This is an implementation of the algorithm for updating the
+% Coefficient matrix of FDDL (fix the dictionary)
+%
+% Please refer to the following paper
+%
+% Meng Yang, Lei Zhang, Xiangchu Feng, and David Zhang,"Fisher Discrimination
+% Dictionary Learning for Sparse Representation", In IEEE Int. Conf. on
+% Computer Vision, 2011.
+% L. Rosasco, A. Verri, M. Santoro, S. Mosci, and S. Villa. Iterative
+% Projection Methods for Structured Sparsity Regularization. MIT Technical
+% Reports, MIT-CSAIL-TR-2009-050,CBCL-282, 2009.
+% J. Bioucas-Dias, M. Figueiredo, ?A new TwIST: two-step iterative shrinkage
+% /thresholding  algorithms for image restoration?, IEEE Transactions on
+% Image Processing, December 2007.
+%----------------------------------------------------------------------
+%
+%  Inputs :   (1) ipts :    the structre of input data
+%                    .D     the dictionary
+%                    .X     the training data
[…]

### 2018-05-15 igovti_fcs/_/pca (copy).py

@@ -1,0 +1,84 @@
+# Create a signal with only 2 useful dimensions
+from sklearn import decomposition
+from sklearn import preprocessing
+from numpy import genfromtxt
+import matplotlib.pyplot as plt
+import numpy as np
+
+org = genfromtxt('org', delimiter='\t')
+fcs = genfromtxt('fcs', delimiter='\t')
+data = genfromtxt('data', delimiter='\t')
+convdata = genfromtxt('convdata', delimiter='\t')
+normalizer = preprocessing.Normalizer().fit(data)
+normdata = normalizer.transform(data)
+
+# raw 
+index = np.power(org[:,1] * 20, 2)
+pca = decomposition.PCA()
+pca.fit(data)
+np.savetxt('eigenvalues.txt', pca.explained_variance_)
+np.savetxt('eigenvalues_ratio.txt', pca.explained_variance_ratio_)
+pcs = pca.explained_variance_ratio_
+pca.n_components = 2
+reduced = pca.fit_transform(data)
+# PCs_FCS
+plt.scatter(pcs * 100, fcs[:,1], s=pcs * 10000, c=fcs[:,1], alpha=0.1)
+plt.title('Classification of PCs into FCS or not (Raw Data)')
+plt.xlabel('PCs (%)')
+plt.ylabel('FCS (1 = True, -1 = False)')
+plt.grid()
+plt.savefig('raw_pcs_fcs.png', format='png')
+plt.clf()
+# Best Institutions_PC1_PC2
+plt.scatter(reduced[:,0], reduced[:,1], s=index, c=index, alpha=0.5)
+plt.title('Two Principal Components (Raw Data) ordered by OrgClassification')
+plt.xlabel('PC1')
+plt.ylabel('PC2')
+plt.grid()
+plt.savefig('raw_pcs.png', format='png')
+plt.clf()
+
+
+# converted
+index = np.power(org[:,1] * 20, 2)
+pca = decomposition.PCA()
+pca.fit(convdata)
+pcs = pca.explained_variance_ratio_
+pca.n_components = 2
+reduced = pca.fit_transform(convdata)
+plt.scatter(reduced[:,0], reduced[:,1], s=index, c=index, alpha=0.5)
+plt.title('Two Principal Components (Converted Data)')
+plt.xlabel('PC1')
+plt.ylabel('PC2')
+plt.grid()
[…]

### 2018-05-15 igovti_fcs/_/un.py

@@ -1,0 +1,12 @@
+from sklearn import cluster, datasets
+
+iris = datasets.load_iris()
+x_iris = iris.data
+y_iris = iris.target
+
+k_means = cluster.KMeans(n_clusters=3)
+k_means.fit(x_iris) 
+
+print(iris)
+print(k_means.labels_[::10])
+print(y_iris[::10])

### 2018-05-15 igovti_fcs/igovti_fcs_rfe.py

@@ -1,0 +1,45 @@
+# coding:utf-8
+## Recursive Feature Elimination to identify the f features more importants for iGovTI classification
+from sklearn import preprocessing
+from sklearn import datasets
+from sklearn.feature_selection import RFE
+from sklearn.svm import SVC
+from numpy import genfromtxt
+
+## settings
+csf = 54
+success = 0.0;
+fcs_success = 0.0;
+
+## loading data
+print('\n# loading data')
+data = genfromtxt('data/class_data_raw', delimiter='\t')
+target = genfromtxt('data/class_target', delimiter='\t')
+qualitative_fcs = genfromtxt('data/class_fcs', delimiter='\t', dtype=None)
+
+## create a base classifier
+svm = SVC(kernel='linear')
+
+## create the RFE model and select csf features, which are the critical success factors
+rfe = RFE(svm, csf)
+rfe = rfe.fit(data, target)
+
+## get results
+for i in range(len(rfe.support_)):
+	if rfe.support_[i] == qualitative_fcs[i]:
+		success = success + 1
+		if (rfe.support_[i] == True):
+			fcs_success = fcs_success + 1
+
+## success ratio
+print('\n# success ratio')
+print success/len(rfe.support_)
+print fcs_success/54
+
+## success ratio
+print('\n# support')
+print rfe.support_
+
+## success ratio
+print('\n# ranking')
+print rfe.ranking_

### 2018-05-15 network-attack-detection/adaptive-multivariate-gaussian/agg.py

@@ -1,0 +1,83 @@
+import binet_features as bf # our feature extractors for CTU-13
+import pandas as pd
+import os
+import sys # for argv
+import datetime as dt
+
+# The data structure to hold our feature extraction functions, which will get applied to each aggregation of the datasets.
+extractors = {
+    'Label'   : [bf.label_atk_v_norm,
+                 bf.n_attack,
+                 bf.background_flow_count,
+                 bf.normal_flow_count,
+                 bf.n_conn,
+                ],
+    'Dport'   : [bf.n_dports_gt1024,
+                 bf.n_dports_lt1024
+                ],
+    'Sport'   : [bf.n_sports_gt1024,
+                 bf.n_sports_lt1024,
+                ],
+    'Dur'     : [bf.avg_duration,
+                ],
+    'SrcAddr' : [bf.n_s_a_p_address,
+                 bf.n_s_b_p_address,
+                 bf.n_s_c_p_address,
+                 bf.n_s_na_p_address,
+                 bf.n_s_ipv6,
+                ],
+    'DstAddr' : [bf.n_d_a_p_address,
+                 bf.n_d_b_p_address,
+                 bf.n_d_c_p_address,
+                 bf.n_d_na_p_address,
+                 bf.n_d_ipv6,
+                ],
+    'Proto'   : [bf.n_tcp,
+                 bf.n_icmp,
+                 bf.n_udp,
+                ],
+}
+
+def get_files(path):
+    # here it's okay if the file is not found
+    # we would want the program to except on this
+    # case, as without the files there is nothing to do.
+    directory = os.fsencode(path)
+    return os.listdir(directory)
+
+def agg_time(interval='1S'):
+    # iterate the files in the dataset directory
+    pickle_dir = os.path.join('/media/thiago/ubuntu/datasets/network/','stratosphere-botnet-2011/CTU-13-Dataset/','ctu13pickles/')
+    path = os.path.join('/media/thiago/ubuntu/datasets/network/','stratosphere-botnet-2011/CTU-13-Dataset/')
+
+    for f in get_files(path):
+        strPath = os.path.join(pickle_dir, f.decode('utf-8') + '_pick')
+        if strPath.endswith('binetflow_pick'):
+            print('Working on ', strPath)
+            if not os.path.isfile(strPath):
+                # read input file
+                df = pd.read_csv(os.path.join(path, f.decode('utf-8')))
+
+                # chop off the ms from the start time and reindex the df to use start time as a datetime object index.
[…]

### 2018-05-15 network-attack-detection/adaptive-multivariate-gaussian/binet_features.py

@@ -1,0 +1,163 @@
+import ipaddress
+import numpy as np
+
+def classify_ip(ip):
+    '''
+    str ip - ip address string to attempt to classify.
+    treat ipv6 addresses as N/A
+    '''
+    try:
+        ip_addr = ipaddress.ip_address(ip)
+        if isinstance(ip_addr,ipaddress.IPv6Address):
+            return 'ipv6'
+        elif isinstance(ip_addr,ipaddress.IPv4Address):
+            # split on .
+            octs = ip_addr.exploded.split('.')
+            if 0 < int(octs[0]) < 127: return 'A'
+            elif 127 < int(octs[0]) < 192: return 'B'
+            elif 191 < int(octs[0]) < 224: return 'C'
+            else: return 'N/A'
+    except ValueError:
+        return 'N/A'
+
+def avg_duration(x):
+    return np.average(x)
+
+def n_dports_gt1024(x):
+    count = 0
+    for p in x:
+        if p>1024: count += 1
+    return count
+n_dports_gt1024.__name__ = 'n_dports>1024'
+
+def n_dports_lt1024(x):
+    count = 0
+    for p in x:
+        if p<1024: count += 1
+    return count
+n_dports_lt1024.__name__ = 'n_dports<1024'
+
+def n_sports_gt1024(x):
+    count = 0
+    for p in x:
+        if p>1024: count += 1
+    return count
+n_sports_gt1024.__name__ = 'n_sports>1024'
+
+def n_sports_lt1024(x):
+    count = 0
+    for p in x:
+        if p<1024: count += 1
+    return count
+n_sports_lt1024.__name__ = 'n_sports<1024'
+
+def label_atk_v_norm(x):
+    for l in x:
+        if 'Botnet' in l: return 'Attack'
+    return 'Normal'
+label_atk_v_norm.__name__ = 'label'
+
+def background_flow_count(x):
+    count = 0
+    for l in x:
+        if 'Background' in l: count += 1
+    return count
+
+def normal_flow_count(x):
+    if x.size == 0: return 0
+    count = 0
+    for l in x:
+        if 'Normal' in l: count += 1
+    return count
+
+def n_conn(x):
+    return x.size
+
+def n_tcp(x):
+    count = 0
+    for p in x:
+        if p == 'tcp': count += 1
+    return count
+
+def n_udp(x):
+    count = 0
+    for p in x:
+        if p == 'udp': count += 1
+    return count
+
+def n_icmp(x):
+    count = 0
+    for p in x:
[…]

### 2018-05-15 network-attack-detection/adaptive-multivariate-gaussian/plots.py

@@ -1,0 +1,102 @@
+import pandas as pd
+import numpy as np
+import os
+import sys
+import datetime as dt
+import matplotlib
+matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+import seaborn as sns
+import binet_features as bf
+
+from functools import reduce
+import ipaddress
+
+heatmaps     = (True if 'heatmaps' in sys.argv else False)
+scatter_mats = (True if 'scatter_mats' in sys.argv else False)
+
+dataset_path = os.path.join('/media/thiago/ubuntu/datasets/network/','stratosphere-botnet-2011/CTU-13-Dataset/')
+directory = os.fsencode(dataset_path)
+
+files = os.listdir(directory)
+
+# The datastructure to hold our feature extraction functions, which will get applied to each aggregation of the datasets.
+extractors = {
+    'Label'   : [bf.label_atk_v_norm,
+                 bf.n_attack,
+                 bf.background_flow_count,
+                 bf.normal_flow_count,
+                 bf.n_conn,
+                ],
+    'Dport'   : [bf.n_dports_gt1024,
+                 bf.n_dports_lt1024
+                ],
+    'Sport'   : [bf.n_sports_gt1024,
+                 bf.n_sports_lt1024,
+                ],
+    'Dur'     : [bf.avg_duration,
+                ],
+    'SrcAddr' : [bf.n_s_a_p_address,
+                 bf.n_s_b_p_address,
+                 bf.n_s_c_p_address,
+                 bf.n_s_na_p_address,
+                 bf.n_s_ipv6,
+                ],
+    'DstAddr' : [bf.n_d_a_p_address,
+                 bf.n_d_b_p_address,
+                 bf.n_d_c_p_address,
+                 bf.n_d_na_p_address,
+                 bf.n_d_ipv6,
+                ],
+    'Proto'   : [bf.n_tcp,
+                 bf.n_icmp,
+                 bf.n_udp,
+                ],
+}
+
+if heatmaps or scatter_mats:
+    print('Reading dataset...')
+    i = 0
+    for f in files:
+        df = pd.read_csv(os.path.join(directory,f).decode('utf8'), low_memory=False)
+
+        df['StartTime'] = df['StartTime'].apply(lambda x: x[:19])
+        df['StartTime'] = pd.to_datetime(df['StartTime'])
+        df = df.set_index('StartTime')
+
+        # replace NaN with a negative port number
+        df['Dport'] = df['Dport'].fillna('-1')
[…]

### 2018-05-15 network-attack-detection/adaptive-multivariate-gaussian/sample.py

@@ -1,0 +1,87 @@
+import os.path
+import sys
+import pandas as pd
+
+# create a sample of the data we will be working with, using pandas.
+basepath = os.path.dirname(__file__)
+dataset_path = os.path.abspath(os.path.join(basepath,"..","CTU-13-Dataset"))
+
+# generate a summary of the dataset
+# There are 13 files in the directory, for a total size of about 2.5GB
+# dictionary to hold information about the dataset.
+dataset_info = {}
+directory = os.fsencode(dataset_path)
+
+def binet_summary_to_md(summaries):
+    '''
+    Expects a dict of summaries of binet flow files.
+    it will spit out a markdown formatted table of the summaries.
+    '''
+    template_header = \
+    '|Scen.|Total Flows|Botnet Flows|Normal Flows|C&C Flows|Background Flows|\n' + \
+    '|---|---|---|---|---|---|\n'
+    
+    # This is hacky...
+    template_row = \
+    '|{}|{}|{}({:.2f}%)|{}({:.2f}%)|{}({:.2f}%)|{}({:.2f}%)|\n'
+
+    output = template_header
+    i = 1
+    for s in summaries:
+        output = output + template_row.format(\
+            i,\
+            str(s['total_flows']),\
+            str(s['botnet_flows']), (s['botnet_flows'] / s['total_flows'] * 100),\
+            str(s['normal_flows']), (s['normal_flows'] / s['total_flows'] * 100),\
+            str(s['cc_flows']), (s['cc_flows'] / s['total_flows'] * 100),\
+            str(s['bg_flows']), (s['bg_flows'] / s['total_flows'] * 100)\
+        )
+        i = i + 1
+    return output
+
+def summarize_binet_file(f):
+    '''
+    @arg f - pandas dataframe representing a parsed CTU-13 binetflow file.
+
+    Each file has the following information associated with it;
+    Total Flows, Botnet Flows, Normal Flows, C&C Flows, Background Flows
+    '''
+    file_summary = {
+        'total_flows': 0,
+        'botnet_flows': 0,
+        'normal_flows': 0,
+        'cc_flows': 0,
+        'bg_flows':0
+    }
+
+    file_summary['total_flows']  = f.size
+    file_summary['botnet_flows'] = f[f['Label'].str.contains('Botnet')].size
+    file_summary['normal_flows'] = f[f['Label'].str.contains('Normal')].size
+    file_summary['cc_flows']     = f[(f['Label'].str.contains('Botnet')) 
[…]

### 2018-05-15 tensor/rank3_CPD.r

@@ -1,0 +1,99 @@
+message("## load libraries")
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(rTensor)
+
+## declare function to convert case to data_frame
+case123_to_df <- function(case123, i) {
+  as_data_frame(case123) %>%
+  mutate(space_index = space_index) %>%
+  gather(time_index, Value, -space_index) %>%
+  mutate(time_index = as.numeric(gsub("V", "", time_index))) %>%
+  mutate(case = i)
+}
+
+## generate data
+message("## generate data")
+# bell-shaped (normal) spacial component with different means
+space_index <- seq(-1, 1, l = 100)
+case1 <- matrix(rep(dnorm(space_index, mean = 0,    sd = 0.3), 10), 100, 100)
+case2 <- matrix(rep(dnorm(space_index, mean = 0.5,  sd = 0.3), 10), 100, 100)
+case3 <- matrix(rep(dnorm(space_index, mean = -0.5, sd = 0.3), 10), 100, 100)
+## visualize noisy case 1, case 2, and case 3 means
+setEPS()
+postscript("normal_and_mean.eps")
+bind_rows(case123_to_df(case1, "case 1"), case123_to_df(case2, "case 2"), case123_to_df(case3, "case 3")) %>%
+  ggplot(aes(y = space_index, x = time_index, fill = Value)) +
+    geom_tile() +
+    facet_wrap(~case, nrow = 1) +
+    xlab("Time") + 
+    ylab("Space") +
+    theme(legend.position = "bottom")
+# sine-shaped temporal component
+sine_wave <- sin(seq(-4*pi, 4*pi, l = 100))
+sine_mat  <- matrix(rep(sine_wave, each = 100), 100, 100)
+# add sine form
+case1 <- case1 + 0.3 * sine_mat
+case2 <- case2 + 0.6 * sine_mat
+case3 <- case3 + 0.9 * sine_mat
+setEPS()
+postscript("normal_mean_sine.eps")
+bind_rows(case123_to_df(case1, "case 1"), case123_to_df(case2, "case 2"), case123_to_df(case3, "case 3")) %>%
+  ggplot(aes(y = space_index, x = time_index, fill = Value)) +
+    geom_tile() +
+    facet_wrap(~case, nrow = 1) +
+    xlab("Time") + 
+    ylab("Space") +
+    theme(legend.position = "bottom")
+# shifts +0,1 and -0,1 in the temporal component
[…]

### 2018-05-15 tensor/tensor_decomposition_cp.r

@@ -1,0 +1,29 @@
+library(rTensor)
+
+space_index <- seq(-1, 1, l = 100)
+bell_curve  <- dnorm(space_index, mean = 0, sd = 0.5)
+
+case1 <- matrix(rep(bell_curve, 10), 100, 100)
+case2 <- matrix(rep(bell_curve, 10), 100, 100)
+case3 <- matrix(rep(bell_curve, 10), 100, 100)
+case2[ , 51:100] <- case2[ , 51:100] + 0.1
+case3[ , 51:100] <- case3[ , 51:100] - 0.1
+
+X <- array(NA, dim = c(90, 100, 100))
+
+for(i in 1:30) {
+	X[i, , ]    <- case1 + matrix(rnorm(10000, sd = 0.1), 100, 100)
+	X[i+30, , ] <- case2 + matrix(rnorm(10000, sd = 0.1), 100, 100)
+	X[i+60, , ] <- case3 + matrix(rnorm(10000, sd = 0.1), 100, 100)
+}
+
+dim(X)
+
+cp_decomp <- cp(as.tensor(X), num_components = 1)
+
+str(cp_decomp)
+# List of 3
+#  $ : num [1:90, 1] 0.0111 0.0111 0.0111 0.0111 0.0112 ...
+#  $ : num [1:100, 1] -0.00233 -0.00251 -0.00271 -0.00292 -0.00314 ...
+#  $ : num [1:100, 1] -0.00996 -0.00994 -0.00996 -0.00993 -0.00997 ...
+# NULL

### 2019-02-14 docs/papers/2018ICASSP/2018icassp.tex
updating after backup recovery

@@
 % ------
-\title{TENDOR-BASED DICTIONARY LEARNING FOR FRAUD DETECTION FROM UNBALANCED DATA}
+{Tensor-based Discriminative Sensing for Fraud Detection}
 %
@@
 \label{sec:introduction}
+
+Compressive sensing is emergent for signal processing and hás been applied to Fields such as signal restoration, image processing and classification (sparce representation classification src )
+
+Iterative dictionary learning hás been used as a blind technique for feature extraction, improving classification without require feature selection or principal conponent analysis
+
+Big data analytics require techniques to Deal with multidimensional data, making Sense of strucutre and relationship of many dimensios.
+Pca is commonly adopted for dimensionality reduction and feature extraction.
+Tensor hás been adopted for multidimensional data analysis and can be explored for dictionary learning
+
+Adaptive techniques for high dimensional and high thoughput
+
+we propose to apply the tensor deconposition for dictionary learning in order to evidentiate the discriminative sensing of a fraud detection dataset
+
 Dictionary Learning is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
@@
 The least square error is used by the method of the optimal directions (MOD) [27] to update the dictionary iteratively:
+% Dictionaries are either available analytically, or can be learned from a suitable training set.
+While analytic dictionaries permit to capture the global structure of a signal and allow a fast implementation, learned dictionaries often perform better in applications as they are more adapted to the considered class of signals.
 
@@
 Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}.
+
+% In imagery, the numerical burden for (i) learning a dictionary and for (ii) employing the dictionary for reconstruction tasks only allows to deal with relatively small image patches that only capture local image information.
+Separable dictionaries aims at overcoming these drawbacks by allowing a separable structure on the dictionary throughout the learning process.
+On the one hand, this permits larger patch-sizes for the learning phase, on the other hand, the dictionary is applied efficiently in reconstruction tasks.
+The crucial idea is to allow the dictionary to have a separable structure, where separable means that the dictionary D is given by the Kronecker product of two smaller dictionaries A ∈ R h×a and B ∈ R w×b
+% Without loss of generality we choose square image patches with w = h = 8, which is in accordance to the patch-sizes mostly used in the literature
 
@@
 
+\section{Related Works}
+
+Face recognition...
+
+Fisher...
+
+Florian...
+
+Tem que ver outro...
+
 \section{Data Model}
@@
 If we choose a separable 2-D sampling grid, we can write the 2D sparse recovery problem as
+
+Imbalance data where zero is predominante
+
+Data divided into learning and test, with isfraud labels with 1 or 0
 
@@
 
-...
+roc_auc is usually used for classification evaluation, but It os not the best option for imbalance data, and Also because It does not consider the unclassified cases, such as false negatives
+
+Due to the inherent complex characteristics of imbalanced data sets, learning from such data requires new understandings, principles, algorithms, and tools to transform vast amounts of raw data efficiently into information and knowledge representation.
+
+We adopt precision recall curve and compare with logistic regression based on original and pca data
+
+Classification based on reconstruction or other value, such as similarity analysis
+
+Evaluate what metric is more discriminative, mse, tsnr or other
 
@@
 The conclusion goes here.
+
+The importance of the right metric for classification of imbalanced data
+
+Tensor based analysis explores the extructure and relationship of multidimensional data and can improve the discriminative power of imbalanced data for classification
+
+Can be used for feature extraction and used for well known algorithm,  and can be used for discriminative classification based on heuristics
+
+Our proposal is better for imbalanced data than others techniques
 

### 2019-04-20 docs/papers/2019JNCA_m-mcd/2019JNCA_m-mcd.tex
adding initial paper template

--- 2017-01-03 docs/papers/2016JNCA/2016JNCA.tex
@@
 \usepackage{amsthm,amsmath}
+
+% declarations
 \DeclarePairedDelimiter\abs{\lvert}{\rvert}%
@@
 
-\journal{ Journal of Network and Computer Applications}
+\journal{Journal of Network and Computer Applications}
 
@@
 
-\title{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
+\title{Cumulant-based Robust Estimates for Network Attack Detection}
 
@@
 \author[unbaddress]{Thiago P. B. Vieira}
-\author[unbaddress]{Danilo F. Tenório}
 \author[unbaddress,Ilmenauaddress,Fraunhoferaddress]{João Paulo C. L. da Costa}
-\author[ufrgsaddress]{\\Edison P. de Freitas}
 \author[Ilmenauaddress,Fraunhoferaddress]{Giovanni Del Galdo}
-\author[unbaddress]{Rafael T. de Sousa Júnior}
 
@@
 \address[Fraunhoferaddress]{Fraunhofer Institute for Integrated Circuits IIS, Erlangen, Germany}
-\address[ufrgsaddress]{Graduate Program in Electrical Engineering, Federal University of Rio Grande do Sul (UFRGS), 90035-190, Porto Alegre, Brazil}
 
 \begin{abstract}
-Novel schemes for attack detection are crucial to identify adaptive malicious traffic coming from sources that are quickly mobilized by attackers in high throughput communication networks.
+Botnets...
 In this context, signal processing techniques have been applied to attack detection due to their capability to detect anomalies that are previously unknown, i.e.
-blind detection.
-This paper proposes a signal processing framework for the detection and identification of network attacks using concepts of model order selection (MOS), eigenvalues and similarity analysis.
-In order to validate the proposed framework, we consider network traffic datasets that contain malicious activity such as flood and port probing attacks.
-We propose to model the network traffic as a superposition of components, namely, user's operations (legitimate traffic), network service operation not related to the user (noise) and the malicious activity.
-The experiments performed in a real network and also using the DARPA 1998 public dataset show that the proposed blind detection approach achieves satisfactory levels of accuracy in terms of timely detection and identification of TCP/UDP ports under attack.
+blind detection, in contaminated or imbalanced data....
+This paper proposes a cumulant-based robust estimates for network attack detection, by combining kurtosis and skewess with Minimum Covariance Determinant.
+In order to validate the proposed framework, we consider CTU-13 dataset.
+The experiments show that the proposed method achieves satisfactory levels of precisio, recall and F1 score for network attack detection.
 \end{abstract}
@@
 \begin{keyword}
-Network Attack Detection \sep Model Order Selection \sep Eigen Analysis \sep Similarity Analysis
+Network Attack Detection \sep Cumulant \sep Minimum Covariance Determinant \sep Mahalanobis Distance
 \end{keyword}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
 
-In the context of anomaly-based schemes, this work proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
-Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
+Botnets...
+
+In the context of anomaly-based schemes, this work proposes a statistical approach based on cumulants and Mimimum Covariance Determinant (MCD) for network attack detection.
 To the best of our knowledge there is no similar model in the literature.
-The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
-In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack.
-In addition, we also evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset.
-Furthermore, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.
-
-The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
+We evaluate the accuracy of the proposed framework applied to CTU-13 dataset ??, which is a well known network traffic dataset.
+
+The main contributions of the proposed framework are a semi-supervised method for network attack detection, wich does not depend on knowledge of previous attacks and consider only normal network traffic for robust estimates normal traffic...
 
@@
 Section \ref{sec:datamodel} presents the data model and the evaluated datasets.
-Section \ref{sec:prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks.
-Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:Complexity} discusses the computational complexity of the proposed framework and evaluates the required processing time for tested scenarios.
-Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
-The \ref{sec:mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.
-
+Section \ref{sec:c_mcd} describes the proposed framework for blind and automatic detection of flood and probe attacks.
+Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Related Works}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Data Model}
@@
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
-We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz, and the operator $\sim$ that sorts the elements of a vector in ascending order.
-
-This section presents details of the experimental scenario and the selected cases of the DARPA dataset, along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic.
-Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
-Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition.
-Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
-
-\subsection{Analyzed Scenario and Data Collection}
-\label{sec:DataCollection}
-
-The environment of the analyzed scenario is composed of two computers and one router with access to the Internet and to an internal network, where the simulation of legitimate traffic, noise, flood and port scan attacks are performed.
-During the traffic generation, one computer assumes the role of the attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:fig1}.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=6cm, width=9cm]{results/figures/fig09.png}
-     \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
-     \label{fig:fig1}
-\end{figure}
-
-Note that the set of network traffic is modeled as legitimate, noise and malicious traffic, where the victim performs legitimate activities, that can be characterized by web access.
-In many organizations this type of traffic is predominant, since most of corporate services are web-based, such as: web pages, customized web-based systems and cloud services.
-It is possible to characterize the traffic of a DHCP service as an example of noise associated with the transport layer.
-For malicious traffic, three types of networks attacks are evaluated: synflood, fraggle and port scan.
-These attacks are reproduced using well-known security tools, such as Nmap\footnote{http://nmap.org} to port scan, Metasploit\footnote{http://www.metasploit.com} to synflood and Hping\footnote{http://hping.org} to lead the fraggle attack.
-
-A network traffic log is commonly formed by timestamp, protocol, source IP address, source port, destination IP address, destination port and additional information, according to the type of the used transport protocol.
-The following TCP traffic log is presented in order to exemplify the collected data:
-\newline
-\newline
-\texttt{\justify21:00:34.099289 IP 192.168.1.102.34712 > 200.221.2.45.80: Flags [S], seq 2424058224, win 14600, options [mss 1460, sackOK,TS val 244136 ecr 0,nop,wscale 7], length 0}
-\newline
-
-and the following to exemplify UDP traffic log: 
-\newline
-\newline
-\texttt{\justify21:24:42.484858 IP 192.168.1.102.68 > 192.168.1.1.67: BOOTP/DHCP, Request from 00:26:9e:b7:82:be, length 300}
-\newline
-
-In the proposed framework, the goal is to detect the anomalies only taking into account the traffic profile, i.e., specific information such as origin IP or day and time of the attack are not considered.
-Therefore, from the entire log information, we just consider the timestamp (for sequencing), port type and port number.
-
-\subsection{Modeling Data}
-\label{sec:ModelingData}
-
-By modeling the dataset as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{U}), noise (\textbf{N}) and malicious traffic (\textbf{A}), according to the following expression:
-
-\begin{equation}\label{eq:eq01}
-\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{N}^{(q)} + \boldsymbol{A}^{(q)},
-\end{equation}
-where $q$ represents the $q$-th time frame, which is a time grouping of network traffic.
-The matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns.
-Each row represents a communication port (TCP port or UDP port), and each column represents time bins having a approtiate size, such as one minute.
-Each element $x_{m,n}^{(q)}$ stands for the number of times that the port $m$ appears at the $n$-th minute, at the $q$-th time frame.
-
-The legitimate traffic $\boldsymbol{U}^{(q)}$ is characterized by the traffic from user's operations.
-When a user accesses a web page, for example, there is the corresponding TCP/IP traffic to request the page, as well as there is the traffic required to domain name resolution.
-Figure \ref{fig:fig3} depicts an example of the legitimate traffic obtained during experiments.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig03.png}
-     \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
-     \label{fig:fig3}
-\end{figure}
-
-The traffic that is not associated with user's operations and with malicious traffic is modeled as noise $\boldsymbol{N}^{(q)}$.
-The automatic acquisition service of logical IP network address (DHCP) is an example of noise.
-Independently of any user operation, the machine receives an IP address, since it is configured to perform a DHCP address request.
-Figure \ref{fig:fig4} depicts an example of noise in a network traffic, represented by traffic to ports 67 and 68.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig04.png}
-     \caption{Network traffic of user independent operations for network management.}
-     \label{fig:fig4}
-\end{figure}
-
-The traffic coming from a malicious activity, such as a synflood or fraggle attack, is represented by the matrix $\boldsymbol{A}^{(q)}$.
-For this work we only consider the traffic from port scanning and flood attacks.
-
-We define that if the obtained $\#\boldsymbol{A}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\boldsymbol{A}^{(q)} = 0$, then there is no malicious traffic.
-This paper shows how to detect the $\#\boldsymbol{A}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious network traffic.
-
-\subsection{Synflood, Fraggle and Port scan}
-\label{sec:SynfloodFraggleandPortscan}
-
-The network attacks evaluated by this work are: synflood, fraggle and port scan.
-The first two attacks can be qualified as flood or denial of service (DoS) attacks, while the last one can be qualified as probe or port scanning attack.
-DoS attempts to block access to system or network resources and is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks.
-Probe attacks scan the networks to collect information about host, such as IP addresses, ports and services.
-
-With respect to the synflood attacks, the attacker sends a large quantity and concurrent successive SYN requests to a target, in order to consume resources and cause a DoS.
-Figure \ref{fig:fig5} depicts an example of a synflood attack carried out in a real computer network.
-In an interval of ten minutes, more than 210,000 packets are sent as a synflood attack.
-This network traffic behavior can be considered an abnormal behavior of network traffic, especially since it is concentrated in a short period of time and presents similar outstanding traffic during the time under attack.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig05.png}
-     \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
-     \label{fig:fig5}
-\end{figure}
-
-With respect to the fraggle attack, large packets with UDP echo segments are sent to the broadcast address of a network.
-Every packet is modified to have the source address of the victim, in order to implement the source address spoofing technique.
-Therefore, each host receives a huge amount of requests UDP echo and all of them replies to the IP address of the victim, causing a packet flooding aiming a DoS.
-This attack can affect the entire network, since all hosts receive several requests UDP echo and respond with the ICMP protocol, therefore each host acts as an amplifier of the attack.
-This last part of the fraggle attack is not taken into account in this work, because the victim receives ICMP (network layer) packets originated from the hosts that are attacked with flooding packet UDP echo.
-Figure \ref{fig:fig6} depicts an example of the fraggle attack in a real computer network.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig06.png}
-     \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
-     \label{fig:fig6}
-\end{figure}
-
-More than 6,000,000 malicious packets can be counted in an interval of ten minutes, which can be considered an abnormal network traffic, especially due to the concentrated traffic in a short period of time and due to the similarity of the outstanding traffic.
-
-Port scan is the attempt to establish a connection to TCP and UDP ports to identify what services are running or are in the listening state.
-There are several available port scanning techniques, including: TCP SYN scan, TCP ACK scan and UDP scan.
-This work evaluates the use of TCP SYN scan and UDP scan.
-
-In TCP SYN scan, a SYN packet is sent to the destination and two types of responses may occur: SYN/ACK or RST/ACK.
-In the first case, the destination port is in the listening state, in the second case, the destination port is not listening.
-At the end of each port scanning, a RST/ACK packet is sent by the system that is performing the port scan.
-Therefore, a full connection or a complete three-way handshake is never established, which makes the detection of the attack sender more difficult, and requires approaches able to identify probe attacks without connection stablishment.
-The UDP scan technique sends UDP packets to the destination port, and if it responds with a \emph{ICMP port unreachable} message, then it indicates that the scanned port is closed.
-On the other hand, if a message is not received, then the port is considered as open.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=5cm, width=9cm]{results/figures/fig07.png}
-     \caption{Connection attempts in order to identify active ports.}
-     \label{fig:fig7}
-\end{figure}
-
-Figure \ref{fig:fig7} depicts an example of the port scan attack in a real computer network.
-Note that the traffic is composed of two packets for each TCP port and one UDP packet to each port.
-The incoming and outgoing packets analysis, for each port, shows the high correlation and similarity of TCP and UDP traffic during the simulated port scan attack.
-
-\subsection{The DARPA Dataset}
-\label{sec:DarpaDataset}
-
-The DARPA 1998 dataset\footnote{https://www.ll.mit.edu/ideval/data/} includes 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks.
-The attacks in this dataset
-can be grouped into: denial-of-service (DoS); remote to local (R2L), which is characterized by unauthorized access from a
-remote machine; user to root (U2R), which is characterized by unauthorized access to local
-super-user privileges; and probe attack.
-Since the proposed approach focus on flood and probe attack, the analysis concentrates on the attacks that present behaviors similar to flood or probe attack.
-We observe that the most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack.
-One example is the occurrence of a neptune attack which sends 20 only SYN packets, what is a behavior that differs of the expected flood attack behavior.
-Therefore, there were selected the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey,osanaiye2016distributed}, and the cases that scan ports sending just a few packets.
-From the simulated probe attacks, we select the cases that rely on TCP or UDP connections.
-
-The data modeling follows the method described by the Subsection \ref{sec:ModelingData}, with time frames of 20 minutes, packet counting aggregation by minute and considering the traffic to the following ports: 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389, 443.
-
-\section{Proposed Framework for Detection and Identification of Network Attacks}
-\label{sec:prop_getv}
+We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matrix, and the operator $\sim$ that sorts the elements of a vector in ascending order.
+
+This section also presents a description of the data model and data aggregation of the CTU-13 dataset, in Subsection \ref{sec:CTU-13} discusses the use of the CTU-13 dataset for evaluation of the proposed approach.
+
+\subsection{The CTU-13 Dataset}
+\label{sec:CTU-13}
+
+...
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Cumulant-based Minimum Covariance Determinant}
+\label{sec:c_mcd}
 
@@
 	\centering
-     \includegraphics[height=11cm, width=9cm]{results/figures/mos_eigen_similarity.eps}
+     % \includegraphics[height=11cm, width=9cm]{results/figures/mos_eigen_similarity.eps}
      \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
@@
 \begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental.eps}
+     % \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental.eps}
      \caption{Traffic selection for incremental approach.}
@@
 \begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/individualized.eps}
+     % \includegraphics[height=4cm, width=12.4cm]{results/figures/individualized.eps}
      \caption{Traffic selection for individual approach.}
@@
 \begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental_individualized.eps}
+     % \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental_individualized.eps}
      \caption{Traffic selection for incremental individualized approach.}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Experiments and Results}
@@
 	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_synflood.eps} 
+     % \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_synflood.eps} 
      \caption{Eigenvalues of the sample covariance matrix (synflood).}
@@
 	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_fraggle.eps}
+     % \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_fraggle.eps}
      \caption{Eigenvalues of the sample covariance matrix (fraggle).}
@@
 	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_portscan.eps}
+     % \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_portscan.eps}
      \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}
@@
 
-\section{Performance Evaluation}
-\label{sec:Complexity}
-
-This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
-
-\subsection{Complexity Analysis}
-\label{sec:ComplexityAnalysis}
-The EVD, calculated according to (\ref{eq:eq06}), requires the previous calculation of covariance matrix, according to Equations (\ref{eq:eq02}), (\ref{eq:eq03}), (\ref{eq:eq04}) and (\ref{eq:eq05}).
-The covariance matrix calculation is $O(M^2N)$ and the EVD is $O(N^3)$, where $M$ denotes the number of network ports and $N$ denotes the period time.
-Therefore, the computational complexity for all steps for EVD can be represented as $O(M^2N + N^3)$ and yields an $O(N^3)$ upper bound on the worst-case running time for EVD.
-
-EDC and EFT are the MOS schemes that presented accuracy on the evaluation for the network attack detection.
-The computational complexity evaluation for MOS focuses on EDC scheme, since EDC requires less processing time than EFT but presents the same accuracy for the evaluated scenario.
-EDC scheme is $O(Q \log Q + Q + Q \log Q)$ and its worst-case running time can be represented as $O(Q \log Q)$, where $Q$ denotes the number of time frames.
-
-The largest eigenvalue analysis is $O(\hat{d}Q)$, where $\hat{d}$ denotes the number of time frame under attack, according to Algorithm \ref{alg:alg01}.
-Subsequently, the eigen similarity analysis relies on EVD and cosine similarity analysis, which is $O(N^2)$, for $\hat{d}$ time frames, therefore the eigen similarity analysis is which is $O(\hat{d}(M^2N + N^3 + N^2))$ and yields an $O(N^3)$ upper bound on the worst-case running time for eigen similarity analysis.
-
-Therefore, the proposed framework is $O(N^3 + Q \log Q + \hat{d}Q + N^3)$ and its worst-case running time is $O(N^3)$.
-The computational complexity of EVD is predominant in the framework, but the approach splits the data into time frames with period time $N$, which makes possible to limitate the growth of $N$ even for evaluations of cases with total time larger than $N$, reducing the impact caused by the computational complexity of EVD.
-
-\subsection{Processing Time Analysis}
-\label{sec:ProcessingAnalysis}
-For better understanding the scalability and impact of configurations of $N$, $M$ and $Q$, the processing time required for different scenarios of parameter configurations and dataset were evaluated, measuring the processing time of:
-
-\begin{enumerate}
-	\item Eigen analysis based on sample covariance of zero mean;
-	\item Eigen analysis based on sample covariance of zero mean and unitary standard deviation;
-	\item EDC MOS scheme;
-\end{enumerate}
-
-The performance evaluation focus on the main steps, which are discussed in subsection \ref{sec:ComplexityAnalysis} regarding the complexity analysis.
-The data modeling is also a time consuming step, however its processing can be optimized through distributed processing techniques, such as MapReduce, achieving high throughput for packet counting or even for deep packet inspection \cite{Vieira2013}.
-
-The experiments were performed in a desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM, considering: variations on the network traffic time; the frame size denoted as $N$; the number of network ports denoted as $M$; the mean processing time for eigen analysis based on sample covariance of zero mean, denoted as 1-time; the mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation, denoted as 2-time; and the mean processing time for EDC MOS scheme, denoted as 3-time.
-The mean time calculations was obtained from 200 measurement repetitions, in order to obtain reliable values.
-
-\begin{table*}[!t]
-	\caption{Processing time of the main steps for anomaly detection}
-	\label{tab:11}
-	\centering
-	\begin{tabular}{|r|r|r|r|r|r|r|}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Traffic Time\\(hour)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Frame Size\\(min)\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Num.
-Ports\end{tabular}	& \begin{tabular}[x]{@{}l@{}}1-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}2-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}3-time\\(ms)\end{tabular}	\\ \hline
-16	& 10	& 17	& 0.7900	& 0.8100	& 0.0650\\ \hline
-16	& 20	& 17	& 0.5250	& 0.5950	& 0.0100\\ \hline
-16	& 60	& 17	& 0.9700	& 1.1400	& 0.0250\\ \hline
-16	& 120	& 17	& 0.6050	& 0.6100	& 0.0050\\ \hline
-16	& 60	& 34	& 1.2750	& 1.2200	& 0.0050\\ \hline
-16	& 120	& 34	& 1.1200	& 1.1700	& 0.0050\\ \hline
-20	& 10	& 17	& 2.7950	& 2.8950	& 1.1000\\ \hline
-20	& 20	& 17	& 2.0700	& 2.0200	& 0.3500\\ \hline
-20	& 60	& 17	& 1.0250	& 1.0450	& 0.0650\\ \hline
-20	& 120	& 17	& 1.0000	& 1.0700	& 0.0350\\ \hline
-20	& 60	& 34	& 2.9650	& 3.2100	& 0.0400\\ \hline
-20	& 120	& 34	& 2.9950	& 3.1150	& 0.0200\\ \hline
-22	& 10	& 17	& 4.7250	& 4.0850	& 1.4600\\ \hline
-22	& 20	& 17	& 2.3200	& 2.6800	& 0.2450\\ \hline
-22	& 60	& 17	& 1.0700	& 1.1200	& 0.0300\\ \hline
-22	& 120	& 17	& 0.9900	& 1.0500	& 0.0250\\ \hline
-22	& 60	& 34	& 3.0850	& 3.1250	& 0.0650\\ \hline
-22	& 120	& 34	& 2.8100	& 2.9600	& 0.0250\\ \hline
-	\end{tabular}
-\end{table*}
-
-Table \ref{tab:11} presents the measured results.
-The experiment considered traffic time of 16, 20 and 22 hours, according to the selected traffic time per day available by the DARPA dataset.
-Note that the processing time increases according to the increment in traffic time, around 2 or 3 times for 1-time and 2-time, but the worst measured processing time is 4.7250 milliseconds.
-
-According to Table \ref{tab:11}, the processing time increases with the frame size $N$ decreasing, therefore it is possible to evaluate the frame size that produces better identification rates and acceptable processing time.
-The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations, since the significant increase of processing time observed between scenarios considering 17 or 34 ports, with growth between 7 \% and 199 \%.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion and Future Works}
@@
 
-\appendix
-
-\section{Model Order Selection (MOS)}
-\label{sec:mos}
-The model order selection is a key point in many digital signal processing applications, including radar, sonar, communications, channel modeling, medical imaging, among others.
-MOS allows analysis of reduced data set, through separating noise components of the main components, for example.
-Moreover, the model order is crucial for many parameter estimation techniques \cite{da2009comparison}, since the amount of parameters to be estimated depends on the model order.
-
-The model selection procedure chooses the ``best'' model of a finite set of models, according to some criterias \cite{rajan1997model}.
-Therefore, given some data set, it is chosen a model which was evaluated as the best model to describe the specified data set.
-
-The state of the art regarding estimation techniques of model order based on eigenvalues includes: Akaike's Information Theoretic Criterion - AIC \cite{akaike1974new,wax1985detection}; Minimum Description Length - MDL \cite{barron1998minimum,wax1985detection}; Efficient Detection Criterion - EDC \cite{zhao1986detection}; Stein's Unbiased Risk Estimator - SURE \cite{ulfarsson2008rank}; RADOI \cite{radoi2004new} and Exponential Fitting Test - EFT \cite{grouffaud1996some,quinlan2006model,david2011blind}.
-
-In AIC, MDL and EDC techniques, the information criterion is a function of the geometric mean \emph{g(k)} and the arithmetic mean \emph{a(k)} relating to smaller \emph{k} eigenvalues, where \emph{k} is a candidate value for the model order \emph{d} \cite{da2009comparison}.
-
-Basically, the difference between the AIC, MDL and EDC schemes is the penalty function \emph{p(k, N, α)}, so these techniques can be written in general as \cite{da2009comparison}:
-
-\begin{equation}\label{eq:eq1}
-\hat{d} = \argmin\limits_k \hspace{1 mm} J(k), 
-\end{equation}
-
-where
-
-\begin{equation}\label{eq:eq2}
-J(k) = -N(\alpha - k) \hspace{1 mm} {\rm log} \hspace{1 mm} (g(k)/a(k)) + p(k,N,\alpha),
-\end{equation}
-
-where $\hat{d}$ is an estimate \emph{d} of the model order, \emph{N} is the number of samples, \emph{α = M} and means the number of variables of the problem, and 0 ≤ \emph{k} ≤ min[\emph{M}, \emph{N}].
-Penalty functions for AIC, MDL and EDC are given by the Table \ref{tab:tab2}.
-%% ???
-
-\begin{table}[h!]
-  \centering
-  \caption{Penalty functions for the schemes AIC, MDL and EDC}
-  \label{tab:tab2}
-  \begin{tabular}{*2c}
-	\toprule
-	\textbf{Scheme} &  \textbf{Penalty function} \\
-	\textbf{} &  $p(k,N,\alpha)$ \\
-	\midrule
-    AIC	& $k(2\alpha - k)$ \\
-    MDL	& $0.5k(2\alpha - k) \hspace{1 mm} {\rm log} (N)$ \\
-    EDC	& $0.5k(2\alpha - k)\sqrt{N\hspace{1 mm}\rm ln(\rm ln N)}$ \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The Exponential Fitting Test (EFT) can effectively be used in cases where the number of samples \emph{N} is small.
-This technique is based on observations of data contaminated only with white noise, where the profile of eigenvalues can be approximated by an exponential decaying \cite{grouffaud1996some}.
-
-Given $\lambda_i$ be the i-th eigenvalue, the exponential model can be expressed by:
-\begin{equation}\label{eq:eq3}
-E\{\lambda_i\} = E\{\lambda_1\} \cdot q(\alpha,\beta)^{i-1},
-\end{equation}
-
-where $E\{\cdot\} $ is the expectation operator, and it is considered that the eigenvalues are ordered in the that $\lambda_1$ represents the largest eigenvalue. The term \emph{q(α, β)} is defined as:
-
-\begin{equation}\label{eq:eq4}
-q(\alpha,\beta) = \exp\left\{-\sqrt{\frac{30}{\alpha^2 + 2} - \sqrt{\frac{900}{(\alpha^2 + 2)^2} - \frac{720\alpha}{\beta(\alpha^4 + \alpha^2 - 2)}}} \right\},
-\end{equation}
-
-where $0 < q(\alpha,\beta) < 1$. According to \cite{quinlan2006model}, if $M \leq N$, then $\beta = N$.
-
 \section*{References}

### 2019-04-20 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
adding initial paper template

--- 2019-04-20 docs/papers/2019JNCA_m-mcd/2019JNCA_m-mcd.tex
@@
 
-\title{Cumulant-based Robust Estimates for Network Attack Detection}
+\title{Robust PCA and Higher Moments Distances for Botnet Detection}
 
@@
 \author[unbaddress]{Thiago P. B. Vieira}
+\author[unbaddress]{Eduardo Said Calil Vilaça}
 \author[unbaddress,Ilmenauaddress,Fraunhoferaddress]{João Paulo C. L. da Costa}
-\author[Ilmenauaddress,Fraunhoferaddress]{Giovanni Del Galdo}
+\author[unbaddress]{Rafael T. de Sousa Júnior}
 
@@
 \begin{keyword}
-Network Attack Detection \sep Cumulant \sep Minimum Covariance Determinant \sep Mahalanobis Distance
+Botnet Detection \sep Network Anomaly Detection \sep Skewness \sep Kurtosis \sep Robust Principal Component Analysis \sep Mahalanobis Distance
 \end{keyword}
@@
 
-Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}.
-In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time.
-Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
-Recently, signal processing schemes have being applied to the detection of malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest}, showing advances in network traffic analysis.
+Botnets...
 
-Information security may consist of both technical and procedural aspects.
-The former includes equipment and security systems, while the latter corresponds to security rules and recommendations.
-Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks.
-Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.
-Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
+Anomaly detection is the identification of rare and suspicions observations by differing the normal or majority of the data.
 
-Botnets...
+Anomalies can be related to network attack, fraud, defect or medical problems, for examples.
+
+Anomalies are also referred to as outliers, novelties, noise, deviations and exceptions.
+
+Anomalies can be hard to identify and separate from normal data.
+
+Considering that anomalies are rare in comparison to normal events, anomaly detection algorithms have to deal with problems imposed by imbalanced data.
 
@@
 
-Several methods have been proposed for the identification and characterization of malicious activity in computer networks.
-Classical methods typically employ data mining \cite{he2008applying,ghourabi2010data,osanaiye2016distributed} and regular file analysis \cite{raynal2004honeypot} to detect patterns that indicate the presence of specific attacks in network traffic.
-
-Data mining is often used to describe the process of extracting useful information from large databases.
-Multiple methods of data mining are used in \cite{he2008applying,osanaiye2016distributed} to analyze data flow in a network with the aim of identifying characteristics of malicious traffic in large scale environments.
-Researchers have applied data mining techniques in log analysis \cite{ghourabi2010data} to improve intrusion detection performance.
-However, data mining techniques used so far in network analysis require prior collection of large data sets, which is a limitation of several schemes for online analysis \cite{Huang2009}.
-
-Regular file analysis \cite{raynal2004honeypot} consists of traffic analysis for detecting known patterns that indicate the presence of attacks, applying statistical analysis to the study of collected traffic.
-An essential feature of this method is that it depends on prior knowledge of the details of the attacks to be identified, and also depends on previous log collection for traffic analysis and false positives reduction.
-
 Principal Component Analysis (PCA) is a statistical technique commonly used for dimensionality reduction.
@@
 Therefore, in order to automate the analysis of eigenvalues profile, model order selection (MOS) schemes should be incorporated.
-
-Signal processing techniques have been successfully applied to network anomaly detection \cite{Lu2009}.
-Lu and Ghorbani \cite{Lu2009} proposed a network anomaly detection model based on network flow, wavelet approximation, and system identification theory.
-However, their work requires a training method to produce a prediction model for normal daily traffic and presents limitations on identification of behaviors without significant outliers, such as port scan attacks.
-Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
-Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work is not applied to probe and flood attack detection, simultaneously.
-
-The data collected from honeypot systems, such as captured traffic and operating system logs, can be analyzed to obtain information about attack techniques, general trends of threats and exploits.
-Blind automatic detection of malicious traffic techniques have been developed for honeypots in \cite{david2011blind,da2012improved}.
-However, traffic on honeypot is simpler than real network traffic, because there are no running legitimate applications, due to the fact that honeypots emulate behavior of a host within a network to deceive and lure attackers \cite{zakaria2012review}.
-Since honeypots do not generate legitimate traffic, the amount of data captured in honeypots is significantly lower in comparison to a Network Intrusion Detection System (NIDS), which captures and analyzes the largest possible amount of network traffic \cite{david2011blind}.
-MOS for blind identification of malicious activities in honeypots was proposed by us in \cite{david2011blind}, which evaluated criteria for selecting the model order, through simulations and comparing the order of the resulting model with the true model order.
-
-The proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic.
+Vieira ....
+(The proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic.
 The proposed solution is automatic and blind for detection of time frames under probe and flood attacks through MOS and eigen analysis.
-Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.
+Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.)
 
@@
 Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations.
-Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD datasets adopt a traffic aggregation by connections, we consider the use of a experimental scenario on a real network and the DARPA 1998 dataset, which is the source for the creation of the KDD 99 and NSL-KDD datasets.
-Note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD datasets.
+We propose to adopt CTU-13...
+
+Our proposal is ...
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Cumulant-based Minimum Covariance Determinant}
-\label{sec:c_mcd}
+\section{Robust PCA and Higher Moments Distances for Botnet Detection}
+\label{sec:m_rpca}
 

### 2019-04-20 docs/papers/2019_m-mcd_slides/m-mcd.tex
slides for m-mcd

%	Name			:: 	sthlm Beamer Theme  HEAVILY based on the hsrmbeamer theme (Benjamin Weiss)
%	Author			:: 	Mark Hendry Olson (mark@hendryolson.com)
%	Created			::	2013-07-31
%	Updated	    	::	[[April]] 04, 2017 at 16:26:39
%	Version			:: 	2.0.2
%	Email			:: 	hendryolson@gmail.com
%	Website			:: 	http://markolson.se
%	Twitter			:: 	markolsonse
%	Instagram		:: 	markolsonse
%
%	License			:: 	This file may be distributed and/or modified under the
%					GNU Public License.
%
%	Description		::	This presentation is a demonstration of the sthlm beamer
%					theme, which is HEAVILY based on the HSRM beamer theme created by Benjamin Weiss
%					(benjamin.weiss@student.hs-rm.de), which can be found on GitHub
%					<https://github.com/hsrmbeamertheme/hsrmbeamertheme>.  It also borrows heavily
%					from the work of Matthias Vogelgesang, (https://bloerg.net) and his Metropolis Mtheme,
%					<https://github.com/matze/mtheme>.
%
%	Theme			::	newPxFont
%	Options			::	progressbar
%					::	sectionpages
%					::	numfooter
%					::	fullfooter
%					::	dovaligncolumns
%					::	protectframetitle
%					::	greybg
%					::	cblock
%					::	minimal
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%        LOADING DOCUMENT
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\documentclass[newPxFont, numfooter, sectionpages]{beamer}
\usepackage[utf8]{inputenc}
\usetheme{sthlm}
\usepackage{pgfplots}
\pgfplotsset{compat=1.9}
\usepackage{cancel}
\usepackage{amsthm,amsmath}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{supertabular}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{color, colortbl}
\usepackage[backend=bibtex, style=numeric, defernumbers=true]{biblatex}
\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
\addbibresource{references.bib}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%

\definecolor{lightgray}{gray}{0.8}
\definecolor{Gray}{gray}{0.9}

\title{Network Anomaly Detection Based on Robust Moments}
\subtitle{\small{Ph.D. Thesis Presentation}}
\author{\texttt{Thiago P. de B. Vieira \newline \textbf{Advisor:} João Paulo C. L. da Costa}}
\institute{
	\scriptsize{Universidade de Brasília}\\
	\scriptsize{Departamento de Engenharia Elétrica - ENE/FT}\\
	\scriptsize{Programa de Pós-Graduação em Engenharia Elétrica - PPGEE}
}

%document-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\begin{document}

%title-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\maketitle

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Theme Package Requirements
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begingroup
\setbeamercolor{normal text}{fg=\cnDarkGrey,bg=white}

%outline-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section*{Outline}

\begin{frame}{Outline}
	\scriptsize
	\tableofcontents[hideallsubsections]% For longer presentations use hideallsubsections option
\end{frame}

%section-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Motivation}

\begin{frame}[c]{Anomaly Detection}	
	\begin{itemize}
		\item Anomaly detection is the identification of rare and suspicions observations by differing the normal or majority of the data.
		\item Anomalies can be related to network attack, fraud, defect or medical problems, for examples.
		\item Anomalies are also referred to as outliers, novelties, noise, deviations and exceptions.
		\item Anomalies can be hard to identify and separate from normal data.
		\item Considering that anomalies are rare in comparison to normal events, anomaly detection algorithms have to deal with problems imposed by imbalanced data.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Anomaly Detection}
	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=8cm]{figures/raw_dur_dport_pairplot.png}
	     \caption{Nomalies from original data.}
	     \label{fig:fig01}
	\end{figure}
\end{frame}

\begin{frame}[c]{Anomaly Detection}	
	\begin{itemize}
		\item Some techniques can make the data discriminative for better anomaly detection:
		\begin{enumerate}
			\item Data engineering (i.e. sampling \cite{??}, aggregation \cite{acarali2016survey}).
			\item Feature extraction (i.e. decomposition \cite{??}, robust estimates \cite{??}).
		\end{enumerate}
	\end{itemize}
\end{frame}

\begin{frame}[c]{Anomaly Detection}
	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=8cm]{figures/sum_avg_duration_conn_pairplot.png}
	     \caption{Anomalies from summarized data.}
	     \label{fig:fig02}
	\end{figure}
\end{frame}

\begin{frame}[c]{Anomaly Detection based on Robust Estimates}	
	\begin{itemize}
		\item Robust statistics produce methods that are not unduly affected by outliers.
		\item Robust statistical methods have been developed for estimating location, scale, and regression parameters.
		\item Robust estimates from contaminated data can denoise, remove anomalies or even separate them.
		\item Anomaly detection based on robust estimates has been adopted for network anomaly detection, due to good results and its adaptability.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Distance Based Anomaly Detection}	
	\begin{itemize}
		\item The distance between robust estimates and new observations can reveal anomalies.
		\item Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples by, considering the correlations between the variables and their mean values.
		\item \textbf{MD has been used for distance based anomaly detection with robust estimates} in many areas, and is commonly combined to Fast Minimum Covariance Determinant (Fast-MCD) algorithm \cite{rousseeuw1999fast} for robust estimates.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Distance Based Anomaly Detection}
	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=7.6cm]{figures/17_1s_distances.png}
	     \caption{Anomaly detection from robust estimates.}
	     \label{fig:fig03}
	\end{figure}
\end{frame}

\begin{frame}{Network Anomaly Detection Based on Robust Moments}
	\begin{itemize}
	    \item Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority}.
	    \item Learning algorithms for imbalanced data has been a relevant research topic. 
	    \item The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
	    \item Skewness and the kurtosis tests are very sensitive outlier detectors. The properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2008outlier}.
    \end{itemize}
\end{frame}

\begin{frame}{Network Anomaly Detection Based on Robust Moments}
	\begin{itemize}
	    \item We propose a network anomaly detection approach based on robust moment estimates and distance analysis.
    	\begin{itemize}
    	    \item Considering that Fast-MCD is a widely adopted robust estimator of location and scatter for anomaly detection, we propose to extend Fast-MCD for computing robust skewness and kurtosis.
		    \item We also propose to use Mahalanobis Distance for detecting anomalies based on robust skewness and kurtosis, instead of the location-based usage of classical approaches.
		    \item Considering that it is hard to have labeled data in real problems \cite{osanaiye2016distributed}, such as for network attacks or frauds, also considering that known anomalies are constantly changing due to adversarial models or due to lack of regular patterns, we propose a Semi-Supervised approach based on robust moments for network anomaly detection.
	    \end{itemize} 
    \end{itemize}
\end{frame}

%section-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Literature Review}

\begin{frame}[c]{Literature Review}
	\cite{gu2007bothunter} (Bothunter: Detecting malware infection through ids-driven dialog correlation)
	\begin{enumerate}
		\item ...
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{gu2008botminer} (Botminer: Clustering analysis of network traffic for protocol-and structure-independent botnet detection)
	\begin{enumerate}
		\item ...
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{hubert2008outlier} (Outlier detection for skewed data)
	\begin{enumerate}
		\item ...
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{garcia2014empirical} (An empirical comparison of botnet detection methods)
	\begin{enumerate}
		\item ...
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{khattak2015botflex} (BotFlex: A community-driven tool for botnet detection)
	\begin{enumerate}
		\item ...
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{wang2017botnet} (Botnet detection based on anomaly and community detection)
	\begin{enumerate}
		\item ...
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{da2018online} (Online detection of Botnets on Network Flows using Stream Mining)
	\begin{enumerate}
		\item ...
	\end{enumerate}
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{moustafa2019holistic} (A holistic review of Network Anomaly Detection Systems: A comprehensive survey)
	\begin{enumerate}
		\item ...
		\item Network anomaly detection approaches are classified by \cite{moustafa2019holistic} in six categories: classification, clustering, deep learning, knowledge, combination and statistical-based.
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{Phua2004minority} (Minority Report in Fraud Detection: Classification of Skewed Data)
	\begin{enumerate}
		\item ...
		\item \cite{acarali2016survey} (Survey of approaches and features for the identification of HTTP-based botnet traffic)
		\item \cite{lakhina2005mining} (Mining anomalies using traffic feature distributions)
		\item \cite{goldberg2015importance} (The importance of features for statistical anomaly detection)
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{acarali2016survey} (Survey of approaches and features for the identification of HTTP-based botnet traffic)
	\begin{enumerate}
		\item ...
		\item \cite{lakhina2005mining} (Mining anomalies using traffic feature distributions)
		\item \cite{goldberg2015importance} (The importance of features for statistical anomaly detection)
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions)
	\begin{enumerate}
		\item ...
		\item \cite{goldberg2015importance} (The importance of features for statistical anomaly detection)
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Literature Review}
	\cite{goldberg2015importance} (The importance of features for statistical anomaly detection)
	\begin{enumerate}
		\item ...
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Clustering Method for anomaly detection}
	\begin{enumerate}
		\item Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{moustafa2019holistic}.
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Gaussian Mixture Model}
	\begin{enumerate}
		\item GMM is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Mahalanobis Distance}
	\begin{enumerate}
		\item The Mahalanobis Distance is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P. C. Mahalanobis in 1936 \cite{mahalanobis1936generalized}.
		\item It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\hat{S}}$ of $\boldsymbol{X}$
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Mahalanobis Distance}
	\begin{enumerate}
		\item Mahalanobis Distance is defined as		
		\begin{equation}\label{eq:eq01}
			d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}.
		\end{equation}
		\item $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector, also referred as location, of known observations and $\boldsymbol{\hat{S}}$ is the covariance matrix, also referred as scatter, of known observations.
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Mahalanobis Distance}
	\begin{enumerate}
		\item Classical estimates can be so strongly affected by contamination that diagnostic tools, such as the Mahalanobis Distances, become unable to detect the outliers.
		\item Therefore, we need reliable and robust estimators that can resist outliers contaminated data.
		\item Minimum Covariance Determinant (MCD) \cite{rousseeuw1984least} is a robust estimator commonly used for computing Mahalanobis Distance in order to detect anomalies.
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{Mahalanobis Distance}
	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=11cm]{figures/mahalanobis_robust.png}
	     \caption{Mahalanobis Distance for non-robust and robust estimates.}
	     \label{fig:fig04}
	\end{figure}
\end{frame}

\begin{frame}[c]{Mahalanobis Distance}
	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=8cm]{figures/mahalanobis_distances01.png}
	     \caption{Mahalanobis Distance for non-robust and robust estimates.}
	     \label{fig:fig05}
	\end{figure}
\end{frame}

\begin{frame}[c]{MCD - Minimum Covariance Determinant}
	\begin{enumerate}
		\item The MCD estimator is one of the first affine equivariant and highly robust estimators of multivariate location and scatter.
		\begin{enumerate}
			\item Affine equivariant means that when the data are translated or subjected to a linear transformation, the resulting location and scatter will transform accordingly \cite{rousseeuw1984least,rousseeuw1999fast}.
		\end{enumerate}
		\item Its objective is to find $h$ observations (out of $n$) whose covariance matrix has the lowest determinant.
		\begin{enumerate}
			\item The lowest determinant means the lowest distance generalized variance and high distance between the largest eigenvalue and the remaining [ref?].
		\end{enumerate}
		\item MCD depends on random sampling, therefore it is not deterministic.
	\end{enumerate}
\end{frame}

\begin{frame}[c]{MCD - Minimum Covariance Determinant}
	\begin{enumerate}
		\item MCD has been applied in numerous fields (medicine, finance, image analysis, and chemistry) and been used to develop many robust multivariate techniques (such as Robust Principal Component Analysis (RPCA), factor analysis, and multiple regression).
		\item The classical MCD has rarely been applied because it is hard to compute, since it requires the evaluation of all $\binom{n}{h}$ subsets of size $h$. % binomial of n over p
		\item Its main use has started since the construction of the computationally efficient Fast-MCD algorithm \cite{rousseeuw1999fast}.
	\end{enumerate}
\end{frame}

\begin{frame}[c]{Fast-MCD}
	\begin{enumerate}
		\item Fast-MCD is an iterative, approximate and resampling algorithm for the MCD.
		\item It starts by random initial subset of size $p+1$ and performs concentration steps (C-steps) yielding consecutive h-subsets with decreasing covariance matrix determinant.
		\item Only two C-steps are applied to each initial subset, and the 10 results with lowest determinant are kept.
		\item C-steps are carried out until convergence and the best solution is kept.
		\item Fast-MCD also is affine equivariant and not deterministic.
	\end{enumerate}
\end{frame}

\begin{frame}[c]{Fast-MCD}

	\textbf{Theorem 1: C-Step.}

	Consider a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ where $\boldsymbol{X} = \{\boldsymbol{x}_1,...,\boldsymbol{x}_n\}$ of $p$-variate observations. Let $\boldsymbol{h} \subset \{1,...,n\}$ with $|\boldsymbol{h}| = h$, where $\boldsymbol{h}$ is a subset of $h$ indexes from $n$ observations.

	Let the mean $\bar{\boldsymbol{x}} \in \mathbb{R}^{1 \times p}$ be

	\begin{equation}\label{eq:eq02}
		\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} \boldsymbol{x}_j, 
	\end{equation}

	the covariance matrix $\boldsymbol{\hat{S}} \in \mathbb{R}^{p \times p}$ be

	\begin{equation}\label{eq:eq03}
		\boldsymbol{\hat{S}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} (\boldsymbol{x}_j - \bar{\boldsymbol{x}})(\boldsymbol{x}_j - \bar{\boldsymbol{x}})^\prime,
	\end{equation}
\end{frame}

\begin{frame}[c]{Fast-MCD}
	and the relative distance be

	\begin{equation}\label{eq:eq04}
		\boldsymbol{d}(i) = \sqrt{(\boldsymbol{x}_i - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime}, 
	\end{equation}

	for $i = 1,...,n$.

	Let $\boldsymbol{h}_1, \bar{\boldsymbol{x}}_1$ and $\boldsymbol{\hat{S}}_1$ and be the selected observations, the location and scatter of the initial subset, and $\boldsymbol{d}_1(i)$ be the initial distances for $i = 1,...,n$.

	Now take $\boldsymbol{h}_2$ containing the indexes of lowest distances, such that $\{\boldsymbol{d}_1(i); i \in \boldsymbol{h}_2\} = \{(\boldsymbol{d}_1)_{1:n},...,(\boldsymbol{d}_1)_{h:n}\}$, where $(\boldsymbol{d}_1)_{1:n} \leq (\boldsymbol{d}_1)_{2:n} \leq ... \leq (\boldsymbol{d}_1)_{n:n}$ are the ordered distances, and compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$.

\end{frame}

\begin{frame}[c]{Fast-MCD}

	Then, $det(\boldsymbol{\hat{S}}_2) \leq det(\boldsymbol{\hat{S}}_1)$ with equality if and only if $\boldsymbol{\bar{x}}_2 = \boldsymbol{\bar{x}}_1$ and $\boldsymbol{\bar{S}}_2 = \boldsymbol{\hat{S}}_1$

	\begin{enumerate}
		\item The proof is given by Rousseeuw and Driessen \cite{rousseeuw1999fast}.
		\item \label{item:cstep}It is referred to the construction in Theorem 1 as a C-step, where C can be taken to stand for "covariance" or for "concentration".
	\end{enumerate}

\end{frame}

\begin{frame}[c]{Fast-MCD}
	\textbf{The C-step can be algorithmically described as follows:}
	\begin{algorithm}[H]\label{alg:alg01}
		\scriptsize
		\SetAlgoLined
		\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
		Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
		\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
			\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
				Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
			}
			Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
			Sort $\boldsymbol{d}_1( i )$, which yields a index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
			Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
			Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
			Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
		}
		\caption{C-Step}
	\end{algorithm}
\end{frame}

\begin{frame}[c]{Fast-MCD}
	\textbf{Theorem 1} thus provides a partial idea for an algorithm: 
		\begin{itemize}
			\item \textit{Take many initial choices of $\boldsymbol{h}_1$ and apply C-steps to each until convergence (don't have determinant decreasing), and keep the result with lowest determinant}.
		\end{itemize}
\end{frame}

\begin{frame}[c]{Fast-MCD}
	\textbf{Initial subset can be algorithmically described as:}
	\begin{algorithm}[H]\label{alg:alg02}
		\scriptsize
		\SetAlgoLined
		\KwResult{$\boldsymbol{h}_1$}
		Given a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $p$-variate observations\;
		Draw a random $(p + 1)$-subset $\boldsymbol{L}$\;
		Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
		\If{$det(\boldsymbol{\hat{S}}_1) \leq 0$}{
			Add another random observation into $\boldsymbol{L}$\;
			Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
		}
		Compute $\boldsymbol{d}_1^2(i) = (\boldsymbol{x}_i - \boldsymbol{\bar{x}}_1)^\prime \boldsymbol{\hat{S}}_1^{-1}(\boldsymbol{X}_i - \boldsymbol{\bar{x}}_0)$ for $i = 1, ..., n$\;
		Sort them into $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_0(\pi(n))$\;
		Put $\boldsymbol{h}_1 = \{\pi(1), ..., \pi(h)\}$\;
		\caption{Constructing the initial subset}
	\end{algorithm}
\end{frame}

\begin{frame}[c]{Fast-MCD}
	\begin{algorithm}[H]\label{alg:alg03}
		\scriptsize
		\SetAlgoLined
		\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
		Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n \leq 600$\;
		\For{500 times}{
			Construct initial $h$-subset $\boldsymbol{h}_1$ (Algorithm \ref{alg:alg02})\;
			Carry out two C-steps (Algorithm \ref{alg:alg01}) resulting $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$\;
		}		
		Put $\boldsymbol{h}_1$ as the 10 lowest $det(\boldsymbol{\hat{S}_2})$ computed above\;
		\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
			Carry out C-steps (Algorithm \ref{alg:alg01})\;
		}
		\caption{Fast-MCD when $n \leq 600$}
	\end{algorithm}
\end{frame}

\begin{frame}[c]{Fast-MCD}
	\begin{algorithm}[H]\label{alg:alg04}
		\scriptsize
		\SetAlgoLined
		\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
		Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n > 600$\;
		Construct up to 5 disjoint random subsets of size $n_{sub} = 300$\;
		\For{each subset, repeat 100 times}{
			Construct initial $\boldsymbol{h}_1$ of size $h_{sub} = [n_{sub}(h/n)]$\;
			Carry out two C-steps of $n_{sub}$ and $h_{sub}$\;
			Keep the 10 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$\;
		}		
		Pool the subsets, yielding the merged set of $n_{merged} = 1500$\;
		\For{50 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$}{
			Carry out two C-steps of $n_{merged}$ and $h_{merged} = [n_{merged}(h/n)]$\;
			Keep the 10 best $(\boldsymbol{\bar{x}}_{merged}, \boldsymbol{\hat{S}}_{merged})$\;
		}
		\For{the full data set and $m_{full}$ best results}{
			Take several C-steps, using $n$ and $h$\;
			Keep the best final result $(\boldsymbol{\bar{x}}_{full}, \boldsymbol{\hat{S}}_{full})$\;
		}
		\caption{Fast-MCD when $n > 600$}
	\end{algorithm}
\end{frame}

%section-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Network Anomaly Detection Based on Robust Moments}

\begin{frame}{Network Anomaly Detection Based on Robust Moments}
	We propose to extend Fast-MCD for computing robust moments (skewness and kurtosis), in order to improve the anomaly detection in a skewed and imbalanced data set of network traffic with normal and botnet traffic.

	Moments are a set of statistical parameters to measure a distribution. The arithmetic mean is the first general moment, the second is the variance, while skewness (asymmetry) is the third moment and kurtosis (excess) is the fourth moment.

	The general formula of the $r$-th moment can be expressed as:
	\begin{equation}\label{eq:eq05}
		m_r = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^r. 
	\end{equation}
\end{frame}

\begin{frame}{Network Anomaly Detection Based on Robust Moments}
	Therefore, the Skewness is calculated as
	\begin{equation}\label{eq:eq06}
		\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
	\end{equation}

	and Kurtosis as
	\begin{equation}\label{eq:eq07}
		\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2} 
	\end{equation}
\end{frame}

\begin{frame}{Network Anomaly Detection Based on Robust Moments}
	For extending Fast-MCD, We initially modify the C-step Algorithm \ref{alg:alg01} to calculate Skewness and Kurtosis after to minimize $det(\boldsymbol{\hat{S}})$, accordingly to Algorithm \ref{alg:alg05}.

	\begin{algorithm}[H]\label{alg:alg05}
		\scriptsize
		\SetAlgoLined
		\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
		Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
		\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
			\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
				Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
			}
			Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
			Sort $\boldsymbol{d}_1( i )$, which yields a index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
			Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
			Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
			Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
		}
		Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
		\caption{C-Step with higher moments}
	\end{algorithm}
\end{frame}

\begin{frame}[c]{Network Anomaly Detection Based on Robust Moments}
    The Algorithm \ref{alg:alg06} describes the Fast-MCD with higher moments when $n \leq 600$, through the modification of the original Fast-MCD for computing Skewness and Kurtosis from the robust estimate.

	\begin{algorithm}[H]\label{alg:alg06}
		\scriptsize
		\SetAlgoLined
		\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
		Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n \leq 600$\;
		\For{500 times}{
			Construct initial $h$-subset $\boldsymbol{h}_1$ (Algorithm \ref{alg:alg02})\;
			Carry out two C-steps with higher moments (Algorithm \ref{alg:alg05}) resulting $\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$\;
		}		
		Put $\boldsymbol{h}_1$ as the 10 lowest $det(\boldsymbol{\hat{S}_2})$ computed above\;
		\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
			Carry out C-steps with higher moments (Algorithm \ref{alg:alg05})\;
		}
		\caption{Fast-MCD with higher moments when $n \leq 600$}
	\end{algorithm}
\end{frame}

\begin{frame}[c]{Network Anomaly Detection Based on Robust Moments}
	The Algorithm \ref{alg:alg07} presents the Fast-MCD with higher moments when $n > 600$, which modifies the original Fast-MCD for computing of Skewness and Kurtosis from the robust estimate data, through the adoption of the C-Step with higher moments. The Algorithm \ref{alg:alg07} is described below.
\end{frame}

\begin{frame}[c]{Network Anomaly Detection Based on Robust Moments}
	\begin{algorithm}[H]\label{alg:alg07}
		\scriptsize
		\SetAlgoLined
		\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
		Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n > 600$\;
		Construct up to 5 disjoint random subsets of size $n_{sub} = 300$\;
		\For{each subset, repeat 100 times}{
			Construct initial $\boldsymbol{h}_1$ of size $h_{sub} = [n_{sub}(h/n)]$\;
			Carry out two C-steps with higher moments of $n_{sub}$ and $h_{sub}$\;
			Keep the 10 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub}, \boldsymbol{\bar{s}}_{sub}, \boldsymbol{\bar{k}}_{sub})$\;
		}		
		Pool the subsets, yielding the merged set of $n_{merged} = 1500$\;
		\For{50 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$}{
			Carry out two C-steps with higher moments of $n_{merged}$ and $h_{merged} = [n_{merged}(h/n)]$\;
			Keep the 10 best $(\boldsymbol{\bar{x}}_{merged}, \boldsymbol{\hat{S}}_{merged}, \boldsymbol{\bar{s}}_{merged}, \boldsymbol{\bar{k}}_{merged})$\;
		}
		\For{the full data set and $m_{full}$ best results}{
			Take several C-steps, using $n$ and $h$\;
			Keep the best final result $(\boldsymbol{\bar{x}}_{full}, \boldsymbol{\hat{S}}_{full}, \boldsymbol{\bar{s}}_{full}, \boldsymbol{\bar{k}}_{full})$\;
		}
		\caption{Fast-MCD with higher moments when $n > 600$}
	\end{algorithm}
\end{frame}

\begin{frame}{Network Anomaly Detection Based on Robust Moments}
    We also propose to use Mahalanobis Distance for detecting anomalies through an approach based on robust Skewness and Kurtosis, instead of the location and scatter that are used by classical approaches.

    We propose to modify the Equation \ref{eq:eq01} to implement a Skewness-based Mahalanobis Distance, as follows:

    \begin{equation}\label{eq:eq08}
		\boldsymbol{d}(\boldsymbol{s}_{\boldsymbol{x}}, \bar{\boldsymbol{s}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{s}_{\boldsymbol{x}} - \bar{\boldsymbol{s}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{s}_{\boldsymbol{x}} - \bar{\boldsymbol{s}})^\prime}, 
	\end{equation}

	where $\boldsymbol{s}_{\boldsymbol{x}} = \boldsymbol{x} - \boldsymbol{\bar{s}}(\boldsymbol{x})$ and $\boldsymbol{\bar{s}}(\boldsymbol{x})$ is the Skewness of $\boldsymbol{x}$.

\end{frame}

\begin{frame}{Network Anomaly Detection Based on Robust Moments}
    We also propose to modify the Equation \ref{eq:eq04} to implement a Kurtosis-based Mahalanobis Distance, as follows:

    \begin{equation}\label{eq:eq09}
		\boldsymbol{d}(\boldsymbol{k}_{\boldsymbol{x}}, \bar{\boldsymbol{k}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{k}_{\boldsymbol{x}} - \bar{\boldsymbol{k}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{k}_{\boldsymbol{x}} - \bar{\boldsymbol{k}})^\prime}, 
	\end{equation}

	where $\boldsymbol{k}_{\boldsymbol{x}} = \boldsymbol{x} - \boldsymbol{\bar{k}}(\boldsymbol{x})$ and $\boldsymbol{\bar{k}}(\boldsymbol{x})$ is the Kurtosis of $\boldsymbol{x}$.
\end{frame}

\begin{frame}[c]{Network Anomaly Detection with Fast-MCD}
	Fast-MCD is commonly used as \textbf{supervised} or \textbf{unsupervised} algorithm for anomaly detection.
    \begin{itemize}		
		\item \textbf{Supervised} approach:
		\begin{enumerate}
		    \item Compute Fast-MCD for training data with known normal and anomalous data.
		    \item Select the best threshold $t$ for detecting known anomalies according to largest distances $\boldsymbol{d}(i)$.
		    \item Save the best $t, \boldsymbol{\bar{x}}$ and $\boldsymbol{\hat{S}}$.
		    \item Compute $d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}})$ for new observations.
		    \item Classify observations with $\boldsymbol{d}(i)$ higher than $t$ as anomalous.
		\end{enumerate}
		\item \textbf{Unsupervised} approach:
		\begin{enumerate}
			\item Given a known contamination $c$, which is the percentage of anomalies in a data set.
			\item Compute Fast-MCD and $d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}})$ for new observations.
			\item Classify the $c$ observations with largest distances as anomalous.
		\end{enumerate}
	\end{itemize}
\end{frame}

\begin{frame}[c]{Network Anomaly Detection Based on Robust Moments}
    Finally, We propose the following \textbf{Semi Supervised} approach for Fast-MCD based anomaly detection:
	\begin{enumerate}
		\item Compute Fast-MCD for training data containing only known normal observations.
		\item Save the fitted $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$.
		\item Carry out cross validation on known contaminated data in order to identify the contamination $c$ that best classify anomalies and normal data.
		\item Compute $d(\boldsymbol{x},\bar{\boldsymbol{x}}_{normal}, \boldsymbol{\hat{S}}_{normal})$ for new observations.
		\item The $c$ observations with largest $\boldsymbol{d}(i)$ are classified as anomalous.
	\end{enumerate}
\end{frame}

%section-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Data Model}

\begin{frame}[c]{The CTU-13 Dataset}
	Evaluating network anomaly detection solutions remains a challenge due to the lack of up-to-date real-world data sets for training and testing \cite{osanaiye2016distributed}. 
	\begin{itemize}
		\item We propose to use CTU-13 \cite{garcia2014empirical} to evaluate our proposal, considering that the CTU-13 is a data set that aims to have a large capture of real botnet attacks and C\&C traffic mixed with normal and background traffic.
		\item The CTU-13 data set consists in thirteen captures (called scenarios) of different botnet samples. 
		\item On each scenario it was executed a specific malware, which used several protocols and performed different actions.		
	\end{itemize}
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	\begin{itemize}
		\item The botnet traffic is divided into attacks and Command and Control (C\&C) traffic.
		\begin{itemize}
			\item \textbf{Attacks:} Click Fraud, Port Scan, FastFlux, Compiled and Controled by Authors, SPAM and DDOS.
			\item \textbf{C\&C:} IRC, P2P and HTTP.
		\end{itemize}
		\item The data set contains the following features:
		\begin{itemize}
			\item Start Time, Duration, Protocol, Source Address, Source Port, Direction, Destination Address, Destination Port, State, Type of service from source to destination, Type of service from destination to source, Total of Packets, Total of bytes, Total bytes from source to destination, Label as normal or anomalous
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	\begin{table}[h!]
		\tiny
		\caption{CTU-13 Dataset Description}
		\label{tab:tab01}
		\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | }
			\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Scenario\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Malware\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Type\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Total Flows\end{tabular} & \begin{tabular}[x]{@{}l@{}}Malicious\end{tabular} & \begin{tabular}[x]{@{}l@{}}C\&C\end{tabular} & \begin{tabular}[x]{@{}l@{}}Botnet\end{tabular}\\ \hline
				10 & neris & IRC, Spam, CF & 2,824,636 & 40,961 (1.45\%) & 341 (0.01\%) & 40,620 (1.44\%) \\ \hline
				11 & neris & IRC, Spam, CF & 1,808,122 & 20,941 (1.16\%) & 673 (0.04\%) & 20,268 (1.12\%) \\ \hline
				12 & rbot & IRC, PS, US & 4,710,638 & 26,822 (0.57\%) & 63 (0.00\%) & 26,759 (0.57\%) \\ \hline
				15 & rbot & IRC, DDoS, US & 1,121,076 & 2,580 (0.23\%) & 52 (0.00\%) & 2,528 (0.23\%) \\ \hline
				15-2 & virut & Spam, PS,HTTP & 129,832 & 901 (0.69\%) & 24 (0.02\%) & 877 (0.68\%) \\ \hline
				16 & menti & PS & 558,919 & 4,630 (0.83\%) & 199 (0.04\%) & 4,431 (0.79\%) \\ \hline
				16-2 & sogou & HTTP & 114,077 & 63 (0.06\%) & 26 (0.02\%) & 37 (0.03\%) \\ \hline
				16-3 & murlo & PS & 2,954,230 & 6,127 (0.21\%) & 1,074 (0.04\%) & 5,053 (0.17\%) \\ \hline
				17 & neris & IRC, Spam, CF,PS & 2,087,508 & 184,987 (8.86\%) & 2,973 (0.14\%) & 182,014 (8.72\%) \\ \hline
				18 & rbot & IRC, DDoS, US & 1,309,791 & 106,352 (8.12\%) & 33 (0.00\%) & 106,319 (8.12\%) \\ \hline
				18-2 & rbot & IRC, DDoS, US & 107,251 & 8,164 (7.61\%) & 2 (0.00\%) & 8,162 (7.61\%) \\ \hline
				19 & nsys.ay & P2P & 325,471 & 2,168 (0.67\%) & 25 (0.01\%) & 2,143 (0.66\%) \\ \hline
				15-3 & virut & Spam, PS, HTTP & 1,925,149 & 40,003 (2.08\%) & 536 (0.03\%) & 39,467 (2.05\%) \\ \hline
		\end{tabular}
	\end{table}

	Table \ref{tab:tab01} presents the malwares used for botnet attacks, the attack and C\&C types, the total number of flows, the number of malicious flows, which includes the C\&C and botnet flows. Table \ref{tab:tab01} also shows the ratio of C\&C and botnet flows.
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	\begin{itemize}
		\item Table \ref{tab:tab01} shows that CTU-13 is very imbalanced data set, when comparing the amount of malicious to normal traffic.
		\item Malicious observations are beteeen $0.06\%$ up to $8.86\%$ of all flows.
		\item The scenarios have between $107,251$ and $4,710,638$ flows.
		\item The CTU-13 data set is also very skewed. Figure \ref{fig:fig06} exemplifies the skewed distribution of normal and anomalous flows, showing the flow duration of scenario 12.
	\end{itemize}
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=8cm]{figures/raw_distplot_capture20110812_Dur.png}
	     \caption{Distribution of flow duration for scenario 12.}
	     \label{fig:fig06}
	\end{figure}
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	\begin{itemize}
		\item Methods for imbalance learning has attracted growing attention from both academia and industry, due to the restrictions imposed by imbalanced data to machine learning algorithms.
		\item He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning.
		\item Data aggregation is a widely adoted approach for feature engineering \cite{chandrashekar2014survey,acarali2016survey}.
		\item Lakhina et al. \cite{lakhina2005mining} argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
	\end{itemize}
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	To deal with the data imbalance of CTU-13 data set, we adopt a sampling method based on data aggragation \cite{acarali2016survey} for feature generation, in order to obtain less imbalanced and more discrimnative data.	
	\begin{itemize}
		\item Based on related works \cite{lakhina2005mining,chandrashekar2014survey,acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
		\item The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15, 0.25, 1 and 2 seconds, for each scenario of the CTU-13 data set.
	\end{itemize}
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	Following (2 slides) are the selected features:
	\begin{itemize}
		\item \textbf{normal-flow-count:} count of normal (known) flow.
		\item \textbf{background-flow-count :} count of background (unknown) flows.
		\item \textbf{avg-duration:} average duration time of flows.
		\item \textbf{n-conn:} count of flows.
		\item \textbf{n-icmp:} count of ICMP packets.
		\item \textbf{n-tcp:} count of TCP packets.
		\item \textbf{n-udp:} count of UDP packets.
		\item \textbf{n-dports<1024:} count of destinations ports lower than 1024.
		\item \textbf{n-sports<1024:} count of source ports lower than 1024.
		\item \textbf{n-sports>1024:} count of source ports higher than 1024.
	\end{itemize}
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	Following are the remaining selected features:
	\begin{itemize}
		\item \textbf{n-s-a-p-address:} count of class A source address.
		\item \textbf{n-d-a-p-address:} count of class A destination address.
		\item \textbf{n-s-b-p-address:} count of class B source address.
		\item \textbf{n-d-b-p-address:} count of class B destination address.
		\item \textbf{n-s-c-p-address:} count of class C source address.
		\item \textbf{n-d-c-p-address:} count of class C destination address.
		\item \textbf{n-s-na-p-address:} count of class D* source address.
		\item \textbf{n-d-na-p-address:} count of class D* destination address.
	\end{itemize}
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	\begin{itemize}
		\item Figures \ref{fig:fig07} and \ref{fig:fig08} show the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 and 2 seconds.
		\item In comparison to Figure \ref{fig:fig06}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on the average is more discriminative.
	\end{itemize}
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=8cm]{figures/agg_distplot_0_15s_12_avg_duration.png}
	     \caption{Distribution of average flow duration for scenario 12, aggregated by windows of 0.15 seconds.}
	     \label{fig:fig07}
	\end{figure}
\end{frame}

\begin{frame}[c]{The CTU-13 Dataset}
	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=8cm]{figures/agg_distplot_2s_12_avg_duration.png}
	     \caption{Distribution of average flow duration for scenario 12, aggregated by windows of 2 seconds.}
	     \label{fig:fig08}
	\end{figure}
\end{frame}

%section-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Experiment}

\begin{frame}[c]{Experiment}
	\textbf{Experiment Plan:}
	\begin{itemize}
		\item Given a contaminated data set, split it into normal (60\%), cross validation (20\% normal + 50\% anomalous) and test (20\% normal + 50\% anomalous).
		\item Select the range of contamination that shall be evaluated and perform cross validation for each contamination. 
		\item For each contamination $c$:
		\begin{enumerate}
			\item Estimate $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$ from normal data.
			\item Predict anomalies for cross validation data using the $c$, $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$.
			\item Select the best F-measure by cross validation and get the best $c$.
		\end{enumerate}
		\item Test the F-measure of anomaly detection for new observations through $d(\boldsymbol{x},\bar{\boldsymbol{x}}_0, \boldsymbol{\hat{S}}_0)$ and $c$.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Experiment}
	\textbf{Experiment Plan:}
	\begin{itemize}
	    \item Evaluation Measure:
		\begin{itemize}
			\item In anomalous detection problems, where anomalies are rare events, if we classify all observations as normal and apply a accuracy evaluation, we would have high accuracy but poor true-positive detection.
			\item Due to the importance given by F-measure to true-positive detection \cite{moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.
    	\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}[c]{Experiment}
	\textbf{Experiment Plan:}
	\begin{itemize}
		\item The F-measure is the harmonic mean of precision and recall and is defined by 
		\begin{equation}\label{eq:eq10}
			F-measure = 2 * \frac{PR * RC}{PR + RC},				
		\end{equation}
		where $PR$ is the precision, defined by 
		\begin{equation}\label{eq:eq11}
			PR = \frac{True Positive}{True Positive + False Positive}
		\end{equation}
		and $RC$ is the recall, defined by 
		\begin{equation}\label{eq:eq12}
			RC = \frac{True Positive}{True Positive + False Negative}.
		\end{equation}
	\end{itemize}	
\end{frame}

\begin{frame}[c]{Experiment}
	\textbf{Data splitting for training, cross validation and testing}
	\begin{itemize}
		\item Gargia et al. \cite{garcia2014empirical} propose a division of CTU-13's scearios into train and test, for evaluation of network attack detection algorithms. 
		\item However, we decide to evaluate our proposed approach for each scenario individually, considering that in the division proposed by Gargia et al. some bots are present only for training and not for testing. 
		\item Although our proposal does not rely on anomalous data for training, we desire to evaluate our proposal to detect all types of bots captured by CTU-13 data set.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Experiment}
	We propose to compare our approach to clustering and statisctical based approaches (K-means and Gaussian Mixture Model (GMM), respectively).
	\begin{itemize}
		\item As clustering-based method, we selected K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007k,moustafa2019holistic}.
		\item It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}.
	\end{itemize}
\end{frame}

%section-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Results}

\begin{frame}[c]{Results}
	\begin{table}[h!]
	  \centering
	  \scriptsize
	  \caption{Results grouped by algorithm and scenario}
	  \label{tab:tab02}
	  \begin{tabular}{ c|c c|c c|c c|c c|c c }
		\toprule
		\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{S-MCD}} &\multicolumn{2}{c}{\textbf{K-MCD}} &\multicolumn{2}{c}{\textbf{MCD}} &\multicolumn{2}{c}{\textbf{GMM}} &\multicolumn{2}{c}{\textbf{K-Means}}\\ 
				\hhline{~----------}
				&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
		\midrule
			10 &\color{red} 0.71 & 0.03 & 0.66 & 0.02 & 0.34 & 0.52 & 0.34 & 0.46 & 0.20 & 0.25 \\ \hline
			11 & 0.03 & 0.54 & 0.00 & 0.54 & 0.71 &\color{red} 0.73 & 0.35 & 0.48 & 0.13 & 0.15 \\ \hline
			12 & 0.19 & 0.02 & 0.20 & 0.05 & 0.06 &\color{red} 0.31 & 0.01 & 0.01 & 0.09 & 0.11 \\ \hline
			15 & 0.00 & 0.00 & 0.00 & 0.00 & 0.24 &\color{red} 0.32 & 0.14 & 0.22 & 0.09 & 0.11 \\ \hline
			15-2 & 0.04 & 0.85 & 0.67 &\color{red} 0.86 & 0.36 & 0.28 & 0.32 & 0.14 & 0.18 & 0.26 \\ \hline
			15-3 & 0.00 & 0.01 & 0.00 & 0.01 &\color{red} 0.55 & 0.48 & 0.00 & 0.01 & 0.22 & 0.29 \\ \hline
			16 & 0.45 & 0.36 & 0.42 & 0.33 & 0.27 & 0.41 &\color{red} 0.51 & 0.49 & 0.07 & 0.16 \\ \hline
			16-2 & 0.00 & 0.16 & 0.00 &\color{red} 0.20 & 0.04 & 0.04 & 0.02 & 0.08 & 0.04 & 0.03 \\ \hline
			16-3 & 0.00 & 0.05 & 0.00 & 0.07 &\color{red} 0.15 & 0.10 & 0.01 & 0.01 & 0.03 & 0.04 \\ \hline
			17 & 0.03 & 0.03 & 0.07 & 0.03 &\color{red} 0.59 & 0.53 & 0.16 & 0.47 & 0.35 & 0.39 \\ \hline
			18 & 0.63 & 0.10 & 0.63 & 0.08 & 0.70 &\color{red} 0.72 & 0.21 & 0.17 & 0.13 & 0.20 \\ \hline
			18-2 & 0.83 & 0.69 & 0.85 & 0.43 & 0.84 &\color{red} 0.91 & 0.50 & 0.55 & 0.62 & 0.70 \\ \hline
			19 &\color{red} 0.71 & 0.00 & 0.68 & 0.00 & 0.11 & 0.42 & 0.17 & 0.20 & 0.20 & 0.17 \\ \hline
			\rowcolor{Gray} \textbf{avg} & \textbf{0.28} & \textbf{0.22} & \textbf{0.32} & \textbf{0.20} & \textbf{0.38} &\color{red} \textbf{0.44} & \textbf{0.21} & \textbf{0.25} & \textbf{0.18} & \textbf{0.22} \\ 
	    \bottomrule
	  \end{tabular}
	\end{table}
\end{frame}

\begin{frame}[c]{Results}
	\begin{table}[h!]
	  \centering
	  \scriptsize
	  \caption{Results grouped by algorithm and scenario}
	  \label{tab:tab03}
	  \begin{tabular}{ c|c c|c c|c c|c c }
		\toprule
		\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{RPCA}} &\multicolumn{2}{c}{\textbf{S-RPCA}} &\multicolumn{2}{c}{\textbf{S-RPCA2}} &\multicolumn{2}{c}{\textbf{MCD}}\\ 
				\hhline{~--------}
				&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
		\midrule
			10 & 0.93 & 0.93 &\color{red} 0.98 &\color{red} 0.98 &\color{red} 0.99 &\color{red} 1.00 & 0.34 & 0.52 \\ \hline
			11 & 0.96 & 0.94 &\color{red} 0.99 &\color{red} 0.98 &\color{red} 0.99 &\color{red} 1.00 & 0.71 & 0.73 \\ \hline
			12 & 0.32 & 0.19 & 0.00 & 0.00 &\color{red} 0.81 & 0.01 & 0.06 & 0.31 \\ \hline
			15 & 0.75 & 0.73 & 0.86 &\color{red} 0.96 &\color{red} 1.00 &\color{red} 0.96 & 0.24 & 0.32 \\ \hline
			15-2 & 0.40 & 0.57 &\color{red} 0.95 &\color{red} 0.94 & 0.25 & 0.93 & 0.36 & 0.28 \\ \hline
			15-3 & 0.68 & 0.68 &\color{red} 0.96 & 0.70 &\color{red} 0.96 &\color{red} 0.96 & 0.55 & 0.48 \\ \hline
			16 &\color{red} 0.47 &\color{red} 0.47 & 0.01 & 0.01 & 0.01 & 0.00 & 0.27 & 0.41 \\ \hline
			16-2 & 0.07 & 0.10 &\color{red} 0.80 & 0.68 & 0.00 & 0.00 & 0.04 & 0.04 \\ \hline
			16-3 & 0.29 & 0.29 & 0.01 & 0.46 &\color{red} 0.79 & 0.76 & 0.15 & 0.10 \\ \hline
			17 &\color{red} 0.82 & 0.79 &\color{red} 0.81 & 0.78 &\color{red} 0.82 & 0.79 & 0.59 & 0.53 \\ \hline
			18 & 0.87 & 0.89 & 0.97 &\color{red} 0.98 &\color{red} 0.98 &\color{red} 0.98 & 0.70 & 0.72 \\ \hline
			18-2 &\color{red} 0.94 &\color{red} 0.94 &\color{red} 0.91 & 0.86 & 0.09 & 0.10 & 0.84 & 0.91 \\ \hline
			19 & 0.27 & 0.31 &\color{red} 0.95 &\color{red} 0.95 & 0.02 & 0.02 & 0.11 & 0.42 \\ \hline
			\rowcolor{Gray} \textbf{Avg} & \textbf{0.60} & \textbf{0.60} &\color{red} \textbf{0.71} &\color{red} \textbf{0.71} & \textbf{0.59} & \textbf{0.58} & \textbf{0.38} & \textbf{0.44} \\
	    \bottomrule
	  \end{tabular}
	\end{table}
\end{frame}

\begin{frame}[c]{Results}
	Results grouped by algorithm for full CTU-13.
	\begin{enumerate}
		\item ...
	\end{enumerate}
\end{frame}

\begin{frame}[c]{Discussion}
	Discuss results..
	\begin{enumerate}
		\item there are no lots of researches using CTU-13, although it is been growing
		\item poor results for GMM, K-Means and MGM are ok and in accordance to \cite{wang2017botnet}, which also evaluate \cite{gu2007bothunter}
	\end{enumerate}
\end{frame}

%section-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Next Steps}
\begin{frame}[c]{Next Steps}
	\begin{enumerate}
		\item Say that the data is lognormal \cite{benson2010network} and link it to skewness https://www.weibull.com/hotwire/issue47/relbasics47.htm
		\item Test distribution fit again e present results
	    \item Conclude this presentation.
		\item Evaluate eigensim for CTU-13 and others algorithms for Danilo's data set.
		\item Compare to vanilla RPCA (Eduardo is working on it).
		\item Write a paper for JNCA (April).
		\item Write the thesis (May and June).
	\end{enumerate}
\end{frame}

%section-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\begin{frame}[plain, allowframebreaks]{Bibliography}
	\setbeamertemplate{bibliography item}[text]
	\printbibliography
\end{frame}

\end{document}

### 2019-04-21 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
copied from slides to paper

@@
 \usepackage[utf8x]{inputenc}
-\usepackage{algorithm,algorithmic}% http://ctan.org/pkg/algorithms
+\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
 \usepackage{natbib}
@@
 \usepackage{tabularx}
-\usepackage{colortbl}
 \usepackage{multirow}
@@
 
-% declarations
-\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
-\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\newcommand*\justify{%
-	\fontdimen2\font=0.4em% interword space
-	\fontdimen3\font=0.2em% interword stretch
-	\fontdimen4\font=0.1em% interword shrink
-	\fontdimen7\font=0.1em% extra space
-	%\hyphenchar\font=`\-% allowing hyphenation
-}
-
 % table
@@
 \definecolor{Gray}{gray}{0.9}
-
-% Algorithmic modifications
-\makeatletter
-\newcommand{\ALOOP}[1]{\ALC@it\algorithmicloop\ #1%
-  \begin{ALC@loop}}
-\newcommand{\ENDALOOP}{\end{ALC@loop}\ALC@it\algorithmicendloop}
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\newcommand{\algorithmicbreak}{\textbf{break}}
-\newcommand{\BREAK}{\STATE \algorithmicbreak}
-\makeatother
 
@@
 
+Some techniques can make the data discriminative for better anomaly detection: Data engineering (i.e.
+sampling \cite{??}, aggregation \cite{acarali2016survey}); Feature extraction (i.e.
+decomposition \cite{??}, robust estimates \cite{??}).
+
+Robust statistics produce methods that are not unduly affected by outliers.
+
+Robust statistical methods have been developed for estimating location, scale, and regression parameters.
+
+Robust estimates from contaminated data can denoise, remove anomalies or even separate them.
+
+Anomaly detection based on robust estimates has been adopted for network anomaly detection, due to good results and its adaptability.
+
+The distance between robust estimates and new observations can reveal anomalies.
+
+Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples by, considering the correlations between the variables and their mean values.
+
+\textbf{MD has been used for distance based anomaly detection with robust estimates} in many areas, and is commonly combined to Fast Minimum Covariance Determinant (Fast-MCD) algorithm \cite{rousseeuw1999fast} for robust estimates.
+
+Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority}.
+
+Learning algorithms for imbalanced data has been a relevant research topic.
+
+The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
+
+Skewness and the kurtosis tests are very sensitive outlier detectors.
+The properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2008outlier}.
+
+We propose a network anomaly detection approach based on robust moment estimates and distance analysis.
+
+Considering that Fast-MCD is a widely adopted robust estimator of location and scatter for anomaly detection, we propose to extend Fast-MCD for computing robust skewness and kurtosis.
+
+We also propose to use Mahalanobis Distance for detecting anomalies based on robust skewness and kurtosis, instead of the location-based usage of classical approaches.
+
+Considering that it is hard to have labeled data in real problems \cite{osanaiye2016distributed}, such as for network attacks or frauds, also considering that known anomalies are constantly changing due to adversarial models or due to lack of regular patterns, we propose a Semi-Supervised approach based on robust moments for network anomaly detection.
+
 In the context of anomaly-based schemes, this work proposes a statistical approach based on cumulants and Mimimum Covariance Determinant (MCD) for network attack detection.
@@
 
+\cite{gu2007bothunter} (Bothunter: Detecting malware infection through ids-driven dialog correlation)
+
+\cite{gu2008botminer} (Botminer: Clustering analysis of network traffic for protocol-and structure-independent botnet detection)
+
+\cite{hubert2008outlier} (Outlier detection for skewed data)
+
+\cite{garcia2014empirical} (An empirical comparison of botnet detection methods)
+
+\cite{khattak2015botflex} (BotFlex: A community-driven tool for botnet detection)
+
+\cite{wang2017botnet} (Botnet detection based on anomaly and community detection)
+
+\cite{da2018online} (Online detection of Botnets on Network Flows using Stream Mining)
+
+\cite{moustafa2019holistic} (A holistic review of Network Anomaly Detection Systems: A comprehensive survey).
+Network anomaly detection approaches are classified by \cite{moustafa2019holistic} in six categories: classification, clustering, deep learning, knowledge, combination and statistical-based.
+
+\cite{Phua2004minority} (Minority Report in Fraud Detection: Classification of Skewed Data)
+
+\cite{acarali2016survey} (Survey of approaches and features for the identification of HTTP-based botnet traffic)
+
+\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions)
+
+\cite{goldberg2015importance} (The importance of features for statistical anomaly detection)
+
+Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{moustafa2019holistic}.
+
+GMM is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.
+
+The Mahalanobis Distance is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
+C.
+Mahalanobis in 1936 \cite{mahalanobis1936generalized}.
+It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\hat{S}}$ of $\boldsymbol{X}$.
+
+Mahalanobis Distance is defined as		
+	\begin{equation}\label{eq:eq01}
+		d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}.
+	\end{equation}
+$\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector, also referred as location, of known observations and $\boldsymbol{\hat{S}}$ is the covariance matrix, also referred as scatter, of known observations.
+
+Classical estimates can be so strongly affected by contamination that diagnostic tools, such as the Mahalanobis Distances, become unable to detect the outliers.
+Therefore, we need reliable and robust estimators that can resist outliers contaminated data.
+Minimum Covariance Determinant (MCD) \cite{rousseeuw1984least} is a robust estimator commonly used for computing Mahalanobis Distance in order to detect anomalies.
+
+\begin{figure}[h!]
+     \centering
+     \includegraphics[width=11cm]{figures/mahalanobis_robust.png}
+     \caption{Mahalanobis Distance for non-robust and robust estimates.}
+     \label{fig:fig04}
+\end{figure}
+
+\begin{figure}[h!]
+     \centering
+     \includegraphics[width=8cm]{figures/mahalanobis_distances01.png}
+     \caption{Mahalanobis Distance for non-robust and robust estimates.}
+     \label{fig:fig05}
+\end{figure}
+
+The MCD estimator is one of the first affine equivariant and highly robust estimators of multivariate location and scatter.
+
+Affine equivariant means that when the data are translated or subjected to a linear transformation, the resulting location and scatter will transform accordingly \cite{rousseeuw1984least,rousseeuw1999fast}.
+
+Its objective is to find $h$ observations (out of $n$) whose covariance matrix has the lowest determinant.
+
+The lowest determinant means the lowest distance generalized variance and high distance between the largest eigenvalue and the remaining [ref?].
+
+MCD depends on random sampling, therefore it is not deterministic.
+
+MCD has been applied in numerous fields (medicine, finance, image analysis, and chemistry) and been used to develop many robust multivariate techniques (such as Robust Principal Component Analysis (RPCA), factor analysis, and multiple regression).
+
+The classical MCD has rarely been applied because it is hard to compute, since it requires the evaluation of all $\binom{n}{h}$ subsets of size $h$. % binomial of n over p
+
+Its main use has started since the construction of the computationally efficient Fast-MCD algorithm \cite{rousseeuw1999fast}.
+
+Fast-MCD is an iterative, approximate and resampling algorithm for the MCD.
+
+It starts by random initial subset of size $p+1$ and performs concentration steps (C-steps) yielding consecutive h-subsets with decreasing covariance matrix determinant.
+
+Only two C-steps are applied to each initial subset, and the 10 results with lowest determinant are kept.
+
+C-steps are carried out until convergence and the best solution is kept.
+
+Fast-MCD also is affine equivariant and not deterministic.
+
+\textbf{Theorem 1: C-Step.}
+
+Consider a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ where $\boldsymbol{X} = \{\boldsymbol{x}_1,...,\boldsymbol{x}_n\}$ of $p$-variate observations.
+Let $\boldsymbol{h} \subset \{1,...,n\}$ with $|\boldsymbol{h}| = h$, where $\boldsymbol{h}$ is a subset of $h$ indexes from $n$ observations.
+
+Let the mean $\bar{\boldsymbol{x}} \in \mathbb{R}^{1 \times p}$ be
+
+\begin{equation}\label{eq:eq02}
+	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} \boldsymbol{x}_j, 
+\end{equation}
+
+the covariance matrix $\boldsymbol{\hat{S}} \in \mathbb{R}^{p \times p}$ be
+
+\begin{equation}\label{eq:eq03}
+	\boldsymbol{\hat{S}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} (\boldsymbol{x}_j - \bar{\boldsymbol{x}})(\boldsymbol{x}_j - \bar{\boldsymbol{x}})^\prime,
+\end{equation}
+
+and the relative distance be
+
+\begin{equation}\label{eq:eq04}
+	\boldsymbol{d}(i) = \sqrt{(\boldsymbol{x}_i - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime}, 
+\end{equation}
+
+for $i = 1,...,n$.
+
+Let $\boldsymbol{h}_1, \bar{\boldsymbol{x}}_1$ and $\boldsymbol{\hat{S}}_1$ and be the selected observations, the location and scatter of the initial subset, and $\boldsymbol{d}_1(i)$ be the initial distances for $i = 1,...,n$.
+
+Now take $\boldsymbol{h}_2$ containing the indexes of lowest distances, such that $\{\boldsymbol{d}_1(i); i \in \boldsymbol{h}_2\} = \{(\boldsymbol{d}_1)_{1:n},...,(\boldsymbol{d}_1)_{h:n}\}$, where $(\boldsymbol{d}_1)_{1:n} \leq (\boldsymbol{d}_1)_{2:n} \leq ...
+\leq (\boldsymbol{d}_1)_{n:n}$ are the ordered distances, and compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$.
+
+Then, $det(\boldsymbol{\hat{S}}_2) \leq det(\boldsymbol{\hat{S}}_1)$ with equality if and only if $\boldsymbol{\bar{x}}_2 = \boldsymbol{\bar{x}}_1$ and $\boldsymbol{\bar{S}}_2 = \boldsymbol{\hat{S}}_1$
+
+The proof is given by Rousseeuw and Driessen \cite{rousseeuw1999fast}.
+
+\textbf{The C-step can be algorithmically described as follows:}
+\begin{algorithm}[H]\label{alg:alg01}
+	\scriptsize
+	\SetAlgoLined
+	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
+	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
+	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
+		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
+			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
+		}
+		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
+		Sort $\boldsymbol{d}_1( i )$, which yields a index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
+		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
+		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
+		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
+	}
+	\caption{C-Step}
+\end{algorithm}
+
+\textbf{Theorem 1} thus provides a partial idea for an algorithm:
+
+\textit{Take many initial choices of $\boldsymbol{h}_1$ and apply C-steps to each until convergence (don't have determinant decreasing), and keep the result with lowest determinant}.
+
+\textbf{Initial subset can be algorithmically described as:}
+\begin{algorithm}[H]\label{alg:alg02}
+	\scriptsize
+	\SetAlgoLined
+	\KwResult{$\boldsymbol{h}_1$}
+	Given a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $p$-variate observations\;
+	Draw a random $(p + 1)$-subset $\boldsymbol{L}$\;
+	Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
+	\If{$det(\boldsymbol{\hat{S}}_1) \leq 0$}{
+		Add another random observation into $\boldsymbol{L}$\;
+		Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
+	}
+	Compute $\boldsymbol{d}_1^2(i) = (\boldsymbol{x}_i - \boldsymbol{\bar{x}}_1)^\prime \boldsymbol{\hat{S}}_1^{-1}(\boldsymbol{X}_i - \boldsymbol{\bar{x}}_0)$ for $i = 1, ..., n$\;
+	Sort them into $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_0(\pi(n))$\;
+	Put $\boldsymbol{h}_1 = \{\pi(1), ..., \pi(h)\}$\;
+	\caption{Constructing the initial subset}
+\end{algorithm}
+
+\begin{algorithm}[H]\label{alg:alg03}
+	\scriptsize
+	\SetAlgoLined
+	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
+	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n \leq 600$\;
+	\For{500 times}{
+		Construct initial $h$-subset $\boldsymbol{h}_1$ (Algorithm \ref{alg:alg02})\;
+		Carry out two C-steps (Algorithm \ref{alg:alg01}) resulting $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$\;
+	}		
+	Put $\boldsymbol{h}_1$ as the 10 lowest $det(\boldsymbol{\hat{S}_2})$ computed above\;
+	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
+		Carry out C-steps (Algorithm \ref{alg:alg01})\;
+	}
+	\caption{Fast-MCD when $n \leq 600$}
+\end{algorithm}
+
+\begin{algorithm}[H]\label{alg:alg04}
+	\scriptsize
+	\SetAlgoLined
+	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
+	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n > 600$\;
+	Construct up to 5 disjoint random subsets of size $n_{sub} = 300$\;
+	\For{each subset, repeat 100 times}{
+		Construct initial $\boldsymbol{h}_1$ of size $h_{sub} = [n_{sub}(h/n)]$\;
+		Carry out two C-steps of $n_{sub}$ and $h_{sub}$\;
+		Keep the 10 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$\;
+	}		
+	Pool the subsets, yielding the merged set of $n_{merged} = 1500$\;
+	\For{50 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$}{
+		Carry out two C-steps of $n_{merged}$ and $h_{merged} = [n_{merged}(h/n)]$\;
+		Keep the 10 best $(\boldsymbol{\bar{x}}_{merged}, \boldsymbol{\hat{S}}_{merged})$\;
+	}
+	\For{the full data set and $m_{full}$ best results}{
+		Take several C-steps, using $n$ and $h$\;
+		Keep the best final result $(\boldsymbol{\bar{x}}_{full}, \boldsymbol{\hat{S}}_{full})$\;
+	}
+	\caption{Fast-MCD when $n > 600$}
+\end{algorithm}
+
 Our proposal is ...
@@
 
-...
+Evaluating network anomaly detection solutions remains a challenge due to the lack of up-to-date real-world data sets for training and testing \cite{osanaiye2016distributed}.
+
+We propose to use CTU-13 \cite{garcia2014empirical} to evaluate our proposal, considering that the CTU-13 is a data set that aims to have a large capture of real botnet attacks and C\&C traffic mixed with normal and background traffic.
+
+The CTU-13 data set consists in thirteen captures (called scenarios) of different botnet samples.
+
+On each scenario it was executed a specific malware, which used several protocols and performed different actions.
+
+The botnet traffic is divided into attacks and Command and Control (C\&C) traffic.
+
+\textbf{Attacks:} Click Fraud, Port Scan, FastFlux, Compiled and Controled by Authors, SPAM and DDOS.
+
+\textbf{C\&C:} IRC, P2P and HTTP.
+
+The data set contains the following features:
+
+Start Time, Duration, Protocol, Source Address, Source Port, Direction, Destination Address, Destination Port, State, Type of service from source to destination, Type of service from destination to source, Total of Packets, Total of bytes, Total bytes from source to destination, Label as normal or anomalous
+
+\begin{table}[h!]
+	\tiny
+	\caption{CTU-13 Dataset Description}
+	\label{tab:tab01}
+	\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | }
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Scenario\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Malware\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Type\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Total Flows\end{tabular} & \begin{tabular}[x]{@{}l@{}}Malicious\end{tabular} & \begin{tabular}[x]{@{}l@{}}C\&C\end{tabular} & \begin{tabular}[x]{@{}l@{}}Botnet\end{tabular}\\ \hline
+			10 & neris & IRC, Spam, CF & 2,824,636 & 40,961 (1.45\%) & 341 (0.01\%) & 40,620 (1.44\%) \\ \hline
+			11 & neris & IRC, Spam, CF & 1,808,122 & 20,941 (1.16\%) & 673 (0.04\%) & 20,268 (1.12\%) \\ \hline
+			12 & rbot & IRC, PS, US & 4,710,638 & 26,822 (0.57\%) & 63 (0.00\%) & 26,759 (0.57\%) \\ \hline
+			15 & rbot & IRC, DDoS, US & 1,121,076 & 2,580 (0.23\%) & 52 (0.00\%) & 2,528 (0.23\%) \\ \hline
+			15-2 & virut & Spam, PS,HTTP & 129,832 & 901 (0.69\%) & 24 (0.02\%) & 877 (0.68\%) \\ \hline
+			16 & menti & PS & 558,919 & 4,630 (0.83\%) & 199 (0.04\%) & 4,431 (0.79\%) \\ \hline
+			16-2 & sogou & HTTP & 114,077 & 63 (0.06\%) & 26 (0.02\%) & 37 (0.03\%) \\ \hline
+			16-3 & murlo & PS & 2,954,230 & 6,127 (0.21\%) & 1,074 (0.04\%) & 5,053 (0.17\%) \\ \hline
+			17 & neris & IRC, Spam, CF,PS & 2,087,508 & 184,987 (8.86\%) & 2,973 (0.14\%) & 182,014 (8.72\%) \\ \hline
+			18 & rbot & IRC, DDoS, US & 1,309,791 & 106,352 (8.12\%) & 33 (0.00\%) & 106,319 (8.12\%) \\ \hline
+			18-2 & rbot & IRC, DDoS, US & 107,251 & 8,164 (7.61\%) & 2 (0.00\%) & 8,162 (7.61\%) \\ \hline
+			19 & nsys.ay & P2P & 325,471 & 2,168 (0.67\%) & 25 (0.01\%) & 2,143 (0.66\%) \\ \hline
+			15-3 & virut & Spam, PS, HTTP & 1,925,149 & 40,003 (2.08\%) & 536 (0.03\%) & 39,467 (2.05\%) \\ \hline
+	\end{tabular}
+\end{table}
+
+Table \ref{tab:tab01} presents the malwares used for botnet attacks, the attack and C\&C types, the total number of flows, the number of malicious flows, which includes the C\&C and botnet flows.
+Table \ref{tab:tab01} also shows the ratio of C\&C and botnet flows.
+
+Table \ref{tab:tab01} shows that CTU-13 is very imbalanced data set, when comparing the amount of malicious to normal traffic.
+
+Malicious observations are beteeen $0.06\%$ up to $8.86\%$ of all flows.
+
+The scenarios have between $107,251$ and $4,710,638$ flows.
+
+The CTU-13 data set is also very skewed. Figure \ref{fig:fig06} exemplifies the skewed distribution of normal and anomalous flows, showing the flow duration of scenario 12.
+
+\begin{figure}[h!]
+     \centering
+     \includegraphics[width=8cm]{figures/raw_distplot_capture20110812_Dur.png}
+     \caption{Distribution of flow duration for scenario 12.}
+     \label{fig:fig06}
+\end{figure}
+
+Methods for imbalance learning has attracted growing attention from both academia and industry, due to the restrictions imposed by imbalanced data to machine learning algorithms.
+
+He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning.
+
+Data aggregation is a widely adoted approach for feature engineering \cite{chandrashekar2014survey,acarali2016survey}.
+
+Lakhina et al.
+\cite{lakhina2005mining} argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
+
+To deal with the data imbalance of CTU-13 data set, we adopt a sampling method based on data aggragation \cite{acarali2016survey} for feature generation, in order to obtain less imbalanced and more discrimnative data.
+
+Based on related works \cite{lakhina2005mining,chandrashekar2014survey,acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
+The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15, 0.25, 1 and 2 seconds, for each scenario of the CTU-13 data set.
+
+Following (2 slides) are the selected features:
+\textbf{normal-flow-count:} count of normal (known) flow.
+\textbf{background-flow-count :} count of background (unknown) flows.
+\textbf{avg-duration:} average duration time of flows.
+\textbf{n-conn:} count of flows.
+\textbf{n-icmp:} count of ICMP packets.
+\textbf{n-tcp:} count of TCP packets.
+\textbf{n-udp:} count of UDP packets.
+\textbf{n-dports<1024:} count of destinations ports lower than 1024.
+\textbf{n-sports<1024:} count of source ports lower than 1024.
+\textbf{n-sports>1024:} count of source ports higher than 1024.
+
+Following are the remaining selected features:
+	\textbf{n-s-a-p-address:} count of class A source address.
+	\textbf{n-d-a-p-address:} count of class A destination address.
+	\textbf{n-s-b-p-address:} count of class B source address.
+	\textbf{n-d-b-p-address:} count of class B destination address.
+	\textbf{n-s-c-p-address:} count of class C source address.
+	\textbf{n-d-c-p-address:} count of class C destination address.
+	\textbf{n-s-na-p-address:} count of class D* source address.
+	\textbf{n-d-na-p-address:} count of class D* destination address.
+
+Figures \ref{fig:fig07} and \ref{fig:fig08} show the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 and 2 seconds.
+
+In comparison to Figure \ref{fig:fig06}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on the average is more discriminative.
+
+\begin{figure}[h!]
+     \centering
+     \includegraphics[width=8cm]{figures/agg_distplot_0_15s_12_avg_duration.png}
+     \caption{Distribution of average flow duration for scenario 12, aggregated by windows of 0.15 seconds.}
+     \label{fig:fig07}
+\end{figure}
+
+\begin{figure}[h!]
+     \centering
+     \includegraphics[width=8cm]{figures/agg_distplot_2s_12_avg_duration.png}
+     \caption{Distribution of average flow duration for scenario 12, aggregated by windows of 2 seconds.}
+     \label{fig:fig08}
+\end{figure}
 
@@
 
-The proposed attack detection algorithm starts by the data pre-processing of a network traffic log containing IP, ports and timestamp of senders and receivers.
-During this step, the desired information is extracted in order to classify and count packets according to the origin and destination ports, and subsequently this information is grouped by minutes and by time frames.
+We propose to extend Fast-MCD for computing robust moments (skewness and kurtosis), in order to improve the anomaly detection in a skewed and imbalanced data set of network traffic with normal and botnet traffic.
 
-With the data grouped into $Q$ time frames, the framework considers the time variations of the matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the attack.
+Moments are a set of statistical parameters to measure a distribution.
+The arithmetic mean is the first general moment, the second is the variance, while skewness (asymmetry) is the third moment and kurtosis (excess) is the fourth moment.
 
-According to flood and port scan attacks' behavior, flood attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively.
-These characteristics are substantiated by the results obtained through the analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
-
-Therefore, to detect flood attacks, it is necessary to calculate the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
-
-\begin{equation}\label{eq:eq02}
-\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
+The general formula of the $r$-th moment can be expressed as:
+\begin{equation}\label{eq:eq05}
+	m_r = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^r. 
 \end{equation}
 
-The set of obtained vectors $\boldsymbol{y}_{m}^{(q)}$ composes the zero mean matrix $\boldsymbol{Y}^{(q)}$, then the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
-
-\begin{equation}\label{eq:eq03}
-\boldsymbol{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
+Therefore, the Skewness is calculated as
+\begin{equation}\label{eq:eq06}
+	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
 \end{equation}
 
-For the detection of the port scan attack, the main components are not dominated by the variables with large variance.
-Moreover, the portscan traffic presents a highly correlated network traffic.
-In order to exploit such structure, we compute the sample covariance $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have zero mean and unitary standard deviation as follows
-
-\begin{equation}\label{eq:eq04}
-\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
+and Kurtosis as
+\begin{equation}\label{eq:eq07}
+	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2} 
 \end{equation}
 
-The set of vectors $\boldsymbol{z}_{m}^{(q)}$ composes the matrix $\boldsymbol{Z}^{(q)}$, then the sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)}$ can be calculated via
+For extending Fast-MCD, We initially modify the C-step Algorithm \ref{alg:alg01} to calculate Skewness and Kurtosis after to minimize $det(\boldsymbol{\hat{S}})$, accordingly to Algorithm \ref{alg:alg05}.
 
-\begin{equation}\label{eq:eq05}
-\boldsymbol{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
+\begin{algorithm}[H]\label{alg:alg05}
+	\scriptsize
+	\SetAlgoLined
+	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
+	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
+	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
+		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
+			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
+		}
+		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
+		Sort $\boldsymbol{d}_1( i )$, which yields a index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
+		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
+		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
+		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
+	}
+	Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
+	\caption{C-Step with higher moments}
+\end{algorithm}
+
+The Algorithm \ref{alg:alg06} describes the Fast-MCD with higher moments when $n \leq 600$, through the modification of the original Fast-MCD for computing Skewness and Kurtosis from the robust estimate.
+
+\begin{algorithm}[H]\label{alg:alg06}
+	\scriptsize
+	\SetAlgoLined
+	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
+	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n \leq 600$\;
+	\For{500 times}{
+		Construct initial $h$-subset $\boldsymbol{h}_1$ (Algorithm \ref{alg:alg02})\;
+		Carry out two C-steps with higher moments (Algorithm \ref{alg:alg05}) resulting $\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$\;
+	}		
+	Put $\boldsymbol{h}_1$ as the 10 lowest $det(\boldsymbol{\hat{S}_2})$ computed above\;
+	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
+		Carry out C-steps with higher moments (Algorithm \ref{alg:alg05})\;
+	}
+	\caption{Fast-MCD with higher moments when $n \leq 600$}
+\end{algorithm}
+
+The Algorithm \ref{alg:alg07} presents the Fast-MCD with higher moments when $n > 600$, which modifies the original Fast-MCD for computing of Skewness and Kurtosis from the robust estimate data, through the adoption of the C-Step with higher moments.
+The Algorithm \ref{alg:alg07} is described below.
+
+\begin{algorithm}[H]\label{alg:alg07}
+	\scriptsize
+	\SetAlgoLined
+	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
+	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n > 600$\;
+	Construct up to 5 disjoint random subsets of size $n_{sub} = 300$\;
+	\For{each subset, repeat 100 times}{
+		Construct initial $\boldsymbol{h}_1$ of size $h_{sub} = [n_{sub}(h/n)]$\;
+		Carry out two C-steps with higher moments of $n_{sub}$ and $h_{sub}$\;
+		Keep the 10 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub}, \boldsymbol{\bar{s}}_{sub}, \boldsymbol{\bar{k}}_{sub})$\;
+	}		
+	Pool the subsets, yielding the merged set of $n_{merged} = 1500$\;
+	\For{50 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$}{
+		Carry out two C-steps with higher moments of $n_{merged}$ and $h_{merged} = [n_{merged}(h/n)]$\;
+		Keep the 10 best $(\boldsymbol{\bar{x}}_{merged}, \boldsymbol{\hat{S}}_{merged}, \boldsymbol{\bar{s}}_{merged}, \boldsymbol{\bar{k}}_{merged})$\;
+	}
+	\For{the full data set and $m_{full}$ best results}{
+		Take several C-steps, using $n$ and $h$\;
+		Keep the best final result $(\boldsymbol{\bar{x}}_{full}, \boldsymbol{\hat{S}}_{full}, \boldsymbol{\bar{s}}_{full}, \boldsymbol{\bar{k}}_{full})$\;
+	}
+	\caption{Fast-MCD with higher moments when $n > 600$}
+\end{algorithm}
+
+We also propose to use Mahalanobis Distance for detecting anomalies through an approach based on robust Skewness and Kurtosis, instead of the location and scatter that are used by classical approaches.
+
+We propose to modify the Equation \ref{eq:eq01} to implement a Skewness-based Mahalanobis Distance, as follows:
+
+\begin{equation}\label{eq:eq08}
+	\boldsymbol{d}(\boldsymbol{s}_{\boldsymbol{x}}, \bar{\boldsymbol{s}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{s}_{\boldsymbol{x}} - \bar{\boldsymbol{s}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{s}_{\boldsymbol{x}} - \bar{\boldsymbol{s}})^\prime},
 \end{equation}
 
-Once the $\boldsymbol{\hat{R}}_{yy}^{(q)}$ and $\boldsymbol{\hat{R}}_{zz}^{(q)}$ have been obtained for flood and port scan attack detection, respectively, and since the next steps are the same for both sample covariance matrices, we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{C}$.
-Therefore, the following step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:eq06}), in order to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).
+where $\boldsymbol{s}_{\boldsymbol{x}} = \boldsymbol{x} - \boldsymbol{\bar{s}}(\boldsymbol{x})$ and $\boldsymbol{\bar{s}}(\boldsymbol{x})$ is the Skewness of $\boldsymbol{x}$.
 
-\begin{equation}\label{eq:eq06}
-\boldsymbol{C}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
+We also propose to modify the Equation \ref{eq:eq04} to implement a Kurtosis-based Mahalanobis Distance, as follows:
+
+\begin{equation}\label{eq:eq09}
+	\boldsymbol{d}(\boldsymbol{k}_{\boldsymbol{x}}, \bar{\boldsymbol{k}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{k}_{\boldsymbol{x}} - \bar{\boldsymbol{k}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{k}_{\boldsymbol{x}} - \bar{\boldsymbol{k}})^\prime},
 \end{equation}
 
-\begin{equation}\label{eq:eq060}
-\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
-\end{equation}
+where $\boldsymbol{k}_{\boldsymbol{x}} = \boldsymbol{x} - \boldsymbol{\bar{k}}(\boldsymbol{x})$ and $\boldsymbol{\bar{k}}(\boldsymbol{x})$ is the Kurtosis of $\boldsymbol{x}$.
 
-where the operator diag$(\cdot)$ extracts the main diagonal of a matrix.
+Fast-MCD is commonly used as \textbf{supervised} or \textbf{unsupervised} algorithm for anomaly detection.
 
-The eigenvalues should be sorted in descending order, i.e., $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ...
-> \lambda_{m}^{(q)}$.
-Therefore, the largest eigenvalue of the $q$-th time frame evaluated for the attack detect is given by $\lambda_{1}^{(q)}$.
+\textbf{Supervised} approach:
 
-The concatenation of the eigenvalues vector $\boldsymbol{e}^{(q)}$ for $q = 1, \ldots, Q$ is represented by
+Compute Fast-MCD for training data with known normal and anomalous data.
+Select the best threshold $t$ for detecting known anomalies according to largest distances $\boldsymbol{d}(i)$.
+Save the best $t, \boldsymbol{\bar{x}}$ and $\boldsymbol{\hat{S}}$.
+Compute $d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}})$ for new observations.
+Classify observations with $\boldsymbol{d}(i)$ higher than $t$ as anomalous.
 
-\begin{equation}\label{eq:eq07}
-\boldsymbol{E} =
-\begin{bmatrix}
-  \lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
-  \lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
-  \lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
-  \vdots & \vdots & \ddots & \vdots  \\
-  \lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
-\end{bmatrix}.
-\end{equation}
+\textbf{Unsupervised} approach:
 
-Note that since $\lambda_1^{(q)} > \lambda_2^{(q)} > \lambda_3^{(q)} > \cdots > \lambda_{m-1}^{(q)} > \lambda_m^{(q)}$, then the first line of the matrix $\boldsymbol{E}$ contains the largest eigenvalues of each $q$-th time frame, which is the Greatest Eigenvalue Time
-Vector (GETV) \cite{tenorio2013greatest}, denoted as
+Given a known contamination $c$, which is the percentage of anomalies in a data set.
+Compute Fast-MCD and $d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}})$ for new observations.
+Classify the $c$ observations with largest distances as anomalous.
 
-\begin{equation}\label{eq:eq08}
-\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
-\end{equation}
+Finally, We propose the following \textbf{Semi Supervised} approach for Fast-MCD based anomaly detection:
+
+Compute Fast-MCD for training data containing only known normal observations.
+Save the fitted $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$.
+Carry out cross validation on known contaminated data in order to identify the contamination $c$ that best classify anomalies and normal data.
+Compute $d(\boldsymbol{x},\bar{\boldsymbol{x}}_{normal}, \boldsymbol{\hat{S}}_{normal})$ for new observations.
+The $c$ observations with largest $\boldsymbol{d}(i)$ are classified as anomalous.
 
@@
 
-Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$.
-However, the goal here is to detect the variations of the eigenvalues for different values of $q$.
-Therefore, instead of using a certain $q$, the proposed approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame, in order to identify variations and estimate the model order $\hat{d}$, which is the estimated number of time frames under attack.
-Therefore, $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
-Note that some MOS schemes may also require the number of minutes that compose a time frame, as $\hat{d} = \rm{MOS}(\boldsymbol{e}_{\rm max},\emph{Q})$.
-For more information about MOS, we refer to \ref{sec:mos}.
-
-In our previous work \cite{tenorio2013greatest}, the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE schemes are evaluated for synflood and port scan attack detection, showing that EDC and EFT are effective for detecting this kind of attacks.
-The present work extends that evaluation to also analyze the effectiveness of the listed MOS schemes for fraggle attack detection, as shown in Section \ref{sec:experimentalresults}.
-
-\subsection{Eigenvalue Analysis}
-\label{sec:prop_EigenvalueAnalysis}
-
-After applying the MOS schemes to the vector $\sim\boldsymbol{e}_{\rm max}$, we obtain the estimate of the $\#\boldsymbol{A}$.
-For instance, in the case of fraggle, synflood and portscan, if $\hat{d} = 1$, , then $\#\boldsymbol{A} = 1$, which means that during the during the $Q$ time frames one attack is present.
-However, if $\hat{d} = 0$, then $\#\boldsymbol{A} = 0$, and this means that none of these attacks are present.
-Note that $\hat{d}$ can be greater than 1, indicating the presence of more than one attack.
-
-In Subsection \ref{sec:prop_MOSSchemes}, we obtained only if $\hat{d} = 1$ or $\hat{d} = 0$.
-However, if $\hat{d} = 1$, the MOS schemes do not provide any information about the $q$-th time frame under attack.
-The identification of the $q$-th time frame under attack can be carried out through a eigenvalues analysis.
-
-The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed according to Algorithm \ref{alg:alg01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$.
-Algorithm \ref{alg:alg01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$, according to Line 2 of Algorithm \ref{alg:alg01}, and its correspondent index, according to Lines 4 and 5 of Algorithm \ref{alg:alg01}.
-Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$, according to Line 8 of Algorithm \ref{alg:alg01}, and a new iteration is performed until $\boldsymbol{e}_{\rm max} = \{\}$.
-
-\begin{algorithm}[h!]
-	\caption{Detection of Time Frames Under Attack}
-  	\label{alg:alg01}
-	\begin{algorithmic}[1]
-		\show\LOOP
-	    \ALOOP {$f = 1$ until $f == \hat{d}$} 
-	    	    \STATE $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
-    	    	    \ALOOP {$i = 1$ until $i == Q$} 
-				\IF {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
-				    \STATE $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
-				\ENDIF
-        		\ENDALOOP
-	    		\STATE $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
-    		\ENDALOOP
-	\end{algorithmic}
-\end{algorithm}
-
-After the estimation of the $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked.
-To deal with this problem, the adoption of a similarity analysis between legitimate traffic and the traffic of time frames estimated as under attack is evaluated, analysing the effectiveness of cosine similarity to highlight abnormalies inserted by network traffic attacks.
-
 \subsection{Eigen Similarity Analysis}
 \label{sec:prop_EigenSimilarityAnalysis}
-
-Cosine similarity calculates the cosine of the angle between two vectors, which represents the similarity of values between the selected vectors.
-Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\boldsymbol{V}^{(q)}$ against the the most significant eigenvectors of time frame detected as under attack, to analyze similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.
-
-This subsection describes the proposed eigen similarity analysis for detailed attack identification, in complement to the attack estimation carried out through MOS schemes and eigenvalue analysis.
-In Subsection \ref{sec:prop_TimeSimilarityAnalysis} we present the eigen similarity analysis for identification of time under attack.
-Next, in Subsection \ref{sec:prop_PortSimilarityAnalysis}, we show how to apply the eigen similarity analysis in order to identify network ports under attack.
 
@@
 
-For eigen similarity analyis, we evaluate the cosine similarity in order to identify lacks of similarity between legitimate and malicious traffic, as follows
-
-\begin{equation}\label{eq:eq11}
-s_n = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}}},
-\end{equation}
-where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, and $\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of traffic to be performed the flood and port scan attack identification.
-
-The most significant eigenvector $\boldsymbol{v}^{(q)}$, of a time frame $q$ without attack, can be derivated from (\ref{eq:eq06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the evaluated matrix.
-The same calculation shall be performed in order to obtain the target eigenvectors $\boldsymbol{v}_{(n)}$, calculated from a time frame without attack plus minutes of a time frame estimated as under attack, to evaluate the occurrence of network attacks.
-
-The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic whithout attack, from a time frame $q$ composed of $Q$ minutes of legitimate network traffic.
-For the detailed attack identification, each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$, as represented by
-
-\begin{equation}\label{eq:eq12}
-\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
-\end{equation}
-
-The resultant $\boldsymbol{X}_{(n)}$ is necessary to obtain $\boldsymbol{v}_{(n)}$, through (\ref{eq:eq06}), for calculating the similarity degree $s_n$, ranging from 0 to 1, for each $n$-th minute.
-The $s_n$ denotes the absolute similarity degree of the $n$-th minute in comparison to a well-known traffic without attack, detected through MOS schemes and eigenvalue analysis.
-
-The incremental approach for similarity analysis is based on the incremental appending of network traffic into $\boldsymbol{X}^{(q)}$, where the first evaluation is based on (\ref{eq:eq12}) and the subsequent evaluations is based on (\ref{eq:eq13}), incrementally appending each $n$-th minute until $n=N$.
-
-\begin{equation}\label{eq:eq13}
-\boldsymbol{X}_{n} = \{\boldsymbol{X}_{n} | \boldsymbol{x}^{(\hat{q})}_{(n)}\},
-\end{equation}
-
-Figure \ref{fig:fig8} ilustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.
-
-\begin{figure}[h!]
-     % \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental.eps}
-     \caption{Traffic selection for incremental approach.}
-     \label{fig:fig8}
-\end{figure}
-
-The eigen similarity analysis starts at $\boldsymbol{x}^{(3)}_{(41)}$ and is incrementally performed until $\boldsymbol{x}^{(\hat{3})}_{(60)}$, in order to calculate the $s_n$.
-We assume that $s_n < l$ means an attack identification, according the anomaly on similiarity of $s_n$ to a defined limiar $l$.
-Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}$ for eigen similarity analyis, the $s_n$ is calculated according to~(\ref{eq:eq11}).
-
-If $s_n = 1$, then the two eigenvectors are completely similar and no anomaly is detected. Smaller values of $s_n$ mean less similarity and can indicate an anomaly, according to a defined 
-threshold $l$, assuming that if $s_n < l$, then a network attack is identified during the $n$-th minute.
-Therefore, the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack is identified, according to
-
-\begin{equation}\label{eq:eq14}
-  \boldsymbol{\hat{n}_{(n)}}=\left\{
-  \begin{array}{@{}ll@{}}
-    1, & \text{if}\ s_n < l \\
-    0, & \text{otherwise}
-  \end{array}\right.
-\end{equation}
-where $\boldsymbol{\hat{n}}_{(n)}$ denotes a vector of $n$-th minutes detected as under attack.
-
-The eigen similarity analysis can also be applied in an individual fashion, where each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, as shown by Figure \ref{fig:fig9}.
-
-\begin{figure}[h!]
-     % \includegraphics[height=4cm, width=12.4cm]{results/figures/individualized.eps}
-     \caption{Traffic selection for individual approach.}
-     \label{fig:fig9}
-\end{figure}
-
-The incremental and the individual approaches can be combined to obtain the incremental individualized approach, where each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
-Subsequently, $\boldsymbol{X}_n$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:fig2}.
-
-\begin{figure}[h!]
-     % \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental_individualized.eps}
-     \caption{Traffic selection for incremental individualized approach.}
-     \label{fig:fig2}
-\end{figure}
-
-This approach of incremental similarity analysis followed by individual analysis after an attack detection allows to identify the attack period, highlighting the first and last time under attack.
-This identification is possible due to the variation of the most significant eigenvectors, which becomes more significant when compared a traffic under attack against a traffic with no attack, according to results which are discussed in Section \ref{sec:experimentalresults}.
-
 \subsubsection{Port Similarity Analysis}
 \label{sec:prop_PortSimilarityAnalysis}
-
-Given $\boldsymbol{\hat{n}}$, which is the set of $n$-th minutes under attack, it is still necessary to obtain more details about the identified network attack, such as the network ports that are attacked during each $n$-th minute identified as under attack.
-Hence, it is also applied the cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port during a $n$-th minute.
-
-For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack shall be used as reference for similarity analysis against the $\boldsymbol{v}_{(n)}$ identified as under attack, individually evaluating the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.
-
-Therefore, $\boldsymbol{v}^{(q)}$ should be calculated from the last $\boldsymbol{X}^{(q)}$ time frame without attack, and $\boldsymbol{v}_{(m,\hat{n})}$ should be calculated from the same traffic appened of all $n$-th minutes until the identified minute under attack, denoted as $\boldsymbol{X}_n$.
-
-For similarity analyis, each $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, denoted as $x_{(m,n)}$, shall be individually replaced by the traffic of the evaluated $m$-th port of the $\hat{n}$-th minute under attack, denoted as $x^{(\hat{q})}_{(m,\hat{n})}$, in order to identify significant variation on similarity caused by the traffic of the $m$-th port.
-
-This approach for detection of ports under attack via similarity analysis is given by
-
-\begin{equation}\label{eq:eq15}
-  \left\{
-  \begin{array}{@{}ll@{}}
-    x_{(m,n)} = x^{(\hat{q})}_{(m,\hat{n})} \\
-    \\
-    s_{m,\hat{n}} = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(m,\hat{n})}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(m,\hat{n})}}},
-  \end{array}\right.
-\end{equation}
-
-where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, which is used to calculate the $\boldsymbol{v}_{(m,\hat{n})}$ most significant eigenvectors that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.
-
-Once $\boldsymbol{v}^{(q)}$ and $\boldsymbol{v}_{(m,\hat{n})}$ are obtained, then the $s_{m,\hat{n}}$ similarity degree can be calculated in order to identify if the traffic replacement highlights the adition of anomalous traffic by the evaluated $m$-th port during the $\hat{n}$-th minute previously identified as under attack.
-
-This procedure should be repeated for each $m$-th target port of $\boldsymbol{\hat{n}}$, in order to individually identify the network ports under attack during each $\hat{q}$-th time frame.
 
@@
 
-The experiment time is 120 minutes, separated into six time frames, with each time frame corresponding to twenty minutes.
-Therefore, as the time of each sampling period is one minute, then $N = 20$.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, as well as a covariance $\boldsymbol{\hat{R}}_{yy}^{(q)} \in \mathbb{R}^{17 \times 17}$ (calculated via (\ref{eq:eq03})) and a sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)} \in \mathbb{R}^{17x17}$, assuming that $q = 1, 2, 3, 4, 5$ and $6$.
+\textbf{Experiment Plan:}
 
-The simulation started at 21:00h, the first time frame was from 21:00h until 21:20h ($q = 1$), the second was from 21:20h until 21:40h ($q = 2$), the third was from 21:40h to 22:00h ($q = 3$), the fourth was from 22:00h until 22:20h ($q = 4$), the fifth was from 22:20h until 22:40h ($q = 5$), and finally, the sixth was from 22:40h until 23.00h ($q = 6$).
-During the simulation, the victim made legitimate access, and the attacker performed the following attacks: at 21:54h ($q = 3$) was performed a port scan, at the interval ranging from 22:10h to 22:20h ($q = 4$) a synflood attack was simulated, and at the interval from 22:30h to 22:40h ($q = 5$) a fraggle attack was performed.
+	Given a contaminated data set, split it into normal (60\%), cross validation (20\% normal + 50\% anomalous) and test (20\% normal + 50\% anomalous).
+	Select the range of contamination that shall be evaluated and perform cross validation for each contamination. 
+	For each contamination $c$:
 
-\subsection{Largest Eigenvalues Analysis}
-\label{sec:largesteigenvaluesanalysis}
+		Estimate $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$ from normal data.
+		Predict anomalies for cross validation data using the $c$, $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$.
+		Select the best F-measure by cross validation and get the best $c$.
 
-For the evaluation of MOS Schemes accuracy for flood and port scan detection, the framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}.
-Through eigenvalue analysis of traffic with flood or port scan attacks, it is possible to visualize a significant difference between the largest eigenvalues and the remain eigenvalues, which can indicate a relationship between an outlier and time frames under attack.
+	Test the F-measure of anomaly detection for new observations through $d(\boldsymbol{x},\bar{\boldsymbol{x}}_0, \boldsymbol{\hat{S}}_0)$ and $c$.
 
-\begin{figure}[h!]
-	\centering
-     % \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_synflood.eps} 
-     \caption{Eigenvalues of the sample covariance matrix (synflood).}
-     \label{fig:fig10}
-\end{figure}
+\textbf{Experiment Plan:}
 
-Figure \ref{fig:fig10} depicts the eigenvalues calculated from sample covariance matrix of the network traffic used to evaluate the synflood attack identification.
-In Figure \ref{fig:fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 4$) stands out significantly from the other eigenvalues.
+    Evaluation Measure:
 
-Figure \ref{fig:fig11} illustrates the eigenvalues calculated from sample covariance matrix of the matrix used for fraggle attack detection.
-In Figure \ref{fig:fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 5$) stands out significantly from the other eigenvalues, in accordance with the result shown in Figure \ref{fig:fig10} for the synflood attack analysis.
+		In anomalous detection problems, where anomalies are rare events, if we classify all observations as normal and apply a accuracy evaluation, we would have high accuracy but poor true-positive detection.
+		Due to the importance given by F-measure to true-positive detection \cite{moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.
 
-\begin{figure}[h!]
-	\centering
-     % \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_fraggle.eps}
-     \caption{Eigenvalues of the sample covariance matrix (fraggle).}
-     \label{fig:fig11}
-\end{figure}
+\textbf{Experiment Plan:}
 
-Figure \ref{fig:fig12} depicts the eigenvalues calculated from covariance matrix of zero mean and unitary standard deviation variables, of the network traffic matrix evaluated for port scan detection.
+	The F-measure is the harmonic mean of precision and recall and is defined by 
+	\begin{equation}\label{eq:eq10}
+		F-measure = 2 * \frac{PR * RC}{PR + RC},				
+	\end{equation}
+	where $PR$ is the precision, defined by 
+	\begin{equation}\label{eq:eq11}
+		PR = \frac{True Positive}{True Positive + False Positive}
+	\end{equation}
+	and $RC$ is the recall, defined by 
+	\begin{equation}\label{eq:eq12}
+		RC = \frac{True Positive}{True Positive + False Negative}.
+	\end{equation}
 
-\begin{figure}[h!]
-	\centering
-     % \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_portscan.eps}
-     \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}
-     \label{fig:fig12}
-\end{figure}
+\textbf{Data splitting for training, cross validation and testing}
 
-As analyzed for the synflood and fraggle attacks, note that the largest eigenvalue, related to this attack ($q = 3$), stands out significantly from the others eigenvalues.
+	Gargia et al. \cite{garcia2014empirical} propose a division of CTU-13's scearios into train and test, for evaluation of network attack detection algorithms. 
+	However, we decide to evaluate our proposed approach for each scenario individually, considering that in the division proposed by Gargia et al.
+some bots are present only for training and not for testing.
+	Although our proposal does not rely on anomalous data for training, we desire to evaluate our proposal to detect all types of bots captured by CTU-13 data set.
 
-Table \ref{tab:tab3} presents the values of the largest eigenvalues of each time frame $q$-th for port scan, synflood and fraggle detection.
+We propose to compare our approach to clustering and statisctical based approaches (K-means and Gaussian Mixture Model (GMM), respectively).
+
+	As clustering-based method, we selected K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007k,moustafa2019holistic}.
+	It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}.
+
+\subsection{Results}
+\label{sec:results}
 
@@
   \centering
-  \footnotesize
-  \caption{Largest Eigenvalue related to attacks detection}
-  \label{tab:tab3}
-  \begin{tabular}{ c c c c c }
+  \scriptsize
+  \caption{Results grouped by algorithm and scenario}
+  \label{tab:tab02}
+  \begin{tabular}{ c|c c|c c|c c|c c|c c }
 	\toprule
-	\multirow{3}{*}{\textbf{Time Frame} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
-			\hhline{~----}
-		&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
-		&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
+	\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{S-MCD}} &\multicolumn{2}{c}{\textbf{K-MCD}} &\multicolumn{2}{c}{\textbf{MCD}} &\multicolumn{2}{c}{\textbf{GMM}} &\multicolumn{2}{c}{\textbf{K-Means}}\\
+			\hhline{~----------}
+			&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
 	\midrule
-	1 &1887545 &1887545 &1887545 &2,0734 \\
-	2 &2341327 &2341327 &2341327 &2,1451 \\
-	3 &3213867 &3213867 &3213867 &10,0718 \\
-	4 &133238294 &133238294 &731229 &2,1620 \\
-	5 &92384021611 &6367983 &92384021611 &2,4253 \\
-	6 &708335 &708335 &708335 &1,7948 \\
+		10 &\color{red} 0.71 & 0.03 & 0.66 & 0.02 & 0.34 & 0.52 & 0.34 & 0.46 & 0.20 & 0.25 \\ \hline
+		11 & 0.03 & 0.54 & 0.00 & 0.54 & 0.71 &\color{red} 0.73 & 0.35 & 0.48 & 0.13 & 0.15 \\ \hline
+		12 & 0.19 & 0.02 & 0.20 & 0.05 & 0.06 &\color{red} 0.31 & 0.01 & 0.01 & 0.09 & 0.11 \\ \hline
+		15 & 0.00 & 0.00 & 0.00 & 0.00 & 0.24 &\color{red} 0.32 & 0.14 & 0.22 & 0.09 & 0.11 \\ \hline
+		15-2 & 0.04 & 0.85 & 0.67 &\color{red} 0.86 & 0.36 & 0.28 & 0.32 & 0.14 & 0.18 & 0.26 \\ \hline
+		15-3 & 0.00 & 0.01 & 0.00 & 0.01 &\color{red} 0.55 & 0.48 & 0.00 & 0.01 & 0.22 & 0.29 \\ \hline
+		16 & 0.45 & 0.36 & 0.42 & 0.33 & 0.27 & 0.41 &\color{red} 0.51 & 0.49 & 0.07 & 0.16 \\ \hline
+		16-2 & 0.00 & 0.16 & 0.00 &\color{red} 0.20 & 0.04 & 0.04 & 0.02 & 0.08 & 0.04 & 0.03 \\ \hline
+		16-3 & 0.00 & 0.05 & 0.00 & 0.07 &\color{red} 0.15 & 0.10 & 0.01 & 0.01 & 0.03 & 0.04 \\ \hline
+		17 & 0.03 & 0.03 & 0.07 & 0.03 &\color{red} 0.59 & 0.53 & 0.16 & 0.47 & 0.35 & 0.39 \\ \hline
+		18 & 0.63 & 0.10 & 0.63 & 0.08 & 0.70 &\color{red} 0.72 & 0.21 & 0.17 & 0.13 & 0.20 \\ \hline
+		18-2 & 0.83 & 0.69 & 0.85 & 0.43 & 0.84 &\color{red} 0.91 & 0.50 & 0.55 & 0.62 & 0.70 \\ \hline
+		19 &\color{red} 0.71 & 0.00 & 0.68 & 0.00 & 0.11 & 0.42 & 0.17 & 0.20 & 0.20 & 0.17 \\ \hline
+		\rowcolor{Gray} \textbf{avg} & \textbf{0.28} & \textbf{0.22} & \textbf{0.32} & \textbf{0.20} & \textbf{0.38} &\color{red} \textbf{0.44} & \textbf{0.21} & \textbf{0.25} & \textbf{0.18} & \textbf{0.22} \\
     \bottomrule
@@
 
-In Table \ref{tab:tab3}, note the significant variation of the eigenvalues associated with attacks, in comparison to the others.
-At $q = 4$, where the synflood attack occurred, the maximum eigenvalue obtained is approximately 21 times larger than the second one.
-At $q = 5$, where the fraggle attack occurred, the maximum eigenvalue obtained is about 29,000 times larger than the second one.
-At $q = 3$, where the port scan attack occurred, the maximum eigenvalue obtained is approximately 4 times larger than the second one.
-In the last case, for port scan attack detection, although the largest eigenvalue presented no too large variance to the second one, if compared to synflood or fraggle attacks, it clearly deviates from the remaining largest eigenvalues.
-
-These results highlight that all $q$-th time frames where a network attack was simulated, present high significant variance between the largest eigenvalue and the remaining eigenvalues, obtained from sample covariance matrix, for flood detection, or from covariance matrix of zero mean and unitary standard deviation variables, for port scan detection.
-Therefore, we propose to apply the vector of the largest eigenvalues to MOS schemes in order to evaluate their accuracy for identification of time frames under attack, motivated by the fact that it is relevant to apply MOS schemes to automate the attack detection process, taking into account the characteristics of the evaluated eigenvalues.
-
-\subsection{MOS Schemes Evaluation}
-\label{sec:MOSSchemesEvaluation}
-
-In \cite{tenorio2013greatest}, we evaluate the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE MOS schemes \cite{da2009comparison,tenorio2013greatest} for synflood and port scan attack detection.
-In this work we extend that evaluation for fraggle attack detection, applying the same schemes to fraggle attack detection over the traffic presented in Section \ref{sec:datamodel}, as results shown in Table \ref{tab:tab4}.
-
 \begin{table}[h!]
@@
   \scriptsize
-  \caption{MOS schemes applied to port scan and flood detection}
-  \label{tab:tab4}
-  \begin{tabular}{ c c c c c c c c }
+  \caption{Results grouped by algorithm and scenario}
+  \label{tab:tab03}
+  \begin{tabular}{ c|c c|c c|c c|c c }
 	\toprule
-	\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
-			\hhline{~------~}
-		&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
+	\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{RPCA}} &\multicolumn{2}{c}{\textbf{S-RPCA}} &\multicolumn{2}{c}{\textbf{S-RPCA2}} &\multicolumn{2}{c}{\textbf{MCD}}\\ 
+			\hhline{~--------}
+			&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
 	\midrule
-	Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-	Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-	\midrule
-	Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-	Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-	\midrule
-	Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
-	Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
-	\midrule
-	Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
-	Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
+		10 & 0.93 & 0.93 &\color{red} 0.98 &\color{red} 0.98 &\color{red} 0.99 &\color{red} 1.00 & 0.34 & 0.52 \\ \hline
+		11 & 0.96 & 0.94 &\color{red} 0.99 &\color{red} 0.98 &\color{red} 0.99 &\color{red} 1.00 & 0.71 & 0.73 \\ \hline
+		12 & 0.32 & 0.19 & 0.00 & 0.00 &\color{red} 0.81 & 0.01 & 0.06 & 0.31 \\ \hline
+		15 & 0.75 & 0.73 & 0.86 &\color{red} 0.96 &\color{red} 1.00 &\color{red} 0.96 & 0.24 & 0.32 \\ \hline
+		15-2 & 0.40 & 0.57 &\color{red} 0.95 &\color{red} 0.94 & 0.25 & 0.93 & 0.36 & 0.28 \\ \hline
+		15-3 & 0.68 & 0.68 &\color{red} 0.96 & 0.70 &\color{red} 0.96 &\color{red} 0.96 & 0.55 & 0.48 \\ \hline
+		16 &\color{red} 0.47 &\color{red} 0.47 & 0.01 & 0.01 & 0.01 & 0.00 & 0.27 & 0.41 \\ \hline
+		16-2 & 0.07 & 0.10 &\color{red} 0.80 & 0.68 & 0.00 & 0.00 & 0.04 & 0.04 \\ \hline
+		16-3 & 0.29 & 0.29 & 0.01 & 0.46 &\color{red} 0.79 & 0.76 & 0.15 & 0.10 \\ \hline
+		17 &\color{red} 0.82 & 0.79 &\color{red} 0.81 & 0.78 &\color{red} 0.82 & 0.79 & 0.59 & 0.53 \\ \hline
+		18 & 0.87 & 0.89 & 0.97 &\color{red} 0.98 &\color{red} 0.98 &\color{red} 0.98 & 0.70 & 0.72 \\ \hline
+		18-2 &\color{red} 0.94 &\color{red} 0.94 &\color{red} 0.91 & 0.86 & 0.09 & 0.10 & 0.84 & 0.91 \\ \hline
+		19 & 0.27 & 0.31 &\color{red} 0.95 &\color{red} 0.95 & 0.02 & 0.02 & 0.11 & 0.42 \\ \hline
+		\rowcolor{Gray} \textbf{Avg} & \textbf{0.60} & \textbf{0.60} &\color{red} \textbf{0.71} &\color{red} \textbf{0.71} & \textbf{0.59} & \textbf{0.58} & \textbf{0.38} & \textbf{0.44} \\
     \bottomrule
@@
 
-Note that $\hat{d} = 1$, if there is attack, while $\hat{d} > 1$ indicates more than one attack.
-An example of this could be seen for attack detection via EFT for traffic containing synflood and fraggle attacks, showing $\hat{d} = 2$, which indicates the presence of two attacks, as expected by the $d$ real values of Table \ref{tab:tab4}.
+Results grouped by algorithm for full CTU-13.
 
-In Table \ref{tab:tab4}, two MOS schemes outperforms from the others, EDC and EFT.
-Efficient Detection Criterion (EDC) and Exponential Fitting Test (EFT) are the most effective schemes, correctly estimating the number of attacks in comparison to the expected values for effetctive attack detection, as defined by the column of real values in Table \ref{tab:tab4}.
-The AIC and MDL schemes are satisfactory only for port scan detection, however SURE and RADOI schemes did not show effective results for port scan or flood detection.
+\subsection{Discussion}
+\label{sec:discussion}
 
-Although EDC and EFT presented the same accuracy on the evaluation, the EDC scheme requires less processing time than EFT, which is an important criteria to select EDC as the MOS scheme for flood and port scan detection on the remain experiments.
-
-According to Table \ref{tab:tab4}, EDC and EFT estimated correctly the number of attacks of a time frame vector, indicating that ocurred $\hat{d}$ network attacks, but not providing additional details, what highlights the necessity of complementary approaches in order to estimate the time and ports under attack.
-Hence, we propose apply eigen analysis to estimate the $q$-th time frames under attack and eigen similarity analysis to estimate the minutes and ports under attack.
-
-\subsection{Eigenvalue Analysis}
-\label{sec:EigenvalueAnalysis}
-
-According to the results presented in Section \ref{sec:largesteigenvaluesanalysis}, the largest eigenvalue stands out significantly from the others eigenvalues of an evaluated $q$-th time frame.
-This behavior can also be observed in the largest eigenvalues analysis, according to results presented in Table \ref{tab:tab3}, where it is possible to observe that the $\hat{d}$ largest eigen values of the time frames under attacks stand out significantly from the others largest eigenvalues.
-
-Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respectives $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:alg01}.
-
-\subsection{Eigen Similarity Analysis}
-\label{sec:EigenSimilarityAnalysis}
-
-This paper proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$.
-Hence, the proposed framework is applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.
-
-\subsubsection{Time Analysis}
-\label{sec:TimeAnalysis}
-
-Three approaches were evaluated for eigen similarity analysis: incremental, individual and incremental individualized approaches.
-For the incremental individualized approach, each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
-Subsequently, $\boldsymbol{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated.
-For the incremental approach, each $n$-th minute must be incrementally appended into $\boldsymbol{X}^{(q)}$, for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual time similarity analyis.
-For the individual approach, each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, without incremental append, but doing individual appended into $\boldsymbol{X}^{(q)}$ for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual similarity analyis.
-
-Table \ref{tab:tab5} presents the results of the evaluation of three approaches for similarity analysis of eigenvectors for port scan detection.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Port Scan Detection}
-  \label{tab:tab5}
-  \begin{tabular}{ c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~~---~}
-			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
-	\midrule
-	3 &1 &0.9946 &0.9946 &0.9946 &no \\
-	3 &2 &0.9934 &0.9934 &0.9999 &no \\
-	3 &3 &0.9912 &0.9912 &0.9999 &no \\
-	3 &4 &0.9888 &0.9888 &0.9999 &no \\
-	3 &5 &0.9856 &0.9856 &0.9998 &no \\
-	3 &6 &0.9840 &0.9840 &0.9999 &no \\
-	3 &7 &0.9824 &0.9824 &1.0000 &no \\
-	3 &8 &0.9794 &0.9794 &0.9999 &no \\
-	3 &9 &0.9673 &0.9673 &0.9926 &no \\
-	3 &10 &0.9674 &0.9674 &0.9997 &no \\
-	3 &11 &0.9733 &0.9733 &0.9993 &no \\
-	3 &12 &0.9702 &0.9702 &0.9993 &no \\
-	3 &13 &0.9677 &0.9677 &0.9999 &no \\
-	3 &14 &0.9646 &0.9646 &0.9998 &no \\
-	3 &15 &0.0216 &0.0216 &0.0276 &yes \\
-	3 &16 &0.9621 &0.0209 &1.0000 &no \\
-	3 &17 &0.9611 &0.0199 &0.9998 &no \\
-	3 &18 &0.9612 &0.0191 &0.9999 &no \\
-	3 &19 &0.9613 &0.0186 &0.9998 &no \\
-	3 &20 &0.9638 &0.0190 &1.0000 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-Table \ref{tab:tab5} shows the evaluation of the time frame $q=3$, when the port scan attack was simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
-According to the presented results, it is possible to observe the high similarity between network traffic without attack, which was larger than 0.9610 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with the simulated port scan attack ($n=15$), which was lower than 0.0276 for all evaluated approaches.
-
-Comparing the approaches for similarity analysis, it is possible to observe that all approaches highlight the low similarity when evaluated the traffic under attack.
-However, the incremental approach figured out low similarity for times without attack, where $n=16, 17, 18, 19, 20$, what indicates that the incremental approach can produce false posive results.
-This behavior occurs because the incremental approaches appends all selected traffic into the reference traffic for comparison against the original reference traffic, what makes more evident the first lack of similarity but reduces the changing detection capability after an attack detection.
-
-Table \ref{tab:tab6} presents the results of the evaluation of the similarity analysis of eigenvectors for synflood detection.
-It shows the evaluation of the time frame $q=4$, when the synflood attack is simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
-According to the results, it is possible to observe the high similarity between network traffic without attack, which is larger than 0.9907 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with synflood attack (between $n=11$ and $n=20$), which is lower than 0.1244 for all evaluated approaches.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Synflood Detection}
-  \label{tab:tab6}
-  \begin{tabular}{ c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~~---~}
-			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
-	\midrule
-	4 &1 &1.0000 &1.0000 &1.0000 &no \\
-	4 &2 &0.9999 &0.9999 &1.0000 &no \\
-	4 &3 &0.9997 &0.9997 &0.9999 &no \\
-	4 &4 &0.9998 &0.9998 &1.0000 &no \\
-	4 &5 &0.9965 &0.9965 &0.9908 &no \\
-	4 &6 &0.9975 &0.9975 &1.0000 &no \\
-	4 &7 &0.9977 &0.9977 &1.0000 &no \\
-	4 &8 &0.9980 &0.9980 &1.0000 &no \\
-	4 &9 &0.9987 &0.9987 &0.9999 &no \\
-	4 &10 &0.9991 &0.9991 &1.0000 &no \\
-	4 &11 &0.0085 &0.0085 &0.0284 &yes \\
-	4 &12 &0.0162 &0.0120 &0.0343 &yes \\
-	4 &13 &0.0248 &0.0158 &0.0427 &yes \\
-	4 &14 &0.1243 &0.0185 &0.1041 &yes \\
-	4 &15 &0.0082 &0.0162 &0.0103 &yes \\
-	4 &16 &0.0404 &0.0070 &0.0580 &yes \\
-	4 &17 &0.0397 &0.0007 &0.0573 &yes \\
-	4 &18 &0.0408 &0.0042 &0.0584 &yes \\
-	4 &19 &0.0408 &0.0079 &0.0584 &yes \\
-	4 &20 &0.0477 &0.0092 &0.0757 &yes \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The incremental approach produces better results if compared with other evaluated approaches, with lower values and maximum of 0.0185 for times under attack, but this approach presents change detection limitation after the first outlier of similarity, as shown in Table \ref{tab:tab5} for port scan detection.
-
-Comparing the incremental individualized and the individual approaches for eigen similarity analysis, it is possible to observe that the incremental individualized approach obtain lowest values for almost all cases, except for the time $n=14$, where incremental individualized approach identified a larger similarity than the individual approach.
-The incremental individualized appends information about each evaluated traffic, therefore it incorporates traffic behaviors that can reduce the outlier capability detection, as ocurred for the time $n=14$.
-
-Table \ref{tab:tab7} presents the results of the eigen similarity analysis evaluation for fraggle detection.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Fraggle Detection}
-  \label{tab:tab7}
-  \begin{tabular}{ c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~~---~}
-			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
-	\midrule
-	5 &1 &1.0000 &1.0000 &1.0000 &no \\
-	5 &2 &0.9999 &0.9999 &1.0000 &no \\
-	5 &3 &1.0000 &1.0000 &1.0000 &no \\
-	5 &4 &0.9999 &0.9999 &1.0000 &no \\
-	5 &5 &0.9993 &0.9993 &0.9997 &no \\
-	5 &6 &0.9993 &0.9993 &0.9997 &no \\
-	5 &7 &0.9994 &0.9994 &1.0000 &no \\
-	5 &8 &0.9995 &0.9995 &1.0000 &no \\
-	5 &9 &0.9995 &0.9995 &1.0000 &no \\
-	5 &10 &0.9995 &0.9995 &1.0000 &no \\
-	5 &11 &0.0031 &0.0031 &0.0021 &yes \\
-	5 &12 &0.0019 &0.0025 &0.0009 &yes \\
-	5 &13 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &14 &0.0030 &0.0027 &0.0020 &yes \\
-	5 &15 &0.0030 &0.0028 &0.0020 &yes \\
-	5 &16 &0.0012 &0.0025 &0.0002 &yes \\
-	5 &17 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &18 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &19 &0.0030 &0.0027 &0.0020 &yes \\
-	5 &20 &0.0069 &0.0023 &0.0083 &yes \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-For fraggle attack detection, the lack of similarity between legitimate and malicious traffic was more evident than for the evaluation of synflood and port scan detecion.
-This behavior can be explained by the number of packets generated through the fraggle attack simulation, that was significative larger than the number of packets generated during the synflood simulation.
-Considering the three approaches, the largest value for times under attack was 0.0083, while the shortest value for times without attacks was 0.9993.
-
-Therefore, considering the evaluation for port scan, synflood and fraggle detection, the incremental approach can produce false positive results, while the individual and incremental individualized approaches produce quite similar results, even though the individual approach be more simple and require less memory and processing time.
-
-These results highlight the capability of change detection based on similarity between legitimate and malicious traffic from flood or port scan attacks, endorsing the effectiveness and safety for adoption of threshold for attack detection through eigen similarity analysis.
-
-\subsubsection{Port Analysis}
-\label{sec:PortAnalysis}
-
-Given $\hat{N}$, which is the set of estimated $n$-th minutes under attack, it is possible to apply cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port, during a $n$-th minute.
-Therefore, the incremental individualized and individual approaches of eigen similarity analysis were evaluated, for detection of ports under flood and port scan attacks, according to results presented in following tables.
-For this evaluation, the $v$ last most significant eigenvectors without attack was used as reference for similarity analysis against each target port $m$-th.
-
-Table \ref{tab:tab8} presents the results of the evaluation of eigen similarity analysis for detection of ports under port scan attack, showing only the time frame $q=3$ and minute $n=15$, due to the simulated port scan attack occurred only at this time, although the remain time frame has been completely evaluated and presented high similarity to the reference of traffic without network attack.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n=15)}
-  \label{tab:tab8}
-  \begin{tabular}{ c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~--~}
-			&\textbf{Incremental Individualized} &\textbf{Individual}\\
-	\midrule
-	80 &0.9999 &0.9999 &no \\
-	443 &0.9999 &0.9999 &no \\
-	53 &0.9999 &0.9999 &no \\
-	21 &0.9999 &0.9997 &yes \\
-	22 &0.0298 &0.9997 &yes \\
-	23 &0.0298 &0.9997 &yes \\
-	25 &0.0298 &0.9997 &yes \\
-	110 &0.0298 &0.9997 &yes \\
-	143 &0.0298 &0.9997 &yes \\
-	161 &0.0298 &0.9997 &yes \\
-	69 &0.0298 &0.9997 &yes \\
-	123 &0.0298 &0.9997 &yes \\
-	445 &0.0298 &0.9997 &yes \\
-	600 &0.9999 &0.9999 &no \\
-	19 &0.9999 &0.9999 &no \\
-	67 &0.9999 &0.9999 &no \\
-	68 &0.9999 &0.9999 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The incremental individualized approach presented more sensibility to anomaly detection than the individual approach, the former produced the identification of a low similarity of 0.0298 for almost all ports under attack, unless the port 21, although the simulation has attacked this port.
-The individual approach was not able to identify low similarity for ports under attack, resunting in values of 0.9997 for ports with anomalous traffic and 0.9999 for ports without network attack.
-
-For the evaluation of the proposed approaches for identification of ports under synflood and fraggle attack, all minutes of each time frame, in which one attack location was estimated, were analyzed.
-Even though, due to space limitations, only the results of the firt minute where a low similarity was idenfified will be shown such as where $n=11$.
-Nevertheless, the results obtained for the evaluation of traffic without attack presented high similarity to the reference traffic, with similarities close to 0.9999, and the evaluation of the other minutes under attack presented results quite similar to the resuts shown in the Tables \ref{tab:tab9} and \ref{tab:tab10}.
-
-Table \ref{tab:tab9} presents the results of the evaluation of eigen similarity analysis for detection of ports under synflood attack, showing only the time frame $q=4$ and minute $n=11$.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
-  \label{tab:tab9}
-  \begin{tabular}{ c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~--~}
-			&\textbf{Incremental Individualized} &\textbf{Individual}\\
-	\midrule
-	80 &1.0000 &1.0000 &no \\
-	443 &1.0000 &1.0000 &no \\
-	53 &1.0000 &1.0000 &no \\
-	21 &1.0000 &1.0000 &nos \\
-	22 &1.0000 &1.0000 &no \\
-	23 &1.0000 &1.0000 &no \\
-	25 &1.0000 &1.0000 &no \\
-	110 &1.0000 &1.0000 &no \\
-	143 &1.0000 &1.0000 &no \\
-	161 &1.0000 &1.0000 &no \\
-	69 &1.0000 &1.0000 &no \\
-	123 &1.0000 &1.0000 &no \\
-	445 &1.0000 &1.0000 &no \\
-	600 &0.0077 &0.0427 &yes \\
-	19 &1.0000 &1.0000 &no \\
-	67 &1.0000 &1.0000 &no \\
-	68 &1.0000 &1.0000 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-According to results presented in Table \ref{tab:tab9}, both approches identifies low similarity for the traffic of port 600, which is the target port of the simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.
-
-Table \ref{tab:tab10} presents the results of the evaluation of eigen similarity analysis for detection of ports under fraggle attack, showing only the time frame $q=5$ and minute $n=11$.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Fraggle Attack (q=5 and t=11)}
-  \label{tab:tab10}
-  \begin{tabular}{ c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~--~}
-			&\textbf{Incremental Individualized} &\textbf{Individual}\\
-	\midrule
-	80 &1.0000 &1.0000 &no \\
-	443 &1.0000 &1.0000 &no \\
-	53 &1.0000 &1.0000 &no \\
-	21 &1.0000 &1.0000 &no \\
-	22 &1.0000 &1.0000 &no \\
-	23 &1.0000 &1.0000 &no \\
-	25 &1.0000 &1.0000 &no \\
-	110 &1.0000 &1.0000 &no \\
-	143 &1.0000 &1.0000 &no \\
-	161 &1.0000 &1.0000 &no \\
-	69 &1.0000 &1.0000 &no \\
-	123 &1.0000 &1.0000 &no \\
-	445 &1.0000 &1.0000 &no \\
-	600 &1.0000 &1.0000 &no \\
-	19 &0.0031 &0.0004 &yes \\
-	67 &1.0000 &1.0000 &no \\
-	68 &1.0000 &1.0000 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The results for the avaluation of ports under fraggle attack, shown in Table \ref{tab:tab10}, were similar to the results obtained for synflood analysis, with the identification of low similarity for traffic of the port under attack.
-Nevertheless, for fraggle analysis, the individual approach identified the lowest similarity, that is 0.0004 while the incremental individualized approach obtained a similarity of 0.0031.
-
-The incremental individualized approach was able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches presented false positives or low sensibility to eigen similarity analyis for network attack detection.
-This approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.
-
-According to the shown significant lack of similarity between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
-
-\subsection{DARPA Scenario}
-\label{sec:DarpaEvaluation}
-
-This subsection presents a summarized view of results obtained from the application of the proposed framework, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 dataset.
-Since the proposed framework is detailed in Section \ref{sec:prop_getv} and in Subsections \ref{sec:AnalyzedScenario}, \ref{sec:largesteigenvaluesanalysis}, \ref{sec:MOSSchemesEvaluation}, \ref{sec:EigenvalueAnalysis} and \ref{sec:EigenSimilarityAnalysis}, here the focus is on the parameter selection, dataset evaluation and results for flood and probe attack identification.
-
-The DARPA dataset includes 7 weeks of sniffed traffic saved into raw packet data.
-The traffic and the labeled attacks are grouped by week and day, with information of the number and types of attacks per day, but also providing the start time for each labeled attack.
-For this evaluation, an evaluation per day was performed, considering the network traffic of 24 hours split into $Q$ time frames of 60 minutes ($N = 60$) and aggregate by minute and by port number.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ is obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.
-
-Since the proposed framework focus on flood and probe attack detection, only the attacks with behavior similar to flood or probe attack were evaluated.
-Initially all DoS and probe attacks were selected, but it was observed that the most cases of DoS focus on exploit system vulnerabilities instead of flooding attack, and most of probe attacks focus on ICMP intead of port scanning.
-Therefore, for evaluation of the proposed approach for flood and probe attack detection, it is necessary to select cases that implements flood or port scan behaviors.
-The following week-day-attack cases  were selected:
-
-\begin{enumerate}
-	\item week3-thursday-neptune;
-	\item week4-friday-portsweep;
-	\item week5-thursday-neptune;
-	\item week5-thursday-portsweep;
-	\item week5-friday-portsweep;
-	\item week6-wednesday-neptune;
-	\item week6-thursday-neptune;
-	\item week7-wednesday-portsweep.
-\end{enumerate}
-
-\begin{table*}[!t]
-	\caption{Results of the attack detection evaluation}
-	\label{tab:tab12}
-	\centering
-	\begin{tabular}{|c|c|c|c|c|}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Solution\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Attack Type\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Metric\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Result\end{tabular} \\ \hline
-	Proposed Work	&Flooding	&True Positive	&100.00 \%\\ \hline
-	Proposed Work	&Flooding	&False Positive	&60.00 \%\\ \hline
-	Proposed Work	&Flooding	&Misclassification	&50.00 \%\\ \hline
-	Proposed Work	&Probe	&True Positive	&76.92 \%\\ \hline
-	Proposed Work	&Probe	&False Positive	&18.52 \%\\ \hline
-	Proposed Work	&Probe	&Misclassification	&32.73 \%\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&True Positive	&82.00 \%\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&False Positive	&-\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&Misclassification	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Overall	&True Positive	&94.67 \%\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Overall	&False Positive	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Overall	&Misclassification	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&True Positive	&50.00 \%\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&False Positive	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&Misclassification	&-\\ \hline
-	\end{tabular}
-\end{table*}
-
-Table \ref{tab:tab12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}.
-The analysis based on sample covariation of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames.
-The results also show 50.00 \% of misclassification rate , which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
-The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time  from one source to one target, with no variation on IP source or target port.
-This observation corroborates with previous evaluations of the DARPA dataset that highlight issues regarding traffic redundancy.
-
-The analysis based on covariance of zero mean and unitary standard deviation variables was evaluated for port scan attacks, including probe attacks and DoS attacks that send few packets for several ports in order to exploit some vulnerability.
-The results show rates of 76.92 \% for TP detection and 18.52 \% for FP detection from 94 time frames.
-The observed misclassification rate for this scenario is 32.73 \%.
-It was observed that all FN cases are probe attacks with a time delay between scanning one port and start scanning the next port, what can be called as sparse probe attacks.
-Cases with a delay of one minute or more were not detected by the proposed approach.
-
-The performance of detection rate of flooding attacks is compared with the method proposed by Callegari \emph{et al} \cite{Zonglin2009}.
-This work is a statistical method, based on PCA, without training or learning methods, even though it relies on visual analysis for principal components selection.
-The best detection rate of \cite{Zonglin2009} was 82.00 \% for detection of synthetically added flood attacks, while this current proposal obtains 100.00 \% of detection rate for detection of flood attacks of DARPA dataset.
-
-Due to the lack of statistical techniques without training or learning methods for detection of probe attacks, this proposed approach is compared to the Lu and Ghorbani's \cite{Lu2009} proposal, which is a network anomaly detection model based on signal processing techniques that uses DARPA dataset for evaluation.
-The results of \cite{Lu2009} show the best detection rate of 94.67 \% in terms of general attack instance detection, but shows a case case with 50.00 \% of attack instance detection for the portsweep attack.
-The proposed approach presents 76.92 \% of detection rate measured specifically for probe attacks, without the requirement of learning or training methods, in contrast to Lu and Ghorbani's \cite{Lu2009} work.
+		there are no lots of researches using CTU-13, although it is been growing
+		poor results for GMM, K-Means and MGM are ok and in accordance to \cite{wang2017botnet}, which also evaluate \cite{gu2007bothunter}
 
@@
 
+Say that the data is lognormal \cite{benson2010network} and link it to skewness https://www.weibull.com/hotwire/issue47/relbasics47.htm
+Test distribution fit again e present results
+Conclude this presentation.
+Evaluate eigensim for CTU-13 and others algorithms for Danilo's data set.
+Compare to vanilla RPCA (Eduardo is working on it).
+Write a paper for JNCA (April).
+Write the thesis (May and June).
+
 \section*{References}

### 2019-04-21 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
adding texts of literature review

@@
 
-\title{Robust PCA and Higher Moments Distances for Botnet Detection}
+\title{Robust Subspace and Higher Moments Distances for Botnet Detection}
 
@@
 Some techniques can make the data discriminative for better anomaly detection: Data engineering (i.e.
-sampling \cite{??}, aggregation \cite{acarali2016survey}); Feature extraction (i.e.
+sampling \cite{??}, aggregation \cite{callegari2011novel, acarali2016survey}); Feature extraction (i.e.
 decomposition \cite{??}, robust estimates \cite{??}).
@@
 
+PCA is very sensitive to anomalies, thus outliers can corrupt the entire data or can be explored to highlight unespected changes and indicate attacks or frauds \cite{callegari2011novel,Lee2013,vieira2017model}.
+However, PCA is not precise to reveal flows, time or detailed information for each anomlies, and sometimes relies into visual analysis for pricipal component selection.
+
 We propose a network anomaly detection approach based on robust moment estimates and distance analysis.
@@
 Considering that it is hard to have labeled data in real problems \cite{osanaiye2016distributed}, such as for network attacks or frauds, also considering that known anomalies are constantly changing due to adversarial models or due to lack of regular patterns, we propose a Semi-Supervised approach based on robust moments for network anomaly detection.
+
+We believe that the distribution of anomalous and normal traffic are skewed but with differences that permit separate anomalous of normal traffic.
+(Assumption and hitotesis).
+We also believe that aggregated traffic and new features can make the data more discriminative for anomaly detection algorithms.
+We algo believe that normal traffic presents significant distance to anomalous traffic, which permit a distance based sepparation, that can be augmented by means of robust estimation of normal traffic, that can discard noise or irrelevant outliers and can highlight higher outliers when comparing to new observations
 
@@
 
+Our approach train from normal traffic and have significant computational cost during training. For prediction, the cost is basically the cost for Mahalanobis distance computation
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques.
+Callegari \emph{et al} \cite{callegari2011novel} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
+However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
+CAllegari separate normal of anomalies according to principal components (normal) and remaining (anomalies), based on PCA.
+It relies in the assumption that anomalies will cause a large change into anomalous components.
+This is similar to our previous work, but we rely on change into principal eigenvectors.
 
-Callegari \emph{et al} \cite{Zonglin2009} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
-However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
-
+Use to justify the sensitivity of pca to outliers, only.
 Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
@@
 Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations.
-We propose to adopt CTU-13...
+Considering that it is important to create intrusion detection datasets in modern-day computing to address the issues of DARPA/KDD.
+The next section discusses ne such contemporary dataset for network traffic analysis \cite{ahmed2016survey}, we propose to adopt CTU-13...
 
@@
 
-\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions)
+\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions).
+-Argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
+Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning.
+Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms.
+\textbf{To explain feature selection and aggregation}.
 
@@
 
-Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{moustafa2019holistic}.
+Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}.
+We rely on this for our implementation
 
@@
 
-Based on related works \cite{lakhina2005mining,chandrashekar2014survey,acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
+Based on related works \cite{lakhina2005mining, callegari2011novel, chandrashekar2014survey, acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
 The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15, 0.25, 1 and 2 seconds, for each scenario of the CTU-13 data set.

### 2019-04-23 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
robpca

@@
 
-\textbf{MD has been used for distance based anomaly detection with robust estimates} in many areas, and is commonly combined to Fast Minimum Covariance Determinant (Fast-MCD) algorithm \cite{rousseeuw1999fast} for robust estimates.
+\textbf{MD has been used for distance based anomaly detection with robust estimates} in many areas, and is commonly combined to Fast Minimum Covariance Determinant (Fast-MCD) algorithm \cite{rousseeuw1999fastmcd} for robust estimates.
 
@@
 Skewness and the kurtosis tests are very sensitive outlier detectors.
-The properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2008outlier}.
+The properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
 
@@
 
-\cite{hubert2008outlier} (Outlier detection for skewed data)
+ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA to reduce the data dimensions and plot the orthogonal distances versus the score distances, to flag outliers using an outlier map.
+However, ROBPCA flags many points as outlying when the original data is skewed.
+To overcome that issue, \cite{hubert2009robustskewed} propose ROBPCA-AO, which relies on an adjusted outlyingness based on robust skewness for flagging outliers....
 
@@
 C.
-Mahalanobis in 1936 \cite{mahalanobis1936generalized}.
+Mahalanobis in 1936 \cite{mahalanobis1936md}.
 It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\hat{S}}$ of $\boldsymbol{X}$.
@@
 Therefore, we need reliable and robust estimators that can resist outliers contaminated data.
-Minimum Covariance Determinant (MCD) \cite{rousseeuw1984least} is a robust estimator commonly used for computing Mahalanobis Distance in order to detect anomalies.
+Minimum Covariance Determinant (MCD) \cite{rousseeuw1984mcd} is a robust estimator commonly used for computing Mahalanobis Distance in order to detect anomalies.
 
@@
 
-Affine equivariant means that when the data are translated or subjected to a linear transformation, the resulting location and scatter will transform accordingly \cite{rousseeuw1984least,rousseeuw1999fast}.
+Affine equivariant means that when the data are translated or subjected to a linear transformation, the resulting location and scatter will transform accordingly \cite{rousseeuw1984mcd,rousseeuw1999fastmcd}.
 
@@
 
-Its main use has started since the construction of the computationally efficient Fast-MCD algorithm \cite{rousseeuw1999fast}.
+Its main use has started since the construction of the computationally efficient Fast-MCD algorithm \cite{rousseeuw1999fastmcd}.
 
@@
 
-The proof is given by Rousseeuw and Driessen \cite{rousseeuw1999fast}.
+The proof is given by Rousseeuw and Driessen \cite{rousseeuw1999fastmcd}.
 
@@
 
+say that more studies about skewness and subspace learning can be done in order to explore it inside algorithms and non gaussian data
+
 \section*{References}

### 2019-04-24 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
conclude robpca and initiate justificative for imbalance and skewness

@@
 
-Some techniques can make the data discriminative for better anomaly detection: Data engineering (i.e.
-sampling \cite{??}, aggregation \cite{callegari2011novel, acarali2016survey}); Feature extraction (i.e.
-decomposition \cite{??}, robust estimates \cite{??}).
-
 Robust statistics produce methods that are not unduly affected by outliers.
@@
 
-Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority}.
+Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning}.
+Learning algorithms for imbalanced data has been a challenging research topic.
+The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
+He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning.
+Some techniques can make the data discriminative for better discriminative analytics: Data engineering (i.e.
+sampling \cite{he2008learning}, aggregation \cite{callegari2011novel, acarali2016survey}); Feature extraction (i.e.
+decomposition \cite{vaswani2018robust}, robust estimates \cite{zhou2017anomaly}).
+To face a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
 
-Learning algorithms for imbalanced data has been a relevant research topic.
-
-The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
+Some approaches assume a gaussian distribute data, hoever it is not true in real world for many fields....Data skewness...
 
@@
 However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
-CAllegari separate normal of anomalies according to principal components (normal) and remaining (anomalies), based on PCA.
+Callegari separate normal of anomalies according to principal components (normal) and remaining (anomalies), based on PCA.
 It relies in the assumption that anomalies will cause a large change into anomalous components.
@@
 
-ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA to reduce the data dimensions and plot the orthogonal distances versus the score distances, to flag outliers using an outlier map.
+ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
 However, ROBPCA flags many points as outlying when the original data is skewed.
-To overcome that issue, \cite{hubert2009robustskewed} propose ROBPCA-AO, which relies on an adjusted outlyingness based on robust skewness for flagging outliers....
+To overcome that issue, \cite{hubert2009robustskewed} propose ROBPCA-AO, which improve ROBPCA by means of an adjusted outlyingness based on robust skewness.
+In contrast to our proposal, ROBPCA and ROBPCA-AO rely on a previously defined cutoff value for classify anomalous observations, and does not require training.
 
@@
 \textbf{The C-step can be algorithmically described as follows:}
-\begin{algorithm}[H]\label{alg:alg01}
+\begin{algorithm}
+	\label{alg:alg01}
 	\scriptsize
@@
 		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
-		Sort $\boldsymbol{d}_1( i )$, which yields a index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
+		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
 		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
@@
 \textbf{Initial subset can be algorithmically described as:}
-\begin{algorithm}[H]\label{alg:alg02}
+\begin{algorithm}
+	\label{alg:alg02}
 	\scriptsize
@@
 
-\begin{algorithm}[H]\label{alg:alg03}
+\begin{algorithm}
+	\label{alg:alg03}
 	\scriptsize
@@
 
-\begin{algorithm}[H]\label{alg:alg04}
+\begin{algorithm}
+	\label{alg:alg04}
 	\scriptsize
@@
 
-He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning.
-
-Data aggregation is a widely adoted approach for feature engineering \cite{chandrashekar2014survey,acarali2016survey}.
-
 Lakhina et al.
@@
 
-\begin{algorithm}[H]\label{alg:alg05}
+\begin{algorithm}
+	\label{alg:alg05}
 	\scriptsize
@@
 
-\begin{algorithm}[H]\label{alg:alg06}
+\begin{algorithm}
+	\label{alg:alg06}
 	\scriptsize
@@
 
-\begin{algorithm}[H]\label{alg:alg07}
+\begin{algorithm}
+	\label{alg:alg07}
 	\scriptsize
@@
 
-		In anomalous detection problems, where anomalies are rare events, if we classify all observations as normal and apply a accuracy evaluation, we would have high accuracy but poor true-positive detection.
+		In anomalous detection problems, where anomalies are rare events, if we classify all observations as normal and apply an accuracy evaluation, we would have high accuracy but poor true-positive detection.
 		Due to the importance given by F-measure to true-positive detection \cite{moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.

### 2019-04-25 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
moved to botnets references and stop at bothunter

@@
 
-Botnets...
+Botnets have emerging as a key threat for cyber security, considering that Botnets are the key platform for many Internet attacks (i.e.
+spam, distributed denial-of-service (DDoS), identity theft, and phishing).
+Most of the current botnet detection approaches work only on specific botnet command and control (.
+Lots of efforts are directed to avoid attacks from known attackers, however attacks from one source is not real nowadays, since the attacks are usualy from distributed botnets that change their approaches frequently, in order to avoid behavioral or pattern based detecions.
+This changing behavior is known as adversarial model \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,da2018online,Wang2018ddosbotnets}.
 
@@
 Some techniques can make the data discriminative for better discriminative analytics: Data engineering (i.e.
-sampling \cite{he2008learning}, aggregation \cite{callegari2011novel, acarali2016survey}); Feature extraction (i.e.
+sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, acarali2016survey}); Feature extraction (i.e.
 decomposition \cite{vaswani2018robust}, robust estimates \cite{zhou2017anomaly}).
@@
 
+Botnets are becoming one of the most serious threats to Internet security.
+A botnet is a network of compromised machines under the influence of malware (bot) code.
+The botnet is commandeered by a “botmaster” and utilized as “resource” or “platform” for attacks such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration.
+In order for a botmaster to command a botnet, there needs to be a command and control (C&C) channel through which bots receive commands and coordinate attacks and fraudulent activities.
+The C&C channel is the means by which individual bots form a botnet.
+Centralized C&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time.
+In this architecture, each bot logs into an IRC channel, and seeks commands from the botmaster.
+Even today, many botnets are still designed this way.
+Quite a few botnets, though, have begun to use other protocols such as HTTP [8, 14, 24, 39], probably because HTTP-based C&C communications are more stealthy given that Web traffic is generally allowed in most networks \cite{gu2008botminer}.
+
 \cite{gu2007bothunter} (Bothunter: Detecting malware infection through ids-driven dialog correlation)
 
-\cite{gu2008botminer} (Botminer: Clustering analysis of network traffic for protocol-and structure-independent botnet detection)
+\cite{gu2008botminer} (Botminer: Clustering analysis of network traffic for protocol-and structure-independent botnet detection).
+The objective of BotMiner is to detect groups of compromised machines within a monitored network that are part of a botnet.
+This framework monitors both who is talking to whom that may suggest C&C communication activities and who is doing what that may suggest malicious activities, and finds a coordinated group pattern in both kinds of activities.
+The detection framework clusters similar communication activities in the C-plane (C&C communication traffic), clusters similar malicious activities in the A-plane (activity traffic), and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.
 
@@
 
+\cite{gu2007bothunter} (Bothunter: Detecting malware infection through ids-driven dialog correlation).
+The BotHunter method was proposed by Gu et al.
+(2007) to detect the infection and coordination dialog of botnets by matching a state-based infection sequence model.
+It consists of a correlation engine that aims at detecting specific stages of the malware infection process, such as inbound scanning, exploit usage, egg downloading, outbound bot coordination dialog and outbound attack propagation.
+
 \cite{garcia2014empirical} (An empirical comparison of botnet detection methods)
+
+\cite{Wang2018ddosbotnets} present an in-depth analysis based on 50,704 different Internet DDoS attacks launched by 674 botnets from 23 different botnet families.
+Geospatial distribution of the attacking sources follows certain patterns, multiple attacks to the same target also exhibit strong patterns of inter-attack time interval, there is a trend for different botnets to launch DDoS attacks targeting the same victim, simultaneously or in turn.
 
@@
 
-Based on related works \cite{lakhina2005mining, callegari2011novel, chandrashekar2014survey, acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
+Based on related works \cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
 The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15, 0.25, 1 and 2 seconds, for each scenario of the CTU-13 data set.

### 2019-04-27 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
writing

@@
 
-\cite{gu2007bothunter} (Bothunter: Detecting malware infection through ids-driven dialog correlation)
+The BotHunter method was proposed by Gu \emph{et al} \cite{gu2007bothunter} to detect the infection and coordination of botnets by matching sequence model, through a correlation approach for detecting stages of the infection process.
+Gu \emph{et al} \cite{gu2008botminer} presented the BotMiner, which aims to detect groups of compromised machines that are part of a botnet.
+BotMiner monitors communications that may suggest C&C or malicious activities, and finds a coordinated group pattern by means of clusters of similar communication activities, clusters of similar malicious activities, and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.
 
-\cite{gu2008botminer} (Botminer: Clustering analysis of network traffic for protocol-and structure-independent botnet detection).
-The objective of BotMiner is to detect groups of compromised machines within a monitored network that are part of a botnet.
-This framework monitors both who is talking to whom that may suggest C&C communication activities and who is doing what that may suggest malicious activities, and finds a coordinated group pattern in both kinds of activities.
-The detection framework clusters similar communication activities in the C-plane (C&C communication traffic), clusters similar malicious activities in the A-plane (activity traffic), and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.
+\cite{garcia2014empirical} (An empirical comparison of botnet detection methods).
+The goal of this paper is to compare three botnet detection methods using a simple and reproducible methodology, a good dataset and a new error metric.
+This paper evaluates some data sets for network anomaly detection and classify the availability of labels and the presesence of background, normal and botnet traffic.
+The paper also survey some approaches for botnet detection.
+The paper proposes two methods (BClus and CAMNEP) and compare results to BotHunter.
+Considering the lack of available labeled datasets, the paper proposes the CTU-13 dataset, which is composed by normal and background labeled data, with the balance of the dataset like in a real network.
+The training and cross-validation datasets should be approximately 80\% of the dataset.The testing dataset should be approximately 20\% of the dataset.
+None of the botnet families used in the training and crossvalidation dataset should be used in the testing dataset.
+This ensures that the methods can generalize and detect new behaviors.
+\cite{garcia2014empirical} uses a time window separation and an aggregation method.
+Thei compare the output of three different botnet detection methods by executing them over a new, real, labeled and large botnet dataset.
+
+Khattak \emph{et al} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, based on a decision module as a Complex Event Processing (CEP) engine and based on a correlation framework that continuously receives events from the sensor module and correlates them according to rule(s).
 
@@
 
-\cite{gu2007bothunter} (Bothunter: Detecting malware infection through ids-driven dialog correlation).
-The BotHunter method was proposed by Gu et al.
-(2007) to detect the infection and coordination dialog of botnets by matching a state-based infection sequence model.
-It consists of a correlation engine that aims at detecting specific stages of the malware infection process, such as inbound scanning, exploit usage, egg downloading, outbound bot coordination dialog and outbound attack propagation.
-
-\cite{garcia2014empirical} (An empirical comparison of botnet detection methods)
-
 \cite{Wang2018ddosbotnets} present an in-depth analysis based on 50,704 different Internet DDoS attacks launched by 674 botnets from 23 different botnet families.
 Geospatial distribution of the attacking sources follows certain patterns, multiple attacks to the same target also exhibit strong patterns of inter-attack time interval, there is a trend for different botnets to launch DDoS attacks targeting the same victim, simultaneously or in turn.
-
-\cite{khattak2015botflex} (BotFlex: A community-driven tool for botnet detection)
 
@@
 
-The data set contains the following features:
+The data set contains the following features:  Start Time, End Time, Duration, Source IP address, Source Port, Direction, Destination IP address, Destination Port, State, SToS, Total Packets and Total Bytes.
 
 Start Time, Duration, Protocol, Source Address, Source Port, Direction, Destination Address, Destination Port, State, Type of service from source to destination, Type of service from destination to source, Total of Packets, Total of bytes, Total bytes from source to destination, Label as normal or anomalous
+
+Our labeling strategy assigns three different labels: back-ground, botnet and normal.
+The priority to assign the labels is the following: 1.
+Assign the Background label to the whole traffic.
+2.
+Assign the Normal label to the traffic that matches certain filters.
+3.
+Assign the Botnet label to all the traffic that comes from or to any of the known infected IP addresses.
 

### 2019-04-27 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
improvements on literature review

@@
 The experiments show that the proposed method achieves satisfactory levels of precisio, recall and F1 score for network attack detection.
+We apply our method to real-world botnet traffic and compare its performance with other methods.
 \end{abstract}
@@
 Lots of efforts are directed to avoid attacks from known attackers, however attacks from one source is not real nowadays, since the attacks are usualy from distributed botnets that change their approaches frequently, in order to avoid behavioral or pattern based detecions.
-This changing behavior is known as adversarial model \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,da2018online,Wang2018ddosbotnets}.
+This changing behavior is known as adversarial model \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,da2018online,Wang2018ddosbotnetssurvey}.
 
@@
 Some techniques can make the data discriminative for better discriminative analytics: Data engineering (i.e.
-sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, acarali2016survey}); Feature extraction (i.e.
+sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, garcia2014empirical, acarali2016survey}); Feature extraction (i.e.
 decomposition \cite{vaswani2018robust}, robust estimates \cite{zhou2017anomaly}).
-To face a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
+To face a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Works}
+% network anomaly detectiom, botnet detection, botnet detection approaches, approaches based on subspace learning, sampling and aggregation, about the methodos to compare and metrics
+\section{Literature Review}
 \label{sec:relatedworks}
+
+Talk about network anomaly detection...
+
+Bhuyan \emph{et al} \cite{acarali2016survey} provide an overview of facets of network anomaly detection, present attacks normally encountered by network intrusion detection systems, and categorize existing network anomaly detection methods and systems based on the underlying techniques.
+Ahmed Bhuyan \emph{et al} \cite{acarali2016survey} present an analysis of four major categories of anomaly detection techniques which include classification, statistical, information theory and clustering.
+Recently, Moustafa \emph{et al} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), explains cyber-attacks and new solutions for anomaly detection, and provides a benchmark datasets for training and validating approaches for detection.
+
+Botnets are becoming one of the most serious threats to Internet security.
+A botnet is a network of compromised machines under the influence of malware (bot) code.
+The botnet is commandeered by a “botmaster” and utilized as “resource” or “platform” for attacks such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration.
+In order for a botmaster to command a botnet, there needs to be a command and control (C&C) channel through which bots receive commands and coordinate attacks and fraudulent activities.
+The C&C channel is the means by which individual bots form a botnet.
+Centralized C&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time.
+In this architecture, each bot logs into an IRC channel, and seeks commands from the botmaster.
+Even today, many botnets are still designed this way.
+Quite a few botnets, though, have begun to use other protocols such as HTTP [8, 14, 24, 39], probably because HTTP-based C&C communications are more stealthy given that Web traffic is generally allowed in most networks \cite{gu2008botminer}.
+
+Acarali \emph{et al} \cite{acarali2016survey} surveys recent network-based detection approaches for HTTP-based botnets, surveys traffic-based features used to detect bot traffic and presents an abstraction of the main types of features in relation to protocols and OSI layers.
+Wang \emph{et al} \cite{Wang2018ddosbotnetssurvey} present an analysis based on 50,704 different Internet DDoS attacks originated of 674 botnets from 23 different botnet families with a total of 9,026 victim belonging to 1,074 organizations in 186 countries.
+Their analysis reveals that geolocation of the attacking sources follows patterns and enables source prediction, and highlight that multiple attacks to the same target also exhibit strong patterns of inter-attack time interval, also presents that there is a trend for different botnets to launch DDoS attacks targeting the same victim, simultaneously or in turn.
+
+The BotHunter method was proposed by Gu \emph{et al} \cite{gu2007bothunter} to detect the infection and coordination of botnets by matching sequence model, through a correlation approach for detecting stages of the infection process.
+Gu \emph{et al} \cite{gu2008botminer} presented the BotMiner, which aims to detect groups of compromised machines that are part of a botnet.
+BotMiner monitors communications that may suggest C&C or malicious activities, and finds a coordinated group pattern by means of clusters of similar communication activities, clusters of similar malicious activities, and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.
+
+Khattak \emph{et al} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, based on a Complex Event Processing (CEP) engine and on a correlation framework that continuously receives events and correlates them according to rule(s).
+BotFlex's results are compared to BotHunter, but the evaluation relies in a own and not public data set.
+
+\cite{garcia2014empirical} (An empirical comparison of botnet detection methods).
+The goal of this paper is to compare three botnet detection methods using a simple and reproducible methodology, a good dataset and a new error metric.
+This paper evaluates some data sets for network anomaly detection and classify the availability of labels and the presesence of background, normal and botnet traffic.
+The paper also survey some approaches for botnet detection.
+The paper proposes two methods (BClus and CAMNEP) and compare results to BotHunter.
+Considering the lack of available labeled datasets, the paper proposes the CTU-13 dataset, which is composed by normal and background labeled data, with the balance of the dataset like in a real network.
+The training and cross-validation datasets should be approximately 80\% of the dataset.The testing dataset should be approximately 20\% of the dataset.
+None of the botnet families used in the training and crossvalidation dataset should be used in the testing dataset.
+This ensures that the methods can generalize and detect new behaviors.
+\cite{garcia2014empirical} uses a time window separation and an aggregation method.
+Thei compare the output of three different botnet detection methods by executing them over a new, real, labeled and large botnet dataset.
+
+Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks.
+Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
+NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications.
+Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations.
+Considering that it is important to create intrusion detection datasets in modern-day computing to address the issues of DARPA/KDD.
+The next section discusses ne such contemporary dataset for network traffic analysis \cite{ahmed2016survey}, we propose to adopt CTU-13...
+
+Wang and Paschalidis \cite{wang2017botnet} propose an botnet detection approach based on anomaly and community detection, aiming for detecting botnets and identifying bots before the botnet becomes active.
+The first stage detects anomalies by leveraging large deviations of an empirical distribution.
+The second stage detects the bots using ideas from social network community detection in a graph that captures correlations of interactions among nodes over time.
+This work is compared with the BotHunter \cite{gu2007bothunter} on the CTU-13 botnet dataset \cite{garcia2014empirical}.
+
+\cite{da2018online} (Online detection of Botnets on Network Flows using Stream Mining) Ver se podemos calcular o F1 e comparar os resutados, mas não seria justo dado que a abordagem deles nao implementa o mesmo sampling e é supervisionado
 
@@
 
-Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks.
-Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
-NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications.
-Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations.
-Considering that it is important to create intrusion detection datasets in modern-day computing to address the issues of DARPA/KDD.
-The next section discusses ne such contemporary dataset for network traffic analysis \cite{ahmed2016survey}, we propose to adopt CTU-13...
-
-Botnets are becoming one of the most serious threats to Internet security.
-A botnet is a network of compromised machines under the influence of malware (bot) code.
-The botnet is commandeered by a “botmaster” and utilized as “resource” or “platform” for attacks such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration.
-In order for a botmaster to command a botnet, there needs to be a command and control (C&C) channel through which bots receive commands and coordinate attacks and fraudulent activities.
-The C&C channel is the means by which individual bots form a botnet.
-Centralized C&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time.
-In this architecture, each bot logs into an IRC channel, and seeks commands from the botmaster.
-Even today, many botnets are still designed this way.
-Quite a few botnets, though, have begun to use other protocols such as HTTP [8, 14, 24, 39], probably because HTTP-based C&C communications are more stealthy given that Web traffic is generally allowed in most networks \cite{gu2008botminer}.
-
-The BotHunter method was proposed by Gu \emph{et al} \cite{gu2007bothunter} to detect the infection and coordination of botnets by matching sequence model, through a correlation approach for detecting stages of the infection process.
-Gu \emph{et al} \cite{gu2008botminer} presented the BotMiner, which aims to detect groups of compromised machines that are part of a botnet.
-BotMiner monitors communications that may suggest C&C or malicious activities, and finds a coordinated group pattern by means of clusters of similar communication activities, clusters of similar malicious activities, and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.
-
-\cite{garcia2014empirical} (An empirical comparison of botnet detection methods).
-The goal of this paper is to compare three botnet detection methods using a simple and reproducible methodology, a good dataset and a new error metric.
-This paper evaluates some data sets for network anomaly detection and classify the availability of labels and the presesence of background, normal and botnet traffic.
-The paper also survey some approaches for botnet detection.
-The paper proposes two methods (BClus and CAMNEP) and compare results to BotHunter.
-Considering the lack of available labeled datasets, the paper proposes the CTU-13 dataset, which is composed by normal and background labeled data, with the balance of the dataset like in a real network.
-The training and cross-validation datasets should be approximately 80\% of the dataset.The testing dataset should be approximately 20\% of the dataset.
-None of the botnet families used in the training and crossvalidation dataset should be used in the testing dataset.
-This ensures that the methods can generalize and detect new behaviors.
-\cite{garcia2014empirical} uses a time window separation and an aggregation method.
-Thei compare the output of three different botnet detection methods by executing them over a new, real, labeled and large botnet dataset.
-
-Khattak \emph{et al} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, based on a decision module as a Complex Event Processing (CEP) engine and based on a correlation framework that continuously receives events from the sensor module and correlates them according to rule(s).
-
 ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
@@
 
-\cite{Wang2018ddosbotnets} present an in-depth analysis based on 50,704 different Internet DDoS attacks launched by 674 botnets from 23 different botnet families.
-Geospatial distribution of the attacking sources follows certain patterns, multiple attacks to the same target also exhibit strong patterns of inter-attack time interval, there is a trend for different botnets to launch DDoS attacks targeting the same victim, simultaneously or in turn.
-
-\cite{wang2017botnet} (Botnet detection based on anomaly and community detection)
-
-\cite{da2018online} (Online detection of Botnets on Network Flows using Stream Mining)
-
-\cite{moustafa2019holistic} (A holistic review of Network Anomaly Detection Systems: A comprehensive survey).
-Network anomaly detection approaches are classified by \cite{moustafa2019holistic} in six categories: classification, clustering, deep learning, knowledge, combination and statistical-based.
-
 \cite{Phua2004minority} (Minority Report in Fraud Detection: Classification of Skewed Data)
-
-\cite{acarali2016survey} (Survey of approaches and features for the identification of HTTP-based botnet traffic)
 
@@
 
+\cite{garcia2014empirical} proposed CTU-13 dataset and argues that for evaluations of anomaly detection, none of the botnet families used in the training and crossvalidation dataset should be used in the testing dataset, considering that this ensures that the methods can generalize and detect new behaviors.
+However, hour proposal is semi-supervised and does not rely on training data with anomalies.
+Adopting the training and testing approach of \cite{garcia2014empirical}, some botnets malware wouldnt be evaluated, since in Garcia's proposal some botnets are present only for training.
+Therefore, we evaluate each scenario individually.
+
+A. Description of CTU-13 Dataset
+The CTU-13 is a dataset of botnet traffic that was captured
+in the Czech Technical University [24]. It contains 13 scenarios
+with various botnet types. We consider scenarios 1, 2, 6, 8, and
+9 of the dataset.
+• Scenario 1 corresponds to an IRC-based botnet that sent
+spam for almost 6.5 hours.
+• Scenario 2 (2.5-hours) is from the same botnet.
+• Scenario 6 is from a botnet that scanned SMTP (Simple
+Mail Transfer Protocol) servers for 2 hours and connected
+to several RDP (Remote Desktop Protocol) services. Dif-
+ferent with Scenarios 1 and 2, this botnet neither sent
+spam nor did it attack. It’s C&C server used a proprietary
+protocol.
+• Scenario 8 is from a botnet that contacted a lot of Chinese
+C&C hosts and received large amounts of encrypted data.
+The botnet also cracked the passwords of machines during
+the 19-hour attack.
+• Scenario 9 corresponds to a case where 10 local hosts
+were infected by a spamming botnet. More than 600 spams
+were sent over 5 hours.
+For all the scenarios, the authors of the CTU-13 dataset
+convert the captured pcap files to NetFlows and release the
+processed flows. The dataset contains ground-truth labels for
+flows as follows: flows from or to the infected machines
+are labeled as “botnet”; flows from or to noninfected ma-
+chines are labeled as “normal”; all other flows are labeled as
+“background.”
+
 Evaluating network anomaly detection solutions remains a challenge due to the lack of up-to-date real-world data sets for training and testing \cite{osanaiye2016distributed}.
@@
 
-To deal with the data imbalance of CTU-13 data set, we adopt a sampling method based on data aggragation \cite{acarali2016survey} for feature generation, in order to obtain less imbalanced and more discrimnative data.
-
-Based on related works \cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
-The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15, 0.25, 1 and 2 seconds, for each scenario of the CTU-13 data set.
+To deal with the data imbalance of CTU-13 data set, we adopt a sampling method based on data aggragation for feature generation, in order to obtain less imbalanced and more discrimnative data.
+Based on related works \cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
+The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15 and 0.25 seconds, for each scenario of the CTU-13 data set.
 
@@
 In anomalous detection problems, where anomalies are rare events, if we classify all observations as normal and apply an accuracy evaluation, we would have high accuracy but poor true-positive detection.
+
 Due to the importance given by F-measure to true-positive detection \cite{moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.
@@
 
-		there are no lots of researches using CTU-13, although it is been growing
-		poor results for GMM, K-Means and MGM are ok and in accordance to \cite{wang2017botnet}, which also evaluate \cite{gu2007bothunter}
+there are no lots of researches using CTU-13, although it is been growing poor results for GMM, K-Means and MGM are ok and in accordance to \cite{wang2017botnet}, which also evaluate \cite{gu2007bothunter}
 

### 2019-04-27 network-attack-detection/moment-rpca/testrospca.R

@@ -1,92 +1,92 @@
 library(rospca)
 
 test_df <- read.csv("/home/thiago/dev/anomaly-detection/network-attack-detection/data/ctu_13/csv_sum_test/datactu_13pkl_sum015s_10pk1_test_df.csv", header=TRUE, sep=",")
-resRS <- robpca(test_df, k=10, skew=TRUE,ndir=5000)
+resRS <- robpca(test_df, k=19, skew=TRUE,ndir=5000)
 write(resRS$flag.all, file = "robpca_k19_datactu_13pkl_sum015s_10pk1_test_df", sep = "\n")
 test_df <- read.csv("/home/thiago/dev/anomaly-detection/network-attack-detection/data/ctu_13/csv_sum_test/datactu_13pkl_sum025s_10pk1_test_df.csv", header=TRUE, sep=",")
-resRS <- robpca(test_df, k=10, skew=TRUE,ndir=5000)
+resRS <- robpca(test_df, k=19, skew=TRUE,ndir=5000)
 write(resRS$flag.all, file = "robpca_k19_datactu_13pkl_sum025s_10pk1_test_df", sep = "\n")
 
 test_df <- read.csv("/home/thiago/dev/anomaly-detection/network-attack-detection/data/ctu_13/csv_sum_test/datactu_13pkl_sum015s_11pk1_test_df.csv", header=TRUE, sep=",")
-resRS <- robpca(test_df, k=10, skew=TRUE,ndir=5000)
+resRS <- robpca(test_df, k=19, skew=TRUE,ndir=5000)
 write(resRS$flag.all, file = "robpca_k19_datactu_13pkl_sum015s_11pk1_test_df", sep = "\n")
 test_df <- read.csv("/home/thiago/dev/anomaly-detection/network-attack-detection/data/ctu_13/csv_sum_test/datactu_13pkl_sum025s_11pk1_test_df.csv", header=TRUE, sep=",")
-resRS <- robpca(test_df, k=10, skew=TRUE,ndir=5000)
+resRS <- robpca(test_df, k=19, skew=TRUE,ndir=5000)
 write(resRS$flag.all, file = "robpca_k19_datactu_13pkl_sum025s_11pk1_test_df", sep = "\n")
 
[…]

### 2019-04-27 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
reorganization of results in files and conclusion of literature review transcription

@@
 Lots of efforts are directed to avoid attacks from known attackers, however attacks from one source is not real nowadays, since the attacks are usualy from distributed botnets that change their approaches frequently, in order to avoid behavioral or pattern based detecions.
-This changing behavior is known as adversarial model \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,da2018online,Wang2018ddosbotnetssurvey}.
+This changing behavior is known as adversarial model \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
 
@@
 
-Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning}.
+Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}.
 Learning algorithms for imbalanced data has been a challenging research topic.
@@
 To face a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
+For more information about learning methods from imbalanced data, we refer to He and Garcia \cite{he2008learning}.
 
-Some approaches assume a gaussian distribute data, hoever it is not true in real world for many fields....Data skewness...
+\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions).
+-Argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
+Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning.
+Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms.
+\textbf{To explain feature selection and aggregation}.
+Goldberg and Shan \cite{goldberg2015importance} argue that processing monitored signals into features is a more fruitful way to build anomaly detection systems than devising ever more complex statistical tests, based on experience and  on putblished real-life examples of anomaly detectors in use at eBay.
+
+% Say that the data is lognormal \cite{benson2010network} and link it to skewness https://www.weibull.com/hotwire/issue47/relbasics47.htm
+Some approaches assume a gaussian distribute data, hoever it is not true in real world for many types of network traffic, wich usually is more related to skewed and heavy-tailed distributions.
+Benson \emph{et al} \cite{benson2010network} conducted an empirical study of the network traffic in 10 data centers belonging to three different types of organizations, including university, enterprise, and cloud data centers.
+Findings of Benson \emph{et al} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
+This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that assum a guassian distributed data, as well as can indicate characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
 
@@
 The botnet is commandeered by a “botmaster” and utilized as “resource” or “platform” for attacks such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration.
-In order for a botmaster to command a botnet, there needs to be a command and control (C&C) channel through which bots receive commands and coordinate attacks and fraudulent activities.
-The C&C channel is the means by which individual bots form a botnet.
-Centralized C&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time.
+In order for a botmaster to command a botnet, there needs to be a command and control (C\&C) channel through which bots receive commands and coordinate attacks and fraudulent activities.
+The C\&C channel is the means by which individual bots form a botnet.
+Centralized C\&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time.
 In this architecture, each bot logs into an IRC channel, and seeks commands from the botmaster.
 Even today, many botnets are still designed this way.
-Quite a few botnets, though, have begun to use other protocols such as HTTP [8, 14, 24, 39], probably because HTTP-based C&C communications are more stealthy given that Web traffic is generally allowed in most networks \cite{gu2008botminer}.
+Quite a few botnets, though, have begun to use other protocols such as HTTP [8, 14, 24, 39], probably because HTTP-based C\&C communications are more stealthy given that Web traffic is generally allowed in most networks \cite{gu2008botminer}.
 
@@
 Gu \emph{et al} \cite{gu2008botminer} presented the BotMiner, which aims to detect groups of compromised machines that are part of a botnet.
-BotMiner monitors communications that may suggest C&C or malicious activities, and finds a coordinated group pattern by means of clusters of similar communication activities, clusters of similar malicious activities, and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.
+BotMiner monitors communications that may suggest C\&C or malicious activities, and finds a coordinated group pattern by means of clusters of similar communication activities, clusters of similar malicious activities, and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.
 
@@
 
-\cite{da2018online} (Online detection of Botnets on Network Flows using Stream Mining) Ver se podemos calcular o F1 e comparar os resutados, mas não seria justo dado que a abordagem deles nao implementa o mesmo sampling e é supervisionado
+% \cite{da2018online} proposes an approach using the Very Fast Decision Tree, a classification algorithm used on stream mining that can learn incrementally when needed, to identify botnets by observing network flows.
+When evaluating the approach on multiple scenarios with different botnets, we were able to achieve high performance metrics on the majority of scenarios, while using a significantly low number of labelled instances.
+It discard scenarios 5 and 6.
+They does not compute F1, but it is easy to compute it from precision and recall provided.
 
@@
 
-\cite{Phua2004minority} (Minority Report in Fraud Detection: Classification of Skewed Data)
-
-\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions).
--Argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
-Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning.
-Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms.
-\textbf{To explain feature selection and aggregation}.
-
-\cite{goldberg2015importance} (The importance of features for statistical anomaly detection)
+\cite{zhou2017anomaly} (Anomaly detection with robust deep autoencoders)
 
@@
 ferent with Scenarios 1 and 2, this botnet neither sent
-spam nor did it attack. It’s C&C server used a proprietary
+spam nor did it attack. It’s C\&C server used a proprietary
 protocol.
 • Scenario 8 is from a botnet that contacted a lot of Chinese
-C&C hosts and received large amounts of encrypted data.
+C\&C hosts and received large amounts of encrypted data.
 The botnet also cracked the passwords of machines during
@@
 
-Due to the importance given by F-measure to true-positive detection \cite{moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.
+Due to the importance given by F-measure to true-positive detection \cite{powers2011evaluation,moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.
 
@@
 
+Why GMM?
 As clustering-based method, we selected K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007k,moustafa2019holistic}.
@@
 
-Say that the data is lognormal \cite{benson2010network} and link it to skewness https://www.weibull.com/hotwire/issue47/relbasics47.htm
 Test distribution fit again e present results 

### 2019-04-27 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
reorganization of results in files and conclusion of literature review transcription

@@
 
-Botnets have emerging as a key threat for cyber security, considering that Botnets are the key platform for many Internet attacks (i.e.
-spam, distributed denial-of-service (DDoS), identity theft, and phishing).
-Most of the current botnet detection approaches work only on specific botnet command and control (.
-Lots of efforts are directed to avoid attacks from known attackers, however attacks from one source is not real nowadays, since the attacks are usualy from distributed botnets that change their approaches frequently, in order to avoid behavioral or pattern based detecions.
-This changing behavior is known as adversarial model \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
+Several efforts and reasearches aim to avoid network attacks based on known attackers, fingerprints or other kind of pattern.
+However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in oder to detect and avoid unkown attacks or even to deal with adversarial changes of behavior, location and patterns, in order to avoid behavioral or pattern based detecion systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
+To face this problem, it is possible to adopt unsupervised or semi-pervised approches, where it is not necessary known anomalies for trannning models.
 
 Anomaly detection is the identification of rare and suspicions observations by differing the normal or majority of the data.
-
 Anomalies can be related to network attack, fraud, defect or medical problems, for examples.
-
 Anomalies are also referred to as outliers, novelties, noise, deviations and exceptions.
-
 Anomalies can be hard to identify and separate from normal data.
@@
 
+Our approach train from normal traffic and have significant computational cost during training. For prediction, the cost is basically the cost for Mahalanobis distance computation
+
 The main contributions of the proposed framework are a semi-supervised method for network attack detection, wich does not depend on knowledge of previous attacks and consider only normal network traffic for robust estimates normal traffic...
@@
 This paper is organized as follows.
-In Section \ref{sec:relatedworks}, related works are discussed.
+In Section \ref{sec:review}, it is conducted a literature review of network anomaly detection, botnet detection, sampling and aggregation for imbalanced learning and robust based anomaly detection.
 Section \ref{sec:datamodel} presents the data model and the evaluated datasets.
-Section \ref{sec:c_mcd} describes the proposed framework for blind and automatic detection of flood and probe attacks.
+Section \ref{sec:m_rpca} describes the proposed framework for botnet detection based on robust subspace and skewness distance.
 Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
-
-Our approach train from normal traffic and have significant computational cost during training. For prediction, the cost is basically the cost for Mahalanobis distance computation
 
@@
 \section{Literature Review}
-\label{sec:relatedworks}
+\label{sec:review}
 

### 2019-04-28 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on introduction

@@
 
-Several efforts and reasearches aim to avoid network attacks based on known attackers, fingerprints or other kind of pattern.
-However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in oder to detect and avoid unkown attacks or even to deal with adversarial changes of behavior, location and patterns, in order to avoid behavioral or pattern based detecion systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
-To face this problem, it is possible to adopt unsupervised or semi-pervised approches, where it is not necessary known anomalies for trannning models.
+Anomaly detection is the identification of rare and suspicions events by differing the normal or majority of the data.
+Anomalies are also referred to as outliers, novelties, noise, deviations or exceptions, and  can be related to network attack, fraud, defect or medical condiction, for examples \cite{bhuyan2014network,ahmed2016survey}.
+Anomalies can be hard to identify and separate from normal data due to the rare occurrencies of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be robust, high discriminative and able to deal with the imbalanced data problem \cite{he2008learning}.
 
-Anomaly detection is the identification of rare and suspicions observations by differing the normal or majority of the data.
-Anomalies can be related to network attack, fraud, defect or medical problems, for examples.
-Anomalies are also referred to as outliers, novelties, noise, deviations and exceptions.
-Anomalies can be hard to identify and separate from normal data.
-
-Considering that anomalies are rare in comparison to normal events, anomaly detection algorithms have to deal with problems imposed by imbalanced data.
+Several efforts and reasearches aim to avoid network attacks based on known attackers, fingerprints or other patterns.
+However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in oder to detect and avoid unkown attacks or even to deal with adversarial changes of behavior, location and patterns, in order to avoid detection by behavioral or pattern based detecion systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
+To face this problem, it is possible to adopt unsupervised or semi-pervised approches, where it is not necessary known anomalies for trannning models \cite{moustafa2019holistic}.
 

### 2019-04-28 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on introduction

@@
 
-\title{Robust Subspace and Higher Moments Distances for Botnet Detection}
+\title{Higher Moments Distances from Robust Subspace for Botnet Detection}
 
@@
 
-Anomaly detection is the identification of rare and suspicions events by differing the normal or majority of the data.
-Anomalies are also referred to as outliers, novelties, noise, deviations or exceptions, and  can be related to network attack, fraud, defect or medical condiction, for examples \cite{bhuyan2014network,ahmed2016survey}.
-Anomalies can be hard to identify and separate from normal data due to the rare occurrencies of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be robust, high discriminative and able to deal with the imbalanced data problem \cite{he2008learning}.
+Anomaly detection can be defined as the identification of rare and suspicions events by differing the normal or majority of the data.
+Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects \cite{bhuyan2014network,ahmed2016survey}.
+Anomalies can be hard to identify and separate from normal data due to the rare occurrencies of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.
 
 Several efforts and reasearches aim to avoid network attacks based on known attackers, fingerprints or other patterns.
-However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in oder to detect and avoid unkown attacks or even to deal with adversarial changes of behavior, location and patterns, in order to avoid detection by behavioral or pattern based detecion systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
-To face this problem, it is possible to adopt unsupervised or semi-pervised approches, where it is not necessary known anomalies for trannning models \cite{moustafa2019holistic}.
+However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in oder to detect and avoid unkown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detecion systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
+To face this problem, it is possible to adopt unsupervised or semi-pervised approaches for network anomaly detection, where it is not necessary known anomalies for trannning models \cite{moustafa2019holistic}.
 
-Robust statistics produce methods that are not unduly affected by outliers.
+Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}.
+Learning algorithms for imbalanced data has been a challenging research topic, considering that the fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
+Some techniques can make the data discriminative for better results on imbalanced data: Data engineering (i.e.
+sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, garcia2014empirical, acarali2016survey}); Feature extraction (i.e.
+decomposition \cite{vaswani2018robust}, robust estimates \cite{zhou2017anomaly}).
 
-Robust statistical methods have been developed for estimating location, scale, and regression parameters.
+% He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning.
+For more information about learning methods from imbalanced data, we refer to He and Garcia \cite{he2008learning}.
 
-Robust estimates from contaminated data can denoise, remove anomalies or even separate them.
+% Say that the data is lognormal \cite{benson2010network} and link it to skewness https://www.weibull.com/hotwire/issue47/relbasics47.htm
+Some approaches or widely adopted algorithms assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network}.
+Findings of Benson \emph{et al}  \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
+This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
+Therefore, learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
+Skewness and kurtosis are higher moments that their measurement can be very sensitive for outlier detection, considering that the properties of the kurtosis has been used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
 
-Anomaly detection based on robust estimates has been adopted for network anomaly detection, due to good results and its adaptability.
-
-The distance between robust estimates and new observations can reveal anomalies.
+Principal Component Analysis (PCA) is very sensitive to anomalies, thus outliers can corrupt the entire data or can be explored to highlight unespected changes and indicate attacks or frauds \cite{callegari2011novel,Lee2013,vieira2017model}.
+However, PCA is not precise to reveal flows, time or detailed information for each anomlies, and sometimes relies into visual analysis for pricipal component selection.
+However, Robust Principal Component Analysis (RPCA) \cite{candes2011robust} aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\textbf{X}$ into the sum of a low rank matrix, $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix, $\textbf{S}$, which refers to outliers’ matrix.
+Therefore, robust subspace learning has been adopted for network anomlay detection \cite{rousseeuw1984mcd, rousseeuw1999fastmcd, hubert2005robpca,hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
 
 Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples by, considering the correlations between the variables and their mean values.
+MD has been used for distance based anomaly detection with robust estimates in many areas, considering that the distance between robust estimates and new observations can reveal anomalies.
 
-\textbf{MD has been used for distance based anomaly detection with robust estimates} in many areas, and is commonly combined to Fast Minimum Covariance Determinant (Fast-MCD) algorithm \cite{rousseeuw1999fastmcd} for robust estimates.
-
-Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}.
-Learning algorithms for imbalanced data has been a challenging research topic.
-The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
-He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning.
-Some techniques can make the data discriminative for better discriminative analytics: Data engineering (i.e.
-sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, garcia2014empirical, acarali2016survey}); Feature extraction (i.e.
-decomposition \cite{vaswani2018robust}, robust estimates \cite{zhou2017anomaly}).
-To face a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
-For more information about learning methods from imbalanced data, we refer to He and Garcia \cite{he2008learning}.
-
-\cite{lakhina2005mining} (Mining anomalies using traffic feature distributions).
--Argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
-Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning.
-Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms.
-\textbf{To explain feature selection and aggregation}.
-Goldberg and Shan \cite{goldberg2015importance} argue that processing monitored signals into features is a more fruitful way to build anomaly detection systems than devising ever more complex statistical tests, based on experience and  on putblished real-life examples of anomaly detectors in use at eBay.
-
-% Say that the data is lognormal \cite{benson2010network} and link it to skewness https://www.weibull.com/hotwire/issue47/relbasics47.htm
-Some approaches assume a gaussian distribute data, hoever it is not true in real world for many types of network traffic, wich usually is more related to skewed and heavy-tailed distributions.
-Benson \emph{et al} \cite{benson2010network} conducted an empirical study of the network traffic in 10 data centers belonging to three different types of organizations, including university, enterprise, and cloud data centers.
-Findings of Benson \emph{et al} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
-This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that assum a guassian distributed data, as well as can indicate characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
-
-Skewness and the kurtosis tests are very sensitive outlier detectors.
-The properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
-
-PCA is very sensitive to anomalies, thus outliers can corrupt the entire data or can be explored to highlight unespected changes and indicate attacks or frauds \cite{callegari2011novel,Lee2013,vieira2017model}.
-However, PCA is not precise to reveal flows, time or detailed information for each anomlies, and sometimes relies into visual analysis for pricipal component selection.
-
-We propose a network anomaly detection approach based on robust moment estimates and distance analysis.
-
-Considering that Fast-MCD is a widely adopted robust estimator of location and scatter for anomaly detection, we propose to extend Fast-MCD for computing robust skewness and kurtosis.
-
-We also propose to use Mahalanobis Distance for detecting anomalies based on robust skewness and kurtosis, instead of the location-based usage of classical approaches.
+We believe that the skewness of anomalous and normal traffic can reveal characteristcs for improving anomaly detection in imbalanced data.
+The botnet detection algorithms can also be improved by feature generation from aggregated traffic, considering that new features can make the data more discriminative for imbalanced learning problems.
+We also believe that the distance between robust estimates of normal traffic and anomalies can highlight discrepancies and be used for botnet detection.
+Therefore, we propose an approach based on higher moments distances from robust subspace for botnet detection.
 
 Considering that it is hard to have labeled data in real problems \cite{osanaiye2016distributed}, such as for network attacks or frauds, also considering that known anomalies are constantly changing due to adversarial models or due to lack of regular patterns, we propose a Semi-Supervised approach based on robust moments for network anomaly detection.
-
-We believe that the distribution of anomalous and normal traffic are skewed but with differences that permit separate anomalous of normal traffic.
-(Assumption and hitotesis).
-We also believe that aggregated traffic and new features can make the data more discriminative for anomaly detection algorithms.
-We algo believe that normal traffic presents significant distance to anomalous traffic, which permit a distance based sepparation, that can be augmented by means of robust estimation of normal traffic, that can discard noise or irrelevant outliers and can highlight higher outliers when comparing to new observations
-
-In the context of anomaly-based schemes, this work proposes a statistical approach based on cumulants and Mimimum Covariance Determinant (MCD) for network attack detection.
-To the best of our knowledge there is no similar model in the literature.
+We also propose to use Mahalanobis Distance for detecting anomalies based on robust skewness and kurtosis, instead of the location-based usage of classical approaches.
+Our approach train from normal traffic and have significant computational cost during training.
+For prediction, the cost is basically the cost for Mahalanobis distance computation To face the anomaly detection from a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
 We evaluate the accuracy of the proposed framework applied to CTU-13 dataset ??, which is a well known network traffic dataset.
 
-Our approach train from normal traffic and have significant computational cost during training. For prediction, the cost is basically the cost for Mahalanobis distance computation
+To the best of our knowledge there is no similar model in the literature.
 
@@
 Talk about network anomaly detection...
+
+Lakhina \emph{et al} \cite{lakhina2005mining} argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
+Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning.
+Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms.
+\textbf{To explain feature selection and aggregation}.
+Goldberg and Shan \cite{goldberg2015importance} argue that processing monitored signals into features is a more fruitful way to build anomaly detection systems than devising ever more complex statistical tests, based on experience and  on putblished real-life examples of anomaly detectors in use at eBay.
 
@@
 They does not compute F1, but it is easy to compute it from precision and recall provided.
+
+We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for a robust subspace learning
+
+Benson \emph{et al} \cite{benson2010network} conducted an empirical study of the network traffic in 10 data centers belonging to three different types of organizations, including university, enterprise, and cloud data centers.
+Findings of Benson \emph{et al} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
+This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
+Therefore, learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
+Skewness and the kurtosis are very sensitive outlier detectors, considering that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
+
+In contrast to classical PCA, our robust PCA detector does not require having a perfect ground-truth for training; it just trains from the (eventually contaminated) background traffic.
+It is, in fact, an unsupervised learning algorithm.
+Moreover, the addition of feature selection allows matching the specificities of the traffic the detector is operating on, which can vary widely both in terms of the most frequent attacks and the relative weight of licit applications.
+Feature selection does require training with a labeled dataset bu
+
+aims to reduce the sensitivity of PCA to outliers, by means of .
+In particular, RPCA allows for the careful teasing apart of sparse outliers so that the remaining low-dimensional approximation is faithful to the noise-free low-dimensional subspace describing the bulk of the raw data [6, 7, 20].
+Specifically, RPCA splits a data matrix X into a low-rank matrix L and a sparse matrix S such that
+X = L + S.
 

### 2019-04-29 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
abstract and introduction ok

@@
 \begin{abstract}
-Botnets...
-In this context, signal processing techniques have been applied to attack detection due to their capability to detect anomalies that are previously unknown, i.e.
-blind detection, in contaminated or imbalanced data....
-This paper proposes a cumulant-based robust estimates for network attack detection, by combining kurtosis and skewess with Minimum Covariance Determinant.
-In order to validate the proposed framework, we consider CTU-13 dataset.
-The experiments show that the proposed method achieves satisfactory levels of precisio, recall and F1 score for network attack detection.
-We apply our method to real-world botnet traffic and compare its performance with other methods.
+Distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid botnet's attacks.
+This paper proposes an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggretation, in a semi-supervisioned fashion and without the computational cost of new robust subspace learning for new observations.
+Experimental results reveal that the proposed approach outerperforms widely adoted algorithms and a method based on robust estimates and robust skewness for network anomaly detection.
 \end{abstract}
@@
 \begin{keyword}
-Botnet Detection \sep Network Anomaly Detection \sep Skewness \sep Kurtosis \sep Robust Principal Component Analysis \sep Mahalanobis Distance
+Network Anomaly Detection \sep Botnet Detection \sep Higher Moments \sep Robust Principal Component Analysis \sep Mahalanobis Distance
 \end{keyword}
@@
 Several efforts and reasearches aim to avoid network attacks based on known attackers, fingerprints or other patterns.
-However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in oder to detect and avoid unkown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detecion systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
+However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid unkown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detecion systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
 To face this problem, it is possible to adopt unsupervised or semi-pervised approaches for network anomaly detection, where it is not necessary known anomalies for trannning models \cite{moustafa2019holistic}.
@@
 % Say that the data is lognormal \cite{benson2010network} and link it to skewness https://www.weibull.com/hotwire/issue47/relbasics47.htm
-Some approaches or widely adopted algorithms assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network}.
+Some widely adopted algorithms assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network}.
 Findings of Benson \emph{et al}  \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
-This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
+This findings show that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.
 Therefore, learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
-Skewness and kurtosis are higher moments that their measurement can be very sensitive for outlier detection, considering that the properties of the kurtosis has been used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
+Skewness and kurtosis are higher moments that can have their measurement as very sensitive for outlier detection, considering that the properties of the kurtosis has been used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
 
 Principal Component Analysis (PCA) is very sensitive to anomalies, thus outliers can corrupt the entire data or can be explored to highlight unespected changes and indicate attacks or frauds \cite{callegari2011novel,Lee2013,vieira2017model}.
-However, PCA is not precise to reveal flows, time or detailed information for each anomlies, and sometimes relies into visual analysis for pricipal component selection.
-However, Robust Principal Component Analysis (RPCA) \cite{candes2011robust} aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\textbf{X}$ into the sum of a low rank matrix, $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix, $\textbf{S}$, which refers to outliers’ matrix.
-Therefore, robust subspace learning has been adopted for network anomlay detection \cite{rousseeuw1984mcd, rousseeuw1999fastmcd, hubert2005robpca,hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
+However, PCA is not precise to reveal flows, time or detailed information for each anomaly, and sometimes relies into visual analysis for pricipal component selection.
+Robust Principal Component Analysis (RPCA) \cite{candes2011robust} is an extensio of PCA that aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, which refers to outliers’ matrix.
+Therefore, robust subspace learning has receiving a growing atention of researchers aiming the development of network anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd, hubert2005robpca,hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
 
-Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples by, considering the correlations between the variables and their mean values.
-MD has been used for distance based anomaly detection with robust estimates in many areas, considering that the distance between robust estimates and new observations can reveal anomalies.
+Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the correlations between the variables and their mean values.
+MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and new observations can reveal anomalies.
 
-We believe that the skewness of anomalous and normal traffic can reveal characteristcs for improving anomaly detection in imbalanced data.
+We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data.
 The botnet detection algorithms can also be improved by feature generation from aggregated traffic, considering that new features can make the data more discriminative for imbalanced learning problems.
 We also believe that the distance between robust estimates of normal traffic and anomalies can highlight discrepancies and be used for botnet detection.
-Therefore, we propose an approach based on higher moments distances from robust subspace for botnet detection.
+Therefore, we propose an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggretation.
+The proposed approach relies on a robust subspace of normal traffic for estimating higher moments.
+The anomaly detection for new observations evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervisioned fashion and without the computational cost of new robust subspace learning for new observations.
 
-Considering that it is hard to have labeled data in real problems \cite{osanaiye2016distributed}, such as for network attacks or frauds, also considering that known anomalies are constantly changing due to adversarial models or due to lack of regular patterns, we propose a Semi-Supervised approach based on robust moments for network anomaly detection.
-We also propose to use Mahalanobis Distance for detecting anomalies based on robust skewness and kurtosis, instead of the location-based usage of classical approaches.
-Our approach train from normal traffic and have significant computational cost during training.
-For prediction, the cost is basically the cost for Mahalanobis distance computation To face the anomaly detection from a high imbalanced data set, we adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
-We evaluate the accuracy of the proposed framework applied to CTU-13 dataset ??, which is a well known network traffic dataset.
+We adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
+We evaluate the accuracy of the proposed approach for botnet detecion on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
 
-To the best of our knowledge there is no similar model in the literature.
+Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statisctical approaches, which are K-means and Gaussian Mixture Model (GMM) \cite{gaddam2007kmeans,moustafa2019holistic}, respectively.
+We also evaluate the results of ROBPCA \cite{hubert2005robpca}, wich is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
 
-The main contributions of the proposed framework are a semi-supervised method for network attack detection, wich does not depend on knowledge of previous attacks and consider only normal network traffic for robust estimates normal traffic...
+The main contribution of this work is a novel semi-supervised method for botnet detection that obtain results with high rates of attack detection in an imbalanced and skewed data set.
 
 This paper is organized as follows.
-In Section \ref{sec:review}, it is conducted a literature review of network anomaly detection, botnet detection, sampling and aggregation for imbalanced learning and robust based anomaly detection.
-Section \ref{sec:datamodel} presents the data model and the evaluated datasets.
-Section \ref{sec:m_rpca} describes the proposed framework for botnet detection based on robust subspace and skewness distance.
+In Section \ref{sec:review}, it is conducted a literature review of network anomaly detection, botnet detection, sampling and aggregation for imbalanced learning and robust based anomaly detection methods.
+Section \ref{sec:datamodel} presents the data model and the evaluated data set.
+Section \ref{sec:m_rpca} describes the proposed approach for botnet detection.
 Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
@@
 Ahmed Bhuyan \emph{et al} \cite{acarali2016survey} present an analysis of four major categories of anomaly detection techniques which include classification, statistical, information theory and clustering.
-Recently, Moustafa \emph{et al} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), explains cyber-attacks and new solutions for anomaly detection, and provides a benchmark datasets for training and validating approaches for detection.
+Recently, Moustafa \emph{et al} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), explains cyber-attacks and new solutions for anomaly detection, and provides a benchmark data sets for training and validating approaches for detection.
 
@@
 \cite{garcia2014empirical} (An empirical comparison of botnet detection methods).
-The goal of this paper is to compare three botnet detection methods using a simple and reproducible methodology, a good dataset and a new error metric.
+The goal of this paper is to compare three botnet detection methods using a simple and reproducible methodology, a good data set and a new error metric.
 This paper evaluates some data sets for network anomaly detection and classify the availability of labels and the presesence of background, normal and botnet traffic.
@@
 The paper proposes two methods (BClus and CAMNEP) and compare results to BotHunter.
-Considering the lack of available labeled datasets, the paper proposes the CTU-13 dataset, which is composed by normal and background labeled data, with the balance of the dataset like in a real network.
-The training and cross-validation datasets should be approximately 80\% of the dataset.The testing dataset should be approximately 20\% of the dataset.
-None of the botnet families used in the training and crossvalidation dataset should be used in the testing dataset.
+Considering the lack of available labeled data sets, the paper proposes the CTU-13 data set, which is composed by normal and background labeled data, with the balance of the data set like in a real network.
+The training and cross-validation data sets should be approximately 80\% of the data set.The testing data set should be approximately 20\% of the data set.
+None of the botnet families used in the training and crossvalidation data set should be used in the testing data set.
 This ensures that the methods can generalize and detect new behaviors.
 \cite{garcia2014empirical} uses a time window separation and an aggregation method.
-Thei compare the output of three different botnet detection methods by executing them over a new, real, labeled and large botnet dataset.
+Thei compare the output of three different botnet detection methods by executing them over a new, real, labeled and large botnet data set.
 
-Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks.
-Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
-NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications.
-Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations.
-Considering that it is important to create intrusion detection datasets in modern-day computing to address the issues of DARPA/KDD.
-The next section discusses ne such contemporary dataset for network traffic analysis \cite{ahmed2016survey}, we propose to adopt CTU-13...
+Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
+Even though the KDD 99 data set are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data sets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
+NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that redundant data records are removed, in order to avoid biased classifications.
+Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public data sets their evaluations.
+Considering that it is important to create intrusion detection data sets in modern-day computing to address the issues of DARPA/KDD.
+The next section discusses ne such contemporary data set for network traffic analysis \cite{ahmed2016survey}, we propose to adopt CTU-13...
 
@@
 The second stage detects the bots using ideas from social network community detection in a graph that captures correlations of interactions among nodes over time.
-This work is compared with the BotHunter \cite{gu2007bothunter} on the CTU-13 botnet dataset \cite{garcia2014empirical}.
+This work is compared with the BotHunter \cite{gu2007bothunter} on the CTU-13 botnet data set \cite{garcia2014empirical}.
 
@@
 Moreover, the addition of feature selection allows matching the specificities of the traffic the detector is operating on, which can vary widely both in terms of the most frequent attacks and the relative weight of licit applications.
-Feature selection does require training with a labeled dataset bu
+Feature selection does require training with a labeled data set bu
 
@@
 However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques.
-Callegari \emph{et al} \cite{callegari2011novel} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
+Callegari \emph{et al} \cite{callegari2011novel} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a data set with synthetic anomalies added in the data.
 However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
@@
 
-This section also presents a description of the data model and data aggregation of the CTU-13 dataset, in Subsection \ref{sec:CTU-13} discusses the use of the CTU-13 dataset for evaluation of the proposed approach.
+This section also presents a description of the data model and data aggregation of the CTU-13 data set, in Subsection \ref{sec:CTU-13} discusses the use of the CTU-13 data set for evaluation of the proposed approach.
 
-\subsection{The CTU-13 Dataset}
+\subsection{The CTU-13 data set}
 \label{sec:CTU-13}
 
-\cite{garcia2014empirical} proposed CTU-13 dataset and argues that for evaluations of anomaly detection, none of the botnet families used in the training and crossvalidation dataset should be used in the testing dataset, considering that this ensures that the methods can generalize and detect new behaviors.
+\cite{garcia2014empirical} proposed CTU-13 data set and argues that for evaluations of anomaly detection, none of the botnet families used in the training and crossvalidation data set should be used in the testing data set, considering that this ensures that the methods can generalize and detect new behaviors.
 However, hour proposal is semi-supervised and does not rely on training data with anomalies.
@@
 
-A. Description of CTU-13 Dataset
-The CTU-13 is a dataset of botnet traffic that was captured
+A. Description of CTU-13 data set
+The CTU-13 is a data set of botnet traffic that was captured
 in the Czech Technical University [24]. It contains 13 scenarios
 with various botnet types. We consider scenarios 1, 2, 6, 8, and
-9 of the dataset.
+9 of the data set.
 • Scenario 1 corresponds to an IRC-based botnet that sent
@@
 were sent over 5 hours.
-For all the scenarios, the authors of the CTU-13 dataset
+For all the scenarios, the authors of the CTU-13 data set
 convert the captured pcap files to NetFlows and release the
-processed flows. The dataset contains ground-truth labels for
+processed flows. The data set contains ground-truth labels for
 flows as follows: flows from or to the infected machines
@@
 	\tiny
-	\caption{CTU-13 Dataset Description}
+	\caption{CTU-13 data set Description}
 	\label{tab:tab01}
@@
 Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario.
-In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated dataset.
+In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated data set.
 Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario.
-Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 dataset.
+Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 data set.
 
@@
 
-We propose to compare our approach to clustering and statisctical based approaches (K-means and Gaussian Mixture Model (GMM), respectively).
-
 Why GMM?
-As clustering-based method, we selected K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007k,moustafa2019holistic}.
+As clustering-based method, we selected K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007kmeans,moustafa2019holistic}.
 It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}.
@@
 
-Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets.
+Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets.
 Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high thoughput network traffic.

### 2019-04-29 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on literature review

@@
 Some widely adopted algorithms assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network}.
-Findings of Benson \emph{et al}  \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
+Findings of Benson \emph{et al.}  \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
 This findings show that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% network anomaly detectiom, botnet detection, botnet detection approaches, approaches based on subspace learning, sampling and aggregation, about the methodos to compare and metrics
+% network anomaly detectiom, botnet detection, botnet detection approaches, approaches based on subspace learning, feature generation and aggregation, about the methodos to compare and metrics
 \section{Literature Review}
@@
 
-Talk about network anomaly detection...
+% botnet detection
+A botnet is a network of compromised machines under the influence of malware (bot).
+The botnet is commandeered by a “botmaster” and utilized as “resource” or “platform” for attacks, such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration.
+The botmaster coordenate a botnet through a command and control (C\&C) channel where bots receive commands and syncronize attacks and fraudulent activities.
+Centralized C\&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time, but other protocols, such as HTTP, and achitectures, such as Peer-To-Peer, have also been adopted \cite{gu2008botminer}.
 
-Lakhina \emph{et al} \cite{lakhina2005mining} argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
-Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning.
-Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms.
-\textbf{To explain feature selection and aggregation}.
-Goldberg and Shan \cite{goldberg2015importance} argue that processing monitored signals into features is a more fruitful way to build anomaly detection systems than devising ever more complex statistical tests, based on experience and  on putblished real-life examples of anomaly detectors in use at eBay.
+Bhuyan \emph{et al.} \cite{acarali2016survey} provide an overview of facets of network anomaly detection, present attacks normally encountered by network intrusion detection systems, and categorize existing network anomaly detection methods and systems based on the underlying techniques.
+Ahmed Bhuyan \emph{et al.} \cite{acarali2016survey} present an analysis of four major categories of anomaly detection techniques which include classification, statistical, information theory and clustering.
+Moustafa \emph{et al.} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), explains cyber-attacks and new solutions for anomaly detection, and provides a benchmark data sets for training and validating approaches for detection.
 
-Bhuyan \emph{et al} \cite{acarali2016survey} provide an overview of facets of network anomaly detection, present attacks normally encountered by network intrusion detection systems, and categorize existing network anomaly detection methods and systems based on the underlying techniques.
-Ahmed Bhuyan \emph{et al} \cite{acarali2016survey} present an analysis of four major categories of anomaly detection techniques which include classification, statistical, information theory and clustering.
-Recently, Moustafa \emph{et al} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), explains cyber-attacks and new solutions for anomaly detection, and provides a benchmark data sets for training and validating approaches for detection.
-
-Botnets are becoming one of the most serious threats to Internet security.
-A botnet is a network of compromised machines under the influence of malware (bot) code.
-The botnet is commandeered by a “botmaster” and utilized as “resource” or “platform” for attacks such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration.
-In order for a botmaster to command a botnet, there needs to be a command and control (C\&C) channel through which bots receive commands and coordinate attacks and fraudulent activities.
-The C\&C channel is the means by which individual bots form a botnet.
-Centralized C\&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time.
-In this architecture, each bot logs into an IRC channel, and seeks commands from the botmaster.
-Even today, many botnets are still designed this way.
-Quite a few botnets, though, have begun to use other protocols such as HTTP [8, 14, 24, 39], probably because HTTP-based C\&C communications are more stealthy given that Web traffic is generally allowed in most networks \cite{gu2008botminer}.
-
-Acarali \emph{et al} \cite{acarali2016survey} surveys recent network-based detection approaches for HTTP-based botnets, surveys traffic-based features used to detect bot traffic and presents an abstraction of the main types of features in relation to protocols and OSI layers.
-Wang \emph{et al} \cite{Wang2018ddosbotnetssurvey} present an analysis based on 50,704 different Internet DDoS attacks originated of 674 botnets from 23 different botnet families with a total of 9,026 victim belonging to 1,074 organizations in 186 countries.
+Acarali \emph{et al.} \cite{acarali2016survey} surveys network-based detection approaches for HTTP-based botnets, surveys traffic-based features used to detect bot traffic and presents an abstraction of the main types of features in relation to protocols and OSI layers.
+Wang \emph{et al.} \cite{Wang2018ddosbotnetssurvey} present an analysis based on 50,704 different Internet DDoS attacks originated of 674 botnets from 23 different botnet families with a total of 9,026 victim belonging to 1,074 organizations in 186 countries.
 Their analysis reveals that geolocation of the attacking sources follows patterns and enables source prediction, and highlight that multiple attacks to the same target also exhibit strong patterns of inter-attack time interval, also presents that there is a trend for different botnets to launch DDoS attacks targeting the same victim, simultaneously or in turn.
 
-The BotHunter method was proposed by Gu \emph{et al} \cite{gu2007bothunter} to detect the infection and coordination of botnets by matching sequence model, through a correlation approach for detecting stages of the infection process.
-Gu \emph{et al} \cite{gu2008botminer} presented the BotMiner, which aims to detect groups of compromised machines that are part of a botnet.
+The BotHunter was proposed by Gu \emph{et al.} \cite{gu2007bothunter} to detect the infection and coordination of botnets by matching sequence model, through a correlation approach for detecting stages of the infection process.
+Gu \emph{et al.} \cite{gu2008botminer} presented the BotMiner, which aims to detect groups of compromised machines that are part of a botnet.
 BotMiner monitors communications that may suggest C\&C or malicious activities, and finds a coordinated group pattern by means of clusters of similar communication activities, clusters of similar malicious activities, and performs cross cluster correlation to identify the hosts that share both similar communication patterns and similar malicious activity patterns.
 
-Khattak \emph{et al} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, based on a Complex Event Processing (CEP) engine and on a correlation framework that continuously receives events and correlates them according to rule(s).
-BotFlex's results are compared to BotHunter, but the evaluation relies in a own and not public data set.
+Khattak \emph{et al.} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, based on a Complex Event Processing (CEP) engine and on a correlation framework that continuously receives events and correlates them according to rules.
+BotFlex's results are compared to BotHunter \cite{gu2007bothunter}, but the evaluation relies in a own and not public data set.
 
-\cite{garcia2014empirical} (An empirical comparison of botnet detection methods).
-The goal of this paper is to compare three botnet detection methods using a simple and reproducible methodology, a good data set and a new error metric.
-This paper evaluates some data sets for network anomaly detection and classify the availability of labels and the presesence of background, normal and botnet traffic.
-The paper also survey some approaches for botnet detection.
-The paper proposes two methods (BClus and CAMNEP) and compare results to BotHunter.
-Considering the lack of available labeled data sets, the paper proposes the CTU-13 data set, which is composed by normal and background labeled data, with the balance of the data set like in a real network.
-The training and cross-validation data sets should be approximately 80\% of the data set.The testing data set should be approximately 20\% of the data set.
-None of the botnet families used in the training and crossvalidation data set should be used in the testing data set.
-This ensures that the methods can generalize and detect new behaviors.
-\cite{garcia2014empirical} uses a time window separation and an aggregation method.
-Thei compare the output of three different botnet detection methods by executing them over a new, real, labeled and large botnet data set.
+% justificative for the dataset and present who else use it
+Several approaches for network attack detections uses the KDD 99 \cite{ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
+Even though the KDD 99 data set are criticized by the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data sets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
+NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that redundant data records are removed, in order to avoid biased classifications.
 
-Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
-Even though the KDD 99 data set are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data sets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
-NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that redundant data records are removed, in order to avoid biased classifications.
-Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public data sets their evaluations.
-Considering that it is important to create intrusion detection data sets in modern-day computing to address the issues of DARPA/KDD.
-The next section discusses ne such contemporary data set for network traffic analysis \cite{ahmed2016survey}, we propose to adopt CTU-13...
+The objective of Garcia \emph{et al.} \cite{garcia2014empirical} is to compare three botnet detection methods using a simple and reproducible methodology, a good data set and a new error metric.
+This paper evaluates some data sets for network anomaly detection, surveys some approaches for botnet detection and proposes two methods (BClus and CAMNEP) and compare results to BotHunter \cite{gu2007bothunter}.
+Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by normal and background labeled data, in an imbalanced distribution like in a real network.
+The autors recommend that none of the botnet families used in the training and crossvalidation data set should be used in the testing data set, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
+Adopting the training and testing approach of \cite{garcia2014empirical}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
+In contrast, we evaluate each scenario of CTU-13 individually in a semi-supervised approach that does not rely on training data with anomalies.
 
-Wang and Paschalidis \cite{wang2017botnet} propose an botnet detection approach based on anomaly and community detection, aiming for detecting botnets and identifying bots before the botnet becomes active.
+Wang and Paschalidis \cite{wang2017botnet} propose a botnet detection approach based on anomaly and community detection, aiming for detecting botnets and identifying bots before the botnet becomes active.
 The first stage detects anomalies by leveraging large deviations of an empirical distribution.
@@
 
-We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for a robust subspace learning
+Traditional PCA-based anomaly detection models are not suitable for anomaly interpretation, as they judge whether a data instance is an anomaly or not based on the length of its projection on the abnormal subspace spanned by the less significant principal components, and there is no direct mapping between PCA’s dimensionality-reduced subspace and the original feature space \cite{ringberg2007sensitivity}.
+Callegari \emph{et al.} \cite{callegari2011novel} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of normal and anomalies according to principal components (normal) and remaining (anomalies).
+Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
+Vieira \emph{et al.} \cite{vieira2017model} proposed a framework that applies Model Order Selection (MOS) for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
 
-Benson \emph{et al} \cite{benson2010network} conducted an empirical study of the network traffic in 10 data centers belonging to three different types of organizations, including university, enterprise, and cloud data centers.
-Findings of Benson \emph{et al} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
-This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
-Therefore, learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
-Skewness and the kurtosis are very sensitive outlier detectors, considering that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
+Existing approaches [22] added a separated interpretation step to solve this problem by using techniques such as decision trees.
+However such indirect interpretation often failed to reveal the true causes of the anomalies detected by PCA-based methods [17].
+Another recent work [11] proposed the joint sparse PCA (JSPCA) model to identify a low-dimensional approximation of the abnormal subspace, so that all anomalies can be localized onto a small subset of original feature variables.
+However, for individual anomaly interpretation, especially anomalies of different types, we need a more accurate and direct way of interpretation.
 
+% Anomaly detection by subspace larning
+Principal Component Analysis (PCA) is a statistical technique commonly used for dimensionality reduction.
+It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
+PCA has been used in attack detection \cite{almotairi2009technique}.
+However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques.
+
+Use to justify the sensitivity of pca to outliers, only.
+Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
+
+Besides being prone to higher errors and false positives, such human intervention makes PCA not useful for real time applications.
+Therefore, in order to automate the analysis of eigenvalues profile, model order selection (MOS) schemes should be incorporated.
+Vieira ....
+(The proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic.
+The proposed solution is automatic and blind for detection of time frames under probe and flood attacks through MOS and eigen analysis.
+Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.)
+
+% Anomaly detection by robust subspace larning
 In contrast to classical PCA, our robust PCA detector does not require having a perfect ground-truth for training; it just trains from the (eventually contaminated) background traffic.
@@
 X = L + S.
-
-Principal Component Analysis (PCA) is a statistical technique commonly used for dimensionality reduction.
-It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
-PCA has been used in attack detection \cite{almotairi2009technique}.
-However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques.
-Callegari \emph{et al} \cite{callegari2011novel} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a data set with synthetic anomalies added in the data.
-However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
-Callegari separate normal of anomalies according to principal components (normal) and remaining (anomalies), based on PCA.
-It relies in the assumption that anomalies will cause a large change into anomalous components.
-This is similar to our previous work, but we rely on change into principal eigenvectors.
-
-Use to justify the sensitivity of pca to outliers, only.
-Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
-In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
-Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
-
-Besides being prone to higher errors and false positives, such human intervention makes PCA not useful for real time applications.
-Therefore, in order to automate the analysis of eigenvalues profile, model order selection (MOS) schemes should be incorporated.
-Vieira ....
-(The proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic.
-The proposed solution is automatic and blind for detection of time frames under probe and flood attacks through MOS and eigen analysis.
-Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.)
-
-ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
-However, ROBPCA flags many points as outlying when the original data is skewed.
-To overcome that issue, \cite{hubert2009robustskewed} propose ROBPCA-AO, which improve ROBPCA by means of an adjusted outlyingness based on robust skewness.
-In contrast to our proposal, ROBPCA and ROBPCA-AO rely on a previously defined cutoff value for classify anomalous observations, and does not require training.
-
-\cite{zhou2017anomaly} (Anomaly detection with robust deep autoencoders)
-
-Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}.
-We rely on this for our implementation
-
-GMM is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.
-
-The Mahalanobis Distance is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
-C.
-Mahalanobis in 1936 \cite{mahalanobis1936md}.
-It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\hat{S}}$ of $\boldsymbol{X}$.
 
@@
 \end{algorithm}
+
+We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for a robust subspace learning
+
+ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
+However, ROBPCA flags many points as outlying when the original data is skewed.
+To overcome that issue, \cite{hubert2009robustskewed} propose ROBPCA-AO, which improve ROBPCA by means of an adjusted outlyingness based on robust skewness.
+In contrast to our proposal, ROBPCA and ROBPCA-AO rely on a previously defined cutoff value for classify anomalous observations, and does not require training.
+
+\cite{zhou2017anomaly} (Anomaly detection with robust deep autoencoders)
+
+% feature generation and aggregation
+Lakhina \emph{et al.} \cite{lakhina2005mining} argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
+Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning.
+Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms.
+\textbf{To explain feature selection and aggregation}.
+Goldberg and Shan \cite{goldberg2015importance} argue that processing monitored signals into features is a more fruitful way to build anomaly detection systems than devising ever more complex statistical tests, based on experience and  on putblished real-life examples of anomaly detectors in use at eBay.
+
+Benson \emph{et al.} \cite{benson2010network} conducted an empirical study of the network traffic in 10 data centers belonging to three different types of organizations, including university, enterprise, and cloud data centers.
+Findings of Benson \emph{et al.} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
+This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
+Therefore, learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
+Skewness and the kurtosis are very sensitive outlier detectors, considering that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
+
+Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}.
+We rely on this for our implementation
+
+GMM is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.
+
+The Mahalanobis Distance is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
+C.
+Mahalanobis in 1936 \cite{mahalanobis1936md}.
+It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\hat{S}}$ of $\boldsymbol{X}$.
+
+\cite{garcia2014empirical} also uses a time window separation and an aggregation method.
 

### 2019-05-02 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on literature review (robust pca)

@@
 However, PCA is not precise to reveal flows, time or detailed information for each anomaly, and sometimes relies into visual analysis for pricipal component selection.
-Robust Principal Component Analysis (RPCA) \cite{candes2011robust} is an extensio of PCA that aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, which refers to outliers’ matrix.
+Robust Principal Component Analysis (RPCA) \cite{candes2011robust} is an extension of PCA that aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, which refers to outliers’ matrix.
 Therefore, robust subspace learning has receiving a growing atention of researchers aiming the development of network anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd, hubert2005robpca,hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
@@
 
+% network anomaly detectiom
+Bhuyan \emph{et al.} \cite{bhuyan2014network} provide an overview of facets of network anomaly detection, present attacks normally encountered by network intrusion detection systems, and categorize existing network anomaly detection methods and systems based on the underlying techniques.
+Ahmed \emph{et al.} \cite{ahmed2016survey} present an analysis of four major categories of anomaly detection techniques which include classification, statistical, information theory and clustering.
+Moustafa \emph{et al.} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), explains cyber-attacks and new solutions for anomaly detection, and provides a benchmark data sets for training and validating approaches for detection.
+
 % botnet detection
-A botnet is a network of compromised machines under the influence of malware (bot).
-The botnet is commandeered by a “botmaster” and utilized as “resource” or “platform” for attacks, such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration.
+A botnet is a network of bots, that are compromised machines under the influence of malware (bot).
+The botnet is commandeered by a botmaster and used as resource for attacks, such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration.
+We refer to \cite{ahmed2016survey,moustafa2019holistic} for an overview of network attacks.
 The botmaster coordenate a botnet through a command and control (C\&C) channel where bots receive commands and syncronize attacks and fraudulent activities.
@@
 
-Bhuyan \emph{et al.} \cite{acarali2016survey} provide an overview of facets of network anomaly detection, present attacks normally encountered by network intrusion detection systems, and categorize existing network anomaly detection methods and systems based on the underlying techniques.
-Ahmed Bhuyan \emph{et al.} \cite{acarali2016survey} present an analysis of four major categories of anomaly detection techniques which include classification, statistical, information theory and clustering.
-Moustafa \emph{et al.} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), explains cyber-attacks and new solutions for anomaly detection, and provides a benchmark data sets for training and validating approaches for detection.
-
-Acarali \emph{et al.} \cite{acarali2016survey} surveys network-based detection approaches for HTTP-based botnets, surveys traffic-based features used to detect bot traffic and presents an abstraction of the main types of features in relation to protocols and OSI layers.
+Acarali \emph{et al.} \cite{acarali2016survey} surveys network-based detection approaches for HTTP-based botnets, surveys traffic-based features used to detect bot traffic and presents an abstraction of the main types of features related to protocols and OSI layers.
 Wang \emph{et al.} \cite{Wang2018ddosbotnetssurvey} present an analysis based on 50,704 different Internet DDoS attacks originated of 674 botnets from 23 different botnet families with a total of 9,026 victim belonging to 1,074 organizations in 186 countries.
@@
 
-Khattak \emph{et al.} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, based on a Complex Event Processing (CEP) engine and on a correlation framework that continuously receives events and correlates them according to rules.
+Khattak \emph{et al.} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, composed by a Complex Event Processing (CEP) engine and on a correlation framework that continuously receives events and correlates them according to rules.
 BotFlex's results are compared to BotHunter \cite{gu2007bothunter}, but the evaluation relies in a own and not public data set.
@@
 NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that redundant data records are removed, in order to avoid biased classifications.
+However, NSL-KDD dataset maintain the limitations of the KDD 99 regarding volume and reproduction of network traffic similar to the real networks.
 
 The objective of Garcia \emph{et al.} \cite{garcia2014empirical} is to compare three botnet detection methods using a simple and reproducible methodology, a good data set and a new error metric.
-This paper evaluates some data sets for network anomaly detection, surveys some approaches for botnet detection and proposes two methods (BClus and CAMNEP) and compare results to BotHunter \cite{gu2007bothunter}.
+This paper evaluates some data sets for network anomaly detection, surveys some approaches for botnet detection, proposes two methods (BClus and CAMNEP) for botnet detection and compare results to BotHunter \cite{gu2007bothunter}.
 Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by normal and background labeled data, in an imbalanced distribution like in a real network.
@@
 Adopting the training and testing approach of \cite{garcia2014empirical}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
-In contrast, we evaluate each scenario of CTU-13 individually in a semi-supervised approach that does not rely on training data with anomalies.
+In contrast, we propose to evaluate each scenario of CTU-13 individually, in a semi-supervised approach that does not rely on training data with anomalies.
 
@@
 
+% Anomaly detection by subspace larning
 Traditional PCA-based anomaly detection models are not suitable for anomaly interpretation, as they judge whether a data instance is an anomaly or not based on the length of its projection on the abnormal subspace spanned by the less significant principal components, and there is no direct mapping between PCA’s dimensionality-reduced subspace and the original feature space \cite{ringberg2007sensitivity}.
-Callegari \emph{et al.} \cite{callegari2011novel} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of normal and anomalies according to principal components (normal) and remaining (anomalies).
-Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
-Vieira \emph{et al.} \cite{vieira2017model} proposed a framework that applies Model Order Selection (MOS) for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
-
-Existing approaches [22] added a separated interpretation step to solve this problem by using techniques such as decision trees.
-However such indirect interpretation often failed to reveal the true causes of the anomalies detected by PCA-based methods [17].
-Another recent work [11] proposed the joint sparse PCA (JSPCA) model to identify a low-dimensional approximation of the abnormal subspace, so that all anomalies can be localized onto a small subset of original feature variables.
-However, for individual anomaly interpretation, especially anomalies of different types, we need a more accurate and direct way of interpretation.
-
-% Anomaly detection by subspace larning
-Principal Component Analysis (PCA) is a statistical technique commonly used for dimensionality reduction.
-It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
-PCA has been used in attack detection \cite{almotairi2009technique}.
-However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques.
-
-Use to justify the sensitivity of pca to outliers, only.
-Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
-
-Besides being prone to higher errors and false positives, such human intervention makes PCA not useful for real time applications.
-Therefore, in order to automate the analysis of eigenvalues profile, model order selection (MOS) schemes should be incorporated.
-Vieira ....
-(The proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic.
-The proposed solution is automatic and blind for detection of time frames under probe and flood attacks through MOS and eigen analysis.
-Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.)
+Callegari \emph{et al.} \cite{callegari2011novel} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of normal and anomalies according to principal components (normal) and remaining (anomalies).
+Lee \emph{et al.} \cite{Lee2013} presented OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
+Vieira \emph{et al.} \cite{vieira2017model} proposed a framework that applies Model Order Selection (MOS) for detection of time frames under attack and uses similarity analysis to extract details and detect the time and ports under attack.
 
 % Anomaly detection by robust subspace larning
-In contrast to classical PCA, our robust PCA detector does not require having a perfect ground-truth for training; it just trains from the (eventually contaminated) background traffic.
+Robust subspace learning has receiving a growing atention of researchers aiming the development of anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
+The problem of PCA or subspace learning for outlier corrupted data is called robust PCA (RPCA) or robust subspace learning \cite{vaswani2018robust}, which can be defined as the problem of decomposing a given data matrix $\textbf{M}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
+This definition is also referred to as the sparse + low rank (S + LR) formulation \cite{vaswani2018robust}.
+We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for robust a subspace learning review.
+
+RPCA is been mainly applied to computer vision, in problems of robust subspace tracking and robust subspace recovery.
+However, RPCA has also been adoted for general outlier detection \cite{hubert2005robpca,hubert2009robustskewed,cherapanamjeri2017thresholding,zhou2017anomaly,NetflixSurus} and for anomaly detection on network traffic \cite{pascoal2012robust}.
+
+ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
+However, ROBPCA flags many points as outlying when the original data is skewed.
+To overcome that issue, \cite{hubert2009robustskewed} propose ROBPCA-AO, which improve ROBPCA by means of an adjusted outlyingness based on robust skewness.
+In contrast to our proposal, ROBPCA and ROBPCA-AO rely on a previously defined cutoff value for classify anomalous observations, and does not require training.
+
+Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on robust feature selection and robust PCA for network traffic anomaly detection.
+It is an unsupervised learning algorithm that allows matching the specificities of the traffic the detector is operating on, which can vary widely both in terms of the most frequent attacks and the relative weight of licit applications.
+
+Our robust PCA detector does not require having a perfect ground-truth for training; it just trains from the (eventually contaminated) background traffic.
 It is, in fact, an unsupervised learning algorithm.
 Moreover, the addition of feature selection allows matching the specificities of the traffic the detector is operating on, which can vary widely both in terms of the most frequent attacks and the relative weight of licit applications.
-Feature selection does require training with a labeled data set bu
+Feature selection does require training with a labeled data set bu.
+One important strength of our anomaly detector is its adaptability to different network environments and traffic conditions
 
@@
 X = L + S.
+
+Zhou and Paffenroth \cite{zhou2017anomaly} proposed the Robust Deep Autoencoders (RDA), which the central idea is that a RDA inherits the non-linear representation capabilities of autoencoders combined with the anomaly detection capabilities of RPCA.
+Considering that outliers and noise may reduce the quality of representations discovered by deep autoencoders, the proposed model isolates noise and outliers in the input by means of a RPCA approach, and the autoencoder is trained after this isolation.
+RDA was evaluated by the authors for the MNIST data set.
 
@@
 \end{algorithm}
-
-We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for a robust subspace learning
-
-ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
-However, ROBPCA flags many points as outlying when the original data is skewed.
-To overcome that issue, \cite{hubert2009robustskewed} propose ROBPCA-AO, which improve ROBPCA by means of an adjusted outlyingness based on robust skewness.
-In contrast to our proposal, ROBPCA and ROBPCA-AO rely on a previously defined cutoff value for classify anomalous observations, and does not require training.
-
-\cite{zhou2017anomaly} (Anomaly detection with robust deep autoencoders)
 

### 2019-05-02 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on literature review (robust pca)

@@
 Traditional PCA-based anomaly detection models are not suitable for anomaly interpretation, as they judge whether a data instance is an anomaly or not based on the length of its projection on the abnormal subspace spanned by the less significant principal components, and there is no direct mapping between PCA’s dimensionality-reduced subspace and the original feature space \cite{ringberg2007sensitivity}.
+However, to overcome the above mentioned limitations, some approaches based on PCA have been proposed for anomaly detection.
 Callegari \emph{et al.} \cite{callegari2011novel} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of normal and anomalies according to principal components (normal) and remaining (anomalies).
@@
 Robust subspace learning has receiving a growing atention of researchers aiming the development of anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
-The problem of PCA or subspace learning for outlier corrupted data is called robust PCA (RPCA) or robust subspace learning \cite{vaswani2018robust}, which can be defined as the problem of decomposing a given data matrix $\textbf{M}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
+The problem of PCA or subspace learning for outlier corrupted data is called robust PCA (RPCA) or robust subspace learning \cite{vaswani2018robust}, which can be defined as the decomposition of a given data matrix $\textbf{M}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
 This definition is also referred to as the sparse + low rank (S + LR) formulation \cite{vaswani2018robust}.
-We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for robust a subspace learning review.
+We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for more details regarding robust subspace learning.
 
@@
 However, RPCA has also been adoted for general outlier detection \cite{hubert2005robpca,hubert2009robustskewed,cherapanamjeri2017thresholding,zhou2017anomaly,NetflixSurus} and for anomaly detection on network traffic \cite{pascoal2012robust}.
-
 ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
 However, ROBPCA flags many points as outlying when the original data is skewed.
-To overcome that issue, \cite{hubert2009robustskewed} propose ROBPCA-AO, which improve ROBPCA by means of an adjusted outlyingness based on robust skewness.
+Therefore, Hubert \emph{et al.} \cite{hubert2009robustskewed} proposed ROBPCA-AO, which improves ROBPCA for problems with skewed data, by means of an adjusted outlyingness based on robust skewness.
+Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on a robust mutual information estimator for feature selection and based on RPCA for outlier detection in internet traffic.
+The anomaly detection proposed by Pascoal \emph{et al.} is an unsupervised approach that estimate the first $k$ robust principal components, calculate the score distance and the orthogonal distance, calculate the thresholds and classify new observations accordingly.
+
 In contrast to our proposal, ROBPCA and ROBPCA-AO rely on a previously defined cutoff value for classify anomalous observations, and does not require training.
-
-Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on robust feature selection and robust PCA for network traffic anomaly detection.
-It is an unsupervised learning algorithm that allows matching the specificities of the traffic the detector is operating on, which can vary widely both in terms of the most frequent attacks and the relative weight of licit applications.
 

### 2019-05-02 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
literature review OK

@@
 We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data.
-The botnet detection algorithms can also be improved by feature generation from aggregated traffic, considering that new features can make the data more discriminative for imbalanced learning problems.
+The botnet detection algorithms can also be improved by feature engineering from aggregated traffic, considering that new features can make the data more discriminative for imbalanced learning problems.
 We also believe that the distance between robust estimates of normal traffic and anomalies can highlight discrepancies and be used for botnet detection.
@@
 
-We adopt a sampling method based on data aggregation for discriminative feature generation, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
+We adopt a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
 We evaluate the accuracy of the proposed approach for botnet detecion on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
@@
 Adopting the training and testing approach of \cite{garcia2014empirical}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
-In contrast, we propose to evaluate each scenario of CTU-13 individually, in a semi-supervised approach that does not rely on training data with anomalies.
 
@@
 Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on a robust mutual information estimator for feature selection and based on RPCA for outlier detection in internet traffic.
-The anomaly detection proposed by Pascoal \emph{et al.} is an unsupervised approach that estimate the first $k$ robust principal components, calculate the score distance and the orthogonal distance, calculate the thresholds and classify new observations accordingly.
-
-In contrast to our proposal, ROBPCA and ROBPCA-AO rely on a previously defined cutoff value for classify anomalous observations, and does not require training.
-
-Our robust PCA detector does not require having a perfect ground-truth for training; it just trains from the (eventually contaminated) background traffic.
-It is, in fact, an unsupervised learning algorithm.
-Moreover, the addition of feature selection allows matching the specificities of the traffic the detector is operating on, which can vary widely both in terms of the most frequent attacks and the relative weight of licit applications.
-Feature selection does require training with a labeled data set bu.
-One important strength of our anomaly detector is its adaptability to different network environments and traffic conditions
-
-aims to reduce the sensitivity of PCA to outliers, by means of .
-In particular, RPCA allows for the careful teasing apart of sparse outliers so that the remaining low-dimensional approximation is faithful to the noise-free low-dimensional subspace describing the bulk of the raw data [6, 7, 20].
-Specifically, RPCA splits a data matrix X into a low-rank matrix L and a sparse matrix S such that
-X = L + S.
+The anomaly detection proposed by Pascoal \emph{et al.} is an unsupervised approach that estimate the first $k$ robust principal components, calculate the score and the orthogonal distances, calculate the thresholds and classify new observations accordingly.
 
@@
 
-Mahalanobis Distance is defined as		
-	\begin{equation}\label{eq:eq01}
-		d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}.
-	\end{equation}
-$\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector, also referred as location, of known observations and $\boldsymbol{\hat{S}}$ is the covariance matrix, also referred as scatter, of known observations.
-
-Classical estimates can be so strongly affected by contamination that diagnostic tools, such as the Mahalanobis Distances, become unable to detect the outliers.
-Therefore, we need reliable and robust estimators that can resist outliers contaminated data.
-Minimum Covariance Determinant (MCD) \cite{rousseeuw1984mcd} is a robust estimator commonly used for computing Mahalanobis Distance in order to detect anomalies.
-
-\begin{figure}[h!]
-     \centering
-     \includegraphics[width=11cm]{figures/mahalanobis_robust.png}
-     \caption{Mahalanobis Distance for non-robust and robust estimates.}
-     \label{fig:fig04}
-\end{figure}
-
-\begin{figure}[h!]
-     \centering
-     \includegraphics[width=8cm]{figures/mahalanobis_distances01.png}
-     \caption{Mahalanobis Distance for non-robust and robust estimates.}
-     \label{fig:fig05}
-\end{figure}
-
-The MCD estimator is one of the first affine equivariant and highly robust estimators of multivariate location and scatter.
-
-Affine equivariant means that when the data are translated or subjected to a linear transformation, the resulting location and scatter will transform accordingly \cite{rousseeuw1984mcd,rousseeuw1999fastmcd}.
-
-Its objective is to find $h$ observations (out of $n$) whose covariance matrix has the lowest determinant.
-
-The lowest determinant means the lowest distance generalized variance and high distance between the largest eigenvalue and the remaining [ref?].
-
-MCD depends on random sampling, therefore it is not deterministic.
-
-MCD has been applied in numerous fields (medicine, finance, image analysis, and chemistry) and been used to develop many robust multivariate techniques (such as Robust Principal Component Analysis (RPCA), factor analysis, and multiple regression).
-
-The classical MCD has rarely been applied because it is hard to compute, since it requires the evaluation of all $\binom{n}{h}$ subsets of size $h$. % binomial of n over p
-
-Its main use has started since the construction of the computationally efficient Fast-MCD algorithm \cite{rousseeuw1999fastmcd}.
-
-Fast-MCD is an iterative, approximate and resampling algorithm for the MCD.
-
-It starts by random initial subset of size $p+1$ and performs concentration steps (C-steps) yielding consecutive h-subsets with decreasing covariance matrix determinant.
-
-Only two C-steps are applied to each initial subset, and the 10 results with lowest determinant are kept.
-
-C-steps are carried out until convergence and the best solution is kept.
-
-Fast-MCD also is affine equivariant and not deterministic.
-
-\textbf{Theorem 1: C-Step.}
-
-Consider a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ where $\boldsymbol{X} = \{\boldsymbol{x}_1,...,\boldsymbol{x}_n\}$ of $p$-variate observations.
-Let $\boldsymbol{h} \subset \{1,...,n\}$ with $|\boldsymbol{h}| = h$, where $\boldsymbol{h}$ is a subset of $h$ indexes from $n$ observations.
-
-Let the mean $\bar{\boldsymbol{x}} \in \mathbb{R}^{1 \times p}$ be
-
-\begin{equation}\label{eq:eq02}
-	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} \boldsymbol{x}_j, 
-\end{equation}
-
-the covariance matrix $\boldsymbol{\hat{S}} \in \mathbb{R}^{p \times p}$ be
-
-\begin{equation}\label{eq:eq03}
-	\boldsymbol{\hat{S}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} (\boldsymbol{x}_j - \bar{\boldsymbol{x}})(\boldsymbol{x}_j - \bar{\boldsymbol{x}})^\prime,
-\end{equation}
-
-and the relative distance be
-
-\begin{equation}\label{eq:eq04}
-	\boldsymbol{d}(i) = \sqrt{(\boldsymbol{x}_i - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime}, 
-\end{equation}
-
-for $i = 1,...,n$.
-
-Let $\boldsymbol{h}_1, \bar{\boldsymbol{x}}_1$ and $\boldsymbol{\hat{S}}_1$ and be the selected observations, the location and scatter of the initial subset, and $\boldsymbol{d}_1(i)$ be the initial distances for $i = 1,...,n$.
-
-Now take $\boldsymbol{h}_2$ containing the indexes of lowest distances, such that $\{\boldsymbol{d}_1(i); i \in \boldsymbol{h}_2\} = \{(\boldsymbol{d}_1)_{1:n},...,(\boldsymbol{d}_1)_{h:n}\}$, where $(\boldsymbol{d}_1)_{1:n} \leq (\boldsymbol{d}_1)_{2:n} \leq ...
-\leq (\boldsymbol{d}_1)_{n:n}$ are the ordered distances, and compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$.
-
-Then, $det(\boldsymbol{\hat{S}}_2) \leq det(\boldsymbol{\hat{S}}_1)$ with equality if and only if $\boldsymbol{\bar{x}}_2 = \boldsymbol{\bar{x}}_1$ and $\boldsymbol{\bar{S}}_2 = \boldsymbol{\hat{S}}_1$
-
-The proof is given by Rousseeuw and Driessen \cite{rousseeuw1999fastmcd}.
-
-\textbf{The C-step can be algorithmically described as follows:}
-\begin{algorithm}
-	\label{alg:alg01}
-	\scriptsize
-	\SetAlgoLined
-	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
-	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
-	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
-		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
-			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
-		}
-		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
-		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
-		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
-		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
-		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
-	}
-	\caption{C-Step}
-\end{algorithm}
-
-\textbf{Theorem 1} thus provides a partial idea for an algorithm:
-
-\textit{Take many initial choices of $\boldsymbol{h}_1$ and apply C-steps to each until convergence (don't have determinant decreasing), and keep the result with lowest determinant}.
-
-\textbf{Initial subset can be algorithmically described as:}
-\begin{algorithm}
-	\label{alg:alg02}
-	\scriptsize
-	\SetAlgoLined
-	\KwResult{$\boldsymbol{h}_1$}
-	Given a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $p$-variate observations\;
-	Draw a random $(p + 1)$-subset $\boldsymbol{L}$\;
-	Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
-	\If{$det(\boldsymbol{\hat{S}}_1) \leq 0$}{
-		Add another random observation into $\boldsymbol{L}$\;
-		Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
-	}
-	Compute $\boldsymbol{d}_1^2(i) = (\boldsymbol{x}_i - \boldsymbol{\bar{x}}_1)^\prime \boldsymbol{\hat{S}}_1^{-1}(\boldsymbol{X}_i - \boldsymbol{\bar{x}}_0)$ for $i = 1, ..., n$\;
-	Sort them into $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_0(\pi(n))$\;
-	Put $\boldsymbol{h}_1 = \{\pi(1), ..., \pi(h)\}$\;
-	\caption{Constructing the initial subset}
-\end{algorithm}
-
-\begin{algorithm}
-	\label{alg:alg03}
-	\scriptsize
-	\SetAlgoLined
-	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
-	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n \leq 600$\;
-	\For{500 times}{
-		Construct initial $h$-subset $\boldsymbol{h}_1$ (Algorithm \ref{alg:alg02})\;
-		Carry out two C-steps (Algorithm \ref{alg:alg01}) resulting $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$\;
-	}		
-	Put $\boldsymbol{h}_1$ as the 10 lowest $det(\boldsymbol{\hat{S}_2})$ computed above\;
-	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
-		Carry out C-steps (Algorithm \ref{alg:alg01})\;
-	}
-	\caption{Fast-MCD when $n \leq 600$}
-\end{algorithm}
-
-\begin{algorithm}
-	\label{alg:alg04}
-	\scriptsize
-	\SetAlgoLined
-	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
-	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n > 600$\;
-	Construct up to 5 disjoint random subsets of size $n_{sub} = 300$\;
-	\For{each subset, repeat 100 times}{
-		Construct initial $\boldsymbol{h}_1$ of size $h_{sub} = [n_{sub}(h/n)]$\;
-		Carry out two C-steps of $n_{sub}$ and $h_{sub}$\;
-		Keep the 10 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$\;
-	}		
-	Pool the subsets, yielding the merged set of $n_{merged} = 1500$\;
-	\For{50 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$}{
-		Carry out two C-steps of $n_{merged}$ and $h_{merged} = [n_{merged}(h/n)]$\;
-		Keep the 10 best $(\boldsymbol{\bar{x}}_{merged}, \boldsymbol{\hat{S}}_{merged})$\;
-	}
-	\For{the full data set and $m_{full}$ best results}{
-		Take several C-steps, using $n$ and $h$\;
-		Keep the best final result $(\boldsymbol{\bar{x}}_{full}, \boldsymbol{\hat{S}}_{full})$\;
-	}
-	\caption{Fast-MCD when $n > 600$}
-\end{algorithm}
-
-% feature generation and aggregation
+% feature engineering and aggregation
 Lakhina \emph{et al.} \cite{lakhina2005mining} argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
-Using entropy as a summarization tool, we show that the analysis of feature distributions leads to significant advances on two fronts: (1) it enables highly sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and (2) it enables automatic classification of anomalies via unsupervised learning.
-Anomaly diagnosis framework, comprising both an extension of the subspace method [23] to accommodate multiple data types, and an unsupervised classification technique using simple clustering algorithms.
-\textbf{To explain feature selection and aggregation}.
+Lakhina \emph{et al.} show that the analysis of feature distributions leads to significant advances on sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and leads gains on automatic classification of anomalies via unsupervised learning.
 Goldberg and Shan \cite{goldberg2015importance} argue that processing monitored signals into features is a more fruitful way to build anomaly detection systems than devising ever more complex statistical tests, based on experience and on putblished real-life examples of anomaly detectors in use at eBay.
@@
 This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
-Therefore, learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
-Skewness and the kurtosis are very sensitive outlier detectors, considering that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
+Considering that learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}, higher moments have been adopted for outlier detection and discriminative analysis.
+Skewness and the kurtosis are higher moments that can contribute to sensitive outlier detectors, taking into account that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
 
-Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}.
-We rely on this for our implementation
+We propose an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggretation from the CTU-13 data set.
+The proposed approach relies on a robust subspace of normal traffic for estimating higher moments.
+The anomaly detection for new observations evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervisioned fashion and without the computational cost of new robust subspace learning for new observations.
 
-GMM is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.
-
-The Mahalanobis Distance is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
-C.
-Mahalanobis in 1936 \cite{mahalanobis1936md}.
-It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\hat{S}}$ of $\boldsymbol{X}$.
-
-\cite{garcia2014empirical} also uses a time window separation and an aggregation method.
-
-Our proposal is ...
+We adopt a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
+Finally, We evaluate the accuracy of the proposed approach for botnet detecion on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
 
@@
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
-We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matrix, and the operator $\sim$ that sorts the elements of a vector in ascending order.
 
-This section also presents a description of the data model and data aggregation of the CTU-13 data set, in Subsection \ref{sec:CTU-13} discusses the use of the CTU-13 data set for evaluation of the proposed approach.
+This section also presents a description of the data model in Subsection \ref{sec:CTU-13} and in Subsection \ref{sec:feat_eng} we describe the data aggregation and feature engineering adopted for the CTU-13.
 
@@
 Table \ref{tab:tab01} also shows the ratio of C\&C and botnet flows.
+
+\subsection{feature engineering}
+\label{sec:feat_eng}
 
@@
 
+Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statisctical approaches, which are K-means and Gaussian Mixture Model (GMM) \cite{gaddam2007kmeans,moustafa2019holistic}, respectively.
+We also evaluate the results of ROBPCA \cite{hubert2005robpca}, wich is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
+In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of CTU-13 individually, in a semi-supervised approach that does not rely on training data with anomalies.
+
 This section presents the performed experiments and the acquired results.
@@
 Given a contaminated data set, split it into normal (60\%), cross validation (20\% normal + 50\% anomalous) and test (20\% normal + 50\% anomalous).
+
 Select the range of contamination that shall be evaluated and perform cross validation for each contamination.
+
 For each contamination $c$:
@@
 Estimate $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$ from normal data.
+
 Predict anomalies for cross validation data using the $c$, $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$.
+
 Select the best F-measure by cross validation and get the best $c$.
@@
 Test the F-measure of anomaly detection for new observations through $d(\boldsymbol{x},\bar{\boldsymbol{x}}_0, \boldsymbol{\hat{S}}_0)$ and $c$.
+
+Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}.
+We rely on this for our implementation
+
+GMM is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.
+
+The Mahalanobis Distance is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
+C.
+Mahalanobis in 1936 \cite{mahalanobis1936md}.
+It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\hat{S}}$ of $\boldsymbol{X}$.
+
+\cite{garcia2014empirical} also uses a time window separation and an aggregation method.
 
@@
 \end{document}
+
+% Mahalanobis Distance is defined as		
+% 	\begin{equation}\label{eq:eq01}
+% 		d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}.
+% 	\end{equation}
+% $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector, also referred as location, of known observations and $\boldsymbol{\hat{S}}$ is the covariance matrix, also referred as scatter, of known observations.
+
+% Classical estimates can be so strongly affected by contamination that diagnostic tools, such as the Mahalanobis Distances, become unable to detect the outliers.
+Therefore, we need reliable and robust estimators that can resist outliers contaminated data.
+Minimum Covariance Determinant (MCD) \cite{rousseeuw1984mcd} is a robust estimator commonly used for computing Mahalanobis Distance in order to detect anomalies.
+
+% \begin{figure}[h!]
+%      \centering
+%      \includegraphics[width=11cm]{figures/mahalanobis_robust.png}
+%      \caption{Mahalanobis Distance for non-robust and robust estimates.}
+%      \label{fig:fig04}
+% \end{figure}
+
+% \begin{figure}[h!]
+%      \centering
+%      \includegraphics[width=8cm]{figures/mahalanobis_distances01.png}
+%      \caption{Mahalanobis Distance for non-robust and robust estimates.}
+%      \label{fig:fig05}
+% \end{figure}
+
+% The MCD estimator is one of the first affine equivariant and highly robust estimators of multivariate location and scatter.
+
+% Affine equivariant means that when the data are translated or subjected to a linear transformation, the resulting location and scatter will transform accordingly \cite{rousseeuw1984mcd,rousseeuw1999fastmcd}.
+
+% Its objective is to find $h$ observations (out of $n$) whose covariance matrix has the lowest determinant.
+
+% The lowest determinant means the lowest distance generalized variance and high distance between the largest eigenvalue and the remaining [ref?].
+
+% MCD depends on random sampling, therefore it is not deterministic.
+
+% MCD has been applied in numerous fields (medicine, finance, image analysis, and chemistry) and been used to develop many robust multivariate techniques (such as Robust Principal Component Analysis (RPCA), factor analysis, and multiple regression).
+
+% The classical MCD has rarely been applied because it is hard to compute, since it requires the evaluation of all $\binom{n}{h}$ subsets of size $h$. % binomial of n over p
+
+% Its main use has started since the construction of the computationally efficient Fast-MCD algorithm \cite{rousseeuw1999fastmcd}.
+
+% Fast-MCD is an iterative, approximate and resampling algorithm for the MCD.
+
+% It starts by random initial subset of size $p+1$ and performs concentration steps (C-steps) yielding consecutive h-subsets with decreasing covariance matrix determinant.
+
+% Only two C-steps are applied to each initial subset, and the 10 results with lowest determinant are kept.
+
+% C-steps are carried out until convergence and the best solution is kept.
+
+% Fast-MCD also is affine equivariant and not deterministic.
+
+% \textbf{Theorem 1: C-Step.}
+
+% Consider a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ where $\boldsymbol{X} = \{\boldsymbol{x}_1,...,\boldsymbol{x}_n\}$ of $p$-variate observations.
+Let $\boldsymbol{h} \subset \{1,...,n\}$ with $|\boldsymbol{h}| = h$, where $\boldsymbol{h}$ is a subset of $h$ indexes from $n$ observations.
+
+% Let the mean $\bar{\boldsymbol{x}} \in \mathbb{R}^{1 \times p}$ be
+
+% \begin{equation}\label{eq:eq02}
+% 	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} \boldsymbol{x}_j, 
+% \end{equation}
+
+% the covariance matrix $\boldsymbol{\hat{S}} \in \mathbb{R}^{p \times p}$ be
+
+% \begin{equation}\label{eq:eq03}
+% 	\boldsymbol{\hat{S}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} (\boldsymbol{x}_j - \bar{\boldsymbol{x}})(\boldsymbol{x}_j - \bar{\boldsymbol{x}})^\prime,
+% \end{equation}
+
+% and the relative distance be
+
+% \begin{equation}\label{eq:eq04}
+% 	\boldsymbol{d}(i) = \sqrt{(\boldsymbol{x}_i - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime}, 
+% \end{equation}
+
+% for $i = 1,...,n$.
+
+% Let $\boldsymbol{h}_1, \bar{\boldsymbol{x}}_1$ and $\boldsymbol{\hat{S}}_1$ and be the selected observations, the location and scatter of the initial subset, and $\boldsymbol{d}_1(i)$ be the initial distances for $i = 1,...,n$.
+
+% Now take $\boldsymbol{h}_2$ containing the indexes of lowest distances, such that $\{\boldsymbol{d}_1(i); i \in \boldsymbol{h}_2\} = \{(\boldsymbol{d}_1)_{1:n},...,(\boldsymbol{d}_1)_{h:n}\}$, where $(\boldsymbol{d}_1)_{1:n} \leq (\boldsymbol{d}_1)_{2:n} \leq ...
+\leq (\boldsymbol{d}_1)_{n:n}$ are the ordered distances, and compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$.
+
+% Then, $det(\boldsymbol{\hat{S}}_2) \leq det(\boldsymbol{\hat{S}}_1)$ with equality if and only if $\boldsymbol{\bar{x}}_2 = \boldsymbol{\bar{x}}_1$ and $\boldsymbol{\bar{S}}_2 = \boldsymbol{\hat{S}}_1$
+
+% The proof is given by Rousseeuw and Driessen \cite{rousseeuw1999fastmcd}.
+
+% \textbf{The C-step can be algorithmically described as follows:}
+% \begin{algorithm}
+% 	\label{alg:alg01}
+% 	\scriptsize
+% 	\SetAlgoLined
+% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
+% 	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
+% 	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
+% 		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
+% 			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
+% 		}
+% 		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
+% 		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
+% 		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
+% 		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
+% 		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
+% 	}
+% 	\caption{C-Step}
+% \end{algorithm}
+
+% \textbf{Theorem 1} thus provides a partial idea for an algorithm:
+
+% \textit{Take many initial choices of $\boldsymbol{h}_1$ and apply C-steps to each until convergence (don't have determinant decreasing), and keep the result with lowest determinant}.
+
+% \textbf{Initial subset can be algorithmically described as:}
+% \begin{algorithm}
+% 	\label{alg:alg02}
+% 	\scriptsize
+% 	\SetAlgoLined
+% 	\KwResult{$\boldsymbol{h}_1$}
+% 	Given a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $p$-variate observations\;
+% 	Draw a random $(p + 1)$-subset $\boldsymbol{L}$\;
+% 	Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
+% 	\If{$det(\boldsymbol{\hat{S}}_1) \leq 0$}{
+% 		Add another random observation into $\boldsymbol{L}$\;
+% 		Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
+% 	}
+% 	Compute $\boldsymbol{d}_1^2(i) = (\boldsymbol{x}_i - \boldsymbol{\bar{x}}_1)^\prime \boldsymbol{\hat{S}}_1^{-1}(\boldsymbol{X}_i - \boldsymbol{\bar{x}}_0)$ for $i = 1, ..., n$\;
+% 	Sort them into $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_0(\pi(n))$\;
+% 	Put $\boldsymbol{h}_1 = \{\pi(1), ..., \pi(h)\}$\;
+% 	\caption{Constructing the initial subset}
+% \end{algorithm}
+
+% \begin{algorithm}
+% 	\label{alg:alg03}
+% 	\scriptsize
+% 	\SetAlgoLined
+% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
+% 	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n \leq 600$\;
+% 	\For{500 times}{
+% 		Construct initial $h$-subset $\boldsymbol{h}_1$ (Algorithm \ref{alg:alg02})\;
+% 		Carry out two C-steps (Algorithm \ref{alg:alg01}) resulting $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$\;
+% 	}		
+% 	Put $\boldsymbol{h}_1$ as the 10 lowest $det(\boldsymbol{\hat{S}_2})$ computed above\;
+% 	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
+% 		Carry out C-steps (Algorithm \ref{alg:alg01})\;
+% 	}
+% 	\caption{Fast-MCD when $n \leq 600$}
+% \end{algorithm}
+
+% \begin{algorithm}
+% 	\label{alg:alg04}
+% 	\scriptsize
+% 	\SetAlgoLined
+% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
+% 	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n > 600$\;
+% 	Construct up to 5 disjoint random subsets of size $n_{sub} = 300$\;
+% 	\For{each subset, repeat 100 times}{
+% 		Construct initial $\boldsymbol{h}_1$ of size $h_{sub} = [n_{sub}(h/n)]$\;
+% 		Carry out two C-steps of $n_{sub}$ and $h_{sub}$\;
+% 		Keep the 10 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$\;
+% 	}		
+% 	Pool the subsets, yielding the merged set of $n_{merged} = 1500$\;
+% 	\For{50 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$}{
+% 		Carry out two C-steps of $n_{merged}$ and $h_{merged} = [n_{merged}(h/n)]$\;
+% 		Keep the 10 best $(\boldsymbol{\bar{x}}_{merged}, \boldsymbol{\hat{S}}_{merged})$\;
+% 	}
+% 	\For{the full data set and $m_{full}$ best results}{
+% 		Take several C-steps, using $n$ and $h$\;
+% 		Keep the best final result $(\boldsymbol{\bar{x}}_{full}, \boldsymbol{\hat{S}}_{full})$\;
+% 	}
+% 	\caption{Fast-MCD when $n > 600$}
+% \end{algorithm}

### 2019-05-03 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on data model

@@
 Wang \emph{et al.} \cite{Wang2018ddosbotnetssurvey} present an analysis based on 50,704 different Internet DDoS attacks originated of 674 botnets from 23 different botnet families with a total of 9,026 victim belonging to 1,074 organizations in 186 countries.
-Their analysis reveals that geolocation of the attacking sources follows patterns and enables source prediction, and highlight that multiple attacks to the same target also exhibit strong patterns of inter-attack time interval, also presents that there is a trend for different botnets to launch DDoS attacks targeting the same victim, simultaneously or in turn.
+Their analysis reveals that geolocation of the attacking sources follows patterns and enables source prediction, and highlights that multiple attacks to the same target also exhibit strong patterns of inter-attack time interval, also presents that there is a trend for different botnets to launch DDoS attacks targeting the same victim, simultaneously or in turn.
 
@@
 
-Wang and Paschalidis \cite{wang2017botnet} propose a botnet detection approach based on anomaly and community detection, aiming for detecting botnets and identifying bots before the botnet becomes active.
+Wang and Paschalidis \cite{wang2017botnet} proposed a botnet detection approach based on anomaly and community detection, aiming for detecting botnets and identifying bots before the botnet becomes active.
 The first stage detects anomalies by leveraging large deviations of an empirical distribution.
@@
 Traditional PCA-based anomaly detection models are not suitable for anomaly interpretation, as they judge whether a data instance is an anomaly or not based on the length of its projection on the abnormal subspace spanned by the less significant principal components, and there is no direct mapping between PCA’s dimensionality-reduced subspace and the original feature space \cite{ringberg2007sensitivity}.
-However, to overcome the above mentioned limitations, some approaches based on PCA have been proposed for anomaly detection.
-Callegari \emph{et al.} \cite{callegari2011novel} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of normal and anomalies according to principal components (normal) and remaining (anomalies).
+However, to overcome the above mentioned limitations, some approaches based on PCA have been proposed for network anomaly detection.
+Callegari \emph{et al.} \cite{callegari2011novel} proposed a PCA-based method for identifying the network traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of normal and anomalies according to principal components (normal) and remaining (anomalies).
 Lee \emph{et al.} \cite{Lee2013} presented OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
@@
 % Anomaly detection by robust subspace larning
-Robust subspace learning has receiving a growing atention of researchers aiming the development of anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
 The problem of PCA or subspace learning for outlier corrupted data is called robust PCA (RPCA) or robust subspace learning \cite{vaswani2018robust}, which can be defined as the decomposition of a given data matrix $\textbf{M}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
@@
 We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for more details regarding robust subspace learning.
+Robust subspace learning has receiving a growing atention of researchers aiming the development of anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
 
@@
 Therefore, Hubert \emph{et al.} \cite{hubert2009robustskewed} proposed ROBPCA-AO, which improves ROBPCA for problems with skewed data, by means of an adjusted outlyingness based on robust skewness.
+
 Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on a robust mutual information estimator for feature selection and based on RPCA for outlier detection in internet traffic.
 The anomaly detection proposed by Pascoal \emph{et al.} is an unsupervised approach that estimate the first $k$ robust principal components, calculate the score and the orthogonal distances, calculate the thresholds and classify new observations accordingly.
-
 	Zhou and Paffenroth \cite{zhou2017anomaly} proposed the Robust Deep Autoencoders (RDA), which the central idea is that a RDA inherits the non-linear representation capabilities of autoencoders combined with the anomaly detection capabilities of RPCA.
@@
 Lakhina \emph{et al.} \cite{lakhina2005mining} argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
-Lakhina \emph{et al.} show that the analysis of feature distributions leads to significant advances on sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, and leads gains on automatic classification of anomalies via unsupervised learning.
-Goldberg and Shan \cite{goldberg2015importance} argue that processing monitored signals into features is a more fruitful way to build anomaly detection systems than devising ever more complex statistical tests, based on experience and on putblished real-life examples of anomaly detectors in use at eBay.
+Lakhina \emph{et al.} show that the analysis of feature distributions leads to significant advances on sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, while the anomaly detection approach leads gains on automatic classification of anomalies via unsupervised learning.
+Goldberg and Shan \cite{goldberg2015importance} argue, based on experience and on putblished real-life examples of anomaly detectors in use at eBay, that processing monitored signals into features is a more fruitful for anomaly detection systems than devising complex statistical tests.
 
@@
 This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
-Considering that learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}, higher moments have been adopted for outlier detection and discriminative analysis.
-Skewness and the kurtosis are higher moments that can contribute to sensitive outlier detectors, taking into account that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
+
+Considering that learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}, higher moments can be used for outlier detection and discriminative analysis of skewed data.
+Skewness and the kurtosis are higher moments that can contribute to sensitive outlier detectors, taking into account that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection in skewed data sets \cite{hubert2009robustskewed}.
 
@@
 The proposed approach relies on a robust subspace of normal traffic for estimating higher moments.
-The anomaly detection for new observations evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervisioned fashion and without the computational cost of new robust subspace learning for new observations.
+Thus, we propose that the anomaly detection for new observations should evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervisioned fashion and without the computational cost of new robust subspace learning for new observations.
 
@@
 
-\cite{garcia2014empirical} proposed CTU-13 data set and argues that for evaluations of anomaly detection, none of the botnet families used in the training and crossvalidation data set should be used in the testing data set, considering that this ensures that the methods can generalize and detect new behaviors.
-However, hour proposal is semi-supervised and does not rely on training data with anomalies.
-Adopting the training and testing approach of \cite{garcia2014empirical}, some botnets malware wouldnt be evaluated, since in Garcia's proposal some botnets are present only for training.
-Therefore, we evaluate each scenario individually.
+The CTU-13 is a data set of botnet traffic that was captured in the Czech Technical University \cite{garcia2014empirical}.
+It contains 13 scenarios with network traffic of botnet malwares, that are: neris, rbot, virut, menti, sogou, nsys.ay and murlo.
 
-A. Description of CTU-13 data set
-The CTU-13 is a data set of botnet traffic that was captured
-in the Czech Technical University [24]. It contains 13 scenarios
-with various botnet types. We consider scenarios 1, 2, 6, 8, and
-9 of the data set.
-• Scenario 1 corresponds to an IRC-based botnet that sent
-spam for almost 6.5 hours.
-• Scenario 2 (2.5-hours) is from the same botnet.
-• Scenario 6 is from a botnet that scanned SMTP (Simple
-Mail Transfer Protocol) servers for 2 hours and connected
-to several RDP (Remote Desktop Protocol) services. Dif-
-ferent with Scenarios 1 and 2, this botnet neither sent
-spam nor did it attack. It’s C\&C server used a proprietary
-protocol.
-• Scenario 8 is from a botnet that contacted a lot of Chinese
-C\&C hosts and received large amounts of encrypted data.
-The botnet also cracked the passwords of machines during
-the 19-hour attack.
-• Scenario 9 corresponds to a case where 10 local hosts
-were infected by a spamming botnet. More than 600 spams
-were sent over 5 hours.
-For all the scenarios, the authors of the CTU-13 data set
-convert the captured pcap files to NetFlows and release the
-processed flows. The data set contains ground-truth labels for
-flows as follows: flows from or to the infected machines
-are labeled as “botnet”; flows from or to noninfected ma-
-chines are labeled as “normal”; all other flows are labeled as
-“background.”
+The types of attacks and C\&C present in CTU-13 data set are:
+\begin{itemize}
+	\item \textbf{Attacks:} Click Fraud (CF), Port Scan (PS), Compiled and Controled by Authors (US), SPAM and DDOS;
+	\item \textbf{C\&C:} IRC, P2P and HTTP.
+\end{itemize}
 
-Evaluating network anomaly detection solutions remains a challenge due to the lack of up-to-date real-world data sets for training and testing \cite{osanaiye2016distributed}.
+For all the scenarios, the authors of the CTU-13 data set convert the captured pcap files to NetFlows and release the processed flows.
+The data set contains ground-truth labels for flows as follows: flows from or to the infected machines are labeled as “botnet”; flows from or to noninfected machines are labeled as “normal”; all other flows are labeled as “background.”
 
-We propose to use CTU-13 \cite{garcia2014empirical} to evaluate our proposal, considering that the CTU-13 is a data set that aims to have a large capture of real botnet attacks and C\&C traffic mixed with normal and background traffic.
+The data set orignally contains the following features:  Start Time, End Time, Duration, Source IP address, Source Port, Direction, Destination IP address, Destination Port, State, SToS, Total of Packets and Total of Bytes.
 
-The CTU-13 data set consists in thirteen captures (called scenarios) of different botnet samples.
-
-On each scenario it was executed a specific malware, which used several protocols and performed different actions.
-
-The botnet traffic is divided into attacks and Command and Control (C\&C) traffic.
-
-\textbf{Attacks:} Click Fraud, Port Scan, FastFlux, Compiled and Controled by Authors, SPAM and DDOS.
-
-\textbf{C\&C:} IRC, P2P and HTTP.
-
-The data set contains the following features:  Start Time, End Time, Duration, Source IP address, Source Port, Direction, Destination IP address, Destination Port, State, SToS, Total Packets and Total Bytes.
-
-Start Time, Duration, Protocol, Source Address, Source Port, Direction, Destination Address, Destination Port, State, Type of service from source to destination, Type of service from destination to source, Total of Packets, Total of bytes, Total bytes from source to destination, Label as normal or anomalous
-
-Our labeling strategy assigns three different labels: back-ground, botnet and normal.
-The priority to assign the labels is the following: 1.
-Assign the Background label to the whole traffic.
-2.
-Assign the Normal label to the traffic that matches certain filters.
-3.
-Assign the Botnet label to all the traffic that comes from or to any of the known infected IP addresses.
+Table \ref{tab:tab01} presents the malwares used for botnet attacks, the attack and C\&C types, the total number of flows, the number of malicious flows, which includes the C\&C and botnet flows.
+Table \ref{tab:tab01} also shows the ratio of C\&C and botnet flows.
 
@@
 			15 & rbot & IRC, DDoS, US & 1,121,076 & 2,580 (0.23\%) & 52 (0.00\%) & 2,528 (0.23\%) \\ \hline
-			15-2 & virut & Spam, PS,HTTP & 129,832 & 901 (0.69\%) & 24 (0.02\%) & 877 (0.68\%) \\ \hline
+			15-2 & virut & Spam, PS, HTTP & 129,832 & 901 (0.69\%) & 24 (0.02\%) & 877 (0.68\%) \\ \hline
 			16 & menti & PS & 558,919 & 4,630 (0.83\%) & 199 (0.04\%) & 4,431 (0.79\%) \\ \hline
@@
 			16-3 & murlo & PS & 2,954,230 & 6,127 (0.21\%) & 1,074 (0.04\%) & 5,053 (0.17\%) \\ \hline
-			17 & neris & IRC, Spam, CF,PS & 2,087,508 & 184,987 (8.86\%) & 2,973 (0.14\%) & 182,014 (8.72\%) \\ \hline
+			17 & neris & IRC, Spam, CF, PS & 2,087,508 & 184,987 (8.86\%) & 2,973 (0.14\%) & 182,014 (8.72\%) \\ \hline
 			18 & rbot & IRC, DDoS, US & 1,309,791 & 106,352 (8.12\%) & 33 (0.00\%) & 106,319 (8.12\%) \\ \hline
@@
 \end{table}
-
-Table \ref{tab:tab01} presents the malwares used for botnet attacks, the attack and C\&C types, the total number of flows, the number of malicious flows, which includes the C\&C and botnet flows.
-Table \ref{tab:tab01} also shows the ratio of C\&C and botnet flows.
 
@@
 
-Following (2 slides) are the selected features:
-\textbf{normal-flow-count:} count of normal (known) flow.
-\textbf{background-flow-count :} count of background (unknown) flows.
-\textbf{avg-duration:} average duration time of flows.
-\textbf{n-conn:} count of flows.
-\textbf{n-icmp:} count of ICMP packets.
-\textbf{n-tcp:} count of TCP packets.
-\textbf{n-udp:} count of UDP packets.
-\textbf{n-dports<1024:} count of destinations ports lower than 1024.
-\textbf{n-sports<1024:} count of source ports lower than 1024.
-\textbf{n-sports>1024:} count of source ports higher than 1024.
-
-Following are the remaining selected features:
-\textbf{n-s-a-p-address:} count of class A source address.
-\textbf{n-d-a-p-address:} count of class A destination address.
-\textbf{n-s-b-p-address:} count of class B source address.
-\textbf{n-d-b-p-address:} count of class B destination address.
-\textbf{n-s-c-p-address:} count of class C source address.
-\textbf{n-d-c-p-address:} count of class C destination address.
-\textbf{n-s-na-p-address:} count of class D* source address.
-\textbf{n-d-na-p-address:} count of class D* destination address.
+\begin{itemize}
+	\item \textbf{normal-flow-count:} count of normal (known) flow.
+	\item \textbf{background-flow-count :} count of background (unknown) flows.
+	\item \textbf{avg-duration:} average duration time of flows.
+	\item \textbf{n-conn:} count of flows.
+	\item \textbf{n-icmp:} count of ICMP packets.
+	\item \textbf{n-tcp:} count of TCP packets.
+	\item \textbf{n-udp:} count of UDP packets.
+	\item \textbf{n-dports<1024:} count of destinations ports lower than 1024.
+	\item \textbf{n-sports<1024:} count of source ports lower than 1024.
+	\item \textbf{n-sports>1024:} count of source ports higher than 1024.
+	\item \textbf{n-s-a-p-address:} count of class A source address.
+	\item \textbf{n-d-a-p-address:} count of class A destination address.
+	\item \textbf{n-s-b-p-address:} count of class B source address.
+	\item \textbf{n-d-b-p-address:} count of class B destination address.
+	\item \textbf{n-s-c-p-address:} count of class C source address.
+	\item \textbf{n-d-c-p-address:} count of class C destination address.
+	\item \textbf{n-s-na-p-address:} count of class D* source address.
+	\item \textbf{n-d-na-p-address:} count of class D* destination address.
+\end{itemize}
 
@@
 % \end{algorithm}
+
+% \cite{garcia2014empirical} proposed CTU-13 data set and argues that for evaluations of anomaly detection, none of the botnet families used in the training and crossvalidation data set should be used in the testing data set, considering that this ensures that the methods can generalize and detect new behaviors.
+However, hour proposal is semi-supervised and does not rely on training data with anomalies.
+Adopting the training and testing approach of \cite{garcia2014empirical}, some botnets malware wouldnt be evaluated, since in Garcia's proposal some botnets are present only for training.
+Therefore, we evaluate each scenario individually.
+
+% \cite{wang2017botnet}
+% A. Description of CTU-13 data set
+% The CTU-13 is a data set of botnet traffic that was captured
+% in the Czech Technical University [24]. It contains 13 scenarios
+% with various botnet types. We consider scenarios 1, 2, 6, 8, and
+% 9 of the data set.
+% • Scenario 1 corresponds to an IRC-based botnet that sent spam for almost 6.5 hours.
+% • Scenario 2 (2.5-hours) is from the same botnet.
+% • Scenario 6 is from a botnet that scanned SMTP (Simple Mail Transfer Protocol) servers for 2 hours and connected to several RDP (Remote Desktop Protocol) services.
+Dif-erent with Scenarios 1 and 2, this botnet neither sent spam nor did it attack.
+It’s C\&C server used a proprietary protocol.
+% • Scenario 8 is from a botnet that contacted a lot of Chinese C\&C hosts and received large amounts of encrypted data. The botnet also cracked the passwords of machines during the 19-hour attack.
+% • Scenario 9 corresponds to a case where 10 local hosts were infected by a spamming botnet. More than 600 spams were sent over 5 hours.

### 2019-05-03 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on data model

@@
 
-We adopt a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
+We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
 We evaluate the accuracy of the proposed approach for botnet detecion on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
@@
 Robust subspace learning has receiving a growing atention of researchers aiming the development of anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
-
 RPCA is been mainly applied to computer vision, in problems of robust subspace tracking and robust subspace recovery.
@@
 
-We adopt a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
+We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
 Finally, We evaluate the accuracy of the proposed approach for botnet detecion on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
@@
 
-The types of attacks and C\&C present in CTU-13 data set are:
+The types of network attacks and C\&C present in CTU-13 data set are:
 \begin{itemize}
@@
 
-The data set orignally contains the following features:  Start Time, End Time, Duration, Source IP address, Source Port, Direction, Destination IP address, Destination Port, State, SToS, Total of Packets and Total of Bytes.
+The data set orignally contains the following features:  Start Time, End Time, Duration, Source IP Address, Source Port, Direction, Destination IP Address, Destination Port, State, SToS, Total of Packets and Total of Bytes.
 
-Table \ref{tab:tab01} presents the malwares used for botnet attacks, the attack and C\&C types, the total number of flows, the number of malicious flows, which includes the C\&C and botnet flows.
-Table \ref{tab:tab01} also shows the ratio of C\&C and botnet flows.
+Table \ref{tab:tab01} presents an overview organized by scenario and shows the malwares used for botnet attacks, the types of attacks and C\&C types, the total number of flows, the number of malicious flows, which includes the C\&C, and finally shows the number of C\&C and attack flows separatelly.
 
@@
 	\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | }
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Scenario\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Malware\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Type\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Total Flows\end{tabular} & \begin{tabular}[x]{@{}l@{}}Malicious\end{tabular} & \begin{tabular}[x]{@{}l@{}}C\&C\end{tabular} & \begin{tabular}[x]{@{}l@{}}Botnet\end{tabular}\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Scenario\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Malware\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Type\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Total Flows\end{tabular} & \begin{tabular}[x]{@{}l@{}}Malicious\end{tabular} & \begin{tabular}[x]{@{}l@{}}C\&C\end{tabular} & \begin{tabular}[x]{@{}l@{}}Attack\end{tabular}\\ \hline
 			10 & neris & IRC, Spam, CF & 2,824,636 & 40,961 (1.45\%) & 341 (0.01\%) & 40,620 (1.44\%) \\ \hline
@@
 
-\subsection{feature engineering}
-\label{sec:feat_eng}
+The CTU-13 aims to reproduce the percentual of C\&C and attack flows according to real cases, where the number of malicious flows are very low in comparison to normal and background flows, resulting into a very imbalanced and challenging data set for botnet detection.
 
-Table \ref{tab:tab01} shows that CTU-13 is very imbalanced data set, when comparing the amount of malicious to normal traffic.
-
-Malicious observations are beteeen $0.06\%$ up to $8.86\%$ of all flows.
-
-The scenarios have between $107,251$ and $4,710,638$ flows.
-
-The CTU-13 data set is also very skewed. Figure \ref{fig:fig06} exemplifies the skewed distribution of normal and anomalous flows, showing the flow duration of scenario 12.
+An Exploratory Data Analysis (EDA) reveals that distributions of original features of the CTU-13 are skewed and presents an overlaping between normal and anomalous flows.
+Following we present the Figure \ref{fig:fig01}, that presents the distribution of flow duration for scenario 12, and exemplifies the data skewness overlaping.
+We adopt a sampling method based on data aggregation for obtaining more discriminative features, in accordance to important approaches for network anomaly detection \cite{lakhina2005mining, callegari2011novel, garcia2014empirical, chandrashekar2014survey,acarali2016survey, vieira2017model}.
 
@@
      \caption{Distribution of flow duration for scenario 12.}
-     \label{fig:fig06}
+     \label{fig:fig01}
 \end{figure}
 
-Methods for imbalance learning has attracted growing attention from both academia and industry, due to the restrictions imposed by imbalanced data to machine learning algorithms.
+\subsection{feature engineering}
+\label{sec:feat_eng}
 
-Lakhina et al.
-\cite{lakhina2005mining} argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
-
-To deal with the data imbalance of CTU-13 data set, we adopt a sampling method based on data aggragation for feature generation, in order to obtain less imbalanced and more discrimnative data.
-Based on related works \cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey} and experience \cite{vieira2017model, galibus2017offline}, we selected some features for improving the anomaly detection.
-The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15 and 0.25 seconds, for each scenario of the CTU-13 data set.
+Based on related works \cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey}, previous experience \cite{vieira2017model, galibus2017offline} and experimental evaluation of our proposal for botnet detection, we selected some features for improving the anomaly detection by means of a method based on higher moments distances from robust subspace.
+Following we list the proposed features:
 
@@
 	\item \textbf{avg-duration:} average duration time of flows.
-	\item \textbf{n-conn:} count of flows.
-	\item \textbf{n-icmp:} count of ICMP packets.
-	\item \textbf{n-tcp:} count of TCP packets.
-	\item \textbf{n-udp:} count of UDP packets.
 	\item \textbf{n-dports<1024:} count of destinations ports lower than 1024.
 	\item \textbf{n-sports<1024:} count of source ports lower than 1024.
-	\item \textbf{n-sports>1024:} count of source ports higher than 1024.
 	\item \textbf{n-s-a-p-address:} count of class A source address.	
-	\item \textbf{n-d-a-p-address:} count of class A destination address.
 	\item \textbf{n-s-b-p-address:} count of class B source address.	
-	\item \textbf{n-d-b-p-address:} count of class B destination address.
 	\item \textbf{n-s-c-p-address:} count of class C source address.	
-	\item \textbf{n-d-c-p-address:} count of class C destination address.
 	\item \textbf{n-s-na-p-address:} count of class D* source address.	
-	\item \textbf{n-d-na-p-address:} count of class D* destination address.
 \end{itemize}
+
+The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15 and 0.25 seconds, for each scenario of the CTU-13 data set.
 

### 2019-05-04 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on data model

@@
 Distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid botnet's attacks.
-This paper proposes an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggretation, in a semi-supervisioned fashion and without the computational cost of new robust subspace learning for new observations.
-Experimental results reveal that the proposed approach outerperforms widely adoted algorithms and a method based on robust estimates and robust skewness for network anomaly detection.
+This paper proposes an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggregation, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations.
+Experimental results reveal that the proposed approach outperforms widely adopted algorithms and a method based on robust estimates and robust skewness for network anomaly detection.
 \end{abstract}
@@
 Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects \cite{bhuyan2014network,ahmed2016survey}.
-Anomalies can be hard to identify and separate from normal data due to the rare occurrencies of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.
+Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.
 
-Several efforts and reasearches aim to avoid network attacks based on known attackers, fingerprints or other patterns.
-However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid unkown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detecion systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
-To face this problem, it is possible to adopt unsupervised or semi-pervised approaches for network anomaly detection, where it is not necessary known anomalies for trannning models \cite{moustafa2019holistic}.
+Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or other patterns.
+However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid unknown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detection systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
+To face this problem, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, where it is not necessary known anomalies for training models \cite{moustafa2019holistic}.
 
@@
 This findings show that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.
-Therefore, learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
+Therefore, learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
 Skewness and kurtosis are higher moments that can have their measurement as very sensitive for outlier detection, considering that the properties of the kurtosis has been used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
 
-Principal Component Analysis (PCA) is very sensitive to anomalies, thus outliers can corrupt the entire data or can be explored to highlight unespected changes and indicate attacks or frauds \cite{callegari2011novel,Lee2013,vieira2017model}.
-However, PCA is not precise to reveal flows, time or detailed information for each anomaly, and sometimes relies into visual analysis for pricipal component selection.
+Principal Component Analysis (PCA) is very sensitive to anomalies, thus outliers can corrupt the entire data or can be explored to highlight unexpected changes and indicate attacks or frauds \cite{callegari2011novel,Lee2013,vieira2017model}.
+However, PCA is not precise to reveal flows, time or detailed information for each anomaly, and sometimes relies into visual analysis for principal component selection.
 Robust Principal Component Analysis (RPCA) \cite{candes2011robust} is an extension of PCA that aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, which refers to outliers’ matrix.
-Therefore, robust subspace learning has receiving a growing atention of researchers aiming the development of network anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd, hubert2005robpca,hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
+Therefore, robust subspace learning has receiving a growing attention of researchers aiming the development of network anomaly detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd, hubert2005robpca,hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
 
@@
 We also believe that the distance between robust estimates of normal traffic and anomalies can highlight discrepancies and be used for botnet detection.
-Therefore, we propose an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggretation.
+Therefore, we propose an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggregation.
 The proposed approach relies on a robust subspace of normal traffic for estimating higher moments.
-The anomaly detection for new observations evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervisioned fashion and without the computational cost of new robust subspace learning for new observations.
+The anomaly detection for new observations evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations.
 
-We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
-We evaluate the accuracy of the proposed approach for botnet detecion on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
+We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adopted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
+We evaluate the accuracy of the proposed approach for botnet detection on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
 
-Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statisctical approaches, which are K-means and Gaussian Mixture Model (GMM) \cite{gaddam2007kmeans,moustafa2019holistic}, respectively.
-We also evaluate the results of ROBPCA \cite{hubert2005robpca}, wich is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
+Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches, which are K-means and Gaussian Mixture Model (GMM) \cite{gaddam2007kmeans,moustafa2019holistic}, respectively.
+We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
 
@@
 A botnet is a network of bots, that are compromised machines under the influence of malware (bot).
-The botnet is commandeered by a botmaster and used as resource for attacks, such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information exfiltration.
+The botnet is commandeered by a botmaster and used as resource for attacks, such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information ex-filtration.
 We refer to \cite{ahmed2016survey,moustafa2019holistic} for an overview of network attacks.
-The botmaster coordenate a botnet through a command and control (C\&C) channel where bots receive commands and syncronize attacks and fraudulent activities.
-Centralized C\&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time, but other protocols, such as HTTP, and achitectures, such as Peer-To-Peer, have also been adopted \cite{gu2008botminer}.
+The botmaster coordinate a botnet through a command and control (C\&C) channel where bots receive commands and synchronize attacks and fraudulent activities.
+Centralized C\&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time, but other protocols, such as HTTP, and architectures, such as Peer-To-Peer, have also been adopted \cite{gu2008botminer}.
 
@@
 % justificative for the dataset and present who else use it
-Several approaches for network attack detections uses the KDD 99 \cite{ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
+Several approaches for network attack detection uses the KDD 99 \cite{ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
 Even though the KDD 99 data set are criticized by the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data sets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
 NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that redundant data records are removed, in order to avoid biased classifications.
-However, NSL-KDD dataset maintain the limitations of the KDD 99 regarding volume and reproduction of network traffic similar to the real networks.
+However, NSL-KDD data set maintain the limitations of the KDD 99 regarding volume and reproduction of network traffic similar to the real networks.
 
@@
 Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by normal and background labeled data, in an imbalanced distribution like in a real network.
-The autors recommend that none of the botnet families used in the training and crossvalidation data set should be used in the testing data set, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
+The authors recommend that none of the botnet families used in the training and cross-validation data set should be used in the testing data set, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
 Adopting the training and testing approach of \cite{garcia2014empirical}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
@@
 We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for more details regarding robust subspace learning.
-Robust subspace learning has receiving a growing atention of researchers aiming the development of anomlay detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
+Robust subspace learning has receiving a growing attention of researchers aiming the development of anomaly detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
 RPCA is been mainly applied to computer vision, in problems of robust subspace tracking and robust subspace recovery.
-However, RPCA has also been adoted for general outlier detection \cite{hubert2005robpca,hubert2009robustskewed,cherapanamjeri2017thresholding,zhou2017anomaly,NetflixSurus} and for anomaly detection on network traffic \cite{pascoal2012robust}.
+However, RPCA has also been adopted for general outlier detection \cite{hubert2005robpca,hubert2009robustskewed,cherapanamjeri2017thresholding,zhou2017anomaly,NetflixSurus} and for anomaly detection on network traffic \cite{pascoal2012robust}.
 ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
@@
 Lakhina \emph{et al.} show that the analysis of feature distributions leads to significant advances on sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, while the anomaly detection approach leads gains on automatic classification of anomalies via unsupervised learning.
-Goldberg and Shan \cite{goldberg2015importance} argue, based on experience and on putblished real-life examples of anomaly detectors in use at eBay, that processing monitored signals into features is a more fruitful for anomaly detection systems than devising complex statistical tests.
+Goldberg and Shan \cite{goldberg2015importance} argue, based on experience and on published real-life examples of anomaly detectors in use at eBay, that processing monitored signals into features is a more fruitful for anomaly detection systems than devising complex statistical tests.
 
@@
 Findings of Benson \emph{et al.} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
-This findings highligh that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
+This findings highlight that the skeweness and heavy-tailed distributions can impact algorithms that rely on gaussian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
 
-Considering that learning methods for imbalanced and skewed data have been atracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}, higher moments can be used for outlier detection and discriminative analysis of skewed data.
+Considering that learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}, higher moments can be used for outlier detection and discriminative analysis of skewed data.
 Skewness and the kurtosis are higher moments that can contribute to sensitive outlier detectors, taking into account that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection in skewed data sets \cite{hubert2009robustskewed}.
 
-We propose an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggretation from the CTU-13 data set.
+We propose an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggregation from the CTU-13 data set.
 The proposed approach relies on a robust subspace of normal traffic for estimating higher moments.
-Thus, we propose that the anomaly detection for new observations should evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervisioned fashion and without the computational cost of new robust subspace learning for new observations.
+Thus, we propose that the anomaly detection for new observations should evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations.
 
-We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adoted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
+We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adopted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
 Finally, We evaluate the accuracy of the proposed approach for botnet detecion on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
@@
 \begin{itemize}
-	\item \textbf{Attacks:} Click Fraud (CF), Port Scan (PS), Compiled and Controled by Authors (US), SPAM and DDOS;
+	\item \textbf{Attacks:} Click Fraud (CF), Port Scan (PS), Compiled and Controlled by Authors (US), SPAM and DDOS;
 	\item \textbf{C\&C:} IRC, P2P and HTTP.
@@
 For all the scenarios, the authors of the CTU-13 data set convert the captured pcap files to NetFlows and release the processed flows.
-The data set contains ground-truth labels for flows as follows: flows from or to the infected machines are labeled as “botnet”; flows from or to noninfected machines are labeled as “normal”; all other flows are labeled as “background.”
+The data set contains ground-truth labels for flows as follows: flows from or to the infected machines are labeled as “botnet”; flows from or to non infected machines are labeled as “normal”; all other flows are labeled as “background.”
 
-The data set orignally contains the following features:  Start Time, End Time, Duration, Source IP Address, Source Port, Direction, Destination IP Address, Destination Port, State, SToS, Total of Packets and Total of Bytes.
+The data set originally contains the following features:  Start Time, End Time, Duration, Source IP Address, Source Port, Direction, Destination IP Address, Destination Port, State, SToS, Total of Packets and Total of Bytes.
 
-Table \ref{tab:tab01} presents an overview organized by scenario and shows the malwares used for botnet attacks, the types of attacks and C\&C types, the total number of flows, the number of malicious flows, which includes the C\&C, and finally shows the number of C\&C and attack flows separatelly.
+Table \ref{tab:tab01} presents an overview organized by scenario and shows the malwares used for botnet attacks, the types of attacks and C\&C types, the total number of flows, the number of malicious flows, which includes the C\&C, and finally shows the number of C\&C and attack flows separately.
 
@@
 
-The CTU-13 aims to reproduce the percentual of C\&C and attack flows according to real cases, where the number of malicious flows are very low in comparison to normal and background flows, resulting into a very imbalanced and challenging data set for botnet detection.
+The CTU-13 aims to reproduce the percentage of C\&C and attack flows according to real cases, where the number of malicious flows are very low in comparison to normal and background flows, resulting into a very imbalanced and challenging data set for botnet detection.
 
-An Exploratory Data Analysis (EDA) reveals that distributions of original features of the CTU-13 are skewed and presents an overlaping between normal and anomalous flows.
-Following we present the Figure \ref{fig:fig01}, that presents the distribution of flow duration for scenario 12, and exemplifies the data skewness overlaping.
+An Exploratory Data Analysis (EDA) reveals that distributions of original features of the CTU-13 are skewed and presents an overlapping between normal and anomalous flows.
+Following we present the Figure \ref{fig:fig01}, that presents the distribution of flow duration for scenario 12, and exemplifies the data skewness overlapping.
 We adopt a sampling method based on data aggregation for obtaining more discriminative features, in accordance to important approaches for network anomaly detection \cite{lakhina2005mining, callegari2011novel, garcia2014empirical, chandrashekar2014survey,acarali2016survey, vieira2017model}.
@@
 
-Based on related works \cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey}, previous experience \cite{vieira2017model, galibus2017offline} and experimental evaluation of our proposal for botnet detection, we selected some features for improving the anomaly detection by means of a method based on higher moments distances from robust subspace.
-Following we list the proposed features:
+Based on related works \cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey}, previous experience \cite{vieira2017model, galibus2017offline} and cross validation  of our proposal for botnet detection, we selected the following features for improving the anomaly detection by means of a method based on higher moments distances from robust subspace.
 
 \begin{itemize}
-	\item \textbf{normal-flow-count:} count of normal (known) flow.
-	\item \textbf{background-flow-count :} count of background (unknown) flows.
-	\item \textbf{avg-duration:} average duration time of flows.
-	\item \textbf{n-dports<1024:} count of destinations ports lower than 1024.
-	\item \textbf{n-sports<1024:} count of source ports lower than 1024.
-	\item \textbf{n-s-a-p-address:} count of class A source address.	
-	\item \textbf{n-s-b-p-address:} count of class B source address.	
-	\item \textbf{n-s-c-p-address:} count of class C source address.	
-	\item \textbf{n-s-na-p-address:} count of class D* source address.	
+	\item \textbf{background-flow-count :} Count of unknown flows.
+	\item \textbf{normal-flow-count:} Count of known flows.
+	\item \textbf{avg-duration:} Average time of flows.
+	\item \textbf{n-dports<1024:} Count of destinations ports lower than 1024.
+	\item \textbf{n-sports<1024:} Count of source ports lower than 1024.
+	\item \textbf{n-s-a-p-address:} Count of class A source address.
+	\item \textbf{n-s-b-p-address:} Count of class B source address.
+	\item \textbf{n-s-c-p-address:} Count of class C source address.
+	\item \textbf{n-s-na-p-address:} Count of source address that are not class A, B or C.
 \end{itemize}
 
-The features are generated by a data aggregation into a selected window size. We evalute our proposal for windows with 0.15 and 0.25 seconds, for each scenario of the CTU-13 data set.
+Since the selected features are counting and average calculations, the features are generated through a data aggregation by time window.
+We evaluate time windows of 0.15 and 0.25 seconds, taking into account that lower windows reduce the meantime between the data collection and results of the botnet detection approach.
 
-Figures \ref{fig:fig07} and \ref{fig:fig08} show the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 and 2 seconds.
-
-In comparison to Figure \ref{fig:fig06}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on the average is more discriminative.
+Figure \ref{fig:fig02} shows the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 and 2 seconds.
+In comparison to Figure \ref{fig:fig01}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on the average is more discriminative.
 
@@
      \includegraphics[width=8cm]{figures/agg_distplot_0_15s_12_avg_duration.png}
-     \caption{Distribution of average flow duration for scenario 12, aggregated by windows of 0.15 seconds.}
-     \label{fig:fig07}
+     \caption{Average flow duration for scenario 12, aggregated by 0.15 seconds.}
+     \label{fig:fig02}
 \end{figure}
 
-\begin{figure}[h!]
-     \centering
-     \includegraphics[width=8cm]{figures/agg_distplot_2s_12_avg_duration.png}
-     \caption{Distribution of average flow duration for scenario 12, aggregated by windows of 2 seconds.}
-     \label{fig:fig08}
-\end{figure}
+The data is modeled into the matrix $\textbf{M}$ with rows as observations and columns the features.
 
@@
 
-Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statisctical approaches, which are K-means and Gaussian Mixture Model (GMM) \cite{gaddam2007kmeans,moustafa2019holistic}, respectively.
-We also evaluate the results of ROBPCA \cite{hubert2005robpca}, wich is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
+Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches, which are K-means and Gaussian Mixture Model (GMM) \cite{gaddam2007kmeans,moustafa2019holistic}, respectively.
+We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
 In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of CTU-13 individually, in a semi-supervised approach that does not rely on training data with anomalies.
@@
 
-	Gargia et al. \cite{garcia2014empirical} propose a division of CTU-13's scearios into train and test, for evaluation of network attack detection algorithms. 
+Gargia et al. \cite{garcia2014empirical} propose a division of CTU-13's scenarios into train and test, for evaluation of network attack detection algorithms. 
 However, we decide to evaluate our proposed approach for each scenario individually, considering that in the division proposed by Gargia et al.
@@
 Why GMM?
+
 As clustering-based method, we selected K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007kmeans,moustafa2019holistic}.
@@
 Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets.
-Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high thoughput network traffic.
-Additionally, future research can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.
+Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring02       rch can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.
 

### 2019-05-12 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
changes during the qcon

@@
 
-\title{Higher Moments Distances from Robust Subspace for Botnet Detection}
+\title{Moment Distances from Robust Subspace for Botnet Detection}
 
@@
 This paper proposes an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggregation, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations.
-Experimental results reveal that the proposed approach outperforms widely adopted algorithms and a method based on robust estimates and robust skewness for network anomaly detection.
+Experimental results reveal that the proposed approach outperforms widely adopted algorithms and a method based on robust skewness for a public data set of botnet attacks.
 \end{abstract}
@@
 \begin{keyword}
-Network Anomaly Detection \sep Botnet Detection \sep Higher Moments \sep Robust Principal Component Analysis \sep Mahalanobis Distance
+Network Anomaly Detection \sep Botnet Detection \sep Moments \sep Robust Principal Component Analysis \sep Mahalanobis Distance
 \end{keyword}
@@
 Learning algorithms for imbalanced data has been a challenging research topic, considering that the fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
-Some techniques can make the data discriminative for better results on imbalanced data: Data engineering (i.e.
+Some techniques can make the data discriminative for better results on the imbalanced learning problem: Data engineering (i.e.
 sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, garcia2014empirical, acarali2016survey}); Feature extraction (i.e.
@@
 Findings of Benson \emph{et al.}  \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
-This findings show that the skeweness and heavy-tailed distributions can impact algorithms that rely on guassian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.
+This findings show that the skeweness and heavy-tailed distributions can impact algorithms that rely on gaussian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.
 Therefore, learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
@@
 The botnet is commandeered by a botmaster and used as resource for attacks, such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information ex-filtration.
-We refer to \cite{ahmed2016survey,moustafa2019holistic} for an overview of network attacks.
+We refer to \cite{ahmed2016survey} and \cite{moustafa2019holistic} for an overview of network attacks.
 The botmaster coordinate a botnet through a command and control (C\&C) channel where bots receive commands and synchronize attacks and fraudulent activities.
@@
 
-% justificative for the dataset and present who else use it
+% justification for the data set and present who else use it
 Several approaches for network attack detection uses the KDD 99 \cite{ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
@@
 
-The objective of Garcia \emph{et al.} \cite{garcia2014empirical} is to compare three botnet detection methods using a simple and reproducible methodology, a good data set and a new error metric.
+The objective of Garcia \emph{et al.} \cite{garcia2014empirical} is to compare three botnet detection methods using a simple and reproducible methodology, and a good data set and a new error metric.
 This paper evaluates some data sets for network anomaly detection, surveys some approaches for botnet detection, proposes two methods (BClus and CAMNEP) for botnet detection and compare results to BotHunter \cite{gu2007bothunter}.
-Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by normal and background labeled data, in an imbalanced distribution like in a real network.
+Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by attack, normal and background labeled data, in an imbalanced distribution like in a real network.
 The authors recommend that none of the botnet families used in the training and cross-validation data set should be used in the testing data set, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
-Adopting the training and testing approach of \cite{garcia2014empirical}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
+Adopting the training and testing approach proposed by Garcia \emph{et al.}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
 
@@
 
-% Anomaly detection by robust subspace larning
-The problem of PCA or subspace learning for outlier corrupted data is called robust PCA (RPCA) or robust subspace learning \cite{vaswani2018robust}, which can be defined as the decomposition of a given data matrix $\textbf{M}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
+% Anomaly detection by robust subspace learning
+The problem of PCA or subspace learning for outlier corrupted data is called robust PCA (RPCA) or robust subspace learning \cite{vaswani2018robust}, which can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
 This definition is also referred to as the sparse + low rank (S + LR) formulation \cite{vaswani2018robust}.
 We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for more details regarding robust subspace learning.
+
 Robust subspace learning has receiving a growing attention of researchers aiming the development of anomaly detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
@@
 However, RPCA has also been adopted for general outlier detection \cite{hubert2005robpca,hubert2009robustskewed,cherapanamjeri2017thresholding,zhou2017anomaly,NetflixSurus} and for anomaly detection on network traffic \cite{pascoal2012robust}.
-ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of location and scatter matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
+ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of mean and covariance matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
 However, ROBPCA flags many points as outlying when the original data is skewed.
@@
 Lakhina \emph{et al.} \cite{lakhina2005mining} argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
-Lakhina \emph{et al.} show that the analysis of feature distributions leads to significant advances on sensitive detection of a wide range of anomalies, augmenting detections by volume-based methods, while the anomaly detection approach leads gains on automatic classification of anomalies via unsupervised learning.
+Lakhina \emph{et al.} show that the analysis of feature distributions leads to significant advances on sensitive detection of a wide range of anomalies, augmenting detection by volume-based methods, while the anomaly detection approach leads gains on automatic classification of anomalies via unsupervised learning.
 Goldberg and Shan \cite{goldberg2015importance} argue, based on experience and on published real-life examples of anomaly detectors in use at eBay, that processing monitored signals into features is a more fruitful for anomaly detection systems than devising complex statistical tests.
@@
 Considering that learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}, higher moments can be used for outlier detection and discriminative analysis of skewed data.
-Skewness and the kurtosis are higher moments that can contribute to sensitive outlier detectors, taking into account that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection in skewed data sets \cite{hubert2009robustskewed}.
+The skewness and kurtosis are higher moments that can contribute to sensitive outlier detectors, taking into account that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection in skewed data sets \cite{hubert2009robustskewed}.
 
@@
 We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adopted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
-Finally, We evaluate the accuracy of the proposed approach for botnet detecion on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
+Finally, We evaluate the accuracy of the proposed approach for botnet detection on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
 
@@
 
-The data set originally contains the following features:  Start Time, End Time, Duration, Source IP Address, Source Port, Direction, Destination IP Address, Destination Port, State, SToS, Total of Packets and Total of Bytes.
+The data set originally contains the following features by flows:  Start Time, End Time, Duration, Source IP Address, Source Port, Direction, Destination IP Address, Destination Port, State, Type of service from source to destination, Type of service from destination to source, Total of Packets and Total of Bytes.
 
@@
 An Exploratory Data Analysis (EDA) reveals that distributions of original features of the CTU-13 are skewed and presents an overlapping between normal and anomalous flows.
-Following we present the Figure \ref{fig:fig01}, that presents the distribution of flow duration for scenario 12, and exemplifies the data skewness overlapping.
-We adopt a sampling method based on data aggregation for obtaining more discriminative features, in accordance to important approaches for network anomaly detection \cite{lakhina2005mining, callegari2011novel, garcia2014empirical, chandrashekar2014survey,acarali2016survey, vieira2017model}.
+Following we present the Figure \ref{fig:fig01}, that presents the distribution of flow duration for scenario 12, and exemplifies the data skewness and overlapping.
 
@@
 \end{figure}
+
+Figure \ref{fig:fig02} shows a boxplot of flow duration on scenario 12, for normal and anomalous classes, in order to highlight the overlapping between classes.
+
+\begin{figure}[h!]
+     \centering
+     \includegraphics[width=8cm]{figures/agg_boxplot_0_15s_12_avg_duration.png}
+     \caption{Flow duration for scenario 12, aggregated by 0.15 seconds.}
+     \label{fig:fig02}
+\end{figure}
+
+We adopt a sampling method based on data aggregation for obtaining more discriminative features, in accordance to important approaches for network anomaly detection \cite{lakhina2005mining, callegari2011novel, garcia2014empirical, chandrashekar2014survey,acarali2016survey, vieira2017model}.
 
@@
 
-Figure \ref{fig:fig02} shows the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 and 2 seconds.
-In comparison to Figure \ref{fig:fig01}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on the average is more discriminative.
+Figure \ref{fig:fig03} shows the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 seconds.
+In comparison to Figure \ref{fig:fig01}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on the average flow duration is more discriminative and shows a better separation between normal and anomalous classes.
 
@@
      \caption{Average flow duration for scenario 12, aggregated by 0.15 seconds.}
-     \label{fig:fig02}
+     \label{fig:fig03}
 \end{figure}
 
-The data is modeled into the matrix $\textbf{M}$ with rows as observations and columns the features.
+The data is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features. Each scenario
 
@@
 
-This section describes the proposed technique to detect synflood, fraggle and port scan, according to Figure \ref{fig:fig80}, which represents the overview of the proposed framework for detection and identification of network attacks.
-In Subsection \ref{sec:prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvelue for each $q$-th time frame.
-Next, in Subsection \ref{sec:prop_MOSSchemes}, we show how to apply the eiganvalues on the MOS scheme in order to detect the attack.
-In Subsection \ref{sec:prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:prop_EigenSimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.
+This section describes the proposed approach for botnet detection by means of moment distances from robust estimates.
+We propose to learn a robust subspace $\textbf{L}$ from normal traffic $\textbf{X}_n$ for computing robust moments, i.e.
+the mean $\textbf{l}$, skewness $\textbf{s}$ and kurtosis $\textbf{k}$, in order to evaluate the distance $\textbf{d}$ of new observations $\textbf{X}_o$ of these moments and classify as anomalous the observations with the largest distances.
 
-\begin{figure}[h!]
-	\centering
-     % \includegraphics[height=11cm, width=9cm]{results/figures/mos_eigen_similarity.eps}
-     \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
-     \label{fig:fig80}
-\end{figure}
+Initially, the data $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ is split into normal data $\textbf{X}_n$ for training from normal data, and for $\textbf{X}_v$ and $\textbf{X}_t$ for respectively cross-validation and testing from contaminated data
 
-\subsection{Largest Eigenvalue by Time Frames}
-\label{sec:prop_LargestEigenvaluebyTimeFrames}
+It was shown that the RPCA problem can be solved via convex optimization that minimizes a combination of the nuclear norm and the $l_1$-norm, and that the method of augmented Lagrange multipliers (ALM) has been adopted to solve this convex problem.
 
-We propose to extend Fast-MCD for computing robust moments (skewness and kurtosis), in order to improve the anomaly detection in a skewed and imbalanced data set of network traffic with normal and botnet traffic.
+Robust PCA aims to recover a low-rank matrix $\textbf{L}$ from corrupted measurements $\textbf{M} = \textbf{L} + \textbf{S}$.
+This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP),[1] Stable PCP,[2] Quantized PCP,[3] Block based PCP,[4] and Local PCP.[5] Then, optimization methods are used such as the Augmented Lagrange Multiplier Method (ALM[6]), Alternating Direction Method (ADM[7]), Fast Alternating Minimization (FAM[8]) or Iteratively Reweighted Least Squares (IRLS [9][10][11]).
+
+Therefore we have $\textbf{L}$...
 

### 2019-05-13 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on data model

@@
 % table
+\usepackage{makecell}
 \usepackage{color, colortbl}
@@
 
-The CTU-13 is a data set of botnet traffic that was captured in the Czech Technical University \cite{garcia2014empirical}.
-It contains 13 scenarios with network traffic of botnet malwares, that are: neris, rbot, virut, menti, sogou, nsys.ay and murlo.
+The CTU-13 is a data set of botnet traffic that was captured in the Czech Technical University \cite{garcia2014empirical}, by means of a testbed and execution of malwares in a real network.
+The CTU-13 data set contains 13 scenarios with network traffic of botnet malwares, that are: neris, rbot, virut, menti, sogou, nsys.ay and murlo.
+The types of C\&C and attack traffic present in CTU-13 data set are:
 
-The types of network attacks and C\&C present in CTU-13 data set are:
 \begin{itemize}
@@
 
+We refer to Garcia \cite{garcia2014identifying} and Garcia \emph{et al.} \cite{garcia2014empirical} for a detailed description of the performed attacks and C\&C flows, also including the topology of the adopted testbed, rules for classifing normal flows, and an analysis of behaviors or patterns of the malware's traffic.
+
 For all the scenarios, the authors of the CTU-13 data set convert the captured pcap files to NetFlows and release the processed flows.
-The data set contains ground-truth labels for flows as follows: flows from or to the infected machines are labeled as “botnet”; flows from or to non infected machines are labeled as “normal”; all other flows are labeled as “background.”
+The data set contains ground-truth labels for flows as follows: flows from or to the infected machines are labeled as “botnet”; flows from or to well known and controlled machines are labeled as “normal”; all other flows are labeled as “background.”
 
@@
 
-Table \ref{tab:tab01} presents an overview organized by scenario and shows the malwares used for botnet attacks, the types of attacks and C\&C types, the total number of flows, the number of malicious flows, which includes the C\&C, and finally shows the number of C\&C and attack flows separately.
+Table \ref{tab:tab01} presents an overview organized by scenario and shows the malwares used for botnet attacks, the types of attacks and C\&C types, the total number of flows, the number of malicious flows, which includes flows of C\&C and attacks, and finally shows the number of normal flows.
 
@@
 	\label{tab:tab01}
-	\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | }
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Scenario\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Malware\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Type\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Total Flows\end{tabular} & \begin{tabular}[x]{@{}l@{}}Malicious\end{tabular} & \begin{tabular}[x]{@{}l@{}}C\&C\end{tabular} & \begin{tabular}[x]{@{}l@{}}Attack\end{tabular}\\ \hline
-			10 & neris & IRC, Spam, CF & 2,824,636 & 40,961 (1.45\%) & 341 (0.01\%) & 40,620 (1.44\%) \\ \hline
-			11 & neris & IRC, Spam, CF & 1,808,122 & 20,941 (1.16\%) & 673 (0.04\%) & 20,268 (1.12\%) \\ \hline
-			12 & rbot & IRC, PS, US & 4,710,638 & 26,822 (0.57\%) & 63 (0.00\%) & 26,759 (0.57\%) \\ \hline
-			15 & rbot & IRC, DDoS, US & 1,121,076 & 2,580 (0.23\%) & 52 (0.00\%) & 2,528 (0.23\%) \\ \hline
-			15-2 & virut & Spam, PS, HTTP & 129,832 & 901 (0.69\%) & 24 (0.02\%) & 877 (0.68\%) \\ \hline
-			16 & menti & PS & 558,919 & 4,630 (0.83\%) & 199 (0.04\%) & 4,431 (0.79\%) \\ \hline
-			16-2 & sogou & HTTP & 114,077 & 63 (0.06\%) & 26 (0.02\%) & 37 (0.03\%) \\ \hline
-			16-3 & murlo & PS & 2,954,230 & 6,127 (0.21\%) & 1,074 (0.04\%) & 5,053 (0.17\%) \\ \hline
-			17 & neris & IRC, Spam, CF, PS & 2,087,508 & 184,987 (8.86\%) & 2,973 (0.14\%) & 182,014 (8.72\%) \\ \hline
-			18 & rbot & IRC, DDoS, US & 1,309,791 & 106,352 (8.12\%) & 33 (0.00\%) & 106,319 (8.12\%) \\ \hline
-			18-2 & rbot & IRC, DDoS, US & 107,251 & 8,164 (7.61\%) & 2 (0.00\%) & 8,162 (7.61\%) \\ \hline
-			19 & nsys.ay & P2P & 325,471 & 2,168 (0.67\%) & 25 (0.01\%) & 2,143 (0.66\%) \\ \hline
-			15-3 & virut & Spam, PS, HTTP & 1,925,149 & 40,003 (2.08\%) & 536 (0.03\%) & 39,467 (2.05\%) \\ \hline
+	\begin{tabular}{| l | l | l | r | r | r | r | r | r | r | r | }
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}\textbf{ID}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\textbf{Bot}\end{tabular}	 &\begin{tabular}[x]{@{}l@{}}\textbf{Type}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\textbf{Total}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Malicious}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{C\&C}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Attack}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Normal}\end{tabular}\\ \hline
+			10 & neris &\makecell[l]{IRC, Spam,\\CF} & 2,824,636 & 40,961 (1.45\%) & 341 (0.01\%) & 40,620 (1.44\%) &30,387 (1.07\%)\\ \hline
+			11 & neris &\makecell[l]{IRC, Spam,\\CF} & 1,808,122 & 20,941 (1.16\%) & 673 (0.04\%) & 20,268 (1.12\%) &9,120 (0.5\%)\\ \hline
+			12 & rbot &\makecell[l]{IRC, PS,\\US} & 4,710,638 & 26,822 (0.57\%) & 63 (0.00\%) & 26,759 (0.57\%) &116,887 (2.48\%)\\ \hline
+			15 & rbot &\makecell[l]{IRC, DDoS,\\US} & 1,121,076 & 2,580 (0.23\%) & 52 (0.00\%) & 2,528 (0.23\%) &25,268 (2.25\%)\\ \hline
+			15-2 & virut &\makecell[l]{Spam, PS,\\HTTP} & 129,832 & 901 (0.69\%) & 24 (0.02\%) & 877 (0.68\%) &4,679 (3.6\%)\\ \hline
+			16 & menti &PS & 558,919 & 4,630 (0.83\%) & 199 (0.04\%) & 4,431 (0.79\%) &7,494 (1.34\%)\\ \hline
+			16-2 & sogou &HTTP & 114,077 & 63 (0.06\%) & 26 (0.02\%) & 37 (0.03\%) &1,677 (1.47\%)\\ \hline
+			16-3 & murlo &PS & 2,954,230 & 6,127 (0.21\%) & 1,074 (0.04\%) & 5,053 (0.17\%) &72,822 (2.46\%)\\ \hline
+			17 & neris &\makecell[l]{IRC, Spam,\\CF, PS} & 2,087,508 & 184,987 (8.86\%) & 2,973 (0.14\%) & 182,014 (8.72\%) &43,340 (1.57\%)\\ \hline
+			18 & rbot &\makecell[l]{IRC, DDoS,\\US} & 1,309,791 & 106,352 (8.12\%) & 33 (0.00\%) & 106,319 (8.12\%) &15,847 (1.2\%)\\ \hline
+			18-2 & rbot &\makecell[l]{IRC, DDoS,\\US} & 107,251 & 8,164 (7.61\%) & 2 (0.00\%) & 8,162 (7.61\%) &2,718 (2.53\%)\\ \hline
+			19 & nsys.ay &P2P & 325,471 & 2,168 (0.67\%) & 25 (0.01\%) & 2,143 (0.66\%) &7,628 (2.35\%)\\ \hline
+			15-3 & virut &\makecell[l]{Spam, PS,\\HTTP} & 1,925,149 & 40,003 (2.08\%) & 536 (0.03\%) & 39,467 (2.05\%) &31,939 (1.65)\\ \hline
 	\end{tabular}
@@
 
-An Exploratory Data Analysis (EDA) reveals that distributions of original features of the CTU-13 are skewed and presents an overlapping between normal and anomalous flows.
+An Exploratory Data Analysis (EDA) reveals that distributions of original features of the CTU-13 are skewed and present an overlapping between normal and anomalous flows.
 Following we present the Figure \ref{fig:fig01}, that presents the distribution of flow duration for scenario 12, and exemplifies the data skewness and overlapping.
@@
 
-Figure \ref{fig:fig02} shows a boxplot of flow duration on scenario 12, for normal and anomalous classes, in order to highlight the overlapping between classes.
-
-\begin{figure}[h!]
-     \centering
-     \includegraphics[width=8cm]{figures/agg_boxplot_0_15s_12_avg_duration.png}
-     \caption{Flow duration for scenario 12, aggregated by 0.15 seconds.}
-     \label{fig:fig02}
-\end{figure}
-
-We adopt a sampling method based on data aggregation for obtaining more discriminative features, in accordance to important approaches for network anomaly detection \cite{lakhina2005mining, callegari2011novel, garcia2014empirical, chandrashekar2014survey,acarali2016survey, vieira2017model}.
+To face data imbalance and skewness, we adopt a feature engineering by means of a sampling method based on data aggregation for obtaining more discriminative features, in accordance to important approaches for network anomaly detection \cite{lakhina2005mining, callegari2011novel, garcia2014empirical, chandrashekar2014survey,acarali2016survey, vieira2017model}.
 
@@
 
-Based on related works \cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey}, previous experience \cite{vieira2017model, galibus2017offline} and cross validation  of our proposal for botnet detection, we selected the following features for improving the anomaly detection by means of a method based on higher moments distances from robust subspace.
+Related works proposed some features for improving network attack detection in various scenarios and approaches\cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey}.
+Inpired by these proposals and based on previous experience \cite{vieira2017model, galibus2017offline} and cross validation, we propose the following features for improving the anomaly detection by means of a method based on higher moments distances from robust subspace.
 
 \begin{itemize}
-	\item \textbf{background-flow-count :} Count of unknown flows.
-	\item \textbf{normal-flow-count:} Count of known flows.
+	\item \textbf{total-flow-count :} Count of all flows.
+	\item \textbf{normal-flow-count:} Count of known flows, from controlled assets of the network, such routers, proxies, switches or machines that are part of the monitoring \cite{garcia2014identifying}.
 	\item \textbf{avg-duration:} Average time of flows.
-	\item \textbf{n-dports<1024:} Count of destinations ports lower than 1024.
-	\item \textbf{n-sports<1024:} Count of source ports lower than 1024.
+	\item \textbf{n-dports$<$1024:} Count of destinations ports lower than 1024.
+	\item \textbf{n-sports$<$1024:} Count of source ports lower than 1024.
 	\item \textbf{n-s-a-p-address:} Count of class A source address.
@@
 	\item \textbf{n-s-c-p-address:} Count of class C source address.
-	\item \textbf{n-s-na-p-address:} Count of source address that are not class A, B or C.
+	\item \textbf{n-s-na-p-address:} Count of source address that are not class A, B or C, such as reserved class D and E addresses, IPV6 or internal address.
 \end{itemize}
 
-Since the selected features are counting and average calculations, the features are generated through a data aggregation by time window.
-We evaluate time windows of 0.15 and 0.25 seconds, taking into account that lower windows reduce the meantime between the data collection and results of the botnet detection approach.
+The proposed features are counting and average calculations, therefore it is necessary to adopt time windows to aggragate flows and calculate the features.
+We evaluate windows of 0.15 and 0.25 seconds, considering account that lower windows reduce the meantime between the data collection and results of the botnet detection approach, as well as permit to have more details to exactly identify malicious flows and attackers.
 
-Figure \ref{fig:fig03} shows the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 seconds.
-In comparison to Figure \ref{fig:fig01}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on the average flow duration is more discriminative and shows a better separation between normal and anomalous classes.
+Figure \ref{fig:fig02} shows the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 seconds.
+In comparison to Figure \ref{fig:fig01}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on aggration and on average flow duration is more discriminative and shows a better separation between normal and anomalous classes.
+Therefore, such data can improve the anomaly detection by learning algorithms.
 
@@
      \caption{Average flow duration for scenario 12, aggregated by 0.15 seconds.}
-     \label{fig:fig03}
+     \label{fig:fig02}
 \end{figure}
 
-The data is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features. Each scenario
+The resulting data set is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features, and is divided into scenarios, where each $i$-th scenario, presented by ID column in Table \ref{tab:tab01}, are represented as $\boldsymbol{X}_i$ and can be denoted as $\boldsymbol{X}_{18-2}$, for example.
 

### 2019-05-14 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on data model

@@
 \usepackage{longtable}
-\usepackage{mathtools}
 \usepackage{listings}
@@
 \usepackage{lipsum}
+\usepackage{mathtools}
 \usepackage{amsthm,amsmath}
@@
 \usepackage{cleveref}
+
+% norm
 
@@
 Section \ref{sec:datamodel} presents the data model and the evaluated data set.
-Section \ref{sec:m_rpca} describes the proposed approach for botnet detection.
+Section \ref{sec:mdrs} describes the proposed approach for botnet detection.
 Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
@@
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
+The Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries.
 
@@
 	\item \textbf{total-flow-count :} Count of all flows.
-	\item \textbf{normal-flow-count:} Count of known flows, from controlled assets of the network, such routers, proxies, switches or machines that are part of the monitoring \cite{garcia2014identifying}.
+	\item \textbf{known-flow-count:} Count of known flows, from controlled assets of the network, such routers, proxies, switches or machines that are part of the monitoring \cite{garcia2014identifying}.
 	\item \textbf{avg-duration:} Average time of flows.
@@
 The resulting data set is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features, and is divided into scenarios, where each $i$-th scenario, presented by ID column in Table \ref{tab:tab01}, are represented as $\boldsymbol{X}_i$ and can be denoted as $\boldsymbol{X}_{18-2}$, for example.
+The normal traffic, required for trainning the semi-supervised approach, is denoted as $\boldsymbol{X}_i^r$
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Robust PCA and Higher Moments Distances for Botnet Detection}
-\label{sec:m_rpca}
+\section{Moment Distances from Robust Subspace for Botnet Detection}
+\label{sec:mdrs}
 
-This section describes the proposed approach for botnet detection by means of moment distances from robust estimates.
-We propose to learn a robust subspace $\textbf{L}$ from normal traffic $\textbf{X}_n$ for computing robust moments, i.e.
-the mean $\textbf{l}$, skewness $\textbf{s}$ and kurtosis $\textbf{k}$, in order to evaluate the distance $\textbf{d}$ of new observations $\textbf{X}_o$ of these moments and classify as anomalous the observations with the largest distances.
+This section describes the proposed approach for botnet detection by means of a disntance analysis between moments computed from a robust subspace and new observations of network traffic.
 
-Initially, the data $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ is split into normal data $\textbf{X}_n$ for training from normal data, and for $\textbf{X}_v$ and $\textbf{X}_t$ for respectively cross-validation and testing from contaminated data
+Robust subspace learning can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
+We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}_i^r$ for computing robust moments, i.e.
+the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^t$ of these moments and classify as anomalous the observations with the largest distances.
 
-It was shown that the RPCA problem can be solved via convex optimization that minimizes a combination of the nuclear norm and the $l_1$-norm, and that the method of augmented Lagrange multipliers (ALM) has been adopted to solve this convex problem.
+Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparce matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$.
+This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}.
+We adotp RPCA with ALM, which can be formulated as
 
-Robust PCA aims to recover a low-rank matrix $\textbf{L}$ from corrupted measurements $\textbf{M} = \textbf{L} + \textbf{S}$.
-This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP),[1] Stable PCP,[2] Quantized PCP,[3] Block based PCP,[4] and Local PCP.[5] Then, optimization methods are used such as the Augmented Lagrange Multiplier Method (ALM[6]), Alternating Direction Method (ADM[7]), Fast Alternating Minimization (FAM[8]) or Iteratively Reweighted Least Squares (IRLS [9][10][11]).
+\begin{equation}\label{eq:eq01}
+	\min_{\boldsymbol{L,S}}\left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
+\end{equation}
+where $\boldsymbol{Y}$ are the Lagrange multipliers and have the constraing $\boldsymbol{C = D - A - E}$...
 
-Therefore we have $\textbf{L}$...
+Initially, the data $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ is split into normal data $\textbf{X}^r$ for rosbust subspace learning, and into $\textbf{X}^t$ for testing the anomaly detection on contaminated data, that can also be viewed as new observations of a production environment.
 
@@
 The general formula of the $r$-th moment can be expressed as:
-\begin{equation}\label{eq:eq05}
+\begin{equation}\label{eq:eq02}
 	m_r = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^r. 
@@
 Therefore, the Skewness is calculated as
-\begin{equation}\label{eq:eq06}
+\begin{equation}\label{eq:eq03}
 	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
@@
 and Kurtosis as
-\begin{equation}\label{eq:eq07}
-	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2} 
+\begin{equation}\label{eq:eq04}
+	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2} .
 \end{equation}
 
-For extending Fast-MCD, We initially modify the C-step Algorithm \ref{alg:alg01} to calculate Skewness and Kurtosis after to minimize $det(\boldsymbol{\hat{S}})$, accordingly to Algorithm \ref{alg:alg05}.
+We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq01} for $\textbf{X}^r$.
+We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies betweem new observations and moments computed from a robust subspace computed by RPCA.
+The classical equation of we modify
 
-\begin{algorithm}
-	\label{alg:alg05}
-	\scriptsize
-	\SetAlgoLined
-	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
-	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
-	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
-		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
-			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
-		}
-		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
-		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
-		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
-		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
-		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
-	}
-	Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
-	\caption{C-Step with higher moments}
-\end{algorithm}
-
-The Algorithm \ref{alg:alg06} describes the Fast-MCD with higher moments when $n \leq 600$, through the modification of the original Fast-MCD for computing Skewness and Kurtosis from the robust estimate.
-
-\begin{algorithm}
-	\label{alg:alg06}
-	\scriptsize
-	\SetAlgoLined
-	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
-	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n \leq 600$\;
-	\For{500 times}{
-		Construct initial $h$-subset $\boldsymbol{h}_1$ (Algorithm \ref{alg:alg02})\;
-		Carry out two C-steps with higher moments (Algorithm \ref{alg:alg05}) resulting $\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$\;
-	}		
-	Put $\boldsymbol{h}_1$ as the 10 lowest $det(\boldsymbol{\hat{S}_2})$ computed above\;
-	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
-		Carry out C-steps with higher moments (Algorithm \ref{alg:alg05})\;
-	}
-	\caption{Fast-MCD with higher moments when $n \leq 600$}
-\end{algorithm}
-
-The Algorithm \ref{alg:alg07} presents the Fast-MCD with higher moments when $n > 600$, which modifies the original Fast-MCD for computing of Skewness and Kurtosis from the robust estimate data, through the adoption of the C-Step with higher moments.
-The Algorithm \ref{alg:alg07} is described below.
-
-\begin{algorithm}
-	\label{alg:alg07}
-	\scriptsize
-	\SetAlgoLined
-	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
-	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n > 600$\;
-	Construct up to 5 disjoint random subsets of size $n_{sub} = 300$\;
-	\For{each subset, repeat 100 times}{
-		Construct initial $\boldsymbol{h}_1$ of size $h_{sub} = [n_{sub}(h/n)]$\;
-		Carry out two C-steps with higher moments of $n_{sub}$ and $h_{sub}$\;
-		Keep the 10 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub}, \boldsymbol{\bar{s}}_{sub}, \boldsymbol{\bar{k}}_{sub})$\;
-	}		
-	Pool the subsets, yielding the merged set of $n_{merged} = 1500$\;
-	\For{50 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$}{
-		Carry out two C-steps with higher moments of $n_{merged}$ and $h_{merged} = [n_{merged}(h/n)]$\;
-		Keep the 10 best $(\boldsymbol{\bar{x}}_{merged}, \boldsymbol{\hat{S}}_{merged}, \boldsymbol{\bar{s}}_{merged}, \boldsymbol{\bar{k}}_{merged})$\;
-	}
-	\For{the full data set and $m_{full}$ best results}{
-		Take several C-steps, using $n$ and $h$\;
-		Keep the best final result $(\boldsymbol{\bar{x}}_{full}, \boldsymbol{\hat{S}}_{full}, \boldsymbol{\bar{s}}_{full}, \boldsymbol{\bar{k}}_{full})$\;
-	}
-	\caption{Fast-MCD with higher moments when $n > 600$}
-\end{algorithm}
-
-We also propose to use Mahalanobis Distance for detecting anomalies through an approach based on robust Skewness and Kurtosis, instead of the location and scatter that are used by classical approaches.
+The classifical MD relies on robust mean and robust covariace matrix, here our contribution is compute the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA, instead of classifical approach, that relies on MCD or Fast-MCD.
 
@@
 where $\boldsymbol{k}_{\boldsymbol{x}} = \boldsymbol{x} - \boldsymbol{\bar{k}}(\boldsymbol{x})$ and $\boldsymbol{\bar{k}}(\boldsymbol{x})$ is the Kurtosis of $\boldsymbol{x}$.
+
+% \begin{algorithm}
+% 	\label{alg:alg01}
+% 	\scriptsize
+% 	\SetAlgoLined
+% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
+% 	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
+% 	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
+% 		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
+% 			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
+% 		}
+% 		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
+% 		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
+% 		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
+% 		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
+% 		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
+% 	}
+% 	Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
+% 	\caption{C-Step with higher moments}
+% \end{algorithm}
+
+Only for evaluation pourpose, we use part of the contaminated data, denoted as $\textbf{X}^v$, for cross-validation, in order to evaluate the contamination rate $c$ of the target data.
+The contamination rate parameter is widely adopted for well stablished outlier detection algorithms \cite{zhao2019pyod}, and refers to the percentage of observations that are classified as anomalous, which can be well known for some areas, or can be computed by cross-validation or can be assumed according to previous observations.
+
+Therefore we have $\textbf{L}$...
 

### 2019-05-14 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
mdrs

@@
 
-The general formula of the $r$-th moment can be expressed as:
+Let the mean $\bar{\boldsymbol{x}} \in \mathbb{R}^{1 \times p}$ be
+
 \begin{equation}\label{eq:eq02}
-	m_r = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^r. 
+	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} \boldsymbol{x}_i, 
 \end{equation}
 
-Therefore, the Skewness is calculated as
+and the covariance matrix $\boldsymbol{\bar{\Sigma}} \in \mathbb{R}^{p \times p}$ be
+
+\begin{equation}\label{eq:eq03}
+	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
+\end{equation}
+
+The general formula of the $u$-th moment can be expressed as
+\begin{equation}\label{eq:eq02}
+	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
+\end{equation}
+
+Therefore, the skewness is calculated by
 \begin{equation}\label{eq:eq03}
@@
 
-and Kurtosis as
+and the kurtosis as
 \begin{equation}\label{eq:eq04}
@@
 We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq01} for $\textbf{X}^r$.
-We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies betweem new observations and moments computed from a robust subspace computed by RPCA.
-The classical equation of we modify
+We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies betweem new observations and the moments calculated from a robust subspace computed by RPCA.
 
-The classifical MD relies on robust mean and robust covariace matrix, here our contribution is compute the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA, instead of classifical approach, that relies on MCD or Fast-MCD.
-
-We propose to modify the Equation \ref{eq:eq01} to implement a Skewness-based Mahalanobis Distance, as follows:
+The classical Mahalanobis Distance is defined as		
+	\begin{equation}\label{eq:eq05}
+		\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
+	\end{equation}
+where $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector of known observations, also referred as location, and $\boldsymbol{\bar{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter.
+The classifical MD is usually relies on robust mean and robust covariace matrix for outlier detection, which are commonly computed by Fast-MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}.
+Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA.
+We also propose to extend the Equation \ref{eq:eq05} to implement a Skewness-based Mahalanobis Distance, as follows:
 
 \begin{equation}\label{eq:eq08}
-	\boldsymbol{d}(\boldsymbol{s}_{\boldsymbol{x}}, \bar{\boldsymbol{s}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{s}_{\boldsymbol{x}} - \bar{\boldsymbol{s}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{s}_{\boldsymbol{x}} - \bar{\boldsymbol{s}})^\prime},
+	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}.
 \end{equation}
 
-where $\boldsymbol{s}_{\boldsymbol{x}} = \boldsymbol{x} - \boldsymbol{\bar{s}}(\boldsymbol{x})$ and $\boldsymbol{\bar{s}}(\boldsymbol{x})$ is the Skewness of $\boldsymbol{x}$.
-
-We also propose to modify the Equation \ref{eq:eq04} to implement a Kurtosis-based Mahalanobis Distance, as follows:
+We finally propose to extend the Equation \ref{eq:eq05} to implement a Kurtosis-based Mahalanobis Distance, as follows:
 
 \begin{equation}\label{eq:eq09}
-	\boldsymbol{d}(\boldsymbol{k}_{\boldsymbol{x}}, \bar{\boldsymbol{k}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{k}_{\boldsymbol{x}} - \bar{\boldsymbol{k}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{k}_{\boldsymbol{x}} - \bar{\boldsymbol{k}})^\prime},
+	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}.
 \end{equation}
 
-where $\boldsymbol{k}_{\boldsymbol{x}} = \boldsymbol{x} - \boldsymbol{\bar{k}}(\boldsymbol{x})$ and $\boldsymbol{\bar{k}}(\boldsymbol{x})$ is the Kurtosis of $\boldsymbol{x}$.
+The distances shall be computed and evaluated separatelly for botnet detection, in order to verify if the scenarios have different patterns that can be explored differently by distances of each moment individually.
+
+The Algorithmm \ref{alg:alg01} presents a overview the experiment
+\begin{algorithm}
+	\label{alg:alg01}
+	\scriptsize
+	\SetAlgoLined
+	\KwResult{$\boldsymbol{\bar{t}}$}
+	Given the data set $\boldsymbol{X}$ split into 13 scenarios $\boldsymbol{X}_i$, and susequently split into normal observations $\boldsymbol{X}_i^n$ and contaminated observations $\boldsymbol{X}_i^t$\;
+	\For{$eacch scenario$}{
+		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
+			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
+		}
+		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
+		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
+		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
+		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
+		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
+	}
+	Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
+	\caption{C-Step with higher moments}
+\end{algorithm}
+
+The expected result $\boldsymbol{\bar{t}}$ refers to the predicted classification of the new observations as normal or anomalous.
 
 % \begin{algorithm}
-% 	\label{alg:alg01}
+% 	\label{alg:alg03}
 % 	\scriptsize
 % 	\SetAlgoLined
-% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$, $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$}
-% 	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
+% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
+% 	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n \leq 600$\;
+% 	\For{500 times}{
+% 		Construct initial $h$-subset $\boldsymbol{h}_1$ (Algorithm \ref{alg:alg02})\;
+% 		Carry out two C-steps (Algorithm \ref{alg:alg01}) resulting $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$\;
+% 	}		
+% 	Put $\boldsymbol{h}_1$ as the 10 lowest $det(\boldsymbol{\hat{S}_2})$ computed above\;
 % 	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
-% 		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
-% 			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
+% 		Carry out C-steps (Algorithm \ref{alg:alg01})\;
 % 	}
-% 		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
-% 		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
-% 		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
-% 		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
-% 		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
-% 	}
-% 	Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
-% 	\caption{C-Step with higher moments}
+% 	\caption{Fast-MCD when $n \leq 600$}
 % \end{algorithm}
@@
 The $c$ observations with largest $\boldsymbol{d}(i)$ are classified as anomalous.
-
-\subsection{MOS Schemes}
-\label{sec:prop_MOSSchemes}
-
-\subsection{Eigen Similarity Analysis}
-\label{sec:prop_EigenSimilarityAnalysis}
-
-\subsubsection{Time Similarity Analysis}
-\label{sec:prop_TimeSimilarityAnalysis}
-
-\subsubsection{Port Similarity Analysis}
-\label{sec:prop_PortSimilarityAnalysis}
 
@@
 
-% Mahalanobis Distance is defined as		
-% 	\begin{equation}\label{eq:eq01}
-% 		d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}.
-% 	\end{equation}
-% $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector, also referred as location, of known observations and $\boldsymbol{\hat{S}}$ is the covariance matrix, also referred as scatter, of known observations.
-
 % Classical estimates can be so strongly affected by contamination that diagnostic tools, such as the Mahalanobis Distances, become unable to detect the outliers.
@@
 Let $\boldsymbol{h} \subset \{1,...,n\}$ with $|\boldsymbol{h}| = h$, where $\boldsymbol{h}$ is a subset of $h$ indexes from $n$ observations.
-
-% Let the mean $\bar{\boldsymbol{x}} \in \mathbb{R}^{1 \times p}$ be
-
-% \begin{equation}\label{eq:eq02}
-% 	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} \boldsymbol{x}_j, 
-% \end{equation}
-
-% the covariance matrix $\boldsymbol{\hat{S}} \in \mathbb{R}^{p \times p}$ be
-
-% \begin{equation}\label{eq:eq03}
-% 	\boldsymbol{\hat{S}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} (\boldsymbol{x}_j - \bar{\boldsymbol{x}})(\boldsymbol{x}_j - \bar{\boldsymbol{x}})^\prime,
-% \end{equation}
 

### 2019-05-15 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on results

@@
 Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the correlations between the variables and their mean values.
+The MD is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
+C.
+Mahalanobis in 1936 \cite{mahalanobis1936md}.
+It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\bar{\Sigma}}$ of $\boldsymbol{X}$.
 MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and new observations can reveal anomalies.
@@
 The Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries.
+We define the operator $[ \mathord{\cdot} ]^n$, which denotes the $n$ largest values of a vector.
 
@@
 
-We refer to Garcia \cite{garcia2014identifying} and Garcia \emph{et al.} \cite{garcia2014empirical} for a detailed description of the performed attacks and C\&C flows, also including the topology of the adopted testbed, rules for classifing normal flows, and an analysis of behaviors or patterns of the malware's traffic.
+We refer to Garcia \cite{garcia2014identifying} and Garcia \emph{et al.} \cite{garcia2014empirical} for a detailed description of the performed attacks and C\&C flows, also including the topology of the adopted testbed, rules for classifying normal flows, and an analysis of behaviors or patterns of the malware's traffic.
 
@@
 Related works proposed some features for improving network attack detection in various scenarios and approaches\cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey}.
-Inpired by these proposals and based on previous experience \cite{vieira2017model, galibus2017offline} and cross validation, we propose the following features for improving the anomaly detection by means of a method based on higher moments distances from robust subspace.
+Inspired by these proposals and based on previous experience \cite{vieira2017model, galibus2017offline} and cross validation, we propose the following features for improving the anomaly detection by means of a method based on higher moments distances from robust subspace.
 
@@
 
-The proposed features are counting and average calculations, therefore it is necessary to adopt time windows to aggragate flows and calculate the features.
+The proposed features are counting and average calculations, therefore it is necessary to adopt time windows to aggregate flows and calculate the features.
 We evaluate windows of 0.15 and 0.25 seconds, considering account that lower windows reduce the meantime between the data collection and results of the botnet detection approach, as well as permit to have more details to exactly identify malicious flows and attackers.
@@
 Figure \ref{fig:fig02} shows the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 seconds.
-In comparison to Figure \ref{fig:fig01}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on aggration and on average flow duration is more discriminative and shows a better separation between normal and anomalous classes.
+In comparison to Figure \ref{fig:fig01}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on aggregation and on average flow duration is more discriminative and shows a better separation between normal and anomalous classes.
 Therefore, such data can improve the anomaly detection by learning algorithms.
@@
 The resulting data set is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features, and is divided into scenarios, where each $i$-th scenario, presented by ID column in Table \ref{tab:tab01}, are represented as $\boldsymbol{X}_i$ and can be denoted as $\boldsymbol{X}_{18-2}$, for example.
-The normal traffic, required for trainning the semi-supervised approach, is denoted as $\boldsymbol{X}_i^r$
+The normal traffic, required for training the semi-supervised approach, is denoted as $\boldsymbol{X}_i^r$
 
@@
 
-This section describes the proposed approach for botnet detection by means of a disntance analysis between moments computed from a robust subspace and new observations of network traffic.
+This section describes the proposed approach for botnet detection by means of a distance analysis between moments computed from a robust subspace and new observations of network traffic.
 
 Robust subspace learning can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
-We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}_i^r$ for computing robust moments, i.e.
-the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^t$ of these moments and classify as anomalous the observations with the largest distances.
+We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}_i^r$ for computing $\textbf{L}_i^r$, $\textbf{S}_i^r$ and the robust moments, i.e.
+the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^t$ to these moments and classify as anomalous the observations with the largest distances.
 
@@
 This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}.
-We adotp RPCA with ALM, which can be formulated as
+We adopt RPCA with ALM, which can be formulated as
 
@@
 \end{equation}
-where $\boldsymbol{Y}$ are the Lagrange multipliers and have the constraing $\boldsymbol{C = D - A - E}$...
-
-Initially, the data $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ is split into normal data $\textbf{X}^r$ for rosbust subspace learning, and into $\textbf{X}^t$ for testing the anomaly detection on contaminated data, that can also be viewed as new observations of a production environment.
+where $\boldsymbol{Y}$ are the Lagrange multipliers and have the constraint $\boldsymbol{C = D - A - E}$...
 
@@
 \begin{equation}\label{eq:eq02}
-	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} \boldsymbol{x}_i, 
+	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i, 
 \end{equation}
@@
 \begin{equation}\label{eq:eq03}
-	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{j\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
+	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
 \end{equation}
@@
 The general formula of the $u$-th moment can be expressed as
-\begin{equation}\label{eq:eq02}
+\begin{equation}\label{eq:eq04}
 	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
@@
 Therefore, the skewness is calculated by
-\begin{equation}\label{eq:eq03}
+\begin{equation}\label{eq:eq05}
 	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
@@
 and the kurtosis as
-\begin{equation}\label{eq:eq04}
-	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2} .
+\begin{equation}\label{eq:eq06}
+	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}.
 \end{equation}
 
-We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq01} for $\textbf{X}^r$.
-We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies betweem new observations and the moments calculated from a robust subspace computed by RPCA.
+We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq01} for $\textbf{X}^r$.
+We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between new observations and the moments calculated from a robust subspace computed by RPCA.
 
 The classical Mahalanobis Distance is defined as		
-	\begin{equation}\label{eq:eq05}
+\begin{equation}\label{eq:eq07}
 	\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
@@
 where $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector of known observations, also referred as location, and $\boldsymbol{\bar{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter.
-The classifical MD is usually relies on robust mean and robust covariace matrix for outlier detection, which are commonly computed by Fast-MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}.
+The classical MD usually relies on robust mean and robust covariance matrix for outlier detection, which are commonly computed by Fast-MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}.
 Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA.
@@
 
-The distances shall be computed and evaluated separatelly for botnet detection, in order to verify if the scenarios have different patterns that can be explored differently by distances of each moment individually.
+The distances $\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})$, $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})$ and $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})$ shall be computed and evaluated separately for botnet detection.
+Therefore, the robust subspace learning from normal data combined to MD can be called md-RPCA when using mean-based MD, sd-RPCA when using skewness-based MD, and kd-RPCA when using kurtosis-based MD.
 
-The Algorithmm \ref{alg:alg01} presents a overview the experiment
+The contamination rate parameter is widely adopted for well established outlier detection algorithms \cite{zhao2019pyod}, and refers to the percentage of observations that are classified as anomalous.
+The contamination rate can be well known for some areas, or can be computed by cross-validation or can be assumed according to previous observations.
+In our proposal, the contamination define the rate of highest distances $\boldsymbol{d}$ that shall be classified as anomalous.
+Therefore, the resulting $\boldsymbol{\bar{t}}$ refers to observations classified as anomalous according to $c$.
+
+The Algorithm \ref{alg:alg01} presents how the proposed Moment Distances from Robust Subspace for Botnet Detection can be algorithmically described.
 \begin{algorithm}
 	\label{alg:alg01}
-	\scriptsize
 	\SetAlgoLined
 	\KwResult{$\boldsymbol{\bar{t}}$}
-	Given the data set $\boldsymbol{X}$ split into 13 scenarios $\boldsymbol{X}_i$, and susequently split into normal observations $\boldsymbol{X}_i^n$ and contaminated observations $\boldsymbol{X}_i^t$\;
-	\For{$eacch scenario$}{
-		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
-			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
-		}
-		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
-		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
-		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
-		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
-		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
-	}
-	Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
-	\caption{C-Step with higher moments}
+	Given $\boldsymbol{X}$ split into $\boldsymbol{X}^n$ and $\boldsymbol{X}_i^t$\;
+	$\min_{\boldsymbol{L}^n, \boldsymbol{S}^n}\left\|\boldsymbol{L}^n\right\|_* + \lambda\left\|\boldsymbol{S}^n\right\|_1 + \langle \boldsymbol{Y}^n, \boldsymbol{X}^n - \boldsymbol{L}^n - \boldsymbol{S}^n  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X}^n - \boldsymbol{L}^n - \boldsymbol{S}^n\right\|_F^2$\;
+	$\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i$, for $\boldsymbol{L}^n$\;
+	$\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime$, for $\boldsymbol{L}^n$\;
+	$\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}}$, for $\boldsymbol{L}^n$\;
+	$\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}$, for $\boldsymbol{L}^n$\;
+	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
+	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}$\;
+	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}$\;
+	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})]^c$\;
+	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})]^c$\;
+	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})]^c$\;
+	\caption{Moment Distances from Robust Subspace}
 \end{algorithm}
 
-The expected result $\boldsymbol{\bar{t}}$ refers to the predicted classification of the new observations as normal or anomalous.
-
-% \begin{algorithm}
-% 	\label{alg:alg03}
-% 	\scriptsize
-% 	\SetAlgoLined
-% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
-% 	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n \leq 600$\;
-% 	\For{500 times}{
-% 		Construct initial $h$-subset $\boldsymbol{h}_1$ (Algorithm \ref{alg:alg02})\;
-% 		Carry out two C-steps (Algorithm \ref{alg:alg01}) resulting $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$\;
-% 	}		
-% 	Put $\boldsymbol{h}_1$ as the 10 lowest $det(\boldsymbol{\hat{S}_2})$ computed above\;
-% 	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
-% 		Carry out C-steps (Algorithm \ref{alg:alg01})\;
-% 	}
-% 	\caption{Fast-MCD when $n \leq 600$}
-% \end{algorithm}
-
-Only for evaluation pourpose, we use part of the contaminated data, denoted as $\textbf{X}^v$, for cross-validation, in order to evaluate the contamination rate $c$ of the target data.
-The contamination rate parameter is widely adopted for well stablished outlier detection algorithms \cite{zhao2019pyod}, and refers to the percentage of observations that are classified as anomalous, which can be well known for some areas, or can be computed by cross-validation or can be assumed according to previous observations.
-
-Therefore we have $\textbf{L}$...
-
-Fast-MCD is commonly used as \textbf{supervised} or \textbf{unsupervised} algorithm for anomaly detection.
-
-\textbf{Supervised} approach:
-
-Compute Fast-MCD for training data with known normal and anomalous data.
-Select the best threshold $t$ for detecting known anomalies according to largest distances $\boldsymbol{d}(i)$.
-Save the best $t, \boldsymbol{\bar{x}}$ and $\boldsymbol{\hat{S}}$.
-Compute $d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}})$ for new observations.
-Classify observations with $\boldsymbol{d}(i)$ higher than $t$ as anomalous.
-
-\textbf{Unsupervised} approach:
-
-Given a known contamination $c$, which is the percentage of anomalies in a data set.
-Compute Fast-MCD and $d(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\hat{S}})$ for new observations.
-Classify the $c$ observations with largest distances as anomalous.
-
-Finally, We propose the following \textbf{Semi Supervised} approach for Fast-MCD based anomaly detection:
-
-Compute Fast-MCD for training data containing only known normal observations.
-Save the fitted $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$.
-Carry out cross validation on known contaminated data in order to identify the contamination $c$ that best classify anomalies and normal data.
-Compute $d(\boldsymbol{x},\bar{\boldsymbol{x}}_{normal}, \boldsymbol{\hat{S}}_{normal})$ for new observations.
-The $c$ observations with largest $\boldsymbol{d}(i)$ are classified as anomalous.
+The steps between 1 and 6 of the Algorithm \ref{alg:alg01} are the training from normal data $\boldsymbol{X}^n$, while the steps between 7 and 12 aim the anomaly detection on new contamianted observations.
+Note that the steps 7 and 10 refers to the md-RPCA, while the steps 8 and 11 refers to the sd-RPCA, and the steps 9 and 12 refers to the kd-RPCA.
+It is also important to observe that the anomaly detection from new observations does not require new robust subspace learning, it is only necessary to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous.
 
@@
 
-Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches, which are K-means and Gaussian Mixture Model (GMM) \cite{gaddam2007kmeans,moustafa2019holistic}, respectively.
-We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
-In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of CTU-13 individually, in a semi-supervised approach that does not rely on training data with anomalies.
-
 This section presents the performed experiments and the acquired results.
 First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
-Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario.
-In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated data set.
-Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario.
-Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 data set.
+Then, Section \ref{sec:results} shows the results of the proposed approach and compares to selected anomaly detection algorithms.
 
@@
 
-\textbf{Experiment Plan:}
+Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training the model, while the latter does not require labeled data or training.
+However, the availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or unkown events.
 
-Given a contaminated data set, split it into normal (60\%), cross validation (20\% normal + 50\% anomalous) and test (20\% normal + 50\% anomalous).
+Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches.
+The proposed approach is semi-supervised and relies on normal data $\boldsymbol{X}^n$ and on contamination rate $c$ for training and anomalous detection, respectively.
+
+Only for evaluation purpose, we use part of the contaminated data, denoted as $\textbf{X}^v$, for cross-validation, in order to estimate the contamination rate $c$ that best detect anomalies from $\textbf{X}^v$.
+However, we assume that the $c$ is well known or can be estimated for real world problems of anomaly detection.
+
+Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}.
+Gaussian Mixture Model (GMM) is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.
+
+This experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches.
+As clustering-based method, we compare to K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007kmeans,moustafa2019holistic}.
+It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, and that algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}.
+Therefore, we compare our proposal to GMM.
+We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
+
+In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach for all botnet malwares of the CTU-13.
+Therefore, the data set and scenarios for our evaluation can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, ...
+, \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to Table \ref{tab:tab01}.
+
+Given a contaminated data set $\boldsymbol{X}$, split it into $\boldsymbol{X}^n$ containing 60\% of the the normal data, and into cross validation $\boldsymbol{X}^v$, composed by 20\% of the normal and 50\% of the anomalous data, and into test $\boldsymbol{X}^t$ containing 20\% of the normal and 50\% of the anomalous data.
+
+Evaluate if the code is correct and implement it as funcions in my code
+
+try other distributions between train cv and test
+
+Explain tha numbers of anomalies and normal for CV and test
 
@@
 Test the F-measure of anomaly detection for new observations through $d(\boldsymbol{x},\bar{\boldsymbol{x}}_0, \boldsymbol{\hat{S}}_0)$ and $c$.
-
-Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}.
-We rely on this for our implementation
-
-GMM is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.
-
-The Mahalanobis Distance is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
-C.
-Mahalanobis in 1936 \cite{mahalanobis1936md}.
-It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\hat{S}}$ of $\boldsymbol{X}$.
 
@@
 \textbf{Data splitting for training, cross validation and testing}
-
-Gargia et al. \cite{garcia2014empirical} propose a division of CTU-13's scenarios into train and test, for evaluation of network attack detection algorithms. 
-However, we decide to evaluate our proposed approach for each scenario individually, considering that in the division proposed by Gargia et al.
-some bots are present only for training and not for testing.
-Although our proposal does not rely on anomalous data for training, we desire to evaluate our proposal to detect all types of bots captured by CTU-13 data set.
-
-Why GMM?
-
-As clustering-based method, we selected K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007kmeans,moustafa2019holistic}.
-It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}.
 
@@
 
-This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
+This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
 Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.
@@
 % • Scenario 9 corresponds to a case where 10 local hosts were infected by a spamming botnet. More than 600 spams were sent over 5 hours.
+
+% Initially, the data $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ is split into normal data $\textbf{X}^r$ for robust subspace learning, and into $\textbf{X}^t$ for testing the anomaly detection on contaminated data, that can also be viewed as new observations of a production environment.

### 2019-05-15 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on results

@@
 
-Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the correlations between the variables and their mean values.
+Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the covariances between the variables and their mean values.
 The MD is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
@@
 
-Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparce matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$.
+Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparse matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$.
 This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}.
-We adopt RPCA with ALM, which can be formulated as
 
+According to Wright et al.
+\cite{rpca_convex_opt}, under rather broad conditions, as long as the error matrix S is sufficiently sparse, it is possible to recover a low-rank matrix by solving the following convex optimization problem:
+
+\begin{equation}\label{eq:pcp}
+	(\boldsymbol{\hat{L}}, \boldsymbol{\hat{S}})\leftarrow min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}
+\end{equation}
+\begin{center} subject to: $\boldsymbol{X} = \boldsymbol{L} + \boldsymbol{S}$ \end{center}
+
+\begin{equation}\label{eq:nuclear_norm}
+    \left \| \boldsymbol{L} \right \|_{*} = \sum_{i} \sigma_{i}(\boldsymbol{L})
+\end{equation}
+
+\begin{equation}\label{eq:l1_norm}
+    \left \| \boldsymbol{S} \right \|_{1} = \sum_{ij} \left | \boldsymbol{S}_{ij} \right |
+\end{equation}
+where $\lambda$ a positive weighting parameter, which determines the sparsity of $\boldsymbol{S}$.
+
+ Before ALM, some methods were proposed to solve that convex optimization problem, such as Iterative Thresholding (IT) and Accelerated Proximal Gradient (APG).
+However, according to \cite{alm} both have scalability problems and require a large number of iterations to converge.
+The Augmented Lagrange Multiplier is proven to have a \emph{Q-linear} convergence speed and experimental results show that ALM is five times faster than APG, which in theory is sub-linear \cite{alm}.
+Furthermore, ALM reaches more accurate results with less iterations.
+
+% We adopt RPCA with ALM, which can be formulated as
+
+Augmented Lagrange Multiplier formula is given by:
 \begin{equation}\label{eq:eq01}
-	\min_{\boldsymbol{L,S}}\left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
+	l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}) = \left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
 \end{equation}
-where $\boldsymbol{Y}$ are the Lagrange multipliers and have the constraint $\boldsymbol{C = D - A - E}$...
+%where $\boldsymbol{Y}$ are the Lagrange multipliers and have the constraint %$\boldsymbol{C = D - A - E}$...
+
+where
+
+\begin{equation}\label{eq:y_cond}
+	\left\{\begin{matrix} \boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k}) \\ \boldsymbol{Y}_{k+1} \in argmin_{\boldsymbol{L,S} \in R^{m \times n}} \{l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}_{k})\} \end{matrix}\right.
+\end{equation}
+$\boldsymbol{Y}$ the multiplier of the linear constraint and $\mu$ the parameter for the violation of the linear constraint \cite{adm}.
+
+We adopt RPCA with ADM, which, generally speaking, is a practical improvement of the classical Augmented Lagrangian method for solving convex programming problem with linear constraints, by fully taking advantage of its high-level separable structure \cite{adm}.
+ADM minimizes $\boldsymbol{L}$ and $\boldsymbol{S}$ variables serially by solving the following problems to generate the new iterate:
+
+\begin{equation}\label{eq:adm_eq}
+	\left\{\begin{matrix}
+	\boldsymbol{L}_{k+1} \in argmin_{\boldsymbol{L} \in R^{m \times n}}\{l(\boldsymbol{L}, \boldsymbol{S}_{k}, \boldsymbol{Y}_{k})\}\\ 
+	\boldsymbol{S}_{k+1} \in argmin_{\boldsymbol{S} \in R^{m \times n}}\{l(\boldsymbol{L}_{k+1}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}\\ 
+	\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k})
+	\end{matrix}\right.
+\end{equation}
 
@@
 The general formula of the $u$-th moment can be expressed as
-\begin{equation}\label{eq:eq04}
+\begin{equation}\label{eq:eq02}
 	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
@@
 Therefore, the skewness is calculated by
-\begin{equation}\label{eq:eq05}
+\begin{equation}\label{eq:eq03}
 	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
@@
 and the kurtosis as
-\begin{equation}\label{eq:eq06}
+\begin{equation}\label{eq:eq04}
 	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}.
@@
 The classical Mahalanobis Distance is defined as		
-\begin{equation}\label{eq:eq07}
+\begin{equation}\label{eq:eq05}
 	\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
@@
 In our proposal, the contamination define the rate of highest distances $\boldsymbol{d}$ that shall be classified as anomalous.
-Therefore, the resulting $\boldsymbol{\bar{t}}$ refers to observations classified as anomalous according to $c$.
+Finally, the resulting $\boldsymbol{\bar{t}}$ refers to the estimated classification of new observations, as normal or anomalous, according to $c$.
 
@@
 	\label{alg:alg01}
+	\scriptsize
 	\SetAlgoLined
@@
 	$\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}$, for $\boldsymbol{L}^n$\;
-	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
+
+	$\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
 	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}$\;
 	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}$\;
-	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})]^c$\;
-	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})]^c$\;
-	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})]^c$\;
-	\caption{Moment Distances from Robust Subspace}
+
+	% Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
+	% Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
+	% Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
+	% Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;	
+	% Compute $\boldsymbol{\bar{s}}_2$ and $\boldsymbol{\bar{k}}_2$ for last $\boldsymbol{h}_2$\;
+	\caption{Moment Distances from Robust Subspace for Botnet Detection}
 \end{algorithm}

### 2019-05-16 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on results

--- 2019-05-15 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
@@
 The Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries.
-We define the operator $[ \mathord{\cdot} ]^n$, which denotes the $n$ largest values of a vector.
+We also define the operator $\langle \mathord{\cdot} \rangle$, that is the standard trace inner product, and $[ \mathord{\cdot} ]^n$, which denotes the $n$ largest values of a vector.
 
@@
      \centering
-     \includegraphics[width=8cm]{figures/raw_distplot_capture20110812_Dur.png}
+     \includegraphics[width=10cm]{figures/raw_distplot_capture20110812_Dur.png}
      \caption{Distribution of flow duration for scenario 12.}
@@
      \centering
-     \includegraphics[width=8cm]{figures/agg_distplot_0_15s_12_avg_duration.png}
+     \includegraphics[width=10cm]{figures/agg_distplot_0_15s_12_avg_duration.png}
      \caption{Average flow duration for scenario 12, aggregated by 0.15 seconds.}
@@
 The resulting data set is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features, and is divided into scenarios, where each $i$-th scenario, presented by ID column in Table \ref{tab:tab01}, are represented as $\boldsymbol{X}_i$ and can be denoted as $\boldsymbol{X}_{18-2}$, for example.
-The normal traffic, required for training the semi-supervised approach, is denoted as $\boldsymbol{X}_i^r$
+The normal traffic, required for training the semi-supervised approach, is denoted as $\boldsymbol{X}^r$
 
@@
 Robust subspace learning can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
-We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}_i^r$ for computing $\textbf{L}_i^r$, $\textbf{S}_i^r$ and the robust moments, i.e.
+We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}^r$ for computing $\textbf{L}^r$, $\textbf{S}^r$ and the robust moments, i.e.
 the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^t$ to these moments and classify as anomalous the observations with the largest distances.
 
-Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparce matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$.
+Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparse matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$.
 This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}.
-We adopt RPCA with ALM, which can be formulated as
 
+According to Wright et al.
+\cite{wright2009robust}, under rather broad conditions, as long as the error matrix S is sufficiently sparse, it is possible to recover a low-rank matrix by solving the following convex optimization problem:
 \begin{equation}\label{eq:eq01}
-	\min_{\boldsymbol{L,S}}\left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
+	(\boldsymbol{\hat{L}}, \boldsymbol{\hat{S}})\leftarrow \min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}
 \end{equation}
-where $\boldsymbol{Y}$ are the Lagrange multipliers and have the constraint $\boldsymbol{C = D - A - E}$...
+\begin{center} subject to: $\boldsymbol{X} = \boldsymbol{L} + \boldsymbol{S}$ \end{center}
+\begin{equation}\label{eq:eq02}
+    \left \| \boldsymbol{L} \right \|_{*} = \sum_{i} \sigma_{i}(\boldsymbol{L})
+\end{equation}
+\begin{equation}\label{eq:eq03}
+    \left \| \boldsymbol{S} \right \|_{1} = \sum_{ij} \left | \boldsymbol{S}_{ij} \right |
+\end{equation}
+where $\lambda$ is a positive weighting parameter, which determines the sparsity of $\boldsymbol{S}$, $\sigma$ denotes the singular values of a matrix.
+
+Before ALM, some methods were proposed to solve that convex optimization problem, such as Iterative Thresholding (IT) and Accelerated Proximal Gradient (APG).
+However, according to \cite{lin2010augmented}, both approaches have scalability problems and require a large number of iterations to converge.
+The Augmented Lagrange Multiplier is proven to have a \emph{Q-linear} convergence speed and experimental results show that ALM is five times faster than APG, which in theory is sub-linear \cite{lin2010augmented}.
+Furthermore, ALM reaches more accurate results with less iterations.
+
+The RPCA with ALM can be formulated:
+\begin{equation}\label{eq:eq04}
+	l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}) = \left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
+\end{equation}
+where $\boldsymbol{Y}$ is the multiplier of the linear constraint, $\mu$ is the penalty parameter for the violation of the linear constraint \cite{yuan2009sparse}.
+Thus, an iterative scheme can be presented as:
+\begin{equation}\label{eq:eq05}
+	\left\{
+		\begin{matrix} 
+			(\boldsymbol{L}_{k+1}, \boldsymbol{S}_{k+1}) \in \operatorname*{argmin}_{\boldsymbol{L,S} \in \mathbb{R}^{m \times n}} \{l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}, \\ 
+			\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k}),
+		\end{matrix}
+	\right.
+\end{equation}
+
+We adopt RPCA with ADM, which, generally speaking, is a practical improvement of the classical Augmented Lagrangian method for solving convex programming problem with linear constraints, by fully taking advantage of its high-level separable structure \cite{yuan2009sparse}.
+ADM minimizes $\boldsymbol{L}$ and $\boldsymbol{S}$ variables serially by solving the following problems to generate the new iterate:
+
+\begin{equation}\label{eq:eq06}
+	\left\{\begin{matrix}
+	\boldsymbol{L}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{L} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}, \boldsymbol{S}_{k}, \boldsymbol{Y}_{k})\}\\ 
+	\boldsymbol{S}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{S} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}_{k+1}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}\\ 
+	\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k})
+	\end{matrix}\right.
+\end{equation}
 
@@
 
-\begin{equation}\label{eq:eq02}
+\begin{equation}\label{eq:eq07}
 	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i, 
@@
 
-\begin{equation}\label{eq:eq03}
+\begin{equation}\label{eq:eq08}
 	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
@@
 The general formula of the $u$-th moment can be expressed as
-\begin{equation}\label{eq:eq04}
+\begin{equation}\label{eq:eq09}
 	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
@@
 Therefore, the skewness is calculated by
-\begin{equation}\label{eq:eq05}
+\begin{equation}\label{eq:eq10}
 	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
@@
 and the kurtosis as
-\begin{equation}\label{eq:eq06}
+\begin{equation}\label{eq:eq11}
 	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}.
@@
 
-We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq01} for $\textbf{X}^r$.
+We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq06} for $\textbf{X}^r$.
 We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between new observations and the moments calculated from a robust subspace computed by RPCA.
@@
 The classical Mahalanobis Distance is defined as		
-\begin{equation}\label{eq:eq07}
+\begin{equation}\label{eq:eq12}
 	\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
@@
 Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA.
-We also propose to extend the Equation \ref{eq:eq05} to implement a Skewness-based Mahalanobis Distance, as follows:
+We also propose to extend the Equation \ref{eq:eq12} to implement a Skewness-based Mahalanobis Distance, as follows:
 
-\begin{equation}\label{eq:eq08}
+\begin{equation}\label{eq:eq13}
 	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}.
@@
 
-We finally propose to extend the Equation \ref{eq:eq05} to implement a Kurtosis-based Mahalanobis Distance, as follows:
+We finally propose to extend the Equation \ref{eq:eq12} to implement a Kurtosis-based Mahalanobis Distance, as follows:
 
-\begin{equation}\label{eq:eq09}
+\begin{equation}\label{eq:eq14}
 	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}.
@@
 The distances $\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})$, $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})$ and $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})$ shall be computed and evaluated separately for botnet detection.
-Therefore, the robust subspace learning from normal data combined to MD can be called md-RPCA when using mean-based MD, sd-RPCA when using skewness-based MD, and kd-RPCA when using kurtosis-based MD.
+Therefore, the robust subspace learning from normal data combined to MD can be called mmd-RPCA when using mean-based MD, smd-RPCA when using skewness-based MD, and kmd-RPCA when using kurtosis-based MD.
 
@@
 	\SetAlgoLined
-	\KwResult{$\boldsymbol{\bar{t}}$}
-	Given $\boldsymbol{X}$ split into $\boldsymbol{X}^n$ and $\boldsymbol{X}_i^t$\;
-	$\min_{\boldsymbol{L}^n, \boldsymbol{S}^n}\left\|\boldsymbol{L}^n\right\|_* + \lambda\left\|\boldsymbol{S}^n\right\|_1 + \langle \boldsymbol{Y}^n, \boldsymbol{X}^n - \boldsymbol{L}^n - \boldsymbol{S}^n  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X}^n - \boldsymbol{L}^n - \boldsymbol{S}^n\right\|_F^2$\;
+	\KwResult{$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$}
+	Given $\boldsymbol{X}$ split into $\boldsymbol{X}^n$ and $\boldsymbol{X}^t$\;
+	\While{not $\min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}$ of $\boldsymbol{X}^n$}{
+		$\boldsymbol{L}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{L} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}, \boldsymbol{S}_{k}, \boldsymbol{Y}_{k})\}$\;
+		$\boldsymbol{S}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{S} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}_{k+1}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}$\;
+		$\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k})$\;
+	}	
 	$\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i$, for $\boldsymbol{L}^n$\;
@@
 	$\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}$, for $\boldsymbol{L}^n$\;
-	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
+	$\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
 	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}$\;
@@
 
-The steps between 1 and 6 of the Algorithm \ref{alg:alg01} are the training from normal data $\boldsymbol{X}^n$, while the steps between 7 and 12 aim the anomaly detection on new contamianted observations.
-Note that the steps 7 and 10 refers to the md-RPCA, while the steps 8 and 11 refers to the sd-RPCA, and the steps 9 and 12 refers to the kd-RPCA.
+The steps between 1 and 10 of the Algorithm \ref{alg:alg01} are the training from normal data $\boldsymbol{X}^n$, while the steps between 11 and 16 aim the anomaly detection on new contamianted observations.
+Note that the steps 11 and 14 refers to the mmd-RPCA, while the steps 12 and 15 refers to the smd-RPCA, and the steps 13 and 16 refers to the kmd-RPCA.
 It is also important to observe that the anomaly detection from new observations does not require new robust subspace learning, it is only necessary to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous.
@@
 Therefore, the data set and scenarios for our evaluation can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, ...
-, \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to Table \ref{tab:tab01}.
+, \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:tab01}.
 
@@
 	The F-measure is the harmonic mean of precision and recall and is defined by 
-	\begin{equation}\label{eq:eq10}
+	\begin{equation}\label{eq:eq15}
 		F-measure = 2 * \frac{PR * RC}{PR + RC},				
@@
 	where $PR$ is the precision, defined by 
-	\begin{equation}\label{eq:eq11}
+	\begin{equation}\label{eq:eq16}
 		PR = \frac{True Positive}{True Positive + False Positive}
@@
 	and $RC$ is the recall, defined by 
-	\begin{equation}\label{eq:eq12}
+	\begin{equation}\label{eq:eq17}
 		RC = \frac{True Positive}{True Positive + False Negative}.

### 2019-05-17 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
working on results

@@
 We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adopted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
-Finally, We evaluate the accuracy of the proposed approach for botnet detection on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
+Finally, We evaluate the F-measure of the proposed approach for botnet detection on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
 
@@
 First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
-Then, Section \ref{sec:results} shows the results of the proposed approach and compares to selected anomaly detection algorithms.
+Section \ref{sec:results} shows the results of the proposed approach, compares to selected anomaly detection algorithms, and presents a discussion about the results.
 
@@
 
-Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training the model, while the latter does not require labeled data or training.
-However, the availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or unkown events.
+Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training anomaly detection models, while the latter does not require labeled data or training.
+However, the availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or even unkown events.
 
 Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches.
-The proposed approach is semi-supervised and relies on normal data $\boldsymbol{X}^n$ and on contamination rate $c$ for training and anomalous detection, respectively.
+The proposed approach is semi-supervised and relies on normal data $\boldsymbol{X}^n$ and on contamination rate $c$ for training and anomaly detection, respectively.
 
@@
 Therefore, we compare our proposal to GMM.
-We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
+We also evaluate the results of ROBPCA for anomaly detection on CTU-13, considering that ROBPCA also relies on robust estimates with adjusted outlyingness based on robust skewness \cite{hubert2009robustskewed}.
 
-In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach for all botnet malwares of the CTU-13.
-Therefore, the data set and scenarios for our evaluation can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, ...
-, \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:tab01}.
+In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach to all botnet malwares of the CTU-13.
 
-Given a contaminated data set $\boldsymbol{X}$, split it into $\boldsymbol{X}^n$ containing 60\% of the the normal data, and into cross validation $\boldsymbol{X}^v$, composed by 20\% of the normal and 50\% of the anomalous data, and into test $\boldsymbol{X}^t$ containing 20\% of the normal and 50\% of the anomalous data.
+Therefore, the full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \\\ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:tab01}.
+In our experiment setup each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^n$ containing 50\% of the the normal data, and into the cross validation $\boldsymbol{X}_i^v$, composed by 25\% of the normal and 50\% of the anomalous data, and into test $\boldsymbol{X}_i^t$ containing 25\% of the normal and 50\% of the anomalous data.
 
-Evaluate if the code is correct and implement it as funcions in my code
+In anomalous detection problems, where anomalies are rare events, if one classify all observations as normal and apply an accuracy evaluation, one would have high accuracy but poor true-positive detection.
+Due to the importance given by the F-measure to true-positive detection, it is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic} and it is the metric used for validation of this experiment.
 
-try other distributions between train cv and test
-
-Explain tha numbers of anomalies and normal for CV and test
-
-Select the range of contamination that shall be evaluated and perform cross validation for each contamination.
-
-For each contamination $c$:
-
-Estimate $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$ from normal data.
-
-Predict anomalies for cross validation data using the $c$, $\boldsymbol{\bar{x}}_{normal}$ and $\boldsymbol{\hat{S}}_{normal}$.
-
-Select the best F-measure by cross validation and get the best $c$.
-
-Test the F-measure of anomaly detection for new observations through $d(\boldsymbol{x},\bar{\boldsymbol{x}}_0, \boldsymbol{\hat{S}}_0)$ and $c$.
-
-\cite{garcia2014empirical} also uses a time window separation and an aggregation method.
-
-\textbf{Experiment Plan:}
-
-Evaluation Measure:
-
-In anomalous detection problems, where anomalies are rare events, if we classify all observations as normal and apply an accuracy evaluation, we would have high accuracy but poor true-positive detection.
-
-Due to the importance given by F-measure to true-positive detection \cite{powers2011evaluation,moustafa2019holistic}, the F-measure is the preferable measure for imbalanced data sets.
-
-\textbf{Experiment Plan:}
-
-	The F-measure is the harmonic mean of precision and recall and is defined by 
+The F-measure is the harmonic mean of precision and recall and is defined by:
 \begin{equation}\label{eq:eq15}
-		F-measure = 2 * \frac{PR * RC}{PR + RC},				
+	F-measure = 2 * \frac{pr * rc}{pr + rc},
 \end{equation}
-	where $PR$ is the precision, defined by 
+where $pr$ is the precision, defined by 
 \begin{equation}\label{eq:eq16}
-		PR = \frac{True Positive}{True Positive + False Positive}
+	pr = \frac{True Positive}{True Positive + False Positive}
 \end{equation}
-	and $RC$ is the recall, defined by 
+and $rc$ is the recall, defined by 
 \begin{equation}\label{eq:eq17}
-		RC = \frac{True Positive}{True Positive + False Negative}.
+	rc = \frac{True Positive}{True Positive + False Negative}.
 \end{equation}
 
-\textbf{Data splitting for training, cross validation and testing}
+It is from $\boldsymbol{X}_i^n$ that we perform the training steps, which are the steps between 1 and 10 described in Algorithm \ref{alg:alg01}, resulting into $\boldsymbol{\bar{x}}$, $\boldsymbol{\bar{\Sigma}}$, $\boldsymbol{\bar{s}}$ and $\boldsymbol{\bar{k}}$.
+The next step of the experiment is the cross validation, which is an empirical testing to identify the contamination $c$ that obtains the best F-measure on anomaly detection, aiming the maximum precision and recall.
+During the cross validation for $\boldsymbol{X}_i^v$ we compute the steps 11, 12 and 13, and iterate over the steps between 14 and 16 for selected values of $c$, and validate the results of $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}^v$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}^v$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}^v$ to the ground truth results for $\boldsymbol{X}_i^v$, in order to identify the $c$ that best classify the anomalies of $\boldsymbol{X}_i^v$.
+
+Finally, we test the proposed approach to anomaly detection on $\boldsymbol{X}_i^t$ by means of the steps between 11 and 16 of Algorithm \ref{alg:alg01}, and compare the results of $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$ to the ground truth results for $\boldsymbol{X}_i^t$.
 
@@
 \label{sec:results}
+
+The Table \ref{tab:tab02} presents the results of the proposed approach and selected methods applied to botnet detection for each scenario of the CTU-13 data set.
+The Table \ref{tab:tab02} shows the F-measure of GMM, K-means, ROBPCA, md-RPCA, kd-RPCA and sd-RPCA by scenario and by an average of all scenarios.
 
@@
   \label{tab:tab02}
-  \begin{tabular}{ c|c c|c c|c c|c c|c c }
+  \begin{tabular}{ c|c c|c c|c c|c c|c c|c c }
 	\toprule
-	\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{S-MCD}} &\multicolumn{2}{c}{\textbf{K-MCD}} &\multicolumn{2}{c}{\textbf{MCD}} &\multicolumn{2}{c}{\textbf{GMM}} &\multicolumn{2}{c}{\textbf{K-Means}}\\
-			\hhline{~----------}
-			&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
+	\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{GMM}} &\multicolumn{2}{c}{\textbf{K-Means}} &\multicolumn{2}{c}{\textbf{ROBPCA}} &\multicolumn{2}{c}{\textbf{md-RPCA}} &\multicolumn{2}{c}{\textbf{kd-RPCA}} &\multicolumn{2}{c}{\textbf{sd-RPCA}}\\
+			\hhline{~------------}
+			&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
 	\midrule
-		10 &\color{red} 0.71 & 0.03 & 0.66 & 0.02 & 0.34 & 0.52 & 0.34 & 0.46 & 0.20 & 0.25 \\ \hline
-		11 & 0.03 & 0.54 & 0.00 & 0.54 & 0.71 &\color{red} 0.73 & 0.35 & 0.48 & 0.13 & 0.15 \\ \hline
-		12 & 0.19 & 0.02 & 0.20 & 0.05 & 0.06 &\color{red} 0.31 & 0.01 & 0.01 & 0.09 & 0.11 \\ \hline
-		15 & 0.00 & 0.00 & 0.00 & 0.00 & 0.24 &\color{red} 0.32 & 0.14 & 0.22 & 0.09 & 0.11 \\ \hline
-		15-2 & 0.04 & 0.85 & 0.67 &\color{red} 0.86 & 0.36 & 0.28 & 0.32 & 0.14 & 0.18 & 0.26 \\ \hline
-		15-3 & 0.00 & 0.01 & 0.00 & 0.01 &\color{red} 0.55 & 0.48 & 0.00 & 0.01 & 0.22 & 0.29 \\ \hline
-		16 & 0.45 & 0.36 & 0.42 & 0.33 & 0.27 & 0.41 &\color{red} 0.51 & 0.49 & 0.07 & 0.16 \\ \hline
-		16-2 & 0.00 & 0.16 & 0.00 &\color{red} 0.20 & 0.04 & 0.04 & 0.02 & 0.08 & 0.04 & 0.03 \\ \hline
-		16-3 & 0.00 & 0.05 & 0.00 & 0.07 &\color{red} 0.15 & 0.10 & 0.01 & 0.01 & 0.03 & 0.04 \\ \hline
-		17 & 0.03 & 0.03 & 0.07 & 0.03 &\color{red} 0.59 & 0.53 & 0.16 & 0.47 & 0.35 & 0.39 \\ \hline
-		18 & 0.63 & 0.10 & 0.63 & 0.08 & 0.70 &\color{red} 0.72 & 0.21 & 0.17 & 0.13 & 0.20 \\ \hline
-		18-2 & 0.83 & 0.69 & 0.85 & 0.43 & 0.84 &\color{red} 0.91 & 0.50 & 0.55 & 0.62 & 0.70 \\ \hline
-		19 &\color{red} 0.71 & 0.00 & 0.68 & 0.00 & 0.11 & 0.42 & 0.17 & 0.20 & 0.20 & 0.17 \\ \hline
-		\rowcolor{Gray} \textbf{avg} & \textbf{0.28} & \textbf{0.22} & \textbf{0.32} & \textbf{0.20} & \textbf{0.38} &\color{red} \textbf{0.44} & \textbf{0.21} & \textbf{0.25} & \textbf{0.18} & \textbf{0.22} \\
+		10 & 0.34 & 0.45 & 0.20 & 0.25 & 0.29 & 0.34 & 0.97 & 0.96 & 0.99 & 0.99 & 0.99 & 0.99 \\ \hline
+		11 & 0.34 & 0.47 & 0.12 & 0.14 & 0.27 & 0.37 & 0.98 & 0.96 & 0.99 & 0.99 & 0.99 & 0.99 \\ \hline
+		12 & 0.00 & 0.00 & 0.09 & 0.11 & 0.08 & 0.11 & 0.86 & 0.80 & 0.00 & 0.00 & 0.94 & 0.90 \\ \hline
+		15 & 0.14 & 0.21 & 0.08 & 0.11 & 0.05 & 0.07 & 0.85 & 0.79 & 1,00 & 0.96 & 0.99 & 0.96 \\ \hline
+		15-2 & 0.32 & 0.14 & 0.18 & 0.25 & 0.21 & 0.30 & 0.85 & 0.82 & 0.98 & 0.99 & 0.98 & 0.99 \\ \hline
+		15-3 & 0.00 & 0.01 & 0.22 & 0.28 & 0.28 & 0.38 & 0.95 & 0.96 & 0.00 & 0.00 & 0.99 & 0.98 \\ \hline
+		16 & 0.50 & 0.48 & 0.06 & 0.16 & 0.06 & 0.10 & 0.85 & 0.83 & 0.00 & 0.00 & 0.96 & 0.00 \\ \hline
+		16-2 & 0.02 & 0.08 & 0.03 & 0.02 & 0.00 & 0.00 & 0.22 & 0.26 & 0.80 & 1,00 & 0.80 & 1,00 \\ \hline
+		16-3 & 0.00 & 0.01 & 0.02 & 0.04 & 0.01 & 0.03 & 0.60 & 0.66 & 0.00 & 0.00 & 0.84 & 0.98 \\ \hline
+		17 & 0.16 & 0.46 & 0.35 & 0.38 & 0.72 & 0.75 & 0.81 & 0.78 & 0.02 & 0.78 & 0.81 & 0.78 \\ \hline
+		18 & 0.21 & 0.17 & 0.13 & 0.19 & 0.13 & 0.15 & 0.91 & 0.88 & 0.97 & 0.98 & 0.97 & 0.98 \\ \hline
+		18-2 & 0.50 & 0.54 & 0.61 & 0.70 & 0.27 & 0.24 & 0.80 & 0.85 & 0.99 & 0.98 & 0.99 & 0.98 \\ \hline
+		19 & 0.16 & 0.20 & 0.19 & 0.17 & 0.10 & 0.13 & 0.91 & 0.91 & 0.96 & 0.99 & 0.96 & 0.99 \\ \hline
+		\rowcolor{Gray} \textbf{avg} &\textbf{0.21} &\textbf{0.25} &\textbf{0.18} & \textbf{0.21} & \textbf{0.19} &\textbf{0.23} &\textbf{0.81} &\textbf{0.80} &\textbf{0.59} &\textbf{0.67} &\textbf{0.94} &\textbf{0.89} \\
     \bottomrule
@@
 
-\begin{table}[h!]
-  \centering
-  \scriptsize
-  \caption{Results grouped by algorithm and scenario}
-  \label{tab:tab03}
-  \begin{tabular}{ c|c c|c c|c c|c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{RPCA}} &\multicolumn{2}{c}{\textbf{S-RPCA}} &\multicolumn{2}{c}{\textbf{S-RPCA2}} &\multicolumn{2}{c}{\textbf{MCD}}\\ 
-			\hhline{~--------}
-			&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
-	\midrule
-		10 & 0.93 & 0.93 &\color{red} 0.98 &\color{red} 0.98 &\color{red} 0.99 &\color{red} 1.00 & 0.34 & 0.52 \\ \hline
-		11 & 0.96 & 0.94 &\color{red} 0.99 &\color{red} 0.98 &\color{red} 0.99 &\color{red} 1.00 & 0.71 & 0.73 \\ \hline
-		12 & 0.32 & 0.19 & 0.00 & 0.00 &\color{red} 0.81 & 0.01 & 0.06 & 0.31 \\ \hline
-		15 & 0.75 & 0.73 & 0.86 &\color{red} 0.96 &\color{red} 1.00 &\color{red} 0.96 & 0.24 & 0.32 \\ \hline
-		15-2 & 0.40 & 0.57 &\color{red} 0.95 &\color{red} 0.94 & 0.25 & 0.93 & 0.36 & 0.28 \\ \hline
-		15-3 & 0.68 & 0.68 &\color{red} 0.96 & 0.70 &\color{red} 0.96 &\color{red} 0.96 & 0.55 & 0.48 \\ \hline
-		16 &\color{red} 0.47 &\color{red} 0.47 & 0.01 & 0.01 & 0.01 & 0.00 & 0.27 & 0.41 \\ \hline
-		16-2 & 0.07 & 0.10 &\color{red} 0.80 & 0.68 & 0.00 & 0.00 & 0.04 & 0.04 \\ \hline
-		16-3 & 0.29 & 0.29 & 0.01 & 0.46 &\color{red} 0.79 & 0.76 & 0.15 & 0.10 \\ \hline
-		17 &\color{red} 0.82 & 0.79 &\color{red} 0.81 & 0.78 &\color{red} 0.82 & 0.79 & 0.59 & 0.53 \\ \hline
-		18 & 0.87 & 0.89 & 0.97 &\color{red} 0.98 &\color{red} 0.98 &\color{red} 0.98 & 0.70 & 0.72 \\ \hline
-		18-2 &\color{red} 0.94 &\color{red} 0.94 &\color{red} 0.91 & 0.86 & 0.09 & 0.10 & 0.84 & 0.91 \\ \hline
-		19 & 0.27 & 0.31 &\color{red} 0.95 &\color{red} 0.95 & 0.02 & 0.02 & 0.11 & 0.42 \\ \hline
-		\rowcolor{Gray} \textbf{Avg} & \textbf{0.60} & \textbf{0.60} &\color{red} \textbf{0.71} &\color{red} \textbf{0.71} & \textbf{0.59} & \textbf{0.58} & \textbf{0.38} & \textbf{0.44} \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-Results grouped by algorithm for full CTU-13.
-
-\subsection{Discussion}
-\label{sec:discussion}
+It is possible to observe that the results of md-RPCA, kd-RPCA and sd-RPCA overcome the remain methods in a comparison by scenario and by average F-measure.
 
@@
 The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
+The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting F-measure of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
 
@@
 
-% Consider a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ where $\boldsymbol{X} = \{\boldsymbol{x}_1,...,\boldsymbol{x}_n\}$ of $p$-variate observations.
-Let $\boldsymbol{h} \subset \{1,...,n\}$ with $|\boldsymbol{h}| = h$, where $\boldsymbol{h}$ is a subset of $h$ indexes from $n$ observations.
+% Consider a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ where $\boldsymbol{X} = \{\boldsymbol{x}_1,\ldots,\boldsymbol{x}_n\}$ of $p$-variate observations.
+Let $\boldsymbol{h} \subset \{1,\ldots,n\}$ with $|\boldsymbol{h}| = h$, where $\boldsymbol{h}$ is a subset of $h$ indexes from $n$ observations.
 
@@
 
-% for $i = 1,...,n$.
+% for $i = 1,\ldots,n$.
 
-% Let $\boldsymbol{h}_1, \bar{\boldsymbol{x}}_1$ and $\boldsymbol{\hat{S}}_1$ and be the selected observations, the location and scatter of the initial subset, and $\boldsymbol{d}_1(i)$ be the initial distances for $i = 1,...,n$.
+% Let $\boldsymbol{h}_1, \bar{\boldsymbol{x}}_1$ and $\boldsymbol{\hat{S}}_1$ and be the selected observations, the location and scatter of the initial subset, and $\boldsymbol{d}_1(i)$ be the initial distances for $i = 1,\ldots,n$.
 
-% Now take $\boldsymbol{h}_2$ containing the indexes of lowest distances, such that $\{\boldsymbol{d}_1(i); i \in \boldsymbol{h}_2\} = \{(\boldsymbol{d}_1)_{1:n},...,(\boldsymbol{d}_1)_{h:n}\}$, where $(\boldsymbol{d}_1)_{1:n} \leq (\boldsymbol{d}_1)_{2:n} \leq ...
-\leq (\boldsymbol{d}_1)_{n:n}$ are the ordered distances, and compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$.
+% Now take $\boldsymbol{h}_2$ containing the indexes of lowest distances, such that $\{\boldsymbol{d}_1(i); i \in \boldsymbol{h}_2\} = \{(\boldsymbol{d}_1)_{1:n},\ldots,(\boldsymbol{d}_1)_{h:n}\}$, where $(\boldsymbol{d}_1)_{1:n} \leq (\boldsymbol{d}_1)_{2:n} \leq \ldots \leq (\boldsymbol{d}_1)_{n:n}$ are the ordered distances, and compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$.
 
@@
 % 		}
-% 		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,..., n$\;
-% 		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_1(\pi(n))$\;
-% 		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), ..., \pi(h)\}$\;
+% 		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,\ldots, n$\;
+% 		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq \ldots \leq \boldsymbol{d}_1(\pi(n))$\;
+% 		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), \ldots, \pi(h)\}$\;
 % 		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
@@
 % 	}
-% 	Compute $\boldsymbol{d}_1^2(i) = (\boldsymbol{x}_i - \boldsymbol{\bar{x}}_1)^\prime \boldsymbol{\hat{S}}_1^{-1}(\boldsymbol{X}_i - \boldsymbol{\bar{x}}_0)$ for $i = 1, ..., n$\;
-% 	Sort them into $\boldsymbol{d}_1(\pi(1)) \leq ... \leq \boldsymbol{d}_0(\pi(n))$\;
-% 	Put $\boldsymbol{h}_1 = \{\pi(1), ..., \pi(h)\}$\;
+% 	Compute $\boldsymbol{d}_1^2(i) = (\boldsymbol{x}_i - \boldsymbol{\bar{x}}_1)^\prime \boldsymbol{\hat{S}}_1^{-1}(\boldsymbol{X}_i - \boldsymbol{\bar{x}}_0)$ for $i = 1, \ldots, n$\;
+% 	Sort them into $\boldsymbol{d}_1(\pi(1)) \leq \ldots \leq \boldsymbol{d}_0(\pi(n))$\;
+% 	Put $\boldsymbol{h}_1 = \{\pi(1), \ldots, \pi(h)\}$\;
 % 	\caption{Constructing the initial subset}

### 2019-05-17 docs/papers/2016JNCA/2016JNCA.tex
compiling version with all chapters
changing what?

@@
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
-We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz, and the operator $\sim$ that sorts the elements of a vector in ascending order.
+We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\uptarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz, and the operator $\sim$ that sorts the elements of a vector in ascending order.
 

### 2019-05-23 docs/thesis-qualification/chapters/4_tensor_dl/tensor_dl.tex
prepare for thesis writing

--- 2017-11-15 docs/thesis/chapters/4_tensor_dl/tensor_dl.tex
@@
 
-Dictionary learning is related to CS but seeks to obtain compact and meaningful signal representations, known as dictionary, from learning algorithms or models.
+Dictionary Learning (DL) is related to CS but seeks to obtain compact and meaningful signal representations, known as dictionary, from learning algorithms or analytical models.
 DL is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
@@
 First is the analytic approach, in which Discrete Cosine Transform basis, wavelets, curvelets and other nonadaptive functions are used as atoms to construct the dictionaries.
-Second is the learning-based approaches, such as the unsupervised learning for dictionary construction [9] and the online dictionary learning [11], [10], which use machine learning methods to construct the dictionary.
-
-In the analytic approach, some pre-defined functions are used to construct the dictionary.
-Curvelets [24], which tracks the shape of the discontinuity set, offers efficient and near-optimal representation of smooth objects.
-Shearlets [25], which is obtained from dilations, action of translations and shear transformations, has nice geometric properties and mathematical properties for image representation.
-Bandelets [26], which specifies the geometry as a vector field, is designed to improve the image compression and noise reduction performance.
-
-In the learning-based approach, machine learning methods are used to construct the dictionary from the training data.
-The least square error is used by the method of the optimal directions (MOD) [27] to update the dictionary iteratively.
+Second is the learning-based approaches, such as the unsupervised learning for dictionary construction [9] and the online dictionary learning \cite{mairal2009online}, which use machine learning methods to construct the dictionary.
+Learning-based approaches usually uses least square error by the method of the optimal directions (MOD) \cite{engan1999method} to update the dictionary iteratively.
 While analytic dictionaries permit to capture the global structure of a signal and allow a fast implementation, learned dictionaries often perform better in applications as they are more adapted to the considered class of signals.
@@
 Multidimensional parameter estimation and learning multidimensional separable dictionaries are growing research problems.
-The crucial idea about separable dictionaries is to allow the dictionary to have a separable structure, where separable means that the dictionary $\textbf{D}$ is given by the Kronecker product of two smaller dictionaries $\boldsymbol{A}^{(1)} \in \mathbb{R}^{h \times a}$ and $\boldsymbol{A}^{(2)} \in \mathbb{R}^{w \times b}$, for example.
+The crucial idea about separable dictionaries is to allow the dictionary to have a separable structure, where separable means that the dictionary $\textbf{A}$ is given by the Kronecker product of two smaller dictionaries $\textbf{A}^{(1)} \in \mathbb{R}^{h \times a}$ and $\textbf{A}^{(2)} \in \mathbb{R}^{w \times b}$, for example.
 Roemer \emph{et al.} \cite{roemer2014tensor} show that the multidimensional dictionary estimation problem can be efficiently formulated in terms of tensors, and that their results outperform existing schemes by exploiting the multilinear structure of the problem.
 
-Considering that big data problems require techniques to deal with multidimensional data in order to make sense of structure and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highlight discriminative structure and relationships of one dataset, this work propose a tensor-based DL for fraud detection from imbalanced data, in order to apply the tensor decomposition for DL methods to highlight the discriminative sensing of a fraud detection dataset.
+Considering that big data problems require techniques to deal with multidimensional data in order to make sense of structure and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highlight the discriminative structure and relationships of one dataset, this work proposes a tensor-based DL for fraud detection from imbalanced data, in order to apply the tensor decomposition for DL methods to highlight the discriminative sensing for fraud detection from a dataset of mobile money transactions.
 
@@
 Section \ref{sec:4_proposal} describes the proposed approach for fraud detection from mobile money transactions.
-Section \ref{sec:2_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.
+Section \ref{sec:4_experiments} discusses the experimental validation, Section \ref{sec:4_results} presents the results, and Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.
 
@@
 Considering that DL aims to obtain meaningful dictionaries, it is possible to evaluate DL algorithms by their performance to generate dictionaries that better represent one target data.
-Therefore, we can compare DL algorithms regarding recovering of a known dictionary.
-We conduct a evaluation of tensor-based DL algorithms for recovering a known dictionary in order to analyze the gains of tensor-based approaches in comparison to well known algorithms.
+Therefore, we can compare DL algorithms and tensor-based DL algorithms regarding recovering of a known dictionary.
+
+In this section we conduct a evaluation of tensor-based DL algorithms for recovering a known separable dictionary in order to analyze the gains of tensor-based approaches in comparison to well known algorithms.
 
@@
 \begin{equation}\label{eq:4_eq01}
-	\boldsymbol{X} = \boldsymbol{A} \cdot \boldsymbol{S} + \boldsymbol{W},
+	\textbf{X} = \textbf{A} \cdot \textbf{S} + \textbf{W},
 \end{equation}
 
-where $\boldsymbol{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\boldsymbol{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary where $N \ll T$, $\boldsymbol{S} \in \mathbb{C}^{N \times T}$ represents the sparse coefficient matrix, $\boldsymbol{W} \in \mathbb{C}^{M \times T}$ is the additive noise, and $M < N < T$.
+where $\textbf{X} \in \mathbb{C}^{M \times T}$ represents $T$ consecutive observations from $M$ features, $\textbf{A} \in \mathbb{C}^{M \times N}$ is the overcomplete dictionary where $N \ll T$, $\textbf{S} \in \mathbb{C}^{N \times T}$ represents the sparse coefficient matrix, $\textbf{W} \in \mathbb{C}^{M \times T}$ is the additive noise, and $M < N < T$.
 
@@
 \begin{equation}\label{eq:4_eq02}
-	\boldsymbol{X} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}) \cdot \boldsymbol{S} + \boldsymbol{W},
+	\textbf{X} = (\textbf{A}^{(1)} \otimes \textbf{A}^{(2)}) \cdot \textbf{S} + \textbf{W},
 \end{equation}
 
-where $\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)})$, $\boldsymbol{A}^{(1)} \in \mathbb{C}^{M_1 \times N_1}$, $\boldsymbol{A}^{(2)} \in \mathbb{C}^{M_2 \times N_2}$, $M = M_1 \times M_2$, and $N = N_1 \times N_2$.
+where $\textbf{A} = (\textbf{A}^{(1)} \otimes \textbf{A}^{(2)})$, $\textbf{A}^{(1)} \in \mathbb{C}^{M_1 \times N_1}$, $\textbf{A}^{(2)} \in \mathbb{C}^{M_2 \times N_2}$, $M = M_1 \times M_2$, and $N = N_1 \times N_2$.
 
-In order to evaluate tensor-based DL algorithms for recovering a known dictionary, we generate two random dictionaries $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ from an i.i.d.
-zero mean Gaussian random process, and calculate the dictionary $\boldsymbol{A}$ through the Kronecker product, according to
+In order to evaluate tensor-based DL algorithms for recovering a known dictionary, we generate two random dictionaries $\textbf{A}^{(1)}$ and $\textbf{A}^{(2)}$ from an i.i.d.
+zero mean Gaussian random process, and calculate the dictionary $\textbf{A}$ through the Kronecker product, according to Equation
 
 \begin{equation}\label{eq:4_eq03}
-	\boldsymbol{A} = (\boldsymbol{A}^{(1)} \otimes \boldsymbol{A}^{(2)}).
+	\textbf{A} = (\textbf{A}^{(1)} \otimes \textbf{A}^{(2)}).
 \end{equation}
 
-Therefore, $\boldsymbol{A}$ is the known dictionary that shall be used for recovery evaluation.
+Therefore, $\textbf{A}$ is the known dictionary that shall be used for recovery evaluation.
 
-Subsequently, we generate a synthetic data set $\boldsymbol{X}$ using the given dictionary $\boldsymbol{A}$, according to Equation \ref{eq:4_eq01}, where $\boldsymbol{S}$ is a random sparse coefficient matrix for what we assume that each column has $K = 5$ non-zero entries, and $\boldsymbol{W}$ is the additive white gaussian noise.
-Thus, we estimate the initial $\boldsymbol{\hat{A}}$ dictionary from a normalized subset of $\boldsymbol{X}$ and make the dictionary separable decomposing the approximation of $\boldsymbol{\hat{A}}$ and generating $\boldsymbol{\hat{A}}^{(1)}$ and $\boldsymbol{\hat{A}}^{(2)}$ according to \cite{van1993approximation}.
-Afterward, we update $\boldsymbol{\hat{A}}$ through
+Subsequently, we generate a synthetic data set $\textbf{X}$ using the given dictionary $\textbf{A}$, according to Equation \ref{eq:4_eq01}, where $\textbf{S}$ is a random sparse coefficient matrix for what we assume that each column has $K = 5$ non-zero entries, and $\textbf{W}$ is the additive white gaussian noise.
+Thus, we estimate the initial $\hat{\textbf{A}}$ dictionary from a normalized subset of $\textbf{X}$ and make the dictionary separable decomposing the approximation of $\hat{\textbf{A}}$ and generating $\hat{\textbf{A}}^{(1)}$ and $\hat{\textbf{A}}^{(2)}$ according to \cite{van1993approximation}.
+Afterward, we update $\hat{\textbf{A}}$ through Equation
 
 \begin{equation}\label{eq:4_eq04}
-	\boldsymbol{\hat{A}} = (\boldsymbol{\hat{A}}^{(1)} \otimes \boldsymbol{\hat{A}}^{(2)}),
+	\hat{\textbf{A}} = (\hat{\textbf{A}}^{(1)} \otimes \hat{\textbf{A}}^{(2)}),
 \end{equation}
 
-and use it for DL experiments using MOD \cite{engan1999method}, K-SVD \cite{aharon2006rm}, RLS-DLA \cite{skretting2010recursive}, T-MOD \cite{roemer2014tensor} and K-HOSVD \cite{roemer2014tensor}.
-Due to its low computational complexity, we employ the Orthogonal Matching Pursuit (OMP) algorithm\cite{davis1997adaptive} for the sparse recovery stage in all the dictionary learning algorithms, assuming that K is known.
+and use it for DL recovery experiments using MOD \cite{engan1999method}, K-SVD \cite{aharon2006rm}, RLS-DLA \cite{skretting2010recursive}, T-MOD \cite{roemer2014tensor} and K-HOSVD \cite{roemer2014tensor}.
+
+Due to its low computational complexity, we employ the Orthogonal Matching Pursuit (OMP) algorithm\cite{davis1997adaptive} for the sparse recovery stage in all the dictionary learning algorithms, assuming that $K$ is known.
 However, other solver for sparse recovery can be used, such as Basis Pursuit \cite{chen2001atomic}, the LASSO algorithm \cite{tibshirani1996regression} or greedy \cite{davis1997adaptive}.
 
-The final estimated dictionaries $\boldsymbol{\hat{A}}$ recovered by each DL algorithm are compared to the known dictionary $\boldsymbol{A}$, through a measure of the distance between two dictionaries.
+The final estimated dictionaries $\hat{\textbf{A}}$ recovered by each DL algorithm are compared to the known dictionary $\textbf{A}$, through a measure of the distance between two dictionaries.
 The recovered dictionary is then compared to the known dictionary, by using the angle between a known atom and a recovered atom as a distance measure.
-The number of identified atoms is computed by comparing these angles to a limit of $\beta_{lim} = 8.11$ degrees, that corresponds to the case where the inner product of the two compared atoms is equal to 0.99, and where each atom has 2-norm equal to 1.
+The number of identified atoms can be computed by comparing these angles to a limit of $\beta_{lim} = 8.11$ degrees, that corresponds to the case where the inner product of the two compared atoms is equal to 0.99, and where each atom has 2-norm equal to 1.
 
-For this experiments we adopt an additive noise with $snr = 20$ and 100 iterations for dictionary learning. All experiments run 100 trials in order to obtain confident results.
+For this experiments we adopt an additive noise with $snr = 20$ and 100 iterations for dictionary learning. All experiments repeats 100 times in order to obtain confident results.
 
-This evaluation extends Roemer \emph{et al.} \cite{roemer2014tensor} including RLS-DLA for comparison, and for adding the evaluation of the degree distribution for atom identification and the evaluation of the atom identification according to the dictionary size variation.
+This evaluation extends Roemer \emph{et al.} \cite{roemer2014tensor} including the comparison to RLS-DLA algorithm, and for adding the evaluation of the degree distribution for atom identification and the evaluation of the atom identification according to the dictionary size variation.
 Additionally, we conduct a experimental evaluation, while Roemer \emph{et al.} \cite{roemer2014tensor} presents numerical results obtained via Monte Carlo simulations.
@@
 It is possible to observe that T-MOD and K-HOSVD are better than MOD and K-SVD, and that the tensor-based methods outperforms MOD and K-SVD significantly for this experiment.
-Finally, Figure \ref{fig:fig1} shows that RLS-DLA recovered almost 40 atoms in $\beta_{lim} = 8.11$ degrees, while other methods present results lower results.
+Finally, Figure \ref{fig:fig1} shows that RLS-DLA recovered almost 40 atoms in $\beta_{lim} = 8.11$ degrees, while other methods present lower results.
 
@@
 	\centering 
-	\includegraphics[width=10cm]{figures/5_20_2000_1000_100.eps}
+	\includegraphics[width=10cm]{figures/ch4/s=5_snr=20_L=2000_noIt=100_N=20_K=50.eps}
 	\caption{Cumulative atom identification per degree rates. $T=2000$, $M=20$, $N=50$}
@@
 It is possible to observe that the tensor-base methods outperform MOD, RLS-DLA and K-SVD in number of identified atoms and the required degrees for identification.
+
 Note that MOD performs quite similar to RLS-DLA, while K-SVD is the worst evaluated case for this experiment.
-Considering the tensor-based methods, T-MOD outperforms K-HOSVD with higher atom identification over lower degrees.
+Considering the tensor-based methods, T-MOD outperforms K-HOSVD with higher atom identification over lower degree rates.
 The experimental results verify that the tensor-based dictionary learning algorithms are able to identify more dictionary atoms when dealing with larger dictionaries and that they converge faster.
@@
 	\centering 
-	\includegraphics[width=10cm]{figures/5_20_2000_24000_100.eps}
+	\includegraphics[width=10cm]{figures/ch4/5_20_2000_24000_100.eps}
 	\caption{Cumulative atom identification per degree rates. $T=2000$, $M=80$, $N=300$}
@@
 
-The results reveals that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases.
+The results also reveals that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases.
 Since dictionary learning and reconstruction error are usually used for sparse representation classification (SRC) problems, it is possible to suppose that tensor-based algorithms can improve results of classification problems, as well as can extract hidden patterns and structure in multidimensional data analytics problems.
@@
 However, the high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes.
-The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms
+The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms.
 
-A key challenge to use SC and DL for classification is how to find proper dictionaries and coefficients that highlight discriminative structure and relationships of one dataset.
-Therefore, we introduce a tensor-based DL method for fraud detection from unbalanced data.
-Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification.
+A key challenge to use CS and DL for classification is how to find proper dictionaries and coefficients that highlight the discriminative structure and relationships of one dataset.
+Therefore, we introduce a tensor-based DL method for fraud detection from imbalanced data.
+Specifically, we apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error of mobile money transactions for fraud classification.
 
@@
 Before this approach the general practice was to use predefined dictionaries (such as Fourier or wavelet transforms).
-However, in certain cases a dictionary that is trained to fit the input data can significantly improve the sparsity, which has applications in data decomposition, compression and analysis and has been used in the fields of image denoising and classification, video and audio processing.
-Sparsity and overcomplete dictionaries have immense applications in image compression, image fusion and inpainting.
+However, in certain cases a dictionary that is trained to fit the input data can significantly improve the sparsity, which has applications in data decomposition, compression and has been used in the fields of image denoising, video and audio processing.
 
 MOD \cite{engan1999method} is a DL algorithm that is commonly used for extensions of Least-Squares (LS) type schemes, also known as iterative least squares dictionary learning algorithm (ILS-DLA) \cite{engan2007family}.
-MOD defines that if we know $\boldsymbol{S}$ in equation \ref{eq:4_eq01}, $\boldsymbol{A}$ can be estimated solving a linear LS problem.
-Additionally, knowing $\boldsymbol{A}$ we obtain $\boldsymbol{S}$ by solving a sparse recovery problem.
-MOD iterates between the two steps, starting with an initial guess of $\boldsymbol{A}$, denoted as $\hat{\boldsymbol{A}}$, which can be estimated, generated by previous dictionary learning or generated analytically.
-Despite - simplicity, the MOD algorithm has been criticized for its slow convergence, which is due to the fact that $\boldsymbol{A}$ and $\boldsymbol{S}$ are updated separately, which may lead to a slow convergence \cite{aharon2006rm}.
+MOD defines that if we know $\textbf{S}$ in Equation \ref{eq:4_eq01}, $\textbf{A}$ can be estimated solving a linear LS problem.
+Additionally, knowing $\textbf{A}$ we obtain $\textbf{S}$ by solving a sparse recovery problem.
+MOD iterates between the two steps, starting with an initial guess of $\textbf{A}$, denoted as $\hat{\textbf{A}}$, which can be estimated, generated by previous dictionary learning or generated analytically.
+Despite its simplicity, the MOD algorithm has been criticized for its slow convergence, which is due to the fact that $\textbf{A}$ and $\textbf{S}$ are updated separately, which may lead to a slow convergence \cite{aharon2006rm}.
 
 K-SVD \cite{aharon2006rm} is a generalization of the K-means algorithm for dictionary learning.
-After the sparse approximation step, the dictionary update is performed by sequentially updating each column of $\boldsymbol{A}$ using SVD to minimize the approximation error.
+After the sparse approximation step, the dictionary update is performed by sequentially updating each column of $\textbf{A}$ using SVD to minimize the approximation error.
 The update step is hence a generalized K-means algorithm since each patch can be represented by multiple atoms and with different weights.
@@
 A simple method to build such a subspace consists in modifying the sparse approximation methods such that the objective function is augmented with a discrimination term that represents the separability properties of the projection subspace.
-This kind of approach typically tries to maximize the variance between the active atoms from $\boldsymbol{A}$ that represent signals in different classes.
-The reduced subspace used for classification is finally formed by the subset of atoms in $\boldsymbol{A}$ whose corresponding coefficients in $\boldsymbol{S}$ are nonzero.
+This kind of classification approach typically tries to maximize the variance between the active atoms from $\textbf{A}$ that represent signals in different classes.
+The reduced subspace used for classification is finally formed by the subset of atoms in $\textbf{A}$ whose corresponding coefficients in $\textbf{S}$ are nonzero.
 The subset selection problem can be interpreted as the inference step in the dictionary learning methods when the objective function is modified to include a discriminative term \cite{tosic2011dictionary}.
@@
 Similarly to data approximation problems, data analysis applications can further benefit from dictionary learning methods.
-The previous section describes subspace selection methods from predefined dictionaries.
 DL algorithms can improve the classification performance, as it leads to a better adaptation of the dictionary by enforcing sparsity in the representation of data in the different classes.
-The atoms in a dictionary $\boldsymbol{A}$ that is computed with dictionary learning methods generally capture the most important constitutive components of the signals.
+The atoms in a dictionary $\textbf{A}$ that is computed with dictionary learning methods generally capture the most important constitutive components of the signals.
 They naturally permit to classify the data into the corresponding linear subspace.
-However, there is no guarantee that the subspace built on a learned dictionary $\boldsymbol{A}$ is truly optimal for classification, as it targets efficient representation but not necessarily class separability \cite{tosic2011dictionary}.
+However, there is no guarantee that the subspace built on a learned dictionary $\textbf{A}$ is truly optimal for classification, as it targets efficient representation but not necessarily class separability \cite{tosic2011dictionary}.
 
@@
 
-Considering that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases, according to subsection \ref{sec:4_motivation_results}, we believe that tensor-based DL algorithms can improve the discriminative sensing for classification classification problems.
-Therefore, we propose use tensor-based DL for learning fraudulent and legitimate data separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics
+Considering that tensor-based algorithms can perform better for dictionary recovery problems when the dictionary size increases, according to subsection \ref{sec:4_motivation_results}, we believe that tensor-based DL algorithms can improve the discriminative sensing for classification problems.
+Therefore, we propose use tensor-based DL for learning fraudulent and legitimate data separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by cosine similarity.
 
@@
 This dataset is synthetically generated by  PaySim \cite{lopez2016paysim}, which is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.
+
 PaySim uses aggregated data from the private dataset to generate a synthetic dataset that resembles the normal operation of transactions and injects malicious behavior to later evaluate the performance of fraud detection methods.
 PaySim simulates mobile money transactions based on a sample of real transactions extracted from one month of financial logs from a mobile money service implemented in an African country.
-The original logs were provided by a multinational company
+The original logs were provided by a multinational company.
 
 The selected dataset is composed by 9 features (variables) and 2.770.408 transactions (observations).
-It is also provided one label that classifies each transaction as fraudulent or not.
-As the majority dataset of anomaly or fraud detection, the dataset of mobile money transactions is imbalanced, where the number of fraudulent transactions is 8.213 against 2.762.196 of normal transactions.
+It is also provided one label that classifies each transaction as fraudulent or legitimate.
+As the majority dataset regarding anomaly or fraud detection, the dataset of mobile money transactions is imbalanced, where the number of fraudulent transactions is 8.213 against 2.762.195 of normal transactions.
 Therefore, the fraud class represents 0,30 \% of the total transactions.
@@
 For evaluation of the proposed classification approach we adopt a data splitting into training and test data, where the training data has 70 \% of the whole dataset.
-We denote the training data as $X_r$ and the test data as $X_e$.
-The dataset can also be divided by normal and fraudulent transactions, where the normal train data is denoted as $X_r^o$ and the fraud train data is denoted as $X_r^f$.
-The same notation of fraud, normal, train and test is applied to qualify $A$ and $S$.
+We denote the training data as $\textbf{X}_r$ and the test data as $\textbf{X}_e$.
+The dataset can also be divided by normal and fraudulent transactions, where the normal train data is denoted as $\textbf{X}_r^o$ and the fraud train data is denoted as $\textbf{X}_r^f$.
+The same notation of fraud, normal, train and test is applied to qualify $\textbf{A}$ and $\textbf{S}$.
 
@@
 % \begin{equation}\label{eq:4_eq04}
-% \boldsymbol{\mathcal{X}} = \boldsymbol{\mathcal{S}} \times_1 \boldsymbol{A}^{(1)} \times_2 \boldsymbol{A}^{(2)} + \boldsymbol{\mathcal{W}},
+% \textbf{\mathcal{X}} = \textbf{\mathcal{S}} \times_1 \textbf{A}^{(1)} \times_2 \textbf{A}^{(2)} + \textbf{\mathcal{W}},
 % \end{equation}
 
-% where $\boldsymbol{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\boldsymbol{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\boldsymbol{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\boldsymbol{X}$, $\boldsymbol{S}$, and $\boldsymbol{W}$ such that $\boldsymbol{X} = [\boldsymbol{\mathcal{X}}]_{(3)}^T$, $\boldsymbol{S} = [\boldsymbol{\mathcal{S}}]_{(3)}^T$, and $\boldsymbol{W} = [\boldsymbol{\mathcal{W}}]_{(3)}^T$, respectively.
+% where $\textbf{\mathcal{X}} \in \mathbb{C}^{M_1 \times M_2 \times T}$, $\textbf{\mathcal{S}} \in \mathbb{C}^{N_1 \times N_2 \times T}$, and $\textbf{\mathcal{W}} \in \mathbb{C}^{M_1 \times M_2 \times T}$ are rearranged versions of the matrices $\textbf{X}$, $\textbf{S}$, and $\textbf{W}$ such that $\textbf{X} = [\textbf{\mathcal{X}}]_{(3)}^T$, $\textbf{S} = [\textbf{\mathcal{S}}]_{(3)}^T$, and $\textbf{W} = [\textbf{\mathcal{W}}]_{(3)}^T$, respectively.
 
-% We consider a randomly generated data in order to evaluate the proposed approach, where initially $\boldsymbol{A}^{(1)}$ and $\boldsymbol{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\boldsymbol{A}$ is obtained.
-In order to be able to evaluate the dictionary reconstruction, the sparse data $\boldsymbol{\mathcal{X}}$ is generated from the dictionary $\boldsymbol{A}$, which is used for further comparisons.
-We adopt 20 as the signal-to-noise ratio (snr) and 5 as the sparseness factor of the data $\boldsymbol{\mathcal{X}}$.
+% We consider a randomly generated data in order to evaluate the proposed approach, where initially $\textbf{A}^{(1)}$ and $\textbf{A}^{(2)}$ are gaussian distributed and randomly generated, then the dictionary $\textbf{A}$ is obtained.
+In order to be able to evaluate the dictionary reconstruction, the sparse data $\textbf{\mathcal{X}}$ is generated from the dictionary $\textbf{A}$, which is used for further comparisons.
+We adopt 20 as the signal-to-noise ratio (snr) and 5 as the sparseness factor of the data $\textbf{\mathcal{X}}$.
 
-% The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\boldsymbol{A}$ from the sparse data $\boldsymbol{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.
+% The dictionary estimation is evaluated by applying the proposed approach to estimate the dictionary $\textbf{A}$ from the sparse data $\textbf{\mathcal{X}}$, according the selected number of sample training and iterations for the optimization problem.
 
@@
 
-We propose a Tensor-based Sparse Representation Classification approach to use tensor-based DL for learning fraudulent and legitimate dictionaries separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics.
+We propose a Tensor-based Sparse Representation Classification approach to use tensor-based DL for learning fraudulent and legitimate dictionaries separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error.
 
 Our proposal requires supervision, where labels of training data are used to classify each transaction as fraud or normal.
-The first step is to split the training dataset into fraud and normal classes, and create the data matrices $\boldsymbol{X}_r^f$ and $\boldsymbol{X}_r^o$, respectively.
-For each data matrix, we randomly select $N$ columns into a matrix and compute its data normalization in order to estimate the initial dictionaries $\boldsymbol{\hat{A}}_r^f$ and $\boldsymbol{\hat{A}}_r^o$.
-Afterward, we apply the dictionaries and the data matrices to the selected DL algorithms for learning the best dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$ that respectively represents fraud and normal classes as basis vectors.
+The first step is to split the training dataset into fraud and normal classes, and create the data matrices $\textbf{X}_r^f$ and $\textbf{X}_r^o$, respectively.
 
-With the trained dictionaries $\boldsymbol{A}_r^f$ and $\boldsymbol{A}_r^o$ and the test data $\boldsymbol{X}_e$ it is possible to process each transaction of $\boldsymbol{X}_e$, denoted as $\boldsymbol{x}_t$, where $t$ represents the transaction index, in order to compute its sparse coding from fraud and normal dictionaries, generating respectively $\boldsymbol{S}_e^f$ and $\boldsymbol{S}_e^o$.
-Subsequently, we compute the signal reconstruction for each test transaction $t$, as $\boldsymbol{x}_e^f = \boldsymbol{A}_r^f \cdot \boldsymbol{S}_e^f$ and $\boldsymbol{x}_e^o = \boldsymbol{A}_r^o \cdot \boldsymbol{S}_e^o$.
+For each data matrix, we randomly select $N$ columns into a matrix and compute its data normalization in order to estimate the initial dictionaries $\hat{\textbf{A}}_r^f$ and $\hat{\textbf{A}}_r^o$.
+Afterward, we apply the dictionaries and the data matrices to the selected DL algorithms for learning the best dictionaries $\hat{\textbf{A}}_r^f$ and $\hat{\textbf{A}}_r^o$ that respectively represents fraud and normal classes as basis vectors.
 
-Finally, we compare $\boldsymbol{x}_t^f$ and $\boldsymbol{x}_t^o$ to $\boldsymbol{x}_t$ in order to evaluate the reconstruction error and to classify the transaction as fraud or normal according to the minimum reconstruction error from $\boldsymbol{x}_t^f$ and $\boldsymbol{x}_t^o$, respectively.
+With the trained dictionaries $\hat{\textbf{A}}_r^f$ and $\hat{\textbf{A}}_r^o$ and the test data $\textbf{X}_e$ it is possible to process each transaction of $\textbf{X}_e$, denoted as $\textbf{x}_t$, where $t$ represents the transaction index, in order to compute its sparse coding from fraud and normal dictionaries, generating respectively $\hat{\textbf{s}}_t^f$ and $\hat{\textbf{s}}_t^o$.
+Subsequently, we compute the signal reconstruction for each test transaction $t$, as $\hat{\textbf{x}}_t^f = \hat{\textbf{A}}_r^f \cdot \hat{\textbf{s}}_t^f$ and $\hat{\textbf{x}}_e^o = \hat{\textbf{A}}_r^o \cdot \hat{\textbf{s}}_e^o$.
+
+Finally, we compare $\hat{\textbf{x}}_t^f$ and $\hat{\textbf{x}}_t^o$ to $\textbf{x}_t$ in order to evaluate the reconstruction error and to classify the transaction as fraud or normal according to the minimum reconstruction error from $\hat{\textbf{x}}_t^f$ and $\hat{\textbf{x}}_t^o$, respectively.
 We adopt the cosine similarity to compare the reconstructed signal to the original one.
@@
 
-Considering that fraud or other kind of malicious events are rare when compared to legitimate events, we propose to evaluate our proposal for undersampled data and for the complete dataset, evaluating the results for training using undersample data ans testing with undersample test, as well as using the undersample data for training and the complete testing data to evaluate the classification accuracy.
+Considering that fraud or other kind of malicious events are rare when compared to legitimate events, we propose to evaluate our proposal for undersampled data and for the complete dataset, evaluating the results for training using undersampled data and testing with undersample test, as well as using the undersampled data for training and the complete testing data to evaluate the classification accuracy.
 
-The approach described in Section \ref{sec:4_proposal} can adopt any DL algorithm, however we propose to evaluate the K-HOSVD in order to analyze how a tensor-based DL algorithm can highlight the discriminative structure and relationships of one dataset.
+The approach described in Section \ref{sec:4_proposal} is generic and can adopt any DL algorithm, however we propose to evaluate the K-HOSVD in order to analyze how a tensor-based DL algorithm can highlight the discriminative structure and relationships of the evaluated dataset.
 We propose to compare the proposed approach against the Support Vector Machines (SVM) algorithm and against the Logistic Regression (LR) algorithm, using features extracted trough Principal Component Analysis (PCA).
@@
 SVMs yield good generalization performance on such problems by directly estimating a decision boundary with maximal separability \cite{akbani2004applying, banerjee2006support}.
+
 LR is the most common method used to model binary response data \cite{hilbe2011logistic}, it is useful for situations in where it is necessary to predict the presence or absence of a characteristic, with binary classification, such as in anomaly or fraud detection \cite{bhattacharyya2011data}.
+PCA \cite{jolliffe2005principal} is a classical feature extraction and data representation technique widely used in the areas of pattern recognition and computer vision, as an approach to extract the most important information of one dataset \cite{yang2004two}.
 
 We adopt the area under the Receiver Operating Characteristic (ROC) curve and the area under the Precision-Recall (PR) curve in order to evaluate our approach for fraud detection from a mobile money transaction dataset, and to compare the results against the logistic regression algorithm applied to the same problem.
+
 The ROC curve makes use of the proportion of two metrics: true positives rate (TPR) and false positives rate (FPR), which are defined as:
@@
 \end{equation}
-
 where $TP$ denotes true positive, $Pc$ is the total positive values, $FP$ denotes the false positive and $Nc$ the total negative values.
@@
 ROC curves are commonly used to present results for binary decision problems in machine learning.
-However, when dealing with highly skewed datasets, PR curves give a more informative picture of an algorithm’s performance \cite{davis2006relationship, he2009learning}.
+However, when dealing with highly skewed datasets, PR curves give a more informative picture of an algorithm’s performance \cite{davis2006relationship, he2009learning}, since it takes into account the false negative values, which are very important for classification of rare positive events.
 The PR curve is defined by plotting Precision Rate (Pr) over the Recall Rate (Rr), which are defined as:
@@
 \begin{equation}\label{eq:4_eq07}
-	Rr = \frac{TP}{TP + FN}.
+	Rr = \frac{TP}{TP + FN},
 \end{equation}
+where FN denotes false negative values.
 
-PR curves exhibit a strong correspondence to ROC curves, where a curve dominates in ROC space if and only if it dominates in PR space.
-However, an algorithm that optimizes the Area Under the Curve (AUC) in the ROC space is not guaranteed to optimize the AUC in PR space.
-Considering that our data is very imbalanced, with many more negatives than positives (frauds) values, we adopt the area under the PR curve for classification evaluation, but we also compute the area under the ROC curve as a secondary classification metric.
+An algorithm that optimizes the Area Under the Curve (AUC) in the ROC space is not guaranteed to optimize the AUC in PR space.
+Therefore, considering that our data is very imbalanced, with many more negatives than positives (frauds) values, we adopt the area under the PR curve for classification evaluation, but we also compute the area under the ROC curve as a secondary classification metric.
 
@@
 
-Table \ref{tab:4_tab1} presents the results for fraud detection from mobile money transactions through SVM, LR with PCA and the proposed K-HOSVD-based classification.
+Table \ref{tab:4_tab1} presents the results for fraud detection from mobile money transactions through SVM, LR with PCA features and the proposed K-HOSVD-based classification.
 
@@
 According to the results, the SVM algorithm presents the best PR-AUC for the scenario that evaluates the generalization performance for fraud detection of imbalanced data.
-However, the PR-AUC of the SVM algorithm was the worst case for training and testing with undersample data.
-The LR algorithm applied to PCA data presents the best results for training and testing with undersample data, although its results for the evaluation of the complete testing data presents a very low PR-AUC.
+However, the PR-AUC of the SVM algorithm was the worst case for training and testing with undersampled data.
+The LR algorithm applied to PCA data presents the best results for training and testing with undersampled data, although its results for the evaluation of the complete testing data presents a very low PR-AUC.
 
-The proposed Tensor-based Sparse Representation Classification, using K-HOSVD presents higher PR-AUC than SVM for training and testing with undersample data, but its results are worst than LR with PCA for the same scenario.
-For training the undersample data and test with complete data, the proposed approach presents better PR-AUC than LR with PCA and the best ROC-AUC.
-Therefore, it is possible to observe that our approach presents average results if compared to two well known and highly adopted classifier algorithms.
+The proposed Tensor-based Sparse Representation Classification using K-HOSVD presents higher PR-AUC than SVM for training and testing with undersampled data, but its results are worst than LR with PCA features for the same scenario.
+For training the undersampled data and test with complete data, the proposed approach presents better PR-AUC than LR with PCA and the best ROC-AUC.
+
+Therefore, it is possible to observe that our approach presents average results if compared to two well known and highly adopted classifier algorithms for fraud detection from mobile money transactions.
 
@@
 
-The results shows how imbalanced data and its instance for fraud detection in mobile money transactions are challenging for classification algorithms, regarding the classification performance for undersampled data and for generalization from undersample data training to complete data testing.
+The results shows how imbalanced data and its instance for fraud detection in mobile money transactions are challenging for classification algorithms, regarding the classification performance for undersampled data and for generalization from undersampled data training to complete data testing.
 
-It is also possible to observe that the proposed tensor-based sparse representation classification, using K-HOSVD, presents average performance and highlights some enhancement possibilities, such as:
+It is also possible to observe that the proposed tensor-based sparse representation classification using K-HOSVD presents average performance and highlights some enhancement possibilities, such as:
 
 \begin{enumerate}
-	\item Extend the proposed approach to adopt one dictionary with basis vectors that represent legitimate and fraudulent data, as a concatenation or as a kronecker product, with the classification according to a sparse code comparison;
-	\item Extend the proposed approach to adopt discriminative discriminative functions for T-MOD and K-HOSVD;
+	\item Extend the proposed approach to adopt one dictionary with basis vectors that represent legitimate and fraudulent data, as a concatenation or as a kronecker product.
+The classification can tries to maximize the variance between the active atoms from $\textbf{A}$ that represent signals in different classes, and the reduced subspace used for classification is formed by the subset of atoms in $\textbf{A}$ whose corresponding coefficients in $\textbf{S}$ are nonzero;
+	\item Extend the proposed approach to adopt discriminative discriminative functions for T-MOD and K-HOSVD, such as the Fisher discrimination criterion, which is imposed on the coding coefficients so that they have small
+within-class scatter but big between-class scatter \cite{yang2011fisher};
 	\item Extend the evaluation for other DL algorithms.

### 2019-05-23 docs/thesis-qualification/chapters/appendix.tex
prepare for thesis writing

--- 2017-09-19 docs/thesis/chapters/3_eigen_similarity/capitulo3.tex
@@
-\chapter{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks} 
-\label{chapter3}
-
-\section{Introduction}
-\label{sec:introduction}
-
-Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}.
-In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time.
-Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
-Recently, signal processing schemes have being applied to the detection of malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest}, showing advances in network traffic analysis.
-
-Information security may consist of both technical and procedural aspects.
-The former includes equipment and security systems, while the latter corresponds to security rules and recommendations.
-Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks.
-Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.
-Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
-
-In the context of anomaly-based schemes, this work proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
-Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
-To the best of our knowledge there is no similar model in the literature.
-The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
-In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack.
-In addition, we also evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset.
-Furthermore, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.
-
-The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
-
-This paper is organized as follows.
-In Section \ref{sec:relatedworks}, related works are discussed.
-Section \ref{sec:datamodel} presents the data model and the evaluated datasets.
-Section \ref{sec:prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks.
-Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:Complexity} discusses the computational complexity of the proposed framework and evaluates the required processing time for tested scenarios.
-Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
-The \ref{sec:mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.
-
-\section{Related Works}
-\label{sec:relatedworks}
-
-Several methods have been proposed for the identification and characterization of malicious activity in computer networks.
-Classical methods typically employ data mining \cite{he2008applying,ghourabi2010data,osanaiye2016distributed} and regular file analysis \cite{raynal2004honeypot} to detect patterns that indicate the presence of specific attacks in network traffic.
-
-Data mining is often used to describe the process of extracting useful information from large databases.
-Multiple methods of data mining are used in \cite{he2008applying,osanaiye2016distributed} to analyze data flow in a network with the aim of identifying characteristics of malicious traffic in large scale environments.
-Researchers have applied data mining techniques in log analysis \cite{ghourabi2010data} to improve intrusion detection performance.
-However, data mining techniques used so far in network analysis require prior collection of large data sets, which is a limitation of several schemes for online analysis \cite{Huang2009}.
-
-Regular file analysis \cite{raynal2004honeypot} consists of traffic analysis for detecting known patterns that indicate the presence of attacks, applying statistical analysis to the study of collected traffic.
-An essential feature of this method is that it depends on prior knowledge of the details of the attacks to be identified, and also depends on previous log collection for traffic analysis and false positives reduction.
-
-Principal Component Analysis (PCA) is a statistical technique commonly used for dimensionality reduction.
-It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
-PCA has been used in attack detection \cite{almotairi2009technique}.
-However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques.
-
-Callegari \emph{et al} \cite{Zonglin2009} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
-However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
-
-Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
-In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
-Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
-
-Besides being prone to higher errors and false positives, such human intervention makes PCA not useful for real time applications.
-Therefore, in order to automate the analysis of eigenvalues profile, model order selection (MOS) schemes should be incorporated.
-
-Signal processing techniques have been successfully applied to network anomaly detection \cite{Lu2009}.
-Lu and Ghorbani \cite{Lu2009} proposed a network anomaly detection model based on network flow, wavelet approximation, and system identification theory.
-However, their work requires a training method to produce a prediction model for normal daily traffic and presents limitations on identification of behaviors without significant outliers, such as port scan attacks.
-Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
-Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work is not applied to probe and flood attack detection, simultaneously.
-
-The data collected from honeypot systems, such as captured traffic and operating system logs, can be analyzed to obtain information about attack techniques, general trends of threats and exploits.
-Blind automatic detection of malicious traffic techniques have been developed for honeypots in \cite{david2011blind,da2012improved}.
-However, traffic on honeypot is simpler than real network traffic, because there are no running legitimate applications, due to the fact that honeypots emulate behavior of a host within a network to deceive and lure attackers \cite{zakaria2012review}.
-Since honeypots do not generate legitimate traffic, the amount of data captured in honeypots is significantly lower in comparison to a Network Intrusion Detection System (NIDS), which captures and analyzes the largest possible amount of network traffic \cite{david2011blind}.
-MOS for blind identification of malicious activities in honeypots was proposed by us in \cite{david2011blind}, which evaluated criteria for selecting the model order, through simulations and comparing the order of the resulting model with the true model order.
-
-The proposed framework does not require either a significant amount of logs to detect attacks, nor prior data collection, in order to make comparisons and evaluate the existence of malicious traffic.
-The proposed solution is automatic and blind for detection of time frames under probe and flood attacks through MOS and eigen analysis.
-Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.
-
-Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks.
-Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
-NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications.
-Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations.
-Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD datasets adopt a traffic aggregation by connections, we consider the use of a experimental scenario on a real network and the DARPA 1998 dataset, which is the source for the creation of the KDD 99 and NSL-KDD datasets.
-Note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD datasets.
-
-\section{Data Model}
-\label{sec:datamodel}
-
-In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
-The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
-We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matriz, and the operator $\sim$ that sorts the elements of a vector in ascending order.
-
-This section presents details of the experimental scenario and the selected cases of the DARPA dataset, along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic.
-Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
-Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition.
-Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
-
-\subsection{Analyzed Scenario and Data Collection}
-\label{sec:DataCollection}
-
-The environment of the analyzed scenario is composed of two computers and one router with access to the Internet and to an internal network, where the simulation of legitimate traffic, noise, flood and port scan attacks are performed.
-During the traffic generation, one computer assumes the role of the attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:fig1}.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=6cm, width=9cm]{results/figures/fig09.png}
-     \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
-     \label{fig:fig1}
-\end{figure}
-
-Note that the set of network traffic is modeled as legitimate, noise and malicious traffic, where the victim performs legitimate activities, that can be characterized by web access.
-In many organizations this type of traffic is predominant, since most of corporate services are web-based, such as: web pages, customized web-based systems and cloud services.
-It is possible to characterize the traffic of a DHCP service as an example of noise associated with the transport layer.
-For malicious traffic, three types of networks attacks are evaluated: synflood, fraggle and port scan.
-These attacks are reproduced using well-known security tools, such as Nmap\footnote{http://nmap.org} to port scan, Metasploit\footnote{http://www.metasploit.com} to synflood and Hping\footnote{http://hping.org} to lead the fraggle attack.
-
-A network traffic log is commonly formed by timestamp, protocol, source IP address, source port, destination IP address, destination port and additional information, according to the type of the used transport protocol.
-The following TCP traffic log is presented in order to exemplify the collected data:
-\newline
-\newline
-\texttt{\justify21:00:34.099289 IP 192.168.1.102.34712 > 200.221.2.45.80: Flags [S], seq 2424058224, win 14600, options [mss 1460, sackOK,TS val 244136 ecr 0,nop,wscale 7], length 0}
-\newline
-
-and the following to exemplify UDP traffic log: 
-\newline
-\newline
-\texttt{\justify21:24:42.484858 IP 192.168.1.102.68 > 192.168.1.1.67: BOOTP/DHCP, Request from 00:26:9e:b7:82:be, length 300}
-\newline
-
-In the proposed framework, the goal is to detect the anomalies only taking into account the traffic profile, i.e., specific information such as origin IP or day and time of the attack are not considered.
-Therefore, from the entire log information, we just consider the timestamp (for sequencing), port type and port number.
-
-\subsection{Modeling Data}
-\label{sec:ModelingData}
-
-By modeling the dataset as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{U}), noise (\textbf{N}) and malicious traffic (\textbf{A}), according to the following expression:
-
-\begin{equation}\label{eq:eq01}
-\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{N}^{(q)} + \boldsymbol{A}^{(q)},
-\end{equation}
-where $q$ represents the $q$-th time frame, which is a time grouping of network traffic.
-The matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns.
-Each row represents a communication port (TCP port or UDP port), and each column represents time bins having a approtiate size, such as one minute.
-Each element $x_{m,n}^{(q)}$ stands for the number of times that the port $m$ appears at the $n$-th minute, at the $q$-th time frame.
-
-The legitimate traffic $\boldsymbol{U}^{(q)}$ is characterized by the traffic from user's operations.
-When a user accesses a web page, for example, there is the corresponding TCP/IP traffic to request the page, as well as there is the traffic required to domain name resolution.
-Figure \ref{fig:fig3} depicts an example of the legitimate traffic obtained during experiments.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig03.png}
-     \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
-     \label{fig:fig3}
-\end{figure}
-
-The traffic that is not associated with user's operations and with malicious traffic is modeled as noise $\boldsymbol{N}^{(q)}$.
-The automatic acquisition service of logical IP network address (DHCP) is an example of noise.
-Independently of any user operation, the machine receives an IP address, since it is configured to perform a DHCP address request.
-Figure \ref{fig:fig4} depicts an example of noise in a network traffic, represented by traffic to ports 67 and 68.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig04.png}
-     \caption{Network traffic of user independent operations for network management.}
-     \label{fig:fig4}
-\end{figure}
-
-The traffic coming from a malicious activity, such as a synflood or fraggle attack, is represented by the matrix $\boldsymbol{A}^{(q)}$.
-For this work we only consider the traffic from port scanning and flood attacks.
-
-We define that if the obtained $\#\boldsymbol{A}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\boldsymbol{A}^{(q)} = 0$, then there is no malicious traffic.
-This paper shows how to detect the $\#\boldsymbol{A}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious network traffic.
-
-\subsection{Synflood, Fraggle and Port scan}
-\label{sec:SynfloodFraggleandPortscan}
-
-The network attacks evaluated by this work are: synflood, fraggle and port scan.
-The first two attacks can be qualified as flood or denial of service (DoS) attacks, while the last one can be qualified as probe or port scanning attack.
-DoS attempts to block access to system or network resources and is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks.
-Probe attacks scan the networks to collect information about host, such as IP addresses, ports and services.
-
-With respect to the synflood attacks, the attacker sends a large quantity and concurrent successive SYN requests to a target, in order to consume resources and cause a DoS.
-Figure \ref{fig:fig5} depicts an example of a synflood attack carried out in a real computer network.
-In an interval of ten minutes, more than 210,000 packets are sent as a synflood attack.
-This network traffic behavior can be considered an abnormal behavior of network traffic, especially since it is concentrated in a short period of time and presents similar outstanding traffic during the time under attack.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig05.png}
-     \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
-     \label{fig:fig5}
-\end{figure}
-
-With respect to the fraggle attack, large packets with UDP echo segments are sent to the broadcast address of a network.
-Every packet is modified to have the source address of the victim, in order to implement the source address spoofing technique.
-Therefore, each host receives a huge amount of requests UDP echo and all of them replies to the IP address of the victim, causing a packet flooding aiming a DoS.
-This attack can affect the entire network, since all hosts receive several requests UDP echo and respond with the ICMP protocol, therefore each host acts as an amplifier of the attack.
-This last part of the fraggle attack is not taken into account in this work, because the victim receives ICMP (network layer) packets originated from the hosts that are attacked with flooding packet UDP echo.
-Figure \ref{fig:fig6} depicts an example of the fraggle attack in a real computer network.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=4.5cm, width=9cm]{results/figures/fig06.png}
-     \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
-     \label{fig:fig6}
-\end{figure}
-
-More than 6,000,000 malicious packets can be counted in an interval of ten minutes, which can be considered an abnormal network traffic, especially due to the concentrated traffic in a short period of time and due to the similarity of the outstanding traffic.
-
-Port scan is the attempt to establish a connection to TCP and UDP ports to identify what services are running or are in the listening state.
-There are several available port scanning techniques, including: TCP SYN scan, TCP ACK scan and UDP scan.
-This work evaluates the use of TCP SYN scan and UDP scan.
-
-In TCP SYN scan, a SYN packet is sent to the destination and two types of responses may occur: SYN/ACK or RST/ACK.
-In the first case, the destination port is in the listening state, in the second case, the destination port is not listening.
-At the end of each port scanning, a RST/ACK packet is sent by the system that is performing the port scan.
-Therefore, a full connection or a complete three-way handshake is never established, which makes the detection of the attack sender more difficult, and requires approaches able to identify probe attacks without connection stablishment.
-The UDP scan technique sends UDP packets to the destination port, and if it responds with a \emph{ICMP port unreachable} message, then it indicates that the scanned port is closed.
-On the other hand, if a message is not received, then the port is considered as open.
-
-\begin{figure}[h!]
-     \centering 
-     \includegraphics[height=5cm, width=9cm]{results/figures/fig07.png}
-     \caption{Connection attempts in order to identify active ports.}
-     \label{fig:fig7}
-\end{figure}
-
-Figure \ref{fig:fig7} depicts an example of the port scan attack in a real computer network.
-Note that the traffic is composed of two packets for each TCP port and one UDP packet to each port.
-The incoming and outgoing packets analysis, for each port, shows the high correlation and similarity of TCP and UDP traffic during the simulated port scan attack.
-
-\subsection{The DARPA Dataset}
-\label{sec:DarpaDataset}
-
-The DARPA 1998 dataset\footnote{https://www.ll.mit.edu/ideval/data/} includes 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks.
-The attacks in this dataset
-can be grouped into: denial-of-service (DoS); remote to local (R2L), which is characterized by unauthorized access from a
-remote machine; user to root (U2R), which is characterized by unauthorized access to local
-super-user privileges; and probe attack.
-Since the proposed approach focus on flood and probe attack, the analysis concentrates on the attacks that present behaviors similar to flood or probe attack.
-We observe that the most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack.
-One example is the occurrence of a neptune attack which sends 20 only SYN packets, what is a behavior that differs of the expected flood attack behavior.
-Therefore, there were selected the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey,osanaiye2016distributed}, and the cases that scan ports sending just a few packets.
-From the simulated probe attacks, we select the cases that rely on TCP or UDP connections.
-
-The data modeling follows the method described by the Subsection \ref{sec:ModelingData}, with time frames of 20 minutes, packet counting aggregation by minute and considering the traffic to the following ports: 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389, 443.
-
-\section{Proposed Framework for Detection and Identification of Network Attacks}
-\label{sec:prop_getv}
-
-This section describes the proposed technique to detect synflood, fraggle and port scan, according to Figure \ref{fig:fig80}, which represents the overview of the proposed framework for detection and identification of network attacks.
-In Subsection \ref{sec:prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvelue for each $q$-th time frame.
-Next, in Subsection \ref{sec:prop_MOSSchemes}, we show how to apply the eiganvalues on the MOS scheme in order to detect the attack.
-In Subsection \ref{sec:prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:prop_EigenSimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=11cm, width=9cm]{results/figures/mos_eigen_similarity.eps}
-     \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
-     \label{fig:fig80}
-\end{figure}
-
-\subsection{Largest Eigenvalue by Time Frames}
-\label{sec:prop_LargestEigenvaluebyTimeFrames}
-
-The proposed attack detection algorithm starts by the data pre-processing of a network traffic log containing IP, ports and timestamp of senders and receivers.
-During this step, the desired information is extracted in order to classify and count packets according to the origin and destination ports, and subsequently this information is grouped by minutes and by time frames.
-
-With the data grouped into $Q$ time frames, the framework considers the time variations of the matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the attack.
-
-According to flood and port scan attacks' behavior, flood attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively.
-These characteristics are substantiated by the results obtained through the analysis based on sample covariation of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
-
-Therefore, to detect flood attacks, it is necessary to calculate the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
-
-\begin{equation}\label{eq:eq02}
-\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
-\end{equation}
-
-The set of obtained vectors $\boldsymbol{y}_{m}^{(q)}$ composes the zero mean matrix $\boldsymbol{Y}^{(q)}$, then the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
-
-\begin{equation}\label{eq:eq03}
-\boldsymbol{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
-\end{equation}
-
-For the detection of the port scan attack, the main components are not dominated by the variables with large variance.
-Moreover, the portscan traffic presents a highly correlated network traffic.
-In order to exploit such structure, we compute the sample covariance $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have zero mean and unitary standard deviation as follows
-
-\begin{equation}\label{eq:eq04}
-\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
-\end{equation}
-
-The set of vectors $\boldsymbol{z}_{m}^{(q)}$ composes the matrix $\boldsymbol{Z}^{(q)}$, then the sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)}$ can be calculated via
-
-\begin{equation}\label{eq:eq05}
-\boldsymbol{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
-\end{equation}
-
-Once the $\boldsymbol{\hat{R}}_{yy}^{(q)}$ and $\boldsymbol{\hat{R}}_{zz}^{(q)}$ have been obtained for flood and port scan attack detection, respectively, and since the next steps are the same for both sample covariance matrices, we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{C}$.
-Therefore, the following step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:eq06}), in order to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).
-
-\begin{equation}\label{eq:eq06}
-\boldsymbol{C}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
-\end{equation}
-
-\begin{equation}\label{eq:eq060}
-\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
-\end{equation}
-
-where the operator diag$(\cdot)$ extracts the main diagonal of a matrix.
-
-The eigenvalues should be sorted in descending order, i.e., $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ...
-> \lambda_{m}^{(q)}$.
-Therefore, the largest eigenvalue of the $q$-th time frame evaluated for the attack detect is given by $\lambda_{1}^{(q)}$.
-
-The concatenation of the eigenvalues vector $\boldsymbol{e}^{(q)}$ for $q = 1, \ldots, Q$ is represented by
-
-\begin{equation}\label{eq:eq07}
-\boldsymbol{E} =
-\begin{bmatrix}
-  \lambda_1^{(1)} & \lambda_1^{(2)} & \lambda_1^{(3)} & \cdots & \lambda_1^{(Q)} \\
-  \lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
-  \lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
-  \vdots & \vdots & \ddots & \vdots  \\
-  \lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
-\end{bmatrix}.
-\end{equation}
-
-Note that since $\lambda_1^{(q)} > \lambda_2^{(q)} > \lambda_3^{(q)} > \cdots > \lambda_{m-1}^{(q)} > \lambda_m^{(q)}$, then the first line of the matrix $\boldsymbol{E}$ contains the largest eigenvalues of each $q$-th time frame, which is the Greatest Eigenvalue Time
-Vector (GETV) \cite{tenorio2013greatest}, denoted as
-
-\begin{equation}\label{eq:eq08}
-\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
-\end{equation}
-
-\subsection{MOS Schemes}
-\label{sec:prop_MOSSchemes}
-
-Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$.
-However, the goal here is to detect the variations of the eigenvalues for different values of $q$.
-Therefore, instead of using a certain $q$, the proposed approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame, in order to identify variations and estimate the model order $\hat{d}$, which is the estimated number of time frames under attack.
-Therefore, $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
-Note that some MOS schemes may also require the number of minutes that compose a time frame, as $\hat{d} = \rm{MOS}(\boldsymbol{e}_{\rm max},\emph{Q})$.
-For more information about MOS, we refer to \ref{sec:mos}.
-
-In our previous work \cite{tenorio2013greatest}, the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE schemes are evaluated for synflood and port scan attack detection, showing that EDC and EFT are effective for detecting this kind of attacks.
-The present work extends that evaluation to also analyze the effectiveness of the listed MOS schemes for fraggle attack detection, as shown in Section \ref{sec:experimentalresults}.
-
-\subsection{Eigenvalue Analysis}
-\label{sec:prop_EigenvalueAnalysis}
-
-After applying the MOS schemes to the vector $\sim\boldsymbol{e}_{\rm max}$, we obtain the estimate of the $\#\boldsymbol{A}$.
-For instance, in the case of fraggle, synflood and portscan, if $\hat{d} = 1$, , then $\#\boldsymbol{A} = 1$, which means that during the during the $Q$ time frames one attack is present.
-However, if $\hat{d} = 0$, then $\#\boldsymbol{A} = 0$, and this means that none of these attacks are present.
-Note that $\hat{d}$ can be greater than 1, indicating the presence of more than one attack.
-
-In Subsection \ref{sec:prop_MOSSchemes}, we obtained only if $\hat{d} = 1$ or $\hat{d} = 0$.
-However, if $\hat{d} = 1$, the MOS schemes do not provide any information about the $q$-th time frame under attack.
-The identification of the $q$-th time frame under attack can be carried out through a eigenvalues analysis.
-
-The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed according to Algorithm \ref{alg:alg01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$.
-Algorithm \ref{alg:alg01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$, according to Line 2 of Algorithm \ref{alg:alg01}, and its correspondent index, according to Lines 4 and 5 of Algorithm \ref{alg:alg01}.
-Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$, according to Line 8 of Algorithm \ref{alg:alg01}, and a new iteration is performed until $\boldsymbol{e}_{\rm max} = \{\}$.
-
-\begin{algorithm}[h!]
-	\caption{Detection of Time Frames Under Attack}
-  	\label{alg:alg01}
-	\begin{algorithmic}[1]
-		\show\LOOP
-	    \ALOOP {$f = 1$ until $f == \hat{d}$} 
-	    	    \STATE $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
-    	    	    \ALOOP {$i = 1$ until $i == Q$} 
-				\IF {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
-				    \STATE $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
-				\ENDIF
-        		\ENDALOOP
-	    		\STATE $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
-    		\ENDALOOP
-	\end{algorithmic}
-\end{algorithm}
-
-After the estimation of the $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked.
-To deal with this problem, the adoption of a similarity analysis between legitimate traffic and the traffic of time frames estimated as under attack is evaluated, analysing the effectiveness of cosine similarity to highlight abnormalies inserted by network traffic attacks.
-
-\subsection{Eigen Similarity Analysis}
-\label{sec:prop_EigenSimilarityAnalysis}
-
-Cosine similarity calculates the cosine of the angle between two vectors, which represents the similarity of values between the selected vectors.
-Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\boldsymbol{V}^{(q)}$ against the the most significant eigenvectors of time frame detected as under attack, to analyze similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.
-
-This subsection describes the proposed eigen similarity analysis for detailed attack identification, in complement to the attack estimation carried out through MOS schemes and eigenvalue analysis.
-In Subsection \ref{sec:prop_TimeSimilarityAnalysis} we present the eigen similarity analysis for identification of time under attack.
-Next, in Subsection \ref{sec:prop_PortSimilarityAnalysis}, we show how to apply the eigen similarity analysis in order to identify network ports under attack.
-
-\subsubsection{Time Similarity Analysis}
-\label{sec:prop_TimeSimilarityAnalysis}
-
-For eigen similarity analyis, we evaluate the cosine similarity in order to identify lacks of similarity between legitimate and malicious traffic, as follows
-
-\begin{equation}\label{eq:eq11}
-s_n = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}}},
-\end{equation}
-where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, and $\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of traffic to be performed the flood and port scan attack identification.
-
-The most significant eigenvector $\boldsymbol{v}^{(q)}$, of a time frame $q$ without attack, can be derivated from (\ref{eq:eq06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the evaluated matrix.
-The same calculation shall be performed in order to obtain the target eigenvectors $\boldsymbol{v}_{(n)}$, calculated from a time frame without attack plus minutes of a time frame estimated as under attack, to evaluate the occurrence of network attacks.
-
-The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic whithout attack, from a time frame $q$ composed of $Q$ minutes of legitimate network traffic.
-For the detailed attack identification, each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$, as represented by
-
-\begin{equation}\label{eq:eq12}
-\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
-\end{equation}
-
-The resultant $\boldsymbol{X}_{(n)}$ is necessary to obtain $\boldsymbol{v}_{(n)}$, through (\ref{eq:eq06}), for calculating the similarity degree $s_n$, ranging from 0 to 1, for each $n$-th minute.
-The $s_n$ denotes the absolute similarity degree of the $n$-th minute in comparison to a well-known traffic without attack, detected through MOS schemes and eigenvalue analysis.
-
-The incremental approach for similarity analysis is based on the incremental appending of network traffic into $\boldsymbol{X}^{(q)}$, where the first evaluation is based on (\ref{eq:eq12}) and the subsequent evaluations is based on (\ref{eq:eq13}), incrementally appending each $n$-th minute until $n=N$.
-
-\begin{equation}\label{eq:eq13}
-\boldsymbol{X}_{n} = \{\boldsymbol{X}_{n} | \boldsymbol{x}^{(\hat{q})}_{(n)}\},
-\end{equation}
-
-Figure \ref{fig:fig8} ilustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.
-
-\begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental.eps}
-     \caption{Traffic selection for incremental approach.}
-     \label{fig:fig8}
-\end{figure}
-
-The eigen similarity analysis starts at $\boldsymbol{x}^{(3)}_{(41)}$ and is incrementally performed until $\boldsymbol{x}^{(\hat{3})}_{(60)}$, in order to calculate the $s_n$.
-We assume that $s_n < l$ means an attack identification, according the anomaly on similiarity of $s_n$ to a defined limiar $l$.
-Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}$ for eigen similarity analyis, the $s_n$ is calculated according to~(\ref{eq:eq11}).
-
-If $s_n = 1$, then the two eigenvectors are completely similar and no anomaly is detected. Smaller values of $s_n$ mean less similarity and can indicate an anomaly, according to a defined 
-threshold $l$, assuming that if $s_n < l$, then a network attack is identified during the $n$-th minute.
-Therefore, the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack is identified, according to
-
-\begin{equation}\label{eq:eq14}
-  \boldsymbol{\hat{n}_{(n)}}=\left\{
-  \begin{array}{@{}ll@{}}
-    1, & \text{if}\ s_n < l \\
-    0, & \text{otherwise}
-  \end{array}\right.
-\end{equation}
-where $\boldsymbol{\hat{n}}_{(n)}$ denotes a vector of $n$-th minutes detected as under attack.
-
-The eigen similarity analysis can also be applied in an individual fashion, where each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, as shown by Figure \ref{fig:fig9}.
-
-\begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/individualized.eps}
-     \caption{Traffic selection for individual approach.}
-     \label{fig:fig9}
-\end{figure}
-
-The incremental and the individual approaches can be combined to obtain the incremental individualized approach, where each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
-Subsequently, $\boldsymbol{X}_n$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:fig2}.
-
-\begin{figure}[h!]
-     \includegraphics[height=4cm, width=12.4cm]{results/figures/incremental_individualized.eps}
-     \caption{Traffic selection for incremental individualized approach.}
-     \label{fig:fig2}
-\end{figure}
-
-This approach of incremental similarity analysis followed by individual analysis after an attack detection allows to identify the attack period, highlighting the first and last time under attack.
-This identification is possible due to the variation of the most significant eigenvectors, which becomes more significant when compared a traffic under attack against a traffic with no attack, according to results which are discussed in Section \ref{sec:experimentalresults}.
-
-\subsubsection{Port Similarity Analysis}
-\label{sec:prop_PortSimilarityAnalysis}
-
-Given $\boldsymbol{\hat{n}}$, which is the set of $n$-th minutes under attack, it is still necessary to obtain more details about the identified network attack, such as the network ports that are attacked during each $n$-th minute identified as under attack.
-Hence, it is also applied the cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port during a $n$-th minute.
-
-For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack shall be used as reference for similarity analysis against the $\boldsymbol{v}_{(n)}$ identified as under attack, individually evaluating the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.
-
-Therefore, $\boldsymbol{v}^{(q)}$ should be calculated from the last $\boldsymbol{X}^{(q)}$ time frame without attack, and $\boldsymbol{v}_{(m,\hat{n})}$ should be calculated from the same traffic appened of all $n$-th minutes until the identified minute under attack, denoted as $\boldsymbol{X}_n$.
-
-For similarity analyis, each $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, denoted as $x_{(m,n)}$, shall be individually replaced by the traffic of the evaluated $m$-th port of the $\hat{n}$-th minute under attack, denoted as $x^{(\hat{q})}_{(m,\hat{n})}$, in order to identify significant variation on similarity caused by the traffic of the $m$-th port.
-
-This approach for detection of ports under attack via similarity analysis is given by
-
-\begin{equation}\label{eq:eq15}
-  \left\{
-  \begin{array}{@{}ll@{}}
-    x_{(m,n)} = x^{(\hat{q})}_{(m,\hat{n})} \\
-    \\
-    s_{m,\hat{n}} = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(m,\hat{n})}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(m,\hat{n})}}},
-  \end{array}\right.
-\end{equation}
-
-where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, which is used to calculate the $\boldsymbol{v}_{(m,\hat{n})}$ most significant eigenvectors that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.
-
-Once $\boldsymbol{v}^{(q)}$ and $\boldsymbol{v}_{(m,\hat{n})}$ are obtained, then the $s_{m,\hat{n}}$ similarity degree can be calculated in order to identify if the traffic replacement highlights the adition of anomalous traffic by the evaluated $m$-th port during the $\hat{n}$-th minute previously identified as under attack.
-
-This procedure should be repeated for each $m$-th target port of $\boldsymbol{\hat{n}}$, in order to individually identify the network ports under attack during each $\hat{q}$-th time frame.
-
-\section{Experiments and Results}
-\label{sec:experimentalresults}
-
-This section presents the performed experiments and the acquired results.
-First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
-Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario.
-In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated dataset.
-Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario.
-Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 dataset.
-
-\subsection{Experimental Scenario}
-\label{sec:AnalyzedScenario}
-
-The experiment time is 120 minutes, separated into six time frames, with each time frame corresponding to twenty minutes.
-Therefore, as the time of each sampling period is one minute, then $N = 20$.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, as well as a covariance $\boldsymbol{\hat{R}}_{yy}^{(q)} \in \mathbb{R}^{17 \times 17}$ (calculated via (\ref{eq:eq03})) and a sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)} \in \mathbb{R}^{17x17}$, assuming that $q = 1, 2, 3, 4, 5$ and $6$.
-
-The simulation started at 21:00h, the first time frame was from 21:00h until 21:20h ($q = 1$), the second was from 21:20h until 21:40h ($q = 2$), the third was from 21:40h to 22:00h ($q = 3$), the fourth was from 22:00h until 22:20h ($q = 4$), the fifth was from 22:20h until 22:40h ($q = 5$), and finally, the sixth was from 22:40h until 23.00h ($q = 6$).
-During the simulation, the victim made legitimate access, and the attacker performed the following attacks: at 21:54h ($q = 3$) was performed a port scan, at the interval ranging from 22:10h to 22:20h ($q = 4$) a synflood attack was simulated, and at the interval from 22:30h to 22:40h ($q = 5$) a fraggle attack was performed.
-
-\subsection{Largest Eigenvalues Analysis}
-\label{sec:largesteigenvaluesanalysis}
-
-For the evaluation of MOS Schemes accuracy for flood and port scan detection, the framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:prop_getv}.
-Through eigenvalue analysis of traffic with flood or port scan attacks, it is possible to visualize a significant difference between the largest eigenvalues and the remain eigenvalues, which can indicate a relationship between an outlier and time frames under attack.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_synflood.eps} 
-     \caption{Eigenvalues of the sample covariance matrix (synflood).}
-     \label{fig:fig10}
-\end{figure}
-
-Figure \ref{fig:fig10} depicts the eigenvalues calculated from sample covariance matrix of the network traffic used to evaluate the synflood attack identification.
-In Figure \ref{fig:fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 4$) stands out significantly from the other eigenvalues.
-
-Figure \ref{fig:fig11} illustrates the eigenvalues calculated from sample covariance matrix of the matrix used for fraggle attack detection.
-In Figure \ref{fig:fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 5$) stands out significantly from the other eigenvalues, in accordance with the result shown in Figure \ref{fig:fig10} for the synflood attack analysis.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_fraggle.eps}
-     \caption{Eigenvalues of the sample covariance matrix (fraggle).}
-     \label{fig:fig11}
-\end{figure}
-
-Figure \ref{fig:fig12} depicts the eigenvalues calculated from covariance matrix of zero mean and unitary standard deviation variables, of the network traffic matrix evaluated for port scan detection.
-
-\begin{figure}[h!]
-	\centering
-     \includegraphics[height=6cm, width=8cm]{results/figures/eigenvalues_portscan.eps}
-     \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}
-     \label{fig:fig12}
-\end{figure}
-
-As analyzed for the synflood and fraggle attacks, note that the largest eigenvalue, related to this attack ($q = 3$), stands out significantly from the others eigenvalues.
-
-Table \ref{tab:tab3} presents the values of the largest eigenvalues of each time frame $q$-th for port scan, synflood and fraggle detection.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Largest Eigenvalue related to attacks detection}
-  \label{tab:tab3}
-  \begin{tabular}{ c c c c c }
-	\toprule
-	\multirow{3}{*}{\textbf{Time Frame} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
-			\hhline{~----}
-		&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
-		&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
-	\midrule
-	1 &1887545 &1887545 &1887545 &2,0734 \\
-	2 &2341327 &2341327 &2341327 &2,1451 \\
-	3 &3213867 &3213867 &3213867 &10,0718 \\
-	4 &133238294 &133238294 &731229 &2,1620 \\
-	5 &92384021611 &6367983 &92384021611 &2,4253 \\
-	6 &708335 &708335 &708335 &1,7948 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-In Table \ref{tab:tab3}, note the significant variation of the eigenvalues associated with attacks, in comparison to the others.
-At $q = 4$, where the synflood attack occurred, the maximum eigenvalue obtained is approximately 21 times larger than the second one.
-At $q = 5$, where the fraggle attack occurred, the maximum eigenvalue obtained is about 29,000 times larger than the second one.
-At $q = 3$, where the port scan attack occurred, the maximum eigenvalue obtained is approximately 4 times larger than the second one.
-In the last case, for port scan attack detection, although the largest eigenvalue presented no too large variance to the second one, if compared to synflood or fraggle attacks, it clearly deviates from the remaining largest eigenvalues.
-
-These results highlight that all $q$-th time frames where a network attack was simulated, present high significant variance between the largest eigenvalue and the remaining eigenvalues, obtained from sample covariance matrix, for flood detection, or from covariance matrix of zero mean and unitary standard deviation variables, for port scan detection.
-Therefore, we propose to apply the vector of the largest eigenvalues to MOS schemes in order to evaluate their accuracy for identification of time frames under attack, motivated by the fact that it is relevant to apply MOS schemes to automate the attack detection process, taking into account the characteristics of the evaluated eigenvalues.
-
-\subsection{MOS Schemes Evaluation}
-\label{sec:MOSSchemesEvaluation}
-
-In \cite{tenorio2013greatest}, we evaluate the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE MOS schemes \cite{da2009comparison,tenorio2013greatest} for synflood and port scan attack detection.
-In this work we extend that evaluation for fraggle attack detection, applying the same schemes to fraggle attack detection over the traffic presented in Section \ref{sec:datamodel}, as results shown in Table \ref{tab:tab4}.
-
-\begin{table}[h!]
-  \centering
-  \scriptsize
-  \caption{MOS schemes applied to port scan and flood detection}
-  \label{tab:tab4}
-  \begin{tabular}{ c c c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
-			\hhline{~------~}
-		&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
-	\midrule
-	Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-	Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-	\midrule
-	Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-	Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-	\midrule
-	Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
-	Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
-	\midrule
-	Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
-	Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-Note that $\hat{d} = 1$, if there is attack, while $\hat{d} > 1$ indicates more than one attack.
-An example of this could be seen for attack detection via EFT for traffic containing synflood and fraggle attacks, showing $\hat{d} = 2$, which indicates the presence of two attacks, as expected by the $d$ real values of Table \ref{tab:tab4}.
-
-In Table \ref{tab:tab4}, two MOS schemes outperforms from the others, EDC and EFT.
-Efficient Detection Criterion (EDC) and Exponential Fitting Test (EFT) are the most effective schemes, correctly estimating the number of attacks in comparison to the expected values for effetctive attack detection, as defined by the column of real values in Table \ref{tab:tab4}.
-The AIC and MDL schemes are satisfactory only for port scan detection, however SURE and RADOI schemes did not show effective results for port scan or flood detection.
-
-Although EDC and EFT presented the same accuracy on the evaluation, the EDC scheme requires less processing time than EFT, which is an important criteria to select EDC as the MOS scheme for flood and port scan detection on the remain experiments.
-
-According to Table \ref{tab:tab4}, EDC and EFT estimated correctly the number of attacks of a time frame vector, indicating that ocurred $\hat{d}$ network attacks, but not providing additional details, what highlights the necessity of complementary approaches in order to estimate the time and ports under attack.
-Hence, we propose apply eigen analysis to estimate the $q$-th time frames under attack and eigen similarity analysis to estimate the minutes and ports under attack.
-
-\subsection{Eigenvalue Analysis}
-\label{sec:EigenvalueAnalysis}
-
-According to the results presented in Section \ref{sec:largesteigenvaluesanalysis}, the largest eigenvalue stands out significantly from the others eigenvalues of an evaluated $q$-th time frame.
-This behavior can also be observed in the largest eigenvalues analysis, according to results presented in Table \ref{tab:tab3}, where it is possible to observe that the $\hat{d}$ largest eigen values of the time frames under attacks stand out significantly from the others largest eigenvalues.
-
-Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respectives $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:alg01}.
-
-\subsection{Eigen Similarity Analysis}
-\label{sec:EigenSimilarityAnalysis}
-
-This paper proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$.
-Hence, the proposed framework is applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.
-
-\subsubsection{Time Analysis}
-\label{sec:TimeAnalysis}
-
-Three approaches were evaluated for eigen similarity analysis: incremental, individual and incremental individualized approaches.
-For the incremental individualized approach, each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
-Subsequently, $\boldsymbol{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated.
-For the incremental approach, each $n$-th minute must be incrementally appended into $\boldsymbol{X}^{(q)}$, for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual time similarity analyis.
-For the individual approach, each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, without incremental append, but doing individual appended into $\boldsymbol{X}^{(q)}$ for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual similarity analyis.
-
-Table \ref{tab:tab5} presents the results of the evaluation of three approaches for similarity analysis of eigenvectors for port scan detection.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Port Scan Detection}
-  \label{tab:tab5}
-  \begin{tabular}{ c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~~---~}
-			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
-	\midrule
-	3 &1 &0.9946 &0.9946 &0.9946 &no \\
-	3 &2 &0.9934 &0.9934 &0.9999 &no \\
-	3 &3 &0.9912 &0.9912 &0.9999 &no \\
-	3 &4 &0.9888 &0.9888 &0.9999 &no \\
-	3 &5 &0.9856 &0.9856 &0.9998 &no \\
-	3 &6 &0.9840 &0.9840 &0.9999 &no \\
-	3 &7 &0.9824 &0.9824 &1.0000 &no \\
-	3 &8 &0.9794 &0.9794 &0.9999 &no \\
-	3 &9 &0.9673 &0.9673 &0.9926 &no \\
-	3 &10 &0.9674 &0.9674 &0.9997 &no \\
-	3 &11 &0.9733 &0.9733 &0.9993 &no \\
-	3 &12 &0.9702 &0.9702 &0.9993 &no \\
-	3 &13 &0.9677 &0.9677 &0.9999 &no \\
-	3 &14 &0.9646 &0.9646 &0.9998 &no \\
-	3 &15 &0.0216 &0.0216 &0.0276 &yes \\
-	3 &16 &0.9621 &0.0209 &1.0000 &no \\
-	3 &17 &0.9611 &0.0199 &0.9998 &no \\
-	3 &18 &0.9612 &0.0191 &0.9999 &no \\
-	3 &19 &0.9613 &0.0186 &0.9998 &no \\
-	3 &20 &0.9638 &0.0190 &1.0000 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-Table \ref{tab:tab5} shows the evaluation of the time frame $q=3$, when the port scan attack was simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
-According to the presented results, it is possible to observe the high similarity between network traffic without attack, which was larger than 0.9610 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with the simulated port scan attack ($n=15$), which was lower than 0.0276 for all evaluated approaches.
-
-Comparing the approaches for similarity analysis, it is possible to observe that all approaches highlight the low similarity when evaluated the traffic under attack.
-However, the incremental approach figured out low similarity for times without attack, where $n=16, 17, 18, 19, 20$, what indicates that the incremental approach can produce false posive results.
-This behavior occurs because the incremental approaches appends all selected traffic into the reference traffic for comparison against the original reference traffic, what makes more evident the first lack of similarity but reduces the changing detection capability after an attack detection.
-
-Table \ref{tab:tab6} presents the results of the evaluation of the similarity analysis of eigenvectors for synflood detection.
-It shows the evaluation of the time frame $q=4$, when the synflood attack is simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
-According to the results, it is possible to observe the high similarity between network traffic without attack, which is larger than 0.9907 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with synflood attack (between $n=11$ and $n=20$), which is lower than 0.1244 for all evaluated approaches.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Synflood Detection}
-  \label{tab:tab6}
-  \begin{tabular}{ c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~~---~}
-			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
-	\midrule
-	4 &1 &1.0000 &1.0000 &1.0000 &no \\
-	4 &2 &0.9999 &0.9999 &1.0000 &no \\
-	4 &3 &0.9997 &0.9997 &0.9999 &no \\
-	4 &4 &0.9998 &0.9998 &1.0000 &no \\
-	4 &5 &0.9965 &0.9965 &0.9908 &no \\
-	4 &6 &0.9975 &0.9975 &1.0000 &no \\
-	4 &7 &0.9977 &0.9977 &1.0000 &no \\
-	4 &8 &0.9980 &0.9980 &1.0000 &no \\
-	4 &9 &0.9987 &0.9987 &0.9999 &no \\
-	4 &10 &0.9991 &0.9991 &1.0000 &no \\
-	4 &11 &0.0085 &0.0085 &0.0284 &yes \\
-	4 &12 &0.0162 &0.0120 &0.0343 &yes \\
-	4 &13 &0.0248 &0.0158 &0.0427 &yes \\
-	4 &14 &0.1243 &0.0185 &0.1041 &yes \\
-	4 &15 &0.0082 &0.0162 &0.0103 &yes \\
-	4 &16 &0.0404 &0.0070 &0.0580 &yes \\
-	4 &17 &0.0397 &0.0007 &0.0573 &yes \\
-	4 &18 &0.0408 &0.0042 &0.0584 &yes \\
-	4 &19 &0.0408 &0.0079 &0.0584 &yes \\
-	4 &20 &0.0477 &0.0092 &0.0757 &yes \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The incremental approach produces better results if compared with other evaluated approaches, with lower values and maximum of 0.0185 for times under attack, but this approach presents change detection limitation after the first outlier of similarity, as shown in Table \ref{tab:tab5} for port scan detection.
-
-Comparing the incremental individualized and the individual approaches for eigen similarity analysis, it is possible to observe that the incremental individualized approach obtain lowest values for almost all cases, except for the time $n=14$, where incremental individualized approach identified a larger similarity than the individual approach.
-The incremental individualized appends information about each evaluated traffic, therefore it incorporates traffic behaviors that can reduce the outlier capability detection, as ocurred for the time $n=14$.
-
-Table \ref{tab:tab7} presents the results of the eigen similarity analysis evaluation for fraggle detection.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Fraggle Detection}
-  \label{tab:tab7}
-  \begin{tabular}{ c c c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~~---~}
-			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
-	\midrule
-	5 &1 &1.0000 &1.0000 &1.0000 &no \\
-	5 &2 &0.9999 &0.9999 &1.0000 &no \\
-	5 &3 &1.0000 &1.0000 &1.0000 &no \\
-	5 &4 &0.9999 &0.9999 &1.0000 &no \\
-	5 &5 &0.9993 &0.9993 &0.9997 &no \\
-	5 &6 &0.9993 &0.9993 &0.9997 &no \\
-	5 &7 &0.9994 &0.9994 &1.0000 &no \\
-	5 &8 &0.9995 &0.9995 &1.0000 &no \\
-	5 &9 &0.9995 &0.9995 &1.0000 &no \\
-	5 &10 &0.9995 &0.9995 &1.0000 &no \\
-	5 &11 &0.0031 &0.0031 &0.0021 &yes \\
-	5 &12 &0.0019 &0.0025 &0.0009 &yes \\
-	5 &13 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &14 &0.0030 &0.0027 &0.0020 &yes \\
-	5 &15 &0.0030 &0.0028 &0.0020 &yes \\
-	5 &16 &0.0012 &0.0025 &0.0002 &yes \\
-	5 &17 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &18 &0.0030 &0.0026 &0.0020 &yes \\
-	5 &19 &0.0030 &0.0027 &0.0020 &yes \\
-	5 &20 &0.0069 &0.0023 &0.0083 &yes \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-For fraggle attack detection, the lack of similarity between legitimate and malicious traffic was more evident than for the evaluation of synflood and port scan detecion.
-This behavior can be explained by the number of packets generated through the fraggle attack simulation, that was significative larger than the number of packets generated during the synflood simulation.
-Considering the three approaches, the largest value for times under attack was 0.0083, while the shortest value for times without attacks was 0.9993.
-
-Therefore, considering the evaluation for port scan, synflood and fraggle detection, the incremental approach can produce false positive results, while the individual and incremental individualized approaches produce quite similar results, even though the individual approach be more simple and require less memory and processing time.
-
-These results highlight the capability of change detection based on similarity between legitimate and malicious traffic from flood or port scan attacks, endorsing the effectiveness and safety for adoption of threshold for attack detection through eigen similarity analysis.
-
-\subsubsection{Port Analysis}
-\label{sec:PortAnalysis}
-
-Given $\hat{N}$, which is the set of estimated $n$-th minutes under attack, it is possible to apply cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port, during a $n$-th minute.
-Therefore, the incremental individualized and individual approaches of eigen similarity analysis were evaluated, for detection of ports under flood and port scan attacks, according to results presented in following tables.
-For this evaluation, the $v$ last most significant eigenvectors without attack was used as reference for similarity analysis against each target port $m$-th.
-
-Table \ref{tab:tab8} presents the results of the evaluation of eigen similarity analysis for detection of ports under port scan attack, showing only the time frame $q=3$ and minute $n=15$, due to the simulated port scan attack occurred only at this time, although the remain time frame has been completely evaluated and presented high similarity to the reference of traffic without network attack.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n=15)}
-  \label{tab:tab8}
-  \begin{tabular}{ c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~--~}
-			&\textbf{Incremental Individualized} &\textbf{Individual}\\
-	\midrule
-	80 &0.9999 &0.9999 &no \\
-	443 &0.9999 &0.9999 &no \\
-	53 &0.9999 &0.9999 &no \\
-	21 &0.9999 &0.9997 &yes \\
-	22 &0.0298 &0.9997 &yes \\
-	23 &0.0298 &0.9997 &yes \\
-	25 &0.0298 &0.9997 &yes \\
-	110 &0.0298 &0.9997 &yes \\
-	143 &0.0298 &0.9997 &yes \\
-	161 &0.0298 &0.9997 &yes \\
-	69 &0.0298 &0.9997 &yes \\
-	123 &0.0298 &0.9997 &yes \\
-	445 &0.0298 &0.9997 &yes \\
-	600 &0.9999 &0.9999 &no \\
-	19 &0.9999 &0.9999 &no \\
-	67 &0.9999 &0.9999 &no \\
-	68 &0.9999 &0.9999 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The incremental individualized approach presented more sensibility to anomaly detection than the individual approach, the former produced the identification of a low similarity of 0.0298 for almost all ports under attack, unless the port 21, although the simulation has attacked this port.
-The individual approach was not able to identify low similarity for ports under attack, resunting in values of 0.9997 for ports with anomalous traffic and 0.9999 for ports without network attack.
-
-For the evaluation of the proposed approaches for identification of ports under synflood and fraggle attack, all minutes of each time frame, in which one attack location was estimated, were analyzed.
-Even though, due to space limitations, only the results of the firt minute where a low similarity was idenfified will be shown such as where $n=11$.
-Nevertheless, the results obtained for the evaluation of traffic without attack presented high similarity to the reference traffic, with similarities close to 0.9999, and the evaluation of the other minutes under attack presented results quite similar to the resuts shown in the Tables \ref{tab:tab9} and \ref{tab:tab10}.
-
-Table \ref{tab:tab9} presents the results of the evaluation of eigen similarity analysis for detection of ports under synflood attack, showing only the time frame $q=4$ and minute $n=11$.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
-  \label{tab:tab9}
-  \begin{tabular}{ c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~--~}
-			&\textbf{Incremental Individualized} &\textbf{Individual}\\
-	\midrule
-	80 &1.0000 &1.0000 &no \\
-	443 &1.0000 &1.0000 &no \\
-	53 &1.0000 &1.0000 &no \\
-	21 &1.0000 &1.0000 &nos \\
-	22 &1.0000 &1.0000 &no \\
-	23 &1.0000 &1.0000 &no \\
-	25 &1.0000 &1.0000 &no \\
-	110 &1.0000 &1.0000 &no \\
-	143 &1.0000 &1.0000 &no \\
-	161 &1.0000 &1.0000 &no \\
-	69 &1.0000 &1.0000 &no \\
-	123 &1.0000 &1.0000 &no \\
-	445 &1.0000 &1.0000 &no \\
-	600 &0.0077 &0.0427 &yes \\
-	19 &1.0000 &1.0000 &no \\
-	67 &1.0000 &1.0000 &no \\
-	68 &1.0000 &1.0000 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-According to results presented in Table \ref{tab:tab9}, both approches identifies low similarity for the traffic of port 600, which is the target port of the simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.
-
-Table \ref{tab:tab10} presents the results of the evaluation of eigen similarity analysis for detection of ports under fraggle attack, showing only the time frame $q=5$ and minute $n=11$.
-
-\begin{table}[h!]
-  \centering
-  \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Fraggle Attack (q=5 and t=11)}
-  \label{tab:tab10}
-  \begin{tabular}{ c c c c }
-	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
-			\hhline{~--~}
-			&\textbf{Incremental Individualized} &\textbf{Individual}\\
-	\midrule
-	80 &1.0000 &1.0000 &no \\
-	443 &1.0000 &1.0000 &no \\
-	53 &1.0000 &1.0000 &no \\
-	21 &1.0000 &1.0000 &no \\
-	22 &1.0000 &1.0000 &no \\
-	23 &1.0000 &1.0000 &no \\
-	25 &1.0000 &1.0000 &no \\
-	110 &1.0000 &1.0000 &no \\
-	143 &1.0000 &1.0000 &no \\
-	161 &1.0000 &1.0000 &no \\
-	69 &1.0000 &1.0000 &no \\
-	123 &1.0000 &1.0000 &no \\
-	445 &1.0000 &1.0000 &no \\
-	600 &1.0000 &1.0000 &no \\
-	19 &0.0031 &0.0004 &yes \\
-	67 &1.0000 &1.0000 &no \\
-	68 &1.0000 &1.0000 &no \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-The results for the avaluation of ports under fraggle attack, shown in Table \ref{tab:tab10}, were similar to the results obtained for synflood analysis, with the identification of low similarity for traffic of the port under attack.
-Nevertheless, for fraggle analysis, the individual approach identified the lowest similarity, that is 0.0004 while the incremental individualized approach obtained a similarity of 0.0031.
-
-The incremental individualized approach was able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches presented false positives or low sensibility to eigen similarity analyis for network attack detection.
-This approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.
-
-According to the shown significant lack of similarity between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
-
-\subsection{DARPA Scenario}
-\label{sec:DarpaEvaluation}
-
-This subsection presents a summarized view of results obtained from the application of the proposed framework, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 dataset.
-Since the proposed framework is detailed in Section \ref{sec:prop_getv} and in Subsections \ref{sec:AnalyzedScenario}, \ref{sec:largesteigenvaluesanalysis}, \ref{sec:MOSSchemesEvaluation}, \ref{sec:EigenvalueAnalysis} and \ref{sec:EigenSimilarityAnalysis}, here the focus is on the parameter selection, dataset evaluation and results for flood and probe attack identification.
-
-The DARPA dataset includes 7 weeks of sniffed traffic saved into raw packet data.
-The traffic and the labeled attacks are grouped by week and day, with information of the number and types of attacks per day, but also providing the start time for each labeled attack.
-For this evaluation, an evaluation per day was performed, considering the network traffic of 24 hours split into $Q$ time frames of 60 minutes ($N = 60$) and aggregate by minute and by port number.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ is obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.
-
-Since the proposed framework focus on flood and probe attack detection, only the attacks with behavior similar to flood or probe attack were evaluated.
-Initially all DoS and probe attacks were selected, but it was observed that the most cases of DoS focus on exploit system vulnerabilities instead of flooding attack, and most of probe attacks focus on ICMP intead of port scanning.
-Therefore, for evaluation of the proposed approach for flood and probe attack detection, it is necessary to select cases that implements flood or port scan behaviors.
-The following week-day-attack cases  were selected:
-
-\begin{enumerate}
-	\item week3-thursday-neptune;
-	\item week4-friday-portsweep;
-	\item week5-thursday-neptune;
-	\item week5-thursday-portsweep;
-	\item week5-friday-portsweep;
-	\item week6-wednesday-neptune;
-	\item week6-thursday-neptune;
-	\item week7-wednesday-portsweep.
-\end{enumerate}
-
-\begin{table*}[!t]
-	\caption{Results of the attack detection evaluation}
-	\label{tab:tab12}
-	\centering
-	\begin{tabular}{|c|c|c|c|c|}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Solution\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Attack Type\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Metric\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Result\end{tabular} \\ \hline
-	Proposed Work	&Flooding	&True Positive	&100.00 \%\\ \hline
-	Proposed Work	&Flooding	&False Positive	&60.00 \%\\ \hline
-	Proposed Work	&Flooding	&Misclassification	&50.00 \%\\ \hline
-	Proposed Work	&Probe	&True Positive	&76.92 \%\\ \hline
-	Proposed Work	&Probe	&False Positive	&18.52 \%\\ \hline
-	Proposed Work	&Probe	&Misclassification	&32.73 \%\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&True Positive	&82.00 \%\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&False Positive	&-\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&Misclassification	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Overall	&True Positive	&94.67 \%\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Overall	&False Positive	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Overall	&Misclassification	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&True Positive	&50.00 \%\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&False Positive	&-\\ \hline
-	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&Misclassification	&-\\ \hline
-	\end{tabular}
-\end{table*}
-
-Table \ref{tab:tab12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}.
-The analysis based on sample covariation of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames.
-The results also show 50.00 \% of misclassification rate , which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
-The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time  from one source to one target, with no variation on IP source or target port.
-This observation corroborates with previous evaluations of the DARPA dataset that highlight issues regarding traffic redundancy.
-
-The analysis based on covariance of zero mean and unitary standard deviation variables was evaluated for port scan attacks, including probe attacks and DoS attacks that send few packets for several ports in order to exploit some vulnerability.
-The results show rates of 76.92 \% for TP detection and 18.52 \% for FP detection from 94 time frames.
-The observed misclassification rate for this scenario is 32.73 \%.
-It was observed that all FN cases are probe attacks with a time delay between scanning one port and start scanning the next port, what can be called as sparse probe attacks.
-Cases with a delay of one minute or more were not detected by the proposed approach.
-
-The performance of detection rate of flooding attacks is compared with the method proposed by Callegari \emph{et al} \cite{Zonglin2009}.
-This work is a statistical method, based on PCA, without training or learning methods, even though it relies on visual analysis for principal components selection.
-The best detection rate of \cite{Zonglin2009} was 82.00 \% for detection of synthetically added flood attacks, while this current proposal obtains 100.00 \% of detection rate for detection of flood attacks of DARPA dataset.
-
-Due to the lack of statistical techniques without training or learning methods for detection of probe attacks, this proposed approach is compared to the Lu and Ghorbani's \cite{Lu2009} proposal, which is a network anomaly detection model based on signal processing techniques that uses DARPA dataset for evaluation.
-The results of \cite{Lu2009} show the best detection rate of 94.67 \% in terms of general attack instance detection, but shows a case case with 50.00 \% of attack instance detection for the portsweep attack.
-The proposed approach presents 76.92 \% of detection rate measured specifically for probe attacks, without the requirement of learning or training methods, in contrast to Lu and Ghorbani's \cite{Lu2009} work.
-
-\section{Performance Evaluation}
-\label{sec:Complexity}
-
-This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.
-
-\subsection{Complexity Analysis}
-\label{sec:ComplexityAnalysis}
-The EVD, calculated according to (\ref{eq:eq06}), requires the previous calculation of covariance matrix, according to Equations (\ref{eq:eq02}), (\ref{eq:eq03}), (\ref{eq:eq04}) and (\ref{eq:eq05}).
-The covariance matrix calculation is $O(M^2N)$ and the EVD is $O(N^3)$, where $M$ denotes the number of network ports and $N$ denotes the period time.
-Therefore, the computational complexity for all steps for EVD can be represented as $O(M^2N + N^3)$ and yields an $O(N^3)$ upper bound on the worst-case running time for EVD.
-
-EDC and EFT are the MOS schemes that presented accuracy on the evaluation for the network attack detection.
-The computational complexity evaluation for MOS focuses on EDC scheme, since EDC requires less processing time than EFT but presents the same accuracy for the evaluated scenario.
-EDC scheme is $O(Q \log Q + Q + Q \log Q)$ and its worst-case running time can be represented as $O(Q \log Q)$, where $Q$ denotes the number of time frames.
-
-The largest eigenvalue analysis is $O(\hat{d}Q)$, where $\hat{d}$ denotes the number of time frame under attack, according to Algorithm \ref{alg:alg01}.
-Subsequently, the eigen similarity analysis relies on EVD and cosine similarity analysis, which is $O(N^2)$, for $\hat{d}$ time frames, therefore the eigen similarity analysis is which is $O(\hat{d}(M^2N + N^3 + N^2))$ and yields an $O(N^3)$ upper bound on the worst-case running time for eigen similarity analysis.
-
-Therefore, the proposed framework is $O(N^3 + Q \log Q + \hat{d}Q + N^3)$ and its worst-case running time is $O(N^3)$.
-The computational complexity of EVD is predominant in the framework, but the approach splits the data into time frames with period time $N$, which makes possible to limitate the growth of $N$ even for evaluations of cases with total time larger than $N$, reducing the impact caused by the computational complexity of EVD.
-
-\subsection{Processing Time Analysis}
-\label{sec:ProcessingAnalysis}
-For better understanding the scalability and impact of configurations of $N$, $M$ and $Q$, the processing time required for different scenarios of parameter configurations and dataset were evaluated, measuring the processing time of:
-
-\begin{enumerate}
-	\item Eigen analysis based on sample covariance of zero mean;
-	\item Eigen analysis based on sample covariance of zero mean and unitary standard deviation;
-	\item EDC MOS scheme;
-\end{enumerate}
-
-The performance evaluation focus on the main steps, which are discussed in subsection \ref{sec:ComplexityAnalysis} regarding the complexity analysis.
-The data modeling is also a time consuming step, however its processing can be optimized through distributed processing techniques, such as MapReduce, achieving high throughput for packet counting or even for deep packet inspection \cite{Vieira2013}.
-
-The experiments were performed in a desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM, considering: variations on the network traffic time; the frame size denoted as $N$; the number of network ports denoted as $M$; the mean processing time for eigen analysis based on sample covariance of zero mean, denoted as 1-time; the mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation, denoted as 2-time; and the mean processing time for EDC MOS scheme, denoted as 3-time.
-The mean time calculations was obtained from 200 measurement repetitions, in order to obtain reliable values.
-
-\begin{table*}[!t]
-	\caption{Processing time of the main steps for anomaly detection}
-	\label{tab:11}
-	\centering
-	\begin{tabular}{|r|r|r|r|r|r|r|}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Traffic Time\\(hour)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Frame Size\\(min)\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Num.
-Ports\end{tabular}	& \begin{tabular}[x]{@{}l@{}}1-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}2-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}3-time\\(ms)\end{tabular}	\\ \hline
-16	& 10	& 17	& 0.7900	& 0.8100	& 0.0650\\ \hline
-16	& 20	& 17	& 0.5250	& 0.5950	& 0.0100\\ \hline
-16	& 60	& 17	& 0.9700	& 1.1400	& 0.0250\\ \hline
-16	& 120	& 17	& 0.6050	& 0.6100	& 0.0050\\ \hline
-16	& 60	& 34	& 1.2750	& 1.2200	& 0.0050\\ \hline
-16	& 120	& 34	& 1.1200	& 1.1700	& 0.0050\\ \hline
-20	& 10	& 17	& 2.7950	& 2.8950	& 1.1000\\ \hline
-20	& 20	& 17	& 2.0700	& 2.0200	& 0.3500\\ \hline
-20	& 60	& 17	& 1.0250	& 1.0450	& 0.0650\\ \hline
-20	& 120	& 17	& 1.0000	& 1.0700	& 0.0350\\ \hline
-20	& 60	& 34	& 2.9650	& 3.2100	& 0.0400\\ \hline
-20	& 120	& 34	& 2.9950	& 3.1150	& 0.0200\\ \hline
-22	& 10	& 17	& 4.7250	& 4.0850	& 1.4600\\ \hline
-22	& 20	& 17	& 2.3200	& 2.6800	& 0.2450\\ \hline
-22	& 60	& 17	& 1.0700	& 1.1200	& 0.0300\\ \hline
-22	& 120	& 17	& 0.9900	& 1.0500	& 0.0250\\ \hline
-22	& 60	& 34	& 3.0850	& 3.1250	& 0.0650\\ \hline
-22	& 120	& 34	& 2.8100	& 2.9600	& 0.0250\\ \hline
-	\end{tabular}
-\end{table*}
-
-Table \ref{tab:11} presents the measured results.
-The experiment considered traffic time of 16, 20 and 22 hours, according to the selected traffic time per day available by the DARPA dataset.
-Note that the processing time increases according to the increment in traffic time, around 2 or 3 times for 1-time and 2-time, but the worst measured processing time is 4.7250 milliseconds.
-
-According to Table \ref{tab:11}, the processing time increases with the frame size $N$ decreasing, therefore it is possible to evaluate the frame size that produces better identification rates and acceptable processing time.
-The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations, since the significant increase of processing time observed between scenarios considering 17 or 34 ports, with growth between 7 \% and 199 \%.
-
-\section{Conclusion and Future Works}
-\label{sec:conclusionandfutureworks}
-
-This paper models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
-
-The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
-
-This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
-Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.
-
-The incremental individualized approach of eigen similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to eigen similarity analyis for network attack detection.
-Therefore, the incremental individualized approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.
-
-According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
-
-Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets.
-Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high thoughput network traffic.
-Additionally, future research can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.
-
 \appendix

### 2019-05-23 docs/thesis/chapters/1_introduction/introduction.tex
prepare for thesis writing

@@
 
-Cloud computing is a new rapidly evolving paradigm in the world of distributed networking and computation.
-The basic features of the cloud environment is providing elastic, on-demand and secure services for the end-users.
-While the first two requirements are rather well conceptualized and supported by the majority of the cloud platforms in use, security is a serious concern of the cloud providers and governmental organizations as well as academia and research community \cite{csa2016,higashi2015,gartner2015}.
-Although for the small and medium-sized enterprises (SME) the cloud environment is often the most cost-effective and easily scalable solution, the security and privacy of the sensitive data in cloud platforms are also not fully conceptualized, leading to obscure and incomplete security paradigms and solutions \cite{galibus2017offline}.
+Anomaly detection can be defined as the identification of rare and suspicions events by differing the normal or majority of the data.
+Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects \cite{bhuyan2014network,ahmed2016survey}.
+Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.
 
-Additional security issues and requirements have to be considered when mobile clients are actively used in corporate cloud environment \cite{yovel2014}.
-Today organizations and enterprises are adopting the Bring-Your-Own-Device (BYOD) paradigm.
-The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms, being the bottleneck of the corporate information security system (ISS).
-While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, the corporate mobile client is usually light-weighted and generally less protected.
+Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or other patterns.
+However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid unknown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detection systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
+To face this problem, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, where it is not necessary known anomalies for training models \cite{moustafa2019holistic}.
 
@@
 Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection.
-
-The field of Mobile Money Transfer (MMT) is a growing market segment and has been motivating investments into security defense and fraud detection.
-Obtaining access to data sets of mobile transactions for research is a very hard task due to the intrinsic private nature of such transactions.
-Scientists and researchers must today spend time and effort in obtaining access to relevant data sets before they can research on such data set.
-However, PaySim \cite{lopez2016paysim} is a financial simulator that simulates mobile money transactions based on an original dataset and provides a synthetic dataset as an approach to such a problem.
 
@@
 
-Considering that imbalanced data can be composed by rare or unknown events, adaptive techniques can be useful in order to improve models and be able to react to changes or unknown patterns and maintain acceptable performance for classification algorithms.
-Adaptive techniques are also important for online data analysis or streaming processing, where an incremental processing approach can be applied to deal with high throughput data analysis.
+Feature engineering and feature selection.
 
-Compressed Sensing (CS) aims to obtain the most relevant information of a dataset, what makes it useful for compression, signal reconstruction, image processing and classification, such as in sparse representation classification (SRC) problems.
-Dictionary learning is related to CS but seeks to obtain compact and meaningful signal representations, known as dictionary, from learning algorithms or models.
-DL is a signal processing technique for sparse representation of signals as basis vectors, learning the representations from training data, as dictionaries.
-DL is based on the principle that some observations can be described by a sparse subset of atoms taken from a redundant dictionary, which represents the causes of some observations of the world.
-The sparse representation in terms of such dictionaries has attracted increased interest for compressive sensing and for solving problems such as denoising, compression, image processing, data decomposition, feature extraction and classification \cite{tosic2011dictionary, zhang2010discriminative, zhu2016coupled,ravishankar2011mr}.
-
-Dictionaries are either available analytically, or can be learned from a suitable training set.
-DL can be formulated as an adaptive algorithm that learns from training data that is updated to represent changes or to adapt to new patterns.
-Iterative dictionary learning has been used as a blind technique for feature extraction, improving classification without require feature selection or principal component analysis
-
-In some applications, the data and its dictionary are multidimensional, e.g., when estimating jointly behavior of users in social networks.
-Computing tensor decompositions of multi-way datasets is particularly useful to extract hidden patterns and structure in multidimensional data analytics problems \cite{kolda2009tensor}.
-Tensor-based algorithms for dictionary learning can improve the performance for cases of multidimensional and separable data, regarding the dictionary identification rating, the required number of training samples and iterations for the optimization problem \cite{roemer2014tensor}.
-Existing dictionary learning schemes can be applied to multidimensional analysis and obtain valuable results.
-Additionally, the performance of tensor-based algorithms for recovering of a known separable dictionary outperform existing schemes when dealing with growing multidimensional datasets.
+Robust PCA and distances between new observations and robust subspaca
 
@@
 We propose to use tensor-based dictionary learning for learning fraudulent and legitimate data separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics.
+
+Hiposesis: Subspace learning by means of decomposition can improve network anomaly detection
 

### 2019-05-23 docs/thesis/chapters/4_m_rpca/m_rpca.tex
prepare for thesis writing

--- 2019-05-17 docs/papers/2019JNCA_m-rpca/2019JNCA_m-rpca.tex
@@
-\documentclass[review]{elsarticle}
+\chapter{Moment Distances from Robust Subspace for Botnet Detection}
+\label{ch:4_m_rpca}
 
-\usepackage{lineno,hyperref}
-\usepackage[mathletters]{ucs}
-\usepackage[utf8x]{inputenc}
-\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
-\usepackage{natbib}
-\usepackage{supertabular}
-\usepackage{fancybox}
-\usepackage{acronym}
-\usepackage{array}
-\usepackage{booktabs}
-\usepackage{graphicx}
-\usepackage{rotating}
-\usepackage{tabularx}
-\usepackage{multirow}
-\usepackage{hhline}
-\usepackage{setspace}
-\usepackage{placeins}
-\usepackage{longtable}
-\usepackage{listings}
-\usepackage{epstopdf}
-\usepackage{subfigure}
-\usepackage{courier}
-\usepackage{amsfonts}
-\usepackage{morefloats}
-\usepackage{lipsum}
-\usepackage{mathtools}
-\usepackage{amsthm,amsmath}
-
-% table
-\usepackage{makecell}
-\usepackage{color, colortbl}
-\definecolor{Gray}{gray}{0.9}
-
-% appendix
-\usepackage[titletoc,title]{appendix}
-\usepackage{hyperref}
-\usepackage{cleveref}
-
-% norm
-
-\modulolinenumbers[5]
-
-\journal{Journal of Network and Computer Applications}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
-%% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
-
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
-%% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
-%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-\begin{frontmatter}
-
-\title{Moment Distances from Robust Subspace for Botnet Detection}
-
-%% Group authors per affiliation:
-\author[unbaddress]{Thiago P. B. Vieira}
-\author[unbaddress]{Eduardo Said Calil Vilaça}
-\author[unbaddress,Ilmenauaddress,Fraunhoferaddress]{João Paulo C. L. da Costa}
-\author[unbaddress]{Rafael T. de Sousa Júnior}
-
-\address[unbaddress]{Department of Electrical Engineering, University of Brasilia (UnB), 70910-900, Brasília-DF, Brazil}
-\address[Ilmenauaddress]{Institute for Information Technology, Ilmenau University of Technology, Ilmenau, Germany}
-\address[Fraunhoferaddress]{Fraunhofer Institute for Integrated Circuits IIS, Erlangen, Germany}
-
-\begin{abstract}
-Distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid botnet's attacks.
-This paper proposes an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggregation, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations.
-Experimental results reveal that the proposed approach outperforms widely adopted algorithms and a method based on robust skewness for a public data set of botnet attacks.
-\end{abstract}
-
-\begin{keyword}
-Network Anomaly Detection \sep Botnet Detection \sep Moments \sep Robust Principal Component Analysis \sep Mahalanobis Distance
-\end{keyword}
-
-\end{frontmatter}
-
-\linenumbers
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sec:introduction}
+\begin{quotation}[]{Paulo Freire}
+No one knows it all. No one is ignorant of everything. We all know something. We are all ignorant of something.
+\end{quotation}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% network anomaly detectiom, botnet detection, botnet detection approaches, approaches based on subspace learning, feature generation and aggregation, about the methodos to compare and metrics
-\section{Literature Review}
-\label{sec:review}
+\section{Related Works}
+\label{sec:4_relatedworks}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Data Model}
-\label{sec:datamodel}
+\label{sec:4_datamodel}
 
-In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $α$, $β$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
+In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $\alpha$, $\beta$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Moment Distances from Robust Subspace for Botnet Detection}
-\label{sec:mdrs}
+\label{sec:4_proposal}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiments and Results}
-\label{sec:experimentalresults}
+\section{Experiments}
+\label{sec:4_experiments}
 
@@
 
-there are no lots of researches using CTU-13, although it is been growing poor results for GMM, K-Means and MGM are ok and in accordance to \cite{wang2017botnet}, which also evaluate \cite{gu2007bothunter}
+% 10: moment are better
+% 11: moment are better
+% 12: kd-RPCA is poor for scenario 12, as well as GMM, K-Means and ROBPCA
+% 15: moment are better
+% 15-2: moment are better
+% 15-3: kd-RPCA is poor for , whilhe ROBPCA obtain 0.38 and other moment based has good results
+% 16: kd-RPCA is poor for , as well as sd-RPCA with 0.25s, while GMM obtain 0.50 and other moment based has good results. Explain based on data
+% 16-2: md-RPCA is poor for , as well as GMM, K-means and ROBPCA, but kd-RPCA and sd-RPCA goes well
+% 16-3: kd_RPCA is poor, as well as GMM, K-means and ROBPCA
+% 17: robpca has high result, higher than kd-RPCA, GMM, K-means, but md and sd are better and equal
+% 18: moment based is better
+% 18-2: high GMM e Kmeans, but moment are even better
+% 19: moment are better
+% there are no lots of researches using CTU-13, although it is been growing poor results for GMM, K-Means and MGM are ok and in accordance to \cite{wang2017botnet}, which also evaluate \cite{gu2007bothunter}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion and Future Works}
-\label{sec:conclusionandfutureworks}
+% a questão é diferenciar aquilo que é malicioso ou não, dentre os outliers ou novelties (esses termos são sinônimos para alguns autores)
+% como eu posso diferenciar isto? Eu suponho que a partir de um modelo de tráfego normal eu possa detectar grandes deviations e classificar estas como ataques
+% a hipotese é esta
+% em outra área (compressive sensing), se utiliza dicionários, que são representações compactas de todas as caracteristicas de um dado
+% no nosso caso, o treinamento nos gera uma especie de dicionário, em que os dados normais são representados de forma robusta
+% onde uma forma robusta seria uma forma menos sensível a outliers.
+Isso pode ser visto como a diferenca entre calcular a média e a mediana de um conjunto.
+A presença de um outlier no conjunto irá corromper a média de uma forma maior que a mediana.
+Desta forma, se considera a mediana como sendo robusta
+
+\section{Results}
+\label{sec:4_results}
+
+\section{Conclusion}
+\label{sec:4_conclusion}
 
@@
 say that more studies about skewness and subspace learning can be done in order to explore it inside algorithms and non gaussian data
-
-\section*{References}
-
-\bibliography{references}
-
-\end{document}
-
-% Classical estimates can be so strongly affected by contamination that diagnostic tools, such as the Mahalanobis Distances, become unable to detect the outliers.
-Therefore, we need reliable and robust estimators that can resist outliers contaminated data.
-Minimum Covariance Determinant (MCD) \cite{rousseeuw1984mcd} is a robust estimator commonly used for computing Mahalanobis Distance in order to detect anomalies.
-
-% \begin{figure}[h!]
-%      \centering
-%      \includegraphics[width=11cm]{figures/mahalanobis_robust.png}
-%      \caption{Mahalanobis Distance for non-robust and robust estimates.}
-%      \label{fig:fig04}
-% \end{figure}
-
-% \begin{figure}[h!]
-%      \centering
-%      \includegraphics[width=8cm]{figures/mahalanobis_distances01.png}
-%      \caption{Mahalanobis Distance for non-robust and robust estimates.}
-%      \label{fig:fig05}
-% \end{figure}
-
-% The MCD estimator is one of the first affine equivariant and highly robust estimators of multivariate location and scatter.
-
-% Affine equivariant means that when the data are translated or subjected to a linear transformation, the resulting location and scatter will transform accordingly \cite{rousseeuw1984mcd,rousseeuw1999fastmcd}.
-
-% Its objective is to find $h$ observations (out of $n$) whose covariance matrix has the lowest determinant.
-
-% The lowest determinant means the lowest distance generalized variance and high distance between the largest eigenvalue and the remaining [ref?].
-
-% MCD depends on random sampling, therefore it is not deterministic.
-
-% MCD has been applied in numerous fields (medicine, finance, image analysis, and chemistry) and been used to develop many robust multivariate techniques (such as Robust Principal Component Analysis (RPCA), factor analysis, and multiple regression).
-
-% The classical MCD has rarely been applied because it is hard to compute, since it requires the evaluation of all $\binom{n}{h}$ subsets of size $h$. % binomial of n over p
-
-% Its main use has started since the construction of the computationally efficient Fast-MCD algorithm \cite{rousseeuw1999fastmcd}.
-
-% Fast-MCD is an iterative, approximate and resampling algorithm for the MCD.
-
-% It starts by random initial subset of size $p+1$ and performs concentration steps (C-steps) yielding consecutive h-subsets with decreasing covariance matrix determinant.
-
-% Only two C-steps are applied to each initial subset, and the 10 results with lowest determinant are kept.
-
-% C-steps are carried out until convergence and the best solution is kept.
-
-% Fast-MCD also is affine equivariant and not deterministic.
-
-% \textbf{Theorem 1: C-Step.}
-
-% Consider a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ where $\boldsymbol{X} = \{\boldsymbol{x}_1,\ldots,\boldsymbol{x}_n\}$ of $p$-variate observations.
-Let $\boldsymbol{h} \subset \{1,\ldots,n\}$ with $|\boldsymbol{h}| = h$, where $\boldsymbol{h}$ is a subset of $h$ indexes from $n$ observations.
-
-% and the relative distance be
-
-% \begin{equation}\label{eq:eq04}
-% 	\boldsymbol{d}(i) = \sqrt{(\boldsymbol{x}_i - \bar{\boldsymbol{x}}) \boldsymbol{\hat{S}}^{-1}(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime}, 
-% \end{equation}
-
-% for $i = 1,\ldots,n$.
-
-% Let $\boldsymbol{h}_1, \bar{\boldsymbol{x}}_1$ and $\boldsymbol{\hat{S}}_1$ and be the selected observations, the location and scatter of the initial subset, and $\boldsymbol{d}_1(i)$ be the initial distances for $i = 1,\ldots,n$.
-
-% Now take $\boldsymbol{h}_2$ containing the indexes of lowest distances, such that $\{\boldsymbol{d}_1(i); i \in \boldsymbol{h}_2\} = \{(\boldsymbol{d}_1)_{1:n},\ldots,(\boldsymbol{d}_1)_{h:n}\}$, where $(\boldsymbol{d}_1)_{1:n} \leq (\boldsymbol{d}_1)_{2:n} \leq \ldots \leq (\boldsymbol{d}_1)_{n:n}$ are the ordered distances, and compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$.
-
-% Then, $det(\boldsymbol{\hat{S}}_2) \leq det(\boldsymbol{\hat{S}}_1)$ with equality if and only if $\boldsymbol{\bar{x}}_2 = \boldsymbol{\bar{x}}_1$ and $\boldsymbol{\bar{S}}_2 = \boldsymbol{\hat{S}}_1$
-
-% The proof is given by Rousseeuw and Driessen \cite{rousseeuw1999fastmcd}.
-
-% \textbf{The C-step can be algorithmically described as follows:}
-% \begin{algorithm}
-% 	\label{alg:alg01}
-% 	\scriptsize
-% 	\SetAlgoLined
-% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
-% 	Given the initial $h-$subset $\boldsymbol{h}_1$ of indexes or the pair $(\boldsymbol{\bar{x}}_1, \boldsymbol{\hat{S}}_1)$\;
-% 	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
-% 		\If{$\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ exist}{
-% 			Put $\boldsymbol{\bar{x}}_1 = \boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_1 = \boldsymbol{\hat{S}}_2$\;
-% 		}
-% 		Compute the distances $\boldsymbol{d}_1( i )$ for $i = 1,\ldots, n$\;
-% 		Sort $\boldsymbol{d}_1( i )$, which yields an index $\boldsymbol{d}_1(\pi(1)) \leq \ldots \leq \boldsymbol{d}_1(\pi(n))$\;
-% 		Put $\boldsymbol{h}_2 = \{ \pi(1), \pi(2), \ldots, \pi(h)\}$\;
-% 		Compute $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$ for $\boldsymbol{h}_2$\;
-% 		Compute $det(\boldsymbol{\hat{S}}_2)$ and $det(\boldsymbol{\hat{S}}_1)$\;			
-% 	}
-% 	\caption{C-Step}
-% \end{algorithm}
-
-% \textbf{Theorem 1} thus provides a partial idea for an algorithm:
-
-% \textit{Take many initial choices of $\boldsymbol{h}_1$ and apply C-steps to each until convergence (don't have determinant decreasing), and keep the result with lowest determinant}.
-
-% \textbf{Initial subset can be algorithmically described as:}
-% \begin{algorithm}
-% 	\label{alg:alg02}
-% 	\scriptsize
-% 	\SetAlgoLined
-% 	\KwResult{$\boldsymbol{h}_1$}
-% 	Given a data set $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $p$-variate observations\;
-% 	Draw a random $(p + 1)$-subset $\boldsymbol{L}$\;
-% 	Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
-% 	\If{$det(\boldsymbol{\hat{S}}_1) \leq 0$}{
-% 		Add another random observation into $\boldsymbol{L}$\;
-% 		Compute $\boldsymbol{\bar{x}}_1$ and $\boldsymbol{\hat{S}}_1$ of $\boldsymbol{L}$ and $det(\boldsymbol{\hat{S}}_1)$\;
-% 	}
-% 	Compute $\boldsymbol{d}_1^2(i) = (\boldsymbol{x}_i - \boldsymbol{\bar{x}}_1)^\prime \boldsymbol{\hat{S}}_1^{-1}(\boldsymbol{X}_i - \boldsymbol{\bar{x}}_0)$ for $i = 1, \ldots, n$\;
-% 	Sort them into $\boldsymbol{d}_1(\pi(1)) \leq \ldots \leq \boldsymbol{d}_0(\pi(n))$\;
-% 	Put $\boldsymbol{h}_1 = \{\pi(1), \ldots, \pi(h)\}$\;
-% 	\caption{Constructing the initial subset}
-% \end{algorithm}
-
-% \begin{algorithm}
-% 	\label{alg:alg03}
-% 	\scriptsize
-% 	\SetAlgoLined
-% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
-% 	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n \leq 600$\;
-% 	\For{500 times}{
-% 		Construct initial $h$-subset $\boldsymbol{h}_1$ (Algorithm \ref{alg:alg02})\;
-% 		Carry out two C-steps (Algorithm \ref{alg:alg01}) resulting $\boldsymbol{\bar{x}}_2$ and $\boldsymbol{\hat{S}}_2$\;
-% 	}		
-% 	Put $\boldsymbol{h}_1$ as the 10 lowest $det(\boldsymbol{\hat{S}_2})$ computed above\;
-% 	\While{$det(\boldsymbol{\hat{S}}_2) < det(\boldsymbol{\hat{S}}_1)$}{
-% 		Carry out C-steps (Algorithm \ref{alg:alg01})\;
-% 	}
-% 	\caption{Fast-MCD when $n \leq 600$}
-% \end{algorithm}
-
-% \begin{algorithm}
-% 	\label{alg:alg04}
-% 	\scriptsize
-% 	\SetAlgoLined
-% 	\KwResult{$\boldsymbol{\bar{x}}_2$, $\boldsymbol{\hat{S}}_2$}
-% 	Given $\boldsymbol{X} \in \mathbb{R}^{n \times p}$, $h < n$, $p \geq 2$ and $n > 600$\;
-% 	Construct up to 5 disjoint random subsets of size $n_{sub} = 300$\;
-% 	\For{each subset, repeat 100 times}{
-% 		Construct initial $\boldsymbol{h}_1$ of size $h_{sub} = [n_{sub}(h/n)]$\;
-% 		Carry out two C-steps of $n_{sub}$ and $h_{sub}$\;
-% 		Keep the 10 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$\;
-% 	}		
-% 	Pool the subsets, yielding the merged set of $n_{merged} = 1500$\;
-% 	\For{50 best $(\boldsymbol{\bar{x}}_{sub}, \boldsymbol{\hat{S}}_{sub})$}{
-% 		Carry out two C-steps of $n_{merged}$ and $h_{merged} = [n_{merged}(h/n)]$\;
-% 		Keep the 10 best $(\boldsymbol{\bar{x}}_{merged}, \boldsymbol{\hat{S}}_{merged})$\;
-% 	}
-% 	\For{the full data set and $m_{full}$ best results}{
-% 		Take several C-steps, using $n$ and $h$\;
-% 		Keep the best final result $(\boldsymbol{\bar{x}}_{full}, \boldsymbol{\hat{S}}_{full})$\;
-% 	}
-% 	\caption{Fast-MCD when $n > 600$}
-% \end{algorithm}
-
-% \cite{garcia2014empirical} proposed CTU-13 data set and argues that for evaluations of anomaly detection, none of the botnet families used in the training and crossvalidation data set should be used in the testing data set, considering that this ensures that the methods can generalize and detect new behaviors.
-However, hour proposal is semi-supervised and does not rely on training data with anomalies.
-Adopting the training and testing approach of \cite{garcia2014empirical}, some botnets malware wouldnt be evaluated, since in Garcia's proposal some botnets are present only for training.
-Therefore, we evaluate each scenario individually.
-
-% \cite{wang2017botnet}
-% A. Description of CTU-13 data set
-% The CTU-13 is a data set of botnet traffic that was captured
-% in the Czech Technical University [24]. It contains 13 scenarios
-% with various botnet types. We consider scenarios 1, 2, 6, 8, and
-% 9 of the data set.
-% • Scenario 1 corresponds to an IRC-based botnet that sent spam for almost 6.5 hours.
-% • Scenario 2 (2.5-hours) is from the same botnet.
-% • Scenario 6 is from a botnet that scanned SMTP (Simple Mail Transfer Protocol) servers for 2 hours and connected to several RDP (Remote Desktop Protocol) services.
-Dif-erent with Scenarios 1 and 2, this botnet neither sent spam nor did it attack.
-It’s C\&C server used a proprietary protocol.
-% • Scenario 8 is from a botnet that contacted a lot of Chinese C\&C hosts and received large amounts of encrypted data. The botnet also cracked the passwords of machines during the 19-hour attack.
-% • Scenario 9 corresponds to a case where 10 local hosts were infected by a spamming botnet. More than 600 spams were sent over 5 hours.
-
-% Initially, the data $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ is split into normal data $\textbf{X}^r$ for robust subspace learning, and into $\textbf{X}^t$ for testing the anomaly detection on contaminated data, that can also be viewed as new observations of a production environment.

### 2019-05-23 docs/thesis/chapters/resumo.tex
prepare for thesis writing

Último passo, por ser um resumo da introduçao, resultados e conclusão.

\begin{keywords}
Discriminative Sensing, Signal Processing, Eigen Similarity Analysis, Tensor Based Dictionary Learning, Critical Factor Analysis, Network Attack Detection, Fraud Detection
\end{keywords}

### 2019-05-27 docs/thesis/chapters/1_introduction/introduction.tex
first deliverable

@@
 Information security may consist of both technical and procedural aspects.
-The former includes equipment and security systems, while the latter corresponds to security rules and recommendations.
+The former includes equipment and security systems, while the latter corresponds to security rules and recommendations of good practices.
 Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks.
 Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.
-Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
 
@@
 Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects \cite{bhuyan2014network,ahmed2016survey}.
+Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
 Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.
 
-Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or other patterns.
+Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or behavioral analysis.
 However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid unknown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detection systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
 To face this problem, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, where it is not necessary known anomalies for training models \cite{moustafa2019holistic}.
-
-The protection scheme used on a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}.
-Therefore, encrypting files and generating keys on a mobile device is not considered a good solution.
-On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}.
-The common practice is the shadow user activities monitoring \cite{yovel2014}.
-However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode.
-When the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools cannot help and the mobile client has to secure itself with its own limited resources.
-Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection.
 
@@
 
-The high availability of raw data increases the challenges related to big data analytics and to imbalanced data, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes.
-The fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms.
+The high availability of raw data increases the challenges related to big data analytics and to imbalanced learning problems, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes.
+The fundamental issue with the imbalanced learning is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms.
 Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection.
 
-Feature engineering and feature selection.
-
-Robust PCA and distances between new observations and robust subspaca
+Some widely adopted algorithms assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network,leon2017probability}.
+It was observed that the skeweness and heavy-tailed distributions can impact algorithms that rely on gaussian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.
 
@@
 
-Considering the above described landscape, this thesis outlines the development and evaluation of approaches based on signal processing for information security analysis, through methods to make the data discriminative and able to identify structures, hidden patterns and the most relevant information for network attack detection, mobile malicious behavior analysis and fraud detection in mobile money transactions.
+Considering the above described landscape, this thesis outlines the development and evaluation of approaches based on subspace learning for network attack detection, through methods to make the data discriminative and able to identify structures, hidden patterns and the most relevant information for anomaly detection.
+In the Subsection \ref{sc:Hypothesis} we present our hypothesis formulation and the Subsection \ref{sc:proposals} describes the proposals to answer the questions and validate the hypotheses.
 
-In the context of anomaly-based schemes, this thesis proposes a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks.
-Inspired by \cite{david2011blind,da2012improved}, this thesis models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
+\subsection{Hypothesis Formulation}
+\label{sc:Hypothesis}
+
+For the experimental evaluation of our proposals, we adopt a methodology based on aspects of GQM (Goal-Question-Metric) template \citep{Basili1994} and define two questions to achieve our goal, which are:
+
+\begin{itemize}
+	\item $Q_1$: Can the analysis of patterns from a learned subspace identify and detect anomalies in network traffic?
+	\item $Q_2$: Can the robust subspace learning improve the anomaly detection in imbalanced and skewed data?
+\end{itemize}
+
+Our testing hypotheses are defined in Table \ref{tab:hypothesis}, that describe the null hypothesis and alternative hypothesis for each previously defined question.
+
+\begin{table}[htb]
+	\centering
+	\caption{Hypotheses to evaluate the defined metrics}
+	\label{tab:hypothesis}
+    \begin{tabular}{|p{6cm}|p{6cm}|c|} \hline
+        \textbf{Alternative Hypothesis}	&\textbf{Null Hypothesis}	&\textbf{Question} 	\\ \hline
+        $H_{1sub}$: It is possible to use subspace learning to identify and detect anomalies in network traffic.
+        &$H_{0sub}$. It is not possible to use subspace learning to identify and detect anomalies in network traffic.	&$Q_1$\\ \hline
+        $H_{1rob.sub}$: The robust subspace learning improves the anomaly detection from imbalanced and skewed data.	&$H_{0rob.sub}$.
+The robust subspace learning does not improves the anomaly detection from imbalanced and skewed data.	&$Q_2$\\ \hline
+    \end{tabular}
+\end{table}
+
+The hypotheses $H_{1num.indct}$ and $H_{0num.indct}$ were defined to evaluate if a subspace learned by singular value decomposition are sensitive to outliers and can be used to detect network attacks.
+We defined the hypotheses $H_{1scale.prop}$ and $H_{0scale.prop}$ to evaluate the feasibility and performance of a robust subspace learned by Robust PCA to network anomaly detection.
+
+The metrics most adopted to evaluate experimental results of network attack detection are the true positive (TP), true negative (TN), false positive (FP) and false negative (FN).
+We adopt these metrics and also adopt the misclassification rate and the F1-score, which is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic}.
+
+\subsection{Proposals}
+\label{sc:problem_statement}
+
+In the context of anomaly-based schemes, this thesis proposes the Eigensim, which is a approach based on subspace learning techniques for detection of malicious traffic in computer networks.
+
+Inspired by \cite{david2011blind,da2012improved}, Eigensim models the network traffic using a signal processing formulation as a composition of three components: signal, artifact and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
 The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
-In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack.
-In addition, we also evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset.
-Furthermore, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.
+In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect detailed time frames under attack.
+We evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 data set \citep{osanaiye2016distributed}, which is a well known network traffic data set.
+Furthermore, this proposed approach is validated into a prove of concept regarding behavioral anomaly detection to detect possible threats to a offline corporate mobile app.
 
-This thesis also proposes an architecture and approach based on user behavior analysis of mobile apps through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats, and to reduce the risks and the harm of the most common security threats to offline corporate mobile apps, which are the expired user misusing password and the intruder attack.
-Additionally, the behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.
-The main target of this proposed solution is to provide a maximum defense at the minimal resource cost.
+Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}.
+Learning algorithms for imbalanced data has been a challenging research topic, considering that the fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
 
-Considering that big data problems require techniques to deal with multidimensional data in order to make sense of structure and relationship of many dimensions, and also considering that a key challenge to use sparse coding and dictionary learning for classification is how to find proper dictionaries and coefficients that highlight discriminative structure and relationships of one dataset, we also propose a tensor-based dictionary learning method for fraud detection from imbalanced data, in order to apply the tensor decomposition for dictionary learning methods to evidentiate the discriminative sensing of a fraud detection dataset.
-Specifically, we propose to apply a sparse representation based classification (SRC) method through learning a tensor-based dictionary and evaluate the reconstruction error for fraud classification of mobile money transactions.
-We propose to use tensor-based dictionary learning for learning fraudulent and legitimate data separately, and apply the learned dictionaries to reconstruct a test signal and classify it as fraud or legitimate, according to the minimum reconstruction error measured by some metrics.
+Some widely adopted algorithms assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network, leon2017probability}.
+Therefore, the skeweness and heavy-tailed distributions can impact algorithms that rely on gaussian distributed data, as well as it can be exploited to reveal characteristics that can improve classifiers for network anomaly detection.
 
-Hiposesis: Subspace learning by means of decomposition can improve network anomaly detection
+We believe that the skewness of anomalous and normal data can highlight features for improving anomaly detection in imbalanced data.
+We also believe that the distance between robust estimates of normal and contaminated data can highlight discrepancies and be used anomaly detection and for network attack detection.
+Therefore, we propose the m-RPCA, which is an approach based on  distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data.
 
@@
 
-We analyze problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques.
+We analyze problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques based on subspace learning.
 The results of the work presented in this thesis provide the following contributions:
@@
 \begin{enumerate}
-	\item We propose an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset;
+	\item We propose an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluate the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 data set;
 	\item We discuss the computational complexity of the proposed framework and evaluate the required processing time for tested scenarios;
-	\item We propose an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture;
-	\item We discuss the processing time of the proposed framework for mobile devices;
-	\item We propose a tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
+	\item We propose an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture, and discuss the processing time of the proposed framework for mobile devices;
+	\item We propose an approach based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data, and evaluate the anomaly detection rate on synthetic and real data sets;
 \end{enumerate}
@@
 This thesis is organized as follows.
-In Chapter \ref{ch:2_mos_eig_sim}, we propose a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
-Chapter \ref{ch:3_mobile} presents an evaluation of an approach and architecture based on user behavior analysis through Model Order Selection (MOS) \cite{tenorio2013greatest}, in order to detect possible threats in a mobile application.
-Chapter \ref{ch:4_tensor_dl} proposes a tensor-based dictionary learning approach for fraud detection from mobile payment transactions.
+In Chapter \ref{ch:2_mos_eig_sim}, we propose the Eigensim, which is an approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
+Chapter \ref{ch:3_mobile} presents a prove of concept regarding the evaluation of an approach and architecture based on user behavior analysis through the Eigensim  \cite{tenorio2013greatest}, in order to detect threats in a mobile application.
+Chapter \ref{ch:4_m_rpca} proposes the m-RPCA, which is an approach based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data.
 Chapter \ref{ch:5_conclusionfuturework} draws the conclusions and the suggestions for future work.

### 2019-05-27 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
prepare for thesis writing
first deliverable

@@
-\chapter{Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
+\chapter{Eigensim - Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
 \label{ch:2_mos_eig_sim}
@@
 
-In the context of anomaly-based schemes, this thesis proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks.
-Inspired by \cite{david2011blind,da2012improved}, this thesis models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
-To the best of our knowledge there is no similar model in the literature.
-The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
+This chapter proposes the Eigensim, which is an approach based on signal processing concepts applied to detection of malicious traffic in computer networks.
+Inspired by \cite{david2011blind,da2012improved}, the Eigensim models the network traffic using a signal processing formulation as a composition of three components: signal, artifact and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
+The Eigensim is based on eigenvalue analysis, model order selection (MOS) and similarity analysis between eigenvectors.
 In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack.
-In addition, we also evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset.
-Furthermore, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.
-
-The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
+In addition, we evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 data set \citep{osanaiye2016distributed}, which is a well known network traffic data set.
+
+The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, i.e.
+in an unsupervised approach, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
 The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
@@
 In Section \ref{sec:2_relatedworks}, related works are discussed.
-Section \ref{sec:2_datamodel} presents the data model and the evaluated datasets.
+Section \ref{sec:2_datamodel} presents the data model and the evaluated data sets.
 Section \ref{sec:2_prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks.
@@
 Section \ref{sec:2_conclusionandfutureworks} draws the conclusions and the suggestions for future work.
-The \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.
+The Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.
 
@@
 Principal Component Analysis (PCA) is a statistical technique commonly used for dimensionality reduction.
-It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
-PCA has been used in attack detection \cite{almotairi2009technique}.
+It uses an orthogonal transformation to convert a set of correlated variables into a new subspace of linearly uncorrelated variables, by means of eigenvalue value decomposition, where the first principal components have the largest variance.
+PCA has been used in attack detection \cite{almotairi2009technique}, considering that this technique is very sensitive to outliers and adopted for classification problems.
 However, PCA requires human intervention in order to identify abnormalities based on the eigenvalues profiles, if used without complementary techniques.
 
-Callegari \emph{et al} \cite{Zonglin2009} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data.
+Callegari \emph{et al} \cite{Zonglin2009} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a data set with synthetic anomalies added in the data.
 However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
 
-Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
-In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
+Lee \emph{et al.} \cite{Lee2013} presented the OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
+In contrast to Lee \emph{et al.}, the Eigensim applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
 Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
-
-Besides being prone to higher errors and false positives, such human intervention makes PCA not useful for real time applications.
-Therefore, in order to automate the analysis of eigenvalues profile, model order selection (MOS) schemes should be incorporated.
 
@@
 However, their work requires a training method to produce a prediction model for normal daily traffic and presents limitations on identification of behaviors without significant outliers, such as port scan attacks.
-Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
+Zonglin \emph{et al} \cite{Zonglin2009} proposed a signal processing method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
 Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work is not applied to probe and flood attack detection, simultaneously.
@@
 Blind automatic detection of malicious traffic techniques have been developed for honeypots in \cite{david2011blind,da2012improved}.
-However, traffic on honeypot is simpler than real network traffic, because there are no running legitimate applications, due to the fact that honeypots emulate behavior of a host within a network to deceive and lure attackers \cite{zakaria2012review}.
+However, traffic on honeypot is simpler than real network traffic, because there are no running legitimate applications, i.e.
+background traffic, due to the fact that honeypots emulate behavior of a host within a network to deceive and lure attackers \cite{zakaria2012review}.
 Since honeypots do not generate legitimate traffic, the amount of data captured in honeypots is significantly lower in comparison to a Network Intrusion Detection System (NIDS), which captures and analyzes the largest possible amount of network traffic \cite{david2011blind}.
@@
 
-Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks.
-Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
-NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications.
-Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations.
-Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD datasets adopt a traffic aggregation by connections, we consider the use of a experimental scenario on a real network and the DARPA 1998 dataset, which is the source for the creation of the KDD 99 and NSL-KDD datasets.
-Note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD datasets.
+Several approaches for network attack detection uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
+Even though the KDD 99 data set are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data sets adopted by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
+
+NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that redundant data records are removed, in order to avoid biased classifications.
+Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public data sets their evaluations.
+Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD data sets adopt a traffic aggregation by connections, we consider the use of a experimental scenario on a real network and the DARPA 1998 data set, which is the source for the creation of the KDD 99 and NSL-KDD datas ets.
+Note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD data sets.
 
@@
 
-This section presents details of the experimental scenario and the selected cases of the DARPA dataset, along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic.
+This section also presents details of the experimental scenario and the selected cases of the DARPA data set, along with a description of the data set model as a signal superposition of signal, artifact and noise, that refer to legitimate, control and malicious traffic.
 Subsection \ref{sec:2_DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
 Subsection \ref{sec:2_ModelingData} presents how network traffic can be modeled as signal superposition.
-Subsection \ref{sec:2_SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:2_DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.
+Subsection \ref{sec:2_SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:2_Darpadata set} discusses the use of the DARPA data set for evaluation of the proposed approach.
 
@@
 
-The environment of the analyzed scenario is composed of two computers and one router with access to the Internet and to an internal network, where the simulation of legitimate traffic, noise, flood and port scan attacks are performed.
-During the traffic generation, one computer assumes the role of the attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:2_fig1}.
-
+The environment of the analyzed scenario is composed by two computers and one router, with access to the Internet and to an internal network.
+In this scenario we performed the simulation of legitimate traffic, control, flood and port scan attacks.
+During the traffic generation, one computer assumes the role of the attacker, while the other is the victim, according to scenario represented by Figure \ref{fig:2.01}.
+The victim generate legitimate traffic by means of ordinary use of Internet resources, and control traffic is constantly generated for purposes of network administration and monitoring.
+
+%TODO - recreate the figure and change the concepts
 \begin{figure}[h!]
@@
      \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
-     \label{fig:2_fig1}
-\end{figure}
-
-Note that the set of network traffic is modeled as legitimate, noise and malicious traffic, where the victim performs legitimate activities, that can be characterized by web access.
-In many organizations this type of traffic is predominant, since most of corporate services are web-based, such as: web pages, customized web-based systems and cloud services.
-It is possible to characterize the traffic of a DHCP service as an example of noise associated with the transport layer.
+     \label{fig:2.01}
+\end{figure}
+
+Note that the set of network traffic is modeled as legitimate, controls and malicious traffic, which refer to signal, artifact and noise, respectively.
+In many organizations the web based traffic is predominant, since most of corporate services are  web pages, customized web-based systems and cloud services.
+It is possible to characterize the traffic of a DHCP service as an example of artifact associated with the transport layer, as well as it would be possible to classify heart beat traffic and other seasonal and controlled traffic as artifact.
 For malicious traffic, three types of networks attacks are evaluated: synflood, fraggle and port scan.
-These attacks are reproduced using well-known security tools, such as Nmap\footnote{http://nmap.org} to port scan, Metasploit\footnote{http://www.metasploit.com} to synflood and Hping\footnote{http://hping.org} to lead the fraggle attack.
+Here we refer to port scan as an attack, according to \cite{ahmed2016survey,moustafa2019holistic}, however it is one approach usually adopted for acquire information in order to perform an attack.
+These attacks are reproduced using well-known security tools, such as Nmap\footnote{http://nmap.org} for port scan, Metasploit\footnote{http://www.metasploit.com} for synflood and Hping\footnote{http://hping.org} to lead the fraggle attack.
 
@@
 
-In the proposed framework, the goal is to detect the anomalies only taking into account the traffic profile, i.e., specific information such as origin IP or day and time of the attack are not considered.
-Therefore, from the entire log information, we just consider the timestamp (for sequencing), port type and port number.
+In the proposed framework, the goal is to detect the anomalies only taking into account the traffic profile, i.e.
+specific information such as origin or destination IP, behavioral pattern or content of the attack are not considered.
+Therefore, IP spoofing or data encryption would not cause impact to the proposed approach and evaluation, since our proposal only relies on the timestamp (for sequencing) and port number.
 
@@
 
-By modeling the dataset as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{U}), noise (\textbf{N}) and malicious traffic (\textbf{A}), according to the following expression:
-
-\begin{equation}\label{eq:eq01}
-	\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{N}^{(q)} + \boldsymbol{A}^{(q)},
-\end{equation}
-where $q$ represents the $q$-th time frame, which is a time grouping of network traffic.
-The matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns.
-Each row represents a communication port (TCP port or UDP port), and each column represents time bins having a appropriate size, such as one minute.
-Each element $x_{m,n}^{(q)}$ stands for the number of times that the port $m$ appears at the $n$-th minute, at the $q$-th time frame.
-
-The legitimate traffic $\boldsymbol{U}^{(q)}$ is characterized by the traffic from user's operations.
-When a user accesses a web page, for example, there is the corresponding TCP/IP traffic to request the page, as well as there is the traffic required to domain name resolution.
-Figure \ref{fig:2_fig3} depicts an example of the legitimate traffic obtained during experiments.
-
+By modeling the data set as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{U}), artifact (\textbf{A}) and malicious traffic (\textbf{N}), according to the following expression:
+
+\begin{equation}\label{eq:2.01}
+	\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{A}^{(q)} + \boldsymbol{N}^{(q)},
+\end{equation}
+where $q$ represents the $q$-th time frame, which is a time aggregation of network traffic.
+The matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns, where each row represents a communication port, and each column represents time bins of a defined size, such as one minute.
+Each element $x_{m,n}^{(q)}$ stands for the number packets that appears at $n$-th minute for the port $m$, during the $q$-th time frame.
+
+The legitimate traffic $\boldsymbol{U}^{(q)}$ is characterized by the traffic from user's ordinary operations, such as when performing access to web pages, as well when there is the traffic required to domain name resolution.
+Figure \ref{fig:2.03} depicts an example of the legitimate traffic simulated during the experiments.
+
+% TODO - Replace
 \begin{figure}[h!]
@@
      \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
-     \label{fig:2_fig3}
-\end{figure}
-
-The traffic that is not associated with user's operations and with malicious traffic is modeled as noise $\boldsymbol{N}^{(q)}$.
-The automatic acquisition service of logical IP network address (DHCP) is an example of noise.
-Independently of any user operation, the machine receives an IP address, since it is configured to perform a DHCP address request.
-Figure \ref{fig:2_fig4} depicts an example of noise in a network traffic, represented by traffic to ports 67 and 68.
-
+     \label{fig:2.03}
+\end{figure}
+
+The traffic that is not associated with user's operations and with malicious traffic is modeled as artifact $\boldsymbol{A}^{(q)}$.
+The acquisition of logical IP network address (DHCP) is an example of artifact, where independently of any user operation, the machine receives an IP address, since it is configured to automatically perform a DHCP address request.
+Figure \ref{fig:2.04} depicts an example of artifact in a network traffic, by means of traffic to ports 67 and 68.
+The traffic coming from a malicious activity, i.e.
+port scanning, synflood or fraggle attacks, is represented by the matrix $\boldsymbol{N}^{(q)}$.
+
+% TODO - Replace
 \begin{figure}[h!]
@@
      \caption{Network traffic of user independent operations for network management.}
-     \label{fig:2_fig4}
-\end{figure}
-
-The traffic coming from a malicious activity, such as a synflood or fraggle attack, is represented by the matrix $\boldsymbol{A}^{(q)}$.
-For this work we only consider the traffic from port scanning and flood attacks.
-
-We define that if the obtained $\#\boldsymbol{A}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\boldsymbol{A}^{(q)} = 0$, then there is no malicious traffic.
-This work shows how to detect the $\#\boldsymbol{A}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious network traffic.
+     \label{fig:2.04}
+\end{figure}
+
+We define that if the $\#\boldsymbol{X}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\boldsymbol{X}^{(q)} = 0$, then there is no malicious traffic.
+We show how to detect the $\#\boldsymbol{X}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious traffic.
 
@@
 The first two attacks can be qualified as flood or denial of service (DoS) attacks, while the last one can be qualified as probe or port scanning attack.
-DoS attempts to block access to system or network resources and is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks.
-Probe attacks scan the networks to collect information about host, such as IP addresses, ports and services.
+
+A DoS is an attempt by an attacker to prevent legitimate access to websites by overwhelming the amount of available bandwidth or resources of the computer system.
+DoS is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks.
+
+Probe attacks scan computer and network systems to collect information about the host, such as open ports, topology, running software or version of technologies, in order to find vulnerabilities.
 
 With respect to the synflood attacks, the attacker sends a large quantity and concurrent successive SYN requests to a target, in order to consume resources and cause a DoS.
-Figure \ref{fig:2_fig5} depicts an example of a synflood attack carried out in a real computer network.
+Figure \ref{fig:2.05} depicts an example of a synflood attack carried out in a real computer network.
 In an interval of ten minutes, more than 210,000 packets are sent as a synflood attack.
@@
 
+% TODO - Replace
 \begin{figure}[h!]
@@
      \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
-     \label{fig:2_fig5}
-\end{figure}
-
-With respect to the fraggle attack, large packets with UDP echo segments are sent to the broadcast address of a network.
+     \label{fig:2.05}
+\end{figure}
+
+Regarding the fraggle attack, large packets with UDP echo segments are sent to the broadcast address of a network.
 Every packet is modified to have the source address of the victim, in order to implement the source address spoofing technique.
 Therefore, each host receives a huge amount of requests UDP echo and all of them replies to the IP address of the victim, causing a packet flooding aiming a DoS.
-This attack can affect the entire network, since all hosts receive several requests UDP echo and respond with the ICMP protocol, therefore each host acts as an amplifier of the attack.
-This last part of the fraggle attack is not taken into account in this work, because the victim receives ICMP (network layer) packets originated from the hosts that are attacked with flooding packet UDP echo.
-Figure \ref{fig:2_fig6} depicts an example of the fraggle attack in a real computer network.
-
+Figure \ref{fig:2.06} depicts an example of the fraggle attack in a real computer network.
+
+The fraggle attack can affect the entire network, since all hosts receive several requests UDP echo and respond with the ICMP protocol, therefore each host acts as an amplifier of the attack.
+This part of the fraggle attack is not taken into account in this work, because the victim receives ICMP (network layer) packets originated from the hosts that are attacked with flooding packet UDP echo.
+
+% TODO - Replace
 \begin{figure}[h!]
@@
      \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
-     \label{fig:2_fig6}
-\end{figure}
-
-More than 6,000,000 malicious packets can be counted in an interval of ten minutes, which can be considered an abnormal network traffic, especially due to the concentrated traffic in a short period of time and due to the similarity of the outstanding traffic.
+     \label{fig:2.06}
+\end{figure}
+
+The Figure \ref{fig:2.06} shows that more than 6,000,000 malicious packets can be counted in an interval of ten minutes, which can be considered an abnormal network traffic, especially due to the concentrated traffic in a short period of time and due to the similarity of the outstanding traffic.
 
@@
 Therefore, a full connection or a complete three-way handshake is never established, which makes the detection of the attack sender more difficult, and requires approaches able to identify probe attacks without connection establishment.
+
 The UDP scan technique sends UDP packets to the destination port, and if it responds with a \emph{ICMP port unreachable} message, then it indicates that the scanned port is closed.
@@
      \caption{Connection attempts in order to identify active ports.}
-     \label{fig:2_fig7}
-\end{figure}
-
-Figure \ref{fig:2_fig7} depicts an example of the port scan attack in a real computer network.
+     \label{fig:2.07}
+\end{figure}
+
+Figure \ref{fig:2.07} depicts an example of the port scan attack in a real computer network.
 Note that the traffic is composed of two packets for each TCP port and one UDP packet to each port.
@@
 
-\subsection{The DARPA Dataset}
-\label{sec:2_DarpaDataset}
-
-The DARPA 1998 dataset\footnote{https://www.ll.mit.edu/ideval/data/} includes 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks.
-The attacks in this dataset can be grouped into: denial-of-service (DoS); remote to local (R2L), which is characterized by unauthorized access from a remote machine; user to root (U2R), which is characterized by unauthorized access to local super-user privileges; and probe attack.
-Since the proposed approach focus on flood and probe attack, the analysis concentrates on the attacks that present behaviors similar to flood or probe attack.
-We observe that the most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack.
-One example is the occurrence of a neptune attack which sends 20 only SYN packets, what is a behavior that differs of the expected flood attack behavior.
+\subsection{The DARPA Data set}
+\label{sec:2_Darpadata set}
+
+The DARPA 1998 data set\footnote{https://www.ll.mit.edu/ideval/data/} includes 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks.
+The attacks in this data set can be grouped into: denial-of-service (DoS); remote to local (R2L), which is characterized by unauthorized access from a remote machine; user to root (U2R), which is characterized by unauthorized access to local super-user privileges; and probe attack.
+Since the proposed approach focus on flood and probe attack, the analysis concentrates on the attacks of the DARPA 1998 data set that present behaviors similar to flood or probe attack.
+
+We observe that the most cases of DoS of DARPA 98 focus on exploit system vulnerabilities instead of on flooding attack.
+One example is the occurrence of a neptune attack which sends 20 SYN packets, what is a behavior that differs of the expected flooding attack behavior.
 Therefore, there were selected the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey,osanaiye2016distributed}, and the cases that scan ports sending just a few packets.
@@
 
-The data modeling follows the method described by the Subsection \ref{sec:2_ModelingData}, with time frames of 20 minutes, packet counting aggregation by minute and considering the traffic to the following ports: 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389, 443.
+The data modeling follows the method described by the Subsection \ref{sec:2_ModelingData}, with time frames of 20 minutes, packet aggregation counting by minute and considering the traffic to the following ports: 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389, 443.
 
@@
 
-This section describes the proposed technique to detect synflood, fraggle and port scan, according to Figure \ref{fig:2_fig80}, which represents the overview of the proposed framework for detection and identification of network attacks.
+This section describes the proposed technique to detect synflood, fraggle and port scan, according to the overview depicted in Figure \ref{fig:2.08}, which represents the proposed framework for detection and identification of network attacks.
 In Subsection \ref{sec:2_prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvalue for each $q$-th time frame.
@@
      \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
-     \label{fig:2_fig80}
+     \label{fig:2.08}
 \end{figure}
@@
 The proposed attack detection algorithm starts by the data pre-processing of a network traffic log containing IP, ports and timestamp of senders and receivers.
-During this step, the desired information is extracted in order to classify and count packets according to the origin and destination ports, and subsequently this information is grouped by minutes and by time frames.
+During this step, the desired information is extracted in order to classify and to count packets according to the origin and destination ports.
+Subsequently, this information is grouped by minutes and by time frames.
 
@@
 
-\begin{equation}\label{eq:eq02}
+\begin{equation}
+\label{eq:2.02}
 \boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
@@
 
-\begin{equation}\label{eq:eq03}
+\begin{equation}\label{eq:2.03}
 \boldsymbol{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
@@
 For the detection of the port scan attack, the main components are not dominated by the variables with large variance.
-Moreover, the portscan traffic presents a highly correlated network traffic.
+Moreover, the port scan traffic presents a highly correlated network traffic between the monitored ports.
 In order to exploit such structure, we compute the sample covariance $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have zero mean and unitary standard deviation as follows
 
-\begin{equation}\label{eq:eq04}
+\begin{equation}
+\label{eq:2.04}
 \boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
@@
 
-\begin{equation}\label{eq:eq05}
+\begin{equation}\label{eq:2.05}
 \boldsymbol{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
@@
 
-Once the $\boldsymbol{\hat{R}}_{yy}^{(q)}$ and $\boldsymbol{\hat{R}}_{zz}^{(q)}$ have been obtained for flood and port scan attack detection, respectively, and since the next steps are the same for both sample covariance matrices, we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{C}$.
-Therefore, the following step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:eq06}), in order to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:eq060}).
-
-\begin{equation}\label{eq:eq06}
-\boldsymbol{C}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
-\end{equation}
-
-\begin{equation}\label{eq:eq060}
+Once the $\boldsymbol{\hat{R}}_{yy}^{(q)}$ and $\boldsymbol{\hat{R}}_{zz}^{(q)}$ have been obtained for flood and port scan detection, respectively, and since the next steps are the same for both sample covariance matrices, we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{\hat{R}}$.
+Therefore, the following step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:2.06}), in order to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:2.06}).
+
+\begin{equation}\label{eq:2.06}
+\boldsymbol{\hat{R}}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
+\end{equation}
+
+\begin{equation}\label{eq:2.060}
 \boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
@@
 
-\begin{equation}\label{eq:eq07}
+\begin{equation}\label{eq:2.07}
 \boldsymbol{E} =
@@
 
-\begin{equation}\label{eq:eq08}
-\boldsymbol{e}_{\rm max} = \boldsymbol{E}\{:,1\} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
+\begin{equation}\label{eq:2.08}
+\boldsymbol{e}_{\rm max} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
 \end{equation}
@@
 Note that some MOS schemes may also require the number of minutes that compose a time frame, as $\hat{d} = \rm{MOS}(\boldsymbol{e}_{\rm max},\emph{Q})$.
-For more information about MOS, we refer to \ref{apx:a_mos}.
+For more information about MOS, we refer to the Appendix \ref{apx:a_mos}.
 
@@
 
-After applying the MOS schemes to the vector $\sim\boldsymbol{e}_{\rm max}$, we obtain the estimate of the $\#\boldsymbol{A}$.
-For instance, in the case of fraggle, synflood and portscan, if $\hat{d} = 1$, , then $\#\boldsymbol{A} = 1$, which means that during the during the $Q$ time frames one attack is present.
-However, if $\hat{d} = 0$, then $\#\boldsymbol{A} = 0$, and this means that none of these attacks are present.
+After applying the MOS schemes to the vector $\sim\boldsymbol{e}_{\rm max}$, we obtain the estimate of the $\#\boldsymbol{X}$.
+For instance, in the case of fraggle, synflood and ports can, if $\hat{d} = 1$, , then $\#\boldsymbol{X} = 1$, which means that during the during the $Q$ time frames one attack is present.
+However, if $\hat{d} = 0$, then $\#\boldsymbol{X} = 0$, and this means that none of these attacks are present.
 Note that $\hat{d}$ can be greater than 1, indicating the presence of more than one attack.
@@
 
-The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed according to Algorithm \ref{alg:2_alg01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$.
-Algorithm \ref{alg:2_alg01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$, according to Line 2 of Algorithm \ref{alg:2_alg01}, and its correspondent index, according to Lines 4 and 5 of Algorithm \ref{alg:2_alg01}.
-Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$, according to Line 8 of Algorithm \ref{alg:2_alg01}, and a new iteration is performed until $\boldsymbol{e}_{\rm max} = \{\}$.
-
-\begin{algorithm}[h!]
+The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed according to Algorithm \ref{alg:2.01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$.
+Algorithm \ref{alg:2.01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$, according to Line 3 of Algorithm \ref{alg:2.01}, and its correspondent index, according to Line 7.
+Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$, according to Line 11 of Algorithm \ref{alg:2.01}, and a new iteration is performed until $\boldsymbol{e}_{\rm max} = []$.
+
+\begin{algorithm}
+	\label{alg:2.01}
+	\SetAlgoLined
+	\KwResult{$\rm{\boldsymbol{\hat{q}}}_{\rm max}$}
+	Given $f = 1$\;
+	\While{$f < \hat{d}$}{
+		$q_{\rm value} = \operatorname*{argmax}_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$\;
+		$i = 1$\;
+    	\While{$i < Q$}{
+    		\If{$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}{
+    			$\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$\;
+    		}
+    		$i = i + 1$\;
+    	}
+    	$\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$\;
+		$f = f + 1$\;
+	}
 	\caption{Detection of Time Frames Under Attack}
-  	\label{alg:2_alg01}
-	\begin{algorithmic}[1]
-		\For{$f = 1$ until $f == \hat{d}$} 
-    	    \State $q_{\rm value} = \argmax\limits_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$
-	    	\For{$i = 1$ until $i == Q$} 
-				\If {$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}
-			    	\State $\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$
-				\EndIf
-    		\EndFor
-    		\State $\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$
-    	\EndFor
-	\end{algorithmic}
 \end{algorithm}
@@
 Cosine similarity calculates the cosine of the angle between two vectors, which represents the similarity of values between the selected vectors.
-Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\boldsymbol{V}^{(q)}$ against the the most significant eigenvectors of time frame detected as under attack, to analyze similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.
+Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\boldsymbol{V}^{(q)}$ against the most significant eigenvectors of time frame detected as under attack, to analyze similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.
 
@@
 \label{sec:2_prop_TimeSimilarityAnalysis}
-
+% ToDo
 For eigen similarity analysis, we evaluate the cosine similarity in order to identify lacks of similarity between legitimate and malicious traffic, as follows
 
-\begin{equation}\label{eq:eq11}
-s_n = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}}},
-\end{equation}
-where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, and $\boldsymbol{v}_{(n)}$ is the most significant eigenvectors obtained after append the target $n$-th minute of traffic to be performed the flood and port scan attack identification.
-
-The most significant eigenvector $\boldsymbol{v}^{(q)}$, of a time frame $q$ without attack, can be derivate from (\ref{eq:eq06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the evaluated matrix.
-The same calculation shall be performed in order to obtain the target eigenvectors $\boldsymbol{v}_{(n)}$, calculated from a time frame without attack plus minutes of a time frame estimated as under attack, to evaluate the occurrence of network attacks.
-
-The reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic without attack, from a time frame $q$ composed of $Q$ minutes of legitimate network traffic.
+\begin{equation}\label{eq:2.11}
+s_n = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}^{(q)}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}^{(q)}}},
+\end{equation}
+where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, according the attack detection described by Algorithm \ref{alg:2.01}, and $\boldsymbol{v}_{(n)}^{(q)}$ is the most significant eigenvectors obtained after append each target $n$-th minute of traffic that needs the identification of flood and port scan attacks.
+
+The most significant eigenvector $\boldsymbol{v}^{(q)}$, of a time frame $q$ without attack, can be computed from (\ref{eq:2.06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the selected time frame $q$.
+The same calculation shall be performed in order to obtain the target eigenvectors $\boldsymbol{v}_{(n)}^{(q)}$, calculated from the time frame without attack appended by selected minutes of a time frame estimated as under attack.
+
+Therefore, the reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic without attack, in a time frame $q$ composed of $Q$ minutes of legitimate network traffic, estimated as normal by Algorithm \ref{alg:2.01}.
 For the detailed attack identification, each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$, as represented by
 
-\begin{equation}\label{eq:eq12}
+\begin{equation}\label{eq:2.12}
 \boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
@@
 
-The resultant $\boldsymbol{X}_{(n)}$ is necessary to obtain $\boldsymbol{v}_{(n)}$, through (\ref{eq:eq06}), for calculating the similarity degree $s_n$, ranging from 0 to 1, for each $n$-th minute.
+Subsequently, the resultant $\boldsymbol{X}_{(n)}$ is used to obtain $\boldsymbol{v}_{(n)}^{(q)}$, through (\ref{eq:2.06}), for calculating the similarity degree $s_n$, ranging from 0 to 1, for each $n$-th minute.
 The $s_n$ denotes the absolute similarity degree of the $n$-th minute in comparison to a well-known traffic without attack, detected through MOS schemes and eigenvalue analysis.
 
-The incremental approach for similarity analysis is based on the incremental appending of network traffic into $\boldsymbol{X}^{(q)}$, where the first evaluation is based on (\ref{eq:eq12}) and the subsequent evaluations is based on (\ref{eq:eq13}), incrementally appending each $n$-th minute until $n=N$.
-
-\begin{equation}\label{eq:eq13}
+We propose three approaches for eigen similarity analysis, which are the incremental, the individual and the incremental individualized.
+
+The incremental approach for eigen similarity analysis is based on the incremental appending of network traffic into $\boldsymbol{X}^{(q)}$, where the first evaluation is based on (\ref{eq:2.12}) and the subsequent evaluations are based on (\ref{eq:2.13}), that denotes the appending of the $n$-th minute $\boldsymbol{x}^{(\hat{q})}_{(n)}$ into $\boldsymbol{X_n}$.
+The incremental approach repeats the Equation \ref{eq:2.13} while $n \leq N$ and compute $\boldsymbol{v}^{(q)}$, $\boldsymbol{v}_{(n)}^{(q)}$ and the Equation \ref{eq:2.11} for each increment.
+
+\begin{equation}\label{eq:2.13}
 \boldsymbol{X}_{n} = \{\boldsymbol{X}_{n} | \boldsymbol{x}^{(\hat{q})}_{(n)}\},
@@
 
-Figure \ref{fig:2_fig8} illustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected.
-
+Figure \ref{fig:2.08} illustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected by Algorithm \ref{alg:2.01}.
+
+%TODO - fix figure that shows s_n > l where should the inverse
 \begin{figure}[h!]
@@
      \caption{Traffic selection for incremental approach.}
-     \label{fig:2_fig8}
-\end{figure}
-
-The eigen similarity analysis starts at $\boldsymbol{x}^{(3)}_{(41)}$ and is incrementally performed until $\boldsymbol{x}^{(\hat{3})}_{(60)}$, in order to calculate the $s_n$.
-We assume that $s_n < l$ means an attack identification, according the anomaly on similarity of $s_n$ to a defined limiar $l$.
-Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}$ for eigen similarity analysis, the $s_n$ is calculated according to~(\ref{eq:eq11}).
-
-If $s_n = 1$, then the two eigenvectors are completely similar and no anomaly is detected. Smaller values of $s_n$ mean less similarity and can indicate an anomaly, according to a defined 
-threshold $l$, assuming that if $s_n < l$, then a network attack is identified during the $n$-th minute.
-Therefore, the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack is identified, according to
-
-\begin{equation}\label{eq:eq14}
+     \label{fig:2.08}
+\end{figure}
+
+For the scenario depicted by Figure \ref{fig:2.08}, the eigen similarity analysis starts at $\boldsymbol{x}^{(3)}_{(41)}$ and is incrementally performed until $\boldsymbol{x}^{(3)}_{(60)}$, in order to calculate the $s_n$.
+We assume that $s_n < l$ means an attack identification, according the anomaly on similarity of $s_n$ compared to a defined limiar $l$.
+
+Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}^{(q)}$ for eigen similarity analysis, the $s_n$ is calculated according to (\ref{eq:2.11}).
+If $s_n = 1$, then the two eigenvectors are completely similar and no anomaly is detected.
+Smaller values of $s_n$ mean less similarity and can indicate an anomaly if $s_n < l$, what denotes that a network attack is identified during the $n$-th minute.
+
+The Equation \ref{eq:2.14} shows how the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack is identified, where
+
+\begin{equation}\label{eq:2.14}
   \boldsymbol{\hat{n}_{(n)}}=\left\{
@@
     0, & \text{otherwise}
-  \end{array}\right.
-\end{equation}
-where $\boldsymbol{\hat{n}}_{(n)}$ denotes a vector of $n$-th minutes detected as under attack.
-
-The eigen similarity analysis can also be applied in an individual fashion, where each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, as shown by Figure \ref{fig:2_fig9}.
-
+  \end{array}\right.,
+\end{equation}
+and $\boldsymbol{\hat{n}}_{(n)}$ denotes a vector of $n$-th minutes detected as under attack.
+
+The eigen similarity analysis can also be applied by means of the individual approach, where each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, as shown by Figure \ref{fig:2.09}.
+In the individual approach there is no incremental appending, therefore only individual $n$-th minute are appended to $\boldsymbol{X_n}$, according to \ref{eq:2.12}, in order to compute $\boldsymbol{v}^{(q)}$, $\boldsymbol{v}_{(n)}^{(q)}$ and the Equation \ref{eq:2.11} for individual $n$-th minute.
+
+%TODO - fix figure that shows s_n > l where should the inverse
 \begin{figure}[h!]
@@
      \caption{Traffic selection for individual approach.}
-     \label{fig:2_fig9}
-\end{figure}
-
-The incremental and the individual approaches can be combined to obtain the incremental individualized approach, where each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
-Subsequently, $\boldsymbol{X}_n$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:2_fig2}.
-
+     \label{fig:2.09}
+\end{figure}
+
+The incremental and the individual approaches can be combined to obtain the incremental individualized approach, where each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}^{(q)}$ to compute similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack, i.e.
+$s_n < l$.
+Subsequently, $\boldsymbol{X}_{n-1}$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:2.02}.
+
+%TODO - fix figure that shows s_n > l where should the inverse
 \begin{figure}[h!]
@@
      \caption{Traffic selection for incremental individualized approach.}
-     \label{fig:2_fig2}
-\end{figure}
-
-This approach of incremental similarity analysis followed by individual analysis after an attack detection allows to identify the attack period, highlighting the first and last time under attack.
+     \label{fig:2.02}
+\end{figure}
+
+The incremental similarity analysis followed by individual analysis after an attack detection allows to identify the attack period, highlighting the first and last time under attack.
 This identification is possible due to the variation of the most significant eigenvectors, which becomes more significant when compared a traffic under attack against a traffic with no attack, according to results which are discussed in Section \ref{sec:2_experimentalresults}.
@@
 
-For detection of ports under attack, the $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack shall be used as reference for similarity analysis against the $\boldsymbol{v}_{(n)}$ identified as under attack, individually evaluating the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.
-
+For detection of ports under attack, the last most significant eigenvectors without attack $\boldsymbol{v}^{(q)}$ shall be used as reference for similarity analysis against the $\boldsymbol{v}_{(n)}^{(q)}$ identified as under attack, and evaluate individually the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.
 Therefore, $\boldsymbol{v}^{(q)}$ should be calculated from the last $\boldsymbol{X}^{(q)}$ time frame without attack, and $\boldsymbol{v}_{(m,\hat{n})}$ should be calculated from the same traffic appended of all $n$-th minutes until the identified minute under attack, denoted as $\boldsymbol{X}_n$.
@@
 
-\begin{equation}\label{eq:eq15}
+\begin{equation}\label{eq:2.15}
   \left\{
@@
 \end{equation}
-
-where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, which is used to calculate the $\boldsymbol{v}_{(m,\hat{n})}$ most significant eigenvectors that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.
+where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute and $q$-th time frame identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, which is used to calculate the $\boldsymbol{v}_{(m,\hat{n})}$ most significant eigenvectors that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.
 
@@
 
-This section presents the performed experiments and the acquired results.
-First, in Section \ref{sec:2_AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
+This section presents the performed experiments and the acquired results for the Eigensim.
+First, in Section \ref{sec:2_AnalyzedScenario}, the experimental scenario adopted in the evaluation is summarized.
 Then, Section \ref{sec:2_largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario.
-In Section \ref{sec:2_MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated dataset.
+In Section \ref{sec:2_MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated data set.
 Section \ref{sec:2_EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:2_EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario.
-Section \ref{sec:2_DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 dataset.
+Section \ref{sec:2_DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 data set.
 
@@
 
-The experiment time is 120 minutes, separated into six time frames, with each time frame corresponding to twenty minutes.
+This experiment consider a simulated scenario of a real network monitored during 120 minutes, that are separated into six time frames of twenty minutes.
 Therefore, as the time of each sampling period is one minute, then $N = 20$.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, as well as a covariance $\boldsymbol{\hat{R}}_{yy}^{(q)} \in \mathbb{R}^{17 \times 17}$ (calculated via (\ref{eq:eq03})) and a sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)} \in \mathbb{R}^{17x17}$, assuming that $q = 1, 2, 3, 4, 5$ and $6$.
+For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, as well as a covariance $\boldsymbol{\hat{R}}_{yy}^{(q)} \in \mathbb{R}^{17 \times 17}$ (calculated via (\ref{eq:2.03})) and a sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)} \in \mathbb{R}^{17x17}$, assuming that $q = 1, 2, 3, 4, 5$ and $6$.
 
@@
      \caption{Eigenvalues of the sample covariance matrix (synflood).}
-     \label{fig:2_fig10}
-\end{figure}
-
-Figure \ref{fig:2_fig10} depicts the eigenvalues calculated from sample covariance matrix of the network traffic used to evaluate the synflood attack identification.
-In Figure \ref{fig:2_fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 4$) stands out significantly from the other eigenvalues.
-
-Figure \ref{fig:2_fig11} illustrates the eigenvalues calculated from sample covariance matrix of the matrix used for fraggle attack detection.
-In Figure \ref{fig:2_fig10}, the largest eigenvalue related to the simulated synflood attack ($q = 5$) stands out significantly from the other eigenvalues, in accordance with the result shown in Figure \ref{fig:2_fig10} for the synflood attack analysis.
+     \label{fig:2.10}
+\end{figure}
+
+Figure \ref{fig:2.10} depicts the eigenvalues calculated from sample covariance matrix of the network traffic used to evaluate the synflood attack identification.
+In Figure \ref{fig:2.10}, the largest eigenvalue related to the simulated synflood attack ($q = 4$) stands out significantly from the other eigenvalues.
+
+Figure \ref{fig:2.11} illustrates the eigenvalues calculated from sample covariance matrix of the matrix used for fraggle attack detection.
+In Figure \ref{fig:2.10}, the largest eigenvalue related to the simulated synflood attack ($q = 5$) stands out significantly from the other eigenvalues, in accordance with the result shown in Figure \ref{fig:2.10} for the synflood attack analysis.
 
@@
      \caption{Eigenvalues of the sample covariance matrix (fraggle).}
-     \label{fig:2_fig11}
-\end{figure}
-
-Figure \ref{fig:2_fig12} depicts the eigenvalues calculated from covariance matrix of zero mean and unitary standard deviation variables, of the network traffic matrix evaluated for port scan detection.
+     \label{fig:2.11}
+\end{figure}
+
+Figure \ref{fig:2.12} depicts the eigenvalues calculated from covariance matrix of zero mean and unitary standard deviation variables, of the network traffic matrix evaluated for port scan detection.
+As analyzed for the synflood and fraggle attacks, note that the largest eigenvalue, related to this attack ($q = 3$), stands out significantly from the others eigenvalues.
 
@@
      \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}
-     \label{fig:2_fig12}
-\end{figure}
-
-As analyzed for the synflood and fraggle attacks, note that the largest eigenvalue, related to this attack ($q = 3$), stands out significantly from the others eigenvalues.
-
-Table \ref{tab:tab3} presents the values of the largest eigenvalues of each time frame $q$-th for port scan, synflood and fraggle detection.
+     \label{fig:2.12}
+\end{figure}
+
+Table \ref{tab:2.03} presents the values of the largest eigenvalues of each time frame $q$-th for port scan, synflood and fraggle detection.
 
@@
   \caption{Largest Eigenvalue related to attacks detection}
-  \label{tab:tab3}
+  \label{tab:2.03}
   \begin{tabular}{ c c c c c }
@@
 
-In Table \ref{tab:tab3}, note the significant variation of the eigenvalues associated with attacks, in comparison to the others.
+In Table \ref{tab:2.03}, note the significant variation of the eigenvalues associated with attacks, in comparison to the others.
 At $q = 4$, where the synflood attack occurred, the maximum eigenvalue obtained is approximately 21 times larger than the second one.
@@
 In \cite{tenorio2013greatest}, we evaluate the accuracy of AIC, MDL, EDC, RADOI, EFT and SURE MOS schemes \cite{da2009comparison,tenorio2013greatest} for synflood and port scan attack detection.
-In this work we extend that evaluation for fraggle attack detection, applying the same schemes to fraggle attack detection over the traffic presented in Section \ref{sec:2_datamodel}, as results shown in Table \ref{tab:tab4}.
+In this work we extend that evaluation for fraggle attack detection, applying the same schemes to fraggle attack detection over the traffic presented in Section \ref{sec:2_datamodel}, as results shown in Table \ref{tab:2.04}.
 
@@
   \caption{MOS schemes applied to port scan and flood detection}
-  \label{tab:tab4}
+  \label{tab:2.04}
   \begin{tabular}{ c c c c c c c c }
@@
 Note that $\hat{d} = 1$, if there is attack, while $\hat{d} > 1$ indicates more than one attack.
-An example of this could be seen for attack detection via EFT for traffic containing synflood and fraggle attacks, showing $\hat{d} = 2$, which indicates the presence of two attacks, as expected by the $d$ real values of Table \ref{tab:tab4}.
-
-In Table \ref{tab:tab4}, two MOS schemes outperforms from the others, EDC and EFT.
-Efficient Detection Criterion (EDC) and Exponential Fitting Test (EFT) are the most effective schemes, correctly estimating the number of attacks in comparison to the expected values for effective attack detection, as defined by the column of real values in Table \ref{tab:tab4}.
+An example of this could be seen for attack detection via EFT for traffic containing synflood and fraggle attacks, showing $\hat{d} = 2$, which indicates the presence of two attacks, as expected by the ground truth values $d$ of Table \ref{tab:2.04}.
+
+In Table \ref{tab:2.04}, two MOS schemes outperforms from the others, which are EDC and EFT.
+Efficient Detection Criterion (EDC) and Exponential Fitting Test (EFT) are the most effective schemes, correctly estimating the number of attacks in comparison to the expected values for effective attack detection, as defined by the column of real values in Table \ref{tab:2.04}.
 The AIC and MDL schemes are satisfactory only for port scan detection, however SURE and RADOI schemes did not show effective results for port scan or flood detection.
@@
 
-According to Table \ref{tab:tab4}, EDC and EFT estimated correctly the number of attacks of a time frame vector, indicating that occurred $\hat{d}$ network attacks, but not providing additional details, what highlights the necessity of complementary approaches in order to estimate the time and ports under attack.
+According to Table \ref{tab:2.04}, EDC and EFT estimated correctly the number of attacks of a time frame vector, indicating that occurred $\hat{d}$ network attacks, but not providing additional details, what highlights the necessity of complementary approaches in order to estimate the time and ports under attack.
 Hence, we propose apply eigen analysis to estimate the $q$-th time frames under attack and eigen similarity analysis to estimate the minutes and ports under attack.
@@
 According to the results presented in Section \ref{sec:2_largesteigenvaluesanalysis}, the largest eigenvalue stands out significantly from the others eigenvalues of an evaluated $q$-th time frame.
-This behavior can also be observed in the largest eigenvalues analysis, according to results presented in Table \ref{tab:tab3}, where it is possible to observe that the $\hat{d}$ largest eigen values of the time frames under attacks stand out significantly from the others largest eigenvalues.
-
-Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respective $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:2_alg01}.
+This behavior can also be observed in the largest eigenvalues analysis, according to results presented in Table \ref{tab:2.03}, where it is possible to observe that the $\hat{d}$ largest eigen values of the time frames under attacks stand out significantly from the others largest eigenvalues.
+
+Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respective $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:2.01}.
 
@@
 Three approaches were evaluated for eigen similarity analysis: incremental, individual and incremental individualized approaches.
-For the incremental individualized approach, each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
+For the incremental individualized approach, each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}^{(q)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
 Subsequently, $\boldsymbol{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated.
-For the incremental approach, each $n$-th minute must be incrementally appended into $\boldsymbol{X}^{(q)}$, for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual time similarity analyis.
-For the individual approach, each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, without incremental append, but doing individual appended into $\boldsymbol{X}^{(q)}$ for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual similarity analyis.
-
-Table \ref{tab:tab5} presents the results of the evaluation of three approaches for similarity analysis of eigenvectors for port scan detection.
+For the incremental approach, each $n$-th minute must be incrementally appended into $\boldsymbol{X}^{(q)}$, for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}^{(q)}$ for individual time similarity analysis.
+For the individual approach, each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, without incremental append, but doing individual appended into $\boldsymbol{X}^{(q)}$ for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}^{(q)}$ for individual similarity analyis.
+
+Table \ref{tab:2.05} presents the results of the evaluation of three approaches for similarity analysis of eigenvectors for port scan detection.
 
@@
   \caption{Eigen Similarity Analysis for Port Scan Detection}
-  \label{tab:tab5}
+  \label{tab:2.05}
   \begin{tabular}{ c c c c c c }
 	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
+	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Ground Truth}}\\ 
 			\hhline{~~---~}
@@
 
-Table \ref{tab:tab5} shows the evaluation of the time frame $q=3$, when the port scan attack was simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
-According to the presented results, it is possible to observe the high similarity between network traffic without attack, which was larger than 0.9610 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with the simulated port scan attack ($n=15$), which was lower than 0.0276 for all evaluated approaches.
+Table \ref{tab:2.05} shows the evaluation of the time frame $q=3$, when the port scan attack was simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
+According to the presented results, it is possible to observe the high similarity between network traffic without attack, which was larger than 0.9610 for all evaluated cases, and emphasize the expressive low similarity when it was evaluated the traffic with the simulated port scan attack ($n=15$), which was lower than 0.0276 for all evaluated approaches.
 
@@
 
-Table \ref{tab:tab6} presents the results of the evaluation of the similarity analysis of eigenvectors for synflood detection.
+Table \ref{tab:2.06} presents the results of the evaluation of the similarity analysis of eigenvectors for synflood detection.
 It shows the evaluation of the time frame $q=4$, when the synflood attack is simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
@@
   \caption{Eigen Similarity Analysis for Synflood Detection}
-  \label{tab:tab6}
+  \label{tab:2.06}
   \begin{tabular}{ c c c c c c }
 	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
+	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Ground Truth}}\\ 
 			\hhline{~~---~}
@@
 
-The incremental approach produces better results if compared with other evaluated approaches, with lower values and maximum of 0.0185 for times under attack, but this approach presents change detection limitation after the first outlier of similarity, as shown in Table \ref{tab:tab5} for port scan detection.
+The incremental approach produces better results if compared with other evaluated approaches, with lower values and maximum of 0.0185 for times under attack, but this approach presents change detection limitation after the first outlier of similarity, in accordance to the results shown in Table \ref{tab:2.05} for port scan detection.
 
@@
 
-Table \ref{tab:tab7} presents the results of the eigen similarity analysis evaluation for fraggle detection.
+Table \ref{tab:2.07} presents the results of the eigen similarity analysis evaluation for fraggle detection.
 
@@
   \caption{Eigen Similarity Analysis for Fraggle Detection}
-  \label{tab:tab7}
+  \label{tab:2.07}
   \begin{tabular}{ c c c c c c }
 	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
+	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Ground Truth}}\\ 
 			\hhline{~~---~}
@@
 
-These results highlight the capability of change detection based on similarity between legitimate and malicious traffic from flood or port scan attacks, endorsing the effectiveness and safety for adoption of threshold for attack detection through eigen similarity analysis.
+These results highlight the capability of change detection based on similarity between legitimate and malicious traffic from flood or port scan attacks, by means of a subspace learned via Eigenvalue Value Decomposition and Model Order Selection,
+endorsing the effectiveness and safety for adoption of threshold for attack detection through eigen similarity analysis.
+Moreover, these results answer positively the question $Q_1$, what rejects the null hypothesis  $H_{0sub}$ and confirms the alternative
+hypothesis $H_{1sub}$.
 
@@
 Therefore, the incremental individualized and individual approaches of eigen similarity analysis were evaluated, for detection of ports under flood and port scan attacks, according to results presented in following tables.
-For this evaluation, the $v$ last most significant eigenvectors without attack was used as reference for similarity analysis against each target port $m$-th.
-
-Table \ref{tab:tab8} presents the results of the evaluation of eigen similarity analysis for detection of ports under port scan attack, showing only the time frame $q=3$ and minute $n=15$, due to the simulated port scan attack occurred only at this time, although the remain time frame has been completely evaluated and presented high similarity to the reference of traffic without network attack.
+For this evaluation, the last most significant eigenvectors without attack $v$ was used as reference for similarity analysis against each target port $m$-th.
+
+Table \ref{tab:2.08} presents the results of the evaluation of eigen similarity analysis for detection of ports under port scan attack, showing only the time frame $q=3$ and minute $n=15$, due to the simulated port scan attack occurred only at this time, although the remain time frame has been completely evaluated and presented high similarity to the reference of traffic without network attack.
 
@@
   \caption{Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n=15)}
-  \label{tab:tab8}
+  \label{tab:2.08}
   \begin{tabular}{ c c c c }
 	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
+	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Ground Truth}}\\ 
 			\hhline{~--~}
@@
 
-For the evaluation of the proposed approaches for identification of ports under synflood and fraggle attack, all minutes of each time frame, in which one attack location was estimated, were analyzed.
-Even though, due to space limitations, only the results of the first minute where a low similarity was identified will be shown such as where $n=11$.
-Nevertheless, the results obtained for the evaluation of traffic without attack presented high similarity to the reference traffic, with similarities close to 0.9999, and the evaluation of the other minutes under attack presented results quite similar to the results shown in the Tables \ref{tab:tab9} and \ref{tab:tab10}.
-
-Table \ref{tab:tab9} presents the results of the evaluation of eigen similarity analysis for detection of ports under synflood attack, showing only the time frame $q=4$ and minute $n=11$.
+For the evaluation of the proposed approaches for identification of ports under synflood and fraggle attack, all minutes of time frames under were analyzed.
+However, due to space limitations, only the results of the first minute, where a low similarity was identified, will be shown.
+Nevertheless, the results obtained for the evaluation of traffic without attack presented high similarity to the reference traffic, with similarities close to 0.9999, and the evaluation of the other minutes under attack presented results quite similar to the results shown in the Tables \ref{tab:2.09} and \ref{tab:2.10}.
+
+Table \ref{tab:2.09} presents the results of the evaluation of eigen similarity analysis for detection of ports under synflood attack, showing only the time frame $q=4$ and minute $n=11$.
 
@@
   \caption{Eigen Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
-  \label{tab:tab9}
+  \label{tab:2.09}
   \begin{tabular}{ c c c c }
 	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
+	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Ground Truth}}\\ 
 			\hhline{~--~}
@@
 
-According to results presented in Table \ref{tab:tab9}, both approaches identifies low similarity for the traffic of port 600, which is the target port of the simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.
-
-Table \ref{tab:tab10} presents the results of the evaluation of eigen similarity analysis for detection of ports under fraggle attack, showing only the time frame $q=5$ and minute $n=11$.
+According to results presented in Table \ref{tab:2.09}, both approaches identifies low similarity for the traffic of port 600, which is the target port of the simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.
+
+Table \ref{tab:2.10} presents the results of the evaluation of eigen similarity analysis for detection of ports under fraggle attack, showing only the time frame $q=5$ and minute $n=11$.
 
@@
   \caption{Eigen Similarity Analysis for Detection of Ports Under Fraggle Attack (q=5 and t=11)}
-  \label{tab:tab10}
+  \label{tab:2.10}
   \begin{tabular}{ c c c c }
 	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
+	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Ground Truth}}\\ 
 			\hhline{~--~}
@@
 
-The results for the evaluation of ports under fraggle attack, shown in Table \ref{tab:tab10}, were similar to the results obtained for synflood analysis, with the identification of low similarity for traffic of the port under attack.
+The results for the evaluation of ports under fraggle attack, shown in Table \ref{tab:2.10}, were similar to the results obtained for synflood analysis, with the identification of low similarity for traffic of the port under attack.
 Nevertheless, for fraggle analysis, the individual approach identified the lowest similarity, that is 0.0004 while the incremental individualized approach obtained a similarity of 0.0031.
@@
 
-This subsection presents a summarized view of results obtained from the application of the proposed framework, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 dataset.
-Since the proposed framework is detailed in Section \ref{sec:2_prop_getv} and in Subsections \ref{sec:2_AnalyzedScenario}, \ref{sec:2_largesteigenvaluesanalysis}, \ref{sec:2_MOSSchemesEvaluation}, \ref{sec:2_EigenvalueAnalysis} and \ref{sec:2_EigenSimilarityAnalysis}, here the focus is on the parameter selection, dataset evaluation and results for flood and probe attack identification.
-
-The DARPA dataset includes 7 weeks of sniffed traffic saved into raw packet data.
+This subsection presents a summarized view of results obtained from the application of the Eigensim, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 data set.
+Since the proposed framework is detailed in Section \ref{sec:2_prop_getv} and in Subsections \ref{sec:2_AnalyzedScenario}, \ref{sec:2_largesteigenvaluesanalysis}, \ref{sec:2_MOSSchemesEvaluation}, \ref{sec:2_EigenvalueAnalysis} and \ref{sec:2_EigenSimilarityAnalysis}, here the focus is on the parameter selection, data set evaluation and results for flood and probe attack identification.
+
+The DARPA data set includes 7 weeks of sniffed traffic saved into raw network packet data, i.e.
+pcap files.
 The traffic and the labeled attacks are grouped by week and day, with information of the number and types of attacks per day, but also providing the start time for each labeled attack.
-For this evaluation, an evaluation per day was performed, considering the network traffic of 24 hours split into $Q$ time frames of 60 minutes ($N = 60$) and aggregate by minute and by port number.
+For this evaluation, an evaluation per day was performed, considering the network traffic of 24 hours split into $Q$ time frames of 60 minute
+s ($N = 60$) and aggregate by minute and by port number.
 For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ is obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.
@@
 	\caption{Results of the attack detection evaluation}
-	\label{tab:tab12}
+	\label{tab:2.12}
 	\centering
@@
 		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Solution\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Attack Type\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Metric\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Result\end{tabular} \\ \hline
-	Proposed Work	&Flooding	&True Positive	&100.00 \%\\ \hline
-	Proposed Work	&Flooding	&False Positive	&60.00 \%\\ \hline
-	Proposed Work	&Flooding	&Misclassification	&50.00 \%\\ \hline
-	Proposed Work	&Probe	&True Positive	&76.92 \%\\ \hline
-	Proposed Work	&Probe	&False Positive	&18.52 \%\\ \hline
-	Proposed Work	&Probe	&Misclassification	&32.73 \%\\ \hline
+	Eigensim	&Flooding	&True Positive	&100.00 \%\\ \hline
+	Eigensim	&Flooding	&False Positive	&60.00 \%\\ \hline
+	Eigensim	&Flooding	&Misclassification	&50.00 \%\\ \hline
+	Eigensim	&Probe	&True Positive	&76.92 \%\\ \hline
+	Eigensim	&Probe	&False Positive	&18.52 \%\\ \hline
+	Eigensim	&Probe	&Misclassification	&32.73 \%\\ \hline
 	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&True Positive	&82.00 \%\\ \hline
@@
 
-Table \ref{tab:tab12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}.
+Table \ref{tab:2.12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}.
 The analysis based on sample covariance of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames.
 The results also show 50.00 \% of misclassification rate, which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
-The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time from one source to one target, with no variation on IP source or target port.
-This observation corroborates with previous evaluations of the DARPA dataset that highlight issues regarding traffic redundancy.
+The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA data set presents high number of packets per time from one source to one target, with no variation on IP source or target port.
+This observation corroborates with previous evaluations of the DARPA data set that highlight issues regarding traffic redundancy.
 
@@
 This work is a statistical method, based on PCA, without training or learning methods, even though it relies on visual analysis for principal components selection.
-The best detection rate of \cite{Zonglin2009} was 82.00 \% for detection of synthetically added flood attacks, while this current proposal obtains 100.00 \% of detection rate for detection of flood attacks of DARPA dataset.
-
-Due to the lack of statistical techniques without training or learning methods for detection of probe attacks, this proposed approach is compared to the Lu and Ghorbani's \cite{Lu2009} proposal, which is a network anomaly detection model based on signal processing techniques that uses DARPA dataset for evaluation.
+The best detection rate of \cite{Zonglin2009} was 82.00 \% for detection of synthetically added flood attacks, while this current proposal obtains 100.00 \% of detection rate for detection of flood attacks of DARPA data set.
+It is important to note that Callegari \emph{et al} \cite{Zonglin2009} did not publish results of false positive and misclassification.
+
+Due to the lack of statistical techniques without training or learning methods for detection of probe attacks, this proposed approach is compared to the Lu and Ghorbani's \cite{Lu2009} proposal, which is a network anomaly detection model based on signal processing techniques that uses DARPA data set for evaluation.
 The results of \cite{Lu2009} show the best detection rate of 94.67 \% in terms of general attack instance detection, but shows a case case with 50.00 \% of attack instance detection for the portsweep attack.
@@
 
-This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:2_fig80} and equations presented in Section \ref{sec:2_prop_getv}.
+This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:2.08} and equations presented in Section \ref{sec:2_prop_getv}.
 
@@
 \label{sec:2_ComplexityAnalysis}
-The EVD, calculated according to (\ref{eq:eq06}), requires the previous calculation of covariance matrix, according to Equations (\ref{eq:eq02}), (\ref{eq:eq03}), (\ref{eq:eq04}) and (\ref{eq:eq05}).
+The EVD, calculated according to (\ref{eq:2.06}), requires the previous calculation of covariance matrix, according to Equations (\ref{eq:2.02}), (\ref{eq:2.03}), (\ref{eq:2.04}) and (\ref{eq:2.05}).
 The covariance matrix calculation is $O(M^2N)$ and the EVD is $O(N^3)$, where $M$ denotes the number of network ports and $N$ denotes the period time.
@@
 
-The largest eigenvalue analysis is $O(\hat{d}Q)$, where $\hat{d}$ denotes the number of time frame under attack, according to Algorithm \ref{alg:2_alg01}.
+The largest eigenvalue analysis is $O(\hat{d}Q)$, where $\hat{d}$ denotes the number of time frame under attack, according to Algorithm \ref{alg:2.01}.
 Subsequently, the eigen similarity analysis relies on EVD and cosine similarity analysis, which is $O(N^2)$, for $\hat{d}$ time frames, therefore the eigen similarity analysis is which is $O(\hat{d}(M^2N + N^3 + N^2))$ and yields an $O(N^3)$ upper bound on the worst-case running time for eigen similarity analysis.
@@
 \label{sec:2_ProcessingAnalysis}
-For better understanding the scalability and impact of configurations of $N$, $M$ and $Q$, the processing time required for different scenarios of parameter configurations and dataset were evaluated, measuring the processing time of:
+For better understanding the scalability and impact of configurations of $N$, $M$ and $Q$, the processing time required for different scenarios of parameter configurations and data set were evaluated, measuring the processing time of:
 
@@
 The data modeling is also a time consuming step, however its processing can be optimized through distributed processing techniques, such as MapReduce, achieving high throughput for packet counting or even for deep packet inspection \cite{Vieira2013}.
+It is also possible to evaluate the adoption of faster SVD algorithms, considering implementations based on truncated or randomized approaches \cite{halko2011finding} that aim reduce the complexity and processing time.
 
@@
 Table \ref{tab:11} presents the measured results.
-The experiment considered traffic time of 16, 20 and 22 hours, according to the selected traffic time per day available by the DARPA dataset.
+The experiment considered traffic time of 16, 20 and 22 hours, according to the selected traffic time per day available by the DARPA data set.
 Note that the processing time increases according to the increment in traffic time, around 2 or 3 times for 1-time and 2-time, but the worst measured processing time is 4.7250 milliseconds.
@@
 
-This work models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
-
-The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
+This work models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, named Eigensim, which is based on subspace decomposition, eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
+
+The Eigensim is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis from a subspace obtained by EVD.
+Therefore, the observed results rejects the null hypothesis  $H_{0sub}$ and confirms the alternative
+hypothesis $H_{1sub}$.
+
 The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
@@
 
-Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets.
-Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high throughput network traffic.
+Future research can be directed to improvements for better false positive rates, as well as to make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets.
+Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high throughput network traffic, as well as it is possible to evaluate approaches with less complexity for EVD \cite{halko2011finding}.
 Additionally, future research can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.

### 2019-05-27 docs/thesis/chapters/4_m_rpca/m_rpca.tex
first deliverable

@@
-\chapter{Moment Distances from Robust Subspace for Botnet Detection}
+\chapter{Moment Distances from Robust Subspace for Network Anomaly Detection}
 \label{ch:4_m_rpca}
@@
 
-Anomaly detection can be defined as the identification of rare and suspicions events by differing the normal or majority of the data.
-Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects \cite{bhuyan2014network,ahmed2016survey}.
-Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.
+Some widely adopted algorithms for anomaly detection assume a gaussian distributed data, however this assumption may not be observed in some real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network}.
 
-Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or other patterns.
-However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid unknown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detection systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
-To face this problem, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, where it is not necessary known anomalies for training models \cite{moustafa2019holistic}.
+Findings of Benson \emph{et al.}  \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
+Leon-Garcia \cite{leon2017probability} also argues that Pareto distribution has been found to capture the behavior of many quantities of interest in the study of Internet behavior.
+Moreover, Benson \emph{et al.}  \cite{benson2010network} observes that the lognormal distribution is the best fit to model arrival processes in a data center.
 
-Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}.
-Learning algorithms for imbalanced data has been a challenging research topic, considering that the fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
-Some techniques can make the data discriminative for better results on the imbalanced learning problem: Data engineering (i.e.
-sampling \cite{he2008learning,gu2008botminer}, aggregation \cite{lakhina2005mining,gu2008botminer,callegari2011novel, garcia2014empirical, acarali2016survey}); Feature extraction (i.e.
-decomposition \cite{vaswani2018robust}, robust estimates \cite{zhou2017anomaly}).
+This findings show that the skeweness and heavy-tailed distributions may be important for network traffic analysis, and can motivate evaluations to study the impact of skewed data into algorithms that rely on gaussian distributed data, as well as it can highlight opportunities to  exploit the skeweness and heavy-tailed distributions in order to obtain accurate classifiers for network anomaly detection.
 
-% He and Garcia \cite{he2008learning} present a classification of imbalanced learning methods as sampling, cost-sensitive learning, kernel-based, and active learning.
-For more information about learning methods from imbalanced data, we refer to He and Garcia \cite{he2008learning}.
+Network anomaly detection problems are usually characterized by imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}.
+However, learning algorithms for imbalanced data has been a challenging research topic, considering that the fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
+Therefore, learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
 
-% Say that the data is lognormal \cite{benson2010network} and link it to skewness https://www.weibull.com/hotwire/issue47/relbasics47.htm
-Some widely adopted algorithms assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network}.
-Findings of Benson \emph{et al.}  \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
-This findings show that the skeweness and heavy-tailed distributions can impact algorithms that rely on gaussian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.
-Therefore, learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
-Skewness and kurtosis are higher moments that can have their measurement as very sensitive for outlier detection, considering that the properties of the kurtosis has been used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection \cite{hubert2009robustskewed}.
+We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data, 
+and that the distance between robust estimates of normal traffic and anomalies can highlight discrepancies and be used for network attack detection.
+Therefore, we propose the md-RPCA, which is an approach based on distances from moments computed from robust subspace learned by Robust Principal Component Analysis, in order to detect anomalies from network traffic.
+The proposed approach relies on a robust subspace of normal traffic for estimating higher moments.
+The anomaly detection from new observations evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervised or unsupervised fashion, without the computational cost of new robust subspace learning for new observations.
 
-Principal Component Analysis (PCA) is very sensitive to anomalies, thus outliers can corrupt the entire data or can be explored to highlight unexpected changes and indicate attacks or frauds \cite{callegari2011novel,Lee2013,vieira2017model}.
-However, PCA is not precise to reveal flows, time or detailed information for each anomaly, and sometimes relies into visual analysis for principal component selection.
-Robust Principal Component Analysis (RPCA) \cite{candes2011robust} is an extension of PCA that aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, which refers to outliers’ matrix.
-Therefore, robust subspace learning has receiving a growing attention of researchers aiming the development of network anomaly detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd, hubert2005robpca,hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
+We evaluate the accuracy of the md-RPCA for anomaly detection on synthetic data set and on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
+The Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches.
 
-Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the covariances between the variables and their mean values.
-The MD is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
-C.
-Mahalanobis in 1936 \cite{mahalanobis1936md}.
-It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\bar{\Sigma}}$ of $\boldsymbol{X}$.
-MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and new observations can reveal anomalies.
+% We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
 
-We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data.
-The botnet detection algorithms can also be improved by feature engineering from aggregated traffic, considering that new features can make the data more discriminative for imbalanced learning problems.
-We also believe that the distance between robust estimates of normal traffic and anomalies can highlight discrepancies and be used for botnet detection.
-Therefore, we propose an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggregation.
-The proposed approach relies on a robust subspace of normal traffic for estimating higher moments.
-The anomaly detection for new observations evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations.
+The main contribution of this work is a novel semi-supervised and unsupervised method for anomaly detection in skewed and imbalanced data, with results of experimental evaluation on synthetic and real data sets.
 
-We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adopted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
-We evaluate the accuracy of the proposed approach for botnet detection on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
-
-Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches, which are K-means and Gaussian Mixture Model (GMM) \cite{gaddam2007kmeans,moustafa2019holistic}, respectively.
-We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
-
-The main contribution of this work is a novel semi-supervised method for botnet detection that obtain results with high rates of attack detection in an imbalanced and skewed data set.
-
-This paper is organized as follows.
-In Section \ref{sec:review}, it is conducted a literature review of network anomaly detection, botnet detection, sampling and aggregation for imbalanced learning and robust based anomaly detection methods.
-Section \ref{sec:datamodel} presents the data model and the evaluated data set.
-Section \ref{sec:mdrs} describes the proposed approach for botnet detection.
-Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work.
+This Chapter is organized as follows.
+In Section \ref{sec:4_relatedworks}, it is conducted a literature review of network anomaly detection, botnet detection, and imbalanced learning.
+Section \ref{sec:4_datamodel} presents the data model and the evaluated data set.
+Section \ref{sec:4_proposal} describes the proposed approach for network attack detection.
+Section \ref{sec:4_experiments} discusses the experimental validation and Section \ref{sec:4_results} presents the results.
+Finally, Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.
 
@@
 
+% TODO - re-write this, it is strange
 % network anomaly detectiom
@@
 
-The objective of Garcia \emph{et al.} \cite{garcia2014empirical} is to compare three botnet detection methods using a simple and reproducible methodology, and a good data set and a new error metric.
+The objective of Garcia \emph{et al.} \cite{garcia2014empirical} is to compare three botnet detection methods using a simple and reproducible methodology, a good data set and a new error metric.
 This paper evaluates some data sets for network anomaly detection, surveys some approaches for botnet detection, proposes two methods (BClus and CAMNEP) for botnet detection and compare results to BotHunter \cite{gu2007bothunter}.
+
 Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by attack, normal and background labeled data, in an imbalanced distribution like in a real network.
-The authors recommend that none of the botnet families used in the training and cross-validation data set should be used in the testing data set, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
+The authors recommend scenarios for training and testing, in the way that none of the botnet families used in the training and cross-validation data set should be used in the testing data set, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
 Adopting the training and testing approach proposed by Garcia \emph{et al.}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
@@
 % Anomaly detection by robust subspace learning
-The problem of PCA or subspace learning for outlier corrupted data is called robust PCA (RPCA) or robust subspace learning \cite{vaswani2018robust}, which can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
+The problem of PCA or subspace learning for outlier corrupted data is called Robust Principal Component Analysis (RPCA) or robust subspace learning \cite{candes2011robust, vaswani2018robust}.
+RPCA aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, which refers to outliers’ matrix.
 This definition is also referred to as the sparse + low rank (S + LR) formulation \cite{vaswani2018robust}.
@@
 
-Robust subspace learning has receiving a growing attention of researchers aiming the development of anomaly detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
 RPCA is been mainly applied to computer vision, in problems of robust subspace tracking and robust subspace recovery.
@@
 
+Robust subspace learning has receiving a growing attention of researchers aiming the development of network anomaly detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd, hubert2005robpca,hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
 Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on a robust mutual information estimator for feature selection and based on RPCA for outlier detection in internet traffic.
@@
 
-% feature engineering and aggregation
-Lakhina \emph{et al.} \cite{lakhina2005mining} argue that the distributions of packet features (IP addresses and ports) observed in flow traces reveals both the presence and the structure of a wide range of anomalies.
-Lakhina \emph{et al.} show that the analysis of feature distributions leads to significant advances on sensitive detection of a wide range of anomalies, augmenting detection by volume-based methods, while the anomaly detection approach leads gains on automatic classification of anomalies via unsupervised learning.
-Goldberg and Shan \cite{goldberg2015importance} argue, based on experience and on published real-life examples of anomaly detectors in use at eBay, that processing monitored signals into features is a more fruitful for anomaly detection systems than devising complex statistical tests.
-
 Benson \emph{et al.} \cite{benson2010network} conducted an empirical study of the network traffic in 10 data centers belonging to three different types of organizations, including university, enterprise, and cloud data centers.
 Findings of Benson \emph{et al.} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
-This findings highlight that the skeweness and heavy-tailed distributions can impact algorithms that rely on gaussian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network traffic analysis.
 
-Considering that learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}, higher moments can be used for outlier detection and discriminative analysis of skewed data.
-The skewness and kurtosis are higher moments that can contribute to sensitive outlier detectors, taking into account that the properties of the kurtosis can be used to identify clusters or outliers \cite{pena2010eigenvectors}, while skewness has been used as adjust factor for outlier detection in skewed data sets \cite{hubert2009robustskewed}.
+Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the covariances between the variables and their mean values.
+The MD is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
+C.
+Mahalanobis in 1936 \cite{mahalanobis1936md}.
+It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\bar{\Sigma}}$ of $\boldsymbol{X}$.
+MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and new observations can reveal anomalies.
 
-We propose an approach based on higher moments distances from robust subspace for botnet detection, applied to new features generated by flow aggregation from the CTU-13 data set.
+We propose an approach based on higher moments distances from robust subspace for botnet detection.
 The proposed approach relies on a robust subspace of normal traffic for estimating higher moments.
-Thus, we propose that the anomaly detection for new observations should evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations.
-
-We propose a sampling method based on data aggregation for discriminative feature engineering, considering that data aggregation is a widely adopted approach for feature engineering \cite{garcia2014empirical, chandrashekar2014survey,acarali2016survey} and network anomaly detection \cite{lakhina2005mining, callegari2011novel, vieira2017model}.
-Finally, We evaluate the F-measure of the proposed approach for botnet detection on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
+Thus, we propose that the anomaly detection for new observations should evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations, or in a unsupervised approach.
 
@@
 
-This section also presents a description of the data model in Subsection \ref{sec:CTU-13} and in Subsection \ref{sec:feat_eng} we describe the data aggregation and feature engineering adopted for the CTU-13.
+This section also presents a description of the synthetic data model in Subsection \ref{sec:4_synthetic} and in Subsection \ref{sec:4_CTU-13} we describe the data model for the CTU-13 data set.
+
+\subsection{The Synthetic data set}
+\label{sec:4_synthetic}
+
+In order to evaluate the null hypothesis $H_{0rob.sub}$, which assumes that the robust subspace learning does not improves the anomaly detection from imbalanced and skewed, we create two synthetic data sets characterized by skewed and heavy tailed distributions.
+We selected Pareto and Lognormal distributions to evaluate the anomaly detection capacity of our proposal, considering that these distributions have been adopted to represent network traffic in the internet and data centers \cite{benson2010network,leon2017probability}.
+
+We adopt a contamination rate of 33\% to reproduce the imbalanced data of anomaly detection problems, and select the gaussian distribution for the addictive anomalies into Pareto and Lognormal distributions.
+Therefore, we generate $\boldsymbol{Y}_p$ and $\boldsymbol{Y}_l$ with 2,400 samples and two features, where 793 samples are gaussian anomalies, also referred as noise or contamination, while 1,607 samples are legitimate signals.
+The data set $\boldsymbol{Y}_p$ is composed by random Pareto distribution with random gaussian noise, while $\boldsymbol{Y}_l$ is composed by random Lognormal distribution with random gaussian noise.
+
+The Figure \ref{fig:4.03} shows the histogram and the PDF of the Pareto distribution for two features of $\boldsymbol{Y}_p$.
+
+\begin{figure}[!htb]
+	\centering
+	\subfigure[Feature 1]{%
+		\includegraphics[width=0.48\textwidth]{figures/pareto1.png}
+		\label{fig:4.01}
+	}
+	\subfigure[Feature 2]{%
+		\includegraphics[width=0.48\textwidth]{figures/pareto2.png}
+		\label{fig:4.02}
+	}
+	\caption[Pareto Distribution]{Pareto Distribution}
+	\label{fig:4.03}
+\end{figure}
+
+The Figure \ref{fig:4.06} shows the histogram and the PDF of the Lognormal distribution for two features of $\boldsymbol{Y}_l$.
+
+\begin{figure}[!htb]
+	\centering
+	\subfigure[Feature 1]{
+		\includegraphics[width=0.48\textwidth]{figures/lognormal3.png}
+		\label{fig:4.04}
+	}
+	\subfigure[Feature 2]{
+		\includegraphics[width=0.48\textwidth]{figures/lognormal4.png}
+		\label{fig:4.05}
+	}
+	\caption[Lognormal Distribution]{Lognormal Distribution}
+	\label{fig:4.06}
+\end{figure}
+
+The Pareto distribution contaminated by gaussian noise is denoted as $\boldsymbol{Y}_p^c$, while $\boldsymbol{Y}_l^c$ denotes the Lognormal distribution contaminated by gaussian noise.
+Taking into account that semi-supervised algorithms requires training from normal data, we also generate $\boldsymbol{Y}_p^t$, while $\boldsymbol{Y}_l^t$, which are data sets of Pareto and Lognormal distributions for training, without contamination and composed by 1,607 samples and two features.
 
 \subsection{The CTU-13 data set}
-\label{sec:CTU-13}
+\label{sec:4_CTU-13}
 
 The CTU-13 is a data set of botnet traffic that was captured in the Czech Technical University \cite{garcia2014empirical}, by means of a testbed and execution of malwares in a real network.
-The CTU-13 data set contains 13 scenarios with network traffic of botnet malwares, that are: neris, rbot, virut, menti, sogou, nsys.ay and murlo.
-The types of C\&C and attack traffic present in CTU-13 data set are:
+The CTU-13 data set contains 13 scenarios with network flows of botnet malwares, that are: neris, rbot, virut, menti, sogou, nsys.ay and murlo.
+The botnet traffic is also classified as attack or command and control (C\&C), while the legitimate flows can also be classified as normal or background.
+It is important to note that the CTU-13 data set can also be formulated as a superposition of signal, artefact and noise, which refer to background, normal a botnet traffic, respectively.
+
+The types of C\&C and attack flows present in CTU-13 data set are:
 
@@
 
-The data set originally contains the following features by flows:  Start Time, End Time, Duration, Source IP Address, Source Port, Direction, Destination IP Address, Destination Port, State, Type of service from source to destination, Type of service from destination to source, Total of Packets and Total of Bytes.
-
-Table \ref{tab:tab01} presents an overview organized by scenario and shows the malwares used for botnet attacks, the types of attacks and C\&C types, the total number of flows, the number of malicious flows, which includes flows of C\&C and attacks, and finally shows the number of normal flows.
+Table \ref{tab:4.01} presents an overview organized by scenario and shows the malwares used for botnet attacks, the types of attacks and C\&C types, the total number of flows, the number of malicious flows, which includes flows of C\&C and attacks, and finally shows the number of normal flows.
 
 \begin{table}[h!]
-	\tiny
+	\scriptsize
 	\caption{CTU-13 data set Description}
-	\label{tab:tab01}
+	\label{tab:4.01}
 	\begin{tabular}{| l | l | l | r | r | r | r | r | r | r | r | }
@@
 
-The CTU-13 aims to reproduce the percentage of C\&C and attack flows according to real cases, where the number of malicious flows are very low in comparison to normal and background flows, resulting into a very imbalanced and challenging data set for botnet detection.
+The CTU-13 data set originally contains the following features:
 
-An Exploratory Data Analysis (EDA) reveals that distributions of original features of the CTU-13 are skewed and present an overlapping between normal and anomalous flows.
-Following we present the Figure \ref{fig:fig01}, that presents the distribution of flow duration for scenario 12, and exemplifies the data skewness and overlapping.
+\begin{itemize}
+	\item Start Time;
+	\item End Time;
+	\item Duration;
+	\item Source IP Address;
+	\item Source Port;
+	\item Direction;
+	\item Destination IP Address;
+	\item Destination Port;
+	\item State;
+	\item Type of service from source to destination;
+	\item Type of service from destination to source;
+	\item Total of Packets;
+	\item Total of Bytes;
+\end{itemize}
 
+Our analysis of the available features leads to discard some features, considering that highly correlated features can bias or not improve the model, and that source or destination IP addresses can insert some false bias into learning models, since they can be changed by IP spoofing or other techniques.
+
+TODO: talk about the feature distribution, that arent similar to pareto or lognormal, but it is not normal too
+
+An Exploratory Data Analysis (EDA) reveals that distributions of some features are skewed and present an overlapping between normal and anomalous flows, as can be seen in  \ref{fig:4.01}, that presents the distribution of transaction state of scenario 16, and exemplifies the data skewness, the overlapping and the long tail distribution of normal data.
 \begin{figure}[h!]
      \centering
-     \includegraphics[width=10cm]{figures/raw_distplot_capture20110812_Dur.png}
+     \includegraphics[width=10cm]{figures/raw_distplot_capture20110816_State.png}
      \caption{Distribution of flow duration for scenario 12.}
-     \label{fig:fig01}
+     \label{fig:4.01}
 \end{figure}
 
-To face data imbalance and skewness, we adopt a feature engineering by means of a sampling method based on data aggregation for obtaining more discriminative features, in accordance to important approaches for network anomaly detection \cite{lakhina2005mining, callegari2011novel, garcia2014empirical, chandrashekar2014survey,acarali2016survey, vieira2017model}.
+Due to the number of available features and the class overlapping, we performed an correlation analysis and cross validation in order to identify the best set of features for network attack detection.
+Therefore, we adopt the following features: State, Type of service from source to destination, Type of service from destination to source, destination port, source port, and the total number of packets.
 
-\subsection{feature engineering}
-\label{sec:feat_eng}
-
-Related works proposed some features for improving network attack detection in various scenarios and approaches\cite{lakhina2005mining,gu2008botminer,callegari2011novel,chandrashekar2014survey,garcia2014empirical,acarali2016survey}.
-Inspired by these proposals and based on previous experience \cite{vieira2017model, galibus2017offline} and cross validation, we propose the following features for improving the anomaly detection by means of a method based on higher moments distances from robust subspace.
-
-\begin{itemize}
-	\item \textbf{total-flow-count :} Count of all flows.
-	\item \textbf{known-flow-count:} Count of known flows, from controlled assets of the network, such routers, proxies, switches or machines that are part of the monitoring \cite{garcia2014identifying}.
-	\item \textbf{avg-duration:} Average time of flows.
-	\item \textbf{n-dports$<$1024:} Count of destinations ports lower than 1024.
-	\item \textbf{n-sports$<$1024:} Count of source ports lower than 1024.
-	\item \textbf{n-s-a-p-address:} Count of class A source address.
-	\item \textbf{n-s-b-p-address:} Count of class B source address.
-	\item \textbf{n-s-c-p-address:} Count of class C source address.
-	\item \textbf{n-s-na-p-address:} Count of source address that are not class A, B or C, such as reserved class D and E addresses, IPV6 or internal address.
-\end{itemize}
-
-The proposed features are counting and average calculations, therefore it is necessary to adopt time windows to aggregate flows and calculate the features.
-We evaluate windows of 0.15 and 0.25 seconds, considering account that lower windows reduce the meantime between the data collection and results of the botnet detection approach, as well as permit to have more details to exactly identify malicious flows and attackers.
-
-Figure \ref{fig:fig02} shows the distribution of average flow duration, for scenario 12, aggregated by windows of 0.15 seconds.
-In comparison to Figure \ref{fig:fig01}, that shows the distribution of flow duration for the same scenario, we can see that the feature based on aggregation and on average flow duration is more discriminative and shows a better separation between normal and anomalous classes.
-Therefore, such data can improve the anomaly detection by learning algorithms.
-
-\begin{figure}[h!]
-     \centering
-     \includegraphics[width=10cm]{figures/agg_distplot_0_15s_12_avg_duration.png}
-     \caption{Average flow duration for scenario 12, aggregated by 0.15 seconds.}
-     \label{fig:fig02}
-\end{figure}
-
-The resulting data set is modeled into the matrix $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ with $n$ rows as observations and $p$ columns of features, and is divided into scenarios, where each $i$-th scenario, presented by ID column in Table \ref{tab:tab01}, are represented as $\boldsymbol{X}_i$ and can be denoted as $\boldsymbol{X}_{18-2}$, for example.
-The normal traffic, required for training the semi-supervised approach, is denoted as $\boldsymbol{X}^r$
+The full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}.
+Each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^n$ containing 50\% of the the normal data, and into $\boldsymbol{X}_i^t$ that is composed by all anomalous flows and ne needed number of normal flows to have a testing data with 33\% of contamination.
 
@@
 
-This section describes the proposed approach for botnet detection by means of a distance analysis between moments computed from a robust subspace and new observations of network traffic.
+This section describes the proposed approach for network attack detection by means of a distance analysis between moments computed from a robust subspace and new observations of network traffic.
 
@@
 \cite{wright2009robust}, under rather broad conditions, as long as the error matrix S is sufficiently sparse, it is possible to recover a low-rank matrix by solving the following convex optimization problem:
-\begin{equation}\label{eq:eq01}
+\begin{equation}\label{eq:4.01}
 	(\boldsymbol{\hat{L}}, \boldsymbol{\hat{S}})\leftarrow \min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}
@@
 \begin{center} subject to: $\boldsymbol{X} = \boldsymbol{L} + \boldsymbol{S}$ \end{center}
-\begin{equation}\label{eq:eq02}
+\begin{equation}\label{eq:4.02}
     \left \| \boldsymbol{L} \right \|_{*} = \sum_{i} \sigma_{i}(\boldsymbol{L})
 \end{equation}
-\begin{equation}\label{eq:eq03}
+\begin{equation}\label{eq:4.03}
     \left \| \boldsymbol{S} \right \|_{1} = \sum_{ij} \left | \boldsymbol{S}_{ij} \right |
@@
 The RPCA with ALM can be formulated:
-\begin{equation}\label{eq:eq04}
+\begin{equation}\label{eq:4.04}
 	l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}) = \left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
@@
 Thus, an iterative scheme can be presented as:
-\begin{equation}\label{eq:eq05}
+\begin{equation}\label{eq:4.05}
 	\left\{
@@
 
-\begin{equation}\label{eq:eq06}
+\begin{equation}\label{eq:4.06}
 	\left\{\begin{matrix}
@@
 
-\begin{equation}\label{eq:eq07}
+\begin{equation}\label{eq:4.07}
 	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i, 
@@
 
-\begin{equation}\label{eq:eq08}
+\begin{equation}\label{eq:4.08}
 	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
@@
 The general formula of the $u$-th moment can be expressed as
-\begin{equation}\label{eq:eq09}
+\begin{equation}\label{eq:4.09}
 	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
@@
 Therefore, the skewness is calculated by
-\begin{equation}\label{eq:eq10}
+\begin{equation}\label{eq:4.10}
 	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
@@
 and the kurtosis as
-\begin{equation}\label{eq:eq11}
+\begin{equation}\label{eq:4.11}
 	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}.
@@
 
-We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:eq06} for $\textbf{X}^r$.
+We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:4.06} for $\textbf{X}^r$.
 We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between new observations and the moments calculated from a robust subspace computed by RPCA.
@@
 The classical Mahalanobis Distance is defined as		
-\begin{equation}\label{eq:eq12}
+\begin{equation}\label{eq:4.12}
 	\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
@@
 Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA.
-We also propose to extend the Equation \ref{eq:eq12} to implement a Skewness-based Mahalanobis Distance, as follows:
+We also propose to extend the Equation \ref{eq:4.12} to implement a Skewness-based Mahalanobis Distance, as follows:
 
-\begin{equation}\label{eq:eq13}
+\begin{equation}\label{eq:4.13}
 	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}.
@@
 
-We finally propose to extend the Equation \ref{eq:eq12} to implement a Kurtosis-based Mahalanobis Distance, as follows:
+We finally propose to extend the Equation \ref{eq:4.12} to implement a Kurtosis-based Mahalanobis Distance, as follows:
 
-\begin{equation}\label{eq:eq14}
+\begin{equation}\label{eq:4.14}
 	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}.
@@
 The distances $\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})$, $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})$ and $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})$ shall be computed and evaluated separately for botnet detection.
-Therefore, the robust subspace learning from normal data combined to MD can be called mmd-RPCA when using mean-based MD, smd-RPCA when using skewness-based MD, and kmd-RPCA when using kurtosis-based MD.
+Therefore, the robust subspace learning from normal data combined to MD can be called md-RPCA when using mean-based MD, sd-RPCA when using skewness-based MD, and kd-RPCA when using kurtosis-based MD.
 
@@
 In our proposal, the contamination define the rate of highest distances $\boldsymbol{d}$ that shall be classified as anomalous.
-Therefore, the resulting $\boldsymbol{\bar{t}}$ refers to observations classified as anomalous according to $c$.
+Therefore, the resulting $\boldsymbol{\bar{t}}$ refers to observations classified as anomalous according to the contamiation $c$.
 
-The Algorithm \ref{alg:alg01} presents how the proposed Moment Distances from Robust Subspace for Botnet Detection can be algorithmically described.
+The Algorithm \ref{alg:4.01} presents how the m-RPCA can be algorithmically described.
 \begin{algorithm}
-	\label{alg:alg01}
+	\label{alg:4.01}
 	\SetAlgoLined
@@
 
-The steps between 1 and 10 of the Algorithm \ref{alg:alg01} are the training from normal data $\boldsymbol{X}^n$, while the steps between 11 and 16 aim the anomaly detection on new contamianted observations.
-Note that the steps 11 and 14 refers to the mmd-RPCA, while the steps 12 and 15 refers to the smd-RPCA, and the steps 13 and 16 refers to the kmd-RPCA.
+The steps between 1 and 10 of the Algorithm \ref{alg:4.01} are the training from normal data $\boldsymbol{X}^n$, while the steps between 11 and 16 aim the anomaly detection on new contaminated observations.
+Note that the steps 11 and 14 refers to the md-RPCA, while the steps 12 and 15 refers to the sd-RPCA, and the steps 13 and 16 refers to the kd-RPCA.
 It is also important to observe that the anomaly detection from new observations does not require new robust subspace learning, it is only necessary to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous.
+
+However, m-RPCA can also be adopted as unsupervised algorithm through changing the lines 1 and 2 of the Algorithm \ref{alg:4.01} to substitute $\boldsymbol{X}^n$ by $\boldsymbol{X}^t$.
+Hence, the robust subspace is learned from contaminated data and used for comparing the distance between higher moments  from robust estimate and moments of new observations.
 
@@
 This section presents the performed experiments and the acquired results.
-First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
-Section \ref{sec:results} shows the results of the proposed approach, compares to selected anomaly detection algorithms, and presents a discussion about the results.
+First, in Section \ref{sec:4_AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
+Section \ref{sec:4_results} shows the results of the proposed approach, compares to selected anomaly detection algorithms, and presents a discussion about the results.
 
 \subsection{Experimental Scenario}
-\label{sec:AnalyzedScenario}
+\label{sec:4_AnalyzedScenario}
 
@@
 
-Therefore, the full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \\\ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:tab01}.
+Therefore, the full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \\\ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}.
 In our experiment setup each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^n$ containing 50\% of the the normal data, and into the cross validation $\boldsymbol{X}_i^v$, composed by 25\% of the normal and 50\% of the anomalous data, and into test $\boldsymbol{X}_i^t$ containing 25\% of the normal and 50\% of the anomalous data.
@@
 The F-measure is the harmonic mean of precision and recall and is defined by:
-\begin{equation}\label{eq:eq15}
+\begin{equation}\label{eq:4.15}
 	F-measure = 2 * \frac{pr * rc}{pr + rc},
@@
 where $pr$ is the precision, defined by 
-\begin{equation}\label{eq:eq16}
+\begin{equation}\label{eq:4.16}
 	pr = \frac{True Positive}{True Positive + False Positive}
@@
 and $rc$ is the recall, defined by 
-\begin{equation}\label{eq:eq17}
+\begin{equation}\label{eq:4.17}
 	rc = \frac{True Positive}{True Positive + False Negative}.
@@
 
-It is from $\boldsymbol{X}_i^n$ that we perform the training steps, which are the steps between 1 and 10 described in Algorithm \ref{alg:alg01}, resulting into $\boldsymbol{\bar{x}}$, $\boldsymbol{\bar{\Sigma}}$, $\boldsymbol{\bar{s}}$ and $\boldsymbol{\bar{k}}$.
+It is from $\boldsymbol{X}_i^n$ that we perform the training steps, which are the steps between 1 and 10 described in Algorithm \ref{alg:4.01}, resulting into $\boldsymbol{\bar{x}}$, $\boldsymbol{\bar{\Sigma}}$, $\boldsymbol{\bar{s}}$ and $\boldsymbol{\bar{k}}$.
 The next step of the experiment is the cross validation, which is an empirical testing to identify the contamination $c$ that obtains the best F-measure on anomaly detection, aiming the maximum precision and recall.
@@
 
-Finally, we test the proposed approach to anomaly detection on $\boldsymbol{X}_i^t$ by means of the steps between 11 and 16 of Algorithm \ref{alg:alg01}, and compare the results of $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$ to the ground truth results for $\boldsymbol{X}_i^t$.
+Finally, we test the proposed approach to anomaly detection on $\boldsymbol{X}_i^t$ by means of the steps between 11 and 16 of Algorithm \ref{alg:4.01}, and compare the results of $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$ to the ground truth results for $\boldsymbol{X}_i^t$.
 
 \subsection{Results}
-\label{sec:results}
+\label{sec:4_results}
 
-The Table \ref{tab:tab02} presents the results of the proposed approach and selected methods applied to botnet detection for each scenario of the CTU-13 data set.
-The Table \ref{tab:tab02} shows the F-measure of GMM, K-means, ROBPCA, md-RPCA, kd-RPCA and sd-RPCA by scenario and by an average of all scenarios.
+The Table \ref{tab:4.02} presents the results of the proposed approach and selected methods applied to botnet detection for each scenario of the CTU-13 data set.
+The Table \ref{tab:4.02} shows the F-measure of GMM, K-means, ROBPCA, md-RPCA, kd-RPCA and sd-RPCA by scenario and by an average of all scenarios.
 
@@
   \caption{Results grouped by algorithm and scenario}
-  \label{tab:tab02}
+  \label{tab:4.02}
   \begin{tabular}{ c|c c|c c|c c|c c|c c|c c }
@@
 say that more studies about skewness and subspace learning can be done in order to explore it inside algorithms and non gaussian data
+
+note that the semi supervised approach requires subspace learning only for training and not for testing, while unsupervised requires subspace learning for each evaluations

### 2019-05-27 docs/thesis/thesis.tex
tensor dl motivation
tensor-related works
improving the approach and experiment description
prepare for thesis writing
first deliverable

@@
 \documentclass[en, oneside, onehalfspacing]{risethesis}
-
+% \usepackage[mathletters]{ucs}
+\usepackage{ucs}
+\usepackage[utf8x]{inputenc}
+% table
+\usepackage{makecell}
+\usepackage{color, colortbl}
+\definecolor{Gray}{gray}{0.9}
+\definecolor{lightgray}{gray}{0.8}
 % packages and configurations
@@
 \usepackage{babel}
-\usepackage{supertabular}
 \usepackage{fancybox}
@@
 \usepackage{array}
-\usepackage{booktabs}
 \usepackage{graphicx}
 \usepackage{rotating}
+\usepackage{booktabs}
+\usepackage{supertabular}
 \usepackage{tabularx}
+\usepackage{longtable}
 \usepackage{color, colortbl}
@@
 \usepackage{placeins}
-\usepackage{longtable}
 \usepackage{amsthm,amsmath}
 \usepackage{mathtools}
-\usepackage{algorithm}% http://ctan.org/pkg/algorithms
-\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
-\usepackage{algpseudocode}
+
+\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
 \usepackage{listings}
@@
 \usepackage{hyperref}
+
+% appendix
+\usepackage{appendix}
+\usepackage{hyperref}
+\usepackage{cleveref}
 
@@
 
-\definecolor{lightgray}{gray}{0.8}
-\definecolor{Gray}{gray}{0.9}
 \renewcommand{\lstlistingname}{Code}
@@
 
-\title{Discriminative Sensing Based on Signal Processing for Information Security Analysis}
-\date{November/2017}
+\title{On The Subspace Learning for Network Attack Detection}
+\date{July/2019}
 
@@
 \include{chapters/3_mobile/mobile}
-\include{chapters/4_tensor_dl/tensor_dl}
+\include{chapters/4_m_rpca/m_rpca}
 \include{chapters/5_conclusion_future_work/conclusion}
@@
 \include{chapters/appendix/b_pca_svm_cfa/pca_svm_cfa}
+\include{chapters/appendix/c_tensor_denoising/c_tensor_denoising}
 

### 2019-06-09 docs/thesis/chapters/1_introduction/introduction.tex
results of robpca

@@
 Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}.
-In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time.
+In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency in resource consumption, processing capacity and response time.
 Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
-Recently, signal processing schemes have being applied to detect malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest, vieira2017model}, showing advances in network traffic analysis.
+Recently, signal processing schemes have been applied to detect malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest, vieira2017model}, showing advances in network traffic analysis.
 
@@
 The former includes equipment and security systems, while the latter corresponds to security rules and recommendations of good practices.
-Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks.
+Intrusion detection and intrusion prevention systems are two types of security systems that are being used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks.
 Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.
 
-Anomaly detection can be defined as the identification of rare and suspicions events by differing the normal or majority of the data.
+Anomaly detection can be defined as the identification of rare and suspicious events by comparing it to the normal data.
 Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects \cite{bhuyan2014network,ahmed2016survey}.
-Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
-Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.
+Additionally, anomaly detection techniques can be categorized by classification, statistical algorithms, information theory and cluster based algorithms, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
+Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events.
+Therefore anomaly detection algorithms have to be highly discriminating, robust to corruption and able to deal with the imbalanced data problem \cite{he2008learning}.
 
 Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or behavioral analysis.
-However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid unknown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detection systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
+However, distributed attacks organized by botnet has increased and demanded the development of counter measures.
+This has been done to detect and avoid unknown attacks or even to deal with adversarial changes of behavior, location and other patterns.
+This has also been done in order to avoid detection by behavioral or pattern based detection systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
 To face this problem, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, where it is not necessary known anomalies for training models \cite{moustafa2019holistic}.

### 2019-06-09 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
results of robpca

@@
 According to Table \ref{tab:11}, the processing time increases with the frame size $N$ decreasing, therefore it is possible to evaluate the frame size that produces better identification rates and acceptable processing time.
-The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations, since the significant increase of processing time observed between scenarios considering 17 or 34 ports, with growth between 7 \% and 199 \%.
+The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations, since the significant increase of processing time observed between scenarios considering 17 or 34 ports, with growth between 7\% and 199\%.
 

### 2019-06-09 docs/thesis/chapters/4_m_rpca/m_rpca.tex
results of robpca

@@
 The Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries.
-We also define the operator $\langle \mathord{\cdot} \rangle$, that is the standard trace inner product, and $[ \mathord{\cdot} ]^n$, which denotes the $n$ largest values of a vector.
+We also define the operator $\langle \mathord{\cdot} \rangle$, that is the standard trace inner product, and $[ \mathord{\cdot} ]^s$, which denotes the $n$ largest values of a vector.
 
@@
 
-In order to evaluate the null hypothesis $H_{0rob.sub}$, which assumes that the robust subspace learning does not improves the anomaly detection from imbalanced and skewed, we create two synthetic data sets characterized by skewed and heavy tailed distributions.
-We selected Pareto and Lognormal distributions to evaluate the anomaly detection capacity of our proposal, considering that these distributions have been adopted to represent network traffic in the internet and data centers \cite{benson2010network,leon2017probability}.
+In order to analyse the null hypotheses $H_{1rob.sub}$ and $H_{0rob.sub}$, which evaluate if the robust subspace learning can improves the anomaly detection from imbalanced and skewed data, we create two synthetic data sets characterized by skewed and heavy tailed distributions.
+We selected Pareto and Lognormal distributions to evaluate the anomaly detection capacity of our proposal, considering that these distributions are skewed and heavy-tailed, and have been adopted to characterize network traffic of the Internet and data centers \cite{benson2010network,leon2017probability}.
+We also evaluate our proposal and the selected algorithms for anomaly detection on gaussian distributed data set, in order to analyse the detection rate for not skewed and not heavy tailed distributions.
 
-We adopt a contamination rate of 33\% to reproduce the imbalanced data of anomaly detection problems, and select the gaussian distribution for the addictive anomalies into Pareto and Lognormal distributions.
-Therefore, we generate $\boldsymbol{Y}_p$ and $\boldsymbol{Y}_l$ with 2,400 samples and two features, where 793 samples are gaussian anomalies, also referred as noise or contamination, while 1,607 samples are legitimate signals.
-The data set $\boldsymbol{Y}_p$ is composed by random Pareto distribution with random gaussian noise, while $\boldsymbol{Y}_l$ is composed by random Lognormal distribution with random gaussian noise.
+The addictive anomalies for Pareto and Lognormal distributions are gaussian distributed, while an uniform noise is used to add anomalies into the gaussian distributed data set.
+We generate $\boldsymbol{Y}_p$, $\boldsymbol{Y}_l$ and $\boldsymbol{Y}_g$ for Pareto, Lognormal and Gaussian multivariate distributions, respectively, with 2 features.
+Each contaminated data set is composed by a number of legitimate observations and the contaminated samples, defined by the adopted $c$.
 
-The Figure \ref{fig:4.03} shows the histogram and the PDF of the Pareto distribution for two features of $\boldsymbol{Y}_p$.
+The Pareto distribution contaminated by gaussian noise is denoted as $\boldsymbol{Y}_p^c$, while $\boldsymbol{Y}_l^c$ denotes the Lognormal distribution contaminated by gaussian noise, and $\boldsymbol{Y}_g^c$ denotes the contaminated gaussian data contaminated by an uniform distribution.
+
+This synthetic data set simulate a total of 2400 events for each scenario, where the number of legitimate observations is simulated according to the desired contamination rate.
+We evaluate contamination rates $c$ between 1\% and 50\%, to simulate the imbalanced data of anomaly detection problems during our experiments.
+
+The Figure \ref{fig:4.01} depicts the distribution of Pareto and gaussian contamination of one variable, as a visual example of $\boldsymbol{Y}_p^c$, while the Figure \ref{fig:4.02} shows one example of Lognormal contaminated by gaussian noise, generated for $\boldsymbol{Y}_l^c$.
 
@@
 	\centering
-	\subfigure[Feature 1]{%
-		\includegraphics[width=0.48\textwidth]{figures/pareto1.png}
+	\subfigure[Pareto and Gaussian Anomalies]{%
+		\includegraphics[width=0.47\textwidth]{figures/ch4/4_Xpg2.png}
 		\label{fig:4.01}
 	}
-	\subfigure[Feature 2]{%
-		\includegraphics[width=0.48\textwidth]{figures/pareto2.png}
+	\subfigure[Lognormal and Gaussian Anomalies]{%
+		\includegraphics[width=0.47\textwidth]{figures/ch4/4_Xlogng2.png}
 		\label{fig:4.02}
 	}
-	\caption[Pareto Distribution]{Pareto Distribution}
+	\caption[Examples of skewed and heavy tailed distributions]{Examples of skewed and heavy tailed distributions}
 	\label{fig:4.03}
 \end{figure}
-
-The Figure \ref{fig:4.06} shows the histogram and the PDF of the Lognormal distribution for two features of $\boldsymbol{Y}_l$.
-
-\begin{figure}[!htb]
-	\centering
-	\subfigure[Feature 1]{
-		\includegraphics[width=0.48\textwidth]{figures/lognormal3.png}
-		\label{fig:4.04}
-	}
-	\subfigure[Feature 2]{
-		\includegraphics[width=0.48\textwidth]{figures/lognormal4.png}
-		\label{fig:4.05}
-	}
-	\caption[Lognormal Distribution]{Lognormal Distribution}
-	\label{fig:4.06}
-\end{figure}
-
-The Pareto distribution contaminated by gaussian noise is denoted as $\boldsymbol{Y}_p^c$, while $\boldsymbol{Y}_l^c$ denotes the Lognormal distribution contaminated by gaussian noise.
-Taking into account that semi-supervised algorithms requires training from normal data, we also generate $\boldsymbol{Y}_p^t$, while $\boldsymbol{Y}_l^t$, which are data sets of Pareto and Lognormal distributions for training, without contamination and composed by 1,607 samples and two features.
 
@@
 
-The CTU-13 is a data set of botnet traffic that was captured in the Czech Technical University \cite{garcia2014empirical}, by means of a testbed and execution of malwares in a real network.
+The CTU-13 \cite{garcia2014empirical} is a data set of botnet traffic that was captured in the Czech Technical University, by means of a testbed and malware execution in a real network.
 The CTU-13 data set contains 13 scenarios with network flows of botnet malwares, that are: neris, rbot, virut, menti, sogou, nsys.ay and murlo.
@@
 
-We refer to Garcia \cite{garcia2014identifying} and Garcia \emph{et al.} \cite{garcia2014empirical} for a detailed description of the performed attacks and C\&C flows, also including the topology of the adopted testbed, rules for classifying normal flows, and an analysis of behaviors or patterns of the malware's traffic.
+We refer to Garcia \cite{garcia2014identifying} and Garcia \emph{et al.} \cite{garcia2014empirical} for a detailed description of the performed attacks and C\&C flows, as well as for more information about the topology of the adopted testbed, rules for classifying normal flows, and an analysis of behaviors or patterns of the malware's traffic.
 
 For all the scenarios, the authors of the CTU-13 data set convert the captured pcap files to NetFlows and release the processed flows.
-The data set contains ground-truth labels for flows as follows: flows from or to the infected machines are labeled as “botnet”; flows from or to well known and controlled machines are labeled as “normal”; all other flows are labeled as “background.”
+The data set provides ground-truth labels for flows as follows: flows from or to the infected machines are labeled as “botnet”; flows from or to well known and controlled machines are labeled as “normal”; all other flows are labeled as “background.”
 
-Table \ref{tab:4.01} presents an overview organized by scenario and shows the malwares used for botnet attacks, the types of attacks and C\&C types, the total number of flows, the number of malicious flows, which includes flows of C\&C and attacks, and finally shows the number of normal flows.
+Table \ref{tab:4.01} presents an overview grouped by scenario, according to the column ID, and shows the malwares used for botnet attacks, the types of attacks or C\&C, the total number of flows, the number of malicious flows which includes flows of C\&C and attacks, and finally shows the number of normal flows.
 
@@
 
-The CTU-13 data set originally contains the following features:
+The full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}.
+In our experiment each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^s$ containing 50\% of the the normal data, and into $\boldsymbol{X}_i^c$ that is composed by all anomalous flows and the necessary number of normal flows to have a testing data with the desired contamination rate.
+
+The CTU-13 data set originally contains the following features for each flow:
 
@@
 	\item Destination Port;
-	\item State;
-	\item Type of service from source to destination;
-	\item Type of service from destination to source;
-	\item Total of Packets;
-	\item Total of Bytes;
+	\item State of TCP flags;
+	\item Destination Type of Service;
+	\item Source Type of Service;
+	\item Total number of Packets;
+	\item Total number of Bytes;
 \end{itemize}
 
-Our analysis of the available features leads to discard some features, considering that highly correlated features can bias or not improve the model, and that source or destination IP addresses can insert some false bias into learning models, since they can be changed by IP spoofing or other techniques.
+Our analysis of the available features leads to discard some features, considering that highly correlated features can bias or not improve the model, and that source or destination IP addresses can insert some false bias into learning models, since they can be changed by IP spoofing.
+Other risk related to adopt IP address for training models is the training model to learn that
+one IP is legitimate and this IP be infected subsequently, which can result into false negative classifications.
 
-TODO: talk about the feature distribution, that arent similar to pareto or lognormal, but it is not normal too
+We conducted and Exploratory Data Analysis (EDA) on the CTU-13 and observed that some features are skewed and present an overlapping between normal and anomalous flows, as can be seen in Figure \ref{fig:4.05} and Figure \ref{fig:4.05}, that present the distributions of TCP state of scenario 16 and the type of services from destination of scenario 10.
 
-An Exploratory Data Analysis (EDA) reveals that distributions of some features are skewed and present an overlapping between normal and anomalous flows, as can be seen in  \ref{fig:4.01}, that presents the distribution of transaction state of scenario 16, and exemplifies the data skewness, the overlapping and the long tail distribution of normal data.
-\begin{figure}[h!]
+\begin{figure}[!htb]
 	\centering
-     \includegraphics[width=10cm]{figures/raw_distplot_capture20110816_State.png}
-     \caption{Distribution of flow duration for scenario 12.}
-     \label{fig:4.01}
+	\subfigure[State of scenario 16.]{%
+		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110816_State.png}
+		\label{fig:4.04}
+	}
+	\subfigure[Destination Type of service of scenario 10.]{%
+		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110810_dTos.png}
+		\label{fig:4.05}
+	}
+	\caption[Skewness and Overlapping]{Example of skewness and overlapping}
+	\label{fig:4.06}
 \end{figure}
 
-Due to the number of available features and the class overlapping, we performed an correlation analysis and cross validation in order to identify the best set of features for network attack detection.
-Therefore, we adopt the following features: State, Type of service from source to destination, Type of service from destination to source, destination port, source port, and the total number of packets.
+However, it is not possible to observe a pattern on distributions of all the features, as depicted by Figures \ref{fig:4.07} and \ref{fig:4.08}, that show the distributions of TCP states of the scenario 10 and source ports of scenario 16, and highlight the distributions of normal and anomalous flows.
 
-The full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}.
-Each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^n$ containing 50\% of the the normal data, and into $\boldsymbol{X}_i^t$ that is composed by all anomalous flows and ne needed number of normal flows to have a testing data with 33\% of contamination.
+\begin{figure}[!htb]
+	\centering
+	\subfigure[State of scenario 10.]{%
+		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110810_State.png}
+		\label{fig:4.07}
+	}
+	\subfigure[Source Port of scenario 16.]{%
+		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110816_Sport.png}
+		\label{fig:4.08}
+	}
+	\caption[Skewness and Overlapping]{Skewness and Overlapping}
+	\label{fig:4.09}
+\end{figure}
+
+Due to the number of available features and the class overlapping between normal and anomalous flows, we performed an correlation analysis and cross validation in order to identify the best set of features for network attack detection.
+Therefore, we adopt the following features: state, destination type of service, destination port, source port, total number of packets, total number of bytes and number of bytes from the source.
 
@@
 Robust subspace learning can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
-We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}^r$ for computing $\textbf{L}^r$, $\textbf{S}^r$ and the robust moments, i.e.
-the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^t$ to these moments and classify as anomalous the observations with the largest distances.
+We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}^s$ for computing $\textbf{L}^s$, $\textbf{S}^s$ and the robust moments, i.e.
+the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^c$ to these moments and classify as anomalous the observations with the largest distances.
 
@@
 Before ALM, some methods were proposed to solve that convex optimization problem, such as Iterative Thresholding (IT) and Accelerated Proximal Gradient (APG).
-However, according to \cite{lin2010augmented}, both approaches have scalability problems and require a large number of iterations to converge.
+However, according to Zhouchen \emph{et al.} \cite{lin2010augmented}, both approaches have scalability problems and require a large number of iterations to converge.
 The Augmented Lagrange Multiplier is proven to have a \emph{Q-linear} convergence speed and experimental results show that ALM is five times faster than APG, which in theory is sub-linear \cite{lin2010augmented}.
@@
 
-We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^r$, after to minimize the Equation \ref{eq:4.06} for $\textbf{X}^r$.
+We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^s$, after to minimize the Equation \ref{eq:4.06} for $\textbf{X}^s$.
 We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between new observations and the moments calculated from a robust subspace computed by RPCA.
@@
 The contamination rate can be well known for some areas, or can be computed by cross-validation or can be assumed according to previous observations.
-In our proposal, the contamination define the rate of highest distances $\boldsymbol{d}$ that shall be classified as anomalous.
-Therefore, the resulting $\boldsymbol{\bar{t}}$ refers to observations classified as anomalous according to the contamiation $c$.
+In our proposal, the contamination define the rate of the largest distances $\boldsymbol{d}$ that shall be classified as anomalous.
+Therefore, the resulting $\boldsymbol{\bar{t}}$ refers to observations classified as anomalous according to the contamination $c$.
 
@@
 	\KwResult{$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$}
-	Given $\boldsymbol{X}$ split into $\boldsymbol{X}^n$ and $\boldsymbol{X}^t$\;
-	\While{not $\min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}$ of $\boldsymbol{X}^n$}{
+	Given $\boldsymbol{X}$ split into $\boldsymbol{X}^s$ and $\boldsymbol{X}^c$\;
+	\While{not $\min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}$ of $\boldsymbol{X}^s$}{
 		$\boldsymbol{L}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{L} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}, \boldsymbol{S}_{k}, \boldsymbol{Y}_{k})\}$\;
@@
 	}
-	$\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i$, for $\boldsymbol{L}^n$\;
-	$\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime$, for $\boldsymbol{L}^n$\;
-	$\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}}$, for $\boldsymbol{L}^n$\;
-	$\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}$, for $\boldsymbol{L}^n$\;
+	$\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i$, for $\boldsymbol{L}^s$\;
+	$\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime$, for $\boldsymbol{L}^s$\;
+	$\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}}$, for $\boldsymbol{L}^s$\;
+	$\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}$, for $\boldsymbol{L}^s$\;
 	$\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
@@
 
-The steps between 1 and 10 of the Algorithm \ref{alg:4.01} are the training from normal data $\boldsymbol{X}^n$, while the steps between 11 and 16 aim the anomaly detection on new contaminated observations.
-Note that the steps 11 and 14 refers to the md-RPCA, while the steps 12 and 15 refers to the sd-RPCA, and the steps 13 and 16 refers to the kd-RPCA.
+The steps between 1 and 10 of the Algorithm \ref{alg:4.01} are the training from normal data $\boldsymbol{X}^s$, while the steps between 11 and 16 aim the anomaly detection from new contaminated observations.
+Note that the steps 11 and 14 refer to the md-RPCA, while the steps 12 and 15 refer to the sd-RPCA, and the steps 13 and 16 refer to the kd-RPCA.
 It is also important to observe that the anomaly detection from new observations does not require new robust subspace learning, it is only necessary to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous.
 
-However, m-RPCA can also be adopted as unsupervised algorithm through changing the lines 1 and 2 of the Algorithm \ref{alg:4.01} to substitute $\boldsymbol{X}^n$ by $\boldsymbol{X}^t$.
-Hence, the robust subspace is learned from contaminated data and used for comparing the distance between higher moments  from robust estimate and moments of new observations.
+The m-RPCA can also be adopted as unsupervised algorithm if the lines 1 and 2 of the Algorithm \ref{alg:4.01} are changed to substitute $\boldsymbol{X}^s$ by $\boldsymbol{X}^c$.
+Hence, the robust subspace is learned from contaminated data and used for comparing the distance between higher moments from robust estimate and moments of the observations of the same data.
+It is important to note that this unsupervised approach requires the computational cost of the subspace learning for any new observation.
 
@@
 
-This section presents the performed experiments and the acquired results.
-First, in Section \ref{sec:4_AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized.
-Section \ref{sec:4_results} shows the results of the proposed approach, compares to selected anomaly detection algorithms, and presents a discussion about the results.
+This section presents the performed experiments on simulated and real data set for anomaly detection.
+First, in Section \ref{sec:4_metric} we present the adopted metric to evaluate imbalanced data in the context of anomaly detection.
+The Section \ref{sec:4_SimulatedScenario} describes the experiment for anomaly detection on simulated skewed and heavy-tailed data set, and the Section \ref{sec:4_CTU-13} presents the experiment for network anomaly detection on CTU-13 data set.
 
-\subsection{Experimental Scenario}
-\label{sec:4_AnalyzedScenario}
-
-Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training anomaly detection models, while the latter does not require labeled data or training.
-However, the availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or even unkown events.
-
-Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches.
-The proposed approach is semi-supervised and relies on normal data $\boldsymbol{X}^n$ and on contamination rate $c$ for training and anomaly detection, respectively.
-
-Only for evaluation purpose, we use part of the contaminated data, denoted as $\textbf{X}^v$, for cross-validation, in order to estimate the contamination rate $c$ that best detect anomalies from $\textbf{X}^v$.
-However, we assume that the $c$ is well known or can be estimated for real world problems of anomaly detection.
-
-Clustering methods for anomaly detection rely on three key assumptions: legitimate data instances often fall into a cluster whereas attacks do not, legitimate data instances are usually located near the closest cluster centroid while anomaly ones are often far away from it, legitimate data instances fall into vast and dense clusters and anomalies into small or spare ones \cite{ahmed2016survey, moustafa2019holistic}.
-Gaussian Mixture Model (GMM) is a statistical-based and parametric approach algorithm that estimates the distribution of the normal class from a training set and is typically based on a set of kernels \cite{moustafa2019holistic}.
-
-This experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches.
-As clustering-based method, we compare to K-means, which is the clustering algorithm most adopted for general purposes and for network anomaly detection \cite{gaddam2007kmeans,moustafa2019holistic}.
-It is observed that network traffic do not belong to a Gaussian distribution \cite{benson2010network,moustafa2019holistic}, and that algorithms based on non-Gaussian distributions, such as GMM, has been adopted for network anomaly detection \cite{moustafa2019holistic}.
-Therefore, we compare our proposal to GMM.
-We also evaluate the results of ROBPCA for anomaly detection on CTU-13, considering that ROBPCA also relies on robust estimates with adjusted outlyingness based on robust skewness \cite{hubert2009robustskewed}.
-
-In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach to all botnet malwares of the CTU-13.
-
-Therefore, the full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \\\ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}.
-In our experiment setup each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^n$ containing 50\% of the the normal data, and into the cross validation $\boldsymbol{X}_i^v$, composed by 25\% of the normal and 50\% of the anomalous data, and into test $\boldsymbol{X}_i^t$ containing 25\% of the normal and 50\% of the anomalous data.
+\subsection{The metric}
+\label{sec:4_metric}
 
 In anomalous detection problems, where anomalies are rare events, if one classify all observations as normal and apply an accuracy evaluation, one would have high accuracy but poor true-positive detection.
-Due to the importance given by the F-measure to true-positive detection, it is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic} and it is the metric used for validation of this experiment.
+Due to the importance given by the F-measure to true-positive detection in scenarios such as the network attack detection, it is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic}.
+Therefore, F1-measure is the metric used for validation of our experiments.
 
@@
 
-It is from $\boldsymbol{X}_i^n$ that we perform the training steps, which are the steps between 1 and 10 described in Algorithm \ref{alg:4.01}, resulting into $\boldsymbol{\bar{x}}$, $\boldsymbol{\bar{\Sigma}}$, $\boldsymbol{\bar{s}}$ and $\boldsymbol{\bar{k}}$.
-The next step of the experiment is the cross validation, which is an empirical testing to identify the contamination $c$ that obtains the best F-measure on anomaly detection, aiming the maximum precision and recall.
-During the cross validation for $\boldsymbol{X}_i^v$ we compute the steps 11, 12 and 13, and iterate over the steps between 14 and 16 for selected values of $c$, and validate the results of $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}^v$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}^v$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}^v$ to the ground truth results for $\boldsymbol{X}_i^v$, in order to identify the $c$ that best classify the anomalies of $\boldsymbol{X}_i^v$.
+This experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection that also rely on contamination rate for anomaly detection, which are: PCA based on the sum of weighted projected distances to the eigenvector hyperplanes \cite{shyu2003novel}; MCD \cite{rousseeuw1984mcd,rousseeuw1999fastmcd}; One-Class Support Vector Machines \cite{scholkopf2001estimating}; Local Outlier Factor (LOF) \cite{breunig2000lof}; k-Nearest Neighbors \cite{angiulli2002fast}; and Isolation Forest \cite{liu2008isolation}.
 
-Finally, we test the proposed approach to anomaly detection on $\boldsymbol{X}_i^t$ by means of the steps between 11 and 16 of Algorithm \ref{alg:4.01}, and compare the results of $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$ to the ground truth results for $\boldsymbol{X}_i^t$.
+% ToDO: We also evaluate the results of ROBPCA for anomaly detection on CTU-13, considering that ROBPCA also relies on robust estimates with adjusted outlyingness based on robust skewness \cite{hubert2009robustskewed}.
 
-\subsection{Results}
-\label{sec:4_results}
+\subsection{Simulated Experiment}
+\label{sec:4_SimulatedScenario}
 
-The Table \ref{tab:4.02} presents the results of the proposed approach and selected methods applied to botnet detection for each scenario of the CTU-13 data set.
-The Table \ref{tab:4.02} shows the F-measure of GMM, K-means, ROBPCA, md-RPCA, kd-RPCA and sd-RPCA by scenario and by an average of all scenarios.
+Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training anomaly detection models, while the latter does not require labeled data or training.
+
+Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches.
+We propose semi-supervised and unsupervised approaches for m-RPCA, where the former relies on normal data $\boldsymbol{X}^s$ for training and on contamination rate $c$ for anomaly detection, while the latter relies only on contaminated data.
+We assume that the $c$ is well known or can be estimated for real world problems of anomaly detection, in accordance to the selected algorithms for comparison, that also rely on contamination rate.
+
+The availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or even unknown events.
+Considering that RPCA has already been adopted to isolate outliers from training data \cite{zhou2017anomaly}, we also propose to evaluate the m-RPCA from a semi-supervised approach based on training from contaminated data.
+
+Therefore, we propose to evaluate the following approaches for m-RPCA:
+\begin{enumerate}
+	\item Semi-supervised;
+	\item Contaminated Semi-supervised;
+	\item Unsupervised.
+\end{enumerate}
+
+For the semi-supervised approach we propose to learn the robust subspace and compute the moments from the normal data set $\boldsymbol{Y}_g$, $\boldsymbol{Y}_p$ and $\boldsymbol{Y}_l$ with Gaussian, Pareto and Lognormal distributions, respectively, and test the anomaly detection for contaminated data set $\boldsymbol{Y}_g^c$, $\boldsymbol{Y}_p^c$ and $\boldsymbol{Y}_l^c$.
+
+In the Contaminated Semi-supervised approach we evaluate the robustness of the m-RPCA for learning from contaminated training data, in order to analyse if m-RPCA can be an alternative for the lack of known normal data.
+Therefore we propose to train the model from a contaminated normal data $\boldsymbol{Y}_g^{c'}$, $\boldsymbol{Y}_p^{c'}$ and $\boldsymbol{Y}_l^{c'}$, with the same contamination rate of the testing data, but without data sharing between training and testing, taking into account that we shall consider a different contamination data but the same contamination rate.
+
+We finally evaluate the unsupervised approach, that relies on the test data sets $\boldsymbol{Y}_g^c$, $\boldsymbol{Y}_p^c$ and $\boldsymbol{Y}_l^c$ for robust subspace learning, and then classify the results according to the distance between the contaminated data $\boldsymbol{Y}_g^c$, $\boldsymbol{Y}_p^c$ and $\boldsymbol{Y}_l^c$, and the moments computed from the learned robust subspace.
+
+\subsection{Experiment for CTU-13}
+\label{sec:4_CTU13Scenario}
+
+In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach to all botnet malwares of the CTU-13.
+
+In our experiment setup each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^s$ containing 50\% of the the normal data, and into test data $\boldsymbol{X}_i^c$ containing 33\% of normal and 67\% of anomalous data.
+
+This experiment consider only the semi-supervised approach due to results of simulated experiments on synthetic data set to be present in Subsection \ref{sec:4_simulated_result}, that shows better results for the semi-supervised approach and highlight that the this approach can obtain good results even when trained with contaminated data set.
+
+\section{Results of Simulated Experiment}
+\label{sec:4_simulated_result}
+
+We initially evaluated the F1-measure of the selected algorithms and m-RPCA approaches for anomaly detection on gaussian distributed legitimate data contaminated by uniform distributed anomalies.
+
+The Figure \ref{fig:4.10} shows the F1-measure over the contamination rate between 1\% and 50\% for m-RPCA approaches and Isolation Forest (IF) \cite{liu2008isolation}, k-Nearest Neighbors (KNN) \cite{angiulli2002fast}, Local Outlier Factor (LOF) \cite{breunig2000lof}, Minimum Covariance Deteminant (MCD) \cite{rousseeuw1999fastmcd}, One-Class Support Vector Machines (OCSVM) \cite{scholkopf2001estimating} and PCA \cite{shyu2003novel}.
+We adopt prefixes to denote the adopted approaches for md-RPCA, sd-RPCA and kd-RPCA, which are ss\_ to denote semi-supervised, css\_ to denote contaminated semi supervised and u\_ for unsupervised.
+
+It is possible to observe that LOF, PCA, css\_kd-RPCA and u\_kd-RPCA present lower performance than the remain algorithms, that obtain results higher than 0.95 in average.
+Note that the css\_kd-RPCA and u\_kd-RPCA are the contaminated semi-supervised and unsupervised versions of kd-RPCA, that obtain worse results than the semi-supervised approach of kd-RPCA.
+
+However, the unsupervised versions of md-RPCA and sd-RPCA presented similar results to widely adopted unsupervised algorithms for outlier detection.
+The results also show that the semi-supervised approach of md-RPCA, sd-RPCA and kd-RPCA obtain high anomaly detection rate and presented similar results to the unsupervised algorithms IF, KNN, MCD and OCSVM.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=10cm]{figures/ch4/gaussian_f1_contamination.png}
+	\caption{Anomaly detection on gaussian distributed with uniform anomalies}
+	\label{fig:4.10}
+\end{figure}
+
+The Figures \ref{fig:4.11} and \ref{fig:4.12} show the results for anomaly detection on skewed and heavy tailed distributions, where the Figure \ref{fig:4.11} presents the results for Pareto with gaussian anomalies, and the Figure \ref{fig:4.12} presents the results for Lognormal with gaussian anomalies.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=10cm]{figures/ch4/pareto_f1_contamination.png}
+	\caption{Anomaly detection on Pareto distributed with gaussian anomalies}
+	\label{fig:4.11}
+\end{figure}
+
+The results for anomaly detection on Pareto with gaussian anomalies, depicted by Figure \ref{fig:4.11}, show that IF, KNN, LOF, OCSVM, css\_md-RPCA and u\_md-RPCA performed worse than the remain algorithms, with results lower than 0.6 even with higher contamination.
+Note that css\_md-RPCA and u\_md-RPCA are mean-based approaches of m-RPCA, however the ss\_md-RPCA presented lower results for lower contamination but achieved more than 0.8 with contamination near of 0.2 or higher.
+
+It is possible to observe in Figure \ref{fig:4.11} that all approaches based on kd-RPCA achieve the best results initially, but the results for the unsupervised and contaminated semi-supervised variate with contamination near of 0.2 or higher, while the ss\_kd-RPCA presents stable results near of 1.0 F1-measure for all evaluated contamination rates.
+
+The approaches based on sd-RPCA obtained anomaly detection higher than 0.8, however the unsupervised and contaminated semi-supervised presented high variation of F-measure and lower results in comparison to the semi-supervised sd-RPCA.
+
+The contaminated semi-supervised approaches of sd-RPCA and kd-RPCA presented results higher than 0.8 and similar to the semi-supervised and unsupervised approaches of sd-RPCA and kd-RPCA.
+These results highlight the resilience of the robust subspace learning even for contaminated training data.
+
+The unsupervised approaches of sd-RPCA and kd-RPCA presented more than 0.8 for all contamination rate, showing better results than the widely adopted unsupervised algorithms for outlier detection.
+However, u\_kd-RPCA and u\_sd-RPCA presented lower results than ss\_kd-RPCA and ss\_sd-RPCA.
+
+Therefore, the semi-supervised approaches of m-RPCA overcome other approaches of m-RPCA and of the selected algorithms for anomaly detection on Pareto distributed data with gaussian anomalies.
+It is possible to highlight the semi-supervised kd-RPCA, which obtained stable results near of 1.0 F-measure for all evaluated contamination.
+
+The results for anomaly detection on Lognormal with gaussian anomalies, depicted by Figure \ref{fig:4.12}, show that IF, KNN, LOF, MCD, OCSVM, css\_md-RPCA and u\_md-RPCA perform worse than the remain algorithms, with results lower than 0.6 even with higher contamination.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=10cm]{figures/ch4/lognormal_f1_contamination.png}
+	\caption{Anomaly detection on Lognormal distributed with gaussian anomalies}
+	\label{fig:4.12}
+\end{figure}
+
+The ss\_md-RPCA and PCA algorithms perform similar, with worse detection rate for lower contamination and better performance with contamination higher than 0.4.
+However, the results of ss\_md-RPCA and PCA are lower than all approaches of sd-RPCA and kd-RPCA.
+
+It is important to note that all mean-based approaches of m-RPCA presented lower results for anomaly detection on Lognormal data, in comparison to approaches based on skewness (sd-RPCA) and kurtosis (kd-RPCA), that achieved the best anomaly detection rates.
+However, the approaches of md-RPCA presented better results than IF, KNN, MCD and OCSVM.
+
+The approaches based on kd-RPCA presented higher detection rate for all contamination, with anomaly detection rate near of 1.0.
+The semi-supervised and contaminated semi-supervised approaches for sd-RPCA performed similarly, but the unsupervised approach of sd-RPCA presented lower anomaly detection for lower contamination.
+
+It is possible to note that the contaminated semi-supervised approaches of kd-RPCA and sd-RPCA performed similar to the semi-supervised approach of the same algorithms.
+These results highlight the resilience of the robust subspace learning even from contaminated training data for anomaly detection on Lognormal data.
+
+Therefore, it is possible to observe that the semi-supervised approaches of m-RPCA overcome other approaches of m-RPCA and all the selected algorithms for anomaly detection on Lognormal distributed data with gaussian anomalies.
+
+Following we present the Tables \ref{tab:4.02},  \ref{tab:4.03} and \ref{tab:4.04} to show the results of the selected algorithms and of our proposals for anomaly detection on gaussian, Pareto and Lognormal distributions, with 10\%, 25\% and 35\% of contamination rate.
 
@@
 	\scriptsize
-  \caption{Results grouped by algorithm and scenario}
+	\caption{Results for Simulated data set with 33\% of contamination}
 	\label{tab:4.02}
-  \begin{tabular}{ c|c c|c c|c c|c c|c c|c c }
+	\begin{tabular}{| l | l | l | l |}
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian + Uniform}\\\textbf{(F-Measure)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto + Gaussian}\\\textbf{(F-Measure)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal + Gaussian}\\\textbf{(F-Measure)}\end{tabular}\\ \hline
+            Isolation Forest	&0.98	&0.44	&0.05\\ \hline
+            KNN	&0.98	&0.08	&0.00\\ \hline
+            LOF	&0.35	&0.40	&0.36\\ \hline
+            MCD	&0.98	&0.02	&0.00\\ \hline
+            One-class SVM	&0.98	&0.24	&0.00\\ \hline
+            PCA	&0.87	&0.76	&0.62\\ \hline
+            Semi-Supervised md-RPCA	&0.98	&0.90	&0.66\\ \hline
+            Semi-Supervised sd-RPCA	&0.98	&0.99	&0.97\\ \hline
+            Semi-Supervised kd-RPCA	&0.97	&0.99	&0.98\\ \hline
+            Contaminated Semi-Supervised md-RPCA	&0.98	&0.55	&0.35\\ \hline
+            Contaminated Semi-Supervised sd-RPCA	&0.98	&0.94	&0.96\\ \hline
+            Contaminated Semi-Supervised kd-RPCA	&0.76	&0.99	&0.98\\ \hline
+            Unsupervised md-RPCA	&0.98	&0.54	&0.31\\ \hline
+            Unsupervised sd-RPCA	&0.98	&0.82	&0.93\\ \hline
+            Unsupervised kd-RPCA	&0.74	&0.86	&0.99\\ \hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[h!]
+    \centering
+	\scriptsize
+	\caption{Results for Simulated data set with 25\% of contamination}
+	\label{tab:4.03}
+	\begin{tabular}{| l | l | l | l |}
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian + Uniform}\\\textbf{(F-Measure)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto + Gaussian}\\\textbf{(F-Measure)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal + Gaussian}\\\textbf{(F-Measure)}\end{tabular}\\ \hline
+            Isolation Forest	&0.97	&0.46	&0.05\\ \hline
+            KNN	&0.97	&0.05	&0.00\\ \hline
+            LOF	&0.27	&0.33	&0.30\\ \hline
+            MCD	&0.97	&0.79	&0.00\\ \hline
+            One-class SVM	&0.97	&0.11	&0.00\\ \hline
+            PCA	&0.91	&0.75	&0.56\\ \hline
+            Semi-Supervised md-RPCA	&0.97	&0.85	&0.49\\ \hline
+            Semi-Supervised sd-RPCA	&0.97	&0.99	&0.96\\ \hline
+            Semi-Supervised kd-RPCA	&0.96	&1.00	&0.98\\ \hline
+            Contaminated Semi-Supervised md-RPCA	&0.97	&0.54	&0.17\\ \hline
+            Contaminated Semi-Supervised sd-RPCA	&0.97	&0.91	&0.95\\ \hline
+            Contaminated Semi-Supervised kd-RPCA	&0.63	&1.00	&0.97\\ \hline
+            Unsupervised md-RPCA	&0.97	&0.53	&0.14\\ \hline
+            Unsupervised sd-RPCA	&0.97	&0.88	&0.91\\ \hline
+            Unsupervised kd-RPCA	&0.62	&0.97	&0.99\\ \hline
+	\end{tabular}
+\end{table}
+
+\begin{table}[h!]
+    \centering
+	\scriptsize
+	\caption{Results for Simulated data set with 10\% of contamination}
+	\label{tab:4.04}
+	\begin{tabular}{| l | l | l | l |}
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian + Uniform}\\\textbf{(F-Measure)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto + Gaussian}\\\textbf{(F-Measure)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal + Gaussian}\\\textbf{(F-Measure)}\end{tabular}\\ \hline
+            Isolation Forest	&0.97	&0.43	&0.00\\ \hline
+            KNN	&0.97	&0.01	&0.00\\ \hline
+            LOF	&0.29	&0.27	&0.24\\ \hline
+            MCD	&0.97	&0.40	&0.00\\ \hline
+            One-class SVM	&0.97	&0.26	&0.00\\ \hline
+            PCA	&0.92	&0.65	&0.06\\ \hline
+            Semi-Supervised md-RPCA	&0.97	&0.59	&0.00\\ \hline
+            Semi-Supervised sd-RPCA	&0.97	&0.98	&0.84\\ \hline
+            Semi-Supervised kd-RPCA	&0.97	&1.00	&0.95\\ \hline
+            Contaminated Semi-Supervised md-RPCA	&0.97	&0.28	&0.00\\ \hline
+            Contaminated Semi-Supervised sd-RPCA	&0.96	&0.90	&0.82\\ \hline
+            Contaminated Semi-Supervised kd-RPCA	&0.48	&1.00	&0.94\\ \hline
+            Unsupervised md-RPCA	&0.97	&0.31	&0.00\\ \hline
+            Unsupervised sd-RPCA	&0.97	&0.93	&0.70\\ \hline
+            Unsupervised kd-RPCA	&0.50	&0.97	&0.98\\ \hline
+	\end{tabular}
+\end{table}
+
+Taking into account the null hypothesis $H_{0rob.sub}$ and the presented results, we can conclude that the robust subspace learning, adopted by md-RPCA, sd-RPCA and kd-RPCA, presented higher anomaly detection from imbalanced and skewed data than widely adopted algorithms for anomaly detection.
+Therefore, the presented results refute the null hypothesis $H_{0rob.sub}$, which defines that the robust subspace learning does not improves the anomaly detection from imbalanced and skewed data.
+
+\section{Results of CTU-13 Experiment}
+\label{sec:4_ctu13_result}
+
+In this section we present the experiment on network anomaly detection from the CTU-13 data set, evaluating the results of md-RPCA, sd-RPCA, kd-RPCA, Isolation Forest (IF) \cite{liu2008isolation}, k-Nearest Neighbors (KNN) \cite{angiulli2002fast}, Local Outlier Factor (LOF) \cite{breunig2000lof}, Minimum Covariance Deteminant (MCD) \cite{rousseeuw1999fastmcd}, One-Class Support Vector Machines (OCSVM) \cite{scholkopf2001estimating} and PCA \cite{shyu2003novel}.
+
+For this experiment we only evaluate the semi-supervised approach of md-RPCA, sd-RPCA and kd-RPCA, considering that the semi-supervised algorithms presented the best results on the simulated experiment and taking into account the observed resilience of the semi-supervised approach when the training data is contaminated.
+Additionally, the semi-supervised approach only requires the robust subspace learning for training, what can indicate less computational cost for network anomaly detection on new observations.
+
+The CTU-13 data set is very imbalanced, with contamination rate between 0.06\% and 8.86\%.
+Therefore we adopted a uniform contamination rate of 33\% for this experiment, considering that the simulated experiment showed better results for contamination higher than 30\% for m-RPCA and for the selected anomaly detection algorithms.
+
+The Table \ref{tab:4.03} present the F-measure of each algorithm for all scenarios of CTU-13, and the last column presents the average F-measure of each algorithm for all scenarios.
+
+\begin{table}[h!]
+  \centering
+  \scriptsize
+  \caption{Network anomaly detection from CTU-13 with 33\% of contamination}
+  \label{tab:4.03}
+  \begin{tabular}{ c|c|c|c|c|c|c|c|c|c|c|c|c|c|c }
 	\toprule
-	\multirow{2}{*}{\textbf{Scenario}}   &\multicolumn{2}{c}{\textbf{GMM}} &\multicolumn{2}{c}{\textbf{K-Means}} &\multicolumn{2}{c}{\textbf{ROBPCA}} &\multicolumn{2}{c}{\textbf{md-RPCA}} &\multicolumn{2}{c}{\textbf{kd-RPCA}} &\multicolumn{2}{c}{\textbf{sd-RPCA}}\\
-			\hhline{~------------}
-			&\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25} &\textbf{0.15} &\textbf{0.25}\\
-	\midrule
-		10 & 0.34 & 0.45 & 0.20 & 0.25 & 0.29 & 0.34 & 0.97 & 0.96 & 0.99 & 0.99 & 0.99 & 0.99 \\ \hline
-		11 & 0.34 & 0.47 & 0.12 & 0.14 & 0.27 & 0.37 & 0.98 & 0.96 & 0.99 & 0.99 & 0.99 & 0.99 \\ \hline
-		12 & 0.00 & 0.00 & 0.09 & 0.11 & 0.08 & 0.11 & 0.86 & 0.80 & 0.00 & 0.00 & 0.94 & 0.90 \\ \hline
-		15 & 0.14 & 0.21 & 0.08 & 0.11 & 0.05 & 0.07 & 0.85 & 0.79 & 1,00 & 0.96 & 0.99 & 0.96 \\ \hline
-		15-2 & 0.32 & 0.14 & 0.18 & 0.25 & 0.21 & 0.30 & 0.85 & 0.82 & 0.98 & 0.99 & 0.98 & 0.99 \\ \hline
-		15-3 & 0.00 & 0.01 & 0.22 & 0.28 & 0.28 & 0.38 & 0.95 & 0.96 & 0.00 & 0.00 & 0.99 & 0.98 \\ \hline
-		16 & 0.50 & 0.48 & 0.06 & 0.16 & 0.06 & 0.10 & 0.85 & 0.83 & 0.00 & 0.00 & 0.96 & 0.00 \\ \hline
-		16-2 & 0.02 & 0.08 & 0.03 & 0.02 & 0.00 & 0.00 & 0.22 & 0.26 & 0.80 & 1,00 & 0.80 & 1,00 \\ \hline
-		16-3 & 0.00 & 0.01 & 0.02 & 0.04 & 0.01 & 0.03 & 0.60 & 0.66 & 0.00 & 0.00 & 0.84 & 0.98 \\ \hline
-		17 & 0.16 & 0.46 & 0.35 & 0.38 & 0.72 & 0.75 & 0.81 & 0.78 & 0.02 & 0.78 & 0.81 & 0.78 \\ \hline
-		18 & 0.21 & 0.17 & 0.13 & 0.19 & 0.13 & 0.15 & 0.91 & 0.88 & 0.97 & 0.98 & 0.97 & 0.98 \\ \hline
-		18-2 & 0.50 & 0.54 & 0.61 & 0.70 & 0.27 & 0.24 & 0.80 & 0.85 & 0.99 & 0.98 & 0.99 & 0.98 \\ \hline
-		19 & 0.16 & 0.20 & 0.19 & 0.17 & 0.10 & 0.13 & 0.91 & 0.91 & 0.96 & 0.99 & 0.96 & 0.99 \\ \hline
-		\rowcolor{Gray} \textbf{avg} &\textbf{0.21} &\textbf{0.25} &\textbf{0.18} & \textbf{0.21} & \textbf{0.19} &\textbf{0.23} &\textbf{0.81} &\textbf{0.80} &\textbf{0.59} &\textbf{0.67} &\textbf{0.94} &\textbf{0.89} \\
+        \textbf{Algorithm}	&\textbf{10}	&\textbf{11}	&\textbf{12}	&\textbf{15}	&\textbf{15-2}	&\textbf{15-3}	&\textbf{16}	&\textbf{16-2}	&\textbf{16-3}	&\textbf{17}	&\textbf{18}	&\textbf{18-2}	&\textbf{19}	&\textbf{Avg} \\ \hline
+        IF	&0.36	&0.34	&0.09	&0.21	&0.40	&0.44	&0.16	&0.34	&0.34	&0.41	&0.12	&0.16	&0.46	&0.29\\ \hline
+        KNN	&0.05	&0.17	&0.01	&0.03	&0.33	&0.23	&0.01	&0.25	&0.03	&0.12	&0.00	&0.00	&0.24	&0.11\\ \hline
+        LOF	&0.15	&0.14	&0.13	&0.17	&0.29	&0.22	&0.29	&0.38	&0.25	&0.24	&0.00	&0.04	&0.38	&0.21\\ \hline
+        MCD	&0.18	&0.29	&0.09	&0.34	&0.79	&0.62	&0.04	&0.58	&0.20	&0.41	&0.20	&0.20	&0.36	&0.33\\ \hline
+        PCA	&0.33	&0.64	&0.69	&0.65	&0.55	&0.62	&0.75	&0.50	&0.77	&0.33	&0.82	&0.01	&0.61	&0.56\\ \hline
+        md-RPCA &0.83	&0.79	&0.95	&0.78	&0.78	&0.87	&0.95	&0.87	&0.80	&0.82	&0.83	&0.82	&0.51	&0.81\\ \hline
+        sd-RPCA &0.25	&0.75	&0.34	&0.64	&0.50	&0.75	&0.86	&0.50	&0.77	&0.33	&0.82	&0.81	&0.21	&0.57\\ \hline
+        kd-RPCA &0.76	&0.76	&0.90	&0.82	&0.57	&0.76	&0.91	&0.50	&0.80	&0.73	&0.83	&0.81	&0.48	&0.74\\ \hline
     \bottomrule
@@
 
-It is possible to observe that the results of md-RPCA, kd-RPCA and sd-RPCA overcome the remain methods in a comparison by scenario and by average F-measure.
+It is possible to observe that md-RPCA, kd-RPCA and sd-RPCA overcome the remain algorithms in average results for all scenarios, according to the column \textbf{avg}.
+The sd-PRCA presented an average of 0.57 while md-RPCA obtained 0.81 and kd-RPCA 0.74 in average.
+The PCA algorithm performed similar to sd-RPCA in average, with results of 0.56 and 0.57, respectively.
+However the results of PCA and sd-RPCA for scenarios 12, 18-2 and 19 presented a large variation.
 
-% 10: moment are better
-% 11: moment are better
-% 12: kd-RPCA is poor for scenario 12, as well as GMM, K-Means and ROBPCA
-% 15: moment are better
-% 15-2: moment are better
-% 15-3: kd-RPCA is poor for , whilhe ROBPCA obtain 0.38 and other moment based has good results
-% 16: kd-RPCA is poor for , as well as sd-RPCA with 0.25s, while GMM obtain 0.50 and other moment based has good results. Explain based on data
-% 16-2: md-RPCA is poor for , as well as GMM, K-means and ROBPCA, but kd-RPCA and sd-RPCA goes well
-% 16-3: kd_RPCA is poor, as well as GMM, K-means and ROBPCA
-% 17: robpca has high result, higher than kd-RPCA, GMM, K-means, but md and sd are better and equal
-% 18: moment based is better
-% 18-2: high GMM e Kmeans, but moment are even better
-% 19: moment are better
-% there are no lots of researches using CTU-13, although it is been growing poor results for GMM, K-Means and MGM are ok and in accordance to \cite{wang2017botnet}, which also evaluate \cite{gu2007bothunter}
+The md-RPCA algorithm presented an anomaly detection rate higher than 0.78 for almost all scenarios, with exception to scenario 19, where the anomaly detection rate of md-RPCA was 0.51.
+The anomalies of the scenario 19 are peer-to-peer botnet traffic generated by nsys.ay, which are related to botnet synchronization and not to network attacks, what can explain the low network detection rate of all evaluated algorithms, where the largest F1-measure was 0.61 achieved by PCA.
 
-% a questão é diferenciar aquilo que é malicioso ou não, dentre os outliers ou novelties (esses termos são sinônimos para alguns autores)
-% como eu posso diferenciar isto? Eu suponho que a partir de um modelo de tráfego normal eu possa detectar grandes deviations e classificar estas como ataques
-% a hipotese é esta
-% em outra área (compressive sensing), se utiliza dicionários, que são representações compactas de todas as caracteristicas de um dado
-% no nosso caso, o treinamento nos gera uma especie de dicionário, em que os dados normais são representados de forma robusta
-% onde uma forma robusta seria uma forma menos sensível a outliers.
-Isso pode ser visto como a diferenca entre calcular a média e a mediana de um conjunto.
-A presença de um outlier no conjunto irá corromper a média de uma forma maior que a mediana.
-Desta forma, se considera a mediana como sendo robusta
+The md-RPCA showed the best anomaly detection for 10 scenarios of a total with 13 scenarios.
+In the scenario 15 the best result was obtained by kd-RPCA, for the 15-2 scenario the best result was for MCD, and PCA was the best algorithm for anomaly detection on scenario 19.
+Even thought md-RPCA not be the best result for scenarios 15, 15-2 and 19, the results of md-RPCA are very close to the best results for scenarios 5 and 15-2.
 
-\section{Results}
-\label{sec:4_results}
+It is important to note that IF, KNN, LOF and MCD algorithms performed poorly for unsupervised network anomaly detection on all scenarios of the CTU-13 data set, with average of 0.29, 0.11, 0.21 and 0.33 respectively.
+From these algorithms, only MCD presented high result for one scenario, which was 0.79 for 15-2.
+
+Unfortunately, was not possible to observe Pareto or Lognormal distributions on features of CTU-13, even though the finds of researches that highlight the fitting between these distributions and Internet traffic \cite{benson2010network, leon2017probability}.
+The CTU-13 data set is very challenging for anomaly detection approaches, due to the high imbalance and large volume of flows.
+However, CTU-13 is one of the up-to-date data sets for network attack detection and is one that provides the data imbalance observed in real anomaly detection problems.
+
+The results of anomaly detection on CTU-13 reveals that m-RPCA algorithms obtain good results for network anomaly detection on contaminated data, overcoming widely adopted algorithms for outlier detection.
 
 \section{Conclusion}
-\label{sec:4_conclusion}
+\label{sec:6_conclusion}
 
-This paper models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
+This work proposed the m-RPCA, which is approaches based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data.
+We evaluated the anomaly detection rate of m-RPCA on simulated data and on the CTU-13 data set, which is composed by network traffic of botnets, attacks and background flows.
 
-The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting F-measure of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
+The m-RPCA can be divided into md-RPCA, sd-RPCA and kd-RPCA algorithms, to denote approaches of m-RPCA based distances of mean, skewness and kurtosis, respectively.
+We also proposed to evaluate m-RPCA for semi-supervised, contaminated semi-supervised and unsupervised methods of anomaly detection.
 
-This paper evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
-Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.
+The experimental results show that moment distances computed from robust subspace can improve the anomaly detection on skewed and imbalanced data set.
+The results also show that the m-RPCA can be adopted for network attack detection, with better results than widely adopted algorithms for outlier detection on the CTU-13 data set.
+Therefore, the observed results rejects the null hypothesis  $H_{0rob.sub}$ and confirms the alternative
+hypothesis $H_{1rob.sub}$.
 
-The incremental individualized approach of eigen similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to eigen similarity analysis for network attack detection.
-Therefore, the incremental individualized approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.
+The results shows that the semi-supervised approach for m-RPCA obtained better results than the contaminated semi-supervised and unsupervised approaches, and that m-RPCA presented good results even when the semi-supervised training was computed from contaminated data.
+This highlights the resilience of the robust subspace learning for the semi-supervised anomaly detection approach to deal with possible lack of known normal data for training.
 
-According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
+The main contributions of this work were: the development and evaluation of novel approaches for anomaly detection on skewed and imbalanced data sets, by means of moments computed from a robust subspace learned by RPCA.
 
-Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets.
-Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring02       rch can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.
-
-Test distribution fit again e present results 
-Conclude this presentation.
-Evaluate eigensim for CTU-13 and others algorithms for Danilo's data set.
-Compare to vanilla RPCA (Eduardo is working on it).
-Write a paper for JNCA (April).
-Write the thesis (May and June).
-
-say that more studies about skewness and subspace learning can be done in order to explore it inside algorithms and non gaussian data
-
-note that the semi supervised approach requires subspace learning only for training and not for testing, while unsupervised requires subspace learning for each evaluations
+Future research can be directed to evaluate the application of the proposed approaches to different data sets and field of anomaly detection, as well as for the extension of m-RPCA to estimate the contamination or to not rely on previously defined contamination.

### 2019-06-09 docs/thesis/chapters/5_conclusion_future_work/conclusion.tex
results of robpca

@@
 The proposed technique is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis, where MOS and eigenvalue analysis are applied to detect time frames under attack.
-In addition, we evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset.
+In addition, we evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 data set.
 
-This thesis also proposes a novel architecture and approach based on user behavior analysis through Model Order Selection (MOS), in order to detect possible threats, and to reduce the risks and the harm of the most common threats.
-The behavioral analysis can indicate well known malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.
+This thesis also proposed a prove of concept of an architecture to evaluate the user behavior analysis on offline mobile clients through Eigensim and Model Order Selection (MOS), in order to implement an anomaly detection module based on user behavioral analysis.
 
-Finally, we proposed a fraud detection approach composed by tensor-based dictionary learning for discriminant analysis of a mobile payment dataset, where the problem of imbalanced data is discussed and it is presented one analysis regarding the classification accuracy of the proposed approach against well known classifiers.
+Finally, we proposed the m-RPCA, which is approaches for anomaly detection on skewed and imbalanced data set, through distances of moments computed from a robust subspace learned by RPCA.
+The m-RPCA was evaluated for anomaly detection on simulated data and on the CTU-13 data set, which is composed by network traffic of botnets, network attacks and background flows.
 
@@
 
-This thesis models the network traffic as a signal processing formulation for applying the framework for detection and identification of network attacks, which is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
+In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time.
+Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
+Recently, signal processing schemes have being applied to detect malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest, vieira2017model}, showing advances in network traffic analysis.
 
-The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
+This thesis models the network traffic as a signal processing formulation for applying the Eigensim, which is a framework for detection and identification of network attacks, that is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
 
-This thesis evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
+The Eigensim was evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and similarity analysis of learned subspace of eigenvalues and eigenvectors.
+The main contributions of Eigensim were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
+
+This thesis evaluated the effectiveness of MOS schemes for network attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
 Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.
 
-The incremental individualized approach of eigen similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to eigen similarity analysis for network attack detection.
-Therefore, the incremental individualized approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.
+Considering the offline mobile security context, an important issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile clients.
+Aware of this problem and its importance, We proposed an architecture based on Eigensim for behavioral anomaly detection in offline corporate mobile apps.
+As prove of concept, a fully working mobile application was developed to test the proposed security solution and acquired results provide evidence that besides achieving the desired security features, the solution also has positive results in terms of performance.
 
-According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
+Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.
 
-Considering the offline mobile security context, an important issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile clients.
-Once a mobile client releases the connection with the corporate cloud, no security measure implemented in the cloud infrastructure assures the protection of sensitive data stored in the mobile device.
-Aware of this problem and its importance, this thesis presented a proposal to address the offline mobile security problem combining a different cryptographic methods.
-Moreover, this proposed approach also prevents malicious user behavior by applying a MOS-based analytic method.
-
-As prove of concept, a fully working mobile application was developed to test the proposed security solution and acquired results provide evidence that besides achieving the desired security features, the solution also has positive results in terms of performance.
-This fact is due to the usage of lightweight operations and the optimized combination of the selected security methods.
-The proposed approach is a practical application to be used in the corporate mobile environment.
-It is implemented as a fully working mobile client and can be used for any type of enterprise.
-Also, part of concept is seed for new security solutions for big data applications.
-
-Finally, this thesis shows that a tensor-based approach can be used for fraud detection from mobile money transactions, with average results when compared to well known classifiers in one scenario with highly skewed datasets.
+This thesis has proposed and evaluated the m-RPCA, which can be divided into md-RPCA, sd-RPCA and kd-RPCA, that are algorithms based on moment distances from robust subspace for anomaly detection.
+These proposed algorithms were evaluated for semi-supervised, contaminated semi-supervised and unsupervised approaches.
+The experimental results show that moment distances computed fro robust subspace can improve the anomaly detection on skewed and imbalanced data set.
+The results also show that the m-RPCA can be adopted for network attack detection, with better results than widely adopted algorithms for outlier detection.
 
@@
 
-We analyze problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques.
+We analyze problems related to detection of anomalies and information security issues, and propose new approaches to improve malicious behavior detection through signal processing techniques and subspace learning.
 The results of the work presented in this thesis provide the following contributions:
@@
 	\item We discussed the processing time of the proposed framework for mobile devices;
-	\item We proposed a tensor-based dictionary learning approach for fraud detection in mobile payment transactions;
+	\item We proposed a approaches for anomaly detection on skewed and imbalanced data set, through distances of moments computed from a robust subspace learned by RPCA;
 	\item We published the following papers reporting our results:
@@
 \begin{itemize}
-	\item Improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel datasets;
+	\item Improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets;
 	\item Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high throughput network traffic;
-	\item Evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis;
-	\item Explore enhancements in the analytics methods as well as to extended the approach to be used by mobile devices with even more severe resources constraints;
+	\item Evaluate the application of the proposed approach to different anomaly detection problems, considering cases that are aware to behavioral analysis;
+	\item Enhance the m-RPCA approaches to estimate the contamination or not rely on previously defined contamination for anomaly detection;
 \end{itemize}

### 2019-06-14 docs/thesis/chapters/1_introduction/introduction.tex
after fix bold notation

@@
     \begin{tabular}{|p{6cm}|p{6cm}|c|} \hline
-        \textbf{Alternative Hypothesis}	&\textbf{Null Hypothesis}	&\textbf{Question} 	\\ \hline
+        \pmb{Alternative Hypothesis}	&\pmb{Null Hypothesis}	&\pmb{Question} 	\\ \hline
         $H_{1sub}$: It is possible to use subspace learning to identify and detect anomalies in network traffic.
@@
 \subsection{Proposals}
-\label{sc:problem_statement}
+\label{sc:proposals}
 

### 2019-06-14 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
after fix bold notation

@@
 
-In this work, scalars are denoted by italic letters ($a, b, A, B, \alpha, \beta$), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
+In this work, scalars are denoted by italic letters ($a, b, A, B, \alpha, \beta$), vectors by lowercase bold letters ($\pmb{a}, \pmb{b}$), matrices by uppercase bold letters ($\pmb{A}, \pmb{B}$), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix $\pmb{A}$.
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
@@
      \centering 
-     \includegraphics[width=11cm]{figures/fig09.png}
+     \includegraphics[width=11cm]{figures/ch2/fig09.png}
      \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
@@
 
-By modeling the data set as a signal superposition, the network traffic (\textbf{X}) can be characterized as a mixture of three components: legitimate traffic (\textbf{U}), artifact (\textbf{A}) and malicious traffic (\textbf{N}), according to the following expression:
+By modeling the data set as a signal superposition, the network traffic ($\pmb{X}$) can be characterized as a mixture of three components: legitimate traffic ($\pmb{U}$), artifact ($\pmb{A}$) and malicious traffic ($\pmb{N}$), according to the following expression:
 
 \begin{equation}\label{eq:2.01}
-	\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{A}^{(q)} + \boldsymbol{N}^{(q)},
+	\pmb{X}^{(q)} = \pmb{U}^{(q)} + \pmb{A}^{(q)} + \pmb{N}^{(q)},
 \end{equation}
 where $q$ represents the $q$-th time frame, which is a time aggregation of network traffic.
-The matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns, where each row represents a communication port, and each column represents time bins of a defined size, such as one minute.
+The matrix $\pmb{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} rows and \emph{N} columns, where each row represents a communication port, and each column represents time bins of a defined size, such as one minute.
 Each element $x_{m,n}^{(q)}$ stands for the number packets that appears at $n$-th minute for the port $m$, during the $q$-th time frame.
 
-The legitimate traffic $\boldsymbol{U}^{(q)}$ is characterized by the traffic from user's ordinary operations, such as when performing access to web pages, as well when there is the traffic required to domain name resolution.
+The legitimate traffic $\pmb{U}^{(q)}$ is characterized by the traffic from user's ordinary operations, such as when performing access to web pages, as well when there is the traffic required to domain name resolution.
 Figure \ref{fig:2.03} depicts an example of the legitimate traffic simulated during the experiments.
@@
      \centering 
-     \includegraphics[width=11cm]{figures/fig03.png}
+     \includegraphics[width=11cm]{figures/ch2/fig03.png}
      \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
@@
 
-The traffic that is not associated with user's operations and with malicious traffic is modeled as artifact $\boldsymbol{A}^{(q)}$.
+The traffic that is not associated with user's operations and with malicious traffic is modeled as artifact $\pmb{A}^{(q)}$.
 The acquisition of logical IP network address (DHCP) is an example of artifact, where independently of any user operation, the machine receives an IP address, since it is configured to automatically perform a DHCP address request.
@@
 The traffic coming from a malicious activity, i.e.
-port scanning, synflood or fraggle attacks, is represented by the matrix $\boldsymbol{N}^{(q)}$.
+port scanning, synflood or fraggle attacks, is represented by the matrix $\pmb{N}^{(q)}$.
 
@@
      \centering 
-     \includegraphics[width=11cm]{figures/fig04.png}
+     \includegraphics[width=11cm]{figures/ch2/fig04.png}
      \caption{Network traffic of user independent operations for network management.}
@@
 
-We define that if the $\#\boldsymbol{X}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\boldsymbol{X}^{(q)} = 0$, then there is no malicious traffic.
-We show how to detect the $\#\boldsymbol{X}^{(q)}$, given only the matrix $\boldsymbol{X}^{(q)}$, in order to identify malicious traffic.
+We define that if the $\#\pmb{X}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\pmb{X}^{(q)} = 0$, then there is no malicious traffic.
+We show how to detect the $\#\pmb{X}^{(q)}$, given only the matrix $\pmb{X}^{(q)}$, in order to identify malicious traffic.
 
@@
      \centering 
-     \includegraphics[width=11cm]{figures/fig05.png}
+     \includegraphics[width=11cm]{figures/ch2/fig05.png}
      \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
@@
      \centering 
-     \includegraphics[width=11cm]{figures/fig06.png}
+     \includegraphics[width=11cm]{figures/ch2/fig06.png}
      \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
@@
      \centering 
-     \includegraphics[width=11cm]{figures/fig07.png}
+     \includegraphics[width=11cm]{figures/ch2/fig07.png}
      \caption{Connection attempts in order to identify active ports.}
@@
 	\centering
-     \includegraphics[width=11cm]{figures/mos_eigen_similarity.eps}
+     \includegraphics[width=11cm]{figures/ch2/mos_eigen_similarity.eps}
      \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
@@
 
-With the data grouped into $Q$ time frames, the framework considers the time variations of the matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the attack.
+With the data grouped into $Q$ time frames, the framework considers the time variations of the matrix $\pmb{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the attack.
 
@@
 
-Therefore, to detect flood attacks, it is necessary to calculate the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the zero mean samples given by
+Therefore, to detect flood attacks, it is necessary to calculate the sample covariance matrix $\hat{\pmb{R}}_{yy}^{(q)}$ of the zero mean samples given by
 
@@
 \label{eq:2.02}
-\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
-\end{equation}
-
-The set of obtained vectors $\boldsymbol{y}_{m}^{(q)}$ composes the zero mean matrix $\boldsymbol{Y}^{(q)}$, then the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
+\pmb{y}_{m}^{(q)} = \pmb{x}_{m}^{(q)} - \bar{\pmb{x}}_{m}^{(q)}.
+\end{equation}
+
+The set of obtained vectors $\pmb{y}_{m}^{(q)}$ composes the zero mean matrix $\pmb{Y}^{(q)}$, then the sample covariance matrix $\hat{\pmb{R}}_{yy}^{(q)}$ can be calculated as follows
 
 \begin{equation}\label{eq:2.03}
-\boldsymbol{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
+\hat{\pmb{R}}_{yy}^{(q)} = \frac{1}{N}\pmb{Y}^{(q)}\pmb{Y}^{(q)^{\rm T}}.
 \end{equation}
@@
 Moreover, the port scan traffic presents a highly correlated network traffic between the monitored ports.
-In order to exploit such structure, we compute the sample covariance $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have zero mean and unitary standard deviation as follows
+In order to exploit such structure, we compute the sample covariance $\hat{\pmb{R}}_{zz}^{(q)}$ whose variables have zero mean and unitary standard deviation as follows
 
@@
 \label{eq:2.04}
-\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
-\end{equation}
-
-The set of vectors $\boldsymbol{z}_{m}^{(q)}$ composes the matrix $\boldsymbol{Z}^{(q)}$, then the sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)}$ can be calculated via
+\pmb{z}_{m}^{(q)} = \frac{\pmb{x}_{m}^{(q)} - \bar{\pmb{x}}_{m}^{(q)}}{\pmb{\sigma}_{m}^{(q)}}.
+\end{equation}
+
+The set of vectors $\pmb{z}_{m}^{(q)}$ composes the matrix $\pmb{Z}^{(q)}$, then the sample covariance matrix $\hat{\pmb{R}}_{zz}^{(q)}$ can be calculated via
 
 \begin{equation}\label{eq:2.05}
-\boldsymbol{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
-\end{equation}
-
-Once the $\boldsymbol{\hat{R}}_{yy}^{(q)}$ and $\boldsymbol{\hat{R}}_{zz}^{(q)}$ have been obtained for flood and port scan detection, respectively, and since the next steps are the same for both sample covariance matrices, we refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{\hat{R}}$.
-Therefore, the following step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:2.06}), in order to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix, according to (\ref{eq:2.06}).
+\hat{\pmb{R}}_{zz}^{(q)} = \frac{1}{N}\pmb{Z}^{(q)}\pmb{Z}^{(q)^{\rm T}}.
+\end{equation}
+
+Once the $\hat{\pmb{R}}_{yy}^{(q)}$ and $\hat{\pmb{R}}_{zz}^{(q)}$ have been obtained for flood and port scan detection, respectively, and since the next steps are the same for both sample covariance matrices, we refer to $\hat{\pmb{R}}_{yy}$ and $\hat{\pmb{R}}_{zz}$ as a matrix $\hat{\pmb{R}}$.
+Therefore, the following step of the algorithm is the eigenvalue decomposition (EVD), calculated according to (\ref{eq:2.06}), in order to obtain the vector of eigenvalues $\pmb{e}^{(q)}$ associated with each matrix, according to (\ref{eq:2.06}).
 
 \begin{equation}\label{eq:2.06}
-\boldsymbol{\hat{R}}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
+\hat{\pmb{R}}^{(q)} = \pmb{V}^{(q)}\pmb{\Lambda}^{(q)}\pmb{V}^{(q)^{\rm T}},
 \end{equation}
@@
 \begin{equation}\label{eq:2.060}
-\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
+\pmb{e}^{(q)} = \rm diag(\pmb{\Lambda}^{(q)}),
 \end{equation}
@@
 
-The concatenation of the eigenvalues vector $\boldsymbol{e}^{(q)}$ for $q = 1, \ldots, Q$ is represented by
+The concatenation of the eigenvalues vector $\pmb{e}^{(q)}$ for $q = 1, \ldots, Q$ is represented by
 
 \begin{equation}\label{eq:2.07}
-\boldsymbol{E} =
+\pmb{E} =
 \begin{bmatrix}
@@
 
-Note that since $\lambda_1^{(q)} > \lambda_2^{(q)} > \lambda_3^{(q)} > \cdots > \lambda_{m-1}^{(q)} > \lambda_m^{(q)}$, then the first line of the matrix $\boldsymbol{E}$ contains the largest eigenvalues of each $q$-th time frame, which is the Greatest Eigenvalue Time
+Note that since $\lambda_1^{(q)} > \lambda_2^{(q)} > \lambda_3^{(q)} > \cdots > \lambda_{m-1}^{(q)} > \lambda_m^{(q)}$, then the first line of the matrix $\pmb{E}$ contains the largest eigenvalues of each $q$-th time frame, which is the Greatest Eigenvalue Time
 Vector (GETV) \cite{tenorio2013greatest}, denoted as
@@
 \begin{equation}\label{eq:2.08}
-\boldsymbol{e}_{\rm max} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
+\pmb{e}_{\rm max} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
 \end{equation}
@@
 
-Traditionally the MOS schemes are applied for the eigenvalues of the vector $\boldsymbol{e}^{(q)}$.
+Traditionally the MOS schemes are applied for the eigenvalues of the vector $\pmb{e}^{(q)}$.
 However, the goal here is to detect the variations of the eigenvalues for different values of $q$.
 Therefore, instead of using a certain $q$, the proposed approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame, in order to identify variations and estimate the model order $\hat{d}$, which is the estimated number of time frames under attack.
-Therefore, $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
-Note that some MOS schemes may also require the number of minutes that compose a time frame, as $\hat{d} = \rm{MOS}(\boldsymbol{e}_{\rm max},\emph{Q})$.
+Therefore, $\pmb{e}_{\rm max}$ is sorted in descending order, producing $\sim\pmb{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\pmb{e}_{\rm max})$.
+Note that some MOS schemes may also require the number of minutes that compose a time frame, as $\hat{d} = \rm{MOS}(\pmb{e}_{\rm max},\emph{Q})$.
 For more information about MOS, we refer to the Appendix \ref{apx:a_mos}.
@@
 
-After applying the MOS schemes to the vector $\sim\boldsymbol{e}_{\rm max}$, we obtain the estimate of the $\#\boldsymbol{X}$.
-For instance, in the case of fraggle, synflood and ports can, if $\hat{d} = 1$, , then $\#\boldsymbol{X} = 1$, which means that during the during the $Q$ time frames one attack is present.
-However, if $\hat{d} = 0$, then $\#\boldsymbol{X} = 0$, and this means that none of these attacks are present.
+After applying the MOS schemes to the vector $\sim\pmb{e}_{\rm max}$, we obtain the estimate of the $\#\pmb{X}$.
+For instance, in the case of fraggle, synflood and ports can, if $\hat{d} = 1$, , then $\#\pmb{X} = 1$, which means that during the during the $Q$ time frames one attack is present.
+However, if $\hat{d} = 0$, then $\#\pmb{X} = 0$, and this means that none of these attacks are present.
 Note that $\hat{d}$ can be greater than 1, indicating the presence of more than one attack.
@@
 
-The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed according to Algorithm \ref{alg:2.01}, where $\rm{\boldsymbol{\hat{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\boldsymbol{e}_{\rm max}$.
-Algorithm \ref{alg:2.01} initially identifies the largest value of $\boldsymbol{e}_{\rm max}$, according to Line 3 of Algorithm \ref{alg:2.01}, and its correspondent index, according to Line 7.
-Subsequently, the largest value is removed of $\boldsymbol{e}_{\rm max}$, according to Line 11 of Algorithm \ref{alg:2.01}, and a new iteration is performed until $\boldsymbol{e}_{\rm max} = []$.
+The largest eigenvalue analysis for estimating the $q$-th time frames that are under attack can be expressed according to Algorithm \ref{alg:2.01}, where $\rm{\hat{\pmb{q}}}_{\rm max} \in \mathbb{R}^{\hat{d}}$ denotes a vector of the $q$-th time frames under attack, which is the $q$-th indexes corresponding to the $\hat{d}$ largest eigenvalues of $\pmb{e}_{\rm max}$.
+Algorithm \ref{alg:2.01} initially identifies the largest value of $\pmb{e}_{\rm max}$, according to Line 3 of Algorithm \ref{alg:2.01}, and its correspondent index, according to Line 7.
+Subsequently, the largest value is removed of $\pmb{e}_{\rm max}$, according to Line 11 of Algorithm \ref{alg:2.01}, and a new iteration is performed until $\pmb{e}_{\rm max} = []$.
 
@@
 	\SetAlgoLined
-	\KwResult{$\rm{\boldsymbol{\hat{q}}}_{\rm max}$}
+	\KwResult{$\rm{\hat{\pmb{q}}}_{\rm max}$}
 	Given $f = 1$\;
 	\While{$f < \hat{d}$}{
-		$q_{\rm value} = \operatorname*{argmax}_{\lambda}  \hspace{1 mm} \boldsymbol{e}_{\rm max}$\;
+		$q_{\rm value} = \operatorname*{argmax}_{\lambda}  \hspace{1 mm} \pmb{e}_{\rm max}$\;
 		$i = 1$\;
     	\While{$i < Q$}{
-    		\If{$\boldsymbol{e}_{\rm max}^{(i)} == q_{\rm value}$}{
-    			$\boldsymbol{\hat{q}}_{\rm max}^{(f)} = i$\;
+    		\If{$\pmb{e}_{\rm max}^{(i)} == q_{\rm value}$}{
+    			$\hat{\pmb{q}}_{\rm max}^{(f)} = i$\;
     		}
@@
     	}
-    	$\boldsymbol{e}_{\rm max} \rightarrow \boldsymbol{\hat{q}}_{\rm max}^{(f)}$\;
+    	$\pmb{e}_{\rm max} \rightarrow \hat{\pmb{q}}_{\rm max}^{(f)}$\;
 		$f = f + 1$\;
@@
 
-After the estimation of the $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected Attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked.
+After the estimation of the $\rm{\hat{\pmb{q}}}_{\rm max}$ time frames under attack, it is necessary to obtain more details of the detected Attacks, such as the $n$-th minutes when the attacks happened and the $m$-th network ports that were attacked.
 To deal with this problem, the adoption of a similarity analysis between legitimate traffic and the traffic of time frames estimated as under attack is evaluated, analyzing the effectiveness of cosine similarity to highlight abnormalities inserted by network traffic attacks.
@@
 Cosine similarity calculates the cosine of the angle between two vectors, which represents the similarity of values between the selected vectors.
-Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\boldsymbol{V}^{(q)}$ against the most significant eigenvectors of time frame detected as under attack, to analyze similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.
+Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\pmb{V}^{(q)}$ against the most significant eigenvectors of time frame detected as under attack, to analyze similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.
 
@@
 \begin{equation}\label{eq:2.11}
-s_n = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}^{(q)}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}^{(q)}}},
-\end{equation}
-where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\boldsymbol{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, according the attack detection described by Algorithm \ref{alg:2.01}, and $\boldsymbol{v}_{(n)}^{(q)}$ is the most significant eigenvectors obtained after append each target $n$-th minute of traffic that needs the identification of flood and port scan attacks.
-
-The most significant eigenvector $\boldsymbol{v}^{(q)}$, of a time frame $q$ without attack, can be computed from (\ref{eq:2.06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the selected time frame $q$.
-The same calculation shall be performed in order to obtain the target eigenvectors $\boldsymbol{v}_{(n)}^{(q)}$, calculated from the time frame without attack appended by selected minutes of a time frame estimated as under attack.
-
-Therefore, the reference eigenvectors $\boldsymbol{v}^{(q)}$ is calculated from the traffic without attack, in a time frame $q$ composed of $Q$ minutes of legitimate network traffic, estimated as normal by Algorithm \ref{alg:2.01}.
-For the detailed attack identification, each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frames shall be individually appended into $\boldsymbol{X}^{(q)}$, as represented by
+s_n = \frac{\abs{\pmb{v}^{(q)} \cdot \pmb{v}_{(n)}^{(q)}}}{\norm{\pmb{v}^{(q)}}\norm{\pmb{v}_{(n)}^{(q)}}},
+\end{equation}
+where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\pmb{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, according the attack detection described by Algorithm \ref{alg:2.01}, and $\pmb{v}_{(n)}^{(q)}$ is the most significant eigenvectors obtained after append each target $n$-th minute of traffic that needs the identification of flood and port scan attacks.
+
+The most significant eigenvector $\pmb{v}^{(q)}$, of a time frame $q$ without attack, can be computed from (\ref{eq:2.06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the selected time frame $q$.
+The same calculation shall be performed in order to obtain the target eigenvectors $\pmb{v}_{(n)}^{(q)}$, calculated from the time frame without attack appended by selected minutes of a time frame estimated as under attack.
+
+Therefore, the reference eigenvectors $\pmb{v}^{(q)}$ is calculated from the traffic without attack, in a time frame $q$ composed of $Q$ minutes of legitimate network traffic, estimated as normal by Algorithm \ref{alg:2.01}.
+For the detailed attack identification, each $\pmb{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\hat{\pmb{q}}}_{\rm max}$ time frames shall be individually appended into $\pmb{X}^{(q)}$, as represented by
 
 \begin{equation}\label{eq:2.12}
-\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\}.
-\end{equation}
-
-Subsequently, the resultant $\boldsymbol{X}_{(n)}$ is used to obtain $\boldsymbol{v}_{(n)}^{(q)}$, through (\ref{eq:2.06}), for calculating the similarity degree $s_n$, ranging from 0 to 1, for each $n$-th minute.
+\pmb{X}_{n} = \{\pmb{X}^{(q)} | \pmb{x}^{(\hat{q})}_{(n)}\}.
+\end{equation}
+
+Subsequently, the resultant $\pmb{X}_{(n)}$ is used to obtain $\pmb{v}_{(n)}^{(q)}$, through (\ref{eq:2.06}), for calculating the similarity degree $s_n$, ranging from 0 to 1, for each $n$-th minute.
 The $s_n$ denotes the absolute similarity degree of the $n$-th minute in comparison to a well-known traffic without attack, detected through MOS schemes and eigenvalue analysis.
@@
 
-The incremental approach for eigen similarity analysis is based on the incremental appending of network traffic into $\boldsymbol{X}^{(q)}$, where the first evaluation is based on (\ref{eq:2.12}) and the subsequent evaluations are based on (\ref{eq:2.13}), that denotes the appending of the $n$-th minute $\boldsymbol{x}^{(\hat{q})}_{(n)}$ into $\boldsymbol{X_n}$.
-The incremental approach repeats the Equation \ref{eq:2.13} while $n \leq N$ and compute $\boldsymbol{v}^{(q)}$, $\boldsymbol{v}_{(n)}^{(q)}$ and the Equation \ref{eq:2.11} for each increment.
+The incremental approach for eigen similarity analysis is based on the incremental appending of network traffic into $\pmb{X}^{(q)}$, where the first evaluation is based on (\ref{eq:2.12}) and the subsequent evaluations are based on (\ref{eq:2.13}), that denotes the appending of the $n$-th minute $\pmb{x}^{(\hat{q})}_{(n)}$ into $\pmb{X_n}$.
+The incremental approach repeats the Equation \ref{eq:2.13} while $n \leq N$ and compute $\pmb{v}^{(q)}$, $\pmb{v}_{(n)}^{(q)}$ and the Equation \ref{eq:2.11} for each increment.
 
 \begin{equation}\label{eq:2.13}
-\boldsymbol{X}_{n} = \{\boldsymbol{X}_{n} | \boldsymbol{x}^{(\hat{q})}_{(n)}\},
-\end{equation}
-
-Figure \ref{fig:2.08} illustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\boldsymbol{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected by Algorithm \ref{alg:2.01}.
+\pmb{X}_{n} = \{\pmb{X}_{n} | \pmb{x}^{(\hat{q})}_{(n)}\},
+\end{equation}
+
+Figure \ref{fig:2.08} illustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\pmb{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected by Algorithm \ref{alg:2.01}.
 
@@
 \begin{figure}[h!]
-     \includegraphics[width=15cm]{figures/incremental.eps}
+     \includegraphics[width=15cm]{figures/ch2/incremental.eps}
      \caption{Traffic selection for incremental approach.}
@@
 
-For the scenario depicted by Figure \ref{fig:2.08}, the eigen similarity analysis starts at $\boldsymbol{x}^{(3)}_{(41)}$ and is incrementally performed until $\boldsymbol{x}^{(3)}_{(60)}$, in order to calculate the $s_n$.
+For the scenario depicted by Figure \ref{fig:2.08}, the eigen similarity analysis starts at $\pmb{x}^{(3)}_{(41)}$ and is incrementally performed until $\pmb{x}^{(3)}_{(60)}$, in order to calculate the $s_n$.
 We assume that $s_n < l$ means an attack identification, according the anomaly on similarity of $s_n$ compared to a defined limiar $l$.
 
-Therefore, after obtaining the most significant eigenvector $\boldsymbol{v}^{(q)}$ and the target eigenvectors $\boldsymbol{v}_{(n)}^{(q)}$ for eigen similarity analysis, the $s_n$ is calculated according to (\ref{eq:2.11}).
+Therefore, after obtaining the most significant eigenvector $\pmb{v}^{(q)}$ and the target eigenvectors $\pmb{v}_{(n)}^{(q)}$ for eigen similarity analysis, the $s_n$ is calculated according to (\ref{eq:2.11}).
 If $s_n = 1$, then the two eigenvectors are completely similar and no anomaly is detected.
@@
 \begin{equation}\label{eq:2.14}
-  \boldsymbol{\hat{n}_{(n)}}=\left\{
+  \hat{\pmb{n}}_{(n)}=\left\{
   \begin{array}{@{}ll@{}}
@@
 \end{equation}
-and $\boldsymbol{\hat{n}}_{(n)}$ denotes a vector of $n$-th minutes detected as under attack.
-
-The eigen similarity analysis can also be applied by means of the individual approach, where each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, as shown by Figure \ref{fig:2.09}.
-In the individual approach there is no incremental appending, therefore only individual $n$-th minute are appended to $\boldsymbol{X_n}$, according to \ref{eq:2.12}, in order to compute $\boldsymbol{v}^{(q)}$, $\boldsymbol{v}_{(n)}^{(q)}$ and the Equation \ref{eq:2.11} for individual $n$-th minute.
+and $\hat{\pmb{n}}_{(n)}$ denotes a vector of $n$-th minutes detected as under attack.
+
+The eigen similarity analysis can also be applied by means of the individual approach, where each $n$-th minute must be individually appended into $\pmb{X}^{(q)}$, as shown by Figure \ref{fig:2.09}.
+In the individual approach there is no incremental appending, therefore only individual $n$-th minute are appended to $\pmb{X_n}$, according to \ref{eq:2.12}, in order to compute $\pmb{v}^{(q)}$, $\pmb{v}_{(n)}^{(q)}$ and the Equation \ref{eq:2.11} for individual $n$-th minute.
 
@@
 \begin{figure}[h!]
-     \includegraphics[width=15cm]{figures/individualized.eps}
+     \includegraphics[width=15cm]{figures/ch2/individualized.eps}
      \caption{Traffic selection for individual approach.}
@@
 
-The incremental and the individual approaches can be combined to obtain the incremental individualized approach, where each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}^{(q)}$ to compute similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack, i.e.
+The incremental and the individual approaches can be combined to obtain the incremental individualized approach, where each minute is incrementally appended into the selected $\pmb{X}^{(q)}$ for obtaining $\pmb{v}_{(n)}^{(q)}$ to compute similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack, i.e.
 $s_n < l$.
-Subsequently, $\boldsymbol{X}_{n-1}$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:2.02}.
+Subsequently, $\pmb{X}_{n-1}$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:2.02}.
 
@@
 \begin{figure}[h!]
-     \includegraphics[width=15cm]{figures/incremental_individualized.eps}
+     \includegraphics[width=15cm]{figures/ch2/incremental_individualized.eps}
      \caption{Traffic selection for incremental individualized approach.}
@@
 
-Given $\boldsymbol{\hat{n}}$, which is the set of $n$-th minutes under attack, it is still necessary to obtain more details about the identified network attack, such as the network ports that are attacked during each $n$-th minute identified as under attack.
+Given $\hat{\pmb{n}}$, which is the set of $n$-th minutes under attack, it is still necessary to obtain more details about the identified network attack, such as the network ports that are attacked during each $n$-th minute identified as under attack.
 Hence, it is also applied the cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port during a $n$-th minute.
 
-For detection of ports under attack, the last most significant eigenvectors without attack $\boldsymbol{v}^{(q)}$ shall be used as reference for similarity analysis against the $\boldsymbol{v}_{(n)}^{(q)}$ identified as under attack, and evaluate individually the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes.
-Therefore, $\boldsymbol{v}^{(q)}$ should be calculated from the last $\boldsymbol{X}^{(q)}$ time frame without attack, and $\boldsymbol{v}_{(m,\hat{n})}$ should be calculated from the same traffic appended of all $n$-th minutes until the identified minute under attack, denoted as $\boldsymbol{X}_n$.
-
-For similarity analysis, each $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, denoted as $x_{(m,n)}$, shall be individually replaced by the traffic of the evaluated $m$-th port of the $\hat{n}$-th minute under attack, denoted as $x^{(\hat{q})}_{(m,\hat{n})}$, in order to identify significant variation on similarity caused by the traffic of the $m$-th port.
+For detection of ports under attack, the last most significant eigenvectors without attack $\pmb{v}^{(q)}$ shall be used as reference for similarity analysis against the $\pmb{v}_{(n)}^{(q)}$ identified as under attack, and evaluate individually the cosine similarity of each $m$-th port of all $\hat{\pmb{n}}$ minutes.
+Therefore, $\pmb{v}^{(q)}$ should be calculated from the last $\pmb{X}^{(q)}$ time frame without attack, and $\pmb{v}_{(m,\hat{n})}$ should be calculated from the same traffic appended of all $n$-th minutes until the identified minute under attack, denoted as $\pmb{X}_n$.
+
+For similarity analysis, each $m$-th port of the last $n$-th minute of $\pmb{X}_n$, denoted as $x_{(m,n)}$, shall be individually replaced by the traffic of the evaluated $m$-th port of the $\hat{n}$-th minute under attack, denoted as $x^{(\hat{q})}_{(m,\hat{n})}$, in order to identify significant variation on similarity caused by the traffic of the $m$-th port.
 
@@
     \\
-    s_{m,\hat{n}} = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(m,\hat{n})}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(m,\hat{n})}}},
+    s_{m,\hat{n}} = \frac{\abs{\pmb{v}^{(q)} \cdot \pmb{v}_{(m,\hat{n})}}}{\norm{\pmb{v}^{(q)}}\norm{\pmb{v}_{(m,\hat{n})}}},
   \end{array}\right.
 \end{equation}
-where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute and $q$-th time frame identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\boldsymbol{X}_n$, which is used to calculate the $\boldsymbol{v}_{(m,\hat{n})}$ most significant eigenvectors that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.
-
-Once $\boldsymbol{v}^{(q)}$ and $\boldsymbol{v}_{(m,\hat{n})}$ are obtained, then the $s_{m,\hat{n}}$ similarity degree can be calculated in order to identify if the traffic replacement highlights the adition of anomalous traffic by the evaluated $m$-th port during the $\hat{n}$-th minute previously identified as under attack.
-
-This procedure should be repeated for each $m$-th target port of $\boldsymbol{\hat{n}}$, in order to individually identify the network ports under attack during each $\hat{q}$-th time frame.
+where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute and $q$-th time frame identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\pmb{X}_n$, which is used to calculate the $\pmb{v}_{(m,\hat{n})}$ most significant eigenvectors that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.
+
+Once $\pmb{v}^{(q)}$ and $\pmb{v}_{(m,\hat{n})}$ are obtained, then the $s_{m,\hat{n}}$ similarity degree can be calculated in order to identify if the traffic replacement highlights the adition of anomalous traffic by the evaluated $m$-th port during the $\hat{n}$-th minute previously identified as under attack.
+
+This procedure should be repeated for each $m$-th target port of $\hat{\pmb{n}}$, in order to individually identify the network ports under attack during each $\hat{q}$-th time frame.
 
@@
 Therefore, as the time of each sampling period is one minute, then $N = 20$.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, as well as a covariance $\boldsymbol{\hat{R}}_{yy}^{(q)} \in \mathbb{R}^{17 \times 17}$ (calculated via (\ref{eq:2.03})) and a sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)} \in \mathbb{R}^{17x17}$, assuming that $q = 1, 2, 3, 4, 5$ and $6$.
+For each time frame $q$, a traffic matrix $\pmb{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, as well as a covariance $\hat{\pmb{R}}_{yy}^{(q)} \in \mathbb{R}^{17 \times 17}$ (calculated via (\ref{eq:2.03})) and a sample covariance matrix $\hat{\pmb{R}}_{zz}^{(q)} \in \mathbb{R}^{17x17}$, assuming that $q = 1, 2, 3, 4, 5$ and $6$.
 
@@
 	\centering
-     \includegraphics[width=11cm]{figures/eigenvalues_synflood.eps} 
+     \includegraphics[width=11cm]{figures/ch2/eigenvalues_synflood.eps} 
      \caption{Eigenvalues of the sample covariance matrix (synflood).}
@@
 	\centering
-     \includegraphics[width=11cm]{figures/eigenvalues_fraggle.eps}
+     \includegraphics[width=11cm]{figures/ch2/eigenvalues_fraggle.eps}
      \caption{Eigenvalues of the sample covariance matrix (fraggle).}
@@
 	\centering
-     \includegraphics[width=11cm]{figures/eigenvalues_portscan.eps}
+     \includegraphics[width=11cm]{figures/ch2/eigenvalues_portscan.eps}
      \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}
@@
 	\toprule
-	\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
+	\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{MOS schemes (estimated model order $\hat{d}$)} &{(d)}\\ 
 			\hhline{~------~}
-		&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
+		&\pmb{AIC} &\pmb{MDL} &\pmb{EDC} &\pmb{RADOI} &\pmb{EFT} &\pmb{SURE}\\
 	\midrule
-	Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-	Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
+	Detection of synflood \\(presence of attack) &2 &1 &\pmb{1} &5 &\pmb{1} &4 &\pmb{1} \\
+	Detection of synflood \\(absence of attack) &1 &1 &\pmb{0} &1 &\pmb{0} &3 &\pmb{0} \\
 	\midrule
-	Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
-	Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
+	Detection of fraggle \\(presence of attack) &1 &1 &\pmb{1} &5 &\pmb{1} &4 &\pmb{1} \\
+	Detection of fraggle \\(absence of attack) &1 &1 &\pmb{0} &1 &\pmb{0} &3 &\pmb{0} \\
 	\midrule
-	Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
-	Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
+	Detection of port scan \\(presence of attack) &1 &1 &\pmb{1} &1 &\pmb{1} &9 &\pmb{1} \\
+	Detection of port scan \\(absence of attack) &0 &0 &\pmb{0} &1 &\pmb{0} &1 &\pmb{0} \\
 	\midrule
-	Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
-	Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
+	Detection of synflood/fraggle \\(presence of attack) &2 &2 &\pmb{2} &5 &\pmb{2} &5 &\pmb{2} \\
+	Detection of synflood/fraggle \\(absence of attack) &1 &1 &\pmb{0} &1 &\pmb{0} &3 &\pmb{0} \\
     \bottomrule
@@
 
-Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respective $q$-th time frames under attack, which is denoted by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:2.01}.
+Therefore, we conclude that the $\hat{d}$ largest eigenvalues correspond to the respective $q$-th time frames under attack, which is denoted by $\rm{\hat{\pmb{q}}}_{\rm max}$ and can be calculated according to Algorithm \ref{alg:2.01}.
 
@@
 
-This work proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$.
+This work proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\hat{\pmb{q}}}_{\rm max}$.
 Hence, the proposed framework is applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.
@@
 Three approaches were evaluated for eigen similarity analysis: incremental, individual and incremental individualized approaches.
-For the incremental individualized approach, each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}^{(q)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
-Subsequently, $\boldsymbol{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated.
-For the incremental approach, each $n$-th minute must be incrementally appended into $\boldsymbol{X}^{(q)}$, for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}^{(q)}$ for individual time similarity analysis.
-For the individual approach, each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, without incremental append, but doing individual appended into $\boldsymbol{X}^{(q)}$ for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}^{(q)}$ for individual similarity analyis.
+For the incremental individualized approach, each minute is incrementally appended into the selected $\pmb{X}^{(q)}$ for obtaining $\pmb{v}_{(n)}^{(q)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
+Subsequently, $\pmb{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated.
+For the incremental approach, each $n$-th minute must be incrementally appended into $\pmb{X}^{(q)}$, for obtaining the next eigenvectors $\pmb{v}_{(n)}^{(q)}$ for individual time similarity analysis.
+For the individual approach, each $n$-th minute must be individually appended into $\pmb{X}^{(q)}$, without incremental append, but doing individual appended into $\pmb{X}^{(q)}$ for obtaining the next eigenvectors $\pmb{v}_{(n)}^{(q)}$ for individual similarity analyis.
 
@@
 	\toprule
-	\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Ground Truth}}\\ 
+	\multirow{2}{*}{\pmb{Time Frame} $q$} &\multirow{2}{*}{\pmb{Time} $n$}   &\multicolumn{3}{c}{\pmb{Similarity Analysis}} &\multirow{2}{*}{\pmb{Ground Truth}}\\ 
 			\hhline{~~---~}
-			& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
+			& &\pmb{Incremental Individualized} &\pmb{Incremental} &\pmb{Individual}\\
 	\midrule
@@
 	\toprule
-	\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Ground Truth}}\\ 
+	\multirow{2}{*}{\pmb{Port} $p$}   &\multicolumn{2}{c}{\pmb{Approaches}} &\multirow{2}{*}{\pmb{Ground Truth}}\\ 
 			\hhline{~--~}
-			&\textbf{Incremental Individualized} &\textbf{Individual}\\
+			&\pmb{Incremental Individualized} &\pmb{Individual}\\
 	\midrule
@@
 s ($N = 60$) and aggregate by minute and by port number.
-For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ is obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.
+For each time frame $q$, a traffic matrix $\pmb{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ is obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.
 

### 2019-06-14 docs/thesis/chapters/4_m_rpca/m_rpca.tex
after fix bold notation

@@
 
-Some widely adopted algorithms for anomaly detection assume a gaussian distributed data, however this assumption may not be observed in some real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network}.
+Some widely adopted algorithms for anomaly detection assume a Gaussian distributed data, however this assumption may not be observed in some real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network, leon2017probability}.
 
@@
 
-This findings show that the skeweness and heavy-tailed distributions may be important for network traffic analysis, and can motivate evaluations to study the impact of skewed data into algorithms that rely on gaussian distributed data, as well as it can highlight opportunities to  exploit the skeweness and heavy-tailed distributions in order to obtain accurate classifiers for network anomaly detection.
+These findings show that the skeweness and heavy-tailed distributions may be important for network traffic analysis, and can motivate researches to evaluate the impact of skewed data into algorithms that rely on Gaussian distributed data.
+These findings can also highlight opportunities to  exploit the skeweness and heavy-tailed distributions to obtain accurate classifiers for network anomaly detection.
 
@@
 
-We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data, 
-and that the distance between robust estimates of normal traffic and anomalies can highlight discrepancies and be used for network attack detection.
-Therefore, we propose the md-RPCA, which is an approach based on distances from moments computed from robust subspace learned by Robust Principal Component Analysis, in order to detect anomalies from network traffic.
-The proposed approach relies on a robust subspace of normal traffic for estimating higher moments.
-The anomaly detection from new observations evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervised or unsupervised fashion, without the computational cost of new robust subspace learning for new observations.
+We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data.
+We also believe that the distance between robust estimates of normal traffic and contaminated observations can highlight discrepancies and be used for network attack detection.
 
-We evaluate the accuracy of the md-RPCA for anomaly detection on synthetic data set and on the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
+Therefore, we propose the m-RPCA, which is an approach based on distances of moments computed from robust subspace learned by Robust Principal Component Analysis (RPCA), in order to detect anomalies from skewed data and network traffic.
+The proposed approach relies on a robust subspace computed from supposed normal observations, for estimating the moments to be used for distance analysis.
+The anomaly detection from new observations evaluate the Mahalanobis distance between the robust moments and the moments of new observations, in a semi-supervised fashion, without the computational cost of new robust subspace learning for new observations.
+The m-RPCA can also be computed as an unsupervised algorithm, with subspace learning from the same contaminated data that is the target of the anomaly detection analysis.
+
+We evaluate the accuracy of the m-RPCA for anomaly detection on simulated data sets, with skewed and heavy-tailed distributions, and for the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
 The Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches.
+We also compare the results to ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
 
-% We also evaluate the results of ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
-
-The main contribution of this work is a novel semi-supervised and unsupervised method for anomaly detection in skewed and imbalanced data, with results of experimental evaluation on synthetic and real data sets.
+The main contribution of this work is the proposal of a novel semi-supervised and unsupervised method for anomaly detection in skewed and imbalanced data, with results of experimental evaluation on simulated and real data sets.
 
@@
 Section \ref{sec:4_proposal} describes the proposed approach for network attack detection.
-Section \ref{sec:4_experiments} discusses the experimental validation and Section \ref{sec:4_results} presents the results.
+Section \ref{sec:4_experiments} discusses the experimental validation, Section \ref{sec:4_simulated_result} presents the results of the simulated scenarios and the Section \ref{sec:4_ctu13_result} describes the results for anomaly detection on CTU-13 data set.
 Finally, Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.
@@
 
-% TODO - re-write this, it is strange
-% network anomaly detectiom
+% network anomaly detection
+Network Anomaly detection has emerging as an important approach for securing communication networks and to deal with the increasing number of network attacks.
 Bhuyan \emph{et al.} \cite{bhuyan2014network} provide an overview of facets of network anomaly detection, present attacks normally encountered by network intrusion detection systems, and categorize existing network anomaly detection methods and systems based on the underlying techniques.
-Ahmed \emph{et al.} \cite{ahmed2016survey} present an analysis of four major categories of anomaly detection techniques which include classification, statistical, information theory and clustering.
-Moustafa \emph{et al.} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), explains cyber-attacks and new solutions for anomaly detection, and provides a benchmark data sets for training and validating approaches for detection.
+Ahmed \emph{et al.} \cite{ahmed2016survey} present an analysis of four major categories of anomaly detection techniques, which include classification, statistical, information theory and clustering.
+Moustafa \emph{et al.} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), describing details of cyber-attacks and new solutions for anomaly detection, and provides a benchmark data sets for training and validating approaches for network anomaly detection.
 
@@
 
-Acarali \emph{et al.} \cite{acarali2016survey} surveys network-based detection approaches for HTTP-based botnets, surveys traffic-based features used to detect bot traffic and presents an abstraction of the main types of features related to protocols and OSI layers.
+Acarali \emph{et al.} \cite{acarali2016survey} surveys network-based detection approaches for HTTP-based botnets, and discuss the traffic-based features used to detect bot traffic and presents an abstraction of the main types of features related to protocols and OSI layers.
 Wang \emph{et al.} \cite{Wang2018ddosbotnetssurvey} present an analysis based on 50,704 different Internet DDoS attacks originated of 674 botnets from 23 different botnet families with a total of 9,026 victim belonging to 1,074 organizations in 186 countries.
@@
 
-% justification for the data set and present who else use it
 Several approaches for network attack detection uses the KDD 99 \cite{ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
 Even though the KDD 99 data set are criticized by the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data sets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
-NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that redundant data records are removed, in order to avoid biased classifications.
-However, NSL-KDD data set maintain the limitations of the KDD 99 regarding volume and reproduction of network traffic similar to the real networks.
+NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that removed the redundant data records, in order to avoid biased classifications.
+However, NSL-KDD data set maintain the limitations of the KDD 99 regarding volume and lacks on reproduction of recent network traffic and attacks.
 
-The objective of Garcia \emph{et al.} \cite{garcia2014empirical} is to compare three botnet detection methods using a simple and reproducible methodology, a good data set and a new error metric.
-This paper evaluates some data sets for network anomaly detection, surveys some approaches for botnet detection, proposes two methods (BClus and CAMNEP) for botnet detection and compare results to BotHunter \cite{gu2007bothunter}.
+The objective of Garcia \emph{et al.} \cite{garcia2014empirical} is to compare three botnet detection methods by means of a simple and reproducible methodology, by a good data set and by a new error metric.
+This paper evaluates some data sets for network anomaly detection, and surveys some approaches for botnet detection, and proposes two methods (BClus and CAMNEP) for botnet detection, comparing results to BotHunter \cite{gu2007bothunter}.
 
-Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by attack, normal and background labeled data, in an imbalanced distribution like in a real network.
-The authors recommend scenarios for training and testing, in the way that none of the botnet families used in the training and cross-validation data set should be used in the testing data set, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
-Adopting the training and testing approach proposed by Garcia \emph{et al.}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
+Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by attack, normal and background labeled data, in an imbalanced distribution like in a real networks.
+The authors recommend scenarios for training and testing in order to avoid the use of traffic from a botnet family for training and testing, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
+Considering the adoption of unsupervised or semi-supervised approaches for anomaly detection, adopting the training and testing approach proposed by Garcia \emph{et al.}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
 
@@
 The second stage detects the bots using ideas from social network community detection in a graph that captures correlations of interactions among nodes over time.
-This work is compared with the BotHunter \cite{gu2007bothunter} on the CTU-13 botnet data set \cite{garcia2014empirical}.
+This work is compared to the BotHunter \cite{gu2007bothunter} for the CTU-13 botnet data set \cite{garcia2014empirical}.
 
-% \cite{da2018online} proposes an approach using the Very Fast Decision Tree, a classification algorithm used on stream mining that can learn incrementally when needed, to identify botnets by observing network flows.
-When evaluating the approach on multiple scenarios with different botnets, we were able to achieve high performance metrics on the majority of scenarios, while using a significantly low number of labelled instances.
-It discard scenarios 5 and 6.
-They does not compute F1, but it is easy to compute it from precision and recall provided.
-
-% Anomaly detection by subspace larning
+% Anomaly detection by subspace learning
 Traditional PCA-based anomaly detection models are not suitable for anomaly interpretation, as they judge whether a data instance is an anomaly or not based on the length of its projection on the abnormal subspace spanned by the less significant principal components, and there is no direct mapping between PCA’s dimensionality-reduced subspace and the original feature space \cite{ringberg2007sensitivity}.
@@
 The problem of PCA or subspace learning for outlier corrupted data is called Robust Principal Component Analysis (RPCA) or robust subspace learning \cite{candes2011robust, vaswani2018robust}.
-RPCA aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, which refers to outliers’ matrix.
-This definition is also referred to as the sparse + low rank (S + LR) formulation \cite{vaswani2018robust}.
+RPCA aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\pmb{X}$ into the sum of a low rank matrix $\pmb{L}$, whose column subspace gives the principal components, and a sparse matrix $\pmb{S}$, which refers to outliers’ matrix.
+This definition is also referred to as the sparse + low rank ($\pmb{S + LR}$) formulation \cite{vaswani2018robust}.
 We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for more details regarding robust subspace learning.
@@
 Therefore, Hubert \emph{et al.} \cite{hubert2009robustskewed} proposed ROBPCA-AO, which improves ROBPCA for problems with skewed data, by means of an adjusted outlyingness based on robust skewness.
+Hubert \emph{et al.} \cite{hubert2009robustskewed} evaluated ROBPCA-AO for real and simulated data sets, but it is not clear if this method was evaluated for very skewed and imbalanced data sets, such as in the network attack detection problem.
 
@@
 
-Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the covariances between the variables and their mean values.
-The MD is a measure of the distance between a vector $\boldsymbol{x}$ and a distribution $\boldsymbol{X}$, introduced by P.
+Mahalanobis Distance (MD) is a generalized distance which is useful for determining the similarity between an unknown sample and a collection of known samples, by considering the covariance between the variables and their mean values.
+The MD is a measure of the distance between a vector $\pmb{x}$ and a distribution $\pmb{X}$, introduced by P.
 C.
 Mahalanobis in 1936 \cite{mahalanobis1936md}.
-It is a multi-dimensional generalization for measuring how many standard deviations away $\boldsymbol{x}$ is from the mean $\boldsymbol{\bar{x}}$ and covariance $\boldsymbol{\bar{\Sigma}}$ of $\boldsymbol{X}$.
+It is a multi-dimensional generalization for measuring how many standard deviations away $\pmb{x}$ is from the mean $\bar{\pmb{x}}$ and covariance $\bar{\pmb{\Sigma}}$ of $\pmb{X}$.
 MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and new observations can reveal anomalies.
 
-We propose an approach based on higher moments distances from robust subspace for botnet detection.
-The proposed approach relies on a robust subspace of normal traffic for estimating higher moments.
-Thus, we propose that the anomaly detection for new observations should evaluate the Mahalanobis distance between the robust higher moments and the higher moments of new observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations, or in a unsupervised approach.
+We propose an approach based on the distances of moments computed from learned robust subspace and new observations, for anomaly detection on skewed and imbalanced data sets, and evaluate our approach for network attack detection.
+The proposed approach relies on learning robust subspace from supposed normal traffic for estimating the moments (mean, skewness and kurtosis).
+Thus, the anomaly detection for new observations should evaluate the Mahalanobis Distance between the robust moments and the moments of new observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations, or in a unsupervised approach, where the robust subspace is learned from the contaminated data and the distance analysis is between the robust moments and the contaminated moments.
 
@@
 
-In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $\alpha$, $\beta$}), vectors by lowercase bold letters (\textbf{a, b}), matrices by uppercase bold letters (\textbf{A, B}), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix \textbf{A}.
+In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $\alpha$, $\beta$}), vectors by lowercase bold letters ($\pmb{a}, \pmb{b}$), matrices by uppercase bold letters ($\pmb{A}$, $\pmb{B}$), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix $\pmb{A}$.
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
@@
 
-This section also presents a description of the synthetic data model in Subsection \ref{sec:4_synthetic} and in Subsection \ref{sec:4_CTU-13} we describe the data model for the CTU-13 data set.
+This section also presents a description of the simulated data model in Subsection \ref{sec:4_simulation} and in Subsection \ref{sec:4_CTU-13} we describe the data model for the CTU-13 data set.
 
-\subsection{The Synthetic data set}
-\label{sec:4_synthetic}
+\subsection{The Simulated data set}
+\label{sec:4_simulation}
 
-In order to analyse the null hypotheses $H_{1rob.sub}$ and $H_{0rob.sub}$, which evaluate if the robust subspace learning can improves the anomaly detection from imbalanced and skewed data, we create two synthetic data sets characterized by skewed and heavy tailed distributions.
-We selected Pareto and Lognormal distributions to evaluate the anomaly detection capacity of our proposal, considering that these distributions are skewed and heavy-tailed, and have been adopted to characterize network traffic of the Internet and data centers \cite{benson2010network,leon2017probability}.
-We also evaluate our proposal and the selected algorithms for anomaly detection on gaussian distributed data set, in order to analyse the detection rate for not skewed and not heavy tailed distributions.
+In order to analyse the null hypotheses $H_{1rob.sub}$ and $H_{0rob.sub}$, which evaluate if the robust subspace learning can improves the anomaly detection from imbalanced and skewed data, we create two simulated data sets characterized by skewed and heavy tailed distributions, and create a simulated Gaussian distributed data set, in order to also analyse the detection rate for not skewed and not heavy tailed distributions.
 
-The addictive anomalies for Pareto and Lognormal distributions are gaussian distributed, while an uniform noise is used to add anomalies into the gaussian distributed data set.
-We generate $\boldsymbol{Y}_p$, $\boldsymbol{Y}_l$ and $\boldsymbol{Y}_g$ for Pareto, Lognormal and Gaussian multivariate distributions, respectively, with 2 features.
-Each contaminated data set is composed by a number of legitimate observations and the contaminated samples, defined by the adopted $c$.
+We selected Pareto (with scale $a=3$, mode $m=1$ and $\mu=1$) and Lognormal (with mean $\mu=0$ and standard deviation $\sigma=1$) distributions, denoted respectively as $\pmb{Y}_p$, while $\pmb{Y}_l$, to simulate normal events and to evaluate the anomaly detection on skewed data performed by our proposals in comparison to widely adopted algorithms for outlier detection, considering that these distributions are skewed and heavy-tailed, and have been adopted to characterize network traffic of the Internet and data centers \cite{benson2010network,leon2017probability}.
+We also adopt the Gaussian distributed data set $\pmb{Y}_g$ to simulate normal events with mean $\mu=0$ and unitary standard deviation $\sigma=1$.
 
-The Pareto distribution contaminated by gaussian noise is denoted as $\boldsymbol{Y}_p^c$, while $\boldsymbol{Y}_l^c$ denotes the Lognormal distribution contaminated by gaussian noise, and $\boldsymbol{Y}_g^c$ denotes the contaminated gaussian data contaminated by an uniform distribution.
-
-This synthetic data set simulate a total of 2400 events for each scenario, where the number of legitimate observations is simulated according to the desired contamination rate.
-We evaluate contamination rates $c$ between 1\% and 50\%, to simulate the imbalanced data of anomaly detection problems during our experiments.
-
-The Figure \ref{fig:4.01} depicts the distribution of Pareto and gaussian contamination of one variable, as a visual example of $\boldsymbol{Y}_p^c$, while the Figure \ref{fig:4.02} shows one example of Lognormal contaminated by gaussian noise, generated for $\boldsymbol{Y}_l^c$.
+The addictive anomalies for Pareto and Lognormal distributions are white Gaussian noise with mean $\mu=0$ and unitary standard deviation $\sigma=1$.
+The Pareto distribution contaminated by Gaussian noise is denoted as $\pmb{Y}_p^c$, while $\pmb{Y}_l^c$ denotes the Lognormal distribution contaminated by Gaussian noise.
+The $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$ are depicted by the Figures \ref{fig:4.01} and \ref{fig:4.02}, in order to provide a visual example one variable of $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$.
 
@@
 	\subfigure[Pareto and Gaussian Anomalies]{%
-		\includegraphics[width=0.47\textwidth]{figures/ch4/4_Xpg2.png}
+		\includegraphics[width=0.48\textwidth]{figures/ch4/4_Xpg2.png}
 		\label{fig:4.01}
@@
 	\subfigure[Lognormal and Gaussian Anomalies]{%
-		\includegraphics[width=0.47\textwidth]{figures/ch4/4_Xlogng2.png}
+		\includegraphics[width=0.48\textwidth]{figures/ch4/4_Xlogng2.png}
 		\label{fig:4.02}
@@
 \end{figure}
+
+An uniform noise between $-6$ and $6$ is used to add anomalies into the Gaussian distributed data set, and $\pmb{Y}_g^c$ denotes the contaminated Gaussian data contaminated by an uniform distribution.
+The figure \ref{fig:4.04} shows an example of the Gaussian distribution of normal observations and the uniform distribution to simulate the addition of anomalies.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=8cm]{figures/ch4/4_Xgu2.png}
+	\caption{Example of Gaussian and Uniform Anomalies}
+	\label{fig:4.04}
+\end{figure}
+
+Each contaminated data set is composed by a number of legitimate observations and contaminated samples.
+We evaluate contamination rates $c$ between 1\% and 50\%, to simulate the imbalanced data of anomaly detection problems during our experiments.
+This data sets simulate a total of 2400 events for each scenario with contaminated Pareto, Lognormal and Gaussian.
+Therefore, the number of legitimate observations is defined according to the contamination rate selected for each evaluation.
 
@@
 \begin{itemize}
-	\item \textbf{Attacks:} Click Fraud (CF), Port Scan (PS), Compiled and Controlled by Authors (US), SPAM and DDOS;
-	\item \textbf{C\&C:} IRC, P2P and HTTP.
+	\item \pmb{Attacks:} Click Fraud (CF), Port Scan (PS), Compiled and Controlled by Authors (US), SPAM and DDOS;
+	\item \pmb{C\&C:} IRC, P2P and HTTP.
 \end{itemize}
@@
 	\begin{tabular}{| l | l | l | r | r | r | r | r | r | r | r | }
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}\textbf{ID}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\textbf{Bot}\end{tabular}	 &\begin{tabular}[x]{@{}l@{}}\textbf{Type}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\textbf{Total}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Malicious}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{C\&C}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Attack}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Normal}\end{tabular}\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}\pmb{ID}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\pmb{Bot}\end{tabular}	 &\begin{tabular}[x]{@{}l@{}}\pmb{Type}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\pmb{Total}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\pmb{Malicious}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\pmb{C\&C}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\pmb{Attack}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\pmb{Normal}\end{tabular}\\ \hline
 			10 & neris &\makecell[l]{IRC, Spam,\\CF} & 2,824,636 & 40,961 (1.45\%) & 341 (0.01\%) & 40,620 (1.44\%) &30,387 (1.07\%)\\ \hline
@@
 
-The full data set and scenarios can be denoted as $\boldsymbol{X} = \{\boldsymbol{X}_{10}, \boldsymbol{X}_{11}, \ldots , \boldsymbol{X}_{18-2}, \boldsymbol{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}.
-In our experiment each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^s$ containing 50\% of the the normal data, and into $\boldsymbol{X}_i^c$ that is composed by all anomalous flows and the necessary number of normal flows to have a testing data with the desired contamination rate.
+The full data set and scenarios can be denoted as $\pmb{X} = \{\pmb{X}_{10}, \pmb{X}_{11}, \ldots , \pmb{X}_{18-2}, \pmb{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}.
+In our experiment each contaminated scenario $\pmb{X}_i$ is split into $\pmb{X}_i^s$ containing 50\% of the the normal data, and into $\pmb{X}_i^c$ that is composed by all anomalous flows and the necessary number of normal flows to have a testing data with the desired contamination rate.
 
@@
 
-We conducted and Exploratory Data Analysis (EDA) on the CTU-13 and observed that some features are skewed and present an overlapping between normal and anomalous flows, as can be seen in Figure \ref{fig:4.05} and Figure \ref{fig:4.05}, that present the distributions of TCP state of scenario 16 and the type of services from destination of scenario 10.
+We conducted and Exploratory Data Analysis (EDA) on the CTU-13 and observed that some features are skewed and present an overlapping between normal and anomalous flows, as can be seen in Figure \ref{fig:4.05} and Figure \ref{fig:4.06}, that present the distributions of TCP state of scenario 16 and the type of services from destination of scenario 10.
 
@@
 		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110816_State.png}
-		\label{fig:4.04}
+		\label{fig:4.05}
 	}
@@
 		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110810_dTos.png}
-		\label{fig:4.05}
+		\label{fig:4.06}
 	}
 	\caption[Skewness and Overlapping]{Example of skewness and overlapping}
-	\label{fig:4.06}
+	\label{fig:4.07}
 \end{figure}
 
-However, it is not possible to observe a pattern on distributions of all the features, as depicted by Figures \ref{fig:4.07} and \ref{fig:4.08}, that show the distributions of TCP states of the scenario 10 and source ports of scenario 16, and highlight the distributions of normal and anomalous flows.
+However, it is not possible to observe a pattern on distributions of all the features and scenarios of CTU-13, as depicted by Figures \ref{fig:4.08} and \ref{fig:4.09}, that show the distributions of TCP states of the scenario 10 and source ports of scenario 16, and highlight the distributions of normal and anomalous flows.
 
@@
 		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110810_State.png}
-		\label{fig:4.07}
+		\label{fig:4.08}
 	}
@@
 		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110816_Sport.png}
-		\label{fig:4.08}
+		\label{fig:4.09}
 	}
 	\caption[Skewness and Overlapping]{Skewness and Overlapping}
-	\label{fig:4.09}
+	\label{fig:4.10}
 \end{figure}
 
-Due to the number of available features and the class overlapping between normal and anomalous flows, we performed an correlation analysis and cross validation in order to identify the best set of features for network attack detection.
+Due to the number of available features and the class overlapping between normal and anomalous flows, we performed an correlation analysis and an empirical cross validation in order to identify the best set of features for network attack detection.
 Therefore, we adopt the following features: state, destination type of service, destination port, source port, total number of packets, total number of bytes and number of bytes from the source.
 
-\section{Moment Distances from Robust Subspace for Botnet Detection}
+\section{Moment Distances from Robust Subspace for Network Attack Detection}
 \label{sec:4_proposal}
 
-This section describes the proposed approach for network attack detection by means of a distance analysis between moments computed from a robust subspace and new observations of network traffic.
+This section describes the proposed approach for network attack detection by means of a distance analysis between moments computed from a robust subspace and moments of new observations of network traffic.
 
-Robust subspace learning can be defined as the decomposition of a given data matrix $\textbf{X}$ into the sum of a low rank matrix $\textbf{L}$, whose column subspace gives the principal components, and a sparse matrix $\textbf{S}$, with outliers or noise.
-We propose to learn a robust subspace $\textbf{L}$ from the normal traffic $\textbf{X}^s$ for computing $\textbf{L}^s$, $\textbf{S}^s$ and the robust moments, i.e.
-the mean $\boldsymbol{\bar{x}}$, skewness $\boldsymbol{\bar{s}}$ and kurtosis $\boldsymbol{\bar{k}}$, in order to evaluate the distance $\boldsymbol{d}$ of new observations $\boldsymbol{X}^c$ to these moments and classify as anomalous the observations with the largest distances.
+Robust subspace learning can be defined as the decomposition of a given data matrix $\pmb{X}$ into the sum of a low rank matrix $\pmb{L}$, whose column subspace gives the principal components, and a sparse matrix $\pmb{S}$, with outliers or noise.
+We propose to learn a robust subspace $\pmb{L}$ from the normal traffic $\pmb{X}^s$ for computing $\pmb{L}^s$, $\pmb{S}^s$ and the robust moments, i.e.
+the mean $\bar{\pmb{x}}$, skewness $\bar{\pmb{s}}$ and kurtosis $\bar{\pmb{k}}$, in order to evaluate the distance $\pmb{d}$ between moments of new observations $\pmb{X}^c$ and moments computed from the learned robust subspace, in order to classify as anomalous the observations with the largest distances.
 
-Robust PCA is a well known method to recover a low-rank matrix $\textbf{L}$ and sparse matrix $\textbf{S}$ from corrupted measurements modeled as $\textbf{X} = \textbf{L} + \textbf{S}$.
+RPCA is a well known method to recover a low-rank matrix $\pmb{L}$ and sparse matrix $\pmb{S}$ from corrupted measurements modeled as $\pmb{X} = \pmb{L} + \pmb{S}$.
 This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}.
@@
 According to Wright et al.
-\cite{wright2009robust}, under rather broad conditions, as long as the error matrix S is sufficiently sparse, it is possible to recover a low-rank matrix by solving the following convex optimization problem:
+\cite{wright2009robust}, under rather broad conditions, as long as the error matrix $\pmb{S}$ is sufficiently sparse, it is possible to recover a low-rank matrix by solving the following convex optimization problem:
 \begin{equation}\label{eq:4.01}
-	(\boldsymbol{\hat{L}}, \boldsymbol{\hat{S}})\leftarrow \min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}
+	(\hat{\pmb{L}}, \hat{\pmb{S}})\leftarrow \min_{\pmb{L},\pmb{S}}\left \| \pmb{L} \right \|_{*} + \lambda \left \| \pmb{S} \right \|_{1}
 \end{equation}
-\begin{center} subject to: $\boldsymbol{X} = \boldsymbol{L} + \boldsymbol{S}$ \end{center}
+\begin{center} subject to: $\pmb{X} = \pmb{L} + \pmb{S}$ \end{center}
 \begin{equation}\label{eq:4.02}
-    \left \| \boldsymbol{L} \right \|_{*} = \sum_{i} \sigma_{i}(\boldsymbol{L})
+    \left \| \pmb{L} \right \|_{*} = \sum_{i} \sigma_{i}(\pmb{L})
 \end{equation}
 \begin{equation}\label{eq:4.03}
-    \left \| \boldsymbol{S} \right \|_{1} = \sum_{ij} \left | \boldsymbol{S}_{ij} \right |
+    \left \| \pmb{S} \right \|_{1} = \sum_{ij} \left | \pmb{S}_{ij} \right |
 \end{equation}
-where $\lambda$ is a positive weighting parameter, which determines the sparsity of $\boldsymbol{S}$, $\sigma$ denotes the singular values of a matrix.
+where $\lambda$ is a positive weighting parameter, which determines the sparsity of $\pmb{S}$, $\sigma$ denotes the singular values of a matrix.
 
@@
 However, according to Zhouchen \emph{et al.} \cite{lin2010augmented}, both approaches have scalability problems and require a large number of iterations to converge.
-The Augmented Lagrange Multiplier is proven to have a \emph{Q-linear} convergence speed and experimental results show that ALM is five times faster than APG, which in theory is sub-linear \cite{lin2010augmented}.
+The Augmented Lagrange Multiplier (ALM) is proven to have a \emph{Q-linear} convergence speed and experimental results show that ALM is five times faster than APG, which in theory is sub-linear \cite{lin2010augmented}.
 Furthermore, ALM reaches more accurate results with less iterations.
@@
 \begin{equation}\label{eq:4.04}
-	l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}) = \left\|\boldsymbol{L}\right\|_* + \lambda\left\|\boldsymbol{S}\right\|_1 + \langle \boldsymbol{Y, X - L - S}  \rangle + \frac{\mu}{2}\left\|\boldsymbol{X - L - S}\right\|_F^2,
+	l(\pmb{L}, \pmb{S}, \pmb{Y}) = \left\|\pmb{L}\right\|_* + \lambda\left\|\pmb{S}\right\|_1 + \langle \pmb{Y}, \pmb{X} - \pmb{L} - \pmb{S}  \rangle + \frac{\mu}{2}\left\|\pmb{X} - \pmb{L} - \pmb{S}\right\|_F^2,
 \end{equation}
-where $\boldsymbol{Y}$ is the multiplier of the linear constraint, $\mu$ is the penalty parameter for the violation of the linear constraint \cite{yuan2009sparse}.
+where $\pmb{Y}$ is the multiplier of the linear constraint, $\mu$ is the penalty parameter for the violation of the linear constraint \cite{yuan2009sparse}.
 Thus, an iterative scheme can be presented as:
@@
 		\begin{matrix} 
-			(\boldsymbol{L}_{k+1}, \boldsymbol{S}_{k+1}) \in \operatorname*{argmin}_{\boldsymbol{L,S} \in \mathbb{R}^{m \times n}} \{l(\boldsymbol{L}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}, \\ 
-			\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k}),
+			(\pmb{L}_{k+1}, \pmb{S}_{k+1}) \in \operatorname*{argmin}_{\pmb{L,S} \in \mathbb{R}^{m \times n}} \{l(\pmb{L}, \pmb{S}, \pmb{Y}_{k})\}, \\ 
+			\pmb{Y}_{k+1} = \pmb{Y}_{k} + \mu(\pmb{X} - \pmb{L}_{k} - \pmb{S}_{k}),
 		\end{matrix}
@@
 
-We adopt RPCA with ADM, which, generally speaking, is a practical improvement of the classical Augmented Lagrangian method for solving convex programming problem with linear constraints, by fully taking advantage of its high-level separable structure \cite{yuan2009sparse}.
-ADM minimizes $\boldsymbol{L}$ and $\boldsymbol{S}$ variables serially by solving the following problems to generate the new iterate:
+We adopt RPCA with ADM, which, generally speaking, is a practical improvement of the classical ALM method for solving convex programming problem with linear constraints, by fully taking advantage of its high-level separable structure \cite{yuan2009sparse}.
+ADM minimizes $\pmb{L}$ and $\pmb{S}$ variables serially by solving the following problems to generate the new iterate:
 
@@
 	\left\{\begin{matrix}
-	\boldsymbol{L}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{L} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}, \boldsymbol{S}_{k}, \boldsymbol{Y}_{k})\}\\ 
-	\boldsymbol{S}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{S} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}_{k+1}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}\\ 
-	\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k})
+	\pmb{L}_{k+1} \in \operatorname*{argmin}_{\pmb{L} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}, \pmb{S}_{k}, \pmb{Y}_{k})\}\\ 
+	\pmb{S}_{k+1} \in \operatorname*{argmin}_{\pmb{S} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}_{k+1}, \pmb{S}, \pmb{Y}_{k})\}\\ 
+	\pmb{Y}_{k+1} = \pmb{Y}_{k} + \mu(\pmb{X} - \pmb{L}_{k} - \pmb{S}_{k})
 	\end{matrix}\right.
@@
 
-Let the mean $\bar{\boldsymbol{x}} \in \mathbb{R}^{1 \times p}$ be
+Let the mean $\bar{\pmb{x}} \in \mathbb{R}^{1 \times p}$ be
 
 \begin{equation}\label{eq:4.07}
-	\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i, 
+	\bar{\pmb{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} \pmb{x}_i, 
 \end{equation}
 
-and the covariance matrix $\boldsymbol{\bar{\Sigma}} \in \mathbb{R}^{p \times p}$ be
+and the covariance matrix $\bar{\pmb{\Sigma}} \in \mathbb{R}^{p \times p}$ be
 
 \begin{equation}\label{eq:4.08}
-	\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime,
+	\bar{\pmb{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} (\pmb{x}_i - \bar{\pmb{x}})(\pmb{x}_i - \bar{\pmb{x}})^\prime,
 \end{equation}
@@
 \begin{equation}\label{eq:4.09}
-	m_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}( x_i - \bar{x})^u.
+	\pmb{m}_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}(\pmb{x}_i - \bar{\pmb{x}})^u.
 \end{equation}
@@
 \begin{equation}\label{eq:4.10}
-	\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}},
+	\bar{\pmb{s}} = \frac{\pmb{m}_3}{\pmb{m}_2^{\frac{3}{2}}},
 \end{equation}
@@
 \begin{equation}\label{eq:4.11}
-	\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}.
+	\bar{\pmb{k}} = \frac{\pmb{m}_4}{\pmb{m}_2^2}.
 \end{equation}
 
-We propose to compute the mean $\boldsymbol{\bar{x}}$, the skewness $\boldsymbol{\bar{s}}$, the kurtosis $\boldsymbol{\bar{k}}$ and the covariance matrix $\boldsymbol{\bar{\Sigma}}$ from the robust subspace $\boldsymbol{L}^s$, after to minimize the Equation \ref{eq:4.06} for $\textbf{X}^s$.
+We propose to compute the mean $\bar{\pmb{x}}$, the skewness $\bar{\pmb{s}}$, the kurtosis $\bar{\pmb{k}}$ and the covariance matrix $\bar{\pmb{\Sigma}}$ from the robust subspace $\pmb{L}^s$, after to minimize the Equation \ref{eq:4.06} for $\pmb{X}^s$.
 We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between new observations and the moments calculated from a robust subspace computed by RPCA.
@@
 \begin{equation}\label{eq:4.12}
-	\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime},
+	\pmb{d}(\pmb{x},\bar{\pmb{x}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{x}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{x}})^\prime},
 \end{equation}
-where $\boldsymbol{x}$ is a vector of a new observation, $\bar{\boldsymbol{x}}$ is the mean vector of known observations, also referred as location, and $\boldsymbol{\bar{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter.
-The classical MD usually relies on robust mean and robust covariance matrix for outlier detection, which are commonly computed by Fast-MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}.
-Here we propose to calculate the MD based on mean and covariance matrix computed from a robust subspace learned by RPCA.
+where $\pmb{x}$ is a vector of a new observation, $\bar{\pmb{x}}$ is the mean vector of known observations, also referred as location, and $\bar{\pmb{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter.
+The classical MD usually relies on robust mean and robust covariance matrix for outlier detection, which are commonly computed by MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}.
+Here we propose to calculate the MD according to Equation \ref{eq:4.12} by means of the mean and covariance matrix computed from a robust subspace learned by RPCA.
+
 We also propose to extend the Equation \ref{eq:4.12} to implement a Skewness-based Mahalanobis Distance, as follows:
@@
 \begin{equation}\label{eq:4.13}
-	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}.
+	\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{s}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{s}})^\prime}.
 \end{equation}
 
-We finally propose to extend the Equation \ref{eq:4.12} to implement a Kurtosis-based Mahalanobis Distance, as follows:
+Finally, we propose to extend the Equation \ref{eq:4.12} to implement a Kurtosis-based Mahalanobis Distance, as follows:
 
 \begin{equation}\label{eq:4.14}
-	\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}.
+	\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{k}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{k}})^\prime}.
 \end{equation}
 
-The distances $\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})$, $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})$ and $\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})$ shall be computed and evaluated separately for botnet detection.
-Therefore, the robust subspace learning from normal data combined to MD can be called md-RPCA when using mean-based MD, sd-RPCA when using skewness-based MD, and kd-RPCA when using kurtosis-based MD.
+The distances $\pmb{d}(\pmb{x},\bar{\pmb{x}}, \bar{\pmb{\Sigma}})$, $\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}})$ and $\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}})$ shall be computed and evaluated separately and independently, for network attack detection.
+Therefore, the robust subspace learning from normal data combined to MD is called as md-RPCA when using mean-based MD, or is called sd-RPCA when using skewness-based MD, or kd-RPCA when using kurtosis-based MD.
 
@@
 The contamination rate can be well known for some areas, or can be computed by cross-validation or can be assumed according to previous observations.
-In our proposal, the contamination define the rate of the largest distances $\boldsymbol{d}$ that shall be classified as anomalous.
-Therefore, the resulting $\boldsymbol{\bar{t}}$ refers to observations classified as anomalous according to the contamination $c$.
+In our proposal, the contamination defines the number of the largest distances $\pmb{d}$ that shall be classified as anomalous.
+The observations classified as legitimate and anomalous, according to the contamination $c$, are denoted by the vector $\bar{\pmb{t}}$, with values of 1 for classifying anomalies and 0 to denote legitimate observation.
 
-The Algorithm \ref{alg:4.01} presents how the m-RPCA can be algorithmically described.
+The Algorithm \ref{alg:4.01} describes all possible steps of m-RPCA and the approaches md-RPCA, sd-RPCA and kd-RPCA, for the semi-supervised learning approach.
+The unsupervised approach adopts the same steps, but adopting the contaminated data for robust subspace learning, and requiring new robust subspace learning for testing anomaly detection on new set of observations.
+
 \begin{algorithm}
@@
 	\SetAlgoLined
-	\KwResult{$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}}$, $\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}}$}
-	Given $\boldsymbol{X}$ split into $\boldsymbol{X}^s$ and $\boldsymbol{X}^c$\;
-	\While{not $\min_{L,S}\left \| \boldsymbol{L} \right \|_{*} + \lambda \left \| \boldsymbol{S} \right \|_{1}$ of $\boldsymbol{X}^s$}{
-		$\boldsymbol{L}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{L} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}, \boldsymbol{S}_{k}, \boldsymbol{Y}_{k})\}$\;
-		$\boldsymbol{S}_{k+1} \in \operatorname*{argmin}_{\boldsymbol{S} \in \mathbb{R}^{m \times n}}\{l(\boldsymbol{L}_{k+1}, \boldsymbol{S}, \boldsymbol{Y}_{k})\}$\;
-		$\boldsymbol{Y}_{k+1} = \boldsymbol{Y}_{k} + \mu(\boldsymbol{X} - \boldsymbol{L}_{k} - \boldsymbol{S}_{k})$\;
+	\KwResult{$\bar{\pmb{t}}_{\bar{\pmb{x}}}$, $\bar{\pmb{t}}_{\bar{\pmb{s}}}$, $\bar{\pmb{t}}_{\bar{\pmb{k}}}$}
+	Given $\pmb{X}$ split into $\pmb{X}^s$ and $\pmb{X}^c$\;
+	\While{not $\min_{L,S}\left \| \pmb{L} \right \|_{*} + \lambda \left \| \pmb{S} \right \|_{1}$ of $\pmb{X}^s$}{
+		$\pmb{L}_{k+1} \in \operatorname*{argmin}_{\pmb{L} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}, \pmb{S}_{k}, \pmb{Y}_{k})\}$\;
+		$\pmb{S}_{k+1} \in \operatorname*{argmin}_{\pmb{S} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}_{k+1}, \pmb{S}, \pmb{Y}_{k})\}$\;
+		$\pmb{Y}_{k+1} = \pmb{Y}_{k} + \mu(\pmb{X} - \pmb{L}_{k} - \pmb{S}_{k})$\;
 	}
-	$\bar{\boldsymbol{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} \boldsymbol{x}_i$, for $\boldsymbol{L}^s$\;
-	$\boldsymbol{\bar{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \boldsymbol{h}} (\boldsymbol{x}_i - \bar{\boldsymbol{x}})(\boldsymbol{x}_i - \bar{\boldsymbol{x}})^\prime$, for $\boldsymbol{L}^s$\;
-	$\boldsymbol{\bar{s}} = \frac{m_3}{m_2^{\frac{3}{2}}}$, for $\boldsymbol{L}^s$\;
-	$\boldsymbol{\bar{k}} = \frac{m_4}{m_2^2}$, for $\boldsymbol{L}^s$\;
-	$\boldsymbol{d}(\boldsymbol{x},\bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{x}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{x}})^\prime}$\;
-	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{s}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{s}})^\prime}$\;
-	$\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}}) = \sqrt{(\boldsymbol{x} - \bar{\boldsymbol{k}}) \boldsymbol{\bar{\Sigma}}^{-1}(\boldsymbol{x} - \bar{\boldsymbol{k}})^\prime}$\;
-	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{x}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{x}}, \boldsymbol{\bar{\Sigma}})]^c$\;
-	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{s}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{s}}, \boldsymbol{\bar{\Sigma}})]^c$\;
-	$\boldsymbol{\bar{t}}_{\bar{\boldsymbol{k}}} = [\boldsymbol{d}(\boldsymbol{x}, \bar{\boldsymbol{k}}, \boldsymbol{\bar{\Sigma}})]^c$\;
+	$\bar{\pmb{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} \pmb{x}_i$, for $\pmb{L}^s$\;
+	$\bar{\pmb{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} (\pmb{x}_i - \bar{\pmb{x}})(\pmb{x}_i - \bar{\pmb{x}})^\prime$, for $\pmb{L}^s$\;
+	$\bar{\pmb{s}} = \frac{\pmb{m}_3}{\pmb{m}_2^{\frac{3}{2}}}$, for $\pmb{L}^s$\;
+	$\bar{\pmb{k}} = \frac{\pmb{m}_4}{\pmb{m}_2^2}$, for $\pmb{L}^s$\;
+	$\pmb{d}(\pmb{x},\bar{\pmb{x}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{x}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{x}})^\prime}$\;
+	$\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{s}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{s}})^\prime}$\;
+	$\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{k}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{k}})^\prime}$\;
+	$\bar{\pmb{t}}_{\bar{\pmb{x}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{x}}, \bar{\pmb{\Sigma}})]^c$\;
+	$\bar{\pmb{t}}_{\bar{\pmb{s}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}})]^c$\;
+	$\bar{\pmb{t}}_{\bar{\pmb{k}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}})]^c$\;
 	\caption{Moment Distances from Robust Subspace}
@@
 
-The steps between 1 and 10 of the Algorithm \ref{alg:4.01} are the training from normal data $\boldsymbol{X}^s$, while the steps between 11 and 16 aim the anomaly detection from new contaminated observations.
-Note that the steps 11 and 14 refer to the md-RPCA, while the steps 12 and 15 refer to the sd-RPCA, and the steps 13 and 16 refer to the kd-RPCA.
-It is also important to observe that the anomaly detection from new observations does not require new robust subspace learning, it is only necessary to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous.
+The steps between 1 and 10 of the Algorithm \ref{alg:4.01} are the training from normal data $\pmb{X}^s$, which are common steps shared by md-RPCA, sd-RPCA and kd-RPCA.
+The steps between 11 and 16 aim the anomaly detection from new contaminated observations, by means of Mahalanobis Distance of robust moments.
+Note that the steps 11 and 14 refer to the steps of md-RPCA for anomaly detection, while the steps 12 and 15 refer to the sd-RPCA, and the steps 13 and 16 refer to the steps of kd-RPCA.
+It is important to highlight that the anomaly detection from new observations does not require new robust subspace learning when adopting the semi-supervised approach, which only requires to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous or legitimate.
 
-The m-RPCA can also be adopted as unsupervised algorithm if the lines 1 and 2 of the Algorithm \ref{alg:4.01} are changed to substitute $\boldsymbol{X}^s$ by $\boldsymbol{X}^c$.
-Hence, the robust subspace is learned from contaminated data and used for comparing the distance between higher moments from robust estimate and moments of the observations of the same data.
-It is important to note that this unsupervised approach requires the computational cost of the subspace learning for any new observation.
+The m-RPCA can also be adopted as unsupervised algorithm if the lines 1 and 2 of the Algorithm \ref{alg:4.01} are changed to substitute $\pmb{X}^s$ by $\pmb{X}^c$.
+Hence, the robust subspace is learned from contaminated data and used for comparing the distance between moments from robust estimate and moments of contaminated observations.
+It is important to note that this unsupervised approach requires the computational cost of computing new subspace learning for any new set of observations.
 
@@
 In anomalous detection problems, where anomalies are rare events, if one classify all observations as normal and apply an accuracy evaluation, one would have high accuracy but poor true-positive detection.
-Due to the importance given by the F-measure to true-positive detection in scenarios such as the network attack detection, it is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic}.
-Therefore, F1-measure is the metric used for validation of our experiments.
+Due to the importance given by the $F_1$ (also referred as F-score or F-measure) to true-positive detection in scenarios such as the network attack detection, it is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic}.
+Therefore, $F_1$ is the metric used for validation of our experiments.
 
-The F-measure is the harmonic mean of precision and recall and is defined by:
+The $F_1$ is the harmonic mean of precision and recall and is defined by:
 \begin{equation}\label{eq:4.15}
-	F-measure = 2 * \frac{pr * rc}{pr + rc},
+	F_1 = 2 * \frac{Precision * Recall}{Precision + Recall},
 \end{equation}
@@
 \begin{equation}\label{eq:4.16}
-	pr = \frac{True Positive}{True Positive + False Positive}
+	Precision = \frac{True Positive}{True Positive + False Positive}
 \end{equation}
@@
 \begin{equation}\label{eq:4.17}
-	rc = \frac{True Positive}{True Positive + False Negative}.
+	Recall = \frac{True Positive}{True Positive + False Negative}.
 \end{equation}
 
-This experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection that also rely on contamination rate for anomaly detection, which are: PCA based on the sum of weighted projected distances to the eigenvector hyperplanes \cite{shyu2003novel}; MCD \cite{rousseeuw1984mcd,rousseeuw1999fastmcd}; One-Class Support Vector Machines \cite{scholkopf2001estimating}; Local Outlier Factor (LOF) \cite{breunig2000lof}; k-Nearest Neighbors \cite{angiulli2002fast}; and Isolation Forest \cite{liu2008isolation}.
+This experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection that also rely on contamination rate for anomaly detection, which are: A PCA approach based on the sum of weighted projected distances to the eigenvector hyperplanes \cite{shyu2003novel}; MCD \cite{rousseeuw1984mcd,rousseeuw1999fastmcd}; One-Class Support Vector Machines \cite{scholkopf2001estimating}; Local Outlier Factor (LOF) \cite{breunig2000lof}; k-Nearest Neighbors \cite{angiulli2002fast}; and Isolation Forest \cite{liu2008isolation}.
 
-% ToDO: We also evaluate the results of ROBPCA for anomaly detection on CTU-13, considering that ROBPCA also relies on robust estimates with adjusted outlyingness based on robust skewness \cite{hubert2009robustskewed}.
+We also compare the results of our proposals to ROBPCA for anomaly detection on simulated and CTU-13 data sets, considering that ROBPCA also relies on robust estimates with adjusted outlyingness based on robust skewness \cite{hubert2009robustskewed}.
 
@@
 Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training anomaly detection models, while the latter does not require labeled data or training.
+Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches.
 
-Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches.
-We propose semi-supervised and unsupervised approaches for m-RPCA, where the former relies on normal data $\boldsymbol{X}^s$ for training and on contamination rate $c$ for anomaly detection, while the latter relies only on contaminated data.
-We assume that the $c$ is well known or can be estimated for real world problems of anomaly detection, in accordance to the selected algorithms for comparison, that also rely on contamination rate.
+We propose semi-supervised and unsupervised approaches for m-RPCA, where the former relies on normal data $\pmb{X}^s$ for training and on contamination rate $c$ for anomaly detection, while the latter relies only on contaminated data $\pmb{X}^c$ for robust subspace learning and rely on contamination rate $c$ to select the largest distances.
+We assume that $c$ is well known or can be estimated for real world problems of anomaly detection, in accordance to the assumption adopted by the well established algorithms \cite{zhao2019pyod} selected for comparison, that also rely on contamination rate $\pmb{X}^c$.
 
 The availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or even unknown events.
-Considering that RPCA has already been adopted to isolate outliers from training data \cite{zhou2017anomaly}, we also propose to evaluate the m-RPCA from a semi-supervised approach based on training from contaminated data.
+Considering that RPCA has already been adopted to isolate outliers from training data \cite{zhou2017anomaly}, we also propose to evaluate the m-RPCA for a semi-supervised approach based on training from contaminated data, in order to evaluate the impact that contaminated data for robust subspace learning can cause in the anomaly detection results.
 
-Therefore, we propose to evaluate the following approaches for m-RPCA:
-\begin{enumerate}
-	\item Semi-supervised;
-	\item Contaminated Semi-supervised;
-	\item Unsupervised.
-\end{enumerate}
+Therefore, we propose to evaluate the following approaches for m-RPCA: Semi-supervised; Contaminated Semi-supervised; and Unsupervised.
 
-For the semi-supervised approach we propose to learn the robust subspace and compute the moments from the normal data set $\boldsymbol{Y}_g$, $\boldsymbol{Y}_p$ and $\boldsymbol{Y}_l$ with Gaussian, Pareto and Lognormal distributions, respectively, and test the anomaly detection for contaminated data set $\boldsymbol{Y}_g^c$, $\boldsymbol{Y}_p^c$ and $\boldsymbol{Y}_l^c$.
+For the semi-supervised approach we propose to learn the robust subspace and compute the moments from the normal data set $\pmb{Y}_g$, $\pmb{Y}_p$ and $\pmb{Y}_l$ with Gaussian, Pareto and Lognormal distributions, respectively, and test the anomaly detection for contaminated data set $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$.
 
-In the Contaminated Semi-supervised approach we evaluate the robustness of the m-RPCA for learning from contaminated training data, in order to analyse if m-RPCA can be an alternative for the lack of known normal data.
-Therefore we propose to train the model from a contaminated normal data $\boldsymbol{Y}_g^{c'}$, $\boldsymbol{Y}_p^{c'}$ and $\boldsymbol{Y}_l^{c'}$, with the same contamination rate of the testing data, but without data sharing between training and testing, taking into account that we shall consider a different contamination data but the same contamination rate.
+For the Contaminated Semi-supervised approach we evaluate the robustness of the m-RPCA approaches for learning from contaminated training data, in order to analyse if m-RPCA can be an alternative for the lack of known normal data.
+Therefore we propose to train the model from a contaminated normal data $\pmb{Y}_g^{c'}$, $\pmb{Y}_p^{c'}$ and $\pmb{Y}_l^{c'}$, with the same contamination rate of the testing data, but without data repetition between training and testing, taking into account that we shall consider a different contaminated data but adopt the same contamination rate for training and testing.
 
-We finally evaluate the unsupervised approach, that relies on the test data sets $\boldsymbol{Y}_g^c$, $\boldsymbol{Y}_p^c$ and $\boldsymbol{Y}_l^c$ for robust subspace learning, and then classify the results according to the distance between the contaminated data $\boldsymbol{Y}_g^c$, $\boldsymbol{Y}_p^c$ and $\boldsymbol{Y}_l^c$, and the moments computed from the learned robust subspace.
+We finally evaluate the unsupervised approach, that relies on the test data sets $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$ for robust subspace learning, and then classify the results according to the distance between moments of the contaminated data $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$, and the moments computed from the learned robust subspace.
 
@@
 
-In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approach to all botnet malwares of the CTU-13.
+In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approaches to all botnet malwares of the CTU-13 data set.
 
-In our experiment setup each contaminated scenario $\boldsymbol{X}_i$ is split into $\boldsymbol{X}_i^s$ containing 50\% of the the normal data, and into test data $\boldsymbol{X}_i^c$ containing 33\% of normal and 67\% of anomalous data.
+In our experiment setup each contaminated scenario $\pmb{X}_i$ of CTU-13 is split into $\pmb{X}_i^s$, containing 50\% of the the normal data, and into test data $\pmb{X}_i^c$ containing 33\% of normal and 67\% of anomalous data.
+We adopt a contamination rate $c$ of 33\% for  experimenting anomaly detection for the CTU-13 data set, considering that this contamination rate presented good results on the simulated experiment for our proposals and for some selected algorithms.
 
-This experiment consider only the semi-supervised approach due to results of simulated experiments on synthetic data set to be present in Subsection \ref{sec:4_simulated_result}, that shows better results for the semi-supervised approach and highlight that the this approach can obtain good results even when trained with contaminated data set.
+This experiment consider only the semi-supervised approach due to results of simulated experiments on simulated data set to be present in Subsection \ref{sec:4_simulated_result}, that shows better results for the semi-supervised approach and highlight that the this approach can obtain good results even when trained with contaminated data set.
 
@@
 
-We initially evaluated the F1-measure of the selected algorithms and m-RPCA approaches for anomaly detection on gaussian distributed legitimate data contaminated by uniform distributed anomalies.
+We adopt prefixes to denote the evaluated approaches for md-RPCA, sd-RPCA and kd-RPCA, which are ss\_ to denote semi-supervised, css\_ to denote contaminated semi-supervised and u\_ for unsupervised.
 
-The Figure \ref{fig:4.10} shows the F1-measure over the contamination rate between 1\% and 50\% for m-RPCA approaches and Isolation Forest (IF) \cite{liu2008isolation}, k-Nearest Neighbors (KNN) \cite{angiulli2002fast}, Local Outlier Factor (LOF) \cite{breunig2000lof}, Minimum Covariance Deteminant (MCD) \cite{rousseeuw1999fastmcd}, One-Class Support Vector Machines (OCSVM) \cite{scholkopf2001estimating} and PCA \cite{shyu2003novel}.
-We adopt prefixes to denote the adopted approaches for md-RPCA, sd-RPCA and kd-RPCA, which are ss\_ to denote semi-supervised, css\_ to denote contaminated semi supervised and u\_ for unsupervised.
-
-It is possible to observe that LOF, PCA, css\_kd-RPCA and u\_kd-RPCA present lower performance than the remain algorithms, that obtain results higher than 0.95 in average.
-Note that the css\_kd-RPCA and u\_kd-RPCA are the contaminated semi-supervised and unsupervised versions of kd-RPCA, that obtain worse results than the semi-supervised approach of kd-RPCA.
-
-However, the unsupervised versions of md-RPCA and sd-RPCA presented similar results to widely adopted unsupervised algorithms for outlier detection.
-The results also show that the semi-supervised approach of md-RPCA, sd-RPCA and kd-RPCA obtain high anomaly detection rate and presented similar results to the unsupervised algorithms IF, KNN, MCD and OCSVM.
+We initially evaluated the $F_1$ of the selected algorithms and m-RPCA approaches for anomaly detection on Gaussian distributed legitimate data contaminated by uniform distributed anomalies.
+The Figure \ref{fig:4.10} shows the $F_1$ over the contamination rate between 1\% and 50\% for m-RPCA approaches and Isolation Forest (IF) \cite{liu2008isolation}, k-Nearest Neighbors (KNN) \cite{angiulli2002fast}, Local Outlier Factor (LOF) \cite{breunig2000lof}, Minimum Covariance Deteminant (MCD) \cite{rousseeuw1999fastmcd}, One-Class Support Vector Machines (OCSVM) \cite{scholkopf2001estimating}, PCA \cite{shyu2003novel} and ROBPCA-AO \cite{hubert2009robustskewed}.
 
@@
 	\includegraphics[width=10cm]{figures/ch4/gaussian_f1_contamination.png}
-	\caption{Anomaly detection on gaussian distributed with uniform anomalies}
+	\caption{Anomaly detection on Gaussian distributed with uniform anomalies}
 	\label{fig:4.10}
@@
 
-The Figures \ref{fig:4.11} and \ref{fig:4.12} show the results for anomaly detection on skewed and heavy tailed distributions, where the Figure \ref{fig:4.11} presents the results for Pareto with gaussian anomalies, and the Figure \ref{fig:4.12} presents the results for Lognormal with gaussian anomalies.
+It is possible to observe in Figure \ref{fig:4.10} that LOF, PCA, css\_kd-RPCA and u\_kd-RPCA presented lower performance than the remain algorithms, that obtain results higher than 0.95 in average.
+The exception is the ROBPCA-AO, that presented high score for lower contamination but decreased with the contamination increasing.
+
+Note that the css\_kd-RPCA and u\_kd-RPCA are the contaminated semi-supervised and unsupervised versions of kd-RPCA, that obtain worse results than the semi-supervised approach of kd-RPCA, for anomaly detection on Gaussian distributed data contaminated by uniform anomalies.
+However, the unsupervised versions of md-RPCA and sd-RPCA presented similar results to widely adopted unsupervised algorithms for outlier detection.
+The results also show that the semi-supervised approach of md-RPCA, sd-RPCA and kd-RPCA obtain high anomaly detection rate and presented similar results to the unsupervised algorithms IF, KNN, MCD and OCSVM.
+
+The Figure \ref{fig:4.11} and \ref{fig:4.12} show the results for anomaly detection on skewed and heavy tailed distributions.
+The Figure \ref{fig:4.11} shows the results for Pareto with Gaussian anomalies.
 
@@
 	\includegraphics[width=10cm]{figures/ch4/pareto_f1_contamination.png}
-	\caption{Anomaly detection on Pareto distributed with gaussian anomalies}
+	\caption{Anomaly detection on Pareto distributed with Gaussian anomalies}
 	\label{fig:4.11}
@@
 
-The results for anomaly detection on Pareto with gaussian anomalies, depicted by Figure \ref{fig:4.11}, show that IF, KNN, LOF, OCSVM, css\_md-RPCA and u\_md-RPCA performed worse than the remain algorithms, with results lower than 0.6 even with higher contamination.
-Note that css\_md-RPCA and u\_md-RPCA are mean-based approaches of m-RPCA, however the ss\_md-RPCA presented lower results for lower contamination but achieved more than 0.8 with contamination near of 0.2 or higher.
+The results for anomaly detection on Pareto with Gaussian anomalies, depicted by Figure \ref{fig:4.11}, show that IF, KNN, LOF, OCSVM, css\_md-RPCA and u\_md-RPCA performed worse than the remain algorithms for the evaluated contamination, with results lower than 0.6 even with higher contamination.
+Note that the approaches of m-RPCA with lower scores are css\_md-RPCA and u\_md-RPCA, which are mean-based approaches.
+However, the ss\_md-RPCA is the mean-based approach that presented lower results for lower contamination but achieved more than 0.8 with contamination near of 0.2 or higher, and achieved results better than MCD and PCA.
 
-It is possible to observe in Figure \ref{fig:4.11} that all approaches based on kd-RPCA achieve the best results initially, but the results for the unsupervised and contaminated semi-supervised variate with contamination near of 0.2 or higher, while the ss\_kd-RPCA presents stable results near of 1.0 F1-measure for all evaluated contamination rates.
+Figure \ref{fig:4.11} shows that ROBPCA-AO presented high scores for low contamination rates, presenting results better than ss\_md-RPCA, MCD and PCA, initially.
+However, the results of ROBPCA-AO decrease drastically with the contamination increasing.
 
-The approaches based on sd-RPCA obtained anomaly detection higher than 0.8, however the unsupervised and contaminated semi-supervised presented high variation of F-measure and lower results in comparison to the semi-supervised sd-RPCA.
+It is possible to observe in Figure \ref{fig:4.11} that all approaches based on kd-RPCA achieved the best results initially, but the results for the unsupervised variate with contamination near of 0.2 or higher, while the ss\_kd-RPCA presents stable results near of 1.0 $F_1$ for all evaluated contamination rates.
 
+All approaches based on sd-RPCA obtained anomaly detection higher than 0.8, however the unsupervised and contaminated semi-supervised presented high variation of $F_1$ and lower results in comparison to the semi-supervised sd-RPCA.
 The contaminated semi-supervised approaches of sd-RPCA and kd-RPCA presented results higher than 0.8 and similar to the semi-supervised and unsupervised approaches of sd-RPCA and kd-RPCA.
@@
 However, u\_kd-RPCA and u\_sd-RPCA presented lower results than ss\_kd-RPCA and ss\_sd-RPCA.
+Therefore, the semi-supervised approaches of m-RPCA overcome other approaches of m-RPCA and overcome the selected algorithms for anomaly detection on Pareto distributed data with Gaussian anomalies.
+It is possible to highlight the semi-supervised kd-RPCA, which obtained stable results near of 1.0 $F_1$ for all evaluated contamination.
 
-Therefore, the semi-supervised approaches of m-RPCA overcome other approaches of m-RPCA and of the selected algorithms for anomaly detection on Pareto distributed data with gaussian anomalies.
-It is possible to highlight the semi-supervised kd-RPCA, which obtained stable results near of 1.0 F-measure for all evaluated contamination.
-
-The results for anomaly detection on Lognormal with gaussian anomalies, depicted by Figure \ref{fig:4.12}, show that IF, KNN, LOF, MCD, OCSVM, css\_md-RPCA and u\_md-RPCA perform worse than the remain algorithms, with results lower than 0.6 even with higher contamination.
+The results for anomaly detection on Lognormal with Gaussian anomalies, depicted by Figure \ref{fig:4.12}, show that IF, KNN, LOF, MCD, OCSVM, css\_md-RPCA, u\_md-RPCA and ROBPCA-AO perform worse than the remain evaluated algorithms, with results lower than 0.6 even with higher contamination.
 
@@
 	\includegraphics[width=10cm]{figures/ch4/lognormal_f1_contamination.png}
-	\caption{Anomaly detection on Lognormal distributed with gaussian anomalies}
+	\caption{Anomaly detection on Lognormal distributed with Gaussian anomalies}
 	\label{fig:4.12}
@@
 However, the results of ss\_md-RPCA and PCA are lower than all approaches of sd-RPCA and kd-RPCA.
+It is important to note that all mean-based approaches of m-RPCA presented lower results for anomaly detection on Lognormal data, in comparison to approaches based on skewness (sd-RPCA) and kurtosis (kd-RPCA), that achieved the best anomaly detection rates.
+However, the approaches of md-RPCA presented better results than IF, KNN, MCD, OCSVM and ROBPCA-AO.
 
-It is important to note that all mean-based approaches of m-RPCA presented lower results for anomaly detection on Lognormal data, in comparison to approaches based on skewness (sd-RPCA) and kurtosis (kd-RPCA), that achieved the best anomaly detection rates.
-However, the approaches of md-RPCA presented better results than IF, KNN, MCD and OCSVM.
-
-The approaches based on kd-RPCA presented higher detection rate for all contamination, with anomaly detection rate near of 1.0.
+The approaches based on kd-RPCA presented higher detection rate for all contamination, with scores near of 1.0.
 The semi-supervised and contaminated semi-supervised approaches for sd-RPCA performed similarly, but the unsupervised approach of sd-RPCA presented lower anomaly detection for lower contamination.
@@
 
-Therefore, it is possible to observe that the semi-supervised approaches of m-RPCA overcome other approaches of m-RPCA and all the selected algorithms for anomaly detection on Lognormal distributed data with gaussian anomalies.
+Therefore, it is possible to observe that the semi-supervised approaches of m-RPCA overcome other approaches of m-RPCA and all the selected algorithms for anomaly detection on Lognormal distributed data with Gaussian anomalies.
 
-Following we present the Tables \ref{tab:4.02},  \ref{tab:4.03} and \ref{tab:4.04} to show the results of the selected algorithms and of our proposals for anomaly detection on gaussian, Pareto and Lognormal distributions, with 10\%, 25\% and 35\% of contamination rate.
+Following we present the Tables \ref{tab:4.02}, \ref{tab:4.03} and \ref{tab:4.04} to show the results of the selected algorithms and of our proposals for anomaly detection on Gaussian, Pareto and Lognormal distributions, with 10\%, 25\% and 33\% of contamination rate.
 
@@
 	\begin{tabular}{| l | l | l | l |}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian + Uniform}\\\textbf{(F-Measure)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto + Gaussian}\\\textbf{(F-Measure)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal + Gaussian}\\\textbf{(F-Measure)}\end{tabular}\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\pmb{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Gaussian + Uniform}\\\pmb{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Pareto + Gaussian}\\\pmb{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\pmb{Lognormal + Gaussian}\\\pmb{($F_1$)}\end{tabular}\\ \hline
             Isolation Forest	&0.98	&0.44	&0.05\\ \hline
@@
             Unsupervised kd-RPCA	&0.74	&0.86	&0.99\\ \hline
+            ROBPCA-AO   &0.00	&0.00	&0.00\\ \hline
 	\end{tabular}
 \end{table}
+
+The results with 33\% of contamination shows high scores of anomaly detection for Gaussian data with uniform anomalies, with exception to LOF and ROBPCA-AO.
+The semi-supervised approaches of m-RPCA presented the highest results for anomaly detection on Pareto data with Gaussian anomalies, and for anomaly detection on Lognormal data with Gaussian anomalies.
 
@@
 	\begin{tabular}{| l | l | l | l |}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian + Uniform}\\\textbf{(F-Measure)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto + Gaussian}\\\textbf{(F-Measure)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal + Gaussian}\\\textbf{(F-Measure)}\end{tabular}\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\pmb{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Gaussian + Uniform}\\\pmb{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Pareto + Gaussian}\\\pmb{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\pmb{Lognormal + Gaussian}\\\pmb{($F_1$)}\end{tabular}\\ \hline
             Isolation Forest	&0.97	&0.46	&0.05\\ \hline
@@
             Unsupervised kd-RPCA	&0.62	&0.97	&0.99\\ \hline
+            ROBPCA-AO	&0.00	&0.00	&0.00\\ \hline
 	\end{tabular}
 \end{table}
+
+The Table \ref{tab:4.03} also shows higher scores of semi-supervised m-RPCA for 25\% of contamination, in comparison to the remain algorithms.
+It is important to note that the highest score for anomaly detection on Lognormal data with Gaussian contamination was the unsupervised kd-RPCA, and that the contaminated semi-supervised approaches presented scores near of the the results for semi-supervised approaches.
 
@@
 	\begin{tabular}{| l | l | l | l |}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian + Uniform}\\\textbf{(F-Measure)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto + Gaussian}\\\textbf{(F-Measure)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal + Gaussian}\\\textbf{(F-Measure)}\end{tabular}\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\pmb{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Gaussian + Uniform}\\\pmb{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Pareto + Gaussian}\\\pmb{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\pmb{Lognormal + Gaussian}\\\pmb{($F_1$)}\end{tabular}\\ \hline
             Isolation Forest	&0.97	&0.43	&0.00\\ \hline
@@
             Unsupervised kd-RPCA	&0.50	&0.97	&0.98\\ \hline
+            ROBPCA-AO	&0.97	&0.87	&0.00\\ \hline
 	\end{tabular}
 \end{table}
+
+The contamination rate of 10\% shown by Table \ref{tab:4.03} shows worse results in comparison to contamination of 25\% and 33\%.
+However, the ROBPCA-AO presented high scores for anomaly detection on Gaussian data, overcoming LOF, css\_kd-RPCA and u\_kd-RPCA.
+ROBPCA also presented better results for anomaly detection on Pareto data in comparison to u\_md-RPCA, css\_md-RCAP, ss\_md-RCAP, PCA, OCSVM, LOF, KNN and IF.
 
@@
 
-In this section we present the experiment on network anomaly detection from the CTU-13 data set, evaluating the results of md-RPCA, sd-RPCA, kd-RPCA, Isolation Forest (IF) \cite{liu2008isolation}, k-Nearest Neighbors (KNN) \cite{angiulli2002fast}, Local Outlier Factor (LOF) \cite{breunig2000lof}, Minimum Covariance Deteminant (MCD) \cite{rousseeuw1999fastmcd}, One-Class Support Vector Machines (OCSVM) \cite{scholkopf2001estimating} and PCA \cite{shyu2003novel}.
+In this section we present the experiment on network anomaly detection from the CTU-13 data set, evaluating the results of md-RPCA, sd-RPCA, kd-RPCA, Isolation Forest (IF) \cite{liu2008isolation}, k-Nearest Neighbors (KNN) \cite{angiulli2002fast}, Local Outlier Factor (LOF) \cite{breunig2000lof}, Minimum Covariance Deteminant (MCD) \cite{rousseeuw1999fastmcd}, One-Class Support Vector Machines (OCSVM) \cite{scholkopf2001estimating}, PCA \cite{shyu2003novel} and ROBPCA-AO \cite{hubert2009robustskewed}.
 
@@
 
-The Table \ref{tab:4.03} present the F-measure of each algorithm for all scenarios of CTU-13, and the last column presents the average F-measure of each algorithm for all scenarios.
-
 \begin{table}[h!]
@@
   \caption{Network anomaly detection from CTU-13 with 33\% of contamination}
-  \label{tab:4.03}
+  \label{tab:4.05}
   \begin{tabular}{ c|c|c|c|c|c|c|c|c|c|c|c|c|c|c }
 	\toprule
-        \textbf{Algorithm}	&\textbf{10}	&\textbf{11}	&\textbf{12}	&\textbf{15}	&\textbf{15-2}	&\textbf{15-3}	&\textbf{16}	&\textbf{16-2}	&\textbf{16-3}	&\textbf{17}	&\textbf{18}	&\textbf{18-2}	&\textbf{19}	&\textbf{Avg} \\ \hline
+        \pmb{Algorithm}	&\pmb{10}	&\pmb{11}	&\pmb{12}	&\pmb{15}	&\pmb{15-2}	&\pmb{15-3}	&\pmb{16}	&\pmb{16-2}	&\pmb{16-3}	&\pmb{17}	&\pmb{18}	&\pmb{18-2}	&\pmb{19}	&\pmb{Avg} \\ \hline
         IF	&0.36	&0.34	&0.09	&0.21	&0.40	&0.44	&0.16	&0.34	&0.34	&0.41	&0.12	&0.16	&0.46	&0.29\\ \hline
@@
         kd-RPCA &0.76	&0.76	&0.90	&0.82	&0.57	&0.76	&0.91	&0.50	&0.80	&0.73	&0.83	&0.81	&0.48	&0.74\\ \hline
+        ROBPCA-AO &0.01	&0.07	&0.00	&0.05	&0.38	&0.11	&0.05	&0.32	&0.03	&0.09	&0.07	&0.09	&0.21	&0.11\\ \hline
     \bottomrule
@@
 
-It is possible to observe that md-RPCA, kd-RPCA and sd-RPCA overcome the remain algorithms in average results for all scenarios, according to the column \textbf{avg}.
+The Table \ref{tab:4.03} present the $F_1$ of each algorithm for all scenarios of CTU-13, and the last column presents the average $F_1$ of each algorithm for all scenarios.
+
+It is possible to observe that md-RPCA, kd-RPCA and sd-RPCA overcome the remain algorithms in average results for all scenarios, according to the column \pmb{avg}.
 The sd-PRCA presented an average of 0.57 while md-RPCA obtained 0.81 and kd-RPCA 0.74 in average.
@@
 The md-RPCA algorithm presented an anomaly detection rate higher than 0.78 for almost all scenarios, with exception to scenario 19, where the anomaly detection rate of md-RPCA was 0.51.
-The anomalies of the scenario 19 are peer-to-peer botnet traffic generated by nsys.ay, which are related to botnet synchronization and not to network attacks, what can explain the low network detection rate of all evaluated algorithms, where the largest F1-measure was 0.61 achieved by PCA.
+The anomalies of the scenario 19 are peer-to-peer botnet traffic generated by nsys.ay, which are related to botnet synchronization and not to network attacks, what can explain the low network detection rate of all evaluated algorithms, where the largest $F_1$ was 0.61 achieved by PCA.
 
 The md-RPCA showed the best anomaly detection for 10 scenarios of a total with 13 scenarios.
-In the scenario 15 the best result was obtained by kd-RPCA, for the 15-2 scenario the best result was for MCD, and PCA was the best algorithm for anomaly detection on scenario 19.
-Even thought md-RPCA not be the best result for scenarios 15, 15-2 and 19, the results of md-RPCA are very close to the best results for scenarios 5 and 15-2.
+In the scenario 15 the best result was obtained by kd-RPCA, for the 15-2 scenario the best result was for MCD, and PCA was the best algorithm for scenario 19.
+Even thought md-RPCA not be the best result for scenarios 15, 15-2 and 19, the results of md-RPCA are very close to the best results for scenarios 15 and 15-2.
 
-It is important to note that IF, KNN, LOF and MCD algorithms performed poorly for unsupervised network anomaly detection on all scenarios of the CTU-13 data set, with average of 0.29, 0.11, 0.21 and 0.33 respectively.
+It is important to note that IF, KNN, LOF, MCD and ROBPCA-AO present the worse results for network attack detection on all scenarios of the CTU-13 data set, with average of 0.29, 0.11, 0.21, 0.33 and 0.11 respectively.
 From these algorithms, only MCD presented high result for one scenario, which was 0.79 for 15-2.
 
-Unfortunately, was not possible to observe Pareto or Lognormal distributions on features of CTU-13, even though the finds of researches that highlight the fitting between these distributions and Internet traffic \cite{benson2010network, leon2017probability}.
 The CTU-13 data set is very challenging for anomaly detection approaches, due to the high imbalance and large volume of flows.
 However, CTU-13 is one of the up-to-date data sets for network attack detection and is one that provides the data imbalance observed in real anomaly detection problems.
+Unfortunately, was not possible to observe Pareto or Lognormal distributions on features of CTU-13, even though the finds of researches that highlight the fitting between these distributions and Internet traffic \cite{benson2010network, leon2017probability}.
 
-The results of anomaly detection on CTU-13 reveals that m-RPCA algorithms obtain good results for network anomaly detection on contaminated data, overcoming widely adopted algorithms for outlier detection.
+The results of anomaly detection on CTU-13 reveals that m-RPCA algorithms obtain good results for network attack detection on contaminated data, overcoming widely adopted algorithms for outlier detection.
 
 \section{Conclusion}
-\label{sec:6_conclusion}
+\label{sec:4_conclusion}
 
 This work proposed the m-RPCA, which is approaches based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data.
-We evaluated the anomaly detection rate of m-RPCA on simulated data and on the CTU-13 data set, which is composed by network traffic of botnets, attacks and background flows.
+We evaluated the anomaly detection rate of m-RPCA for simulated data and for the CTU-13 data set, which is composed by network traffic of botnets, attacks and background flows.
 
-The m-RPCA can be divided into md-RPCA, sd-RPCA and kd-RPCA algorithms, to denote approaches of m-RPCA based distances of mean, skewness and kurtosis, respectively.
+The m-RPCA can be divided into md-RPCA, sd-RPCA and kd-RPCA algorithms, to denote approaches of m-RPCA based on distances of mean, skewness and kurtosis, respectively.
 We also proposed to evaluate m-RPCA for semi-supervised, contaminated semi-supervised and unsupervised methods of anomaly detection.
@@
 
-The main contributions of this work were: the development and evaluation of novel approaches for anomaly detection on skewed and imbalanced data sets, by means of moments computed from a robust subspace learned by RPCA.
+The main contributions of this work were: the development and evaluation of novel approaches for anomaly detection on skewed and imbalanced data sets, by means of moments computed from a robust subspace learned by RPCA; the evaluation of the proposed approaches for anomaly detection on simulated data set and for network attack detection on real botnet traffic data set.
 
-Future research can be directed to evaluate the application of the proposed approaches to different data sets and field of anomaly detection, as well as for the extension of m-RPCA to estimate the contamination or to not rely on previously defined contamination.
+Future research can be directed to evaluate the application of the proposed approaches to different data sets and anomaly detection problems.
+The m-RPCA can also be extended to be online and to learn new subspace in a adaptive fashion, by means of new robust subspace learning from new observations or through robust subspace tracking.

### 2019-06-14 docs/thesis/chapters/5_conclusion_future_work/conclusion.tex
after fix bold notation

@@
 \begin{itemize}
-	\item Improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets;
+	\item Improvements for obtaining better false positive rates, as well as to make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets;
 	\item Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high throughput network traffic;
-	\item Evaluate the application of the proposed approach to different anomaly detection problems, considering cases that are aware to behavioral analysis;
-	\item Enhance the m-RPCA approaches to estimate the contamination or not rely on previously defined contamination for anomaly detection;
+	\item Evaluate the application of the proposed approaches to different anomaly detection problems, considering cases that are aware to behavioral analysis;
+	\item Enhance the m-RPCA approaches to estimate the contamination or to not rely on previously defined contamination for anomaly detection;
+	\item The m-RPCA can also be extended to be online and to learn new subspace in a adaptive fashion, by means of new robust subspace learning from new observations or through robust subspace tracking.
 \end{itemize}

### 2019-06-15 docs/thesis/chapters/4_m_rpca/m_rpca.tex
fixing

@@
 
-Some widely adopted algorithms for anomaly detection assume a Gaussian distributed data, however this assumption may not be observed in some real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network, leon2017probability}.
+Some widely adopted algorithms for anomaly detection assume a Gaussian or symmetric distributed data, however this assumption may not be observed in some real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network, leon2017probability}.
 
@@
 Leon-Garcia \cite{leon2017probability} also argues that Pareto distribution has been found to capture the behavior of many quantities of interest in the study of Internet behavior.
-Moreover, Benson \emph{et al.}  \cite{benson2010network} observes that the lognormal distribution is the best fit to model arrival processes in a data center.
+Moreover, Benson \emph{et al.}  \cite{benson2010network} observes that the Lognormal distribution is the best fit to model arrival processes in a data center.
 
-These findings show that the skeweness and heavy-tailed distributions may be important for network traffic analysis, and can motivate researches to evaluate the impact of skewed data into algorithms that rely on Gaussian distributed data.
-These findings can also highlight opportunities to  exploit the skeweness and heavy-tailed distributions to obtain accurate classifiers for network anomaly detection.
+These findings show that the skewness and heavy-tailed distributions may be important for network traffic analysis, and can motivate researches to evaluate the impact of skewed data into algorithms that rely on Gaussian distributed data.
+These findings can also highlight opportunities to  exploit the skewness and heavy-tailed distributions to obtain accurate classifiers for network anomaly detection.
+
+An outlyingness-approach based on a robust estimator of skewness, combined with robust estimators of location and scale, can be able to flag the outlying measurements.
+When the same methodology would be used with non-robust estimators of location, scale and skewness, the outlyingness-values would be affected by the outliers such that the outlying group could be masked \cite{hubert2009robustskewed}.
 
@@
 
-Therefore, we propose the m-RPCA, which is an approach based on distances of moments computed from robust subspace learned by Robust Principal Component Analysis (RPCA), in order to detect anomalies from skewed data and network traffic.
+Therefore, we propose the m-RPCA, which is an approach based on distances between moments computed from a robust subspace learned by Robust Principal Component Analysis (RPCA) and contaminated observations, in order to detect anomalies from skewed data and network traffic.
 The proposed approach relies on a robust subspace computed from supposed normal observations, for estimating the moments to be used for distance analysis.
-The anomaly detection from new observations evaluate the Mahalanobis distance between the robust moments and the moments of new observations, in a semi-supervised fashion, without the computational cost of new robust subspace learning for new observations.
+The anomaly detection from contaminated observations evaluate the Mahalanobis distance between the robust moments and new contaminated observations, in a semi-supervised fashion, without the computational cost of new robust subspace learning for new observations.
 The m-RPCA can also be computed as an unsupervised algorithm, with subspace learning from the same contaminated data that is the target of the anomaly detection analysis.
@@
 The authors recommend scenarios for training and testing in order to avoid the use of traffic from a botnet family for training and testing, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
-Considering the adoption of unsupervised or semi-supervised approaches for anomaly detection, adopting the training and testing approach proposed by Garcia \emph{et al.}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
+Taking into account the adoption of unsupervised or semi-supervised approaches for anomaly detection, adopting the training and testing approach proposed by Garcia \emph{et al.}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
 
@@
 It is a multi-dimensional generalization for measuring how many standard deviations away $\pmb{x}$ is from the mean $\bar{\pmb{x}}$ and covariance $\bar{\pmb{\Sigma}}$ of $\pmb{X}$.
-MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and new observations can reveal anomalies.
+MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and contaminated observations can reveal anomalies.
 
-We propose an approach based on the distances of moments computed from learned robust subspace and new observations, for anomaly detection on skewed and imbalanced data sets, and evaluate our approach for network attack detection.
+We propose an approach based on the distances between moments computed from learned robust subspace and contaminated observations, for anomaly detection on skewed and imbalanced data sets, and evaluate our approach for network attack detection.
 The proposed approach relies on learning robust subspace from supposed normal traffic for estimating the moments (mean, skewness and kurtosis).
-Thus, the anomaly detection for new observations should evaluate the Mahalanobis Distance between the robust moments and the moments of new observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations, or in a unsupervised approach, where the robust subspace is learned from the contaminated data and the distance analysis is between the robust moments and the contaminated moments.
+Thus, the anomaly detection for contaminated observations should evaluate the Mahalanobis Distance between the robust moments and contaminated observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations, or in a unsupervised approach, where the robust subspace is learned from the contaminated data and the distance analysis is between the robust moments and the contaminated observations.
 
@@
 
-The CTU-13 data set originally contains the following features for each flow:
-
-\begin{itemize}
-	\item Start Time;
-	\item End Time;
-	\item Duration;
-	\item Source IP Address;
-	\item Source Port;
-	\item Direction;
-	\item Destination IP Address;
-	\item Destination Port;
-	\item State of TCP flags;
-	\item Destination Type of Service;
-	\item Source Type of Service;
-	\item Total number of Packets;
-	\item Total number of Bytes;
-\end{itemize}
+The CTU-13 data set originally contains the following features for each flow: Start Time, End Time, Duration, Source IP Address, Source Port.
+Direction, Destination IP Address, Destination Port, State of TCP flags, Destination Type of Service, Source Type of Service, Total number of Packets, Total number of Bytes.
 
@@
 
-This section describes the proposed approach for network attack detection by means of a distance analysis between moments computed from a robust subspace and moments of new observations of network traffic.
+This section describes the proposed approach for network attack detection by means of a distance analysis between moments computed from a robust subspace and contaminated observations of network traffic.
 
@@
 We propose to learn a robust subspace $\pmb{L}$ from the normal traffic $\pmb{X}^s$ for computing $\pmb{L}^s$, $\pmb{S}^s$ and the robust moments, i.e.
-the mean $\bar{\pmb{x}}$, skewness $\bar{\pmb{s}}$ and kurtosis $\bar{\pmb{k}}$, in order to evaluate the distance $\pmb{d}$ between moments of new observations $\pmb{X}^c$ and moments computed from the learned robust subspace, in order to classify as anomalous the observations with the largest distances.
+the mean $\bar{\pmb{x}}$, skewness $\bar{\pmb{s}}$ and kurtosis $\bar{\pmb{k}}$, in order to evaluate the distance $\pmb{d}$ between contaminated observations $\pmb{X}^c$ and moments computed from the learned robust subspace, in order to classify as anomalous the observations with the largest distances.
 
@@
 We propose to compute the mean $\bar{\pmb{x}}$, the skewness $\bar{\pmb{s}}$, the kurtosis $\bar{\pmb{k}}$ and the covariance matrix $\bar{\pmb{\Sigma}}$ from the robust subspace $\pmb{L}^s$, after to minimize the Equation \ref{eq:4.06} for $\pmb{X}^s$.
-We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between new observations and the moments calculated from a robust subspace computed by RPCA.
+We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between contaminated observations and the moments calculated from a robust subspace computed by RPCA.
 
@@
 The m-RPCA can also be adopted as unsupervised algorithm if the lines 1 and 2 of the Algorithm \ref{alg:4.01} are changed to substitute $\pmb{X}^s$ by $\pmb{X}^c$.
-Hence, the robust subspace is learned from contaminated data and used for comparing the distance between moments from robust estimate and moments of contaminated observations.
+Hence, the robust subspace is learned from contaminated data and used for comparing the distance between moments from robust estimate and contaminated observations.
 It is important to note that this unsupervised approach requires the computational cost of computing new subspace learning for any new set of observations.
@@
 The availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or even unknown events.
-Considering that RPCA has already been adopted to isolate outliers from training data \cite{zhou2017anomaly}, we also propose to evaluate the m-RPCA for a semi-supervised approach based on training from contaminated data, in order to evaluate the impact that contaminated data for robust subspace learning can cause in the anomaly detection results.
+Considering that RPCA has already been adopted to isolate outliers from training data \cite{zhou2017anomaly}, we also propose to evaluate the m-RPCA for a contaminated semi-supervised approach based on training from contaminated data, in order to evaluate the impact that contaminated data for robust subspace learning can cause in the anomaly detection results.
 
-Therefore, we propose to evaluate the following approaches for m-RPCA: Semi-supervised; Contaminated Semi-supervised; and Unsupervised.
+Therefore, we propose to evaluate the following approaches for m-RPCA: semi-supervised; contaminated semi-supervised; and unsupervised.
 
@@
 
-We finally evaluate the unsupervised approach, that relies on the test data sets $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$ for robust subspace learning, and then classify the results according to the distance between moments of the contaminated data $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$, and the moments computed from the learned robust subspace.
+We finally evaluate the unsupervised approach, that relies on the test data sets $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$ for robust subspace learning, and then classify the results according to the distance between the contaminated data $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$, and the moments computed from the learned robust subspace.
 
@@
 
-This experiment consider only the semi-supervised approach due to results of simulated experiments on simulated data set to be present in Subsection \ref{sec:4_simulated_result}, that shows better results for the semi-supervised approach and highlight that the this approach can obtain good results even when trained with contaminated data set.
+This experiment consider only the semi-supervised approach due to results of simulated experiments on simulated data set presented in Subsection \ref{sec:4_simulated_result}, that shows better results for the semi-supervised approach and highlight that the this approach can obtain good results even when trained with contaminated data set.
 

### 2019-06-16 docs/thesis/glossaries-lists/acronyms.tex
new template

% Example for Glossary:
% \newglossaryentry{latex}
% {
%     name=latex,
%     description={Is a mark up language specially suited 
%     for scientific documents}
% }
%
% To include acronyms into the list of symbols, do as follow:
% * add [parent=acronymslists] after every "\newacronym";
% * Comment the line "\listadeacronimos" in file "thesis.tex";
% * Remove the comment the following line:
%\newglossaryentry{acronymslists}{name={Acronyms list},description={\glspar},sort=c}

\newglossaryentry{maths}
{
    name=mathematic,
    description={Mathematics is what mathematicians do}
}

% Example for Acronyms:
% \newacronym{label}{Acronym}{Full name}
\newacronym{uti}{UTI}{Unidade de Tratamento Intensivo}
\newacronym{scuba}{SCUBA}{Self-Contained Underwater Breathing Apparatus}
\newacronym{radar}{RADAR}{RAdio Detection And Ranging}
\newacronym[description={Data Link layer/protocols}]{dl}{DL}{Data link}

% In the text:
%     use \gls{label}, for an first time extend name, plus acronyms
%     use \acrshort{label}, for an acronym only
%     use \acrlong{label}, for a full name only
% In the text , for plural:
%     \glspl{duck}
% In the text , for Caps and/or plural:
%     \Gls{duck}
%     \Glspl{duck}

### 2019-06-16 docs/thesis/parts/3_m_rpca.tex
new template

--- 2019-06-15 docs/thesis/chapters/4_m_rpca/m_rpca.tex
@@
 \label{ch:4_m_rpca}
-
-\begin{quotation}[]{Paulo Freire}
-No one knows it all. No one is ignorant of everything. We all know something. We are all ignorant of something.
-\end{quotation}
 
@@
 \begin{figure}[!htb]
-	\centering
-	\subfigure[Pareto and Gaussian Anomalies]{%
-		\includegraphics[width=0.48\textwidth]{figures/ch4/4_Xpg2.png}
-		\label{fig:4.01}
-	}
-	\subfigure[Lognormal and Gaussian Anomalies]{%
-		\includegraphics[width=0.48\textwidth]{figures/ch4/4_Xlogng2.png}
-		\label{fig:4.02}
-	}
-	\caption[Examples of skewed and heavy tailed distributions]{Examples of skewed and heavy tailed distributions}
-	\label{fig:4.03}
+	% \centering
+	% \subfigure[Pareto and Gaussian Anomalies]{%
+	% 	\includegraphics[width=0.48\textwidth]{figs/ch4/4_Xpg2.png}
+	% 	\label{fig:4.01}
+	% }
+	% \subfigure[Lognormal and Gaussian Anomalies]{%
+	% 	\includegraphics[width=0.48\textwidth]{figs/ch4/4_Xlogng2.png}
+	% 	\label{fig:4.02}
+	% }
+	% \caption[Examples of skewed and heavy tailed distributions]{Examples of skewed and heavy tailed distributions}
+	% \label{fig:4.03}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=8cm]{figures/ch4/4_Xgu2.png}
+	\includegraphics[width=8cm]{figs/ch4/4_Xgu2.png}
 	\caption{Example of Gaussian and Uniform Anomalies}
@@
 \begin{figure}[!htb]
-	\centering
-	\subfigure[State of scenario 16.]{%
-		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110816_State.png}
-		\label{fig:4.05}
-	}
-	\subfigure[Destination Type of service of scenario 10.]{%
-		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110810_dTos.png}
-		\label{fig:4.06}
-	}
-	\caption[Skewness and Overlapping]{Example of skewness and overlapping}
-	\label{fig:4.07}
+	% \centering
+	% \subfigure[State of scenario 16.]{%
+	% 	\includegraphics[width=0.47\textwidth]{figs/ch4/raw_distplot_capture20110816_State.png}
+	% 	\label{fig:4.05}
+	% }
+	% \subfigure[Destination Type of service of scenario 10.]{%
+	% 	\includegraphics[width=0.47\textwidth]{figs/ch4/raw_distplot_capture20110810_dTos.png}
+	% 	\label{fig:4.06}
+	% }
+	% \caption[Skewness and Overlapping]{Example of skewness and overlapping}
+	% \label{fig:4.07}
 \end{figure}
@@
 \begin{figure}[!htb]
-	\centering
-	\subfigure[State of scenario 10.]{%
-		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110810_State.png}
-		\label{fig:4.08}
-	}
-	\subfigure[Source Port of scenario 16.]{%
-		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110816_Sport.png}
-		\label{fig:4.09}
-	}
-	\caption[Skewness and Overlapping]{Skewness and Overlapping}
-	\label{fig:4.10}
+	% \centering
+	% \subfigure[State of scenario 10.]{%
+	% 	\includegraphics[width=0.47\textwidth]{figs/ch4/raw_distplot_capture20110810_State.png}
+	% 	\label{fig:4.08}
+	% }
+	% \subfigure[Source Port of scenario 16.]{%
+	% 	\includegraphics[width=0.47\textwidth]{figs/ch4/raw_distplot_capture20110816_Sport.png}
+	% 	\label{fig:4.09}
+	% }
+	% \caption[Skewness and Overlapping]{Skewness and Overlapping}
+	% \label{fig:4.10}
 \end{figure}
@@
 	\label{alg:4.01}
-	\SetAlgoLined
-	\KwResult{$\bar{\pmb{t}}_{\bar{\pmb{x}}}$, $\bar{\pmb{t}}_{\bar{\pmb{s}}}$, $\bar{\pmb{t}}_{\bar{\pmb{k}}}$}
-	Given $\pmb{X}$ split into $\pmb{X}^s$ and $\pmb{X}^c$\;
-	\While{not $\min_{L,S}\left \| \pmb{L} \right \|_{*} + \lambda \left \| \pmb{S} \right \|_{1}$ of $\pmb{X}^s$}{
-		$\pmb{L}_{k+1} \in \operatorname*{argmin}_{\pmb{L} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}, \pmb{S}_{k}, \pmb{Y}_{k})\}$\;
-		$\pmb{S}_{k+1} \in \operatorname*{argmin}_{\pmb{S} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}_{k+1}, \pmb{S}, \pmb{Y}_{k})\}$\;
-		$\pmb{Y}_{k+1} = \pmb{Y}_{k} + \mu(\pmb{X} - \pmb{L}_{k} - \pmb{S}_{k})$\;
-	}
-	$\bar{\pmb{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} \pmb{x}_i$, for $\pmb{L}^s$\;
-	$\bar{\pmb{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} (\pmb{x}_i - \bar{\pmb{x}})(\pmb{x}_i - \bar{\pmb{x}})^\prime$, for $\pmb{L}^s$\;
-	$\bar{\pmb{s}} = \frac{\pmb{m}_3}{\pmb{m}_2^{\frac{3}{2}}}$, for $\pmb{L}^s$\;
-	$\bar{\pmb{k}} = \frac{\pmb{m}_4}{\pmb{m}_2^2}$, for $\pmb{L}^s$\;
-	$\pmb{d}(\pmb{x},\bar{\pmb{x}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{x}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{x}})^\prime}$\;
-	$\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{s}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{s}})^\prime}$\;
-	$\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{k}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{k}})^\prime}$\;
-	$\bar{\pmb{t}}_{\bar{\pmb{x}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{x}}, \bar{\pmb{\Sigma}})]^c$\;
-	$\bar{\pmb{t}}_{\bar{\pmb{s}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}})]^c$\;
-	$\bar{\pmb{t}}_{\bar{\pmb{k}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}})]^c$\;
+	% \SetAlgoLined
+	% \KwResult{$\bar{\pmb{t}}_{\bar{\pmb{x}}}$, $\bar{\pmb{t}}_{\bar{\pmb{s}}}$, $\bar{\pmb{t}}_{\bar{\pmb{k}}}$}
+	% Given $\pmb{X}$ split into $\pmb{X}^s$ and $\pmb{X}^c$\;
+	% \While{not $\min_{L,S}\left \| \pmb{L} \right \|_{*} + \lambda \left \| \pmb{S} \right \|_{1}$ of $\pmb{X}^s$}{
+	% 	$\pmb{L}_{k+1} \in \operatorname*{argmin}_{\pmb{L} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}, \pmb{S}_{k}, \pmb{Y}_{k})\}$\;
+	% 	$\pmb{S}_{k+1} \in \operatorname*{argmin}_{\pmb{S} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}_{k+1}, \pmb{S}, \pmb{Y}_{k})\}$\;
+	% 	$\pmb{Y}_{k+1} = \pmb{Y}_{k} + \mu(\pmb{X} - \pmb{L}_{k} - \pmb{S}_{k})$\;
+	% }
+	% $\bar{\pmb{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} \pmb{x}_i$, for $\pmb{L}^s$\;
+	% $\bar{\pmb{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} (\pmb{x}_i - \bar{\pmb{x}})(\pmb{x}_i - \bar{\pmb{x}})^\prime$, for $\pmb{L}^s$\;
+	% $\bar{\pmb{s}} = \frac{\pmb{m}_3}{\pmb{m}_2^{\frac{3}{2}}}$, for $\pmb{L}^s$\;
+	% $\bar{\pmb{k}} = \frac{\pmb{m}_4}{\pmb{m}_2^2}$, for $\pmb{L}^s$\;
+	% $\pmb{d}(\pmb{x},\bar{\pmb{x}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{x}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{x}})^\prime}$\;
+	% $\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{s}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{s}})^\prime}$\;
+	% $\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{k}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{k}})^\prime}$\;
+	% $\bar{\pmb{t}}_{\bar{\pmb{x}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{x}}, \bar{\pmb{\Sigma}})]^c$\;
+	% $\bar{\pmb{t}}_{\bar{\pmb{s}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}})]^c$\;
+	% $\bar{\pmb{t}}_{\bar{\pmb{k}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}})]^c$\;
 	\caption{Moment Distances from Robust Subspace}
@@
 	\centering
-	\includegraphics[width=10cm]{figures/ch4/gaussian_f1_contamination.png}
+	\includegraphics[width=10cm]{figs/ch4/gaussian_f1_contamination.png}
 	\caption{Anomaly detection on Gaussian distributed with uniform anomalies}
@@
 	\centering
-	\includegraphics[width=10cm]{figures/ch4/pareto_f1_contamination.png}
+	\includegraphics[width=10cm]{figs/ch4/pareto_f1_contamination.png}
 	\caption{Anomaly detection on Pareto distributed with Gaussian anomalies}
@@
 	\centering
-	\includegraphics[width=10cm]{figures/ch4/lognormal_f1_contamination.png}
+	\includegraphics[width=10cm]{figs/ch4/lognormal_f1_contamination.png}
 	\caption{Anomaly detection on Lognormal distributed with Gaussian anomalies}

### 2019-06-16 docs/thesis/parts/abstract.tex
new template

%PORTUGUESE
\resumo{Resumo}{

Resumo em Português, espaçamento de 1.5 requerido pela norma (automático).

}
%ENGLISH
\vspace{2cm}
\resumo{Abstract}{

Traditionally, cyber defense methods can be effective against well known types of attacks, yet may fail against innovative malicious techniques. In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency in resource consumption, processing capacity and response time. Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns. Recently, signal processing schemes have been applied to detect malicious traffic in computer networks, showing advances in network traffic analysis.

Information security may consist of both technical and procedural aspects. The former includes equipment and security systems, while the latter corresponds to security rules and recommendations of good practices. Intrusion detection and intrusion prevention systems are two types of security systems that are being used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks. Such systems can work in the following fashions: signature-based, anomaly-based or hybrid .

Anomaly detection can be defined as the identification of rare and suspicious events by comparing it to the normal data. Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects. Additionally, anomaly detection techniques can be categorized by classification, statistical algorithms, information theory and cluster based algorithms. Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events. Therefore anomaly detection algorithms have to be highly discriminating, robust to corruption and able to deal with the imbalanced data problem.

Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or behavioral analysis. However, distributed attacks organized by botnet has increased and demanded the development of counter measures. This has been done to detect and avoid unknown attacks or even to deal with adversarial changes of behavior, location and other patterns. This has also been done in order to avoid detection by behavioral or pattern based detection systems. To face this problem, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, where it is not necessary known anomalies for training models.

Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate. This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security. Despite the actual high availability of information, the relevant information of some observations is generally of under much reduced dimensionality compared to available data sets. The extraction of relevant information by identifying the generating causes within classes of signals is useful for classification problems and for security analysis.

The high availability of raw data increases the challenges related to big data analytics and to imbalanced learning problems, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes. The fundamental issue with the imbalanced learning is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms. Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection.

Some widely adopted algorithms for anomaly detection assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions. It was observed that the skeweness and heavy-tailed distributions can impact algorithms that rely on gaussian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.

Therefore, as a first important contribution, we propose the Eigensimilarity, which is an approach based on signal processing concepts applied to detection of malicious traffic in computer networks. Eigensimilarity models the network traffic using a signal processing formulation as a composition of three components: signal, artifact and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP). The Eigensimilarity is based on eigenvalue analysis, model order selection (MOS) and similarity analysis between eigenvectors, where MOS and eigenvalue analysis are applied to detect time frames under attack. We evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 data set, which is a well known network traffic data set. The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, i.e. in an unsupervised approach, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.

Some widely adopted algorithms for anomaly detection assume a Gaussian or symmetric distributed data, however this assumption may not be observed in some real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network, leon2017probability}.

Findings of Benson \emph{et al.}  \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}. Leon-Garcia \cite{leon2017probability} also argues that Pareto distribution has been found to capture the behavior of many quantities of interest in the study of Internet behavior. Moreover, Benson \emph{et al.}  \cite{benson2010network} observes that the Lognormal distribution is the best fit to model arrival processes in a data center.

These findings show that the skewness and heavy-tailed distributions may be important for network traffic analysis, and can motivate researches to evaluate the impact of skewed data into algorithms that rely on Gaussian distributed data. These findings can also highlight opportunities to  exploit the skewness and heavy-tailed distributions to obtain accurate classifiers for network anomaly detection.

An outlyingness-approach based on a robust estimator of skewness, combined with robust estimators of location and scale, can be able to flag the outlying measurements. When the same methodology would be used with non-robust estimators of location, scale and skewness, the outlyingness-values would be affected by the outliers such that the outlying group could be masked \cite{hubert2009robustskewed}.

Network anomaly detection problems are usually characterized by imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}. However, learning algorithms for imbalanced data has been a challenging research topic, considering that the fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}. Therefore, learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.

We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data. We also believe that the distance between robust estimates of normal traffic and contaminated observations can highlight discrepancies and be used for network attack detection.

Therefore, we propose the m-RPCA, which is an approach based on distances between moments computed from a robust subspace learned by Robust Principal Component Analysis (RPCA) and contaminated observations, in order to detect anomalies from skewed data and network traffic. The proposed approach relies on a robust subspace computed from supposed normal observations, for estimating the moments to be used for distance analysis. The anomaly detection from contaminated observations evaluate the Mahalanobis distance between the robust moments and new contaminated observations, in a semi-supervised fashion, without the computational cost of new robust subspace learning for new observations. The m-RPCA can also be computed as an unsupervised algorithm, with subspace learning from the same contaminated data that is the target of the anomaly detection analysis.

We evaluate the accuracy of the m-RPCA for anomaly detection on simulated data sets, with skewed and heavy-tailed distributions, and for the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}. The Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches. We also compare the results to ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.

The main contribution of this work is the proposal of a novel semi-supervised and unsupervised method for anomaly detection in skewed and imbalanced data, with results of experimental evaluation on simulated and real data sets.

This Chapter is organized as follows. In Section \ref{sec:4_relatedworks}, it is conducted a literature review of network anomaly detection, botnet detection, and imbalanced learning. Section \ref{sec:4_datamodel} presents the data model and the evaluated data set. Section \ref{sec:4_proposal} describes the proposed approach for network attack detection. Section \ref{sec:4_experiments} discusses the experimental validation, Section \ref{sec:4_simulated_result} presents the results of the simulated scenarios and the Section \ref{sec:4_ctu13_result} describes the results for anomaly detection on CTU-13 data set. Finally, Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.

The protection scheme used in a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}. Therefore, encrypting files and generating keys on a mobile device is not considered a good solution. On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}. The common practice is the shadow user activities monitoring \cite{yovel2014}. However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode. When the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools can not help and the mobile client has to secure itself with its own limited resources. Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection.

Additional security issues and requirements have to be considered when mobile clients are actively used in corporate cloud environment \cite{yovel2014}. Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm. The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS). While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, such as CASB and CAC, the corporate mobile client is usually light-weighted and generally less protected.

This work proposes an architecture and approach for user behavior analysis based on Eigensim \cite{vieira2017model}, in order to detect possible threats in offline corporate mobile applications. The key expiration period is safely incorporated into the proposed system solution in order to enhance security and the behavioral analysis can indicate malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.

The work is structured as follows. Section \ref{sec:3_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world. Section \ref{sec:3_the_client_security_architecture} outlines the mobile security architecture of the proposed solution. Section \ref{sec:3_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security. Section \ref{sec:3_eigensim} explains the use of Eigensim. Section \ref{sec:3_results} discusses the common threat scenarios, the data modeling, the performance analysis and discussion of the practical implementation results. Section \ref{sec:3_conclusion} concludes the chapter.

}

### 2019-06-16 docs/thesis/parts/b_csf_pca_svm.tex
new template

--- 2017-11-15 docs/thesis/chapters/appendix/b_pca_svm_cfa/pca_svm_cfa.tex
@@
      \centering 
-     \includegraphics[height=10cm, width=10cm]{figures/ch2/raw_igovti_covariance.eps}
+     \includegraphics[height=10cm, width=10cm]{figs/b_csf_pca_svm/raw_igovti_covariance.eps}
      \caption{Covariance matrix of iGovTI questions.}
@@
      \centering 
-     \includegraphics[height=8cm, width=11cm]{figures/ch2/raw_variance_ecdf.eps}
+     \includegraphics[height=8cm, width=11cm]{figs/b_csf_pca_svm/raw_variance_ecdf.eps}
      \caption{Empirical CDF of variance.}
@@
      \centering 
-     \includegraphics[height=8cm, width=11cm]{figures/ch2/raw_igovti_ranking_pc2.eps}
+     \includegraphics[height=8cm, width=11cm]{figs/b_csf_pca_svm/raw_igovti_ranking_pc2.eps}
      \caption{iGovTI ranking from 2 principal components.}

### 2019-06-16 docs/thesis/thesis.tex
new template

--- 2017-11-09 docs/thesis/thesis.tex
@@
-% class
-\documentclass[en, oneside, onehalfspacing]{risethesis}
+%% These files were modified in Lyx by:
+%           Lus F. C. Fgueiredo, Fev., 2017.
+%% Git repo for Lyx version are available at:
+%%          https://github.com/lara-unb/DOCUMENTS/tree/master/templates
+%% Latex version for sharelatex and overleaf v2 ported by:
+%%          Vinícius Galvão Guimarães
+%% 
+%% Contributions of:
+%%          Miguel Eduardo Gutierrez Paredes
+%% LyX 2.2.3 created this file.  For more info, see http://www.lyx.org/.
 
-% packages and configurations
-\usepackage{enumitem}
-\usepackage{lineno, hyperref}
-\usepackage[numbers]{natbib}
-\usepackage{babel}
-\usepackage{supertabular}
-\usepackage{fancybox}
-\usepackage{acronym}
-\usepackage{array}
-\usepackage{booktabs}
-\usepackage{graphicx}
-\usepackage{rotating}
-\usepackage{tabularx}
-\usepackage{color, colortbl}
-\usepackage{multirow}
-\usepackage{hhline}
-\usepackage{setspace}
-\usepackage{placeins}
-\usepackage{longtable}
-\usepackage{amsthm,amsmath}
-\usepackage{mathtools}
-\usepackage{algorithm}% http://ctan.org/pkg/algorithms
-\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
-\usepackage{algpseudocode}
-\usepackage{listings}
-\usepackage{epstopdf}
-\usepackage{subfigure}
-\usepackage{courier}
-\usepackage{amsfonts}
-\usepackage{morefloats}
-\usepackage{lipsum}
-\usepackage{cleveref}
-\usepackage{hyperref}
+%\documentclass[11pt,oneside,portuguese]{book}
+\documentclass[12pt,oneside,english]{book}
+\input{setup/packages}
+\input{setup/other}
+\input{setup/theorem_lemma}
+\input{setup/languages}
+\input{setup/listing_def}
+\input{setup/glossaries_setup}
 
-\newtheorem{thm}{Scenario}
-\hypersetup{breaklinks=true}
-\urlstyle{same}
-\Urlmuskip=0mu plus 1mu\relax
-\lstset{numbers=left, 
-	stepnumber=1, 
-	firstnumber=1, 
-	numberstyle=\tiny, 
-	extendedchars=true, 
-	breaklines=true,
-	frame=tb,
-	basicstyle=\footnotesize, 
-	stringstyle=\ttfamily,
-	showstringspaces=false
-}
+\makeatother
+\begin{document}
+\pagenumbering{roman}
 
-\definecolor{lightgray}{gray}{0.8}
-\definecolor{Gray}{gray}{0.9}
-\renewcommand{\lstlistingname}{Code}
-\renewcommand{\lstlistlistingname}{Lista de Listagens}
+\input{setup/title_author_etc}
+\input{parts/dedicated_to}
 
-\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
-\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
-\DeclareMathOperator*{\argmin}{arg\,min}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\newcommand*\justify{%
-	\fontdimen2\font=0.4em% interword space
-	\fontdimen3\font=0.2em% interword stretch
-	\fontdimen4\font=0.1em% interword shrink
-	\fontdimen7\font=0.1em% extra space
-	%\hyphenchar\font=`\-% allowing hyphenation
-}
+\runfrontend{}
 
-\address{BRASÍLIA}
+\import{}{macros}
+\import{parts/}{abstract}
 
-\universitypt{Universidade de Brasília}
-\universityen{Universidade de Brasília}
+\sumario
 
-\departmentpt{Departamento de Engenharia Elétrica - ENE/FT}
-\departmenten{Departamento de Engenharia Elétrica - ENE/FT}
+\listadefiguras
 
-\programpt{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}
-\programen{Progama de Pós-Graduação em Engenharia Elétrica - PPGEE}
+% \listadetabelas
 
-\majorfieldpt{Engenharia Elétrica}
-\majorfielden{Electrical Engineering}
+%-----------------------------------------------
+% Use os simbolos, acronimos e notacoes com as funcoes a seguir
+% To use this acronyms list, check the file glossaries-list/symbos.tex.
+	\listadesimbolos
 
-\title{Discriminative Sensing Based on Signal Processing for Information Security Analysis}
-\date{November/2017}
+% To use this acronyms list, check the file glossaries-list/acronyms.tex.
+	\listadeacronimos
 
-\author{Thiago Pereira de Brito Vieira}
-\adviser{João Paulo Carvalho Lustosa da Costa}
-% \coadviser{?}
+% To use this acronyms list, check the file glossaries-list/notation.tex.
+	\listadenotacoes
+%-----------------------------------------------
 
-\begin{document}
+\markboth{}{}
 
-\frontmatter
-\frontpage
-\presentationpage
+%Main Body
+\mainmatter 
+\setcounter{page}{1} \pagestyle{plain}
 
-\begin{dedicatory}Ficha Catalográfica\end{dedicatory}
+\import{parts/}{1_introduction}
+\import{parts/}{2_eigen_similarity}
+\import{parts/}{3_m_rpca}
+\import{parts/}{4_mobile}
+\import{parts/}{5_conclusion}
 
-\begin{dedicatory}Signatures\end{dedicatory}
+%BIBLIOGRAPHY
+\addcontentsline{toc}{chapter}{REFERENCES}
+% Do not use \bibliographystyle{IEEEtran} with abntex2cite package!!!!!
+% Also, try not to use abntex2cite at all! just comment its package in setup/packages.tex
+%\bibliographystyle{chicago}
+%\bibliographystyle{acm}
+\bibliographystyle{IEEEtran}
 
-\begin{dedicatory}Dedicatory.\end{dedicatory}
+\bibliography{bibliography}
 
-\agradecimentos
-First and foremost, I would like to thank God for giving me the life, strength, knowledge and opportunity to undertake this research study and to ..
-
-(...)
-
-Thank You!!
-
-\begin{epigraph}[]{Confucius}
-	Wherever you go, go with all your heart.
-\end{epigraph}
-
-\resumo
-\input{chapters/resumo}
-
-\abstract
-\input{chapters/abstract}
-
-\tableofcontents
-
-\makeatletter
-\renewcommand{\@thesubfigure}{\thesubfigure:\hskip\subfiglabelskip}
-\makeatother
-\setcounter{lofdepth}{2}
-
-\listoffigures
-\listoftables
-
-\input{chapters/acronyms}
-\mainmatter
-
-\include{chapters/1_introduction/introduction}
-\include{chapters/2_eigen_similarity/eigen_similarity}
-\include{chapters/3_mobile/mobile}
-\include{chapters/4_tensor_dl/tensor_dl}
-\include{chapters/5_conclusion_future_work/conclusion}
-
+%APPENDIX
+\part*{APPENDIX}
 \appendix
-\include{chapters/appendix/a_mos/a_mos}
-\include{chapters/appendix/b_pca_svm_cfa/pca_svm_cfa}
-
-\bibliographystyle{natbib}
-\addcontentsline{toc}{chapter}{Bibliography}
-\bibliography{references}
-
-\clearpage
-\addappheadtotoc
+\import{parts/}{a_mos}
+\import{parts/}{b_csf_pca_svm}
 

### 2019-10-17 docs/thesis/parts/1_introduction.tex

--- 2019-06-14 docs/thesis/chapters/1_introduction/introduction.tex
@@
 \label{ch:1_introduction}
-
-\begin{quotation}[]{Lao Tzu}
-The softest things in the world overcome the hardest things in the world.
-\end{quotation}
-
-\section{Motivation}
-\label{sc:motivation}
 
@@
 
-For the experimental evaluation of our proposals, we adopt a methodology based on aspects of GQM (Goal-Question-Metric) template \citep{Basili1994} and define two questions to achieve our goal, which are:
+For the experimental evaluation of our proposals, we adopt a methodology based on aspects of Goal-Question-Metric (GQM) template \cite{Basili1994} and define two questions to achieve our goal, which are:
 
@@
 In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect detailed time frames under attack.
-We evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 data set \citep{osanaiye2016distributed}, which is a well known network traffic data set.
+We evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 data set \cite{osanaiye2016distributed}, which is a well known network traffic data set.
 Furthermore, this proposed approach is validated into a prove of concept regarding behavioral anomaly detection to detect possible threats to a offline corporate mobile app.

### 2019-10-17 docs/thesis/parts/2_eigen_similarity.tex

--- 2019-06-14 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
@@
 \label{ch:2_mos_eig_sim}
-
-\begin{quotation}[]{Lao Tzu}
-All difficult things have their origin in that which is easy, and great things in that which is small. 
-\end{quotation}
 
@@
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig09.png}
+     \includegraphics[width=11cm]{figs/ch2/fig09.png}
      \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
@@
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig03.png}
+     \includegraphics[width=11cm]{figs/ch2/fig03.png}
      \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
@@
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig04.png}
+     \includegraphics[width=11cm]{figs/ch2/fig04.png}
      \caption{Network traffic of user independent operations for network management.}
@@
 
-We define that if the $\#\pmb{X}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\pmb{X}^{(q)} = 0$, then there is no malicious traffic.
+We define that if the $\#\pmb{X}^{(q)} \neq 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\pmb{X}^{(q)} = 0$, then there is no malicious traffic.
 We show how to detect the $\#\pmb{X}^{(q)}$, given only the matrix $\pmb{X}^{(q)}$, in order to identify malicious traffic.
@@
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig05.png}
+     \includegraphics[width=11cm]{figs/ch2/fig05.png}
      \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
@@
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig06.png}
+     \includegraphics[width=11cm]{figs/ch2/fig06.png}
      \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
@@
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig07.png}
+     \includegraphics[width=11cm]{figs/ch2/fig07.png}
      \caption{Connection attempts in order to identify active ports.}
@@
 	\centering
-     \includegraphics[width=11cm]{figures/ch2/mos_eigen_similarity.eps}
+     \includegraphics[width=11cm]{figs/ch2/mos_eigen_similarity.eps}
      \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
@@
 	\label{alg:2.01}
-	\SetAlgoLined
-	\KwResult{$\rm{\hat{\pmb{q}}}_{\rm max}$}
-	Given $f = 1$\;
-	\While{$f < \hat{d}$}{
-		$q_{\rm value} = \operatorname*{argmax}_{\lambda}  \hspace{1 mm} \pmb{e}_{\rm max}$\;
-		$i = 1$\;
-    	\While{$i < Q$}{
-    		\If{$\pmb{e}_{\rm max}^{(i)} == q_{\rm value}$}{
-    			$\hat{\pmb{q}}_{\rm max}^{(f)} = i$\;
-    		}
-    		$i = i + 1$\;
-    	}
-    	$\pmb{e}_{\rm max} \rightarrow \hat{\pmb{q}}_{\rm max}^{(f)}$\;
-		$f = f + 1$\;
-	}
-	\caption{Detection of Time Frames Under Attack}
+	% \SetAlgoLined
+	% \KwResult{$\rm{\hat{\pmb{q}}}_{\rm max}$}
+	% Given $f = 1$\;
+	% \While{$f < \hat{d}$}{
+	% 	$q_{\rm value} = \operatorname*{argmax}_{\lambda}  \hspace{1 mm} \pmb{e}_{\rm max}$\;
+	% 	$i = 1$\;
+ %    	\While{$i < Q$}{
+ %    		\If{$\pmb{e}_{\rm max}^{(i)} == q_{\rm value}$}{
+ %    			$\hat{\pmb{q}}_{\rm max}^{(f)} = i$\;
+ %    		}
+ %    		$i = i + 1$\;
+ %    	}
+ %    	$\pmb{e}_{\rm max} \rightarrow \hat{\pmb{q}}_{\rm max}^{(f)}$\;
+	% 	$f = f + 1$\;
+	% }
+	% \caption{Detection of Time Frames Under Attack}
 \end{algorithm}
@@
 
-Figure \ref{fig:2.08} illustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\pmb{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected by Algorithm \ref{alg:2.01}.
+Figure \ref{fig:2.88} illustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\pmb{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected by Algorithm \ref{alg:2.01}.
 
@@
 \begin{figure}[h!]
-     \includegraphics[width=15cm]{figures/ch2/incremental.eps}
+     \includegraphics[width=15cm]{figs/ch2/incremental.eps}
      \caption{Traffic selection for incremental approach.}
-     \label{fig:2.08}
+     \label{fig:2.88}
 \end{figure}
@@
 \begin{figure}[h!]
-     \includegraphics[width=15cm]{figures/ch2/individualized.eps}
+     \includegraphics[width=15cm]{figs/ch2/individualized.eps}
      \caption{Traffic selection for individual approach.}
@@
 \begin{figure}[h!]
-     \includegraphics[width=15cm]{figures/ch2/incremental_individualized.eps}
+     \includegraphics[width=15cm]{figs/ch2/incremental_individualized.eps}
      \caption{Traffic selection for incremental individualized approach.}
@@
 	\centering
-     \includegraphics[width=11cm]{figures/ch2/eigenvalues_synflood.eps} 
+     \includegraphics[width=11cm]{figs/ch2/eigenvalues_synflood.eps} 
      \caption{Eigenvalues of the sample covariance matrix (synflood).}
@@
 	\centering
-     \includegraphics[width=11cm]{figures/ch2/eigenvalues_fraggle.eps}
+     \includegraphics[width=11cm]{figs/ch2/eigenvalues_fraggle.eps}
      \caption{Eigenvalues of the sample covariance matrix (fraggle).}
@@
 	\centering
-     \includegraphics[width=11cm]{figures/ch2/eigenvalues_portscan.eps}
+     \includegraphics[width=11cm]{figs/ch2/eigenvalues_portscan.eps}
      \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}

### 2019-10-17 docs/thesis/parts/abstract.tex

@@
 \resumo{Resumo}{
-
 Resumo em Português, espaçamento de 1.5 requerido pela norma (automático).
-
 }
@@
 
-Traditionally, cyber defense methods can be effective against well known types of attacks, yet may fail against innovative malicious techniques.
-In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency in resource consumption, processing capacity and response time.
-Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
-Recently, signal processing schemes have been applied to detect malicious traffic in computer networks, showing advances in network traffic analysis.
+Signal processing techniques are usually applied for adaptive, streaming and real time problems, with high data volume and throughput.
+These techniques are widely adotped for denoising, signal separation and anomaly detection, and has attracted attention to apply signal processing for network attack detection.
+Recently, signal processing schemes have been applied to detect malicious traffic in computer networks, showing advances in network traffic analysis and network attack detection.
+Most of network-wide anomaly traffic detection methods are PCA-based method, which draw on PCA to divide traffic into normal and abnormal subspace, in accordance to anomaly detection problems.
 
-Information security may consist of both technical and procedural aspects.
-The former includes equipment and security systems, while the latter corresponds to security rules and recommendations of good practices.
-Intrusion detection and intrusion prevention systems are two types of security systems that are being used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks.
-Such systems can work in the following fashions: signature-based, anomaly-based or hybrid .
-
-Anomaly detection can be defined as the identification of rare and suspicious events by comparing it to the normal data.
+Anomaly detection can be defined as the identification of rare and suspicious events by comparing they to the normal data.
 Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects.
@@
 
-Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or behavioral analysis.
-However, distributed attacks organized by botnet has increased and demanded the development of counter measures.
-This has been done to detect and avoid unknown attacks or even to deal with adversarial changes of behavior, location and other patterns.
-This has also been done in order to avoid detection by behavioral or pattern based detection systems.
-To face this problem, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, where it is not necessary known anomalies for training models.
-
-Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate.
-This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security.
-Despite the actual high availability of information, the relevant information of some observations is generally of under much reduced dimensionality compared to available data sets.
-The extraction of relevant information by identifying the generating causes within classes of signals is useful for classification problems and for security analysis.
+As a first important contribution, we propose the Eigensimilarity, which is an approach based on signal processing concepts applied to detection of malicious traffic in computer networks.
+Eigensimilarity models the network traffic using a signal processing formulation as a composition of three components: signal, artifact and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
+The Eigensimilarity is based on eigenvalue analysis, model order selection (MOS) and similarity analysis between eigenvectors, where MOS and eigenvalue analysis are applied to detect time frames under attack, while the similarity analysis between eigenvectors aims the detection of the exxact time and network ports under attack.
+We evaluate the accuracy and performance of the proposed framework applied to a simulated scenario and to the DARPA 1998 data set.
+The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately by Eigensimilarity and with great detail in an automatic and blind fashion, i.e.
+in an unsupervised approach.
 
@@
 The fundamental issue with the imbalanced learning is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms.
-Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection.
+Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection...
+link to fraud and skewed data
 
 Some widely adopted algorithms for anomaly detection assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions.
-It was observed that the skeweness and heavy-tailed distributions can impact algorithms that rely on gaussian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.
+Findings indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network.
+These findings show that the skewness and heavy-tailed distributions may be important for network traffic analysis, and can motivate researches to evaluate the impact of skewed data into algorithms that rely on Gaussian distributed data.
 
-Therefore, as a first important contribution, we propose the Eigensimilarity, which is an approach based on signal processing concepts applied to detection of malicious traffic in computer networks.
-Eigensimilarity models the network traffic using a signal processing formulation as a composition of three components: signal, artifact and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
-The Eigensimilarity is based on eigenvalue analysis, model order selection (MOS) and similarity analysis between eigenvectors, where MOS and eigenvalue analysis are applied to detect time frames under attack.
-We evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 data set, which is a well known network traffic data set.
-The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, i.e.
-in an unsupervised approach, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
+We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data, and that the distance between robust estimates of normal traffic and contaminated observations can highlight discrepancies and be used for network attack detection.
+Therefore, we propose the m-RPCA, which is an approach based on distances between moments computed from a robust subspace learned by Robust Principal Component Analysis (RPCA) and contaminated observations, in order to detect anomalies from skewed data and network traffic.
+The anomaly detection from contaminated observations evaluate the Mahalanobis distance between the robust moments and new contaminated observations, in a semi-supervised fashion, but m-RPCA can also be computed as an unsupervised algorithm, with subspace learning from the contaminated data.
 
-Some widely adopted algorithms for anomaly detection assume a Gaussian or symmetric distributed data, however this assumption may not be observed in some real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network, leon2017probability}.
+We evaluate the accuracy of the m-RPCA for anomaly detection on simulated data sets, with skewed and heavy-tailed distributions, and for the CTU-13 data set.
+The Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches, and shows that the distance between robust estimates and contaminated observations can improve the anomaly detection on skewed data and the m-RPCA can be adoted for network attack detection.
 
-Findings of Benson \emph{et al.}  \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
-Leon-Garcia \cite{leon2017probability} also argues that Pareto distribution has been found to capture the behavior of many quantities of interest in the study of Internet behavior.
-Moreover, Benson \emph{et al.}  \cite{benson2010network} observes that the Lognormal distribution is the best fit to model arrival processes in a data center.
-
-These findings show that the skewness and heavy-tailed distributions may be important for network traffic analysis, and can motivate researches to evaluate the impact of skewed data into algorithms that rely on Gaussian distributed data.
-These findings can also highlight opportunities to  exploit the skewness and heavy-tailed distributions to obtain accurate classifiers for network anomaly detection.
-
-An outlyingness-approach based on a robust estimator of skewness, combined with robust estimators of location and scale, can be able to flag the outlying measurements.
-When the same methodology would be used with non-robust estimators of location, scale and skewness, the outlyingness-values would be affected by the outliers such that the outlying group could be masked \cite{hubert2009robustskewed}.
-
-Network anomaly detection problems are usually characterized by imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}.
-However, learning algorithms for imbalanced data has been a challenging research topic, considering that the fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
-Therefore, learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
-
-We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data.
-We also believe that the distance between robust estimates of normal traffic and contaminated observations can highlight discrepancies and be used for network attack detection.
-
-Therefore, we propose the m-RPCA, which is an approach based on distances between moments computed from a robust subspace learned by Robust Principal Component Analysis (RPCA) and contaminated observations, in order to detect anomalies from skewed data and network traffic.
-The proposed approach relies on a robust subspace computed from supposed normal observations, for estimating the moments to be used for distance analysis.
-The anomaly detection from contaminated observations evaluate the Mahalanobis distance between the robust moments and new contaminated observations, in a semi-supervised fashion, without the computational cost of new robust subspace learning for new observations.
-The m-RPCA can also be computed as an unsupervised algorithm, with subspace learning from the same contaminated data that is the target of the anomaly detection analysis.
-
-We evaluate the accuracy of the m-RPCA for anomaly detection on simulated data sets, with skewed and heavy-tailed distributions, and for the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
-The Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches.
-We also compare the results to ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
-
-The main contribution of this work is the proposal of a novel semi-supervised and unsupervised method for anomaly detection in skewed and imbalanced data, with results of experimental evaluation on simulated and real data sets.
-
-This Chapter is organized as follows.
-In Section \ref{sec:4_relatedworks}, it is conducted a literature review of network anomaly detection, botnet detection, and imbalanced learning.
-Section \ref{sec:4_datamodel} presents the data model and the evaluated data set.
-Section \ref{sec:4_proposal} describes the proposed approach for network attack detection.
-Section \ref{sec:4_experiments} discusses the experimental validation, Section \ref{sec:4_simulated_result} presents the results of the simulated scenarios and the Section \ref{sec:4_ctu13_result} describes the results for anomaly detection on CTU-13 data set.
-Finally, Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.
-
-The protection scheme used in a mobile device should be both computationally secure as well as resource-constrained due to battery power limitations \cite{khan2015cloud}.
-Therefore, encrypting files and generating keys on a mobile device is not considered a good solution.
-On the other hand, the protection schemes with good computational qualities lack the security analysis in many cases \cite{khan2014bss}.
-The common practice is the shadow user activities monitoring \cite{yovel2014}.
-However, the mobile device usage stays unprotected in all the proposed scenarios while in offline mode.
-When the mobile client goes offline with the sensitive corporate data on board all powerful cloud-based tools can not help and the mobile client has to secure itself with its own limited resources.
-Moreover, due to the resources constraint, there is a crucial difference in strategy of online and offline mode protection.
-
-Additional security issues and requirements have to be considered when mobile clients are actively used in corporate cloud environment \cite{yovel2014}.
-Today more and more organizations and enterprises are functioning in the Bring-Your-Own-Device (BYOD) paradigm.
-The uncontrolled usage of the mobile devices represents a serious risk to the development of secure SME cloud platforms being the bottleneck of the corporate information security system (ISS).
-While the enterprise cloud infrastructure based on the web interface can be protected by powerful third-party services, such as CASB and CAC, the corporate mobile client is usually light-weighted and generally less protected.
-
-This work proposes an architecture and approach for user behavior analysis based on Eigensim \cite{vieira2017model}, in order to detect possible threats in offline corporate mobile applications.
-The key expiration period is safely incorporated into the proposed system solution in order to enhance security and the behavioral analysis can indicate malicious behaviors, their variations, as well as novel attacks, which present low or high variance in comparison to legitimate user behaviors.
-
-The work is structured as follows.
-Section \ref{sec:3_related_works} analyzes the most common security problems in the mobile cloud environment and the solutions for offline protection in the BYOD world.
-Section \ref{sec:3_the_client_security_architecture} outlines the mobile security architecture of the proposed solution.
-Section \ref{sec:3_offline_mode} presents the detailed scheme of the proposed solution to the problem of offline mobile client security.
-Section \ref{sec:3_eigensim} explains the use of Eigensim.
-Section \ref{sec:3_results} discusses the common threat scenarios, the data modeling, the performance analysis and discussion of the practical implementation results.
-Section \ref{sec:3_conclusion} concludes the chapter.
+Moreover, we propose an architecture and approach to evaluate a prove of concept of Eigensimilarity for malicious behavior detection, in order to detect possible threats in offline corporate mobile applications.
+We propose scenarios, features and approaches for threats by means of Eigensimilarity, and evaluate the processing time required for Eigensimilarity execution in mobile devices.
 

### 2019-10-17 docs/thesis/setup/title_author_etc.tex

\titulolinhas{}

%% [sub-título]{Título original}
\titulolinhas[]{On the Subspace Learning for Network Attack Detection}

\maketitle

%% No caso de haver outros autores (máximo 3), crie mais um "\autores"
\autores{Thiago Pereira de Brito Vieira}

%% {Grau desejado em detalhes}{Tipo de monografia}{Departamento}{Grau desejado}{Sigla do departamento}{Programa do aluno ou departamento novamente]}
\grau{Doutor em Engenharia El\'{e}trica}{Tese de doutorado}{Engenharia El\'{e}trica}{Doutor}{ENE}{Engenharia El\'{e}trica}

\datainfo{Julho}{12}{2019}

%% {Traduzido para outro língua}{Título original}
\titulolinguas{}{On the Subspace Learning for Network Attack Detection}

%% Senão houver co-orientador, deixe o campo [] vazio.
\orientador[Rafael Timóteo de Sousa Júnior, Dr.]{João Paulo Carvalho Lustosa da Costa, Dr.-Ing.}

%% Para adicionar mais membros da banca: Após o último membro, crie uma nova linha (enter) e vá na aba superior esquerda (abaixo de “file”) e selecione “membro da banca”. Para deletar membros, apenas delete o item correspondente
\membrodabanca[Coorientador]{Prof. Dr. Rafael Timóteo de Sousa Júnior \textendash{} ENE/Universidade de Brasília}

\membrodabanca[Membro Interno]{Prof. Dr. William Ferreira Giozza \textendash{} ENE/Universidade de Brasília}

\membrodabanca[Membro Externo]{Prof. Dr. André Lima Férrer de Almeida \textendash{} DETI/Universidade Federal do Ceará}

\membrodabanca[Membro Externo]{Prof. Dr. Jacob Scharcanski \textendash{} INA/Universidade Federal do Rio Grande do Sul}

\catalogonome[Vieira, T. P. B.]{Vieira, Thiago Pereira de Brito}

%% {Publicação N#}{Palavra chave #1}{chave #2}{chave #3}{chave #4}
\catalogoinfo{PPGEE}{Network Attack Detection}{Anomaly Detection}{Subspace Learning}{Robust Subspace Learning}

### 2019-10-17 docs/thesis/parts/abstract.tex
Revert "abstract"
This reverts commit 759f3f56d8e8e954dd6810d9ccfe4ee590c8eac6.

@@
 
+Traditionally, cyber defense methods can be effective against well known attacks, yet may fail against innovative malicious techniques.
+In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency in attack detection, behavioral analysis, resource consumption, processing capacity and response time.
+Moreover, it is crucial to obtain high detection accuracy and capacity to detect malicious patterns without previous knowledge.
+
+Recently, signal processing schemes have been applied to detect malicious traffic in computer networks, showing advances in network traffic analysis and network attack detection.
 Signal processing techniques are usually applied for adaptive, streaming and real time problems, with high data volume and throughput.
 These techniques are widely adotped for denoising, signal separation and anomaly detection, and has attracted attention to apply signal processing for network attack detection.
-Recently, signal processing schemes have been applied to detect malicious traffic in computer networks, showing advances in network traffic analysis and network attack detection.
 Most of network-wide anomaly traffic detection methods are PCA-based method, which draw on PCA to divide traffic into normal and abnormal subspace, in accordance to anomaly detection problems.
@@
 Eigensimilarity models the network traffic using a signal processing formulation as a composition of three components: signal, artifact and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
-The Eigensimilarity is based on eigenvalue analysis, model order selection (MOS) and similarity analysis between eigenvectors, where MOS and eigenvalue analysis are applied to detect time frames under attack, while the similarity analysis between eigenvectors aims the detection of the exxact time and network ports under attack.
+The Eigensimilarity is based on eigenvalue analysis, model order selection (MOS) and similarity analysis between eigenvectors, where MOS and eigenvalue analysis are applied to detect time frames under attack.
 We evaluate the accuracy and performance of the proposed framework applied to a simulated scenario and to the DARPA 1998 data set.
@@
 
+Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate.
+This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security.
+Despite the actual high availability of information, the relevant information of some observations is generally of under much reduced dimensionality compared to available data sets.
+The extraction of relevant information by identifying the generating causes within classes of signals is useful for security analysis and for classification or anomaly detection problems.
+
 The high availability of raw data increases the challenges related to big data analytics and to imbalanced learning problems, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes.
 The fundamental issue with the imbalanced learning is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms.
-Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection...
-link to fraud and skewed data
+Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection.
 

### 2020-04-28 docs/thesis/chapters/1_introduction/introduction.tex
updating thesis latex

--- 2019-05-27 docs/thesis/chapters/1_introduction/introduction.tex
@@
 
-\begin{quotation}[]{Lao Tzu}
-The softest things in the world overcome the hardest things in the world.
-\end{quotation}
+According to Gartner surveys and client feedback, security remains a top concern for business and IT leaders.
+Cybersecurity is also highly visible in the media because of privacy concerns, destructive attacks such as ransomware, fears around IoT hacking, and an increasingly visible effect of cybersecurity on geopolitics \cite{Chuvakin2018}.
 
-\section{Motivation}
-\label{sc:motivation}
+The Whitehouse of the U.S.
+government \cite{whitehouse2018} estimates that malicious cyber activity cost the U.S.
+economy between US\$57 billion and US\$109 billion in 2016.
+According to Bissel \emph{et al.} \cite{BISSELL2019}, the total annual cost of all types of cyberattacks is increasing.
+Malware and Web-based attacks continue to be the most expensive threats for global companies, corresponding to US\$2.6 million and US\$2.2 million, respectively.
+However, it is possible to observe an increasing of 10\% of Denial of Service (DoS) attack, 12\% of botnets, 15\% of malicious insider and an increment of 21\% in occurrences of ransomware, between 2017 and 2018 \cite{BISSELL2019}.
 
-Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}.
-In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time.
+Accenture's study conducted by Bissel \emph{et al.} \cite{BISSELL2019} shows that the extent of the economic value may be at risk if security investments are not made wisely.
+Bissel \emph{et al.} \cite{BISSELL2019} shows that the size of opportunity varies by industry, with high tech subject to the greatest value at risk — US\$753 billion — over the next five years, followed by US\$ 642 billion for life sciences and US\$505 billion for the automotive industry.
+Bissel \emph{et al.} \cite{BISSELL2019} analyze the total technology savings minus total technology spend in cybersecurity by organizations, reporting that security intelligence and threat sharing have been adopted by 67\% of respondents and saving US\$2.26 million.
+Bissel \emph{et al.} \cite{BISSELL2019} also reported that automation, artificial intelligence, and machine learning have been adopted by 38\% of respondents and saving US\$2.09 million, while the cyber and user behavior analytics have been adopted by 32\% of respondents and saving US\$1.72 million.
+
+Additionally, Levi \cite{Levi2019} highlights that Blizzard Entertainment was hit by a Distributed Denial of Service (DDoS) attack in 2017, when the downtime lasted an entire weekend, due to difficulties related to track the root cause of the anomaly.
+According to Levi \cite{Levi2019}, attacks like these can cost an enterprise more than US\$2 million, which makes anomaly detection essential to protecting revenue.
+
+Traditionally, cyber defense methods can be effective against well-known types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}.
+In order to be able to detect and avoid network attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency in resource consumption, processing capacity and response time.
 Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
-Recently, signal processing schemes have being applied to detect malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest, vieira2017model}, showing advances in network traffic analysis.
+Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or behavioral analysis \cite{gu2008botminer, khattak2015botflex}.
 
-Information security may consist of both technical and procedural aspects.
-The former includes equipment and security systems, while the latter corresponds to security rules and recommendations of good practices.
-Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks.
-Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.
+An attacker can create a bot for malicious purposes to generate a DoS attack, to exploit or abuse a target application.
+A botnet is a collection of bots or systems for executing automated tasks, often in the form of compromised hosts, including desktops, mobile devices or things (as in IoT).
+Some recent common examples of botnets include Cyclone, Mirai, Nitol and Sentry MBA \cite{Catucci2019}.
+Distributed attacks organized by botnet has increased and demanded the development of counter measures, in order to detect and avoid unknown attacks or even to deal with adversarial changes of behavior, location and other patterns \cite{wang2017botnet, Wang2018ddosbotnetssurvey}.
 
-Anomaly detection can be defined as the identification of rare and suspicions events by differing the normal or majority of the data.
-Anomalies are also referred to as outliers, novelties, noise or deviations, and can be related to network attacks, frauds or defects \cite{bhuyan2014network,ahmed2016survey}.
-Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
-Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.
+According to Hevesi \cite{Hevesi2019} in a Gartner's publication, Botnets scan the Internet looking for unprotected Internet of Things (IoT) devices — and there are projected to be 25 billion devices online in 2021.
+Mobile network providers are deploying 5G capabilities to their networks, and industry is looking for new ways to leverage the service to bring new capabilities to market.
+Hevesi argues that if just 1\% was able to be added to a botnet and had access to 5G, the potential network throughput would be 250 million devices able to push 10 to 50 Gbps of network traffic in a flooding attack.
 
-Several efforts and researches aim to avoid network attacks based on known attackers, fingerprints or behavioral analysis.
-However, distributed attacks organized by botnet has increasing and demanding the development of counter measures in order to detect and avoid unknown attacks or even to deal with adversarial changes on behavior, location and other patterns, in order to avoid detection by behavioral or pattern based detection systems \cite{gu2008botminer, garcia2014empirical,khattak2015botflex,acarali2016survey,wang2017botnet,Wang2018ddosbotnetssurvey}.
-To face this problem, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, where it is not necessary known anomalies for training models \cite{moustafa2019holistic}.
+Cyber security systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{mudzingwa2012study}.
+A signature or fingerprint can be seen as a sequence of data, behavior and rules which are often unique to known malware or attack types, allowing the identification of attackers that reproduce some signature, such as occurs in less sophisticated viruses and automated toolkits for security exploitation and intrusion.
+However, signature-based systems for attack detection have to deal with adversarial models and techniques that aims to avoid detection based on well-known patterns, such as instruction virtualization, packing, polymorphism, emulation, and metamorphism to write and change malicious codes that can evade the detection \cite{bhuyan2014network}.
 
-Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate.
+Anomaly-based systems for network attack detection focus on finding exceptional, suspicious or rare observations in network traffic that do not conform to the expected legitimate behavior \cite{bhuyan2014network}.
+In a general view, anomalies are referred to as outliers, novelties or deviations, and can be related to defects, diseases, network attacks, intrusion detection for cyber security, fraud detection for credit cards, and military surveillance issues.
+Additionally, anomaly detection techniques can be categorized by classification, statistical algorithms, information theory and cluster based algorithms, according to \cite{ahmed2016survey}.
+
+Hybrid systems for network attack detection exploit benefits of both signature-based and anomaly-based detection techniques, and attempt to detect known as well as unknown attacks \cite{ahmed2016survey}.
+
+%Recent developments in science and technology have enabled the growth and availability of raw data to occur at an explosive rate.
 This has created an immense opportunity for knowledge discovery and data engineering research to play an essential role in a wide range of applications from daily civilian life to national security.
-Despite the actual high availability of information, the relevant information of some observations is generally of under much reduced dimensionality compared to available data sets.
-The extraction of relevant information by identifying the generating causes within classes of signals is useful for classification problems and for security analysis.
+Despite the high availability of information, the relevant information for network anomaly detection from the whole network traffic is lower than the total available traffic, with very imbalanced classes and with susceptibility to imbalanced learning problem.
 
-The high availability of raw data increases the challenges related to big data analytics and to imbalanced learning problems, which corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes.
-The fundamental issue with the imbalanced learning is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms.
-Data analysis of imbalanced data is challenging for classification and prediction problems related to anomaly detection, novelty detection, fraud detection and attack detection.
+Accenture consulting argues that 89\% of survey respondents believe breakthrough technologies, like artificial intelligence, machine or deep learning, user behavior analytics, and blockchain, are essential for securing the future of their organizations \cite{Accenture2018}.
+To face the adversarial model, novel attacks and counter measures of attackers to avoid detection, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, by means of behavioral analysis, where known anomalies are not necessaries for training models \cite{moustafa2019holistic}.
 
-Some widely adopted algorithms assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network,leon2017probability}.
-It was observed that the skeweness and heavy-tailed distributions can impact algorithms that rely on gaussian distributed data, as well as it can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.
+Recently, signal processing schemes have been applied to detect malicious traffic in computer networks by means of unsupervised approaches, showing advances in network traffic analysis \cite{Zonglin2009}, in network attack detection \cite{callegari2011novel}, and in network anomaly detection based on network flows \cite{Lu2009}.
+
+Furthermore, it is possible to observe that anomaly-based and behavioral-based solutions have been adopted for cyber security and attracting investments, as can be seen in the Featurespace case, which is a provider of adaptive behavioral analytics for fraud detection and risk management, that raised US\$32.3 million from a funding round led by Insight Venture Partners and MissionOG \cite{Williams2019}.
+
+Anomalies in the context of network traffic can be hard to identify and separate from legitimate data due to the rare occurrences of anomalies in comparison to legitimate events.
+Therefore, anomaly detection algorithms have to be highly discriminating, robust to corruption and able to deal with the imbalanced data problem \cite{he2008learning}.
+Note that data corruption refers to outliers that can be part of the data, while the imbalanced data problem corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes \cite{he2009learning}, which can be legitimate or malicious classes in network anomaly detection problems.
+
+The imbalanced data can compromise the performance of most standard learning algorithms, creating bias or unfair weight to learn from the majority class and reducing detection capacity of anomalies that are characterized by the minority class.
+Hence, data analysis of imbalanced data is challenging for learning algorithms applied to classification problems of anomaly detection, novelty detection, fraud detection and network attack detection.
+However, anomaly-based algorithms, that rely on unsupervised or semi-supervised approaches, can be alternatives for network anomaly detection from imbalanced data \cite{moustafa2019holistic}.
+
+Some widely adopted algorithms for anomaly detection assume a Gaussian distributed legitimate data \cite{lakhina2005mining}, however this assumption may not be observed in real world problems, such as the case of network traffic analysis \cite{benson2010network}, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{benson2010network, leon2017probability}.
+According to \cite{hubert2009robustskewed}, the skeweness and heavy-tailed distributions can impact algorithms that rely on Gaussian distributed data, and can reveal characteristics that can be exploited in order to obtain accurate classifiers for network anomaly detection.
+
+Therefore, this thesis focus on anomaly-based problems.
+More specifically, we focus on network attack detection and propose behavioral-based approaches for network anomaly detection through subspace learning techniques and similarity analysis.
+
+Firstly, we propose the Eigensimilarity, which is an approach based on signal processing methods applied to detection of probe and flooding attacks in computer networks.
+We present an architecture and approach to evaluate a proof of concept of Eigensimilarity for malicious behavior detection on mobile applications, in order to detect possible threats in offline corporate mobile client.
+Additionally, we propose the Moment-based Robust Principal Component Analysis (m-RPCA), which is an approach based on distances between contaminated observations and moments computed from a robust subspace learned by Robust Principal Component Analysis (RPCA), in order to detect anomalies from imbalanced and skewed data, such as network traffic.
+
+This chapter is organized as follows.
+The problem statement, the hypotheses formulation and the proposed approaches for network attack detection are introduced in Section \ref{sec:1_problems}.
+In Section \ref{sec:1_contributions} we present the main contributions of this thesis and in Section \ref{sec:1_organization} we describe the thesis organization for the next chapters.
 
 \section{Problem Statement}
-\label{sc:problems}
+\label{sec:1_problems}
 
-Considering the above described landscape, this thesis outlines the development and evaluation of approaches based on subspace learning for network attack detection, through methods to make the data discriminative and able to identify structures, hidden patterns and the most relevant information for anomaly detection.
-In the Subsection \ref{sc:Hypothesis} we present our hypothesis formulation and the Subsection \ref{sc:proposals} describes the proposals to answer the questions and validate the hypotheses.
+Considering the previous described landscape, this thesis outlines the development and evaluation of approaches based on subspace learning for network attack detection, through methods to make the data discriminative and able to identify structures, hidden patterns and the most relevant information for anomaly detection.
+In Subsection \ref{sec:1_Hypothesis} we present our hypothesis formulation and in Subsection \ref{sec:1_proposals} we describe the proposed approaches to answer the questions and validate the hypotheses.
 
 \subsection{Hypothesis Formulation}
-\label{sc:Hypothesis}
+\label{sec:1_Hypothesis}
 
-For the experimental evaluation of our proposals, we adopt a methodology based on aspects of GQM (Goal-Question-Metric) template \citep{Basili1994} and define two questions to achieve our goal, which are:
+For the experimental evaluation of our proposals, we adopt a methodology based on aspects of Goal-Question-Metric (GQM) template \citep{Basili1994} and define two questions to achieve our goal, which are:
 
@@
 
-Our testing hypotheses are defined in Table \ref{tab:hypothesis}, that describe the null hypothesis and alternative hypothesis for each previously defined question.
+Our testing hypotheses are defined in Table \ref{tab:hypothesis}, that describe the null hypotheses ($H_1^{(N)}$ and $H_2^{(N)}$) and alternative hypotheses ($H_1^{(A)}$ and $H_2^{(A)}$) for each previously defined question.
 
@@
 	\centering
-	\caption{Hypotheses to evaluate the defined metrics}
+	\caption{Hypotheses to evaluate the defined questions}
 	\label{tab:hypothesis}
@@
         \textbf{Alternative Hypothesis}	&\textbf{Null Hypothesis}	&\textbf{Question} 	\\ \hline
-        $H_{1sub}$: It is possible to use subspace learning to identify and detect anomalies in network traffic.
-        &$H_{0sub}$. It is not possible to use subspace learning to identify and detect anomalies in network traffic.	&$Q_1$\\ \hline
-        $H_{1rob.sub}$: The robust subspace learning improves the anomaly detection from imbalanced and skewed data.	&$H_{0rob.sub}$.
-The robust subspace learning does not improves the anomaly detection from imbalanced and skewed data.	&$Q_2$\\ \hline
+        $H_1^{(A)}$: A subspace learned by eigenvalue decomposition can be used to detect and identify network attacks.
+        &$H_1^{(N)}$. A subspace learned by eigenvalue decomposition can not be used to detect and identify network attacks.	&$Q_1$\\ \hline
+        $H_2^{(A)}$: An approach based on robust subspace learning improves the anomaly detection from imbalanced and skewed data.	&$H_2^{(N)}$.
+An approach based on robust subspace learning does not improves the anomaly detection from imbalanced and skewed data.	&$Q_2$\\ \hline
     \end{tabular}
@@
 
-The hypotheses $H_{1num.indct}$ and $H_{0num.indct}$ were defined to evaluate if a subspace learned by singular value decomposition are sensitive to outliers and can be used to detect network attacks.
-We defined the hypotheses $H_{1scale.prop}$ and $H_{0scale.prop}$ to evaluate the feasibility and performance of a robust subspace learned by Robust PCA to network anomaly detection.
+The hypotheses $H_1^{(A)}$ and $H_1^{(N)}$ are defined to evaluate if a subspace learned by eigenvalue decomposition are sensitive to outliers and can be used to detect network attacks.
+We define the hypotheses $H_2^{(A)}$ and $H_2^{(N)}$ to evaluate if the distance between contaminated data and robust moments learned by Robust Principal Component Analysis (RPCA) can improve the anomaly detection in simulated imbalanced and skewed data.
 
-The metrics most adopted to evaluate experimental results of network attack detection are the true positive (TP), true negative (TN), false positive (FP) and false negative (FN).
+%The metrics most adopted to evaluate experimental results of network attack detection are the true positive (TP), true negative (TN), false positive (FP) and false negative (FN).
 We adopt these metrics and also adopt the misclassification rate and the F1-score, which is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic}.
@@
 \subsection{Proposals}
-\label{sc:problem_statement}
+\label{sec:1_proposals}
 
-In the context of anomaly-based schemes, this thesis proposes the Eigensim, which is a approach based on subspace learning techniques for detection of malicious traffic in computer networks.
+In the context of anomaly-based schemes, this thesis proposes the Eigensimilarity, which is an approach based on subspace learning techniques for detection of malicious traffic in computer networks, by means of eigenvalue analysis, model order selection (MOS) and a similarity analysis between eigenvectors of estimated legitimate observations and observations of a time frame estimated as under attack.
 
-Inspired by \cite{david2011blind,da2012improved}, Eigensim models the network traffic using a signal processing formulation as a composition of three components: signal, artifact and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
-The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
 In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect detailed time frames under attack.
-We evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 data set \citep{osanaiye2016distributed}, which is a well known network traffic data set.
-Furthermore, this proposed approach is validated into a prove of concept regarding behavioral anomaly detection to detect possible threats to a offline corporate mobile app.
+We evaluate the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 data set \citep{osanaiye2016distributed}, which is a well-known network traffic data set.
+Furthermore, this proposed approach is evaluated by a proof of concept regarding behavioral anomaly detection to detect possible threats to an offline corporate mobile app.
 
-Network anomaly detection problems are usually characterized by skewed and imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}.
-Learning algorithms for imbalanced data has been a challenging research topic, considering that the fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
+The skewness of anomalous and legitimate data can highlight features for improving anomaly detection in imbalanced data, and the distance between robust estimates of legitimate observations and contaminated data can be used for network attack detection.
+Therefore, we propose the Moment-based Robust Principal Component Analysis (m-RPCA), which is an approach based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data.
+We evaluate the results of m-RPCA for anomaly detection on simulated imbalanced and skewed data, and evaluate the results of m-RPCA for network attack detection on CTU-13 data set.
 
-Some widely adopted algorithms assume a gaussian distributed data, however this assumption may not be observed in real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network, leon2017probability}.
-Therefore, the skeweness and heavy-tailed distributions can impact algorithms that rely on gaussian distributed data, as well as it can be exploited to reveal characteristics that can improve classifiers for network anomaly detection.
+Eigensimilarity and m-RPCA are frameworks for network attack detection by means of anomaly-based analysis from network traffic.
+In a simplified architecture for network attack detection shown by Figure \ref{fig:1.01}, Eigensimilarity and m-RPCA are deployed as a module of an application firewall, which is responsible for secure and protect application communications, but can also work on network level.
 
-We believe that the skewness of anomalous and normal data can highlight features for improving anomaly detection in imbalanced data.
-We also believe that the distance between robust estimates of normal and contaminated data can highlight discrepancies and be used anomaly detection and for network attack detection.
-Therefore, we propose the m-RPCA, which is an approach based on  distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data.
+\begin{figure}[h!]
+	\centering
+     \includegraphics[width=15cm]{figures/ch1/architecture.eps} 
+     \caption{Simplified Architecture for Network Attack Detection.}
+     \label{fig:1.01}
+\end{figure}
+
+An application firewall can implement behavioral analysis in order to detect attacks against application services, such as Spam or Click Fraud (CF), or can detect anomalies in network level, such as a DoS attack based on Ping flood.
+
+The Figure \ref{fig:1.01} depicts the flow of network traffic between legitimate or malicious users to a corporate network.
+All income traffic shall be received and distributed by the load balancer, to be evaluated by an application firewall, which is is responsible for network attack detection.
+Therefore, the application firewall implements a high throughput traffic analyzer, to capture and parse the network traffic for further analysis, by means of Eigensimilarity and m-RPCA.
+The Figure \ref{fig:1.01} also depicts the use of Eigensimilarity as a module of an offline corporate mobile app, as a proof of concept regarding behavioral anomaly detection from user activities.
 
 \section{Contributions}
-\label{sc:contributions}
+\label{sec:1_contributions}
 
-We analyze problems related to detection of information security issues and propose new approaches to improve malicious behavior detection through signal processing techniques based on subspace learning.
-The results of the work presented in this thesis provide the following contributions:
+We analyze problems related to detection of information security issues in network traffic and propose new approaches to improve malicious behavior detection through signal processing techniques based on subspace learning.
+The results of the work presented in this thesis provide the following publications and contributions:
 
-\begin{enumerate}
-	\item We propose an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluate the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 data set;
-	\item We discuss the computational complexity of the proposed framework and evaluate the required processing time for tested scenarios;
-	\item We propose an architecture and techniques for offline behavioral analysis of a corporate mobile client security architecture, and discuss the processing time of the proposed framework for mobile devices;
-	\item We propose an approach based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data, and evaluate the anomaly detection rate on synthetic and real data sets;
+\begin{enumerate}[label*=\arabic*.]
+	\item T.
+P.
+B.
+Vieira, D.
+F.
+Tenório, J.
+P.
+C.
+da Costa, E.
+P.
+de Freitas, G.
+Del Galdo, and R.
+T.de Sousa Júnior, “Model order selection and eigen similarity based framework for detection and identification of network attacks, ” \textit{Journal of Network and ComputerApplications}, vol.
+90, pp.
+26–41, 2017 \citeme{vieira2017model}.
+	\begin{enumerate}[label*=\arabic*.]
+		\item We propose the Eigensimilarity, which is an approach based on eigenvector similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluate the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 data set;
+		\item We discuss the computational complexity of the Eigensimilarity and evaluate the required processing time for tested scenarios;
+	\end{enumerate}
+	\item T.
+P.
+B.
+Vieira, J.
+P.
+C.
+L.
+da Costa, E.
+S.
+C.
+Vilaça, E.
+S.
+Gualberto, and R.
+T.de Sousa Júnior, “Moment distances from robust subspace for network attack detection, ” \textit{Journal of Network and Computer Applications}, To Appear \citeme{vieira2019moment}.
+	\begin{enumerate}[label*=\arabic*.]
+		\item We propose the m-RPCA, which is an approach based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data, and evaluate the anomaly detection and network attack detection rates on simulated and real data sets.
+	\end{enumerate}
+	\item T.
+Galibus, T.
+P.
+B.
+Vieira, E.
+P.
+de Freitas, R.
+d.
+O.
+Albuquerque, R.
+T.de Sousa Júnior, V.
+Krasnoproshin, A.
+Zaleski, H.
+Vissia, G.
+del Galdoet al., “Of-fline mode for corporate mobile client security architecture, ” \textit{Mobile Networks and Applications}, pp.
+1–17, 2017 \citeme{galibus2017offline}.
+	\begin{enumerate}[label*=\arabic*.]
+		\item We propose an architecture and implement a proof of concept for offline behavioral analysis of a corporate mobile client, and discuss the processing time of the Eigensimilarity for mobile devices;
+	\end{enumerate}
+	\item K.
+H.
+C.
+Ramos, R.
+T.
+de Sousa Junior, T.
+P.
+B.
+Vieira, and J.
+P.
+C.
+L.
+da Costa, “Discovering critical success factors for information technologies governance through bibliometric analysis of research publications in this domain, ” \textit{International Information Institute (Tokyo).
+Information}, vol.
+19, no.
+6B, p.
+2193, 2016.
+\citeme{ramos2016information}.
+	\item K.
+H.
+C.
+Ramos, T.
+P.
+B.
+Vieira, J.
+P.
+C.
+L.
+da Costa, and R.
+T.
+de Sousa Júnior, “Multidimensional analysis of critical success factors for it governance within the Brazilian federal public administration in the Light of External Auditing Data".
+\textit{12th International CONTECSI}, 2015 \citeme{ramos2015multidimensional}.
+	\begin{enumerate}[label*=\arabic*.]
+		\item We propose a critical factors analysis based on Principal Component Analysis (PCA) for visual discriminant analysis, and presenting an approach based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM), in order to identify the Critical Success Factors (CSF) for IT governance.
+	\end{enumerate}
 \end{enumerate}
@@
 \section{Thesis Organization}
-\label{sc:organization}
+\label{sec:1_organization}
 
 This thesis is organized as follows.
-In Chapter \ref{ch:2_mos_eig_sim}, we propose the Eigensim, which is an approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
-Chapter \ref{ch:3_mobile} presents a prove of concept regarding the evaluation of an approach and architecture based on user behavior analysis through the Eigensim  \cite{tenorio2013greatest}, in order to detect threats in a mobile application.
-Chapter \ref{ch:4_m_rpca} proposes the m-RPCA, which is an approach based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data.
-Chapter \ref{ch:5_conclusionfuturework} draws the conclusions and the suggestions for future work.
-Furthermore, the Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes and the Appendix \ref{apx:b_csf_fs} presents a critical factors analysis based on Principal Component Analysis (PCA) for visual discriminant analysis, and presents an approach based on Recursive Feature Elimination (RFE) combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of Brazilian public organizations, in order to identify the Critical Success Factors (CSF) for IT governance of the public sector according to TCU.
+In Chapter \ref{ch:2_mos_eig_sim}, we propose the Eigensimilarity, which is an approach based on signal processing techniques for detection of malicious traffic in computer networks, based on eigenvalue analysis, model order selection (MOS) and similarity analysis.
+In Chapter \ref{ch:3_mobile} we present a proof of concept regarding the evaluation of an approach and architecture based on user behavior analysis through the Eigensimilarity \citeme{vieira2017model}, in order to detect threats in a mobile application.
+The m-RPCA is proposed in Chapter \ref{ch:4_m_rpca}, where is presented the proposed approach based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data.
+In Chapter \ref{ch:5_conclusionfuturework} we draw the conclusions and the suggestions for future work.
+Furthermore, in Appendix \ref{apx:b_csf_fs} we present a critical factors analysis based on Principal Component Analysis (PCA), Recursive Feature Elimination (RFE) and Support Vector Machine (SVM), in order to identify the Critical Success Factors (CSF) for IT governance.

### 2020-04-28 docs/thesis/chapters/2_eigen_similarity/eigen_similarity.tex
updating thesis latex

@@
-\chapter{Eigensim - Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks}
+\chapter{Eigensimilarity based Framework for Detection and Identification of Network Attacks}
 \label{ch:2_mos_eig_sim}
 
-\begin{quotation}[]{Lao Tzu}
-All difficult things have their origin in that which is easy, and great things in that which is small. 
-\end{quotation}
-
-Traditionally, cyber defense methods can be effective against ordinary and conventional types of attacks, yet may fail against innovative malicious techniques \cite{lakhina2005mining}.
-In order to be able to detect and avoid novel attacks and their variations, it is necessary to develop or improve techniques to achieve efficiency on resource consumption, processing capacity and response time.
-Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
-Recently, signal processing schemes have being applied to the detection of malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest}, showing advances in network traffic analysis.
-
-Information security may consist of both technical and procedural aspects.
-The former includes equipment and security systems, while the latter corresponds to security rules and recommendations.
-Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks.
-Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}.
-Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.
-
-This chapter proposes the Eigensim, which is an approach based on signal processing concepts applied to detection of malicious traffic in computer networks.
-Inspired by \cite{david2011blind,da2012improved}, the Eigensim models the network traffic using a signal processing formulation as a composition of three components: signal, artifact and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP).
-The Eigensim is based on eigenvalue analysis, model order selection (MOS) and similarity analysis between eigenvectors.
-In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack.
-In addition, we evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 data set \citep{osanaiye2016distributed}, which is a well known network traffic data set.
-
-The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, i.e.
-in an unsupervised approach, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis.
-The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via eigen similarity analysis.
+According to \cite{Hevesi2019}, at times, organizations relate their security measures to confidentiality and integrity, ignoring availability.
+However, according to \cite{BISSELL2019}, the global average annual cost of DoS attack was US\$1.7 million in 2018.
+
+The Denial of Service (DoS) attack attempts to deny access to system or network resources, attacking the availability of the service, by means of flooding techniques for consuming the available resources, or by means of subtle approaches which send small amount of data that can cause failure and unavailability of the system.
+In 2017 the top motivation behind DDoS attacks was criminals demonstrating attack capabilities, with gaming and criminal extortion attempts in second and third place, respectively \cite{cisco2019}.
+
+With application layer Distributed Denial of Service (DDoS) attacks continuing to rise, vendors have begun adding DDoS mitigation features into solutions that protect web applications \cite{Hevesi2019}.
+The leading solutions for monitoring and mitigation of DDoS attacks in application layer start with behavioral analytics and supplement with signature-based detection for known malicious application attacks \cite{Hevesi2019}.
+
+Probe attacks aim to scan the networks to identify running services, open ports, running services and vulnerabilities that can be exploited.
+A probe attack is considered the first step in an attack to compromise a host or network.
+Although no specific damage is caused by these attacks, they are considered serious threats by \cite{ahmed2016survey} and \cite{moustafa2019holistic}.
+Note that we refer to port scan as an attack, according to \cite{ahmed2016survey}.
+
+Anomaly-based systems for network attack detection focus on finding exceptional, suspicious or rare observations in network traffic that do not conform to the expected normal behavior \cite{bhuyan2014network}.
+It is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, by means of behavioral analysis, where known anomalies for training models are not necessary \cite{moustafa2019holistic}.
+
+Recently, signal processing schemes have been applied for network anomaly detection in computer networks by means of unsupervised approaches, showing advances in network traffic analysis \cite{Zonglin2009}, in network attack detection \cite{callegari2011novel}, and in network anomaly detection based on network flows \cite{Lu2009}.
+
+Considering that a subspace learned by eigenvalue decomposition can highlight anomalies for network attack detection by means of model order selection (MOS), and that the comparison between the principal eigenvectors can reveal anomalies when comparing legitimate and anomalous observations.
+In this chapter we propose the Eigensimilarity, which is an approach based on signal processing concepts applied to detection of malicious traffic in computer networks, with focus on DoS and portscan attacks.
+
+The Eigensimilarity is based on eigenvalue analysis, model order selection (MOS) and similarity analysis between the principal eigenvectors.
+In contrast to \cite{david2011blind, da2012improved, tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack, and the analysis of similarity between the principal eigenvectors are used for precise identification of the time and ports under attack.
+In addition, we evaluate the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 data set \citep{osanaiye2016distributed}, which is a well-known network traffic data set.
+
+The performed experiments show that synflood, fraggle and port scan attacks can be accurately detected and with great detail in an automatic and blind fashion, i.e.
+in an unsupervised approach that does not require data for training, applying signal processing concepts for traffic modeling and through approaches based on MOS and principal eigenvector similarity analysis.
+The main contributions of the proposed framework are the capability to blindly detect time frames under network attack via MOS and eigen analysis, and the detailed identification of the network attack via principal eigenvector similarity analysis.
 
 This chapter is organized as follows.
-In Section \ref{sec:2_relatedworks}, related works are discussed.
-Section \ref{sec:2_datamodel} presents the data model and the evaluated data sets.
-Section \ref{sec:2_prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks.
-Section \ref{sec:2_experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:2_Complexity} discusses the computational complexity of the proposed framework and evaluates the required processing time for tested scenarios.
-Section \ref{sec:2_conclusionandfutureworks} draws the conclusions and the suggestions for future work.
-The Appendix \ref{apx:a_mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.
+In Section \ref{sec:2_relatedworks} the related works are discussed.
+We present the data model and the evaluated data sets in Section \ref{sec:2_datamodel} and the proposed framework for blind and automatic detection of flood and probe attacks is described in Section \ref{sec:2_prop_getv}.
+The discussion of the experimental validation and results are presented in Section \ref{sec:2_experimentalresults}, while the discussion of the computational complexity and evaluation of the required processing time for tested scenarios are presented in Section \ref{sec:2_Complexity}.
+In Section \ref{sec:2_conclusionandfutureworks} we draw the conclusions and the suggestions for future work.
 
@@
 Researchers have applied data mining techniques in log analysis \cite{ghourabi2010data} to improve intrusion detection performance.
-However, data mining techniques used so far in network analysis require prior collection of large data sets, which is a limitation of several schemes for online analysis \cite{Huang2009}.
+However, data mining techniques used so far in network analysis require prior collection of large data sets, which is a limitation of several schemes for online analysis \cite{he2008applying}.
 
@@
 
-Callegari \emph{et al} \cite{Zonglin2009} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a data set with synthetic anomalies added in the data.
-However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
+Callegari \emph{et al.} \cite{callegari2011novel} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a data set with synthetic anomalies added in the data.
+However, Callegari \emph{et al.} \cite{callegari2011novel} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.
 
 Lee \emph{et al.} \cite{Lee2013} presented the OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
-In contrast to Lee \emph{et al.}, the Eigensim applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
-Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
+In contrast to Lee \emph{et al.} \cite{Lee2013}, the Eigensimilarity applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack.
+Additionally, Lee \emph{et al.} \cite{Lee2013} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.
 
@@
 However, their work requires a training method to produce a prediction model for normal daily traffic and presents limitations on identification of behaviors without significant outliers, such as port scan attacks.
-Zonglin \emph{et al} \cite{Zonglin2009} proposed a signal processing method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
-Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work is not applied to probe and flood attack detection, simultaneously.
-
-The data collected from honeypot systems, such as captured traffic and operating system logs, can be analyzed to obtain information about attack techniques, general trends of threats and exploits.
-Blind automatic detection of malicious traffic techniques have been developed for honeypots in \cite{david2011blind,da2012improved}.
+Zonglin \emph{et al.} \cite{Zonglin2009} proposed a signal processing method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies.
+Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work is not applied for probe attack detection and does not evaluate results for attack detection on DARPA data set.
+
+The data collected from honeypot systems, such as captured traffic and operating system logs, can be analyzed to obtain information about attack techniques, general trends of threats and exploits \cite{david2011parallel}.
+Blind automatic detection of malicious traffic techniques have been developed for honeypots in \cite{david2011blind, da2012improved}.
 However, traffic on honeypot is simpler than real network traffic, because there are no running legitimate applications, i.e.
@@
 Since honeypots do not generate legitimate traffic, the amount of data captured in honeypots is significantly lower in comparison to a Network Intrusion Detection System (NIDS), which captures and analyzes the largest possible amount of network traffic \cite{david2011blind}.
+
 MOS for blind identification of malicious activities in honeypots was proposed by us in \cite{david2011blind}, which evaluated criteria for selecting the model order, through simulations and comparing the order of the resulting model with the true model order.
@@
 The proposed solution is automatic and blind for detection of time frames under probe and flood attacks through MOS and eigen analysis.
-Moreover, we apply eigen similarity analysis to identify details of time and ports under network attacks.
-
-Several approaches for network attack detection uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
-Even though the KDD 99 data set are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data sets adopted by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
-
+Moreover, we apply principal eigenvector similarity analysis to identify details of time and ports under network attacks.
+
+Ahmed \emph{et al.} \cite{ahmed2016survey} highlight a observed lack of publicly available labeled data set for network anomaly detection, and discuss recent research related to overcome the issues of publicly available network intrusion evaluation data sets, including a list of current data sets and an evaluation of attack types and available labels.
+According to \cite{ahmed2016survey}, the available data sets present limitations on availability of labeled traffic for classification of background, legitimate and malicious traffic.
+The data imbalance between legitimate and malicious traffic is a important concern for real world problems of network attack, but the principal data sets usually adopt a balanced distribution between legitimate and malicious traffic, what make the attack detection more feasible for general classifiers and machine learning algorithms than for anomaly detection algorithms.
+Moreover, according to Ahmed \emph{et al.} \cite{ahmed2016survey}, the most adopted data set are not up-to-date with current or novel attacks, tools, operational systems and network topology.
+
+Several approaches for network attack detection uses the KDD 99 \cite{ji2016multi, ahmed2016survey, osanaiye2016distributed, bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
+Even though the KDD 99 data set are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data sets adopted by researchers \cite{osanaiye2016distributed, bhuyan2014network}.
 NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that redundant data records are removed, in order to avoid biased classifications.
 Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public data sets their evaluations.
-Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD data sets adopt a traffic aggregation by connections, we consider the use of a experimental scenario on a real network and the DARPA 1998 data set, which is the source for the creation of the KDD 99 and NSL-KDD datas ets.
+
+Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD data sets adopt a traffic aggregation by connections, we consider the use of an experimental scenario on a real network and the DARPA 1998 data set, which is the source for the creation of the KDD 99 and NSL-KDD data sets.
 Note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD data sets.
@@
 The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
+The $l_2$ norm is denoted by $\norm{\cdot}$.
 We define the operator $\rm diag(\cdot)$ that returns the vector of the main diagonal of a given matrix, the operator $\rightarrow$, which denotes the deletion of a given element from a set and the operator $\#$, that returns the rank of a matrix, and the operator $\sim$ that sorts the elements of a vector in ascending order.
 
-This section also presents details of the experimental scenario and the selected cases of the DARPA data set, along with a description of the data set model as a signal superposition of signal, artifact and noise, that refer to legitimate, control and malicious traffic.
-Subsection \ref{sec:2_DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks.
-Subsection \ref{sec:2_ModelingData} presents how network traffic can be modeled as signal superposition.
-Subsection \ref{sec:2_SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:2_Darpadata set} discusses the use of the DARPA data set for evaluation of the proposed approach.
+This section also presents details of the experimental scenario and the selected cases of the DARPA data set.
+In Subsection \ref{sec:2_DataCollection} we describe the environment and scenario adopted in order to reproduce flood and probe attacks.
+In Subsection \ref{sec:2_ModelingData} are presented how network traffic can be modeled as signal superposition.
+We detail, in Subsection \ref{sec:2_SynfloodFraggleandPortscan}, the traffic of synflood, fraggle and port scan attacks, and in Subsection \ref{sec:2_Darpadata set} are discussed the use of the DARPA data set for evaluation of the proposed approach.
 
@@
 
-%TODO - recreate the figure and change the concepts
 \begin{figure}[h!]
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig09.png}
-     \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
+     \includegraphics[width=9cm]{figures/ch2/scenario.eps}
+     \caption{Scenario to reproduce legitimate traffic, flood and port scan.}
      \label{fig:2.01}
@@
 
-Note that the set of network traffic is modeled as legitimate, controls and malicious traffic, which refer to signal, artifact and noise, respectively.
 In many organizations the web based traffic is predominant, since most of corporate services are  web pages, customized web-based systems and cloud services.
-It is possible to characterize the traffic of a DHCP service as an example of artifact associated with the transport layer, as well as it would be possible to classify heart beat traffic and other seasonal and controlled traffic as artifact.
+It is possible to characterize the traffic of a Dynamic Host Configuration Protocol (DHCP) service as an example of legitimate associated with the application layer, as well as it would be possible to classify seasonal and controlled traffic as legitimate traffic.
 For malicious traffic, three types of networks attacks are evaluated: synflood, fraggle and port scan.
-Here we refer to port scan as an attack, according to \cite{ahmed2016survey,moustafa2019holistic}, however it is one approach usually adopted for acquire information in order to perform an attack.
+Here we refer to port scan as an attack, according to \cite{ahmed2016survey, moustafa2019holistic}, however it is one approach usually adopted for acquire information in order to perform an attack.
 These attacks are reproduced using well-known security tools, such as Nmap\footnote{http://nmap.org} for port scan, Metasploit\footnote{http://www.metasploit.com} for synflood and Hping\footnote{http://hping.org} to lead the fraggle attack.
@@
 specific information such as origin or destination IP, behavioral pattern or content of the attack are not considered.
-Therefore, IP spoofing or data encryption would not cause impact to the proposed approach and evaluation, since our proposal only relies on the timestamp (for sequencing) and port number.
+Therefore, IP spoofing or data encryption would not cause impact to the proposed approach and evaluation, since our proposal only relies on the timestamp (for sequencing) and destination port number.
 
@@
 
-By modeling the data set as a signal superposition, the network traffic ($\pmb{X}$) can be characterized as a mixture of three components: legitimate traffic ($\pmb{U}$), artifact ($\pmb{A}$) and malicious traffic ($\pmb{N}$), according to the following expression:
+By modeling the data set as a signal superposition, the network traffic ($\pmb{X}$) can be characterized as a mixture of two components: legitimate traffic ($\pmb{U}$) and malicious traffic ($\pmb{N}$), according to the following expression:
 
 \begin{equation}\label{eq:2.01}
-	\pmb{X}^{(q)} = \pmb{U}^{(q)} + \pmb{A}^{(q)} + \pmb{N}^{(q)},
+	\pmb{X}^{(q)} = \pmb{U}^{(q)} + \pmb{N}^{(q)},
 \end{equation}
@@
 
-The legitimate traffic $\pmb{U}^{(q)}$ is characterized by the traffic from user's ordinary operations, such as when performing access to web pages, as well when there is the traffic required to domain name resolution.
-Figure \ref{fig:2.03} depicts an example of the legitimate traffic simulated during the experiments.
-
-% TODO - Replace
+The legitimate traffic $\pmb{U}^{(q)}$ is characterized by user's ordinary operations and by legitimate traffic that are automatically generated for network management and for background services.
+The access to web pages by users or the name resolution by means of the Domain Name System (DNS) are examples of legitimate traffic generated by user operations, as can be seen in Figure \ref{fig:2.03}, which depicts the legitimate traffic simulated to reproduce user's operations during the experiments.
+
 \begin{figure}[h!]
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig03.png}
-     \caption{Traffic from user's operations, that can be characterized by web access, traffic of well-known applications or network protocols.}
+     \includegraphics[width=11cm]{figures/ch2/user_traffic.eps}
+     \caption{Traffic from user's operations.}
      \label{fig:2.03}
@@
 
-The traffic that is not associated with user's operations and with malicious traffic is modeled as artifact $\pmb{A}^{(q)}$.
-The acquisition of logical IP network address (DHCP) is an example of artifact, where independently of any user operation, the machine receives an IP address, since it is configured to automatically perform a DHCP address request.
-Figure \ref{fig:2.04} depicts an example of artifact in a network traffic, by means of traffic to ports 67 and 68.
-The traffic coming from a malicious activity, i.e.
-port scanning, synflood or fraggle attacks, is represented by the matrix $\pmb{N}^{(q)}$.
-
-% TODO - Replace
+The Figure \ref{fig:2.04} depicts an example of legitimate traffic of user independent operations, by means of traffic to ports 67 and 68, where it is possible to observe a low amount of packets.
+The acquisition of logical IP network address by means of DHCP is an example of legitimate traffic not associated with a user, where independently of any user operation, the machine receives an IP address, since it is configured to automatically perform a DHCP address request.
+
 \begin{figure}[h!]
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig04.png}
+     \includegraphics[width=11cm]{figures/ch2/auto_traffic.eps}
      \caption{Network traffic of user independent operations for network management.}
@@
 
-We define that if the $\#\pmb{X}^{(q)} ≠ 0$, then there is malicious traffic in the evaluated time frame $q$, on the other hand, if the $\#\pmb{X}^{(q)} = 0$, then there is no malicious traffic.
+The traffic coming from a malicious activity, i.e.
+port scanning, synflood or fraggle attacks, is represented by the matrix $\pmb{N}^{(q)}$.
+We define that if the rank of $\pmb{X}^{(q)}$ is not zero, according to $\#\pmb{X}^{(q)} \neq 0$, which denotes that the rank of the $q$-th $\pmb{X}$ is different of 0, then there is malicious traffic in the evaluated time frame $q$.
+On the other hand, if the $\#\pmb{X}^{(q)} = 0$, then there is no malicious traffic in time frame $q$.
 We show how to detect the $\#\pmb{X}^{(q)}$, given only the matrix $\pmb{X}^{(q)}$, in order to identify malicious traffic.
@@
 DoS is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks.
-
 Probe attacks scan computer and network systems to collect information about the host, such as open ports, topology, running software or version of technologies, in order to find vulnerabilities.
@@
 
-% TODO - Replace
 \begin{figure}[h!]
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig05.png}
+     \includegraphics[width=11cm]{figures/ch2/synflood_traffic.eps}
      \caption{A large quantity of SYN requests to a target, in order to cause a DoS.}
@@
 Therefore, each host receives a huge amount of requests UDP echo and all of them replies to the IP address of the victim, causing a packet flooding aiming a DoS.
-Figure \ref{fig:2.06} depicts an example of the fraggle attack in a real computer network.
-
-The fraggle attack can affect the entire network, since all hosts receive several requests UDP echo and respond with the ICMP protocol, therefore each host acts as an amplifier of the attack.
-This part of the fraggle attack is not taken into account in this work, because the victim receives ICMP (network layer) packets originated from the hosts that are attacked with flooding packet UDP echo.
-
-% TODO - Replace
+
+The Figure \ref{fig:2.06} depicts an example of the fraggle attack in a real computer network, and shows that more than 6,000,000 malicious packets can be counted in an interval of ten minutes, which can be considered an abnormal network traffic, due to the concentrated traffic in a short period of time and due to the similarity of the outstanding traffic.
+
+The fraggle attack can affect the entire network, since all hosts receive several requests UDP echo and respond with the  Internet Control Message Protocol (ICMP), therefore each host acts as an amplifier of the attack.
+This part of the fraggle attack is not taken into account in this work, because the victim receives ICMP packets originated from the hosts that are attacked with flooding packet UDP echo.
+
 \begin{figure}[h!]
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig06.png}
+     \includegraphics[width=11cm]{figures/ch2/fraggle_traffic.eps}
      \caption{Large amount of “UDP echo” requests and replies, causing packet flooding.}
@@
 \end{figure}
-
-The Figure \ref{fig:2.06} shows that more than 6,000,000 malicious packets can be counted in an interval of ten minutes, which can be considered an abnormal network traffic, especially due to the concentrated traffic in a short period of time and due to the similarity of the outstanding traffic.
 
@@
      \centering 
-     \includegraphics[width=11cm]{figures/ch2/fig07.png}
+     \includegraphics[width=11cm]{figures/ch2/portscam_traffic.eps}
      \caption{Connection attempts in order to identify active ports.}
@@
 
-Figure \ref{fig:2.07} depicts an example of the port scan attack in a real computer network.
-Note that the traffic is composed of two packets for each TCP port and one UDP packet to each port.
+Figure \ref{fig:2.07} depicts an example of port scan attack in a real computer network.
+We simulated a traffic with two packets for each TCP port and one UDP packet to each port.
 The incoming and outgoing packets analysis, for each port, shows the high correlation and similarity of TCP and UDP traffic during the simulated port scan attack.
@@
 
-We observe that the most cases of DoS of DARPA 98 focus on exploit system vulnerabilities instead of on flooding attack.
+The most cases of DoS of DARPA 98 focus on exploit system vulnerabilities instead of on flooding attack.
 One example is the occurrence of a neptune attack which sends 20 SYN packets, what is a behavior that differs of the expected flooding attack behavior.
-Therefore, there were selected the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey,osanaiye2016distributed}, and the cases that scan ports sending just a few packets.
+Therefore, there were selected the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey, osanaiye2016distributed}, and the cases that scan ports sending just a few packets.
 From the simulated probe attacks, we select the cases that rely on TCP or UDP connections.
@@
 
-This section describes the proposed technique to detect synflood, fraggle and port scan, according to the overview depicted in Figure \ref{fig:2.08}, which represents the proposed framework for detection and identification of network attacks.
+According to the overview depicted in Figure \ref{fig:2.08}, we present in this section the proposed framework for detection and identification of network attacks.
+
 In Subsection \ref{sec:2_prop_LargestEigenvaluebyTimeFrames} we present the steps for extraction of the largest eigenvalue for each $q$-th time frame.
-Next, in Subsection \ref{sec:2_prop_MOSSchemes}, we show how to apply the eigenvalues on the MOS scheme in order to detect the attack.
-In Subsection \ref{sec:2_prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:2_prop_EigenSimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.
+Next, in Subsection \ref{sec:2_mos} are presented the mathematical concepts and examples of state-of-the-art MOS schemes, and how to apply the eigenvalues on the MOS scheme in order to detect the attack.
+In Subsection \ref{sec:2_prop_EigenvalueAnalysis}, we present the eigenvalue analysis to identify the time frames detected as under attack, and the Subsection \ref{sec:2_prop_EigensimilarityAnalysis} describes the similarity analysis evaluated for detailed attack identification.
 
@@
 	\centering
-     \includegraphics[width=11cm]{figures/ch2/mos_eigen_similarity.eps}
-     \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
+     \includegraphics[width=12cm]{figures/ch2/mos_eigen_similarity.eps}
+     \caption{Overview of the framework for detection and identification of network attacks.}
      \label{fig:2.08}
@@
 
-The proposed attack detection algorithm starts by the data pre-processing of a network traffic log containing IP, ports and timestamp of senders and receivers.
-During this step, the desired information is extracted in order to classify and to count packets according to the origin and destination ports.
+The proposed attack detection algorithm starts by the data preprocessing of a network traffic log containing IP, ports and timestamp of senders and receivers.
+During this step, the desired information is extracted in order to count packets according to the destination ports by time.
 Subsequently, this information is grouped by minutes and by time frames.
 
-With the data grouped into $Q$ time frames, the framework considers the time variations of the matrix $\pmb{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$, in order to detect the attack.
+With the data grouped into $Q$ time frames, the framework is initially applied to each matrix $\pmb{X}^{(q)} \in \mathbb{R}^{M\times{N}}$, with $q = 1, \ldots, Q$.
+According to Subsection \ref{sec:2_ModelingData}, $q$ represents the $q$-th time frame of aggregated network traffic.
+The matrix $\pmb{X}^{(q)} \in \mathbb{R}^{M \times N}$ consists of \emph{M} network ports and \emph{N} time bins.
+We adopt time bins with 1 minute of aggregated traffic and time frames with 20 observations.
+We assume that the count of ports is defined according to $M < N$, and adopt 17 network ports for our evaluation.
+Hence, we have $\pmb{X}^{(q)} \in \mathbb{R}^{17\times{20}}$ and each element $x_{m,n}^{(q)}$ stands for the number of packets at $n$-th minute for the port $m$, during the $q$-th time frame.
+
+The time frame size is an important concern to define the sampling size for estimating the sample covariance matrix and the eigenvalue decomposition, considering that if the sample size \emph{N} is small and the number of variables \emph{M} is large, the empirical estimators of covariance and correlation can be unstable and the empirical estimate of the covariance matrix becomes singular, i.e.
+it cannot be inverted to compute the precision matrix.
+
+However, high-dimension, low-sample-size (HDLSS) data are emerging in many areas, such as genetic, imaging, text classification, finance and face recognition.
+Thus, many methods of shrinkage and regularization have been proposed to improve the stability for estimation of the covariance matrix \cite{chen2011robust} and eigenvalues \cite{yata2010effective}.
 
 According to flood and port scan attacks' behavior, flood attacks and port scan attacks can be characterized as covariance aware attack \citep{jin2004covariance} and correlation aware attack \citep{lakhina2005mining}, respectively.
-These characteristics are substantiated by the results obtained through the analysis based on sample covariance of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:2_experimentalresults}, which shows that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
+These characteristics are substantiated by the results obtained through the analysis based on sample covariance of zero mean variables and on covariance of zero mean and unitary standard deviation variables, described in Section \ref{sec:2_experimentalresults}.
+
+The results in Section \ref{sec:2_experimentalresults} show that the main components of flood attacks are dominated by the variables with more variance and that the traffic associated with port scan attack does not generate many logs, however, it presents high covariance of zero mean and unitary standard deviation variables.
 
@@
 
-\subsection{MOS Schemes}
-\label{sec:2_prop_MOSSchemes}
-
-Traditionally the MOS schemes are applied for the eigenvalues of the vector $\pmb{e}^{(q)}$.
-However, the goal here is to detect the variations of the eigenvalues for different values of $q$.
-Therefore, instead of using a certain $q$, the proposed approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame, in order to identify variations and estimate the model order $\hat{d}$, which is the estimated number of time frames under attack.
+\subsection{Model Order Selection (MOS)}
+\label{sec:2_mos}
+
+The model order selection is a key point in many digital signal processing applications, including radar, sonar, communications, channel modeling, medical imaging, among others \cite{da2011subspace, da2011multi, xiong2017bayesian}.
+MOS allows analysis of reduced data set, through separating noise components of the main components, for example.
+Moreover, the model order is crucial for many parameter estimation techniques \cite{da2007enhanced, da2009comparison}, since the amount of parameters to be estimated depends on the model order.
+
+The model selection procedure chooses the ``best'' model of a finite set of models, according to some criteria \cite{rajan1997model}.
+Therefore, given some data set, it is chosen a model which was evaluated as the best model to describe the specified data set.
+
+The state of the art regarding estimation techniques of model order based on eigenvalues includes: Akaike's Information Theoretic Criterion - AIC \cite{akaike1974new,wax1985detection}; Minimum Description Length - MDL \cite{barron1998minimum,wax1985detection}; Efficient Detection Criterion - EDC \cite{zhao1986detection}; Stein's Unbiased Risk Estimator - SURE \cite{ulfarsson2008rank}; RADOI \cite{radoi2004new} and Exponential Fitting Test - EFT \cite{grouffaud1996some, quinlan2006model, david2011blind}.
+
+In AIC, MDL and EDC techniques, the information criterion is a function of the geometric mean $g(k)$ and the arithmetic mean $a(k)$ relating to smaller $k$ eigenvalues, where $k$ is a candidate value for the model order $d$ \cite{da2009comparison}.
+
+Basically, the difference between the AIC, MDL and EDC schemes is the penalty function $p(k, N, \alpha)$, so these techniques can be written in general as \cite{da2009comparison}:
+
+\begin{equation}\label{eq:eq1}
+  \hat{d} = \argmin\limits_k \hspace{1 mm} J(k), 
+\end{equation}
+
+where
+
+\begin{equation}\label{eq:eq2}
+  J(k) = -N(\alpha - k) \hspace{1 mm} {\rm log} \hspace{1 mm} (g(k)/a(k)) + p(k,N,\alpha),
+\end{equation}
+
+where $\hat{d}$ is an estimate $d$ of the model order, $N$ is the number of samples, $\alpha = M$ and means the number of variables of the problem, and $0 \leqslant k \leqslant min[M, N]$.
+
+Penalty functions for AIC, MDL and EDC are given by the Table \ref{tab:mos_penalty_functions}.
+
+\begin{table}[h!]
+  \centering
+  \caption{Penalty functions for the schemes AIC, MDL and EDC}
+  \label{tab:mos_penalty_functions}
+  \begin{tabular}{*2c}
+	\toprule
+	\textbf{Scheme} &  \textbf{Penalty function} \\
+	\textbf{} &  $p(k,N,\alpha)$ \\
+	\midrule
+    AIC	& $k(2\alpha - k)$ \\
+    MDL	& $0.5k(2\alpha - k) \hspace{1 mm} {\rm log} (N)$ \\
+    EDC	& $0.5k(2\alpha - k)\sqrt{N\hspace{1 mm}\rm ln(\rm ln N)}$ \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+The Exponential Fitting Test (EFT) can effectively be used in cases where the number of samples $N$ is small.
+This technique is based on observations of data contaminated only with white noise, where the profile of eigenvalues can be approximated by an exponential decaying \cite{grouffaud1996some}.
+
+Given $\lambda_i$ be the i-th eigenvalue, the exponential model can be expressed by:
+\begin{equation}\label{eq:eq3}
+  E\{\lambda_i\} = E\{\lambda_1\} \cdot q(\alpha,\beta)^{i-1},
+\end{equation}
+
+where $E\{\cdot\} $ is the expectation operator, and it is considered that the eigenvalues are ordered in the that $\lambda_1$ represents the largest eigenvalue.
+The term $q(\alpha, \beta)$ is defined as:
+
+\begin{equation}\label{eq:eq4}
+  q(\alpha,\beta) = \exp\left\{-\sqrt{\frac{30}{\alpha^2 + 2} - \sqrt{\frac{900}{(\alpha^2 + 2)^2} - \frac{720\alpha}{\beta(\alpha^4 + \alpha^2 - 2)}}} \right\},
+\end{equation}
+
+where $0 < q(\alpha,\beta) < 1$. According to \cite{quinlan2006model}, if $M \leq N$, then $\beta = N$.
+
+Traditionally, the MOS schemes are applied for the eigenvalues, that are the vector $\pmb{e}^{(q)}$ when considering the application of MOS schemes for Eigensimilarity.
+However, the goal of our proposal is to detect the variations of the eigenvalues for different values of $q$.
+
+Instead of using a certain $q$, the proposed approach applies MOS schemes for a vector of the largest eigenvalues of each $q$-th time frame, in order to identify variations and estimate the model order $\hat{d}$, which is the estimated number of time frames under attack.
 Therefore, $\pmb{e}_{\rm max}$ is sorted in descending order, producing $\sim\pmb{e}_{\rm max}$, that is used as input parameter for MOS schemes, according to $\hat{d} = \rm{MOS}(\sim\pmb{e}_{\rm max})$.
 Note that some MOS schemes may also require the number of minutes that compose a time frame, as $\hat{d} = \rm{MOS}(\pmb{e}_{\rm max},\emph{Q})$.
-For more information about MOS, we refer to the Appendix \ref{apx:a_mos}.
 
@@
 
-After applying the MOS schemes to the vector $\sim\pmb{e}_{\rm max}$, we obtain the estimate of the $\#\pmb{X}$.
-For instance, in the case of fraggle, synflood and ports can, if $\hat{d} = 1$, , then $\#\pmb{X} = 1$, which means that during the during the $Q$ time frames one attack is present.
-However, if $\hat{d} = 0$, then $\#\pmb{X} = 0$, and this means that none of these attacks are present.
-Note that $\hat{d}$ can be greater than 1, indicating the presence of more than one attack.
-
-In Subsection \ref{sec:2_prop_MOSSchemes}, we obtained only if $\hat{d} = 1$ or $\hat{d} = 0$.
-However, if $\hat{d} = 1$, the MOS schemes do not provide any information about the $q$-th time frame under attack.
-The identification of the $q$-th time frame under attack can be carried out through a eigenvalues analysis.
+After applying the MOS schemes to the vector $\sim\pmb{e}_{\rm max}$, we obtain the rank estimate of the $\#\pmb{X}$.
+For instance, in the case of fraggle, synflood and ports can, if $\hat{d} = 1$, then $\#\pmb{X} = 1$ indicate a estimate of rank 1 for $\pmb{X}$, which means that during the during the $Q$ time frames one time frame $q$ is under attack.
+However, if $\hat{d} = 0$, then $\#\pmb{X} = 0$, and this means that none of the $Q$ time frames is under attack.
+Note that $\hat{d}$ can be greater than 1, indicating the presence of more than one attacked time frame.
+
+In Subsection \ref{sec:2_mos}, we obtained only if $\hat{d} = 1$ or $\hat{d} = 0$, estimating the number of time frames under attack.
+However, if $\hat{d} > 1$, the MOS schemes does not identify the $q$-th attacked time frames.
+The identification of the $q$-th time frame under attack can be carried out through an eigenvalues analysis.
 
@@
 
-\subsection{Eigen Similarity Analysis}
-\label{sec:2_prop_EigenSimilarityAnalysis}
-
+\subsection{Principal Eigenvector Similarity Analysis}
+\label{sec:2_prop_EigensimilarityAnalysis}
+
+The eigenvector corresponding to the eigenvalue of largest magnitude is called the principal eigenvector or dominant eigenvector.
 Cosine similarity calculates the cosine of the angle between two vectors, which represents the similarity of values between the selected vectors.
-Therefore, cosine similarity can be used to evaluate the variation of the most significant eigenvectors of $\pmb{V}^{(q)}$ against the most significant eigenvectors of time frame detected as under attack, to analyze similarity changes into the most significant eigenvectors caused by the insertion of anomalous traffic \cite{Lee2013}.
-
-This subsection describes the proposed eigen similarity analysis for detailed attack identification, in complement to the attack estimation carried out through MOS schemes and eigenvalue analysis.
-In Subsection \ref{sec:2_prop_TimeSimilarityAnalysis} we present the eigen similarity analysis for identification of time under attack.
-Next, in Subsection \ref{sec:2_prop_PortSimilarityAnalysis}, we show how to apply the eigen similarity analysis in order to identify network ports under attack.
+Therefore, cosine similarity can be used to evaluate the difference between the principal eigenvector $\pmb{V}^{(q)}$ against the principal eigenvector of one time frame detected as under attack, in order to analyze similarity changes between the principal eigenvectors with the largest variance caused by the insertion of anomalous traffic \cite{Lee2013} in a time frame.
+
+The cosine similarity is a measure between two vectors and does not consider the distance between points or features of a vector.
+Therefore, the computed angle between eigenvectors does not reveal element-wise distance neither the contribution of each component for the measured distance, but can be used to identify the principal eigenvector of observations that deviates from a reference principal eigenvector and can identify network attacks.
+After the identification of the time under attack by means of the cosine similarity between principal eigenvectors, the cosine similarity can also be applied for identification of components that insert the deviation between the principal eigenvectors, according explained in subsection \ref{sec:2_prop_PortSimilarityAnalysis}.
+
+This subsection describes the proposed principal eigenvector similarity analysis for detailed attack identification, in complement to the attack estimation carried out through MOS schemes and eigenvalue analysis.
+In Subsection \ref{sec:2_prop_TimeSimilarityAnalysis} we present the principal eigenvector similarity analysis for identification of time under attack.
+Next, in Subsection \ref{sec:2_prop_PortSimilarityAnalysis}, we show how to apply the principal eigenvector similarity analysis in order to identify network ports under attack.
 
@@
 \label{sec:2_prop_TimeSimilarityAnalysis}
-% ToDo
-For eigen similarity analysis, we evaluate the cosine similarity in order to identify lacks of similarity between legitimate and malicious traffic, as follows
+
+For principal eigenvector similarity analysis, we evaluate the cosine similarity in order to identify lacks of similarity between legitimate and malicious traffic, as follows:
 
@@
 \end{equation}
-where $s_n$ denotes the absolute similarity degree of the $n$-th minute, $\pmb{v}^{(q)}$ is the most significant eigenvectors of a selected set of minutes without network attack, according the attack detection described by Algorithm \ref{alg:2.01}, and $\pmb{v}_{(n)}^{(q)}$ is the most significant eigenvectors obtained after append each target $n$-th minute of traffic that needs the identification of flood and port scan attacks.
-
-The most significant eigenvector $\pmb{v}^{(q)}$, of a time frame $q$ without attack, can be computed from (\ref{eq:2.06}) and selected according to the eigenvector of the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the selected time frame $q$.
-The same calculation shall be performed in order to obtain the target eigenvectors $\pmb{v}_{(n)}^{(q)}$, calculated from the time frame without attack appended by selected minutes of a time frame estimated as under attack.
-
-Therefore, the reference eigenvectors $\pmb{v}^{(q)}$ is calculated from the traffic without attack, in a time frame $q$ composed of $Q$ minutes of legitimate network traffic, estimated as normal by Algorithm \ref{alg:2.01}.
+where $s_n$ denotes the absolute similarity degree between the $n$-th minute and the reference principal eigenvector, given by an inner product that measures the cosine of the angle between two vectors, where the $l_2$ norm is denoted by $\norm{\cdot}$ and $\abs{\cdot}$ denotes the absolute value.
+
+In (\ref{eq:2.11}), the $\pmb{v}^{(q)}$ denotes the principal eigenvector of a selected set of minutes without network attack, according the attack detection described by Algorithm \ref{alg:2.01}, and $\pmb{v}_{(n)}^{(q)}$ is the principal eigenvector obtained after append each target $n$-th minute of traffic that needs the identification of flood and port scan attacks.
+
+The principal eigenvector $\pmb{v}^{(q)}$ of a time frame $q$ without attack can be computed from (\ref{eq:2.06}) and selected according to the eigenvector corresponding to the largest eigenvalue $\lambda_1^{(q)}$, which is the principal component of the selected time frame $q$.
+The same calculation shall be performed in order to obtain the target principal eigenvector $\pmb{v}_{(n)}^{(q)}$, calculated from the time frame without attack appended by selected minutes of a time frame estimated as under attack.
+
+Therefore, the principal eigenvector $\pmb{v}^{(q)}$ is calculated from the traffic without attack, in a time frame $q$ composed of $Q$ minutes of legitimate network traffic, estimated as normal by Algorithm \ref{alg:2.01}.
 For the detailed attack identification, each $\pmb{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\hat{\pmb{q}}}_{\rm max}$ time frames shall be individually appended into $\pmb{X}^{(q)}$, as represented by
@@
 
-We propose three approaches for eigen similarity analysis, which are the incremental, the individual and the incremental individualized.
-
-The incremental approach for eigen similarity analysis is based on the incremental appending of network traffic into $\pmb{X}^{(q)}$, where the first evaluation is based on (\ref{eq:2.12}) and the subsequent evaluations are based on (\ref{eq:2.13}), that denotes the appending of the $n$-th minute $\pmb{x}^{(\hat{q})}_{(n)}$ into $\pmb{X_n}$.
-The incremental approach repeats the Equation \ref{eq:2.13} while $n \leq N$ and compute $\pmb{v}^{(q)}$, $\pmb{v}_{(n)}^{(q)}$ and the Equation \ref{eq:2.11} for each increment.
+We propose three approaches for principal eigenvector similarity analysis, which are the incremental, the individual and the incremental individualized.
+
+The incremental approach for principal eigenvector similarity analysis is based on the incremental appending of network traffic into $\pmb{X}^{(q)}$, where the first evaluation is based on (\ref{eq:2.12}) and the subsequent evaluations are based on (\ref{eq:2.13}), that denotes the appending of the $n$-th minute $\pmb{x}^{(\hat{q})}_{(n)}$ into $\pmb{X_n}$.
+The incremental approach repeats the (\ref{eq:2.13}) while $n \leq N$ and compute $\pmb{v}^{(q)}$, $\pmb{v}_{(n)}^{(q)}$ and the (\ref{eq:2.11}) for each increment.
 
@@
 
-Figure \ref{fig:2.08} illustrates the network traffic selection for the incremental approach of eigen similarity analysis, where the $\pmb{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected by Algorithm \ref{alg:2.01}.
+Figure \ref{fig:2.88} illustrates the network traffic selection for the incremental approach of principal eigenvector similarity analysis, where the $\pmb{X}^{(1)}$ is chosen as reference for similarity analysis of the $m$-th minutes of the time frame $q=3$, where one network attack was previously detected by Algorithm \ref{alg:2.01}.
 
@@
      \caption{Traffic selection for incremental approach.}
-     \label{fig:2.08}
+     \label{fig:2.88}
 \end{figure}
 
-For the scenario depicted by Figure \ref{fig:2.08}, the eigen similarity analysis starts at $\pmb{x}^{(3)}_{(41)}$ and is incrementally performed until $\pmb{x}^{(3)}_{(60)}$, in order to calculate the $s_n$.
-We assume that $s_n < l$ means an attack identification, according the anomaly on similarity of $s_n$ compared to a defined limiar $l$.
-
-Therefore, after obtaining the most significant eigenvector $\pmb{v}^{(q)}$ and the target eigenvectors $\pmb{v}_{(n)}^{(q)}$ for eigen similarity analysis, the $s_n$ is calculated according to (\ref{eq:2.11}).
-If $s_n = 1$, then the two eigenvectors are completely similar and no anomaly is detected.
+For the scenario depicted by Figure \ref{fig:2.08}, the principal eigenvector similarity analysis starts at $\pmb{x}^{(3)}_{(41)}$ and is incrementally performed until $\pmb{x}^{(3)}_{(60)}$, in order to calculate the $s_n$.
+We assume that $s_n < l$ means an attack identification, according the anomaly on similarity of $s_n$ compared to a defined threshold $l$.
+
+Therefore, after obtaining the principal eigenvector $\pmb{v}^{(q)}$ and the target principal eigenvector $\pmb{v}_{(n)}^{(q)}$ for principal eigenvector similarity analysis, the $s_n$ is calculated according to (\ref{eq:2.11}).
+If $s_n = 1$, then the two principal eigenvectors are completely similar and no anomaly is detected.
 Smaller values of $s_n$ mean less similarity and can indicate an anomaly if $s_n < l$, what denotes that a network attack is identified during the $n$-th minute.
 
-The Equation \ref{eq:2.14} shows how the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack is identified, where
+The (\ref{eq:2.14}) shows how the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if an attack is identified, where
 
@@
 
-The eigen similarity analysis can also be applied by means of the individual approach, where each $n$-th minute must be individually appended into $\pmb{X}^{(q)}$, as shown by Figure \ref{fig:2.09}.
-In the individual approach there is no incremental appending, therefore only individual $n$-th minute are appended to $\pmb{X_n}$, according to \ref{eq:2.12}, in order to compute $\pmb{v}^{(q)}$, $\pmb{v}_{(n)}^{(q)}$ and the Equation \ref{eq:2.11} for individual $n$-th minute.
+The principal eigenvector similarity analysis can also be applied by means of the individual approach, where each $n$-th minute must be individually appended into $\pmb{X}^{(q)}$, as shown by Figure \ref{fig:2.09}.
+In the individual approach there is no incremental appending, therefore only individual $n$-th minute are appended to $\pmb{X_n}$, according to (\ref{eq:2.12}), in order to compute $\pmb{v}^{(q)}$, $\pmb{v}_{(n)}^{(q)}$ and the (\ref{eq:2.11}) for individual $n$-th minute.
 
@@
 $s_n < l$.
-Subsequently, $\pmb{X}_{n-1}$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:2.02}.
+Subsequently, $\pmb{X}_{n-1}$ becomes the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated, as shown in Figure \ref{fig:2.22}.
 
@@
      \caption{Traffic selection for incremental individualized approach.}
-     \label{fig:2.02}
+     \label{fig:2.22}
 \end{figure}
@@
 The incremental similarity analysis followed by individual analysis after an attack detection allows to identify the attack period, highlighting the first and last time under attack.
-This identification is possible due to the variation of the most significant eigenvectors, which becomes more significant when compared a traffic under attack against a traffic with no attack, according to results which are discussed in Section \ref{sec:2_experimentalresults}.
+This identification is possible due to the similarity variation between the principal eigenvectors, which highlight lacks of similarity when compared the principal eigenvector of a traffic under attack against the principal eigenvector of a traffic with no attack, according to results which are discussed in Section \ref{sec:2_experimentalresults}.
 
@@
 Given $\hat{\pmb{n}}$, which is the set of $n$-th minutes under attack, it is still necessary to obtain more details about the identified network attack, such as the network ports that are attacked during each $n$-th minute identified as under attack.
-Hence, it is also applied the cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port during a $n$-th minute.
-
-For detection of ports under attack, the last most significant eigenvectors without attack $\pmb{v}^{(q)}$ shall be used as reference for similarity analysis against the $\pmb{v}_{(n)}^{(q)}$ identified as under attack, and evaluate individually the cosine similarity of each $m$-th port of all $\hat{\pmb{n}}$ minutes.
+Hence, it is also applied the cosine similarity analysis to identify variation of the principal eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port during a $n$-th minute.
+
+For detection of ports under attack, the last principal eigenvector without attack $\pmb{v}^{(q)}$ shall be used as reference for similarity analysis against the $\pmb{v}_{(n)}^{(q)}$ identified as under attack, and evaluate individually the cosine similarity of each $m$-th port of all $\hat{\pmb{n}}$ minutes.
 Therefore, $\pmb{v}^{(q)}$ should be calculated from the last $\pmb{X}^{(q)}$ time frame without attack, and $\pmb{v}_{(m,\hat{n})}$ should be calculated from the same traffic appended of all $n$-th minutes until the identified minute under attack, denoted as $\pmb{X}_n$.
@@
 \end{equation}
-where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute and $q$-th time frame identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\pmb{X}_n$, which is used to calculate the $\pmb{v}_{(m,\hat{n})}$ most significant eigenvectors that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.
-
-Once $\pmb{v}^{(q)}$ and $\pmb{v}_{(m,\hat{n})}$ are obtained, then the $s_{m,\hat{n}}$ similarity degree can be calculated in order to identify if the traffic replacement highlights the adition of anomalous traffic by the evaluated $m$-th port during the $\hat{n}$-th minute previously identified as under attack.
+where $x^{(\hat{q})}_{(m,\hat{n})}$ denotes the $m$-th port of the selected $n$-th minute and $q$-th time frame identified as under attack and $x_{(m,n)}$ denotes the $m$-th port of the last $n$-th minute of $\pmb{X}_n$, which is used to calculate the $\pmb{v}_{(m,\hat{n})}$ principal eigenvector that contains the traffic of the $m$-th port of the $\hat{n}$-th minute identified as under attack.
+
+Once $\pmb{v}^{(q)}$ and $\pmb{v}_{(m,\hat{n})}$ are obtained, then the $s_{m,\hat{n}}$ similarity degree can be calculated in order to identify if the traffic replacement highlights the addition of anomalous traffic by the evaluated $m$-th port during the $\hat{n}$-th minute previously identified as under attack.
 
@@
 
-This section presents the performed experiments and the acquired results for the Eigensim.
+This section presents the performed experiments and the acquired results for the Eigensimilarity.
 First, in Section \ref{sec:2_AnalyzedScenario}, the experimental scenario adopted in the evaluation is summarized.
-Then, Section \ref{sec:2_largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario.
-In Section \ref{sec:2_MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated data set.
-Section \ref{sec:2_EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:2_EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario.
-Section \ref{sec:2_DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 data set.
+Then, in Section \ref{sec:2_largesteigenvaluesanalysis} the results of the largest eigenvalue analysis by time frames for the experimental scenario are shown.
+In Section \ref{sec:2_MOSSchemesEvaluation} we describe the results of the evaluated MOS schemes for attack detection in the simulated data set.
+In Section \ref{sec:2_EigenvalueAnalysis} we present the results of the eigenvalue analysis for identification of time frames under attack.
+In Section \ref{sec:2_EigensimilarityAnalysis} we show the results of similarity analysis for detailed flood and port scan identification for the experimental scenario.
+In Section \ref{sec:2_DarpaEvaluation} we present the results of the proposed framework for flood and probe attack detection in the DARPA 1998 data set.
 
@@
 
-This experiment consider a simulated scenario of a real network monitored during 120 minutes, that are separated into six time frames of twenty minutes.
+This experiment considers a simulated scenario of a real network monitored during 120 minutes, that are separated into six time frames of twenty minutes.
 Therefore, as the time of each sampling period is one minute, then $N = 20$.
@@
 For the evaluation of MOS Schemes accuracy for flood and port scan detection, the framework defines that it is necessary to obtain the largest eigenvalue of each time frame, through eigen decomposition from a covariance of zero mean variables or covariance matrix of zero mean and unitary standard deviation variables, calculated from the evaluated traffic, as described in Section \ref{sec:2_prop_getv}.
+
 Through eigenvalue analysis of traffic with flood or port scan attacks, it is possible to visualize a significant difference between the largest eigenvalues and the remain eigenvalues, which can indicate a relationship between an outlier and time frames under attack.
 
+Figure \ref{fig:2.10} depicts the eigenvalues calculated from sample covariance matrix of the network traffic used to evaluate the synflood attack identification.
+In Figure \ref{fig:2.10}, the largest eigenvalue related to the simulated synflood attack ($q = 4$) stands out significantly from the other eigenvalues.
+
 \begin{figure}[h!]
 	\centering
-     \includegraphics[width=11cm]{figures/ch2/eigenvalues_synflood.eps} 
+     \includegraphics[width=9cm]{figures/ch2/eigenvalues_synflood.eps} 
      \caption{Eigenvalues of the sample covariance matrix (synflood).}
@@
 
-Figure \ref{fig:2.10} depicts the eigenvalues calculated from sample covariance matrix of the network traffic used to evaluate the synflood attack identification.
-In Figure \ref{fig:2.10}, the largest eigenvalue related to the simulated synflood attack ($q = 4$) stands out significantly from the other eigenvalues.
-
 Figure \ref{fig:2.11} illustrates the eigenvalues calculated from sample covariance matrix of the matrix used for fraggle attack detection.
@@
 	\centering
-     \includegraphics[width=11cm]{figures/ch2/eigenvalues_fraggle.eps}
+     \includegraphics[width=9cm]{figures/ch2/eigenvalues_fraggle.eps}
      \caption{Eigenvalues of the sample covariance matrix (fraggle).}
@@
 	\centering
-     \includegraphics[width=11cm]{figures/ch2/eigenvalues_portscan.eps}
+     \includegraphics[width=9cm]{figures/ch2/eigenvalues_portscan.eps}
      \caption{Eigenvalues of the covariance matrix of zero mean and unitary standard deviation (port scan).}
@@
   \centering
-  % \footnotesize
+  \footnotesize
   \caption{Largest Eigenvalue related to attacks detection}
@@
 
+Note that $\hat{d} = 1$, if there is one attack, while $\hat{d} > 1$ indicates more than one attack.
+An example of this could be seen for attack detection via EFT for traffic containing synflood and fraggle attacks, showing $\hat{d} = 2$, which indicates the presence of two attacks, as expected by the ground truth values $d$ of Table \ref{tab:2.04}.
+
 \begin{table}[h!]
   \centering
-  % \scriptsize
+  \footnotesize
   \caption{MOS schemes applied to port scan and flood detection}
@@
 			\hhline{~------~}
-		&\pmb{AIC} &\pmb{MDL} &\pmb{EDC} &\pmb{RADOI} &\pmb{EFT} &\pmb{SURE}\\
+		&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
 	\midrule
-	Detection of synflood \\(presence of attack) &2 &1 &\pmb{1} &5 &\pmb{1} &4 &\pmb{1} \\
-	Detection of synflood \\(absence of attack) &1 &1 &\pmb{0} &1 &\pmb{0} &3 &\pmb{0} \\
+	Detection of synflood \\(presence of attack) &2 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
+	Detection of synflood \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
 	\midrule
-	Detection of fraggle \\(presence of attack) &1 &1 &\pmb{1} &5 &\pmb{1} &4 &\pmb{1} \\
-	Detection of fraggle \\(absence of attack) &1 &1 &\pmb{0} &1 &\pmb{0} &3 &\pmb{0} \\
+	Detection of fraggle \\(presence of attack) &1 &1 &\textbf{1} &5 &\textbf{1} &4 &\textbf{1} \\
+	Detection of fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
 	\midrule
-	Detection of port scan \\(presence of attack) &1 &1 &\pmb{1} &1 &\pmb{1} &9 &\pmb{1} \\
-	Detection of port scan \\(absence of attack) &0 &0 &\pmb{0} &1 &\pmb{0} &1 &\pmb{0} \\
+	Detection of port scan \\(presence of attack) &1 &1 &\textbf{1} &1 &\textbf{1} &9 &\textbf{1} \\
+	Detection of port scan \\(absence of attack) &0 &0 &\textbf{0} &1 &\textbf{0} &1 &\textbf{0} \\
 	\midrule
-	Detection of synflood/fraggle \\(presence of attack) &2 &2 &\pmb{2} &5 &\pmb{2} &5 &\pmb{2} \\
-	Detection of synflood/fraggle \\(absence of attack) &1 &1 &\pmb{0} &1 &\pmb{0} &3 &\pmb{0} \\
+	Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{2} &5 &\textbf{2} &5 &\textbf{2} \\
+	Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{0} &1 &\textbf{0} &3 &\textbf{0} \\
     \bottomrule
@@
 
-Note that $\hat{d} = 1$, if there is attack, while $\hat{d} > 1$ indicates more than one attack.
-An example of this could be seen for attack detection via EFT for traffic containing synflood and fraggle attacks, showing $\hat{d} = 2$, which indicates the presence of two attacks, as expected by the ground truth values $d$ of Table \ref{tab:2.04}.
-
 In Table \ref{tab:2.04}, two MOS schemes outperforms from the others, which are EDC and EFT.
@@
 According to Table \ref{tab:2.04}, EDC and EFT estimated correctly the number of attacks of a time frame vector, indicating that occurred $\hat{d}$ network attacks, but not providing additional details, what highlights the necessity of complementary approaches in order to estimate the time and ports under attack.
-Hence, we propose apply eigen analysis to estimate the $q$-th time frames under attack and eigen similarity analysis to estimate the minutes and ports under attack.
+Hence, we propose apply eigen analysis to estimate the $q$-th time frames under attack and principal eigenvector similarity analysis to estimate the minutes and ports under attack.
 
@@
 
-\subsection{Eigen Similarity Analysis}
-\label{sec:2_EigenSimilarityAnalysis}
-
-This work proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\hat{\pmb{q}}}_{\rm max}$.
+\subsection{Principal Eigenvector Similarity Analysis}
+\label{sec:2_EigensimilarityAnalysis}
+
+In order to analyze the hypotheses $H_1^{(N)}$ and $H_1^{(A)}$, which evaluate if a subspace learned by eigenvalue decomposition can be used to detect and identify network attacks, we propose to apply principal eigenvector similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\hat{\pmb{q}}}_{\rm max}$.
 Hence, the proposed framework is applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.
@@
 
-Three approaches were evaluated for eigen similarity analysis: incremental, individual and incremental individualized approaches.
+Three approaches were evaluated for principal eigenvector similarity analysis: incremental, individual and incremental individualized approaches.
 For the incremental individualized approach, each minute is incrementally appended into the selected $\pmb{X}^{(q)}$ for obtaining $\pmb{v}_{(n)}^{(q)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack.
 Subsequently, $\pmb{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated.
-For the incremental approach, each $n$-th minute must be incrementally appended into $\pmb{X}^{(q)}$, for obtaining the next eigenvectors $\pmb{v}_{(n)}^{(q)}$ for individual time similarity analysis.
-For the individual approach, each $n$-th minute must be individually appended into $\pmb{X}^{(q)}$, without incremental append, but doing individual appended into $\pmb{X}^{(q)}$ for obtaining the next eigenvectors $\pmb{v}_{(n)}^{(q)}$ for individual similarity analyis.
-
-Table \ref{tab:2.05} presents the results of the evaluation of three approaches for similarity analysis of eigenvectors for port scan detection.
+For the incremental approach, each $n$-th minute must be incrementally appended into $\pmb{X}^{(q)}$, for obtaining the next principal eigenvector $\pmb{v}_{(n)}^{(q)}$ for individual time similarity analysis.
+For the individual approach, each $n$-th minute must be individually appended into $\pmb{X}^{(q)}$, without incremental append, but doing individual appended into $\pmb{X}^{(q)}$ for obtaining the next principal eigenvector $\pmb{v}_{(n)}^{(q)}$ for individual similarity analysis.
+
+Table \ref{tab:2.05} presents the results of the evaluation of three approaches for similarity analysis of principal eigenvectors for port scan detection.
+Table \ref{tab:2.05} shows the evaluation of the time frame $q=3$, when the port scan attack was simulated, considering the incremental individualized, incremental and individual approaches for principal eigenvector similarity analysis.
+According to the presented results, it is possible to observe the high similarity between network traffic without attack, which was larger than 0.9610 for all evaluated cases, and emphasize the expressive low similarity when it was evaluated the traffic with the simulated port scan attack ($n=15$), which was lower than 0.0276 for all evaluated approaches.
 
@@
   \footnotesize
-  \caption{Eigen Similarity Analysis for Port Scan Detection}
+  \caption{Principal eigenvector similarity analysis for port scan detection}
   \label{tab:2.05}
@@
 
-Table \ref{tab:2.05} shows the evaluation of the time frame $q=3$, when the port scan attack was simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
-According to the presented results, it is possible to observe the high similarity between network traffic without attack, which was larger than 0.9610 for all evaluated cases, and emphasize the expressive low similarity when it was evaluated the traffic with the simulated port scan attack ($n=15$), which was lower than 0.0276 for all evaluated approaches.
-
 Comparing the approaches for similarity analysis, it is possible to observe that all approaches highlight the low similarity when evaluated the traffic under attack.
@@
 
-Table \ref{tab:2.06} presents the results of the evaluation of the similarity analysis of eigenvectors for synflood detection.
-It shows the evaluation of the time frame $q=4$, when the synflood attack is simulated, considering the incremental individualized, incremental and individual approaches for eigen similarity analysis.
+Table \ref{tab:2.06} presents the results of the evaluation of the similarity analysis of principal eigenvectors for synflood detection.
+It shows the evaluation of the time frame $q=4$, when the synflood attack is simulated, considering the incremental individualized, incremental and individual approaches for principal eigenvector similarity analysis.
 According to the results, it is possible to observe the high similarity between network traffic without attack, which is larger than 0.9907 for all evaluated cases, and emphasize the expressive low similarity when evaluated the traffic with synflood attack (between $n=11$ and $n=20$), which is lower than 0.1244 for all evaluated approaches.
@@
   \footnotesize
-  \caption{Eigen Similarity Analysis for Synflood Detection}
+  \caption{Principal eigenvector similarity analysis for synflood detection}
   \label{tab:2.06}
@@
 
-Comparing the incremental individualized and the individual approaches for eigen similarity analysis, it is possible to observe that the incremental individualized approach obtain lowest values for almost all cases, except for the time $n=14$, where incremental individualized approach identified a larger similarity than the individual approach.
+Comparing the incremental individualized and the individual approaches for principal eigenvector similarity analysis, it is possible to observe that the incremental individualized approach obtain lowest values for almost all cases, except for the time $n=14$, where incremental individualized approach identified a larger similarity than the individual approach.
 The incremental individualized appends information about each evaluated traffic, therefore it incorporates traffic behaviors that can reduce the outlier capability detection, as occurred for the time $n=14$.
 
-Table \ref{tab:2.07} presents the results of the eigen similarity analysis evaluation for fraggle detection.
+Table \ref{tab:2.07} presents the results of the principal eigenvector similarity analysis evaluation for fraggle detection.
+For fraggle attack detection, the lack of similarity between legitimate and malicious traffic was more evident than for the evaluation of synflood and port scan detection.
+This behavior can be explained by the number of packets generated through the fraggle attack simulation, that was significantly larger than the number of packets generated during the synflood simulation.
+Considering the three approaches, the largest value for times under attack was 0.0083, while the shortest value for times without attacks was 0.9993.
 
@@
   \footnotesize
-  \caption{Eigen Similarity Analysis for Fraggle Detection}
+  \caption{Principal eigenvector similarity analysis for fraggle detection}
   \label{tab:2.07}
@@
 
-For fraggle attack detection, the lack of similarity between legitimate and malicious traffic was more evident than for the evaluation of synflood and port scan detection.
-This behavior can be explained by the number of packets generated through the fraggle attack simulation, that was significantly larger than the number of packets generated during the synflood simulation.
-Considering the three approaches, the largest value for times under attack was 0.0083, while the shortest value for times without attacks was 0.9993.
-
 Therefore, considering the evaluation for port scan, synflood and fraggle detection, the incremental approach can produce false positive results, while the individual and incremental individualized approaches produce quite similar results, even though the individual approach be more simple and require less memory and processing time.
 
-These results highlight the capability of change detection based on similarity between legitimate and malicious traffic from flood or port scan attacks, by means of a subspace learned via Eigenvalue Value Decomposition and Model Order Selection,
-endorsing the effectiveness and safety for adoption of threshold for attack detection through eigen similarity analysis.
-Moreover, these results answer positively the question $Q_1$, what rejects the null hypothesis  $H_{0sub}$ and confirms the alternative
-hypothesis $H_{1sub}$.
+These results highlight the capability of change detection based on similarity between legitimate and malicious traffic from flood or port scan attacks, by means of a subspace learned via EVD and MOS,
+endorsing the effectiveness and safety for adoption of threshold for attack detection through principal eigenvector similarity analysis.
+
+Moreover, these results answer positively the question $Q_1$, which ask "Can the analysis of patterns from a learned subspace identify and detect anomalies in network traffic?".
+The results rejects the null hypothesis $H_1^{(N)}$ and confirms the alternative
+hypothesis $H_1^{(A)}$, which argues that "A subspace learned by eigenvalue decomposition can be used to detect and identify network attacks".
 
@@
 
-Given $\hat{N}$, which is the set of estimated $n$-th minutes under attack, it is possible to apply cosine similarity analysis to identify variation of the most significant eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port, during a $n$-th minute.
-Therefore, the incremental individualized and individual approaches of eigen similarity analysis were evaluated, for detection of ports under flood and port scan attacks, according to results presented in following tables.
-For this evaluation, the last most significant eigenvectors without attack $v$ was used as reference for similarity analysis against each target port $m$-th.
-
-Table \ref{tab:2.08} presents the results of the evaluation of eigen similarity analysis for detection of ports under port scan attack, showing only the time frame $q=3$ and minute $n=15$, due to the simulated port scan attack occurred only at this time, although the remain time frame has been completely evaluated and presented high similarity to the reference of traffic without network attack.
+Given $\hat{N}$, which is the set of estimated $n$-th minutes under attack, it is possible to apply cosine similarity analysis to identify variation between the principal eigenvectors, caused by the insertion of anomalous network traffic by a selected $m$-th port, during a $n$-th minute.
+Therefore, the incremental individualized and individual approaches of principal eigenvector similarity analysis were evaluated, for detection of ports under flood and port scan attacks, according to results presented in following tables.
+For this evaluation, the last principal eigenvector without attack $\pmb{v}$ was used as reference for similarity analysis against each target port $m$-th.
+
+Table \ref{tab:2.08} presents the results of the evaluation of principal eigenvector similarity analysis for detection of ports under port scan attack, showing only the time frame $q=3$ and minute $n=15$, due to the simulated port scan attack occurred only at this time, although the remain time frame has been completely evaluated and presented high similarity to the reference of traffic without network attack.
 
@@
   \centering
-  % \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n=15)}
+  \footnotesize
+  \caption{Principal eigenvector similarity analysis for detection of ports under port scan attack (q=3 and n=15)}
   \label{tab:2.08}
@@
 For the evaluation of the proposed approaches for identification of ports under synflood and fraggle attack, all minutes of time frames under were analyzed.
-However, due to space limitations, only the results of the first minute, where a low similarity was identified, will be shown.
+However, due to space limitations, only the results of the first minute, where a low similarity was identified, are shown.
 Nevertheless, the results obtained for the evaluation of traffic without attack presented high similarity to the reference traffic, with similarities close to 0.9999, and the evaluation of the other minutes under attack presented results quite similar to the results shown in the Tables \ref{tab:2.09} and \ref{tab:2.10}.
 
-Table \ref{tab:2.09} presents the results of the evaluation of eigen similarity analysis for detection of ports under synflood attack, showing only the time frame $q=4$ and minute $n=11$.
+Table \ref{tab:2.09} presents the results of the evaluation of principal eigenvector similarity analysis for detection of ports under synflood attack, showing only the time frame $q=4$ and minute $n=11$.
 
@@
   \centering
-  % \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
+  \footnotesize
+  \caption{Principal eigenvector similarity analysis for detection of ports under synflood attack (q=4 and n=11)}
   \label{tab:2.09}
@@
 	53 &1.0000 &1.0000 &no \\
-	21 &1.0000 &1.0000 &nos \\
+	21 &1.0000 &1.0000 &no \\
 	22 &1.0000 &1.0000 &no \\
@@
 
-According to results presented in Table \ref{tab:2.09}, both approaches identifies low similarity for the traffic of port 600, which is the target port of the simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through eigen similarity analysis assisted by threshold definition.
-
-Table \ref{tab:2.10} presents the results of the evaluation of eigen similarity analysis for detection of ports under fraggle attack, showing only the time frame $q=5$ and minute $n=11$.
+According to results presented in Table \ref{tab:2.09}, both approaches identifies low similarity for the traffic of port 600, which is the target port of the simulated synflood attack, but the incremental individualized approach identifies the lowest similarity and presents better sensibility to identification of synflood attack through principal eigenvector similarity analysis assisted by threshold definition.
+
+Table \ref{tab:2.10} presents the results of the evaluation of principal eigenvector similarity analysis for detection of ports under fraggle attack, showing only the time frame $q=5$ and minute $n=11$.
 
@@
   \centering
-  % \footnotesize
-  \caption{Eigen Similarity Analysis for Detection of Ports Under Fraggle Attack (q=5 and t=11)}
+  \footnotesize
+  \caption{Principal eigenvector similarity analysis for detection of ports under fraggle attack (q=5 and t=11)}
   \label{tab:2.10}
@@
 
-The incremental individualized approach was able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches presented false positives or low sensibility to eigen similarity analyis for network attack detection.
+The incremental individualized approach was able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches presented false positives or low sensibility to principal eigenvector similarity analysis for network attack detection.
 This approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.
 
-According to the shown significant lack of similarity between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
+According to the shown significant lack of similarity between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through principal eigenvector similarity analysis.
 
@@
 
-This subsection presents a summarized view of results obtained from the application of the Eigensim, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 data set.
-Since the proposed framework is detailed in Section \ref{sec:2_prop_getv} and in Subsections \ref{sec:2_AnalyzedScenario}, \ref{sec:2_largesteigenvaluesanalysis}, \ref{sec:2_MOSSchemesEvaluation}, \ref{sec:2_EigenvalueAnalysis} and \ref{sec:2_EigenSimilarityAnalysis}, here the focus is on the parameter selection, data set evaluation and results for flood and probe attack identification.
+This subsection presents a summarized view of results obtained from the application of the Eigensimilarity, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 data set.
+Since the proposed framework is detailed in Section \ref{sec:2_prop_getv} and in Subsections \ref{sec:2_AnalyzedScenario}, \ref{sec:2_largesteigenvaluesanalysis}, \ref{sec:2_MOSSchemesEvaluation}, \ref{sec:2_EigenvalueAnalysis} and \ref{sec:2_EigensimilarityAnalysis}, here the focus is on the parameter selection, data set evaluation and results for flood and probe attack identification.
 
@@
 
-\begin{table*}[!t]
+Table \ref{tab:2.12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}.
+
+The analysis based on sample covariance of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames.
+The results also show 50.00 \% of misclassification rate, which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
+The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA data set presents high number of packets per time from one source to one target, with no variation on IP source or target port.
+This observation corroborates with previous evaluations of the DARPA data set that highlight issues regarding traffic redundancy.
+
+\begin{table*}[!h]
 	\caption{Results of the attack detection evaluation}
 	\label{tab:2.12}
+	\footnotesize
 	\centering
@@
 		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Solution\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Attack Type\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Metric\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Result\end{tabular} \\ \hline
-	Eigensim	&Flooding	&True Positive	&100.00 \%\\ \hline
-	Eigensim	&Flooding	&False Positive	&60.00 \%\\ \hline
-	Eigensim	&Flooding	&Misclassification	&50.00 \%\\ \hline
-	Eigensim	&Probe	&True Positive	&76.92 \%\\ \hline
-	Eigensim	&Probe	&False Positive	&18.52 \%\\ \hline
-	Eigensim	&Probe	&Misclassification	&32.73 \%\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&True Positive	&82.00 \%\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&False Positive	&-\\ \hline
-	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&Misclassification	&-\\ \hline
+	Eigensimilarity	&Flooding	&True Positive	&100.00 \%\\ \hline
+	Eigensimilarity	&Flooding	&False Positive	&60.00 \%\\ \hline
+	Eigensimilarity	&Flooding	&Misclassification	&50.00 \%\\ \hline
+	Eigensimilarity	&Probe	&True Positive	&76.92 \%\\ \hline
+	Eigensimilarity	&Probe	&False Positive	&18.52 \%\\ \hline
+	Eigensimilarity	&Probe	&Misclassification	&32.73 \%\\ \hline
+	Callegari \emph{et al.} \cite{callegari2011novel}	&Flooding	&True Positive	&82.00 \%\\ \hline
+	Callegari \emph{et al.} \cite{callegari2011novel}	&Flooding	&False Positive	&-\\ \hline
+	Callegari \emph{et al.} \cite{callegari2011novel}	&Flooding	&Misclassification	&-\\ \hline
 	Lu and Ghorbani \cite{Lu2009}	&Overall	&True Positive	&94.67 \%\\ \hline
@@
 
-Table \ref{tab:2.12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}.
-The analysis based on sample covariance of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames.
-The results also show 50.00 \% of misclassification rate, which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed.
-The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA data set presents high number of packets per time from one source to one target, with no variation on IP source or target port.
-This observation corroborates with previous evaluations of the DARPA data set that highlight issues regarding traffic redundancy.
-
 The analysis based on covariance of zero mean and unitary standard deviation variables was evaluated for port scan attacks, including probe attacks and DoS attacks that send few packets for several ports in order to exploit some vulnerability.
@@
 
-The performance of detection rate of flooding attacks is compared with the method proposed by Callegari \emph{et al} \cite{Zonglin2009}.
+The performance of detection rate of flooding attacks is compared with the method proposed by Callegari \emph{et al.} \cite{callegari2011novel}.
 This work is a statistical method, based on PCA, without training or learning methods, even though it relies on visual analysis for principal components selection.
-The best detection rate of \cite{Zonglin2009} was 82.00 \% for detection of synthetically added flood attacks, while this current proposal obtains 100.00 \% of detection rate for detection of flood attacks of DARPA data set.
-It is important to note that Callegari \emph{et al} \cite{Zonglin2009} did not publish results of false positive and misclassification.
+The best detection rate of \cite{callegari2011novel} was 82.00 \% for detection of synthetically added flood attacks, while this current proposal obtains 100.00 \% of detection rate for detection of flood attacks of DARPA data set.
+It is important to note that Callegari \emph{et al.} \cite{callegari2011novel} did not publish results of false positive and misclassification.
 
@@
 
-This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:2.08} and equations presented in Section \ref{sec:2_prop_getv}.
+This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and principal eigenvector similarity analysis, according to Figure \ref{fig:2.08} and equations presented in Section \ref{sec:2_prop_getv}.
 
@@
 \label{sec:2_ComplexityAnalysis}
-The EVD, calculated according to (\ref{eq:2.06}), requires the previous calculation of covariance matrix, according to Equations (\ref{eq:2.02}), (\ref{eq:2.03}), (\ref{eq:2.04}) and (\ref{eq:2.05}).
+The EVD, calculated according to (\ref{eq:2.06}), requires the previous calculation of covariance matrix, according to (\ref{eq:2.02}), (\ref{eq:2.03}), (\ref{eq:2.04}) and (\ref{eq:2.05}).
 The covariance matrix calculation is $O(M^2N)$ and the EVD is $O(N^3)$, where $M$ denotes the number of network ports and $N$ denotes the period time.
@@
 The largest eigenvalue analysis is $O(\hat{d}Q)$, where $\hat{d}$ denotes the number of time frame under attack, according to Algorithm \ref{alg:2.01}.
-Subsequently, the eigen similarity analysis relies on EVD and cosine similarity analysis, which is $O(N^2)$, for $\hat{d}$ time frames, therefore the eigen similarity analysis is which is $O(\hat{d}(M^2N + N^3 + N^2))$ and yields an $O(N^3)$ upper bound on the worst-case running time for eigen similarity analysis.
+Subsequently, the principal eigenvector similarity analysis relies on EVD and cosine similarity analysis, which is $O(N^2)$, for $\hat{d}$ time frames, therefore the principal eigenvector similarity analysis is which is $O(\hat{d}(M^2N + N^3 + N^2))$ and yields an $O(N^3)$ upper bound on the worst-case running time for principal eigenvector similarity analysis.
 
@@
 The performance evaluation focus on the main steps, which are discussed in subsection \ref{sec:2_ComplexityAnalysis} regarding the complexity analysis.
-The data modeling is also a time consuming step, however its processing can be optimized through distributed processing techniques, such as MapReduce, achieving high throughput for packet counting or even for deep packet inspection \cite{Vieira2013}.
+The data modeling is also a time consuming step, however its processing can be optimized through distributed processing techniques, such as MapReduce, achieving high throughput for packet counting or even for deep packet inspection \cite{vieira2012measuring, vieira2012evaluating, vieira2013evaluating, vieira2013approach}.
 It is also possible to evaluate the adoption of faster SVD algorithms, considering implementations based on truncated or randomized approaches \cite{halko2011finding} that aim reduce the complexity and processing time.
 
-The experiments were performed in a desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM, considering: variations on the network traffic time; the frame size denoted as $N$; the number of network ports denoted as $M$; the mean processing time for eigen analysis based on sample covariance of zero mean, denoted as 1-time; the mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation, denoted as 2-time; and the mean processing time for EDC MOS scheme, denoted as 3-time.
+The experiments were performed in a desktop computer with an Intel Core i7-4510U 2.00GHz and 16 GB of RAM, considering: variations on the network traffic time; the frame size denoted as $N$; the number of network ports denoted as $M$; the mean processing time for eigen analysis based on sample covariance of zero mean, denoted as 1-time; the mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation, denoted as 2-time; and the mean processing time for EDC MOS scheme, denoted as 3-time.
 The mean time calculations was obtained from 200 measurement repetitions, in order to obtain reliable values.
 
-\begin{table*}[!t]
+Table \ref{tab:11} presents the measured results.
+The experiment considered traffic time of 16, 20 and 22 hours, according to the selected traffic time per day available by the DARPA data set.
+Note that the processing time increases according to the increment in traffic time, around 2 or 3 times for 1-time and 2-time, but the worst measured processing time is 4.7250 milliseconds.
+
+\begin{table*}[!h]
 	\caption{Processing time of the main steps for anomaly detection}
 	\label{tab:11}
+	\footnotesize
 	\centering
@@
 
-Table \ref{tab:11} presents the measured results.
-The experiment considered traffic time of 16, 20 and 22 hours, according to the selected traffic time per day available by the DARPA data set.
-Note that the processing time increases according to the increment in traffic time, around 2 or 3 times for 1-time and 2-time, but the worst measured processing time is 4.7250 milliseconds.
-
 According to Table \ref{tab:11}, the processing time increases with the frame size $N$ decreasing, therefore it is possible to evaluate the frame size that produces better identification rates and acceptable processing time.
@@
 
-This work models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, named Eigensim, which is based on subspace decomposition, eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
-
-The Eigensim is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis from a subspace obtained by EVD.
-Therefore, the observed results rejects the null hypothesis  $H_{0sub}$ and confirms the alternative
-hypothesis $H_{1sub}$.
-
-The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
-
-This work evaluated the effectiveness of MOS schemes for fraggle attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
-Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.
-
-The incremental individualized approach of eigen similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to eigen similarity analysis for network attack detection.
+This work models the network traffic as a signal processing formulation for applying to the framework for detection and identification of network attacks, named Eigensimilarity, which is based on subspace decomposition, eigenvalue analysis, model order selection (MOS) and principal eigenvector similarity analysis.
+
+The Eigensimilarity is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, without training, applying signal processing concepts for traffic modeling and through approaches based on MOS and principal eigenvector similarity analysis from a subspace obtained by EVD.
+Therefore, the observed results rejects the null hypothesis $H_1^{(N)}$ and confirms the alternative
+hypothesis $H_1^{(a)}$.
+
+The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an principal eigenvector similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
+
+The incremental individualized approach of principal eigenvector similarity analysis, is able to detect low similarity for all evaluated scenarios and types of network attack, while the other approaches present false positives or low sensibility to principal eigenvector similarity analysis for network attack detection.
 Therefore, the incremental individualized approach is able to gradually and incrementally adapt to network traffic changing, preserving the sensibility to identify outliers or anomalies by time or network port, and reducing the occurrence of false positives.
 
-According to the significant similarity difference between legitimate and malicious traffic, it is possible to adopt safe thresholds for flood and port scan detection through eigen similarity analysis.
+The principal eigenvector similarity analysis is applied for each dimension individually, for the time dimension initially and after for port dimension, in order do identify lack of similarity.
+Future work can be directed to pairwise similarity analysis, where the two dimensions are not evaluated separately, or for a principal component analysis with indication of element-wise anomalies, such as the Robust Principal Component Analysis (RPCA).
+
+Considering the reported limitations for the DARPA and KDD data sets, and the lack of available labeled data sets for network attack detection, future work can be directed to evaluate our proposed framework for novel and up-to-date data sets of network attack detection.
 
@@
 Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high throughput network traffic, as well as it is possible to evaluate approaches with less complexity for EVD \cite{halko2011finding}.
-Additionally, future research can evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.
+Future research can also evaluate the application of the proposed approach to different attack types and domains, considering cases that are aware to behavioral analysis.
+
+We adopted a data modeling based on two dimensions, which are the time and the network ports.
+Future works can evaluate improvements given by a multidimensional modeling and adoption of tensor-based approaches \cite{da2008robust, gomes2019tensor}, in order to evaluate complex patterns that can be revealed by multidimensional analysis and tensor-based decomposition.
+It is possible to consider the addition of a seasonal dimension for our proposed data modeling, in order to evaluate the the relationship between the selected season, such as week or month, to the analysis of amount of packets per network port and time.

### 2020-04-28 docs/thesis/chapters/4_m_rpca/m_rpca.tex
updating thesis latex

@@
-\chapter{Moment Distances from Robust Subspace for Network Anomaly Detection}
+\chapter{Robust Framework for Detection of Network Attacks in Imbalanced Traffic}
 \label{ch:4_m_rpca}
 
-\begin{quotation}[]{Paulo Freire}
-No one knows it all. No one is ignorant of everything. We all know something. We are all ignorant of something.
-\end{quotation}
+According to \cite{distilnetworks2019}, 37.9\% of all Internet traffic of 2018 was not from human activities, but from bots, that can be classified as good bots, that perform legitimate operations, and bad bots, which perform malicious activities, such as DDoS attacks, probe attack or frauds.
+In 2018, bad bots accounted for 20.4\% of all website traffic.
 
-Some widely adopted algorithms for anomaly detection assume a Gaussian or symmetric distributed data, however this assumption may not be observed in some real world problems, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{lakhina2005mining,benson2010network, leon2017probability}.
+According to \cite{BISSELL2019} the proportion of cost on discovery activities for cybersecurity has increased steadily since 2015.
+Forensic cyber and user behavior analytics also present an opportunity for cost savings — US\$1.72 million — with discovery and investigation activities.
+However, only 32\% of organizations have deployed these technologies enterprise-wide.
+Accenture consulting argues that 89\% of survey respondents believe breakthrough technologies, like machine learning and user behavior analytics, are essential to securing the future of their organizations \cite{Accenture2018}.
+
+To face the adversarial model, network attacks and counter measures of attackers to avoid detection, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, by means of behavioral analysis, where it is not necessary known anomalies for training models \cite{moustafa2019holistic}.
+
+Anomalies in the context of network traffic can be hard to identify and separate from legitimate data due to the rare occurrences of anomalies in comparison to legitimate events.
+Therefore, anomaly detection algorithms have to be highly discriminating, robust to corruption and able to deal with the imbalanced data problem \cite{he2008learning}.
+Note that data corruption refers to outliers that can be part of the data but not be considered malicious.
+The imbalanced data problem corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes \cite{he2009learning}, which can be legitimate or malicious classes in network anomaly detection problems.
+
+Traditionally adopted algorithms for anomaly detection assume a Gaussian or symmetric distributed data \cite{lakhina2005mining}, however this assumption may not be observed in some real world problems \cite{benson2010network}, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{leon2017probability}.
 
@@
 
-These findings show that the skewness and heavy-tailed distributions may be important for network traffic analysis, and can motivate researches to evaluate the impact of skewed data into algorithms that rely on Gaussian distributed data.
-These findings can also highlight opportunities to  exploit the skewness and heavy-tailed distributions to obtain accurate classifiers for network anomaly detection.
+The findings presented by Benson \emph{et al.} \cite{benson2010network} and Leon-Garcia \cite{leon2017probability} show that the skewness and heavy-tailed distributions may be important for network traffic analysis, and can motivate researches to evaluate the impact of skewed data into algorithms that rely on Gaussian distributed data for network anomaly detection.
+Additionally, the fitting of network traffic to skewed and heavy-tailed distributions can indicate opportunities to exploit properties and characteristics of the skewness and heavy-tailed distributions to obtain improved classifiers for network anomaly detection.
 
-An outlyingness-approach based on a robust estimator of skewness, combined with robust estimators of location and scale, can be able to flag the outlying measurements.
-When the same methodology would be used with non-robust estimators of location, scale and skewness, the outlyingness-values would be affected by the outliers such that the outlying group could be masked \cite{hubert2009robustskewed}.
+Network anomaly detection problems are usually characterized by imbalanced data \cite{Phua2004minority}.
+However, learning algorithms for imbalanced data has been a challenging research topic, considering that the imbalanced data can compromise the performance of most standard learning algorithms, creating bias or unfair weight to learn from the majority class and reducing detection capacity of anomalies that are characterized by the minority class \cite{he2008learning}.
+Therefore, learning methods for imbalanced and skewed data have attracted attention of researchers \cite{hubert2009robustskewed}.
 
-Network anomaly detection problems are usually characterized by imbalanced data \cite{Phua2004minority,he2008learning,benson2010network}.
-However, learning algorithms for imbalanced data has been a challenging research topic, considering that the fundamental issue with the imbalanced learning problem is the ability of imbalanced data to significantly compromise the performance of most standard learning algorithms \cite{he2008learning}.
-Therefore, learning methods for imbalanced and skewed data have been attracted attention of researchers \cite{Phua2004minority,hubert2009robustskewed}.
+Robust subspace learning has been attracting a growing attention of researchers aiming the development of network attack detection systems \cite{pascoal2012robust, zhou2017anomaly} that rely on behavioral analysis.
+An outlyingness-approach based on a robust estimator of skewness, combined with robust estimators of location and scale, can be able to flag the outlying measurements \cite{hubert2009robustskewed}.
+According to Hubert \emph{et al.} \cite{hubert2009robustskewed}, when the same methodology would be used with non-robust estimators of location, scale and skewness, the outlyingness-values would be affected by the outliers such that the outlying group could be masked.
 
-We believe that the skewness of anomalous and normal traffic can highlight features for improving anomaly detection in imbalanced data.
-We also believe that the distance between robust estimates of normal traffic and contaminated observations can highlight discrepancies and be used for network attack detection.
+Considering that the skewness of anomalous and legitimate traffic can highlight features for improving anomaly detection and network attack detection in imbalanced data, and considering that the distance between robust estimates can be used for network attack detection, we propose the Moment-based Robust Principal Component Analysis (m-RPCA), which is an approach based on distances between moments computed from a robust subspace learned by Robust Principal Component Analysis (RPCA) from contaminated observations, in order to detect anomalies from skewed data and network traffic.
 
-Therefore, we propose the m-RPCA, which is an approach based on distances between moments computed from a robust subspace learned by Robust Principal Component Analysis (RPCA) and contaminated observations, in order to detect anomalies from skewed data and network traffic.
-The proposed approach relies on a robust subspace computed from supposed normal observations, for estimating the moments to be used for distance analysis.
-The anomaly detection from contaminated observations evaluate the Mahalanobis distance between the robust moments and new contaminated observations, in a semi-supervised fashion, without the computational cost of new robust subspace learning for new observations.
+The proposed approach relies on a robust subspace computed from supposed legitimate observations, for estimating the moments to be used for distance analysis.
+The anomaly detection from contaminated observations evaluate the Mahalanobis distance between the robust moments and new contaminated observations, in a semi-supervised fashion, without the computational cost of new robust subspace learning for anomaly detection from new observations.
 The m-RPCA can also be computed as an unsupervised algorithm, with subspace learning from the same contaminated data that is the target of the anomaly detection analysis.
 
-We evaluate the accuracy of the m-RPCA for anomaly detection on simulated data sets, with skewed and heavy-tailed distributions, and for the CTU-13 data set \cite{garcia2014empirical}, which is a large data set of normal, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
-The Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection based on clustering and statistical approaches.
-We also compare the results to ROBPCA \cite{hubert2005robpca}, which is a method that also relies on robust estimates with adjusted outlyingness based on robust skewness.
+We evaluate the accuracy of the m-RPCA for anomaly detection on simulated data sets, with skewed and heavy-tailed distributions, and for network attack detection on CTU-13 data set \cite{garcia2014empirical}, which is a large data set of legitimate, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
+The Experimental evaluation compares our proposal to standard and widely adopted algorithms for anomaly detection, which are based on clustering and statistical approaches, and to ROBPCA \cite{hubert2005robpca}, which is a anomaly detection method that relies on robust estimates with adjusted outlyingness based on robust skewness.
 
-The main contribution of this work is the proposal of a novel semi-supervised and unsupervised method for anomaly detection in skewed and imbalanced data, with results of experimental evaluation on simulated and real data sets.
+The main contribution of this work is the proposal of a novel semi-supervised and unsupervised method for anomaly detection in skewed and imbalanced data, with results showing improvements through an experimental evaluation on simulated skewed and heavy-tailed distributions and on real data set with legitimate network traffic and botnet attacks.
 
-This Chapter is organized as follows.
-In Section \ref{sec:4_relatedworks}, it is conducted a literature review of network anomaly detection, botnet detection, and imbalanced learning.
-Section \ref{sec:4_datamodel} presents the data model and the evaluated data set.
-Section \ref{sec:4_proposal} describes the proposed approach for network attack detection.
-Section \ref{sec:4_experiments} discusses the experimental validation, Section \ref{sec:4_simulated_result} presents the results of the simulated scenarios and the Section \ref{sec:4_ctu13_result} describes the results for anomaly detection on CTU-13 data set.
-Finally, Section \ref{sec:4_conclusion} draws the conclusions and the suggestions for future work.
+This chapter is organized as follows.
+In Section \ref{sec:3_relatedworks}, a literature review about network anomaly detection, botnet detection, and imbalanced learning is conducted.
+We present in Section \ref{sec:3_datamodel} the data model and the evaluated data set.
+In Section \ref{sec:3_proposal} it is described the proposed approach for network attack detection.
+We discuss in Section \ref{sec:3_experiments} the experimental validation, and in Section \ref{sec:3_simulated_result} are presented the results of the simulated scenarios and in Section \ref{sec:3_ctu13_result} we describe the results for anomaly detection on CTU-13 data set.
+Finally, in Section \ref{sec:3_conclusion} we draw the conclusions and the suggestions for future work.
 
 \section{Related Works}
-\label{sec:4_relatedworks}
+\label{sec:3_relatedworks}
 
@@
 
+According to \cite{distilnetworks2019}, 37.9\% of all Internet traffic of 2018 was not from human activities, but from bots, that can be classified as good bots, that perform legitimate operations, and bad bots, which are responsible by malicious activities, such as DDoS, probe attack or frauds.
+
 Acarali \emph{et al.} \cite{acarali2016survey} surveys network-based detection approaches for HTTP-based botnets, and discuss the traffic-based features used to detect bot traffic and presents an abstraction of the main types of features related to protocols and OSI layers.
@@
 Khattak \emph{et al.} \cite{khattak2015botflex} proposed BotFlex, which is a network-based tool for botnet detection, composed by a Complex Event Processing (CEP) engine and on a correlation framework that continuously receives events and correlates them according to rules.
-BotFlex's results are compared to BotHunter \cite{gu2007bothunter}, but the evaluation relies in a own and not public data set.
+BotFlex's results are compared to BotHunter \cite{gu2007bothunter}, but the evaluation relies in an own and not public data set.
 
@@
 
-Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by attack, normal and background labeled data, in an imbalanced distribution like in a real networks.
+Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by attack, legitimate and background labeled data, in an imbalanced distribution like in a real networks.
 The authors recommend scenarios for training and testing in order to avoid the use of traffic from a botnet family for training and testing, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
@@
 However, to overcome the above mentioned limitations, some approaches based on PCA have been proposed for network anomaly detection.
-Callegari \emph{et al.} \cite{callegari2011novel} proposed a PCA-based method for identifying the network traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of normal and anomalies according to principal components (normal) and remaining (anomalies).
+
+Callegari \emph{et al.} \cite{callegari2011novel} proposed a PCA-based method for identifying the network traffic flows responsible for an anomaly detected at the aggregate level, by means of a separation of legitimate and anomaly observations according to principal components (legitimate) and remaining (anomalies).
 Lee \emph{et al.} \cite{Lee2013} presented OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
-Vieira \emph{et al.} \cite{vieira2017model} proposed a framework that applies Model Order Selection (MOS) for detection of time frames under attack and uses similarity analysis to extract details and detect the time and ports under attack.
+Vieira \emph{et al.} \citeme{vieira2017model} proposed a framework that applies Model Order Selection (MOS) for detection of time frames under attack and uses similarity analysis to extract details and detect the time and ports under attack.
 
@@
 RPCA aims to be resilient to outliers by means of a robust subspace learning \cite{vaswani2018robust} for outlier corrupted data, decomposing a given data matrix $\pmb{X}$ into the sum of a low rank matrix $\pmb{L}$, whose column subspace gives the principal components, and a sparse matrix $\pmb{S}$, which refers to outliers’ matrix.
-This definition is also referred to as the sparse + low rank ($\pmb{S + LR}$) formulation \cite{vaswani2018robust}.
 We refer to \cite{lerman2018overview} and \cite{vaswani2018robust} for more details regarding robust subspace learning.
 
-RPCA is been mainly applied to computer vision, in problems of robust subspace tracking and robust subspace recovery.
+RPCA has been mainly applied to computer vision, in problems of robust subspace tracking and robust subspace recovery.
 However, RPCA has also been adopted for general outlier detection \cite{hubert2005robpca,hubert2009robustskewed,cherapanamjeri2017thresholding,zhou2017anomaly,NetflixSurus} and for anomaly detection on network traffic \cite{pascoal2012robust}.
@@
 
-Robust subspace learning has receiving a growing attention of researchers aiming the development of network anomaly detection systems \cite{rousseeuw1984mcd, rousseeuw1999fastmcd, hubert2005robpca,hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
+Robust subspace learning has received a growing attention of researchers aiming the development of network anomaly detection systems \cite{ hubert2009robustskewed, pascoal2012robust, zhou2017anomaly}, considering outlier-robust methods and sparse-corruption methods \cite{lerman2018overview}.
 Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on a robust mutual information estimator for feature selection and based on RPCA for outlier detection in internet traffic.
 The anomaly detection proposed by Pascoal \emph{et al.} is an unsupervised approach that estimate the first $k$ robust principal components, calculate the score and the orthogonal distances, calculate the thresholds and classify new observations accordingly.
+
 Zhou and Paffenroth \cite{zhou2017anomaly} proposed the Robust Deep Autoencoders (RDA), which the central idea is that a RDA inherits the non-linear representation capabilities of autoencoders combined with the anomaly detection capabilities of RPCA.
@@
 Mahalanobis in 1936 \cite{mahalanobis1936md}.
-It is a multi-dimensional generalization for measuring how many standard deviations away $\pmb{x}$ is from the mean $\bar{\pmb{x}}$ and covariance $\bar{\pmb{\Sigma}}$ of $\pmb{X}$.
+MD is a multi-dimensional generalization for measuring how many standard deviations away $\pmb{x}$ is from the mean $\pmb{\mu}$ and covariance $\hat{\pmb{\Sigma}}$ of $\pmb{X}$.
 MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and contaminated observations can reveal anomalies.
@@
 We propose an approach based on the distances between moments computed from learned robust subspace and contaminated observations, for anomaly detection on skewed and imbalanced data sets, and evaluate our approach for network attack detection.
-The proposed approach relies on learning robust subspace from supposed normal traffic for estimating the moments (mean, skewness and kurtosis).
-Thus, the anomaly detection for contaminated observations should evaluate the Mahalanobis Distance between the robust moments and contaminated observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations, or in a unsupervised approach, where the robust subspace is learned from the contaminated data and the distance analysis is between the robust moments and the contaminated observations.
+The proposed approach relies on learning robust subspace from supposed legitimate traffic for estimating the moments (mean, skewness and kurtosis).
+
+Thus, the anomaly detection for contaminated observations should evaluate the distance between the robust moments and contaminated observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations, or in an unsupervised approach, where the robust subspace is learned from the contaminated data and the distance analysis is between the robust moments and the contaminated observations.
 
 \section{Data Model}
-\label{sec:4_datamodel}
+\label{sec:3_datamodel}
 
-In this paper, scalars are denoted by italic letters (\emph{a, b, A, B, $\alpha$, $\beta$}), vectors by lowercase bold letters ($\pmb{a}, \pmb{b}$), matrices by uppercase bold letters ($\pmb{A}$, $\pmb{B}$), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix $\pmb{A}$.
-The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
-The Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries.
-We also define the operator $\langle \mathord{\cdot} \rangle$, that is the standard trace inner product, and $[ \mathord{\cdot} ]^s$, which denotes the $n$ largest values of a vector.
+In this chapter, we adopt the data model introduced in Section \ref{sec:2_datamodel} and define that the Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries.
+We also define the operator $\langle \mathord{\cdot} \rangle$, that is the standard trace inner product, and the operator $[ \mathord{\cdot} ]^c$, which denotes the indexes of the $c$ largest values of a vector.
 
-This section also presents a description of the simulated data model in Subsection \ref{sec:4_simulation} and in Subsection \ref{sec:4_CTU-13} we describe the data model for the CTU-13 data set.
+This section also presents a description of the simulated data model in Subsection \ref{sec:3_simulation} and in Subsection \ref{sec:3_CTU-13} we describe the data model for the CTU-13 data set.
 
 \subsection{The Simulated data set}
-\label{sec:4_simulation}
+\label{sec:3_simulation}
 
-In order to analyse the null hypotheses $H_{1rob.sub}$ and $H_{0rob.sub}$, which evaluate if the robust subspace learning can improves the anomaly detection from imbalanced and skewed data, we create two simulated data sets characterized by skewed and heavy tailed distributions, and create a simulated Gaussian distributed data set, in order to also analyse the detection rate for not skewed and not heavy tailed distributions.
+In order to analyze the hypotheses $H_2^{(N)}$ and $H_2^{(A)}$, which evaluate if the robust subspace learning can improves the anomaly detection from imbalanced and skewed data, we create two simulated data sets characterized by skewed and heavy tailed distributions, and create a simulated Gaussian distributed data set, in order to also analyze the detection rate for not skewed and not heavy tailed distributions.
 
-We selected Pareto (with scale $a=3$, mode $m=1$ and $\mu=1$) and Lognormal (with mean $\mu=0$ and standard deviation $\sigma=1$) distributions, denoted respectively as $\pmb{Y}_p$, while $\pmb{Y}_l$, to simulate normal events and to evaluate the anomaly detection on skewed data performed by our proposals in comparison to widely adopted algorithms for outlier detection, considering that these distributions are skewed and heavy-tailed, and have been adopted to characterize network traffic of the Internet and data centers \cite{benson2010network,leon2017probability}.
-We also adopt the Gaussian distributed data set $\pmb{Y}_g$ to simulate normal events with mean $\mu=0$ and unitary standard deviation $\sigma=1$.
+We selected Pareto (with scale $a=3$, mode $m=1$ and $\mu=1$) and Lognormal (with mean $\mu=0$ and standard deviation $\sigma=1$) distributions, denoted respectively as $\pmb{Y}_p$, while $\pmb{Y}_l$, to simulate legitimate events and to evaluate the anomaly detection on skewed data performed by our proposals in comparison to widely adopted algorithms for outlier detection, considering that these distributions are skewed and heavy-tailed, and have been adopted to characterize network traffic of the Internet and data centers \cite{benson2010network,leon2017probability}.
+We also adopt the Gaussian distributed data set $\pmb{Y}_g$ to simulate legitimate events with mean $\mu=0$ and unitary standard deviation $\sigma=1$.
 
@@
 An uniform noise between $-6$ and $6$ is used to add anomalies into the Gaussian distributed data set, and $\pmb{Y}_g^c$ denotes the contaminated Gaussian data contaminated by an uniform distribution.
-The figure \ref{fig:4.04} shows an example of the Gaussian distribution of normal observations and the uniform distribution to simulate the addition of anomalies.
+The figure \ref{fig:4.04} shows an example of the Gaussian distribution of legitimate observations and the uniform distribution to simulate the addition of anomalies.
 
@@
 \subsection{The CTU-13 data set}
-\label{sec:4_CTU-13}
+\label{sec:3_CTU-13}
 
@@
 The CTU-13 data set contains 13 scenarios with network flows of botnet malwares, that are: neris, rbot, virut, menti, sogou, nsys.ay and murlo.
-The botnet traffic is also classified as attack or command and control (C\&C), while the legitimate flows can also be classified as normal or background.
-It is important to note that the CTU-13 data set can also be formulated as a superposition of signal, artefact and noise, which refer to background, normal a botnet traffic, respectively.
+The botnet traffic is also classified as attack or command and control (C\&C), while the legitimate flows can also be classified as legitimate or background.
 
@@
 \begin{itemize}
-	\item \pmb{Attacks:} Click Fraud (CF), Port Scan (PS), Compiled and Controlled by Authors (US), SPAM and DDOS;
+	\item \pmb{Attacks:} Click Fraud (CF), Port Scan (PS), Fast Flux (FF), SPAM and DDoS;
 	\item \pmb{C\&C:} IRC, P2P and HTTP.
@@
 
-We refer to Garcia \cite{garcia2014identifying} and Garcia \emph{et al.} \cite{garcia2014empirical} for a detailed description of the performed attacks and C\&C flows, as well as for more information about the topology of the adopted testbed, rules for classifying normal flows, and an analysis of behaviors or patterns of the malware's traffic.
+We refer to Garcia \cite{garcia2014identifying} and Garcia \emph{et al.} \cite{garcia2014empirical} for a detailed description of the performed attacks and C\&C flows, as well as for more information about the topology of the adopted testbed, rules for classifying legitimate flows, and an analysis of behaviors or patterns of the malware's traffic.
 
 For all the scenarios, the authors of the CTU-13 data set convert the captured pcap files to NetFlows and release the processed flows.
-The data set provides ground-truth labels for flows as follows: flows from or to the infected machines are labeled as “botnet”; flows from or to well known and controlled machines are labeled as “normal”; all other flows are labeled as “background.”
+The data set provides ground-truth labels for flows as follows: flows from or to the infected machines are labeled as “botnet”; flows from or to well-known and controlled machines are labeled as “normal”; all other flows are labeled as “background.”
 
-Table \ref{tab:4.01} presents an overview grouped by scenario, according to the column ID, and shows the malwares used for botnet attacks, the types of attacks or C\&C, the total number of flows, the number of malicious flows which includes flows of C\&C and attacks, and finally shows the number of normal flows.
+Table \ref{tab:4.01} presents an overview grouped by scenario, according to the column ID, and shows the malwares used for botnet attacks, the types of attacks or C\&C, the total number of flows, the number of malicious flows which includes flows of C\&C and attacks, and finally shows the number of legitimate flows.
 
@@
 	\scriptsize
-	\caption{CTU-13 data set Description}
+	\caption{CTU-13 data set description}
 	\label{tab:4.01}
 	\begin{tabular}{| l | l | l | r | r | r | r | r | r | r | r | }
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}\pmb{ID}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\pmb{Bot}\end{tabular}	 &\begin{tabular}[x]{@{}l@{}}\pmb{Type}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\pmb{Total}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\pmb{Malicious}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\pmb{C\&C}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\pmb{Attack}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\pmb{Normal}\end{tabular}\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}\textbf{ID}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\textbf{Bot}\end{tabular}	 &\begin{tabular}[x]{@{}l@{}}\textbf{Type}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\textbf{Total}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Malicious}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{C\&C}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Attack}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Normal}\end{tabular}\\ \hline
 			10 & neris &\makecell[l]{IRC, Spam,\\CF} & 2,824,636 & 40,961 (1.45\%) & 341 (0.01\%) & 40,620 (1.44\%) &30,387 (1.07\%)\\ \hline
 			11 & neris &\makecell[l]{IRC, Spam,\\CF} & 1,808,122 & 20,941 (1.16\%) & 673 (0.04\%) & 20,268 (1.12\%) &9,120 (0.5\%)\\ \hline
-			12 & rbot &\makecell[l]{IRC, PS,\\US} & 4,710,638 & 26,822 (0.57\%) & 63 (0.00\%) & 26,759 (0.57\%) &116,887 (2.48\%)\\ \hline
-			15 & rbot &\makecell[l]{IRC, DDoS,\\US} & 1,121,076 & 2,580 (0.23\%) & 52 (0.00\%) & 2,528 (0.23\%) &25,268 (2.25\%)\\ \hline
+			12 & rbot &\makecell[l]{IRC, PS} & 4,710,638 & 26,822 (0.57\%) & 63 (0.00\%) & 26,759 (0.57\%) &116,887 (2.48\%)\\ \hline
+			15 & rbot &\makecell[l]{IRC, DDoS} & 1,121,076 & 2,580 (0.23\%) & 52 (0.00\%) & 2,528 (0.23\%) &25,268 (2.25\%)\\ \hline
 			15-2 & virut &\makecell[l]{Spam, PS,\\HTTP} & 129,832 & 901 (0.69\%) & 24 (0.02\%) & 877 (0.68\%) &4,679 (3.6\%)\\ \hline
@@
 			17 & neris &\makecell[l]{IRC, Spam,\\CF, PS} & 2,087,508 & 184,987 (8.86\%) & 2,973 (0.14\%) & 182,014 (8.72\%) &43,340 (1.57\%)\\ \hline
-			18 & rbot &\makecell[l]{IRC, DDoS,\\US} & 1,309,791 & 106,352 (8.12\%) & 33 (0.00\%) & 106,319 (8.12\%) &15,847 (1.2\%)\\ \hline
-			18-2 & rbot &\makecell[l]{IRC, DDoS,\\US} & 107,251 & 8,164 (7.61\%) & 2 (0.00\%) & 8,162 (7.61\%) &2,718 (2.53\%)\\ \hline
+			18 & rbot &\makecell[l]{IRC, DDoS} & 1,309,791 & 106,352 (8.12\%) & 33 (0.00\%) & 106,319 (8.12\%) &15,847 (1.2\%)\\ \hline
+			18-2 & rbot &\makecell[l]{IRC, DDoS} & 107,251 & 8,164 (7.61\%) & 2 (0.00\%) & 8,162 (7.61\%) &2,718 (2.53\%)\\ \hline
 			19 & nsys.ay &P2P & 325,471 & 2,168 (0.67\%) & 25 (0.01\%) & 2,143 (0.66\%) &7,628 (2.35\%)\\ \hline
@@
 The full data set and scenarios can be denoted as $\pmb{X} = \{\pmb{X}_{10}, \pmb{X}_{11}, \ldots , \pmb{X}_{18-2}, \pmb{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}.
-In our experiment each contaminated scenario $\pmb{X}_i$ is split into $\pmb{X}_i^s$ containing 50\% of the the normal data, and into $\pmb{X}_i^c$ that is composed by all anomalous flows and the necessary number of normal flows to have a testing data with the desired contamination rate.
+In our experiment each contaminated scenario $\pmb{X}_i$ is split into $\pmb{X}_i^s$ containing 50\% of the legitimate data, and into $\pmb{X}_i^c$ that is composed by all anomalous flows and the necessary number of legitimate flows to have a testing data with the desired contamination rate.
 
@@
 
-We conducted and Exploratory Data Analysis (EDA) on the CTU-13 and observed that some features are skewed and present an overlapping between normal and anomalous flows, as can be seen in Figure \ref{fig:4.05} and Figure \ref{fig:4.06}, that present the distributions of TCP state of scenario 16 and the type of services from destination of scenario 10.
+We conducted and Exploratory Data Analysis (EDA) on the CTU-13 and observed that some features are skewed and present high overlapping between legitimate and anomalous flows, as can be seen in Figure \ref{fig:4.05} and Figure \ref{fig:4.06}, that present the distributions of TCP state of scenario 16 and the type of services from destination of scenario 10.
 
@@
 
-However, it is not possible to observe a pattern on distributions of all the features and scenarios of CTU-13, as depicted by Figures \ref{fig:4.08} and \ref{fig:4.09}, that show the distributions of TCP states of the scenario 10 and source ports of scenario 16, and highlight the distributions of normal and anomalous flows.
+However, it is not possible to observe a pattern on distributions of all the features and scenarios of CTU-13, as depicted by Figures \ref{fig:4.08} and \ref{fig:4.09}, that show the distributions of TCP states of the scenario 10 and source ports of scenario 16, and highlight the distributions of legitimate and anomalous flows.
 
@@
 
-Due to the number of available features and the class overlapping between normal and anomalous flows, we performed an correlation analysis and an empirical cross validation in order to identify the best set of features for network attack detection.
+Due to the number of available features and the class overlapping between legitimate and anomalous flows, we performed an correlation analysis and an empirical cross validation in order to identify the best set of features for network attack detection.
 Therefore, we adopt the following features: state, destination type of service, destination port, source port, total number of packets, total number of bytes and number of bytes from the source.
@@
 \section{Moment Distances from Robust Subspace for Network Attack Detection}
-\label{sec:4_proposal}
+\label{sec:3_proposal}
 
@@
 
-Robust subspace learning can be defined as the decomposition of a given data matrix $\pmb{X}$ into the sum of a low rank matrix $\pmb{L}$, whose column subspace gives the principal components, and a sparse matrix $\pmb{S}$, with outliers or noise.
-We propose to learn a robust subspace $\pmb{L}$ from the normal traffic $\pmb{X}^s$ for computing $\pmb{L}^s$, $\pmb{S}^s$ and the robust moments, i.e.
-the mean $\bar{\pmb{x}}$, skewness $\bar{\pmb{s}}$ and kurtosis $\bar{\pmb{k}}$, in order to evaluate the distance $\pmb{d}$ between contaminated observations $\pmb{X}^c$ and moments computed from the learned robust subspace, in order to classify as anomalous the observations with the largest distances.
+Robust subspace learning can be defined as the decomposition of a given data matrix $\pmb{X} \in \mathbb{R}^{M \times N}$, with rows representing observations and columns representing features.
+Note that in Section \ref{sec:2_ModelingData} we introduced $\pmb{X}$ for a packet level analysis, modeled as a matrix of communication port by time.
+However, in this chapter we adopt a flow based analysis, where $\pmb{X}$ is modeled as a matrix of flows by features, and $\pmb{X}$ is decomposed into the sum of a low rank matrix $\pmb{L} \in \mathbb{R}^{M \times N}$, whose column subspace gives the robust principal components without outliers and noise, and a sparse matrix $\pmb{S} \in \mathbb{R}^{M \times N}$, with element-wise outliers or noise.
 
-RPCA is a well known method to recover a low-rank matrix $\pmb{L}$ and sparse matrix $\pmb{S}$ from corrupted measurements modeled as $\pmb{X} = \pmb{L} + \pmb{S}$.
+Even though robust subspace learning has been adopted for network anomaly detection by means of highlights from the matrix $\pmb{S}$, it was shown that $\pmb{S}$ can indicate noise and outliers that can not be classified as malicious \cite{vaswani2018robust, lerman2018overview}, resulting into false positive classifications and requiring complementary approaches in order to obtain precise network anomaly detection \cite{zhou2017anomaly}.
+
+Therefore, for network anomaly detection, we propose to learn a robust subspace from the legitimate traffic $\pmb{X}^s$ for computing $\pmb{L}^s$ and $\pmb{S}^s$, followed by computing the robust moments, i.e.
+the mean $\pmb{\mu}$, skewness $\pmb{\epsilon}$ and kurtosis $\pmb{\kappa}$, in order to evaluate the distance $\pmb{d}$ between contaminated observations $\pmb{X}^c$ and robust moments.
+The largest distances are classified as anomalous and indicate network attacks, denoted as $\pmb{N}$ by the data model $\pmb{X} = \pmb{U} + \pmb{N}$, introduced by (\ref{eq:2.01}), where $\pmb{U}$ denotes the legitimate network traffic.
+
+RPCA is a well-known method to recover a low-rank matrix $\pmb{L}$ and sparse matrix $\pmb{S}$ from corrupted measurements modeled as $\pmb{X} = \pmb{L} + \pmb{S}$.
 This decomposition in low-rank and sparse matrices can be achieved by techniques such as Principal Component Pursuit method (PCP), and by optimization methods, such as the Augmented Lagrange Multiplier Method (ALM), Alternating Direction Method (ADM), Fast Alternating Minimization (FAM) or Iteratively Reweighted Least Squares (IRLS) \cite{candes2011robust,vaswani2018robust,lerman2018overview}.
@@
 \end{equation}
-where $\lambda$ is a positive weighting parameter, which determines the sparsity of $\pmb{S}$, $\sigma$ denotes the singular values of a matrix.
+where $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix, $\lambda$ is a positive weighting parameter, which determines the sparsity of $\pmb{S}$, and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries, and $\sigma$ denotes the singular values of a matrix.
 
@@
 
-The RPCA with ALM can be formulated:
+The RPCA with ALM can be formulated as:
 \begin{equation}\label{eq:4.04}
@@
 \end{equation}
-where $\pmb{Y}$ is the multiplier of the linear constraint, $\mu$ is the penalty parameter for the violation of the linear constraint \cite{yuan2009sparse}.
+where $\pmb{Y}$ is the multiplier of the linear constraint and $\mu$ is the penalty parameter for the violation of the linear constraint \cite{yuan2009sparse}.
 Thus, an iterative scheme can be presented as:
@@
 		\begin{matrix} 
-			(\pmb{L}_{k+1}, \pmb{S}_{k+1}) \in \operatorname*{argmin}_{\pmb{L,S} \in \mathbb{R}^{m \times n}} \{l(\pmb{L}, \pmb{S}, \pmb{Y}_{k})\}, \\ 
+			(\pmb{L}_{k+1}, \pmb{S}_{k+1}) \in \argmin\limits_{\pmb{L,S} \in \mathbb{R}^{m \times n}} \{l(\pmb{L}, \pmb{S}, \pmb{Y}_{k})\}, \\ 
 			\pmb{Y}_{k+1} = \pmb{Y}_{k} + \mu(\pmb{X} - \pmb{L}_{k} - \pmb{S}_{k}),
@@
 	\left\{\begin{matrix}
-	\pmb{L}_{k+1} \in \operatorname*{argmin}_{\pmb{L} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}, \pmb{S}_{k}, \pmb{Y}_{k})\}\\ 
-	\pmb{S}_{k+1} \in \operatorname*{argmin}_{\pmb{S} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}_{k+1}, \pmb{S}, \pmb{Y}_{k})\}\\ 
+	\pmb{L}_{k+1} \in \argmin\limits_{\pmb{L} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}, \pmb{S}_{k}, \pmb{Y}_{k})\}\\ 
+	\pmb{S}_{k+1} \in \argmin\limits_{\pmb{S} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}_{k+1}, \pmb{S}, \pmb{Y}_{k})\}\\ 
 	\pmb{Y}_{k+1} = \pmb{Y}_{k} + \mu(\pmb{X} - \pmb{L}_{k} - \pmb{S}_{k})
@@
 Moments are a set of statistical parameters to measure a distribution.
-The arithmetic mean is the first general moment, the second is the variance, while skewness (asymmetry) is the third moment and kurtosis (excess) is the fourth moment.
+The arithmetic mean is the first general moment, the second is the variance, while skewness (asymmetry) is the third moment and kurtosis (tailedness) is the fourth moment \cite{reis2016audio}.
 
-Let the mean $\bar{\pmb{x}} \in \mathbb{R}^{1 \times p}$ be
+Let the mean $\pmb{\mu} \in \mathbb{R}^{1 \times N}$ be
 
 \begin{equation}\label{eq:4.07}
-	\bar{\pmb{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} \pmb{x}_i, 
+	\pmb{\mu} = \displaystyle\frac{1}{M}\displaystyle\sum_{i = 1}^{M} \pmb{x}_i,
 \end{equation}
 
-and the covariance matrix $\bar{\pmb{\Sigma}} \in \mathbb{R}^{p \times p}$ be
+where $M$ is the number of samples, and let the sample covariance matrix $\hat{\pmb{\Sigma}} \in \mathbb{R}^{N \times N}$ be
 
 \begin{equation}\label{eq:4.08}
-	\bar{\pmb{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} (\pmb{x}_i - \bar{\pmb{x}})(\pmb{x}_i - \bar{\pmb{x}})^\prime,
+	\hat{\pmb{\Sigma}} = \displaystyle\frac{1}{N-1}\displaystyle\sum_{i = 1}^{N} (\pmb{x}_i - \pmb{\mu})(\pmb{x}_i - \pmb{\mu})^T,
 \end{equation}
 
-The general formula of the $u$-th moment can be expressed as
+According to Zwillinger and Kokoska \cite{zwillinger1999crc}, the general expression for the $p$-th moment $\pmb{m}_p \in \mathbb{R}^{1 \times N}$ about the mean $\pmb{\mu}$ is given by
 \begin{equation}\label{eq:4.09}
-	\pmb{m}_u = \displaystyle\frac{1}{n}\displaystyle\sum_{i = 1}^{n}(\pmb{x}_i - \bar{\pmb{x}})^u.
+	\pmb{m}_p = \displaystyle\frac{1}{M}\displaystyle\sum_{i = 1}^{M}(\pmb{x}_i - \pmb{\mu})^p.
 \end{equation}
 
-Therefore, the skewness is calculated by
+Therefore, Zwillinger and Kokoska \cite{zwillinger1999crc} present that the skewness $\pmb{\epsilon} \in \mathbb{R}^{1 \times N}$ about the mean $\pmb{\mu}$ is calculated by
 \begin{equation}\label{eq:4.10}
-	\bar{\pmb{s}} = \frac{\pmb{m}_3}{\pmb{m}_2^{\frac{3}{2}}},
+	\pmb{\epsilon} = \frac{\pmb{m}_3}{\pmb{m}_2^{\frac{3}{2}}},
 \end{equation}
 
-and the kurtosis as
+and the kurtosis $\pmb{\kappa} \in \mathbb{R}^{1 \times N}$ is given as
 \begin{equation}\label{eq:4.11}
-	\bar{\pmb{k}} = \frac{\pmb{m}_4}{\pmb{m}_2^2}.
+	\pmb{\kappa} = \frac{\pmb{m}_4}{\pmb{m}_2^2}.
 \end{equation}
 
-We propose to compute the mean $\bar{\pmb{x}}$, the skewness $\bar{\pmb{s}}$, the kurtosis $\bar{\pmb{k}}$ and the covariance matrix $\bar{\pmb{\Sigma}}$ from the robust subspace $\pmb{L}^s$, after to minimize the Equation \ref{eq:4.06} for $\pmb{X}^s$.
+We propose to compute the mean $\pmb{\mu}$, the skewness $\pmb{\epsilon}$, the kurtosis $\pmb{\kappa}$ and the covariance matrix $\hat{\pmb{\Sigma}}$ from the robust subspace $\pmb{L}^s$, after to minimize the (\ref{eq:4.06}) for $\pmb{X}^s$.
 We also propose to compute the Mahalanobis Distance (MD) for detecting anomalies between contaminated observations and the moments calculated from a robust subspace computed by RPCA.
@@
 \begin{equation}\label{eq:4.12}
-	\pmb{d}(\pmb{x},\bar{\pmb{x}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{x}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{x}})^\prime},
+	\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \pmb{\mu}) \hat{\pmb{\Sigma}}^{-1}(\pmb{x} - \pmb{\mu})^T},
 \end{equation}
-where $\pmb{x}$ is a vector of a new observation, $\bar{\pmb{x}}$ is the mean vector of known observations, also referred as location, and $\bar{\pmb{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter.
+where $\pmb{x}$ is a vector of a new observations, $\pmb{\mu}$ is the mean vector of known observations, also referred as location, and $\hat{\pmb{\Sigma}}$ is the covariance matrix of known observations, also referred as scatter.
 The classical MD usually relies on robust mean and robust covariance matrix for outlier detection, which are commonly computed by MCD \cite{rousseeuw1984mcd, rousseeuw1999fastmcd}.
-Here we propose to calculate the MD according to Equation \ref{eq:4.12} by means of the mean and covariance matrix computed from a robust subspace learned by RPCA.
+Here we propose to compute the Robust-Mean Distance $\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}})$ according to (\ref{eq:4.12}), but adopting the mean $\pmb{\mu}$ and covariance matrix $\hat{\pmb{\Sigma}}$ calculated from a robust subspace $\pmb{L}^s$ learned by RPCA.
 
-We also propose to extend the Equation \ref{eq:4.12} to implement a Skewness-based Mahalanobis Distance, as follows:
+We also propose to extend (\ref{eq:4.12}) to implement the Robust-Skewness Distance $\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}})$, as follows:
 
 \begin{equation}\label{eq:4.13}
-	\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{s}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{s}})^\prime}.
+	\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \pmb{\epsilon}) \hat{\pmb{\Sigma}}^{-1}(\pmb{x} - \pmb{\epsilon})^T}.
 \end{equation}
 
-Finally, we propose to extend the Equation \ref{eq:4.12} to implement a Kurtosis-based Mahalanobis Distance, as follows:
+Finally, we propose to extend (\ref{eq:4.12}) to implement the Robust-Kurtosis Distance $\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}})$, as follows:
 
 \begin{equation}\label{eq:4.14}
-	\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{k}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{k}})^\prime}.
+	\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \pmb{\kappa}) \hat{\pmb{\Sigma}}^{-1}(\pmb{x} - \pmb{\kappa})^T}.
 \end{equation}
 
-The distances $\pmb{d}(\pmb{x},\bar{\pmb{x}}, \bar{\pmb{\Sigma}})$, $\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}})$ and $\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}})$ shall be computed and evaluated separately and independently, for network attack detection.
-Therefore, the robust subspace learning from normal data combined to MD is called as md-RPCA when using mean-based MD, or is called sd-RPCA when using skewness-based MD, or kd-RPCA when using kurtosis-based MD.
+The distances $\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}})$, $\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}})$ and $\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}})$ shall be computed and evaluated separately and independently, for network attack detection based on robust mean, robust skewness and robust kurtosis, respectively.
 
-The contamination rate parameter is widely adopted for well established outlier detection algorithms \cite{zhao2019pyod}, and refers to the percentage of observations that are classified as anomalous.
-The contamination rate can be well known for some areas, or can be computed by cross-validation or can be assumed according to previous observations.
-In our proposal, the contamination defines the number of the largest distances $\pmb{d}$ that shall be classified as anomalous.
-The observations classified as legitimate and anomalous, according to the contamination $c$, are denoted by the vector $\bar{\pmb{t}}$, with values of 1 for classifying anomalies and 0 to denote legitimate observation.
+Therefore, the robust subspace $\pmb{L}^s$ learned from legitimate data $\pmb{X}^s$ followed by the Robust-Mean Distance $\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}})$ is called as Mean Distance of Robust Principal Component Analysis (md-RPCA), or is called Skewness Distance of Robust Principal Component Analysis (sd-RPCA) when followed by Robust-Skewness Distance, given by $\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}})$, or Kurtosis Distance of Robust Principal Component Analysis (kd-RPCA) when followed by Robust-Kurtosis Distance, given by $\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}})$.
+
+The contamination rate, denoted by $c$, is a parameter traditionally adopted by well established outlier detection algorithms \cite{zhao2019pyod}, and refers to the percentage rate of observations that are known as anomalous.
+The contamination rate can be well-known for some areas, or can be computed by cross-validation \cite{arlot2010survey} or can be assumed according to previous observations.
+In our proposal, the contamination defines the number of the largest distances $\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}})$, $\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}})$ or $\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}})$ that shall be classified as anomalous.
+The observations classified as legitimate and anomalous, according to the contamination $c$, are denoted by the vector $\hat{\pmb{t}}$, with values of 1 for observations classified as anomalous and 0 to denote legitimate observations.
 
@@
 	\SetAlgoLined
-	\KwResult{$\bar{\pmb{t}}_{\bar{\pmb{x}}}$, $\bar{\pmb{t}}_{\bar{\pmb{s}}}$, $\bar{\pmb{t}}_{\bar{\pmb{k}}}$}
+	\KwResult{$\hat{\pmb{t}}_{\mu}$, $\hat{\pmb{t}}_{\epsilon}$, $\hat{\pmb{t}}_{\kappa}$}
 	Given $\pmb{X}$ split into $\pmb{X}^s$ and $\pmb{X}^c$\;
-	\While{not $\min_{L,S}\left \| \pmb{L} \right \|_{*} + \lambda \left \| \pmb{S} \right \|_{1}$ of $\pmb{X}^s$}{
-		$\pmb{L}_{k+1} \in \operatorname*{argmin}_{\pmb{L} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}, \pmb{S}_{k}, \pmb{Y}_{k})\}$\;
-		$\pmb{S}_{k+1} \in \operatorname*{argmin}_{\pmb{S} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}_{k+1}, \pmb{S}, \pmb{Y}_{k})\}$\;
+	\While{\rm not $\min_{L,S}\left \| \pmb{L} \right \|_{*} + \lambda \left \| \pmb{S} \right \|_{1}$ \rm from $\pmb{X}^s$}{
+		$\pmb{L}_{k+1} \in \argmin\limits_{\pmb{L} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}, \pmb{S}_{k}, \pmb{Y}_{k})\}$\;
+		$\pmb{S}_{k+1} \in \argmin\limits_{\pmb{S} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}_{k+1}, \pmb{S}, \pmb{Y}_{k})\}$\;
 		$\pmb{Y}_{k+1} = \pmb{Y}_{k} + \mu(\pmb{X} - \pmb{L}_{k} - \pmb{S}_{k})$\;
 	}
-	$\bar{\pmb{x}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} \pmb{x}_i$, for $\pmb{L}^s$\;
-	$\bar{\pmb{\Sigma}} = \displaystyle\frac{1}{h}\displaystyle\sum_{i\in \pmb{h}} (\pmb{x}_i - \bar{\pmb{x}})(\pmb{x}_i - \bar{\pmb{x}})^\prime$, for $\pmb{L}^s$\;
-	$\bar{\pmb{s}} = \frac{\pmb{m}_3}{\pmb{m}_2^{\frac{3}{2}}}$, for $\pmb{L}^s$\;
-	$\bar{\pmb{k}} = \frac{\pmb{m}_4}{\pmb{m}_2^2}$, for $\pmb{L}^s$\;
-	$\pmb{d}(\pmb{x},\bar{\pmb{x}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{x}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{x}})^\prime}$\;
-	$\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{s}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{s}})^\prime}$\;
-	$\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \bar{\pmb{k}}) \bar{\pmb{\Sigma}}^{-1}(\pmb{x} - \bar{\pmb{k}})^\prime}$\;
-	$\bar{\pmb{t}}_{\bar{\pmb{x}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{x}}, \bar{\pmb{\Sigma}})]^c$\;
-	$\bar{\pmb{t}}_{\bar{\pmb{s}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{s}}, \bar{\pmb{\Sigma}})]^c$\;
-	$\bar{\pmb{t}}_{\bar{\pmb{k}}} = [\pmb{d}(\pmb{x}, \bar{\pmb{k}}, \bar{\pmb{\Sigma}})]^c$\;
+	$\pmb{\mu} = \displaystyle\frac{1}{M}\displaystyle\sum_{i = 1}^{M} \pmb{x}_i$, from $\pmb{L}$\;
+	$\hat{\pmb{\Sigma}} = \displaystyle\frac{1}{N-1}\displaystyle\sum_{i = 1}^{N} (\pmb{x}_i - \pmb{\mu})(\pmb{x}_i - \pmb{\mu})^T$, \rm from $\pmb{L}$\;
+	$\pmb{\epsilon} = \frac{\pmb{m}_3}{\pmb{m}_2^{\frac{3}{2}}}$, \rm from $\pmb{L}$\;
+	$\pmb{\kappa} = \frac{\pmb{m}_4}{\pmb{m}_2^2}$, \rm from $\pmb{L}$\;
+	$\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \pmb{\mu}) \hat{\pmb{\Sigma}}^{-1}(\pmb{x} - \pmb{\mu})^T}$, \rm from $\pmb{X}^c$\;
+	$\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \pmb{\epsilon}) \hat{\pmb{\Sigma}}^{-1}(\pmb{x} - \pmb{\epsilon})^T}$, \rm from $\pmb{X}^c$\;
+	$\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \pmb{\kappa}) \hat{\pmb{\Sigma}}^{-1}(\pmb{x} - \pmb{\kappa})^T}$, \rm from $\pmb{X}^c$\;
+	$\hat{\pmb{t}}_{\mu} = [\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}})]^c$\;
+	$\hat{\pmb{t}}_{\epsilon} = [\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}})]^c$\;
+	$\hat{\pmb{t}}_{\kappa} = [\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}})]^c$\;
 	\caption{Moment Distances from Robust Subspace}
@@
 
-The steps between 1 and 10 of the Algorithm \ref{alg:4.01} are the training from normal data $\pmb{X}^s$, which are common steps shared by md-RPCA, sd-RPCA and kd-RPCA.
+The steps between 1 and 10 of the Algorithm \ref{alg:4.01} are the training from legitimate data $\pmb{X}^s$, which are common steps shared by md-RPCA, sd-RPCA and kd-RPCA.
 The steps between 11 and 16 aim the anomaly detection from new contaminated observations, by means of Mahalanobis Distance of robust moments.
 Note that the steps 11 and 14 refer to the steps of md-RPCA for anomaly detection, while the steps 12 and 15 refer to the sd-RPCA, and the steps 13 and 16 refer to the steps of kd-RPCA.
-It is important to highlight that the anomaly detection from new observations does not require new robust subspace learning when adopting the semi-supervised approach, which only requires to compute a moment-based Mahalanobis distance to classify the $c$ largest distances as anomalous or legitimate.
+It is important to highlight that the anomaly detection from new observations does not require new robust subspace learning when adopting the semi-supervised approach, which only requires to compute a moment-based distance to classify the $c$ largest distances as anomalous or legitimate.
 
@@
 \section{Experiments}
-\label{sec:4_experiments}
+\label{sec:3_experiments}
 
 This section presents the performed experiments on simulated and real data set for anomaly detection.
-First, in Section \ref{sec:4_metric} we present the adopted metric to evaluate imbalanced data in the context of anomaly detection.
-The Section \ref{sec:4_SimulatedScenario} describes the experiment for anomaly detection on simulated skewed and heavy-tailed data set, and the Section \ref{sec:4_CTU-13} presents the experiment for network anomaly detection on CTU-13 data set.
+First, in Section \ref{sec:3_metric} we present the adopted metric to evaluate imbalanced data in the context of anomaly detection.
+In Section \ref{sec:3_SimulatedScenario} we describe the experiment for anomaly detection on simulated skewed and heavy-tailed data set, and we present in Section \ref{sec:3_CTU-13} the experiment for network anomaly detection on CTU-13 data set.
 
 \subsection{The metric}
-\label{sec:4_metric}
+\label{sec:3_metric}
 
-In anomalous detection problems, where anomalies are rare events, if one classify all observations as normal and apply an accuracy evaluation, one would have high accuracy but poor true-positive detection.
+In anomalous detection problems, where anomalies are rare events, if one classify all observations as legitimate and apply an accuracy evaluation, one would have high accuracy but poor true-positive detection.
 Due to the importance given by the $F_1$ (also referred as F-score or F-measure) to true-positive detection in scenarios such as the network attack detection, it is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic}.
@@
 
-The $F_1$ is the harmonic mean of precision and recall and is defined by:
+The $F_1$ is the harmonic mean of precision and recall, where $p_{\rm p}$ denotes the precision and recall is denoted as $r_{\rm c}$. The $F_1$ is given by
 \begin{equation}\label{eq:4.15}
-	F_1 = 2 * \frac{Precision * Recall}{Precision + Recall},
+	F_1 = 2 \cdot \frac{p_{\rm r} \cdot r_{\rm c}}{p_{\rm r} + r_{\rm c}}.
 \end{equation}
-where $pr$ is the precision, defined by 
+
+Precision can be seen as a measure of exactness or accuracy, that relies on true positive and false positive measures, denoted by $t_{\rm p}$ and $f_{\rm p}$, respectively.
+The precision is defined by
 \begin{equation}\label{eq:4.16}
-	Precision = \frac{True Positive}{True Positive + False Positive}
+	p_r = \frac{t_{\rm p}}{t_{\rm p} + f_{\rm p}}.
 \end{equation}
-and $rc$ is the recall, defined by 
+
+Recall is a measure of completeness, to calculate proportion of actual positives was identified correctly, by means of the true positive and false negative measures, denoted by $t_{\rm p}$ and $f_{\rm n}$.
+The recall is defined by
 \begin{equation}\label{eq:4.17}
-	Recall = \frac{True Positive}{True Positive + False Negative}.
+	r_c = \frac{t_{\rm p}}{t_{\rm p} + f_{\rm n}}.
 \end{equation}
@@
 \subsection{Simulated Experiment}
-\label{sec:4_SimulatedScenario}
+\label{sec:3_SimulatedScenario}
 
-Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled normal and anomalous data for training anomaly detection models, while the latter does not require labeled data or training.
-Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on normal data for training and that non-malicious data can be obtained from historical information and from rule-based approaches.
+Anomaly detection algorithms usually rely on supervised or unsupervised methods, where the former requires labeled legitimate and anomalous data for training anomaly detection models, while the latter does not require labeled data or training.
+Semi-supervised algorithms are an alternative for the anomaly detection problem, considering that this method only relies on legitimate data for training and that non-malicious data can be obtained from historical information and from rule-based approaches.
 
-We propose semi-supervised and unsupervised approaches for m-RPCA, where the former relies on normal data $\pmb{X}^s$ for training and on contamination rate $c$ for anomaly detection, while the latter relies only on contaminated data $\pmb{X}^c$ for robust subspace learning and rely on contamination rate $c$ to select the largest distances.
-We assume that $c$ is well known or can be estimated for real world problems of anomaly detection, in accordance to the assumption adopted by the well established algorithms \cite{zhao2019pyod} selected for comparison, that also rely on contamination rate $\pmb{X}^c$.
+We propose semi-supervised and unsupervised approaches for m-RPCA, where the former relies on legitimate data $\pmb{X}^s$ for training and on contamination rate $c$ for anomaly detection, while the latter relies only on contaminated data $\pmb{X}^c$ for robust subspace learning and relies on contamination rate $c$ to select the largest distances.
+We assume that $c$ is well-known or can be estimated for real world problems of anomaly detection, in accordance to the assumption adopted by the well established algorithms \cite{zhao2019pyod} selected for comparison, that also rely on contamination rate $\pmb{X}^c$.
 
@@
 
-Therefore, we propose to evaluate the following approaches for m-RPCA: semi-supervised; contaminated semi-supervised; and unsupervised.
+Therefore, we propose to evaluate the following approaches for m-RPCA: semi-supervised, contaminated semi-supervised, and unsupervised.
 
-For the semi-supervised approach we propose to learn the robust subspace and compute the moments from the normal data set $\pmb{Y}_g$, $\pmb{Y}_p$ and $\pmb{Y}_l$ with Gaussian, Pareto and Lognormal distributions, respectively, and test the anomaly detection for contaminated data set $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$.
+For the semi-supervised approach we propose to learn the robust subspace and compute the moments from the legitimate data sets $\pmb{Y}_g$, $\pmb{Y}_p$ and $\pmb{Y}_l$ with Gaussian, Pareto and Lognormal distributions, respectively, and test the anomaly detection for contaminated data sets $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$.
 
-For the Contaminated Semi-supervised approach we evaluate the robustness of the m-RPCA approaches for learning from contaminated training data, in order to analyse if m-RPCA can be an alternative for the lack of known normal data.
-Therefore we propose to train the model from a contaminated normal data $\pmb{Y}_g^{c'}$, $\pmb{Y}_p^{c'}$ and $\pmb{Y}_l^{c'}$, with the same contamination rate of the testing data, but without data repetition between training and testing, taking into account that we shall consider a different contaminated data but adopt the same contamination rate for training and testing.
+For the Contaminated Semi-supervised approach we evaluate the robustness of the m-RPCA approaches for learning from contaminated training data, in order to analyze if m-RPCA can be an alternative for the lack of known legitimate data.
+Therefore we propose to train the model from a contaminated legitimate data $\pmb{Y}_g^{c'}$, $\pmb{Y}_p^{c'}$ and $\pmb{Y}_l^{c'}$, with the same contamination rate of the testing data, but without data repetition between training and testing, taking into account that we shall consider a different contaminated data but adopt the same contamination rate for training and testing.
 
@@
 \subsection{Experiment for CTU-13}
-\label{sec:4_CTU13Scenario}
+\label{sec:3_CTU13Scenario}
 
@@
 
-In our experiment setup each contaminated scenario $\pmb{X}_i$ of CTU-13 is split into $\pmb{X}_i^s$, containing 50\% of the the normal data, and into test data $\pmb{X}_i^c$ containing 33\% of normal and 67\% of anomalous data.
+In our experiment setup each contaminated scenario $\pmb{X}_i$ of CTU-13 is split into $\pmb{X}_i^s$, containing 50\% of the legitimate data, and into test data $\pmb{X}_i^c$ containing 33\% of legitimate and 67\% of anomalous data.
 We adopt a contamination rate $c$ of 33\% for  experimenting anomaly detection for the CTU-13 data set, considering that this contamination rate presented good results on the simulated experiment for our proposals and for some selected algorithms.
 
-This experiment consider only the semi-supervised approach due to results of simulated experiments on simulated data set presented in Subsection \ref{sec:4_simulated_result}, that shows better results for the semi-supervised approach and highlight that the this approach can obtain good results even when trained with contaminated data set.
+This experiment consider only the semi-supervised approach due to results of simulated experiments on simulated data set presented in Subsection \ref{sec:3_simulated_result}, that shows better results for the semi-supervised approach and highlight that the this approach can obtain good results even when trained with contaminated data set.
 
 \section{Results of Simulated Experiment}
-\label{sec:4_simulated_result}
+\label{sec:3_simulated_result}
 
@@
 
-We initially evaluated the $F_1$ of the selected algorithms and m-RPCA approaches for anomaly detection on Gaussian distributed legitimate data contaminated by uniform distributed anomalies.
+We evaluated the $F_1$ of the selected algorithms and m-RPCA approaches for anomaly detection on Gaussian distributed legitimate data contaminated by uniform distributed anomalies.
 The Figure \ref{fig:4.10} shows the $F_1$ over the contamination rate between 1\% and 50\% for m-RPCA approaches and Isolation Forest (IF) \cite{liu2008isolation}, k-Nearest Neighbors (KNN) \cite{angiulli2002fast}, Local Outlier Factor (LOF) \cite{breunig2000lof}, Minimum Covariance Deteminant (MCD) \cite{rousseeuw1999fastmcd}, One-Class Support Vector Machines (OCSVM) \cite{scholkopf2001estimating}, PCA \cite{shyu2003novel} and ROBPCA-AO \cite{hubert2009robustskewed}.
@@
 	\centering
-	\includegraphics[width=10cm]{figures/ch4/gaussian_f1_contamination.png}
+	\includegraphics[width=12cm]{figures/ch4/gaussian_f1_contamination.eps}
 	\caption{Anomaly detection on Gaussian distributed with uniform anomalies}
@@
 	\centering
-	\includegraphics[width=10cm]{figures/ch4/pareto_f1_contamination.png}
+	\includegraphics[width=12cm]{figures/ch4/pareto_f1_contamination.eps}
 	\caption{Anomaly detection on Pareto distributed with Gaussian anomalies}
@@
 	\centering
-	\includegraphics[width=10cm]{figures/ch4/lognormal_f1_contamination.png}
+	\includegraphics[width=12cm]{figures/ch4/lognormal_f1_contamination.eps}
 	\caption{Anomaly detection on Lognormal distributed with Gaussian anomalies}
@@
 
-Following we present the Tables \ref{tab:4.02}, \ref{tab:4.03} and \ref{tab:4.04} to show the results of the selected algorithms and of our proposals for anomaly detection on Gaussian, Pareto and Lognormal distributions, with 10\%, 25\% and 33\% of contamination rate.
+Following we present the Tables \ref{tab:4.02}, \ref{tab:4.03} and \ref{tab:4.04} to show the results of the selected algorithms and our proposals for anomaly detection on Gaussian, Pareto and Lognormal distributions, with 10\%, 25\% and 33\% of contamination rate.
+The results are the $F_1$ for each scenario and algorithm, sorted by best result by algorithm for anomaly detection on Gaussian, Pareto and Lognormal distributions.
 
@@
 	\begin{tabular}{| l | l | l | l |}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\pmb{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Gaussian + Uniform}\\\pmb{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Pareto + Gaussian}\\\pmb{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\pmb{Lognormal + Gaussian}\\\pmb{($F_1$)}\end{tabular}\\ \hline
-            Isolation Forest	&0.98	&0.44	&0.05\\ \hline
-            KNN	&0.98	&0.08	&0.00\\ \hline
-            LOF	&0.35	&0.40	&0.36\\ \hline
-            MCD	&0.98	&0.02	&0.00\\ \hline
-            One-class SVM	&0.98	&0.24	&0.00\\ \hline
-            PCA	&0.87	&0.76	&0.62\\ \hline
-            Semi-Supervised md-RPCA	&0.98	&0.90	&0.66\\ \hline
-            Semi-Supervised sd-RPCA	&0.98	&0.99	&0.97\\ \hline
-            Semi-Supervised kd-RPCA	&0.97	&0.99	&0.98\\ \hline
-            Contaminated Semi-Supervised md-RPCA	&0.98	&0.55	&0.35\\ \hline
-            Contaminated Semi-Supervised sd-RPCA	&0.98	&0.94	&0.96\\ \hline
-            Contaminated Semi-Supervised kd-RPCA	&0.76	&0.99	&0.98\\ \hline
-            Unsupervised md-RPCA	&0.98	&0.54	&0.31\\ \hline
-            Unsupervised sd-RPCA	&0.98	&0.82	&0.93\\ \hline
-            Unsupervised kd-RPCA	&0.74	&0.86	&0.99\\ \hline
-            ROBPCA-AO   &0.00	&0.00	&0.00\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian} +\\\textbf{Uniform} \textbf{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}\\ \hline
+        	Proposed Semi-Supervised sd-RPCA	&0.98	&0.99	&0.97	\\ \hline
+        	Proposed Semi-Supervised kd-RPCA	&0.97	&0.99	&0.98	\\ \hline
+        	Proposed Contaminated Semi-Supervised sd-RPCA	&0.98	&0.94	&0.96	\\ \hline
+        	Proposed Unsupervised sd-RPCA	&0.98	&0.82	&0.93	\\ \hline
+        	Proposed Contaminated Semi-Supervised kd-RPCA	&0.76	&0.99	&0.98	\\ \hline
+        	Proposed Unsupervised kd-RPCA	&0.74	&0.86	&0.99	\\ \hline
+        	Proposed Semi-Supervised md-RPCA	&0.98	&0.90	&0.66	\\ \hline
+        	PCA \cite{shyu2003novel}	&0.87	&0.76	&0.62	\\ \hline
+        	Proposed Contaminated Semi-Supervised md-RPCA	&0.98	&0.55	&0.35	\\ \hline
+        	Proposed Unsupervised md-RPCA	&0.98	&0.54	&0.31	\\ \hline
+        	Isolation Forest \cite{liu2008isolation}	&0.98	&0.44	&0.05	\\ \hline
+        	One-class SVM \cite{scholkopf2001estimating}	&0.98	&0.24	&0.00	\\ \hline
+        	LOF \cite{breunig2000lof}	&0.35	&0.40	&0.36	\\ \hline
+        	KNN \cite{angiulli2002fast}	&0.98	&0.08	&0.00	\\ \hline
+        	MCD \cite{rousseeuw1999fastmcd}	&0.98	&0.02	&0.00	\\ \hline
+        	ROBPCA-AO \cite{hubert2009robustskewed}		&0.00	&0.00	&0.00	\\ \hline
 	\end{tabular}
@@
 
-The results with 33\% of contamination shows high scores of anomaly detection for Gaussian data with uniform anomalies, with exception to LOF and ROBPCA-AO.
-The semi-supervised approaches of m-RPCA presented the highest results for anomaly detection on Pareto data with Gaussian anomalies, and for anomaly detection on Lognormal data with Gaussian anomalies.
+The results with 33\% of contamination shows high scores of anomaly detection for Gaussian data with uniform anomalies by the evaluated algorithms, with exception to LOF and ROBPCA-AO.
+The semi-supervised approaches of m-RPCA presented the highest results for anomaly detection on Pareto data with Gaussian anomalies, and for anomaly detection on Lognormal data with Gaussian anomalies, as well as for Gaussian data with uniform anomalies.
 
@@
 	\begin{tabular}{| l | l | l | l |}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\pmb{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Gaussian + Uniform}\\\pmb{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Pareto + Gaussian}\\\pmb{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\pmb{Lognormal + Gaussian}\\\pmb{($F_1$)}\end{tabular}\\ \hline
-            Isolation Forest	&0.97	&0.46	&0.05\\ \hline
-            KNN	&0.97	&0.05	&0.00\\ \hline
-            LOF	&0.27	&0.33	&0.30\\ \hline
-            MCD	&0.97	&0.79	&0.00\\ \hline
-            One-class SVM	&0.97	&0.11	&0.00\\ \hline
-            PCA	&0.91	&0.75	&0.56\\ \hline
-            Semi-Supervised md-RPCA	&0.97	&0.85	&0.49\\ \hline
-            Semi-Supervised sd-RPCA	&0.97	&0.99	&0.96\\ \hline
-            Semi-Supervised kd-RPCA	&0.96	&1.00	&0.98\\ \hline
-            Contaminated Semi-Supervised md-RPCA	&0.97	&0.54	&0.17\\ \hline
-            Contaminated Semi-Supervised sd-RPCA	&0.97	&0.91	&0.95\\ \hline
-            Contaminated Semi-Supervised kd-RPCA	&0.63	&1.00	&0.97\\ \hline
-            Unsupervised md-RPCA	&0.97	&0.53	&0.14\\ \hline
-            Unsupervised sd-RPCA	&0.97	&0.88	&0.91\\ \hline
-            Unsupervised kd-RPCA	&0.62	&0.97	&0.99\\ \hline
-            ROBPCA-AO	&0.00	&0.00	&0.00\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian} +\\\textbf{Uniform} \textbf{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}\\ \hline
+            Proposed Semi-Supervised kd-RPCA    &0.96  &1.00  &0.98  \\ \hline
+            Proposed Semi-Supervised sd-RPCA    &0.97  &0.99  &0.96  \\ \hline
+            Proposed Contaminated Semi-Supervised sd-RPCA   &0.97  &0.91  &0.95  \\ \hline
+            Proposed Unsupervised sd-RPCA &0.97  &0.88  &0.91  \\ \hline
+            Proposed Contaminated Semi-Supervised kd-RPCA   &0.63  &1.00  &0.97  \\ \hline
+            Proposed Unsupervised kd-RPCA &0.62  &0.97  &0.99  \\ \hline
+            Proposed Semi-Supervised md-RPCA    &0.97  &0.85  &0.49  \\ \hline
+            PCA \cite{shyu2003novel}      &0.91  &0.75  &0.56  \\ \hline
+            MCD \cite{rousseeuw1999fastmcd}     &0.97  &0.79  &0.00  \\ \hline
+            Proposed Contaminated Semi-Supervised md-RPCA   &0.97  &0.54  &0.17  \\ \hline
+            Proposed Unsupervised md-RPCA &0.97  &0.53  &0.14  \\ \hline
+            Isolation Forest \cite{liu2008isolation}  &0.97  &0.46  &0.05  \\ \hline
+            One-class SVM \cite{scholkopf2001estimating}    &0.97  &0.11  &0.00  \\ \hline
+            KNN \cite{angiulli2002fast}   &0.97  &0.05  &0.00  \\ \hline
+            LOF \cite{breunig2000lof}     &0.27  &0.33  &0.30  \\ \hline
+            ROBPCA-AO \cite{hubert2009robustskewed}   &0.00  &0.00  &0.00  \\ \hline
 	\end{tabular}
@@
 
-The Table \ref{tab:4.03} also shows higher scores of semi-supervised m-RPCA for 25\% of contamination, in comparison to the remain algorithms.
-It is important to note that the highest score for anomaly detection on Lognormal data with Gaussian contamination was the unsupervised kd-RPCA, and that the contaminated semi-supervised approaches presented scores near of the the results for semi-supervised approaches.
+The Table \ref{tab:4.03} also shows higher scores of semi-supervised m-RPCA for 25\% of contamination, in comparison to the remain compared algorithms.
+It is important to note that the highest score for anomaly detection on Lognormal data with Gaussian contamination was the unsupervised kd-RPCA, with 0.99 $F_1$ score, and that the contaminated semi-supervised approaches of sd-RPCA and kd-RPCA presented scores near of the results for semi-supervised approaches in scenarios with Gaussian, Pareto and Lognormal distributions.
+
+The contamination rate of 10\% shown by Table \ref{tab:4.04} shows worse results in comparison to contamination of 25\% and 33\%.
+However, the ROBPCA-AO presented high scores for anomaly detection on Gaussian data, overcoming LOF, css\_kd-RPCA and u\_kd-RPCA.
+ROBPCA also presented better results for anomaly detection on Pareto data in comparison to u\_md-RPCA, css\_md-RCAP, ss\_md-RCAP, PCA, OCSVM, LOF, KNN and IF.
 
@@
 	\begin{tabular}{| l | l | l | l |}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\pmb{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Gaussian + Uniform}\\\pmb{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\pmb{Pareto + Gaussian}\\\pmb{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\pmb{Lognormal + Gaussian}\\\pmb{($F_1$)}\end{tabular}\\ \hline
-            Isolation Forest	&0.97	&0.43	&0.00\\ \hline
-            KNN	&0.97	&0.01	&0.00\\ \hline
-            LOF	&0.29	&0.27	&0.24\\ \hline
-            MCD	&0.97	&0.40	&0.00\\ \hline
-            One-class SVM	&0.97	&0.26	&0.00\\ \hline
-            PCA	&0.92	&0.65	&0.06\\ \hline
-            Semi-Supervised md-RPCA	&0.97	&0.59	&0.00\\ \hline
-            Semi-Supervised sd-RPCA	&0.97	&0.98	&0.84\\ \hline
-            Semi-Supervised kd-RPCA	&0.97	&1.00	&0.95\\ \hline
-            Contaminated Semi-Supervised md-RPCA	&0.97	&0.28	&0.00\\ \hline
-            Contaminated Semi-Supervised sd-RPCA	&0.96	&0.90	&0.82\\ \hline
-            Contaminated Semi-Supervised kd-RPCA	&0.48	&1.00	&0.94\\ \hline
-            Unsupervised md-RPCA	&0.97	&0.31	&0.00\\ \hline
-            Unsupervised sd-RPCA	&0.97	&0.93	&0.70\\ \hline
-            Unsupervised kd-RPCA	&0.50	&0.97	&0.98\\ \hline
-            ROBPCA-AO	&0.97	&0.87	&0.00\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian} +\\\textbf{Uniform} \textbf{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}\\ \hline
+            Proposed Semi-Supervised kd-RPCA    &0.97  &1.00  &0.95  \\ \hline
+            Proposed Semi-Supervised sd-RPCA    &0.97  &0.98  &0.84  \\ \hline
+            Proposed Contaminated Semi-Supervised sd-RPCA   &0.96  &0.90  &0.82  \\ \hline
+            Proposed Unsupervised sd-RPCA &0.97  &0.93  &0.70  \\ \hline
+            Proposed Unsupervised kd-RPCA &0.50  &0.97  &0.98  \\ \hline
+            Proposed Contaminated Semi-Supervised kd-RPCA   &0.48  &1.00  &0.94  \\ \hline
+            ROBPCA-AO \cite{hubert2009robustskewed}   &0.97  &0.87  &0.00  \\ \hline
+            PCA \cite{shyu2003novel}      &0.92  &0.65  &0.06  \\ \hline
+            Proposed Semi-Supervised md-RPCA    &0.97  &0.59  &0.00  \\ \hline
+            Isolation Forest \cite{liu2008isolation}  &0.97  &0.43  &0.00  \\ \hline
+            MCD \cite{rousseeuw1999fastmcd}     &0.97  &0.40  &0.00  \\ \hline
+            Proposed Unsupervised md-RPCA &0.97  &0.31  &0.00  \\ \hline
+            Proposed Contaminated Semi-Supervised md-RPCA   &0.97  &0.28  &0.00  \\ \hline
+            One-class SVM \cite{scholkopf2001estimating}    &0.97  &0.26  &0.00  \\ \hline
+            KNN \cite{angiulli2002fast}   &0.97  &0.01  &0.00  \\ \hline
+            LOF \cite{breunig2000lof}     &0.29  &0.27  &0.24  \\ \hline
 	\end{tabular}
@@
 
-The contamination rate of 10\% shown by Table \ref{tab:4.03} shows worse results in comparison to contamination of 25\% and 33\%.
-However, the ROBPCA-AO presented high scores for anomaly detection on Gaussian data, overcoming LOF, css\_kd-RPCA and u\_kd-RPCA.
-ROBPCA also presented better results for anomaly detection on Pareto data in comparison to u\_md-RPCA, css\_md-RCAP, ss\_md-RCAP, PCA, OCSVM, LOF, KNN and IF.
-
-Taking into account the null hypothesis $H_{0rob.sub}$ and the presented results, we can conclude that the robust subspace learning, adopted by md-RPCA, sd-RPCA and kd-RPCA, presented higher anomaly detection from imbalanced and skewed data than widely adopted algorithms for anomaly detection.
-Therefore, the presented results refute the null hypothesis $H_{0rob.sub}$, which defines that the robust subspace learning does not improves the anomaly detection from imbalanced and skewed data.
+Taking into account the null hypothesis $H_2^{(N)}$ and the presented results, we can conclude that the robust subspace learning, adopted by md-RPCA, sd-RPCA and kd-RPCA, presented higher anomaly detection from imbalanced and skewed data than widely adopted algorithms for anomaly detection.
+Therefore, the presented results refute the null hypothesis $H_2^{(N)}$, which defines that the robust subspace learning does not improves the anomaly detection from imbalanced and skewed data.
 
 \section{Results of CTU-13 Experiment}
-\label{sec:4_ctu13_result}
+\label{sec:3_ctu13_result}
 
@@
 The CTU-13 data set is very imbalanced, with contamination rate between 0.06\% and 8.86\%.
-Therefore we adopted a uniform contamination rate of 33\% for this experiment, considering that the simulated experiment showed better results for contamination higher than 30\% for m-RPCA and for the selected anomaly detection algorithms.
+Therefore we adopted an uniform contamination rate of 33\% for this experiment, considering that the simulated experiment showed better results for contamination higher than 30\% for m-RPCA and for the selected anomaly detection algorithms.
 
@@
 	\toprule
-        \pmb{Algorithm}	&\pmb{10}	&\pmb{11}	&\pmb{12}	&\pmb{15}	&\pmb{15-2}	&\pmb{15-3}	&\pmb{16}	&\pmb{16-2}	&\pmb{16-3}	&\pmb{17}	&\pmb{18}	&\pmb{18-2}	&\pmb{19}	&\pmb{Avg} \\ \hline
-        IF	&0.36	&0.34	&0.09	&0.21	&0.40	&0.44	&0.16	&0.34	&0.34	&0.41	&0.12	&0.16	&0.46	&0.29\\ \hline
-        KNN	&0.05	&0.17	&0.01	&0.03	&0.33	&0.23	&0.01	&0.25	&0.03	&0.12	&0.00	&0.00	&0.24	&0.11\\ \hline
-        LOF	&0.15	&0.14	&0.13	&0.17	&0.29	&0.22	&0.29	&0.38	&0.25	&0.24	&0.00	&0.04	&0.38	&0.21\\ \hline
-        MCD	&0.18	&0.29	&0.09	&0.34	&0.79	&0.62	&0.04	&0.58	&0.20	&0.41	&0.20	&0.20	&0.36	&0.33\\ \hline
-        PCA	&0.33	&0.64	&0.69	&0.65	&0.55	&0.62	&0.75	&0.50	&0.77	&0.33	&0.82	&0.01	&0.61	&0.56\\ \hline
-        md-RPCA &0.83	&0.79	&0.95	&0.78	&0.78	&0.87	&0.95	&0.87	&0.80	&0.82	&0.83	&0.82	&0.51	&0.81\\ \hline
-        sd-RPCA &0.25	&0.75	&0.34	&0.64	&0.50	&0.75	&0.86	&0.50	&0.77	&0.33	&0.82	&0.81	&0.21	&0.57\\ \hline
-        kd-RPCA &0.76	&0.76	&0.90	&0.82	&0.57	&0.76	&0.91	&0.50	&0.80	&0.73	&0.83	&0.81	&0.48	&0.74\\ \hline
-        ROBPCA-AO &0.01	&0.07	&0.00	&0.05	&0.38	&0.11	&0.05	&0.32	&0.03	&0.09	&0.07	&0.09	&0.21	&0.11\\ \hline
+        \textbf{Algorithm}	&\textbf{10}	&\textbf{11}	&\textbf{12}	&\textbf{15}	&\textbf{15-2}	&\textbf{15-3}	&\textbf{16}	&\textbf{16-2}	&\textbf{16-3}	&\textbf{17}	&\textbf{18}	&\textbf{18-2}	&\textbf{19}	&\textbf{Avg} \\ \hline
+        md-RPCA     &0.83  &0.79  &0.95  &0.78  &0.78  &0.87  &0.95  &0.87  &0.80  &0.82  &0.83  &0.82  &0.51  &0.81  \\ \hline
+        kd-RPCA     &0.76  &0.76  &0.90  &0.82  &0.57  &0.76  &0.91  &0.50  &0.80  &0.73  &0.83  &0.81  &0.48  &0.74  \\ \hline
+        sd-RPCA     &0.25  &0.75  &0.34  &0.64  &0.50  &0.75  &0.86  &0.50  &0.77  &0.33  &0.82  &0.81  &0.21  &0.57  \\ \hline
+        PCA   &0.33  &0.64  &0.69  &0.65  &0.55  &0.62  &0.75  &0.50  &0.77  &0.33  &0.82  &0.01  &0.61  &0.56  \\ \hline
+        MCD   &0.18  &0.29  &0.09  &0.34  &0.79  &0.62  &0.04  &0.58  &0.20  &0.41  &0.20  &0.20  &0.36  &0.33  \\ \hline
+        IF    &0.36  &0.34  &0.09  &0.21  &0.40  &0.44  &0.16  &0.34  &0.34  &0.41  &0.12  &0.16  &0.46  &0.29  \\ \hline
+        LOF   &0.15  &0.14  &0.13  &0.17  &0.29  &0.22  &0.29  &0.38  &0.25  &0.24  &0.00  &0.04  &0.38  &0.21  \\ \hline
+        KNN   &0.05  &0.17  &0.01  &0.03  &0.33  &0.23  &0.01  &0.25  &0.03  &0.12  &0.00  &0.00  &0.24  &0.11  \\ \hline
+        ROBPCA-AO   &0.01  &0.07  &0.00  &0.05  &0.38  &0.11  &0.05  &0.32  &0.03  &0.09  &0.07  &0.09  &0.21  &0.11  \\ \hline
     \bottomrule
@@
 
-It is possible to observe that md-RPCA, kd-RPCA and sd-RPCA overcome the remain algorithms in average results for all scenarios, according to the column \pmb{avg}.
+It is possible to observe that md-RPCA, kd-RPCA and sd-RPCA overcome the remain algorithms in average results for all scenarios, according to the column \textbf{avg}.
 The sd-PRCA presented an average of 0.57 while md-RPCA obtained 0.81 and kd-RPCA 0.74 in average.
@@
 \section{Conclusion}
-\label{sec:4_conclusion}
+\label{sec:3_conclusion}
 
@@
 The results also show that the m-RPCA can be adopted for network attack detection, with better results than widely adopted algorithms for outlier detection on the CTU-13 data set.
-Therefore, the observed results rejects the null hypothesis  $H_{0rob.sub}$ and confirms the alternative
-hypothesis $H_{1rob.sub}$.
+Therefore, the observed results rejects the null hypothesis $H_1^{(N)}$ and confirms the alternative
+hypothesis $H_2^{(A)}$.
 
-The results shows that the semi-supervised approach for m-RPCA obtained better results than the contaminated semi-supervised and unsupervised approaches, and that m-RPCA presented good results even when the semi-supervised training was computed from contaminated data.
-This highlights the resilience of the robust subspace learning for the semi-supervised anomaly detection approach to deal with possible lack of known normal data for training.
+The results show that the semi-supervised approach for m-RPCA obtained better results than the contaminated semi-supervised and unsupervised approaches, and that m-RPCA presented good results even when the semi-supervised training was computed from contaminated data.
+This highlights the resilience of the robust subspace learning for the semi-supervised anomaly detection approach to deal with possible lack of known legitimate data for training.
 
@@
 Future research can be directed to evaluate the application of the proposed approaches to different data sets and anomaly detection problems.
-The m-RPCA can also be extended to be online and to learn new subspace in a adaptive fashion, by means of new robust subspace learning from new observations or through robust subspace tracking.
+The m-RPCA can also be extended to be online and to learn new subspace in an adaptive fashion, by means of new robust subspace learning from new observations or through robust subspace tracking.
+
+In this work we adopt a multivariate analysis based on subspace learning from matrix decomposition, but a multidimensional approach can be evaluated in order to exploit the relationship between three or mode dimensions.
+Therefore, future works can evaluate the contribution of a robust tensor-based data modeling \cite{da2008robust, gomes2017tensor}, tensor decomposition \cite{gomes2019tensor} and multidimensional analysis for m-RPCA in problems of anomaly detection and network attack detection.
+
+Future works can also consider to evaluate the sparse matrix $\pmb{S}$ and its sparsity \cite{liu2017tensor} for m-RPCA, in order to improve the accuracy and to be able to have element-wise anomaly detection.

### 2020-04-28 docs/thesis/chapters/5_conclusion_future_work/conclusion.tex
updating thesis latex

@@
 
-\begin{quotation}[]{Chico Xavier}
-Though nobody can go back and make a new beginning, anyone can start over and make a new ending.
-\end{quotation}
-
 % falar sobre uma contextualização de forma bem resumida'
-
-In the context of anomaly-based schemes, this thesis proposed a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks.
+In the context of anomaly-based schemes, in thesis  we propose propose a statistical approach based on signal processing techniques for detection of malicious traffic in computer networks.
 The proposed technique is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis, where MOS and eigenvalue analysis are applied to detect time frames under attack.
@@
 
-This thesis also proposed a prove of concept of an architecture to evaluate the user behavior analysis on offline mobile clients through Eigensim and Model Order Selection (MOS), in order to implement an anomaly detection module based on user behavioral analysis.
-
-Finally, we proposed the m-RPCA, which is approaches for anomaly detection on skewed and imbalanced data set, through distances of moments computed from a robust subspace learned by RPCA.
+We propose the m-RPCA, which an approach for anomaly detection on skewed and imbalanced data set, through distances of moments computed from a robust subspace learned by RPCA.
 The m-RPCA was evaluated for anomaly detection on simulated data and on the CTU-13 data set, which is composed by network traffic of botnets, network attacks and background flows.
 
+Finally, it is proposed a proof of concept of an architecture to evaluate the user behavior analysis on offline mobile clients through Eigensimilarity and Model Order Selection (MOS), in order to implement an anomaly detection module based on user behavioral analysis.
+
+This chapter is organized as follows.
+In Section \ref{sec:5_conc_conclusion}, we discuss the conclusion remarks, in Section \ref{sec:5_conc_contributions} it is present the main contributions and in \ref{sec:5_conc_futurework} we propose opportunities for future work.
+
 \section{Conclusion}
-\label{sc:conc_conclusion}
+\label{sec:5_conc_conclusion}
 
@@
 Moreover, it is crucial to obtain high detection accuracy and capacity to detect variations of malicious patterns.
-Recently, signal processing schemes have being applied to detect malicious traffic in computer networks \cite{Lu2009,Huang2009,Zonglin2009,david2011blind,da2012improved,tenorio2013greatest, vieira2017model}, showing advances in network traffic analysis.
+Recently, signal processing schemes have being applied to detect malicious traffic in computer networks, showing advances in network traffic analysis.
 
-This thesis models the network traffic as a signal processing formulation for applying the Eigensim, which is a framework for detection and identification of network attacks, that is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
+This thesis models the network traffic as a signal processing formulation for applying the Eigensimilarity, which is a framework for detection and identification of network attacks, that is based on eigenvalue analysis, model order selection (MOS) and eigen similarity analysis.
 
-The Eigensim was evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and similarity analysis of learned subspace of eigenvalues and eigenvectors.
-The main contributions of Eigensim were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
+The Eigensimilarity was evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and similarity analysis of learned subspace of eigenvalues and eigenvectors.
+The main contributions of Eigensimilarity were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.
 
-This thesis evaluated the effectiveness of MOS schemes for network attack detection, extending our previous work \cite{tenorio2013greatest} and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
+This thesis evaluated the effectiveness of MOS schemes for network attack detection, extending our previous work and showing that the analysis of the largest eigenvalues by time frames can be applied to detect the number of port scanning, and flood attacks, but still requiring more information for detailed attack detection.
 Therefore, we proposed a novel approach for detailed network attack detection, based on eigen similarity analysis.
@@
 Considering the offline mobile security context, an important issue faced by corporations that use cloud-based systems is how to provide security mechanisms to support offline corporate mobile clients.
-Aware of this problem and its importance, We proposed an architecture based on Eigensim for behavioral anomaly detection in offline corporate mobile apps.
-As prove of concept, a fully working mobile application was developed to test the proposed security solution and acquired results provide evidence that besides achieving the desired security features, the solution also has positive results in terms of performance.
+Aware of this problem and its importance, we proposed an architecture based on Eigensimilarity for behavioral anomaly detection in offline corporate mobile apps.
+As proof of concept, a fully working mobile application was developed to test the proposed security solution and acquired results provide evidence that besides achieving the desired security features, the solution also has positive results in terms of performance.
 
-Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem \cite{he2008learning}.
+Anomalies can be hard to identify and separate from legitimate data due to the rare occurrences of anomalies in comparison to legitimate events, therefore anomaly detection algorithms have to be high discriminative, robust to contamination and able to deal with the imbalanced data problem.
 
@@
 \section{Contributions}
-\label{sc:conc_contributions}
+\label{sec:5_conc_contributions}
 
 We analyze problems related to detection of anomalies and information security issues, and propose new approaches to improve malicious behavior detection through signal processing techniques and subspace learning.
-The results of the work presented in this thesis provide the following contributions:
+The results of the work presented in this thesis provided the following contributions:
 
 \begin{enumerate}
-	\item We proposed an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 dataset;
+	\item We proposed an approach based on eigen similarity analysis for extracting detailed information about accurate time and network ports under network attack, and evaluated the accuracy and performance of the proposed framework applied to an experimental scenario and to the DARPA 1998 data set;
 	\item We discussed the computational complexity of the proposed framework and evaluated the required processing time for tested scenarios;
@@
 	\item We discussed the processing time of the proposed framework for mobile devices;
-	\item We proposed a approaches for anomaly detection on skewed and imbalanced data set, through distances of moments computed from a robust subspace learned by RPCA;
-	\item We published the following papers reporting our results:
-	\begin{enumerate}
-		\item T.
-P.
-B.
-Vieira, D.
-F.
-Ten\'orio, J.
-P.
-C.
-L.
-da Costa, E.
-P.
-de Freitas, G.
-Del Galdo, and R.
-T.
-de Sousa J\'unior, \textit{Model Order Selection and Eigen Similarity based Framework for Detection and Identication of Network Attacks}.
-Journal of Networking and Computer Applications (JNCA), Vol 90, Jul 2017, Pages 26–41 \cite{vieira2017model}.
-		\item T.
-Galibus, T.
-P.
-B.
-Vieira, E.
-P.
-de Freitas, R.
-Albuquerque, J.
-P.
-C.
-L.
-da Costa, R.
-T.
-de Sousa Jr, V.
-Krasnoproshin, A.
-Zaleski, H.
-Vissia, and G.
-del Galdo, \textit{Offline Mode for Corporate Mobile Client Security Architecture}.
-Mobile Networks and Applications, Springer, Mar 2017, Pages 1-17 \cite{galibus2017offline}.
-		\item K.
-H.
-C.
-Ramos, R.
-T.
-Sousa Jr, T.
-P.
-B.
-Vieira, and J.
-P.
-C.
-L.
-da Costa, \textit{Discovering Critical Success Factors for Information Technologies Governance through Bibliometric Analysis of Research Publications in This Domain}.
-Information (Yamaguchi), 2016 \cite{ramos2016information}.
-		\item K.
-H.
-C.
-Ramos, T.
-P.
-B.
-Vieira, J.
-P.
-C.
-L.
-da Costa, and R.
-T.
-Sousa Jr., \textit{Análise Multidimensional de Fatores Críticos de Sucesso em Governança de TI na Administração Pública Federal à Luz dos Dados de Controle Externo}.
-in Proc.
-12th International Conference on Management of Technology and Information Systems (CONTECSI), 2015, São Paulo, Brazil \cite{ramos2015}.
-	\end{enumerate}
+	\item We proposed approaches for network anomaly detection on skewed and imbalanced data set, through distances of moments computed from a robust subspace learned by RPCA;
 \end{enumerate}
@@
 \section{Future Work}
-\label{sc:conc_futurework}
+\label{sec:5_conc_futurework}
 
-Because of time constraints imposed on the Ph.D.
-degree, this thesis addresses some problems, but some problems are still open and others are emerging from current results.
-Thus, the following issues should be investigated as future work:
+This thesis addresses some problems, but some problems are still open and others are emerging from current results. Thus, the following issues should be investigated as future work:
 
 \begin{itemize}
-	\item Improvements for obtaining better false positive rates, as well as to make the proposed framework able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets;
-	\item Distributed or parallel processing can also be evaluated to analyze the scalability and processing capacity for monitoring high throughput network traffic;
+	\item Improvements for obtaining better false positive rates, as well as to make the Eigensimilarity able to identify sparse probe attacks or subtle behaviors, such as exfiltration or covert communication, considering the evaluation of a flow-based analysis and novel data sets;
 	\item Evaluate the application of the proposed approaches to different anomaly detection problems, considering cases that are aware to behavioral analysis;
 	\item Enhance the m-RPCA approaches to estimate the contamination or to not rely on previously defined contamination for anomaly detection;
-	\item The m-RPCA can also be extended to be online and to learn new subspace in a adaptive fashion, by means of new robust subspace learning from new observations or through robust subspace tracking.
+	\item The m-RPCA can also be extended to be online and to learn new subspace in an adaptive fashion, by means of new robust subspace learning from new observations or through robust subspace tracking;
+	\item Evaluate improvements given by a multidimensional modeling and adoption of tensor-based approaches for subspace learning, in order to analyze complex patterns that can be revealed by multidimensional analysis and tensor-based decomposition.
 \end{itemize}

### 2020-04-28 docs/thesis/chapters/abstract.tex
updating thesis latex

The cost of all types of cyberattacks is increasing for global organizations. The Whitehouse of the U.S. government estimates that malicious cyber activity cost the U.S. economy between US\$57 billion and US\$109 billion in 2016. Recently, it is possible to observe an increasing in numbers of Denial of Service (DoS), botnets, malicious insider and ransomware attacks.

Accenture consulting argues that 89\% of survey respondents believe breakthrough technologies, like artificial intelligence, machine learning and user behavior analytics, are essential for securing their organizations. To face adversarial models, novel network attacks and counter measures of attackers to avoid detection, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, by means of behavioral analysis, where known anomalies are not necessaries for training models.

Signal processing schemes have been applied to detect malicious traffic in computer networks through unsupervised approaches, showing advances in network traffic analysis, in network attack detection, and in network intrusion detection systems.

Anomalies can be hard to identify and separate from normal data due to the rare occurrences of anomalies in comparison to normal events. The imbalanced data can compromise the performance of most standard learning algorithms, creating bias or unfair weight to learn from the majority class and reducing detection capacity of anomalies that are characterized by the minority class. Therefore, anomaly detection algorithms have to be highly discriminating, robust to corruption and able to deal with the imbalanced data problem.

Some widely adopted algorithms for anomaly detection assume a Gaussian distributed data for legitimate observations, however this assumption may not be observed in network traffic, which is usually characterized by skewed and heavy-tailed distributions.

As a first important contribution, we propose the Eigensimilarity, which is an approach based on signal processing concepts applied to detection of malicious traffic in computer networks. We evaluate the accuracy and performance of the proposed framework applied to a simulated scenario and to the DARPA 1998 data set. The performed experiments show that synflood, fraggle and port scan attacks can be detected accurately by Eigensimilarity and with great detail, in an automatic and blind fashion, i.e. in an unsupervised approach.

Considering that the skewness improves anomaly detection in imbalanced and skewed data, such as network traffic, we propose the Moment-based Robust Principal Component Analysis (m-RPCA) for network attack detection. The m-RPCA is a framework based on distances between contaminated observations and moments computed from a robust subspace learned by Robust Principal Component Analysis (RPCA), in order to detect anomalies from skewed data and network traffic. We evaluate the accuracy of the m-RPCA for anomaly detection on simulated data sets, with skewed and heavy-tailed distributions, and for the CTU-13 data set. The Experimental evaluation compares our proposal to widely adopted algorithms for anomaly detection and shows that the distance between robust estimates and contaminated observations can improve the anomaly detection on skewed data and the network attack detection.

Moreover, we propose an architecture and approach to evaluate a proof of concept of Eigensimilarity for malicious behavior detection on mobile applications, in order to detect possible threats in offline corporate mobile client. We propose scenarios, features and approaches for threat analysis by means of Eigensimilarity, and evaluate the processing time required for Eigensimilarity execution in mobile devices.

\begin{keywords}
Anomaly Detection, Network Attack Detection, Imbalanced Data, Principal Component Analysis (PCA), Eigenvector Similarity, Robust Principal Component Analysis (RPCA).
\end{keywords}

### 2020-04-28 docs/thesis/chapters/appendix/a_pca_svm_cfa/pca_svm_cfa.tex
updating thesis latex

--- 2017-11-15 docs/thesis/chapters/appendix/b_pca_svm_cfa/pca_svm_cfa.tex
@@
 The identification of the most significant information for one problem is referred to as feature selection by the signal processing and data mining areas, as well as it can be formulated as a principal component problem, which is a widely adopted signal processing technique for data visualization and feature extraction.
-Feature selection aims to select a subset of relevant information from a larger dataset, in order to improve: data visualization and data understanding, storage requirements, dimensionality, processing time, discriminative sensing, and to overcome overfitting problems to improve prediction and classification performance \cite{chandrashekar2014survey}.
+Feature selection aims to select a subset of relevant information from a larger data set, in order to improve: data visualization and data understanding, storage requirements, dimensionality, processing time, discriminative sensing, and to overcome overfitting problems to improve prediction and classification performance \cite{chandrashekar2014survey}.
 
@@
 
-We propose a critical factors analysis based on Principal Component Analysis (PCA) for visual discriminant analysis and based on RFE combined with Support Vector Machine (SVM) \cite{hearst1998support}, applied to the survey that evaluates the IT governance of Brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
-Results show how PCA can make the data discriminative and that SVM is the classifier that best performs and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of Brazilian public sector.
-Finally, SVM is used to highlight the more significant features identified by RFE, which are similar to CSFs previously identified by a qualitative analysis of the same dataset.
+We propose a critical factors analysis based on Principal Component Analysis (PCA) for visual analysis, and based on RFE combined with Support Vector Machine (SVM) \cite{hearst1998support} for classification and identification of critical success factors, applied to the survey that evaluates the IT governance of Brazilian public organizations, in order to identify the CSF for IT governance of the public sector according to TCU.
+Results show how PCA can help the data visualization and the discriminative visual analysis, and that SVM is the classifier that best performs on classification and obtains an accuracy of 91.42\% to learn and classify according to TCU's IT governance evaluation of Brazilian public sector.
+Finally, SVM is used to highlight the more significant features identified by RFE, which are similar to CSFs previously identified by a qualitative analysis of the same data set.
 
@@
 In Section \ref{sec:b_relatedworks}, related works are discussed.
-Section \ref{sec:b_datamodel} presents the data model and the evaluated datasets.
+Section \ref{sec:b_datamodel} presents the data model and the evaluated data sets.
 Section \ref{sec:b_csf_fs} describes the proposed approach for critical success factors analysis.
@@
 
-Ramos \emph{et al} \cite{ramos2016information} propose an overview regarding the evolution of scientific research on IT Governance critical success factors within the domain of public administration.
+Ramos \emph{et al.} \citeme{ramos2016information, ramos2015multidimensional} propose an overview regarding the evolution of scientific research on IT Governance critical success factors within the domain of public administration.
 By means of bibliometric analysis it was investigated seminal works regarding this theme, considering the characteristic key words found during our analysis.
@@
 
-Guyon \emph{et al} \cite{guyon2002gene} propose a method of gene selection utilizing Support Vector Machine methods based on Recursive Feature Elimination (RFE) and demonstrate that the selected genes yield better classification performance and are biologically relevant to cancer.
+Guyon \emph{et al.} \cite{guyon2002gene} propose a method of gene selection utilizing Support Vector Machine methods based on Recursive Feature Elimination (RFE) and demonstrate that the selected genes yield better classification performance and are biologically relevant to cancer.
 The proposed method eliminates gene redundancy automatically and yields better and more compact gene subsets.
@@
 The Brazilian Federal Court of Accounts (TCU, in Portuguese) surveys data regarding IT practices of Brazilian public organizations in order to audit IT governance.
-The dataset with a consolidate view about the answers for this survey the IT governance index is called iGovTI.
+The data set with a consolidate view about the answers for this survey the IT governance index is called iGovTI.
 The iGovTI is composed by 201 multiple choice questions, used for ranking according to their IT governance, submitted to 349 organizations.
 The TCU computes the IT governance index and classifies the IT governance of each organization.
-Additionally, Ramos \emph{et al} \cite{ramos2016information} classifies each question regarding its relevance for IT governance through a qualitative analysis, and identify the CSFs for selected IT managers regarding IT governance.
+Additionally, Ramos \emph{et al.} \citeme{ramos2016information} classifies each question regarding its relevance for IT governance through a qualitative analysis, and identify the CSFs for selected IT managers regarding IT governance.
 
@@
 
-PCA is a statistical technique commonly used for signal denoising, blind source separation, data compression, data visualization, feature extraction and dimensionality reduction, where a reduced number of features is extracted retaining as much information as possible \cite{jolliffe1986principal}.
+PCA is a statistical technique commonly used for signal denoising, data compression, data visualization, feature extraction and dimensionality reduction, where a reduced number of features is extracted retaining as much information as possible \cite{jolliffe1986principal}.
 It uses an orthogonal transformation to convert a set of correlated variables into a set of linearly uncorrelated variables, where the first principal components have the largest variance.
@@
 Considering that PCA combines attributes and creates new ones, with measurements from all of the original variables, it is hard to identify what original variables are most relevant.
-Therefore, it is still necessary an additional method to reveal what are the CSF for iGovTI.
+Therefore, we do not adopt PCA for classification and identification of CSF, thus it is still necessary an additional method to reveal what are the CSF for iGovTI.
 For this problem, we propose a CSF analysis based on RFE.
@@
 
-Initially, it is necessary to identify one classifier that can identify the iGovTI classification of one organization according to a training dataset.
+Initially, it is necessary to identify one classifier that can identify the iGovTI classification of one organization according to a training data set.
 Therefore, we propose an algorithm evaluation in order to identify which one presents the best accuracy for iGovTI classification.
-The selected algorithm should be used by RFE to identify the CSF, that should be compared to results of the CSF qualitative analysis conducted by Ramos \emph{et al} \cite{ramos2016information}.
+The selected algorithm should be used by RFE to identify the CSF, that should be compared to results of the CSF qualitative analysis conducted by Ramos \emph{et al.} \citeme{ramos2016information}.
 
@@
 Given an external estimator or classifier, that assigns weights to features, RFE is able to select features by recursively considering smaller and smaller sets of features.
-First, the estimator is trained on the initial set of features and the importance of each feature is obtained either through a coefficient attribute or through a feature importances attribute.
+First, the estimator is trained on the initial set of features and the importance of each feature is obtained either through a coefficient attribute or through a feature importance attribute.
 Then, the least important features are pruned from current set of features.
@@
 The variables with the least effect on training errors or largest weights indicates more relevance for a classifier.
-Therefore, we propose to assume that the selected most relevant features are the CSF for iGovTI and also propose to validate this assumption against the results of the CSF qualitative analysis conducted by Ramos \emph{et al} \cite{ramos2016information}.
+Therefore, we propose to assume that the selected most relevant features are the CSF for iGovTI and also propose to validate this assumption against the results of the CSF qualitative analysis conducted by Ramos \emph{et al.} \citeme{ramos2016information}.
 
@@
 
-For the visual discriminant analysis we initially we compute the zero mean and the the covariance matrix $\mathbf{C}$ is computed from $\mathbf{X}$.The results are shown by Figure \ref{fig:ch2_fig1}.
+For the visual discriminant analysis we initially we compute the zero mean and the the covariance matrix $\mathbf{C}$ is computed from $\mathbf{X}$.The results are shown by Figure \ref{fig:b_fig1}.
 
@@
      \centering 
-     \includegraphics[height=10cm, width=10cm]{figures/ch2/raw_igovti_covariance.eps}
+     \includegraphics[height=10cm, width=10cm]{figures/apx_b/raw_igovti_covariance-eps-converted-to.pdf}
      \caption{Covariance matrix of iGovTI questions.}
-     \label{fig:ch2_fig1}
+     \label{fig:b_fig1}
 \end{figure}
 
-Figure \ref{fig:ch2_fig1} presents the covariance matrix of 201 questions of iGovTI, where is possible to observe low covariance for the majority questions and high variance for just a few questions.
+Figure \ref{fig:b_fig1} presents the covariance matrix of 201 questions of iGovTI, where is possible to observe low covariance for the majority questions and high variance for just a few questions.
 
@@
 Considering that the first principal components have the largest variance indicated by their eigenvalues, it is necessary to evaluate the variance distribution of the measured eigenvalues.
-Therefore, we present the Figure \ref{fig:ch2_fig2}, wich shows the variance ECDF of the iGovTI questions and shows that 97\% of the questions have variance lesser than 2, while just 3\% of the questions have variance between 2 and 11.
+Therefore, we present the Figure \ref{fig:b_fig2}, which shows the variance ECDF of the iGovTI questions and shows that 97\% of the questions have variance lesser than 2, while just 3\% of the questions have variance between 2 and 11.
 This result indicates that just a few principal components concentrates the most significant information and motivates the evaluation regarding the relationship between the principal components and the iGovTI classification.
@@
      \centering 
-     \includegraphics[height=8cm, width=11cm]{figures/ch2/raw_variance_ecdf.eps}
+     \includegraphics[height=8cm, width=11cm]{figures/apx_b/raw_variance_ecdf.eps}
      \caption{Empirical CDF of variance.}
-     \label{fig:ch2_fig2}
+     \label{fig:b_fig2}
 \end{figure}
@@
 Finally, we select the two largest eigenvalues and their correspondent components in order to evaluate the relationship between the iGovTI classification for 349 organizations and the two most informative variables.
-The Figure \ref{fig:ch2_fig3} shows the scatter diagram that plots the organizations with higher iGovTI with larger circumferences and colors near of red, while organizations with lower iGovTI have colors near of blue and shorter circumferences.
+The Figure \ref{fig:b_fig3} shows the scatter diagram that plots the organizations with higher iGovTI with larger circumferences and colors near of red, while organizations with lower iGovTI have colors near of blue and shorter circumferences.
 It is possible to observe that organizations with higher iGovTI are concentrated in the top left quadrant, with a visual separation of organizations with lower iGovTI, but without a clear distance separating the classes.
@@
      \centering 
-     \includegraphics[height=8cm, width=11cm]{figures/ch2/raw_igovti_ranking_pc2.eps}
+     \includegraphics[height=8cm, width=11cm]{figures/apx_b/raw_igovti_ranking_pc2.eps}
      \caption{iGovTI ranking from 2 principal components.}
-     \label{fig:ch2_fig3}
+     \label{fig:b_fig3}
 \end{figure}
@@
 We perform an algorithm evaluation in order to identify the classifier with the highest accuracy for iGovTI classification.
-The evaluated dataset is composed of 201 questions (features) and 349 organizations (observations), where each organization is classified according to iGovTI classes, which are initial, intermediate and enhanced.
-We divide the dataset into train and test, with the training data equivalent to 90\% of the whole dataset, while the test has 10 \% of the whole dataset.
+The evaluated data set is composed of 201 questions (features) and 349 organizations (observations), where each organization is classified according to iGovTI classes, which are initial, intermediate and enhanced.
+We divide the data set into train and test, with the training data equivalent to 90\% of the whole data set, while the test has 10 \% of the whole data set.
 
-The Table \ref{tab:ch2_tab1} present the selected classification algorithms and the measured accuracy for iGovTI classification.
-According to results, SVM \cite{hearst1998support} is the algorithm with highest classification accuracy, with 91.42\%, while the second is Elastic Net \cite{zou2005regularization}, with 83.15\% of accuracy.
-The results for iGovTI classification corroborates previous evaluations that highlight the advantages of SVM for classifications from small datasets \cite{guyon2002gene}.
-Additionally, our results indicate that SVM is the best classifier for RFE, which corroborates the previous research of Guyon \emph{et al} \cite{guyon2002gene}, where is proposed SVM-RFE, that is a method of gene selection utilizing SVM methods based on RFE.
+The Table \ref{tab:b_tab1} present the selected classification algorithms and the measured accuracy for iGovTI classification.
+According to results, SVM \cite{hearst1998support, reis2016esprit} is the algorithm with highest classification accuracy, with 91.42\%, while the second is Elastic Net \cite{zou2005regularization}, with 83.15\% of accuracy.
+The results for iGovTI classification corroborates previous evaluations that highlight the advantages of SVM for classifications from small data sets \cite{guyon2002gene}.
+Additionally, our results indicate that SVM is the best classifier for RFE, which corroborates the previous research of Guyon \emph{et al.} \cite{guyon2002gene}, where is proposed SVM-RFE, that is a method of gene selection utilizing SVM methods based on RFE.
 
@@
 	\caption{Evaluation of classification accuracy for iGovTI}
-  	\label{tab:ch2_tab1}
+  	\label{tab:b_tab1}
 	\centering
@@
 The next step is to use the selected algorithm as validator for recursive feature elimination in order to identify the CSF for iGovTI.
-Additionally, the RFE algorithm requires the number of desired most significants features.
-We adopt 54 for this variable, considering that is the number of CSF identified by Ramos \emph{et al} \cite{ramos2016information} in a qualitative analysis, that is the proposed target to validate the results of the CSF identification for iGovTI.
+Additionally, the RFE algorithm requires the number of desired most significant features.
+We adopt 54 for this variable, considering that is the number of CSF identified by Ramos \emph{et al.} \citeme{ramos2016information} in a qualitative analysis, that is the proposed target to validate the results of the CSF identification for iGovTI.
 
-The RFE algorithm presents an ranking according to the significance of each feature, with an accuracy of 69,9\% for CSF classification when compared to the CSF identified by Ramos \emph{et al} \cite{ramos2016information} in a qualitative analysis.
+The RFE algorithm presents an ranking according to the significance of each feature, with an accuracy of 69,9\% for CSF classification when compared to the CSF identified by Ramos \emph{et al.} \citeme{ramos2016information} in a qualitative analysis.
 
@@
 
-Finally, SVM is used to select the more significant features identified by RFE, which are 69,9\% similar to CSFs previously identified by a qualitative analysis of the same dataset.
+Finally, SVM is used to select the more significant features identified by RFE, which are 69,9\% similar to CSFs previously identified by a qualitative analysis of the same data set.

### 2020-04-28 docs/thesis/chapters/resumo.tex
updating thesis latex

@@
-Último passo, por ser um resumo da introduçao, resultados e conclusão.
+% No abstract nao temos espaço para muitos detalhes.
+Certamente é importante vender as contribuições, mas  o abstract pode ser resumido ou então podemos manter o tamanho, mas não detalhando demais em aspectos estatísticos
+% Esta motivação inicial precisam melhorar. Por favor procure investimentos e informações de ataques e custos de ataques para motivar melhor. 
+% Você fez um levantamento provando isto? Por favor inclua a estatística da quantidade de artigos para cada técnica para podemos afirmar isto. 
+% Por favor colocar algo mais formal. 
+% Melhorar também o texto aqui. Nao parece um texto científico. Está muito informal e a escrita precisa melhorar.
+
+O custo com todos os tipos de ciberataques tem crescido nas organizações.
+A casa branca do goveno norte americano estima que atividades cibernéticas maliciosas custaram em 2016 um valor entre US\$57 bilhões e US\$109 bilhões para a economia norte americana.
+Recentemente, é possível observar um crescimento no número de ataques de negação de serviço, botnets, invasões e ransomware.
+
+A Accenture argumenta que 89\% dos entrevistados em uma pesquisa acreditam que tecnologias como inteligência artificial, aprendizagem de máquina e análise baseada em comportamentos, são essenciais para a segurança das organizações.
+É possível adotar abordagens semi-supervisionada e não-supervisionadas para implementar análises baseadas em comportamentos, que podem ser aplicadas na detecção de anomalias em tráfego de rede, sem a ncessidade de dados de ataques para treinamento.
+
+Esquemas de processamento de sinais têm sido aplicados na detecção de tráfegos maliciosos em redes de computadores, através de abordagens não-supervisionadas que mostram ganhos na detecção de ataques de rede e na detecção e anomalias.
+
+A detecção de anomalias pode ser desafiadora em cenários de dados desbalanceados, que são casos com raras ocorrências de anomalias em comparação com o número de eventos normais.
+O desbalanceamento entre classes pode comprometer o desempenho de algoritmos traficionais de classificação, através de um viés para a classe predominante, motivando o desenvolvimento de algoritmos para detecção de anomalias em dados desbalanceados.
+
+Alguns algoritmos amplamente utilizados na detecção de anomalias assumem que observações legítimas seguem uma distribuição Gaussiana.
+Entretanto, esta suposição pode não ser observada na análise de tráfego de rede, que tem suas variáveis usualmente caracterizadas por distribuições assimétricas
+ou de cauda pesada.
+Desta forma, algoritmos de detecção de anomalias têm atraído pesquisas para se tornarem mais discriminativos em distribuições assimétricas, como também para se tornarem mais robustos à corrupção e capazes de lidar com problemas causados pelo desbalanceamento de dados.
+
+Como uma primeira contribuição, foi proposta a Autosimilaridade (\textit{Eigensimilarity} em inglês), que é uma abordagem baseada em conceitos de processamento de sinais com o objetivo de detectar tráfego malicioso em redes de computadores.
+Foi avaliada a acurácia e o desempenho da abordagem proposta através de cenários simulados e dos dados do DARPA 1998.
+Os experimentos mostram que Autosimilaridade detecta os ataques synflood, fraggle e varredura de portas com precisão, com detalhes e de uma forma automática e cega, i.e.
+em uma abordagem não-supervisionada.
+
+Considerando que a assimetria de distribuições de dados podem melhorar a detecção de anomalias em dados desbalanceados e assimétricos, como no caso de tráfego de rede, foi proposta a Análise Robusta de Componentes Principais baseada em Momentos (ARCP-m), que é uma abordagem baseada em distâncias entre observações contaminadas e momentos calculados a partir subespaços robustos aprendidos através da Análise Robusta de Componentes Principais (ARCP), com o objetivo de detectar anomalias em dados assimétricos e em tráfego de rede.
+
+Foi avaliada a acurácia do ARCP-m para detecção de anomalias em dados simulados, com distribuições assimétricas e de cauda pesada, como também para os dados do CTU-13.
+Os experimentos comparam nossa proposta com algoritmos amplamente utilizados para detecção de anomalias e mostra que a distância entre estimativas robustas e observações contaminadas pode melhorar a detecção de anomalias em dados assimétricos e a detecção de ataques de rede.
+
+Adicionalmente, foi proposta uma arquitetura e abordagem para avaliar uma prova de conceito da Autosimilaridade para a detecção de comportamentos maliciosos em aplicações móveis corporativas.
+Neste sentido, foram propostos cenários, variáveis e abordagem para a análise de ameaças, como também foi avaliado o tempo de processamento necessário para a execução do Autosimilaridade em dispositivos móveis.
 
 \begin{keywords}
-Discriminative Sensing, Signal Processing, Eigen Similarity Analysis, Tensor Based Dictionary Learning, Critical Factor Analysis, Network Attack Detection, Fraud Detection
+Detecção de Anomalias, Detecção de Ataques de Rede, Dados Desbalanceados, Seleção de Ordem do Modelo (SOM), Similaridade de Autovetores, Análise Robusta de Componentes Principais (ARCP). 
 \end{keywords}

### 2020-04-28 docs/thesis/thesis.tex
updating thesis latex

--- 2019-05-27 docs/thesis/thesis.tex
@@
 \usepackage{hyperref}
+\usepackage{titlesec}
+\setcounter{secnumdepth}{4}
+
+\usepackage{multibib}
+\newcites{me}{Bibliography (Own Publications)}
 
@@
 \author{Thiago Pereira de Brito Vieira}
-\adviser{João Paulo Carvalho Lustosa da Costa}
-% \coadviser{?}
+\adviser{Prof. Dr.-Ing. João Paulo C. Lustosa da Costa}
+\coadviser{Prof. Dr. Rafael Timóteo de Sousa Júnior}
 
@@
 
-\begin{dedicatory}Signatures\end{dedicatory}
+% \begin{dedicatory}Signatures\end{dedicatory}
 
-\begin{dedicatory}Dedicatory.\end{dedicatory}
+\begin{dedicatory}Eu dedico esta tese a João Emídio, meu pai, pela alegria que demonstrou ao me ver ingressar neste desafio e pela felicidade que teria tido ao ver os resultados obtidos.\end{dedicatory}
 
 \agradecimentos
-First and foremost, I would like to thank God for giving me the life, strength, knowledge and opportunity to undertake this research study and to ..
+Primeiramente eu gostaria de agradecer a Deus pela vida, saúde e todas oportunidades criadas em minha vida.
 
-(...)
+Agradeço aos meus pais, João e Ana, e ao meu irmão André, por todo o amor, carinho e incentivos para que eu possa sempre buscar crescimento pessoal e profissional, além de sempre me apoiarem nas minhas decisões e se mostrarem preocupados e empenhados em me ajudar a alcançar meus objetivos.
 
-Thank You!!
+Agradeço à Alynne, minha esposa, por todo o amor e paciência durante todo nosso relacionamento, principalmente nestes intensos anos de doutorado, em que foram essenciais suas palavras de apoio nos momentos difíceis e sua descontração para me dar mais engergia e vontade de seguir em frente.
 
-\begin{epigraph}[]{Confucius}
-	Wherever you go, go with all your heart.
-\end{epigraph}
+Agradeço ao meu filho Lucas, pela alegria e energia positiva que sempre me fortaleceu, mesmo nos momentos em que eu não pude estar ao seu lado, em virture da dedicação necessária aos estudos e produção científica.
+
+Agradeço a todos que me orientaram e forneceram algum ensinamento durante o doutorado, em especial ao Prof.
+João Paulo Lustosa pelo acolhimento, apoio, orientações, cobranças e todos os importantes ensinamentos durante estes meses.
+Agradeço ao Prof.
+Rafael Timóteo por todos os ensinamentos e orientações em momentos importantes da minha pesquisa, pelos artigos que produzimos em conjunto, por aceitar ser meu coorientador e por presidir a banca da minha defesa de tese.
+Também agradeço ao Eduardo Kalil, pelo apoio, dúvidas e provocações que me ajudaram a visualizar melhorias importantes para o nosso trabalho.
+
+Agradeço ao Prof.
+Jacob, ao Prof.
+André, ao Prof.
+Willian, ao Prof.
+Flávio e ao Ricardo por aceitarem fazer parte da banca da minha qualificação e de defesa de tese, como também pelas valiosas críticas construtivas e contribuições para a mehoria do meu trabalho.
+
+Agradeço à Agência Nacional de Telecomunicações (Anatel), por permitir e proporcionar mais um aprendizado na minha vida.
+Gostaria de agradecer especialmente a Guilherme Chehab e Maria Lúcia Valadares por autorizarem o desafio de cursar um doutorado em concomitância com as atividades profissionais que desempenho.
+Agradeço a Eder Gualberto, pelo apoio ao longo destes anos desafiadores.
+
+Agradeço a todos os amigos que fiz durante o doutorado, que contribuiram para que estes dias desafiadores fossem mais agradáveis.
+Finalmente, gostaria de agradecer a todos aqueles que colaboraram direta ou indiretamente na realização deste trabalho.
+
+Muito Obrigado!!!
 
@@
 \appendix
-\include{chapters/appendix/a_mos/a_mos}
-\include{chapters/appendix/b_pca_svm_cfa/pca_svm_cfa}
-\include{chapters/appendix/c_tensor_denoising/c_tensor_denoising}
+\include{chapters/appendix/a_pca_svm_cfa/pca_svm_cfa}
 
-\bibliographystyle{natbib}
+\bibliographystyleme{IEEEtran}
+\bibliographyme{references}
+
+\bibliographystyle{IEEEtran}
 \addcontentsline{toc}{chapter}{Bibliography}

### 2020-04-28 docs/thesis_slides/m-mcd_slides.tex
updating thesis_slides latex

%	Name			:: 	sthlm Beamer Theme  HEAVILY based on the hsrmbeamer theme (Benjamin Weiss)
%	Author			:: 	Mark Hendry Olson (mark@hendryolson.com)
%	Created			::	2013-07-31
%	Updated	    	::	[[April]] 04, 2017 at 16:26:39
%	Version			:: 	2.0.2
%	Email			:: 	hendryolson@gmail.com
%	Website			:: 	http://markolson.se
%	Twitter			:: 	markolsonse
%	Instagram		:: 	markolsonse
%
%	License			:: 	This file may be distributed and/or modified under the
%					GNU Public License.
%
%	Description		::	This presentation is a demonstration of the sthlm beamer
%					theme, which is HEAVILY based on the HSRM beamer theme created by Benjamin Weiss
%					(benjamin.weiss@student.hs-rm.de), which can be found on GitHub
%					<https://github.com/hsrmbeamertheme/hsrmbeamertheme>.  It also borrows heavily
%					from the work of Matthias Vogelgesang, (https://bloerg.net) and his Metropolis Mtheme,
%					<https://github.com/matze/mtheme>.
%
%	Theme			::	newPxFont
%	Options			::	progressbar
%					::	sectionpages
%					::	numfooter
%					::	fullfooter
%					::	dovaligncolumns
%					::	protectframetitle
%					::	greybg
%					::	cblock
%					::	minimal
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%
%        LOADING DOCUMENT
%
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=

\documentclass[newPxFont, numfooter, sectionpages]{beamer}
\usepackage[utf8]{inputenc}
\usetheme{sthlm}
\usepackage{pgfplots}
\pgfplotsset{compat=1.9}
\usepackage{cancel}
\usepackage{amsthm,amsmath}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{supertabular}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{color, colortbl}
\usepackage[backend=bibtex, style=numeric, defernumbers=true]{biblatex}
\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
\addbibresource{references.bib}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand*\justify{%
	\fontdimen2\font=0.4em% interword space
	\fontdimen3\font=0.2em% interword stretch
	\fontdimen4\font=0.1em% interword shrink
	\fontdimen7\font=0.1em% extra space
	%\hyphenchar\font=`\-% allowing hyphenation
}

\definecolor{lightgray}{gray}{0.8}
\definecolor{Gray}{gray}{0.9}

\title{On The Subspace Learning for Network AttackDetection}
\subtitle{\small{Ph.D. Thesis Presentation}}
\author{
    \texttt{
        Thiago P. de B. Vieira
    }
    \texttt{
        \newline \textbf{Advisor:} Prof. Dr.-Ing. João Paulo C. L. da Costa
        \newline \textbf{Co-Advisor:} Prof. Dr. Rafael T. de Sousa Júnior
    }
}

\institute{
	\scriptsize{Universidade de Brasília}\\
	\scriptsize{Departamento de Engenharia Elétrica - ENE/FT}\\
	\scriptsize{Programa de Pós-Graduação em Engenharia Elétrica - PPGEE}
}

%document-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\begin{document}

%title-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\maketitle

%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
%	FRAME: Theme Package Requirements
%-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
\begingroup
\setbeamercolor{normal text}{fg=\cnDarkGrey,bg=white}

%outline-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section*{Outline}

\begin{frame}{Outline}
	\scriptsize
	\tableofcontents[hideallsubsections]% For longer presentations use hideallsubsections option
\end{frame}

%introduction-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Introduction}

\begin{frame}[c]{Motivation}
	\begin{itemize}
		\item \textbf{Security remains a top concern} for business and IT leaders \cite{Chuvakin2018};
		\item It is possible to observe an increasing of \textbf{10\%} of Denial of Service (\textbf{DoS}) attack, \textbf{12\% of botnets} between \textbf{2017 and 2018} \cite{BISSELL2019};
		\item \textbf{89\%} of survey respondents believe that artificial \textbf{intelligence, machine learning and user behavior analytics} are essential for \textbf{securing organizations} \cite{Accenture2018};
	\end{itemize}
\end{frame}

\begin{frame}[c]{Motivation}
	\begin{itemize}
	    \item \textbf{Anomaly-based} and \textbf{behavioral-based} solutions have been adopted for \textbf{cyber security} and attracting \textbf{investments};
		\item \textbf{User behavior analytics} have been adopted by \textbf{32\%} of top global companies and \textbf{saving \$1.72 million} \cite{BISSELL2019};
		\item It is possible to adopt \textbf{unsupervised or semi-supervised} approaches for \textbf{network anomaly detection}, by means of \textbf{behavioral analysis};
		\item Widely adopted \textbf{algorithms} for anomaly detection assume a \textbf{Gaussian} distributed data \cite{lakhina2005mining}, however this assumption \textbf{may not be observed in network traffic analysis};
	\end{itemize}
\end{frame}

\begin{frame}[c]{Problem Statement}
    We define two questions:
    \begin{itemize}
    	\item $Q_1$: Can the analysis of \textbf{patterns} from a learned \textbf{subspace} identify and detect \textbf{anomalies} in network traffic?
    	\item $Q_2$: Can the \textbf{robust subspace} learning improve the \textbf{anomaly} detection in \textbf{imbalanced and skewed} data?
    \end{itemize}
\end{frame}

\begin{frame}[c]{Problem Statement}
    \begin{table}[htb]
    	\centering
    	\small
    	\caption{Hypotheses to evaluate the defined questions}
    	\label{tab:hypothesis}
        \begin{tabular}{|p{4cm}|p{4cm}|c|} \hline
            \textbf{Alternative Hypothesis}	&\textbf{Null Hypothesis}	&\textbf{Question} 	\\ \hline
            $H_1^{(A)}$: A \textbf{subspace} learned by eigenvalue decomposition can be used to detect and \textbf{identify network attacks}.
            &$H_1^{(N)}$. A subspace learned by eigenvalue decomposition can \textbf{not} be used to detect and identify network attacks.	&$Q_1$\\ \hline
            $H_2^{(A)}$: An approach based on \textbf{robust subspace} learning \textbf{improves} the \textbf{anomaly detection} from \textbf{imbalanced and skewed} data.	&$H_2^{(N)}$. An approach based on  robust subspace learning does \textbf{not} improves the anomaly detection from imbalanced and skewed data.	&$Q_2$\\ \hline
        \end{tabular}
    \end{table}
\end{frame}

\begin{frame}[c]{Proposals}
    \begin{itemize}
    	\item \textbf{Eigensimilarity}: a framework based on subspace learning for network attack detection, by means of \textbf{eigenvalue analysis, model order selection (MOS) and a similarity analysis between eigenvectors};
    	\item \textbf{Moment-based Robust Principal Component Analysis (m-RPCA)}: a framework based on\textbf{ distances of moments} computed from a \textbf{robust subspace} learned by RPCA, for \textbf{anomaly detection} on \textbf{imbalanced and skewed} data;
    	\item An \textbf{architecture} and \textbf{proof of concept} for \textbf{offline behavioral analysis} of a \textbf{app}.
    \end{itemize}
\end{frame}

\begin{frame}[c]{Proposals}
    \begin{figure}[h!]
    	     \centering
    	     \includegraphics[width=11cm]{figures/ch1/architecture.eps}
    	     \caption{Simplified Architecture for Network Attack Detection.}
    	     \label{fig:2_fig1}
    	\end{figure}
\end{frame}

\begin{frame}[c]{Contributions}
    \begin{itemize}
    	\item T. P. B. Vieira, D. F. Tenório, J. P. C. da Costa, E. P. de Freitas, G. Del Galdo, and R. T.de Sousa Júnior, “\textbf{Model order selection and eigen similarity based framework for detection and identification of network attacks}\textbf{}”, \textit{Journal of Network and ComputerApplications}, vol. 90, pp. 26–41, 2017 \cite{vieira2017model}.
    	\item T. Galibus, T. P. B. Vieira, E. P. de Freitas, R. d. O. Albuquerque, R. T.de Sousa Júnior, V. Krasnoproshin, A. Zaleski, H. Vissia, G. del Galdoet al., “\textbf{Offline mode for corporate mobile client security architecture}”, \textit{Mobile Networks and Applications}, pp. 1–17, 2017 \cite{galibus2017offline}.
    \end{itemize}
\end{frame}

\begin{frame}[c]{Contributions - Appendix}
    \begin{itemize}
    	\item K. H. C. Ramos, T. P. B. Vieira, J. P. C. L. da Costa, and R. T. de Sousa Júnior, “\textbf{Multidimensional analysis of critical success factors for it governance within the Brazilian federal public administration in the Light of External Auditing Data}". \textit{12th International CONTECSI}, 2015
    	\cite{ramos2015}.
    	\item T. P. B. Vieira, J. P. C. L. da Costa, E. S. C. Vilaça, E. S. Gualberto, and R. T.de Sousa Júnior, “\textbf{Moment distances from robust subspace for network attack detection}”, \textit{Journal of Network and Computer Applications}, \textbf{To Appear} \cite{vieira2019moment}.
    \end{itemize}
\end{frame}

%Eigensimilarity-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Eigensimilarity based Framework for Detection and Identification of Network Attacks}

\begin{frame}[c]{Introduction}
	\begin{itemize}
		\item \textbf{DDoS} attacks are continuing to \textbf{rise};
		\item \textbf{Solutions} for monitoring and mitigation of \textbf{DDoS} in \textbf{application layer} start with \textbf{behavioral analytics} and supplement with \textbf{signature-based} detection \cite{Hevesi2019}.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Introduction}
	\begin{itemize}
        \item \textbf{Probe} attacks aim to \textbf{scan} networks to identify \textbf{opportunities};
        \item A \textbf{probe attack} is considered the \textbf{first step} in an attack;
        \item \textbf{Probe} attacks are considered \textbf{serious threats} by \cite{ahmed2016survey} and \cite{moustafa2019holistic}.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Introduction}
    We propose the Eigensimilarity:
	\begin{itemize}
		\item \textbf{MOS} and \textbf{eigenvalue analysis} are applied to detect \textbf{time frames under attack};
		\item \textbf{Eigenvector similarity analysis} is used for extracting detailed time and network ports under attack;
		\item We evaluate the \textbf{computational complexity} and the required \textbf{processing time}.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Related works}
	\begin{itemize}
		\item Callegari \emph{et al} \cite{Zonglin2009} propose a \textbf{PCA-based} method for identifying the traffic flows responsible for an anomaly detected at the aggregate level. However, Callegari \emph{et al} \textbf{focus on flood attack detection, not addressing probe} attack detection, and their approach \textbf{relies on visual analysis}.
		\item Lu and Ghorbani \cite{Lu2009} proposed a \textbf{network anomaly} detection model based on \textbf{network flow, wavelet approximation, and system identification theory}. However, their work requires a \textbf{training} and presents \textbf{limitations} on identification of behaviors without \textbf{significant outliers}.
	\end{itemize}
\end{frame}

% \begin{frame}[c]{Related works}
% 	\begin{itemize}
% 		\item Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA, determines the \textbf{anomaly} according to the variation of \textbf{eigenvector by similarity analysis and over sampling}. We apply \textbf{MOS for attack detection} and \textbf{similarity analysis for identification} of time and ports under attack;
% 	\end{itemize}
% \end{frame}

\begin{frame}{data sets}
	We focus on \textbf{probe} and \textbf{flooding} attacks for two scenarios: 
	\begin{itemize}
		\item A \textbf{synthetic data set} as a signal superposition of legitimate traffic and malicious traffic;
		\item Selected cases of \textbf{DARPA data set} that reproduce probe and flooding attacks.
	\end{itemize}
\end{frame}

\begin{frame}{Data Model}
	Network traffic as a signal superposition matrix:
	\begin{equation}\label{eq:eq01}
		\boldsymbol{X}^{(q)} = \boldsymbol{U}^{(q)} + \boldsymbol{N}^{(q)},
	\end{equation}
	\begin{itemize}
	    \item $\boldsymbol{X}^{(q)} \in \mathbb{R}^{M \times N}$ has \emph{M} rows and \emph{N} columns;
		\item The data is divided into \textbf{$Q$ time frames};
		\item Each \textbf{row} is a \textbf{TCP/UDP port}, and each \textbf{column} represents \textbf{time} bins (minute);
		\item Each element $x_{m,n}^{(q)}$ stands for the \textbf{number of packets for port $m$ at the $n$-th minute}, at the $q$-th time frame.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Synthetic data set}
	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=7cm]{figures/ch2/scenario.eps}
	     \caption{Scenario for \textbf{legitimate, synflood, fraggle and port scan attack traffic}.}
	     \label{fig:2_fig1}
	\end{figure}
\end{frame}

\begin{frame}[c]{Synthetic data set}
	\begin{figure}[h!]
	     \centering
	     \includegraphics[width=9cm]{figures/ch2/user_operations.eps}
	     \caption{Scenario with \textbf{legitimate} traffic.}
	     \label{fig:2_fig1}
	\end{figure}
\end{frame}

\begin{frame}[c]{Synthetic data set}

	\begin{figure}[h!]
	     \centering 
	     \includegraphics[width=9cm]{figures/ch2/synflood_traffic.eps}
	     \caption{A large quantity of \textbf{SYN} requests, in order to cause a DoS.}
	     \label{fig:2_fig5}
	\end{figure}

\end{frame}

\begin{frame}{DARPA data set}
	\begin{itemize}
	    \item \textbf{7 weeks} of sniffed traffic saved into raw TCPDUMP packet data, from \textbf{inside and outside} origins, with \textbf{labeled} attacks;
	    \item We \textbf{select the cases} that reproduce \textbf{probe and flooding} attacks, apply a \textbf{packet level} analysis.
	\end{itemize}
\end{frame}

\begin{frame}{Eigensimilarity}
	\begin{figure}[h!]
		\centering
	     \includegraphics[width=7cm]{figures/ch2/mos_eigen_similarity.eps}
	     \caption{Overview of The Framework for Detection and Identification of Network Attacks.}
	     \label{fig:2_fig80}
	\end{figure}
\end{frame}

% \begin{frame}{Eigensimilarity}
% 	For \textbf{flooding} detection, calculate the \textbf{sample covariance matrix} $\boldsymbol{\hat{R}}_{yy}^{(q)}$ of the \textbf{zero mean samples} given by
% 	\begin{equation}\label{eq:eq02}
% 		\boldsymbol{y}_{m}^{(q)} = \boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}.
% 	\end{equation}
% 	the sample covariance matrix $\boldsymbol{\hat{R}}_{yy}^{(q)}$ can be calculated as follows
% 	\begin{equation}\label{eq:eq03}
% 		\boldsymbol{\hat{R}}_{yy}^{(q)} = \frac{1}{N}\boldsymbol{Y}^{(q)}\boldsymbol{Y}^{(q)^{\rm T}}.
% 	\end{equation}
% \end{frame}

% \begin{frame}{Eigensimilarity}
% 	For \textbf{probing} detection, compute the \textbf{sample covariance} $\boldsymbol{\hat{R}}_{zz}^{(q)}$ whose variables have \textbf{zero mean and unitary standard deviation} as follows
% 	\begin{equation}\label{eq:eq04}
% 		\boldsymbol{z}_{m}^{(q)} = \frac{\boldsymbol{x}_{m}^{(q)} - \bar{\boldsymbol{x}}_{m}^{(q)}}{\boldsymbol{\sigma}_{m}^{(q)}}.
% 	\end{equation}
% 	The sample covariance matrix $\boldsymbol{\hat{R}}_{zz}^{(q)}$ can be calculated via 
% 	\begin{equation}\label{eq:eq05}
% 		\boldsymbol{\hat{R}}_{zz}^{(q)} = \frac{1}{N}\boldsymbol{Z}^{(q)}\boldsymbol{Z}^{(q)^{\rm T}}.
% 	\end{equation}
% \end{frame}

% \begin{frame}{Eigensimilarity}
% 	We refer to $\boldsymbol{\hat{R}}_{yy}$ and $\boldsymbol{\hat{R}}_{zz}$ as a matrix $\boldsymbol{\hat{R}}$
% 	Compute eigenvalue decomposition (EVD) to obtain the vector of eigenvalues $\boldsymbol{e}^{(q)}$ associated with each matrix (time frame $q$), according to (\ref{eq:eq060}).
% 	\begin{equation}\label{eq:eq06}
% 	\boldsymbol{\hat{R}}^{(q)} = \boldsymbol{V}^{(q)}\boldsymbol{\Lambda}^{(q)}\boldsymbol{V}^{(q)^{\rm T}},
% 	\end{equation}
% 	\begin{equation}\label{eq:eq060}
% 	\boldsymbol{e}^{(q)} = \rm diag(\boldsymbol{\Lambda}^{(q)}),
% 	\end{equation}
% 	The eigenvalues should be sorted in descending order, i.e., $\lambda_{1}^{(q)} > \lambda_{2}^{(q)} > \lambda_{3}^{(q)} > ... > \lambda_{m}^{(q)}$.
% \end{frame}

% \begin{frame}{Eigensimilarity}
% 	\begin{equation}\label{eq:eq07}
% 		\boldsymbol{E} =
% 		\begin{bmatrix}
% 			\color{red}\lambda_1^{(1)} & \color{red}\lambda_1^{(2)} & \color{red}\lambda_1^{(3)} & \cdots & \color{red}\lambda_1^{(Q)} \\
% 			\lambda_2^{(1)} & \lambda_2^{(2)} & \lambda_2^{(3)} & \cdots & \lambda_2^{(Q)} \\
% 			\lambda_3^{(1)} & \lambda_3^{(2)} & \lambda_3^{(3)} & \cdots & \lambda_3^{(Q)} \\
% 			\vdots & \vdots & \ddots & \vdots  \\
% 			\lambda_m^{(1)} & \lambda_m^{(2)} & \lambda_m^{(3)} & \cdots & \lambda_m^{(Q)} \\
% 		\end{bmatrix}.
% 	\end{equation}
% 	\begin{equation}\label{eq:eq08}
% 		\boldsymbol{e}_{\rm max} = [ \lambda_1^{(1)}, \lambda_1^{(2)} ... \lambda_1^{(Q)}]
% 	\end{equation}
% \end{frame}

% \begin{frame}{MOS Scheme}
% 	\begin{itemize}
% 		\item $\boldsymbol{e}_{\rm max}$ is sorted in descending order, producing $\sim\boldsymbol{e}_{\rm max}$, that is used as input parameter for MOS schemes;
% 		\item $\hat{d} = \rm{MOS}(\sim\boldsymbol{e}_{\rm max})$.
% 		\item $\hat{d}$ is the estimated \textbf{number of time frames under attack}.
% 	\end{itemize}
% \end{frame}

% \begin{frame}{Eigenvalue Analysis}
% 	\textbf{Eigenvalue analysis} identifies the \textbf{time frames} $q$ under attack.
% 	\begin{figure}[h!]
% 	     \centering 
% 	     \includegraphics[width=8cm]{figures/ch2/alg.png}
% 	     \label{fig:2_fig91}
% 	\end{figure}
% \end{frame}

\begin{frame}{Eigenvector Similarity Analysis}
	\textbf{Eigenvector Similarity} identifies the attacked \textbf{time and ports}.
	\begin{itemize}
	    \item \textbf{Low similarity} between the most significant eigenvectors can highlight anomalies inserted by \textbf{network attacks};
	\end{itemize}
	\begin{equation}
		\label{eq:eq11}
		s_n = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(n)}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(n)}}};
	\end{equation}	
	\begin{itemize}
	    \item $s_n$ denotes the absolute \textbf{similarity} degree of the $n$-th minute;
	    \item $\boldsymbol{v}^{(q)}$ is the \textbf{legitimate} most significant eigenvectors;
	    \item $\boldsymbol{v}_{(n)}$ is the most \textbf{significant eigenvectors} obtained after append the target $n$-th minute for \textbf{testing}.
	\end{itemize}
\end{frame}

\begin{frame}{Eigenvector Similarity Analysis}
	\begin{itemize}
	    \item $\boldsymbol{X}^{(q)}$ is a time frame estimated as \textbf{legitimate};
	    \item Each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ vector of each $n$-th minutes of the estimated $\rm{\boldsymbol{\hat{q}}}_{\rm max}$ time frame shall be individually appended into $\boldsymbol{X}^{(q)}$, as
	\end{itemize}
	\begin{equation}\label{eq:eq12}
		\boldsymbol{X}_{n} = \{\boldsymbol{X}^{(q)} | \boldsymbol{x}^{(\hat{q})}_{(n)}\};
	\end{equation}
	\begin{itemize}
	    \item $\boldsymbol{v}_{(q)}$ is computed from $\boldsymbol{X}_{(q)}$;
	    \item $\boldsymbol{v}_{(n)}$ is computed from $\boldsymbol{X}_{(n)}$;
	    \item $\boldsymbol{s}_{(n)}$ is computed from $\boldsymbol{v}_{(q)}$ and $\boldsymbol{v}_{(n)}$;
	    \item Low $\boldsymbol{s}_{(n)}$ identifies a $n$-th\textbf{ minute under attack}.
	\end{itemize}
\end{frame}

\begin{frame}{Eigenvector Similarity Analysis}
	We propose and evaluate \textbf{three approaches} for eigenvector similarity analysis: 
	\begin{itemize}
		\item \textbf{incremental:} concatenates $\boldsymbol{x}^{(\hat{q})}_{(n)}$ incrementally into $\boldsymbol{X}^{(q)}$;
		\item \textbf{individual:} concatenates each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ individually into $\boldsymbol{X}^{(q)}$;
		\item \textbf{incremental individualized:} concatenates $\boldsymbol{x}^{(\hat{q})}_{(n)}$ incrementally into $\boldsymbol{X}^{(q)}$ until attack detection. Afterwards, concatenates each $\boldsymbol{x}^{(\hat{q})}_{(n)}$ individually into $\boldsymbol{X}^{(q)}$.
	\end{itemize}
\end{frame}

\begin{frame}{Eigenvector Similarity Analysis}
	\begin{figure}[h!]
		\centering
	    \includegraphics[width=11.5cm]{figures/ch2/incremental.eps}
	    \caption{Traffic selection for incremental approach.}
	    \label{fig:2_fig8}
	\end{figure}
\end{frame}

\begin{frame}{Eigenvector Similarity Analysis}
	\begin{figure}[h!]
	     \includegraphics[width=11.5cm]{figures/ch2/individualized.eps}
	     \caption{Traffic selection for individual approach.}
	     \label{fig:2_fig9}
	\end{figure}
\end{frame}

\begin{frame}{Eigenvector Similarity Analysis}
	\begin{figure}[h!]
	     \includegraphics[width=11.5cm]{figures/ch2/incremental_individualized.eps}
	     \caption{Traffic selection for incremental individualized approach.}
	     \label{fig:2_fig2}
	\end{figure}
\end{frame}

\begin{frame}{Eigenvector Similarity Analysis}
	\textbf{Detection of ports under attack}:
	\begin{itemize}
		\item The $\boldsymbol{v}^{(q)}$ last most significant eigenvectors without attack is compared against the $\boldsymbol{v}_{(n)}$ identified as under attack;
		\item Evaluate the cosine similarity of each $m$-th port of all $\boldsymbol{\hat{n}}$ minutes identified as under attack.
	\end{itemize}
	\begin{equation}\label{eq:eq15}
		\left\{
		\begin{array}{@{}ll@{}}
			x_{(m,n)} = x^{(\hat{q})}_{(m,\hat{n})} \\
			\\
			s_{m,\hat{n}} = \frac{\abs{\boldsymbol{v}^{(q)} \cdot \boldsymbol{v}_{(m,\hat{n})}}}{\norm{\boldsymbol{v}^{(q)}}\norm{\boldsymbol{v}_{(m,\hat{n})}}},
		\end{array}\right.
	\end{equation}
	\begin{itemize}
		\item Low $\boldsymbol{s}_{m,\hat{n}}$ identifies a $m$-th \textbf{port under attack} during the $n$-th minute.
	\end{itemize}
\end{frame}

\begin{frame}{Results - synthetic data set}
	\begin{table}[h!]
	  \centering
	  \scriptsize
	  \caption{Largest Eigenvalue related to attacks detection}
	  \label{tab:tab3}
	  \begin{tabular}{ c c c c c }
		\toprule
		\multirow{3}{*}{\textbf{Time Frame} $q$} &\multicolumn{4}{c }{\textbf{Vectors GETV}}\\ 
				\hhline{~----}
			&\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}	 &\textbf{Detection of}\\
			&\textbf{\emph{synflood/fraggle}}	 &\textbf{\emph{synflood}}	 &\textbf{\emph{fraggle}}	 &\textbf{\emph{port scan}}\\
		\midrule
		1 &1887545 &1887545 &1887545 &2,0734 \\
		2 &2341327 &2341327 &2341327 &2,1451 \\
		3 &3213867 &3213867 &3213867 &\color{red}10,0718 \\
		4 &\color{red}133238294 &\color{red}133238294 &731229 &2,1620 \\
		5 &\color{red}92384021611 &6367983 &\color{red}92384021611 &2,4253 \\
		6 &708335 &708335 &708335 &1,7948 \\
	    \bottomrule
	  \end{tabular}
	\end{table}
\end{frame}

\begin{frame}{Results - synthetic data set}
	\begin{table}[h!]
	  \centering
	  \tiny
	  \caption{MOS schemes applied to port scan and flood detection}
	  \label{tab:tab4}
	  \begin{tabular}{ c c c c c c c c }
		\toprule
		\multirow{2}{*}{\textbf{Type of analysis} $q$} &\multicolumn{6}{c}{\textbf{MOS schemes (estimated model order $\hat{d}$)}} &{\textbf{(d)}}\\ 
				\hhline{~------~}
			&\textbf{AIC} &\textbf{MDL} &\textbf{EDC} &\textbf{RADOI} &\textbf{EFT} &\textbf{SURE}\\
		\midrule
		Detection of synflood \\(presence of attack) &2 &1 &\textbf{\color{red}1} &5 &\textbf{\color{red}1} &4 &\textbf{\color{red}1} \\
		Detection of synflood \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
		\midrule
		Detection of fraggle \\(presence of attack) &1 &1 &\textbf{\color{red}1} &5 &\textbf{\color{red}1} &4 &\textbf{\color{red}1} \\
		Detection of fraggle \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
		\midrule
		Detection of port scan \\(presence of attack) &1 &1 &\textbf{\color{red}1} &1 &\textbf{\color{red}1} &9 &\textbf{\color{red}1} \\
		Detection of port scan \\(absence of attack) &0 &0 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} \\
		\midrule
		Detection of synflood/fraggle \\(presence of attack) &2 &2 &\textbf{\color{red}2} &5 &\textbf{\color{red}2} &5 &\textbf{\color{red}2} \\
		Detection of synflood/fraggle \\(absence of attack) &1 &1 &\textbf{\color{red}0} &1 &\textbf{\color{red}0} &3 &\textbf{\color{red}0} \\
	    \bottomrule
	  \end{tabular}
	\end{table}
\end{frame}

\begin{frame}{Results - synthetic data set}
	\begin{table}[h!]
	  \centering
	  \tiny
	  \caption{Eigenvector Similarity Analysis for Port Scan Detection}
	  \label{tab:tab5}
	  \begin{tabular}{ c c c c c c }
		\toprule
		\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
				\hhline{~~---~}
				& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
		\midrule
		3 &1 &0.9946 &0.9946 &0.9946 &no \\
		3 &2 &0.9934 &0.9934 &0.9999 &no \\
		3 &3 &0.9912 &0.9912 &0.9999 &no \\
		3 &4 &0.9888 &0.9888 &0.9999 &no \\
		3 &5 &0.9856 &0.9856 &0.9998 &no \\
		3 &6 &0.9840 &0.9840 &0.9999 &no \\
		3 &7 &0.9824 &0.9824 &1.0000 &no \\
		3 &8 &0.9794 &0.9794 &0.9999 &no \\
		3 &9 &0.9673 &0.9673 &0.9926 &no \\
		3 &10 &0.9674 &0.9674 &0.9997 &no \\
		3 &11 &0.9733 &0.9733 &0.9993 &no \\
		3 &12 &0.9702 &0.9702 &0.9993 &no \\
		3 &13 &0.9677 &0.9677 &0.9999 &no \\
		3 &14 &0.9646 &0.9646 &0.9998 &no \\
		3 &15 &\color{red}0.0216 &\color{red}0.0216 &\color{red}0.0276 &\color{red}yes \\
		3 &16 &0.9621 &\color{red}0.0209 &1.0000 &no \\
		3 &17 &0.9611 &\color{red}0.0199 &0.9998 &no \\
		3 &18 &0.9612 &\color{red}0.0191 &0.9999 &no \\
		3 &19 &0.9613 &\color{red}0.0186 &0.9998 &no \\
		3 &20 &0.9638 &\color{red}0.0190 &1.0000 &no \\
	    \bottomrule
	  \end{tabular}
	\end{table}	
\end{frame}

%\begin{frame}{Results - synthetic data set}
%	\begin{table}[h!]
%	  \centering
%	  \tiny
%	  \caption{Eigenvector Similarity Analysis for Synflood Detection}
%	  \label{tab:tab6}
%	  \begin{tabular}{ c c c c c c }
%		\toprule
%		\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
%				\hhline{~~---~}
%				& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
%		\midrule
%		4 &1 &1.0000 &1.0000 &1.0000 &no \\
%		4 &2 &0.9999 &0.9999 &1.0000 &no \\
%		4 &3 &0.9997 &0.9997 &0.9999 &no \\
%		4 &4 &0.9998 &0.9998 &1.0000 &no \\
%		4 &5 &0.9965 &0.9965 &0.9908 &no \\
%		4 &6 &0.9975 &0.9975 &1.0000 &no \\
%		4 &7 &0.9977 &0.9977 &1.0000 &no \\
%		4 &8 &0.9980 &0.9980 &1.0000 &no \\
%		4 &9 &0.9987 &0.9987 &0.9999 &no \\
%		4 &10 &0.9991 &0.9991 &1.0000 &no \\
%		4 &11 &\color{red}0.0085 &\color{red}0.0085 &\color{red}0.0284 &\color{red}yes \\
%		4 &12 &\color{red}0.0162 &\color{red}0.0120 &\color{red}0.0343 &\color{red}yes \\
%		4 &13 &\color{red}0.0248 &\color{red}0.0158 &\color{red}0.0427 &\color{red}yes \\
%		4 &14 &\color{red}0.1243 &\color{red}0.0185 &\color{red}0.1041 &\color{red}yes \\
%		4 &15 &\color{red}0.0082 &\color{red}0.0162 &\color{red}0.0103 &\color{red}yes \\
%		4 &16 &\color{red}0.0404 &\color{red}0.0070 &\color{red}0.0580 &\color{red}yes \\
%		4 &17 &\color{red}0.0397 &\color{red}0.0007 &\color{red}0.0573 &\color{red}yes \\
%		4 &18 &\color{red}0.0408 &\color{red}0.0042 &\color{red}0.0584 &\color{red}yes \\
%		4 &19 &\color{red}0.0408 &\color{red}0.0079 &\color{red}0.0584 &\color{red}yes \\
%		4 &20 &\color{red}0.0477 &\color{red}0.0092 &\color{red}0.0757 &\color{red}yes \\
%	    \bottomrule
%	  \end{tabular}
%	\end{table}	
%\end{frame}

%\begin{frame}{Results - synthetic data set}
%	\begin{table}[h!]
%	  \centering
%	  \tiny
%	  \caption{Eigenvector Similarity Analysis for Fraggle Detection}
%	  \label{tab:tab7}
%	  \begin{tabular}{ c c c c c c }
%		\toprule
%		\multirow{2}{*}{\textbf{Time Frame} $q$} &\multirow{2}{*}{\textbf{Time} $n$}   &\multicolumn{3}{c}{\textbf{Similarity Analysis}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
%				\hhline{~~---~}
%				& &\textbf{Incremental Individualized} &\textbf{Incremental} &\textbf{Individual}\\
%		\midrule
%		5 &1 &1.0000 &1.0000 &1.0000 &no \\
%		5 &2 &0.9999 &0.9999 &1.0000 &no \\
%		5 &3 &1.0000 &1.0000 &1.0000 &no \\
%		5 &4 &0.9999 &0.9999 &1.0000 &no \\
%		5 &5 &0.9993 &0.9993 &0.9997 &no \\
%		5 &6 &0.9993 &0.9993 &0.9997 &no \\
%		5 &7 &0.9994 &0.9994 &1.0000 &no \\
%		5 &8 &0.9995 &0.9995 &1.0000 &no \\
%		5 &9 &0.9995 &0.9995 &1.0000 &no \\
%		5 &10 &0.9995 &0.9995 &1.0000 &no \\
%		5 &11 &\color{red}0.0031 &\color{red}0.0031 &\color{red}0.0021 &\color{red}yes \\
%		5 &12 &\color{red}0.0019 &\color{red}0.0025 &\color{red}0.0009 &\color{red}yes \\
%		5 &13 &\color{red}0.0030 &\color{red}0.0026 &\color{red}0.0020 &\color{red}yes \\
%		5 &14 &\color{red}0.0030 &\color{red}0.0027 &\color{red}0.0020 &\color{red}yes \\
%		5 &15 &\color{red}0.0030 &\color{red}0.0028 &\color{red}0.0020 &\color{red}yes \\
%		5 &16 &\color{red}0.0012 &\color{red}0.0025 &\color{red}0.0002 &\color{red}yes \\
%		5 &17 &\color{red}0.0030 &\color{red}0.0026 &\color{red}0.0020 &\color{red}yes \\
%		5 &18 &\color{red}0.0030 &\color{red}0.0026 &\color{red}0.0020 &\color{red}yes \\
%		5 &19 &\color{red}0.0030 &\color{red}0.0027 &\color{red}0.0020 &\color{red}yes \\
%		5 &20 &\color{red}0.0069 &\color{red}0.0023 &\color{red}0.0083 &\color{red}yes \\
%	    \bottomrule
%	  \end{tabular}
%	\end{table}
%\end{frame}

%\begin{frame}{Results - synthetic data set}
%	\begin{table}[h!]
%	  \centering
%	  \tiny
%	  \caption{Eigenvector Similarity Analysis for Detection of Ports Under Port Scan Attack (q=3 and n%=15)}
%	  \label{tab:tab8}
%	  \begin{tabular}{ c c c c }
%		\toprule
%		\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
%				\hhline{~--~}
%				&\textbf{Incremental Individualized} &\textbf{Individual}\\
%		\midrule
%		80 &0.9999 &0.9999 &no \\
%		443 &0.9999 &0.9999 &no \\
%		53 &0.9999 &0.9999 &no \\
%		21 &0.9999 &0.9997 &\color{red}yes \\
%		22 &\color{red}0.0298 &0.9997 &\color{red}yes \\
%		23 &\color{red}0.0298 &0.9997 &\color{red}yes \\
%		25 &\color{red}0.0298 &0.9997 &\color{red}yes \\
%		110 &\color{red}0.0298 &0.9997 &\color{red}yes \\
%		143 &\color{red}0.0298 &0.9997 &\color{red}yes \\
%		161 &\color{red}0.0298 &0.9997 &\color{red}yes \\
%		69 &\color{red}0.0298 &0.9997 &\color{red}yes \\
%		123 &\color{red}0.0298 &0.9997 &\color{red}yes \\
%		445 &\color{red}0.0298 &0.9997 &\color{red}yes \\
%		600 &0.9999 &0.9999 &no \\
%		19 &0.9999 &0.9999 &no \\
%		67 &0.9999 &0.9999 &no \\
%		68 &0.9999 &0.9999 &no \\
%	    \bottomrule
%	  \end{tabular}
%	\end{table}
%\end{frame}

%\begin{frame}{Results - synthetic data set}
%	\begin{table}[h!]
%	  \centering
%	  \tiny
%	  \caption{Eigenvector Similarity Analysis for Detection of Ports Under Synflood Attack (q=4 and n=11)}
%	  \label{tab:tab9}
%	  \begin{tabular}{ c c c c }
%		\toprule
%		\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
%				\hhline{~--~}
%				&\textbf{Incremental Individualized} &\textbf{Individual}\\
%		\midrule
%		80 &1.0000 &1.0000 &no \\
%		443 &1.0000 &1.0000 &no \\
%		53 &1.0000 &1.0000 &no \\
%		21 &1.0000 &1.0000 &no \\
%		22 &1.0000 &1.0000 &no \\
%		23 &1.0000 &1.0000 &no \\
%		25 &1.0000 &1.0000 &no \\
%		110 &1.0000 &1.0000 &no \\
%		143 &1.0000 &1.0000 &no \\
%		161 &1.0000 &1.0000 &no \\
%		69 &1.0000 &1.0000 &no \\
%		123 &1.0000 &1.0000 &no \\
%		445 &1.0000 &1.0000 &no \\
%		600 &\color{red}0.0077 &\color{red}0.0427 &\color{red}yes \\
%		19 &1.0000 &1.0000 &no \\
%		67 &1.0000 &1.0000 &no \\
%		68 &1.0000 &1.0000 &no \\
%	    \bottomrule
%	  \end{tabular}
%	\end{table}
%\end{frame}

%\begin{frame}{Results - synthetic data set}
%	\begin{table}[h!]
%	  \centering
%	  \tiny
%	  \caption{Eigenvector Similarity Analysis for Detection of Ports Under Fraggle Attack (q=5 and t=11)}
%	  \label{tab:tab10}
%	  \begin{tabular}{ c c c c }
%		\toprule
%		\multirow{2}{*}{\textbf{Port} $p$}   &\multicolumn{2}{c}{\textbf{Approaches}} &\multirow{2}{*}{\textbf{Attack?}}\\ 
%				\hhline{~--~}
%				&\textbf{Incremental Individualized} &\textbf{Individual}\\
%		\midrule
%		80 &1.0000 &1.0000 &no \\
%		443 &1.0000 &1.0000 &no \\
%		53 &1.0000 &1.0000 &no \\
%		21 &1.0000 &1.0000 &no \\
%		22 &1.0000 &1.0000 &no \\
%		23 &1.0000 &1.0000 &no \\
%		25 &1.0000 &1.0000 &no \\
%		110 &1.0000 &1.0000 &no \\
%		143 &1.0000 &1.0000 &no \\
%		161 &1.0000 &1.0000 &no \\
%		69 &1.0000 &1.0000 &no \\
%		123 &1.0000 &1.0000 &no \\
%		445 &1.0000 &1.0000 &no \\
%		600 &1.0000 &1.0000 &no \\
%		19 &\color{red}0.0031 &\color{red}0.0004 &\color{red}yes \\
%		67 &1.0000 &1.0000 &no \\
%		68 &1.0000 &1.0000 &no \\
%	    \bottomrule
%	  \end{tabular}
%	\end{table}
%\end{frame}

\begin{frame}{Results - synthetic data set}
	For \textbf{time} attack detection:
	\begin{itemize}
		\item \textbf{High similarity} between network traffic \textbf{without attack} (0.9610) and \textbf{low similarity} with \textbf{attack} (0.0276);
		\item The \textbf{incremental} approach produces \textbf{false positive} results.
	\end{itemize}
	For \textbf{port} attack detection:
	\begin{itemize}
		\item The \textbf{incremental individualized} has more \textbf{sensibility} to anomaly detection;
		\item The \textbf{individual} approach \textbf{was not able} to identify low similarity for ports under attack;
	\end{itemize}
\end{frame}

\begin{frame}{Results - synthetic data set}
    Hypothesis evaluation:
	\begin{itemize}
		\item The results answer positively the question $Q_1$ ("\textbf{Can the analysis of patterns from a learned subspace identify and detect anomalies in network traffic?}");
		\item The results rejects the null hypothesis $H_1^{(N)}$ and confirms the alternative hypothesis $H_1^{(A)}$ ("\textbf{A subspace learned by eigenvalue decomposition can be used to detect and identify network attacks}").
	\end{itemize}
\end{frame}

\begin{frame}{Results - DARPA data set}
	\begin{table}[!t]
		\caption{Results of the attack detection evaluation}
		\label{tab:tab12}
		\centering
		\scriptsize
		\begin{tabular}{|c|c|c|c|c|}
			\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Solution\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Attack Type\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Metric\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Result\end{tabular} \\ \hline
			Proposed Work	&Flooding	&True Positive	&100.00 \%\\ \hline
			Proposed Work	&Flooding	&False Positive	&60.00 \%\\ \hline
			Proposed Work	&Flooding	&Misclassification	&50.00 \%\\ \hline
			Proposed Work	&Probe	&True Positive	&76.92 \%\\ \hline
			Proposed Work	&Probe	&False Positive	&18.52 \%\\ \hline
			Proposed Work	&Probe	&Misclassification	&32.73 \%\\ \hline
			Callegari \emph{et al}	&Flooding	&True Positive	&82.00 \%\\ \hline
			Callegari \emph{et al}	&Flooding	&False Positive	&-\\ \hline
			Callegari \emph{et al}	&Flooding	&Misclassification	&-\\ \hline
			Lu and Ghorbani	&Overall	&True Positive	&94.67 \%\\ \hline
			Lu and Ghorbani	&Overall	&False Positive	&-\\ \hline
			Lu and Ghorbani	&Overall	&Misclassification	&-\\ \hline
			Lu and Ghorbani	&Portsweep	&True Positive	&50.00 \%\\ \hline
			Lu and Ghorbani	&Portsweep	&False Positive	&-\\ \hline
			Lu and Ghorbani	&Portsweep	&Misclassification	&-\\ \hline
		\end{tabular}
	\end{table}
\end{frame}

\begin{frame}{Results - DARPA data set}
	% Misclassification is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$;
	\begin{itemize}
		\item \textbf{High FP and misclassification} due to \textbf{massive legitimate traffic} (such as an attack) sometimes;
		\item Callegari \emph{et al} is based on PCA, without training
		\begin{itemize}
			\item \textbf{82 \% while we obtain 100 \%};
			\item \textbf{FP and misclassification was not evaluated}.
		\end{itemize}
		\item Lu and Ghorbani's is based on signal processing and uses DARPA
		\begin{itemize}
			\item 94.67 \% and 50.00 \%, while we obtain \textbf{100 \% and 76.92 \%};
			\item \textbf{FP and misclassification was not evaluated}.
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Performance Evaluation}
	Complexity Analysis:
	\begin{itemize}
		\item The worst-case running time is $O(N^3)$;
		\item The computational complexity of \textbf{EVD is predominant};
		\item \textbf{However}, the approach \textbf{splits the data into time frames} with period time $N$, which makes possible to \textbf{limit the growth of $N$}.
	\end{itemize}
\end{frame}

%\begin{frame}{Performance Evaluation}
%	\begin{table}[!t]
%		\caption{Processing time of the main steps for anomaly detection}
%		\label{tab:11}
%		\centering
%		\tiny
%		\begin{tabular}{|r|r|r|r|r|r|r|}
%			\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Traffic Time\\(hour)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Frame Size\\(min)\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Num. Ports\end{tabular}	& \begin{tabular}[x]{@{}l@{}}1-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}2-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}3-time\\(ms)\end{tabular}	\\ \hline
%			16	& 10	& 17	& 0.7900	& 0.8100	& 0.0650\\ \hline
%			16	& 20	& 17	& 0.5250	& 0.5950	& 0.0100\\ \hline
%			16	& 60	& 17	& 0.9700	& 1.1400	& 0.0250\\ \hline
%			16	& 120	& 17	& 0.6050	& 0.6100	& 0.0050\\ \hline
%			16	& 60	& 34	& 1.2750	& 1.2200	& 0.0050\\ \hline
%			16	& 120	& 34	& 1.1200	& 1.1700	& 0.0050\\ \hline
%			20	& 10	& 17	& 2.7950	& 2.8950	& 1.1000\\ \hline
%			20	& 20	& 17	& 2.0700	& 2.0200	& 0.3500\\ \hline
%			20	& 60	& 17	& 1.0250	& 1.0450	& 0.0650\\ \hline
%			20	& 120	& 17	& 1.0000	& 1.0700	& 0.0350\\ \hline
%			20	& 60	& 34	& 2.9650	& 3.2100	& 0.0400\\ \hline
%			20	& 120	& 34	& 2.9950	& 3.1150	& 0.0200\\ \hline
%			22	& 10	& 17	& 4.7250	& 4.0850	& 1.4600\\ \hline
%			22	& 20	& 17	& 2.3200	& 2.6800	& 0.2450\\ \hline
%			22	& 60	& 17	& 1.0700	& 1.1200	& 0.0300\\ \hline
%			22	& 120	& 17	& 0.9900	& 1.0500	& 0.0250\\ \hline
%			22	& 60	& 34	& 3.0850	& 3.1250	& 0.0650\\ \hline
%			22	& 120	& 34	& 2.8100	& 2.9600	& 0.0250\\ \hline
%		\end{tabular}
%	\end{table}
%\end{frame}

\begin{frame}{Performance Evaluation}
	Processing Time Analysis:
	\begin{itemize}
		\item The processing time \textbf{increases according to traffic time}, but the \textbf{worst} processing time is \textbf{4.7250 milliseconds};
		% \item The processing time \textbf{increases} with the frame size $N$ \textbf{decreasing};
		\item The \textbf{number of ports} $M$ increases the \textbf{processing time}.
	\end{itemize}
\end{frame}

%m-RPCA-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Robust Framework for Detection of Network Attacks in Imbalanced Traffic}

\begin{frame}{Introduction}
	\begin{itemize}
		\item \textbf{37.9\%} of all Internet traffic is from \textbf{bots};
		\item \textbf{Anomalies} in network traffic can be \textbf{hard} to identify from legitimate data, due to the \textbf{rare} occurrences;
		\item \textbf{Network anomaly detection} is usually applied to \textbf{imbalanced} data \cite{Phua2004minority};
	\end{itemize}
\end{frame}

\begin{frame}{Introduction}
	\begin{itemize}
		\item \textbf{Imbalanced} data can \textbf{compromise} the performance of most standard \textbf{learning algorithms};
		\item \textbf{Network traffic} is usually characterized by \textbf{skewed and heavy-tailed} distributions \cite{leon2017probability};
		\item Traditionally adopted \textbf{algorithms for anomaly detection} assume a \textbf{Gaussian} or symmetric distributed data \cite{lakhina2005mining}.
	\end{itemize}
\end{frame}

\begin{frame}{Introduction}
    \textbf{Moment-based Robust Principal Component Analysis (m-RPCA)}:
	\begin{itemize}
		\item A \textbf{robust subspace} is computed from \textbf{legitimate observations}, for estimating \textbf{robust moments};
		\item The \textbf{Mahalanobis distance} between the \textbf{robust moments} and \textbf{new contaminated observations} can estimate \textbf{anomalies} in a \textbf{semi-supervised and unsupervised} fashion;
		\item We \textbf{evaluate} m-RPCA for anomaly detection on \textbf{simulated data sets} and for network attack detection on \textbf{CTU-13 data set} \cite{garcia2014empirical};
		\item We \textbf{compare} our proposal to \textbf{widely adopted algorithms} for anomaly detection and to \textbf{ROBPCA}.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Related works}
	\begin{itemize}
	    \item Hubert \emph{et al.} \cite{hubert2009robustskewed} proposed \textbf{ROBPCA-AO}, which improves ROBPCA for problems with skewed data, by means of an adjusted outlyingness based on robust skewness;
		\item Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on a robust mutual information estimator for feature selection and based on \textbf{RPCA for outlier detection in internet traffic};
		\item Zhou and Paffenroth \cite{zhou2017anomaly} proposed the \textbf{Robust Deep Autoencoders (RDA)}, which inherits the \textbf{non-linear representation} capabilities of \textbf{autoencoders} combined with the \textbf{outlier separation of RPCA}. 
	\end{itemize}
\end{frame}

\begin{frame}[c]{Data Set - Simulated}
    \textbf{We propose to evaluate m-RPCA for anomaly detection on simulated data}:
	\begin{itemize}
		\item Two data sets with \textbf{skewed and heavy tailed distributions}:
		\begin{itemize}
    		\item Pareto and Gaussian anomalies;
    		\item Lognormal and Gaussian anomalies;
		\end{itemize}
		\item Gaussian distributed data set, to evaluate m-RPCA for general purposes:
    		\begin{itemize}
        		\item Gaussian and uniform anomalies.
    		\end{itemize}
	\end{itemize}
\end{frame}

% \begin{frame}[c]{Data Set - Simulated}
%     \begin{figure}[h!]
%     	\centering
%     	\includegraphics[width=9cm]{figures/ch4/4_Xpg2.png}
%     	\caption{Example of Pareto and Gaussian Anomalies}
%     	\label{fig:4.04}
%     \end{figure}
% \end{frame}

% \begin{frame}[c]{Data Set - Simulated}
%     \begin{figure}[h!]
%     	\centering
%     		\includegraphics[width=9cm]{figures/ch4/4_Xlogng2.png}
%     	\caption{Example of Lognormal and Gaussian Anomalies}
%     	\label{fig:4.04}
%     \end{figure}
% \end{frame}

% \begin{frame}[c]{Data Set - Simulated}
%     \begin{figure}[h!]
%     	\centering
%     	\includegraphics[width=9cm]{figures/ch4/4_Xgu2.png}
%     	\caption{Example of Gaussian and Uniform Anomalies}
%     	\label{fig:4.04}
%     \end{figure}
% \end{frame}

\begin{frame}[c]{Data Set - CTU-13}
    \textbf{We propose to evaluate m-RPCA for network attack detection on CTU-13}:
    \begin{itemize}
    	\item \textbf{Botnet traffic} captured from a real network (\textbf{testbed}) of Czech Technical University (CTU);
    	\item Normal, Background, attack or command and control (\textbf{C\&C}) traffic;
        \item The data set is composed by\textbf{13 scenarios}/files:
            \begin{itemize}
                \item Malwares: neris, rbot, virut, menti, sogou, nsys.ay and murlo;	
            	\item Attacks: Click Fraud (CF), Port Scan (PS), Fast Flux (FF), SPAM and DDoS;
            	\item C\&C: IRC, P2P and HTTP.
            \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}[c]{Data Set - CTU-13}
    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=11cm]{figures/ch4/ctu13.png}
    	\caption{CTU-13 data set description}
    	\label{fig:4.04}
    \end{figure}
\end{frame}

% \begin{frame}[c]{Data Set - CTU-13}
%     \begin{figure}[h!]
%     	\centering
%     	\includegraphics[width=9cm]{figures/ch4/raw_distplot_capture20110816_State.png}
%     	\caption{State of scenario 16}
%     	\label{fig:4.04}
%     \end{figure}
% \end{frame}

% \begin{frame}[c]{Data Set - CTU-13}
%     \begin{figure}[h!]
%     	\centering
%     	\includegraphics[width=9cm]{figures/ch4/raw_distplot_capture20110816_Sport.png}
%     	\caption{Source Port of scenario 16}
%     	\label{fig:4.04}
%     \end{figure}
% \end{frame}

\begin{frame}[c]{Data Set - CTU-13}
    Selected features by cross validation: 
    \begin{itemize}
        \item state
        \item destination type of service
        \item destination port
        \item source port
        \item total number of packets
        \item total number of bytes
        \item number of bytes from the source;
    \end{itemize}
\end{frame}

\begin{frame}[c]{m-RPCA}
    Robust Subspace Learning
	\begin{enumerate}
		\item  $\pmb{X}$ is \textbf{decomposed} into the sum of a \textbf{low rank} matrix $\pmb{L} \in \mathbb{R}^{M \times N}$, whose column subspace gives the \textbf{robust principal components without outliers}, and a sparse matrix $\pmb{S} \in \mathbb{R}^{M \times N}$, with \textbf{element-wise outliers}.
		\item RPCA is a well-known method to recover a low-rank matrix $\pmb{L}$ and sparse matrix $\pmb{S}$ from corrupted measurements modeled as $\pmb{X} = \pmb{L} + \pmb{S}$;
	\end{enumerate}	
\end{frame}

\begin{frame}[c]{m-RPCA}
	\begin{enumerate}
		\item According to Wright et al. \cite{wright2009robust}, it is possible to recover a low-rank matrix by solving the following convex optimization problem:
	\end{enumerate}	
    \begin{equation}\label{eq:4.01}
    	(\hat{\pmb{L}}, \hat{\pmb{S}})\leftarrow \min_{\pmb{L},\pmb{S}}\left \| \pmb{L} \right \|_{*} + \lambda \left \| \pmb{S} \right \|_{1}
    \end{equation}
    \begin{center} subject to: $\pmb{X} = \pmb{L} + \pmb{S}$ \end{center}
    \begin{equation}\label{eq:4.02}
        \left \| \pmb{L} \right \|_{*} = \sum_{i} \sigma_{i}(\pmb{L})
    \end{equation}
    \begin{equation}\label{eq:4.03}
        \left \| \pmb{S} \right \|_{1} = \sum_{ij} \left | \pmb{S}_{ij} \right |
    \end{equation}
\end{frame}

%\begin{frame}[c]{m-RPCA}
%    where:
%    \begin{itemize}
%        \item $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix;
%        \item $\lambda$ is a positive weighting parameter, which determines the sparsity of $\pmb{S}$; 
%        \item $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries;
%        \item $\sigma$ denotes the singular values of a matrix. 
%    \end{itemize}
%\end{frame}

%\begin{frame}[c]{m-RPCA}
%    The RPCA with ALM can be formulated as:
%    \begin{equation}\label{eq:4.04}
%    	l(\pmb{L}, \pmb{S}, \pmb{Y}) = \left\|\pmb{L}\right\|_* + \lambda\left\|\pmb{S}\right\|_1 + \langle \pmb{Y}, \pmb{X} - \pmb{L} - \pmb{S}  \rangle + \frac{\mu}{2}\left\|\pmb{X} - \pmb{L} - \pmb{S}\right\|_F^2,
%    \end{equation}
%    where $\pmb{Y}$ is the multiplier of the linear constraint and $\mu$ is the penalty parameter for the violation of the linear constraint \cite{yuan2009sparse}.
%	\begin{itemize}
%		\item An iterative scheme can be presented as:
%	\end{itemize}
%    \begin{equation}\label{eq:4.05}
%    	\left\{
%    		\begin{matrix} 
%    			(\pmb{L}_{k+1}, \pmb{S}_{k+1}) \in \argmin\limits_{\pmb{L,S} \in \mathbb{R}^{m \times n}} \{l(\pmb{L}, \pmb{S}, \pmb{Y}_{k})\}, \\ 
%    			\pmb{Y}_{k+1} = \pmb{Y}_{k} + \mu(\pmb{X} - \pmb{L}_{k} - \pmb{S}_{k}),
%    		\end{matrix}
%    	\right.
%    \end{equation}
%\end{frame}

% \begin{frame}[c]{m-RPCA}
%     \begin{itemize}
% 		\item We adopt R\textbf{PCA with ADM}, which is a improvement of ALM for solving convex programming problem with linear constraints, taking advantage of its high-level separable structure \cite{yuan2009sparse};
% 		\item ADM minimizes $\pmb{L}$ and $\pmb{S}$ iteratively by:
% 	\end{itemize}
%     \begin{equation}\label{eq:4.06}
%     	\left\{\begin{matrix}
%     	\pmb{L}_{k+1} \in \argmin\limits_{\pmb{L} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}, \pmb{S}_{k}, \pmb{Y}_{k})\}\\ 
%     	\pmb{S}_{k+1} \in \argmin\limits_{\pmb{S} \in \mathbb{R}^{m \times n}}\{l(\pmb{L}_{k+1}, \pmb{S}, \pmb{Y}_{k})\}\\ 
%     	\pmb{Y}_{k+1} = \pmb{Y}_{k} + \mu(\pmb{X} - \pmb{L}_{k} - \pmb{S}_{k})
%     	\end{matrix}\right.
%     \end{equation}
% \end{frame}

\begin{frame}[c]{m-RPCA}
    The proposal:
    \begin{itemize}
		\item To learn a \textbf{robust subspace} from the legitimate traffic $\pmb{X}^s$ (semi-supervised) or from contaminated traffic $\pmb{X}^c$ (unsupervised), for computing $\pmb{L}^s$ and $\pmb{S}^s$;
		\item To compute the \textbf{robust moments} (the mean $\pmb{\mu}$, skewness $\pmb{\epsilon}$ and kurtosis $\pmb{\kappa}$) from $\pmb{L}^s$
		\item To compute and evaluate the distance $\pmb{d}$ between contaminated observations $\pmb{X}^c$ and the \textbf{robust moments}.
	\end{itemize}
\end{frame}

\begin{frame}[c]{m-RPCA}
    Moments are a set of statistical parameters to measure a distribution.
	\begin{itemize}
		\item 1st moment: The arithmetic mean;
		\item 2nd moment: The variance;
		\item 3rd moment: The skewness (asymmetry);
		\item 4th moment: kurtosis (excess).
	\end{itemize}

    The general expression for the $p$-th moment about the mean $\mu$ is given by
    \begin{equation}\label{eq:4.09}
    	m_p = \displaystyle\frac{1}{M}\displaystyle\sum_{i = 1}^{M}(x_i - \mu)^p.
    \end{equation}
\end{frame}

%\begin{frame}[c]{m-RPCA}
%    The skewness is calculated by
%    \begin{equation}\label{eq:4.10}
%    	\epsilon = \frac{m_3}{m_2^{\frac{3}{2}}},
%    \end{equation}
%    
%    The kurtosis is given as
%    \begin{equation}\label{eq:4.11}
%    	\kappa = \frac{m_4}{m_2^2}.
%    \end{equation}
%    
%    We compute the mean $\pmb{\mu}$, the skewness $\pmb{\epsilon} \in \mathbb{R}^{1 \times N}$, the kurtosis $\pmb{\kappa} \in \mathbb{R}^{1 \times N}$ and the covariance matrix $\hat{\pmb{\Sigma}}$ from the robust subspace $\pmb{L}^s$.
%\end{frame}

\begin{frame}[c]{m-RPCA}
    The classical Mahalanobis Distance is defined as		
    \begin{equation}\label{eq:4.12}
    	\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \pmb{\mu}) \hat{\pmb{\Sigma}}^{-1}(\pmb{x} - \pmb{\mu})^T}.
    \end{equation}
\end{frame}

\begin{frame}[c]{m-RPCA}
    We adopt the mean $\pmb{\mu}$ and covariance matrix $\hat{\pmb{\Sigma}}$ calculated from $\pmb{L}^s$ learned by RPCA for $\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}})$.

    We propose the Robust-Skewness Mahalanobis Distance $\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}})$:

    \begin{equation}\label{eq:4.13}
    	\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \pmb{\epsilon}) \hat{\pmb{\Sigma}}^{-1}(\pmb{x} - \pmb{\epsilon})^T}.
    \end{equation}

    We propose the Robust-Kurtosis Mahalanobis Distance $\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}})$:

    \begin{equation}\label{eq:4.14}
    	\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}}) = \sqrt{(\pmb{x} - \pmb{\kappa}) \hat{\pmb{\Sigma}}^{-1}(\pmb{x} - \pmb{\kappa})^T}.
    \end{equation}
\end{frame}

\begin{frame}[c]{m-RPCA}
	\begin{itemize}
		\item The\textbf{ contamination rate} $c$ is traditionally adopted by well established \textbf{outlier detection algorithms} \cite{zhao2019pyod};
		\item The contamination defines the number of the largest distances $\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}})$, $\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}})$ or $\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}})$ that shall be classified as anomalous.
	\end{itemize}
\end{frame}

\begin{frame}[c]{m-RPCA}
    \textbf{Mean Distance of Robust Principal Component Analysis (md-RPCA)}:
	\begin{itemize}
		\item $\pmb{d}(\pmb{x}, \pmb{\mu}, \hat{\pmb{\Sigma}})$ from the robust subspace $\pmb{L}^s$ learned from legitimate data $\pmb{X}^s$ or contaminated data $\pmb{X}^c$;
	\end{itemize}

	\textbf{Skewness Distance of Robust Principal Component Analysis (sd-RPCA)}:
	\begin{itemize}
		\item $\pmb{d}(\pmb{x}, \pmb{\epsilon}, \hat{\pmb{\Sigma}})$ from the robust subspace $\pmb{L}^s$ learned from legitimate data $\pmb{X}^s$ or contaminated data $\pmb{X}^c$;
	\end{itemize}

	\textbf{Kurtosis Distance of Robust Principal Component Analysis (kd-RPCA)}:
	\begin{itemize}
		\item $\pmb{d}(\pmb{x}, \pmb{\kappa}, \hat{\pmb{\Sigma}})$ from the robust subspace $\pmb{L}^s$ learned from legitimate data $\pmb{X}^s$ or contaminated data $\pmb{X}^c$.
	\end{itemize}
\end{frame}

%\begin{frame}[c]{m-RPCA}
%	\begin{figure}[h!]
%	     \centering
%	     \includegraphics[width=6.2cm]{figures/ch4/alg2.png}
%	     \label{fig:fig06}
%	\end{figure}
%\end{frame}

\begin{frame}[c]{Experiments}
    \textbf{$F_1$} is the \textbf{preferable} measure for \textbf{imbalanced} data sets \cite{powers2011evaluation,moustafa2019holistic}.

    The $F_1$ is the harmonic mean of precision and recall:
    \begin{equation}\label{eq:4.15}
    	F_1 = 2 \cdot \frac{p_{\rm r} \cdot r_{\rm c}}{p_{\rm r} + r_{\rm c}}.
    \end{equation}

    Precision can be seen as accuracy:
    \begin{equation}\label{eq:4.16}
    	p_r = \frac{t_{\rm p}}{t_{\rm p} + f_{\rm p}}.
    \end{equation}

    Recall calculates proportion of actual positives identified correctly:
    \begin{equation}\label{eq:4.17}
    	r_c = \frac{t_{\rm p}}{t_{\rm p} + f_{\rm n}}.
    \end{equation}
\end{frame}

\begin{frame}[c]{Experiments - Simulated data set}
    For the \textbf{simulated data set}, we evaluate m-RPCA for:

    \textbf{Semi-supervised Scenario}
	\begin{itemize}
		\item $\pmb{L}^s$ and moments from the legitimate $\pmb{X}$;
		\item Anomaly detection for contaminated $\pmb{X}^c$.
	\end{itemize}

    \textbf{Contaminated semi-supervised Scenario}
    \begin{itemize}
		\item $\pmb{L}^s$ and moments from the contaminated legitimate $\pmb{X}^{c'}$;
		\item Anomaly detection for contaminated $\pmb{X}^c$.
	\end{itemize}

    \textbf{Unsupervised Scenario}
    \begin{itemize}
		\item $\pmb{L}^s$ and moments from the contaminated $\pmb{X}^c$;
		\item Anomaly detection for contaminated $\pmb{X}^c$.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Experiments - Simulated data set}
    \textbf{Contamination} rates $c$ between \textbf{1\% and 50\% }for each \textbf{scenario} and for each \textbf{data distribution};

    \textbf{m-RPCA is compared to}:
    \begin{itemize}
        \item Isolation Forest (IF) \cite{liu2008isolation};
        \item k-Nearest Neighbors (KNN) \cite{angiulli2002fast};
        \item Local Outlier Factor (LOF) \cite{breunig2000lof};
        \item Minimum Covariance Deteminant (MCD) \cite{rousseeuw1999fastmcd};
        \item One-Class Support Vector Machines (OCSVM) \cite{scholkopf2001estimating};
        \item PCA \cite{shyu2003novel};
        \item ROBPCA-AO \cite{hubert2009robustskewed}.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Results - Gaussian with uniform anomalies}
    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=8cm]{figures/ch4/gaussian_f1_contamination.eps}
    	\caption{Anomaly detection on Gaussian distributed with uniform anomalies}
    	\label{fig:4.10}
    \end{figure}
\end{frame}

\begin{frame}[c]{Results - Gaussian with uniform anomalies}
	\begin{itemize}
		\item \textbf{LOF and contaminated kd-RPCA} present \textbf{lower $F_1$} than remain algorithms;
		\item \textbf{ROBPCA-AO} present high $F_1$ for lower contamination, but \textbf{decrease};
		\item \textbf{u\_md-RPCA and u\_sd-RPCA} presented \textbf{similar} results to other \textbf{unsupervised algorithms} for outlier detection;
		\item \textbf{Semi-supervised m-RPCA} obtain $F_1$ and \textbf{similar} results to \textbf{IF, KNN, MCD and OCSVM};
		\item \textbf{md-RPCA} approaches present $F_1$ \textbf{similar} to the \textbf{best} results.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Results - Pareto with Gaussian anomalies}
    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=8cm]{figures/ch4/pareto_f1_contamination.eps}
    	\caption{Anomaly detection on Pareto distributed with Gaussian anomalies}
    	\label{fig:4.11}
    \end{figure}
\end{frame}

\begin{frame}[c]{Results - Pareto with Gaussian anomalies}
	\begin{itemize}
		\item \textbf{kd-RPCA and sd-RPCA} obtained more than \textbf{0.8} $F_1$;
		\item The \textbf{contaminated semi-supervised of sd-RPCA and kd-RPCA} presented results \textbf{similar} to the \textbf{semi-supervised and unsupervised} approaches;
		\item The \textbf{unsupervised sd-RPCA and kd-RPCA} presented \textbf{better} results than widely adopted \textbf{unsupervised algorithms for outlier detection};
		\item The \textbf{semi-supervised kd-RPCA and sd-RPCA} obtained the \textbf{best} results.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Results - Lognormal with Gaussian anomalies}
    \begin{figure}[h!]
    	\centering
    	\includegraphics[width=8cm]{figures/ch4/lognormal_f1_contamination.eps}
    	\caption{Anomaly detection on Lognormal distributed with Gaussian anomalies}
    	\label{fig:4.12}
    \end{figure}
\end{frame}

\begin{frame}[c]{Results - Lognormal with Gaussian anomalies}
	\begin{itemize}
		\item \textbf{mean-based} approaches of \textbf{m-RPCA} presented \textbf{low} $F_1$;
		\item IF, KNN, LOF, MCD, OCSVM, PCA and ROBPCA were \textbf{worse than sd-RPCA and kd-RPCA};
		\item kd-RPCA performed better than sd-RPCA, with $F_1$ higher than 0.9.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Results - Simulated data set}
    Conclusion:
    \begin{itemize}
        \item We can observe that mean-based approaches of m-RPCA are competitive for gaussian distributed data;
        \item The \textbf{robust subspace learning} of by md-RPCA, sd-RPCA and kd-RPCA, presented \textbf{higher anomaly detection} from \textbf{imbalanced and skewed} data than \textbf{widely adopted algorithms}. 
        \item The results refute the null hypothesis $H_2^{(N)}$ (\textbf{"An approach based on robust subspace learning does not improves the anomaly detection from imbalanced and skewed data"}).
	\end{itemize}
\end{frame}

\begin{frame}[c]{Experiments - CTU-13 data set}
    \textbf{Semi-supervised} approach of md-RPCA, sd-RPCA and kd-RPCA.

    Uniform \textbf{contamination} rate of \textbf{33\%}.

    m-RPCA is compared to:
    \begin{itemize}
        \item Isolation Forest (IF) \cite{liu2008isolation};
        \item k-Nearest Neighbors (KNN) \cite{angiulli2002fast};
        \item Local Outlier Factor (LOF) \cite{breunig2000lof};
        \item Minimum Covariance Deteminant (MCD) \cite{rousseeuw1999fastmcd};
        \item One-Class Support Vector Machines (OCSVM) \cite{scholkopf2001estimating};
        \item PCA \cite{shyu2003novel};
        \item ROBPCA-AO \cite{hubert2009robustskewed}.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Results - CTU-13 data set}
    \begin{table}[h!]
      \centering
      \tiny
    %   \caption{Network anomaly detection from CTU-13 with 33\% of contamination}
      \label{tab:4.05}
      \begin{tabular}{ c|c|c|c|c|c|c|c }
    	\toprule
            \textbf{Algorithm}	&\textbf{10}	&\textbf{11}	&\textbf{12}	&\textbf{15}	&\textbf{15-2}	&\textbf{15-3}	&\textbf{16}	\\ \hline
            IF	&0.36	&0.34	&0.09	&0.21	&0.40	&0.44	&0.16\\ \hline
            KNN	&0.05	&0.17	&0.01	&0.03	&0.33	&0.23	&0.01\\ \hline
            LOF	&0.15	&0.14	&0.13	&0.17	&0.29	&0.22	&0.29\\ \hline
            MCD	&0.18	&0.29	&0.09	&0.34	&\color{red}\textbf{0.79}	&0.62	&0.04\\ \hline
            PCA	&0.33	&0.64	&0.69	&0.65	&0.55	&0.62	&0.75\\ \hline
            \color{red}\textbf{md-RPCA} &\color{red}\textbf{0.83}	&\color{red}\textbf{0.79}	&\color{red}\textbf{0.95}	&\color{blue}0.78	&\color{blue}0.78	&\color{red}\textbf{0.87}	&\color{red}\textbf{0.95}\\ \hline
            sd-RPCA &0.25	&0.75	&0.34	&0.64	&0.50	&0.75	&0.86\\ \hline
            kd-RPCA &0.76	&0.76	&0.90	&\color{red}\textbf{0.82}	&0.57	&0.76	&0.91\\ \hline
            ROBPCA-AO &0.01	&0.07	&0.00	&0.05	&0.38	&0.11	&0.05\\ \hline
        \bottomrule
      \end{tabular}
    \end{table}

        \begin{table}[h!]
      \centering
      \tiny
    %   \caption{Network anomaly detection from CTU-13 with 33\% of contamination}
      \label{tab:4.05}
      \begin{tabular}{ c|c|c|c|c|c|c|c }
    	\toprule
            \textbf{Algorithm}	&\textbf{16-2}	&\textbf{16-3}	&\textbf{17}	&\textbf{18}	&\textbf{18-2}	&\textbf{19}	&\textbf{Avg} \\ \hline
            IF	&0.34	&0.34	&0.41	&0.12	&0.16	&0.46	&0.29\\ \hline
            KNN	&0.25	&0.03	&0.12	&0.00	&0.00	&0.24	&0.11\\ \hline
            LOF	&0.38	&0.25	&0.24	&0.00	&0.04	&0.38	&0.21\\ \hline
            MCD	&0.58	&0.20	&0.41	&0.20	&0.20	&0.36	&0.33\\ \hline
            PCA	&0.50	&0.77	&0.33	&0.82	&0.01	&\color{red}\textbf{0.61}	&0.56\\ \hline
            \color{red}\textbf{md-RPCA} &\color{red}\textbf{0.87}	&\color{red}\textbf{0.80}	&\color{red}\textbf{0.82}	&\color{red}\textbf{0.83}	&\color{red}\textbf{0.82}	&\color{blue}0.51	&\color{red}\textbf{0.81}\\ \hline
            sd-RPCA &0.50	&0.77	&0.33	&0.82	&0.81	&0.21	&0.57\\ \hline
            kd-RPCA &0.50	&0.80	&0.73	&0.83	&0.81	&0.48	&0.74\\ \hline
            ROBPCA-AO &0.32	&0.03	&0.09	&0.07	&0.09	&0.21	&0.11\\ \hline
        \bottomrule
      \end{tabular}
    \end{table}
\end{frame}

\begin{frame}[c]{Results - CTU-13 data set}
	\begin{itemize}
		\item \textbf{md-RPCA, kd-RPCA and sd-RPCA overcome} the remain algorithms in \textbf{average} results;
		\item \textbf{md-RPCA} obtained the \textbf{best average} and the \textbf{best $F_1$ in 10 scenarios};
		\item The anomalies of the \textbf{scenario 19} are peer-to-peer botnet traffic generated by nsys.ay, which are related to botnet synchronization and \textbf{not to network attacks}.
	\end{itemize}
\end{frame}

%offline-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Eigensimilarity for Anomaly Detection in Offline Mobile Client}
\label{Blocks}

\begin{frame}[c]{Introduction}
	\begin{itemize}
		\item According to \cite{khan2013towards, khan2015cloud}, the \textbf{state-of-the-art mobile cloud security models do not consider} the problem of the \textbf{offline security mode};
		\item Providers of the mobile security \textbf{avoid} the problem by completely \textbf{forbidding} the offline access.
	\end{itemize}
\end{frame}

\begin{frame}[c]{Introduction}
	In this work we present:
	\begin{itemize}
	    \item An \textbf{architecture and proof of concept of Eigensimilarity} for malicious behavior detection on mobile app;
		\item The common \textbf{threat} scenarios for \textbf{file cloud apps};
		\item The \textbf{performance} for \textbf{mobile} offline processing.
	\end{itemize}
\end{frame}

% \begin{frame}[c]{The mobile security architecture}
% 	\begin{figure}[h!]
% 		\centering
% 		\includegraphics[width=5cm]{figures/ch3/fig01.eps}
% 		\caption{The core set of functions and protocols of the mobile cloud security infrastructure}
% 		\label{fig:3_01}
% 	\end{figure}
% \end{frame}

\begin{frame}[c]{Proposal for offline mobile security}
	\begin{figure}[h!]
		\centering
		\includegraphics[width=11cm]{figures/ch3/fig03.eps}
		\caption{Proposed Architecture for Offline Mobile Security}
		\label{fig:3_03}
	\end{figure}
\end{frame}

\begin{frame}[c]{Offline Behavioral Analysis}
	\begin{figure}[h!]
		\centering
		\includegraphics[width=7cm]{figures/ch3/fig06.eps}
		\caption{The Threat Intelligence Manager Workflow}
		\label{fig:3_06}
	\end{figure}
\end{frame}

\begin{frame}[c]{MOS for Threat Intelligence}
	\begin{itemize}
		\item Largest eigenvalue analysis and MOS (Previously presented);
		\item Applied to user behavior analysis;
		\begin{itemize}
			\item We model file operations as signals (matrices);
			\item We select features that can highlight malicious behaviors;
			\item We implement and evaluate into mobile devices.
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}[c]{Common threat scenarios}
	\begin{enumerate}
		\item An attacker uses a \textbf{valid password} to perform operations on a \textbf{bulk} of files.;
		\item Usage of \textbf{expired password} to perform unauthorized \textbf{operations};
	\end{enumerate}
\end{frame}

\begin{frame}[c]{Data Modeling for Behavioral Analysis}
	Selected features:
	\begin{itemize}
		\item File Access (Time and file system location);
		\item File Update (Time and file system location);
		\item File Download (Start time, end time and file system location);
		\item File Upload (Start time, end time and file system location).
	\end{itemize}
\end{frame}

% \begin{frame}[c]{Result: Data Modeling and Eigenvalue Analysis}
% 	\begin{table}
% 		\tiny
% 		\label{tab:3_04}
% 		\centering
% 		\begin{tabular}{|l|l|l|l|l|l|l|l|}
% 			\hline \rowcolor{Gray} Device	& \begin{tabular}[x]{@{}l@{}}Log Size\\(MB)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Window\\(min)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Modeling\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Avg. Eig.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Std. Eig.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Eig. Min.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Eig. Max.\\(ms)\end{tabular}	\\ \hline
% 			Galaxy GT-I9300	& 6	& 60	& 107	& 209.52	& 18.58	& 183	& 276	\\ \hline
% 			Galaxy GT-I9300	& 6	& 40	& 115	& 227.26	& 18.13	& 191	& 289	\\ \hline
% 			Galaxy GT-I9300	& 6	& 20	& 89	& 268.14	& 21.94	& 229	& 315	\\ \hline
% 			Galaxy GT-I9300	& 6	& 10	& 90	& 347.42	& 24.11	& 304	& 421	\\ \hline
% 			Galaxy GT-I9300	& 4.1	& 60	& 20	& 60.90	& 15.19	& 37	& 106	\\ \hline
% 			Galaxy GT-I9300	& 4.1	& 40	& 20	& 68.72	& 15.71	& 43	& 114	\\ \hline
% 			Galaxy GT-I9300	& 4.1	& 20	& 34	& 89.04	& 16.78	& 54	& 133	\\ \hline
% 			Galaxy GT-I9300	& 4.1	& 10	& 21	& 117.24	& 14.36	& 96	& 171	\\ \hline
% 			Galaxy GT-I9300	& 1.4	& 60	& 10	& 159.82	& 15.82	& 125	& 197	\\ \hline
% 			Galaxy GT-I9300	& 1.4	& 40	& 10	& 168.06	& 15.90	& 139	& 220	\\ \hline
% 			Galaxy GT-I9300	& 1.4	& 20	& 11	& 204.4	& 20.46	& 176	& 269	\\ \hline
% 			Galaxy GT-I9300	& 1.4	& 10	& 13	& 259.00	& 21.34	& 220	& 315	\\ \hline
% 			Galaxy Tab SM-T800	& 6	& 60	& 7	& 59.30	& 6.55	& 54	& 74	\\ \hline
% 			Galaxy Tab SM-T800	& 6	& 40	& 8	& 62.56	& 7.05	& 56	& 80	\\ \hline
% 			Galaxy Tab SM-T800	& 6	& 20	& 10	& 73.28	& 8.59	& 65	& 95	\\ \hline
% 			Galaxy Tab SM-T800	& 6	& 10	& 8	& 93.48	& 9.13	& 83	& 130	\\ \hline
% 			Galaxy Tab SM-T800	& 4.1	& 60	& 11	& 18.64	& 4.51	& 16	& 38	\\ \hline
% 			Galaxy Tab SM-T800	& 4.1	& 40	& 11	& 19.64	& 5.12	& 17	& 38	\\ \hline
% 			Galaxy Tab SM-T800	& 4.1	& 20	& 12	& 25.12	& 5.55	& 21	& 46	\\ \hline
% 			Galaxy Tab SM-T800	& 4.1	& 10	& 12	& 32.32	& 7.29	& 27	& 55	\\ \hline
% 			Galaxy Tab SM-T800	& 1.4	& 60	& 4	& 49.08	& 6.01	& 42	& 62	\\ \hline
% 			Galaxy Tab SM-T800	& 1.4	& 40	& 5	& 51.42	& 7.36	& 44	& 74	\\ \hline
% 			Galaxy Tab SM-T800	& 1.4	& 20	& 5	& 51.12	& 7.80	& 54	& 91	\\ \hline
% 			Galaxy Tab SM-T800	& 1.4	& 10	& 7	& 75.24	& 7.71	& 65	& 90	\\ \hline
% 		\end{tabular}
% 	\end{table}
% \end{frame}

% \begin{frame}[c]{Result: EDC MOS}
% 	\begin{table}[!t]
% 		\tiny
% 		\label{tab:3_05}
% 		\centering
% 		\begin{tabular}{|l|l|l|l|l|l|l|l|}
% 			\hline \rowcolor{Gray} Device	& \begin{tabular}[x]{@{}l@{}}Log Size\\(MB)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Window\\(min)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Avg. EDC.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Std. EDC.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Min. EDC.\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Max. EDC.\\(ms)\end{tabular}	\\ \hline
% 			Galaxy GT-I9300	& 6	& 60	& 5.27	& 4.04	& 3	& 20	\\ \hline
% 			Galaxy GT-I9300	& 6	& 40	& 10.78	& 6.37	& 6	& 34	\\ \hline
% 			Galaxy GT-I9300	& 6	& 20	& 32.62	& 12.44	& 21	& 88	\\ \hline
% 			Galaxy GT-I9300	& 6	& 10	& 115.08	& 17.45	& 88	& 158	\\ \hline
% 			Galaxy GT-I9300	& 4.1	& 60	& 5.68	& 4.18	& 3	& 23	\\ \hline
% 			Galaxy GT-I9300	& 4.1	& 40	& 10.76	& 5.31	& 7	& 27	\\ \hline
% 			Galaxy GT-I9300	& 4.1	& 20	& 37.58	& 10.30	& 23	& 61	\\ \hline
% 			Galaxy GT-I9300	& 4.1	& 10	& 125.98	& 18.56	& 101	& 191	\\ \hline
% 			Galaxy GT-I9300	& 1.4	& 60	& 4.92	& 3.49	& 3	& 17	\\ \hline
% 			Galaxy GT-I9300	& 1.4	& 40	& 9.00	& 4.23	& 6	& 25	\\ \hline
% 			Galaxy GT-I9300	& 1.4	& 20	& 30.14	& 9.21	& 19	& 62	\\ \hline
% 			Galaxy GT-I9300	& 1.4	& 10	& 100.62	& 15.83	& 69	& 163	\\ \hline
% 			Galaxy Tab SM-T800	& 6	& 60	& 1.84	& 0.65	& 1	& 3	\\ \hline
% 			Galaxy Tab SM-T800	& 6	& 40	& 3.26	& 1.24	& 2	& 7	\\ \hline
% 			Galaxy Tab SM-T800	& 6	& 20	& 10.90	& 2.40	& 9	& 21	\\ \hline
% 			Galaxy Tab SM-T800	& 6	& 10	& 41.86	& 7.33	& 34	& 60	\\ \hline
% 			Galaxy Tab SM-T800	& 4.1	& 60	& 1.85	& 0.60	& 1	& 3	\\ \hline
% 			Galaxy Tab SM-T800	& 4.1	& 40	& 3.62	& 1.10	& 2	& 8	\\ \hline
% 			Galaxy Tab SM-T800	& 4.1	& 20	& 12.04	& 2.79	& 9	& 22	\\ \hline
% 			Galaxy Tab SM-T800	& 4.1	& 10	& 40.16	& 6.48	& 35	& 60	\\ \hline
% 			Galaxy Tab SM-T800	& 1.4	& 60	& 1.98	& 0.89	& 1	& 6	\\ \hline
% 			Galaxy Tab SM-T800	& 1.4	& 40	& 3.30	& 1.16	& 2	& 7	\\ \hline
% 			Galaxy Tab SM-T800	& 1.4	& 20	& 10.48	& 2.90	& 8	& 21	\\ \hline
% 			Galaxy Tab SM-T800	& 1.4	& 10	& 34.52	& 4.08	& 30	& 45	\\ \hline
% 		\end{tabular}
% 	\end{table}
% \end{frame}

\begin{frame}[c]{Results - synthetic data set}
	\begin{itemize}
		\item The \textbf{highest eigenvalue analysis} time: \textbf{421ms} with average of \textbf{347.42ms};
		\item The \textbf{lower window} size leads to the \textbf{larger eigenvalue analysis time};
		\item The \textbf{MOS} processing time \textbf{increases} with the \textbf{window size decreasing};
		\item The \textbf{highest MOS} processing time: lower than \textbf{200s};
		\item This results represent an \textbf{acceptable processing time} for \textbf{anomaly detection in mobile clients}.
	\end{itemize}
\end{frame}

%section-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\section{Conclusion}

\begin{frame}[c]{Conclusion}
	\begin{itemize}
		\item \textbf{Synflood, fraggle and port scan} attacks can be \textbf{detected} accurately and with great detail by \textbf{Eigensimilarity}, in an automatic and \textbf{blind} fashion;
		\item As \textbf{proof of concept}, a mobile app was developed to test the \textbf{security solution}, with positive \textbf{results} in terms of \textbf{performance};
	\end{itemize}
\end{frame}

\begin{frame}[c]{Conclusion}
	\begin{itemize}
		\item \textbf{m-RPCA} can \textbf{improve the anomaly detection on skewed and imbalanced data set}. 
		\item \textbf{m-RPCA} can be adopted for \textbf{network attack detection}, with \textbf{better} results than \textbf{widely adopted algorithms for outlier detection}; 
	\end{itemize}
\end{frame}

%section-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
\begin{frame}[plain, allowframebreaks]{Bibliography}
	\setbeamertemplate{bibliography item}[text]
	\printbibliography
\end{frame}

\end{document}

### 2020-04-28 docs/papers/2019JNCA_m-rpca/main.tex
updating JNCA paper

--- 2020-04-28 docs/thesis/chapters/4_m_rpca/m_rpca.tex
@@
-\chapter{Robust Framework for Detection of Network Attacks in Imbalanced Traffic}
-\label{ch:4_m_rpca}
+\documentclass[review]{elsarticle}
 
-According to \cite{distilnetworks2019}, 37.9\% of all Internet traffic of 2018 was not from human activities, but from bots, that can be classified as good bots, that perform legitimate operations, and bad bots, which perform malicious activities, such as DDoS attacks, probe attack or frauds.
-In 2018, bad bots accounted for 20.4\% of all website traffic.
+\usepackage{lineno,hyperref}
+\usepackage[mathletters]{ucs}
+\usepackage[utf8x]{inputenc}
+\usepackage[linesnumbered, ruled, vlined]{algorithm2e}
+\usepackage{natbib}
+\usepackage{supertabular}
+\usepackage{fancybox}
+\usepackage{acronym}
+\usepackage{array}
+\usepackage{booktabs}
+\usepackage{graphicx}
+\usepackage{rotating}
+\usepackage{tabularx}
+\usepackage{multirow}
+\usepackage{hhline}
+\usepackage{setspace}
+\usepackage{placeins}
+\usepackage{longtable}
+\usepackage{listings}
+\usepackage{epstopdf}
+\usepackage{subfigure}
+\usepackage{courier}
+\usepackage{amsfonts}
+\usepackage{morefloats}
+\usepackage{lipsum}
+\usepackage{mathtools}
+\usepackage{amsthm,amsmath}
 
-According to \cite{BISSELL2019} the proportion of cost on discovery activities for cybersecurity has increased steadily since 2015.
-Forensic cyber and user behavior analytics also present an opportunity for cost savings — US\$1.72 million — with discovery and investigation activities.
-However, only 32\% of organizations have deployed these technologies enterprise-wide.
-Accenture consulting argues that 89\% of survey respondents believe breakthrough technologies, like machine learning and user behavior analytics, are essential to securing the future of their organizations \cite{Accenture2018}.
+% table
+\usepackage{makecell}
+\usepackage{color, colortbl}
+\definecolor{Gray}{gray}{0.9}
 
+% appendix
+\usepackage[titletoc,title]{appendix}
+\usepackage{hyperref}
+\usepackage{cleveref}
+
+% norm
+
+\modulolinenumbers[5]
+
+\journal{Journal of Network and Computer Applications}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+%% Elsevier bibliography styles
+%%%%%%%%%%%%%%%%%%%%%%%
+%% To change the style, put a % in front of the second line of the current style and
+%% remove the % from the second line of the style you would like to use.
+%%%%%%%%%%%%%%%%%%%%%%%
+
+%% Numbered
+%\bibliographystyle{model1-num-names}
+
+%% Numbered without titles
+%\bibliographystyle{model1a-num-names}
+
+%% Harvard
+%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+
+%% Vancouver numbered
+%\usepackage{numcompress}\bibliographystyle{model3-num-names}
+
+%% Vancouver name/year
+%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
+
+%% APA style
+%\bibliographystyle{model5-names}\biboptions{authoryear}
+
+%% AMA style
+%\usepackage{numcompress}\bibliographystyle{model6-num-names}
+
+%% `Elsevier LaTeX' style
+\bibliographystyle{elsarticle-num}
+%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
+
+\begin{frontmatter}
+
+\title{Moment Distances from Robust Subspace for Botnet Detection}
+
+%% Group authors per affiliation:
+\author[unbaddress]{Thiago P. B. Vieira}
+\author[unbaddress]{Eduardo S. C. Vilaça}
+\author[unbaddress,Ilmenauaddress,Fraunhoferaddress]{João Paulo C. L. da Costa}
+\author[unbaddress]{Éder S. Gualberto}
+\author[unbaddress]{Rafael T. de Sousa Júnior}
+
+\address[unbaddress]{Department of Electrical Engineering, University of Brasilia (UnB), 70910-900, Brasília-DF, Brazil}
+\address[Ilmenauaddress]{Institute for Information Technology, Ilmenau University of Technology, Ilmenau, Germany}
+\address[Fraunhoferaddress]{Fraunhofer Institute for Integrated Circuits IIS, Erlangen, Germany}
+
+\begin{abstract}
+The U.S.
+government estimates that malicious cyber activity cost the U.S.
+economy between US\$57 billion and US\$109 billion in 2016.
+Accenture argues that 89\% of survey respondents believe breakthrough technologies, like artificial intelligence, machine learning and user behavior analytics, are essential for securing organizations.
+The imbalanced data can compromise the performance of standard learning algorithms, creating bias or unfair weight to learn from the majority class and reducing detection capacity of anomalies.
+Some widely adopted algorithms for anomaly detection assume a Gaussian distributed data for legitimate observations, but this assumption may not be observed in network traffic, which is usually characterized by skewed and heavy-tailed distributions.
+We propose the Moment-based Robust Principal Component Analysis (m-RPCA) for botnet attack detection, which is a framework based on distances between contaminated observations and moments computed from a robust subspace learned by RPCA, to detect anomalies and network attacks.
+We evaluate m-RPCA on simulated data and on CTU-13 data set, and results show that m-RPCA can improve the botnet attack detection and the anomaly detection on skewed data, in comparison to anomaly detection algorithms.
+\end{abstract}
+
+\begin{keyword}
+Anomaly Detection, Botnet Attack Detection, Skewed Data, Imbalanced Data, Robust Principal Component Analysis (RPCA).
+\end{keyword}
+
+\end{frontmatter}
+
+\linenumbers
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+\label{sec:introduction}
+
+According to \cite{distilnetworks2019}, 37.9\% of all Internet traffic of 2018 was not from human activities, but from bots, that can be classified as good or bad bots, which accounted for 20.4\% of all website traffic.
+A botnet is a collection of bots or systems for executing automated tasks, often in the form of compromised hosts, including desktops, mobile devices or things of the Internet of Things (IoT).
+Some examples of botnets include Cyclone, Mirai, Nitol and Sentry MBA \cite{Catucci2019}.
+
+According to Hevesi \cite{Hevesi2019} in a Gartner's publication, botnets scan the Internet looking for unprotected IoT devices, which are projected to be 25 billion devices in 2021.
+Hevesi argues that if 1\% of IoT devices is added to a botnet and had access to 5G, the potential throughput would be 50 Gbps in a flooding attack.
+
+Distributed attacks organized by botnet has increased and demanded the development of counter measures to detect and avoid unknown attacks or even to deal with adversarial changes of behavior, location and other patterns \cite{wang2017botnet, Wang2018ddosbotnetssurvey}.
 To face the adversarial model, network attacks and counter measures of attackers to avoid detection, it is possible to adopt unsupervised or semi-supervised approaches for network anomaly detection, by means of behavioral analysis, where it is not necessary known anomalies for training models \cite{moustafa2019holistic}.
@@
 Anomalies in the context of network traffic can be hard to identify and separate from legitimate data due to the rare occurrences of anomalies in comparison to legitimate events.
-Therefore, anomaly detection algorithms have to be highly discriminating, robust to corruption and able to deal with the imbalanced data problem \cite{he2008learning}.
-Note that data corruption refers to outliers that can be part of the data but not be considered malicious.
-The imbalanced data problem corresponds to data sets exhibiting significant imbalances of classes or rare events of some classes \cite{he2009learning}, which can be legitimate or malicious classes in network anomaly detection problems.
-
-Traditionally adopted algorithms for anomaly detection assume a Gaussian or symmetric distributed data \cite{lakhina2005mining}, however this assumption may not be observed in some real world problems \cite{benson2010network}, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{leon2017probability}.
+Additionally, widely adopted algorithms for anomaly detection assume a Gaussian or symmetric distributed data \cite{lakhina2005mining}, however this assumption may not be observed in some real world problems \cite{benson2010network}, such as the case of network traffic analysis, where network traffic features are usually more characterized by skewed and heavy-tailed distributions \cite{leon2017probability}.
 
@@
 
-The findings presented by Benson \emph{et al.} \cite{benson2010network} and Leon-Garcia \cite{leon2017probability} show that the skewness and heavy-tailed distributions may be important for network traffic analysis, and can motivate researches to evaluate the impact of skewed data into algorithms that rely on Gaussian distributed data for network anomaly detection.
-Additionally, the fitting of network traffic to skewed and heavy-tailed distributions can indicate opportunities to exploit properties and characteristics of the skewness and heavy-tailed distributions to obtain improved classifiers for network anomaly detection.
+The findings presented by Benson \emph{et al.} \cite{benson2010network} and Leon-Garcia \cite{leon2017probability} show that the skewness and heavy-tailed distributions may be important for network traffic analysis, and can motivate researches to evaluate the impact of skewed data into algorithms that rely on Gaussian data for network anomaly detection.
+Additionally, the fitting of network traffic to skewed and heavy-tailed distributions can indicate opportunities to exploit characteristics of the skewness and heavy-tailed distributions to improve classifiers for network anomaly detection.
 
 Network anomaly detection problems are usually characterized by imbalanced data \cite{Phua2004minority}.
-However, learning algorithms for imbalanced data has been a challenging research topic, considering that the imbalanced data can compromise the performance of most standard learning algorithms, creating bias or unfair weight to learn from the majority class and reducing detection capacity of anomalies that are characterized by the minority class \cite{he2008learning}.
+However, learning algorithms for imbalanced data has been a challenging topic, considering that imbalanced data can compromise the performance of most standard learning algorithms, creating bias or unfair weight to learn from the majority class and reducing the detection capacity \cite{he2008learning}.
 Therefore, learning methods for imbalanced and skewed data have attracted attention of researchers \cite{hubert2009robustskewed}.
 
-Robust subspace learning has been attracting a growing attention of researchers aiming the development of network attack detection systems \cite{pascoal2012robust, zhou2017anomaly} that rely on behavioral analysis.
+Robust subspace learning has been attracting a growing attention of researchers aiming the development of network attack detection systems that rely on behavioral analysis \cite{pascoal2012robust, zhou2017anomaly}.
 An outlyingness-approach based on a robust estimator of skewness, combined with robust estimators of location and scale, can be able to flag the outlying measurements \cite{hubert2009robustskewed}.
@@
 
-Considering that the skewness of anomalous and legitimate traffic can highlight features for improving anomaly detection and network attack detection in imbalanced data, and considering that the distance between robust estimates can be used for network attack detection, we propose the Moment-based Robust Principal Component Analysis (m-RPCA), which is an approach based on distances between moments computed from a robust subspace learned by Robust Principal Component Analysis (RPCA) from contaminated observations, in order to detect anomalies from skewed data and network traffic.
+Considering that the skewness of anomalous and legitimate traffic can highlight features for improving anomaly detection and network attack detection in imbalanced data, and considering that the distance between robust estimates can be used for network attack detection, we propose the Moment-based Robust Principal Component Analysis (m-RPCA), which is a framework based on distances between moments computed from a robust subspace learned by Robust Principal Component Analysis (RPCA) from contaminated observations, to detect anomalies from skewed data and network traffic.
 
-The proposed approach relies on a robust subspace computed from supposed legitimate observations, for estimating the moments to be used for distance analysis.
-The anomaly detection from contaminated observations evaluate the Mahalanobis distance between the robust moments and new contaminated observations, in a semi-supervised fashion, without the computational cost of new robust subspace learning for anomaly detection from new observations.
+The m-RPCA relies on a robust subspace computed from supposed legitimate observations, for estimating the moments to be used for distance analysis.
+The anomaly detection from contaminated observations evaluate the Mahalanobis distance between the robust moments and new contaminated observations, in a semi-supervised fashion.
 The m-RPCA can also be computed as an unsupervised algorithm, with subspace learning from the same contaminated data that is the target of the anomaly detection analysis.
 
-We evaluate the accuracy of the m-RPCA for anomaly detection on simulated data sets, with skewed and heavy-tailed distributions, and for network attack detection on CTU-13 data set \cite{garcia2014empirical}, which is a large data set of legitimate, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for anomaly detection systems \cite{osanaiye2016distributed}.
-The Experimental evaluation compares our proposal to standard and widely adopted algorithms for anomaly detection, which are based on clustering and statistical approaches, and to ROBPCA \cite{hubert2005robpca}, which is a anomaly detection method that relies on robust estimates with adjusted outlyingness based on robust skewness.
+We evaluate the accuracy of the m-RPCA for anomaly detection on simulated data set, with skewed and heavy-tailed distributions, and for botnet attack detection on CTU-13 data set \cite{garcia2014empirical}, which is a large data set of legitimate, background and botnet traffic that has been adopted to deal with the lack of up-to-date real-world data sets for network attack detection \cite{osanaiye2016distributed}.
+The experimental evaluation compares m-RPCA to standard and widely adopted algorithms for anomaly detection, which are based on clustering and statistical approaches, and to ROBPCA-AO \cite{hubert2005robpca}, which is an anomaly detection method that relies on robust estimates with adjusted outlyingness based on robust skewness.
 
-The main contribution of this work is the proposal of a novel semi-supervised and unsupervised method for anomaly detection in skewed and imbalanced data, with results showing improvements through an experimental evaluation on simulated skewed and heavy-tailed distributions and on real data set with legitimate network traffic and botnet attacks.
+The main contributions of this work are a novel semi-supervised and unsupervised method for anomaly detection in skewed and imbalanced data, with results showing improvements in experimental evaluation on simulated skewed and heavy-tailed data and on real data set with legitimate and botnet traffic.
 
-This chapter is organized as follows.
-In Section \ref{sec:3_relatedworks}, a literature review about network anomaly detection, botnet detection, and imbalanced learning is conducted.
-We present in Section \ref{sec:3_datamodel} the data model and the evaluated data set.
-In Section \ref{sec:3_proposal} it is described the proposed approach for network attack detection.
-We discuss in Section \ref{sec:3_experiments} the experimental validation, and in Section \ref{sec:3_simulated_result} are presented the results of the simulated scenarios and in Section \ref{sec:3_ctu13_result} we describe the results for anomaly detection on CTU-13 data set.
-Finally, in Section \ref{sec:3_conclusion} we draw the conclusions and the suggestions for future work.
+This paper is organized as follows.
+In Section \ref{sec:review}, a literature review about network anomaly detection, botnet detection, and imbalanced learning is conducted.
+We present in Section \ref{sec:datamodel} the data model and the evaluated data set.
+In Section \ref{sec:mdrs} it is described the proposed framework for network attack detection.
+In Section \ref{sec:experimentalresults} we discuss the experimental validation, the results for simulated scenarios and the results for botnet attack detection on CTU-13 data set.
+Finally, in Section \ref{sec:conclusionandfutureworks} we draw the conclusions and the suggestions for future work.
 
-\section{Related Works}
-\label{sec:3_relatedworks}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% network anomaly detectiom, botnet detection, botnet detection approaches, approaches based on subspace learning, feature generation and aggregation, about the methodos to compare and metrics
+\section{Literature Review}
+\label{sec:review}
 
@@
 Bhuyan \emph{et al.} \cite{bhuyan2014network} provide an overview of facets of network anomaly detection, present attacks normally encountered by network intrusion detection systems, and categorize existing network anomaly detection methods and systems based on the underlying techniques.
-Ahmed \emph{et al.} \cite{ahmed2016survey} present an analysis of four major categories of anomaly detection techniques, which include classification, statistical, information theory and clustering.
-Moustafa \emph{et al.} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), describing details of cyber-attacks and new solutions for anomaly detection, and provides a benchmark data sets for training and validating approaches for network anomaly detection.
+Ahmed \emph{et al.} \cite{ahmed2016survey} present an analysis of four major categories of anomaly detection techniques, which include classification, statistical, information theory and clustering approaches.
+Moustafa \emph{et al.} \cite{moustafa2019holistic} discuss aspects of anomaly-based Network Intrusion Detection Systems (NIDSs), describing details of cyber-attacks and new solutions for anomaly detection, and provides a benchmark data set for training and validating approaches for network anomaly detection.
 
 % botnet detection
+According to \cite{distilnetworks2019}, 37.9\% of all Internet traffic of 2018 was not from human activities, but from bots, that can be classified as good bots or bad bots.
 A botnet is a network of bots, that are compromised machines under the influence of malware (bot).
-The botnet is commandeered by a botmaster and used as resource for attacks, such as distributed denial-of-service (DDoS) attacks, and fraudulent activities such as spam, phishing, identity theft, and information ex-filtration.
+The botnet is commandeered by a botmaster and used as resource for attacks, such as distributed denial-of-service (DDoS) attacks, and fraudulent activities, such as spam, phishing, identity theft, and information ex-filtration.
 We refer to \cite{ahmed2016survey} and \cite{moustafa2019holistic} for an overview of network attacks.
@@
 Centralized C\&C structures using the Internet Relay Chat (IRC) protocol have been utilized by botmasters for a long time, but other protocols, such as HTTP, and architectures, such as Peer-To-Peer, have also been adopted \cite{gu2008botminer}.
-
-According to \cite{distilnetworks2019}, 37.9\% of all Internet traffic of 2018 was not from human activities, but from bots, that can be classified as good bots, that perform legitimate operations, and bad bots, which are responsible by malicious activities, such as DDoS, probe attack or frauds.
 
@@
 
-Several approaches for network attack detection uses the KDD 99 \cite{ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data sets for accuracy and performance evaluation, due to their availability and labeled attacks.
-Even though the KDD 99 data set are criticized by the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data sets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
-NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 data set that removed the redundant data records, in order to avoid biased classifications.
-However, NSL-KDD data set maintain the limitations of the KDD 99 regarding volume and lacks on reproduction of recent network traffic and attacks.
+Several approaches for network attack detection uses the KDD 99 \cite{ahmed2016survey,osanaiye2016distributed,bhuyan2014network} data set for accuracy and performance evaluation, due to their public availability and labeled traffic.
+Even though the KDD 99 are criticized by the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled data set in use by researchers \cite{osanaiye2016distributed,bhuyan2014network}.
+NSL-KDD \cite{tavallaee2009detailed} data set is the refined version of the KDD 99 that removed the redundant data records, to avoid biased classifications.
+However, NSL-KDD maintain the limitations of the KDD 99 regarding volume and lacks on reproduction of recent network traffic and attacks.
 
-The objective of Garcia \emph{et al.} \cite{garcia2014empirical} is to compare three botnet detection methods by means of a simple and reproducible methodology, by a good data set and by a new error metric.
-This paper evaluates some data sets for network anomaly detection, and surveys some approaches for botnet detection, and proposes two methods (BClus and CAMNEP) for botnet detection, comparing results to BotHunter \cite{gu2007bothunter}.
+Garcia \emph{et al.} \cite{garcia2014empirical} compare three botnet detection methods by means of a simple and reproducible methodology, by a good data set and by a new error metric.
+This paper evaluates data sets for network anomaly detection, and surveys approaches for botnet detection, and proposes two methods (BClus and CAMNEP) for botnet detection, comparing results to BotHunter \cite{gu2007bothunter}.
 
-Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by attack, legitimate and background labeled data, in an imbalanced distribution like in a real networks.
-The authors recommend scenarios for training and testing in order to avoid the use of traffic from a botnet family for training and testing, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
-Taking into account the adoption of unsupervised or semi-supervised approaches for anomaly detection, adopting the training and testing approach proposed by Garcia \emph{et al.}, some botnet malwares wouldn't be tested, since in the author's proposal some botnets are present only for training.
+Considering the lack of available labeled data sets, Garcia \emph{et al.} \cite{garcia2014empirical} proposes the CTU-13 data set, which is composed by attack, legitimate and background labeled data, in an imbalanced distribution and according to real network traffics.
+The authors recommend scenarios for training and testing to avoid the use of traffic from a botnet family for training and testing, aiming to ensure that the evaluated methods can generalize and detect new behaviors.
+Taking into account the adoption of unsupervised or semi-supervised approaches for anomaly detection, if adopted the training and testing approach proposed by Garcia \emph{et al.}, some botnet malwares wouldn't be tested, since in the author's proposal, some botnets are present only for training.
 
@@
 The second stage detects the bots using ideas from social network community detection in a graph that captures correlations of interactions among nodes over time.
-This work is compared to the BotHunter \cite{gu2007bothunter} for the CTU-13 botnet data set \cite{garcia2014empirical}.
+This work is compared to the BotHunter \cite{gu2007bothunter} for the CTU-13 data set \cite{garcia2014empirical}.
 
 % Anomaly detection by subspace learning
-Traditional PCA-based anomaly detection models are not suitable for anomaly interpretation, as they judge whether a data instance is an anomaly or not based on the length of its projection on the abnormal subspace spanned by the less significant principal components, and there is no direct mapping between PCA’s dimensionality-reduced subspace and the original feature space \cite{ringberg2007sensitivity}.
-However, to overcome the above mentioned limitations, some approaches based on PCA have been proposed for network anomaly detection.
+According to Ringberg \emph{et al.} \cite{ringberg2007sensitivity}, traditional PCA-based anomaly detection models are not suitable for anomaly interpretation, as they judge whether a data instance is an anomaly or not based on the length of its projection on the abnormal subspace spanned by the less significant principal components, and there is no direct mapping between PCA’s dimensionality-reduced subspace and the original subspace \cite{ringberg2007sensitivity}.
+To overcome the above mentioned limitations, some approaches based on PCA have been proposed for network anomaly detection.
 
@@
 Lee \emph{et al.} \cite{Lee2013} presented OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling.
-Vieira \emph{et al.} \citeme{vieira2017model} proposed a framework that applies Model Order Selection (MOS) for detection of time frames under attack and uses similarity analysis to extract details and detect the time and ports under attack.
+Vieira \emph{et al.} \cite{vieira2017model} proposed a framework that applies Model Order Selection (MOS) for detection of time frames under attack and uses similarity analysis to extract details and detect the time and ports under attack.
 
@@
 However, RPCA has also been adopted for general outlier detection \cite{hubert2005robpca,hubert2009robustskewed,cherapanamjeri2017thresholding,zhou2017anomaly,NetflixSurus} and for anomaly detection on network traffic \cite{pascoal2012robust}.
-ROBPCA \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of mean and covariance matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
+ROBPCA-AO \cite{hubert2005robpca} intends to identify outliers using PCA from robust estimates of mean and covariance matrix, to reduce the data dimensions and plotting the orthogonal distances versus the robust score distances, to flag an outlier map.
 However, ROBPCA flags many points as outlying when the original data is skewed.
 Therefore, Hubert \emph{et al.} \cite{hubert2009robustskewed} proposed ROBPCA-AO, which improves ROBPCA for problems with skewed data, by means of an adjusted outlyingness based on robust skewness.
-Hubert \emph{et al.} \cite{hubert2009robustskewed} evaluated ROBPCA-AO for real and simulated data sets, but it is not clear if this method was evaluated for very skewed and imbalanced data sets, such as in the network attack detection problem.
+Hubert \emph{et al.} \cite{hubert2009robustskewed} evaluated ROBPCA-AO for real and simulated data set, but it is not clear if this method was evaluated for very skewed and imbalanced data set, such as in the network attack detection problem.
 
@@
 Pascoal \emph{et al.} \cite{pascoal2012robust} proposed an approach based on a robust mutual information estimator for feature selection and based on RPCA for outlier detection in internet traffic.
-The anomaly detection proposed by Pascoal \emph{et al.} is an unsupervised approach that estimate the first $k$ robust principal components, calculate the score and the orthogonal distances, calculate the thresholds and classify new observations accordingly.
+The anomaly detection proposed by Pascoal \emph{et al.} \cite{pascoal2012robust} is an unsupervised approach that estimate the first $k$ robust principal components, calculate the score and the orthogonal distances, calculate the thresholds and classify new observations accordingly.
 
 Zhou and Paffenroth \cite{zhou2017anomaly} proposed the Robust Deep Autoencoders (RDA), which the central idea is that a RDA inherits the non-linear representation capabilities of autoencoders combined with the anomaly detection capabilities of RPCA.
-Considering that outliers and noise may reduce the quality of representations discovered by deep autoencoders, the proposed model isolates noise and outliers in the input by means of a RPCA approach, and the autoencoder is trained after this isolation.
+Considering that outliers and noise may reduce the quality of representations discovered by deep autoencoders, the proposed model isolates noise and outliers in the input by means of a RPCA, and the autoencoder is trained after this isolation.
 RDA was evaluated by the authors for the MNIST data set.
@@
 Benson \emph{et al.} \cite{benson2010network} conducted an empirical study of the network traffic in 10 data centers belonging to three different types of organizations, including university, enterprise, and cloud data centers.
-Findings of Benson \emph{et al.} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
+Findings of Benson \emph{et al.} \cite{benson2010network} indicate that certain positive skewed and heavy-tailed distributions can model data center switch traffic, and highlights a difference between the data center environment and the wide area network, where the long-tailed Pareto distribution typically shows the best fit \cite{benson2010network}.
 
@@
 MD is a multi-dimensional generalization for measuring how many standard deviations away $\pmb{x}$ is from the mean $\pmb{\mu}$ and covariance $\hat{\pmb{\Sigma}}$ of $\pmb{X}$.
-MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the Mahalanobis Distance between robust estimates and contaminated observations can reveal anomalies.
+MD has been used for distance based anomaly detection with robust estimates in many areas, assuming that the MD between robust estimates and contaminated observations can reveal anomalies.
 
-We propose an approach based on the distances between moments computed from learned robust subspace and contaminated observations, for anomaly detection on skewed and imbalanced data sets, and evaluate our approach for network attack detection.
-The proposed approach relies on learning robust subspace from supposed legitimate traffic for estimating the moments (mean, skewness and kurtosis).
+We propose a framework based on the distances between moments computed from learned robust subspace and contaminated observations, for anomaly detection on skewed and imbalanced data, and for botnet attack detection.
 
-Thus, the anomaly detection for contaminated observations should evaluate the distance between the robust moments and contaminated observations, in a semi-supervised fashion and without the computational cost of new robust subspace learning for new observations, or in an unsupervised approach, where the robust subspace is learned from the contaminated data and the distance analysis is between the robust moments and the contaminated observations.
+The proposed semi-supervised approach relies on learning robust subspace from supposed legitimate traffic for estimating the moments (mean, skewness and kurtosis), and the anomaly detection is computed from contaminated observations by means of the distance between the robust moments and contaminated observations, without the computational cost of new robust subspace learning to detect anomalies on new observations.
+The m-RPCA can also be adopted as an unsupervised approach, where the robust subspace is learned from the contaminated data and the distance analysis is computed between the robust moments and the contaminated observations.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Data Model}
-\label{sec:3_datamodel}
+\label{sec:datamodel}
 
-In this chapter, we adopt the data model introduced in Section \ref{sec:2_datamodel} and define that the Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries.
+In this paper, scalars are denoted by italic letters ($a, b, A, B, \alpha, \beta$), vectors by lowercase bold letters ($\pmb{a}, \pmb{b}$), matrices by uppercase bold letters ($\pmb{A}, \pmb{B}$), and $a_i,_j$ denotes the (\emph{i, j}) elements of the matrix $\pmb{A}$.
+The superscripts \textsuperscript{T} and \textsuperscript{-1} are used for matrix transposition and matrix inversion, respectively.
+We define that the Frobenius norm is denoted as $\left\| \mathord{\cdot} \right\|_F$, while $\left\| \mathord{\cdot} \right\|_*$ denotes the nuclear norm of a matrix and $\left\| \mathord{\cdot} \right\|_1$ means the sum of the absolute values of matrix entries.
 We also define the operator $\langle \mathord{\cdot} \rangle$, that is the standard trace inner product, and the operator $[ \mathord{\cdot} ]^c$, which denotes the indexes of the $c$ largest values of a vector.
 
-This section also presents a description of the simulated data model in Subsection \ref{sec:3_simulation} and in Subsection \ref{sec:3_CTU-13} we describe the data model for the CTU-13 data set.
+This section also presents a description of the simulated data model in Subsection \ref{sec:3_simulation} and in Subsection \ref{sec:3_CTU-13} we describe the CTU-13 data set.
 
@@
 
-In order to analyze the hypotheses $H_2^{(N)}$ and $H_2^{(A)}$, which evaluate if the robust subspace learning can improves the anomaly detection from imbalanced and skewed data, we create two simulated data sets characterized by skewed and heavy tailed distributions, and create a simulated Gaussian distributed data set, in order to also analyze the detection rate for not skewed and not heavy tailed distributions.
+We create two simulated data sets characterized by skewed and heavy tailed distributions, and create a simulated Gaussian distributed data set, to also analyze the detection rate of m-RPCA for not skewed and not heavy tailed distribution.
 
-We selected Pareto (with scale $a=3$, mode $m=1$ and $\mu=1$) and Lognormal (with mean $\mu=0$ and standard deviation $\sigma=1$) distributions, denoted respectively as $\pmb{Y}_p$, while $\pmb{Y}_l$, to simulate legitimate events and to evaluate the anomaly detection on skewed data performed by our proposals in comparison to widely adopted algorithms for outlier detection, considering that these distributions are skewed and heavy-tailed, and have been adopted to characterize network traffic of the Internet and data centers \cite{benson2010network,leon2017probability}.
-We also adopt the Gaussian distributed data set $\pmb{Y}_g$ to simulate legitimate events with mean $\mu=0$ and unitary standard deviation $\sigma=1$.
+We selected Pareto (with scale $a=3$, mode $m=1$ and $\mu=1$) and Lognormal (with mean $\mu=0$ and standard deviation $\sigma=1$) distributions, denoted respectively as $\pmb{Y}_p$, while $\pmb{Y}_l$, to simulate legitimate observations and to evaluate the anomaly detection on skewed data, by m-RPCA in comparison to widely adopted algorithms for outlier detection.
+The simulated anomalies for Pareto and Lognormal distributions are white Gaussian noise with mean $\mu=0$ and unitary standard deviation $\sigma=1$.
+Note that Pareto and Lognormal distributions are skewed and heavy-tailed, and have been adopted to characterize network traffic of the Internet and data centers \cite{benson2010network,leon2017probability}.
 
-The addictive anomalies for Pareto and Lognormal distributions are white Gaussian noise with mean $\mu=0$ and unitary standard deviation $\sigma=1$.
-The Pareto distribution contaminated by Gaussian noise is denoted as $\pmb{Y}_p^c$, while $\pmb{Y}_l^c$ denotes the Lognormal distribution contaminated by Gaussian noise.
-The $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$ are depicted by the Figures \ref{fig:4.01} and \ref{fig:4.02}, in order to provide a visual example one variable of $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$.
+We adopt the Gaussian distributed data $\pmb{Y}_g$ to simulate legitimate observations, with mean $\mu=0$ and unitary standard deviation $\sigma=1$.
+An uniform noise between $-6$ and $6$ is used to add anomalies into the Gaussian distributed data, where $\pmb{Y}_g^c$ denotes the Gaussian data contaminated by an uniform distribution.
+
+The Figure \ref{fig:4.01} shows a visual example of $\pmb{Y}_p^c$ distribution, where the Pareto distribution $\pmb{Y}_p$ is adopted for the majority class, and the Gaussian distribution simulate anomalies in comparison to the legitimate distribution, composing the Pareto distribution contaminated by Gaussian noise $\pmb{Y}_p^c$.
 
@@
 	\centering
-	\subfigure[Pareto and Gaussian Anomalies]{%
-		\includegraphics[width=0.48\textwidth]{figures/ch4/4_Xpg2.png}
+	\includegraphics[width=0.8\textwidth]{figures/4_Xpg2.png}
+	\caption{Pareto distribution and Gaussian contamination}
 	\label{fig:4.01}
-	}
-	\subfigure[Lognormal and Gaussian Anomalies]{%
-		\includegraphics[width=0.48\textwidth]{figures/ch4/4_Xlogng2.png}
-		\label{fig:4.02}
-	}
-	\caption[Examples of skewed and heavy tailed distributions]{Examples of skewed and heavy tailed distributions}
-	\label{fig:4.03}
 \end{figure}
 
-An uniform noise between $-6$ and $6$ is used to add anomalies into the Gaussian distributed data set, and $\pmb{Y}_g^c$ denotes the contaminated Gaussian data contaminated by an uniform distribution.
-The figure \ref{fig:4.04} shows an example of the Gaussian distribution of legitimate observations and the uniform distribution to simulate the addition of anomalies.
+The $\pmb{Y}_l^c$ denotes the Lognormal distribution contaminated by Gaussian noise, accordingly to depicted by to Figure \ref{fig:4.02}, where it is possible to observe the Lognormal legitimate distribution $\pmb{Y}_l$ and the addictive gaussian noise to simulate the anomalies.Z
+
+\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.8\textwidth]{figures/4_Xlogng2.png}=
+	\caption{Lognormal distribution and Gaussian contamination}
+	\label{fig:4.02}
+\end{figure}
+
+The figure \ref{fig:4.03} shows an example of the Gaussian distribution $\pmb{Y}_g$ of legitimate observations and the uniform distribution to simulate the addition of anomalies, composing the Gaussian distribution contaminated by uniform noise $\pmb{Y}_g^c$.
 
@@
 	\centering
-	\includegraphics[width=8cm]{figures/ch4/4_Xgu2.png}
+	\includegraphics[width=0.8\textwidth]{figures/4_Xgu2.png}
 	\caption{Example of Gaussian and Uniform Anomalies}
-	\label{fig:4.04}
+	\label{fig:4.03}
 \end{figure}
@@
 We evaluate contamination rates $c$ between 1\% and 50\%, to simulate the imbalanced data of anomaly detection problems during our experiments.
-This data sets simulate a total of 2400 events for each scenario with contaminated Pareto, Lognormal and Gaussian.
+This data set simulate a total of 2400 events for each scenario with contaminated Pareto, Lognormal or Gaussian distribution.
 Therefore, the number of legitimate observations is defined according to the contamination rate selected for each evaluation.
@@
 \begin{table}[h!]
-	\scriptsize
+	\tiny
 	\caption{CTU-13 data set description}
@@
 		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}\textbf{ID}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\textbf{Bot}\end{tabular}	 &\begin{tabular}[x]{@{}l@{}}\textbf{Type}\end{tabular}	&\begin{tabular}[x]{@{}l@{}}\textbf{Total}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Malicious}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{C\&C}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Attack}\end{tabular} &\begin{tabular}[x]{@{}l@{}}\textbf{Normal}\end{tabular}\\ \hline
-			10 & neris &\makecell[l]{IRC, Spam,\\CF} & 2,824,636 & 40,961 (1.45\%) & 341 (0.01\%) & 40,620 (1.44\%) &30,387 (1.07\%)\\ \hline
-			11 & neris &\makecell[l]{IRC, Spam,\\CF} & 1,808,122 & 20,941 (1.16\%) & 673 (0.04\%) & 20,268 (1.12\%) &9,120 (0.5\%)\\ \hline
-			12 & rbot &\makecell[l]{IRC, PS} & 4,710,638 & 26,822 (0.57\%) & 63 (0.00\%) & 26,759 (0.57\%) &116,887 (2.48\%)\\ \hline
-			15 & rbot &\makecell[l]{IRC, DDoS} & 1,121,076 & 2,580 (0.23\%) & 52 (0.00\%) & 2,528 (0.23\%) &25,268 (2.25\%)\\ \hline
-			15-2 & virut &\makecell[l]{Spam, PS,\\HTTP} & 129,832 & 901 (0.69\%) & 24 (0.02\%) & 877 (0.68\%) &4,679 (3.6\%)\\ \hline
+			10 & neris &\makecell[l]{IRC,\\Spam,\\CF} & 2,824,636 & 40,961 (1.45\%) & 341 (0.01\%) & 40,620 (1.44\%) &30,387 (1.07\%)\\ \hline
+			11 & neris &\makecell[l]{IRC,\\Spam,\\CF} & 1,808,122 & 20,941 (1.16\%) & 673 (0.04\%) & 20,268 (1.12\%) &9,120 (0.5\%)\\ \hline
+			12 & rbot &\makecell[l]{IRC,\\PS} & 4,710,638 & 26,822 (0.57\%) & 63 (0.00\%) & 26,759 (0.57\%) &116,887 (2.48\%)\\ \hline
+			15 & rbot &\makecell[l]{IRC,\\DDoS} & 1,121,076 & 2,580 (0.23\%) & 52 (0.00\%) & 2,528 (0.23\%) &25,268 (2.25\%)\\ \hline
+			15-2 & virut &\makecell[l]{Spam,\\PS,\\HTTP} & 129,832 & 901 (0.69\%) & 24 (0.02\%) & 877 (0.68\%) &4,679 (3.6\%)\\ \hline
 			16 & menti &PS & 558,919 & 4,630 (0.83\%) & 199 (0.04\%) & 4,431 (0.79\%) &7,494 (1.34\%)\\ \hline
@@
 			16-3 & murlo &PS & 2,954,230 & 6,127 (0.21\%) & 1,074 (0.04\%) & 5,053 (0.17\%) &72,822 (2.46\%)\\ \hline
-			17 & neris &\makecell[l]{IRC, Spam,\\CF, PS} & 2,087,508 & 184,987 (8.86\%) & 2,973 (0.14\%) & 182,014 (8.72\%) &43,340 (1.57\%)\\ \hline
-			18 & rbot &\makecell[l]{IRC, DDoS} & 1,309,791 & 106,352 (8.12\%) & 33 (0.00\%) & 106,319 (8.12\%) &15,847 (1.2\%)\\ \hline
-			18-2 & rbot &\makecell[l]{IRC, DDoS} & 107,251 & 8,164 (7.61\%) & 2 (0.00\%) & 8,162 (7.61\%) &2,718 (2.53\%)\\ \hline
+			17 & neris &\makecell[l]{IRC,\\Spam,\\CF,\\PS} & 2,087,508 & 184,987 (8.86\%) & 2,973 (0.14\%) & 182,014 (8.72\%) &43,340 (1.57\%)\\ \hline
+			18 & rbot &\makecell[l]{IRC,\\DDoS} & 1,309,791 & 106,352 (8.12\%) & 33 (0.00\%) & 106,319 (8.12\%) &15,847 (1.2\%)\\ \hline
+			18-2 & rbot &\makecell[l]{IRC,\\DDoS} & 107,251 & 8,164 (7.61\%) & 2 (0.00\%) & 8,162 (7.61\%) &2,718 (2.53\%)\\ \hline
 			19 & nsys.ay &P2P & 325,471 & 2,168 (0.67\%) & 25 (0.01\%) & 2,143 (0.66\%) &7,628 (2.35\%)\\ \hline
-			15-3 & virut &\makecell[l]{Spam, PS,\\HTTP} & 1,925,149 & 40,003 (2.08\%) & 536 (0.03\%) & 39,467 (2.05\%) &31,939 (1.65)\\ \hline
+			15-3 & virut &\makecell[l]{Spam,\\PS,\\HTTP} & 1,925,149 & 40,003 (2.08\%) & 536 (0.03\%) & 39,467 (2.05\%) &31,939 (1.65)\\ \hline
 	\end{tabular}
@@
 
-The full data set and scenarios can be denoted as $\pmb{X} = \{\pmb{X}_{10}, \pmb{X}_{11}, \ldots , \pmb{X}_{18-2}, \pmb{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}.
+The full data set and scenarios can be denoted as $\pmb{X} = \{\pmb{X}_{10}, \pmb{X}_{11}, \ldots , \pmb{X}_{18-2},\\ \pmb{X}_{19}\}$, in accordance to IDs presented in Table \ref{tab:4.01}.
 In our experiment each contaminated scenario $\pmb{X}_i$ is split into $\pmb{X}_i^s$ containing 50\% of the legitimate data, and into $\pmb{X}_i^c$ that is composed by all anomalous flows and the necessary number of legitimate flows to have a testing data with the desired contamination rate.
@@
 	\subfigure[State of scenario 16.]{%
-		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110816_State.png}
+		\includegraphics[width=0.47\textwidth]{figures/raw_distplot_capture20110816_State.png}
 		\label{fig:4.05}
@@
 	\subfigure[Destination Type of service of scenario 10.]{%
-		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110810_dTos.png}
+		\includegraphics[width=0.47\textwidth]{figures/raw_distplot_capture20110810_dTos.png}
 		\label{fig:4.06}
@@
 	\subfigure[State of scenario 10.]{%
-		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110810_State.png}
+		\includegraphics[width=0.47\textwidth]{figures/raw_distplot_capture20110810_State.png}
 		\label{fig:4.08}
@@
 	\subfigure[Source Port of scenario 16.]{%
-		\includegraphics[width=0.47\textwidth]{figures/ch4/raw_distplot_capture20110816_Sport.png}
+		\includegraphics[width=0.47\textwidth]{figures/raw_distplot_capture20110816_Sport.png}
 		\label{fig:4.09}
@@
 
-Due to the number of available features and the class overlapping between legitimate and anomalous flows, we performed an correlation analysis and an empirical cross validation in order to identify the best set of features for network attack detection.
+Due to the number of available features and the class overlapping between legitimate and anomalous flows, we performed an correlation analysis and an empirical cross validation to identify the best set of features for network attack detection.
 Therefore, we adopt the following features: state, destination type of service, destination port, source port, total number of packets, total number of bytes and number of bytes from the source.
 
-\section{Moment Distances from Robust Subspace for Network Attack Detection}
-\label{sec:3_proposal}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Moment Distances from Robust Subspace for Botnet Detection}
+\label{sec:mdrs}
 
@@
 Robust subspace learning can be defined as the decomposition of a given data matrix $\pmb{X} \in \mathbb{R}^{M \times N}$, with rows representing observations and columns representing features.
-Note that in Section \ref{sec:2_ModelingData} we introduced $\pmb{X}$ for a packet level analysis, modeled as a matrix of communication port by time.
-However, in this chapter we adopt a flow based analysis, where $\pmb{X}$ is modeled as a matrix of flows by features, and $\pmb{X}$ is decomposed into the sum of a low rank matrix $\pmb{L} \in \mathbb{R}^{M \times N}$, whose column subspace gives the robust principal components without outliers and noise, and a sparse matrix $\pmb{S} \in \mathbb{R}^{M \times N}$, with element-wise outliers or noise.
+In this work we adopt a flow based analysis, where $\pmb{X}$ is modeled as a matrix of flows by features, and $\pmb{X}$ is decomposed into the sum of a low rank matrix $\pmb{L} \in \mathbb{R}^{M \times N}$, whose column subspace gives the robust principal components without outliers and noise, and a sparse matrix $\pmb{S} \in \mathbb{R}^{M \times N}$, with element-wise outliers or noise.
 
-Even though robust subspace learning has been adopted for network anomaly detection by means of highlights from the matrix $\pmb{S}$, it was shown that $\pmb{S}$ can indicate noise and outliers that can not be classified as malicious \cite{vaswani2018robust, lerman2018overview}, resulting into false positive classifications and requiring complementary approaches in order to obtain precise network anomaly detection \cite{zhou2017anomaly}.
+Even though robust subspace learning has been adopted for network anomaly detection by means of highlights from the matrix $\pmb{S}$, it was shown that $\pmb{S}$ can indicate noise and outliers that can not be classified as malicious \cite{vaswani2018robust, lerman2018overview}, resulting into false positive classifications and requiring complementary approaches to obtain precise network anomaly detection \cite{zhou2017anomaly}.
 
 Therefore, for network anomaly detection, we propose to learn a robust subspace from the legitimate traffic $\pmb{X}^s$ for computing $\pmb{L}^s$ and $\pmb{S}^s$, followed by computing the robust moments, i.e.
-the mean $\pmb{\mu}$, skewness $\pmb{\epsilon}$ and kurtosis $\pmb{\kappa}$, in order to evaluate the distance $\pmb{d}$ between contaminated observations $\pmb{X}^c$ and robust moments.
-The largest distances are classified as anomalous and indicate network attacks, denoted as $\pmb{N}$ by the data model $\pmb{X} = \pmb{U} + \pmb{N}$, introduced by (\ref{eq:2.01}), where $\pmb{U}$ denotes the legitimate network traffic.
+the mean $\pmb{\mu}$, skewness $\pmb{\epsilon}$ and kurtosis $\pmb{\kappa}$, to evaluate the distance $\pmb{d}$ between contaminated observations $\pmb{X}^c$ and robust moments.
+The largest distances are classified as anomalous and indicate network attacks, denoted as $\pmb{N}$ by the data model $\pmb{X} = \pmb{U} + \pmb{N}$, where $\pmb{U}$ denotes the legitimate network traffic.
 
@@
 where $\pmb{Y}$ is the multiplier of the linear constraint and $\mu$ is the penalty parameter for the violation of the linear constraint \cite{yuan2009sparse}.
+
 Thus, an iterative scheme can be presented as:
@@
 	\hat{\pmb{\Sigma}} = \displaystyle\frac{1}{N-1}\displaystyle\sum_{i = 1}^{N} (\pmb{x}_i - \pmb{\mu})(\pmb{x}_i - \pmb{\mu})^T,
+\end{equation}
+
+\begin{equation}\label{eq:4.08.1}
+	\frac{1}{N-1} \sum_{i = 1}^{N} I_x1 (x_i - \mu ) _x2 (x_i - \mu) 
 \end{equation}
@@
 
-\section{Experiments}
-\label{sec:3_experiments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments and Results}
+\label{sec:experimentalresults}
 
@@
 In anomalous detection problems, where anomalies are rare events, if one classify all observations as legitimate and apply an accuracy evaluation, one would have high accuracy but poor true-positive detection.
-Due to the importance given by the $F_1$ (also referred as F-score or F-measure) to true-positive detection in scenarios such as the network attack detection, it is the preferable measure for imbalanced data sets \cite{powers2011evaluation,moustafa2019holistic}.
+Due to the importance given by the $F_1$ (also referred as F-score or F-measure) to true-positive detection in scenarios such as the network attack detection, it is the preferable measure for imbalanced data \cite{powers2011evaluation,moustafa2019holistic}.
 Therefore, $F_1$ is the metric used for validation of our experiments.
@@
 
-We also compare the results of our proposals to ROBPCA for anomaly detection on simulated and CTU-13 data sets, considering that ROBPCA also relies on robust estimates with adjusted outlyingness based on robust skewness \cite{hubert2009robustskewed}.
+We also compare the results of our proposals to ROBPCA-AO for anomaly detection on simulated and CTU-13 data set, considering that ROBPCA-AO also relies on robust estimates with adjusted outlyingness based on robust skewness \cite{hubert2009robustskewed}.
 
@@
 The availability of labeled data is a challenging concern in real world problems of anomalous detection, where anomalies are rare or even unknown events.
-Considering that RPCA has already been adopted to isolate outliers from training data \cite{zhou2017anomaly}, we also propose to evaluate the m-RPCA for a contaminated semi-supervised approach based on training from contaminated data, in order to evaluate the impact that contaminated data for robust subspace learning can cause in the anomaly detection results.
+Considering that RPCA has already been adopted to isolate outliers from training data \cite{zhou2017anomaly}, we also propose to evaluate the m-RPCA for a contaminated semi-supervised approach based on training from contaminated data, to evaluate the impact that contaminated data for robust subspace learning can cause in the anomaly detection results.
 
@@
 
-For the semi-supervised approach we propose to learn the robust subspace and compute the moments from the legitimate data sets $\pmb{Y}_g$, $\pmb{Y}_p$ and $\pmb{Y}_l$ with Gaussian, Pareto and Lognormal distributions, respectively, and test the anomaly detection for contaminated data sets $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$.
+For the semi-supervised approach we propose to learn the robust subspace and compute the moments from the legitimate data $\pmb{Y}_g$, $\pmb{Y}_p$ and $\pmb{Y}_l$ with Gaussian, Pareto and Lognormal distributions, respectively, and test the anomaly detection for contaminated data $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$.
 
-For the Contaminated Semi-supervised approach we evaluate the robustness of the m-RPCA approaches for learning from contaminated training data, in order to analyze if m-RPCA can be an alternative for the lack of known legitimate data.
+For the Contaminated Semi-supervised approach we evaluate the robustness of the m-RPCA approaches for learning from contaminated training data, to analyze if m-RPCA can be an alternative for the lack of known legitimate data.
 Therefore we propose to train the model from a contaminated legitimate data $\pmb{Y}_g^{c'}$, $\pmb{Y}_p^{c'}$ and $\pmb{Y}_l^{c'}$, with the same contamination rate of the testing data, but without data repetition between training and testing, taking into account that we shall consider a different contaminated data but adopt the same contamination rate for training and testing.
 
-We finally evaluate the unsupervised approach, that relies on the test data sets $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$ for robust subspace learning, and then classify the results according to the distance between the contaminated data $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$, and the moments computed from the learned robust subspace.
+We finally evaluate the unsupervised approach, that relies on the test data $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$ for robust subspace learning, and then classify the results according to the distance between the contaminated data $\pmb{Y}_g^c$, $\pmb{Y}_p^c$ and $\pmb{Y}_l^c$, and the moments computed from the learned robust subspace.
 
@@
 
-In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, in order to evaluate our proposed approaches to all botnet malwares of the CTU-13 data set.
+In contrast to Garcia \emph{et al.} \cite{garcia2014empirical}, we propose to evaluate each scenario of the CTU-13 data set individually, in a semi-supervised approach that does not rely on training data with labeled anomalies, to evaluate our proposed approaches to all botnet malwares of the CTU-13 data set.
 
@@
 
-\section{Results of Simulated Experiment}
+\subsection{Results of Simulated Experiment}
 \label{sec:3_simulated_result}
@@
 
-We evaluated the $F_1$ of the selected algorithms and m-RPCA approaches for anomaly detection on Gaussian distributed legitimate data contaminated by uniform distributed anomalies.
+Initially, we evaluated the $F_1$ of the selected algorithms and m-RPCA approaches for anomaly detection on Gaussian distributed legitimate data contaminated by uniform distributed anomalies.
 The Figure \ref{fig:4.10} shows the $F_1$ over the contamination rate between 1\% and 50\% for m-RPCA approaches and Isolation Forest (IF) \cite{liu2008isolation}, k-Nearest Neighbors (KNN) \cite{angiulli2002fast}, Local Outlier Factor (LOF) \cite{breunig2000lof}, Minimum Covariance Deteminant (MCD) \cite{rousseeuw1999fastmcd}, One-Class Support Vector Machines (OCSVM) \cite{scholkopf2001estimating}, PCA \cite{shyu2003novel} and ROBPCA-AO \cite{hubert2009robustskewed}.
@@
 	\centering
-	\includegraphics[width=12cm]{figures/ch4/gaussian_f1_contamination.eps}
+	\includegraphics[width=10cm]{figures/gaussian_f1_contamination.eps}
 	\caption{Anomaly detection on Gaussian distributed with uniform anomalies}
@@
 
-It is possible to observe in Figure \ref{fig:4.10} that LOF, PCA, css\_kd-RPCA and u\_kd-RPCA presented lower performance than the remain algorithms, that obtain results higher than 0.95 in average.
+It is possible to observe in Figure \ref{fig:4.10} that LOF, PCA, css\_kd-RPCA and u\_kd-RPCA presented lower performance than the remain algorithms, that obtain results higher than 0.95 in average, for anomaly detection on Gaussian distributed data.
 The exception is the ROBPCA-AO, that presented high score for lower contamination but decreased with the contamination increasing.
@@
 However, the unsupervised versions of md-RPCA and sd-RPCA presented similar results to widely adopted unsupervised algorithms for outlier detection.
-The results also show that the semi-supervised approach of md-RPCA, sd-RPCA and kd-RPCA obtain high anomaly detection rate and presented similar results to the unsupervised algorithms IF, KNN, MCD and OCSVM.
+The results also show that the semi-supervised approach of md-RPCA, sd-RPCA and kd-RPCA obtain high anomaly detection rate on Gaussian distributed data and presented similar results to the unsupervised algorithms IF, KNN, MCD and OCSVM.
 
-The Figure \ref{fig:4.11} and \ref{fig:4.12} show the results for anomaly detection on skewed and heavy tailed distributions.
-The Figure \ref{fig:4.11} shows the results for Pareto with Gaussian anomalies.
+The Figure \ref{fig:4.11} and \ref{fig:4.12} show the results for anomaly detection on skewed and heavy tailed distributions, presenting results for anomaly detection on Pareto and Lognormal distributions, respectively.
 
@@
 	\centering
-	\includegraphics[width=12cm]{figures/ch4/pareto_f1_contamination.eps}
+	\includegraphics[width=10cm]{figures/pareto_f1_contamination.eps}
 	\caption{Anomaly detection on Pareto distributed with Gaussian anomalies}
@@
 
-The results for anomaly detection on Pareto with Gaussian anomalies, depicted by Figure \ref{fig:4.11}, show that IF, KNN, LOF, OCSVM, css\_md-RPCA and u\_md-RPCA performed worse than the remain algorithms for the evaluated contamination, with results lower than 0.6 even with higher contamination.
+The Figure \ref{fig:4.11} depics the results for anomaly detection on Pareto legitimate data with Gaussian anomalies, and shows that IF, KNN, LOF, OCSVM, css\_md-RPCA and u\_md-RPCA performed worse than the remain algorithms for the evaluated contamination, with results lower than 0.6 even with higher contamination.
 Note that the approaches of m-RPCA with lower scores are css\_md-RPCA and u\_md-RPCA, which are mean-based approaches.
@@
 
-Figure \ref{fig:4.11} shows that ROBPCA-AO presented high scores for low contamination rates, presenting results better than ss\_md-RPCA, MCD and PCA, initially.
+Figure \ref{fig:4.11} shows that ROBPCA-AO presented high $F_1$ score for low contamination rates, presenting better results than ss\_md-RPCA, MCD and PCA for anomaly detection on Pareto distributed data, initially.
 However, the results of ROBPCA-AO decrease drastically with the contamination increasing.
+It is also possible to observe in Figure \ref{fig:4.11} that all approaches based on kd-RPCA initially achieved the best results for anomaly detection on Pareto distributed data, but the results for the unsupervised variate with contamination near of 0.2 or higher, while the ss\_kd-RPCA presents stable results, near of 1.0 $F_1$, for all evaluated contamination rates.
 
-It is possible to observe in Figure \ref{fig:4.11} that all approaches based on kd-RPCA achieved the best results initially, but the results for the unsupervised variate with contamination near of 0.2 or higher, while the ss\_kd-RPCA presents stable results near of 1.0 $F_1$ for all evaluated contamination rates.
-
-All approaches based on sd-RPCA obtained anomaly detection higher than 0.8, however the unsupervised and contaminated semi-supervised presented high variation of $F_1$ and lower results in comparison to the semi-supervised sd-RPCA.
-The contaminated semi-supervised approaches of sd-RPCA and kd-RPCA presented results higher than 0.8 and similar to the semi-supervised and unsupervised approaches of sd-RPCA and kd-RPCA.
+All approaches based on sd-RPCA obtained a $F_1$ score higher than 0.8 for anomaly detection on Pareto distributed data, however the unsupervised and contaminated semi-supervised presented high variation of $F_1$ and lower results in comparison to the semi-supervised sd-RPCA.
+The contaminated semi-supervised approaches of sd-RPCA and kd-RPCA presented higher results than 0.8 and similar to the semi-supervised and unsupervised approaches of sd-RPCA and kd-RPCA.
 These results highlight the resilience of the robust subspace learning even for contaminated training data.
@@
 However, u\_kd-RPCA and u\_sd-RPCA presented lower results than ss\_kd-RPCA and ss\_sd-RPCA.
+
 Therefore, the semi-supervised approaches of m-RPCA overcome other approaches of m-RPCA and overcome the selected algorithms for anomaly detection on Pareto distributed data with Gaussian anomalies.
-It is possible to highlight the semi-supervised kd-RPCA, which obtained stable results near of 1.0 $F_1$ for all evaluated contamination.
+Finally, it is possible to highlight the semi-supervised kd-RPCA, which obtained stable results near of 1.0 $F_1$ for all evaluated contamination.
 
-The results for anomaly detection on Lognormal with Gaussian anomalies, depicted by Figure \ref{fig:4.12}, show that IF, KNN, LOF, MCD, OCSVM, css\_md-RPCA, u\_md-RPCA and ROBPCA-AO perform worse than the remain evaluated algorithms, with results lower than 0.6 even with higher contamination.
+The results for anomaly detection on Lognormal legitimate data with Gaussian anomalies is depicted by Figure \ref{fig:4.12}, and show that IF, KNN, LOF, MCD, OCSVM, css\_md-RPCA, u\_md-RPCA and ROBPCA-AO perform worse than the remain evaluated algorithms, with results lower than 0.6, even with higher contamination.
 
@@
 	\centering
-	\includegraphics[width=12cm]{figures/ch4/lognormal_f1_contamination.eps}
+	\includegraphics[width=10cm]{figures/lognormal_f1_contamination.eps}
 	\caption{Anomaly detection on Lognormal distributed with Gaussian anomalies}
@@
 
-The ss\_md-RPCA and PCA algorithms perform similar, with worse detection rate for lower contamination and better performance with contamination higher than 0.4.
+The ss\_md-RPCA and PCA algorithms perform similar for anomaly detection on Lognormal distributed data, with worse detection rate for lower contamination and better performance with contamination higher than 0.4.
 However, the results of ss\_md-RPCA and PCA are lower than all approaches of sd-RPCA and kd-RPCA.
+
 It is important to note that all mean-based approaches of m-RPCA presented lower results for anomaly detection on Lognormal data, in comparison to approaches based on skewness (sd-RPCA) and kurtosis (kd-RPCA), that achieved the best anomaly detection rates.
@@
 The approaches based on kd-RPCA presented higher detection rate for all contamination, with scores near of 1.0.
-The semi-supervised and contaminated semi-supervised approaches for sd-RPCA performed similarly, but the unsupervised approach of sd-RPCA presented lower anomaly detection for lower contamination.
+The semi-supervised and contaminated semi-supervised approaches for sd-RPCA performed similarly, but the unsupervised approach of sd-RPCA presented lower anomaly detection on Lognormal distributed data, for lower contamination.
 
 It is possible to note that the contaminated semi-supervised approaches of kd-RPCA and sd-RPCA performed similar to the semi-supervised approach of the same algorithms.
-These results highlight the resilience of the robust subspace learning even from contaminated training data for anomaly detection on Lognormal data.
+These results highlight the resilience of the robust subspace learning, even from contaminated training data.
 
-Therefore, it is possible to observe that the semi-supervised approaches of m-RPCA overcome other approaches of m-RPCA and all the selected algorithms for anomaly detection on Lognormal distributed data with Gaussian anomalies.
+Finally, it is possible to observe that the semi-supervised approaches of m-RPCA overcome other approaches of m-RPCA and all the selected algorithms for anomaly detection on Lognormal distributed data with Gaussian anomalies.
 
-Following we present the Tables \ref{tab:4.02}, \ref{tab:4.03} and \ref{tab:4.04} to show the results of the selected algorithms and our proposals for anomaly detection on Gaussian, Pareto and Lognormal distributions, with 10\%, 25\% and 33\% of contamination rate.
-The results are the $F_1$ for each scenario and algorithm, sorted by best result by algorithm for anomaly detection on Gaussian, Pareto and Lognormal distributions.
+Following we present the Tables \ref{tab:4.02}, \ref{tab:4.03} and \ref{tab:4.04} to show the results of the selected algorithms and m-RPCA for anomaly detection on Gaussian, Pareto and Lognormal distributions, with 10\%, 25\% and 33\% of contamination rate.
+The columns present the $F_1$ score for each scenario, sorted by the best $F_1$.
 
@@
 	\begin{tabular}{| l | l | l | l |}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian} +\\\textbf{Uniform} \textbf{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian} \\+\textbf{Uniform} \end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto}+\\\textbf{Gaussian} \end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal} \\+\textbf{Gaussian} \end{tabular}\\ \hline
         	Proposed Semi-Supervised sd-RPCA	&0.98	&0.99	&0.97	\\ \hline
@@
         	One-class SVM \cite{scholkopf2001estimating}	&0.98	&0.24	&0.00	\\ \hline
-        	LOF \cite{breunig2000lof}	&0.35	&0.40	&0.36	\\ \hline
         	KNN \cite{angiulli2002fast}	&0.98	&0.08	&0.00	\\ \hline
         	MCD \cite{rousseeuw1999fastmcd}	&0.98	&0.02	&0.00	\\ \hline
+        	LOF \cite{breunig2000lof}	&0.35	&0.40	&0.36	\\ \hline
         	ROBPCA-AO \cite{hubert2009robustskewed}		&0.00	&0.00	&0.00	\\ \hline
@@
 
-The results with 33\% of contamination shows high scores of anomaly detection for Gaussian data with uniform anomalies by the evaluated algorithms, with exception to LOF and ROBPCA-AO.
+The results for 33\% of contamination show high scores of anomaly detection for Gaussian data with uniform anomalies for the evaluated algorithms, with exception of LOF and ROBPCA-AO.
 The semi-supervised approaches of m-RPCA presented the highest results for anomaly detection on Pareto data with Gaussian anomalies, and for anomaly detection on Lognormal data with Gaussian anomalies, as well as for Gaussian data with uniform anomalies.
@@
 	\begin{tabular}{| l | l | l | l |}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian} +\\\textbf{Uniform} \textbf{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian} \\+\textbf{Uniform} \end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto}+\\\textbf{Gaussian} \end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal} \\+\textbf{Gaussian} \end{tabular}\\ \hline
             Proposed Semi-Supervised kd-RPCA    &0.96  &1.00  &0.98  \\ \hline
@@
 
-The Table \ref{tab:4.03} also shows higher scores of semi-supervised m-RPCA for 25\% of contamination, in comparison to the remain compared algorithms.
-It is important to note that the highest score for anomaly detection on Lognormal data with Gaussian contamination was the unsupervised kd-RPCA, with 0.99 $F_1$ score, and that the contaminated semi-supervised approaches of sd-RPCA and kd-RPCA presented scores near of the results for semi-supervised approaches in scenarios with Gaussian, Pareto and Lognormal distributions.
+The Table \ref{tab:4.03} also shows higher $F_1$ scores of semi-supervised m-RPCA for 25\% of contamination, in comparison to the remain evaluated algorithms.
+It is important to note that the highest $F_1$ score for anomaly detection on Lognormal distributed data with Gaussian contamination was the unsupervised kd-RPCA, with 0.99 $F_1$ score, and note that the contaminated semi-supervised approaches of sd-RPCA and kd-RPCA presented $F_1$ scores near of the results for semi-supervised approaches in scenarios with Gaussian, Pareto and Lognormal legitimate data.
 
-The contamination rate of 10\% shown by Table \ref{tab:4.04} shows worse results in comparison to contamination of 25\% and 33\%.
+The results for contamination rate of 10\% shown by Table \ref{tab:4.04} shows worse $F_1$ score in comparison to contamination of 25\% and 33\%.
 However, the ROBPCA-AO presented high scores for anomaly detection on Gaussian data, overcoming LOF, css\_kd-RPCA and u\_kd-RPCA.
-ROBPCA also presented better results for anomaly detection on Pareto data in comparison to u\_md-RPCA, css\_md-RCAP, ss\_md-RCAP, PCA, OCSVM, LOF, KNN and IF.
+ROBPCA-AO also presented better results for anomaly detection on Pareto data in comparison to u\_md-RPCA, css\_md-RCAP, ss\_md-RCAP, PCA, OCSVM, LOF, KNN and IF.
 
@@
 	\begin{tabular}{| l | l | l | l |}
-		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian} +\\\textbf{Uniform} \textbf{($F_1$)}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal} +\\\textbf{Gaussian} \textbf{($F_1$)}\end{tabular}\\ \hline
+		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}c@{}}\textbf{Algorithm}\end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Gaussian} \\+\textbf{Uniform} \end{tabular}	&\begin{tabular}[x]{@{}c@{}}\textbf{Pareto}+\\\textbf{Gaussian} \end{tabular}	 &\begin{tabular}[x]{@{}c@{}}\textbf{Lognormal} \\+\textbf{Gaussian} \end{tabular}\\ \hline
             Proposed Semi-Supervised kd-RPCA    &0.97  &1.00  &0.95  \\ \hline
@@
 
-Taking into account the null hypothesis $H_2^{(N)}$ and the presented results, we can conclude that the robust subspace learning, adopted by md-RPCA, sd-RPCA and kd-RPCA, presented higher anomaly detection from imbalanced and skewed data than widely adopted algorithms for anomaly detection.
-Therefore, the presented results refute the null hypothesis $H_2^{(N)}$, which defines that the robust subspace learning does not improves the anomaly detection from imbalanced and skewed data.
+Therefore, it is possible to observe that the md-RPCA, sd-RPCA and kd-RPCA approaches, which are based on robust subspace learning, presented higher anomaly detection from imbalanced and skewed data, in comparison to widely adopted algorithms for anomaly detection.
 
-\section{Results of CTU-13 Experiment}
+\subsection{Results of CTU-13 Experiment}
 \label{sec:3_ctu13_result}
@@
 
-For this experiment we only evaluate the semi-supervised approach of md-RPCA, sd-RPCA and kd-RPCA, considering that the semi-supervised algorithms presented the best results on the simulated experiment and taking into account the observed resilience of the semi-supervised approach when the training data is contaminated.
+For this experiment we only evaluate the semi-supervised approach of md-RPCA, sd-RPCA and kd-RPCA, considering that these semi-supervised algorithms presented the best results on the simulated experiment and taking into account the observed resilience of the semi-supervised approach when the training data is contaminated.
 Additionally, the semi-supervised approach only requires the robust subspace learning for training, what can indicate less computational cost for network anomaly detection on new observations.
@@
 
-\begin{table}[h!]
+The Table \ref{tab:4.03} present the $F_1$ of each algorithm for all scenarios of CTU-13, sorted by the average $F_1$ of each algorithm for all scenarios, which is presented in the last column.
+
+\begin{table}[t]
   \centering
-  \scriptsize
+  \tiny
   \caption{Network anomaly detection from CTU-13 with 33\% of contamination}
@@
 
-The Table \ref{tab:4.03} present the $F_1$ of each algorithm for all scenarios of CTU-13, and the last column presents the average $F_1$ of each algorithm for all scenarios.
-
 It is possible to observe that md-RPCA, kd-RPCA and sd-RPCA overcome the remain algorithms in average results for all scenarios, according to the column \textbf{avg}.
 The sd-PRCA presented an average of 0.57 while md-RPCA obtained 0.81 and kd-RPCA 0.74 in average.
-The PCA algorithm performed similar to sd-RPCA in average, with results of 0.56 and 0.57, respectively.
+The PCA based algorithm performed similar to sd-RPCA in average, with results of 0.56 and 0.57, respectively.
 However the results of PCA and sd-RPCA for scenarios 12, 18-2 and 19 presented a large variation.
@@
 The md-RPCA showed the best anomaly detection for 10 scenarios of a total with 13 scenarios.
-In the scenario 15 the best result was obtained by kd-RPCA, for the 15-2 scenario the best result was for MCD, and PCA was the best algorithm for scenario 19.
-Even thought md-RPCA not be the best result for scenarios 15, 15-2 and 19, the results of md-RPCA are very close to the best results for scenarios 15 and 15-2.
+For the scenario 15, the best result was obtained by kd-RPCA, for the 15-2 scenario the best result was for MCD, and PCA was the best algorithm for scenario 19.
+Even thought md-RPCA not be the best result for scenarios 15, 15-2 and 19, the results of md-RPCA are very close to the best results obtained for scenarios 15 and 15-2.
 
@@
 However, CTU-13 is one of the up-to-date data sets for network attack detection and is one that provides the data imbalance observed in real anomaly detection problems.
-Unfortunately, was not possible to observe Pareto or Lognormal distributions on features of CTU-13, even though the finds of researches that highlight the fitting between these distributions and Internet traffic \cite{benson2010network, leon2017probability}.
+Unfortunately, was not possible to observe Pareto or Lognormal distributions on features of CTU-13, even though the finds of researchers that highlight the fitting between these distributions and Internet traffic \cite{benson2010network, leon2017probability}.
 
-The results of anomaly detection on CTU-13 reveals that m-RPCA algorithms obtain good results for network attack detection on contaminated data, overcoming widely adopted algorithms for outlier detection.
+The results of anomaly detection on CTU-13 reveals that m-RPCA algorithms obtain good results for botnet attack detection on contaminated data, overcoming widely adopted algorithms for outlier detection.
 
-\section{Conclusion}
-\label{sec:3_conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion and Future Works}
+\label{sec:conclusionandfutureworks}
 
-This work proposed the m-RPCA, which is approaches based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data.
-We evaluated the anomaly detection rate of m-RPCA for simulated data and for the CTU-13 data set, which is composed by network traffic of botnets, attacks and background flows.
+This paper proposed the m-RPCA, which a framework based on distances of moments computed from a robust subspace learned by RPCA, for anomaly detection on imbalanced and skewed data.
+The m-RPCA can be divided into md-RPCA, sd-RPCA and kd-RPCA algorithms, to denote approaches of m-RPCA based on distances of mean, skewness and kurtosis, respectively.
 
-The m-RPCA can be divided into md-RPCA, sd-RPCA and kd-RPCA algorithms, to denote approaches of m-RPCA based on distances of mean, skewness and kurtosis, respectively.
-We also proposed to evaluate m-RPCA for semi-supervised, contaminated semi-supervised and unsupervised methods of anomaly detection.
+We proposed approaches of m-RPCA for semi-supervised, contaminated semi-supervised and unsupervised methods of anomaly detection, and evaluated the anomaly detection score of m-RPCA for simulated data and for the CTU-13 data set, which is composed by network traffic of botnets, attacks and background flows.
 
@@
 The results also show that the m-RPCA can be adopted for network attack detection, with better results than widely adopted algorithms for outlier detection on the CTU-13 data set.
-Therefore, the observed results rejects the null hypothesis $H_1^{(N)}$ and confirms the alternative
-hypothesis $H_2^{(A)}$.
 
-The results show that the semi-supervised approach for m-RPCA obtained better results than the contaminated semi-supervised and unsupervised approaches, and that m-RPCA presented good results even when the semi-supervised training was computed from contaminated data.
-This highlights the resilience of the robust subspace learning for the semi-supervised anomaly detection approach to deal with possible lack of known legitimate data for training.
+Future research can be directed to evaluate the application of the proposed approaches to different data sets and to anomaly detection problems.
+The m-RPCA can also be extended to be online and to learn new subspace in an adaptive fashion, by means of new robust subspace learning from new observations or through robust subspace tracking.
+Future works can also consider to evaluate the sparse matrix $\pmb{S}$ and its sparsity \cite{liu2017tensor} for m-RPCA, to improve the accuracy and to be able to have element-wise anomaly detection.
 
-The main contributions of this work were: the development and evaluation of novel approaches for anomaly detection on skewed and imbalanced data sets, by means of moments computed from a robust subspace learned by RPCA; the evaluation of the proposed approaches for anomaly detection on simulated data set and for network attack detection on real botnet traffic data set.
+\bibliography{references}
 
-Future research can be directed to evaluate the application of the proposed approaches to different data sets and anomaly detection problems.
-The m-RPCA can also be extended to be online and to learn new subspace in an adaptive fashion, by means of new robust subspace learning from new observations or through robust subspace tracking.
-
-In this work we adopt a multivariate analysis based on subspace learning from matrix decomposition, but a multidimensional approach can be evaluated in order to exploit the relationship between three or mode dimensions.
-Therefore, future works can evaluate the contribution of a robust tensor-based data modeling \cite{da2008robust, gomes2017tensor}, tensor decomposition \cite{gomes2019tensor} and multidimensional analysis for m-RPCA in problems of anomaly detection and network attack detection.
-
-Future works can also consider to evaluate the sparse matrix $\pmb{S}$ and its sparsity \cite{liu2017tensor} for m-RPCA, in order to improve the accuracy and to be able to have element-wise anomaly detection.
+\end{document}
